### 2018-01-31 upgm.tex
[upgm] very short intro, heteromorphisms, representability

\documentclass{article}

\usepackage[margin=1in]{geometry}
\usepackage{tikz-cd}
\usepackage{amssymb}
\usepackage{amsmath}

\begin{document}

\newtheorem{theorem}{Theorem}
\newtheorem{definition}{Definition}
\newcommand{\pto}{\nrightarrow}
\newcommand{\pfrom}{\nleftarrow}
\newcommand{\vcat}{\mathcal}
\newcommand{\cat}{\mathbb}

\title{Universal Properties in Generalized Multicategories}
\author{Max S. New}

\maketitle

In this paper, we seek to develop some theory of right and left
universal properties at the abstract level of generalized
multicategories.
%
Our goal is to ease the development of new type theories with sound
and complete categorical semantics based on Martin-Lof's notion of
judgmental structure.
%
In effect we are abstracting the Curry-Howard-Lambek correspondence as
follows:

\begin{tabular}{|c|c|}
  \hline
  Syntax/Type Theory & Semantics/Category Theory\\ \hline \hline
  Judgmental Structure & Generalized Multicategory\\ \hline
  Negative Type & Representable Generic Right Universal Property\\ \hline
  Positive Type & Representable Generic Left Universal Property\\ \hline  
\end{tabular}

In the process we will provide definitions and theorems that should
free the working semanticist from proving some tedious theorems and
provide guidance for the development of new type theories.
%
In particular, our notions of generic right and left-universal
property lead to a transparent definition of the category of models in
which the type theory presents the initial model.
% TODO: actually see if we *can* prove the following LOL
Further, we provide a quite abstract ``presentation parsimony''
theorem that shows that type theories presenting functorial
constructions can \emph{derive} the functoriality from the
introduction and elimination rules.

\section{Horizontal Profunctors and Representability}

In ordinary category theory there is no need for a separate notion of
right- and left-universal property since categories satisfy perfect
duality: the opposite of a category is a category and so a
left-universal property is defined as a right-universal property in
the opposite category.
%
However multicategories are assymetric by default, the source is given
by a context and the target by a single object, so the opposite of a
multicategory is not a multicategory but a poycategory.
%
It is conceivable that the more general framework of polycategories
will restore this duality, but for now we will work with the
asymmetric notion, which will help to develop intuition in the worst
case.

Despite the lack of perfect duality, we can abstract over some of the
commonalities between right and left universal properties.
%
Right universal properties are right modules and left universal
properties are left modules, but rather than being modules between
categories or between multicategories, the modules will have one side
be a category and the other a multicategory.
%
This structure is generally modelled by a horizontal profunctor
between the virtual equipments of categories and multicategories.

\begin{definition}[Horizontal Profunctor]
  Given virtual double categories $\vcat{V},\vcat{U}$, a
  \emph{horizontal profunctor} $H : \vcat V \pfrom \vcat U$ is for
  each $v \in \vcat V, u \in \vcat U$ a set $H(v,u)$ of ``horizontal
  heteromorphisms'' from $u$ to $v$, so written $v \pfrom u$.
  Additionally a notion of ``hetero-2-cell'' with shapes:

  \begin{tikzcd}
v_n \arrow[dd, "f"] & \cdots \arrow[l, "R_n"] & v_0 \arrow[l, "R_1"] & u_0 \arrow[l, "h"] & \cdots \arrow[l, "Q_1"] & u_m \arrow[l, "Q_m"] \arrow[dd, "g"] \\
 &  &  &  &  &  \\
v' &  &  &  &  & u' \arrow[lllll, "h'"]
  \end{tikzcd}

  that can be composed with other hetero-2-cells and 2-cells in $\vcat
  V, \vcat U$ in the evident way, satisfying the evident unitality and
  associativity laws.
\end{definition}

We draw diagrams for horizontal profunctors the same way we draw them
for virtual double categories, this is justified by the following
definition, in which they are just diagrams in a virtual double category:

\begin{definition}[Cograph of a Horizontal Profunctor]
  For any horizontal profunctor $H : \vcat V \pfrom \vcat U$ we
  construct the cograph $V +_{H} U$ to have as objects $V_0 + U_0$,
  vertical arrows only within $V,U$ and horizontal morphisms defined
  as:
  \begin{align*}
    (V +_{H} U)(v,v') &= V(v,v')\\
    (V +_{H} U)(u,u') &= U(u,u')\\
    (V +_{H} U)(v,u) &= H(v,u)\\
    (V +_{H} U)(u,v) &= \emptyset
  \end{align*}
  With 2-cells inherited from $V,U,H$.
\end{definition}

I'm not sure if it's possible to define a virtual equipment of virtual
double categories, double functors and \emph{horizontal} profunctors,
but we will need at least the following very simple kind of 2-cell in
this imagined equipment of the following shape:

\begin{tikzcd}
 & W \arrow[ldd, "F_V"] \arrow[rdd, "F_U"] &  \\
 & F &  \\
V &  & U \arrow[ll, "H"]
\end{tikzcd}

\begin{definition}[Parameterized Element of a Profunctor]
  Let $\vcat {V,U}$ be virtual double categories, $\cat W$ be a
  category and let $H \vcat V \pfrom \vcat U$ be a profunctor.

  Then a $\cat W$-parameterized element of $H$ consists of
  \begin{enumerate}
  \item Functors $F_U : \cat W \to (\vcat U)_v$ and $F_{\vcat V} :
    \cat W \to (\vcat V)_v$.
  \item For each $w \in W$, a heteromorphism $F(w) : F_{\vcat V} w
    \pfrom F_{\vcat W} w$
  \item For each $f : w \to w'$, a 2-cell
    \begin{tikzcd}
F_{V} w \arrow[dd, "F_{V} f"] &  & F_{U} w \arrow[ll, "Fw"] \arrow[dd, "F_{U}f"] \\
 & Ff &  \\
F_{V} w' &  & F_{U} w' \arrow[ll, "Fw'"]
    \end{tikzcd}
    such that the assignment is functorial (preserve identity,
    composition).
  \end{enumerate}
\end{definition}

Now we can define a general notion of representability in a
profunctor.
%
Representability is a formulation of a (parameterized) type having a
universal property: the horizontal morphism is the universal property
and the 2-cell means ``maps into/out of this object look like the
universal property''.
%
In a virtual double category representability is a relationship
between horizontal and vertical morphisms with the same boundary, but
our horizontal profunctor doesn't give us a notion of vertical
morphism.
%
It might be possible to develop a more general definition of
profunctor, but instead for now we will limit to the following case:

\begin{definition}{Representable Heteromorphism}
  Let $H : \vcat V \pfrom \vcat U$ be a horizontal profunctor and
  $\bar H : \vcat U \to \vcat V$ be a functor with a $(\vcat U)_v$
  parameterized element of $H$:
  \[
  \begin{tikzcd}
 & U_v \arrow[ldd, "\bar H"'] \arrow[rdd, "id"] &  \\
 & \bar H &  \\
V &  & U \arrow[ll, "H"]
\end{tikzcd}
  \]

  Then a representation of a heteromorphism $h : u \pfrom v$ consists
  of a vertical arrow $\bar h : v \to \bar H u$ and a 2-cell
  \[
  \begin{tikzcd}
u \arrow[dd, "id"] &  & v \arrow[dd, "\bar h"] \arrow[ll, "h"] \\
 & \bar h &  \\
u &  & \bar H u \arrow[ll, "\bar H u"]
  \end{tikzcd}
  \]
  that is cartesian in $\vcat V +_{H} \vcat U$.
\end{definition}

Below, we think of all heteromorphisms as expressing
\emph{specifications} for vertical morphisms in this way, which is why
it is thought of as a ``property'' that an object may or may not have.

\section{Concrete and Generic Right Universal Properties}

\section{Left Universal Properties}

\end{document}

%% Local Variables:
%% compile-command: "pdflatex upgm.tex"
%% End:

### 2018-01-31 upgm.tex
right modules: problem(!) we need vertical and horizontal heteros

@@
 \newcommand{\cat}{\mathbb}
+\newcommand{\vtkmnd}{\mathbb{K}\text{Mod}(\vcat{V},T)}
+\newcommand{\rmod}{\text{RMod}}
 
@@
 
+A right universal property is a specification for a parameterized
+object by defining what maps \emph{into} the object look like.
+%
+These are simpler to formalize for generalized multicategories because
+we have less interaction with the context as given by the monad $T$.
+%
+If our generalized multicategories are defined as kleisli monoids in
+$\vcat V$ with monad $T$, then a right universal property on a
+``multicategory'' $\vcat M \in \vtkmnd$ will be parameterized by a
+``category'' $\cat C \in \vcat V$ and is defined as an element of the
+following profunctor:
+
+\begin{definition}{Profunctor of Right Modules}
+  The profunctor of category-parameterized right modules on
+  multicategories $\rmod : \vtkmnd \pfrom \vcat V$ has as elements
+  $\vcat M \pfrom \cat C$ \emph{right modules} which consist of
+  \begin{enumerate}
+  \item A profunctor $Q : T \vcat M \pfrom \cat C$ in $\vcat V$.
+  \item An action
+    \[
+\begin{tikzcd}
+T^2M_0 \arrow[dd, "\mu"] & T M_0 \arrow[l, "T(M_1)"] & C \arrow[dd, "id"] \arrow[l, "Q"] \\
+ & Qcut &  \\
+TM_0 &  & C \arrow[ll, "Q"]
+\end{tikzcd}
+\]
+  \item That is associative:
+    \[
+    \begin{tikzcd}
+T^3M_0 \arrow[dd, "\mu"] &  & T^2M_0 \arrow[dd, "\mu"] \arrow[ll, "T^2(M_1)"] & T M_0 \arrow[l, "T(M_1)"] & C \arrow[dd, "id"] \arrow[l, "Q"] &  & T^3M_0 \arrow[dd, "\mu"] & T^2M_0 \arrow[l] & TM_0 \arrow[l] \arrow[dd, "id"] & C \arrow[l, "Q"] \arrow[dd, "id"] \\
+ & \mu &  & Qcut &  &  &  & T(M_1) &  &  \\
+T^2M_0 \arrow[dd, "\mu"] &  & TM_0 \arrow[ll, "T(M_1)"] &  & C \arrow[ll, "Q"] \arrow[dd, "id"] & = & T^2M_0 \arrow[dd, "\mu"] &  & TM_0 \arrow[ll, "T(M_1)"] & C \arrow[l, "Q"] \arrow[dd, "id"] \\
+ &  & Qcut &  &  &  &  &  & Qcut &  \\
+TM_0 &  &  &  & C \arrow[llll, "Q"] &  & TM_0 &  &  & C \arrow[lll, "Q"]
+\end{tikzcd}
+    \]
+  \item And unital:
+    \[\begin{tikzcd}
+ & TM_0 \arrow[ld, "T\eta"] \arrow[d, "id"] &  & C \arrow[ll, "Q"] \arrow[d, "id"] &  &  &  \\
+T^2M_0 \arrow[dd, "\mu"] & TM_0 \arrow[l, "M_1"] &  & C \arrow[ll, "Q"] \arrow[dd, "id"] & = & TM_0 \arrow[d, "id"] & C \arrow[d, "id"] \arrow[l, "Q"] \\
+ &  & Qcut &  &  & TM_0 & C \arrow[l, "Q"] \\
+TM_0 &  &  & C \arrow[lll, "Q"] &  &  & 
+    \end{tikzcd}\]
+  \end{enumerate}
+
+  And a 2-cell with shape 
+  \[\begin{tikzcd}
+{\vcat M}_n \arrow[dd, "f"] & \cdots \arrow[l, "R_n"] & {\vcat M}_0 \arrow[l, "R_1"] & {\cat C}_0 \arrow[l, "Q"] & \cdots \arrow[l, "P_1"] &  {\cat C}_m \arrow[l, "P_m"] \arrow[dd, "g"] \\
+ &  &  &  &  &  \\
+{\vcat M}' &  &  &  &  & {\cat C}' \arrow[lllll, "Q'"]
+  \end{tikzcd}\]
+  is given by a 2-cell in $\vcat V$ with shape:
+  \[\begin{tikzcd}
+T^{n+1}\vcat{M}_n \arrow[d, "\mu^n"] & \cdots \arrow[l, "T^nR_n"] & T\vcat{M}_0 \arrow[l, "TR_1"] & \cat C_0 \arrow[l, "Q"] & \cdots \arrow[l, "P_1"] & \cat C_m \arrow[l, "P_m"] \arrow[dd, "g"] \\
+T\vcat{M}_n \arrow[d, "Tf"] &  &  &  &  &  \\
+T\vcat{M}' &  &  &  &  & \cat C' \arrow[lllll, "Q'"]
+\end{tikzcd}
+  \]
+
+  with composition defined in the obvious way and unitality and
+  associativity properties of 2-cells following from the unitality and
+  associativity properties of the actions.
+\end{definition}
+
+\subsection{Examples}
+
+The most basic is the identity module which is given by $M_1 : TM_0
+\pfrom M_0$ itself. We will use this for representability next.
+
+If $\vcat V$ is cartesian or even virtually cartesian, we can define
+the \emph{product} property as $\Delta^{*}(M_1\times M_1) : TM_0
+\pfrom M_0\times M_0$ and the \emph{cartesian unit} as $!^*(Id_{1}) :
+TM_0 \pfrom 1$.
+
+We will need some technology from the next section to define a general
+notion of \emph{closed} multicategory, but we can define the property
+of $A \to B$ concretely in many different multicategories.
+
+\subsection{Representability}
+
+As noted above, the hom of a generalized multicategory can be seen as
+a module $M_1 : T M_0 \pfrom M_0$ and we can make this into a
+parameterized element of $\rmod$ of shape:
+  \[
+  \begin{tikzcd}
+ & \vtkmnd_v \arrow[ldd,"id"] \arrow[rdd, "-_{0}"] &  \\
+ & -_1 &  \\
+\vtkmnd &  & \vcat V \arrow[ll, "\rmod"]
+\end{tikzcd}
+  \]
+and then we can define a representable negative universal property as
+just a representable right module with respect to $-_1$.
+
+In more detail a representation of $Q : \vcat M \pfrom \cat C$ is a
+2-cell:
+  \[
+  \begin{tikzcd}
+TM_0 \arrow[dd, "id"] &  & C \arrow[dd, "G_Q"] \arrow[ll, "Q"] \\
+ & I_Q &  \\
+TM_0 &  & M_1 \arrow[ll, "M_1"]
+  \end{tikzcd}
+  \]
+that is cartesian in the appropriate sense.
+%
+Returning to type-theoretic intuition, this square is the
+\emph{introduction rule}: above the line we have the ``formal''
+product or functin etc, and below the line we have the
+``internalized'' notion.
+%
+The associativity rule gives the definition of substitution for the
+introductin rule.
+%
+We can define an ``elimination'' rule using the cartesian property of
+$I_Q$, except \emph{we can't we need a notion of vertical morphism from cat to multicat(!)}:
+
+\[\begin{tikzcd}
+C \arrow[d, "F"] &  & C \arrow[ll, "\to_C"] \arrow[dd, "id"] &  & C \arrow[dd, "F"] &  & C \arrow[dd, "F"] \arrow[ll, "=_C"] \\
+M_0 \arrow[d, "\eta"] & \epsilon &  &  &  & F &  \\
+T(M_0) \arrow[dd, "id", two heads, tail] &  & C \arrow[ll, "|-^P"] \arrow[dd, "F"] & = & M_0 \arrow[dd, "\eta"] &  & M_0 \arrow[dd, "id"] \arrow[ll, "=_{M_0}"] \\
+ & \iota &  &  &  & Var &  \\
+T(M_0) &  & M_0 \arrow[ll, "\vdash"'] &  & T(M_0) &  & M_0 \arrow[ll, "|-"]
+\end{tikzcd}
+\]
+
 \section{Left Universal Properties}

### 2018-02-01 upgm.tex
most of what I have on left universal props, need 2-cells

@@
 \newcommand{\rmod}{\text{RMod}}
+\newcommand{\lmod}{\text{LMod}}
 
@@
   \begin{enumerate}
-  \item A profunctor $Q : T \vcat M \pfrom \cat C$ in $\vcat V$.
+  \item A profunctor $Q : T \vcat M_0 \pfrom \cat C$ in $\vcat V$.
   \item An action
@@
 
+For left-universal properties, consider sums as an illustrative
+example.
+%
+In a category $\cat C$, a sum/coproduct $A + B$ is defined by the property that
+\[ \cat C(A + B, C) \equiv \cat C(A,C) \times \cat C(B,C) \]
+%
+However, a sum in a multicategory $\vcat C$ should \emph{not} just be
+a sum in the underlying category (unlike for right universal
+properties).
+%
+Instead it is what's called a \emph{distributive} coproduct, which can
+be defined manually by some laws relating compositions of the
+coproduct with the monoidal structure.
+%
+We want to find an abstract setting that captures these distributivity
+properties, which already have a type-theoretic explanation: the
+equivalence happens ``in an arbitrary context'':
+\[ \vcat C(\Gamma,A + B; C) \equiv \vcat C(\Gamma, A; C) \times C(\Gamma,B;C)\]
+this can be defined using fibrations but we will try a different,
+hopefully more direct tactic.
+
+We define left universal properties as saying that maps out of a
+context with the object are given by a certain module. For this to
+make sense we need a concept of ``context with an object missing'',
+and after McBride and Combinatorialists we will call this the
+``derivative'' of the context. Since the shape of contexts is given by
+a monad, we would like a concept that applies to an arbitrary monad,
+not just polynomial functors. I haven't nailed it down but it looks
+like we need the following at least.
+
+The microcoms principle is here inescapable and so we need an extra
+condition of $\vcat V$ being a \emph{monoidal} virtual double
+category, which we will here weaken to a \emph{virtually monoidal}
+double category which might be called a ``virtual multi-double
+category'' or something like that.
+
+\begin{definition}[Derivative of a Monad]
+  Let $T$ be a monad on a virtually monoidal virtual equipment $\vcat
+  V$. A \emph{derivative} of $T$ is an endofunctor $dT : \vcat V \to
+  \vcat V$ such that/with
+  \begin{enumerate}
+  \item $dT$ is a vertical monoid in $\vcat V$, i.e. there are natural
+    monoidal product $m_A : dT(A),dT(A) \to dT(A)$ and unit $e_A :
+    \cdot \to dT(A)$.
+  \item $dT$ is a left-module of $T$, i.e. there is a vertical action
+    $d\mu : dT \circ T \Rightarrow dT$ that interacts with $\mu,\eta$
+    in the obvious way.
+  \item A ``plugging in'' morphism $dT,Id \Rightarrow T$, i.e., $l :
+    dT(A),A \to TA$.
+  \item $\lambda l : dT \to (Id \to T)$ is a monoid homomorphism
+    where $(Id \to T)$ uses kleisli composition
+  \item The two ways of plugging $dT(T A), A \to T A$ are equal:
+    \[\begin{tikzcd}
+ & dT(TA),A \arrow[ld, "{dT(\eta),id}"'] \arrow[rd, "{id,\eta}"] &  \\
+dT(A),A \arrow[rdd, "l_A"] &  & dT(TA),TA \arrow[d, "l_{TA}"] \\
+ &  & T^2A \arrow[ld, "\mu"] \\
+ & TA & 
+    \end{tikzcd}\]
+  \item Possibly other laws...
+  \end{enumerate}
+\end{definition}
+What other conditions we have no clue unfortunately. For instance all
+of the above are satisfied by the trivial functor $dT(A) = 1$
+
+Given a derivative $dT$ of $T$, we define the profunctor of
+\emph{left} modules as follows:
+
+\begin{definition}[Profunctor of Left Modules]
+  The profunctor of category-parameterized left modules on
+  multicategories $\lmod : \vcat V \pfrom \vtkmnd$ has as elements
+  $\cat C \pfrom \vcat M$ \emph{left modules} which consist of
+  \begin{enumerate}
+  \item A profunctor $P : dT(\vcat M_0) \otimes \cat C \pfrom \vcat M_0$ in $\vcat V$
+  \item An action \[\begin{tikzcd}
+dT(TM_0) \times (dT(M_0) \times C) \arrow[dd, "l"] &  & dT(M_0) \times C \arrow[ll, "dT(M_1) \times P"] &  & M_0 \arrow[dd, "id"] \arrow[ll, "P"] \\
+ &  & Pcut &  &  \\
+dT(M_0) \times C &  &  &  & M_0 \arrow[llll, "P"]
+  \end{tikzcd}\]
+  \item That is associative
+  \item Unital
+  \item 2-cells are defined as...
+  \end{enumerate}
+\end{definition}
+
 \end{document}

### 2018-02-01 scratch.hs
some code for defining "free derivatives"

@@ -1,0 +1,21 @@
+{-# LANGUAGE GADTs #-}
+{-# LANGUAGE KindSignatures #-}
+module Main where
+
+import Control.Monad (join)
+main = print ()
+
+data DT (m :: * -> *) a where
+  E :: DT m a
+  M :: DT m a -> DT m a -> DT m a
+  A :: DT m (m a) -> DT m a
+
+instance Functor m => Functor (DT m) where
+  fmap f E = E
+  fmap f (M g d) = M (fmap f g) (fmap f d)
+  fmap f (A gg)  = A (fmap (fmap f) gg)
+
+plug :: Monad m => DT m a -> a -> m a
+plug E       x = return x
+plug (M g d) x = plug g =<< plug d x
+plug (A gg)  x = join (plug gg (return x))

### 2018-02-07 upgm.tex
[gen. multicats] some small edits based on recent thoughts

@@
   \]
-that is cartesian in the appropriate sense.
+that is cartesian in the appropriate sense and a module
+\emph{homomorphism} in that the introduction rule turns $Q$-cuts into
+$M$-cuts.
 %
@@
 %
-The associativity rule gives the definition of substitution for the
+The homomorphism property gives the definition of substitution for the
 introductin rule.
@@
 \[ \vcat C(\Gamma,A + B; C) \equiv \vcat C(\Gamma, A; C) \times C(\Gamma,B;C)\]
-this can be defined using fibrations but we will try a different,
-hopefully more direct tactic.
+the usual approach is to use fibrations, and our approach is similar,
+but more general and we hope more conceptually direct.
 

### 2020-02-07 universal-properties/upgm.tex
delete failed agda syntax attempt and add note about polynomials

--- 2018-02-07 upgm.tex
@@
 
+\section{Polynomial Multicategories}
+
+$T$-multicategories for an arbitrary monad on a virtual equipment are
+maybe too general. Things get a bit more concrete if we consider the
+case where $T$ is a \emph{polynomial monad}, and we can get a toy
+theory by considering polynomials in \emph{one} variable, i.e., spans.
+
+In this case, we consider the double category of polynomials in some
+rich category $E$, whose 2-cells are given by cartesian
+morphisms. Then we can define a \emph{flavor} of multicategory to just
+be some (horizontal) monad $T : I \pto I$ in that double category, and
+can define a $T$-multicategory to be a monad in the slice over
+$T$. This makes our monad $T$ much more concrete: the object $I$ is an
+object of \emph{sorts}, each ``shape'' gives us a \emph{judgment} of
+our multicategory, and the positions in the shape tell us what the
+free variables are. The monad structure gives us identity judgments
+and the different cuts available between judgments (i.e., all of
+them).
+
 \end{document}

### 2020-02-13 impl/src/Grammar.hs
module stuff, re-doing impl

@@ -1,254 +1,276 @@
 module Grammar where
 
--- | A Program is a sequence of signature and module definitions.
-newtype Program = Program { _defs :: [TopLevel] }
+-- | A Program is just a module
+type Program = ModuleBody
+
+-- | A Module is a sequence of statements
+newtype ModuleBody = ModuleBody { _defs :: [Decl ScopedExp] }
   deriving (Show, Read, Eq)
-data TopLevel = TLSig SigDef | TLMod ModDef
+
+-- | A statement is a definition of a signature, module or term or an
+-- assertion of an equality (between transformations?)
+data ScopedExp
+  = ScSig SigExp
+  | ScModule ModExp
+  | ScSet SetExp
+  | ScFun ScopedEltExp
+  | ScSpan ScopedSpanExp
+  | ScTrans ScopedTransExp
+  | ScAssert ScopedProofExp
   deriving (Show, Read, Eq)
 
-nameOf :: TopLevel -> String
-nameOf (TLSig sigdef) = _sigDefName sigdef
-nameOf (TLMod moddef) = _modDefName moddef
+data Decl a = Decl { _name :: String , _defn :: a }
+  deriving (Show, Read, Eq)
+  
+type Ctx = [Decl ScopedExp]
 
-type CheckingEnv = [(ModName, SigExp)]
+-- First order lambda calculus
+data SigExp
+  = SigBase GroundSigExp
+  | SigApp GroundSigExp [Exp]
+  deriving (Show, Read, Eq)
 
-data SigDef = SigDef { _sigDefName :: SigName
-                     , _sigLambda  :: SigLambda
-                     }
+data GroundSigExp
+  = GSigVar String
+  | GSigLam SigLambda
   deriving (Show, Read, Eq)
 
 data SigLambda = SigLam
-  { _sigLamArgs :: [(ModName, SigExp)]
-  , _sigBody :: [SigDecl]
+  { _sigLamArgs :: [Decl Generator]
+  , _sigBody :: [Decl Generator]
   }
   deriving (Show, Read, Eq)
 
--- | An expression that denotes a signature, must be the application
--- of a defined signature to concrete module arguments
-type SigExp = SigApp
-
-seapp = SigApp
-
-data SigApp = SigApp { _sigCtor :: HOSigExp, _sigAppArgs :: [ModExp] }
+data Generator
+  = GenMod SigExp -- "abstract module"
+  | GenSet -- "base type"
[…]

### 2020-02-13 impl/src/Parse.hs

@@ -1,5 +1,6 @@
 module Parse where
 
+
 import Control.Monad
 import Control.Monad.Identity
 import Text.Parsec
@@ -12,15 +13,17 @@ type Parse = Parsec String ()
 
 sexp_lang :: Monad m => GenLanguageDef String () m
 sexp_lang = P.LanguageDef
-  { P.commentStart = ""
-  , P.commentEnd = ""
-  , P.commentLine = "--"
-  , P.nestedComments = False
+  { P.commentStart = "(*"
+  , P.commentEnd = "*)"
+  , P.commentLine = ""
+  , P.nestedComments = True
   , P.identStart = identChar
   , P.identLetter = identChar <|> digit <|> char '\''
   , P.opStart = mzero
   , P.opLetter = mzero
-  , P.reservedNames = ["module", "signature", "where", "end", "set", "span", "fun", "term"]
+  , P.reservedNames = ["module", "signature", "where", "set", "span", "fun", "term", "trans",
+                       "def-sig", "def-mod", "def-set", "def-fun", "def-span", "def-trans", "def-proof",
+                       "." ]
   , P.reservedOpNames = []
   , P.caseSensitive = True }
   where
@@ -39,105 +42,202 @@ symbol = P.symbol lexer
 comma = P.comma lexer
 
 program :: Parse Program
-program = Program <$> many (TLSig <$> signature <|> TLMod <$> modul)
+program = ModuleBody <$> moduleBody
 
-signature :: Parse SigDef
-signature = do
-  reserved "signature"
-  (name, args) <- multi_app sigName parameter
-  reserved "where"
-  bod  <- semiSep sigdecl
-  reserved "end"
-  return $ SigDef name (SigLam args bod)
+moduleBody :: Parse [Decl ScopedExp]
+moduleBody = many modulDecl
 
-modul :: Parse ModDef
-modul = do
-  reserved "module"
-  (name, args) <- multi_app modName parameter
-  colon
-  osig <- sigExp
-  reserved "where"
-  bod <- semiSep moddecl
-  reserved "end"
-  return $ ModDef name $ ModLam args osig bod
-  
-parameter :: Parse (ModName, SigExp)
-parameter = do
-  name <- identifier
-  colon
-  sig  <- sigExp
[…]

### 2020-02-13 impl/src/TypeCheck.hs

@@ -7,7 +7,6 @@ import Data.Foldable
 import Data.Traversable
 import Grammar
 
-
 type TC a = Either String a
 
 typeCheckProg :: Program -> TC ()

### 2020-02-14 impl/src/Grammar.hs
much more impl

@@ -1,4 +1,16 @@
+{-# LANGUAGE GeneralizedNewtypeDeriving #-}
+{-# LANGUAGE DeriveFunctor #-}
 module Grammar where
+
+import qualified Text.Parsec as Parsec
+
+-- | Raw S-expressions, the output of the parser.
+data ParsedSExp = ParsedSExp { _loc :: Parsec.SourcePos, _sexp :: SExp }
+  deriving (Show)
+data SExp
+  = SEAtom String
+  | SEList [ParsedSExp]
+  deriving (Show)
 
 -- | A Program is just a module
 type Program = ModuleBody
@@ -11,7 +23,7 @@ newtype ModuleBody = ModuleBody { _defs :: [Decl ScopedExp] }
 -- assertion of an equality (between transformations?)
 data ScopedExp
   = ScSig SigExp
-  | ScModule ModExp
+  | ScMod ModExp
   | ScSet SetExp
   | ScFun ScopedEltExp
   | ScSpan ScopedSpanExp
@@ -20,14 +32,38 @@ data ScopedExp
   deriving (Show, Read, Eq)
 
 data Decl a = Decl { _name :: String , _defn :: a }
-  deriving (Show, Read, Eq)
-  
-type Ctx = [Decl ScopedExp]
+  deriving (Show, Read, Eq, Functor)
+
+
+type Ctx = [Decl DeforGen]
+data DeforGen
+  = DGExp ScopedExp
+  | DGGen Generator
+
+lookupDecl :: String -> [Decl DeforGen] -> Maybe DeforGen
+lookupDecl s xs = lookup s (map declToPair xs)
+  where
+    declToPair :: Decl a -> (String, a)
+    declToPair d = (_name d, _defn d)
+
+isSig :: DeforGen -> Bool
+isSig (DGExp (ScSig _)) = True
+isSig _                 = False
+
+isMod :: DeforGen -> Bool
+isMod (DGExp (ScMod _)) = True
+isMod (DGGen (GenMod _))    = True
+isMod _                 = False
+
+isSet :: DeforGen -> Bool
+isSet (DGExp (ScSet _)) = True
+isSet (DGGen GenSet)    = True
+isSet _                 = False
 
 -- First order lambda calculus
 data SigExp
   = SigBase GroundSigExp
-  | SigApp GroundSigExp [Exp]
+  | SigApp GroundSigExp [AnyExp]
   deriving (Show, Read, Eq)
 
 data GroundSigExp
[…]

### 2020-02-14 impl/src/Parse.hs

@@ -1,6 +1,5 @@
 module Parse where
 
-
 import Control.Monad
 import Control.Monad.Identity
 import Text.Parsec
@@ -21,13 +20,11 @@ sexp_lang = P.LanguageDef
   , P.identLetter = identChar <|> digit <|> char '\''
   , P.opStart = mzero
   , P.opLetter = mzero
-  , P.reservedNames = ["module", "signature", "where", "set", "span", "fun", "term", "trans",
-                       "def-sig", "def-mod", "def-set", "def-fun", "def-span", "def-trans", "def-proof",
-                       "." ]
+  , P.reservedNames = []
   , P.reservedOpNames = []
   , P.caseSensitive = True }
   where
-    identChar = letter <|> char '-' <|> char '_'
+    identChar = letter <|> char '-' <|> char '_' <|> char '>' <|> char '<' <|> char '=' <|> char '*' <|> char '~'
 
 lexer :: P.GenTokenParser String () Identity
 lexer = P.makeTokenParser sexp_lang
@@ -41,203 +38,16 @@ semiSep = P.semiSep lexer
 symbol = P.symbol lexer
 comma = P.comma lexer
 
-program :: Parse Program
-program = ModuleBody <$> moduleBody
+program = sexpList
 
-moduleBody :: Parse [Decl ScopedExp]
-moduleBody = many modulDecl
+sexpList :: Parse [ParsedSExp]
+sexpList = many parsedSexp
 
-modulDecl :: Parse (Decl ScopedExp)
-modulDecl
-  = decl "def-sig" (ScSig <$> sigExp)
-  <|> decl "def-mod" (ScModule <$> modExp)
-  <|> decl "def-set" (ScSet <$> setExp)
-  <|> decl "def-fun" (ScFun <$> scopedEltExp)
-  <|> decl "def-span" (ScSpan <$> scopedSpanExp)
-  <|> decl "def-trans" (ScTrans <$> scopedTransExp)
-  <|> decl "def-proof" (ScAssert <$> scopedProofExp)
-
-decl :: String -> Parse a -> Parse (Decl a)
-decl defA parseA = parens $ reserved defA >>
-  Decl <$> identifier <*> parseA
-
-sigExp :: Parse SigExp
-sigExp =
-  SigBase <$> groundSigExp
-  <|> (parens $ SigApp <$> groundSigExp <*> many anyExp)
-  where
-    groundSigExp :: Parse GroundSigExp
[…]

### 2020-02-14 impl/src/TypeCheck.hs

@@ -1,257 +1,180 @@
+{-# LANGUAGE LambdaCase #-}
+{-# LANGUAGE GeneralizedNewtypeDeriving #-}
 module TypeCheck where
 
+import Control.Applicative
+import Data.Functor
 import Control.Monad
+import Control.Monad.Except
 import Control.Monad.Identity
+import Control.Monad.Reader
+import Control.Monad.State
 import Data.Monoid
 import Data.Foldable
 import Data.Traversable
 import Grammar
 
-type TC a = Either String a
+-- | Typechecking here is combined with parsing the sexpression,
+-- unfortunately. The effects are as follows:
+-- 1. We can throw type/parse errors
+-- 2. We accumulate bindings in the Ctx
+-- 3. We consume the s-expression as we go
 
-typeCheckProg :: Program -> TC ()
-typeCheckProg (Program []) = return ()
-typeCheckProg (Program (tl : rest)) = do
-  typeCheckTL tl
-  typeCheckProg (subst tl (Program rest))
+-- we're either parsing a sequence, or inspecting a single sexp
+-- either err (s -> (s,a))
+-- s -> (s, either err a)
+-- s -> either err (s, a)
 
-typeCheckTL :: TopLevel -> TC ()
-typeCheckTL (TLSig (SigDef name lam)) = typeCheckSigLam lam
-typeCheckTL (TLMod (ModDef name lam)) = typeCheckModLam lam
+newtype TCS a = TCS { runTCS :: StateT ([ParsedSExp], Ctx) (ExceptT ErrMsg Identity) a }
+  deriving (Functor,Applicative,Monad,Alternative,MonadError ErrMsg, MonadState ([ParsedSExp], Ctx))
 
-typeCheckSigLam :: SigLambda -> TC ()
-typeCheckSigLam (SigLam sigDefArgs sigDefBody) = do
-  typeCheckSigArgs sigDefArgs
-  typeCheckSigBody sigDefBody sigDefArgs
+newtype TC a = TC { runTC :: ReaderT ParsedSExp (StateT Ctx (ExceptT ErrMsg Identity)) a }
+  deriving (Functor,Applicative,Monad,Alternative,MonadError ErrMsg, MonadReader ParsedSExp, MonadState Ctx)
 
-typeCheckSigArgs :: [(ModName, SigExp)] -> TC ()
-typeCheckSigArgs args = loop args []
+type ErrMsg = String
+
[…]

### 2020-02-14 impl/test/Test.hs

@@ -2,6 +2,7 @@ import Data.Foldable
 
 import Control.Applicative
 import Control.Monad
+import Data.Either
 import Data.Functor
 import Data.List (intercalate)
 import Data.Traversable
@@ -19,572 +20,599 @@ main = void $ runTestTT tests
                      , tcTests
                      ]
 
-tcTests = "Type Checking Tests" ~: test [ goodTests, badTests ]
+
+
+tcTests = "Type Checking Tests" ~: test [ goodTests
+                                        , badTests
+                                        ]
 goodTests = "Successful Type Checks" ~: test
-  [ typeChecks (Program good1) ~? "category signature"
-  , typeChecks (Program good2) ~? "functor signature"
-  , typeChecks (Program good3) ~? "transformation signature"
-  , typeChecks (Program goodset) ~? "sets signature"
-  , typeChecks (Program goodfun) ~? "function signature"
-  , typeChecks (Program goodextfun) ~? "module deref in signature"
-  , typeChecks modTest1 ~? "module with identity function"
-  , typeChecks modTest2 ~? "module with supplied function"
-  , typeChecks modTest3 ~? "module with composition of functions"
-  , typeChecks modTest4 ~? "module with big composition of function"
-  , typeChecks spanmod0 ~? "span signatures"
-  , typeChecks spanmod1 ~? "eta expanded span"
-  , typeChecks spanmod2 ~? "span with restriction"
-  , typeChecks spanmod3 ~? "span with more restriction"
+  [ typeChecks "" ~? "empty program"
+  , typeChecks "(def-sig S (sig ()))" ~? "simpl signature"
+  , typeChecks "(def-sig S (sig ((set X)) (set Y)))" ~? "more signature"
+  , typeChecks "(def-sig S (sig ((set X)) (fun f X X)))" ~?  "more signature"
+  , typeChecks "(def-sig S (sig ((set X) (fun f X X)) ))" ~? "more signature"
+  , typeChecks "(def-sig S (sig ((set X) (set Y)) (fun f Y X)))" ~? "more signature"
+  , typeChecks "(def-sig S (sig ((set X) (set Y) (fun f X Y)) (span R X X)))" ~? "more signature"
   ]
+--   , typeChecks (Program good2) ~? "functor signature"
+--   , typeChecks (Program good3) ~? "transformation signature"
[…]

### 2020-02-16 impl/src/Grammar.hs
not working but some progress

@@ -60,14 +60,35 @@ isSet (DGExp (ScSet _)) = True
 isSet (DGGen GenSet)    = True
 isSet _                 = False
 
--- First order lambda calculus
+getFunTy :: DeforGen -> Maybe EltTy
+getFunTy (DGExp (ScFun scoped)) = return . eltScopeToTy . _eltscope $ scoped
+getFunTy (DGGen (GenFun ty)) = return $ ty
+getFunTy _ = Nothing
+
+-- | Semantic Signature Values.
+--   going to use HOAS
+--   either a ground signature,
+--   or parameterized by a specific type of
+--   term (which includes modules)
+-- data SigVal
+--   = SVSig [Decl Generator]
+--   | SVPSet (SetVal -> SigVal)
+  -- | FunVal -> SigVal
+  -- | SpanVal -> SigVal
+  -- | TransVal -> SigVal
+  -- 
+  
+
 data SigExp
   = SigBase GroundSigExp
   | SigApp GroundSigExp [AnyExp]
   deriving (Show, Read, Eq)
 
+type SigVal = [Decl Generator]
+
 data GroundSigExp
   = GSigVar String
+  | GSigVal SigVal
   | GSigLam SigLambda
   deriving (Show, Read, Eq)
 
@@ -116,7 +137,7 @@ data GroundModExp
 
 data ModLambda = ModLam
   { _modLamParams :: [Decl Generator]
-  , _modOSig  :: SigExp
+  , _modOSig      :: [Decl Generator]
   , _modBody :: ModuleBody
   }
   deriving (Show, Read, Eq)
@@ -148,7 +169,12 @@ data EltScope
 data EltTy = EltTy { _eltdomty :: SetExp, _eltcodty :: SetExp }
   deriving (Show, Read, Eq)
 
-type ScopedEltExp = (EltScope, EltExp)
+eltScopeToTy :: EltScope -> EltTy
+eltScopeToTy es = EltTy (_eltvarty . _eltinp $ es) (_eltty $ es)
+
+data ScopedEltExp =
+  ScopedEltExp { _eltscope :: EltScope , _eltExp :: EltExp }
+  deriving (Show, Read, Eq)
 
 data SpanExp
   = SpanEApp ModDeref EltExp EltExp

### 2020-02-16 impl/src/TypeCheck.hs

@@ -112,28 +112,37 @@ sideeffect k m = do
 
 decl :: TC (Decl ScopedExp)
 decl = list . sideeffect (declare . fmap DGExp) $
-      tcDecl "def-sig" (ScSig <$> sigExp)
-  <|> tcDecl "def-mod" (ScMod <$> modExp)
+      tcDecl "def-sig" (ScSig <$> only sigExp)
+  <|> tcDecl "def-mod" (ScMod <$> only modExp)
+  <|> tcDecl "def-set" (ScSet <$> only setExp)
+  <|> tcDecl "def-fun" (ScFun <$> scopedEltExp)
+  -- (def-fun f (x A) B
+  --   e)
+--  <|> tcDecl "def-fun" ()
+  -- TODO: fun, span, trans, assertion
   where
     tcDecl :: String -> TCS a -> TCS (Decl a)
     tcDecl def p = (tcHd (atomEq def) >> Decl <$> tcHd anyAtom <*> p)
 
--- (sig (param ...) param ...)
+-- (sig () param ...) -- sig value
+-- (sig (param param ...) param ...) -- sig lambda
 -- or S
 -- or (S ...)
 -- or (. C ...)
 
-sigExp :: TCS SigExp
-sigExp = only (sigBase <|> sigApp)
+sigExp :: TC SigExp
+sigExp = sigBase <|> sigApp
   where
-    sigBase = SigBase <$> (GSigVar <$> sigVar <|> GSigLam <$> sigLam)
+    sigBase = SigBase <$> (GSigVar <$> sigVar <|> GSigVal <$> sigVal <|> GSigLam <$> sigLam)
     sigVar  = do
       var <- anyAtom
       ctx <- get
       case fmap isSig $ lookupDecl var ctx of
         Just True -> return var
         _ -> empty
-    sigLam = localize $ list $ tcHd (atomEq "sig") >> SigLam <$> tcHd (list params) <*> params
+    sigLam = localize . list $ tcHd (atomEq "psig") >> SigLam <$> tcHd (list params) <*> params
+    sigVal = localize . list $ tcHd (atomEq "sig") >> params
+
     sigApp :: TC SigExp
     sigApp = empty -- TODO: signature application
 
@@ -141,17 +150,34 @@ sigExp = only (sigBase <|> sigApp)
 -- or C
 -- or (. C ...)
 -- or (M anyExpr ...)
-modExp :: TCS ModExp -- | TODO: module lambda
[…]

### 2020-02-16 impl/test/Test.hs

@@ -27,12 +27,16 @@ tcTests = "Type Checking Tests" ~: test [ goodTests
                                         ]
 goodTests = "Successful Type Checks" ~: test
   [ typeChecks "" ~? "empty program"
-  , typeChecks "(def-sig S (sig ()))" ~? "simpl signature"
-  , typeChecks "(def-sig S (sig ((set X)) (set Y)))" ~? "more signature"
-  , typeChecks "(def-sig S (sig ((set X)) (fun f X X)))" ~?  "more signature"
-  , typeChecks "(def-sig S (sig ((set X) (fun f X X)) ))" ~? "more signature"
-  , typeChecks "(def-sig S (sig ((set X) (set Y)) (fun f Y X)))" ~? "more signature"
-  , typeChecks "(def-sig S (sig ((set X) (set Y) (fun f X Y)) (span R X X)))" ~? "more signature"
+  , typeChecks "(def-sig S (sig))" ~? "simpl signature"
+  , typeChecks "(def-sig S (psig ((set X)) (set Y)))" ~? "more signature"
+  , typeChecks "(def-sig S (psig ((set X)) (fun f X X)))" ~?  "more signature"
+  , typeChecks "(def-sig S (psig ((set X) (fun f X X)) ))" ~? "more signature"
+  , typeChecks "(def-sig S (psig ((set X) (set Y)) (fun f Y X)))" ~? "more signature"
+  , typeChecks "(def-sig S (psig ((set X) (set Y) (fun f X Y)) (span R X X)))" ~? "more signature"
+  , typeChecks "(def-mod M (mod () (sig)))" ~? "world's smallest module"
+  , typeChecks "(def-mod M (mod ((set X)) (sig (set Y)) (def-set Y X)))" ~? "world's smallest module"
+  , typeChecks "(def-mod ID (mod ((set X)) (def-fun id (x X) X x))" ~? "id fun"
+  , typeChecks "(def-mod SELFCOMP (mod ((set X) (fun f X X)) (def-fun g (x X) X (f x)))" ~? "endo-comp"
   ]
 --   , typeChecks (Program good2) ~? "functor signature"
 --   , typeChecks (Program good3) ~? "transformation signature"
@@ -50,12 +54,14 @@ goodTests = "Successful Type Checks" ~: test
 --   ]
 badTests = "Type Checking Failures" ~: test
   [ not (typeChecks "()") ~? "empty decl"
[…]

### 2020-02-18 impl/src/Grammar.hs
impl sets, funs

@@ -7,19 +7,20 @@ import qualified Text.Parsec as Parsec
 -- | Raw S-expressions, the output of the parser.
 data ParsedSExp = ParsedSExp { _loc :: Parsec.SourcePos, _sexp :: SExp }
   deriving (Show)
+
 data SExp
   = SEAtom String
   | SEList [ParsedSExp]
   deriving (Show)
 
--- | A Program is just a module
+-- | A Program is just a module body
 type Program = ModuleBody
 
--- | A Module is a sequence of statements
-newtype ModuleBody = ModuleBody { _defs :: [Decl ScopedExp] }
+-- | A Module is a sequence of definitions
+newtype ModuleBody = ModuleBody { _defs :: [SynDecl ScopedExp] }
   deriving (Show, Read, Eq)
 
--- | A statement is a definition of a signature, module or term or an
+-- | A definition is a definition of a signature, module or term or an
 -- assertion of an equality (between transformations?)
 data ScopedExp
   = ScSig SigExp
@@ -31,39 +32,65 @@ data ScopedExp
   | ScAssert ScopedProofExp
   deriving (Show, Read, Eq)
 
-data Decl a = Decl { _name :: String , _defn :: a }
+data Decl name a = Decl { _name :: name , _defn :: a }
   deriving (Show, Read, Eq, Functor)
 
+type SynDecl = Decl String
 
-type Ctx = [Decl DeforGen]
-data DeforGen
-  = DGExp ScopedExp
-  | DGGen Generator
+type Ctx = [SynDecl SemVal]
 
-lookupDecl :: String -> [Decl DeforGen] -> Maybe DeforGen
+-- | 
+data SemVal
+  = SemSet SetNF
+  | SemFun SetNF -- domain set
+           SetNF -- codomain set
+           (EltNF -> EltNF) -- Yoneda-embedding of the term
+  | SemSpan SetNF -- contravariant dependency
+            SetNF -- covariant dependency
+            (EltNF -> EltNF -> SpanNF) -- Yoneda embedding of the term
+-- | SemTrans ??
+-- | SemProof ??
+  -- | SemSig -- TODO
+  | SemMod -- TODO
+
+type SetNF = DBRef
+data EltNF
+  = ENFId
+  | ENFFunApp DBRef EltNF
+  deriving (Show, Eq)
+
[…]

### 2020-02-18 impl/src/TypeCheck.hs

@@ -20,18 +20,32 @@ import Grammar
 -- 2. We accumulate bindings in the Ctx
 -- 3. We consume the s-expression as we go
 
--- we're either parsing a sequence, or inspecting a single sexp
--- either err (s -> (s,a))
--- s -> (s, either err a)
--- s -> either err (s, a)
+-- The basic process is that we take in an S-expression, produce an
+-- AST for the purposes of parsing but in the context we produce
+-- semantic values for the purposes of simple equality checking
 
+-- There are two monads involved
+-- One for parsing a sequence of S-expressions
 newtype TCS a = TCS { runTCS :: StateT ([ParsedSExp], Ctx) (ExceptT ErrMsg Identity) a }
   deriving (Functor,Applicative,Monad,Alternative,MonadError ErrMsg, MonadState ([ParsedSExp], Ctx))
 
+-- And one for parsing a single S-expression
 newtype TC a = TC { runTC :: ReaderT ParsedSExp (StateT Ctx (ExceptT ErrMsg Identity)) a }
   deriving (Functor,Applicative,Monad,Alternative,MonadError ErrMsg, MonadReader ParsedSExp, MonadState Ctx)
 
 type ErrMsg = String
+
+ctx :: TCS Ctx
+ctx = snd <$> get
+
+popDecl :: TCS (SynDecl SemVal)
+popDecl = do
+  (p, ctx) <- get
+  case ctx of
+    [] -> throwError $ "unbound variable"
+    (d:rest) -> do
+      put (p, rest)
+      return d
 
 typeCheck :: TCS a -> [ParsedSExp] -> Either String a
 typeCheck (TCS m) exps =
@@ -41,7 +55,7 @@ typeCheckOne :: TC a -> ParsedSExp -> Either String a
 typeCheckOne (TC m) exp =
   fmap fst $ runIdentity $ runExceptT $ runStateT (runReaderT m exp) []
 
-localize :: TC a -> TC a
+localize :: (MonadState s m) => m a -> m a
 localize m = do
   s <- get
   x <- m
@@ -101,7 +115,7 @@ program = moduleBody
 moduleBody :: TCS ModuleBody
 moduleBody = ModuleBody <$> (several decl)
 
-declare :: Decl DeforGen -> TCS ()
+declare :: SynDecl SemVal -> TCS ()
 declare d = modify (fmap (d:))
 
[…]

### 2020-02-18 impl/test/Test.hs

@@ -27,16 +27,17 @@ tcTests = "Type Checking Tests" ~: test [ goodTests
                                         ]
 goodTests = "Successful Type Checks" ~: test
   [ typeChecks "" ~? "empty program"
-  , typeChecks "(def-sig S (sig))" ~? "simpl signature"
-  , typeChecks "(def-sig S (psig ((set X)) (set Y)))" ~? "more signature"
-  , typeChecks "(def-sig S (psig ((set X)) (fun f X X)))" ~?  "more signature"
-  , typeChecks "(def-sig S (psig ((set X) (fun f X X)) ))" ~? "more signature"
-  , typeChecks "(def-sig S (psig ((set X) (set Y)) (fun f Y X)))" ~? "more signature"
-  , typeChecks "(def-sig S (psig ((set X) (set Y) (fun f X Y)) (span R X X)))" ~? "more signature"
   , typeChecks "(def-mod M (mod () (sig)))" ~? "world's smallest module"
-  , typeChecks "(def-mod M (mod ((set X)) (sig (set Y)) (def-set Y X)))" ~? "world's smallest module"
-  , typeChecks "(def-mod ID (mod ((set X)) (def-fun id (x X) X x))" ~? "id fun"
-  , typeChecks "(def-mod SELFCOMP (mod ((set X) (fun f X X)) (def-fun g (x X) X (f x)))" ~? "endo-comp"
+  , typeChecks "(def-mod S (mod ((set X) (set Y)) (sig)))" ~? "sets in params"
+  , typeChecks "(def-mod M (mod ((set X)) (sig (set Y)) (def-set Y X)))" ~? "definition of a set"
+  , typeChecks "(def-mod S (mod ((set X) (fun f X X)) (sig)))" ~?  "fun in params"
+  , typeChecks "(def-mod S (mod ((set X) (set Y) (fun f Y X)) (sig)))" ~? "more fun in params"
+  , typeChecks "(def-mod ID (mod ((set X)) (sig) ))" ~? "id fun sig"
+  , typeChecks "(def-mod ID (mod ((set X)) (sig) (def-fun id (x X) X x)))" ~? "id fun"
+  , typeChecks "(def-mod SELFCOMP (mod ((set X) (fun f X X)) (sig) (def-fun g (x X) X (f x))))" ~? "endo-comp"
[…]

### 2020-02-18 impl/src/TypeCheck.hs
implement basic spans NEEEEED error messages ASAP

@@ -140,29 +140,48 @@ denote = \case
   (ScMod _) -> return SemMod
   (ScSet e) -> SemSet <$> denoteSet e
   (ScFun e) -> (\(dom,cod,f) -> SemFun dom cod f) <$> denoteFun e
+  (ScSpan e) -> (\(contra,co,spanF) -> SemSpan contra co spanF) <$> denoteSpan e
 
 denoteSet :: SetExp -> TCS SetNF
 denoteSet mdref = do
-  (Decl _ (SemSet dbref)) <- resolveRef mdref
-  return dbref
+  (Decl _ val) <- resolveRef mdref
+  case val of
+    (SemSet dbref) -> return dbref
+    _ -> throwError $ "expected a set but got...something else"
 
 denoteFun :: ScopedEltExp -> TCS (SetNF, SetNF, (EltNF -> EltNF))
 denoteFun (ScopedEltExp (EltScope (TypedEltVar x domTyExp) codTyExp) e) = do
   dom <- denoteSet domTyExp
   cod <- denoteSet codTyExp
-  f <- loop x dom cod e
+  f <- denoteFunOpen x dom cod e
   return (dom,cod,f)
   where
-    loop :: String -> SetNF -> SetNF -> EltExp -> TCS (EltNF -> EltNF)
-    loop x dom cod (EEVar y) = do
+
+denoteFunOpen :: String -> SetNF -> SetNF -> EltExp -> TCS (EltNF -> EltNF)
+denoteFunOpen x dom cod (EEVar y) = do
       guard $ x == y
       guard $ dom == cod
       return (\nf -> nf)
-    loop x dom cod (EEApp ref e) = do
-      (Decl _ (SemFun dom' cod' f)) <- resolveRef ref
-      guard $ cod == cod'
-      g <- loop x dom dom' e
-      return (f . g)
+denoteFunOpen x dom cod (EEApp ref e) = do
+      (Decl _ val) <- resolveRef ref
+      case val of
+        (SemFun dom' cod' f) -> do
+          guard $ cod == cod'
+          g <- denoteFunOpen x dom dom' e
+          return (f . g)
+        _ -> throwError $ "expected a function but got...something else"
+
+denoteSpan :: ScopedSpanExp -> TCS (SetNF, SetNF, (EltNF -> EltNF -> SpanNF))
[…]

### 2020-02-18 impl/test/Test.hs

@@ -37,7 +37,10 @@ goodTests = "Successful Type Checks" ~: test
   , typeChecks "(def-mod SELFCOMP (mod ((set X) (fun f X X)) (sig) (def-fun g (x X) X (f x))))" ~? "endo-comp"
   , typeChecks "(def-mod MANYSELFCOMP (mod ((set X) (fun f X X)) (sig) (def-fun g (x X) X (f (f (f (f x)))))))" ~? "many endo-comp"
   , typeChecks "(def-mod COMP (mod ((set X) (set Y) (fun f X Y) (set Z) (fun g Y Z)) (sig) (def-fun h (x X) Z (g (f x))))))" ~? "many endo-comp"
-  -- , typeChecks "(def-mod S (mod ((set X) (set Y) (fun f X Y) (span R X X))))" ~? "span in params"
+  , typeChecks "(def-mod S (mod ((set X) (set Y) (span R X Y)) (sig)))" ~? "span in params"
+  , typeChecks "(def-mod S (mod ((set X) (set Y) (span R X X)) (sig)))" ~? "span in params"
+  , typeChecks "(def-mod S (mod ((set X) (set Y) (fun f X Y) (span R X Y)) (sig) (def-span Q (x X) (y Y) (R x y))))" ~? "span eta"
+  , typeChecks "(def-mod S (mod ((set X) (set Y) (fun f X Y) (span R X Y)) (sig) (def-span Q (x X) (x' X) (R x (f x')))))" ~? "span subst"
   ]
 --   , typeChecks (Program good2) ~? "functor signature"
 --   , typeChecks (Program good3) ~? "transformation signature"
@@ -62,6 +65,8 @@ badTests = "Type Checking Failures" ~: test
   , not (typeChecks "(def-mod M (mod ((set X)) (sig ) (def-set Y Y)))") ~? "set out of scope"
   , not (typeChecks "(def-mod M (mod ((set X) (set Y)) (sig) (def-fun f (x X) Y x)))") ~? "type error in fun def"
   , not (typeChecks "(def-mod M (mod ((set X) (set Y) (fun f X Y) (fun g X Y)) (sig) (def-fun h (x X) Y (g (f x)))))") ~? "type error in fun comp"
+  , not (typeChecks "(def-mod ID (mod ((set X) (fun g X X)) (sig) (def-fun id (x g) g x)))") ~? "fun used as a set"
[…]

### 2020-02-18 impl/src/Grammar.hs
implement transformation symbols

@@ -1,5 +1,6 @@
 {-# LANGUAGE GeneralizedNewtypeDeriving #-}
 {-# LANGUAGE DeriveFunctor #-}
+{-# LANGUAGE DeriveTraversable #-}
 module Grammar where
 
 import qualified Text.Parsec as Parsec
@@ -48,10 +49,22 @@ data SemVal
   | SemSpan SetNF -- contravariant dependency
             SetNF -- covariant dependency
             (EltNF -> EltNF -> SpanNF) -- Yoneda embedding of the term
--- | SemTrans ??
+  | SemTrans (NEList SetNF) -- set indices
+             [SpanNF] -- span domains
+             SpanNF   -- codomain
+             ([TransNF] -> TransNF)
 -- | SemProof ??
   -- | SemSig -- TODO
   | SemMod -- TODO
+
+quoteSemFun :: (EltNF -> EltNF) -> EltNF
+quoteSemFun f = f ENFId
+
+quoteSemSpan :: (EltNF -> EltNF -> SpanNF) -> SpanNF
+quoteSemSpan r = r ENFId ENFId
+
+quoteSemTrans :: ([TransNF] -> TransNF) -> TransNF
+quoteSemTrans t = t (repeat TNFId)
 
 type SetNF = DBRef
 data EltNF
@@ -62,10 +75,10 @@ data EltNF
 data SpanNF = SNFSpanApp { _spanSymb :: DBRef, _contraElt :: EltNF, _covarElt :: EltNF }
   deriving (Show, Eq)
 
--- A context
--- data NForGen
---   = NGNF  ScopedNF
---   | NGGen Generator
+data TransNF
+  = TNFId
+  | TNFApp DBRef [TransNF]
+  deriving (Show, Eq)
 
 lookupDecl :: (Eq name) => name -> [Decl name a] -> Maybe a
 lookupDecl s xs = lookup s (map declToPair xs)
@@ -140,6 +153,30 @@ data ModDeref
   | MDSelect ModExp [String]
   deriving (Show, Read, Eq)
 
+data NEList a
+  = Done a
+  | Cons a (NEList a)
+  deriving (Show, Read, Eq, Functor, Foldable, Traversable)
+
+toNE :: [a] -> Maybe (NEList a)
+toNE = foldr cons Nothing
+  where cons :: a -> Maybe (NEList a) -> Maybe (NEList a)
+        cons x   Nothing = Just (Done x)
+        cons x (Just xs) = Just (Cons x xs)
+
[…]

### 2020-02-18 impl/src/TypeCheck.hs

@@ -33,7 +33,9 @@ newtype TCS a = TCS { runTCS :: StateT ([ParsedSExp], Ctx) (ExceptT ErrMsg Ident
 newtype TC a = TC { runTC :: ReaderT ParsedSExp (StateT Ctx (ExceptT ErrMsg Identity)) a }
   deriving (Functor,Applicative,Monad,Alternative,MonadError ErrMsg, MonadReader ParsedSExp, MonadState Ctx)
 
-type ErrMsg = String
+type ErrMsg = First String
+
+mkErr = First . Just
 
 ctx :: TCS Ctx
 ctx = snd <$> get
@@ -42,18 +44,25 @@ popDecl :: TCS (SynDecl SemVal)
 popDecl = do
   (p, ctx) <- get
   case ctx of
-    [] -> throwError $ "unbound variable"
+    [] -> throwError . mkErr $ "unbound variable"
     (d:rest) -> do
       put (p, rest)
       return d
 
 typeCheck :: TCS a -> [ParsedSExp] -> Either String a
 typeCheck (TCS m) exps =
-  fmap fst $ runIdentity $ runExceptT $ runStateT m (exps, [])
+  lErrToStr .  fmap fst $ runIdentity $ runExceptT $ runStateT m (exps, [])
+  
 
 typeCheckOne :: TC a -> ParsedSExp -> Either String a
 typeCheckOne (TC m) exp =
-  fmap fst $ runIdentity $ runExceptT $ runStateT (runReaderT m exp) []
+  lErrToStr . fmap fst $ runIdentity $ runExceptT $ runStateT (runReaderT m exp) []
+
+lErrToStr = either (Left . errToStr) Right
+
+errToStr :: ErrMsg -> String
+errToStr (First Nothing) = "internal error: no cases"
+errToStr (First (Just s)) = s
 
 localize :: (MonadState s m) => m a -> m a
 localize m = do
@@ -66,16 +75,19 @@ localize m = do
 tcHd :: TC a -> TCS a
 tcHd p = TCS . StateT $ \(sexps, ctx) ->
   case sexps of
-    [] -> throwError "expected another thing but there was nothing left :("
+    [] -> empty
     (x:xs) -> do
       (a, ctx) <- runStateT (runReaderT (runTC p) $ x) ctx
       return (a, (xs,ctx))
+
+mayHd :: TC a -> TCS (Maybe a)
+mayHd p = Just <$> tcHd p <|> return Nothing
 
[…]

### 2020-02-18 impl/test/Test.hs

@@ -26,21 +26,29 @@ tcTests = "Type Checking Tests" ~: test [ goodTests
                                         , badTests
                                         ]
 goodTests = "Successful Type Checks" ~: test
-  [ typeChecks "" ~? "empty program"
-  , typeChecks "(def-mod M (mod () (sig)))" ~? "world's smallest module"
-  , typeChecks "(def-mod S (mod ((set X) (set Y)) (sig)))" ~? "sets in params"
-  , typeChecks "(def-mod M (mod ((set X)) (sig (set Y)) (def-set Y X)))" ~? "definition of a set"
-  , typeChecks "(def-mod S (mod ((set X) (fun f X X)) (sig)))" ~?  "fun in params"
-  , typeChecks "(def-mod S (mod ((set X) (set Y) (fun f Y X)) (sig)))" ~? "more fun in params"
-  , typeChecks "(def-mod ID (mod ((set X)) (sig) ))" ~? "id fun sig"
-  , typeChecks "(def-mod ID (mod ((set X)) (sig) (def-fun id (x X) X x)))" ~? "id fun"
-  , typeChecks "(def-mod SELFCOMP (mod ((set X) (fun f X X)) (sig) (def-fun g (x X) X (f x))))" ~? "endo-comp"
-  , typeChecks "(def-mod MANYSELFCOMP (mod ((set X) (fun f X X)) (sig) (def-fun g (x X) X (f (f (f (f x)))))))" ~? "many endo-comp"
-  , typeChecks "(def-mod COMP (mod ((set X) (set Y) (fun f X Y) (set Z) (fun g Y Z)) (sig) (def-fun h (x X) Z (g (f x))))))" ~? "many endo-comp"
-  , typeChecks "(def-mod S (mod ((set X) (set Y) (span R X Y)) (sig)))" ~? "span in params"
-  , typeChecks "(def-mod S (mod ((set X) (set Y) (span R X X)) (sig)))" ~? "span in params"
-  , typeChecks "(def-mod S (mod ((set X) (set Y) (fun f X Y) (span R X Y)) (sig) (def-span Q (x X) (y Y) (R x y))))" ~? "span eta"
-  , typeChecks "(def-mod S (mod ((set X) (set Y) (fun f X Y) (span R X Y)) (sig) (def-span Q (x X) (x' X) (R x (f x')))))" ~? "span subst"
+  [ assertTC "empty program" ""
[…]

### 2020-02-18 impl/src/Grammar.hs
parsing transformations

@@ -258,13 +258,20 @@ data SpanTy
   = SpanTy { _spancontraty :: SetExp, _spancoty :: SetExp }
   deriving (Show, Read, Eq)
 
-data TransExp = TransExp -- TODO
+data TransExp =
+  TransEVar String
+  | TransEApp ModDeref [TransExp]
+  -- TODO: more stuff like hom eliminator etc
   deriving (Show, Read, Eq)
 
-data TransScope = TransScope
+data SpanVar = SpanVar String SpanExp
   deriving (Show, Read, Eq)
 
-type ScopedTransExp = (TransScope, TransExp)
+data TransScope = TransScope [TypedEltVar] [SpanVar] SpanExp
+  deriving (Show, Read, Eq)
+
+data ScopedTransExp = ScopedTransExp TransScope TransExp
+  deriving (Show, Read, Eq)
 
 data TransTy = TransTy { _eltVars :: [TypedEltVar], _domSpans :: [SpanExp], _codSpan :: SpanExp }
   deriving (Show, Read, Eq)

### 2020-02-18 impl/src/TypeCheck.hs

@@ -153,6 +153,7 @@ denote = \case
   (ScSet e) -> SemSet <$> denoteSet e
   (ScFun e) -> (\(dom,cod,f) -> SemFun dom cod f) <$> denoteFun e
   (ScSpan e) -> (\(contra,co,spanF) -> SemSpan contra co spanF) <$> denoteSpan e
+  (ScTrans e) -> (\(indices, doms, cod, transF) -> SemTrans indices doms cod transF) <$> denoteTrans e
 
 denoteSet :: SetExp -> TCS SetNF
 denoteSet mdref = do
@@ -209,6 +210,9 @@ denoteSpanStringOpen (Cons (x, contraset) vars) (e : es) = do
 denoteSpanStringOpen (Done _) (e : es) = throwError . mkErr $ "not enough indices for the inputs of a transformation"
 denoteSpanStringOpen (Cons _ _) [] = throwError . mkErr $ "too many indices for the inputs of a transformation"
 
+denoteTrans :: ScopedTransExp -> TCS (NEList SetNF, [SpanNF], SpanNF, [TransNF] -> TransNF)
+denoteTrans = empty -- TODO
+
 denoteAndDeclare :: SynDecl ScopedExp -> TCS ()
 denoteAndDeclare declaration = do
   v <- denote $ _defn declaration
@@ -221,7 +225,7 @@ decl = list . sideeffect denoteAndDeclare $
   <|> tcSynDecl "def-set" (ScSet <$> only setExp)
   <|> tcSynDecl "def-fun" (ScFun <$> scopedEltExp)
   <|> tcSynDecl "def-span" (ScSpan <$> scopedSpanExp)
-
+  <|> tcSynDecl "def-trans" (ScTrans <$> scopedTransExp)
   -- (def-trans t (a b c d) (Rab Sbc Qcd) Tad bod)
   -- TODO: trans, assertion
   where
@@ -291,6 +295,10 @@ eltExp = EEVar <$> anyAtom
 
 typedEltVar :: TC TypedEltVar
 typedEltVar = list $ TypedEltVar <$> tcHd anyAtom <*> tcHd setExp
+
+spanVar :: TC SpanVar
+spanVar = list $ SpanVar <$> tcHd anyAtom <*> tcHd spanExp
+
 -- (x A) (y B) t
 scopedSpanExp :: TCS ScopedSpanExp      
 scopedSpanExp = (,) <$> spanScope <*> only spanExp
[…]

### 2020-02-19 impl/src/Grammar.hs
refactor, progress on transformations

@@ -3,7 +3,10 @@
 {-# LANGUAGE DeriveTraversable #-}
 module Grammar where
 
+import Data.Bifunctor
 import qualified Text.Parsec as Parsec
+
+import Util
 
 -- | Raw S-expressions, the output of the parser.
 data ParsedSExp = ParsedSExp { _loc :: Parsec.SourcePos, _sexp :: SExp }
@@ -43,19 +46,35 @@ type Ctx = [SynDecl SemVal]
 -- | 
 data SemVal
   = SemSet SetNF
-  | SemFun SetNF -- domain set
-           SetNF -- codomain set
-           (EltNF -> EltNF) -- Yoneda-embedding of the term
-  | SemSpan SetNF -- contravariant dependency
-            SetNF -- covariant dependency
-            (EltNF -> EltNF -> SpanNF) -- Yoneda embedding of the term
-  | SemTrans (NEList SetNF) -- set indices
-             [SpanNF] -- span domains
-             SpanNF   -- codomain
-             ([TransNF] -> TransNF)
+  | SemFun ScopedSemFun
+  | SemSpan ScopedSemSpan
+  | SemTrans ScopedSemTrans
 -- | SemProof ??
   -- | SemSig -- TODO
   | SemMod -- TODO
+
+data ScopedSemFun = ScopedSemFun { _scfunDom :: SetNF, _scfunCod :: SetNF, _scfun :: EltNF -> EltNF }
+data ScopedSemSpan = ScopedSemSpan { _scspContra :: SetNF, _scspCovar :: SetNF, _scspan :: EltNF -> EltNF -> SpanNF }
+
+type NamedSemTransCtx = ConsStar (String, SetNF, String, SpanNF) (String, SetNF)
+type SemTransCtx = ConsStar (SetNF, SpanNF) SetNF
+
+ctxIndices :: SemTransCtx -> NEList SetNF
+ctxIndices = consStartoNE . first fst
+
+ctxSpans :: SemTransCtx -> [SpanNF]
+ctxSpans = allAs . first snd
+
+ctxUnName :: NamedSemTransCtx -> SemTransCtx
+ctxUnName = bimap (\(_,a,_,r) -> (a,r)) snd
+
+boundary :: SemTransCtx -> (SetNF, SetNF)
+boundary = firstAndLast . ctxIndices
+
+data ScopedSemTrans
+  = ScopedSemTrans { _sctransCtx  :: SemTransCtx
[…]

### 2020-02-19 impl/src/TypeCheck.hs

@@ -1,9 +1,11 @@
 {-# LANGUAGE LambdaCase #-}
 {-# LANGUAGE GeneralizedNewtypeDeriving #-}
+{-# LANGUAGE OverloadedStrings #-}
 module TypeCheck where
 
 import Control.Applicative
 import Data.Functor
+import Data.String
 import Control.Monad
 import Control.Monad.Except
 import Control.Monad.Identity
@@ -12,7 +14,9 @@ import Control.Monad.State
 import Data.Monoid
 import Data.Foldable
 import Data.Traversable
+
 import Grammar
+import Util
 
 -- | Typechecking here is combined with parsing the sexpression,
 -- unfortunately. The effects are as follows:
@@ -33,9 +37,11 @@ newtype TCS a = TCS { runTCS :: StateT ([ParsedSExp], Ctx) (ExceptT ErrMsg Ident
 newtype TC a = TC { runTC :: ReaderT ParsedSExp (StateT Ctx (ExceptT ErrMsg Identity)) a }
   deriving (Functor,Applicative,Monad,Alternative,MonadError ErrMsg, MonadReader ParsedSExp, MonadState Ctx)
 
-type ErrMsg = First String
+newtype ErrMsg = ErrMsg { _getMsg :: First String }
+  deriving (Show, Read, Eq, Monoid)
 
-mkErr = First . Just
+instance IsString ErrMsg where
+  fromString = ErrMsg . First . Just
 
 ctx :: TCS Ctx
 ctx = snd <$> get
@@ -44,7 +50,7 @@ popDecl :: TCS (SynDecl SemVal)
 popDecl = do
   (p, ctx) <- get
   case ctx of
-    [] -> throwError . mkErr $ "unbound variable"
+    [] -> throwError $ "unbound variable"
     (d:rest) -> do
       put (p, rest)
       return d
@@ -61,8 +67,8 @@ typeCheckOne (TC m) exp =
 lErrToStr = either (Left . errToStr) Right
 
 errToStr :: ErrMsg -> String
-errToStr (First Nothing) = "internal error: no cases"
-errToStr (First (Just s)) = s
+errToStr (ErrMsg (First Nothing)) = "internal error: no cases"
+errToStr (ErrMsg (First (Just s))) = s
 
 localize :: (MonadState s m) => m a -> m a
 localize m = do
@@ -87,7 +93,7 @@ mayHd p = Just <$> tcHd p <|> return Nothing
 done :: TCS ()
 done = fst <$> get >>= \case
   [] -> return ()
[…]

### 2020-02-19 impl/src/Util.hs

@@ -1,0 +1,50 @@
+{-# LANGUAGE GeneralizedNewtypeDeriving #-}
+{-# LANGUAGE DeriveFunctor #-}
+{-# LANGUAGE DeriveTraversable #-}
+module Util where
+
+import Data.Bifunctor
+
+data NEList a
+  = Done a
+  | Cons a (NEList a)
+  deriving (Show, Read, Eq, Functor, Foldable, Traversable)
+
+toNE :: [a] -> Maybe (NEList a)
+toNE = foldr cons Nothing
+  where cons :: a -> Maybe (NEList a) -> Maybe (NEList a)
+        cons x   Nothing = Just (Done x)
+        cons x (Just xs) = Just (Cons x xs)
+
+neFold :: (a -> b -> b) -> (a -> b) -> NEList a -> b
+neFold cons done (Done x)    = done x
+neFold cons done (Cons x xs) = cons x (neFold cons done xs)
+
+neHd :: NEList a -> a
+neHd (Done x)   = x
+neHd (Cons x _) = x
+
+firstAndLast :: NEList a -> (a, a)
+firstAndLast = neFold cons done
+  where done x = (x,x)
+        cons first (_,last) = (first, last)
+
+data ConsStar a b
+  = DoneB b
+  | ConsA a (ConsStar a b)
+  deriving (Show, Eq)
+
+instance Bifunctor ConsStar where
+  bimap fa fb = go
+    where go (DoneB b) = DoneB (fb b)
+          go (ConsA a rest) = ConsA (fa a) (go rest)
+
+allAs :: ConsStar a b -> [a]
+allAs (DoneB _)      = []
+allAs (ConsA a rest) = a : allAs rest
+
+
+consStartoNE :: ConsStar a a -> NEList a
+consStartoNE (DoneB a) = Done a
+consStartoNE (ConsA a rest) = Cons a (consStartoNE rest)
+

### 2020-02-19 impl/test/Test.hs

@@ -49,6 +49,8 @@ goodTests = "Successful Type Checks" ~: test
     "(def-mod S (mod ((set Ob) (span Mor Ob Ob) (trans id ((X Ob)) () (Mor X X)) (trans comp ((X Ob) (Y Ob) (Z Ob)) ((Mor X Y) (Mor Y Z)) (Mor X Z)))))"
   , assertTC "RUP"
     "(def-mod S (mod ((set A) (span HomA A A) (set B) (fun G B A) (span M A B) (trans counit ((b B)) () (M (G b) b)) (trans intro ((a A) (b B)) ((M a b)) (HomA a (G b))))))"
+  , assertTC "ID-trans"
+  "(def-mod ID-trans (mod ((set A) (span R A A)) (trans id (a A) (x (R a a)) (R a a) x)))"
   ]
 --   , typeChecks (Program good2) ~? "functor signature"
 --   , typeChecks (Program good3) ~? "transformation signature"

### 2020-02-19 impl/src/Grammar.hs
get some very basic transformation checking

@@ -76,6 +76,8 @@ data ScopedSemTrans
                    , _sctransCod     :: SpanNF   -- codomain
                    , _sctrans        ::([TransNF] -> TransNF) }
 
+type SemTransSubst = [TransNF] -> [TransNF]
+
 quoteSemFun :: (EltNF -> EltNF) -> EltNF
 quoteSemFun f = f ENFId
 
@@ -262,13 +264,13 @@ data TransExp =
 data SpanVar = SpanVar String SpanExp
   deriving (Show, Read, Eq)
 
-data TransScope = TransScope [TypedEltVar] [SpanVar] SpanExp
+data TransScope = TransScope (NEList TypedEltVar) [SpanVar] SpanExp
   deriving (Show, Read, Eq)
 
 data ScopedTransExp = ScopedTransExp TransScope TransExp
   deriving (Show, Read, Eq)
 
-data TransTy = TransTy { _eltVars :: [TypedEltVar], _domSpans :: [SpanExp], _codSpan :: SpanExp }
+data TransTy = TransTy { _eltVars :: (NEList TypedEltVar), _domSpans :: [SpanExp], _codSpan :: SpanExp }
   deriving (Show, Read, Eq)
 
 

### 2020-02-19 impl/src/TypeCheck.hs

@@ -4,6 +4,7 @@
 module TypeCheck where
 
 import Control.Applicative
+import Control.Lens
 import Data.Functor
 import Data.String
 import Control.Monad
@@ -37,11 +38,11 @@ newtype TCS a = TCS { runTCS :: StateT ([ParsedSExp], Ctx) (ExceptT ErrMsg Ident
 newtype TC a = TC { runTC :: ReaderT ParsedSExp (StateT Ctx (ExceptT ErrMsg Identity)) a }
   deriving (Functor,Applicative,Monad,Alternative,MonadError ErrMsg, MonadReader ParsedSExp, MonadState Ctx)
 
-newtype ErrMsg = ErrMsg { _getMsg :: First String }
+newtype ErrMsg = ErrMsg { _getMsg :: Last String }
   deriving (Show, Read, Eq, Monoid)
 
 instance IsString ErrMsg where
-  fromString = ErrMsg . First . Just
+  fromString = ErrMsg . Last . Just
 
 ctx :: TCS Ctx
 ctx = snd <$> get
@@ -67,8 +68,8 @@ typeCheckOne (TC m) exp =
 lErrToStr = either (Left . errToStr) Right
 
 errToStr :: ErrMsg -> String
-errToStr (ErrMsg (First Nothing)) = "internal error: no cases"
-errToStr (ErrMsg (First (Just s))) = s
+errToStr (ErrMsg (Last Nothing)) = "internal error: no cases"
+errToStr (ErrMsg (Last (Just s))) = s
 
 localize :: (MonadState s m) => m a -> m a
 localize m = do
@@ -94,10 +95,9 @@ done :: TCS ()
 done = fst <$> get >>= \case
   [] -> return ()
   _  -> throwError $ "there's some junk here"
-  
 
-only :: TC a -> TCS a
-only p = tcHd p <* done
+single :: TC a -> TCS a
+single p = tcHd p <* done
   
 -- "hanndle" the reader effect using the state effect?
 list :: TCS a -> TC a
@@ -110,6 +110,12 @@ list k = TC . ReaderT $ \sexp -> StateT $ \ctx -> case _sexp sexp of
   
 several :: TC a -> TCS [a]
 several tc = ([] <$ done) <|> ((:) <$> tcHd tc <*> several tc)
+
+atLeastOne :: TC a -> TCS (NEList a)
[…]

### 2020-02-19 impl/test/Test.hs

@@ -29,7 +29,7 @@ goodTests = "Successful Type Checks" ~: test
   [ assertTC "empty program" ""
   , assertTC "world's smallest module" "(def-mod M (mod () (sig)))"
   , assertTC "sets in params" "(def-mod S (mod ((set X) (set Y)) (sig)))"
-  , assertTC "definition of a set" "(def-mod M (mod ((set X)) (sig (set Y)) (def-set Y X)))"
+  , assertTC "definition of a set" "(def-mod M (mod ((set X)) (def-set Y X)))"
   , assertTC "fun in params"  "(def-mod S (mod ((set X) (fun f X X)) (sig)))"
   , assertTC "more fun in params" "(def-mod S (mod ((set X) (set Y) (fun f Y X)) (sig)))"
   , assertTC "id fun sig" "(def-mod ID (mod ((set X)) (sig) ))"
@@ -49,8 +49,12 @@ goodTests = "Successful Type Checks" ~: test
     "(def-mod S (mod ((set Ob) (span Mor Ob Ob) (trans id ((X Ob)) () (Mor X X)) (trans comp ((X Ob) (Y Ob) (Z Ob)) ((Mor X Y) (Mor Y Z)) (Mor X Z)))))"
   , assertTC "RUP"
     "(def-mod S (mod ((set A) (span HomA A A) (set B) (fun G B A) (span M A B) (trans counit ((b B)) () (M (G b) b)) (trans intro ((a A) (b B)) ((M a b)) (HomA a (G b))))))"
+  , assertTC "ID-trans-params"
+    "(def-mod ID-trans (mod ((set A) (span R A A)) (sig) ))"
   , assertTC "ID-trans"
-  "(def-mod ID-trans (mod ((set A) (span R A A)) (trans id (a A) (x (R a a)) (R a a) x)))"
+    "(def-mod ID-trans (mod ((set A) (span R A A)) (sig) (def-trans id ((a A) (a' A)) ((x (R a a'))) (R a a') x)))"
+  -- , assertTC "trans eta"
+  --   "(def-mod trans-eta (mod ((set A) (set B) (span R A B) (trans foo () () ()) )))"
   ]
 --   , typeChecks (Program good2) ~? "functor signature"
 --   , typeChecks (Program good3) ~? "transformation signature"
@@ -79,6 +83,14 @@ badTests = "Type Checking Failures" ~: test
[…]

### 2020-02-19 impl/src/TypeCheck.hs
finish first impl of transformations, but not yet tested thoroughly

@@ -262,20 +262,24 @@ denoteTrans ctx codSpan (TransEApp r subTerms) =
     return $ transF . subst
   _ -> throwError $ "expected a previously defined transformation, but got...something else"
 
-denoteTransSubst :: NamedSemTransCtx -> EltNF -> EltNF -> SemTransCtx -> [TransExp] -> TCS SemTransSubst
-denoteTransSubst domCtx contraFun covarFun codCtx [] = case (domCtx, codCtx) of
-  (ConsA _ _, _) -> throwError $ "unused variables"
-  (DoneB _, ConsA _ _) -> throwError $ "transformation applied to too few arguments"
-  (DoneB _, DoneB _)   -> do guard $ contraFun == covarFun -- TODO: write an error message, when does this happen?
-                             return $ \_ -> []
-denoteTransSubst domCtx contraFun covarFun codCtx (e:es) = case codCtx of
+denoteTransSubst :: NamedSemTransCtx -> EltNF -> EltNF -> SemTransCtx -> [TransExp] -> TCS ([TransNF] -> [TransNF])
+denoteTransSubst domCtx contraFun covarFun codCtx es = do
+  (leftoverCtx, covarFun', composableSubst) <- denoteTransSubstComp domCtx contraFun codCtx es
+  case (leftoverCtx, covarFun == covarFun') of
+    (ConsA _ _, _) -> throwError $ "unused variables"
+    (_, False) -> throwError "some kind of mismatch, can this even happen?" -- TODO: write a better error message, when does this happen?
+    (DoneB _, True) -> return $ evalState composableSubst
+
+denoteTransSubstComp :: NamedSemTransCtx -> EltNF -> SemTransCtx -> [TransExp] -> TCS (NamedSemTransCtx,EltNF, State [TransNF] [TransNF])
+denoteTransSubstComp domCtx contraFun codCtx [] = case codCtx of
+  ConsA _ _ -> throwError $ "transformation applied to too few arguments"
+  DoneB _   -> return (domCtx, contraFun, return [])
+denoteTransSubstComp domCtx contraFun codCtx (e:es) = case codCtx of
   DoneB _ -> throwError $ "transformation applied to too many arguments"
   ConsA (_,cod) codCtx -> do
     (domCtx, contraFun, prefixEater) <- denoteTransSubstCons domCtx contraFun cod e
-    suffixEater <- denoteTransSubst domCtx contraFun covarFun codCtx es
-    return $ \inps ->
[…]

### 2020-02-19 impl/test/Test.hs

@@ -53,8 +53,12 @@ goodTests = "Successful Type Checks" ~: test
     "(def-mod ID-trans (mod ((set A) (span R A A)) (sig) ))"
   , assertTC "ID-trans"
     "(def-mod ID-trans (mod ((set A) (span R A A)) (sig) (def-trans id ((a A) (a' A)) ((x (R a a'))) (R a a') x)))"
-  -- , assertTC "trans eta"
-  --   "(def-mod trans-eta (mod ((set A) (set B) (span R A B) (trans foo () () ()) )))"
+  , assertTC "trans eta empty"
+    "(def-mod trans-eta (mod ((set A) (span R A A) (trans foo ((a A)) () (R a a))) (sig) (def-trans bar ((a A)) () (R a a) (foo))))"
+  , assertTC "trans eta one-arg"
+    "(def-mod trans-eta (mod ((set A) (set B) (span Q A B) (span R A B) (trans foo ((a A) (b B)) ((Q a b)) (R a b))) (sig) (def-trans bar ((a A) (b B)) ((x (Q a b))) (R a b) (foo x))))"
+  , assertTC "trans eta inst"
+    "(def-mod trans-eta (mod ((set A) (set B) (fun F A B) (span R B B) (trans foo ((b B)) () (R b b))) (sig) (def-trans bar ((a A)) () (R (F a) (F a)) (foo))))"
   ]
 --   , typeChecks (Program good2) ~? "functor signature"
 --   , typeChecks (Program good3) ~? "transformation signature"
@@ -91,7 +95,16 @@ badTests = "Type Checking Failures" ~: test
   ~? "improper duplication of indices"
   , not (typeChecks "(def-mod ID-trans (mod ((set A) (span R A A)) (sig) (def-trans id ((a A) (a' A)) ((x (R a a))) (R a a') x)))")
   ~? "more improper duplication of indices"
-  -- , not (typeChecks "(def-mod S (() (set X)))(def-sig T (sig () (fun R X X)))") ~? "locality of scope"
+  , not (typeChecks  
+    "(def-mod trans-eta (mod ((set A) (set B) (span Q A B) (span R A B) (trans foo ((a A) (b B)) ((Q a b)) (R a b))) (sig) (def-trans bar ((a A) (b B)) ((x (Q a b))) (R a b) (foo))))")
+    ~? "unused var"
+  , not (typeChecks  
[…]

### 2022-02-03 paper/lipics-v2021-sample-article.tex
init fscd draft

\documentclass[a4paper,UKenglish,cleveref, autoref, thm-restate]{lipics-v2021}
%This is a template for producing LIPIcs articles. 
%See lipics-v2021-authors-guidelines.pdf for further information.
%for A4 paper format use option "a4paper", for US-letter use option "letterpaper"
%for british hyphenation rules use option "UKenglish", for american hyphenation rules use option "USenglish"
%for section-numbered lemmas etc., use "numberwithinsect"
%for enabling cleveref support, use "cleveref"
%for enabling autoref support, use "autoref"
%for anonymousing the authors (e.g. for double-blind review), add "anonymous"
%for enabling thm-restate support, use "thm-restate"
%for enabling a two-column layout for the author/affilation part (only applicable for > 6 authors), use "authorcolumns"
%for producing a PDF according the PDF/A standard, add "pdfa"

%\pdfoutput=1 %uncomment to ensure pdflatex processing (mandatatory e.g. to submit to arXiv)
%\hideLIPIcs  %uncomment to remove references to LIPIcs series (logo, DOI, ...), e.g. when preparing a pre-final version to be uploaded to arXiv or another public repository

%\graphicspath{{./graphics/}}%helpful if your graphic files are in another directory

\bibliographystyle{plainurl}% the mandatory bibstyle

\title{A Simple Type Theory for Category Theory} %TODO Please add

%\titlerunning{Dummy short title} %TODO optional, please use if title is longer than one line

\author{Max S. New}
       {University of Michigan Computer Science \& Engineering, Ann Arbor, Michigan, USA \and \url{http://www.maxsnew.com} }
       {maxsnew@umich.edu}
       {}
       {}
       %% {https://orcid.org/0000-0002-1825-0097}
       %% {(Optional) author-specific funding acknowledgements}%TODO mandatory, please use full name
       %% ; only 1 author per \author macro; first two parameters are mandatory, other parameters ca
       %% n be empty. Please provide at least the name of the affiliation and the country. The full address is optional. Use additional curly braces to indicate the correct name splitting when the last name consists of multiple name parts.

\author{Daniel R. Licata}{Wesleyan University, Middletown, Connecticut, USA}
       {dlicata@wesleyan.edu}{}{}

\authorrunning{M.S. New and D.R. Licata} %TODO mandatory. First: Use abbreviated first/middle names. Second (only in severe cases): Use first author plus 'et al.'

\Copyright{Max S. New and Daniel R. Licata} %TODO mandatory, please use full first names. LIPIcs license is "CC-BY";  http://creativecommons.org/licenses/by/3.0/

\ccsdesc[100]{\textcolor{red}{Replace ccsdesc macro with valid one}} %TODO mandatory: Please choose ACM 2012 classifications from https://dl.acm.org/ccs/ccs_flat.cfm

\keywords{Dummy keyword} %TODO mandatory; please add comma-separated list of keywords

\category{} %optional, e.g. invited paper

\relatedversion{} %optional, e.g. full version hosted on arXiv, HAL, or other respository/website
%\relatedversiondetails[linktext={opt. text shown instead of the URL}, cite=DBLP:books/mk/GrayR93]{Classification (e.g. Full Version, Extended Version, Previous Version}{URL to related version} %linktext and cite are optional

%\supplement{}%optional, e.g. related research data, source code, ... hosted on a repository like zenodo, figshare, GitHub, ...
%\supplementdetails[linktext={opt. text shown instead of the URL}, cite=DBLP:books/mk/GrayR93, subcategory={Description, Subcategory}, swhid={Software Heritage Identifier}]{General Classification (e.g. Software, Dataset, Model, ...)}{URL to related version} %linktext, cite, and subcategory are optional

%\funding{(Optional) general funding statement \dots}%optional, to capture a funding statement, which applies to all authors. Please enter author specific funding statements as fifth argument of the \author macro.

%% \acknowledgements{I want to thank \dots}%optional

%\nolinenumbers %uncomment to disable line numbering

%Editor-only macros:: begin (do not touch as author)%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\EventEditors{John Q. Open and Joan R. Access}
\EventNoEds{2}
\EventLongTitle{42nd Conference on Very Important Topics (CVIT 2016)}
\EventShortTitle{CVIT 2016}
\EventAcronym{CVIT}
\EventYear{2016}
\EventDate{December 24--27, 2016}
\EventLocation{Little Whinging, United Kingdom}
\EventLogo{}
\SeriesVolume{42}
\ArticleNo{23}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

\maketitle

%TODO mandatory: add short abstract of the document
\begin{abstract}
  We present a domain-specific type theory for formal category theory
  constructions and proofs. The type theory axiomatizes notions of
  category, functor, profunctor and a generalized form of natural
  transformations. By restricting the type theory to a form of ordred
  logic, we guarantee that all mappings between categories are
  functorial and all transformations are natural by construction, with
  no separate proof necessary. We demonstrate the utility of the type
  theory by proving a form of the Yoneda lemma which reads as the
  classical proof, but with all naturality side-conditions
  automatically satisfied.

  We show that the type theory is sound and complete for virtual
  equipments, a kind of double category which is known to model not
  just ordinary categories but enriched and internal categories as
  well. This means that proofs and constructions in our theory hold
  not just for ordinary categories but also for a broad class of
  generalized categories. We demonstrate that the type theory
  simplifies working with virtual equipments compared to their
  algebraic description in the literature.
\end{abstract}

\section{Introduction}

Category theory is a branch of mathematics that formalizes the
algebraic structure of function composition. The high level of
abstraction allows for succinct proofs relevant to structures across a
broad variety of mathematical domains. Category theory is of
particular interest in the theory of programming languages and proof
assistants, where it is commonly used to aid in the design of typed
languages, proof assistants and functional programming abstractions.

Novices in category theory are often overwhelmed by the number of
naturality and functoriality side conditions that abound in otherwise
simple proofs in category theory. For instance consider the proof of
the Yoneda lemma, often the first ``big'' proof in an introductory
category theory text. ... Here the reader is left to themselves to
show that the ``evaluation at identity'' morphism and its inverse are
actually natural transformations. They will then work through a
diagram chase to determine that this naturality condition indeed
holds.

Experienced category theorists have no trouble with this, as they are
familiar with a great many principles for constructing transformations
that are guaranteed to be natural. Consider an analogy with
calculus. When a mathematician writes a formula such as

$$\sin(f(x) - g(x))$$

They do not need to spend time with an explicit $\epsilon,\delta$
proof that the function is continuous or furthermore smooth. Instead,
they follow syntactic principles that compositions of continuous
functions are continuous.

There are similar syntactic notions for guaranteeing that
transformations are natural.

\subsection{Profunctors}

The syntax of Judgment Theory gives not just a syntax for the familiar
trio of category, functor and natural transformation, but additionally
a concept that is not usually highlighted in introductory category
theory texts: the profunctor. A simple intuition is that profunctors
are to functors as relations are to functions. Indeed, the syntax for
judgment theory has a straightforward semantics in the category of
sets and relations.

The reason profunctors are not commonly introduced is that they are
seen as an auxiliary notion: a profunctor $R : C \relto D$ is
typically defined as a functor $C^{\textrm{op}}\times D \to
\textrm{Set}$. However this definition requires two concepts that are
somewhat problematic: the opposite category $C^\textrm{op}$ and the
category of (small) sets $\textrm{Set}$. However, these notions do not
necessarily exist in all of our intended applications. First, we are
interested in semantics where are objects are \emph{multicategories}
where morphisms have multiple inputs but only a single output. In this
case the opposite of a multicategory is not a multicategory at all but
instead some kind of ``co-multicategory'' with multiple outputs but
one input. Second, having a universe object like $\textrm{Set}$ is also

\subsection{Applications to Enriched, Internal and Multicategory Theory}

The syntax of judgment theory is an internal language for
\emph{virtual equipments}. These are a kind of joint generalization of
double categories and multicategories.

Notably, the definition of a category in \secref{sec:internal} can be
applied as an extremely general notion of category, including internal
categories and enriched categories. This general notion is often
called a ``generalized multi-category'' and encompasses an astonishing
range of mathematical structures, including categories, topological
spaces, metric spaces, etc.

\section{Syntax of Judgment Theory}

\subsection{Positive Multiplicatives}

\subsection{Negative Multiplicatives}

\subsection{Additives}

\section{Synthetic Category Theory}

\section{Internal Category Theory}

\subsection{Duality}

\section{Semantics in Virtual Equipments}

\section{Related Work}

Cartesian Bicategories.

Directed Type theory

Fibrational approaches to logical relations

Double-cat-like syntax: Gradual Type Theory, System P of parametricity, cubical stuff?

String Diagrams?

%% \section{Typesetting instructions -- Summary}
%% \label{sec:typesetting-summary}

%% LIPIcs is a series of open access high-quality conference proceedings across all fields in informatics established in cooperation with Schloss Dagstuhl. 
%% In order to do justice to the high scientific quality of the conferences that publish their proceedings in the LIPIcs series, which is ensured by the thorough review process of the respective events, we believe that LIPIcs proceedings must have an attractive and consistent layout matching the standard of the series.
%% Moreover, the quality of the metadata, the typesetting and the layout must also meet the requirements of other external parties such as indexing service, DOI registry, funding agencies, among others. The guidelines contained in this document serve as the baseline for the authors, editors, and the publisher to create documents that meet as many different requirements as possible.

%% Please comply with the following instructions when preparing your article for a LIPIcs proceedings volume. 
%% \paragraph*{Minimum requirements}

%% \begin{itemize}
%% \item Use pdflatex and an up-to-date \LaTeX{} system.
%% \item Use further \LaTeX{} packages and custom made macros carefully and only if required.
%% \item Use the provided sectioning macros: \verb+\section+, \verb+\subsection+, \verb+\subsubsection+, \linebreak \verb+\paragraph+, \verb+\paragraph*+, and \verb+\subparagraph*+.
%% \item Provide suitable graphics of at least 300dpi (preferably in PDF format).
%% \item Use BibTeX and keep the standard style (\verb+plainurl+) for the bibliography.
%% \item Please try to keep the warnings log as small as possible. Avoid overfull \verb+\hboxes+ and any kind of warnings/errors with the referenced BibTeX entries.
%% \item Use a spellchecker to correct typos.
%% \end{itemize}

%% \paragraph*{Mandatory metadata macros}
%% Please set the values of the metadata macros carefully since the information parsed from these macros will be passed to publication servers, catalogues and search engines.
%% Avoid placing macros inside the metadata macros. The following metadata macros/environments are mandatory:
%% \begin{itemize}
%% \item \verb+\title+ and, in case of long titles, \verb+\titlerunning+.
%% \item \verb+\author+, one for each author, even if two or more authors have the same affiliation.
%% \item \verb+\authorrunning+ and \verb+\Copyright+ (concatenated author names)\\
%% The \verb+\author+ macros and the \verb+\Copyright+ macro should contain full author names (especially with regard to the first name), while \verb+\authorrunning+ should contain abbreviated first names.
%% \item \verb+\ccsdesc+ (ACM classification, see \url{https://www.acm.org/publications/class-2012}).
%% \item \verb+\keywords+ (a comma-separated list of keywords).
%% \item \verb+\relatedversion+ (if there is a related version, typically the ``full version''); please make sure to provide a persistent URL, e.\,g., at arXiv.
%% \item \verb+\begin{abstract}...\end{abstract}+ .
%% \end{itemize}

%% \paragraph*{Please do not \ldots} %Do not override the \texttt{\seriesstyle}-defaults}
%% Generally speaking, please do not override the \texttt{lipics-v2021}-style defaults. To be more specific, a short checklist also used by Dagstuhl Publishing during the final typesetting is given below.
%% In case of \textbf{non-compliance} with these rules Dagstuhl Publishing will remove the corresponding parts of \LaTeX{} code and \textbf{replace it with the \texttt{lipics-v2021} defaults}. In serious cases, we may reject the LaTeX-source and expect the corresponding author to revise the relevant parts.
%% \begin{itemize}
%% \item Do not use a different main font. (For example, the \texttt{times} package is forbidden.)
%% \item Do not alter the spacing of the \texttt{lipics-v2021.cls} style file.
%% \item Do not use \verb+enumitem+ and \verb+paralist+. (The \texttt{enumerate} package is preloaded, so you can use
%%  \verb+\begin{enumerate}[(a)]+ or the like.)
%% \item Do not use ``self-made'' sectioning commands (e.\,g., \verb+\noindent{\bf My+ \verb+Paragraph}+).
%% \item Do not hide large text blocks using comments or \verb+\iffalse+ $\ldots$ \verb+\fi+ constructions. 
%% \item Do not use conditional structures to include/exclude content. Instead, please provide only the content that should be published -- in one file -- and nothing else.
%% \item Do not wrap figures and tables with text. In particular, the package \texttt{wrapfig} is not supported.
%% \item Do not change the bibliography style. In particular, do not use author-year citations. (The
%% \texttt{natbib} package is not supported.)
%% \end{itemize}

%% \enlargethispage{\baselineskip}

%% This is only a summary containing the most relevant details. Please read the complete document ``LIPIcs: Instructions for Authors and the \texttt{lipics-v2021} Class'' for all details and don't hesitate to contact Dagstuhl Publishing (\url{mailto:publishing@dagstuhl.de}) in case of questions or comments:
%% \href{http://drops.dagstuhl.de/styles/lipics-v2021/lipics-v2021-authors/lipics-v2021-authors-guidelines.pdf}{\texttt{http://drops.dagstuhl.de/styles/lipics-v2021/\newline lipics-v2021-authors/lipics-v2021-authors-guidelines.pdf}}

%% \section{Lorem ipsum dolor sit amet}

%% Lorem ipsum dolor sit amet, consectetur adipiscing elit \cite{DBLP:journals/cacm/Knuth74}. Praesent convallis orci arcu, eu mollis dolor. Aliquam eleifend suscipit lacinia. Maecenas quam mi, porta ut lacinia sed, convallis ac dui. Lorem ipsum dolor sit amet, consectetur adipiscing elit. Suspendisse potenti. Donec eget odio et magna ullamcorper vehicula ut vitae libero. Maecenas lectus nulla, auctor nec varius ac, ultricies et turpis. Pellentesque id ante erat. In hac habitasse platea dictumst. Curabitur a scelerisque odio. Pellentesque elit risus, posuere quis elementum at, pellentesque ut diam. Quisque aliquam libero id mi imperdiet quis convallis turpis eleifend.

%% \begin{lemma}[Lorem ipsum]
%% \label{lemma:lorem}
%% Vestibulum sodales dolor et dui cursus iaculis. Nullam ullamcorper purus vel turpis lobortis eu tempus lorem semper. Proin facilisis gravida rutrum. Etiam sed sollicitudin lorem. Proin pellentesque risus at elit hendrerit pharetra. Integer at turpis varius libero rhoncus fermentum vitae vitae metus.
%% \end{lemma}

%% \begin{proof}
%% Cras purus lorem, pulvinar et fermentum sagittis, suscipit quis magna.

%% \proofsubparagraph*{Just some paragraph within the proof.}
%% Nam liber tempor cum soluta nobis eleifend option congue nihil imperdiet doming id quod mazim placerat facer possim assum. Lorem ipsum dolor sit amet, consectetuer adipiscing elit, sed diam nonummy nibh euismod tincidunt ut laoreet dolore magna aliquam erat volutpat.
%% \begin{claim}
%% content...
%% \end{claim}
%% \begin{claimproof}
%% content...
%%     \begin{enumerate}
%%         \item abc abc abc \claimqedhere{}
%%     \end{enumerate}
%% \end{claimproof}

%% \end{proof}

%% \begin{corollary}[Curabitur pulvinar, \cite{DBLP:books/mk/GrayR93}]
%% \label{lemma:curabitur}
%% Nam liber tempor cum soluta nobis eleifend option congue nihil imperdiet doming id quod mazim placerat facer possim assum. Lorem ipsum dolor sit amet, consectetuer adipiscing elit, sed diam nonummy nibh euismod tincidunt ut laoreet dolore magna aliquam erat volutpat.
%% \end{corollary}

%% \begin{proposition}\label{prop1}
%% This is a proposition
%% \end{proposition}

%% \autoref{prop1} and \cref{prop1} \ldots

%% \subsection{Curabitur dictum felis id sapien}

%% Curabitur dictum \cref{lemma:curabitur} felis id sapien \autoref{lemma:curabitur} mollis ut venenatis tortor feugiat. Curabitur sed velit diam. Integer aliquam, nunc ac egestas lacinia, nibh est vehicula nibh, ac auctor velit tellus non arcu. Vestibulum lacinia ipsum vitae nisi ultrices eget gravida turpis laoreet. Duis rutrum dapibus ornare. Nulla vehicula vulputate iaculis. Proin a consequat neque. Donec ut rutrum urna. Morbi scelerisque turpis sed elit sagittis eu scelerisque quam condimentum. Pellentesque habitant morbi tristique senectus et netus et malesuada fames ac turpis egestas. Aenean nec faucibus leo. Cras ut nisl odio, non tincidunt lorem. Integer purus ligula, venenatis et convallis lacinia, scelerisque at erat. Fusce risus libero, convallis at fermentum in, dignissim sed sem. Ut dapibus orci vitae nisl viverra nec adipiscing tortor condimentum \cite{DBLP:journals/cacm/Dijkstra68a}. Donec non suscipit lorem. Nam sit amet enim vitae nisl accumsan pretium.

%% \begin{lstlisting}[caption={Useless code.},label=list:8-6,captionpos=t,float,abovecaptionskip=-\medskipamount]
%% for i:=maxint to 0 do 
%% begin 
%%     j:=square(root(i));
%% end;
%% \end{lstlisting}

%% \subsection{Proin ac fermentum augue}

%% Proin ac fermentum augue. Nullam bibendum enim sollicitudin tellus egestas lacinia euismod orci mollis. Nulla facilisi. Vivamus volutpat venenatis sapien, vitae feugiat arcu fringilla ac. Mauris sapien tortor, sagittis eget auctor at, vulputate pharetra magna. Sed congue, dui nec vulputate convallis, sem nunc adipiscing dui, vel venenatis mauris sem in dui. Praesent a pretium quam. Mauris non mauris sit amet eros rutrum aliquam id ut sapien. Nulla aliquet fringilla sagittis. Pellentesque eu metus posuere nunc tincidunt dignissim in tempor dolor. Nulla cursus aliquet enim. Cras sapien risus, accumsan eu cursus ut, commodo vel velit. Praesent aliquet consectetur ligula, vitae iaculis ligula interdum vel. Integer faucibus faucibus felis.

%% \begin{itemize}
%% \item Ut vitae diam augue. 
%% \item Integer lacus ante, pellentesque sed sollicitudin et, pulvinar adipiscing sem. 
%% \item Maecenas facilisis, leo quis tincidunt egestas, magna ipsum condimentum orci, vitae facilisis nibh turpis et elit. 
%% \end{itemize}

%% \begin{remark}
%% content...
%% \end{remark}

%% \section{Pellentesque quis tortor}

%% Nec urna malesuada sollicitudin. Nulla facilisi. Vivamus aliquam tempus ligula eget ornare. Praesent eget magna ut turpis mattis cursus. Aliquam vel condimentum orci. Nunc congue, libero in gravida convallis \cite{DBLP:conf/focs/HopcroftPV75}, orci nibh sodales quam, id egestas felis mi nec nisi. Suspendisse tincidunt, est ac vestibulum posuere, justo odio bibendum urna, rutrum bibendum dolor sem nec tellus.

%% \begin{lemma} [Quisque blandit tempus nunc]
%% Sed interdum nisl pretium non. Mauris sodales consequat risus vel consectetur. Aliquam erat volutpat. Nunc sed sapien ligula. Proin faucibus sapien luctus nisl feugiat convallis faucibus elit cursus. Nunc vestibulum nunc ac massa pretium pharetra. Nulla facilisis turpis id augue venenatis blandit. Cum sociis natoque penatibus et magnis dis parturient montes, nascetur ridiculus mus.
%% \end{lemma}

%% Fusce eu leo nisi. Cras eget orci neque, eleifend dapibus felis. Duis et leo dui. Nam vulputate, velit et laoreet porttitor, quam arcu facilisis dui, sed malesuada risus massa sit amet neque.

%% \section{Morbi eros magna}

%% Morbi eros magna, vestibulum non posuere non, porta eu quam. Maecenas vitae orci risus, eget imperdiet mauris. Donec massa mauris, pellentesque vel lobortis eu, molestie ac turpis. Sed condimentum convallis dolor, a dignissim est ultrices eu. Donec consectetur volutpat eros, et ornare dui ultricies id. Vivamus eu augue eget dolor euismod ultrices et sit amet nisi. Vivamus malesuada leo ac leo ullamcorper tempor. Donec justo mi, tempor vitae aliquet non, faucibus eu lacus. Donec dictum gravida neque, non porta turpis imperdiet eget. Curabitur quis euismod ligula.

%% %%
%% %% Bibliography
%% %%

%% %% Please use bibtex,

\bibliography{lipics-v2021-sample-article}

\appendix

\section{Additional Rules}

%% \section{Styles of lists, enumerations, and descriptions}\label{sec:itemStyles}

%% List of different predefined enumeration styles:

%% \begin{itemize}
%% \item \verb|\begin{itemize}...\end{itemize}|
%% \item \dots
%% \item \dots
%% %\item \dots
%% \end{itemize}

%% \begin{enumerate}
%% \item \verb|\begin{enumerate}...\end{enumerate}|
%% \item \dots
%% \item \dots
%% %\item \dots
%% \end{enumerate}

%% \begin{alphaenumerate}
%% \item \verb|\begin{alphaenumerate}...\end{alphaenumerate}|
%% \item \dots
%% \item \dots
%% %\item \dots
%% \end{alphaenumerate}

%% \begin{romanenumerate}
%% \item \verb|\begin{romanenumerate}...\end{romanenumerate}|
%% \item \dots
%% \item \dots
%% %\item \dots
%% \end{romanenumerate}

%% \begin{bracketenumerate}
%% \item \verb|\begin{bracketenumerate}...\end{bracketenumerate}|
%% \item \dots
%% \item \dots
%% %\item \dots
%% \end{bracketenumerate}

%% \begin{description}
%% \item[Description 1] \verb|\begin{description} \item[Description 1]  ...\end{description}|
%% \item[Description 2] Fusce eu leo nisi. Cras eget orci neque, eleifend dapibus felis. Duis et leo dui. Nam vulputate, velit et laoreet porttitor, quam arcu facilisis dui, sed malesuada risus massa sit amet neque.
%% \item[Description 3]  \dots
%% %\item \dots
%% \end{description}

%% \cref{testenv-proposition} and \autoref{testenv-proposition} ...

%% \section{Theorem-like environments}\label{sec:theorem-environments}

%% List of different predefined enumeration styles:

%% \begin{theorem}\label{testenv-theorem}
%% Fusce eu leo nisi. Cras eget orci neque, eleifend dapibus felis. Duis et leo dui. Nam vulputate, velit et laoreet porttitor, quam arcu facilisis dui, sed malesuada risus massa sit amet neque.
%% \end{theorem}

%% \begin{lemma}\label{testenv-lemma}
%% Fusce eu leo nisi. Cras eget orci neque, eleifend dapibus felis. Duis et leo dui. Nam vulputate, velit et laoreet porttitor, quam arcu facilisis dui, sed malesuada risus massa sit amet neque.
%% \end{lemma}

%% \begin{corollary}\label{testenv-corollary}
%% Fusce eu leo nisi. Cras eget orci neque, eleifend dapibus felis. Duis et leo dui. Nam vulputate, velit et laoreet porttitor, quam arcu facilisis dui, sed malesuada risus massa sit amet neque.
%% \end{corollary}

%% \begin{proposition}\label{testenv-proposition}
%% Fusce eu leo nisi. Cras eget orci neque, eleifend dapibus felis. Duis et leo dui. Nam vulputate, velit et laoreet porttitor, quam arcu facilisis dui, sed malesuada risus massa sit amet neque.
%% \end{proposition}

%% \begin{conjecture}\label{testenv-conjecture}
%% Fusce eu leo nisi. Cras eget orci neque, eleifend dapibus felis. Duis et leo dui. Nam vulputate, velit et laoreet porttitor, quam arcu facilisis dui, sed malesuada risus massa sit amet neque.
%% \end{conjecture}

%% \begin{observation}\label{testenv-observation}
%% Fusce eu leo nisi. Cras eget orci neque, eleifend dapibus felis. Duis et leo dui. Nam vulputate, velit et laoreet porttitor, quam arcu facilisis dui, sed malesuada risus massa sit amet neque.
%% \end{observation}

%% \begin{exercise}\label{testenv-exercise}
%% Fusce eu leo nisi. Cras eget orci neque, eleifend dapibus felis. Duis et leo dui. Nam vulputate, velit et laoreet porttitor, quam arcu facilisis dui, sed malesuada risus massa sit amet neque.
%% \end{exercise}

%% \begin{definition}\label{testenv-definition}
%% Fusce eu leo nisi. Cras eget orci neque, eleifend dapibus felis. Duis et leo dui. Nam vulputate, velit et laoreet porttitor, quam arcu facilisis dui, sed malesuada risus massa sit amet neque.
%% \end{definition}

%% \begin{example}\label{testenv-example}
%% Fusce eu leo nisi. Cras eget orci neque, eleifend dapibus felis. Duis et leo dui. Nam vulputate, velit et laoreet porttitor, quam arcu facilisis dui, sed malesuada risus massa sit amet neque.
%% \end{example}

%% \begin{note}\label{testenv-note}
%% Fusce eu leo nisi. Cras eget orci neque, eleifend dapibus felis. Duis et leo dui. Nam vulputate, velit et laoreet porttitor, quam arcu facilisis dui, sed malesuada risus massa sit amet neque.
%% \end{note}

%% \begin{note*}
%% Fusce eu leo nisi. Cras eget orci neque, eleifend dapibus felis. Duis et leo dui. Nam vulputate, velit et laoreet porttitor, quam arcu facilisis dui, sed malesuada risus massa sit amet neque.
%% \end{note*}

%% \begin{remark}\label{testenv-remark}
%% Fusce eu leo nisi. Cras eget orci neque, eleifend dapibus felis. Duis et leo dui. Nam vulputate, velit et laoreet porttitor, quam arcu facilisis dui, sed malesuada risus massa sit amet neque.
%% \end{remark}

%% \begin{remark*}
%% Fusce eu leo nisi. Cras eget orci neque, eleifend dapibus felis. Duis et leo dui. Nam vulputate, velit et laoreet porttitor, quam arcu facilisis dui, sed malesuada risus massa sit amet neque.
%% \end{remark*}

%% \begin{claim}\label{testenv-claim}
%% Fusce eu leo nisi. Cras eget orci neque, eleifend dapibus felis. Duis et leo dui. Nam vulputate, velit et laoreet porttitor, quam arcu facilisis dui, sed malesuada risus massa sit amet neque.
%% \end{claim}

%% \begin{claim*}\label{testenv-claim2}
%% Fusce eu leo nisi. Cras eget orci neque, eleifend dapibus felis. Duis et leo dui. Nam vulputate, velit et laoreet porttitor, quam arcu facilisis dui, sed malesuada risus massa sit amet neque.
%% \end{claim*}

%% \begin{proof}
%% Fusce eu leo nisi. Cras eget orci neque, eleifend dapibus felis. Duis et leo dui. Nam vulputate, velit et laoreet porttitor, quam arcu facilisis dui, sed malesuada risus massa sit amet neque.
%% \end{proof}

%% \begin{claimproof}
%% Fusce eu leo nisi. Cras eget orci neque, eleifend dapibus felis. Duis et leo dui. Nam vulputate, velit et laoreet porttitor, quam arcu facilisis dui, sed malesuada risus massa sit amet neque.
%% \end{claimproof}

\end{document}

### 2022-02-04 paper/lipics-v2021-sample-article.tex
work on fscd version

@@
 \documentclass[a4paper,UKenglish,cleveref, autoref, thm-restate]{lipics-v2021}
+
+\usepackage[margin=1in]{geometry}
+\usepackage{tikz-cd}
+\usepackage{amssymb}
+\usepackage{amsmath}
+\usepackage{amsthm}
+\usepackage{mathtools}
+\usepackage{stmaryrd}
+\usepackage{mathpartir}
+\usepackage{bbm}
+\usepackage{cleveref}
+
+\DeclareMathOperator*{\colim}{colim}
+
 %This is a template for producing LIPIcs articles. 
@@
 \begin{document}
+%% \newtheorem{theorem}{Theorem}
+%% \newtheorem{lemma}{Lemma}
+%% \newtheorem{construction}{Construction}
+%% \newtheorem{definition}{Definition}
+\newcommand{\pto}{\nrightarrow}
+\newcommand{\pfrom}{\nleftarrow}
+\newcommand{\vcat}{\mathcal}
+\newcommand{\cat}{\mathbbm}
+\newcommand{\isaSet}{\,\,\textrm{Set}}
+\newcommand{\Set}{\textrm{Set}}
+\newcommand{\vtkmnd}{\mathbb{K}\text{Mod} (\vcat{V},T)}
+\newcommand{\rmod}{\text{RMod}}
+\newcommand{\lmod}{\text{LMod}}
+\newcommand{\varr}[2]{\text{Fun}{#1}{#2}}
+\newcommand{\harr}[2]{\text{Rel}{#1}{#2}}
+\newcommand{\harrapp}[3]{#1(#2;#3)}
+
+\newcommand{\jnctx}{\curlyveedownarrow}
+
+\newcommand{\id}{\textrm{id}}
+\newcommand{\for}{\textrm{for}\,}
+\newcommand{\when}{\textrm{when}\,}
+\newcommand{\lett}{\textrm{let}\,}
+%% \newcommand{\sort}{\,\,\textrm{sort}}
+\newcommand{\Sort}{\textrm{Sort}}
+\newcommand{\isadtctx}{\,\,\textrm{type context}}
+\newcommand{\isavectx}{\,\,\textrm{cat context}}
+%% \newcommand{\obctx}{\,\,\textrm{object context}}
+\newcommand{\boundary}{\,\,\textrm{boundary}}
+\newcommand{\prof}{\,\,\textrm{span}}
+\newcommand{\subst}{\,\,\textrm{subst}}
+\newcommand{\sigctx}{\,\,\textrm{sig-ctx}}
+\newcommand{\sig}{\,\,\textrm{sig}}
+\newcommand{\pipe}{\mathrel{|}}
+
+\newcommand{\punitinXfromYtoZ}[3]{#2 \mathop{\to_{#1}} #3}
+\newcommand{\punitrefl}[1]{\textrm{id}_{#1}}
+\newcommand{\punitelimtoYwithkontZ}[2]{\textrm{ind}_{\to}(#1,#2)}
+
+\newcommand{\odotexists}[1]{\mathop{\overset{\exists #1}\odot}}
+\newcommand{\tensorexistsXwithYandZ}[3]{#2 \odotexists{#1} #3}
+\newcommand{\tensorintroofXandY}[2]{\textrm{pair}_\odot(#1,#2)}
+\newcommand{\tensorelimXtoYwithkontZ}[3]{\textrm{ind}_{\to}(#1,#2,#3)}
+\newcommand{\tensorelimfromWXtoYwithkontZ}[4]{\textrm{ind}_{\odot}(#1,#2,#3,#4)}
+\newcommand{\tensorelimVatWtoXtoYwithkontZ}[5]{\textrm{ind}_{\to}(#4,#2,#1,#3,#5)}
+
+\newcommand{\tlwith}[1]{\mathop{\prescript{#1}{}\triangleleft}}
+\newcommand{\tlforall}[1]{\tlwith{\forall #1}}
+
+\newcommand{\trwith}[1]{\mathop{\triangleright^{#1}}}
+\newcommand{\trforall}[1]{\trwith {\forall #1}}
+
+\newcommand{\homrallXYtoZ}[3]{#2 \trforall {#1} #3}
+
+\newcommand{\homlallXYtoZ}[3]{#2 \tlforall #1 #3}
+
+\newcommand{\homrlambdaXatYdotZ}[3]{\lambda^\triangleright (#1, #2). #3}
+\newcommand{\homllambdaXatYdotZ}[3]{\lambda^\triangleleft (#2, #1). #3}
+
+\newcommand{\homrappXtoYatZ}[3]{#1 \trwith{#3} #2}
+\newcommand{\homlappXtoYatZ}[3]{#2 \tlwith{#3} #1}
+
+\newcommand{\homunary}[4]{\pendallXdotY {#1} {\homrallXYtoZ {#3} {#2} {#4}}}
+
+\newcommand{\lambdaunary}[4]{\pendlambdaXdotY {#1} {\homrlambdaXatYdotZ {#2} {#3} {#4}}}
+\newcommand{\lambdabinary}[6]{\lambdaunary{#1}{#2}{#3}{\homrlambdaXatYdotZ{#4}{#5}{#6}}}
+\newcommand{\lambdatrinary}[8]{\lambdabinary{#1}{#2}{#3}{#4}{#5}{\homrlambdaXatYdotZ{#6}{#7}{#8}}}
+
+\newcommand{\appunary}[4]{\homrappXtoYatZ {\pendappXtoY {#1} {#2}} {#3} {#4}}
+\newcommand{\appbinary}[6]{\homrappXtoYatZ {\appunary{#1}{#2}{#3}{#4}}{#5}{#6}}
+\newcommand{\apptrinary}[8]{\homrappXtoYatZ {\appbinary{#1}{#2}{#3}{#4}{#5}{#6}}{#7}{#8}}
+
+\newcommand{\pendallXdotY}[2]{\forall #1. #2}
+\newcommand{\pendlambdaXdotY}[2]{\lambda #1. #2}
+\newcommand{\pendappXtoY}[2]{#1^{#2}}
+
+\newcommand{\equalizeVofWbyXeqYatZ}[5]{\{ {#1} : {#2} \pipe {#3} = {#4} : {#5} \}}
 
@@
 The reason profunctors are not commonly introduced is that they are
-seen as an auxiliary notion: a profunctor $R : C \relto D$ is
+seen as an auxiliary notion: a profunctor $R : C \pto D$ is
 typically defined as a functor $C^{\textrm{op}}\times D \to
@@
 
-Notably, the definition of a category in \secref{sec:internal} can be
+Notably, the definition of a category in \cref{sec:internal} can be
 applied as an extremely general notion of category, including internal
@@
 
-\section{Syntax of Judgment Theory}
-
-\subsection{Positive Multiplicatives}
-
-\subsection{Negative Multiplicatives}
-
-\subsection{Additives}
+\section{Syntax of SCT}
+
+\begin{figure}
+  \begin{mathpar}
+    \begin{array}{rcl}
+      \textrm{Type} & A,B,C ::= &
+      \sum_{X:A} B \pipe \prod_{X:A} B \pipe \Id A M N \\
+      \Cat \pipe \varr{\cat C}{\cat D}\pipe \harr {\cat C} {\cat D}\pipe \pendallXdotY {\alpha:{\cat C}} R\\
+      \textrm{Term} & L,M,N ::= & X \pipe \lambda X:M. N \pipe M\,N \pipe (M,N) \pipe \pi_1 M \pipe \pi_2 M\\
+      &&\lceil \cat C \rceil \pipe \vlambda {\alpha:\cat C} a \pipe \hlambda {\alpha:\cat C}{\beta: \cat D} R \ldots \\
+      \textrm{Categories} &\cat C, \cat D, \cat E ::= & M \pipe \cat C \times \cat D \pipe \cat 1\\
+      \textrm{Objects} & a, b, c ::= & \alpha \pipe M a \ldots\\
+      \textrm{Sets} & P,Q,R ::= & \ldots \\
+      \textrm{Transformations} & s,t,u & ::= & \ldots \\
+      \\
+      \textrm{Type Context} & \Gamma ::= & \cdot \pipe \Gamma, X : A\\
+      \textrm{Cat Context} & \Phi,\Psi ::= & \cdot \pipe \alpha : \cat C \pipe \Phi , x : P , \Psi\\
+    \end{array}
+  \end{mathpar}
+  \caption{Simple Category Theory Syntactic Forms}
+\end{figure}
+
+\subsection{Judgmental Structure}
+
+The following figures present the basic formation, primitive and
+admissible equality rules of our type theory, an internal language for
+virtual double categories with restrictions.
+%
+The type theory has 4 central concepts, which we will refer to
+sometimes using type theoretic/set theoretic and at other times
+category-theoretic terminology.
+%
+In type theoretic terminology, judgment theory has \emph{sorts} $\cat
+C, \cat D, \cat E$, and \emph{types} $A,B,C$ which have a given sort
+and are also \emph{parameterized} by an object of some other sort, and
+so can also be thought of as \emph{functions} with input and output
+sorts.
+%
+Next there are \emph{judgments} $P,Q,R,S$ which are parameterized by
+two variables.
+%
+In $\alpha^o:\cat C;\beta : \cat D \pipe R$, we say $R$ depends on
+$\alpha$ \emph{contravariantly} and in $\beta$ \emph{covariantly}.
+%
+Finally, we have \emph{terms} $s,t,u$ that are elements of some
+judgment $R$ but are also parameterized by elements of some other
+judgments which are given by a \emph{context} $\Phi$, which is a
+``compatible string'' of judgments where the covariant variable of one
+judgment is the same as the contravariant variable of the next.
+%
+Since these terms are parameterized by other terms, we can see these
+as \emph{inference rules} for the judgments.
+
+At other times we will want to think of these using set-theoretic
+terminology, in which case we have sets $\cat C,\cat D, \cat E$,
+functions/elements $A,B,C$, \emph{spans} $R,\ldots$ and morphisms of
+spans $s,t,u$.
+%
+Finally when we develop multicategory theory we will use
+category-theoretic terminology, in which the structure is most
+familiar and we have categories, objects/functors,
+\emph{profunctors}/bimodules and 2-cells/homomorphisms of bimodules.
+
+Figure\ref{fig:formation} introduces the formation rules for the basic
+constructs. Sorts give a syntax for objects of the double
+category. Types have an output sort and a sorted input variable, and
+model vertical arrows. Next, judgments have two sorted variables, one
+contravariant and one covariant, which can be the same
+variable. Judgments model the horizontal arrows. Next are contexts,
+whose formation rule is the same as judgments, and which give a syntax
+for ``compatible strings'' of judgments. Then terms have an input
+context and an output judgment, which must have the same covariant and
+contravaiant variables, modeling ``globular'' 2-cells. Finally, we
+have substitutions for terms, which are typed with respect to an input
+context, an output context and two types, forming a kind of
+square. These model the ``compatible string'' of 2-cells that arise in
+the composition laws.
+
+\begin{figure}
+  \begin{mathpar}
+    \inferrule
+        {}
+        {\Gamma \isadtctx}
+
+    \inferrule
+    {\Gamma \isadtctx}
+    {\Gamma \vdash \Xi \boundary}
+
+    \inferrule
+    {\Gamma \isadtctx \and \Gamma \vdash \Xi \boundary}
+    {\Gamma \pipe \Xi \vdash S \isaSet}
+
+    \inferrule*[right=Object/Type-formation]
+    {\Gamma \vdash \cat C : \Sort \and \Gamma \vdash \cat D : \Sort}
+    {\Gamma \pipe \alpha : \cat C \vdash A : \cat D}\\
+
+    \inferrule*[right=Context-formation]
+    {\Gamma \isadtctx}
+    {\Gamma \vdash \Phi \isavectx}
+
+    \inferrule*[right=Term formation]
+    {\Gamma \vdash \Phi \isavectx\and
+     \Gamma\pipe \underline \Phi \vdash R \isaSet}
+    {\Gamma \pipe \Phi \vdash M : R}
+
+    \inferrule
+    {\Delta \isadtctx \and \Gamma \isadtctx}
+    {\gamma : \Delta \to \Gamma}
+
+    \inferrule*[right=Boundary Map formation]
+    {\Gamma \vdash Z \boundary \and \Gamma \vdash \Xi \boundary}
+    {\Gamma \vdash \xi : Z \to \Xi}
+
+    \inferrule*[right=Cat Subst formation]
+    {\Gamma \vdash \Psi \isavectx \and \Gamma \vdash \Phi \isavectx}
+    {\Gamma \vdash \phi : \Psi \to \Phi}
+  \end{mathpar}
+  \caption{Formation Rules}
+  \label{fig:formation}
+\end{figure}
+
+Next, Figure~\ref{fig:basic-rules} gives the basic structural rules of
+judgment theory, which model composition and restrictions in a virtual
+double category. First, we have varialbes and substitution for types,
+modeling identity and composition of vertical arrows. Next, we have
+restriction of a judgment to more specific types, which is given by
+substitution of types for the judgment's parameter variables. Next, we
+have rules for constructing an ``empty'' context, which is just a
+
+\begin{figure}
+  \begin{mathpar}
+    \inferrule{}{\cdot \isadtctx}
+
+    \inferrule
+    {\Gamma \isadtctx \and \Gamma\pipe \cdot \vdash A \isaSet}
+    {\Gamma , x : A \isadtctx}\\
+
+    \inferrule{}{\Gamma \vdash \cdot \boundary}
+
+    \inferrule{\Gamma \vdash C : \Sort}{\Gamma \vdash \alpha: C \boundary}
+
+    \inferrule{\Gamma \vdash C : \Sort\and\Gamma\vdash D : \Sort}{\Gamma \vdash \alpha: C;\beta:D \boundary}\\
+
+    \inferrule{}{\Gamma \vdash \cdot \isavectx}
+
+    \inferrule{\Gamma \vdash C : \Sort}{\Gamma \vdash \alpha: C \isavectx}
+
+    \inferrule
+    {\Gamma \vdash \Phi \isavectx\and
+    \Gamma \vdash \Psi \isavectx \and
+    \Gamma \pipe d^+\Phi, d^- \Psi \vdash R \isaSet}
+    {\Gamma \vdash \Phi, x:R, \Psi \isavectx}
+
+    (\Phi,x:R,\Psi),y:S,\Sigma = \Phi,(x:R,\Psi,y:S,\Sigma)
+  \end{mathpar}
+  \begin{align*}
+    \underline \cdot &= \cdot\\
+    \underline {\alpha:C} &= \alpha:C\\
+    \underline{\Phi,x:R,\Psi} &= d^-\Phi; d^+{\Psi}
+  \end{align*}
+  \begin{align*}
+    d^{\pm}(\cdot) &\quad\text{undefined}\\
+    d^{\pm}(\alpha:C) &= \alpha : C\\
+    d^{-}(\Phi,x:R,\Psi) &= d^-\Phi\\
+    d^{+}(\Phi,x:R,\Psi) &= d^-\Psi
+  \end{align*}
+  \caption{Basic Rules for Contexts}
+\end{figure}
+
+\begin{figure}
+  \begin{mathpar}
+    \inferrule
+    {}
+    {\Gamma\pipe \cdot \vdash \Sort \isaSet}
+
+    \inferrule
+    {x:S \in \Gamma}
+    {\Gamma \pipe \cdot \vdash x : S}
+
+    \inferrule
+    {\Gamma\pipe\cdot \vdash C : \Sort}
+    {\Gamma \pipe \alpha:\cat C \vdash \alpha : \cat C}
+
+    \inferrule
+    {\Gamma \pipe \alpha:C,\beta:D \vdash R \isaSet}
+    {\Gamma\pipe \alpha:C, x:R, \beta: D \vdash x : R}
+
+    \inferrule{\forall x:A \in \Gamma. \Delta \vdash \gamma(x) : A}{\gamma : \Delta \to \Gamma}\\
+
+    \inferrule*%% [right=NullaryBoundaryMap]
+               {\Gamma \isadtctx}{\Gamma \vdash \cdot : \cdot \to \cdot}\and
+    \inferrule*{\Gamma \pipe \alpha':C' \vdash A : C}{\Gamma \vdash A/\alpha : \alpha':C' \to \alpha:C}
+
+    \inferrule*
+    {\Gamma \pipe \alpha':C' \vdash A : C\and
+      \Gamma \pipe \beta':D' \vdash B : D
+    }
+    {\Gamma \vdash (A/\alpha;B/\beta) : (\alpha':C';\beta':D') \to (\alpha:C;\beta':D')}\\
+
+    \inferrule*%% [right=NullCtxMap]
+               {\Gamma \isadtctx}{\Gamma \vdash \cdot : \cdot \to \cdot}
+
+    \inferrule*{\Gamma \pipe \alpha':C' \vdash A : C}{\Gamma \vdash A/\alpha : \alpha':C' \to \alpha:C}
+
+    \inferrule
+    {\Psi = \Psi_1 \jnctx \Psi_M \jnctx \Psi_2\and
+      \Gamma \vdash \phi_1 : \Psi_1 \to \Phi_1\and
+      \Gamma \pipe \Psi_M \vdash M : R[d^+\phi_1;d^-\phi_2] \and
+      \Gamma \vdash \phi_2 : \Psi_2 \to \Phi_2
+      }
+    {\Gamma \vdash \phi_1,M/x,\phi_2 : \Psi \to \Phi_1,x:R,\Phi_2}
+  \end{mathpar}
+  \caption{Basic Rules of Judgment Theory}
+  \label{fig:basic-rules}
+\end{figure}
+
+\begin{figure}
+  \begin{mathpar}
+    \inferrule
+    {\Gamma \vdash \Phi \isavectx \and
+    \Gamma \vdash \Psi \isavectx \and
+    d^+\Phi = d^- \Psi}
+    {\Gamma \vdash \Phi \jnctx \Psi \isavectx}
+
+    \inferrule
+    {\Gamma \vdash \phi : \Phi' \to \Phi \and
+     \Gamma \vdash \psi : \Psi' \to \Psi\and
+     d^+\phi = d^- \psi}
+    {\Gamma \vdash \phi \jnctx \psi : \Phi' \jnctx \Psi' \to \Phi \jnctx \Psi}
+
+    \inferrule*[right=ObSubst]
+    {\alpha : \cat C \vdash B : \cat D\and
+    \beta : \cat D \vdash A : \cat E}
+    {\alpha : \cat C \vdash A[B/\beta] : \cat E}\\
+
+    \inferrule*[right=Everything Subst]{\gamma : \Delta \to \Gamma \and \Gamma J}{\Delta J}
+
+    \inferrule{\Gamma \vdash \xi : Z \to \Xi\and \Gamma \pipe \Xi \vdash R \isaSet}{\Gamma \pipe Z \vdash R[\xi] \isaSet}
+
+    \inferrule{\Gamma \vdash \phi : \Psi \to \Phi\and \Gamma \pipe \Phi \vdash M : R}{\Gamma \pipe \Psi \vdash M[\phi] : R[\underline \phi]}
+  \end{mathpar}
+  \begin{align*}
+    (\Phi \jnctx \Psi) \jnctx \Sigma &= \Phi \jnctx (\Psi \jnctx \Sigma)\\
+    \Phi \jnctx \alpha:C &= \Phi\\
+    \alpha:C \jnctx \Psi &= \Psi\\\\
+    d^\pm(A/\alpha) &= A/\alpha\\
+    d^+(\phi_1,M/x,\phi_2) &= d^+ \phi_2\\
+    d^-(\phi_1,M/x,\phi_2) &= d^- \phi_1\\\\
+    \underline \cdot &= \cdot\\
+    \underline {A/\alpha} &= A/\alpha\\
+    \underline {\phi_1,M/x,\phi_2} &= d^-\phi_1;d^+\phi_2
+  \end{align*}
+  \caption{Admissible Rules}
+\end{figure}
+
+\begin{figure}
+  \begin{mathpar}
+    \begin{array}{rcl}
+      x[\gamma] &=& \gamma(x)\\
+      J[\gamma][\delta] & = & J[\gamma[\delta]]\\
+      \gamma[\delta](x) &=& \gamma(x)[\delta]\\
+
+      A[\alpha/\alpha] &=& A\\
+      \alpha[A/\alpha] &=& A\\
+      A[B/\beta][C/\gamma] & = & A[B[C/\gamma]/\beta]\\
+      A[B/\beta][\gamma] &=& A[\gamma][B[\gamma]/\beta]\\\\
+
+      R[\cdot] &=& R\\
+      R[\alpha/\alpha] &=& R\\
+      R[\alpha/\alpha,\beta/\beta] &=& R\\\\
+      R[\xi][\xi'] & = & R[\xi[\xi']]\\
+      R[\xi][\gamma] &=& R[\gamma][\xi[\gamma]]\\\\
+
+      \cdot[\xi] &=& \cdot\\
+      (A/\alpha)[B/\beta] &=& A[B/\beta]/\alpha\\
+      (A/\alpha;B/\beta)[\xi] &=& A[d^-\xi]/\alpha,B[d^+\xi]/\beta\\
+      \cdot[\gamma] &=& \cdot\\
+      (A/\alpha)[\gamma] &=& A[\gamma]/\alpha\\
+      (A/\alpha;B/\beta)[\gamma] &=& A[\gamma]/\alpha;B[\gamma]/\beta\\\\
+
+      \cdot[\gamma] &=& \cdot\\
+      (A/\alpha)[\gamma] &=& A[\gamma]/\alpha\\
+      (\phi_1,M/x,\phi_2)[\gamma] &=& \phi_1[\gamma],M[\gamma]/x,\phi_2[\gamma]
+
+      TODO: cat substitution id/assoc
+    \end{array}
+  \end{mathpar}  
+  \caption{Algebraic Properties}
+  \label{fig:equations}
+\end{figure}
+
+Now that we have established a judgmental structure, we move on to the
+multiplicative connectives, starting with the easier negative
+connectives and then the more difficult positives.
+
+\subsection{Left and Right Hom of Profunctors}
+
+First, we study the two ``internal homs'' of judgments. Since the
+``monoidal'' structure of judgments is not symmetric, we get two
+distinct notions internalizing concatenating on the left and the right
+of the context. In ordinary category theory, these profunctors can be
+constructed using an \emph{end} and the hom of functor
+categories. Given two profunctors $P : C^o \times E \to \Set$ and $R :
+D^o \times E \to Set$, their hom is given as
+\[ \int_{e \in E} R(-,e) \Rightarrow P(e,=) \]
+
+Type theoretically this is a kind of multi-hom where we are
+quantifying both over an indexing sort variable $e$ and a profunctor
+variable of type $R$. We write the covariant and contravariant
+versions of these connectives as $\homrallXYtoZ \gamma R S$ and
+$\homlallXYtoZ \alpha R S$.
+
+The syntax is given in Figure~\ref{fig:cotensor}.
+%
+First, the formation rule for covariant cotensor formation requires
+that the covariant variables for $R,P$ are the same, and the
+contravariant variables match the boundary of the overall
+judgment. Note that we use $d^+\Xi,d^-\Xi$ to allow for the case that
+the boundary is a single variable $\Xi = \alpha:\cat C$.
+%
+Next, the substitution rule is straightforward.
+%
+Then we have the elimination form, first written in a simple form with
+no substitution built in.
+%
+This says that given a variable of the cotensor judgment, and another
+variable of judgment $R$, we can apply these to get a term of judgment
+$P$. As a diagram this is written as follows:
+
+\[ % https://tikzcd.yichuanshen.de/#N4Igdg9gJgpgziAXAbVABwnAlgFyxMJZABgBpiBdUkANwEMAbAVxiRAGEQBfU9TXfIRQBGclVqMWbACLdeIDNjwEiAJjHV6zVohABROXyWCiZYeK1TdnHkYEqU685sk793cTCgBzeEVAAZgBOEAC2SGQgOBBIohLabAA6iQEQQYwMAATJ3nShoXQAdJkACtmJOEFYdGDeDDD1ATiZAEqGIMFhsdTRSOrxViBtth0h4YgAzD0xiAAsLgm6Je2d45G9k9QMWGBuUBBMAEb1INQAFjB0UGyQu6dRdFgMNwSsI6t900jzIBdXL3ctjs9gdjqweo9nrpbm8KFwgA
+\begin{tikzcd}
+C \arrow[r, "\forall \gamma. P \triangleleft R"] \arrow[d, no head, Rightarrow] & D \arrow[r, "R"] & E \arrow[d, no head, Rightarrow] \\
+C \arrow[rr, "P"]                                                               &                  & E                               
+\end{tikzcd}\]
+
+The universal property of this judgment is that this is the universal
+2-cell of this type, in that any 2-cell
+
+\[ % https://tikzcd.yichuanshen.de/#N4Igdg9gJgpgziAXAbVABwnAlgFyxMJZABgBpiBdUkANwEMAbAVxiRAGEByEAX1PUy58hFACZyVWoxZsAIr34gM2PASIBmCdXrNWiEAFEFAlcKJkAjJJ0z97Y0sGqRyTVe3S9hh8qFqUFlpSumy8kjBQAObwRKAAZgBOEAC2SIEgOBBI4sG2IABKDokpSJoZWYgALB4h+gAKRUmpiGTlpdQMWGBecBCdUCDUABYwdAOIYEwMDNQ4dFgMbJDdjSWIOZlI1SAjY0sErB1dXlAQTABGDIcZ84v6y6x88U1IrZuIAKxPIMXNH7MVdKdFb6KB0OAjAbDUbjcAHMI8IA
+\begin{tikzcd}
+C' \arrow[d] \arrow[r] & {} \arrow[r, no head, dashed] & D \arrow[r, "R"] & E \arrow[d, no head, Rightarrow] \\
+C \arrow[rrr, "P"]     &                               &                  & E                               
+\end{tikzcd}\]
+
+Exhibits a unique factorization:
+
+\[ % https://tikzcd.yichuanshen.de/#N4Igdg9gJgpgziAXAbVABwnAlgFyxMJZABgBoBGAXVJADcBDAGwFcYkQBhEAX1PU1z5CKAEwVqdJq3YARHnxAZseAkQDM4mgxZtEIAKLz+yoUTIiJ26Xq69jg1Sg0WtU3QaOKBK4SVLFLN3YOAHJPJQdfcn9AnXZw71NRGNc4vTk7LxNHZA0A1OsPbgkYKABzeCJQADMAJwgAWyRokBwIJDFJNJAAJU86xqQNVvbEABYC9wAFfvqmxDIRoZpGLDB3KAhmACNGNhoACxh6KHZIdZAaHHosRjOCNkyB+c62pAmQI5P7i5W1ja2u32rRudz050eCmeSAArFdRgA2J5zJAI+FIADsfwueig9DgR1Oh2Op3BD1mg0QWKWiBaqxxIE2Oz2l0+JJ+wOutw5FPmcJpxGRlMWb1pk3YAB0JdUILUmIwAARSsr0BoNegAOgVUyVEpwtSw9DAZT2e2qOAVfSF82pooAHOK9FaoSjEA6aZ16QDmcCvqTwOSrqCecVuEA
+\begin{tikzcd}
+C' \arrow[r] \arrow[d]                                                           & {} \arrow[r, no head, dashed] & D \arrow[d, no head, Rightarrow] \arrow[r, "R"] & E \arrow[d, no head, Rightarrow] \\
+C \arrow[d, no head, Rightarrow] \arrow[rr, "\forall \gamma. P \triangleleft R"] &                               & D \arrow[r, "R"]                                & E \arrow[d, no head, Rightarrow] \\
+C \arrow[rrr, "P"]                                                               &                               &                                                 & E                               
+\end{tikzcd}\]
+
+We give elimination rules and introduction rules in a style similar to
+$\lambda$-calculus, with corresponding $\beta\eta$ rules.
+
+\begin{figure}
+  \begin{mathpar}
+      \inferrule*[right=Covariant Hom formation]
+      {\Gamma \pipe d^+\Xi; \alpha : \cat C \vdash  R \and \Gamma \pipe  d^-\Xi; \alpha : \cat C \vdash P}
+      {\Gamma \pipe \Xi \vdash \homrallXYtoZ \alpha R P}
+
+      (\homrallXYtoZ \alpha R P)[\xi] = \homrallXYtoZ \alpha {R[d^+\xi;\alpha/\alpha]} {P[\alpha/\alpha;d^-\xi]}
+
+      \inferrule*[right=Covariant Hom introduction]
+      {\Gamma \pipe \Phi, x : R, \alpha : \cat C \vdash M : P}
+      {\Gamma \pipe \Phi \vdash
+        \homrlambdaXatYdotZ {x : R} {\alpha : \cat C} M : \homrallXYtoZ \alpha R P}
+
+     (\homrlambdaXatYdotZ {x} {\alpha} M)[\phi]
+      = \homrlambdaXatYdotZ {x} {\alpha} M[\phi,x/x,\gamma/\gamma]
+
+      %% \inferrule*[right=Covar Hom elim simple]
+      %% {}
+      %% {\alpha: \cat C, f:\forall \gamma. P \triangleleft R, \beta : \cat D, x : R, \gamma : \cat E \vdash f\triangleleft (x, \gamma) : P}
+
+      \inferrule*[right=Covariant Hom elimination]
+      {\Phi = \Phi_f \jnctx \Phi_a\and
+       \Gamma\pipe\Phi_f \vdash M : \homrallXYtoZ \alpha R P \and
+       \Phi_a \vdash N : R[A/\alpha]
+        }
+      {\Gamma\pipe \Phi \vdash \homrappXtoYatZ M N A : P[A/\alpha]}\\
+
+      \homrappXtoYatZ {(\homrlambdaXatYdotZ x \alpha M)} N A = M[N/x,A/\alpha] (\beta)
+      %% (\lambda^\triangleleft (x, \gamma). M)\triangleleft (x, \gamma) = M (\beta)
+
+      \inferrule
+      {\Gamma\pipe\Phi, x:R, \alpha : \cat C \vdash \homrappXtoYatZ M x \alpha = \homrappXtoYatZ N x \alpha : P}
+      {\Gamma\pipe\Phi \vdash M = N : \homrallXYtoZ \alpha R P}
+
+      \inferrule*[right=Contra Hom formation]
+      {\Gamma \pipe \alpha:\cat C; d^-\Xi\vdash R \and \Gamma \pipe \alpha : \cat C; d^+\Xi\vdash P}
+      {\Gamma \pipe \Xi \vdash \homlallXYtoZ \alpha R P}
+
+      \inferrule*[right=Contra Hom introduction]
+      {\Gamma\pipe\alpha:\cat C, x : R, \Phi \vdash M : P}
+      {\Gamma\pipe\Phi \vdash \homllambdaXatYdotZ {x : R} {\alpha : \cat C} M : \homlallXYtoZ \alpha R P}
+
+     (\homllambdaXatYdotZ {x} {\alpha} M)[\phi]
+      = \homllambdaXatYdotZ {x} {\alpha} M[\phi,x/x,\gamma/\gamma]
+
+      \inferrule*[right=Contra Hom elimination]
+      {\Phi = \Phi_a \jnctx \Phi_f\and
+       \Gamma \pipe \Phi_a \vdash N : R[A/\alpha] \and
+       \Gamma \pipe \Phi_f \vdash M : \homlallXYtoZ \alpha R P}
+      {\Gamma \pipe \Phi \vdash \homlappXtoYatZ M N A : P[A/\alpha]}\\
+
+      \inferrule*[Right=$\beta$]
+                 {}
+                 {\homlappXtoYatZ {(\homllambdaXatYdotZ x \alpha M)} N A = M[A/\alpha,N/x]}
+
+      \inferrule*[right=$\eta$]
+      {\Gamma \pipe \alpha:\cat C, x:R, \Phi \vdash \homlappXtoYatZ M x \alpha = \homlappXtoYatZ N x \alpha : P}
+      {\Gamma \pipe \Phi \vdash M = N : \homlallXYtoZ \alpha R P}
+  \end{mathpar}
+  \caption{Hom of Relations/Profunctors}
+  \label{fig:hom}
+\end{figure}
+
+Hom judgments are useful not just in applications (Kan extensions, the
+Yoneda lemma), but also for purely syntactic reasons. In particular, I
+don't see how to give a sensible, complete natural deduction-style
+syntax for the other multiplicative connectives (Unit (i.e., the hom
+profunctor) and Tensor (profunctor composition) without relying on the
+presence of homs. The key property that the hom allows is that given
+any variable in context of a term $\Phi,\alpha,x,\beta,\Psi \vdash M :
+Q$ we can represent this term as a ``single-variable term'' by moving
+$\Phi$ and $\Psi$ to the right of the turnstile:
+
+\[ \alpha, x, \beta \vdash \lambda \Phi. \lambda \Psi M. : \Phi \triangleright Q \triangleleft \Psi \]
+
+This allows us to describe restricted positive eliminations that only
+work on single variable contexts. We give an admissible ``syntax
+sugar'' in Figure~\ref{fig:multi-hom}.
+
+Any $\Phi_l,\beta:C,\Phi_r \vdash M : R$ can be represented as a term
+$\beta:C \vdash M' : \Phi_r \triangleright_{\beta:C} R \triangleleft_{\beta:C} $
+
+Say we have a term in context
+\[ \alpha,x:P,\beta,y:Q,\gamma \vdash M : R \]
+Using $\triangleright\triangleleft$, we can produce an equivalent term
+of a different output type whose context is one of the single
+variables $\alpha,\beta$ or $\gamma$ in the context.
+
+We can make it $\alpha$ by using $\triangleleft$ twice:
+\[ \alpha \vdash \lambda^\triangleleft (x:P,\beta). \lambda^\triangleleft (y:Q, \gamma). M : \forall \beta. (\forall \gamma. R \triangleleft Q) \triangleleft P \]
+
+Or $\gamma$ by using $\triangleright$ twice:
+\[ \gamma \vdash\lambda^\triangleright (\beta, y:Q).  \lambda^\triangleright (\alpha, x:P). M : \forall \gamma. Q \triangleright \forall \alpha. P \triangleright R \]
+
+And we have two different ways to make the context $\beta$, either we use $\triangleright$ and then $\triangleleft$:
+\[ \beta \vdash\lambda^\triangleright (\alpha, x:P).  \lambda^\triangleleft (y:Q, \gamma). M : \forall \alpha. P \triangleright (\forall \gamma. R \triangleleft Q) \]
+or we do the opposite order:
+\[ \beta \vdash \lambda^\triangleleft (y:Q, \gamma). \lambda^\triangleright (\alpha, x:P). M : \forall \gamma. (\forall \alpha. P \triangleright R) \triangleleft Q \]
+
+In general we can iterate this process. That is given any $\Phi,\Psi$
+with $d^+\Phi = d^-\Psi = \beta:D$, we can take a term $\Phi \jnctx
+\Psi \vdash M : R$ and produce an equivalent term $\beta:D \vdash M :
+\Phi \triangleright R \triangleleft \Psi$. To formalize this, we show
+how we can create ``multi-hom'' jumbo connectives $\Phi \triangleright
+R$ and $R \triangleleft \Psi$, where the boundary conditions on
+$\Phi,\Psi$ are analogous to the ordinary case. These are given in
+Figure\ref{multi-hom}.
+
+\begin{figure}
+  \begin{align*}
+    P \triangleleft \gamma:E &= P\\
+    P \triangleleft (\Psi,x:R,\gamma:E) &= (\forall \gamma: E. P \triangleleft R) \triangleleft \Psi\\
+    %% \cdot \triangleright_{\alpha : C} P &= P\\
+    %% (x:R,\beta:D,\Phi_r)\triangleright_{\alpha:C} P  &= \Phi_R \triangleright_{\beta:D} \forall \beta. R \triangleright P
+  \end{align*}
+  \begin{mathpar}
+    \inferrule
+    {\Gamma \pipe d^+\Xi; \gamma: E \vdash P \isaSet\and
+     \Gamma \pipe d^-\Xi; \gamma:E \vdash \Psi \isavectx}
+    {\Gamma \pipe \Xi \vdash P \triangleleft \Psi}
+
+    \inferrule*[right=Covariant Multi-Hom introduction]
+    {\Gamma\pipe\Phi\jnctx \Psi \vdash M : P}
+    {\Gamma\pipe\Phi \vdash \lambda^\triangleleft \Psi. M : P \triangleleft \Psi}\and
+
+    \inferrule*[right=Covar Multi-Hom elim]
+    {~}
+    {\Gamma \pipe f:P\triangleleft \Psi,\Psi \vdash \textrm{app}\triangleleft_\Psi : P}
+
+    %%   (\lambda^\triangleleft \Psi. M)[\phi] = \lambda^\triangleleft \Psi. M[\phi,\Psi]\\
+    %%   \lambda^\triangleleft \gamma. M = M\\
+    %%   \lambda^\triangleleft (\Psi,x:R,\gamma:\cat E). M = \lambda^\triangleleft \Psi. \lambda^\triangleleft (x:R,\gamma:\cat E). M\\
+
+    %%   (\lambda^\triangleleft \gamma. M)[\phi] = M[\phi] = \lambda^\triangleleft \gamma. M[\phi]\\
+    %%   (\lambda^\triangleleft (\Psi,x:r,\gamma:\cat E). M)[\phi] =
+    %%   (\lambda^\triangleleft \Psi. \lambda^\triangleleft (x:R,\gamma:\cat E). M)[\phi]
+    %%   = \lambda^\triangleleft \Psi. (\lambda^\triangleleft (x:R,\gamma:\cat E). M)[\phi, \Psi]
+    %%   = \lambda^\triangleleft \Psi. \lambda^\triangleleft (x:R,\gamma:\cat E). M[\phi,\Psi,x:R,\gamma]
+
+    %%   \inferrule*[right=Covariant Multi-Hom elimination]
+    %%   {\Phi_1 \vdash M : P\triangleleft \Phi_2 \and \phi_2 : \Psi \vdash \Phi_2 \subst(\xi)}
+    %%   {\Phi_1,\Psi \vdash M\triangleleft \phi_2 : P[d^+\xi]}\\
+
+    %%   M \triangleleft (A/\gamma) = M[A/\gamma]\\
+    %%   M \triangleleft (\phi, N/x,A/\gamma) =  M \triangleleft (N, A) \triangleleft \phi
+
+    %%   (M \triangleleft \phi_2)[\phi_1,\psi] = M[\phi_1] \triangleleft [\phi_2[\psi]]\\
+    %%   TODO: prove
+
+    %%   TODO: (\lambda^\triangleleft \Phi. M) \triangleleft \phi = M[\phi]
+    %%   TODO: M = \lambda^\triangleleft \Phi. M\triangleleft \Phi
+
+    %%   TODO: contravariant by duality
+    %%   %% \inferrule*[right=Contra Multi-Hom formation]
+    %%   %% {\Phi \isavectx(\alpha:\cat C; d^-\Xi) \and P \prof(\alpha : \cat C; d^+\Xi)}
+    %%   %% {\Phi \triangleright P \prof(\xi)}
+
+    %%   %% \alpha \triangleright P = P\\
+    %%   %% \alpha:\cat C, x:R, \Phi \triangleright P = \forall \alpha. R \triangleright \Phi \triangleright P\\
+  \end{mathpar}
+  \begin{align*}
+    \lambda^\triangleleft \gamma:E. M &= M\\
+    \lambda^\triangleleft (\Psi,x:R,\gamma:E). M &= \lambda^\triangleleft \Psi. \lambda^\triangleleft (x:R,\gamma:E). M\\\\
+    \textrm{app}\triangleleft_{\gamma:E} &= f\\
+    \textrm{app}\triangleleft_{\Psi,x:R\gamma:E} &= \textrm{app}\triangleleft_\Psi \triangleleft (x,\gamma) \\
+  \end{align*}
+  \caption{Multi-Hom}
+  \label{multi-hom}
+\end{figure}
+
+\subsection{Unit and Tensor Profunctors}
+
+While the cotensors of the previous section work quite smoothly since
+they are negative connectives in an intuitionistic-like (many formulas
+on the left of the sequent and only one on the right), the positive
+connectives, the unit and tensor, are trickier.
+
+First, to internalize $\cdot_{\alpha : \cat C}$ we can define the
+\emph{unit} $A \to_{\cat C} A'$, which in set-theoretic settings we
+might call the \emph{identity/equality judgment} $A \equiv_{\cat C}
+A'$.
+%
+First the formation rule says given two objects $A,A'$ in the same
+category $\cat C$, we can form the hom-set from $A$ to $A'$ $A
+\to_{\cat C} A'$.
+%
+The introduction rule says that given any object $A$, we can define
+the identity arrow from $A$ to itself $\id(A) : A \to_{\cat C} A$.
+%
+Without knowing more about the category in question, these are the
+only arrows we know exist.
+%
+
+The elimination form is trickiest to define in a natural deduction style.
+%
+First, let's consider the universal property of the Hom. From
+Crutwell-Shulman, we want a 2-cell
+
+\[% https://tikzcd.yichuanshen.de/#N4Igdg9gJgpgziAXAbVABwnAlgFyxMJZARgBoAGAXVJADcBDAGwFcYkQBBEAX1PU1z5CKcqQBM1Ok1bsuvftjwEiY8ZIYs2iTjz4gMioUTLF10rSABOMAGaMekmFADm8IqBuWIAWyRkQOBBIqlKa7AASProeXr6IogFBiP6MWGAWUBDMAEaMbDQAFjD0UOyQ6SA0OPRY9trlbPIgnlHxVUkhqRXamTl5lSBFJWUE+QE1deCjDtxAA
+\begin{tikzcd}
+                    & A \arrow[ldd, no head, Rightarrow] \arrow[rdd, no head, Rightarrow] &   \\
+                    & refl                                                                &   \\
+A \arrow[rr, "Hom"] &                                                                     & A
+\end{tikzcd}\]
+
+That is universal/opcartesian in that for any 2-cell $t$:
+
+\[ % https://tikzcd.yichuanshen.de/#N4Igdg9gJgpgziAXAbVABwnAlgFyxMJZABgBpiBdUkANwEMAbAVxiRACEQBfU9TXfIRQBGclVqMWbbrxAZseAkQBMY6vWatEIAIIy+CwUQAsaiZrZ6eBgUpQA2Mxqnb2Acn1z+ioclXLxZy0QAGFPeVtfUwD1SWCAEXDvIxQAZlJhQLi2HA8ucRgoAHN4IlAAMwAnCABbJDIQHAgkYWsQKtqW6iakVRAGLDBgqDo4AAtCkGoJuig2SCHPDrrEPp7EdP7B4LgIAbnpmFmkMCYGBmoGOgAjGAYABWS7fphynCnzFxAACU625aQm3Wpi2i20I3Gk0Ox20C1Y-2qKwa6wArAjOogQet7OiVijus1EI5PsEAErcChcIA
+\begin{tikzcd}
+B \arrow[r] \arrow[rrdd] & {} \arrow[r, no head, dashed] & A \arrow[rr, "Hom"] &    & A \arrow[rr, no head, dashed] &  & B' \arrow[lldd] \\
+                         &                               &                     & t' &                               &  &                 \\
+                         &                               & C \arrow[rr, "R"]   &    & D                             &  &                
+\end{tikzcd}\]
+
+there exists a $t'$ which extends $t$ in that the following is equal
+to $t$:
+\[% https://tikzcd.yichuanshen.de/#N4Igdg9gJgpgziAXAbVABwnAlgFyxMJZABgBoAmAXVJADcBDAGwFcYkQAhEAX1PU1z5CKAIwVqdJq3Y8+IDNjwEi5cTQYs2iEAEFZ-RUKIAWNZM3s9vA4OUoAbGY3TtHAOT75ApcOSrjEs5aIADCngq2vqYB6lLBACLh3kYoAMykqYFx7Dge1l6GdsjpxFkW2lZyET5EZKWx5SBJhb5i9eYuTfnVKciO7UEy3clF6SJlnQBOMABmjDwSMFAA5vBEoDOTEAC2SGQgOBBIIvmbO8c0h0iqIIxYYMFQ9HAAFksgNG-0UOyQD55nXaIG5XRDpW73YJwCB3H6fGDfJBgZiMRg0Rj0ABGMEYAAURsJbrMcB8OsEABLnU5bIHg0GmCH-bRPV7veGI7R-NjU86IfaggCsPKBDNB9mFSAFlyOiEcZPYACUATSkAAOaXXdGQ9hQCDMTGMNjsn6cghGg70LDzU3-CWIdUHGXgu5MkC6-WG0lfE3gM2knCW62+21yQFIACcGr5dsjjuO+xdUJhWDhIG9SJRaItVt+frtIn5MrEjKTsK9CJNyNRl0DuZDGxViALUYdiZ1zzeqfTNu5ocbDtBImLbeZHbZaYrdd7Dd5Q6jDJHbr1BvN3eD5oDOZ7C24QA
+\begin{tikzcd}
+{} \arrow[dd] \arrow[r]  & {} \arrow[dd] \arrow[rr, no head, dashed] &                     & A \arrow[ldd, no head, Rightarrow] \arrow[rdd, no head, Rightarrow] \arrow[rrr, no head, dashed] &                               &  & {} \arrow[dd, no head, Rightarrow] \\
+                         &                                           &                     & refl                                                                                             &                               &  &                                    \\
+B \arrow[r] \arrow[rrdd] & {} \arrow[r, no head, dashed]             & A \arrow[rr, "Hom"] &                                                                                                  & A \arrow[rr, no head, dashed] &  & B' \arrow[lldd]                    \\
+                         &                                           &                     & t'                                                                                               &                               &  &                                    \\
+                         &                                           & C \arrow[rr, "R"]   &                                                                                                  & D                             &  &
+\end{tikzcd}\]
+
+Clearly, our introduction rule models this first 2-cell. Furthermore,
+the construction in the second condition is modeled faithfully by the
+following sequent rule in the figure. This is a J-like induction rule:
+to construct a term using an input of the Hom type, it is sufficient
+to consider the case in which the input is the identity morphism.
+%
+However, building in a substitution the usual way we would (pushing
+the substitution into the continuation $N$) doesn't work properly
+because we can't necessarily extract a substitution for the
+continuation from the substitution for the the whole term. A
+substitution for the whole term would be of the form
+\[\phi,D_1/\delta_1,M/f,D_2/\delta_2,\psi\]
+The the candidate for the continuation's substitution would be
+$\phi,?,\psi$ but there is not necessarily something to go in the
+middle here since $D_1,D_2$ do not necessarily have the same
+\emph{input variable}. However, if we assume the existence already of
+the cotensor products the problem can be simplified. Instead of
+allowing for a general elimination of a hom term anywhere in the
+context, we can instead make a remarkably restricted rule: we can only
+eliminate a Hom term when it is the \emph{sole} input in the context:
+
+\[ \inferrule*[right=Hom Elim Very Simple]{\alpha:\cat C \vdash N : Q[\alpha/\alpha_1,\alpha/\alpha_2]}{\alpha_1:\cat C, f : \alpha_1 \to_{\cat C} \alpha_2, \alpha_2:\cat C \vdash J(\alpha_1.\alpha_2.
+Q, \alpha.
+N, \alpha_1, f, \alpha_2) : Q}\]
+
+While this rule looks very impoverished at first sight, it has the
+virtue of easily supporting a built-in cut:
+
+\[ \inferrule*[right=Hom Elim VSimple Subst]{
+  Q \prof(\alpha_1:\cat C; \alpha_2:\cat C)\and
+  \Phi \vdash M : A_1 \to_{\cat C} A_2\and
+  \alpha:\cat C \vdash N : Q[\alpha/\alpha_1,\alpha/\alpha_2]}{\Phi \vdash J(\alpha_1.\alpha_2. Q, \alpha. N, A_1, M, A_2) : Q[A_1/\alpha_1, A_2/\alpha_2]}\]
+
+However, this rule on its own clearly is not enough to support our
+desired universal property shown above diagrammatically. To prove our
+completeness theorem, we will need to show that this rule is enough to
+give us the desired universal property \emph{assuming the presence of
+  left and right cotensors}.
+
+So first, we must show that the ``proper'' sequent rule
+
+\[ \inferrule*[right=Hom-Left (Sequent Style)]
+    {\Phi_1[\alpha/\alpha_1], \alpha:\cat C, \Phi_2[\alpha/\alpha_2]\vdash Q[\alpha/\alpha_1,\alpha/\alpha_2]}
+    {\Phi_1, \alpha_1 : \cat C, f : \alpha_1 \to_{\cat C} \alpha_2, \alpha_2 : \cat C, \Phi_2\vdash : Q}
+\]
+
+is admissible.
+
+Which we can sketch as follows:
+
+\[
+\inferrule*
+    {\inferrule*{\inferrule*{\Phi_1[\alpha/\alpha_1],\alpha:\cat C, \Phi_2[\alpha/\alpha_1] \vdash Q[\alpha/\alpha_1;\alpha/\alpha_2]}{\alpha:\cat C \vdash \Phi_1[\alpha/\alpha_1] \triangleright Q[\alpha/\alpha_1;\alpha/\alpha_2] \triangleleft \Phi_2[\alpha/\alpha_2]}}{\alpha_1 : \cat C, f : \alpha_1 \to_{\cat C} \alpha_2, \alpha_2 : \cat C\vdash \Phi_1 \triangleright Q \triangleleft \Phi_2}}
+    {\Phi_1, \alpha_1 : \cat C, f : \alpha_1 \to_{\cat C} \alpha_2, \alpha_2 : \cat C, \Phi_2\vdash Q}
+\]
+
+Which using our syntax sugar is given by the term:
+\begin{mathpar}
+  \inferrule*
+    {\inferrule*
+      {\inferrule*
+        {\Phi_1[\alpha/\alpha_1],\alpha:\cat C, \Phi_2[\alpha/\alpha_1] \vdash N : Q[\alpha/\alpha_1;\alpha/\alpha_2]}
+        {\alpha:\cat C \vdash \lambda^\triangleright \Phi_1[\alpha/\alpha_1].
+\lambda^\triangleleft \Phi_2[\alpha/\alpha_2].
+N : \Phi_1[\alpha/\alpha_1] \triangleright Q[\alpha/\alpha_1;\alpha/\alpha_2] \triangleleft \Phi_2[\alpha/\alpha_2]}}
+      {\alpha_1 : \cat C, f : \alpha_1 \to_{\cat C} \alpha_2, \alpha_2 : \cat C\vdash J(\alpha_1.\alpha_2.\Phi_1 \triangleright Q \triangleleft \Phi_2, \alpha.
+\lambda^\triangleright \Phi_1[\alpha/\alpha_1].
+\lambda^\triangleleft \Phi_2[\alpha/\alpha_2].
+N, \alpha_1,f,\alpha_2) : \Phi_1 \triangleright Q \triangleleft \Phi_2}}
+    {\Phi_1, \alpha_1 : \cat C, f : \alpha_1 \to_{\cat C} \alpha_2, \alpha_2 : \cat C, \Phi_2\vdash
+      \Phi_1 \triangleright (J(\alpha_1.\alpha_2.\Phi_1 \triangleright Q \triangleleft \Phi_2, \alpha.
+\lambda^\triangleright \Phi_1[\alpha/\alpha_1].
+\lambda^\triangleleft \Phi_2[\alpha/\alpha_2].
+N, \alpha_1,f,\alpha_2)) \triangleleft \Phi_2 : Q}
+\end{mathpar}
+
+We need to show $\beta$ and $\eta$ hold.
+
+First, $\beta$.
+
+\begin{align*}
+  &(\Phi_1 \triangleright (J_\to(\alpha_1.\alpha_2.\Phi_1 \triangleright Q \triangleleft \Phi_2, \alpha.
+\lambda^\triangleright \Phi_1[\alpha/\alpha_1].
+\lambda^\triangleleft \Phi_2[\alpha/\alpha_2].
+N, \alpha_1,f,\alpha_2)) \triangleleft \Phi_2) [\Phi_1[A/\alpha_1],A/\alpha_1,\id(A),A/\alpha_2,\Phi_2[A/\alpha_2]] \\
+  &=
+  \Phi_1[A/\alpha_1] \triangleright (J_\to(\alpha_1.\alpha_2.\Phi_1 \triangleright Q \triangleleft \Phi_2, \alpha.
+\lambda^\triangleright \Phi_1[\alpha/\alpha_1].
+\lambda^\triangleleft \Phi_2[\alpha/\alpha_2].
+N, A,\id(A), A)) \triangleleft \Phi_2[A/\alpha_2]\\
+  &=
+  \Phi_1[A/\alpha_1] \triangleright (\lambda^\triangleright \Phi_1[A/\alpha_1]. \lambda^\triangleleft \Phi_2[A/\alpha_2]. N[A/\alpha]) \triangleleft \Phi_2[A/\alpha_2]\\
+  &=  (\lambda^\triangleleft \Phi_2[A/\alpha_2]. N[A/\alpha]) \triangleleft \Phi_2[A/\alpha_2]\\
+  &=  N[A/\alpha]
+\end{align*}
+
+\begin{figure}
+  \begin{mathpar}
+    \inferrule*[right=Unit-Formation]
+    {\Gamma \vdash C : \Sort
+     \Gamma \pipe d^-\Xi \vdash A : \cat C\and \Gamma \pipe d^+ \Xi \vdash A' : \cat C}
+    {\Gamma \pipe \Xi \vdash \punitinXfromYtoZ {C} A {A'}}
+
+    %% (A \to A')[\xi] = A[d^-\xi] \to A'[d^+\xi]
+
+    \inferrule*[right=Unit-Introduction]
+    {\Gamma\pipe\cdot\vdash C : \Sort}
+    {\Gamma \pipe\cdot \vdash \punitrefl C : \pendallXdotY {\alpha:C} \punitinXfromYtoZ C \alpha {\alpha}}
+
+    \inferrule*[right=Unit Elim]{
+      \Gamma\pipe\cdot \vdash Q : \harr C C\and
+      \Gamma\pipe\cdot\vdash N : \pendallXdotY {\alpha:C} {\harrapp Q \alpha \alpha}}
+    {\Gamma\pipe\cdot \vdash \punitelimtoYwithkontZ Q N :
+      \pendallXdotY {\alpha} {\homrallXYtoZ {\alpha'} {\punitinXfromYtoZ C {\alpha}{\alpha'}} {\harrapp Q {\alpha}{\alpha'}}}
+      }
+
+    %% J_\to(\alpha_1.\alpha_2. Q, \alpha. N, A_1, M, A_2)[\phi] = J_\to(\alpha_1.\alpha_2. Q, \alpha. N, A_1[d^-\phi], M[\phi], A_2[d^+\phi])
+
+    %% \inferrule*[right=Unit$\beta$]{\beta:\cat D \vdash A : \cat C\and
+    %%   \alpha:\cat C \vdash N : Q[\alpha/\alpha_1,\alpha/\alpha_2]
+    %% }{\beta : \cat D \vdash J_\to(\alpha_1.\alpha_2. Q, \alpha. N, A, \id(A), A) = N[A/\alpha] : Q[A/\alpha_1,A/\alpha_2]}
+    %%
+
+    \homrappXtoYatZ {\pendappXtoY {\punitelimtoYwithkontZ Q N} A} {\pendappXtoY {\punitrefl C} A} A = \pendappXtoY N A
+
+    \inferrule*[right=Unit$\eta$]
+    {\alpha_1 : \cat C, f : \alpha_1 \to_{\cat C} \alpha_2, \alpha_2:\cat C \vdash M : \harrapp Q {\alpha_1} {\alpha_2}}
+    {\alpha_1 : \cat C, f : \alpha_1 \to_{\cat C} \alpha_2, \alpha_2:\cat C \vdash M =
+      \homrappXtoYatZ {\pendappXtoY {\punitelimtoYwithkontZ Q {\pendlambdaXdotY \alpha M[\alpha/\alpha_1,\pendappXtoY {\punitrefl C} \alpha/f,\alpha/\alpha_2]}} {\alpha_1}} f {\alpha_2}
+    : \harrapp Q {\alpha_1}{\alpha_2}}
+  \end{mathpar}  
+  \caption{Unit Primitive Rules}
+\end{figure}
+
+\begin{figure}
+  \begin{mathpar}
+    \inferrule*[right=Full Unit Elim*]
+    {\Phi_1[\alpha/\alpha_1], \alpha:\cat C, \Phi_2[\alpha/\alpha_2]\vdash N : Q[\alpha/\alpha_1,\alpha/\alpha_2]}
+    {\Phi_1, \alpha_1 : \cat C, f : \alpha_1 \to_{\cat C} \alpha_2, \alpha_2 : \cat C, \Phi_2\vdash J_\to'(\Phi_1, \alpha_1 : \cat C, f : \alpha_1 \to_{\cat C} \alpha_2, \alpha_2 : \cat C, \Phi_2, Q, \alpha.
+N) : Q}
+
+    J_\to'(\Phi_1, \alpha_1 : \cat C, f : \alpha_1 \to_{\cat C} \alpha_2, \alpha_2 : \cat C, \Phi_2, Q, \alpha. N) =
+    \Phi_1 \triangleright (J(\alpha_1.\alpha_2.\Phi_1 \triangleright Q \triangleleft \Phi_2, \alpha.
+\lambda^\triangleright \Phi_1[\alpha/\alpha_1].
+\lambda^\triangleleft \Phi_2[\alpha/\alpha_2].
+N, \alpha_1,f,\alpha_2)) \triangleleft \Phi_2
+
+    \inferrule*[right=Full Unit $\beta$*]
+    {}
+    {J_\to'(\Phi_1, \alpha_1 : \cat C, f : \alpha_1 \to_{\cat C} \alpha_2, \alpha_2 : \cat C, \Phi_2, Q, \alpha.
+N)[\Phi_1[\alpha/\alpha_1],\alpha/\alpha_1, \id(\alpha)/f, \alpha/\alpha_2, \Phi_2[\alpha/\alpha_2]]
+    = N}
+
+    (\Phi_1 \triangleright J(\alpha_1.\alpha_2.\Phi_1 \triangleright Q \triangleleft \Phi_2, \alpha.
+\lambda^\triangleright \Phi_1[\alpha/\alpha_1].
+\lambda^\triangleleft \Phi_2[\alpha/\alpha_2].
+N, \alpha_1,f,\alpha_2) \triangleleft \Phi_2)[\Phi_1[\alpha/\alpha_1],\alpha/\alpha_1, \id(\alpha)/f, \alpha/\alpha_2, \Phi_2[\alpha/\alpha_2]]\\
+    = \Phi_1[\alpha/\alpha_1] \triangleright J(\alpha_1.\alpha_2.\Phi_1 \triangleright Q \triangleleft \Phi_2, \alpha.
+\lambda^\triangleright \Phi_1[\alpha/\alpha_1].
+\lambda^\triangleleft \Phi_2[\alpha/\alpha_2].
+N, \alpha,\id(\alpha),\alpha) \triangleleft \Phi_2[\alpha/\alpha_2]\\
+    = \Phi_1[\alpha/\alpha_1] \triangleright
+    (\lambda^\triangleright \Phi_1[\alpha/\alpha_1]. \lambda^\triangleleft \Phi_2[\alpha/\alpha_2]. N)
+    \triangleleft \Phi_2[\alpha/\alpha_2]\\
+    = N
+
+    \inferrule*[right=Unit$\eta$]
+    {\Phi_1, \alpha_1 : \cat C, f : \alpha_1 \to_{\cat C} \alpha_2, \alpha_2 : \cat C, \Phi_2 \vdash M : Q}
+    {M =
+      J_\to'(\Phi_1, \alpha_1 : \cat C, f : \alpha_1 \to_{\cat C} \alpha_2, \alpha_2 : \cat C, \Phi_2, Q, \alpha.
+M[\Phi_1[\alpha/\alpha_1],\alpha/\alpha_1,\id(\alpha), \alpha/\alpha_2,\Phi_2[\alpha/\alpha_2]])}
+
+    M
+    =(\triangleleft\beta\triangleright\beta)
+    \Phi_1 \triangleright
+    (\lambda^\triangleright \Phi_1. \lambda^\triangleleft \Phi_2. M)
+    \triangleleft \Phi_2
+    = \Phi_1 \triangleright
+    J_\to(\alpha_1.\alpha_2.Q, \alpha.
+\lambda^\triangleright \Phi_1[\alpha/\alpha_1].
+\lambda^\triangleleft \Phi_2[\alpha/\alpha_2].
+M[\alpha/\alpha_1, \id(\alpha)/f,\alpha/\alpha_2], \alpha_1, f, \alpha_2)
+    \triangleleft \Phi_2
+    =
+    J_\to'(\Phi_1, \alpha_1 : \cat C, f : \alpha_1 \to_{\cat C} \alpha_2, \alpha_2 : \cat C, \Phi_2, Q, \alpha. N)
+  \end{mathpar}  
+  \caption{Admissible Unit Universal Property Rules}
+\end{figure}
+
+Next, we can internalize the $,\alpha,$ as a positive connective as a
+kind of ``existentially quantified tensor product'', abstracting the
+composition of profunctors.
+
+The ordinary formula for this is that for profunctors $R : C^o \times
+D \to \Set$ and $S : D^o \times E \to \Set$, the composite is given by
+a coend:
+
+\[ \int^{d \in D} R(-,d) \otimes S(d,=) \]
+
+We will instead use slightly more logical notation, writing the coend as
+an existential quantifier, justified by the analogy between the two
+and the structure of the introduction/elimination forms in our syntax
+in particular.
+\begin{figure}
+\begin{mathpar}
+  \inferrule*[right=Tensor formation]
+  {\Gamma \pipe d^-\Xi; \beta:D\vdash P \isaSet\and
+    \Gamma \pipe \beta:D;d^+\Xi \vdash Q}
+  {\Gamma \pipe \Xi \vdash \tensorexistsXwithYandZ {\beta:D} P Q}
+
+  %% \inferrule*[right=Tensor intro function]
+  %% {}
+  %% {\alpha : \cat C, q : Q, \beta : \cat D, r : R, \gamma : \cat E \vdash (q, \beta,r) : \exists \beta : \cat D. Q \odot R}
+
+  \inferrule*[right=Tensor introduction]
+  {\Gamma \vdash P : \harr C D \and
+    \Gamma \vdash Q : \harr D E}
+  {\Gamma\pipe\cdot\vdash \tensorintroofXandY P Q : \pendallXdotY {\alpha} \homrallXYtoZ \beta {\harrapp P \alpha \beta} {\homrallXYtoZ \gamma {\harrapp Q \beta \gamma} {(\tensorexistsXwithYandZ \beta {\harrapp P \alpha \beta}{\harrapp Q \beta \gamma})}}}
+
+  %% \inferrule*[right=Tensor introduction]
+  %% {\Phi = \Psi_M \jnctx \Psi_N \and
+  %%  \Gamma \pipe d^+\Psi_M \vdash B : D\and
+  %%  \Gamma \pipe \Psi_M \vdash M : P[;B/\beta] \and
+  %%  \Gamma \pipe \Psi_N \vdash N : Q[B/\beta;]}
+  %% {\Gamma\pipe\Phi\vdash \tensorintroatXwithYandZ B M N : \tensorexistsXwithYandZ {\beta : D} P Q}
+
+  %% \tensorintroatXwithYandZ B M N [\phi,B'/\beta',\psi] = \tensorintroatXwithYandZ {B[B'\beta]} {M[\phi]} {N[\psi]}
+
+  %% \inferrule*[right=Tensor simpl]
+  %% {\Gamma\pipe\cdot \vdash R : \harr {C} {E}\and
+  %%  \Gamma\pipe\cdot \vdash N : \pendallXdotY {\alpha:C} {\homrallXYtoZ {\beta:D} P {\homrallXYtoZ {\gamma:E} Q {\harrapp R \alpha\gamma}}}}
+  %% {\Gamma\pipe\alpha:C,p:\tensorexistsXwithYandZ {\beta:D} P Q,\gamma:E \vdash\tensorelimVatWtoXtoYwithkontZ p \alpha \gamma R N : \harrapp R A G}
+
+  \inferrule*[right=Tensor-elim]
+  {\Gamma \pipe \cdot \vdash P : \harr C D\and\Gamma \pipe \cdot \vdash Q : \harr D E\and
+   \Gamma \pipe \cdot \vdash R : \harr C E\and
+   \Gamma \pipe \cdot \vdash N : \pendallXdotY \alpha {\homrallXYtoZ \beta {\harrapp P \alpha \beta} {\homrallXYtoZ \gamma {\harrapp Q \beta \gamma}}}   
+  }
+  {\Gamma\pipe\cdot \vdash \tensorelimfromWXtoYwithkontZ P Q R N : \pendallXdotY \alpha {\homrallXYtoZ \gamma {\tensorexistsXwithYandZ \beta {\harrapp P \alpha \beta}{\harrapp Q \beta \gamma}} {\harrapp R \alpha \gamma}}}
+
+  %% \inferrule*[right=Tensor elim]
+  %% {\Gamma\pipe\cdot \vdash R : \harr {C} {E}\and
+  %%  \Gamma\pipe\Phi \vdash M : \tensorexistsXwithYandZ {\beta:D} {P[A/\alpha;]} {Q[;G/\gamma]}\and
+  %%  \Gamma\pipe\cdot \vdash N : \pendallXdotY {\alpha:C} {\homrallXYtoZ {\beta:D} P {\homrallXYtoZ {\gamma:E} Q {\harrapp R \alpha\gamma}}}}
+  %% {\Gamma\pipe\Phi \vdash\tensorelimVatWtoXtoYwithkontZ M A G R N : \harrapp R A G}
+
+  %% J_\odot(\alpha. \gamma. S, \alpha.q.\beta.r.\gamma.N, A, M, G)[\phi] =
+  %% J_\odot(\alpha. \gamma. S, \alpha.q.\beta.r.\gamma.N, A[d^-\phi], M[\phi], G[d^+\phi])
+
+  %% \inferrule*[right=Tensor elim Sequent]
+  %% {\Phi,\alpha : \cat C, q : Q, \beta : \cat D, r: R, \gamma : \cat E,\Psi \vdash N : S}
+  %% {\Phi, \alpha : \cat C, x : \exists \beta. Q \odot R, \gamma : \cat E,\Psi \vdash J_\odot(x;q.\beta.r. N) : S}
+
+  %% \inferrule*[right=Tensor elim Subst]
+  %% {\Phi,\alpha : \cat C, q : Q, \beta : \cat D, r: R, \gamma : \cat E,\Phi' \vdash N : S
+  %% \and
+  %% \phi : \Psi \to \Phi,\alpha : \cat C, q : Q, \beta : \cat D, r: R, \gamma : \cat E,\Phi'}
+  %% {\Psi  \vdash J_\odot(x;q.\beta.r. N)[\phi] : S}
+
+  \inferrule*[right=Tensor $\beta$]
+  {}
+  {\homrappXtoYatZ {\pendappXtoY{\tensorelimfromWXtoYwithkontZ P Q R N} A} {(\homrappXtoYatZ{\homrappXtoYatZ{\pendappXtoY {\tensorintroofXandY P Q} A} f B} g G)} G =
+    \homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY N A} f B} g G}
+
+  %% \inferrule*[right=Tensor $\beta$]
+  %% {}{J_\odot(\alpha.\gamma.S, \alpha.q.\beta.r.\gamma.N,A,(M,B,N),G) = N[A/\alpha,M/q,B/\beta,N/r,G/\gamma]}
+
+  \inferrule*[right=Tensor$\eta$]
+  {\alpha : \cat C, x : \tensorexistsXwithYandZ \beta P Q, \gamma : \cat E \vdash M : R}
+  {\alpha : \cat C, x : \exists \beta. Q \odot R, \gamma : \cat E \vdash M =
+    \homrappXtoYatZ {\pendappXtoY{\tensorelimfromWXtoYwithkontZ P Q R
+        {\pendlambdaXdotY \alpha {\homrlambdaXatYdotZ p \beta {\homrlambdaXatYdotZ q \gamma {M[\homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {\tensorintroofXandY P Q} \alpha} p \beta} q \gamma/x]}}}}
+      } \alpha} x \gamma : S}
+\end{mathpar}
+\caption{Tensor Primitive Rules}
+\label{fig:tensor}
+\end{figure}
+First, the syntax for tensor is based on the coend formula that gives
+the definition in $Cat$, but using logical syntax $\exists$ rather
+than integrals because I don't like integral syntax and the syntax
+matches existential quantification in type theory.
+%
+Note that $\exists \beta. ~Q \odot R$ is the application of a single
+connective to $Q$ and $R$, and not separable into a quantifier
+$\exists$ connective and a product connective $\odot$.
+%
+Introduction, elimination and $\beta\eta$ are similar to a combination
+of existential quantification and the tensor product in linear logic.
+%
+The introduction rule is justified by the fact that semantically what
+we have is a transformation
+\[ \textrm{pair} : (\alpha : \cat C,Q,\beta:\cat D,R,\gamma:\cat E) \Rightarrow \exists \beta. R \odot Q \]
+And the introduction rule is just building in a substitution to this
+transformation.
+%
+
+For the elim rule we can pull a similar trick as with the Hom
+type. First, the rule where there's only one variable.
+\[ \inferrule*[right=Tensor elim Simple]
+  {\alpha : \cat C, q : Q, \beta : \cat D, r: R, \gamma : \cat E \vdash N : S}
+  {\alpha : \cat C, x : \exists \beta. Q \odot R, \gamma : \cat E \vdash J_\odot(x, q.\beta.r. N) : S}
+\]
+
+\begin{figure}
+\begin{mathpar}
+    \inferrule*[right=Full Tensor Elim*]
+    {\Phi_1, \alpha:\cat C, q:Q, \beta:\cat D, r:R, \gamma:\cat E \Phi_2\vdash N : S}
+    {\Phi_1, \alpha : \cat C, x : \exists \beta:\cat D. Q \odot R, \gamma : \cat E, \Phi_2
+      \vdash J_\odot'(\Phi_1, \alpha : \cat C, x : \exists \beta:\cat D. Q \odot R, \gamma : \cat E, \Phi_2; S;q.\beta.r.N) : S}
+
+    J_\odot'(\Phi_1, \alpha : \cat C, x : \exists \beta:\cat D. Q \odot R, \gamma : \cat E, \Phi_2;S; q.\beta.r.N)
+    = \Phi_1 \triangleright J_\odot(\alpha.\gamma.\Phi_1 \triangleright S \triangleleft \Phi_2, \alpha.q.\beta.r.\gamma.
+\lambda^\triangleleft \Phi_1.
+\lambda^\triangleright\Phi_2.
+N, \alpha, x, \gamma) \triangleleft \Phi_2
+
+    J_\odot'(\Phi_1, \alpha : \cat C, x : \exists \beta:\cat D. Q \odot R, \gamma : \cat E, \Phi_2; q.\beta.r.N)[\Phi_1,\alpha/\alpha,(q,\beta,r)/x,\gamma/\gamma,\Phi_2] = N
+
+  \inferrule*[right=Full Tensor elim*]{ S \prof(\alpha:\cat C; \gamma : \cat E)\and
+    \Phi \vdash M : \exists \beta. Q[A/\alpha] \odot R[G/\gamma]\and
+    \alpha : \cat C, q : Q, \beta : \cat D, r: R, \gamma : \cat E \vdash N : S}
+  {\Phi \vdash J_\odot(\alpha. \gamma. S, \alpha.q.\beta.r.\gamma.N, A, M, G) : S[A/\alpha, G/\gamma]}
+
+  J_\odot(\alpha. \gamma. S, \alpha.q.\beta.r.\gamma.N, A, M, G)[\phi] =
+  J_\odot(\alpha. \gamma. S, \alpha.q.\beta.r.\gamma.N, A[d^-\phi], M[\phi], G[d^+\phi])
+
+  %% \inferrule*[right=Tensor elim Sequent]
+  %% {\Phi,\alpha : \cat C, q : Q, \beta : \cat D, r: R, \gamma : \cat E,\Psi \vdash N : S}
+  %% {\Phi, \alpha : \cat C, x : \exists \beta. Q \odot R, \gamma : \cat E,\Psi \vdash J_\odot(x;q.\beta.r. N) : S}
+
+  %% \inferrule*[right=Tensor elim Subst]
+  %% {\Phi,\alpha : \cat C, q : Q, \beta : \cat D, r: R, \gamma : \cat E,\Phi' \vdash N : S
+  %% \and
+  %% \phi : \Psi \to \Phi,\alpha : \cat C, q : Q, \beta : \cat D, r: R, \gamma : \cat E,\Phi'}
+  %% {\Psi  \vdash J_\odot(x;q.\beta.r. N)[\phi] : S}
+
+  \inferrule*[right=Tensor $\beta$]
+  {}{J_\odot(\alpha.\gamma.S, \alpha.q.\beta.r.\gamma.N,A,(M,B,N),G) = N[A/\alpha,M/q,B/\beta,N/r,G/\gamma]}
+
+  \inferrule*[right=Tensor$\eta$]
+  {\alpha : \cat C, x : \exists \beta. Q \odot R, \gamma : \cat E \vdash M : S}
+  {\alpha : \cat C, x : \exists \beta. Q \odot R, \gamma : \cat E \vdash M =
+    J(\alpha.\gamma.S, \alpha.q.\beta.r.\gamma. M[\alpha/\alpha,(q,\beta,r)/x,\gamma/\gamma])
+    : S}
+\end{mathpar}
+\caption{Tensor Admissible Universal Property Rules}
+\label{fig:tensor}
+\end{figure}
+
+\subsection{Dependent Types}
+
+Next, we cover the basic dependent type theory structure: $\Pi,\Sigma$
+and identity types.
+
+\begin{figure}
+  \begin{mathpar}
+  \inferrule
+  {\Gamma \pipe\cdot \vdash S \isaSet\and \Gamma \pipe\cdot \vdash T \isaSet}
+  {\Gamma\pipe\cdot \vdash \prod_{x:S} T \isaSet}
+
+  \inferrule
+  {\Gamma \pipe \cdot \vdash M : \prod_{x:S} T\and
+    \Gamma \pipe \cdot \vdash N : S}
+  {\Gamma \pipe \cdot \vdash M N : T[N/x]}
+
+  \inferrule
+  {\Gamma,x:S \pipe \cdot \vdash M : T}
+  {\Gamma \pipe \cdot \vdash \lambda x:S. M : \prod_{x:S} T}
+
+  \inferrule
+  {\Gamma \pipe\cdot \vdash S \isaSet\and \Gamma \pipe\cdot \vdash T \isaSet}
+  {\Gamma\pipe\cdot \vdash \sum_{x:S} T \isaSet}
+
+  \inferrule
+  {~}
+  {\Gamma\pipe\cdot \vdash \top \isaSet}
+
+  \inferrule
+  {\Gamma \pipe \cdot \vdash M : S\and \Gamma \pipe \cdot \vdash N : S}
+  {\Gamma \pipe \cdot \vdash \textrm{Id}_S M N \isaSet}
+
+  \inferrule
+  {\Gamma \pipe \cdot \vdash M : S}
+  {\Gamma \pipe \cdot \vdash \textrm{refl}(M) : \textrm{Id}_S M M}
+
+  \inferrule
+  {\Gamma,x:S,y:S,p:\textrm{Id}_S x y \pipe \cdot \vdash C \isaSet\and
+   \Gamma \pipe \cdot \vdash M_k : \prod_{x:S} C[x/y,\textrm{refl}(x)/p]}
+  {\Gamma \pipe \cdot \vdash \textrm{J}(x,y,p.C,M) : \prod_{x:S,y:S,p:\textrm{Id}_S x y} C}
+  \end{mathpar}
+  \caption{Ordinary Dependent Type Structure}
+\end{figure}
+
+\subsection{Internalizing Virtual Equipment Structure}
+
+\begin{figure}
+  \begin{mathpar}
+    \inferrule
+    {\Gamma \vdash \cat C : \Sort \and\Gamma \vdash \cat D : \Sort}
+    {\Gamma\pipe\cdot \vdash \varr{\cat C}{\cat D} \isaSet}
+
+    \inferrule
+    {\Gamma \pipe \alpha:\cat C \vdash A : \cat D}
+    {\Gamma \vdash \lambda \alpha:\cat C. A : \varr{\cat C}{\cat D}}
+
+    \inferrule
+    {\Gamma \pipe \alpha:\cat C \vdash A : \cat D\and
+      \Gamma\pipe\cdot \vdash M : \varr{\cat D}{\cat E}}
+    {\Gamma\pipe \alpha:\cat C \vdash M A : \cat E}
+
+    (\lambda\alpha:C. A) B = A[B/\alpha]
+
+    \inferrule*[right=VArr $\eta$]
+    {\Gamma \pipe\cdot \vdash M : \varr C D}
+    {\Gamma \pipe\cdot \vdash M = (\lambda \alpha:C. M \alpha) : \varr C D}
+    \\
+
+    \inferrule
+    {\Gamma \vdash \cat C : \Sort \and\Gamma \vdash \cat D : \Sort}
+    {\Gamma\pipe\cdot \vdash \harr{\cat C}{\cat D} \isaSet}
+
+    \inferrule
+    {\Gamma \pipe \alpha:\cat C;\beta:\cat D \vdash R \,\isaSet}
+    {\Gamma \vdash \lambda \alpha:\cat C;\beta:\cat D. R : \harr{\cat C}{\cat D}}
+
+    \inferrule
+    {\Gamma \vdash M : \harr{\cat C}{\cat D}\and
+     \Gamma \pipe d^- \Xi \vdash A : \cat C\and
+     \Gamma \pipe d^+ \Xi \vdash B : \cat C}
+    {\Gamma \pipe \Xi \vdash M A; B \isaSet}
+
+    (\lambda \alpha:C;\beta:D. R)A;B = R[A/\alpha;B/\beta]
+
+    \inferrule*[right=HArr $\eta$]
+    {\Gamma \pipe \cdot \vdash M : \harr C D}
+    {\Gamma \pipe \cdot \vdash M = (\lambda \alpha;\beta. M \alpha;\beta) : \harr C D}\\
+
+    \inferrule
+    {\Gamma \pipe \alpha:\cat C\vdash R \isaSet}
+    {\Gamma \pipe \cdot \vdash \pendallXdotY {\alpha:{\cat C}} R \isaSet}
+
+    \inferrule
+    {\Gamma \pipe \alpha:\cat C\vdash M : R}
+    {\Gamma \pipe \cdot \vdash \pendlambdaXdotY \alpha M : \pendallXdotY \alpha R}
+
+    \inferrule
+    {\Gamma\pipe\cdot \pipe \cdot \vdash M : \forall \alpha:\cat C. R\and
+      \Gamma \pipe \beta:\cat D \vdash A : \cat C}
+    {\Gamma \pipe \beta:\cat D\pipe\beta:\cat D \vdash \pendappXtoY M A : R[A/\alpha]}
+
+    \pendappXtoY {(\pendlambdaXdotY \alpha M)} A = M[A/\alpha]
+
+    \inferrule*[right=2Cell$\eta$]
+    {\Gamma \pipe \cdot \vdash M : \pendallXdotY \alpha R }
+    {\Gamma \pipe \cdot \vdash M = \pendlambdaXdotY \alpha {\pendappXtoY M \alpha} : \pendallXdotY \alpha R }
+  \end{mathpar}
+  \caption{Internalizing VE Structure}
+\end{figure}
+
+\subsection{Equalizers}
+
+To encode the concept of internal hom of profunctors, kan extensions
+and (co-)limits, we need to define the \emph{equalizer} of horizontal
+arrows. Specifically, the universal property that we need is an
+equalizer in the category whose objects are horizontal arrows (for
+fixed endpoints) and whose arrows are the 2-cells.
+
+\begin{figure}
+  \begin{mathpar}
+    \inferrule*[right=Equalizer-formation]
+    {\Gamma \pipe \cdot \vdash P : \harr\, C\,D\and
+     \Gamma \pipe \cdot \vdash Q : \harr\, C \, D\\\\
+     \Gamma \pipe \cdot \vdash \phi : \homunary \alpha {\harrapp P \alpha \beta} \beta {\harrapp Q \alpha \beta}\and
+     \Gamma \pipe \cdot \vdash \psi : \homunary \alpha {\harrapp P \alpha \beta} \beta {\harrapp Q \alpha \beta}\and}
+    {\Gamma \pipe \cdot \vdash \textrm{Equalizer}(C,D,P,\phi,\psi) : \harr\, C\, D}\and
+
+    \inferrule
+    {\Gamma \pipe \Phi \vdash M : \harrapp{\textrm{Equalizer}(C,D,P,\phi,\psi)} A B}
+    {\Gamma \pipe \Phi \vdash \textrm{forget}(M) : \harrapp P A B}
+
+    \inferrule
+    {\Gamma \pipe \Phi \vdash M : \harrapp P A B \and \appunary \phi A M B = \appunary \psi A M B}
+    {\Gamma \pipe \Phi \vdash \textrm{equalizing}_{\phi,\psi}(M) : \harrapp{\textrm{Equalizer}(C,D,P,\phi,\psi)} A B}
+
+    TODO: obvious equations
+  \end{mathpar}
+  \caption{Equalizers (Logical Framework style)}
+\end{figure}
+
+\begin{figure}
+  \begin{mathpar}
+    \inferrule*[right=Equalizer-formation]
+    {\Gamma \pipe \Xi \vdash P \isaSet\and\Gamma \pipe \Xi \vdash Q \isaSet\and
+      \Gamma\pipe d^-\Xi, p : P, d^+ \Xi \vdash \phi : Q\and
+      \Gamma\pipe d^-\Xi, p : P, d^+ \Xi \vdash \psi : Q }
+    {\Gamma \pipe \Xi \vdash \equalizeVofWbyXeqYatZ p P \phi \psi Q : \harr\, C\, D}\and
+
+    \inferrule*[right=Equalizer-formation]
+    {\Gamma \pipe \alpha:C;\beta:D \vdash P \isaSet\and\Gamma \pipe \alpha:C;\beta:D \vdash Q \isaSet\and
+      \Gamma \pipe d^-\Xi \vdash A : C\and
+      \Gamma \pipe d^+\Xi \vdash B : D\\\\
+      \Gamma\pipe \cdot \vdash \phi : \homunary \alpha P \beta Q\and
+      \Gamma\pipe \cdot \vdash \psi : \homunary \alpha P \beta Q
+    }
+    {\Gamma \pipe \Xi \vdash \equalizeVofWbyXeqYatZ p {P[A/\alpha;B/\beta]} {\appunary \phi A p B} {\appunary \psi A p B} {Q[A/\alpha;B/\beta]} : \harr\, C\, D}\and
+
+    \inferrule
+    {\Gamma \pipe \Phi \vdash M : \equalizeVofWbyXeqYatZ p P \phi \psi Q}
+    {\Gamma \pipe \Phi \vdash \pi_{eq}\,M : P}
+
+    \inferrule
+    {\Gamma \pipe \Phi \vdash M : \harrapp P A B \and \appunary \phi A M B = \appunary \psi A M B}
+    {\Gamma \pipe \Phi \vdash \textrm{equalizing}_{\phi,\psi}(M) : \harrapp{\textrm{Equalizer}(C,D,P,\phi,\psi)} A B}
+
+    TODO: equations
+  \end{mathpar}
+  \caption{Equalizers (Natural Deduction Style)}
+\end{figure}
+
+\subsection{Products}
+
+We add in parallel two different notions of product: products of sorts
+and products of spans, which have the universal property in their
+respective categories.
+
+\begin{figure}
+  \begin{mathpar}
+    \inferrule{}{\Gamma \vdash 1 : \Sort}\and
+    \inferrule*{}{\Gamma\pipe\alpha:C \vdash () : 1}\and
+    \inferrule*{\Gamma\pipe\alpha:C \vdash A : 1}{A = ()}
+    \inferrule{}{\Gamma\pipe \Xi \vdash 1 \isaSet}\and
+    \inferrule{}{\Gamma\pipe \Phi \vdash () :  1}\and
+    \inferrule{\Gamma\pipe \Phi \vdash M :  1}{M = ()}\\\\
+
+    \inferrule{\Gamma \vdash C : \Sort \and \Gamma \vdash D : \Sort}
+              {\Gamma \vdash C \times D : \Sort}\and
+
+    \inferrule{\forall i \in \{1,2\}.~ \Gamma \pipe \alpha:C \vdash A_i : D_i}{\Gamma \pipe \alpha:C \vdash (A_1,A_2) : D_1 \times D_2}
+
+    \inferrule{\Gamma \pipe \alpha:C \vdash A : D_1 \times D_2}{\Gamma \pipe \alpha:C \vdash \pi_i A : D_i}
+
+    \pi_i(A_1,A_2) = A_i\\
+    \inferrule{\Gamma\pipe \alpha:C \vdash A :  D_1\times D_2}{A = (\pi_1 A,\pi_2 A)}\\\\
+
+    \inferrule{\Gamma\pipe\Xi \vdash R \isaSet \and \Gamma\pipe\Xi \vdash S\isaSet}
+              {\Gamma\pipe\Xi \vdash R \times S\isaSet}\and
+
+    \inferrule{\forall i \in \{1,2\}.~ \Gamma \pipe\Phi \vdash M_i : R_i}{\Gamma \pipe\Phi \vdash (M_1,M_2) : R_1 \times R_2}
+
+    \inferrule{\Gamma \pipe\Phi\vdash M : R_1 \times R_2}{\Gamma \pipe\Phi \vdash \pi_i M : R_i}
+
+    \pi_i(M_1,M_2) = M_i\\
+    \inferrule{\Gamma\pipe\Phi \vdash M :  R_1\times R_2}{M = (\pi_1 M,\pi_2 M)}
+  \end{mathpar}
+  \caption{Products}
+  \label{fig:products}
+\end{figure}
+
+One additional connective we might desire is a ``parallel'' product of
+sets $R \times' S : \harr {C_1 \times C_2} {D_1 \times D_2}$ where $R
+: \harr {C_1} {D_1}$ and $R : \harr {C_2} {D_2}$. We don't need to
+define this as a primitive though because it is expressible in terms
+of the two kinds of products we have above:
+
+\[
+\inferrule
+{\Gamma \pipe \alpha : C_i; \beta : D_i \vdash R_i \isaSet}
+{\Gamma \pipe \alpha : C_1\times C_2; \beta : D_1 \times D_2
+  R_1[\pi_1\alpha/\alpha;\pi_1\beta/\beta] \times R_2[\pi_2\alpha/\alpha;\pi_2\beta/\beta]\isaSet}
+\]
+
+The empty products give us access now to several traditional notions
+of category theory that we couldn't before define, such as objects and
+presheaves of either variance:
+
+$$\textrm{Object}\,C = \varr 1 C$$
+$$\textrm{Presheaf}\,C = \harr 1 C$$
+$$\textrm{coPresheaf}\,C = \harr C 1$$
+
+However, we find that in practice we don't have much use for these
+concepts as most theorems about them easily generalize to more general
+theorems about arbitrary functors or profunctors.
+
+%% To define internal ends and coends, we need to go beyond the usual
+%% simple types and add equalizers (a form of subsets) and co-equalizers
+%% (a form of quotients) of judgments. Here's a sketch of the syntax for
+%% each. These make type checking very difficult as they result in
+%% obligations for proving equality of certain terms, which may or may
+%% not be decidable.
+
+%% \begin{mathpar}
+%%   \inferrule*[right=Equalizer Formation]
+%%   {\alpha;\beta\,|\, R \and
+%%   x : R \vdash t : S x:R \vdash u : S}
+%%   {\alpha;\beta\,|\,\{ x : R \,|\, t = u \}}
+
+%%   \inferrule*[right=Equalizer Intro]
+%%   {\Phi \vdash r : R \and t[r/x] = u[r/x]}
+%%   {\Phi \vdash \textrm{assert}(x.t = u)(r) : \{ x : R \,|\, t = u \}}
+
+%%   \inferrule*[right=Equalizer Elim]
+%%   {\Phi \vdash r : \{ x : R \,|\, t = u \}}
+%%   {\Phi \vdash \textrm{forget}(r) : R}
+
+%%   t[\textrm{forget}(r)/x] = u[\textrm{forget(r)}/x]
+
+%%   \textrm{assert}(x.t = u)(\textrm{forget}(r)) = r
+%%   \textrm{forget}(\textrm{assert}(x.t = u)(r)) = r
+%% \end{mathpar}
+
+%% \begin{mathpar}
+%%   \inferrule*[right=Co-equalizer Formation]
+%%   {\alpha;\beta\,|\, R \and
+%%   x : S \vdash t : R x:S \vdash u : R}
+%%   {\alpha;\beta\,|\,R/(\forall x:S. t = u)}
+
+%%   \inferrule*[right=Coequalizer Intro]
+%%   {\Phi \vdash r : R}
+%%   {\Phi \vdash \textrm{equate}(x:S.t=u)(r) : R/(\forall x.t=u)}
+
+%%   \inferrule*[right=Coequalizer Elim]
+%%   {\Psi \vdash r : R/(x.t = u) \and
+%%     \Phi,y:R,\Phi' \vdash s : S \and
+%%     s[t/y] = s[u/y]}
+%%   {\Phi,\Psi,\Phi' \vdash \textrm{lett} \textrm{equate}(y) = r.~s : S}
+
+%%   \textrm{equate}(x:S.t = u)(t[\gamma]) = \textrm{equate}(x:S.t = u)(u[\gamma])
+
+%%   \cdots
+%% \end{mathpar}
+
+%% \subsection{Positive Multiplicatives}
+
+%% \subsection{Negative Multiplicatives}
+
+%% \subsection{Additives}
 
 \section{Synthetic Category Theory}
+Next, we look at Synthetic category theory in a virtual equipment. In
+this section we will use category-theoretic terminology for judgment
+theory: categories, functors/objects, profunctors and transformations.
+
+\subsection{Profunctor Isomorphisms}
+
+Here are some isomorphisms of profunctors that should be provable
+synthetically.
+%
+Unit laws:
+\begin{enumerate}
+\item $\tensorexistsXwithYandZ {\beta} {P} {\punitinXfromYtoZ {D} {\beta} {\beta'}} \cong P[\beta'/\beta]$
+\item $\tensorexistsXwithYandZ {\alpha} {\punitinXfromYtoZ {C} {\alpha'} {\alpha}} {P} \cong P[\alpha'/\alpha]$
+\item $\homrallXYtoZ {\beta} {\punitinXfromYtoZ {D} {\beta'} {\beta}} {P}$
+\item $\homlallXYtoZ {\alpha} {\punitinXfromYtoZ {C} {\alpha} {\alpha'}} {P}$
+\end{enumerate}
+
+Associativity laws:
+\begin{enumerate}
+\item $\tensorexistsXwithYandZ {\beta} {P} {(\tensorexistsXwithYandZ \gamma Q R)} \cong \tensorexistsXwithYandZ {\gamma} {(\tensorexistsXwithYandZ \beta P Q)} R$
+\item $\homrallXYtoZ {\gamma} {\tensorexistsXwithYandZ \beta P Q} S \cong \homrallXYtoZ \beta P {\homrallXYtoZ \gamma Q S}$
+\item $\homlallXYtoZ {\alpha} {\tensorexistsXwithYandZ \beta P Q} S \cong \homlallXYtoZ \beta Q {\homlallXYtoZ \alpha P S}$
+\item $\homrallXYtoZ \gamma Q {(\homlallXYtoZ \alpha P S)} \cong \homlallXYtoZ \alpha P {(\homrallXYtoZ \gamma Q S)}$
+\end{enumerate}
+
+\subsection{Constructions: Categories, Functors and Natural Transformations}
+
+Hom sets don't provide primitive syntax for composition of morphisms
+in a category. As our first synthetic construction, we can
+\emph{define} composition using the elimination form.
+%
+\begin{definition}[Composition of Morphisms]
+  We define composition of morphisms as follows
+  \[
+  f : \beta \to_{\cat C} \beta', g : \beta' \to_{\cat C} \beta'' \vdash g \circ f :=
+  \when f = \id_{\beta}.~ \when g = \id_{\beta}.~ \id_{\beta} : \beta \to_{\cat c} \beta''
+  \]
+\end{definition}
+It should follow by $\beta\eta$ that this is associative and unital.
+
+Next, let's prove synthetically that every type $\alpha : \cat C
+\vdash A : \cat D$ is in fact functorial.
+\begin{definition}
+  There is a morphism
+  \[ \alpha^o:\cat C;\alpha':\cat C \vdash f : \alpha \to_{\cat C} \alpha' \vdash
+  \when f = \id_{\alpha}.~ \id_{A} : A[\alpha] \to_{\cat D} A[\alpha']
+  \]
+  which we call $A[f/\alpha]$.
+
+  Note that
+  \[ A[\id_\alpha/\alpha] = \id_{A} \]
+  \[ A[f;g] = A[f];A[g] \]
+\end{definition}
+Note the similarity to the proof that all functions are congruences in
+MLTT/HoTT.
+
+Next we can define natural transformations as specific kinds of
+terms whose type is $(F \alpha \to G \alpha')$.
+\begin{definition}
+  A natural transformation $\phi$ from $\alpha : \cat C \vdash A : \cat D$ to
+  $A'[\alpha]$ of the same type is a term
+  \[ \alpha\pipe \cdot \vdash \phi_\alpha : A[\alpha] \to_{\cat D} A'[\alpha] \]
+
+  Note that
+  \[ \alpha;\alpha' \pipe f : \alpha \to \alpha' \vdash
+  f \circ \phi_{\alpha} = \phi_{\alpha'} \circ f
+  \]
+  Because if $f$ is the identity, then both sides are the same.
+\end{definition}
+
+\subsection{Adjoint Functors}
+
+There are two common definitions of adjoint functors, one in terms of
+Hom-sets and one in terms of units and counits.
+
+\begin{definition}[Hom Iso Adjunction]
+  A functor $\alpha : \cat C \vdash L : \cat D$ is left adjoint to a
+  functor $\beta : \cat D \vdash R : \cat C$ if there is an isomorphism
+  \[ (L[\alpha] \to_{\cat D} \beta) \cong (\alpha \to_{\cat C} R[\beta]) \]
+\end{definition}
+
+\begin{definition}[Unit/Counit Adjunction]
+  Alternatively, $L$ is left adjoint to $R$ if there exist
+  transformations
+  \[ \eta : \alpha \to_{\cat C} R[L[\alpha]/\beta] \] % unit of the monad 1 -> RL 
+  \[ \epsilon : L[R[\beta]/\alpha] \to_{\cat D} \beta  \] % counit of the comonad LR -> 1
+  satisfying the triangle/zig-zag identities
+  \[ \id_{R[\beta]} = R[\epsilon_{\beta}] \circ \eta_{R[\beta]} \]
+  \[ \id_{L[\alpha]} = \epsilon_{L[\alpha]} \circ L[\eta_{\alpha}] \]
+\end{definition}
+
+\begin{definition}[Universal Morphism Adjunction]
+  A functor $\alpha: C \vdash L : D$ is a left adjoint if there exists a 
+\end{definition}
+
+We should be able to synthetically prove these are equivalent.
+
+\subsection{Tensors and CoTensors/Powers}
+
+David Jaz Myers tells me that to do all of formal category theory in
+an equipment, you just need to have hom sets, tensors and cotensors.
+
+As another example synthetic theorem, we can prove the covariant and
+contravariant \emph{co-yoneda lemmas}, which are precisely the
+statement that Hom is the unit of the tensor product.
+\begin{lemma}[CoYoneda]
+  We can define isomorphisms
+  \[ \exists \beta. (\alpha \to_{\cat C} \beta) \odot Q \equiv Q[\alpha/\beta] \]
+  \[ \exists \beta. P \odot (\beta \to_{\cat C} \gamma) \equiv P[\gamma/\beta] \]
+  The forward direction is defined by considering the case that the
+  function is the identity and the backward direction by using the
+  identity.
+\end{lemma}
+\begin{proof}
+  The terms are
+  \[ f : \exists \beta. (\alpha\to_{\cat C} \beta) \odot Q \vdash
+  \lett (\beta,g,h) = f.~ \when g = \id_{\alpha}.~ h : Q[\alpha/\beta]
+  \]
+  And the inverse
+  \[ h : Q[\alpha/\beta] \vdash (\alpha, \id_{\alpha}, h) : \exists \beta. (\alpha\to_{\cat C}\beta)\odot Q\]
+
+  One direction of iso follows by $\beta$, the other by $\eta$.
+\end{proof}
+
+And we get another synthetic theorem: the Yoneda lemma, which says
+that Hom is the unit for the cotensor product.
+\begin{lemma}[Synthetic Yoneda]
+  We can define isomorphisms
+  \[ \forall \alpha.~ (\alpha \to \beta) \triangleright P[\alpha;\gamma] \cong P[\beta/\alpha;\gamma] \]
+  \[ \forall \gamma.~ P[\alpha;\gamma] \triangleleft (\beta \to \gamma) \cong P[\alpha;\beta/\gamma]\]
+\end{lemma}
+\begin{proof}
+  Forward direction
+  \[ f : \forall \alpha.~(\alpha\to\beta)\triangleright P \vdash
+  f(\beta,\id_{\beta})^\triangleright : P[\beta/\alpha]
+  \]
+  Backwards
+  \[ x : P[\beta/\alpha]\vdash
+  \lambda^\triangleright(\alpha,\phi : \alpha \to \beta).~\when \phi = \id_{\beta}.~ x : \forall \alpha. (\alpha\to \beta) \triangleright P
+  \]
+  One direction is by $\beta$, the other $\eta$.
+\end{proof}
+
+\subsection{Right (Left) Adjoints preserve (co-)Limits}
+
+Let's prove that classic theorem!
+
+\begin{theorem}
+  Let $\alpha : \cat C \vdash A \cat D$ be an arbitrary functor and
+  $\beta : \cat D \vdash R : \cat E$ be a functor with a left adjoint
+  $L$. Then, if the limit $\lim_\alpha A$ exists, then $\lim_\alpha
+  R[A/\beta]$ also exists and is
+  \[ \lim_\alpha R[A/\beta] = R[\lim_\alpha A] \]
+\end{theorem}
+\begin{proof}
+  We need to give an isomorphism
+  \[ \gamma \to R[\lim_\alpha A/\beta] \cong PRan_{()}{\alpha. R[A/\beta]} = \forall \alpha. (\gamma \to R[A/\beta]) \triangleleft (() \to_1 ()) \]
+  By assumption we have an isomorphism
+  \[ l : \beta \to \lim_\alpha A \cong PRan_{()}{\alpha. A} = \forall \alpha. (\beta \to A[\alpha]) \triangleleft (()\to_1 ())\]
+  TODO.
+\end{proof}
+
+%% \subsection{Additive Facts}
+
+%% I think the following isomorphisms should hold, which connect additive
+%% connectives on objects with the analogous connective on profunctors:
+%% \[ () \to_1 () \cong () \]
+%% \[ (\pi_1 \alpha, \pi_2 \alpha) \to_{\cat C \times \cat D} (\pi_1 \beta, \pi_2 \beta) \cong
+%% ((\pi_1 \alpha) \to_{\cat C} (\pi_1 \beta)) \times ((\pi_2 \alpha) \to_{\cat C} (\pi_2 \beta)) \]
 
@@
 
+Next, we introduce the notion of category internal to our syntactic
+virtual equipment and the \emph{monoids and modules} translation that
+shows that, with the inclusion of hom sets, our syntax is already
+suitable for doing syntactic category theory.
+
+The benefits of our syntax become clear when we define internal
+categories, functors, profunctors and transformations: the definitions
+read exactly as the classical definitions.
+\begin{definition}[Internal Category]
+  An \emph{(internal) category} consists of\footnote{the following is
+    easily interpreted as a $\Sigma$ type}:
+  \begin{enumerate}
+  \item A sort $C_0 : \Sort$
+  \item A span $C_1 : \harr{C_0}{ C_0}$
+  \item Identity arrows
+    $C_r : \pendallXdotY {\alpha :  C_0} {\harrapp{C_1} \alpha \alpha}$
+  \item Composition of arrows
+    \[ C_t : \pendallXdotY \alpha {\homrallXYtoZ \beta {\harrapp{C_1} \alpha \beta} {\homrallXYtoZ \gamma {\harrapp{C_1} \beta \gamma} {\harrapp{C_1} \alpha \gamma}}}\]
+
+    %% \[ C_t^\triangleright : \forall \alpha :  C_0. \forall \beta. ({C_1} \beta \alpha) \triangleright \forall \gamma. ({C_1} \gamma \beta) \triangleright {C_1} \gamma \alpha \]
+
+    %% or
+
+    %% \[ C_t^\triangleleft : \forall \alpha. \forall \beta. (\forall \gamma. {C_1} \alpha \gamma \triangleleft {C_1} \beta \gamma) \triangleleft {C_1} \alpha \beta
+    %% \]
+
+    %% or using tensors and covariant homs
+
+    %% \[ C_t^\odot : \forall \alpha. \forall \gamma. (\exists \beta. {C_1}\alpha\beta\odot {C_1}\beta\gamma)\triangleright {C_1}\alpha \gamma \]
+
+    %% or using tensors and contravariant homs
+
+    %% \[ C_t^\odot : \forall \gamma. \forall \alpha. {C_1}\alpha\gamma \triangleleft (\exists \beta. {C_1}\alpha\beta\odot {C_1}\beta\gamma)\]
+  \item Satisfying right unitality, i.e., for any $\Phi \vdash f :
+    \harrapp {C_1} A B$, we'd like $f; r(B) = f$. We can formalize this
+    using an equality of two transformations of type $\pendallXdotY
+    \alpha {\homrallXYtoZ \beta {\harrapp{C_1} \alpha \beta} {\harrapp{C_1} \alpha \beta}}$
+
+    \[ \pendlambdaXdotY \alpha {\homrlambdaXatYdotZ f \beta {\homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {C_t} \alpha} {f} \beta} {\pendappXtoY {C_r} \beta} \beta}}
+    = \pendlambdaXdotY \alpha {\homrlambdaXatYdotZ f \beta f} \]
+
+    These kinds of equalities between 2-cells are very common in
+    internal category theory, so to save space we will suppress the
+    sequence of $\lambda$s that start such equalities. So the above can simply be written as
+
+    \[ { {\homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {C_t} \alpha} {f} \beta} {\pendappXtoY {C_r} \beta} \beta}} = { f}\]
+
+  \item and left unitality (an equality of the same type as right unitality):
+
+    \[ { {\homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {C_t} \alpha} {\pendappXtoY {C_r} \alpha} \alpha} f \beta}} = { f}\]
+  \item and associativity, which is an equality at type
+    \[ \pendallXdotY \alpha {\homrallXYtoZ \beta {\harrapp{C_1} \alpha \beta} {\homrallXYtoZ \gamma {\harrapp{C_1} \beta \gamma} {\homrallXYtoZ \delta {\harrapp{C_1} \gamma \delta} {\harrapp{C_1} \alpha \delta}}}}\]
+
+    \[ \homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {C_t}\alpha} f \beta} {(\homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {C_t}\beta} g \gamma} h \delta)} \delta
+    =
+    \homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {C_t}\alpha} {(\homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {C_t}\alpha} f \beta} g \gamma)} \gamma} h \delta \]
+  \end{enumerate}
+\end{definition}
+
+\begin{definition}[Internal Functor]
+  A functor $F$ between internal categories $C$ and $D$ consists of
+  \begin{enumerate}
+  \item a map of objects $F_0 : \varr {C_0} {D_0}$
+  \item a map of arrows
+    $F_1 : \pendallXdotY {\alpha}
+    {\homrallXYtoZ \beta {\harrapp{C_1} \alpha \beta} {\harrapp{D_1} {F_0 \alpha}{F_0 \beta}}}$
+  \item That preserves identity. This is given by equality between the
+    following terms of type $\forall \alpha:{{C_0}}. D_1 {(F_0 \alpha)}{(F_0\alpha)}$:
+    \[ \homrappXtoYatZ {\pendappXtoY {F_1}\alpha} {\pendappXtoY {C_r} \alpha} \alpha = \pendappXtoY {D_r}{F_0\alpha} \]
+  \item And preserves composition.
+    This is given by an equality between the following terms of type
+
+    \[ \pendallXdotY \alpha
+    \homrallXYtoZ {\beta} {\harrapp{C_1} \alpha \beta} {\homrallXYtoZ \gamma {\harrapp{C_1} \beta \gamma} {\harrapp{D_1} {F_0 \alpha} {F_0 \gamma}}}\]
+
+    First, we can compose in $C$, and then map using $F$:
+    \[ \homrappXtoYatZ {\pendappXtoY{F_1} \alpha} {(\homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {C_t} \alpha} f \beta} g \gamma)}\gamma\]
+
+    or we can map the morphisms with $F$, and then compose them in $D$:
+    \[ \homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {D_t} {(F_0 \alpha)}} {(\homrappXtoYatZ {\pendappXtoY {F_1} \alpha} f \beta)} {(F_0 \beta)}} {(\homrappXtoYatZ {\pendappXtoY {F_1} \beta} g \gamma)} {{F_0} \gamma}\]
+  \end{enumerate}
+\end{definition}
+
+We can immediately define identity and composition of functors.
+
+\begin{definition}
+  For any $C : \textrm{Cat}$, we can define an identity functor
+  $\textrm{Id}_C : \textrm{Functor} C C$ by
+  \begin{enumerate}
+  \item $\textrm{Id}_0 = \lambda \alpha:C_0. \alpha$
+  \item $\textrm{Id}_1 = \pendlambdaXdotY \alpha
+    {\homrlambdaXatYdotZ f {\alpha'} f}$
+  \item Preservation of identity requires proving
+    \[
+    \lambda \alpha. \pendappXtoY {C_r}{\textrm{Id}_0\alpha} \]
+    is equal to
+    \[ \lambda \alpha.
+    \homrappXtoYatZ {\pendappXtoY {\textrm{Id}_1} \alpha} {(\pendappXtoY {C_r}\alpha)} \alpha\]
+    Both sides simply $\beta$ reduce to $ \lambda \alpha. \pendappXtoY {C_r} \alpha$
+    and the result follows by reflexivity.
+  \item Preservation of composition requires proving (under a $\lambda \alpha. \homrlambdaXatYdotZ f \beta \homrlambdaXatYdotZ g \gamma$)
+    \[
+    \homrappXtoYatZ {\pendappXtoY {{\textrm{Id}}_1} \alpha} {(\homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {C_t} \alpha} f \beta} g \gamma)}\gamma
+    = \homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {C_t} {{\textrm{Id}}_0 \alpha}} {(\homrappXtoYatZ {\pendappXtoY {{\textrm{Id}}_1} \alpha} f \beta)} {{\textrm{Id}}_0 \beta}} {(\homrappXtoYatZ {\pendappXtoY{{\textrm{Id}}_1} \beta} g \gamma)} {{\textrm{Id}}_0 \gamma}\]
+    Both sides $\beta$ reduce to
+
+    \[ \homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {C_t} \alpha} f \beta} g \gamma\]
+    so the result holds by reflexivity.
+  \end{enumerate}
+\end{definition}
+
+\begin{definition}
+  Given categories $C,D,E$ and functors $F : \textrm{Functor} C D$ and
+  $G : \textrm{Functor} D E$, we define the composition $G \circ F :
+  \textrm{Functor} C E$ as
+  \begin{enumerate}
+  \item $(G \circ F)_0 = \lambda \alpha. {G_0 ({F_0} \alpha)}$
+  \item $(G \circ F)_1 =
+    \pendlambdaXdotY \alpha {
+    \homrlambdaXatYdotZ f \beta
+    {\homrappXtoYatZ {\pendappXtoY {G_1} {F_0 \alpha}} {\homrappXtoYatZ {\pendappXtoY {F_1} \alpha} f \beta} {F_0 \beta}}}$
+  \item Preservation of identity requires proving
+    \[ \lambda \alpha. \pendappXtoY {E_r} {(G \circ F)_0\alpha} \]
+    is equal to
+    \[ \lambda \alpha.
+    \homrappXtoYatZ {\pendappXtoY {(G \circ F)_1} \alpha} {\pendappXtoY {C_r}\alpha} \alpha \]
+
+    Expanding definitions and $\beta$ reducing we need to show
+
+    \[ \pendlambdaXdotY\alpha {\pendappXtoY {E_r}{G_0 (F_0 \alpha)}} \]
+    is equal to
+    \[ \pendlambdaXdotY \alpha {\homrappXtoYatZ {\pendappXtoY {G_1} {F_0 \alpha}} {(\homrappXtoYatZ {\pendappXtoY {F_1} \alpha} {\pendappXtoY {C_r} \alpha} \alpha)} {F_0 \alpha}}\]
+
+    Proceeding from right to left applying transitivity:
+    \begin{align*}
+      &\pendlambdaXdotY \alpha {\homrappXtoYatZ {\pendappXtoY {G_1} {F_0 \alpha}} {(\homrappXtoYatZ {\pendappXtoY {F_1} \alpha} {\pendappXtoY {C_r} \alpha} \alpha)} {F_0 \alpha}}\\
+      &=\pendlambdaXdotY \alpha {\homrappXtoYatZ {\pendappXtoY {G_1} {F_0 \alpha}} {(\pendappXtoY{(\pendlambdaXdotY \alpha {\homrappXtoYatZ {\pendappXtoY {F_1} \alpha} {\pendappXtoY {C_r} \alpha} \alpha})}\alpha)}} {F_0 \alpha}\\
+      &=\pendlambdaXdotY \alpha {\homrappXtoYatZ {\pendappXtoY {G_1} {F_0 \alpha}} {(\pendappXtoY{(\pendlambdaXdotY \alpha {{\pendappXtoY {D_r} {F_0 \alpha}}})}\alpha)}} {F_0 \alpha}\tag{*}\\
+      &=\pendlambdaXdotY \alpha {\homrappXtoYatZ {\pendappXtoY {G_1} {F_0 \alpha}} {\pendappXtoY {D_r} {F_0 \alpha}}} {F_0 \alpha}\\
+      &= \pendlambdaXdotY \alpha
+      \pendappXtoY {(\pendlambdaXdotY \beta {{\homrappXtoYatZ {\pendappXtoY {G_1} \beta} {\pendappXtoY {D_r} \beta} \beta}})} {F_0 \alpha}\\
+      &= \pendlambdaXdotY \alpha
+      \pendappXtoY {(\pendlambdaXdotY \beta \pendappXtoY {E_r} {G_0 \beta})} {F_0 \alpha}\tag{*}\\
+      &= \pendlambdaXdotY \alpha \pendappXtoY {E_r} {G_0 (F_0 \beta)}\\
+    \end{align*}
+
+    Where the two marked lines follow from congruence using the
+    assumptions about $F,G$.
+
+  \item Preservation of Composition.
+
+    \[ \homrappXtoYatZ {\pendappXtoY{(G \circ F)_1} \alpha} {(\homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {C_t} \alpha} f \beta} g \gamma)}\gamma\]
+    equals
+
+    \[ \homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {E_t} {(G \circ F)_0 \alpha}} {(\homrappXtoYatZ {\pendappXtoY {(G\circ F)_1} \alpha} f \beta)} {(G\circ F)_0 \beta}} {(\homrappXtoYatZ {\pendappXtoY {(G\circ F)_1} \beta} g \gamma)} {{F_0} \gamma}\]
+
+    Reducing, we need to show
+    \[ \homrappXtoYatZ {\pendappXtoY{G_1} {F_0 \alpha}} {(\homrappXtoYatZ {\pendappXtoY {F_1} \alpha} {(\homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {C_t}\alpha} f {\beta}} g {\gamma})} \gamma)} {F_0 \gamma} \]
+
+    is equal to
+    \[ \homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {E_t} {G_0(F_0\alpha)}}
+      {({\homrappXtoYatZ {\pendappXtoY {G_1} {F_0 \alpha}} {(\homrappXtoYatZ {\pendappXtoY {F_1} \alpha} f \beta)} {F_0 \beta}})}
+      {G_0(F_0\beta)}}
+       {({\homrappXtoYatZ {\pendappXtoY {G_1} {F_0 \beta}} {(\homrappXtoYatZ {\pendappXtoY {F_1} \beta} g \gamma)} {F_0 \gamma}})}
+       {{F_0} \gamma}\]
+
+    Going from left to right:
+    \begin{align*}
+      &\homrappXtoYatZ {\pendappXtoY{G_1} {F_0 \alpha}} {(\homrappXtoYatZ {\pendappXtoY {F_1} \alpha} {(\homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {C_t}\alpha} f {\beta}} g {\gamma})} \gamma)} {F_0 \gamma}\\
+      &=
+      \homrappXtoYatZ
+      {\pendappXtoY{G_1} {F_0 \alpha}}
+      {(({\homrappXtoYatZ {\homrappXtoYatZ{\pendappXtoY{(\pendlambdaXdotY \alpha {\homrlambdaXatYdotZ f \beta {\homrlambdaXatYdotZ g \gamma {\homrappXtoYatZ {\pendappXtoY {F_1} \alpha} {(\homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {C_t}\alpha} f {\beta}} g {\gamma})} \gamma}}})}\alpha} f \beta} g \gamma}))}
+      {F_0 \gamma}\\
+      &=
+      \homrappXtoYatZ
+      {\pendappXtoY{G_1} {F_0 \alpha}}
+      {(({\homrappXtoYatZ {\homrappXtoYatZ{\pendappXtoY{(\pendlambdaXdotY \alpha {\homrlambdaXatYdotZ f \beta {\homrlambdaXatYdotZ g \gamma {\homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {D_t} {(F_0 \alpha)}} {(\homrappXtoYatZ {\pendappXtoY {F_1} \alpha} f \beta)} {(F_0 \beta)}} {(\homrappXtoYatZ {\pendappXtoY {F_1} \beta} g \gamma)} {{F_0} \gamma}}}})}\alpha} f \beta} g \gamma}))}
+      {F_0 \gamma}\tag{*}\\
+      &=
+      \homrappXtoYatZ
+      {\pendappXtoY{G_1} {F_0 \alpha}}
+      {\homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {D_t} {(F_0 \alpha)}} {(\homrappXtoYatZ {\pendappXtoY {F_1} \alpha} f \beta)} {(F_0 \beta)}} {(\homrappXtoYatZ {\pendappXtoY {F_1} \beta} g \gamma)} {{F_0} \gamma}}
+      {F_0 \gamma}\\
+      &=
+      {(({\homrappXtoYatZ {\homrappXtoYatZ{\pendappXtoY{(\pendlambdaXdotY {\alpha'} {\homrlambdaXatYdotZ {f'} {\beta'} {\homrlambdaXatYdotZ {g'} {\gamma'} {{\homrappXtoYatZ {\pendappXtoY {G_1} {\alpha'}} {(\homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {D_t}{\alpha'}} {f'} {\beta'}} {g'} {\gamma'})} {\gamma'}}}}})}{F_0\alpha}} {(\homrappXtoYatZ {\pendappXtoY {F_1} \alpha} f \beta)} {F_0\beta}} {(\homrappXtoYatZ {\pendappXtoY {F_1} \beta} g \gamma)} {F_0\gamma}}))}\\
+      &=
+      {(({\homrappXtoYatZ {\homrappXtoYatZ{\pendappXtoY{(\pendlambdaXdotY {\alpha'} {\homrlambdaXatYdotZ {f'} {\beta'} {\homrlambdaXatYdotZ {g'} {\gamma'} {{\homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {E_t} {(G_0 {\alpha'})}} {(\homrappXtoYatZ {\pendappXtoY {G_1} {\alpha'}} {f'} {\beta'})} {(G_0 \beta')}} {(\homrappXtoYatZ {\pendappXtoY {G_1} {\beta'}} {g'} {\gamma'})} {{G_0} \gamma'}}}}})}{F_0\alpha}} {(\homrappXtoYatZ {\pendappXtoY {F_1} \alpha} f \beta)} {F_0\beta}} {(\homrappXtoYatZ {\pendappXtoY {F_1} \beta} g \gamma)} {F_0\gamma}}))}\\
+      &= \homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {E_t} {G_0(F_0\alpha)}}
+      {({\homrappXtoYatZ {\pendappXtoY {G_1} {F_0 \alpha}} {(\homrappXtoYatZ {\pendappXtoY {F_1} \alpha} f \beta)} {F_0 \beta}})}
+      {G_0(F_0\beta)}}
+       {({\homrappXtoYatZ {\pendappXtoY {G_1} {F_0 \beta}} {(\homrappXtoYatZ {\pendappXtoY {F_1} \beta} g \gamma)} {F_0 \gamma}})}
+       {{F_0} \gamma}
+    \end{align*}
+  \end{enumerate}
+\end{definition}
+
+\begin{definition}[Internal Profunctor]
+  A profunctor $R$ between internal categories $\cat C$ and $\cat D$
+  consists of
+  \begin{enumerate}
+  \item A span $R_s : \harr {C_0} {D_0}$.
+  \item A ternary ``mixed composition'' action:
+    $R_t :\pendallXdotY \alpha {\homrallXYtoZ {\alpha'} {\harrapp {C_1} {\alpha}{\alpha'}} {\homrallXYtoZ {\beta'} {\harrapp {R_s} {\alpha'}{\beta'}} {\homrallXYtoZ {\beta} {\harrapp {D_1} {\beta'} {\beta}} {\harrapp {R_s} {\alpha}{\beta}}}}}$
+
+  \item Satisfying unitality (under a $\pendlambdaXdotY \alpha {\homrlambdaXatYdotZ f \beta {}}$):
+    \[
+    f = \homrappXtoYatZ {\homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {R_t} \alpha} {\pendappXtoY {C_r} \alpha} \alpha}f \beta} {\pendappXtoY {D_r} \beta} \beta
+    \]
+
+  \item And associativity (both with a large implicit
+    $\pendlambdaXdotY \alpha {\homrlambdaXatYdotZ g \beta \homrlambdaXatYdotZ {g'} {\beta'}
+    \homrlambdaXatYdotZ {r} {\beta''}\homrlambdaXatYdotZ {f'} {\alpha''}\homrlambdaXatYdotZ {f} {\alpha'}}$)
+
+    % f : a -> a'
+    % f' : a' -> a''
+    % r : a'' -> b''
+    % g' : b'' -> b'
+    % g : b' -> b
+    %%
+    %
+    % f o (f' o r o g') o g
+    %% and
+    % (f o f') o r o (g' o g)
+
+    \[ \homrappXtoYatZ {\homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {R_t} {\alpha}} {f} {\alpha'}}{({\homrappXtoYatZ {\homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {R_t} {\alpha'}} {f'} {\alpha''}}r {\beta''}} {g'} {\beta'}})} {\beta'}} {g} {\beta}\]
+    is equal to
+    \[ \homrappXtoYatZ
+       {\homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {R_t} {\alpha}} {(\homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {C_t} {\alpha}} {f} {\alpha'}} {f'} {\alpha''})} {\alpha''}} r {\beta''}}
+       {(\homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {D_t} {\beta''}} {g'} {\beta'}} g {\beta})} {\beta}\]
+  \end{enumerate}
+\end{definition}
+
+We can immediately define the Hom profunctor for any category $C$ as
+$(\text{Hom}_C R)_s = C_1$ with actions given by $C_t$.
+
+Next, we show that Profunctors can be restricted by functors.
+
+\begin{definition}[Restriction of Profunctors]
+  If $C, D, C', D' : \text{Cat}$, and $F : \text{Functor} C C', G : \text{Functor} D D'$ and
+  $R : \text{Prof} C' D'$, then we can define a profunctor $R[F;G] : \text{Prof} C D$ by
+  \begin{enumerate}
+  \item $R[F;G]_s = \lambda (\alpha:C,\beta:D). R_s (F_0 \alpha) (G_0 \beta)$
+  \item with composition action given by
+
+    \[ R[F;G]_t =
+    \pendlambdaXdotY \alpha {\homrlambdaXatYdotZ f {\alpha'} {\homrlambdaXatYdotZ r {\beta'} {\homrlambdaXatYdotZ g {\beta} {\homrappXtoYatZ {\homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {R_t} {F_0 \alpha}} {(\homrappXtoYatZ {\pendappXtoY {F_1} {\alpha}} f {\alpha'})} {F_0 \alpha'}} r {G_0 \beta'}} {(\homrappXtoYatZ {\pendappXtoY {G_1} {\beta'}} g \beta)} {G_0 \beta}}}} }
+ \]
+  %% \item $\cat C'$ composition:
+  %%   \[ \lambda \alpha. \lambda^\triangleleft (\beta, (f : C_1 \alpha \alpha',\alpha',\alpha', r : R_s \alpha' \beta)). R_l(F_1 f, r) \]
+  %% \item $\cat D'$ composition:
+  %%   \[ \lambda \alpha. \lambda^\triangleleft ('\beta, (r : R_s \alpha \alpha',\alpha', \beta',h : D_1 \beta' \beta)). R_r(r, G_1 h) \]
+  \end{enumerate}
+  Identity and associativity properties should follow. TODO
+\end{definition}
+
+\begin{definition}[Homomorphism]
+  Given $R : \text{Prof} C D$ and $S : \text{Prof} C D$, a
+  homomorphism $m$ from $R$ to $S$ consists of
+  \begin{enumerate}
+  \item $m_f : \pendallXdotY \alpha {\homrallXYtoZ \beta {\harrapp R \alpha \beta}{\harrapp S \alpha \beta}}$
+    % m_f (R_t f r g) = S_t f (m_f r) g
+  \item Satisfying naturality (under a $\ldots$)
+
+    \[
+    \homrappXtoYatZ {\pendappXtoY{m_f} \alpha} {(\homrappXtoYatZ {\homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {R_t} \alpha} f {\alpha'}} r {\beta'}} g \beta)} \beta
+    =
+    \homrappXtoYatZ {\homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {S_t} \alpha} {f} {\alpha'}} {(\homrappXtoYatZ {\pendappXtoY {m_f}{\alpha'}} r {\beta'})} {\beta'}} g \beta
+    \]
+  \end{enumerate}
+\end{definition}
+
+\begin{definition}
+  An isomorphism between profunctors $R$ and $S$ is a mutually inverse
+  pair of homomorphisms.
+\end{definition}
+
+Additionally, we can define natural transformations between functors,
+but more generally, we can define a natural element of any
+endo-profunctor, where natural transformations from $F$ to $G$ are the
+same as natural elements of $\text{Hom}(F-,G=)$.
+
+\begin{definition}[Natural Element]
+  A natural element $\phi$ of a profunctor $R : \text{Prof}\,C C$ is a term
+  \[ \phi : \pendallXdotY \alpha {\harrapp R \alpha \alpha} \]
+  that is natural in that
+  \[
+  \homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {C_t} \alpha} {\pendappXtoY \phi \alpha} \alpha} f {\alpha'}
+  = \homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {C_t } \alpha} f {\alpha'}} {\pendappXtoY \phi {\alpha'}}{\alpha'}
+  \]
+\end{definition}
+
+\begin{definition}[Natural Transformation]
+  A natural transformation from $F : \text{Functor} \,C\, D$ and $G :
+  \text{Functor}\,C\,D$ is a natural element of $\text{Hom}(D)[F;G]$
+\end{definition}
+
+Composition of natural transformations is then given by $C_t$.
+
+Natural elements are closely related to homomorphisms: a natural
+element equivalent to a homomorphism out of a Hom profunctor. This is
+our first instance of a ``Yoneda'' principle that the behavior of some
+structure that varies naturally with the morphisms of a category is
+fully determined by its action on the identity morphisms.
+
+\begin{theorem}
+  Natural elements of $R : \text{Prof}\, C C$ are equivalent to
+  homomorphisms from $\text{Hom}_C$ to $R$.
+\end{theorem}
+\begin{enumerate}
+\item Given a natural element $\phi$ of $R$, we can define a
+  homomorphism by
+  \[ \pendlambdaXdotY \alpha {\homrlambdaXatYdotZ f {\alpha'} \homrappXtoYatZ {\homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {R_t} \alpha} {\pendappXtoY {C_r} \alpha} \alpha}{\pendappXtoY \phi \alpha} {\alpha}} f {\alpha'}} \]
+\item Given a homomorphism $m$ from $\text{Hom}(C)$ to $R$, define the
+  element as
+  \[ \pendlambdaXdotY \alpha {\homrappXtoYatZ {\pendappXtoY {m_f} \alpha} {\pendappXtoY {C_r}\alpha} \alpha}  \]
+
+  We should be able to show this is an isomorphism in the logic TODO!
+\end{enumerate}
+
+%% This last definition is probably unfamiliar to the reader that is not
+%% experienced with profunctors.
+%% %
+%% However, we can show that the traditional definition is equivalent to
+%% a natural transformation $\alpha : \cat C_0 \pipe \cdot \vdash
+%% t_{\alpha} : F(\alpha) \to_{\cat D} G(\alpha')$.
+
+%% \begin{definition}[Traditional Natural Transformation]
+%%   A traditional natural transformation from $\alpha : \cat C_0 \vdash
+%%   F(\alpha) : \cat D_0$ to $\alpha : \cat \vdash G(\alpha') : \cat
+%%   D_0$ consists of
+%%   \begin{enumerate}
+%%   \item A term $\cdot \vdash t_{\alpha} : F(\alpha) \to_{\cat D} G(\alpha)$
+%%   \item Such that
+%%     \[ \phi : \alpha \to_{\cat C} \alpha' \vdash t_{\alpha'} \circ F_1(\phi) \equiv G_1(\phi) \circ t_{\alpha} : F(\alpha) \to_{\cat D} G(\alpha')
+%%     \]
+%%   \end{enumerate}
+%% \end{definition}
+
+%% \begin{theorem}[Equivalent Presentations of Natural Transformations]
+%%   There is a bijection between traditional natural transformations $t
+%%   : F \to G$ and profunctor homomorphisms $h : \alpha \to_{\cat C}
+%%   \alpha' \vdash F(\alpha) \to_{\cat D} G(\alpha')$ given by
+%%   \begin{enumerate}
+%%   \item From a natural transformation $t$ we define a homomorphism
+%%     $h(t)$ by composing on the $F$ side. The choice of this side is
+%%     arbitrary and we could just as easily have used $G$ since they are
+%%     equal by naturality.
+%%     \[ \phi : \alpha \to_{\cat C} \alpha' \vdash t_{\alpha'} \circ F_1(\phi) : F(\alpha) \to_{\cat D} G_1(\alpha') \]
+%%   \item From a homomorphism $\phi.h$, we define a natural transformation $t(h)$ as
+%%     \[ \alpha \pipe \cdot \vdash h[{\id_{\alpha}}/\phi] \]
+%%   \end{enumerate}
+%% \end{theorem}
+%% \begin{proof}
+%%   First we show that the maps $h(\cdot),t(\cdot)$ actually produce
+%%   homomorphisms and natural transformations.
+%%   \begin{enumerate}
+%%   \item On the left, we need to show $h(t)_{\phi} \circ F_1(\psi) =
+%%     h(t)_{|phi \circ \psi}$. Expanding the definition, we just need
+%%     associativity and functoriality of $F_1$:
+%%     \begin{align*}
+%%       h(t)_{\phi} \circ F_1(\psi) &\equiv (t_{\alpha'} \circ F_1(\phi)) \circ F_1(\psi)\\
+%%       &\equiv t_{\alpha'} \circ F_1(\phi \circ \psi)\\
+%%       &\equiv h(t)_{\phi \circ \psi}
+%%     \end{align*}
+%%     on the right, we use naturality:
+%%     \begin{align*}
+%%       G_1(\theta) \circ h(t)_{\phi} &\equiv G_1(\theta) \circ (t_{\alpha'} \circ F_1(\phi))\\
+%%       &\equiv (t_{\alpha''} \circ F_1(\theta)) \circ F_1(\phi)\\
+%%       &\equiv t_{\alpha''} \circ F_1(\theta \circ \phi)\\
+%%       &\equiv t_{\theta \circ \phi}
+%%     \end{align*}
+%%   \item In the other direction, to prove naturality:
+%%     \begin{align*}
+%%       h[\id_\alpha/\phi] \circ F_1(\psi) &\equiv h[\id_\alpha \circ \psi/\phi]\\
+%%       &\equiv h[\psi/\phi]\\
+%%       &\equiv h[\psi \circ \id_{\alpha'}/\phi]\\
+%%       &\equiv G_1(\psi) \circ h[\id_{\alpha}/\phi]
+%%     \end{align*}
+%%   \end{enumerate}
+%%   Next we seek to prove that the functions are mutually inverse.
+%%   \begin{enumerate}
+%%   \item Round trip for homomorphisms:
+%%     \begin{align*}
+%%       h(t(h))_{\phi} &= t(h)_{\alpha} \circ F_1[\phi]\\
+%%       &= h[\id_{\alpha}] \circ F_1[\phi]\\
+%%       &= h[\id \circ \phi]\\
+%%       &= h[\phi]
+%%     \end{align*}
+%%   \item Round trip for natural transformations uses the fact that
+%%     functors preserve identity.
+%%     \begin{align*}
+%%       t(h(t))_{\alpha} &= h(t)_{\id_{\alpha}}\\
+%%       &= t_{\alpha} \circ F_1[\id_{\alpha}]\\
+%%       &= t_{\alpha} \circ \id_{F_0\alpha}\\
+%%       &= t_{\alpha}
+%%     \end{align*}
+%%   \end{enumerate}
+%% \end{proof}
+
+\subsection{Universal Properties}
+
+Since we have profunctors as a primitive notion, the most natural
+presentation of universal properties is given by the notion of a
+functor representing a profunctor.
+
+First, we have the ``algebraic'' notion of a functor representing a
+profunctor, i.e., when a profunctor $R : C\pto D$ is equivalent to
+$Hom_D(-,G=)$.
+\begin{definition}[Right Representability]
+  Let $C, D : \text{Cat}$ and $Q : \textrm{Prof}\,C D$ be a profunctor
+  between them. Then a \emph{right representation} of $Q$ consists of
+  \begin{enumerate}
+  \item A functor $G : \textrm{Functor}\, D C$
+  \item An isomorphism $I,I^-$ between $Q$ and $\text{Hom}(C)[Id;G]$
+  \end{enumerate}
+\end{definition}
+
+This is a nice definition for applications, but a lot of the structure
+(homomorphism, functoriality of G) is determined by other components
+of the structure. This can be simplified to the notion of a
+``universal morphism''.
+
+\begin{definition}[Universal Morphism]
+  A profunctor $Q : \textrm{Prof}\, C D$ is right representable when
+  there exists
+  \begin{enumerate}
+  \item A function on objects $G_0 : \varr {D_0} {C_0}$
+  \item A ``universal morphism'', i.e., a term $\epsilon :
+    \pendallXdotY \beta {\harrapp Q {G_0 \beta} \beta}$ (not
+    presupposed to be natural).
+  \item An ``introduction rule'', i.e., a term $I : \pendallXdotY
+    \alpha {\homrallXYtoZ \beta {\harrapp {Q_s}\alpha \beta} {\harrapp
+        {C_1} \alpha {G_0\beta}}}$ (not presupposed to satisfy the
+    homomorphism condition).
+  \item Such that composition with $\epsilon$ is a left and right
+    inverse to $I$. In more detail, we have two equalities.
+
+    \begin{enumerate}
+    \item First, the $\beta$ rule states that composing the universal
+      morphism with the introduction rule applied to a morphism $q : \harrapp {Q} \alpha {\beta}$ is
+      equal to the original $q$:
+
+      \[ q = \homrappXtoYatZ {\homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {Q_t} \alpha} {(\homrappXtoYatZ {\pendappXtoY I \alpha} {q} \beta)} {G_0 \beta}} {\pendappXtoY \epsilon \beta} {\beta}} {\pendappXtoY {D_r} \beta} {\beta} \]
+
+    \item Second, the $\eta$ rule says that every morphism $f :
+      \harrapp {C_1} \alpha {G_0 \beta}$ can be reconstructed using
+      $I$ after postcomposing with $\epsilon$:
+
+      \[ f = \homrappXtoYatZ {\pendappXtoY I \alpha} {({\homrappXtoYatZ {\homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {Q_t} \alpha} {f} {G_0 \beta}} {\pendappXtoY \epsilon \beta} {\beta}} {\pendappXtoY {D_r} \beta} {\beta}  })} {\beta}\]
+
+      As a special case, if $f = \pendappXtoY {C_r} {G_0 \beta}$, this shows that $\homrappXtoYatZ {\pendappXtoY I {G_0 \beta}} {\pendappXtoY \epsilon \beta} \beta = \pendappXtoY {C_r} {G_0 \beta}$
+    \end{enumerate}
+  \end{enumerate}
+\end{definition}
+
+These two notions turn out to be equivalent because the ``extra''
+information in the hom isomorphism is in fact fully determined by the
+data of a universal morphism: the functoriality of $G$ and the
+naturality of $I,I_{inv}$.
+%
+The proof can be broken up into 3 conceptual steps:
+\begin{enumerate}
+\item First,, the functoriality of $G$ is derivable from $I,\epsilon$.
+\item Next, $I_{inv}$ is a homomorphism automatically since composition
+  with a morphism is automatically natural, where $I_{inv}$ is defined as
+  \[\pendlambdaXdotY \alpha {\homrlambdaXatYdotZ f {\beta} \homrappXtoYatZ {\homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {Q_t} \alpha} f \beta}{\pendappXtoY \epsilon \beta} {\beta}} {\pendappXtoY {D_r }\beta} {\beta}} \]
+
+  \item Finally, $I$ is a homomorphism because the inverse of a
+    natural transformation is always natural.
+\end{enumerate}
+
+%% Part of this equivalence is the \emph{Yoneda Lemma}, which we state
+%% now in its internal, parameterized form.
+%% The usual non-parameterized Yoneda lemma can be formed if we have a
+%% unit sort.
+
+\begin{lemma}[Universal Constructions are functorial]
+  For any universal construction $(G_0,\epsilon,I)$ for $Q :
+  \text{Prof}\,C\,D$, we can extend $G_0$ to a functor by defining
+
+  \[ G_1 = \pendlambdaXdotY \beta {\homrlambdaXatYdotZ f {\beta'}
+    {\homrappXtoYatZ {\pendappXtoY I {G_0\beta}} {(\homrappXtoYatZ {\homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {Q_t} {G_0\beta}} {\pendappXtoY {C_r} {G_0\beta}} {G_0\beta}} {\pendappXtoY \epsilon {\beta}} \beta} {f} {\beta'})} {\beta'}}
+  }\]
+
+  Furthermore, $I,I^-1$ are homomorphisms with respect to the induced
+  profunctor structure.
+\end{lemma}
+\begin{proof}
+  \begin{enumerate}
+  \item To prove $G_1$ preserves identity we need to show
+    \[ {\homrappXtoYatZ {\pendappXtoY I {G_0\beta}} {(\homrappXtoYatZ {\homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {Q_t} {G_0\beta}} {\pendappXtoY {C_r} {G_0\beta}} {G_0\beta}} {\pendappXtoY \epsilon {\beta}} \beta} {\pendappXtoY {D_r} {\beta}} {\beta})} {\beta}} = \pendappXtoY{C_r} {G_0\beta}\]
+    which is precisely the $\eta$ principle.
+
+  \item Next, we show $I_{inv}$ is a homomorphism from $\text{Hom}(C)[\text{Id}(C);G]$ to $Q$.
+    All of this is under the binders
+    \[\lambdatrinary {\alpha : C} {f : \harrapp {C_s} \alpha {\alpha'}}
+    {\alpha' : C} {g : \harrapp{C_s} {\alpha'} {G_0 \beta'}}
+    {\beta' : D} {h : \harrapp {D_s} {\beta'} {\beta}}
+    {}\]
+
+    % given
+    % f : a -> a' in C
+    % g : a' -> G b'
+    % h : b' -> b
+
+    %% I_{inv} (f ;C (g ;C G_1 h)) = Q_t f (I_{inv} g) h
+
+    \begin{align*}
+      & \appunary
+      {I_{inv}}
+      \alpha
+      {(\appbinary {C_t} \alpha f {\alpha'} {(
+          \appbinary {C_t} {\alpha'} g {G_0\beta'}
+          {(\appunary {G_1} {\beta'} h \beta)} {G_0 \beta})} {G_0 \beta})}
+      {\beta}\\
+      &=
+      \apptrinary {Q_t} \alpha
+      {(\appbinary {C_t} \alpha f {\alpha'} {(
+          \appbinary {C_t} {\alpha'} g {G_0\beta'}
+          {(\appunary {G_1} {\beta'} h \beta)} {G_0 \beta})} {G_0 \beta})} \beta
+      {\pendappXtoY \epsilon \beta} \beta
+      {\pendappXtoY {D_r} \beta} {\beta}\\
+      &=
+      \apptrinary {Q_t} \alpha
+      {(\appbinary {C_t} \alpha f {\alpha'} {(
+          \appbinary {C_t} {\alpha'} g {G_0\beta'}
+          {(\appunary {G_1} {\beta'} h \beta)} {G_0 \beta})} {G_0 \beta})} \beta
+      {\pendappXtoY \epsilon \beta} \beta
+      {(\appbinary {D_t} \beta {\pendappXtoY {D_r} \beta} \beta {\pendappXtoY {D_r} \beta} \beta)} {\beta}\\
+      &=
+      \apptrinary {Q_t} \alpha
+      {(\appbinary {C_t} \alpha {(\appbinary {C_t} \alpha f {\alpha'} g {G_0 \beta'})} {G_0 \beta'}
+        {(\appunary {G_1} {\beta'} h \beta)} {G_0 \beta})} \beta
+      {\pendappXtoY \epsilon \beta} \beta
+      {(\appbinary {D_t} \beta {\pendappXtoY {D_r} \beta} \beta {\pendappXtoY {D_r} \beta} \beta)} {\beta}\\
+      &=
+      \apptrinary {Q_t} \alpha
+      {(\appbinary {C_t} \alpha f {\alpha'} g {G_0 \beta'})} {G_0 \beta'}
+      {(\apptrinary {Q_t} {G_0 \beta'}
+        {(\appunary {G_1} {\beta'} h {\beta})} {G_0 \beta}
+        {\pendappXtoY \epsilon \beta} \beta
+        {\pendappXtoY {D_r} \beta} \beta)} \beta
+      {\pendappXtoY {D_r} \beta} \beta
+      \\
+      &=
+      \apptrinary {Q_t} \alpha
+      {(\appbinary {C_t} \alpha f {\alpha'} g {G_0 \beta'})} {G_0 \beta'}
+      {(\appunary {I_{inv}} {G_0 \beta'} {(\appunary {G_1} {\beta'} h \beta)} \beta)} \beta
+      {\pendappXtoY {D_r} \beta} \beta
+      \\
+      &=
+      \apptrinary {Q_t} \alpha
+      {(\appbinary {C_t} \alpha f {\alpha'} g {G_0 \beta'})} {G_0 \beta'}
+      {(\appunary {I_{inv}} {G_0 \beta'} {(\appunary {I} {G_0\beta'} {(\apptrinary {Q_t} {G_0 \beta'} {\pendappXtoY {C_r} {G_0 \beta'}} {G_0 \beta'} {\pendappXtoY \epsilon {\beta'}} {\beta'} h \beta)} \beta)} \beta)} \beta
+      {\pendappXtoY {D_r} \beta} \beta
+      \\
+      &=
+      \apptrinary {Q_t} \alpha
+      {(\appbinary {C_t} \alpha f {\alpha'} g {G_0 \beta'})} {G_0 \beta'}
+      {(\apptrinary {Q_t} {G_0 \beta'} {\pendappXtoY {C_r} {G_0 \beta'}} {G_0 \beta'} {\pendappXtoY \epsilon {\beta'}} {\beta'} h \beta)} \beta
+      {\pendappXtoY {D_r} \beta} \beta
+      \\
+      &=
+      \apptrinary {Q_t} \alpha
+      {(\appbinary {C_t} \alpha {(\appbinary {C_t} \alpha f {\alpha'} g {G_0 \beta'})} {G_0 \beta'}
+        {\pendappXtoY {C_r} {G_0 \beta'}} {G_0 \beta'})} {G_0 \beta'}
+      {\pendappXtoY \epsilon {\beta'}} {\beta'}
+      {(\appbinary {D_t} {\beta'} h \beta {\pendappXtoY {D_r} \beta} \beta)} \beta\\
+      &=
+      \apptrinary {Q_t} \alpha
+      {(\appbinary {C_t} \alpha f {\alpha'} g {G_0 \beta'})} {G_0 \beta'}
+      {\pendappXtoY \epsilon {\beta'}} {\beta'}
+      {(\appbinary {D_t} {\beta'} {\pendappXtoY {D_r} {\beta'}} {\beta'} h \beta )} \beta\\
+      &=
+      \apptrinary {Q_t} \alpha
+      f {\alpha'}
+      {(\apptrinary {Q_t} {\alpha'} g {G_0 \beta'} {\pendappXtoY \epsilon {\beta'}} {\beta'}{\pendappXtoY {D_r} {\beta'}} {\beta'})} {\beta'}
+      h \beta\\
+      &= \homrappXtoYatZ {\homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {Q_t} \alpha} f {\alpha'}} {(\homrappXtoYatZ {\pendappXtoY {I_{inv}} {\alpha'}} {g} {\beta'})} {\beta'}} {h} {\beta}\\
+    \end{align*}
+
+  \item To prove $G_1$ preserves composition given $f : \harrapp {D_1} {\beta} {\beta'}$, $f' : \harrapp {D_1} {\beta'} {\beta''}$:
+    \begin{align*}
+      & \appunary {G_1} {\beta} {\appbinary {D_t} {\beta} f {\beta'} {f'} {\beta''}} {\beta'}\\
+      &= \appunary {I} {G_0 \beta} {(\apptrinary {Q_t} {G_0 \beta} {\pendappXtoY {C_r} {G_0 \beta}} {G_0 \beta} {\pendappXtoY \epsilon {\beta}} {\beta} {({\appbinary {D_t} {\beta} f {\beta'} {f'} {\beta''}})} {\beta''})} {\beta''}\\
+      &= \appunary {I} {G_0 \beta} {(\apptrinary {Q_t} {G_0 \beta} {(\appbinary {C_t} {G_0 \beta} {\pendappXtoY {C_r} {G_0 \beta}} {G_0 \beta} {\pendappXtoY {C_r} {G_0 \beta}} {G_0 \beta})} {G_0 \beta} {\pendappXtoY \epsilon {\beta}} {\beta} {({\appbinary {D_t} {\beta} f {\beta'} {f'} {\beta''}})} {\beta''})} {\beta''}\\
+      &=
+      \appunary {I} {G_0 \beta}
+      {(\apptrinary {Q_t} {G_0 \beta}
+        {\pendappXtoY {C_r} {G_0 \beta}} {G_0 \beta}
+        {(\apptrinary {Q_t} {G_0 \beta}
+          {\pendappXtoY {C_r} {G_0 \beta}} {G_0 \beta}
+          {\pendappXtoY \epsilon {\beta}} {\beta}
+          f {\beta'})} {\beta'}
+        {f'} {\beta''}
+        )}
+      {\beta''}\\
+      &=
+      \appbinary {C_t} {G_0 \beta}
+      {\pendappXtoY {C_r} {G_0 \beta}} {G_0 \beta}
+      {(\appbinary {C_t} {G_0 \beta}
+        {(\appunary {I} {G_0 \beta}
+        {(\apptrinary {Q_t} {G_0 \beta}
+          {\pendappXtoY {C_r} {G_0 \beta}} {G_0 \beta}
+          {\pendappXtoY \epsilon {\beta}} {\beta}
+          f {\beta'})} {\beta'})} {G_0 \beta'}
+        {(\appunary {G_1} {\beta'} {f'} {\beta''})} {G_0 \beta''})}
+      {G_0 \beta''}\\
+      &=
+      \appbinary {C_t} {G_0 \beta}
+        {(\appbinary {C_t} {G_0 \beta}
+          {\pendappXtoY {C_r} {G_0 \beta}} {G_0 \beta}
+          {(\appbinary {C_t} {G_0 \beta}
+            {(\appunary {I} {G_0 \beta} {\pendappXtoY \epsilon {\beta}} {\beta})} {G_0 \beta}
+            {(\appunary {G_1} \beta f {\beta'})} {G_0 \beta'})} {G_0 \beta'})} {G_0 \beta'}
+        {(\appunary {G_1} {\beta'} {f'} {\beta''})} {G_0 \beta''}\\
+      &=
+      \appbinary {C_t} {G_0 \beta}
+        {(\appbinary {C_t} {G_0 \beta}
+            {(\appunary {I} {G_0 \beta} {\pendappXtoY \epsilon {\beta}} {\beta})} {G_0 \beta}
+            {(\appunary {G_1} \beta f {\beta'})} {G_0 \beta'})} {G_0 \beta'}
+        {(\appunary {G_1} {\beta'} {f'} {\beta''})} {G_0 \beta''}\\
+      &=
+      \appbinary {C_t} {G_0 \beta}
+        {(\appbinary {C_t} {G_0 \beta}
+            {\pendappXtoY {C_r} {G_0 \beta}} {G_0 \beta}
+            {(\appunary {G_1} \beta f {\beta'})} {G_0 \beta'})} {G_0 \beta'}
+        {(\appunary {G_1} {\beta'} {f'} {\beta''})} {G_0 \beta''}\\
+      &=
+      \appbinary {C_t} {G_0 \beta}
+      {(\appunary {G_1} \beta f {\beta'})} {G_0 \beta'}
+        {(\appunary {G_1} {\beta'} {f'} {\beta''})} {G_0 \beta''}\\
+      %% &\homrappXtoYatZ {\pendappXtoY {G_1} \beta} {(\homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {D_t} {\beta}} {f} {\beta'}} {f'} {\beta''})} {\beta''}\\
+      %% &={\homrappXtoYatZ {\pendappXtoY I {G_0\beta}} {(\homrappXtoYatZ {\homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {Q_t} {G_0\beta}} {\pendappXtoY {C_r} {G_0\beta}} {G_0\beta}} {\pendappXtoY \epsilon {\beta}} \beta} {(\homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {D_t} {\beta}} {f} {\beta'}} {f'} {\beta''})} {\beta''})} {\beta''}}\\
+      %% &={\homrappXtoYatZ {\pendappXtoY I {G_0\beta}} {(\homrappXtoYatZ {\homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {Q_t} {G_0\beta}} {({\pendappXtoY {C_r} {G_0\beta}})} {G_0\beta}} {\pendappXtoY \epsilon {\beta}} \beta} {(\homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {D_t} {\beta}} {f} {\beta'}} {f'} {\beta''})} {\beta''})} {\beta''}}\\
+    \end{align*}
+    %% \[
+    %%   {\homrappXtoYatZ {\pendappXtoY I {G_0\beta}} {(\homrappXtoYatZ {\homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {Q_t} {G_0\beta}} {\pendappXtoY {C_r} {G_0\beta}} {G_0\beta}} {\pendappXtoY \epsilon {\beta}} \beta} {(\homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {D_t} {\beta}} {f} {\beta'}} {f'} {\beta''})} {\beta''})} {\beta''}}
+    %% \]
+  \end{enumerate}
+\end{proof}
+
+\begin{lemma}[Inverse of homomorphism is a homomorphism]
+  Let $\phi$ be a homomorphism from $R$ to $S : \text{Prof} C D$ and
+  let $\psi : \pendallXdotY \alpha {\homrallXYtoZ \beta {\harrapp S
+      \alpha \beta}{\harrapp R \alpha \beta}}$. If $\psi$ is a
+  two-sided inverse to $\phi_f$ then $\psi$ is a homomorphism.
+\end{lemma}
+\begin{proof}
+  We need to show the homomorphism condition for $\psi$:
+  \[ \homrappXtoYatZ {\pendappXtoY \psi \alpha} {(\homrappXtoYatZ {\homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {S_t} \alpha} f {\alpha'}} s {\beta'}} g {\beta})} {\beta}
+    = \homrappXtoYatZ {\homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {R_t} \alpha} f {\alpha'}} {(\homrappXtoYatZ {\pendappXtoY \psi {\alpha'}} s {\beta'})} {\beta'}} g \beta
+    \]
+
+    \begin{align*}
+      \homrappXtoYatZ {\pendappXtoY \psi \alpha} {(\homrappXtoYatZ {\homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {S_t} \alpha} f {\alpha'}} s {\beta'}} g {\beta})} {\beta}
+      &= \homrappXtoYatZ {\pendappXtoY \psi \alpha} {(\homrappXtoYatZ {\homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {S_t} \alpha} f {\alpha'}} {(\homrappXtoYatZ {\pendappXtoY {\phi_f} {\alpha'}} {(\homrappXtoYatZ {\pendappXtoY \psi {\alpha'}} s {\beta'})} {\beta'})} {\beta'}} g {\beta})} {\beta}\\
+      &= \homrappXtoYatZ {\pendappXtoY \psi \alpha}
+      {(\homrappXtoYatZ {\pendappXtoY {\phi_f} \alpha} {(\homrappXtoYatZ {\homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {R_t} \alpha} f {\alpha'}} {(\homrappXtoYatZ {\pendappXtoY \psi {\alpha'}} s {\beta'})} {\beta'}} g \beta)} \beta)} {\beta}\\
+      & = \homrappXtoYatZ {\homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {R_t} \alpha} f {\alpha'}} {(\homrappXtoYatZ {\pendappXtoY \psi {\alpha'}} s {\beta'})} {\beta'}} g \beta
+    \end{align*}
+\end{proof}
+
+\subsection{Self-Interpretation: Hom of Internal Profunctors}
+
+We can define a version of the hom $\homrallXYtoZ \alpha P Q$ on
+\emph{internal} profunctors using a combination of this connective,
+and an \emph{equalizer} to enforce a homomorphism condition.
+
+\begin{definition}[Internal Covariant Hom of Profunctors]
+  Let $P : \textrm{Prof}\,D\,E$ and $Q : \textrm{Prof}\,C\,E$. Define
+  the covariant Hom $P \triangleright Q : \textrm{Prof}\,C\,D$ by
+  \[ (P \triangleright Q)_1 = \lambda (\alpha;\beta).
+\equalizeVofWbyXeqYatZ \phi {\homrallXYtoZ \gamma {\harrapp {P_1} \alpha \gamma} {\harrapp {Q_1} \beta \gamma}} {\psi_{dom}} {\psi_{cod}} {\homrallXYtoZ {\gamma'} {\harrapp {P_1 }\beta {\gamma'}} {\homrallXYtoZ {\gamma} {\harrapp {E_1} {\gamma'} \gamma} {\harrapp {Q_1} \alpha \gamma}}}\]
+  where $\psi_{dom}$ and $\chi_{cod}$ are  as
+  \[ \psi_{dom} =\homrlambdaXatYdotZ p {\gamma'} {\homrlambdaXatYdotZ e \gamma
+    {\homrappXtoYatZ {\phi} {(\apptrinary {P_t} \beta {\pendappXtoY {D_r} \beta} \beta p {\gamma'} e \gamma)} \gamma}}
+  \]
+  \[
+  \psi_{cod} = \homrlambdaXatYdotZ p {\gamma'} {\homrlambdaXatYdotZ e \gamma
+    {\apptrinary {Q_t} \alpha {\pendappXtoY {C_r} \alpha} \alpha {(\homrappXtoYatZ \phi {p} {\gamma'})} {\gamma'} e \gamma}}
+  \]
+\end{definition}
+
+We can easily define a contravariant hom $P \triangleleft Q$ by a dual
+definition (TODO).
+
+\subsection{Limits and Kan Extensions}
+
+We can then use our notion of Hom of profunctors
+
+The right Kan extension of $F$ along $J$ is the best (maximal) way to
+fill in the following square:
+\[% https://tikzcd.yichuanshen.de/#N4Igdg9gJgpgziAXAbVABwnAlgFyxMJZABgBpiBdUkANwEMAbAVxiRAGEQBfU9TXfIRQBGclVqMWbACLdeIDNjwEiZYePrNWiDgHJu4mFADm8IqABmAJwgBbJGRA4ISAEzVNUnQCk5lm-aIjs5IohJabABiBlxAA
+\begin{tikzcd}
+C \arrow[d, "J"] \arrow[r, "F"] & D \\
+C'                              &  
+\end{tikzcd}\]
+I.e. a functor $Ran_J F : C' \to D$ with a natural transformation
+$Ran_J F \Rightarrow F$ that is universal something something.
+
+If $D$ is complete, there is a formula using ends for this. Using hom
+of profunctors, we can directly adapt this to define a
+\emph{profunctor} that is a universal filler. We can then say that a
+right Kan extension is precisely a functor that represents this
+profunctor.
+
+\begin{definition}
+  Let $C,C',D$ be categories and let $F : \textrm{Functor}\,C\,D$ and
+  $J: \textrm{Functor}\,C'\,C$ be functors. Then the right Kan
+  extension specification profunctor is a profunctor
+  $\textrm{RanSpec}_J\,F : \textrm{Prof}\,D\,C$ defined as
+
+  \[ \textrm{RanSpec}_J\,F = \textrm{Hom}(C')[;J] \triangleright \textrm{Hom}(D)[;F] \]
+
+  And a right Kan extension is a functor
+  $\textrm{Ran}_JF:\textrm{Functor}\,C'\,D$ that is a
+  right-representation of $\textrm{RanSpec}_J F$.
+\end{definition}
+
+Writing this out more pointfully, we get that the hom set
+$\harrapp{D_1} {\beta}{\textrm{Ran}_JF\, \alpha}$ is an equalizer of
+the following spans:
+\[ \homrallXYtoZ {\alpha':C'} {\harrapp {C_1'} {\alpha'}{J_0\alpha}}{\harrapp {D_1} {\beta} {F_0 \alpha'}}\]
+
+If we have a unit and therefore unit category and functor, then we can
+now define limits as a special case of Kan extensions.
+\begin{definition}
+  The limit of a functor $J : \textrm{Functor}I\,C$ is a right Kan
+  extension $\textrm{Ran}_E J$ where $E$ is the unique functor $E :
+  \textrm{Functor}I\,1$.
+\end{definition}
+
+Unraveling this definition, we get that $\harrapp{C_1} \alpha {\lim_{i} J(i)}$ is a coherent family (i.e., equalizer) of
+\[  \homrallXYtoZ {i:I} {1} {\harrapp{C_1} \alpha {J_0(i)}}  \]
+with the equalizing condition being that for any $p : \homrallXYtoZ
+{i:I} {1} {\harrapp{C_1} \alpha {J_0(i)}}$ and $f : \harrapp{I_1} {i'}{i}$ and $g : \harrapp{1}{pt}{E_0 i'}$
+\[ \homrappXtoYatZ {p} {(\appbinary {1_t} {pt} {\pendappXtoY {1_r} {pt}} {pt} {(\appbinary {1_t} {pt} g {E_0 i'} {(\appunary {E_1} {i'} f {i})} {i})} {i})} {i}
+= \appbinary {C_t} {\alpha} {(\homrappXtoYatZ {p} g {i'})} {J_0 i'} {(\appbinary {C_t} {J_0 i'} {(\appunary {J_1} i f {i'})} {J_0 i} {\pendappXtoY {C_r} {J_0 i}} {J_0 i})} {J_0 i}
+\]
+We can greatly simplify using the $\eta$ principle for $1$ and
+eliminating some reflexivities to reduce this to
+\[ \homrappXtoYatZ {p} {()} {i} =
+\appbinary {C_t} {\alpha} {(\homrappXtoYatZ {p} {()} {i'})} {J_0 i'} {{(\appunary {J_1} i f {i'})}} {J_0 i}
+\]
+which reads exactly as the commuting condition for a cone.
+
+We can similarly define \emph{left} Kan extensions by a dual
+formula. Left Kan extensions are the ``minimal'' solution to the same
+filling problem as right Kan extensions. I.e., they have a universal
+morphism out of them. They can be described by a coend formula, but in
+our profunctor calculus, they are easiest to describe using a similar
+formula to right Kan extensions, but using the contravariant hom of
+profunctors.
+
+\begin{definition}
+  Let $C,C',D$ be categories and let $F : \textrm{Functor}\,C\,D$ and
+  $J: \textrm{Functor}\,C'\,C$ be functors. Then the left Kan
+  extension specification profunctor is a profunctor
+  $\textrm{LanSpec}_J\,F : \textrm{Prof}\,C\,D$ defined as
+
+  \[ \textrm{RanSpec}_J\,F = \textrm{Hom}(D)[F;] \triangleleft \textrm{Hom}(C')[J;]  \]
+  And a left Kan extension is a functor $\textrm{Lan}_J F :
+  \textrm{Functor}\,C'\,D$ that represents this profunctor on the
+  left.
+\end{definition}
+
+Unraveling the definition we get the hom span $\harrapp {D_1}{\textrm{Ran}_JF {\alpha'}} \beta$ is an equalizer of
+\[ \homlallXYtoZ \alpha {\harrapp {C_1'} {J_0 \alpha} {\alpha'}} {\harrapp{D_1} {F_0 \alpha} \beta}\]
+
+\begin{definition}
+  The colimit of a functor $J : \textrm{Functor}I\,C$ is a left Kan
+  extension $\textrm{Lan}_E J$ where $E$ is the unique functor $E :
+  \textrm{Functor}I\,1$.
+\end{definition}
+
+We can similarly unravel this definition to see that this looks
+essentially like the classical definition of colimits as universal
+cocones.
+
+%% We can also define a left Kan extensions using the contravariant
+%% cotensors.  Normally the left Kan extension is defined using tensors,
+%% but the definition using profunctors doesn't look right to me. Let's
+%% work backwards from the definition of colimit to see why.
+
+%% Let $\alpha : \cat C \vdash A : \cat D$ be a functor we want the
+%% colimit of. The UP of a colimit is that maps out of it are equivalent
+%% to cocones from $A$. Semi-formally this looks like
+%% \[ (\colim_{\alpha} A \to_{\cat D} \beta) \cong \forall \alpha:\cat C.~ A \to_{\cat D} \beta \]
+%% This looks a lot like a cotensor with a trivial domain:
+%% \[ (\colim_{\alpha} A \to_{\cat D} \beta) \cong \forall \alpha. (() \to_{1} \gamma) \triangleright (A \to_{\cat D} \beta) \]
+%% Suggesting that we define the left Kan extension profunctor as a
+%% contravariant cotensor as follows.
+%% \begin{definition}
+%%   Let $\cat C,\cat C',\cat D$ be cats and let $\alpha:\cat C \vdash A
+%%   : \cat D$ and $\alpha : \cat C \vdash J : \cat C'$ be functors.
+%%   Then the left Kan extension profunctor $\gamma^o:\cat C'; \beta:\cat
+%%   D \pipe (PLan_J A)[\gamma;\beta]$ is defined as
+%%   \[
+%%   \gamma^o:\cat C'; \beta:\cat D \pipe (PLan_J A)[\gamma;\beta]
+%%   := \forall \alpha : \cat C.~ (J \to_{\cat C'} \gamma) \triangleright (A \to_{\cat D} \beta)
+%%   \]
+%%   Then a left Kan extension of $A$ along $J$ is a functor $\gamma :
+%%   \cat C' \vdash Lan_J A : \cat D$ that represents $PLan_J A$ in that
+%%   \[ (Lan_J A \to_{\cat D} \beta) \cong (LPan_J A)[\gamma;\beta] \]
+%% \end{definition}
+
+%% \begin{definition}
+  %% Let $\cat C,\cat C',\cat D$ be cats and let $\alpha:\cat C \vdash
+  %% : \cat D$ and $\alpha : \cat C \vdash J : \cat C'$ be functors.
+  %% Then the left Kan extension profunctor $\beta^o:\cat D; \gamma:\cat
+  %% C' \pipe (PLan_J F)[\beta;\gamma]$ is defined as
+%%   \[
+%%   \beta^o:\cat D; \gamma:\cat C' \pipe (PLan_J A)[\beta;\gamma] :=
+%%   \exists \alpha:\cat C. (\beta \to_{\cat D} A[\alpha]) \odot (J[\alpha] \to_{\cat C'} \gamma)
+%%   \]
+%%   And a left Kan extension is a functor $\gamma :\cat C' \vdash Lan_J
+%%   A : \cat D$ that represents $PLan_J A$ in that
+%%   \[ (\beta \to_{\cat D} Lan_J A[\alpha]) \cong (PLan_J A)[\beta;\alpha] \]
+%% \end{definition}
+
+%% \begin{lemma}[Parameterized Yoneda Lemma]
+%%   Given categories $\cat C, \cat D$ and a profunctor $\alpha^o:\cat
+%%   C;\beta : \cat D \vdash \alpha \to_{Q} \beta$ and a function
+%%   $\beta : \cat D \vdash G[\beta] : \cat C$, then there is an
+%%   isomorphism between
+%%   \begin{enumerate}
+%%   \item Terms $\cdot \vdash t : G[\beta] \to_{Q} \beta$
+%%   \item left-homomorphisms $\phi : \alpha \to_{\cat C} G[\beta] \vdash
+%%     u : \alpha \to_{Q} \beta$ i.e. they satisfy homomorphism on the
+%%     left (the other side doesn't make sense if $G$ is not a functor):
+%%     $u[\phi \circ \psi] = u[\phi] \circ \psi$
+%%   \end{enumerate}
+%% \end{lemma}
+%% \begin{proof}
+%%   First, the constructions
+%%   \begin{enumerate}
+%%   \item Given $t$, we define $u(t)$ to be
+%%     \[\phi : \alpha \to_{\cat C} G[\beta] \vdash t \circ_{Q} \phi : \alpha \to_{Q} \beta\]
+
+%%     Which is a left-homomorphism because:
+%%     \[ (t \circ_{Q} \phi) \circ_{Q} \psi = t \circ_{Q} (\phi \circ_{\cat C} \psi)\]
+%%   \item Given $u$, we define $t(u)$ to be
+%%     \[ \cdot \vdash u[\id_{G[\beta]}/\phi] : G[\beta] \to_{Q} \beta \]
+%%   \end{enumerate}
+%%   Now we show they are inverse
+%%   \begin{enumerate}
+%%   \item First, the very easy case, $\cdot \vdash t(u(t)) : G[\beta] \to_{Q} \beta$:
+%%     \begin{align*}
+%%       t \circ_{Q} \id_{G[\beta]} &= t
+%%     \end{align*}
+%%     by the fact that $Q$ is a profunctor.
+%%   \item Next, $\phi : \alpha \to_{\cat C} G[\beta] \vdash u(t(u)) : \alpha \to_{Q} \beta$:
+%%     \begin{align*}
+%%       u[\id_{G[\beta]}/\phi] \circ_{Q} \phi &= u[\id \circ_{\cat C} \phi/\phi]\\
+%%       &= u[\phi/\phi]\\
+%%       &= u
+%%     \end{align*}
+%%     which uses the fact that $u$ is a left-homomorphism.
+%%   \end{enumerate}
+%% \end{proof}
+
+%% \begin{theorem}[Negative Types and Right Representables are Equivalent]
+%%   For any profunctor $\alpha^o: \cat C; \beta : \cat D \pipe \alpha
+%%   \to_{Q} \beta$, there is a bijection between
+%%   \begin{enumerate}
+%%   \item Right Represntables $\beta : \cat D \vdash G[\beta] : \cat C$ for $Q$.
+%%   \item Negative Types $\beta : \cat D \vdash N[\beta] : \cat C$
+%%     representing $Q$.
+%%   \end{enumerate}
+%% \end{theorem}
+%% \begin{proof}
+%%   \begin{enumerate}
+%%   \item Given a right representable $G[\beta]$, we set $N(G)[\beta] =
+%%     G[\beta]$, the introduction rule is the same and the elimination
+%%     rule is given as in the Yoneda lemma and the inverse property
+%%     follows by the Yoneda isomorphism and the fact that $I_G^{-1}$ is
+%%     an inverse for $I_G$.
+%%   \item Given a negative type $N[\beta]$ we set $G(N)[\beta] =
+%%     N[\beta]$, the introduction rule is the same and the inverse
+%%     introduction rule is given as in the Yoneda lemma.  We need to
+%%     show that $N[\beta]$ has the structure of a functor and that
+%%     $I_{N}$ and $I_{N}^{-1}$ are homomorphisms on the right with
+%%     respect to that structure.  Define $N_1$ as follows:
+%%     \begin{mathpar}
+%%       \inferrule
+%%           {\beta_1^o:\cat D;\beta_2:\cat D \vdash \psi \circ \epsilon : N[\beta_1/\beta] \to_{Q} \beta_2\\
+%%             \phi : N[\beta_1/\beta] \to_{Q} \beta_2 \vdash I_N[\phi] : N[\beta_1/\beta] \to_{\cat C} N[\beta_2/\beta]
+%%           }
+%%           {\beta_1^o:\cat D;\beta_2:\cat D\pipe \psi : \beta_1 \to_{\cat D} \beta_2 \vdash I_N[\psi \circ \epsilon_{\beta_1}/\phi] : N[\beta_1/\beta] \to_{\cat C} N[\beta_2/\beta]}
+%%     \end{mathpar}
+%%     We need to show functoriality.
+%%     For identity we have
+%%     \begin{align*}
+%%       I_N[\id_{\beta} \circ \epsilon_{\beta}/\phi] &= I_N[\epsilon_{\beta}]\\
+%%       &= I_N[\epsilon \circ \id_{N[\beta]}]\\
+%%       &= \id_{N[\beta]} \tag{$\eta$}
+%%     \end{align*}
+%%     For composition,
+%%     \begin{align*}
+%%       N_1[\theta] \circ N_1[\psi]
+%%       &= I_N[\theta \circ \epsilon] \circ I_N[\psi \circ \epsilon]\\
+%%       &= I_N[(\theta \circ \epsilon) \circ I_N[\psi \circ \epsilon]]\tag{left homomorphism}\\
+%%       &= I_N[\theta \circ (\epsilon \circ I_N[\psi \circ \epsilon])]\\
+%%       &= I_N[\theta \circ (\psi \circ \epsilon)] \tag{$\beta$}\\
+%%       &= I_N[(\theta \circ \psi) \circ \epsilon]\\
+%%       &= N_1[\theta \circ \psi]
+%%     \end{align*}
+%%     Next we show $I_N$ is a right homomorphism, the typing here is
+%%     $\phi : \alpha \to_{Q} \beta_1, \theta : \beta_1 \to_{\cat D}
+%%     \beta_2$.
+%%     \begin{align*}
+%%       N_1[\theta] \circ I_N[\phi] &= I_N[\theta \circ \epsilon] \circ I_N[\phi]\\
+%%       &= I_N[\theta \circ \epsilon \circ I_N[\phi]]\tag{left homomorphism}\\
+%%       &= I_N[\theta \circ \phi] \tag{$\beta$}
+%%     \end{align*}
+%%     and to show $I_N^{-1}$ is a right homomorphism under $\phi : \alpha \to_{\cat C}N[\beta_1], \theta : \beta_1 \to_{\cat D} \beta_2$
+%%     \begin{align*}
+%%       I_N^{-1}[N_1[\theta] \circ \phi] &= \epsilon \circ (N_1[\theta] \circ \phi) \tag{def}\\
+%%       &= \epsilon \circ (I_N[\theta \circ \epsilon] \circ \phi)\tag{def}\\
+%%       &= (\epsilon \circ I_N[\theta \circ \epsilon]) \circ \phi\tag{assoc}\\
+%%       &= (\theta \circ \epsilon) \circ \phi\tag{$\beta$}\\
+%%       &= \theta \circ (\epsilon \circ \phi)\tag{assoc}\\
+%%       &= \theta \circ I_N^{-1}[\phi] \tag{def}
+%%     \end{align*}
+%%   \end{enumerate}
+
+%%   Next we need to show that this is a \emph{bijection}.
+%%   \begin{enumerate}
+%%   \item Starting with a negative type $N[\beta]$, we recover the
+%%     original $\epsilon$ by the Yoneda lemma.
+%%   \item Starting with a right representable $G[\beta]$, we recover the
+%%     original $I_N^{-1}$ by the Yoneda lemma. We need to show that the
+%%     action of $G$ on arrows is the same as the one defined from $G$
+%%     viewed as a negative type, which we will call $G_1'$:
+%%     \begin{align*}
+%%       G_1'[\theta] &= I_G[\theta \circ \epsilon]\\
+%%       &= G_1[\theta] \circ I_G[\epsilon]\tag{right homomorphism}\\
+%%       &= G_1[\theta] \circ I_G[\epsilon \circ \id]\\
+%%       &= G_1[\theta] \circ \id\\
+%%       &= G_1[\theta]
+%%     \end{align*}
+%%   \end{enumerate}
+%% \end{proof}
+
+%% \subsection{Adjunctions by Universal Morphism}
+
+%% As an example of formal category theory in our syntax, let's prove the
+%% equivalence between two definitions of an adjunction: one by universal
+%% morphism and the other by natural hom isomorphism.
+%% %
+%% Though the definitions and proof will look like we are doing ordinary
+%% category theory, since it is syntactic it immediately applies to all
+%% models, including enriched and internal category theory.
+
+%% We start with the usual definition of an adjunction using
+%% hom-isomorphism and a profunctor.
+%% \begin{definition}
+%%   Let $C,D$ be categories and $c^o:C_0,d:D_0 \vdash H[c,d]$ a
+%%   profunctor between them. A \emph{left adjoint} to $H$ consists of
+%%   \begin{enumerate}
+%%   \item A functor $c:C \vdash F[c] : D$
+%%   \item A natural isomorphism $c,d \pipe h : H[c,d] \vdash
+%%     F_{ind}[h]: D_1[F_0[c],d]$. Call the inverse $g. F_{ind}^{-1}[g]$.
+%%   \end{enumerate}
+%% \end{definition}
+
+%% \begin{definition}
+%%   Let $c:C_0,d:D_0 \vdash H[c,d]$ be a profunctor.
+%%   A \emph{universal $H$-heteromorphism from $C$} consists of
+%%   \begin{enumerate}
+%%   \item For every $c:C_0$, an $H$-morphism $\eta_c : H(c,F_0[c])$ for
+%%     some function $c:C_0 \vdash F_0[c] : D_0$.
+%%   \item Such that every $h : H[c,d]$ factors through $\eta_c$ in that
+%%     there is a unique $h : H[c,d] \vdash F_{ind}[h] : D[F_0[c],d]$
+%%     such that $h = F_{ind}[h] \circ \eta_c$.  Uniqueness means
+%%     validity of the following rule:
+%%     \begin{mathpar}
+%%       \inferrule
+%%       {g:D[F_0[c],d] \vdash g \circ \eta = h}
+%%       {g:D[F_0[c],d] \vdash g = F_{ind}[h]}
+%%     \end{mathpar}
+%%     (Note that we do not require $F_{ind}$ to be natural in any sense)
+%%   \end{enumerate}
+%% \end{definition}
+
+%% It is easy to see that a left adjoint gives a universal heteromorphism
+%% by a Yoneda-like argument:
+%% \begin{construction}
+%%   Let $F$ be a left adjoint to $H$. Then we define a left universal
+%%   $H$-heteromorphism by
+%%   \begin{enumerate}
+%%   \item $c \vdash \eta_c = F_{ind}^{-1}[\id[F_0[c]]]: H[c,F_0[c]]$
+%%   \item Every $h : H[c,d]$ factorizes through $\eta$ because
+%%     \begin{align*}
+%%       F_{ind}[h] \circ F_{ind}^{-1}[\id] &= F_{ind}^{-1}[F_{ind}[h] \circ \id] \tag{naturality}\\
+%%       &= F_{ind}^{-1}[F_{ind}[h]] \tag{unit}\\
+%%       &= h \tag{isomorphism}
+%%     \end{align*}
+%%     This factorization is unique because if $g \circ F^{-1}[\id] = h$,
+%%     then
+%%     \begin{align*}
+%%       F_{ind}[h] &= F_{ind}[g \circ F^{-1}[\id]] \tag{congruence}\\
+%%       &= g \circ F_{ind}[F^{-1}[\id]] \tag{naturality}\\
+%%       &= g \circ \id \tag{iso}\\
+%%       &= g \tag{unit}
+%%     \end{align*}
+%%   \end{enumerate}
+%% \end{construction}
+
+%% But the utility of the universal morphism definition is that the
+%% universal morphism extracted from a left adjoint is enough to uniquely
+%% reconstruct the rest of the structure, which includes (1)
+%% functoriality of $F$ and (2) naturality of $F_{ind}$.
+%% \begin{construction}
+%%   Let $(F_0,\eta,F_{ind}[h])$ be a left universal $H$-morphism.
+%%   We extend this to a left adjoint as follows.
+%%   \begin{enumerate}
+%%   \item First, $F$ acts on morphisms by
+%%     \begin{mathpar}
+%%       \inferrule
+%%       {
+%%         \inferrule
+%%          {f : C[c,c'] \and \eta_{c'} : H[c',F_0[c']]}
+%%          {f : C[c,c'] \vdash \eta_{c'} \circ f : H[c,F_0[c']]}
+%%       }
+%%       {f : C[c,c'] \vdash F_{ind}[\eta_{c'} \circ f] : D[F_0[c],F_0[c']]}
+%%     \end{mathpar}
+%%     This preserves identity
+%%     \begin{mathpar}
+%%       \inferrule
+%%       {
+%%         \inferrule
+%%         {\id \circ \eta = \eta}
+%%         {\id = F_{ind}[\eta]}
+%%       }
+%%       {\id = F_{ind}[\eta \circ \id]}
+%%     \end{mathpar}
+%%     And composition
+%%     \begin{mathpar}
+%%       \inferrule
+%%       {F_{ind}[\eta \circ f] \circ F_{ind}[\eta \circ g] \circ \eta = \eta \circ f \circ g}
+%%       {F_{ind}[\eta \circ f] \circ F_{ind}[\eta \circ g] = F_{ind}[\eta \circ f \circ g]}
+%%     \end{mathpar}
+%%     Which follows by a couple uses of the factorization:
+%%     \begin{align*}
+%%       F_{ind}[\eta \circ f] \circ F_{ind}[\eta \circ g] \circ \eta
+%%       &= F_{ind}[\eta \circ f] \circ \eta \circ g \\
+%%       &= \eta \circ f \circ g
+%%     \end{align*}
+%%   \item Next, we show that $h : H[c,d] \vdash F_{ind}[h] :
+%%     D[F_0[c],d]$ is a natural transformation. As a first step, we show
+%%     that $\eta_c : H[c,F_0[c]]$ is natural in $c$ with the action of
+%%     $F$ on morphisms we have just defined.
+%%     \[ \eta \circ f = F[f] \circ \eta \]
+%%     expanding the definition, this is precisely the factorization
+%%     property:
+%%     \[ \eta \circ f = F_{ind}[\eta \circ f] \circ \eta \]
+
+%%     Next, to show that $F_{ind}$ is natural, we need to show that
+%%     (when it is well-defined):
+%%     \[F_{ind}[g \circ h \circ f] = g \circ F_{ind}[h] \circ F[f] \]
+%%     By the factorization property, the following sufficeth:
+%%     \begin{align*}
+%%      g \circ F_{ind}[h] \circ F[f] \circ \eta
+%%      &= g \circ F_{ind}[h] \circ \eta \circ f \tag{naturality of $\eta$}\\
+%%      &= g \circ h \circ f \tag{factorization}
+%%     \end{align*}
+%%   \item Finally, we define $F_{ind}^{-1}[g]$ as
+%%     \[ g : D[F_0[c],d] \vdash F_{ind}^{-1}[g] = g \circ \eta : H[c,d] \]
+%%     And we need to confirm that it is (1) a section of $F_{ind}$ (2) a
+%%     retraction of $F_{ind}$ and (3) natural.
+%%     \begin{enumerate}
+%%     \item For retraction we need to show $h \vdash F_{ind}[h] \circ
+%%       \eta = h$ which is precisely the factorization property.
+%%     \item For section, we need to show $g \vdash F_{ind}[g \circ \eta]
+%%       = g$. Immediately the uniquenes principle says its sufficient to
+%%       show $g \circ \eta = g \circ \eta$.
+%%     \item For naturality, we need to show
+%%       \[ F_{ind}^{-1}[k \circ g \circ F[f]] = k \circ F_{ind}^{-1}[g] \circ f \]
+%%       expanding the definition this is just:
+%%       \[ k \circ g \circ F[f] \circ \eta = k \circ g \circ \eta \circ f \]
+%%       which is just naturality of $\eta$.
+%%     \end{enumerate}
+%%   \end{enumerate}
+%% \end{construction}
+
+\subsection{Products and Units of Internal Categories}
+
+%% While we have now proved a beautiful internal theorem in great
+%% generality that has many useful instances in different models, we are
+%% hard pressed to actually \emph{instantiate} the theorem in the syntax
+%% itself because we can't really come up with any useful universal properties!
+%% The reason is that Judgment Theory itself is too bare-bones.
+
+%% For instance how would we say that a category $\cat C$ has a
+%% \emph{terminal object}? Well, it would have to be in the first place a
+%% \emph{single} object in $\cat C$, but so far our syntax only enables
+%% us to talk about functions and functors. Well an object of $\cat C$ is
+%% the same as a function from the unit sort, but so far our syntax only
+%% has base sorts.
+
+%% So we see that defining the terminal object, \emph{necessitates} some
+%% notion of terminal object in our ``metatheory'': Judgment Theory.
+%% This mysterious and pervasive aspect of higher category theory is
+%% called the \emph{microcosm principle} because it says that in order to
+%% discuss a property of a single object in a single category (the
+%% microcosm), we need the same sort of object at the meta-level (the
+%% macrocosm). The adage here is ``As above, so below''.
+
+%% That might all sound very spiritual, but we'll see that it's actually
+%% quite a nice heuristic for formalizing universal properties in
+%% internal category theory.
+%% %
+%% In the spirit of this heuristic, instead of first giving a bunch of
+%% new type constructions for Judgment Theory and then applying them to
+%% define universal properties, we will consider them in pairs, first
+%% above, and then below.
+
+First we consider the terminal sort and then the idea of a terminal
+object in an internal category.
+
+Next, the terminal sort trivially has a category structure using the
+trivial judgment as its hom set:
+
+\begin{definition}[Terminal Category]
+  The terminal category $\cat 1$ is defined as having
+  \begin{enumerate}
+  \item Its object sort as $1$
+  \item Its hom set as $\lambda \alpha:1;\beta:1. 1 : \harr 1 1$
+  \item Its identity arrow is given by $\lambda \alpha: 1. () : 1$
+  \item Composition is defined by $\phi : 1, \psi : 1 \vdash () : 1$
+  \item Associativity and unitality are trivial by \textsc{1-$\eta$}
+  \end{enumerate}
+\end{definition}
+
+\begin{definition}[Terminal Object Specification]
+  In judgment type theory with a terminal sort/judgment, we can define
+  for any category $\cat C$ a profunctor $\alpha^o:\cat C;\beta : \cat
+  1 \pipe 1$, which specifies a terminal object.
+\end{definition}
+
+Now let's instantiate the definition for the terminal object defined
+as a negative type and as a representable and see how it reproduces
+the standard type-theoretic and category-theoretic definitions.
+
+A representation of $\alpha^o:\cat C; \beta : \cat 1 \pipe 1$ consists
+of a functor $\beta : \cat 1 \vdash 1_{\cat C} : \cat C$, an introduction rule
+\[ \phi : 1 \vdash I_{1}[\phi] : \alpha \to_{\cat C} 1_{\cat C}[\beta] \]
+satisfying $I_{1}[\phi] \circ \gamma =I_{1}[\phi \circ \gamma]$ and an
+inverse $\psi : \alpha \to_{\cat C} 1_{\cat C}[\beta] \vdash
+I_{1}^{-1}[\psi] : 1$ which by \textsc{1-$\eta$} is equal to $()$ and
+therefore every $t : \alpha \to_{\cat C} 1_{\cat C}[\beta]$ is equal
+to $I_{1}[\phi]$. In light of this it would be appropriate to write
+$I_1[\phi]$ as $()$ and then the left-homomorphism property is $()
+\circ \gamma = ()$, which is precisely the definition of substitution.
+
+This alternate syntax makes even more sense in light of the negative
+type presentation. A negative type satisfying the terminal object
+specification consists of a functor $1_{\cat C}$ as above and an
+introduction rule as above and an elimination form $\vdash \epsilon :
+1$, which is trivial.
+
+First the product category
+\begin{definition}[Product category]
+  For any categories $\cat C_1$, $\cat C_2$ we can form the product
+  category whose sort of objects is $\cat C_1 \times \cat C_2$, whose
+  arrow judgment is $\alpha^o : \cat C_1 \times \cat C_2;\alpha \pipe
+  (\pi_1 \alpha \to_{\cat C_1} \pi_1 \alpha) \times (\pi_2 \alpha
+  \to_{\cat C_2} \pi_2 \alpha)$. Identity is given by $\alpha^o : \cat
+  C_1 \times \cat C_2;\alpha\pipe\cdot \vdash (\id_{\pi_1 \alpha},
+  \id_{\pi_2 \alpha}) : (\pi_1 \alpha \to_{\cat C_1} \pi_1 \alpha)
+  \times (\pi_2 \alpha \to_{\cat C_2} \pi_2 \alpha)$ and composition
+  by
+  \[
+  \phi %: (\pi_1 \alpha_0 \to_{\cat C_1} \pi_1 \alpha_1) \times (\pi_2 \alpha_0 \to_{\cat C_2} \pi_2 \alpha_1)
+  ,
+  \psi %: (\pi_1 \alpha_1 \to_{\cat C_1} \pi_1 \alpha_2) \times (\pi_2 \alpha_1 \to_{\cat C_2} \pi_2 \alpha_2)
+  \vdash ((\pi_1 \phi) \circ (\pi_1 \psi), (\pi_2 \psi) \circ (\pi_2 \phi)) :
+  (\pi_1 \alpha_0 \to_{\cat C_1} \pi_1 \alpha_2) \times (\pi_2 \alpha_0 \to_{\cat C_2} \pi_2 \alpha_2)  
+  \]
+  unitality, associativity follow from the same properties of $\cat C_1,\cat C_2$
+\end{definition}
+
+\begin{definition}[Product Functor Specification]
+  For any category $\cat C$, we can specify the product functor by
+  \begin{enumerate}
+  \item The span is $\alpha^o:\cat C; \beta : \cat C \times \cat C \pipe (\alpha \to_{\cat C} \pi_1 \beta) \times (\alpha \to_{\cat C} \pi_2 \beta)$
+  \item Left Composition is defined as
+    \[ \phi, \psi \vdash (\pi_1 \phi \circ \psi, \pi_2 \phi \circ \psi)\]
+  \item Left unitality:
+    \begin{align*}
+      (\pi_1 \phi \circ \id, \pi_2 \phi \circ \id)
+      &= (\pi_1 \phi, \pi_2 \phi)\\
+      &= \phi\tag{$\times-eta$}
+    \end{align*}
+  \item Left associativity:
+    \begin{align*}
+      (\phi \circ \psi) \circ \psi'
+      &= (\pi_1 \phi \circ \psi, \pi_2 \phi \circ \psi) \circ \psi'\\
+      &= (\pi_1 (\pi_1 \phi \circ \psi, \pi_2 \phi \circ \psi) \circ \psi',
+      \pi_2 (\pi_1 \phi \circ \psi, \pi_2 \phi \circ \psi) \circ \psi')\\
+      &= ((\pi_1 \phi \circ \psi) \circ \psi', (\pi_2 \phi \circ \psi) \circ \psi')\\
+      &= (\pi_1 \phi \circ (\psi \circ \psi'), \pi_2 \phi \circ (\psi \circ \psi'))\\
+      &= \phi \circ (\psi \circ \psi')
+    \end{align*}
+  \item Right composition is defined as
+    \[ \phi, \theta \vdash (\pi_1 \theta \circ \pi_1 \phi, \pi_2 \theta \circ \pi_2 \phi) \]
+  \item Right unitality:
+    \begin{align*}
+      (\pi_1 (\id,\id) \circ \pi_1 \phi, \pi_2 (\id,\id) \circ \pi_2 \phi)
+      &= (\id \circ \pi_1 \phi, \id \circ \pi_2 \phi) \tag{$\times-\beta$}\\
+      &= (\pi_1 \phi, \id \circ \pi_2 \phi)\\
+      &= \phi \tag{$\times-\eta$}
+    \end{align*}
+  \item Right associativity
+    \begin{align*}
+      \theta' \circ (\theta \circ \phi)
+      &= \theta' \circ (\pi_1 \theta \circ \pi_1 \phi, \pi_2 \theta \circ \pi_2 \phi)\\
+      &= (\pi_1 \theta' \circ \pi_1(\pi_1 \theta \circ \pi_1 \phi, \pi_2 \theta \circ \pi_2 \phi),
+      \pi_2 \theta' \circ \pi_2(\pi_1 \theta \circ \pi_1 \phi, \pi_2 \theta \circ \pi_2 \phi))\\
+      &= (\pi_1 \theta' \circ (\pi_1 \theta \circ \pi_1 \phi), \pi_2 \theta' \circ (\pi_2 \theta \circ \pi_2 \phi))\\
+      &= ((\pi_1 \theta' \circ \pi_1 \theta) \circ \pi_1 \phi, (\pi_2 \theta' \circ \pi_2 \theta) \circ \pi_2 \phi)\\
+      &= (\pi_1(\theta' \circ \theta) \circ \pi_1 \phi, \pi_2(\theta' \circ \theta) \circ \pi_2 \phi)\\
+      &= (\theta' \circ \theta) \circ \phi \\
+    \end{align*}
+  \end{enumerate}
+\end{definition}
+
+\subsection{Discrete and Codiscrete Categories}
+
+If we have terminal judgments, then we can define the
+\emph{codiscrete} category for any sort $\cat C$.
+\begin{definition}
+  The codiscrete category $\textrm{CoDisc} \cat C$ is defined as
+  \begin{enumerate}
+  \item Object sort is $\cat C$
+  \item Morphisms judgment is $1$
+  \item Identity and composition are given by the unique terms of
+    those types, making associativity and unitality trivial to verify.
+  \end{enumerate}
+\end{definition}
+
+If we have \emph{identity} objects, which we write as $\equiv_A$, then we
+can define discrete categories.
+\begin{definition}
+  The discrete category $\textrm{Disc} \cat C$ is defined as
+  \begin{enumerate}
+  \item Object sort is $\cat C$
+  \item Morphisms judgment is $\alpha;\alpha' \pipe \alpha \equiv \alpha'$
+  \item Identity and composition are defined as in the section on
+    synthetic category theory.
+  \end{enumerate}
+\end{definition}
+
+\subsection{Internal Tensor, Cotensor}
+
+We should be able to define internal \emph{tensor} using synthetic
+tensor and coequalizers and also internal \emph{cotensor} using
+synthetic cotensor and equalizers. These look exactly like the
+set-theoretic definitions.
+\begin{definition}
+  Let $R$ be an internal profunctor from $C$ to $D$ and $S$ be an
+  internal profunctor from $D$ to $E$.
+
+  Then we can define the internal tensor product $R \odot S$, a
+  profunctor from $C$ to $E$ as
+  \begin{enumerate}
+  \item The underlying span can be defined as a quotient of the synthetic tensor product
+    \[ (R \odot S)[\alpha;\gamma] = \exists \beta:\cat D. R \odot S/Rcomp = Scomp \]
+
+    Where R-comp and S-comp have as source the span
+    \[ \exists \beta:\cat D.~ R[\alpha;\beta] \odot (\exists \beta':\cat D. Hom_C(\beta,\beta') \odot S[\beta';\gamma]) \]
+    or the isomorphic
+    \[ \exists \beta':\cat D.~ (\exists \beta:\cat D. R[\alpha;\beta] \odot Hom_C(\beta,\beta')) \odot S[\beta';\gamma] \]
+    where R-comp composes the $C$-morphism with the $R$-morphism and
+    S-comp the $S$-morphism.
+  \item Composition is defined using composition in $R,S$.
+  \end{enumerate}
+\end{definition}
+Note that $R$-comp and $S$-comp are each constructed using the
+
+\begin{definition}
+  Let $R$ be an internal profunctor from $C$ to $D$ and $S$ be an
+  internal profunctor from $C$ to $E$. Then the internal contravariant
+  cotensor product $R \triangleright S$ is defined as
+  \[ (R \triangleright S)[\beta;\gamma] = \{ p : \forall \alpha:\cat C.~ R[\alpha;\beta] \triangleright S[\alpha;\gamma] 
+  \pipe R-comp = S-comp \}  \]
+  R-comp and S-comp have as codomain the span
+  \[ \forall \alpha:\cat C.~ R[\alpha;\beta] \triangleright \forall \alpha':\cat C.~ Hom_C(\alpha',\alpha) \triangleright S[\alpha';\gamma] \]
+  or the isomorphic
+  \[ \forall \alpha':\cat C.~ (\exists \alpha. Hom_C(\alpha',\alpha) \odot R[\alpha;\beta]) \triangleright S[\alpha';\gamma]
+  \]
+\end{definition}
+
+We can also construct the \emph{free profunctor} from a span using tensor:
+
+\begin{definition}
+  Let $\cat C$ and $\cat D$ be internal categories and let $R[\alpha;\beta]$ be a
+  span from $Ob(\cat C)$ to $Ob(\cat D)$. Then the free profunctor
+  extending $R$ to be an internal profunctor from $\cat C$ to $\cat D$ is
+  \[ F_{\cat C,\cat D}R[\alpha';\beta'] = \exists \alpha. Hom_C(\alpha',\alpha) \odot (\exists \beta. R[\alpha;\beta] \odot Hom_C(\beta,\beta')) \]
+  or the isomorphic
+  \[ F_{\cat C,\cat D}R[\alpha';\beta'] = \exists \beta. (\exists \alpha. Hom_C(\alpha',\alpha) \odot R[\alpha;\beta]) \odot Hom_C(\beta,\beta') \]
+\end{definition}
+
+And maybe we can even define the co-free profunctor from a span using cotensor?
+\begin{definition}
+  Let $\cat C$ and $\cat D$ be internal categories and let $R[\alpha;\beta]$ be a
+  span from $Ob(\cat C)$ to $Ob(\cat D)$. Then the co-free profunctor
+  extending $R$ to be an internal profunctor from $\cat C$ to $\cat D$ is
+  \[ G_{\cat C,\cat D}R[\alpha';\beta'] =
+  \forall \alpha. Hom_C(\alpha,\alpha') \triangleright (\forall \beta. R[\alpha;\beta] \triangleleft Hom_D(\beta',\beta)) \]
+  or the isomorphic
+  \[ G_{\cat C,\cat D}R[\alpha';\beta'] =
+  \forall \beta. 
+  (\forall \alpha. Hom_C(\alpha,\alpha') \triangleright R[\alpha;\beta]) \triangleleft Hom_D(\beta',\beta)) \]
+\end{definition}
+
 \subsection{Duality}
 
-\section{Semantics in Virtual Equipments}
+\section{Semantics}
+
+A model of our theory is given by a virtual double category with
+restrictions, units, tensors and homs. We informally describe the
+semantics of each judgment now.
+
+A context $\Gamma \isadtctx$ denotes a set.
+%
+A set with empty boundary $\Gamma \pipe \cdot \vdash S \isaSet$ denotes a
+set for every $\gamma:\Gamma$.
+%
+The type $\Gamma \pipe \cdot \vdash \Sort \isaSet$ denotes the set of
+objects of the virtual equipment (for any $\gamma:\Gamma$).
+%
+A term $\Gamma \pipe \alpha:C \vdash A : D$ denotes a vertical arrow
+from $C \gamma$ to $D \gamma$ for every $\gamma:\Gamma$.
+%
+A term $\Gamma \pipe \alpha:C; \beta : D \vdash R \isaSet$ denotes for
+each $\gamma:\Gamma$ a horizontal arrow from $C\gamma$ to $D\gamma$.
+%
+A term $\Gamma \pipe \alpha:C \vdash R \isaSet$ denotes for each
+$\gamma$, a horizontal arrow from $C\gamma$ to itself.
+%
+A non-$\cdot$ context $\Gamma \vdash \Phi\isavectx$ denotes a
+``composable string'' of horizontal arrows from $d^-\Phi \gamma$ to
+$d^+ \Phi \gamma$.
+%
+A term $\Gamma \pipe \cdot \vdash M : S$ denotes a function from
+$\gamma :\Gamma$ to $S \gamma$ (section of the bundle blah blah).
+%
+A term $\Gamma \pipe \Phi \vdash M : R$ for non-empty $\Phi$ denotes a
+``globular'' 2-cell in the virtual equipment for every
+$\gamma:\Gamma$.
+
+Next, substitutions model the forms of composition in the system.
+Substitution $A[B/\alpha]$ denotes composition of vertical arrows
+(with variable being the identity). Substitution $R[A/\alpha;B/\beta]$
+denotes restriction of a horizontal arrow $R$ along the vertical
+arrows $A,B$. And substitution $M[\phi]$ denotes composition of
+2-cells, where some restrictions are of course involved, and again a
+variable is the identity (globular) 2-cell.
+
+Next, we would also like to prove some kind of \emph{completeness}
+theorem for our calculus, at least the part relevant to virtual
+equipments. One way to do this would be to prove a kind of
+parameterized completeness theorem, where we treat the $\Gamma$ like a
+signature, and show that the syntax is complete for every fixed
+$\Gamma$. This amounts to proving that for any fixed $\Gamma$, the
+judgments above present a virtual equipment with a trivial model
+(i.e., each judgment is interpreted as itself). This shows that the
+syntax is closed under all constructions that are present in every
+model.
+
+Here is the outline of how we expect we should be able to prove
+this. For any fixed $\Gamma$, the objects of our double category are
+the terms $\Gamma \vdash C : \Sort$. A vertical arrow from $C$ to $D$
+is a term $\Gamma\pipe \alpha:C \vdash A : D$. A horizontal arrow from
+$C$ to $D$ is a term $\Gamma \pipe \alpha:C;\beta:D \vdash R
+\isaSet$. From any composable string of such $R$s we can construct a
+context $\Gamma \vdash \Phi \isavectx$ and then a 2-cell from $\Phi$ to
+$R$ along $A,B$ is given by a term $\Gamma \pipe \Phi \vdash
+R[A/\alpha;B/\beta]$. Compositions and restriction are modeled by the
+appropriate notion of substitution and units/tensors/homs are modeled
+by the corresponding type formers. There are a lot of details to be
+verified, especially the equations.
 

### 2022-02-05 paper/lipics-v2021-sample-article.tex
first subsection of syntax :$

@@
 
-\usepackage[margin=1in]{geometry}
 \usepackage{tikz-cd}
@@
 \newcommand{\isaSet}{\,\,\textrm{Set}}
+\newcommand{\isaTy}{\,\,\textrm{Type}}
+\newcommand{\isaCat}{\,\,\textrm{Cat}}
 \newcommand{\Set}{\textrm{Set}}
@@
 \newcommand{\isadtctx}{\,\,\textrm{type context}}
-\newcommand{\isavectx}{\,\,\textrm{cat context}}
+\newcommand{\isavectx}{\,\,\textrm{trans. context}}
 %% \newcommand{\obctx}{\,\,\textrm{object context}}
-\newcommand{\boundary}{\,\,\textrm{boundary}}
+\newcommand{\boundary}{\,\,\textrm{set context}}
 \newcommand{\prof}{\,\,\textrm{span}}
@@
 \newcommand{\pendappXtoY}[2]{#1^{#2}}
+
+\newcommand{\Cat}{\textrm{Cat}}
+\newcommand{\Id}[3]{\textrm{Id} #1\,#2\,#3}
+\newcommand{\vlambda}[1]{\lambda^{F}{#1}.}
+\newcommand{\hlambda}[2]{\lambda^{P}{#1};{#2}.}
 
@@
 
+The syntactic forms of Simple Category Theory (SCT) are given in
+\cref{fig:syntax}.
+%
+First, we have the types and terms of the logic over our category
+theory. Types include the usual dependent sum, product and identity
+types of dependent type theory, as well as several forms that allow us
+to abstract over the category theoretic constructions: a type $\Cat$
+of categories, a type $\varr{\cat C}{\cat D}$ for functors from $\cat
+C$ to $\cat D$, a type $\harr{\cat C}{\cat D}$ for profunctors
+contravariant in $\cat C$ and covariant in $\cat D$, and finally a
+type $\pendallXdotY {\alpha:{\cat C}} R$ that, when combined with
+other constructions on profunctors, allows for abstraction over
+arbitrary natural transformations.
+%
+Terms $L,M,N$ describe the elements of our logic's types, and include
+the usual constructions as well as forms for defining categories,
+functors, profunctors and natural transformations.
+%
+Next, we have a syntax for categories, which are written using a
+blackboard font $\cat C, \cat D, \cat E$. Categories include terms $M$
+of the category type (allowing for abstraction), in addition to
+primitives for product and unit categories.
+%
+Next, we have a syntax for objects of a category, which, because they
+are typed with respect to an input variable, provide a syntax for
+functors.
+%
+The objects we can define are object variables, application of a
+functor $M a$, and pairs of objects (for the product category) and the
+unique object of the unit category.
+%
+Next, $P,Q,R$ stand for \emph{sets}, but similar to objects, they are
+always typed with respect to a contravariant and covariant variable,
+and so denote profunctors.
+%
+These include first the multiplicatives: the hom set
+$\punitinXfromYtoZ {\cat C} a b$ representing the morphisms from $a$
+to $b$ in $\cat C$, a tensor product $\tensorexistsXwithYandZ
+{\beta:\cat D} P Q$, and two kinds of ``ordered function types''
+$\homrallXYtoZ \beta P Q$ and $\homlallXYtoZ \alpha R S$. 
+%
+We also include application of a profunctor term $M$ to two objects,
+and cartesian product and unit sets.
+%
+Next, we have transformations $s,t,u$, which are typed with respect
+to (0 or more) input and (one) output set.
+
+Finally, we have three forms of contexts, type contexts $\Gamma$, set
+contexts $\Xi$ and transformation contexts $\Phi$. A type context
+$\Gamma$ is the familiar notion of context from dependent type
+theory. We use uppercase $X,Y,Z$ for term variables to distinguish
+them from set variables. Every set expression $P$ must depend on an
+object of one category covariantly and an object of one category
+contravariantly, however these are allowed to be the same
+object. These two scenarios are captured by the notion of set context
+$\Xi$. The set context $\alpha:\cat C; \beta: \cat D$ gives the common
+case where the contravariant variable $\alpha$ is different from the
+covariant variable $\beta$, and the set context $\alpha: \cat C$ alone
+gives the case where the profunctor depends on the same variable both
+covariantly and contravariantly. Note that we write object variables
+as greek letters $\alpha, \beta$. Transformation contexts $\Phi$
+describe the inputs to a natural transformation. The input variables
+of a natural transformation $x: R$ stand for elements of sets, but as
+just described, sets always have covariant and contravariant
+dependencies. For this reason, transformation contexts include
+category variables $\alpha: \cat C$ interleaved with the set variables
+$x : R$, a typical context $\Phi$ would look like $\alpha_1:\cat C_1,
+x_1:R_1, \alpha_2: \cat C_2, x_2 : R_2, \alpha_3: \cat C_3$, where
+$R_1$ is parameterized by $\alpha_1$ contravariantly and $\alpha_2$
+covariantly and $R_2$ is similarly parameterized by
+$\alpha_2;\alpha_3$. Note that we consider transformation contexts up
+to associativity in that $(\Phi, x:R, \Psi), y : S, \Sigma$ should be
+considered equal to $\Phi, x:R, (\Psi, y: S, \Sigma)$.  In this
+calculus there is no single ``empty'' transformation context, but
+rather the empty context is one that consists of a single category
+variable $\alpha:\cat C$ and no set variables. Each form of context
+has a corresponding form of substitution, denoted using the lower-case
+form of the same greek letters.
+
 \begin{figure}
   \begin{mathpar}
-    \begin{array}{rcl}
-      \textrm{Type} & A,B,C ::= &
+    \begin{array}{rccl}
+      \textrm{Type} & A,B,C & ::= &
       \sum_{X:A} B \pipe \prod_{X:A} B \pipe \Id A M N \\
-      \Cat \pipe \varr{\cat C}{\cat D}\pipe \harr {\cat C} {\cat D}\pipe \pendallXdotY {\alpha:{\cat C}} R\\
-      \textrm{Term} & L,M,N ::= & X \pipe \lambda X:M. N \pipe M\,N \pipe (M,N) \pipe \pi_1 M \pipe \pi_2 M\\
-      &&\lceil \cat C \rceil \pipe \vlambda {\alpha:\cat C} a \pipe \hlambda {\alpha:\cat C}{\beta: \cat D} R \ldots \\
-      \textrm{Categories} &\cat C, \cat D, \cat E ::= & M \pipe \cat C \times \cat D \pipe \cat 1\\
-      \textrm{Objects} & a, b, c ::= & \alpha \pipe M a \ldots\\
-      \textrm{Sets} & P,Q,R ::= & \ldots \\
+      &&& \Cat \pipe \varr{\cat C}{\cat D}\pipe \harr {\cat C} {\cat D}\pipe \pendallXdotY {\alpha:{\cat C}} R\\
+      \textrm{Term} & L,M,N & ::= & X \pipe \lambda X:M. N \pipe M\,N \pipe (M,N) \pipe \pi_1 M \pipe \pi_2 M\\
+      &&&\lceil \cat C \rceil \pipe \vlambda {\alpha:\cat C} a \pipe \hlambda {\alpha:\cat C}{\beta: \cat D} R \ldots \\
+      \textrm{Categories} &\cat C, \cat D, \cat E & ::= & M \pipe \cat C \times \cat D \pipe \cat 1\\
+      \textrm{Objects} & a, b, c & ::= & \alpha \pipe M a \pipe (a,b) \pipe () \ldots\\
+      \textrm{Sets} & P,Q,R & ::= & \punitinXfromYtoZ {\cat C} a b \pipe \tensorexistsXwithYandZ {\beta} P Q \pipe \homrallXYtoZ \beta P Q \pipe \homlallXYtoZ \alpha R S \\
+      &&& M(a,b) \pipe 1 \pipe P \times Q\\
       \textrm{Transformations} & s,t,u & ::= & \ldots \\
       \\
-      \textrm{Type Context} & \Gamma ::= & \cdot \pipe \Gamma, X : A\\
-      \textrm{Cat Context} & \Phi,\Psi ::= & \cdot \pipe \alpha : \cat C \pipe \Phi , x : P , \Psi\\
+      \textrm{Type Context} & \Gamma, \Delta & ::= & \cdot \pipe \Gamma, X : A\\
+      \textrm{Category Context} & \Xi & ::= & \alpha:\cat C \pipe \alpha:\cat C;\beta: \cat D\\
+      \textrm{Set Context} & \Phi,\Psi & ::= & \alpha : \cat C \pipe \Phi , x : P , \Psi\\
+      \textrm{Term Substitution} & \gamma, \delta & ::= & \cdot \pipe \gamma, M/X\\
+      \textrm{Object Substitution} & \xi, \zeta & ::= & a/\alpha \pipe a/\alpha;b/\beta\\
+      \textrm{Transformation Substitution} & \phi, \psi & ::= & a/\alpha \pipe \phi,s/x, \psi      
     \end{array}
@@
   \caption{Simple Category Theory Syntactic Forms}
+  \label{fig:syntax}
 \end{figure}
 
-\subsection{Judgmental Structure}
-
-The following figures present the basic formation, primitive and
-admissible equality rules of our type theory, an internal language for
-virtual double categories with restrictions.
-%
-The type theory has 4 central concepts, which we will refer to
-sometimes using type theoretic/set theoretic and at other times
-category-theoretic terminology.
-%
-In type theoretic terminology, judgment theory has \emph{sorts} $\cat
-C, \cat D, \cat E$, and \emph{types} $A,B,C$ which have a given sort
-and are also \emph{parameterized} by an object of some other sort, and
-so can also be thought of as \emph{functions} with input and output
-sorts.
-%
-Next there are \emph{judgments} $P,Q,R,S$ which are parameterized by
-two variables.
-%
-In $\alpha^o:\cat C;\beta : \cat D \pipe R$, we say $R$ depends on
-$\alpha$ \emph{contravariantly} and in $\beta$ \emph{covariantly}.
-%
-Finally, we have \emph{terms} $s,t,u$ that are elements of some
-judgment $R$ but are also parameterized by elements of some other
-judgments which are given by a \emph{context} $\Phi$, which is a
-``compatible string'' of judgments where the covariant variable of one
-judgment is the same as the contravariant variable of the next.
-%
-Since these terms are parameterized by other terms, we can see these
-as \emph{inference rules} for the judgments.
-
-At other times we will want to think of these using set-theoretic
-terminology, in which case we have sets $\cat C,\cat D, \cat E$,
-functions/elements $A,B,C$, \emph{spans} $R,\ldots$ and morphisms of
-spans $s,t,u$.
-%
-Finally when we develop multicategory theory we will use
-category-theoretic terminology, in which the structure is most
-familiar and we have categories, objects/functors,
-\emph{profunctors}/bimodules and 2-cells/homomorphisms of bimodules.
-
-Figure\ref{fig:formation} introduces the formation rules for the basic
-constructs. Sorts give a syntax for objects of the double
-category. Types have an output sort and a sorted input variable, and
-model vertical arrows. Next, judgments have two sorted variables, one
-contravariant and one covariant, which can be the same
-variable. Judgments model the horizontal arrows. Next are contexts,
-whose formation rule is the same as judgments, and which give a syntax
-for ``compatible strings'' of judgments. Then terms have an input
-context and an output judgment, which must have the same covariant and
-contravaiant variables, modeling ``globular'' 2-cells. Finally, we
-have substitutions for terms, which are typed with respect to an input
-context, an output context and two types, forming a kind of
-square. These model the ``compatible string'' of 2-cells that arise in
-the composition laws.
+Next, \cref{fig:formation} shows the \emph{formation} rules for these
+various syntactic forms. All other forms are typed with respect to a
+well-formed type context $\Gamma$. Objects are typed additionally with
+an input object variable $\alpha : \cat C$ and an output category
+$\cat D$; in the semantics these objects are modeled as functors. A
+set $S$ is typed with respect to a set context $\Xi$ to describe its
+covariant/contravariant dependence on some input objects. These are
+semantically modeled as profunctors. A transformation $s$ has a
+context of transformation variable and a single set output. To be well
+formed, the context and set must be parameterized by the same
+contravariant and covariant object variables. To ensure this, we use a
+coercion operation $\underline \Phi$ that erases everything in the
+context but the left-most and right-most object variables.
 
@@
   \begin{mathpar}
-    \inferrule
+    \inferrule*[right=TyCtxForm]
         {}
@@
 
-    \inferrule
+    \inferrule*[right=Cat-form]
+    {\Gamma \isadtctx}
+    {\Gamma\vdash \cat C \isaCat}
+
+    \inferrule*[right=Ob-form]
+    {\Gamma \vdash \cat C\isaCat\and \Gamma \vdash \cat D \isaCat}
+    {\Gamma \pipe \alpha : \cat C \vdash a : \cat D}\\
+
+    \inferrule*[right=SetCtxForm]
     {\Gamma \isadtctx}
@@
 
-    \inferrule
+    \inferrule*[right=SetForm]
     {\Gamma \isadtctx \and \Gamma \vdash \Xi \boundary}
@@
 
-    \inferrule*[right=Object/Type-formation]
-    {\Gamma \vdash \cat C : \Sort \and \Gamma \vdash \cat D : \Sort}
-    {\Gamma \pipe \alpha : \cat C \vdash A : \cat D}\\
-
-    \inferrule*[right=Context-formation]
+    \inferrule*[right=TransCtxForm]
     {\Gamma \isadtctx}
@@
 
-    \inferrule*[right=Term formation]
+    \inferrule*[right=TransFormation]
     {\Gamma \vdash \Phi \isavectx\and
      \Gamma\pipe \underline \Phi \vdash R \isaSet}
-    {\Gamma \pipe \Phi \vdash M : R}
+    {\Gamma \pipe \Phi \vdash s : R}
 
@@
 
-Next, Figure~\ref{fig:basic-rules} gives the basic structural rules of
-judgment theory, which model composition and restrictions in a virtual
-double category. First, we have varialbes and substitution for types,
-modeling identity and composition of vertical arrows. Next, we have
-restriction of a judgment to more specific types, which is given by
-substitution of types for the judgment's parameter variables. Next, we
-have rules for constructing an ``empty'' context, which is just a
+%% The following figures present the basic formation, primitive and
+%% admissible equality rules of our type theory, an internal language for
+%% virtual double categories with restrictions.
+%% %
+%% The type theory has 4 central concepts, which we will refer to
+%% sometimes using type theoretic/set theoretic and at other times
+%% category-theoretic terminology.
+%% %
+%% In type theoretic terminology, judgment theory has \emph{sorts} $\cat
+%% C, \cat D, \cat E$, and \emph{types} $A,B,C$ which have a given sort
+%% and are also \emph{parameterized} by an object of some other sort, and
+%% so can also be thought of as \emph{functions} with input and output
+%% sorts.
+%% %
+%% Next there are \emph{judgments} $P,Q,R,S$ which are parameterized by
+%% two variables.
+%% %
+%% In $\alpha^o:\cat C;\beta : \cat D \pipe R$, we say $R$ depends on
+%% $\alpha$ \emph{contravariantly} and in $\beta$ \emph{covariantly}.
+%% %
+%% Finally, we have \emph{terms} $s,t,u$ that are elements of some
+%% judgment $R$ but are also parameterized by elements of some other
+%% judgments which are given by a \emph{context} $\Phi$, which is a
+%% ``compatible string'' of judgments where the covariant variable of one
+%% judgment is the same as the contravariant variable of the next.
+%% %
+%% Since these terms are parameterized by other terms, we can see these
+%% as \emph{inference rules} for the judgments.
+
+%% At other times we will want to think of these using set-theoretic
+%% terminology, in which case we have sets $\cat C,\cat D, \cat E$,
+%% functions/elements $A,B,C$, \emph{spans} $R,\ldots$ and morphisms of
+%% spans $s,t,u$.
+%% %
+%% Finally when we develop multicategory theory we will use
+%% category-theoretic terminology, in which the structure is most
+%% familiar and we have categories, objects/functors,
+%% \emph{profunctors}/bimodules and 2-cells/homomorphisms of bimodules.
+
+%% Figure\ref{fig:formation} introduces the formation rules for the basic
+%% constructs. Sorts give a syntax for objects of the double
+%% category. Types have an output sort and a sorted input variable, and
+%% model vertical arrows. Next, judgments have two sorted variables, one
+%% contravariant and one covariant, which can be the same
+%% variable. Judgments model the horizontal arrows. Next are contexts,
+%% whose formation rule is the same as judgments, and which give a syntax
+%% for ``compatible strings'' of judgments. Then terms have an input
+%% context and an output judgment, which must have the same covariant and
+%% contravaiant variables, modeling ``globular'' 2-cells. Finally, we
+%% have substitutions for terms, which are typed with respect to an input
+%% context, an output context and two types, forming a kind of
+%% square. These model the ``compatible string'' of 2-cells that arise in
+%% the composition laws.
+
+%% Next, Figure~\ref{fig:basic-rules} gives the basic structural rules of
+%% judgment theory, which model composition and restrictions in a virtual
+%% double category. First, we have varialbes and substitution for types,
+%% modeling identity and composition of vertical arrows. Next, we have
+%% restriction of a judgment to more specific types, which is given by
+%% substitution of types for the judgment's parameter variables. Next, we
+%% have rules for constructing an ``empty'' context, which is just a
+In typical natural-deduction style, all syntactic forms typed in
+context admit an action of substitution. For terms, this is the
+ordinary type theoretic substitution. Naturally objects $\alpha:\cat C
+\vdash a : \cat D$ can be substituted for object variables $\beta :
+\cat D$ in other objects. We can also substitute objects into
+\emph{sets}. For instance, if we have a hom set in a category
+$\punitinXfromYtoZ {\cat C} \alpha {\alpha'}$, and we can define
+objects $a$ and $a'$ in the category $\cat C$, then we can substitute
+those in to form the hom set between $a$ and $a'$: $\punitinXfromYtoZ
+{\cat C} a {a'}$. This semantically is a ``restriction'' of a
+profunctor along two functors, and modeling this operation simply as a
+substitution considerably simplifies reasoning using profunctors.
+%
+Finally we have the action of substitution on natural
+transformations. A natural transformation is indexed by set variables
+$x:R$, whose sets are each parameterized by a contravariant and a
+covariant category variable $\alpha;\beta$. We can think of natural
+transformations as \emph{polymorphic} in the categories involved, and
+so when we make a transformation substitution, we substitute not just
+the set variables, but we also \emph{instantiate} the polymorphic
+category variables with objects.
+
+\cref{fig:contexts} details the typing rules for contexts and the
+typing for variables and the admissible actions of substitution (the
+definition of substitution is straightforward and found in the
+appendix). the rules for term and object variables are as usual.
+%
+An object substitution into a single object context $\alpha:\cat C$
+consists of an object $a : \cat C$ parameterized by an object of
+another category.
+%
+An object substitution into a pair of co- and contra-variant variables
+$\alpha:\cat C;\beta:\cat D$ is a pair of an object substituted for
+$\alpha$ and one for $\beta$, each typed with respect to the
+appropriate variable of an object context $\Xi$.
+%
+Note that this allows for $\Xi$ to consist of either two variables or
+a single variable.
+%
+When
+constructing a transformation context $\Phi, x : R, \Psi$, we need to
+ensure that the contravariant dependence of $R$ matches the covariant
+dependence of $\Phi$ and vice-versa for $\Psi$. We define this with
+auxiliary metafunctions $d^+,d^-$ which extract the covariant and
+contravariant object variable of a transformation context.
+%
+The rules for transformation variables and substitutions define the
+use of variables to be \emph{linear} (each variables is used exactly
+once) and \emph{ordered}, they occur in the term in the same order
+that they appear in the context.
+%
+The rule for a transformation variables says a variable $x$ can only
+be used when it is the sole variable in the context.
+%
+A transformation subsitution into an ``empty'' transformation context
+$\alpha: \cat C$ is the same as an object substitution into
+$\alpha:\cat C$.
+%
+The rule for constructing a transformation substitution into a
+non-empty context $\Phi_1,x:R,\Phi_2$ says we can construct a
+substitution by constructing substitutions $\phi_1,\phi_2$ for
+$\Phi_1$ and $\Phi_2$ as well as a transformation $s$ into an
+instantiation of $R$ where the input contexts come from splitting the
+input context $\Psi$ into three components, and ensuring that the
+instantiation of $R$ agrees with how $\phi_1,\phi_2$ instantiate the
+variables. This uses two new metafunctions. First, $\Phi \jnctx \Psi$
+is an operation for concatenating two contexts when $d^+(\Phi) =
+d^-(\Psi)$. Next, we extend the boundary maps $d^\pm$ to substitutions
+to extract their object substitutions.
 
@@
     \inferrule
-    {\Gamma \isadtctx \and \Gamma\pipe \cdot \vdash A \isaSet}
-    {\Gamma , x : A \isadtctx}\\
-
-    \inferrule{}{\Gamma \vdash \cdot \boundary}
-
-    \inferrule{\Gamma \vdash C : \Sort}{\Gamma \vdash \alpha: C \boundary}
-
-    \inferrule{\Gamma \vdash C : \Sort\and\Gamma\vdash D : \Sort}{\Gamma \vdash \alpha: C;\beta:D \boundary}\\
-
-    \inferrule{}{\Gamma \vdash \cdot \isavectx}
-
-    \inferrule{\Gamma \vdash C : \Sort}{\Gamma \vdash \alpha: C \isavectx}
+    {\Gamma \isadtctx \and \Gamma \vdash A \isaTy}
+    {\Gamma , X : A \isadtctx}\\
+
+    \inferrule
+    {X : A \in \Gamma}
+    {\Gamma \vdash X : A}
+
+    \inferrule
+    {}
+    {\alpha: \cat C \vdash \alpha: \cat C}
+
+    \inferrule{\Gamma \vdash \cat C : \isaCat}{\Gamma \vdash \alpha: \cat C \boundary}
+
+    \inferrule{\Gamma \vdash \cat C : \isaCat\and
+      \Gamma\vdash \cat D : \isaCat}{\Gamma \vdash \alpha: \cat C;\beta:\cat D \boundary}\\
+
+    \inferrule*
+        {\Gamma \pipe \alpha':\cat C' \vdash A : \cat C}
+        {\Gamma \vdash a/\alpha : \alpha':\cat C' \to \alpha:\cat C}
+
+    \inferrule*
+    {\Gamma \pipe d^-\Xi \vdash a : \cat C\and
+     \Gamma \pipe d^+\Xi \vdash b : \cat D
+    }
+    {\Gamma \vdash (a/\alpha;b/\beta) : \Xi \to (\alpha:\cat C;\beta:\cat D)}\\
+
+    \inferrule
+        {\Gamma \vdash \cat C : \isaCat}
+        {\Gamma \vdash \alpha: \cat C \isavectx}
 
@@
     \Gamma \vdash \Psi \isavectx \and
-    \Gamma \pipe d^+\Phi, d^- \Psi \vdash R \isaSet}
+    \Gamma \pipe d^+\Phi; d^- \Psi \vdash R \isaSet}
     {\Gamma \vdash \Phi, x:R, \Psi \isavectx}
 
-    (\Phi,x:R,\Psi),y:S,\Sigma = \Phi,(x:R,\Psi,y:S,\Sigma)
+    \inferrule
+    {}
+    {\Gamma \pipe \alpha:\cat C, x : R, \beta: \cat D \vdash x : R}
+
+    \inferrule
+    {\Psi = \Psi_1 \jnctx \Psi_M \jnctx \Psi_2\and
+      \Gamma \vdash \phi_1 : \Psi_1 \to \Phi_1\and
+      \Gamma \pipe \Psi_s \vdash s : R[d^+\phi_1;d^-\phi_2] \and
+      \Gamma \vdash \phi_2 : \Psi_2 \to \Phi_2
+      }
+    {\Gamma \vdash \phi_1,t/x,\phi_2 : \Psi \to \Phi_1,x:R,\Phi_2}
   \end{mathpar}
   \begin{align*}
-    \underline \cdot &= \cdot\\
-    \underline {\alpha:C} &= \alpha:C\\
+    \underline{\alpha:C} &= \alpha:C\\
     \underline{\Phi,x:R,\Psi} &= d^-\Phi; d^+{\Psi}
@@
   \begin{align*}
-    d^{\pm}(\cdot) &\quad\text{undefined}\\
-    d^{\pm}(\alpha:C) &= \alpha : C\\
+    d^{\pm}(\alpha:\cat C) &= \alpha : \cat C\\
     d^{-}(\Phi,x:R,\Psi) &= d^-\Phi\\
@@
   \end{align*}
-  \caption{Basic Rules for Contexts}
+    \begin{align*}
+    d^{\pm}(a/\alpha) &= a/\alpha\\
+    d^{-}(\phi,t/x,\psi) &= d^-\phi\\
+    d^{+}(\phi,t/x,\psi) &= d^+\psi\\
+  \end{align*}
+  \caption{Basic Rules of Judgment Theory}
+  \label{fig:contexts}
 \end{figure}
 
-\begin{figure}
-  \begin{mathpar}
-    \inferrule
-    {}
-    {\Gamma\pipe \cdot \vdash \Sort \isaSet}
-
-    \inferrule
-    {x:S \in \Gamma}
-    {\Gamma \pipe \cdot \vdash x : S}
-
-    \inferrule
-    {\Gamma\pipe\cdot \vdash C : \Sort}
-    {\Gamma \pipe \alpha:\cat C \vdash \alpha : \cat C}
-
-    \inferrule
-    {\Gamma \pipe \alpha:C,\beta:D \vdash R \isaSet}
-    {\Gamma\pipe \alpha:C, x:R, \beta: D \vdash x : R}
-
-    \inferrule{\forall x:A \in \Gamma. \Delta \vdash \gamma(x) : A}{\gamma : \Delta \to \Gamma}\\
-
-    \inferrule*%% [right=NullaryBoundaryMap]
-               {\Gamma \isadtctx}{\Gamma \vdash \cdot : \cdot \to \cdot}\and
-    \inferrule*{\Gamma \pipe \alpha':C' \vdash A : C}{\Gamma \vdash A/\alpha : \alpha':C' \to \alpha:C}
-
-    \inferrule*
-    {\Gamma \pipe \alpha':C' \vdash A : C\and
-      \Gamma \pipe \beta':D' \vdash B : D
-    }
-    {\Gamma \vdash (A/\alpha;B/\beta) : (\alpha':C';\beta':D') \to (\alpha:C;\beta':D')}\\
-
-    \inferrule*%% [right=NullCtxMap]
-               {\Gamma \isadtctx}{\Gamma \vdash \cdot : \cdot \to \cdot}
-
-    \inferrule*{\Gamma \pipe \alpha':C' \vdash A : C}{\Gamma \vdash A/\alpha : \alpha':C' \to \alpha:C}
-
-    \inferrule
-    {\Psi = \Psi_1 \jnctx \Psi_M \jnctx \Psi_2\and
-      \Gamma \vdash \phi_1 : \Psi_1 \to \Phi_1\and
-      \Gamma \pipe \Psi_M \vdash M : R[d^+\phi_1;d^-\phi_2] \and
-      \Gamma \vdash \phi_2 : \Psi_2 \to \Phi_2
-      }
-    {\Gamma \vdash \phi_1,M/x,\phi_2 : \Psi \to \Phi_1,x:R,\Phi_2}
-  \end{mathpar}
-  \caption{Basic Rules of Judgment Theory}
-  \label{fig:basic-rules}
-\end{figure}
-
-\begin{figure}
-  \begin{mathpar}
-    \inferrule
-    {\Gamma \vdash \Phi \isavectx \and
-    \Gamma \vdash \Psi \isavectx \and
-    d^+\Phi = d^- \Psi}
-    {\Gamma \vdash \Phi \jnctx \Psi \isavectx}
-
-    \inferrule
-    {\Gamma \vdash \phi : \Phi' \to \Phi \and
-     \Gamma \vdash \psi : \Psi' \to \Psi\and
-     d^+\phi = d^- \psi}
-    {\Gamma \vdash \phi \jnctx \psi : \Phi' \jnctx \Psi' \to \Phi \jnctx \Psi}
-
-    \inferrule*[right=ObSubst]
-    {\alpha : \cat C \vdash B : \cat D\and
-    \beta : \cat D \vdash A : \cat E}
-    {\alpha : \cat C \vdash A[B/\beta] : \cat E}\\
-
-    \inferrule*[right=Everything Subst]{\gamma : \Delta \to \Gamma \and \Gamma J}{\Delta J}
-
-    \inferrule{\Gamma \vdash \xi : Z \to \Xi\and \Gamma \pipe \Xi \vdash R \isaSet}{\Gamma \pipe Z \vdash R[\xi] \isaSet}
-
-    \inferrule{\Gamma \vdash \phi : \Psi \to \Phi\and \Gamma \pipe \Phi \vdash M : R}{\Gamma \pipe \Psi \vdash M[\phi] : R[\underline \phi]}
-  \end{mathpar}
-  \begin{align*}
-    (\Phi \jnctx \Psi) \jnctx \Sigma &= \Phi \jnctx (\Psi \jnctx \Sigma)\\
-    \Phi \jnctx \alpha:C &= \Phi\\
-    \alpha:C \jnctx \Psi &= \Psi\\\\
-    d^\pm(A/\alpha) &= A/\alpha\\
-    d^+(\phi_1,M/x,\phi_2) &= d^+ \phi_2\\
-    d^-(\phi_1,M/x,\phi_2) &= d^- \phi_1\\\\
-    \underline \cdot &= \cdot\\
-    \underline {A/\alpha} &= A/\alpha\\
-    \underline {\phi_1,M/x,\phi_2} &= d^-\phi_1;d^+\phi_2
-  \end{align*}
-  \caption{Admissible Rules}
-\end{figure}
-
-\begin{figure}
-  \begin{mathpar}
-    \begin{array}{rcl}
-      x[\gamma] &=& \gamma(x)\\
-      J[\gamma][\delta] & = & J[\gamma[\delta]]\\
-      \gamma[\delta](x) &=& \gamma(x)[\delta]\\
-
-      A[\alpha/\alpha] &=& A\\
-      \alpha[A/\alpha] &=& A\\
-      A[B/\beta][C/\gamma] & = & A[B[C/\gamma]/\beta]\\
-      A[B/\beta][\gamma] &=& A[\gamma][B[\gamma]/\beta]\\\\
-
-      R[\cdot] &=& R\\
-      R[\alpha/\alpha] &=& R\\
-      R[\alpha/\alpha,\beta/\beta] &=& R\\\\
-      R[\xi][\xi'] & = & R[\xi[\xi']]\\
-      R[\xi][\gamma] &=& R[\gamma][\xi[\gamma]]\\\\
-
-      \cdot[\xi] &=& \cdot\\
-      (A/\alpha)[B/\beta] &=& A[B/\beta]/\alpha\\
-      (A/\alpha;B/\beta)[\xi] &=& A[d^-\xi]/\alpha,B[d^+\xi]/\beta\\
-      \cdot[\gamma] &=& \cdot\\
-      (A/\alpha)[\gamma] &=& A[\gamma]/\alpha\\
-      (A/\alpha;B/\beta)[\gamma] &=& A[\gamma]/\alpha;B[\gamma]/\beta\\\\
-
-      \cdot[\gamma] &=& \cdot\\
-      (A/\alpha)[\gamma] &=& A[\gamma]/\alpha\\
-      (\phi_1,M/x,\phi_2)[\gamma] &=& \phi_1[\gamma],M[\gamma]/x,\phi_2[\gamma]
-
-      TODO: cat substitution id/assoc
-    \end{array}
-  \end{mathpar}  
-  \caption{Algebraic Properties}
-  \label{fig:equations}
-\end{figure}
-
-Now that we have established a judgmental structure, we move on to the
-multiplicative connectives, starting with the easier negative
-connectives and then the more difficult positives.
-
-\subsection{Left and Right Hom of Profunctors}
+Now that we have established a judgmental structure, in typical type
+theoretic fashion, our type constructs follow naturally as
+internalizing the judgments.
+
+\subsection{Negative Multiplicatives}
 
@@
 
+TODO: define substitution
+\begin{figure}
+  \begin{mathpar}
+    \inferrule
+    {}
+    {\Gamma\pipe \cdot \vdash \Sort \isaSet}
+
+    \inferrule
+    {x:S \in \Gamma}
+    {\Gamma \pipe \cdot \vdash x : S}
+
+    \inferrule
+    {\Gamma\pipe\cdot \vdash C : \Sort}
+    {\Gamma \pipe \alpha:\cat C \vdash \alpha : \cat C}
+
+    \inferrule
+    {\Gamma \pipe \alpha:C,\beta:D \vdash R \isaSet}
+    {\Gamma\pipe \alpha:C, x:R, \beta: D \vdash x : R}
+
+    \inferrule{\forall x:A \in \Gamma. \Delta \vdash \gamma(x) : A}{\gamma : \Delta \to \Gamma}\\
+
+    \inferrule*%% [right=NullaryBoundaryMap]
+               {\Gamma \isadtctx}{\Gamma \vdash \cdot : \cdot \to \cdot}\and
+    \inferrule*{\Gamma \pipe \alpha':C' \vdash A : C}{\Gamma \vdash A/\alpha : \alpha':C' \to \alpha:C}
+
+    \inferrule*
+    {\Gamma \pipe \alpha':C' \vdash A : C\and
+      \Gamma \pipe \beta':D' \vdash B : D
+    }
+    {\Gamma \vdash (A/\alpha;B/\beta) : (\alpha':C';\beta':D') \to (\alpha:C;\beta':D')}\\
+
+    \inferrule*%% [right=NullCtxMap]
+               {\Gamma \isadtctx}{\Gamma \vdash \cdot : \cdot \to \cdot}
+
+    \inferrule*{\Gamma \pipe \alpha':C' \vdash A : C}{\Gamma \vdash A/\alpha : \alpha':C' \to \alpha:C}
+
+    \inferrule
+    {\Psi = \Psi_1 \jnctx \Psi_M \jnctx \Psi_2\and
+      \Gamma \vdash \phi_1 : \Psi_1 \to \Phi_1\and
+      \Gamma \pipe \Psi_M \vdash M : R[d^+\phi_1;d^-\phi_2] \and
+      \Gamma \vdash \phi_2 : \Psi_2 \to \Phi_2
+      }
+    {\Gamma \vdash \phi_1,M/x,\phi_2 : \Psi \to \Phi_1,x:R,\Phi_2}
+  \end{mathpar}
+  \caption{Substitutions in SCT}
+  \label{fig:basic-rules}
+\end{figure}
+
+\begin{figure}
+  \begin{mathpar}
+    \inferrule
+    {}
+    {\cdot : \cdot \to \cdot}
+
+    \inferrule
+    {\gamma : \Delta \to \Gamma \and \Delta \vdash M : A[\gamma]}
+    {\gamma,M/X : \Delta \to \Gamma,X:A}
+
+    \inferrule
+    {\Gamma \vdash \Phi \isavectx \and
+    \Gamma \vdash \Psi \isavectx \and
+    d^+\Phi = d^- \Psi}
+    {\Gamma \vdash \Phi \jnctx \Psi \isavectx}
+
+    \inferrule
+    {\Gamma \vdash \phi : \Phi' \to \Phi \and
+     \Gamma \vdash \psi : \Psi' \to \Psi\and
+     d^+\phi = d^- \psi}
+    {\Gamma \vdash \phi \jnctx \psi : \Phi' \jnctx \Psi' \to \Phi \jnctx \Psi}
+
+    \inferrule*[right=ObSubst]
+    {\alpha : \cat C \vdash B : \cat D\and
+    \beta : \cat D \vdash A : \cat E}
+    {\alpha : \cat C \vdash A[B/\beta] : \cat E}\\
+
+    \inferrule*[right=Everything Subst]{\gamma : \Delta \to \Gamma \and \Gamma J}{\Delta J}
+
+    \inferrule{\Gamma \vdash \xi : Z \to \Xi\and \Gamma \pipe \Xi \vdash R \isaSet}{\Gamma \pipe Z \vdash R[\xi] \isaSet}
+
+    \inferrule{\Gamma \vdash \phi : \Psi \to \Phi\and \Gamma \pipe \Phi \vdash M : R}{\Gamma \pipe \Psi \vdash M[\phi] : R[\underline \phi]}
+  \end{mathpar}
+  \begin{align*}
+    (\Phi \jnctx \Psi) \jnctx \Sigma &= \Phi \jnctx (\Psi \jnctx \Sigma)\\
+    \Phi \jnctx \alpha:C &= \Phi\\
+    \alpha:C \jnctx \Psi &= \Psi\\\\
+    d^\pm(A/\alpha) &= A/\alpha\\
+    d^+(\phi_1,M/x,\phi_2) &= d^+ \phi_2\\
+    d^-(\phi_1,M/x,\phi_2) &= d^- \phi_1\\\\
+    \underline \cdot &= \cdot\\
+    \underline {A/\alpha} &= A/\alpha\\
+    \underline {\phi_1,M/x,\phi_2} &= d^-\phi_1;d^+\phi_2
+  \end{align*}
+  \caption{Admissible Rules}
+\end{figure}
+
+\begin{figure}
+  \begin{mathpar}
+    \begin{array}{rcl}
+      x[\gamma] &=& \gamma(x)\\
+      J[\gamma][\delta] & = & J[\gamma[\delta]]\\
+      \gamma[\delta](x) &=& \gamma(x)[\delta]\\
+
+      A[\alpha/\alpha] &=& A\\
+      \alpha[A/\alpha] &=& A\\
+      A[B/\beta][C/\gamma] & = & A[B[C/\gamma]/\beta]\\
+      A[B/\beta][\gamma] &=& A[\gamma][B[\gamma]/\beta]\\\\
+
+      R[\cdot] &=& R\\
+      R[\alpha/\alpha] &=& R\\
+      R[\alpha/\alpha,\beta/\beta] &=& R\\\\
+      R[\xi][\xi'] & = & R[\xi[\xi']]\\
+      R[\xi][\gamma] &=& R[\gamma][\xi[\gamma]]\\\\
+
+      \cdot[\xi] &=& \cdot\\
+      (A/\alpha)[B/\beta] &=& A[B/\beta]/\alpha\\
+      (A/\alpha;B/\beta)[\xi] &=& A[d^-\xi]/\alpha,B[d^+\xi]/\beta\\
+      \cdot[\gamma] &=& \cdot\\
+      (A/\alpha)[\gamma] &=& A[\gamma]/\alpha\\
+      (A/\alpha;B/\beta)[\gamma] &=& A[\gamma]/\alpha;B[\gamma]/\beta\\\\
+
+      \cdot[\gamma] &=& \cdot\\
+      (A/\alpha)[\gamma] &=& A[\gamma]/\alpha\\
+      (\phi_1,M/x,\phi_2)[\gamma] &=& \phi_1[\gamma],M[\gamma]/x,\phi_2[\gamma]
+
+      TODO: cat substitution id/assoc
+    \end{array}
+  \end{mathpar}  
+  \caption{Algebraic Properties}
+  \label{fig:equations}
+\end{figure}
+
 \section{Additional Rules}

### 2022-02-06 paper/lipics-v2021-sample-article.tex
ends and funs

@@
 
-\newcommand{\homlallXYtoZ}[3]{#2 \tlforall #1 #3}
+\newcommand{\homlallXYtoZ}[3]{#3 \tlforall #1 #2}
 
@@
 theoretic fashion, our type constructs follow naturally as
-internalizing the judgments.
+internalizing the judgments. Though the positive multiplicatives are
+more familiar, we start by describing the negative multiplicatives,
+since they are better behaved, and are useful to have already defined
+when discussing the positives.
+
+\subsection{Types}
+
+Finally, we describe the \emph{end}, $\pendallXdotY {\alpha} P$, which
+when combined with the left or right hom types described in the next
+section allows for us to internalize arbitrary transformation sequents
+$\Phi \vdash P$ as a type.
+%
+This type is well formed when $\alpha$ is the covariant \emph{and}
+contravariant variable of $P$.
+%
+An element of the type is a polymorphic element of $P$, varying in
+$\alpha$.
+%
+The introduction form is naturally a $\lambda$ abstracting over
+$\alpha:\cat C$ and the elimination form instantiates $\alpha$ with an
+arbitrary object $a : \cat C$.
+%
+We also include the necessary $\beta\eta$ equalities.
+
+\begin{figure}
+  \begin{mathpar}
+    \inferrule
+    {\Gamma \pipe \alpha:\cat C\vdash P \isaSet}
+    {\Gamma \vdash \pendallXdotY {\alpha:{\cat C}} P \isaTy}
+
+    \inferrule
+    {\Gamma \pipe \alpha:\cat C\vdash t : P}
+    {\Gamma \vdash \pendlambdaXdotY \alpha t : \pendallXdotY \alpha P}
+
+    \inferrule
+    {\Gamma \vdash M : \pendallXdotY \alpha P\and
+     \Gamma \pipe \beta:\cat D \vdash a : \cat C}
+    {\Gamma \pipe \beta:\cat D \vdash \pendappXtoY M a : P[a/\alpha]}
+
+    \pendappXtoY {(\pendlambdaXdotY \alpha t)} a = t[a/\alpha]
+
+    \inferrule*[right=2Cell$\eta$]
+    {\Gamma \vdash M : \pendallXdotY \alpha P }
+    {\Gamma \vdash M = \pendlambdaXdotY \alpha {\pendappXtoY M \alpha} : \pendallXdotY \alpha P}
+  \end{mathpar}
+  \caption{End}
+  \label{fig:end}
+\end{figure}
+
+Semantically, this type corresponds to taking the end of a profunctor
+$P : {\cat C}^{op} \times {\cat C} \to \Set$, typically written
+\[ \int^{\alpha:\cat C} P \alpha \alpha \]
 
@@
 
-First, we study the two ``internal homs'' of judgments. Since the
-``monoidal'' structure of judgments is not symmetric, we get two
-distinct notions internalizing concatenating on the left and the right
-of the context. In ordinary category theory, these profunctors can be
-constructed using an \emph{end} and the hom of functor
-categories. Given two profunctors $P : C^o \times E \to \Set$ and $R :
-D^o \times E \to Set$, their hom is given as
-\[ \int_{e \in E} R(-,e) \Rightarrow P(e,=) \]
-
-Type theoretically this is a kind of multi-hom where we are
-quantifying both over an indexing sort variable $e$ and a profunctor
-variable of type $R$. We write the covariant and contravariant
-versions of these connectives as $\homrallXYtoZ \gamma R S$ and
-$\homlallXYtoZ \alpha R S$.
-
-The syntax is given in Figure~\ref{fig:cotensor}.
-%
-First, the formation rule for covariant cotensor formation requires
-that the covariant variables for $R,P$ are the same, and the
-contravariant variables match the boundary of the overall
-judgment. Note that we use $d^+\Xi,d^-\Xi$ to allow for the case that
-the boundary is a single variable $\Xi = \alpha:\cat C$.
-%
-Next, the substitution rule is straightforward.
-%
-Then we have the elimination form, first written in a simple form with
-no substitution built in.
-%
-This says that given a variable of the cotensor judgment, and another
-variable of judgment $R$, we can apply these to get a term of judgment
-$P$. As a diagram this is written as follows:
-
-\[ % https://tikzcd.yichuanshen.de/#N4Igdg9gJgpgziAXAbVABwnAlgFyxMJZABgBpiBdUkANwEMAbAVxiRAGEQBfU9TXfIRQBGclVqMWbACLdeIDNjwEiAJjHV6zVohABROXyWCiZYeK1TdnHkYEqU685sk793cTCgBzeEVAAZgBOEAC2SGQgOBBIohLabAA6iQEQQYwMAATJ3nShoXQAdJkACtmJOEFYdGDeDDD1ATiZAEqGIMFhsdTRSOrxViBtth0h4YgAzD0xiAAsLgm6Je2d45G9k9QMWGBuUBBMAEb1INQAFjB0UGyQu6dRdFgMNwSsI6t900jzIBdXL3ctjs9gdjqweo9nrpbm8KFwgA
-\begin{tikzcd}
-C \arrow[r, "\forall \gamma. P \triangleleft R"] \arrow[d, no head, Rightarrow] & D \arrow[r, "R"] & E \arrow[d, no head, Rightarrow] \\
-C \arrow[rr, "P"]                                                               &                  & E                               
-\end{tikzcd}\]
-
-The universal property of this judgment is that this is the universal
-2-cell of this type, in that any 2-cell
-
-\[ % https://tikzcd.yichuanshen.de/#N4Igdg9gJgpgziAXAbVABwnAlgFyxMJZABgBpiBdUkANwEMAbAVxiRAGEByEAX1PUy58hFACZyVWoxZsAIr34gM2PASIBmCdXrNWiEAFEFAlcKJkAjJJ0z97Y0sGqRyTVe3S9hh8qFqUFlpSumy8kjBQAObwRKAAZgBOEAC2SIEgOBBI4sG2IABKDokpSJoZWYgALB4h+gAKRUmpiGTlpdQMWGBecBCdUCDUABYwdAOIYEwMDNQ4dFgMbJDdjSWIOZlI1SAjY0sErB1dXlAQTABGDIcZ84v6y6x88U1IrZuIAKxPIMXNH7MVdKdFb6KB0OAjAbDUbjcAHMI8IA
-\begin{tikzcd}
-C' \arrow[d] \arrow[r] & {} \arrow[r, no head, dashed] & D \arrow[r, "R"] & E \arrow[d, no head, Rightarrow] \\
-C \arrow[rrr, "P"]     &                               &                  & E                               
-\end{tikzcd}\]
-
-Exhibits a unique factorization:
-
-\[ % https://tikzcd.yichuanshen.de/#N4Igdg9gJgpgziAXAbVABwnAlgFyxMJZABgBoBGAXVJADcBDAGwFcYkQBhEAX1PU1z5CKAEwVqdJq3YARHnxAZseAkQDM4mgxZtEIAKLz+yoUTIiJ26Xq69jg1Sg0WtU3QaOKBK4SVLFLN3YOAHJPJQdfcn9AnXZw71NRGNc4vTk7LxNHZA0A1OsPbgkYKABzeCJQADMAJwgAWyRokBwIJDFJNJAAJU86xqQNVvbEABYC9wAFfvqmxDIRoZpGLDB3KAhmACNGNhoACxh6KHZIdZAaHHosRjOCNkyB+c62pAmQI5P7i5W1ja2u32rRudz050eCmeSAArFdRgA2J5zJAI+FIADsfwueig9DgR1Oh2Op3BD1mg0QWKWiBaqxxIE2Oz2l0+JJ+wOutw5FPmcJpxGRlMWb1pk3YAB0JdUILUmIwAARSsr0BoNegAOgVUyVEpwtSw9DAZT2e2qOAVfSF82pooAHOK9FaoSjEA6aZ16QDmcCvqTwOSrqCecVuEA
-\begin{tikzcd}
-C' \arrow[r] \arrow[d]                                                           & {} \arrow[r, no head, dashed] & D \arrow[d, no head, Rightarrow] \arrow[r, "R"] & E \arrow[d, no head, Rightarrow] \\
-C \arrow[d, no head, Rightarrow] \arrow[rr, "\forall \gamma. P \triangleleft R"] &                               & D \arrow[r, "R"]                                & E \arrow[d, no head, Rightarrow] \\
-C \arrow[rrr, "P"]                                                               &                               &                                                 & E                               
-\end{tikzcd}\]
-
-We give elimination rules and introduction rules in a style similar to
-$\lambda$-calculus, with corresponding $\beta\eta$ rules.
+Since our calculus of natural transformations is ordered linear, we
+have two distinct ``function sets'', internalizing concatenation on
+the left or right of the context. The typing rules for these are given
+in \cref{fig:cotensor}. The two types are the \emph{right hom}
+$\homrallXYtoZ \alpha R P$ and the \emph{left hom} $\homlallXYtoZ
+\alpha R P$. As a type, we can think of these both as a kind of
+polymorphic function type in that in addition to parameterizing an
+element of the set $P$ by an element of the set $R$, they also
+quantify over an object $\alpha$ that is in $P$ and $R$.
+
+These behave like function types, but their typing is somewhat
+involved. First consider the right hom $\homrallXYtoZ \alpha R
+P$. This acts very much like a polymorphic function type $\forall
+\alpha. R \to P$, but with restrictions on the object
+variables. First, the object variable $\alpha$ must occur covariantly
+in $P$ and contravariantly in $R$. Next, the contravariant variable of
+the type is the contravariant variable of $P$ and the covariant
+variable of the type is the \emph{contravariant} variable of $R$, with
+the flip in variance intuitively justified since $R$ occurs in the
+negative position of the function. The placement of the $\forall
+\alpha$ on the right side of the triangle is intended to be suggestive
+of the fact that $\alpha$ occurs \emph{covariantly} in the input and
+output sets $R,P$ and also that the corresponding $\lambda$ introduces
+variables on the right.
+%
+As you would expect from a polymorphic function type, the introduction
+form is a $\lambda$ that abstracts over both a transformation variable
+and an object variable, which are concatenated on the right of the
+context, aligning with the condition that the contravariant variable
+of $R$ match the covariant variable of the original context.
+%
+The elimination rule is a combined instantiation and application form,
+$\homrappXtoYatZ s t a$ which instantiates the object $\alpha$ to $a$
+and passes an element $t : R[a/\alpha]$ as input, yielding an element
+$P[a/\alpha]$ as output. Finally, we have the expected $\beta$ and
+$\eta$ reasoning principles that we would have for a function type.
+
+The left hom $\homlallXYtoZ \alpha R P$ dually abstracts over context
+extension on the left. This time, the variable $\alpha$ occurs
+\emph{contravariantly} in the input and output sets, and the covariant
+dependence of the function set matches the covariant dependence of the
+ouput set, while the contravariant dependence is the covariant
+dependence of the input. In our notation, we place the $\forall\alpha$
+on the left to be suggestive of the fact that it is contravariant in
+$R$ and $P$. Next, the introduction form is again a polymorphic
+$\lambda$, but this time we write the object variable $\alpha$ first,
+and then the set variable $x$, to match the structure of the
+context. The elimination form is similarly an application, but to
+agree with the ordered use of variables, this time we place the
+argument to the left and the function to the right.
 
@@
 
-      (\homrallXYtoZ \alpha R P)[\xi] = \homrallXYtoZ \alpha {R[d^+\xi;\alpha/\alpha]} {P[\alpha/\alpha;d^-\xi]}
-
       \inferrule*[right=Covariant Hom introduction]
-      {\Gamma \pipe \Phi, x : R, \alpha : \cat C \vdash M : P}
+      {\Gamma \pipe \Phi, x : R, \alpha : \cat C \vdash t : P}
       {\Gamma \pipe \Phi \vdash
-        \homrlambdaXatYdotZ {x : R} {\alpha : \cat C} M : \homrallXYtoZ \alpha R P}
-
-     (\homrlambdaXatYdotZ {x} {\alpha} M)[\phi]
-      = \homrlambdaXatYdotZ {x} {\alpha} M[\phi,x/x,\gamma/\gamma]
+        \homrlambdaXatYdotZ {x : R} {\alpha : \cat C} t : \homrallXYtoZ \alpha R P}
 
@@
       {\Phi = \Phi_f \jnctx \Phi_a\and
-       \Gamma\pipe\Phi_f \vdash M : \homrallXYtoZ \alpha R P \and
-       \Phi_a \vdash N : R[A/\alpha]
+       \Gamma\pipe\Phi_f \vdash s : \homrallXYtoZ {\alpha:\cat C} R P \and
+       d^+\Phi_a \vdash a : \cat C\and
+       \Phi_a \vdash t : R[a/\alpha]
         }
-      {\Gamma\pipe \Phi \vdash \homrappXtoYatZ M N A : P[A/\alpha]}\\
-
-      \homrappXtoYatZ {(\homrlambdaXatYdotZ x \alpha M)} N A = M[N/x,A/\alpha] (\beta)
+      {\Gamma\pipe \Phi \vdash \homrappXtoYatZ s t a : P[a/\alpha]}\\
+
+      \homrappXtoYatZ {(\homrlambdaXatYdotZ x \alpha s)} t a = s[t/x,a/\alpha] (\beta)
       %% (\lambda^\triangleleft (x, \gamma). M)\triangleleft (x, \gamma) = M (\beta)
@@
       \inferrule
-      {\Gamma\pipe\Phi, x:R, \alpha : \cat C \vdash \homrappXtoYatZ M x \alpha = \homrappXtoYatZ N x \alpha : P}
-      {\Gamma\pipe\Phi \vdash M = N : \homrallXYtoZ \alpha R P}
+      {\Gamma\pipe\Phi \vdash s : \homrallXYtoZ \alpha R P}
+      {s = \homrlambdaXatYdotZ {x}{\alpha} \homrappXtoYatZ s x \alpha}
+
+      %% \inferrule
+      %% {\Gamma\pipe\Phi, x:R, \alpha : \cat C \vdash \homrappXtoYatZ s x \alpha = \homrappXtoYatZ t x \alpha : P}
+      %% {\Gamma\pipe\Phi \vdash s = t : \homrallXYtoZ \alpha R P}
 
@@
       \inferrule*[right=Contra Hom introduction]
-      {\Gamma\pipe\alpha:\cat C, x : R, \Phi \vdash M : P}
-      {\Gamma\pipe\Phi \vdash \homllambdaXatYdotZ {x : R} {\alpha : \cat C} M : \homlallXYtoZ \alpha R P}
-
-     (\homllambdaXatYdotZ {x} {\alpha} M)[\phi]
-      = \homllambdaXatYdotZ {x} {\alpha} M[\phi,x/x,\gamma/\gamma]
+      {\Gamma\pipe\alpha:\cat C, x : R, \Phi \vdash t : P}
+      {\Gamma\pipe\Phi \vdash \homllambdaXatYdotZ {x : R} {\alpha : \cat C} t : \homlallXYtoZ \alpha R P}
 
@@
       {\Phi = \Phi_a \jnctx \Phi_f\and
-       \Gamma \pipe \Phi_a \vdash N : R[A/\alpha] \and
-       \Gamma \pipe \Phi_f \vdash M : \homlallXYtoZ \alpha R P}
-      {\Gamma \pipe \Phi \vdash \homlappXtoYatZ M N A : P[A/\alpha]}\\
+        \Gamma \pipe d^-\Phi_a \vdash a : \cat C\and
+       \Gamma \pipe \Phi_a \vdash s : R[a/\alpha] \and
+       \Gamma \pipe \Phi_f \vdash t : \homlallXYtoZ {\alpha:\cat C} R P}
+      {\Gamma \pipe \Phi \vdash \homlappXtoYatZ t s a : P[a/\alpha]}\\
 
@@
                  {}
-                 {\homlappXtoYatZ {(\homllambdaXatYdotZ x \alpha M)} N A = M[A/\alpha,N/x]}
+                 {\homlappXtoYatZ {(\homllambdaXatYdotZ x \alpha t)} s a = t[a/\alpha,s/x]}
 
       \inferrule*[right=$\eta$]
-      {\Gamma \pipe \alpha:\cat C, x:R, \Phi \vdash \homlappXtoYatZ M x \alpha = \homlappXtoYatZ N x \alpha : P}
-      {\Gamma \pipe \Phi \vdash M = N : \homlallXYtoZ \alpha R P}
+      {\Gamma \pipe \Phi \vdash t : \homlallXYtoZ \alpha R P}
+      {\Gamma \pipe \Phi \vdash t = \homllambdaXatYdotZ x \alpha \homlappXtoYatZ t x \alpha : \homlallXYtoZ \alpha R P}
+
+      %% \inferrule*[right=$\eta$]
+      %% {\Gamma \pipe \alpha:\cat C, x:R, \Phi \vdash \homlappXtoYatZ M x \alpha = \homlappXtoYatZ N x \alpha : P}
+      %% {\Gamma \pipe \Phi \vdash M = N : \homlallXYtoZ \alpha R P}
   \end{mathpar}
@@
 
-Hom judgments are useful not just in applications (Kan extensions, the
-Yoneda lemma), but also for purely syntactic reasons. In particular, I
-don't see how to give a sensible, complete natural deduction-style
-syntax for the other multiplicative connectives (Unit (i.e., the hom
-profunctor) and Tensor (profunctor composition) without relying on the
-presence of homs. The key property that the hom allows is that given
-any variable in context of a term $\Phi,\alpha,x,\beta,\Psi \vdash M :
-Q$ we can represent this term as a ``single-variable term'' by moving
-$\Phi$ and $\Psi$ to the right of the turnstile:
-
-\[ \alpha, x, \beta \vdash \lambda \Phi. \lambda \Psi M. : \Phi \triangleright Q \triangleleft \Psi \]
-
-This allows us to describe restricted positive eliminations that only
-work on single variable contexts. We give an admissible ``syntax
-sugar'' in Figure~\ref{fig:multi-hom}.
-
-Any $\Phi_l,\beta:C,\Phi_r \vdash M : R$ can be represented as a term
-$\beta:C \vdash M' : \Phi_r \triangleright_{\beta:C} R \triangleleft_{\beta:C} $
-
-Say we have a term in context
-\[ \alpha,x:P,\beta,y:Q,\gamma \vdash M : R \]
-Using $\triangleright\triangleleft$, we can produce an equivalent term
-of a different output type whose context is one of the single
-variables $\alpha,\beta$ or $\gamma$ in the context.
-
-We can make it $\alpha$ by using $\triangleleft$ twice:
-\[ \alpha \vdash \lambda^\triangleleft (x:P,\beta). \lambda^\triangleleft (y:Q, \gamma). M : \forall \beta. (\forall \gamma. R \triangleleft Q) \triangleleft P \]
-
-Or $\gamma$ by using $\triangleright$ twice:
-\[ \gamma \vdash\lambda^\triangleright (\beta, y:Q).  \lambda^\triangleright (\alpha, x:P). M : \forall \gamma. Q \triangleright \forall \alpha. P \triangleright R \]
-
-And we have two different ways to make the context $\beta$, either we use $\triangleright$ and then $\triangleleft$:
-\[ \beta \vdash\lambda^\triangleright (\alpha, x:P).  \lambda^\triangleleft (y:Q, \gamma). M : \forall \alpha. P \triangleright (\forall \gamma. R \triangleleft Q) \]
-or we do the opposite order:
-\[ \beta \vdash \lambda^\triangleleft (y:Q, \gamma). \lambda^\triangleright (\alpha, x:P). M : \forall \gamma. (\forall \alpha. P \triangleright R) \triangleleft Q \]
-
-In general we can iterate this process. That is given any $\Phi,\Psi$
-with $d^+\Phi = d^-\Psi = \beta:D$, we can take a term $\Phi \jnctx
-\Psi \vdash M : R$ and produce an equivalent term $\beta:D \vdash M :
-\Phi \triangleright R \triangleleft \Psi$. To formalize this, we show
-how we can create ``multi-hom'' jumbo connectives $\Phi \triangleright
-R$ and $R \triangleleft \Psi$, where the boundary conditions on
-$\Phi,\Psi$ are analogous to the ordinary case. These are given in
-Figure\ref{multi-hom}.
-
-\begin{figure}
-  \begin{align*}
-    P \triangleleft \gamma:E &= P\\
-    P \triangleleft (\Psi,x:R,\gamma:E) &= (\forall \gamma: E. P \triangleleft R) \triangleleft \Psi\\
-    %% \cdot \triangleright_{\alpha : C} P &= P\\
-    %% (x:R,\beta:D,\Phi_r)\triangleright_{\alpha:C} P  &= \Phi_R \triangleright_{\beta:D} \forall \beta. R \triangleright P
-  \end{align*}
-  \begin{mathpar}
-    \inferrule
-    {\Gamma \pipe d^+\Xi; \gamma: E \vdash P \isaSet\and
-     \Gamma \pipe d^-\Xi; \gamma:E \vdash \Psi \isavectx}
-    {\Gamma \pipe \Xi \vdash P \triangleleft \Psi}
-
-    \inferrule*[right=Covariant Multi-Hom introduction]
-    {\Gamma\pipe\Phi\jnctx \Psi \vdash M : P}
-    {\Gamma\pipe\Phi \vdash \lambda^\triangleleft \Psi. M : P \triangleleft \Psi}\and
-
-    \inferrule*[right=Covar Multi-Hom elim]
-    {~}
-    {\Gamma \pipe f:P\triangleleft \Psi,\Psi \vdash \textrm{app}\triangleleft_\Psi : P}
-
-    %%   (\lambda^\triangleleft \Psi. M)[\phi] = \lambda^\triangleleft \Psi. M[\phi,\Psi]\\
-    %%   \lambda^\triangleleft \gamma. M = M\\
-    %%   \lambda^\triangleleft (\Psi,x:R,\gamma:\cat E). M = \lambda^\triangleleft \Psi. \lambda^\triangleleft (x:R,\gamma:\cat E). M\\
-
-    %%   (\lambda^\triangleleft \gamma. M)[\phi] = M[\phi] = \lambda^\triangleleft \gamma. M[\phi]\\
-    %%   (\lambda^\triangleleft (\Psi,x:r,\gamma:\cat E). M)[\phi] =
-    %%   (\lambda^\triangleleft \Psi. \lambda^\triangleleft (x:R,\gamma:\cat E). M)[\phi]
-    %%   = \lambda^\triangleleft \Psi. (\lambda^\triangleleft (x:R,\gamma:\cat E). M)[\phi, \Psi]
-    %%   = \lambda^\triangleleft \Psi. \lambda^\triangleleft (x:R,\gamma:\cat E). M[\phi,\Psi,x:R,\gamma]
-
-    %%   \inferrule*[right=Covariant Multi-Hom elimination]
-    %%   {\Phi_1 \vdash M : P\triangleleft \Phi_2 \and \phi_2 : \Psi \vdash \Phi_2 \subst(\xi)}
-    %%   {\Phi_1,\Psi \vdash M\triangleleft \phi_2 : P[d^+\xi]}\\
-
-    %%   M \triangleleft (A/\gamma) = M[A/\gamma]\\
-    %%   M \triangleleft (\phi, N/x,A/\gamma) =  M \triangleleft (N, A) \triangleleft \phi
-
-    %%   (M \triangleleft \phi_2)[\phi_1,\psi] = M[\phi_1] \triangleleft [\phi_2[\psi]]\\
-    %%   TODO: prove
-
-    %%   TODO: (\lambda^\triangleleft \Phi. M) \triangleleft \phi = M[\phi]
-    %%   TODO: M = \lambda^\triangleleft \Phi. M\triangleleft \Phi
-
-    %%   TODO: contravariant by duality
-    %%   %% \inferrule*[right=Contra Multi-Hom formation]
-    %%   %% {\Phi \isavectx(\alpha:\cat C; d^-\Xi) \and P \prof(\alpha : \cat C; d^+\Xi)}
-    %%   %% {\Phi \triangleright P \prof(\xi)}
-
-    %%   %% \alpha \triangleright P = P\\
-    %%   %% \alpha:\cat C, x:R, \Phi \triangleright P = \forall \alpha. R \triangleright \Phi \triangleright P\\
-  \end{mathpar}
-  \begin{align*}
-    \lambda^\triangleleft \gamma:E. M &= M\\
-    \lambda^\triangleleft (\Psi,x:R,\gamma:E). M &= \lambda^\triangleleft \Psi. \lambda^\triangleleft (x:R,\gamma:E). M\\\\
-    \textrm{app}\triangleleft_{\gamma:E} &= f\\
-    \textrm{app}\triangleleft_{\Psi,x:R\gamma:E} &= \textrm{app}\triangleleft_\Psi \triangleleft (x,\gamma) \\
-  \end{align*}
-  \caption{Multi-Hom}
-  \label{multi-hom}
-\end{figure}
-
-\subsection{Unit and Tensor Profunctors}
-
-While the cotensors of the previous section work quite smoothly since
-they are negative connectives in an intuitionistic-like (many formulas
-on the left of the sequent and only one on the right), the positive
-connectives, the unit and tensor, are trickier.
-
-First, to internalize $\cdot_{\alpha : \cat C}$ we can define the
-\emph{unit} $A \to_{\cat C} A'$, which in set-theoretic settings we
+In category theory, these profunctors can be constructed using an
+\emph{end} and the hom of functor categories\cite{somecats}. Given two
+profunctors $P : C^o \times E \to \Set$ and $R : D^o \times E \to
+Set$, their hom is given as
+\[ \int_{e \in E} R(-,e) \Rightarrow P(e,=) \]
+
+%% The syntax is given in Figure~\ref{fig:cotensor}.
+%% %
+%% First, the formation rule for covariant cotensor formation requires
+%% that the covariant variables for $R,P$ are the same, and the
+%% contravariant variables match the boundary of the overall
+%% judgment. Note that we use $d^+\Xi,d^-\Xi$ to allow for the case that
+%% the boundary is a single variable $\Xi = \alpha:\cat C$.
+%% %
+%% Next, the substitution rule is straightforward.
+%% %
+%% Then we have the elimination form, first written in a simple form with
+%% no substitution built in.
+%% %
+%% This says that given a variable of the cotensor judgment, and another
+%% variable of judgment $R$, we can apply these to get a term of judgment
+%% $P$. As a diagram this is written as follows:
+
+%% \[ % https://tikzcd.yichuanshen.de/#N4Igdg9gJgpgziAXAbVABwnAlgFyxMJZABgBpiBdUkANwEMAbAVxiRAGEQBfU9TXfIRQBGclVqMWbACLdeIDNjwEiAJjHV6zVohABROXyWCiZYeK1TdnHkYEqU685sk793cTCgBzeEVAAZgBOEAC2SGQgOBBIohLabAA6iQEQQYwMAATJ3nShoXQAdJkACtmJOEFYdGDeDDD1ATiZAEqGIMFhsdTRSOrxViBtth0h4YgAzD0xiAAsLgm6Je2d45G9k9QMWGBuUBBMAEb1INQAFjB0UGyQu6dRdFgMNwSsI6t900jzIBdXL3ctjs9gdjqweo9nrpbm8KFwgA
+%% \begin{tikzcd}
+%% C \arrow[r, "\forall \gamma. P \triangleleft R"] \arrow[d, no head, Rightarrow] & D \arrow[r, "R"] & E \arrow[d, no head, Rightarrow] \\
+%% C \arrow[rr, "P"]                                                               &                  & E                               
+%% \end{tikzcd}\]
+
+%% The universal property of this judgment is that this is the universal
+%% 2-cell of this type, in that any 2-cell
+
+%% \[ % https://tikzcd.yichuanshen.de/#N4Igdg9gJgpgziAXAbVABwnAlgFyxMJZABgBpiBdUkANwEMAbAVxiRAGEByEAX1PUy58hFACZyVWoxZsAIr34gM2PASIBmCdXrNWiEAFEFAlcKJkAjJJ0z97Y0sGqRyTVe3S9hh8qFqUFlpSumy8kjBQAObwRKAAZgBOEAC2SIEgOBBI4sG2IABKDokpSJoZWYgALB4h+gAKRUmpiGTlpdQMWGBecBCdUCDUABYwdAOIYEwMDNQ4dFgMbJDdjSWIOZlI1SAjY0sErB1dXlAQTABGDIcZ84v6y6x88U1IrZuIAKxPIMXNH7MVdKdFb6KB0OAjAbDUbjcAHMI8IA
+%% \begin{tikzcd}
+%% C' \arrow[d] \arrow[r] & {} \arrow[r, no head, dashed] & D \arrow[r, "R"] & E \arrow[d, no head, Rightarrow] \\
+%% C \arrow[rrr, "P"]     &                               &                  & E                               
+%% \end{tikzcd}\]
+
+%% Exhibits a unique factorization:
+
+%% \[ % https://tikzcd.yichuanshen.de/#N4Igdg9gJgpgziAXAbVABwnAlgFyxMJZABgBoBGAXVJADcBDAGwFcYkQBhEAX1PU1z5CKAEwVqdJq3YARHnxAZseAkQDM4mgxZtEIAKLz+yoUTIiJ26Xq69jg1Sg0WtU3QaOKBK4SVLFLN3YOAHJPJQdfcn9AnXZw71NRGNc4vTk7LxNHZA0A1OsPbgkYKABzeCJQADMAJwgAWyRokBwIJDFJNJAAJU86xqQNVvbEABYC9wAFfvqmxDIRoZpGLDB3KAhmACNGNhoACxh6KHZIdZAaHHosRjOCNkyB+c62pAmQI5P7i5W1ja2u32rRudz050eCmeSAArFdRgA2J5zJAI+FIADsfwueig9DgR1Oh2Op3BD1mg0QWKWiBaqxxIE2Oz2l0+JJ+wOutw5FPmcJpxGRlMWb1pk3YAB0JdUILUmIwAARSsr0BoNegAOgVUyVEpwtSw9DAZT2e2qOAVfSF82pooAHOK9FaoSjEA6aZ16QDmcCvqTwOSrqCecVuEA
+%% \begin{tikzcd}
+%% C' \arrow[r] \arrow[d]                                                           & {} \arrow[r, no head, dashed] & D \arrow[d, no head, Rightarrow] \arrow[r, "R"] & E \arrow[d, no head, Rightarrow] \\
+%% C \arrow[d, no head, Rightarrow] \arrow[rr, "\forall \gamma.
+P \triangleleft R"] &                               & D \arrow[r, "R"]                                & E \arrow[d, no head, Rightarrow] \\
+%% C \arrow[rrr, "P"]                                                               &                               &                                                 & E                               
+%% \end{tikzcd}\]
+
+%% We give elimination rules and introduction rules in a style similar to
+%% $\lambda$-calculus, with corresponding $\beta\eta$ rules.
+
+These hom types are useful both for category-theoretic applications
+(Kan extensions, the Yoneda lemma), but also for purely syntactic
+reasons.  The universal property of these function types allows us, by
+iterated hom to express any sequent $\Phi \vdash R$ as an equivalent
+one with an ``empty'' context, i.e., a single object variable $d^\pm
+\Phi$. Then using the end type defined in the previous section, we can
+internalize that judgment as a type in our ambient dependent type
+theory.
+
+As an example, any sequent $\Gamma \pipe \alpha, x: R, \beta, y : S, \gamma \vdash
+P$ is equivalent to either of
+\[ \Gamma \vdash \pendallXdotY\alpha \homrallXYtoZ \beta R \homrallXYtoZ \gamma S P 
+\qquad\qquad\qquad \Gamma \vdash \pendallXdotY \gamma \homlallXYtoZ \beta S {\homlallXYtoZ \alpha R P}\]
+This property is useful in giving a simple syntax to positive types.
+
+\subsection{Unit and Tensor of Profunctors}
+
+While the hom sets of the previous section work quite smoothly since
+they are negative connectives in a natural deduction system, the positive
+connectives, the unit and tensor, are trickier to define.
+
+First, the multiplicative unit is the set $\punitinXfromYtoZ {\cat C}
+a {a'}$, that is the hom set of the category $\cat C$. The rules for
+this type act like .
+
+we can define the
+\emph{unit} , which in set-theoretic settings we
 might call the \emph{identity/equality judgment} $A \equiv_{\cat C}
@@
 
-    \inferrule
-    {\Gamma \pipe \alpha:\cat C\vdash R \isaSet}
-    {\Gamma \pipe \cdot \vdash \pendallXdotY {\alpha:{\cat C}} R \isaSet}
-
-    \inferrule
-    {\Gamma \pipe \alpha:\cat C\vdash M : R}
-    {\Gamma \pipe \cdot \vdash \pendlambdaXdotY \alpha M : \pendallXdotY \alpha R}
-
-    \inferrule
-    {\Gamma\pipe\cdot \pipe \cdot \vdash M : \forall \alpha:\cat C. R\and
-      \Gamma \pipe \beta:\cat D \vdash A : \cat C}
-    {\Gamma \pipe \beta:\cat D\pipe\beta:\cat D \vdash \pendappXtoY M A : R[A/\alpha]}
-
-    \pendappXtoY {(\pendlambdaXdotY \alpha M)} A = M[A/\alpha]
-
-    \inferrule*[right=2Cell$\eta$]
-    {\Gamma \pipe \cdot \vdash M : \pendallXdotY \alpha R }
-    {\Gamma \pipe \cdot \vdash M = \pendlambdaXdotY \alpha {\pendappXtoY M \alpha} : \pendallXdotY \alpha R }
   \end{mathpar}
@@
 
-TODO: define substitution
+TODO: complete substitution
+\begin{figure}
+  \begin{mathpar}
+      (\homrallXYtoZ \alpha R P)[\xi] = \homrallXYtoZ \alpha {R[d^+\xi;\alpha/\alpha]} {P[\alpha/\alpha;d^-\xi]}
+
+     (\homrlambdaXatYdotZ {x} {\alpha} M)[\phi]
+      = \homrlambdaXatYdotZ {x} {\alpha} M[\phi,x/x,\gamma/\gamma]
+
+     (\homllambdaXatYdotZ {x} {\alpha} M)[\phi]
+      = \homllambdaXatYdotZ {x} {\alpha} M[\phi,x/x,\gamma/\gamma]
+
+  \end{mathpar}
+  \caption{Substitution}
+
+\end{figure}
+
 \begin{figure}
@@
 
+The key property that the hom allows is that given
+any variable in context of a term $\Phi,\alpha,x,\beta,\Psi \vdash M :
+Q$ we can represent this term as a ``single-variable term'' by moving
+$\Phi$ and $\Psi$ to the right of the turnstile:
+
+\[ \alpha, x, \beta \vdash \lambda \Phi. \lambda \Psi M. : \Phi \triangleright Q \triangleleft \Psi \]
+
+This allows us to describe restricted positive eliminations that only
+work on single variable contexts. We give an admissible ``syntax
+sugar'' in Figure~\ref{fig:multi-hom}.
+
+Any $\Phi_l,\beta:C,\Phi_r \vdash M : R$ can be represented as a term
+$\beta:C \vdash M' : \Phi_r \triangleright_{\beta:C} R \triangleleft_{\beta:C} $
+
+Say we have a term in context
+\[ \alpha,x:P,\beta,y:Q,\gamma \vdash M : R \]
+Using $\triangleright\triangleleft$, we can produce an equivalent term
+of a different output type whose context is one of the single
+variables $\alpha,\beta$ or $\gamma$ in the context.
+
+We can make it $\alpha$ by using $\triangleleft$ twice:
+\[ \alpha \vdash \lambda^\triangleleft (x:P,\beta). \lambda^\triangleleft (y:Q, \gamma). M : \forall \beta. (\forall \gamma. R \triangleleft Q) \triangleleft P \]
+
+Or $\gamma$ by using $\triangleright$ twice:
+\[ \gamma \vdash\lambda^\triangleright (\beta, y:Q).  \lambda^\triangleright (\alpha, x:P). M : \forall \gamma. Q \triangleright \forall \alpha. P \triangleright R \]
+
+And we have two different ways to make the context $\beta$, either we use $\triangleright$ and then $\triangleleft$:
+\[ \beta \vdash\lambda^\triangleright (\alpha, x:P).  \lambda^\triangleleft (y:Q, \gamma). M : \forall \alpha. P \triangleright (\forall \gamma. R \triangleleft Q) \]
+or we do the opposite order:
+\[ \beta \vdash \lambda^\triangleleft (y:Q, \gamma). \lambda^\triangleright (\alpha, x:P). M : \forall \gamma. (\forall \alpha. P \triangleright R) \triangleleft Q \]
+
+In general we can iterate this process. That is given any $\Phi,\Psi$
+with $d^+\Phi = d^-\Psi = \beta:D$, we can take a term $\Phi \jnctx
+\Psi \vdash M : R$ and produce an equivalent term $\beta:D \vdash M :
+\Phi \triangleright R \triangleleft \Psi$. To formalize this, we show
+how we can create ``multi-hom'' jumbo connectives $\Phi \triangleright
+R$ and $R \triangleleft \Psi$, where the boundary conditions on
+$\Phi,\Psi$ are analogous to the ordinary case. These are given in
+Figure\ref{multi-hom}.
+
+\begin{figure}
+  \begin{align*}
+    P \triangleleft \gamma:E &= P\\
+    P \triangleleft (\Psi,x:R,\gamma:E) &= (\forall \gamma: E. P \triangleleft R) \triangleleft \Psi\\
+    %% \cdot \triangleright_{\alpha : C} P &= P\\
+    %% (x:R,\beta:D,\Phi_r)\triangleright_{\alpha:C} P  &= \Phi_R \triangleright_{\beta:D} \forall \beta. R \triangleright P
+  \end{align*}
+  \begin{mathpar}
+    \inferrule
+    {\Gamma \pipe d^+\Xi; \gamma: E \vdash P \isaSet\and
+     \Gamma \pipe d^-\Xi; \gamma:E \vdash \Psi \isavectx}
+    {\Gamma \pipe \Xi \vdash P \triangleleft \Psi}
+
+    \inferrule*[right=Covariant Multi-Hom introduction]
+    {\Gamma\pipe\Phi\jnctx \Psi \vdash M : P}
+    {\Gamma\pipe\Phi \vdash \lambda^\triangleleft \Psi. M : P \triangleleft \Psi}\and
+
+    \inferrule*[right=Covar Multi-Hom elim]
+    {~}
+    {\Gamma \pipe f:P\triangleleft \Psi,\Psi \vdash \textrm{app}\triangleleft_\Psi : P}
+
+    %%   (\lambda^\triangleleft \Psi. M)[\phi] = \lambda^\triangleleft \Psi. M[\phi,\Psi]\\
+    %%   \lambda^\triangleleft \gamma. M = M\\
+    %%   \lambda^\triangleleft (\Psi,x:R,\gamma:\cat E). M = \lambda^\triangleleft \Psi. \lambda^\triangleleft (x:R,\gamma:\cat E). M\\
+
+    %%   (\lambda^\triangleleft \gamma. M)[\phi] = M[\phi] = \lambda^\triangleleft \gamma. M[\phi]\\
+    %%   (\lambda^\triangleleft (\Psi,x:r,\gamma:\cat E). M)[\phi] =
+    %%   (\lambda^\triangleleft \Psi. \lambda^\triangleleft (x:R,\gamma:\cat E). M)[\phi]
+    %%   = \lambda^\triangleleft \Psi. (\lambda^\triangleleft (x:R,\gamma:\cat E). M)[\phi, \Psi]
+    %%   = \lambda^\triangleleft \Psi. \lambda^\triangleleft (x:R,\gamma:\cat E). M[\phi,\Psi,x:R,\gamma]
+
+    %%   \inferrule*[right=Covariant Multi-Hom elimination]
+    %%   {\Phi_1 \vdash M : P\triangleleft \Phi_2 \and \phi_2 : \Psi \vdash \Phi_2 \subst(\xi)}
+    %%   {\Phi_1,\Psi \vdash M\triangleleft \phi_2 : P[d^+\xi]}\\
+
+    %%   M \triangleleft (A/\gamma) = M[A/\gamma]\\
+    %%   M \triangleleft (\phi, N/x,A/\gamma) =  M \triangleleft (N, A) \triangleleft \phi
+
+    %%   (M \triangleleft \phi_2)[\phi_1,\psi] = M[\phi_1] \triangleleft [\phi_2[\psi]]\\
+    %%   TODO: prove
+
+    %%   TODO: (\lambda^\triangleleft \Phi. M) \triangleleft \phi = M[\phi]
+    %%   TODO: M = \lambda^\triangleleft \Phi. M\triangleleft \Phi
+
+    %%   TODO: contravariant by duality
+    %%   %% \inferrule*[right=Contra Multi-Hom formation]
+    %%   %% {\Phi \isavectx(\alpha:\cat C; d^-\Xi) \and P \prof(\alpha : \cat C; d^+\Xi)}
+    %%   %% {\Phi \triangleright P \prof(\xi)}
+
+    %%   %% \alpha \triangleright P = P\\
+    %%   %% \alpha:\cat C, x:R, \Phi \triangleright P = \forall \alpha. R \triangleright \Phi \triangleright P\\
+  \end{mathpar}
+  \begin{align*}
+    \lambda^\triangleleft \gamma:E. M &= M\\
+    \lambda^\triangleleft (\Psi,x:R,\gamma:E). M &= \lambda^\triangleleft \Psi. \lambda^\triangleleft (x:R,\gamma:E). M\\\\
+    \textrm{app}\triangleleft_{\gamma:E} &= f\\
+    \textrm{app}\triangleleft_{\Psi,x:R\gamma:E} &= \textrm{app}\triangleleft_\Psi \triangleleft (x,\gamma) \\
+  \end{align*}
+  \caption{Multi-Hom}
+  \label{multi-hom}
+\end{figure}
+
 %% \section{Styles of lists, enumerations, and descriptions}\label{sec:itemStyles}

### 2022-02-06 paper/lipics-v2021-sample-article.tex
synthetic progress

@@
 \newcommand{\tensorintroofXandY}[2]{\textrm{pair}_\odot(#1,#2)}
+\newcommand{\tensorintroatXwithYandZ}[3]{(#2,#1,#3)}
 \newcommand{\tensorelimXtoYwithkontZ}[3]{\textrm{ind}_{\to}(#1,#2,#3)}
@@
 
-\newcommand{\homlallXYtoZ}[3]{#3 \tlforall #1 #2}
+\newcommand{\homlallXYtoZ}[3]{#3 \tlforall {#1} #2}
 
@@
 \newcommand{\homunary}[4]{\pendallXdotY {#1} {\homrallXYtoZ {#3} {#2} {#4}}}
+\newcommand{\hombinary}[6]{\pendallXdotY {#1} {\homrallXYtoZ {#3} {#2} \homrallXYtoZ {#5} {#4} {#6}}}
 
@@
       \textrm{Type Context} & \Gamma, \Delta & ::= & \cdot \pipe \Gamma, X : A\\
-      \textrm{Category Context} & \Xi & ::= & \alpha:\cat C \pipe \alpha:\cat C;\beta: \cat D\\
+      \textrm{Category Context} & \Xi, Z & ::= & \alpha:\cat C \pipe \alpha:\cat C;\beta: \cat D\\
       \textrm{Set Context} & \Phi,\Psi & ::= & \alpha : \cat C \pipe \Phi , x : P , \Psi\\
@@
 First, the multiplicative unit is the set $\punitinXfromYtoZ {\cat C}
-a {a'}$, that is the hom set of the category $\cat C$. The rules for
-this type act like .
-
-we can define the
-\emph{unit} , which in set-theoretic settings we
-might call the \emph{identity/equality judgment} $A \equiv_{\cat C}
-A'$.
-%
-First the formation rule says given two objects $A,A'$ in the same
-category $\cat C$, we can form the hom-set from $A$ to $A'$ $A
-\to_{\cat C} A'$.
-%
-The introduction rule says that given any object $A$, we can define
-the identity arrow from $A$ to itself $\id(A) : A \to_{\cat C} A$.
-%
-Without knowing more about the category in question, these are the
-only arrows we know exist.
-%
-
-The elimination form is trickiest to define in a natural deduction style.
-%
-First, let's consider the universal property of the Hom. From
-Crutwell-Shulman, we want a 2-cell
-
-\[% https://tikzcd.yichuanshen.de/#N4Igdg9gJgpgziAXAbVABwnAlgFyxMJZARgBoAGAXVJADcBDAGwFcYkQBBEAX1PU1z5CKcqQBM1Ok1bsuvftjwEiY8ZIYs2iTjz4gMioUTLF10rSABOMAGaMekmFADm8IqBuWIAWyRkQOBBIqlKa7AASProeXr6IogFBiP6MWGAWUBDMAEaMbDQAFjD0UOyQ6SA0OPRY9trlbPIgnlHxVUkhqRXamTl5lSBFJWUE+QE1deCjDtxAA
-\begin{tikzcd}
-                    & A \arrow[ldd, no head, Rightarrow] \arrow[rdd, no head, Rightarrow] &   \\
-                    & refl                                                                &   \\
-A \arrow[rr, "Hom"] &                                                                     & A
-\end{tikzcd}\]
-
-That is universal/opcartesian in that for any 2-cell $t$:
-
-\[ % https://tikzcd.yichuanshen.de/#N4Igdg9gJgpgziAXAbVABwnAlgFyxMJZABgBpiBdUkANwEMAbAVxiRACEQBfU9TXfIRQBGclVqMWbbrxAZseAkQBMY6vWatEIAIIy+CwUQAsaiZrZ6eBgUpQA2Mxqnb2Acn1z+ioclXLxZy0QAGFPeVtfUwD1SWCAEXDvIxQAZlJhQLi2HA8ucRgoAHN4IlAAMwAnCABbJDIQHAgkYWsQKtqW6iakVRAGLDBgqDo4AAtCkGoJuig2SCHPDrrEPp7EdP7B4LgIAbnpmFmkMCYGBmoGOgAjGAYABWS7fphynCnzFxAACU625aQm3Wpi2i20I3Gk0Ox20C1Y-2qKwa6wArAjOogQet7OiVijus1EI5PsEAErcChcIA
-\begin{tikzcd}
-B \arrow[r] \arrow[rrdd] & {} \arrow[r, no head, dashed] & A \arrow[rr, "Hom"] &    & A \arrow[rr, no head, dashed] &  & B' \arrow[lldd] \\
-                         &                               &                     & t' &                               &  &                 \\
-                         &                               & C \arrow[rr, "R"]   &    & D                             &  &                
-\end{tikzcd}\]
-
-there exists a $t'$ which extends $t$ in that the following is equal
-to $t$:
-\[% https://tikzcd.yichuanshen.de/#N4Igdg9gJgpgziAXAbVABwnAlgFyxMJZABgBoAmAXVJADcBDAGwFcYkQAhEAX1PU1z5CKAIwVqdJq3Y8+IDNjwEi5cTQYs2iEAEFZ-RUKIAWNZM3s9vA4OUoAbGY3TtHAOT75ApcOSrjEs5aIADCngq2vqYB6lLBACLh3kYoAMykqYFx7Dge1l6GdsjpxFkW2lZyET5EZKWx5SBJhb5i9eYuTfnVKciO7UEy3clF6SJlnQBOMABmjDwSMFAA5vBEoDOTEAC2SGQgOBBIIvmbO8c0h0iqIIxYYMFQ9HAAFksgNG-0UOyQD55nXaIG5XRDpW73YJwCB3H6fGDfJBgZiMRg0Rj0ABGMEYAAURsJbrMcB8OsEABLnU5bIHg0GmCH-bRPV7veGI7R-NjU86IfaggCsPKBDNB9mFSAFlyOiEcZPYACUATSkAAOaXXdGQ9hQCDMTGMNjsn6cghGg70LDzU3-CWIdUHGXgu5MkC6-WG0lfE3gM2knCW62+21yQFIACcGr5dsjjuO+xdUJhWDhIG9SJRaItVt+frtIn5MrEjKTsK9CJNyNRl0DuZDGxViALUYdiZ1zzeqfTNu5ocbDtBImLbeZHbZaYrdd7Dd5Q6jDJHbr1BvN3eD5oDOZ7C24QA
-\begin{tikzcd}
-{} \arrow[dd] \arrow[r]  & {} \arrow[dd] \arrow[rr, no head, dashed] &                     & A \arrow[ldd, no head, Rightarrow] \arrow[rdd, no head, Rightarrow] \arrow[rrr, no head, dashed] &                               &  & {} \arrow[dd, no head, Rightarrow] \\
-                         &                                           &                     & refl                                                                                             &                               &  &                                    \\
-B \arrow[r] \arrow[rrdd] & {} \arrow[r, no head, dashed]             & A \arrow[rr, "Hom"] &                                                                                                  & A \arrow[rr, no head, dashed] &  & B' \arrow[lldd]                    \\
-                         &                                           &                     & t'                                                                                               &                               &  &                                    \\
-                         &                                           & C \arrow[rr, "R"]   &                                                                                                  & D                             &  &
-\end{tikzcd}\]
-
-Clearly, our introduction rule models this first 2-cell. Furthermore,
-the construction in the second condition is modeled faithfully by the
-following sequent rule in the figure. This is a J-like induction rule:
-to construct a term using an input of the Hom type, it is sufficient
-to consider the case in which the input is the identity morphism.
-%
-However, building in a substitution the usual way we would (pushing
-the substitution into the continuation $N$) doesn't work properly
-because we can't necessarily extract a substitution for the
-continuation from the substitution for the the whole term. A
-substitution for the whole term would be of the form
-\[\phi,D_1/\delta_1,M/f,D_2/\delta_2,\psi\]
-The the candidate for the continuation's substitution would be
-$\phi,?,\psi$ but there is not necessarily something to go in the
-middle here since $D_1,D_2$ do not necessarily have the same
-\emph{input variable}. However, if we assume the existence already of
-the cotensor products the problem can be simplified. Instead of
-allowing for a general elimination of a hom term anywhere in the
-context, we can instead make a remarkably restricted rule: we can only
-eliminate a Hom term when it is the \emph{sole} input in the context:
-
-\[ \inferrule*[right=Hom Elim Very Simple]{\alpha:\cat C \vdash N : Q[\alpha/\alpha_1,\alpha/\alpha_2]}{\alpha_1:\cat C, f : \alpha_1 \to_{\cat C} \alpha_2, \alpha_2:\cat C \vdash J(\alpha_1.\alpha_2.
-Q, \alpha.
-N, \alpha_1, f, \alpha_2) : Q}\]
-
-While this rule looks very impoverished at first sight, it has the
-virtue of easily supporting a built-in cut:
-
-\[ \inferrule*[right=Hom Elim VSimple Subst]{
-  Q \prof(\alpha_1:\cat C; \alpha_2:\cat C)\and
-  \Phi \vdash M : A_1 \to_{\cat C} A_2\and
-  \alpha:\cat C \vdash N : Q[\alpha/\alpha_1,\alpha/\alpha_2]}{\Phi \vdash J(\alpha_1.\alpha_2. Q, \alpha. N, A_1, M, A_2) : Q[A_1/\alpha_1, A_2/\alpha_2]}\]
-
-However, this rule on its own clearly is not enough to support our
-desired universal property shown above diagrammatically. To prove our
-completeness theorem, we will need to show that this rule is enough to
-give us the desired universal property \emph{assuming the presence of
-  left and right cotensors}.
-
-So first, we must show that the ``proper'' sequent rule
-
-\[ \inferrule*[right=Hom-Left (Sequent Style)]
-    {\Phi_1[\alpha/\alpha_1], \alpha:\cat C, \Phi_2[\alpha/\alpha_2]\vdash Q[\alpha/\alpha_1,\alpha/\alpha_2]}
-    {\Phi_1, \alpha_1 : \cat C, f : \alpha_1 \to_{\cat C} \alpha_2, \alpha_2 : \cat C, \Phi_2\vdash : Q}
-\]
-
-is admissible.
-
-Which we can sketch as follows:
-
-\[
-\inferrule*
-    {\inferrule*{\inferrule*{\Phi_1[\alpha/\alpha_1],\alpha:\cat C, \Phi_2[\alpha/\alpha_1] \vdash Q[\alpha/\alpha_1;\alpha/\alpha_2]}{\alpha:\cat C \vdash \Phi_1[\alpha/\alpha_1] \triangleright Q[\alpha/\alpha_1;\alpha/\alpha_2] \triangleleft \Phi_2[\alpha/\alpha_2]}}{\alpha_1 : \cat C, f : \alpha_1 \to_{\cat C} \alpha_2, \alpha_2 : \cat C\vdash \Phi_1 \triangleright Q \triangleleft \Phi_2}}
-    {\Phi_1, \alpha_1 : \cat C, f : \alpha_1 \to_{\cat C} \alpha_2, \alpha_2 : \cat C, \Phi_2\vdash Q}
-\]
-
-Which using our syntax sugar is given by the term:
-\begin{mathpar}
-  \inferrule*
-    {\inferrule*
-      {\inferrule*
-        {\Phi_1[\alpha/\alpha_1],\alpha:\cat C, \Phi_2[\alpha/\alpha_1] \vdash N : Q[\alpha/\alpha_1;\alpha/\alpha_2]}
-        {\alpha:\cat C \vdash \lambda^\triangleright \Phi_1[\alpha/\alpha_1].
-\lambda^\triangleleft \Phi_2[\alpha/\alpha_2].
-N : \Phi_1[\alpha/\alpha_1] \triangleright Q[\alpha/\alpha_1;\alpha/\alpha_2] \triangleleft \Phi_2[\alpha/\alpha_2]}}
-      {\alpha_1 : \cat C, f : \alpha_1 \to_{\cat C} \alpha_2, \alpha_2 : \cat C\vdash J(\alpha_1.\alpha_2.\Phi_1 \triangleright Q \triangleleft \Phi_2, \alpha.
-\lambda^\triangleright \Phi_1[\alpha/\alpha_1].
-\lambda^\triangleleft \Phi_2[\alpha/\alpha_2].
-N, \alpha_1,f,\alpha_2) : \Phi_1 \triangleright Q \triangleleft \Phi_2}}
-    {\Phi_1, \alpha_1 : \cat C, f : \alpha_1 \to_{\cat C} \alpha_2, \alpha_2 : \cat C, \Phi_2\vdash
-      \Phi_1 \triangleright (J(\alpha_1.\alpha_2.\Phi_1 \triangleright Q \triangleleft \Phi_2, \alpha.
-\lambda^\triangleright \Phi_1[\alpha/\alpha_1].
-\lambda^\triangleleft \Phi_2[\alpha/\alpha_2].
-N, \alpha_1,f,\alpha_2)) \triangleleft \Phi_2 : Q}
-\end{mathpar}
-
-We need to show $\beta$ and $\eta$ hold.
-
-First, $\beta$.
-
-\begin{align*}
-  &(\Phi_1 \triangleright (J_\to(\alpha_1.\alpha_2.\Phi_1 \triangleright Q \triangleleft \Phi_2, \alpha.
-\lambda^\triangleright \Phi_1[\alpha/\alpha_1].
-\lambda^\triangleleft \Phi_2[\alpha/\alpha_2].
-N, \alpha_1,f,\alpha_2)) \triangleleft \Phi_2) [\Phi_1[A/\alpha_1],A/\alpha_1,\id(A),A/\alpha_2,\Phi_2[A/\alpha_2]] \\
-  &=
-  \Phi_1[A/\alpha_1] \triangleright (J_\to(\alpha_1.\alpha_2.\Phi_1 \triangleright Q \triangleleft \Phi_2, \alpha.
-\lambda^\triangleright \Phi_1[\alpha/\alpha_1].
-\lambda^\triangleleft \Phi_2[\alpha/\alpha_2].
-N, A,\id(A), A)) \triangleleft \Phi_2[A/\alpha_2]\\
-  &=
-  \Phi_1[A/\alpha_1] \triangleright (\lambda^\triangleright \Phi_1[A/\alpha_1]. \lambda^\triangleleft \Phi_2[A/\alpha_2]. N[A/\alpha]) \triangleleft \Phi_2[A/\alpha_2]\\
-  &=  (\lambda^\triangleleft \Phi_2[A/\alpha_2]. N[A/\alpha]) \triangleleft \Phi_2[A/\alpha_2]\\
-  &=  N[A/\alpha]
-\end{align*}
+{a_1} {a_2}$, semantically, the hom set of the category $\cat C$. The
+rules for this type act like an identity type in a directed type
+theory. The introduction rule defines an identity morphism of an
+object in any category. The elimination rule is then the appropriate
+analogue of the $J$ rule of the identity type. Given any profunctor
+$Q$ from $\cat C$ to itself, we can construct an element $\harrapp Q
+{\alpha_1}{\alpha_2}$ for any objects in $\cat C$ if we specify a
+natural element $\harrapp Q \alpha \alpha$ for any $\alpha$ in $\cat
+C$. The reason this is sufficient, semantically, is that all
+profunctors must be functorial in their two indices. In fact, as we
+will see in the section on synthetic category theory, this induction
+principle is how we can show internally that profunctors are
+functorial in their indices. We choose to express this elimination
+form using the right hom type as its syntax more closely matches the
+usual syntax of function types, but this choice is arbitrary, and we
+would get an equivalent calculus had we used the left hom, or added
+both.
+
+Next, we continue with the tensor product $\tensorexistsXwithYandZ
+\beta P Q$. Adjoint to the right/left hom sets, this provides a form
+of \emph{existentially quantified pair}. The formation rule states the
+set is well-formed when $\beta$ is the covariant variable of $P$ and
+the contravariant variable of $Q$, and, its contravariant dependence
+is the same as $P$, while its covariant dependence is the same as $Q$.
+%
+The introduction rule is then a tuple $\tensorintroatXwithYandZ b s t$
+with an instantiation $b$ of the quantified variable $\beta$. Note
+that the use of $d^+\Psi_s$ in the rule could be replaced by
+$d^-\Psi_t$ since they must be equal by the side condition $\Phi =
+\jnctx \Psi_s \Psi_t$.
+%
+The elimination rule says to construct a transformation taking a pair
+as input it is sufficient to construct one that takes in a $P$ and $Q$
+for any $\beta$, directly corresponding to the adjoint relationship
+between the tensor and the right hom.
+%
+Finally, we have the expected $\beta\eta$ equations.
 
@@
     \inferrule*[right=Unit-Formation]
-    {\Gamma \vdash C : \Sort
-     \Gamma \pipe d^-\Xi \vdash A : \cat C\and \Gamma \pipe d^+ \Xi \vdash A' : \cat C}
-    {\Gamma \pipe \Xi \vdash \punitinXfromYtoZ {C} A {A'}}
+    {\Gamma \vdash \cat C \isaCat\and
+     \Gamma \pipe d^-\Xi \vdash a_1 : \cat C\and \Gamma \pipe d^+ \Xi \vdash a_2 : \cat C}
+    {\Gamma \pipe \Xi \vdash \punitinXfromYtoZ {\cat C} {a_1} {a_2}}
 
@@
     \inferrule*[right=Unit-Introduction]
-    {\Gamma\pipe\cdot\vdash C : \Sort}
-    {\Gamma \pipe\cdot \vdash \punitrefl C : \pendallXdotY {\alpha:C} \punitinXfromYtoZ C \alpha {\alpha}}
+    {\Gamma\pipe\beta: \cat D \vdash a : \cat C}
+    {\Gamma \pipe\cdot \vdash \punitrefl a : \punitinXfromYtoZ {\cat C} a a}
 
     \inferrule*[right=Unit Elim]{
-      \Gamma\pipe\cdot \vdash Q : \harr C C\and
-      \Gamma\pipe\cdot\vdash N : \pendallXdotY {\alpha:C} {\harrapp Q \alpha \alpha}}
-    {\Gamma\pipe\cdot \vdash \punitelimtoYwithkontZ Q N :
-      \pendallXdotY {\alpha} {\homrallXYtoZ {\alpha'} {\punitinXfromYtoZ C {\alpha}{\alpha'}} {\harrapp Q {\alpha}{\alpha'}}}
+      \Gamma\vdash Q : \harr {\cat C} {\cat C}\and
+      \Gamma\vdash t : \pendallXdotY {\alpha:\cat C} {\harrapp Q \alpha \alpha}}
+    {\Gamma \vdash \punitelimtoYwithkontZ Q N :
+      \pendallXdotY {\alpha_1} {\homrallXYtoZ {\alpha_2} {\punitinXfromYtoZ {\cat C} {\alpha_1}{\alpha_2}} {\harrapp Q {\alpha_1}{\alpha_2}}}
       }
@@
 
-    \homrappXtoYatZ {\pendappXtoY {\punitelimtoYwithkontZ Q N} A} {\pendappXtoY {\punitrefl C} A} A = \pendappXtoY N A
+    \appunary {\punitelimtoYwithkontZ Q N} a {\punitrefl a} a = \pendappXtoY N a
 
     \inferrule*[right=Unit$\eta$]
-    {\alpha_1 : \cat C, f : \alpha_1 \to_{\cat C} \alpha_2, \alpha_2:\cat C \vdash M : \harrapp Q {\alpha_1} {\alpha_2}}
-    {\alpha_1 : \cat C, f : \alpha_1 \to_{\cat C} \alpha_2, \alpha_2:\cat C \vdash M =
-      \homrappXtoYatZ {\pendappXtoY {\punitelimtoYwithkontZ Q {\pendlambdaXdotY \alpha M[\alpha/\alpha_1,\pendappXtoY {\punitrefl C} \alpha/f,\alpha/\alpha_2]}} {\alpha_1}} f {\alpha_2}
-    : \harrapp Q {\alpha_1}{\alpha_2}}
-  \end{mathpar}  
-  \caption{Unit Primitive Rules}
-\end{figure}
-
-\begin{figure}
-  \begin{mathpar}
-    \inferrule*[right=Full Unit Elim*]
-    {\Phi_1[\alpha/\alpha_1], \alpha:\cat C, \Phi_2[\alpha/\alpha_2]\vdash N : Q[\alpha/\alpha_1,\alpha/\alpha_2]}
-    {\Phi_1, \alpha_1 : \cat C, f : \alpha_1 \to_{\cat C} \alpha_2, \alpha_2 : \cat C, \Phi_2\vdash J_\to'(\Phi_1, \alpha_1 : \cat C, f : \alpha_1 \to_{\cat C} \alpha_2, \alpha_2 : \cat C, \Phi_2, Q, \alpha.
-N) : Q}
-
-    J_\to'(\Phi_1, \alpha_1 : \cat C, f : \alpha_1 \to_{\cat C} \alpha_2, \alpha_2 : \cat C, \Phi_2, Q, \alpha. N) =
-    \Phi_1 \triangleright (J(\alpha_1.\alpha_2.\Phi_1 \triangleright Q \triangleleft \Phi_2, \alpha.
-\lambda^\triangleright \Phi_1[\alpha/\alpha_1].
-\lambda^\triangleleft \Phi_2[\alpha/\alpha_2].
-N, \alpha_1,f,\alpha_2)) \triangleleft \Phi_2
-
-    \inferrule*[right=Full Unit $\beta$*]
-    {}
-    {J_\to'(\Phi_1, \alpha_1 : \cat C, f : \alpha_1 \to_{\cat C} \alpha_2, \alpha_2 : \cat C, \Phi_2, Q, \alpha.
-N)[\Phi_1[\alpha/\alpha_1],\alpha/\alpha_1, \id(\alpha)/f, \alpha/\alpha_2, \Phi_2[\alpha/\alpha_2]]
-    = N}
-
-    (\Phi_1 \triangleright J(\alpha_1.\alpha_2.\Phi_1 \triangleright Q \triangleleft \Phi_2, \alpha.
-\lambda^\triangleright \Phi_1[\alpha/\alpha_1].
-\lambda^\triangleleft \Phi_2[\alpha/\alpha_2].
-N, \alpha_1,f,\alpha_2) \triangleleft \Phi_2)[\Phi_1[\alpha/\alpha_1],\alpha/\alpha_1, \id(\alpha)/f, \alpha/\alpha_2, \Phi_2[\alpha/\alpha_2]]\\
-    = \Phi_1[\alpha/\alpha_1] \triangleright J(\alpha_1.\alpha_2.\Phi_1 \triangleright Q \triangleleft \Phi_2, \alpha.
-\lambda^\triangleright \Phi_1[\alpha/\alpha_1].
-\lambda^\triangleleft \Phi_2[\alpha/\alpha_2].
-N, \alpha,\id(\alpha),\alpha) \triangleleft \Phi_2[\alpha/\alpha_2]\\
-    = \Phi_1[\alpha/\alpha_1] \triangleright
-    (\lambda^\triangleright \Phi_1[\alpha/\alpha_1]. \lambda^\triangleleft \Phi_2[\alpha/\alpha_2]. N)
-    \triangleleft \Phi_2[\alpha/\alpha_2]\\
-    = N
-
-    \inferrule*[right=Unit$\eta$]
-    {\Phi_1, \alpha_1 : \cat C, f : \alpha_1 \to_{\cat C} \alpha_2, \alpha_2 : \cat C, \Phi_2 \vdash M : Q}
-    {M =
-      J_\to'(\Phi_1, \alpha_1 : \cat C, f : \alpha_1 \to_{\cat C} \alpha_2, \alpha_2 : \cat C, \Phi_2, Q, \alpha.
-M[\Phi_1[\alpha/\alpha_1],\alpha/\alpha_1,\id(\alpha), \alpha/\alpha_2,\Phi_2[\alpha/\alpha_2]])}
-
-    M
-    =(\triangleleft\beta\triangleright\beta)
-    \Phi_1 \triangleright
-    (\lambda^\triangleright \Phi_1. \lambda^\triangleleft \Phi_2. M)
-    \triangleleft \Phi_2
-    = \Phi_1 \triangleright
-    J_\to(\alpha_1.\alpha_2.Q, \alpha.
-\lambda^\triangleright \Phi_1[\alpha/\alpha_1].
-\lambda^\triangleleft \Phi_2[\alpha/\alpha_2].
-M[\alpha/\alpha_1, \id(\alpha)/f,\alpha/\alpha_2], \alpha_1, f, \alpha_2)
-    \triangleleft \Phi_2
-    =
-    J_\to'(\Phi_1, \alpha_1 : \cat C, f : \alpha_1 \to_{\cat C} \alpha_2, \alpha_2 : \cat C, \Phi_2, Q, \alpha. N)
-  \end{mathpar}  
-  \caption{Admissible Unit Universal Property Rules}
-\end{figure}
-
-Next, we can internalize the $,\alpha,$ as a positive connective as a
-kind of ``existentially quantified tensor product'', abstracting the
-composition of profunctors.
-
-The ordinary formula for this is that for profunctors $R : C^o \times
-D \to \Set$ and $S : D^o \times E \to \Set$, the composite is given by
-a coend:
-
-\[ \int^{d \in D} R(-,d) \otimes S(d,=) \]
-
-We will instead use slightly more logical notation, writing the coend as
-an existential quantifier, justified by the analogy between the two
-and the structure of the introduction/elimination forms in our syntax
-in particular.
-\begin{figure}
-\begin{mathpar}
+    {\Gamma \vdash M : \pendallXdotY {\alpha_1} \homrallXYtoZ {\alpha_2} {\punitinXfromYtoZ {\cat C} {\alpha_1} {\alpha_2}} {\harrapp Q {\alpha_1} {\alpha_2}}}
+    {M = {\punitelimtoYwithkontZ Q {\appunary M \alpha {\punitrefl \alpha} \alpha}}}
+    %% {}
+    %% {\alpha_1 : \cat C, f : \alpha_1 \to_{\cat C} \alpha_2, \alpha_2:\cat C \vdash M : \harrapp Q {\alpha_1} {\alpha_2}}
+    %% {\alpha_1 : \cat C, f : \alpha_1 \to_{\cat C} \alpha_2, \alpha_2:\cat C \vdash M =
+    %%   \homrappXtoYatZ {\pendappXtoY {\punitelimtoYwithkontZ Q {\pendlambdaXdotY \alpha M[\alpha/\alpha_1,\pendappXtoY {\punitrefl C} \alpha/f,\alpha/\alpha_2]}} {\alpha_1}} f {\alpha_2}
+    %% : \harrapp Q {\alpha_1}{\alpha_2}}
+
     \inferrule*[right=Tensor formation]
     {\Gamma \pipe d^-\Xi; \beta:D\vdash P \isaSet\and
-    \Gamma \pipe \beta:D;d^+\Xi \vdash Q}
+     \Gamma \pipe \beta:D;d^+\Xi \vdash Q \isaSet}
     {\Gamma \pipe \Xi \vdash \tensorexistsXwithYandZ {\beta:D} P Q}
@@
 
+  %% \inferrule*[right=Tensor introduction]
+  %% {\Gamma \vdash P : \harr C D \and
+  %%   \Gamma \vdash Q : \harr D E}
+  %% {\Gamma\pipe\cdot\vdash \tensorintroofXandY P Q : \pendallXdotY {\alpha} \homrallXYtoZ \beta {\harrapp P \alpha \beta} {\homrallXYtoZ \gamma {\harrapp Q \beta \gamma} {(\tensorexistsXwithYandZ \beta {\harrapp P \alpha \beta}{\harrapp Q \beta \gamma})}}}
+
   \inferrule*[right=Tensor introduction]
-  {\Gamma \vdash P : \harr C D \and
-    \Gamma \vdash Q : \harr D E}
-  {\Gamma\pipe\cdot\vdash \tensorintroofXandY P Q : \pendallXdotY {\alpha} \homrallXYtoZ \beta {\harrapp P \alpha \beta} {\homrallXYtoZ \gamma {\harrapp Q \beta \gamma} {(\tensorexistsXwithYandZ \beta {\harrapp P \alpha \beta}{\harrapp Q \beta \gamma})}}}
-
-  %% \inferrule*[right=Tensor introduction]
-  %% {\Phi = \Psi_M \jnctx \Psi_N \and
-  %%  \Gamma \pipe d^+\Psi_M \vdash B : D\and
-  %%  \Gamma \pipe \Psi_M \vdash M : P[;B/\beta] \and
-  %%  \Gamma \pipe \Psi_N \vdash N : Q[B/\beta;]}
-  %% {\Gamma\pipe\Phi\vdash \tensorintroatXwithYandZ B M N : \tensorexistsXwithYandZ {\beta : D} P Q}
+  {\Phi = \Psi_s \jnctx \Psi_t \and
+   \Gamma \pipe d^+\Psi_s \vdash b : \cat D\and
+   \Gamma \pipe \Psi_s \vdash s : P[b/\beta] \and
+   \Gamma \pipe \Psi_t \vdash t : Q[b/\beta]}
+  {\Gamma\pipe\Phi\vdash \tensorintroatXwithYandZ b s t : \tensorexistsXwithYandZ {\beta : \cat D} P Q}
 
@@
   \inferrule*[right=Tensor-elim]
-  {\Gamma \pipe \cdot \vdash P : \harr C D\and\Gamma \pipe \cdot \vdash Q : \harr D E\and
-   \Gamma \pipe \cdot \vdash R : \harr C E\and
-   \Gamma \pipe \cdot \vdash N : \pendallXdotY \alpha {\homrallXYtoZ \beta {\harrapp P \alpha \beta} {\homrallXYtoZ \gamma {\harrapp Q \beta \gamma}}}   
-  }
-  {\Gamma\pipe\cdot \vdash \tensorelimfromWXtoYwithkontZ P Q R N : \pendallXdotY \alpha {\homrallXYtoZ \gamma {\tensorexistsXwithYandZ \beta {\harrapp P \alpha \beta}{\harrapp Q \beta \gamma}} {\harrapp R \alpha \gamma}}}
+  {\Gamma \pipe \vdash P : \harr {\cat C} {\cat D}\and\Gamma \vdash Q : \harr {\cat D}{\cat E}\and
+   \Gamma \vdash R : \harr {\cat C}{\cat E}\and
+   \Gamma \vdash N : \hombinary \alpha {\harrapp P \alpha \beta} \beta {\harrapp Q \beta \gamma} \gamma {\harrapp R \alpha \gamma}}
+  {\Gamma \vdash \tensorelimfromWXtoYwithkontZ P Q R N : \homunary \alpha {(\tensorexistsXwithYandZ \beta {\harrapp P \alpha \beta} {\harrapp Q \beta \gamma})} {\gamma} {\harrapp R \alpha \gamma}}
 
@@
   {}
-  {\homrappXtoYatZ {\pendappXtoY{\tensorelimfromWXtoYwithkontZ P Q R N} A} {(\homrappXtoYatZ{\homrappXtoYatZ{\pendappXtoY {\tensorintroofXandY P Q} A} f B} g G)} G =
-    \homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY N A} f B} g G}
+  {\appunary {\tensorelimfromWXtoYwithkontZ P Q R N} a {\tensorintroatXwithYandZ b s t} c = \appbinary N a s b t c}
+  %% {\homrappXtoYatZ {\pendappXtoY{\tensorelimfromWXtoYwithkontZ P Q R N} A} {(\homrappXtoYatZ{\homrappXtoYatZ{\pendappXtoY {\tensorintroofXandY P Q} A} f B} g G)} G =
+  %%   \homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY N A} f B} g G}
 
@@
   \inferrule*[right=Tensor$\eta$]
-  {\alpha : \cat C, x : \tensorexistsXwithYandZ \beta P Q, \gamma : \cat E \vdash M : R}
-  {\alpha : \cat C, x : \exists \beta. Q \odot R, \gamma : \cat E \vdash M =
-    \homrappXtoYatZ {\pendappXtoY{\tensorelimfromWXtoYwithkontZ P Q R
-        {\pendlambdaXdotY \alpha {\homrlambdaXatYdotZ p \beta {\homrlambdaXatYdotZ q \gamma {M[\homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {\tensorintroofXandY P Q} \alpha} p \beta} q \gamma/x]}}}}
-      } \alpha} x \gamma : S}
+  {\Gamma \vdash N : \homunary \alpha {(\tensorexistsXwithYandZ \beta {\harrapp P \alpha \beta}{\harrapp Q \beta \gamma})} \gamma {\harrapp R \alpha \gamma}}
+  {N = \tensorelimfromWXtoYwithkontZ P Q R {\lambdabinary \alpha x \beta y \gamma \appunary N \alpha {\tensorintroatXwithYandZ \beta x y} \gamma}}
+
+  %% \inferrule*[right=Tensor$\eta$]
+  %% {\alpha : \cat C, x : \tensorexistsXwithYandZ \beta P Q, \gamma : \cat E \vdash M : R}
+  %% {\alpha : \cat C, x : \exists \beta. Q \odot R, \gamma : \cat E \vdash M =
+  %%   \homrappXtoYatZ {\pendappXtoY{\tensorelimfromWXtoYwithkontZ P Q R
+  %%       {\pendlambdaXdotY \alpha {\homrlambdaXatYdotZ p \beta {\homrlambdaXatYdotZ q \gamma {M[\homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {\tensorintroofXandY P Q} \alpha} p \beta} q \gamma/x]}}}}
+  %%     } \alpha} x \gamma : S}
   \end{mathpar}  
-\caption{Tensor Primitive Rules}
-\label{fig:tensor}
+  \caption{Unit, Tensor of Profunctors}
 \end{figure}
-First, the syntax for tensor is based on the coend formula that gives
-the definition in $Cat$, but using logical syntax $\exists$ rather
-than integrals because I don't like integral syntax and the syntax
-matches existential quantification in type theory.
-%
-Note that $\exists \beta. ~Q \odot R$ is the application of a single
-connective to $Q$ and $R$, and not separable into a quantifier
-$\exists$ connective and a product connective $\odot$.
-%
-Introduction, elimination and $\beta\eta$ are similar to a combination
-of existential quantification and the tensor product in linear logic.
-%
-The introduction rule is justified by the fact that semantically what
-we have is a transformation
-\[ \textrm{pair} : (\alpha : \cat C,Q,\beta:\cat D,R,\gamma:\cat E) \Rightarrow \exists \beta. R \odot Q \]
-And the introduction rule is just building in a substitution to this
-transformation.
-%
-
-For the elim rule we can pull a similar trick as with the Hom
-type. First, the rule where there's only one variable.
-\[ \inferrule*[right=Tensor elim Simple]
-  {\alpha : \cat C, q : Q, \beta : \cat D, r: R, \gamma : \cat E \vdash N : S}
-  {\alpha : \cat C, x : \exists \beta. Q \odot R, \gamma : \cat E \vdash J_\odot(x, q.\beta.r. N) : S}
-\]
-
-\begin{figure}
-\begin{mathpar}
-    \inferrule*[right=Full Tensor Elim*]
-    {\Phi_1, \alpha:\cat C, q:Q, \beta:\cat D, r:R, \gamma:\cat E \Phi_2\vdash N : S}
-    {\Phi_1, \alpha : \cat C, x : \exists \beta:\cat D. Q \odot R, \gamma : \cat E, \Phi_2
-      \vdash J_\odot'(\Phi_1, \alpha : \cat C, x : \exists \beta:\cat D. Q \odot R, \gamma : \cat E, \Phi_2; S;q.\beta.r.N) : S}
-
-    J_\odot'(\Phi_1, \alpha : \cat C, x : \exists \beta:\cat D. Q \odot R, \gamma : \cat E, \Phi_2;S; q.\beta.r.N)
-    = \Phi_1 \triangleright J_\odot(\alpha.\gamma.\Phi_1 \triangleright S \triangleleft \Phi_2, \alpha.q.\beta.r.\gamma.
-\lambda^\triangleleft \Phi_1.
-\lambda^\triangleright\Phi_2.
-N, \alpha, x, \gamma) \triangleleft \Phi_2
-
-    J_\odot'(\Phi_1, \alpha : \cat C, x : \exists \beta:\cat D. Q \odot R, \gamma : \cat E, \Phi_2; q.\beta.r.N)[\Phi_1,\alpha/\alpha,(q,\beta,r)/x,\gamma/\gamma,\Phi_2] = N
-
-  \inferrule*[right=Full Tensor elim*]{ S \prof(\alpha:\cat C; \gamma : \cat E)\and
-    \Phi \vdash M : \exists \beta. Q[A/\alpha] \odot R[G/\gamma]\and
-    \alpha : \cat C, q : Q, \beta : \cat D, r: R, \gamma : \cat E \vdash N : S}
-  {\Phi \vdash J_\odot(\alpha. \gamma. S, \alpha.q.\beta.r.\gamma.N, A, M, G) : S[A/\alpha, G/\gamma]}
-
-  J_\odot(\alpha. \gamma. S, \alpha.q.\beta.r.\gamma.N, A, M, G)[\phi] =
-  J_\odot(\alpha. \gamma. S, \alpha.q.\beta.r.\gamma.N, A[d^-\phi], M[\phi], G[d^+\phi])
-
-  %% \inferrule*[right=Tensor elim Sequent]
-  %% {\Phi,\alpha : \cat C, q : Q, \beta : \cat D, r: R, \gamma : \cat E,\Psi \vdash N : S}
-  %% {\Phi, \alpha : \cat C, x : \exists \beta. Q \odot R, \gamma : \cat E,\Psi \vdash J_\odot(x;q.\beta.r. N) : S}
-
-  %% \inferrule*[right=Tensor elim Subst]
-  %% {\Phi,\alpha : \cat C, q : Q, \beta : \cat D, r: R, \gamma : \cat E,\Phi' \vdash N : S
-  %% \and
-  %% \phi : \Psi \to \Phi,\alpha : \cat C, q : Q, \beta : \cat D, r: R, \gamma : \cat E,\Phi'}
-  %% {\Psi  \vdash J_\odot(x;q.\beta.r. N)[\phi] : S}
-
-  \inferrule*[right=Tensor $\beta$]
-  {}{J_\odot(\alpha.\gamma.S, \alpha.q.\beta.r.\gamma.N,A,(M,B,N),G) = N[A/\alpha,M/q,B/\beta,N/r,G/\gamma]}
-
-  \inferrule*[right=Tensor$\eta$]
-  {\alpha : \cat C, x : \exists \beta. Q \odot R, \gamma : \cat E \vdash M : S}
-  {\alpha : \cat C, x : \exists \beta. Q \odot R, \gamma : \cat E \vdash M =
-    J(\alpha.\gamma.S, \alpha.q.\beta.r.\gamma. M[\alpha/\alpha,(q,\beta,r)/x,\gamma/\gamma])
-    : S}
-\end{mathpar}
-\caption{Tensor Admissible Universal Property Rules}
-\label{fig:tensor}
-\end{figure}
+
+%% Next, we can internalize the $,\alpha,$ as a positive connective as a
+%% kind of ``existentially quantified tensor product'', abstracting the
+%% composition of profunctors.
+
+%% The ordinary formula for this is that for profunctors $R : C^o \times
+%% D \to \Set$ and $S : D^o \times E \to \Set$, the composite is given by
+%% a coend:
+
+%% \[ \int^{d \in D} R(-,d) \otimes S(d,=) \]
+
+%% We will instead use slightly more logical notation, writing the coend as
+%% an existential quantifier, justified by the analogy between the two
+%% and the structure of the introduction/elimination forms in our syntax
+%% in particular.
+%% \begin{figure}
+%% \begin{mathpar}
+%% \end{mathpar}
+%% \caption{Tensor Primitive Rules}
+%% \label{fig:tensor}
+%% \end{figure}
+
+Semantically, the tensor product is the composition of profunctors,
+which is typically written as a coend
+\[ \int_{\beta} P(-,\beta) \times Q(\beta,=) \]
 
@@
 
-\subsection{Equalizers}
-
-To encode the concept of internal hom of profunctors, kan extensions
-and (co-)limits, we need to define the \emph{equalizer} of horizontal
-arrows. Specifically, the universal property that we need is an
-equalizer in the category whose objects are horizontal arrows (for
-fixed endpoints) and whose arrows are the 2-cells.
-
-\begin{figure}
-  \begin{mathpar}
-    \inferrule*[right=Equalizer-formation]
-    {\Gamma \pipe \cdot \vdash P : \harr\, C\,D\and
-     \Gamma \pipe \cdot \vdash Q : \harr\, C \, D\\\\
-     \Gamma \pipe \cdot \vdash \phi : \homunary \alpha {\harrapp P \alpha \beta} \beta {\harrapp Q \alpha \beta}\and
-     \Gamma \pipe \cdot \vdash \psi : \homunary \alpha {\harrapp P \alpha \beta} \beta {\harrapp Q \alpha \beta}\and}
-    {\Gamma \pipe \cdot \vdash \textrm{Equalizer}(C,D,P,\phi,\psi) : \harr\, C\, D}\and
-
-    \inferrule
-    {\Gamma \pipe \Phi \vdash M : \harrapp{\textrm{Equalizer}(C,D,P,\phi,\psi)} A B}
-    {\Gamma \pipe \Phi \vdash \textrm{forget}(M) : \harrapp P A B}
-
-    \inferrule
-    {\Gamma \pipe \Phi \vdash M : \harrapp P A B \and \appunary \phi A M B = \appunary \psi A M B}
-    {\Gamma \pipe \Phi \vdash \textrm{equalizing}_{\phi,\psi}(M) : \harrapp{\textrm{Equalizer}(C,D,P,\phi,\psi)} A B}
-
-    TODO: obvious equations
-  \end{mathpar}
-  \caption{Equalizers (Logical Framework style)}
-\end{figure}
-
-\begin{figure}
-  \begin{mathpar}
-    \inferrule*[right=Equalizer-formation]
-    {\Gamma \pipe \Xi \vdash P \isaSet\and\Gamma \pipe \Xi \vdash Q \isaSet\and
-      \Gamma\pipe d^-\Xi, p : P, d^+ \Xi \vdash \phi : Q\and
-      \Gamma\pipe d^-\Xi, p : P, d^+ \Xi \vdash \psi : Q }
-    {\Gamma \pipe \Xi \vdash \equalizeVofWbyXeqYatZ p P \phi \psi Q : \harr\, C\, D}\and
-
-    \inferrule*[right=Equalizer-formation]
-    {\Gamma \pipe \alpha:C;\beta:D \vdash P \isaSet\and\Gamma \pipe \alpha:C;\beta:D \vdash Q \isaSet\and
-      \Gamma \pipe d^-\Xi \vdash A : C\and
-      \Gamma \pipe d^+\Xi \vdash B : D\\\\
-      \Gamma\pipe \cdot \vdash \phi : \homunary \alpha P \beta Q\and
-      \Gamma\pipe \cdot \vdash \psi : \homunary \alpha P \beta Q
-    }
-    {\Gamma \pipe \Xi \vdash \equalizeVofWbyXeqYatZ p {P[A/\alpha;B/\beta]} {\appunary \phi A p B} {\appunary \psi A p B} {Q[A/\alpha;B/\beta]} : \harr\, C\, D}\and
-
-    \inferrule
-    {\Gamma \pipe \Phi \vdash M : \equalizeVofWbyXeqYatZ p P \phi \psi Q}
-    {\Gamma \pipe \Phi \vdash \pi_{eq}\,M : P}
-
-    \inferrule
-    {\Gamma \pipe \Phi \vdash M : \harrapp P A B \and \appunary \phi A M B = \appunary \psi A M B}
-    {\Gamma \pipe \Phi \vdash \textrm{equalizing}_{\phi,\psi}(M) : \harrapp{\textrm{Equalizer}(C,D,P,\phi,\psi)} A B}
-
-    TODO: equations
-  \end{mathpar}
-  \caption{Equalizers (Natural Deduction Style)}
-\end{figure}
-
-\subsection{Products}
-
-We add in parallel two different notions of product: products of sorts
-and products of spans, which have the universal property in their
-respective categories.
-
-\begin{figure}
-  \begin{mathpar}
-    \inferrule{}{\Gamma \vdash 1 : \Sort}\and
-    \inferrule*{}{\Gamma\pipe\alpha:C \vdash () : 1}\and
-    \inferrule*{\Gamma\pipe\alpha:C \vdash A : 1}{A = ()}
-    \inferrule{}{\Gamma\pipe \Xi \vdash 1 \isaSet}\and
-    \inferrule{}{\Gamma\pipe \Phi \vdash () :  1}\and
-    \inferrule{\Gamma\pipe \Phi \vdash M :  1}{M = ()}\\\\
-
-    \inferrule{\Gamma \vdash C : \Sort \and \Gamma \vdash D : \Sort}
-              {\Gamma \vdash C \times D : \Sort}\and
-
-    \inferrule{\forall i \in \{1,2\}.~ \Gamma \pipe \alpha:C \vdash A_i : D_i}{\Gamma \pipe \alpha:C \vdash (A_1,A_2) : D_1 \times D_2}
-
-    \inferrule{\Gamma \pipe \alpha:C \vdash A : D_1 \times D_2}{\Gamma \pipe \alpha:C \vdash \pi_i A : D_i}
-
-    \pi_i(A_1,A_2) = A_i\\
-    \inferrule{\Gamma\pipe \alpha:C \vdash A :  D_1\times D_2}{A = (\pi_1 A,\pi_2 A)}\\\\
-
-    \inferrule{\Gamma\pipe\Xi \vdash R \isaSet \and \Gamma\pipe\Xi \vdash S\isaSet}
-              {\Gamma\pipe\Xi \vdash R \times S\isaSet}\and
-
-    \inferrule{\forall i \in \{1,2\}.~ \Gamma \pipe\Phi \vdash M_i : R_i}{\Gamma \pipe\Phi \vdash (M_1,M_2) : R_1 \times R_2}
-
-    \inferrule{\Gamma \pipe\Phi\vdash M : R_1 \times R_2}{\Gamma \pipe\Phi \vdash \pi_i M : R_i}
-
-    \pi_i(M_1,M_2) = M_i\\
-    \inferrule{\Gamma\pipe\Phi \vdash M :  R_1\times R_2}{M = (\pi_1 M,\pi_2 M)}
-  \end{mathpar}
-  \caption{Products}
-  \label{fig:products}
-\end{figure}
-
-One additional connective we might desire is a ``parallel'' product of
-sets $R \times' S : \harr {C_1 \times C_2} {D_1 \times D_2}$ where $R
-: \harr {C_1} {D_1}$ and $R : \harr {C_2} {D_2}$. We don't need to
-define this as a primitive though because it is expressible in terms
-of the two kinds of products we have above:
-
-\[
-\inferrule
-{\Gamma \pipe \alpha : C_i; \beta : D_i \vdash R_i \isaSet}
-{\Gamma \pipe \alpha : C_1\times C_2; \beta : D_1 \times D_2
-  R_1[\pi_1\alpha/\alpha;\pi_1\beta/\beta] \times R_2[\pi_2\alpha/\alpha;\pi_2\beta/\beta]\isaSet}
-\]
-
-The empty products give us access now to several traditional notions
-of category theory that we couldn't before define, such as objects and
-presheaves of either variance:
-
-$$\textrm{Object}\,C = \varr 1 C$$
-$$\textrm{Presheaf}\,C = \harr 1 C$$
-$$\textrm{coPresheaf}\,C = \harr C 1$$
-
-However, we find that in practice we don't have much use for these
-concepts as most theorems about them easily generalize to more general
-theorems about arbitrary functors or profunctors.
-
-%% To define internal ends and coends, we need to go beyond the usual
-%% simple types and add equalizers (a form of subsets) and co-equalizers
-%% (a form of quotients) of judgments. Here's a sketch of the syntax for
-%% each. These make type checking very difficult as they result in
-%% obligations for proving equality of certain terms, which may or may
-%% not be decidable.
-
+%% \subsection{Equalizers}
+
+%% To encode the concept of internal hom of profunctors, kan extensions
+%% and (co-)limits, we need to define the \emph{equalizer} of horizontal
+%% arrows. Specifically, the universal property that we need is an
+%% equalizer in the category whose objects are horizontal arrows (for
+%% fixed endpoints) and whose arrows are the 2-cells.
+
+%% \begin{figure}
 %%   \begin{mathpar}
-%%   \inferrule*[right=Equalizer Formation]
-%%   {\alpha;\beta\,|\, R \and
-%%   x : R \vdash t : S x:R \vdash u : S}
-%%   {\alpha;\beta\,|\,\{ x : R \,|\, t = u \}}
-
-%%   \inferrule*[right=Equalizer Intro]
-%%   {\Phi \vdash r : R \and t[r/x] = u[r/x]}
-%%   {\Phi \vdash \textrm{assert}(x.t = u)(r) : \{ x : R \,|\, t = u \}}
-
-%%   \inferrule*[right=Equalizer Elim]
-%%   {\Phi \vdash r : \{ x : R \,|\, t = u \}}
-%%   {\Phi \vdash \textrm{forget}(r) : R}
-
-%%   t[\textrm{forget}(r)/x] = u[\textrm{forget(r)}/x]
-
-%%   \textrm{assert}(x.t = u)(\textrm{forget}(r)) = r
-%%   \textrm{forget}(\textrm{assert}(x.t = u)(r)) = r
+%%     \inferrule*[right=Equalizer-formation]
+%%     {\Gamma \pipe \cdot \vdash P : \harr\, C\,D\and
+%%      \Gamma \pipe \cdot \vdash Q : \harr\, C \, D\\\\
+%%      \Gamma \pipe \cdot \vdash \phi : \homunary \alpha {\harrapp P \alpha \beta} \beta {\harrapp Q \alpha \beta}\and
+%%      \Gamma \pipe \cdot \vdash \psi : \homunary \alpha {\harrapp P \alpha \beta} \beta {\harrapp Q \alpha \beta}\and}
+%%     {\Gamma \pipe \cdot \vdash \textrm{Equalizer}(C,D,P,\phi,\psi) : \harr\, C\, D}\and
+
+%%     \inferrule
+%%     {\Gamma \pipe \Phi \vdash M : \harrapp{\textrm{Equalizer}(C,D,P,\phi,\psi)} A B}
+%%     {\Gamma \pipe \Phi \vdash \textrm{forget}(M) : \harrapp P A B}
+
+%%     \inferrule
+%%     {\Gamma \pipe \Phi \vdash M : \harrapp P A B \and \appunary \phi A M B = \appunary \psi A M B}
+%%     {\Gamma \pipe \Phi \vdash \textrm{equalizing}_{\phi,\psi}(M) : \harrapp{\textrm{Equalizer}(C,D,P,\phi,\psi)} A B}
+
+%%     TODO: obvious equations
 %%   \end{mathpar}
-
+%%   \caption{Equalizers (Logical Framework style)}
+%% \end{figure}
+
+%% \begin{figure}
 %%   \begin{mathpar}
-%%   \inferrule*[right=Co-equalizer Formation]
-%%   {\alpha;\beta\,|\, R \and
-%%   x : S \vdash t : R x:S \vdash u : R}
-%%   {\alpha;\beta\,|\,R/(\forall x:S. t = u)}
-
-%%   \inferrule*[right=Coequalizer Intro]
-%%   {\Phi \vdash r : R}
-%%   {\Phi \vdash \textrm{equate}(x:S.t=u)(r) : R/(\forall x.t=u)}
-
-%%   \inferrule*[right=Coequalizer Elim]
-%%   {\Psi \vdash r : R/(x.t = u) \and
-%%     \Phi,y:R,\Phi' \vdash s : S \and
-%%     s[t/y] = s[u/y]}
-%%   {\Phi,\Psi,\Phi' \vdash \textrm{lett} \textrm{equate}(y) = r.~s : S}
-
-%%   \textrm{equate}(x:S.t = u)(t[\gamma]) = \textrm{equate}(x:S.t = u)(u[\gamma])
-
-%%   \cdots
+%%     \inferrule*[right=Equalizer-formation]
+%%     {\Gamma \pipe \Xi \vdash P \isaSet\and\Gamma \pipe \Xi \vdash Q \isaSet\and
+%%       \Gamma\pipe d^-\Xi, p : P, d^+ \Xi \vdash \phi : Q\and
+%%       \Gamma\pipe d^-\Xi, p : P, d^+ \Xi \vdash \psi : Q }
+%%     {\Gamma \pipe \Xi \vdash \equalizeVofWbyXeqYatZ p P \phi \psi Q : \harr\, C\, D}\and
+
+%%     \inferrule*[right=Equalizer-formation]
+%%     {\Gamma \pipe \alpha:C;\beta:D \vdash P \isaSet\and\Gamma \pipe \alpha:C;\beta:D \vdash Q \isaSet\and
+%%       \Gamma \pipe d^-\Xi \vdash A : C\and
+%%       \Gamma \pipe d^+\Xi \vdash B : D\\\\
+%%       \Gamma\pipe \cdot \vdash \phi : \homunary \alpha P \beta Q\and
+%%       \Gamma\pipe \cdot \vdash \psi : \homunary \alpha P \beta Q
+%%     }
+%%     {\Gamma \pipe \Xi \vdash \equalizeVofWbyXeqYatZ p {P[A/\alpha;B/\beta]} {\appunary \phi A p B} {\appunary \psi A p B} {Q[A/\alpha;B/\beta]} : \harr\, C\, D}\and
+
+%%     \inferrule
+%%     {\Gamma \pipe \Phi \vdash M : \equalizeVofWbyXeqYatZ p P \phi \psi Q}
+%%     {\Gamma \pipe \Phi \vdash \pi_{eq}\,M : P}
+
+%%     \inferrule
+%%     {\Gamma \pipe \Phi \vdash M : \harrapp P A B \and \appunary \phi A M B = \appunary \psi A M B}
+%%     {\Gamma \pipe \Phi \vdash \textrm{equalizing}_{\phi,\psi}(M) : \harrapp{\textrm{Equalizer}(C,D,P,\phi,\psi)} A B}
+
+%%     TODO: equations
 %%   \end{mathpar}
-
-%% \subsection{Positive Multiplicatives}
-
-%% \subsection{Negative Multiplicatives}
-
-%% \subsection{Additives}
+%%   \caption{Equalizers (Natural Deduction Style)}
+%% \end{figure}
+
+%% \subsection{Products}
+
+%% We add in parallel two different notions of product: products of sorts
+%% and products of spans, which have the universal property in their
+%% respective categories.
+
+%% \begin{figure}
+%%   \begin{mathpar}
+%%     \inferrule{}{\Gamma \vdash 1 : \Sort}\and
+%%     \inferrule*{}{\Gamma\pipe\alpha:C \vdash () : 1}\and
+%%     \inferrule*{\Gamma\pipe\alpha:C \vdash A : 1}{A = ()}
+%%     \inferrule{}{\Gamma\pipe \Xi \vdash 1 \isaSet}\and
+%%     \inferrule{}{\Gamma\pipe \Phi \vdash () :  1}\and
+%%     \inferrule{\Gamma\pipe \Phi \vdash M :  1}{M = ()}\\\\
+
+%%     \inferrule{\Gamma \vdash C : \Sort \and \Gamma \vdash D : \Sort}
+%%               {\Gamma \vdash C \times D : \Sort}\and
+
+%%     \inferrule{\forall i \in \{1,2\}.~ \Gamma \pipe \alpha:C \vdash A_i : D_i}{\Gamma \pipe \alpha:C \vdash (A_1,A_2) : D_1 \times D_2}
+
+%%     \inferrule{\Gamma \pipe \alpha:C \vdash A : D_1 \times D_2}{\Gamma \pipe \alpha:C \vdash \pi_i A : D_i}
+
+%%     \pi_i(A_1,A_2) = A_i\\
+%%     \inferrule{\Gamma\pipe \alpha:C \vdash A :  D_1\times D_2}{A = (\pi_1 A,\pi_2 A)}\\\\
+
+%%     \inferrule{\Gamma\pipe\Xi \vdash R \isaSet \and \Gamma\pipe\Xi \vdash S\isaSet}
+%%               {\Gamma\pipe\Xi \vdash R \times S\isaSet}\and
+
+%%     \inferrule{\forall i \in \{1,2\}.~ \Gamma \pipe\Phi \vdash M_i : R_i}{\Gamma \pipe\Phi \vdash (M_1,M_2) : R_1 \times R_2}
+
+%%     \inferrule{\Gamma \pipe\Phi\vdash M : R_1 \times R_2}{\Gamma \pipe\Phi \vdash \pi_i M : R_i}
+
+%%     \pi_i(M_1,M_2) = M_i\\
+%%     \inferrule{\Gamma\pipe\Phi \vdash M :  R_1\times R_2}{M = (\pi_1 M,\pi_2 M)}
+%%   \end{mathpar}
+%%   \caption{Products}
+%%   \label{fig:products}
+%% \end{figure}
+
+%% One additional connective we might desire is a ``parallel'' product of
+%% sets $R \times' S : \harr {C_1 \times C_2} {D_1 \times D_2}$ where $R
+%% : \harr {C_1} {D_1}$ and $R : \harr {C_2} {D_2}$. We don't need to
+%% define this as a primitive though because it is expressible in terms
+%% of the two kinds of products we have above:
+
+%% \[
+%% \inferrule
+%% {\Gamma \pipe \alpha : C_i; \beta : D_i \vdash R_i \isaSet}
+%% {\Gamma \pipe \alpha : C_1\times C_2; \beta : D_1 \times D_2
+%%   R_1[\pi_1\alpha/\alpha;\pi_1\beta/\beta] \times R_2[\pi_2\alpha/\alpha;\pi_2\beta/\beta]\isaSet}
+%% \]
+
+%% The empty products give us access now to several traditional notions
+%% of category theory that we couldn't before define, such as objects and
+%% presheaves of either variance:
+
+%% $$\textrm{Object}\,C = \varr 1 C$$
+%% $$\textrm{Presheaf}\,C = \harr 1 C$$
+%% $$\textrm{coPresheaf}\,C = \harr C 1$$
+
+%% However, we find that in practice we don't have much use for these
+%% concepts as most theorems about them easily generalize to more general
+%% theorems about arbitrary functors or profunctors.
+
+%% %% To define internal ends and coends, we need to go beyond the usual
+%% %% simple types and add equalizers (a form of subsets) and co-equalizers
+%% %% (a form of quotients) of judgments. Here's a sketch of the syntax for
+%% %% each. These make type checking very difficult as they result in
+%% %% obligations for proving equality of certain terms, which may or may
+%% %% not be decidable.
+
+%% %% \begin{mathpar}
+%% %%   \inferrule*[right=Equalizer Formation]
+%% %%   {\alpha;\beta\,|\, R \and
+%% %%   x : R \vdash t : S x:R \vdash u : S}
+%% %%   {\alpha;\beta\,|\,\{ x : R \,|\, t = u \}}
+
+%% %%   \inferrule*[right=Equalizer Intro]
+%% %%   {\Phi \vdash r : R \and t[r/x] = u[r/x]}
+%% %%   {\Phi \vdash \textrm{assert}(x.t = u)(r) : \{ x : R \,|\, t = u \}}
+
+%% %%   \inferrule*[right=Equalizer Elim]
+%% %%   {\Phi \vdash r : \{ x : R \,|\, t = u \}}
+%% %%   {\Phi \vdash \textrm{forget}(r) : R}
+
+%% %%   t[\textrm{forget}(r)/x] = u[\textrm{forget(r)}/x]
+
+%% %%   \textrm{assert}(x.t = u)(\textrm{forget}(r)) = r
+%% %%   \textrm{forget}(\textrm{assert}(x.t = u)(r)) = r
+%% %% \end{mathpar}
+
+%% %% \begin{mathpar}
+%% %%   \inferrule*[right=Co-equalizer Formation]
+%% %%   {\alpha;\beta\,|\, R \and
+%% %%   x : S \vdash t : R x:S \vdash u : R}
+%% %%   {\alpha;\beta\,|\,R/(\forall x:S. t = u)}
+
+%% %%   \inferrule*[right=Coequalizer Intro]
+%% %%   {\Phi \vdash r : R}
+%% %%   {\Phi \vdash \textrm{equate}(x:S.t=u)(r) : R/(\forall x.t=u)}
+
+%% %%   \inferrule*[right=Coequalizer Elim]
+%% %%   {\Psi \vdash r : R/(x.t = u) \and
+%% %%     \Phi,y:R,\Phi' \vdash s : S \and
+%% %%     s[t/y] = s[u/y]}
+%% %%   {\Phi,\Psi,\Phi' \vdash \textrm{lett} \textrm{equate}(y) = r.~s : S}
+
+%% %%   \textrm{equate}(x:S.t = u)(t[\gamma]) = \textrm{equate}(x:S.t = u)(u[\gamma])
+
+%% %%   \cdots
+%% %% \end{mathpar}
+
+%% %% \subsection{Positive Multiplicatives}
+
+%% %% \subsection{Negative Multiplicatives}
+
+%% %% \subsection{Additives}
 
 \section{Synthetic Category Theory}
-Next, we look at Synthetic category theory in a virtual equipment. In
-this section we will use category-theoretic terminology for judgment
-theory: categories, functors/objects, profunctors and transformations.
+
+Next, we look at synthetic category theory constructions in SCT.
 
@@
 
-Here are some isomorphisms of profunctors that should be provable
-synthetically.
-%
-Unit laws:
+We start our study of synthetic category theory constructions with
+some isomorphisms that show the interaction between the 5
+multiplicative connectives: right/left hom, end, unit and tensor.
+
+First, we fix some terminology
+\begin{lemma}
+  Let $P, Q : \harr {\cat C} {\cat D}$.
+  A homomorphism from $P$ to $Q$ is a term of type
+  \[ \homunary {\alpha} {\harrapp P \alpha \beta} {\beta} {\harrapp Q \alpha \beta} \]
+  Given homomorphisms $M : P \multimap Q$, $N : Q \multimap R$ we can define their composite as
+  \[ (N \circ M) = \lambdaunary {\alpha} {x:\harrapp P \alpha \beta} \beta {\appunary \alpha N \beta {\appunary \alpha M \beta x}} \]
+  This notion of composition is associative and unital using the
+  $\beta\eta$ equations for the end/right hom connectives, with unit
+  $\lambdaunary {\alpha} x \beta x$.
+
+  An isomorphism between $P, Q : \harr {\cat C}{\cat D}$ is then a
+  homomorphism $P \multimap Q$ with an inverse $Q \multimap P$.
+\end{lemma}
+
+Our first synthetic theorem is that we have that the unit, i.e., the
+hom set, is the unit for tensor and hom.
+%
+The theorem for hom is the synthetic \emph{Yoneda lemma} and the
+theorem for tensor is the synthetic \emph{Co-Yoneda lemma}.
+\begin{lemma}[{Yoneda and Co-Yoneda}]
+  \label{lem:yo-and-co-yo}
+  Let $P : \harr {\cat C} {\cat D}$. Then the following sets are (naturally) isomorphic to $\harrapp P \alpha \beta$
   \begin{enumerate}
-\item $\tensorexistsXwithYandZ {\beta} {P} {\punitinXfromYtoZ {D} {\beta} {\beta'}} \cong P[\beta'/\beta]$
-\item $\tensorexistsXwithYandZ {\alpha} {\punitinXfromYtoZ {C} {\alpha'} {\alpha}} {P} \cong P[\alpha'/\alpha]$
-\item $\homrallXYtoZ {\beta} {\punitinXfromYtoZ {D} {\beta'} {\beta}} {P}$
-\item $\homlallXYtoZ {\alpha} {\punitinXfromYtoZ {C} {\alpha} {\alpha'}} {P}$
+  \item $\tensorexistsXwithYandZ {\beta'} {\harrapp P \alpha {\beta'}} {(\punitinXfromYtoZ {\cat D} {\beta'} {\beta})}$
+  \item $\tensorexistsXwithYandZ {\alpha'} {(\punitinXfromYtoZ {C} {\alpha} {\alpha'})} {\harrapp P {\alpha'}{\beta}}$
+  \item $\homrallXYtoZ {\beta'} {(\punitinXfromYtoZ {D} {\beta'} {\beta})} {\harrapp P {\alpha} {\beta'}}$
+  \item $\homlallXYtoZ {{\alpha'}} {(\punitinXfromYtoZ {C} {\alpha'} {\alpha})} {\harrapp P {{\alpha'}} {\beta}}$
   \end{enumerate}
-
-Associativity laws:
+\end{lemma}
+
+Next, we have various associativities between the homs, tensor and end
+constructions. Semantically, these are related to so-called
+\emph{Fubini} theorems about interchange of ends.
+
+\begin{lemma}[{Fubini}]
+  \label{lem:fubini}
+  The following isomorphisms hold for any $P : \harr {\cat C}{\cat D},
+  P' : \harr {\cat C}{\cat D} Q :\harr {\cat D}{\cat E}, R : \harr
+  {\cat E} {\cat F}, S : \harr {\cat C} {\cat E}$. where $\alpha:\cat
+  C, \beta:\cat D, \gamma : \cat E, \delta: \cat F$
   \begin{enumerate}
-\item $\tensorexistsXwithYandZ {\beta} {P} {(\tensorexistsXwithYandZ \gamma Q R)} \cong \tensorexistsXwithYandZ {\gamma} {(\tensorexistsXwithYandZ \beta P Q)} R$
-\item $\homrallXYtoZ {\gamma} {\tensorexistsXwithYandZ \beta P Q} S \cong \homrallXYtoZ \beta P {\homrallXYtoZ \gamma Q S}$
-\item $\homlallXYtoZ {\alpha} {\tensorexistsXwithYandZ \beta P Q} S \cong \homlallXYtoZ \beta Q {\homlallXYtoZ \alpha P S}$
-\item $\homrallXYtoZ \gamma Q {(\homlallXYtoZ \alpha P S)} \cong \homlallXYtoZ \alpha P {(\homrallXYtoZ \gamma Q S)}$
+  \item $\tensorexistsXwithYandZ {\beta} {\harrapp P \alpha \beta} {(\tensorexistsXwithYandZ \gamma {\harrapp Q \beta \gamma} {\harrapp R \gamma \delta})} \cong \tensorexistsXwithYandZ {\gamma} {(\tensorexistsXwithYandZ \beta {\harrapp P \alpha \beta} {\harrapp Q \beta \gamma})} {\harrapp R \gamma \delta}$
+  \item $\homrallXYtoZ {\gamma} {(\tensorexistsXwithYandZ \beta {\harrapp P \alpha \beta} {\harrapp Q \beta \gamma})} {\harrapp S \alpha \gamma} \cong \homrallXYtoZ \beta {\harrapp P \alpha \beta} {\homrallXYtoZ \gamma {\harrapp Q \beta \gamma} {\harrapp S \alpha \gamma}}$
+  \item $\homlallXYtoZ {\alpha} {(\tensorexistsXwithYandZ \beta {\harrapp P \alpha \beta} {\harrapp Q \beta \gamma})} {\harrapp S \alpha \gamma} \cong \homlallXYtoZ \beta {\harrapp Q \beta \gamma} {\homlallXYtoZ \alpha {\harrapp P \alpha \beta} {\harrapp S \alpha \gamma}}$
+  \item $\homrallXYtoZ \gamma {\harrapp Q \beta \gamma} {(\homlallXYtoZ \alpha {\harrapp P \alpha \beta} {\harrapp S \alpha \gamma})} \cong \homlallXYtoZ \alpha {\harrapp P \alpha \beta} {(\homrallXYtoZ \gamma {\harrapp Q \beta \gamma} {\harrapp S \alpha \gamma})}$
+  \item $\pendallXdotY \alpha \homrallXYtoZ \beta P {P'} \cong \pendallXdotY \beta \homlallXYtoZ \alpha P {P'}$
   \end{enumerate}  
-
-\subsection{Constructions: Categories, Functors and Natural Transformations}
+\end{lemma}
+
+\subsection{Composition, Functoriality, Profunctoriality, Naturality}
 
@@
 in a category. As our first synthetic construction, we can
-\emph{define} composition using the elimination form.
+\emph{define} composition using the elimination form for the unit.
 %
 \begin{definition}[Composition of Morphisms]
-  We define composition of morphisms as follows
+  We define composition of morphisms as the following term of type $\hombinary {\alpha_1} {\punitinXfromYtoZ {\cat C}{\alpha_1} {\alpha_2}} {\alpha_2} {\punitinXfromYtoZ {\cat C}{\alpha_2} {\alpha_3}} {\alpha_3} {\punitinXfromYtoZ {\cat C}{\alpha_1} {\alpha_3}}$:
   \[
-  f : \beta \to_{\cat C} \beta', g : \beta' \to_{\cat C} \beta'' \vdash g \circ f :=
-  \when f = \id_{\beta}.~ \when g = \id_{\beta}.~ \id_{\beta} : \beta \to_{\cat c} \beta''
+  \punitelimtoYwithkontZ {\hlambda {\alpha_1}{\alpha_2} \homrallXYtoZ {\alpha_3} {\punitinXfromYtoZ {\cat C} {\alpha_2} {\alpha_3}} {\punitinXfromYtoZ {\cat C} {\alpha_1} {\alpha_3}}} {\lambdaunary {\alpha} {f : \punitinXfromYtoZ {\cat C}{\alpha} {\alpha_3}} {\alpha_3} f}
   \]
 \end{definition}
+This has the same flavor as a directed type theory: to construct the
+composite it is sufficient to define the action on the identity
+morphism, which is of course the identity for composition.
+%
+Note that we gave a slight bias to the definition to follow by
+induction on the \emph{first} function, whereas we could make the
+following alternative definition, naturally of type
+$\pendallXdotY {\alpha_2} \homrallXYtoZ {\alpha_3} {\punitinXfromYtoZ {\cat C}{\alpha_2} {\alpha_3}}
+{(\homlallXYtoZ {{\alpha_1}} {\punitinXfromYtoZ {\cat C}{\alpha_1} {\alpha_2}}{\punitinXfromYtoZ {\cat C}{\alpha_1} {\alpha_3}})}$:
+\[
+  \punitelimtoYwithkontZ {\hlambda {\alpha_2}{\alpha_3} \homlallXYtoZ {{\alpha_1}} {\punitinXfromYtoZ {\cat C}{\alpha_1} {\alpha_2}}{\punitinXfromYtoZ {\cat C}{\alpha_1} {\alpha_3}} } {\lambdaunary {\alpha} {g : \punitinXfromYtoZ {\cat C}{\alpha_1}{\alpha}} {\alpha} g}
+\]
+
 It should follow by $\beta\eta$ that this is associative and unital.
@@
 
+\subsection{Limits and Kan Extensions}
+
+We can then use our notion of Hom of profunctors
+
+The right Kan extension of $F$ along $J$ is the best (maximal) way to
+fill in the following square:
+\[% https://tikzcd.yichuanshen.de/#N4Igdg9gJgpgziAXAbVABwnAlgFyxMJZABgBpiBdUkANwEMAbAVxiRAGEQBfU9TXfIRQBGclVqMWbACLdeIDNjwEiZYePrNWiDgHJu4mFADm8IqABmAJwgBbJGRA4ISAEzVNUnQCk5lm-aIjs5IohJabABiBlxAA
+\begin{tikzcd}
+C \arrow[d, "J"] \arrow[r, "F"] & D \\
+C'                              &  
+\end{tikzcd}\]
+I.e. a functor $Ran_J F : C' \to D$ with a natural transformation
+$Ran_J F \Rightarrow F$ that is universal something something.
+
+If $D$ is complete, there is a formula using ends for this. Using hom
+of profunctors, we can directly adapt this to define a
+\emph{profunctor} that is a universal filler. We can then say that a
+right Kan extension is precisely a functor that represents this
+profunctor.
+
+\begin{definition}
+  Let $C,C',D$ be categories and let $F : \textrm{Functor}\,C\,D$ and
+  $J: \textrm{Functor}\,C'\,C$ be functors. Then the right Kan
+  extension specification profunctor is a profunctor
+  $\textrm{RanSpec}_J\,F : \textrm{Prof}\,D\,C$ defined as
+
+  \[ \textrm{RanSpec}_J\,F = \textrm{Hom}(C')[;J] \triangleright \textrm{Hom}(D)[;F] \]
+
+  And a right Kan extension is a functor
+  $\textrm{Ran}_JF:\textrm{Functor}\,C'\,D$ that is a
+  right-representation of $\textrm{RanSpec}_J F$.
+\end{definition}
+
+Writing this out more pointfully, we get that the hom set
+$\harrapp{D_1} {\beta}{\textrm{Ran}_JF\, \alpha}$ is an equalizer of
+the following spans:
+\[ \homrallXYtoZ {\alpha':C'} {\harrapp {C_1'} {\alpha'}{J_0\alpha}}{\harrapp {D_1} {\beta} {F_0 \alpha'}}\]
+
+If we have a unit and therefore unit category and functor, then we can
+now define limits as a special case of Kan extensions.
+\begin{definition}
+  The limit of a functor $J : \textrm{Functor}I\,C$ is a right Kan
+  extension $\textrm{Ran}_E J$ where $E$ is the unique functor $E :
+  \textrm{Functor}I\,1$.
+\end{definition}
+
+Unraveling this definition, we get that $\harrapp{C_1} \alpha {\lim_{i} J(i)}$ is a coherent family (i.e., equalizer) of
+\[  \homrallXYtoZ {i:I} {1} {\harrapp{C_1} \alpha {J_0(i)}}  \]
+with the equalizing condition being that for any $p : \homrallXYtoZ
+{i:I} {1} {\harrapp{C_1} \alpha {J_0(i)}}$ and $f : \harrapp{I_1} {i'}{i}$ and $g : \harrapp{1}{pt}{E_0 i'}$
+\[ \homrappXtoYatZ {p} {(\appbinary {1_t} {pt} {\pendappXtoY {1_r} {pt}} {pt} {(\appbinary {1_t} {pt} g {E_0 i'} {(\appunary {E_1} {i'} f {i})} {i})} {i})} {i}
+= \appbinary {C_t} {\alpha} {(\homrappXtoYatZ {p} g {i'})} {J_0 i'} {(\appbinary {C_t} {J_0 i'} {(\appunary {J_1} i f {i'})} {J_0 i} {\pendappXtoY {C_r} {J_0 i}} {J_0 i})} {J_0 i}
+\]
+We can greatly simplify using the $\eta$ principle for $1$ and
+eliminating some reflexivities to reduce this to
+\[ \homrappXtoYatZ {p} {()} {i} =
+\appbinary {C_t} {\alpha} {(\homrappXtoYatZ {p} {()} {i'})} {J_0 i'} {{(\appunary {J_1} i f {i'})}} {J_0 i}
+\]
+which reads exactly as the commuting condition for a cone.
+
+We can similarly define \emph{left} Kan extensions by a dual
+formula. Left Kan extensions are the ``minimal'' solution to the same
+filling problem as right Kan extensions. I.e., they have a universal
+morphism out of them. They can be described by a coend formula, but in
+our profunctor calculus, they are easiest to describe using a similar
+formula to right Kan extensions, but using the contravariant hom of
+profunctors.
+
+\begin{definition}
+  Let $C,C',D$ be categories and let $F : \textrm{Functor}\,C\,D$ and
+  $J: \textrm{Functor}\,C'\,C$ be functors. Then the left Kan
+  extension specification profunctor is a profunctor
+  $\textrm{LanSpec}_J\,F : \textrm{Prof}\,C\,D$ defined as
+
+  \[ \textrm{RanSpec}_J\,F = \textrm{Hom}(D)[F;] \triangleleft \textrm{Hom}(C')[J;]  \]
+  And a left Kan extension is a functor $\textrm{Lan}_J F :
+  \textrm{Functor}\,C'\,D$ that represents this profunctor on the
+  left.
+\end{definition}
+
+Unraveling the definition we get the hom span $\harrapp {D_1}{\textrm{Ran}_JF {\alpha'}} \beta$ is an equalizer of
+\[ \homlallXYtoZ \alpha {\harrapp {C_1'} {J_0 \alpha} {\alpha'}} {\harrapp{D_1} {F_0 \alpha} \beta}\]
+
+\begin{definition}
+  The colimit of a functor $J : \textrm{Functor}I\,C$ is a left Kan
+  extension $\textrm{Lan}_E J$ where $E$ is the unique functor $E :
+  \textrm{Functor}I\,1$.
+\end{definition}
+
+We can similarly unravel this definition to see that this looks
+essentially like the classical definition of colimits as universal
+cocones.
+
 \subsection{Right (Left) Adjoints preserve (co-)Limits}
@@
   \begin{enumerate}
-  \item A sort $C_0 : \Sort$
+  \item A sort $C_0 \isaCat$
   \item A span $C_1 : \harr{C_0}{ C_0}$
@@
 
+We can define an internal functor to be a pair of an object function
+and a function on arrows that preserves identity and 
 \begin{definition}[Internal Functor]
@@
 %% \end{proof}
+
+\subsection{Internal Category Translation}
 
@@
 
-\subsection{Self-Interpretation: Hom of Internal Profunctors}
-
-We can define a version of the hom $\homrallXYtoZ \alpha P Q$ on
-\emph{internal} profunctors using a combination of this connective,
-and an \emph{equalizer} to enforce a homomorphism condition.
-
-\begin{definition}[Internal Covariant Hom of Profunctors]
-  Let $P : \textrm{Prof}\,D\,E$ and $Q : \textrm{Prof}\,C\,E$. Define
-  the covariant Hom $P \triangleright Q : \textrm{Prof}\,C\,D$ by
-  \[ (P \triangleright Q)_1 = \lambda (\alpha;\beta).
+%% \subsection{Self-Interpretation: Hom of Internal Profunctors}
+
+%% We can define a version of the hom $\homrallXYtoZ \alpha P Q$ on
+%% \emph{internal} profunctors using a combination of this connective,
+%% and an \emph{equalizer} to enforce a homomorphism condition.
+
+%% \begin{definition}[Internal Covariant Hom of Profunctors]
+%%   Let $P : \textrm{Prof}\,D\,E$ and $Q : \textrm{Prof}\,C\,E$. Define
+%%   the covariant Hom $P \triangleright Q : \textrm{Prof}\,C\,D$ by
+%%   \[ (P \triangleright Q)_1 = \lambda (\alpha;\beta).
 \equalizeVofWbyXeqYatZ \phi {\homrallXYtoZ \gamma {\harrapp {P_1} \alpha \gamma} {\harrapp {Q_1} \beta \gamma}} {\psi_{dom}} {\psi_{cod}} {\homrallXYtoZ {\gamma'} {\harrapp {P_1 }\beta {\gamma'}} {\homrallXYtoZ {\gamma} {\harrapp {E_1} {\gamma'} \gamma} {\harrapp {Q_1} \alpha \gamma}}}\]
-  where $\psi_{dom}$ and $\chi_{cod}$ are  as
-  \[ \psi_{dom} =\homrlambdaXatYdotZ p {\gamma'} {\homrlambdaXatYdotZ e \gamma
-    {\homrappXtoYatZ {\phi} {(\apptrinary {P_t} \beta {\pendappXtoY {D_r} \beta} \beta p {\gamma'} e \gamma)} \gamma}}
-  \]
-  \[
-  \psi_{cod} = \homrlambdaXatYdotZ p {\gamma'} {\homrlambdaXatYdotZ e \gamma
-    {\apptrinary {Q_t} \alpha {\pendappXtoY {C_r} \alpha} \alpha {(\homrappXtoYatZ \phi {p} {\gamma'})} {\gamma'} e \gamma}}
-  \]
-\end{definition}
-
-We can easily define a contravariant hom $P \triangleleft Q$ by a dual
-definition (TODO).
-
-\subsection{Limits and Kan Extensions}
-
-We can then use our notion of Hom of profunctors
-
-The right Kan extension of $F$ along $J$ is the best (maximal) way to
-fill in the following square:
-\[% https://tikzcd.yichuanshen.de/#N4Igdg9gJgpgziAXAbVABwnAlgFyxMJZABgBpiBdUkANwEMAbAVxiRAGEQBfU9TXfIRQBGclVqMWbACLdeIDNjwEiZYePrNWiDgHJu4mFADm8IqABmAJwgBbJGRA4ISAEzVNUnQCk5lm-aIjs5IohJabABiBlxAA
-\begin{tikzcd}
-C \arrow[d, "J"] \arrow[r, "F"] & D \\
-C'                              &  
-\end{tikzcd}\]
-I.e. a functor $Ran_J F : C' \to D$ with a natural transformation
-$Ran_J F \Rightarrow F$ that is universal something something.
-
-If $D$ is complete, there is a formula using ends for this. Using hom
-of profunctors, we can directly adapt this to define a
-\emph{profunctor} that is a universal filler. We can then say that a
-right Kan extension is precisely a functor that represents this
-profunctor.
-
-\begin{definition}
-  Let $C,C',D$ be categories and let $F : \textrm{Functor}\,C\,D$ and
-  $J: \textrm{Functor}\,C'\,C$ be functors. Then the right Kan
-  extension specification profunctor is a profunctor
-  $\textrm{RanSpec}_J\,F : \textrm{Prof}\,D\,C$ defined as
-
-  \[ \textrm{RanSpec}_J\,F = \textrm{Hom}(C')[;J] \triangleright \textrm{Hom}(D)[;F] \]
-
-  And a right Kan extension is a functor
-  $\textrm{Ran}_JF:\textrm{Functor}\,C'\,D$ that is a
-  right-representation of $\textrm{RanSpec}_J F$.
-\end{definition}
-
-Writing this out more pointfully, we get that the hom set
-$\harrapp{D_1} {\beta}{\textrm{Ran}_JF\, \alpha}$ is an equalizer of
-the following spans:
-\[ \homrallXYtoZ {\alpha':C'} {\harrapp {C_1'} {\alpha'}{J_0\alpha}}{\harrapp {D_1} {\beta} {F_0 \alpha'}}\]
-
-If we have a unit and therefore unit category and functor, then we can
-now define limits as a special case of Kan extensions.
-\begin{definition}
-  The limit of a functor $J : \textrm{Functor}I\,C$ is a right Kan
-  extension $\textrm{Ran}_E J$ where $E$ is the unique functor $E :
-  \textrm{Functor}I\,1$.
-\end{definition}
-
-Unraveling this definition, we get that $\harrapp{C_1} \alpha {\lim_{i} J(i)}$ is a coherent family (i.e., equalizer) of
-\[  \homrallXYtoZ {i:I} {1} {\harrapp{C_1} \alpha {J_0(i)}}  \]
-with the equalizing condition being that for any $p : \homrallXYtoZ
-{i:I} {1} {\harrapp{C_1} \alpha {J_0(i)}}$ and $f : \harrapp{I_1} {i'}{i}$ and $g : \harrapp{1}{pt}{E_0 i'}$
-\[ \homrappXtoYatZ {p} {(\appbinary {1_t} {pt} {\pendappXtoY {1_r} {pt}} {pt} {(\appbinary {1_t} {pt} g {E_0 i'} {(\appunary {E_1} {i'} f {i})} {i})} {i})} {i}
-= \appbinary {C_t} {\alpha} {(\homrappXtoYatZ {p} g {i'})} {J_0 i'} {(\appbinary {C_t} {J_0 i'} {(\appunary {J_1} i f {i'})} {J_0 i} {\pendappXtoY {C_r} {J_0 i}} {J_0 i})} {J_0 i}
-\]
-We can greatly simplify using the $\eta$ principle for $1$ and
-eliminating some reflexivities to reduce this to
-\[ \homrappXtoYatZ {p} {()} {i} =
-\appbinary {C_t} {\alpha} {(\homrappXtoYatZ {p} {()} {i'})} {J_0 i'} {{(\appunary {J_1} i f {i'})}} {J_0 i}
-\]
-which reads exactly as the commuting condition for a cone.
-
-We can similarly define \emph{left} Kan extensions by a dual
-formula. Left Kan extensions are the ``minimal'' solution to the same
-filling problem as right Kan extensions. I.e., they have a universal
-morphism out of them. They can be described by a coend formula, but in
-our profunctor calculus, they are easiest to describe using a similar
-formula to right Kan extensions, but using the contravariant hom of
-profunctors.
-
-\begin{definition}
-  Let $C,C',D$ be categories and let $F : \textrm{Functor}\,C\,D$ and
-  $J: \textrm{Functor}\,C'\,C$ be functors. Then the left Kan
-  extension specification profunctor is a profunctor
-  $\textrm{LanSpec}_J\,F : \textrm{Prof}\,C\,D$ defined as
-
-  \[ \textrm{RanSpec}_J\,F = \textrm{Hom}(D)[F;] \triangleleft \textrm{Hom}(C')[J;]  \]
-  And a left Kan extension is a functor $\textrm{Lan}_J F :
-  \textrm{Functor}\,C'\,D$ that represents this profunctor on the
-  left.
-\end{definition}
-
-Unraveling the definition we get the hom span $\harrapp {D_1}{\textrm{Ran}_JF {\alpha'}} \beta$ is an equalizer of
-\[ \homlallXYtoZ \alpha {\harrapp {C_1'} {J_0 \alpha} {\alpha'}} {\harrapp{D_1} {F_0 \alpha} \beta}\]
-
-\begin{definition}
-  The colimit of a functor $J : \textrm{Functor}I\,C$ is a left Kan
-  extension $\textrm{Lan}_E J$ where $E$ is the unique functor $E :
-  \textrm{Functor}I\,1$.
-\end{definition}
-
-We can similarly unravel this definition to see that this looks
-essentially like the classical definition of colimits as universal
-cocones.
+%%   where $\psi_{dom}$ and $\chi_{cod}$ are  as
+%%   \[ \psi_{dom} =\homrlambdaXatYdotZ p {\gamma'} {\homrlambdaXatYdotZ e \gamma
+%%     {\homrappXtoYatZ {\phi} {(\apptrinary {P_t} \beta {\pendappXtoY {D_r} \beta} \beta p {\gamma'} e \gamma)} \gamma}}
+%%   \]
+%%   \[
+%%   \psi_{cod} = \homrlambdaXatYdotZ p {\gamma'} {\homrlambdaXatYdotZ e \gamma
+%%     {\apptrinary {Q_t} \alpha {\pendappXtoY {C_r} \alpha} \alpha {(\homrappXtoYatZ \phi {p} {\gamma'})} {\gamma'} e \gamma}}
+%%   \]
+%% \end{definition}
+
+%% We can easily define a contravariant hom $P \triangleleft Q$ by a dual
+%% definition (TODO).
 
@@
 
-\subsection{Duality}
+\subsection{Duality Principle}
+
+One common source of category theoretic arguments is the duality
+principle: since the opposite of any category is also a category, any
+statement that quantifies over all categories is equally true of their
+opposites. Furthermore in a formal language of categories, any
+statement can be dualized to a statement where domains and codomains
+are swapped and composition is flipped.
+%
+A similar duality principle holds in SCT.
+%
+The basic idea is to simply reverse the orientation of all \emph{set}
+expressions by swapping which variable is covariant and which is
+contravariant.
+%
+This has the effect of reversing the order of the context $\Phi$.
+
+We can define this as a simple translation $\cdot^* : SCT \to SCT$
+defined on syntactic forms of SCT.
+
+\begin{figure}
+    \begin{align*}
+      (\harr P Q)^* &= \harr {P^*} {Q^*}\\
+      (\alpha:\cat C)^* &= \alpha: \cat C^*\\
+      (\alpha:\cat C; \beta:\cat D)^* &= \beta:\cat D^*; \alpha:\cat C^*\\
+      (\Phi,x:R,\Psi)^* &= \Psi^*,x:R^*, \Phi^*\\
+      (\homrallXYtoZ \alpha P Q)^* &= \homlallXYtoZ \alpha {P^*} {Q^*}\\
+      (\homlallXYtoZ \alpha P Q)^* &= \homrallXYtoZ \alpha {P^*} {Q^*}\\
+      (\homrlambdaXatYdotZ x \alpha s)^* &= \homllambdaXatYdotZ x \alpha s^*\\
+      (\homrappXtoYatZ s t a)^* &= \homlappXtoYatZ {s^*} {t^*} {a^*}\\
+      (\tensorexistsXwithYandZ \beta P Q)^* &=
+      \tensorexistsXwithYandZ \beta {Q^*}{P^*}\\
+      (\tensorintroatXwithYandZ b s t)^* &= \tensorintroatXwithYandZ {b^*} {t^*} {s^*}\\
+      (\punitinXfromYtoZ {\cat C} {a_1}{a_2})^* &= \punitinXfromYtoZ {\cat C^*} {a_1^*}{a_2^*}
+    \end{align*}
+
+  \caption{Duality Involution}
+\end{figure}
+
+TODO: example of applying the principle in our calculus.
 
 \section{Semantics}
+
+In this section we establish our desired semantic soundness theorem,
+that all constructions in the theory can be interpreted in any virtual
+equipment with the corresponding universal properties.
+%
+This proof is straightforward, since our constructions are explicitly
+modeled on the universal properties, with the main difference being
+that our notion of restriction is strict.
+
+%% \section{Conservativity of Homs, Tensors}
+
+%% The soundness theorem is not entirely sufficient for our goals,
+%% however.
+%% %
+%% In particular, we are interested in applying the internal category
+%% constructions to models that only have units, and do not necessarily
+%% admit left/right homs or tensor products of horizontal arrows.
+%% %
+%% To do so, we now prove a \emph{conservativity} result for the calculus
+%% with homs and tensors over a simpler calculus whose only profunctor
+%% constructor is the unit.
+%% %
+%% We must take care in defining the more restricted calculus, however.
+%% %
+%% If we simply remove homs and tensors from the calculus, then there is
+%% no way to internalize the transformations as types, and so no way to
+%% actually define internal categories!
+%% %
+%% For example, if we remove the hom sets, there is no way to abstract
+%% over the composition transformation:
+%% \[\alpha_1:\cat C, f : \harrapp {\cat C_1}{\alpha_1}{\alpha_2}, \alpha_2:\cat C, g : \harrapp{\cat C_1}{\alpha_2}{\alpha_3}, \alpha_3:\cat C \vdash \textrm{compose} : \harrapp{\cat C_1}{\alpha_1}{\alpha_3}
+%% \]
+%% %
+%% So in addition to removing the hom sets and tensors, we add a
+%% ``jumbo'' combined form of the hom set and end type, described in
+%% \cref{fig:jumbo}.
+%% %
+%% This type directly internalizes a transformation judgment of arbitrary
+%% large context, as opposed to the end type which internalized only
+%% transformations from an ``empty'' context.
+%% %
+
+%% \begin{figure}
+%%   \begin{mathpar}
+%%     \inferrule*[right=JumboHomForm]
+%%     {\Gamma \vdash \Phi \isavectx \and \Gamma \pipe \underline \Phi \vdash R \isaSet}
+%%     {\Gamma \vdash \Phi \triangleright R}
+
+%%     \inferrule*[right=JumboHomI]
+%%     {\Gamma \pipe \Phi \vdash s : R}
+%%     {\Gamma \vdash \lambda \Phi. s : \Phi \triangleright R}
+
+%%     \inferrule*[right=JumboHomE]
+%%     {\Gamma \vdash M : \Phi \triangleright R\and
+%%       \Gamma \vdash \phi : \Psi \to \Phi}
+%%     {\Gamma \pipe \Psi \vdash M \triangleright \phi : R[\underline \phi]}
+
+%%     \inferrule*[right=Jumbo$\beta$]
+%%     {}
+%%     {\lambda \Phi. M \triangleright (\Phi) = M}
+
+%%     \inferrule*[right=Jumbo$\eta$]
+%%     {\Gamma \vdash M : \Phi \triangleright R}
+%%     {M = \lambda \Phi. M \triangleright \Phi}
+%%   \end{mathpar}
+%%   \caption{Jumbo End/Homs}
+%%   \label{fig:jumbo}
+%% \end{figure}
+
+%% %
+%% To demonstrate this, we prove a \emph{conservativity} result for the
+%% calculus with homs and tensors over a calculus containing only units,
+%% and a ``jumbo'' internalization
 
@@
 
-%% \section{Styles of lists, enumerations, and descriptions}\label{sec:itemStyles}
-
-%% List of different predefined enumeration styles:
-
-%% \begin{itemize}
-%% \item \verb|\begin{itemize}...\end{itemize}|
-%% \item \dots
-%% \item \dots
-%% %\item \dots
-%% \end{itemize}
-
-%% \begin{enumerate}
-%% \item \verb|\begin{enumerate}...\end{enumerate}|
-%% \item \dots
-%% \item \dots
-%% %\item \dots
-%% \end{enumerate}
-
-%% \begin{alphaenumerate}
-%% \item \verb|\begin{alphaenumerate}...\end{alphaenumerate}|
-%% \item \dots
-%% \item \dots
-%% %\item \dots
-%% \end{alphaenumerate}
-
-%% \begin{romanenumerate}
-%% \item \verb|\begin{romanenumerate}...\end{romanenumerate}|
-%% \item \dots
-%% \item \dots
-%% %\item \dots
-%% \end{romanenumerate}
-
-%% \begin{bracketenumerate}
-%% \item \verb|\begin{bracketenumerate}...\end{bracketenumerate}|
-%% \item \dots
-%% \item \dots
-%% %\item \dots
-%% \end{bracketenumerate}
-
-%% \begin{description}
-%% \item[Description 1] \verb|\begin{description} \item[Description 1]  ...\end{description}|
-%% \item[Description 2] Fusce eu leo nisi.
-Cras eget orci neque, eleifend dapibus felis.
-Duis et leo dui.
-Nam vulputate, velit et laoreet porttitor, quam arcu facilisis dui, sed malesuada risus massa sit amet neque.
-%% \item[Description 3]  \dots
-%% %\item \dots
-%% \end{description}
-
-%% \cref{testenv-proposition} and \autoref{testenv-proposition} ...
-
-%% \section{Theorem-like environments}\label{sec:theorem-environments}
-
-%% List of different predefined enumeration styles:
-
-%% \begin{theorem}\label{testenv-theorem}
-%% Fusce eu leo nisi. Cras eget orci neque, eleifend dapibus felis. Duis et leo dui. Nam vulputate, velit et laoreet porttitor, quam arcu facilisis dui, sed malesuada risus massa sit amet neque.
-%% \end{theorem}
-
-%% \begin{lemma}\label{testenv-lemma}
-%% Fusce eu leo nisi. Cras eget orci neque, eleifend dapibus felis. Duis et leo dui. Nam vulputate, velit et laoreet porttitor, quam arcu facilisis dui, sed malesuada risus massa sit amet neque.
-%% \end{lemma}
-
-%% \begin{corollary}\label{testenv-corollary}
-%% Fusce eu leo nisi. Cras eget orci neque, eleifend dapibus felis. Duis et leo dui. Nam vulputate, velit et laoreet porttitor, quam arcu facilisis dui, sed malesuada risus massa sit amet neque.
-%% \end{corollary}
-
-%% \begin{proposition}\label{testenv-proposition}
-%% Fusce eu leo nisi. Cras eget orci neque, eleifend dapibus felis. Duis et leo dui. Nam vulputate, velit et laoreet porttitor, quam arcu facilisis dui, sed malesuada risus massa sit amet neque.
-%% \end{proposition}
-
-%% \begin{conjecture}\label{testenv-conjecture}
-%% Fusce eu leo nisi. Cras eget orci neque, eleifend dapibus felis. Duis et leo dui. Nam vulputate, velit et laoreet porttitor, quam arcu facilisis dui, sed malesuada risus massa sit amet neque.
-%% \end{conjecture}
-
-%% \begin{observation}\label{testenv-observation}
-%% Fusce eu leo nisi. Cras eget orci neque, eleifend dapibus felis. Duis et leo dui. Nam vulputate, velit et laoreet porttitor, quam arcu facilisis dui, sed malesuada risus massa sit amet neque.
-%% \end{observation}
-
-%% \begin{exercise}\label{testenv-exercise}
-%% Fusce eu leo nisi. Cras eget orci neque, eleifend dapibus felis. Duis et leo dui. Nam vulputate, velit et laoreet porttitor, quam arcu facilisis dui, sed malesuada risus massa sit amet neque.
-%% \end{exercise}
-
-%% \begin{definition}\label{testenv-definition}
-%% Fusce eu leo nisi. Cras eget orci neque, eleifend dapibus felis. Duis et leo dui. Nam vulputate, velit et laoreet porttitor, quam arcu facilisis dui, sed malesuada risus massa sit amet neque.
-%% \end{definition}
-
-%% \begin{example}\label{testenv-example}
-%% Fusce eu leo nisi. Cras eget orci neque, eleifend dapibus felis. Duis et leo dui. Nam vulputate, velit et laoreet porttitor, quam arcu facilisis dui, sed malesuada risus massa sit amet neque.
-%% \end{example}
-
-%% \begin{note}\label{testenv-note}
-%% Fusce eu leo nisi. Cras eget orci neque, eleifend dapibus felis. Duis et leo dui. Nam vulputate, velit et laoreet porttitor, quam arcu facilisis dui, sed malesuada risus massa sit amet neque.
-%% \end{note}
-
-%% \begin{note*}
-%% Fusce eu leo nisi. Cras eget orci neque, eleifend dapibus felis. Duis et leo dui. Nam vulputate, velit et laoreet porttitor, quam arcu facilisis dui, sed malesuada risus massa sit amet neque.
-%% \end{note*}
-
-%% \begin{remark}\label{testenv-remark}
-%% Fusce eu leo nisi. Cras eget orci neque, eleifend dapibus felis. Duis et leo dui. Nam vulputate, velit et laoreet porttitor, quam arcu facilisis dui, sed malesuada risus massa sit amet neque.
-%% \end{remark}
-
-%% \begin{remark*}
-%% Fusce eu leo nisi. Cras eget orci neque, eleifend dapibus felis. Duis et leo dui. Nam vulputate, velit et laoreet porttitor, quam arcu facilisis dui, sed malesuada risus massa sit amet neque.
-%% \end{remark*}
-
-%% \begin{claim}\label{testenv-claim}
-%% Fusce eu leo nisi. Cras eget orci neque, eleifend dapibus felis. Duis et leo dui. Nam vulputate, velit et laoreet porttitor, quam arcu facilisis dui, sed malesuada risus massa sit amet neque.
-%% \end{claim}
-
-%% \begin{claim*}\label{testenv-claim2}
-%% Fusce eu leo nisi. Cras eget orci neque, eleifend dapibus felis. Duis et leo dui. Nam vulputate, velit et laoreet porttitor, quam arcu facilisis dui, sed malesuada risus massa sit amet neque.
-%% \end{claim*}
-
-%% \begin{proof}
-%% Fusce eu leo nisi. Cras eget orci neque, eleifend dapibus felis. Duis et leo dui. Nam vulputate, velit et laoreet porttitor, quam arcu facilisis dui, sed malesuada risus massa sit amet neque.
-%% \end{proof}
-
-%% \begin{claimproof}
-%% Fusce eu leo nisi. Cras eget orci neque, eleifend dapibus felis. Duis et leo dui. Nam vulputate, velit et laoreet porttitor, quam arcu facilisis dui, sed malesuada risus massa sit amet neque.
-%% \end{claimproof}
+\section{Unit and Tensor}
+we can define the
+\emph{unit} , which in set-theoretic settings we
+might call the \emph{identity/equality judgment} $A \equiv_{\cat C}
+A'$.
+%
+First the formation rule says given two objects $A,A'$ in the same
+category $\cat C$, we can form the hom-set from $A$ to $A'$ $A
+\to_{\cat C} A'$.
+%
+The introduction rule says that given any object $A$, we can define
+the identity arrow from $A$ to itself $\id(A) : A \to_{\cat C} A$.
+%
+Without knowing more about the category in question, these are the
+only arrows we know exist.
+%
+
+The elimination form is trickiest to define in a natural deduction style.
+%
+First, let's consider the universal property of the Hom. From
+Crutwell-Shulman, we want a 2-cell
+
+\[% https://tikzcd.yichuanshen.de/#N4Igdg9gJgpgziAXAbVABwnAlgFyxMJZARgBoAGAXVJADcBDAGwFcYkQBBEAX1PU1z5CKcqQBM1Ok1bsuvftjwEiY8ZIYs2iTjz4gMioUTLF10rSABOMAGaMekmFADm8IqBuWIAWyRkQOBBIqlKa7AASProeXr6IogFBiP6MWGAWUBDMAEaMbDQAFjD0UOyQ6SA0OPRY9trlbPIgnlHxVUkhqRXamTl5lSBFJWUE+QE1deCjDtxAA
+\begin{tikzcd}
+                    & A \arrow[ldd, no head, Rightarrow] \arrow[rdd, no head, Rightarrow] &   \\
+                    & refl                                                                &   \\
+A \arrow[rr, "Hom"] &                                                                     & A
+\end{tikzcd}\]
+
+That is universal/opcartesian in that for any 2-cell $t$:
+
+\[ % https://tikzcd.yichuanshen.de/#N4Igdg9gJgpgziAXAbVABwnAlgFyxMJZABgBpiBdUkANwEMAbAVxiRACEQBfU9TXfIRQBGclVqMWbbrxAZseAkQBMY6vWatEIAIIy+CwUQAsaiZrZ6eBgUpQA2Mxqnb2Acn1z+ioclXLxZy0QAGFPeVtfUwD1SWCAEXDvIxQAZlJhQLi2HA8ucRgoAHN4IlAAMwAnCABbJDIQHAgkYWsQKtqW6iakVRAGLDBgqDo4AAtCkGoJuig2SCHPDrrEPp7EdP7B4LgIAbnpmFmkMCYGBmoGOgAjGAYABWS7fphynCnzFxAACU625aQm3Wpi2i20I3Gk0Ox20C1Y-2qKwa6wArAjOogQet7OiVijus1EI5PsEAErcChcIA
+\begin{tikzcd}
+B \arrow[r] \arrow[rrdd] & {} \arrow[r, no head, dashed] & A \arrow[rr, "Hom"] &    & A \arrow[rr, no head, dashed] &  & B' \arrow[lldd] \\
+                         &                               &                     & t' &                               &  &                 \\
+                         &                               & C \arrow[rr, "R"]   &    & D                             &  &                
+\end{tikzcd}\]
+
+there exists a $t'$ which extends $t$ in that the following is equal
+to $t$:
+\[% https://tikzcd.yichuanshen.de/#N4Igdg9gJgpgziAXAbVABwnAlgFyxMJZABgBoAmAXVJADcBDAGwFcYkQAhEAX1PU1z5CKAIwVqdJq3Y8+IDNjwEi5cTQYs2iEAEFZ-RUKIAWNZM3s9vA4OUoAbGY3TtHAOT75ApcOSrjEs5aIADCngq2vqYB6lLBACLh3kYoAMykqYFx7Dge1l6GdsjpxFkW2lZyET5EZKWx5SBJhb5i9eYuTfnVKciO7UEy3clF6SJlnQBOMABmjDwSMFAA5vBEoDOTEAC2SGQgOBBIIvmbO8c0h0iqIIxYYMFQ9HAAFksgNG-0UOyQD55nXaIG5XRDpW73YJwCB3H6fGDfJBgZiMRg0Rj0ABGMEYAAURsJbrMcB8OsEABLnU5bIHg0GmCH-bRPV7veGI7R-NjU86IfaggCsPKBDNB9mFSAFlyOiEcZPYACUATSkAAOaXXdGQ9hQCDMTGMNjsn6cghGg70LDzU3-CWIdUHGXgu5MkC6-WG0lfE3gM2knCW62+21yQFIACcGr5dsjjuO+xdUJhWDhIG9SJRaItVt+frtIn5MrEjKTsK9CJNyNRl0DuZDGxViALUYdiZ1zzeqfTNu5ocbDtBImLbeZHbZaYrdd7Dd5Q6jDJHbr1BvN3eD5oDOZ7C24QA
+\begin{tikzcd}
+{} \arrow[dd] \arrow[r]  & {} \arrow[dd] \arrow[rr, no head, dashed] &                     & A \arrow[ldd, no head, Rightarrow] \arrow[rdd, no head, Rightarrow] \arrow[rrr, no head, dashed] &                               &  & {} \arrow[dd, no head, Rightarrow] \\
+                         &                                           &                     & refl                                                                                             &                               &  &                                    \\
+B \arrow[r] \arrow[rrdd] & {} \arrow[r, no head, dashed]             & A \arrow[rr, "Hom"] &                                                                                                  & A \arrow[rr, no head, dashed] &  & B' \arrow[lldd]                    \\
+                         &                                           &                     & t'                                                                                               &                               &  &                                    \\
+                         &                                           & C \arrow[rr, "R"]   &                                                                                                  & D                             &  &
+\end{tikzcd}\]
+
+Clearly, our introduction rule models this first 2-cell. Furthermore,
+the construction in the second condition is modeled faithfully by the
+following sequent rule in the figure. This is a J-like induction rule:
+to construct a term using an input of the Hom type, it is sufficient
+to consider the case in which the input is the identity morphism.
+%
+However, building in a substitution the usual way we would (pushing
+the substitution into the continuation $N$) doesn't work properly
+because we can't necessarily extract a substitution for the
+continuation from the substitution for the the whole term. A
+substitution for the whole term would be of the form
+\[\phi,D_1/\delta_1,M/f,D_2/\delta_2,\psi\]
+The the candidate for the continuation's substitution would be
+$\phi,?,\psi$ but there is not necessarily something to go in the
+middle here since $D_1,D_2$ do not necessarily have the same
+\emph{input variable}. However, if we assume the existence already of
+the cotensor products the problem can be simplified. Instead of
+allowing for a general elimination of a hom term anywhere in the
+context, we can instead make a remarkably restricted rule: we can only
+eliminate a Hom term when it is the \emph{sole} input in the context:
+
+\[ \inferrule*[right=Hom Elim Very Simple]{\alpha:\cat C \vdash N : Q[\alpha/\alpha_1,\alpha/\alpha_2]}{\alpha_1:\cat C, f : \alpha_1 \to_{\cat C} \alpha_2, \alpha_2:\cat C \vdash J(\alpha_1.\alpha_2.
+Q, \alpha.
+N, \alpha_1, f, \alpha_2) : Q}\]
+
+While this rule looks very impoverished at first sight, it has the
+virtue of easily supporting a built-in cut:
+
+\[ \inferrule*[right=Hom Elim VSimple Subst]{
+  Q \prof(\alpha_1:\cat C; \alpha_2:\cat C)\and
+  \Phi \vdash M : A_1 \to_{\cat C} A_2\and
+  \alpha:\cat C \vdash N : Q[\alpha/\alpha_1,\alpha/\alpha_2]}{\Phi \vdash J(\alpha_1.\alpha_2. Q, \alpha. N, A_1, M, A_2) : Q[A_1/\alpha_1, A_2/\alpha_2]}\]
+
+However, this rule on its own clearly is not enough to support our
+desired universal property shown above diagrammatically. To prove our
+completeness theorem, we will need to show that this rule is enough to
+give us the desired universal property \emph{assuming the presence of
+  left and right cotensors}.
+
+So first, we must show that the ``proper'' sequent rule
+
+\[ \inferrule*[right=Hom-Left (Sequent Style)]
+    {\Phi_1[\alpha/\alpha_1], \alpha:\cat C, \Phi_2[\alpha/\alpha_2]\vdash Q[\alpha/\alpha_1,\alpha/\alpha_2]}
+    {\Phi_1, \alpha_1 : \cat C, f : \alpha_1 \to_{\cat C} \alpha_2, \alpha_2 : \cat C, \Phi_2\vdash : Q}
+\]
+
+is admissible.
+
+Which we can sketch as follows:
+
+\[
+\inferrule*
+    {\inferrule*{\inferrule*{\Phi_1[\alpha/\alpha_1],\alpha:\cat C, \Phi_2[\alpha/\alpha_1] \vdash Q[\alpha/\alpha_1;\alpha/\alpha_2]}{\alpha:\cat C \vdash \Phi_1[\alpha/\alpha_1] \triangleright Q[\alpha/\alpha_1;\alpha/\alpha_2] \triangleleft \Phi_2[\alpha/\alpha_2]}}{\alpha_1 : \cat C, f : \alpha_1 \to_{\cat C} \alpha_2, \alpha_2 : \cat C\vdash \Phi_1 \triangleright Q \triangleleft \Phi_2}}
+    {\Phi_1, \alpha_1 : \cat C, f : \alpha_1 \to_{\cat C} \alpha_2, \alpha_2 : \cat C, \Phi_2\vdash Q}
+\]
+
+Which using our syntax sugar is given by the term:
+\begin{mathpar}
+  \inferrule*
+    {\inferrule*
+      {\inferrule*
+        {\Phi_1[\alpha/\alpha_1],\alpha:\cat C, \Phi_2[\alpha/\alpha_1] \vdash N : Q[\alpha/\alpha_1;\alpha/\alpha_2]}
+        {\alpha:\cat C \vdash \lambda^\triangleright \Phi_1[\alpha/\alpha_1].
+\lambda^\triangleleft \Phi_2[\alpha/\alpha_2].
+N : \Phi_1[\alpha/\alpha_1] \triangleright Q[\alpha/\alpha_1;\alpha/\alpha_2] \triangleleft \Phi_2[\alpha/\alpha_2]}}
+      {\alpha_1 : \cat C, f : \alpha_1 \to_{\cat C} \alpha_2, \alpha_2 : \cat C\vdash J(\alpha_1.\alpha_2.\Phi_1 \triangleright Q \triangleleft \Phi_2, \alpha.
+\lambda^\triangleright \Phi_1[\alpha/\alpha_1].
+\lambda^\triangleleft \Phi_2[\alpha/\alpha_2].
+N, \alpha_1,f,\alpha_2) : \Phi_1 \triangleright Q \triangleleft \Phi_2}}
+    {\Phi_1, \alpha_1 : \cat C, f : \alpha_1 \to_{\cat C} \alpha_2, \alpha_2 : \cat C, \Phi_2\vdash
+      \Phi_1 \triangleright (J(\alpha_1.\alpha_2.\Phi_1 \triangleright Q \triangleleft \Phi_2, \alpha.
+\lambda^\triangleright \Phi_1[\alpha/\alpha_1].
+\lambda^\triangleleft \Phi_2[\alpha/\alpha_2].
+N, \alpha_1,f,\alpha_2)) \triangleleft \Phi_2 : Q}
+\end{mathpar}
+
+We need to show $\beta$ and $\eta$ hold.
+
+First, $\beta$.
+
+\begin{align*}
+  &(\Phi_1 \triangleright (J_\to(\alpha_1.\alpha_2.\Phi_1 \triangleright Q \triangleleft \Phi_2, \alpha.
+\lambda^\triangleright \Phi_1[\alpha/\alpha_1].
+\lambda^\triangleleft \Phi_2[\alpha/\alpha_2].
+N, \alpha_1,f,\alpha_2)) \triangleleft \Phi_2) [\Phi_1[A/\alpha_1],A/\alpha_1,\id(A),A/\alpha_2,\Phi_2[A/\alpha_2]] \\
+  &=
+  \Phi_1[A/\alpha_1] \triangleright (J_\to(\alpha_1.\alpha_2.\Phi_1 \triangleright Q \triangleleft \Phi_2, \alpha.
+\lambda^\triangleright \Phi_1[\alpha/\alpha_1].
+\lambda^\triangleleft \Phi_2[\alpha/\alpha_2].
+N, A,\id(A), A)) \triangleleft \Phi_2[A/\alpha_2]\\
+  &=
+  \Phi_1[A/\alpha_1] \triangleright (\lambda^\triangleright \Phi_1[A/\alpha_1]. \lambda^\triangleleft \Phi_2[A/\alpha_2]. N[A/\alpha]) \triangleleft \Phi_2[A/\alpha_2]\\
+  &=  (\lambda^\triangleleft \Phi_2[A/\alpha_2]. N[A/\alpha]) \triangleleft \Phi_2[A/\alpha_2]\\
+  &=  N[A/\alpha]
+\end{align*}
+
+\begin{figure}
+  \begin{mathpar}
+    \inferrule*[right=Full Unit Elim*]
+    {\Phi_1[\alpha/\alpha_1], \alpha:\cat C, \Phi_2[\alpha/\alpha_2]\vdash N : Q[\alpha/\alpha_1,\alpha/\alpha_2]}
+    {\Phi_1, \alpha_1 : \cat C, f : \alpha_1 \to_{\cat C} \alpha_2, \alpha_2 : \cat C, \Phi_2\vdash J_\to'(\Phi_1, \alpha_1 : \cat C, f : \alpha_1 \to_{\cat C} \alpha_2, \alpha_2 : \cat C, \Phi_2, Q, \alpha.
+N) : Q}
+
+    J_\to'(\Phi_1, \alpha_1 : \cat C, f : \alpha_1 \to_{\cat C} \alpha_2, \alpha_2 : \cat C, \Phi_2, Q, \alpha. N) =
+    \Phi_1 \triangleright (J(\alpha_1.\alpha_2.\Phi_1 \triangleright Q \triangleleft \Phi_2, \alpha.
+\lambda^\triangleright \Phi_1[\alpha/\alpha_1].
+\lambda^\triangleleft \Phi_2[\alpha/\alpha_2].
+N, \alpha_1,f,\alpha_2)) \triangleleft \Phi_2
+
+    \inferrule*[right=Full Unit $\beta$*]
+    {}
+    {J_\to'(\Phi_1, \alpha_1 : \cat C, f : \alpha_1 \to_{\cat C} \alpha_2, \alpha_2 : \cat C, \Phi_2, Q, \alpha.
+N)[\Phi_1[\alpha/\alpha_1],\alpha/\alpha_1, \id(\alpha)/f, \alpha/\alpha_2, \Phi_2[\alpha/\alpha_2]]
+    = N}
+
+    (\Phi_1 \triangleright J(\alpha_1.\alpha_2.\Phi_1 \triangleright Q \triangleleft \Phi_2, \alpha.
+\lambda^\triangleright \Phi_1[\alpha/\alpha_1].
+\lambda^\triangleleft \Phi_2[\alpha/\alpha_2].
+N, \alpha_1,f,\alpha_2) \triangleleft \Phi_2)[\Phi_1[\alpha/\alpha_1],\alpha/\alpha_1, \id(\alpha)/f, \alpha/\alpha_2, \Phi_2[\alpha/\alpha_2]]\\
+    = \Phi_1[\alpha/\alpha_1] \triangleright J(\alpha_1.\alpha_2.\Phi_1 \triangleright Q \triangleleft \Phi_2, \alpha.
+\lambda^\triangleright \Phi_1[\alpha/\alpha_1].
+\lambda^\triangleleft \Phi_2[\alpha/\alpha_2].
+N, \alpha,\id(\alpha),\alpha) \triangleleft \Phi_2[\alpha/\alpha_2]\\
+    = \Phi_1[\alpha/\alpha_1] \triangleright
+    (\lambda^\triangleright \Phi_1[\alpha/\alpha_1]. \lambda^\triangleleft \Phi_2[\alpha/\alpha_2]. N)
+    \triangleleft \Phi_2[\alpha/\alpha_2]\\
+    = N
+
+    \inferrule*[right=Unit$\eta$]
+    {\Phi_1, \alpha_1 : \cat C, f : \alpha_1 \to_{\cat C} \alpha_2, \alpha_2 : \cat C, \Phi_2 \vdash M : Q}
+    {M =
+      J_\to'(\Phi_1, \alpha_1 : \cat C, f : \alpha_1 \to_{\cat C} \alpha_2, \alpha_2 : \cat C, \Phi_2, Q, \alpha.
+M[\Phi_1[\alpha/\alpha_1],\alpha/\alpha_1,\id(\alpha), \alpha/\alpha_2,\Phi_2[\alpha/\alpha_2]])}
+
+    M
+    =(\triangleleft\beta\triangleright\beta)
+    \Phi_1 \triangleright
+    (\lambda^\triangleright \Phi_1. \lambda^\triangleleft \Phi_2. M)
+    \triangleleft \Phi_2
+    = \Phi_1 \triangleright
+    J_\to(\alpha_1.\alpha_2.Q, \alpha.
+\lambda^\triangleright \Phi_1[\alpha/\alpha_1].
+\lambda^\triangleleft \Phi_2[\alpha/\alpha_2].
+M[\alpha/\alpha_1, \id(\alpha)/f,\alpha/\alpha_2], \alpha_1, f, \alpha_2)
+    \triangleleft \Phi_2
+    =
+    J_\to'(\Phi_1, \alpha_1 : \cat C, f : \alpha_1 \to_{\cat C} \alpha_2, \alpha_2 : \cat C, \Phi_2, Q, \alpha. N)
+  \end{mathpar}  
+  \caption{Admissible Unit Universal Property Rules}
+\end{figure}
+
+First, the syntax for tensor is based on the coend formula that gives
+the definition in $Cat$, but using logical syntax $\exists$ rather
+than integrals because I don't like integral syntax and the syntax
+matches existential quantification in type theory.
+%
+Note that $\exists \beta. ~Q \odot R$ is the application of a single
+connective to $Q$ and $R$, and not separable into a quantifier
+$\exists$ connective and a product connective $\odot$.
+%
+Introduction, elimination and $\beta\eta$ are similar to a combination
+of existential quantification and the tensor product in linear logic.
+%
+The introduction rule is justified by the fact that semantically what
+we have is a transformation
+\[ \textrm{pair} : (\alpha : \cat C,Q,\beta:\cat D,R,\gamma:\cat E) \Rightarrow \exists \beta. R \odot Q \]
+And the introduction rule is just building in a substitution to this
+transformation.
+%
+
+For the elim rule we can pull a similar trick as with the Hom
+type. First, the rule where there's only one variable.
+\[ \inferrule*[right=Tensor elim Simple]
+  {\alpha : \cat C, q : Q, \beta : \cat D, r: R, \gamma : \cat E \vdash N : S}
+  {\alpha : \cat C, x : \exists \beta. Q \odot R, \gamma : \cat E \vdash J_\odot(x, q.\beta.r. N) : S}
+\]
+
+\begin{figure}
+\begin{mathpar}
+    \inferrule*[right=Full Tensor Elim*]
+    {\Phi_1, \alpha:\cat C, q:Q, \beta:\cat D, r:R, \gamma:\cat E \Phi_2\vdash N : S}
+    {\Phi_1, \alpha : \cat C, x : \exists \beta:\cat D. Q \odot R, \gamma : \cat E, \Phi_2
+      \vdash J_\odot'(\Phi_1, \alpha : \cat C, x : \exists \beta:\cat D. Q \odot R, \gamma : \cat E, \Phi_2; S;q.\beta.r.N) : S}
+
+    J_\odot'(\Phi_1, \alpha : \cat C, x : \exists \beta:\cat D. Q \odot R, \gamma : \cat E, \Phi_2;S; q.\beta.r.N)
+    = \Phi_1 \triangleright J_\odot(\alpha.\gamma.\Phi_1 \triangleright S \triangleleft \Phi_2, \alpha.q.\beta.r.\gamma.
+\lambda^\triangleleft \Phi_1.
+\lambda^\triangleright\Phi_2.
+N, \alpha, x, \gamma) \triangleleft \Phi_2
+
+    J_\odot'(\Phi_1, \alpha : \cat C, x : \exists \beta:\cat D. Q \odot R, \gamma : \cat E, \Phi_2; q.\beta.r.N)[\Phi_1,\alpha/\alpha,(q,\beta,r)/x,\gamma/\gamma,\Phi_2] = N
+
+  \inferrule*[right=Full Tensor elim*]{ S \prof(\alpha:\cat C; \gamma : \cat E)\and
+    \Phi \vdash M : \exists \beta. Q[A/\alpha] \odot R[G/\gamma]\and
+    \alpha : \cat C, q : Q, \beta : \cat D, r: R, \gamma : \cat E \vdash N : S}
+  {\Phi \vdash J_\odot(\alpha. \gamma. S, \alpha.q.\beta.r.\gamma.N, A, M, G) : S[A/\alpha, G/\gamma]}
+
+  J_\odot(\alpha. \gamma. S, \alpha.q.\beta.r.\gamma.N, A, M, G)[\phi] =
+  J_\odot(\alpha. \gamma. S, \alpha.q.\beta.r.\gamma.N, A[d^-\phi], M[\phi], G[d^+\phi])
+
+  %% \inferrule*[right=Tensor elim Sequent]
+  %% {\Phi,\alpha : \cat C, q : Q, \beta : \cat D, r: R, \gamma : \cat E,\Psi \vdash N : S}
+  %% {\Phi, \alpha : \cat C, x : \exists \beta. Q \odot R, \gamma : \cat E,\Psi \vdash J_\odot(x;q.\beta.r. N) : S}
+
+  %% \inferrule*[right=Tensor elim Subst]
+  %% {\Phi,\alpha : \cat C, q : Q, \beta : \cat D, r: R, \gamma : \cat E,\Phi' \vdash N : S
+  %% \and
+  %% \phi : \Psi \to \Phi,\alpha : \cat C, q : Q, \beta : \cat D, r: R, \gamma : \cat E,\Phi'}
+  %% {\Psi  \vdash J_\odot(x;q.\beta.r. N)[\phi] : S}
+
+  \inferrule*[right=Tensor $\beta$]
+  {}{J_\odot(\alpha.\gamma.S, \alpha.q.\beta.r.\gamma.N,A,(M,B,N),G) = N[A/\alpha,M/q,B/\beta,N/r,G/\gamma]}
+
+  \inferrule*[right=Tensor$\eta$]
+  {\alpha : \cat C, x : \exists \beta. Q \odot R, \gamma : \cat E \vdash M : S}
+  {\alpha : \cat C, x : \exists \beta. Q \odot R, \gamma : \cat E \vdash M =
+    J(\alpha.\gamma.S, \alpha.q.\beta.r.\gamma. M[\alpha/\alpha,(q,\beta,r)/x,\gamma/\gamma])
+    : S}
+\end{mathpar}
+\caption{Tensor Admissible Universal Property Rules}
+\label{fig:tensor}
+\end{figure}
 

### 2022-02-07 paper/lipics-v2021-sample-article.tex
more synthetic progress

@@
 \end{lemma}
+\begin{proof}
+  TODO
+\end{proof}
 
@@
 constructions. Semantically, these are related to so-called
-\emph{Fubini} theorems about interchange of ends.
-
+\emph{Fubini} theorems about interchange of (co)-ends.
 \begin{lemma}[{Fubini}]
@@
 \end{lemma}
-
+\begin{proof}
+  TODO
+\end{proof}
 \subsection{Composition, Functoriality, Profunctoriality, Naturality}
@@
 \begin{definition}[Composition of Morphisms]
-  We define composition of morphisms as the following term of type $\hombinary {\alpha_1} {\punitinXfromYtoZ {\cat C}{\alpha_1} {\alpha_2}} {\alpha_2} {\punitinXfromYtoZ {\cat C}{\alpha_2} {\alpha_3}} {\alpha_3} {\punitinXfromYtoZ {\cat C}{\alpha_1} {\alpha_3}}$:
+  We define composition of morphisms as the following term of type $\Pi_{\cat C:\Cat}\hombinary {\alpha_1} {\punitinXfromYtoZ {\cat C}{\alpha_1} {\alpha_2}} {\alpha_2} {\punitinXfromYtoZ {\cat C}{\alpha_2} {\alpha_3}} {\alpha_3} {\punitinXfromYtoZ {\cat C}{\alpha_1} {\alpha_3}}$:
   \[
+  \lambda \cat C.
 \punitelimtoYwithkontZ {\hlambda {\alpha_1}{\alpha_2} \homrallXYtoZ {\alpha_3} {\punitinXfromYtoZ {\cat C} {\alpha_2} {\alpha_3}} {\punitinXfromYtoZ {\cat C} {\alpha_1} {\alpha_3}}} {\lambdaunary {\alpha} {f : \punitinXfromYtoZ {\cat C}{\alpha} {\alpha_3}} {\alpha_3} f}
   \]
+  We call this term $\textrm{comp}$
 \end{definition}
@@
 
+\begin{lemma}[Equivalence of Composition Definitions]
+  TODO
+\end{lemma}
+
 It should follow by $\beta\eta$ that this is associative and unital.
 
-Next, let's prove synthetically that every type $\alpha : \cat C
-\vdash A : \cat D$ is in fact functorial.
+Next, let's prove synthetically that every functor $F : \varr{\cat
+  C}{\cat D}$ is functorial in this notion of composition.
 \begin{definition}
-  There is a morphism
-  \[ \alpha^o:\cat C;\alpha':\cat C \vdash f : \alpha \to_{\cat C} \alpha' \vdash
-  \when f = \id_{\alpha}.~ \id_{A} : A[\alpha] \to_{\cat D} A[\alpha']
+  For any $F : \varr {\cat C} {\cat D}$, we can define a morphism $F_1
+  : \homunary {\alpha_1} {\punittinXfromYtoZ {\cat C} {\alpha_1}
+    {\alpha_2}} {\alpha_2} {\punittinXfromYtoZ {\cat D} {F \alpha_1}
+    {F \alpha_2}}$ as
+  \[ F_1 = \punitelimtoYwithkontZ {\hlambda {\alpha_1}{\alpha_2} \punittinXfromYtoZ {\cat D} {F \alpha_1}{F \alpha_2}}
+  {\pendlambdaXdotY \alpha \punitrefl {\cat D} {F \alpha}}
   \]
-  which we call $A[f/\alpha]$.
-
   Note that
@@
 
-Next we can define natural transformations as specific kinds of
-terms whose type is $(F \alpha \to G \alpha')$.
+And similarly, every profunctor $R : \harr {\cat C}{\cat D}$ is
+\emph{pro-functorial}.
 \begin{definition}
-  A natural transformation $\phi$ from $\alpha : \cat C \vdash A : \cat D$ to
-  $A'[\alpha]$ of the same type is a term
-  \[ \alpha\pipe \cdot \vdash \phi_\alpha : A[\alpha] \to_{\cat D} A'[\alpha] \]
-
-  Note that
-  \[ \alpha;\alpha' \pipe f : \alpha \to \alpha' \vdash
-  f \circ \phi_{\alpha} = \phi_{\alpha'} \circ f
+  For any $R : \varr {\cat C} {\cat D}$, we can define left and right
+  actions of composition:
+  \[ \homunary {\alpha_1} {\tensorexistsXwithYandZ {\alpha_2} {(\punitinXfromYtoZ {\cat C} {\alpha_1}{\alpha_2})} {\tensorexistsXwithYandZ {\alpha_3} {(\harrapp R {\alpha_2} {\alpha_3})}}{(\punitinXfromYtoZ {\cat C} {\alpha_3}{\alpha_4})}} {\alpha_4} {\harrapp R {\alpha_1}{\alpha_4}}
   \]
-  Because if $f$ is the identity, then both sides are the same.
+  Furthermore we can show that they are functorial in that the induced functions
+  TODO
+  are equal
+\end{definition}
+
+Next we can define natural transformations as specific kinds of terms
+whose type is $\pendallXdotY {\alpha} \punitinXfromYtoZ {\cat D} {F
+  \alpha}{G \alpha}$. In fact we can generalize this to any
+endo-profunctor $P$:
+\begin{definition}[Natural element]
+  A natural element of an endo-profunctor $P : \harr {\cat C}{\cat C}$
+  is a term of type $\pendallXdotY {\alpha} \harrapp P \alpha \alpha$.
+
+  A natural transformation is the special case where $P =
+  \punitinXfromYtoZ {\cat D} {F \alpha}{G \alpha}$ for functors $F, G
+  : \varr {\cat C}{\cat D}$.
+
+  This is natural in the sense that the incuded terms
+  \[ \homunary {\alpha_1}{\punitinXfromYtoZ {} {\alpha_1}{\alpha_2}} {\alpha_2} {\harrapp P {\alpha_1}{\alpha_2}}  \]
+  and
+  \[ \pendallXdotY {\alpha_2} \homlallXYtoZ {\alpha_1} {\punitinXfromYtoZ {} {\alpha_1}{\alpha_2}} {\harrapp P {\alpha_1}{\alpha_2}}\]
+
+  are equal under the associativity equivalence between these two
+  types.
 \end{definition}
@@
 \section{Internal Category Theory}
+
+A downside of working synthetically is that since everything we define
+must be functorial/natural, we cannot prove any theorems about
+functions that are not assumed to be functorial or transformations
+that are not assumed ot be natural.
+%
+These kinds of constructions are surprisingly common in basic category theory.
+%
+For instance, when defining a left adjoint $F$ to a given right
+adjoint functor $U$, the functoriality of $F$ is often not assumed,
+rather we are given that each $F a$ has a certain universal mapping
+property and this can then be used to show the action of $F$ on
+objects extends to an action on morphisms.
+%
+This definition is quite important for applications of category theory
+to type theory: in type theory when a type constructor is defined it
+is given introduction, elimination and $\beta\eta$ rules, but no
+primitive syntax is given for a functorial action.
+%
+The fact that functoriality is determined from the universal mapping
+property of the type is then a key step in proving the type theory is
+sound and complete with respect to models which are expected to have
+certain functors on them.
+
+As another example, a common lemma states that an \emph{possibly
+unnatural} transformation is in fact natural if it is the inverse of a
+natural transformation.
+%
+Neither of these examples can be stated as synthetic theorems, as it
+is impossible to quantify over ``possibly unnatural transformations''
+or ``functions on objects''.
+%
+Inside the calculus, we cannot ``see'' that categories have an
+underlying set of objects.
+%
+Fortunately there is a way to use our calculus in a different manner
+to study these constructions: by defining categories \emph{internal
+to} SCT.
+%
+This reproduces classical constructions, since there is a model where
+a category $\cat C$ is interpreted simply as a set, functors as
+functions and profunctors $\harr {\cat C}{\cat D}$ as \emph{spans},
+i.e. sets $R$ with (source and target) projections $\cat C \leftarrow
+R \rightarrow \cat D$.
+%
+The proofs read like the classical proofs (with some care additional
+attention on the ordering of variable usage), but by virtue of being
+syntactic automatically apply to any notion of category internal to a
+virtual equipment, which includes classical internal categories,
+enriched categories and a broad range of generalized multicategory.
+
+We note that a synthetic approach to model this situation may be
+possible were we to extend our calculus with a \emph{modality} $|\cat
+C|$ that takes a category to the discrete category on the set of
+objects. We discuss this in \cref{sec:discussion}.
 

### 2022-02-08 paper/lipics-v2021-sample-article.tex
work on adjunctions

@@
   \label{lem:yo-and-co-yo}
-  Let $P : \harr {\cat C} {\cat D}$. Then the following sets are (naturally) isomorphic to $\harrapp P \alpha \beta$
+  Let $P : \harr {\cat C} {\cat D}$.
+
+  For any $\gamma : \cat E \vdash b : \cat D$, the following are naturally isomorphic to $\harrapp P \alpha b$:
   \begin{enumerate}
-  \item $\tensorexistsXwithYandZ {\beta'} {\harrapp P \alpha {\beta'}} {(\punitinXfromYtoZ {\cat D} {\beta'} {\beta})}$
-  \item $\tensorexistsXwithYandZ {\alpha'} {(\punitinXfromYtoZ {C} {\alpha} {\alpha'})} {\harrapp P {\alpha'}{\beta}}$
-  \item $\homrallXYtoZ {\beta'} {(\punitinXfromYtoZ {D} {\beta'} {\beta})} {\harrapp P {\alpha} {\beta'}}$
-  \item $\homlallXYtoZ {{\alpha'}} {(\punitinXfromYtoZ {C} {\alpha'} {\alpha})} {\harrapp P {{\alpha'}} {\beta}}$
+  \item $\homrallXYtoZ {\beta} {(\punitinXfromYtoZ {D} {\beta} {b})} {\harrapp P {\alpha} {\beta}}$
+  \item $\tensorexistsXwithYandZ {\beta} {\harrapp P \alpha {\beta}} {(\punitinXfromYtoZ {\cat D} {\beta} {b})}$
+  \end{enumerate}
+
+  For any $\gamma : \cat E \vdash a : \cat C$, the following are naturally isomorphic to $\harrapp P a \beta$:
+  \begin{enumerate}
+  \item $\homlallXYtoZ {{\alpha}} {(\punitinXfromYtoZ {C} {\alpha} {a})} {\harrapp P {{\alpha}} {\beta}}$
+  \item $\tensorexistsXwithYandZ {\alpha} {(\punitinXfromYtoZ {C} {a} {\alpha})} {\harrapp P {\alpha}{\beta}}$
   \end{enumerate}
@@
 \begin{proof}
-  TODO
+  TODO: what's better, build the substitution in or prove the
+  substitution can be built in afterwards? The one building in the
+  substitution obscures the connection to the normal tensor/hom/unit
+  relationship, but the one with the built in subst is obviously more
+  general.
 \end{proof}
@@
 
-There are two common definitions of adjoint functors, one in terms of
-Hom-sets and one in terms of units and counits.
-
-\begin{definition}[Hom Iso Adjunction]
-  A functor $\alpha : \cat C \vdash L : \cat D$ is left adjoint to a
-  functor $\beta : \cat D \vdash R : \cat C$ if there is an isomorphism
-  \[ (L[\alpha] \to_{\cat D} \beta) \cong (\alpha \to_{\cat C} R[\beta]) \]
+Next, we demonstrate that the two common definitions of adjoint
+functors are equivalent: isomorphism of hom-sets and units and
+co-units.
+%
+The proof goes through most easily, and utilizes duality most
+effectively, if we factorize the construction through an intermediate
+notion of when a functor \emph{represents} a profunctor.
+
+\begin{definition}[Functor Represents a Profunctor]
+  A functor $G : \varr{\cat D}{\cat C}$ right-represents a profunctor $R :
+  \harr {\cat C}{\cat D}$ when there is an isomorphism of profunctors:
+  \[ \punitinXfromYtoZ {\cat C} {\alpha} {G \beta} \cong \harrapp R {\alpha}{\beta}\]
+
+  We say a functor $F : \varr{\cat C}{\cat D}$ \emph{left-represents} a profunctor $R : \harr {\cat C}{\cat D}$ when there is an isomorphism of profunctors
+  \[ \punitinXfromYtoZ {\cat D} {F \alpha} {\beta} \cong \harrapp R {\alpha}{\beta}\]
 \end{definition}
 
-\begin{definition}[Unit/Counit Adjunction]
-  Alternatively, $L$ is left adjoint to $R$ if there exist
-  transformations
-  \[ \eta : \alpha \to_{\cat C} R[L[\alpha]/\beta] \] % unit of the monad 1 -> RL 
-  \[ \epsilon : L[R[\beta]/\alpha] \to_{\cat D} \beta  \] % counit of the comonad LR -> 1
-  satisfying the triangle/zig-zag identities
+This is a parameterized generalization of the notion of an object
+representing a functor $R : C^{op} \to Set$.
+These notions are precisely dual. $F$ left-represents a profunctor $R$
+precisely if it right-represents it in the opposite SCT.
+
+Note that trivially any functor $G$ right-represents
+$\hlambda\alpha\beta \punitinXfromYtoZ {\cat D}{\alpha} {G \beta}$ and
+dually, any functor $F$ left-represents
+$\hlambda\alpha\beta\punitinXfromYtoZ {\cat C}{F\alpha}{\beta}$.
+
+Then we can show that any right-representation isomorphism is
+determined by a co-unit:
+\begin{lemma}[(Co)-unit of a Representation]
+  If $G: \varr {\cat D}{\cat C}$ right-represents $R : \harr {\cat C}{\cat D}$, then there is a natural element
+
+  \[ \pendallXdotY \beta \harrapp R {G\beta} \beta \]
+
+  Such that the half of the isomorphism
+
+  \[ \punitinXfromYtoZ {\cat C} {\alpha} {G \beta} \multimap \harrapp R {\alpha}{\beta} \]
+
+  is defined by composition with the natural element.
+
+  Dually, any $F$ left-representing isomorphism with $R$ is determined
+  by a term of type $\pendallXdotY \alpha \harrapp R {\alpha}
+  {F\alpha}$.
+\end{lemma}
+\begin{proof}
+  A homomorphism $\punitinXfromYtoZ {\cat C} {\alpha} {G \beta} \multimap \harrapp R {\alpha}{\beta}$ is by definition a term of type
+  \[ \pendallXdotY \alpha \homrallXYtoZ {\beta} {(\punitinXfromYtoZ {\cat C} \alpha {G \beta})} {\harrapp R \alpha \beta} \]
+  By \cref{lem:fubini}, this type is isomorphic to 
+  \[ \pendallXdotY \beta \homlallXYtoZ {\alpha} {(\punitinXfromYtoZ {\cat C} \alpha {G \beta})} {\harrapp R \alpha \beta} \]
+  By \cref{lem:yo-and-co-yo}, this type is isomorphic to
+  \[ \pendallXdotY \beta {{\harrapp R {(G \beta)} \beta}} \]
+
+  Therefore this half of the isomorphism is determined by a natural
+  element. The construction in the proof of \cref{lem:yo-and-co-yo} is
+  indeed by composition (i.e., unit elimination).
+
+  The $F$ case is precisely dual.
+\end{proof}
+
+Combining these two, we get the unit-counit definition for adjoint
+functors
+
+\begin{definition}[Adjunction by unit/co-unit]
+  We say $G : \varr {\cat D}{\cat C}$ is right adjoint to $F : \varr
+  {\cat C}{\cat D}$ if there is a unit
+  \[ \eta : \pendallXdotY \alpha \punitinXfromYtoZ \alpha {G(F(\alpha))}\]
+  and a co-unit
+  \[ \epsilon : \pendallXdotY \beta \punitinXfromYtoZ {F(G(\beta))} \beta \]
+
+  satisfying the triangle identities...
   \[ \id_{R[\beta]} = R[\epsilon_{\beta}] \circ \eta_{R[\beta]} \]
@@
 
-\begin{definition}[Universal Morphism Adjunction]
-  A functor $\alpha: C \vdash L : D$ is a left adjoint if there exists a 
-\end{definition}
-
-We should be able to synthetically prove these are equivalent.
-
-\subsection{Tensors and CoTensors/Powers}
-
-David Jaz Myers tells me that to do all of formal category theory in
-an equipment, you just need to have hom sets, tensors and cotensors.
-
-As another example synthetic theorem, we can prove the covariant and
-contravariant \emph{co-yoneda lemmas}, which are precisely the
-statement that Hom is the unit of the tensor product.
-\begin{lemma}[CoYoneda]
-  We can define isomorphisms
-  \[ \exists \beta. (\alpha \to_{\cat C} \beta) \odot Q \equiv Q[\alpha/\beta] \]
-  \[ \exists \beta. P \odot (\beta \to_{\cat C} \gamma) \equiv P[\gamma/\beta] \]
-  The forward direction is defined by considering the case that the
-  function is the identity and the backward direction by using the
-  identity.
-\end{lemma}
+\begin{theorem}[Adjunctions]
+  $G$ is right adjoint to $F$ if and only if there is a natural isomorphism
+  $\punitinXfromYtoZ {\cat D} {(F \alpha)} \beta \cong \punitinXfromYtoZ {\cat D} \alpha {G \beta}$
+\end{theorem}
 \begin{proof}
-  The terms are
-  \[ f : \exists \beta. (\alpha\to_{\cat C} \beta) \odot Q \vdash
-  \lett (\beta,g,h) = f.~ \when g = \id_{\alpha}.~ h : Q[\alpha/\beta]
-  \]
-  And the inverse
-  \[ h : Q[\alpha/\beta] \vdash (\alpha, \id_{\alpha}, h) : \exists \beta. (\alpha\to_{\cat C}\beta)\odot Q\]
-
-  One direction of iso follows by $\beta$, the other by $\eta$.
-\end{proof}
-
-And we get another synthetic theorem: the Yoneda lemma, which says
-that Hom is the unit for the cotensor product.
-\begin{lemma}[Synthetic Yoneda]
-  We can define isomorphisms
-  \[ \forall \alpha.~ (\alpha \to \beta) \triangleright P[\alpha;\gamma] \cong P[\beta/\alpha;\gamma] \]
-  \[ \forall \gamma.~ P[\alpha;\gamma] \triangleleft (\beta \to \gamma) \cong P[\alpha;\beta/\gamma]\]
-\end{lemma}
-\begin{proof}
-  Forward direction
-  \[ f : \forall \alpha.~(\alpha\to\beta)\triangleright P \vdash
-  f(\beta,\id_{\beta})^\triangleright : P[\beta/\alpha]
-  \]
-  Backwards
-  \[ x : P[\beta/\alpha]\vdash
-  \lambda^\triangleright(\alpha,\phi : \alpha \to \beta).~\when \phi = \id_{\beta}.~ x : \forall \alpha. (\alpha\to \beta) \triangleright P
-  \]
-  One direction is by $\beta$, the other $\eta$.
+  First, by \cref{lem:co-unit}, the forward direction of the
+  isomorphism is equivalent to providing a term
+  \[ \eta : \pendallXdotY \alpha \punitinXfromYtoZ \alpha {G(F(\alpha)) } \]
+
+  Second, the backward direction of the isomorphism is equivalent to
+  providing a term
+  \[ \epsilon : \pendallXdotY \beta \punitinXfromYtoZ {F(G(\beta))} \beta \]
+
+  And the triangle laws state exactly that the induced functions are
+  isomorphisms.
 \end{proof}

### 2022-02-08 paper/lipics-v2021-sample-article.tex
start working on kan extensions, fix errors

@@
   For any $F : \varr {\cat C} {\cat D}$, we can define a morphism $F_1
-  : \homunary {\alpha_1} {\punittinXfromYtoZ {\cat C} {\alpha_1}
-    {\alpha_2}} {\alpha_2} {\punittinXfromYtoZ {\cat D} {F \alpha_1}
+  : \homunary {\alpha_1} {\punitinXfromYtoZ {\cat C} {\alpha_1}
+    {\alpha_2}} {\alpha_2} {\punitinXfromYtoZ {\cat D} {F \alpha_1}
     {F \alpha_2}}$ as
-  \[ F_1 = \punitelimtoYwithkontZ {\hlambda {\alpha_1}{\alpha_2} \punittinXfromYtoZ {\cat D} {F \alpha_1}{F \alpha_2}}
+  \[ F_1 = \punitelimtoYwithkontZ {\hlambda {\alpha_1}{\alpha_2} \punitinXfromYtoZ {\cat D} {F \alpha_1}{F \alpha_2}}
   {\pendlambdaXdotY \alpha \punitrefl {\cat D} {F \alpha}}
@@
 
-\subsection{Limits and Kan Extensions}
-
-We can then use our notion of Hom of profunctors
-
-The right Kan extension of $F$ along $J$ is the best (maximal) way to
+\subsection{(co)-Limits and Kan Extensions}
+
+As a further example of the synthetic approach to universal
+properties, we consider \emph{Kan extensions}, which are a
+generalization of limits, and their dual \emph{Kan lifts}, which
+generalize co-limits.
+
+The right Kan extension of $F$ along $J$ is the ``best'' (maximal) way to
 fill in the following square:
@@
   Let $C,C',D$ be categories and let $F : \textrm{Functor}\,C\,D$ and
-  $J: \textrm{Functor}\,C'\,C$ be functors. Then the right Kan
+  $J: \textrm{Functor}\,C'\,C$ be functors. Then a right Kan extension
+  is a functor $\textrm{Ran}_JF :\varr {\cat C'}{\cat D}$ that
+  right-represents the following profunctor:
+  \[ \]
+
+  Then the right Kan
   extension specification profunctor is a profunctor

### 2022-02-09 paper/lipics-v2021-sample-article.tex
kan extensions

@@
 
-\subsection{(co)-Limits and Kan Extensions}
+\subsection{Kan Extensions/Lifts}
 
 As a further example of the synthetic approach to universal
-properties, we consider \emph{Kan extensions}, which are a
-generalization of limits, and their dual \emph{Kan lifts}, which
-generalize co-limits.
+properties, we consider right and left \emph{Kan extensions}, which
+are generalizations of limits and colimits, respectively.
 
@@
 \begin{tikzcd}
-C \arrow[d, "J"] \arrow[r, "F"] & D \\
-C'                              &  
+\cat C \arrow[d, "J"] \arrow[r, "F"] & \cat D \\
+\cat C'                              &  
 \end{tikzcd}\]
-I.e. a functor $Ran_J F : C' \to D$ with a natural transformation
-$Ran_J F \Rightarrow F$ that is universal something something.
-
-If $D$ is complete, there is a formula using ends for this. Using hom
-of profunctors, we can directly adapt this to define a
-\emph{profunctor} that is a universal filler. We can then say that a
-right Kan extension is precisely a functor that represents this
-profunctor.
-
-\begin{definition}
-  Let $C,C',D$ be categories and let $F : \textrm{Functor}\,C\,D$ and
-  $J: \textrm{Functor}\,C'\,C$ be functors. Then a right Kan extension
-  is a functor $\textrm{Ran}_JF :\varr {\cat C'}{\cat D}$ that
-  right-represents the following profunctor:
-  \[ \]
-
-  Then the right Kan
-  extension specification profunctor is a profunctor
-  $\textrm{RanSpec}_J\,F : \textrm{Prof}\,D\,C$ defined as
-
-  \[ \textrm{RanSpec}_J\,F = \textrm{Hom}(C')[;J] \triangleright \textrm{Hom}(D)[;F] \]
-
-  And a right Kan extension is a functor
-  $\textrm{Ran}_JF:\textrm{Functor}\,C'\,D$ that is a
-  right-representation of $\textrm{RanSpec}_J F$.
+I.e. a functor $\textrm{Ran}_J F : \cat C' \to \cat D$ with a
+universal natural transformation $\textrm{Ran}_J F \circ J \Rightarrow
+F$. A left kan extension $\textrm{Lan}_J F$ is the same but with a
+natural transformation $F \Rightarrow \textrm{Lan}_J F \circ J $.
+
+We can express the concept in SCT as representing a certain profunctor
+constructed using the profunctor hom:
+\begin{definition}[Kan Extensions]
+  Given functors $F : \varr {\cat C}{\cat D}$ and $J : \varr {\cat
+    C}{\cat C'}$, a right Kan extension is a functor $\textrm{Ran}_J F
+  : \varr {\cat C'}{\cat D} $that right-represents the profunctor
+  \[ \hlambda {\beta:\cat D}{\alpha': \cat C} \homrallXYtoZ {\alpha}{(\punitinXfromYtoZ {\cat C'}{\alpha'}{J\alpha})} {\punitinXfromYtoZ {\cat D}{\beta}{F \alpha}} \]
+
+  A left Kan extensions is a functor $\textrm{Lan}_J F : \varr {\cat
+    C'}{\cat D}$ is a functor satisfying the dual property: it
+  \emph{left}-represents the profunctor
+  \[\hlambda {\alpha':\cat C}{\beta:\cat D} \homlallXYtoZ {\alpha}{(\punitinXfromYtoZ {\cat C'}{J\alpha}{\alpha'})} {\punitinXfromYtoZ {\cat D}{F \alpha}{\beta}}
+  \]
 \end{definition}
 
-Writing this out more pointfully, we get that the hom set
-$\harrapp{D_1} {\beta}{\textrm{Ran}_JF\, \alpha}$ is an equalizer of
-the following spans:
-\[ \homrallXYtoZ {\alpha':C'} {\harrapp {C_1'} {\alpha'}{J_0\alpha}}{\harrapp {D_1} {\beta} {F_0 \alpha'}}\]
-
-If we have a unit and therefore unit category and functor, then we can
-now define limits as a special case of Kan extensions.
-\begin{definition}
-  The limit of a functor $J : \textrm{Functor}I\,C$ is a right Kan
-  extension $\textrm{Ran}_E J$ where $E$ is the unique functor $E :
-  \textrm{Functor}I\,1$.
-\end{definition}
-
-Unraveling this definition, we get that $\harrapp{C_1} \alpha {\lim_{i} J(i)}$ is a coherent family (i.e., equalizer) of
-\[  \homrallXYtoZ {i:I} {1} {\harrapp{C_1} \alpha {J_0(i)}}  \]
-with the equalizing condition being that for any $p : \homrallXYtoZ
-{i:I} {1} {\harrapp{C_1} \alpha {J_0(i)}}$ and $f : \harrapp{I_1} {i'}{i}$ and $g : \harrapp{1}{pt}{E_0 i'}$
-\[ \homrappXtoYatZ {p} {(\appbinary {1_t} {pt} {\pendappXtoY {1_r} {pt}} {pt} {(\appbinary {1_t} {pt} g {E_0 i'} {(\appunary {E_1} {i'} f {i})} {i})} {i})} {i}
-= \appbinary {C_t} {\alpha} {(\homrappXtoYatZ {p} g {i'})} {J_0 i'} {(\appbinary {C_t} {J_0 i'} {(\appunary {J_1} i f {i'})} {J_0 i} {\pendappXtoY {C_r} {J_0 i}} {J_0 i})} {J_0 i}
+The limit of a functor $F : \varr{\cat C}{\cat D}$ is a special case
+of a Kan extension where $J : \varr{\cat C}{1}$ is the unique functor
+to the terminal category. Similarly, the colimit of $F$ is the left
+Kan extension with respect to the same $J$.
+%
+Specializing to this case, the representability for limits means:
+\[ \punitinXfromYtoZ {\cat D}{\beta}{\textrm{lim} F} \cong \homrallXYtoZ {\alpha}{(\punitinXfromYtoZ {\cat 1}{\alpha'}{()})}{\punitinXfromYtoZ {\cat D}{\beta}{F \alpha}} \]
+And if we view a morphism in $\cat 1$ as being trivial, then this
+essentially states that mapping $\beta$ to $\textrm{lim} F$ is the
+same thing as constructing a cone: for all $\alpha :\cat C$, a
+morphism from $\beta$ to $F \alpha$, naturally in $\alpha$.
+
+The usual formulation of Kan extensions is in terms of a universal
+morphism.
+%
+We can reproduce this formulation using our established theory of
+equivalences for homs of profunctors.
+\begin{lemma}
+  One half of the representability isomorphism for a right Kan
+  extension $\textrm{Ran}_J F$ is determined by a natural
+  transformation
+  \[ \pendallXdotY \alpha \punitinXfromYtoZ {\cat D}{(\textrm{Ran}_JF)(J\alpha)}{F\alpha} \]
+
+  Dually, one half of the representability isomorphism for a left Kan
+  extension $\textrm{Lan}_J F$ is determined by a natural
+  transformation
+  \[ \pendallXdotY \alpha \punitinXfromYtoZ {\cat D}{F\alpha}{(\textrm{Lan}_JF)(J\alpha)} \]
+\end{lemma}
+\begin{proof}
+  By \cref{lem:unit-repr}, one half of representability is determined by a natural element
+  \[ \pendallXdotY {\alpha'} \homrallXYtoZ {\alpha}{(\punitinXfromYtoZ {\cat C'}{\alpha'}{J\alpha})} {\punitinXfromYtoZ {\cat D}{(\textrm{Ran}_J F)\alpha'}{F \alpha}} \]
+  which, by \cref{lem:fubini} is isomorphic to
+  \[
+  \pendallXdotY {\alpha} \homlallXYtoZ {\alpha'}{(\punitinXfromYtoZ {\cat C'}{\alpha'}{J\alpha})}{\punitinXfromYtoZ {\cat D}{(\textrm{Ran}_J F)\alpha'}{F \alpha}}
   \]
-We can greatly simplify using the $\eta$ principle for $1$ and
-eliminating some reflexivities to reduce this to
-\[ \homrappXtoYatZ {p} {()} {i} =
-\appbinary {C_t} {\alpha} {(\homrappXtoYatZ {p} {()} {i'})} {J_0 i'} {{(\appunary {J_1} i f {i'})}} {J_0 i}
-\]
-which reads exactly as the commuting condition for a cone.
-
-We can similarly define \emph{left} Kan extensions by a dual
-formula. Left Kan extensions are the ``minimal'' solution to the same
-filling problem as right Kan extensions. I.e., they have a universal
-morphism out of them. They can be described by a coend formula, but in
-our profunctor calculus, they are easiest to describe using a similar
-formula to right Kan extensions, but using the contravariant hom of
-profunctors.
-
-\begin{definition}
-  Let $C,C',D$ be categories and let $F : \textrm{Functor}\,C\,D$ and
-  $J: \textrm{Functor}\,C'\,C$ be functors. Then the left Kan
-  extension specification profunctor is a profunctor
-  $\textrm{LanSpec}_J\,F : \textrm{Prof}\,C\,D$ defined as
-
-  \[ \textrm{RanSpec}_J\,F = \textrm{Hom}(D)[F;] \triangleleft \textrm{Hom}(C')[J;]  \]
-  And a left Kan extension is a functor $\textrm{Lan}_J F :
-  \textrm{Functor}\,C'\,D$ that represents this profunctor on the
-  left.
-\end{definition}
-
-Unraveling the definition we get the hom span $\harrapp {D_1}{\textrm{Ran}_JF {\alpha'}} \beta$ is an equalizer of
-\[ \homlallXYtoZ \alpha {\harrapp {C_1'} {J_0 \alpha} {\alpha'}} {\harrapp{D_1} {F_0 \alpha} \beta}\]
-
-\begin{definition}
-  The colimit of a functor $J : \textrm{Functor}I\,C$ is a left Kan
-  extension $\textrm{Lan}_E J$ where $E$ is the unique functor $E :
-  \textrm{Functor}I\,1$.
-\end{definition}
-
-We can similarly unravel this definition to see that this looks
-essentially like the classical definition of colimits as universal
-cocones.
-
-\subsection{Right (Left) Adjoints preserve (co-)Limits}
-
-Let's prove that classic theorem!
-
-\begin{theorem}
-  Let $\alpha : \cat C \vdash A \cat D$ be an arbitrary functor and
-  $\beta : \cat D \vdash R : \cat E$ be a functor with a left adjoint
-  $L$. Then, if the limit $\lim_\alpha A$ exists, then $\lim_\alpha
-  R[A/\beta]$ also exists and is
-  \[ \lim_\alpha R[A/\beta] = R[\lim_\alpha A] \]
+  which, by \cref{lem:yo-and-co-yo}, is isomorphic to 
+  \[ \pendallXdotY \alpha \punitinXfromYtoZ {\cat D}{(\textrm{Ran}_JF)(J\alpha)}{F\alpha} \]
+  by taking $P = \hlambda {\alpha'}{\alpha} {\punitinXfromYtoZ {\cat D}{(\textrm{Ran}_J F)\alpha'}{F \alpha}}$.
+
+  The proof for left Kan extensions follows by \cref{lem:duality}.
+\end{proof}
+
+With the definition based on representability, we can easily prove the
+generalization of the famous theorem, that right adjoints preserve not
+just limits but all right Kan extensions
+
+\begin{theorem}[Right/Left Adjoints preserve Right/Left Kan Extensions]
+  Let $K : \varr{\cat C}{\cat D}$ and $J : \varr{\cat C}{\cat C'}$.
+
+  If the right Kan extension $\textrm{Ran}_J K$ exists, and if $G :
+  \varr {\cat D}{\cat D'}$ has a left adjoint $F : \varr{\cat D'}{\cat
+    D}$, then the right Kan extension $\textrm{Ran}_J G\circ K$ exists
+  and is given by $G \circ \textrm{Ran}_J K$.
+
+  Similarly, if the left Kan extension $\textrm{Lan}_J K$ exists and
+  if $F : \varr{\cat D}{\cat D'}$ has a right adjoint $G : \varr{\cat
+    D'}{\cat D}$, then the left Kan extension $\textrm{Lan}_J F\circ
+  K$ exists and is given by $F \circ \textrm{Ran}_J K$.
 \end{theorem}
 \begin{proof}
-  We need to give an isomorphism
-  \[ \gamma \to R[\lim_\alpha A/\beta] \cong PRan_{()}{\alpha. R[A/\beta]} = \forall \alpha. (\gamma \to R[A/\beta]) \triangleleft (() \to_1 ()) \]
-  By assumption we have an isomorphism
-  \[ l : \beta \to \lim_\alpha A \cong PRan_{()}{\alpha. A} = \forall \alpha. (\beta \to A[\alpha]) \triangleleft (()\to_1 ())\]
-  TODO.
+  By the composition of the following natural isomorphisms:
+  \begin{align*}
+    \punitinXfromYtoZ {\cat D'} {\beta'} {G ((\textrm{Ran}_J K) \alpha')} &\cong \punitinXfromYtoZ {\cat D}{F \beta'} {(\textrm{Ran}_J K) \alpha'} \tag{Adjunction}\\
+    &\cong \homrallXYtoZ {\alpha}{(\punitinXfromYtoZ {\cat C'}{\alpha'}{J\alpha})} {\punitinXfromYtoZ {\cat D}{F \beta'}{K \alpha}} \tag{Ran assumption, \cref{lem:natural-restriction}}\\
+    &\cong \homrallXYtoZ {\alpha}{(\punitinXfromYtoZ {\cat C'}{\alpha'}{J\alpha})} {\punitinXfromYtoZ {\cat D}{\beta'}{G(K \alpha)}} \tag{Adjunction}\\
+  \end{align*}
 \end{proof}
-
-%% \subsection{Additive Facts}
-
-%% I think the following isomorphisms should hold, which connect additive
-%% connectives on objects with the analogous connective on profunctors:
-%% \[ () \to_1 () \cong () \]
-%% \[ (\pi_1 \alpha, \pi_2 \alpha) \to_{\cat C \times \cat D} (\pi_1 \beta, \pi_2 \beta) \cong
-%% ((\pi_1 \alpha) \to_{\cat C} (\pi_1 \beta)) \times ((\pi_2 \alpha) \to_{\cat C} (\pi_2 \beta)) \]
 

### 2022-02-11 paper/lipics-v2021-sample-article.tex
discussion

@@
 
-\section{Related Work}
+\section{Discussion}
+
+\paragraph{Directed Type Theory}
 
@@
 String Diagrams?
+
+\paragraph{Semantic extensions}
+
+\paragraph{Extensions}
+
+While we have demonstrated that many common constructions in category
+theory can be expressed using the simple combination of homs, tensors
+and units, there are several that are out of reach for SCT due to its
+restrictive judgmental structure.
+
+First, while we can add products to the calculus, adding in a
+\emph{functor category} ${\cat C}^{\cat D}$ would necessitate adding a
+syntax for functors of multiple arguments, to allow for the lambda rule:
+\[ \inferrule {\alpha:\cat C, \beta:\cat D \vdash e : \cat E} {\alpha:\cat C \vdash \lambda \beta:\cat D. e : \cat E} \]
+This would then require also some extension to profunctors for
+allowing 0 or more covariant/contravariant dependencies as well.
+%
+In addition, we could consider extending our semantics with a notion
+of \emph{opposite category}, with the property that a covariant
+dependence of a profunctor on a variable in a category should be
+equivalent to a contravariant dependence on a variable in the opposite
+category.
+%
+This could possibly be done by introducing a modal distinction
+$\alpha:^{+}\cat C$ for covariant dependence and $\alpha:^{-}\cat C$
+for contravariant dependence, with the opposite category being a
+modality that internalizes flipping the variance as a type
+constructor.
+%
+With these two extensions, possibly, the category contexts for
+functors and profunctors could be unified as one notion of a sequence
+of object variables with variances.
+%
+With this unification, it would then be simpler to incorporate a third
+extension: a universe object, corresponding to an object
+$\textrm{Set}$, which would allow us to internalize the profunctor
+judgment in the functor judgment, modeling that in classical category
+theory, profunctors are the same as certain functors to the category
+of Sets.
+%
+While all of these extensions seem plausible, they all move away to
+the simple ordered type theory structure that allowed us to define
+units, tensor and hom profunctors so easily. It is unclear how to
+define these types in this extended structure. As a simple example, if
+profunctors have many variables of mixed variance, rather than exactly
+one of each variance, it is unclear what the appropriate
+generalization of the transformation context $\Phi$ should be.
+
+An orthogonal extension to these would be to consider adding a
+\emph{discretization} modality $|\cat C|$ that gives the discrete
+category on the underlying set of objects of $\cat C$. This would give
+us the ability to formalize concepts like ``function on objects'' or
+``unnatural transformation'' without the need to define internal
+categories. This could be modeled by adding a modal distinction
+$\alpha:^{\#}\cat C$ for ``crisp'' dependence on an object of $\cat
+C$, i.e., dependence merely on the object of $\cat C$ with no
+functoriality/naturality conditions. This would then be internalized
+by the discretization modality.
 

### 2022-02-11 paper/lipics-v2021-sample-article.tex
abstract update

@@
 \begin{abstract}
-  We present a domain-specific type theory for formal category theory
-  constructions and proofs. The type theory axiomatizes notions of
+  We present a domain-specific type theory for constructions and
+  proofs in category theory. The type theory axiomatizes notions of
   category, functor, profunctor and a generalized form of natural
-  transformations. By restricting the type theory to a form of ordred
+  transformations. By restricting the type theory to a form of ordered
   logic, we guarantee that all mappings between categories are
   functorial and all transformations are natural by construction, with
-  no separate proof necessary. We demonstrate the utility of the type
-  theory by proving a form of the Yoneda lemma which reads as the
-  classical proof, but with all naturality side-conditions
-  automatically satisfied.
-
-  We show that the type theory is sound and complete for virtual
-  equipments, a kind of double category which is known to model not
-  just ordinary categories but enriched and internal categories as
-  well. This means that proofs and constructions in our theory hold
-  not just for ordinary categories but also for a broad class of
-  generalized categories. We demonstrate that the type theory
-  simplifies working with virtual equipments compared to their
-  algebraic description in the literature.
+  no separate proof necessary. Important category theoretic proofs
+  such as the Yoneda lemma and Co-yoneda lemma become simple type
+  theoretic proofs about the relationship between unit, tensor and
+  (ordered) function types.
+
+  We show that the type theory is sound for models called
+  \emph{virtual equipments}, a kind of double category which is known
+  to model not just ordinary categories but enriched, internal and
+  generalized multi-categories as well. This means that while proofs
+  and constructions in our theory read like typical category-theoretic
+  constructions, the proofs automatically apply to these generalized
+  categories. Our type theory simplifies arguments over the algebraic
+  description of virtual equipments by modeling the algebraic
+  operation of restriction of profunctors along functors as instead
+  the strict syntactic substitution.
 \end{abstract}

### 2022-02-11 paper/lipics-v2021-sample-article.tex
intro update

@@
 
-Experienced category theorists have no trouble with this, as they are
-familiar with a great many principles for constructing transformations
-that are guaranteed to be natural. Consider an analogy with
-calculus. When a mathematician writes a formula such as
-
-$$\sin(f(x) - g(x))$$
-
-They do not need to spend time with an explicit $\epsilon,\delta$
-proof that the function is continuous or furthermore smooth. Instead,
-they follow syntactic principles that compositions of continuous
-functions are continuous.
-
-There are similar syntactic notions for guaranteeing that
-transformations are natural.
-
-\subsection{Profunctors}
+More experienced users of category theory in programming language
+theory find that for modeling computational effects such as general
+recursion or indexed-state, or modeling logics for parametricity,
+gradual typing, or step-indexed logical relations involves not just
+ordinary category theory, but more general notions such as enriched,
+internal or multi-categories. However, model construction theorems and
+especially soundness and completeness theorems in these areas will
+require their own, even more complex versions of the same common
+lemmas of category theory, in addition to the domain-specific results
+such as existence of fixed points.
+
+At this level of generality, the proofs can be difficult to
+understand, as they often involve some combination of the original
+category theoretic proof with some additional side-conditions that
+everythinng is in fact appropriately step-indexed, continuous etc.
+%
+The solution to taming this complexity is to prove these general
+theorems at a higher level of generality, for instance, proving them
+in some abstract 2-category of enriched/internal categories, functors
+and transformations.
+%
+For proofs in generalized category theory specifically, the notion of
+a \emph{virtual equipment} has been shown to be occupy an algebraic
+sweet-spot of being complex enough to express most basic constructions
+(functors, profunctors, natural transformations, adjunctions, kan
+extensions, etc) but also general enough to encompass internal,
+enriched and ``generalized multi-categories'', which include ordinary
+flavors of multi-categories like the cartesian multi-categories that
+model sequents in natural deduction, but even structures such as
+topological spaces.
+%
+However, there are two difficulties with this approach.
+%
+First, the abstract arguments can be difficult to understand, and can
+seem quite hard to connect to the original argument for ordinary
+category theory.
+%
+Second, the algebraic structure of a virtual equipment is quite
+unwieldy to manipulate. The equivalent of a 2-cell is a kind of
+``multi-2-cell'' with many inputs and a single output, in
+diagrammatic form these are rendered as
+
+\[ % https://tikzcd.yichuanshen.de/#N4Igdg9gJgpgziAXAbVABwnAlgFyxMJZABgBpiBdUkANwEMAbAVxiRAGEB9YkAX1PSZc+QigCM5KrUYs2XMXwEgM2PASIAmSdXrNWiDpw2LBqkUQDM26XrYAdOwygQcCfqeHqUAFmu7ZBlzAYADUYrwmykJqoiSkYlL++iCRKp6xvgk6Msl8UjBQAObwRKAAZgBOEAC2SBIgOBBIWjYBIABKnAruIJU1zdSNSFatyZ3GPX21iCNDiL6jbJ2Ek1XTZA1NiPVJS9yRU0gbcwCs2bYGAGIHa0hnm0gAbOdtAMo3-fODW8+LBgDieV4QA
+\begin{tikzcd}
+C_0 \arrow[r, "R_0"] \arrow[d, "F"] & C_1 \arrow[r, "R_1"] & C_2 \arrow[r, "R_2"] & \ldots \arrow[r, "R_n"] & C_{n+1} \arrow[d, "G"] \\
+{} \arrow[rrrr, "S"]                &                      &                      &                         & {}                    
+\end{tikzcd} \]
+
+This 2-cell has one output pro-arrow, many input pro-arrows which must
+match the source from one to the target object of the next and a left
+and right ``frame'' arrow. These can be composed in a manner similar
+to multi-categories, where you can compose this with many input
+2-cells, one for each input pro-arrow.
+%
+These then have complex associativity and unit rules and side
+conditions that say the framing arrows compose alongside the 2-cells.
+
+We solve both of these issues by developing an alternative \emph{type
+theoretic syntax} for the constructions in a virtual equipment.
+%
+In this syntax the ``multi-2-cell'' above instead becomes a natural
+transformation judgment
+\[ \Phi \vdash t : R[F(x);G(y)] \]
+We write natural transformations in a natural deduction style: a
+transformation is written as a ``parameterized'' element of the output
+profunctor. Furthermore the restriction operators for vertical arrows
+into horizontal arrows is naturally modeled by a syntactic
+substitution.
+%
+This interpretation of restriction as substitution both matches the
+semantic intuition for the restriction operation as well as greatly
+simplifies reasoning about composition of transformations.
+
+%% Experienced category theorists have no trouble with this, as they are
+%% familiar with a great many principles for constructing transformations
+%% that are guaranteed to be natural. Consider an analogy with
+%% calculus. When a mathematician writes a formula such as
+
+%% $$\sin(f(x) - g(x))$$
+
+%% They do not need to spend time with an explicit $\epsilon,\delta$
+%% proof that the function is continuous or furthermore smooth. Instead,
+%% they follow syntactic principles that compositions of continuous
+%% functions are continuous.
+
+%% There are similar syntactic notions for guaranteeing that
+%% transformations are natural.
+
+\subsection{Category Theory with Profunctors}
 
@@
 spaces, metric spaces, etc.
+
+\subsection{Paper Overview}
+
+To demonstrate the utility of SCT, we proceed as follows
+\begin{enumerate}
+\item In \cref{sec:syntax} we introduce the syntax of Simple Category
+  Theory.
+\item In \cref{sec:synthetic} we show how common theorems of category
+  theory such as the (co)-Yoneda lemma, equivalence of different
+  formulations of adjunctions, and preservation of limits can all be
+  cleanly reproduced in SCT.
+\item In \cref{sec:internal}, we show how we can use SCT to define
+  \emph{internal categories}, which allows us to prove additional
+  theorems that are not formalizable as synthetic theorems.
+\item In \cref{sec:semantics}, we show that the syntax of SCT is sound
+  for models in an arbitrary virtual double category with
+  corresponding universal constructions.
+\item In \cref{sec:discussion}, we discuss related type theories and
+  potential extensions to model multi-variable and higher-order
+  categorical constructions.
+\end{enumerate}
 
@@
 
+\begin{definition}[Psuedo Double Category]
+
+\end{definition}
+
+\begin{definition}[Restriction]
+
+\end{definition}
+
+\begin{definition}[Unit]
+
+\end{definition}
+
 %% \section{Conservativity of Homs, Tensors}

### 2022-02-13 paper/lipics-v2021-sample-article.tex
some semantics

@@
 
-In this section we establish our desired semantic soundness theorem,
-that all constructions in the theory can be interpreted in any virtual
-equipment with the corresponding universal properties.
-%
-This proof is straightforward, since our constructions are explicitly
-modeled on the universal properties, with the main difference being
-that our notion of restriction is strict.
-
-\begin{definition}[Psuedo Double Category]
-
+In this section we establish our desired semantic soundness and
+completeness theorems for models in virtual equipments with the
+corresponding universal properties.
+%
+The first component of this is \emph{soundness}: every category,
+object, set and transformation denote an object, vertical arrow,
+horizontal arrow and 2-cell, respectively in any representable virtual
+equipment, such that all equations in the theory are satisfied.
+%
+We combine this with a set theoretic semantics for types $A$ and
+contexts $\Gamma$ to get a simple semantics of our logic over a
+virtual equipment.
+%
+The second component of this is the \emph{completeness} theorem:
+informally, the theory contains every category, object, set and
+transformation that \emph{must} exist in any virtual equipment, and
+further equations between syntactic forms hold if and only if they
+hold in every model.
+%
+While sounding more ambitious, the completeness theorem is
+typically simpler as it just amounts to showing that the syntax itself
+forms a virtual equipment.
+
+\subsection{Notion of Model}
+
+\begin{definition}[Virtual Double Category]
+  A virtual double category $V$ consists of
+  \begin{enumerate}
+    \item A set $V_0$ of objects.
+    \item For each $\cat C, \cat D \in V_0$, a set $V_v(\cat C, \cat
+      D)$ of vertical arrows, with composition and unit operations
+      satisfying the category axioms.
+    \item A set $V_h$ of horizontal arrows, with $d^+,d^- : V_h \to
+      V_0$.
+    \item For each composable string $\Phi$ of horizontal arrows
+      starting at $\cat C$ and ending at $\cat D$, and horizontal
+      arrow $R$, a set $V_2(\Phi; R)$ of 2-cells, with projection
+      functions $d^{\pm} : V_2(\Phi;R) \to V_v(d^\pm \Phi, d^\pm R)$.
+
+    \item For each $R \in V_h$, an identity arrow $\id_R \in
+      V_2(d^-R,R,d^+ R; R)$ with $d^\pm \id_R = \id_{d^\pm R}$.
+
+    \item For any $2$-cell $\alpha \in V_2(\Phi;R)$ and composable
+      string $\phi$ of 2-cells into $\Phi$, a composite $\alpha \circ
+      \phi$.
+
+    \item Composition should be unital in that:
+      \[ \id_R \circ [d^-\alpha, \alpha, d^+\alpha] = \alpha  \]
+      and
+      \[ \alpha \circ [\vec \id] = \alpha\]
+    \item And associative in that
+      \[ (\alpha \circ [\beta_1,\ldots]) \circ [\gamma_{11},\ldots,\gamma_{21},\ldots,] = \alpha \circ [(\beta_1 \circ \gamma_{11},\ldots), (\beta_2 \circ \gamma_{21},\ldots)]\]
+  \end{enumerate}
 \end{definition}
@@
 \begin{definition}[Restriction]
-
+  Let $R \in V_h$ and $f \in V_v(\cat C, d^-R)$ and $g \in V_v(\cat D,
+  d^+R)$, then a restriction of $R$ along $f,g$ is a horizontal arrow
+  $R(f,g) \in V_h$ with $d^-R(f,g)= \cat C$ and $d^+R(f,g)= \cat D$
+  with a (op?)-cartesian 2-cell $\alpha \in V_2(R(f,g),R)$ with
+  $d^-\alpha = f$ and $d^+\alpha = g$.
 \end{definition}
@@
 \begin{definition}[Unit]
-
+  A unit $U_{\cat C}\in V_h$ with $d^\pm U_{\cat C} = \cat C$ and an op-cartesian 2-cell
+  \[ r : V_h(\cat C, U_{\cat C}) \]
+  with $d^\pm r = \id_{\cat C}$.
+\end{definition}
+
+\begin{definition}[Tensor]
+  If $R,Q \in V_h$ with $d^+R = d^-Q$, a tensor is a horizontal arrow
+  $R \odot Q \in V_h$ with $d^+ (R\odot Q) = d^+ R$ and $d^-(R \odot
+  Q) = d^- Q$ with an op-cartesian 2-cell
+  \[ p : V_h(R,Q;R \odot Q) \]
+\end{definition}
+
+\begin{definition}[Hom]
+
+\end{definition}
+
+\subsection{Soundness}
+
+\subsection{Completeness}
+
+\begin{definition}
+  For any $\Gamma \isadtctx$, we construct a virtual double category
+  $V^\Gamma$.
+  \begin{enumerate}
+  \item $V^\Gamma_v$ is the set of terms $\Gamma \vdash \cat C \isaCat$.
+  \item $V^\Gamma_v(\cat C, \cat D)$ is the set of objects $\Gamma
+    \pipe \alpha:\cat C \vdash a : \cat D$.
+  \item $V_h$ is the set of triples $(\cat C, \cat D, \Gamma \pipe \alpha:\cat C;\beta:\cat D \vdash R \isaSet)$
+    where $d^\pm$ are the projections
+  \item $V^\Gamma_h(\vec R; S)$ is the set of triples $(a,b,\Gamma
+    \pipe \Phi \vdash t : S[a;b])$ where $\Phi$ is the context
+    corresponding to $\vec R$ and $d^\pm$ are the obvious projections.
+
+  \item The identity morphisms are $(\cat C, \cat D, R)$ is $\Gamma\pipe\alpha:\cat C, x:R\beta:\cat D \vdash x:R$
+  \item Composition $(a,b,t) \circ [(a_1,),\ldots]$ is given by
+    substitution in each component: $(a[a_1],b[a_n], t[\phi])$.
+  \item associativity and unit rules for composition follow by the
+    definition of substitution.
+  \end{enumerate}
 \end{definition}

### 2022-02-15 paper/lipics-v2021-sample-article.tex
start the interpretation part

@@
 
-\subsection{Soundness}
+\subsection{Category of Signatures and Models}
+
+Our goal is to establish that the syntax of SCT (with whichever type
+constructors we choose) presents a free virtual double category (with
+the appropriate universal properties). To start we define appropriate
+categories of signatures and double categories.
+
+\begin{definition}
+  Define the category Sig to have as objects well-formed contexts
+  $\Gamma$ and morphisms $\gamma : \Delta \to \Gamma$ morphisms of
+  contexts, with the composition as substitution and identity as a
+  context morphism of all variable references.
+\end{definition}
+
+Second, there is a category of virtual double categories and functors.
+
+\begin{definition}
+  A morphism of virtual equipments with blah is a functor of virtual
+  double categories that strictly preserves the universal properties.
+\end{definition}
+
+Next, we define a profunctor between the category of signatures and
+the category of virtual equipments.
+
+\begin{definition}
+  Given a well-formed signature $\Gamma$ and a virtual equipment
+  $\mathcal V$, we mutually define the following notions:
+
+  \begin{itemize}
+  \item The set $\sem{\Gamma}_{\mathcal V}$
+  \item For any $\Gamma \vdash A$ and $\gamma \in \sem{\Gamma}_{\mathcal V}$, a set $\sem{A}_{\mathcal V} \gamma$
+  \item For any $\Gamma \vdash M : A$ and $\gamma \in \sem{\Gamma}_{\mathcal V}$, an element $\sem{M}_{\mathcal V} \in \sem{A}_{\mathcal V} \gamma$
+  \item categories...
+  \item functors...
+  \item profunctors...
+  \item trans contexts,
+  \item transformations
+  \end{itemize}
+
+  An interpretation of a signature $\Gamma$ in a virtual equipment
+  $\mathcal V$, $\textrm{Interp}(\Gamma,\mathcal V)$ is an element of
+  the set $\sem{\Gamma}_{\mathcal V}$ where we mutually define the set
+
+  \begin{align*}
+    \sem{\cdot}_{\mathcal V} &= 1\\
+    \sem{\Gamma, x:A}_{\mathcal V} &= \sum_{\gamma \in \sem{\Gamma}_{\mathcal V}} \sem{A}_{\mathcal V}\gamma\\
+  \end{align*}
+
+  with an 
+  \begin{align*}
+    \sem{\Gamma, x:A}_{\mathcal V} &= \sem{\Gamma}_{\mathcal V} \times \sem{A}_{\mathcal V}\\
+    \sem{\Gamma}_{\mathcal V} &= 1\\\\
+    \sem{\Sort}_{\mathcal V} &= \mathcal V_0\\
+    \sem{\varr{\cat C}{\cat D}}_{\mathcal V} &= \mathcal V_v(\sem{\cat C})
+  \end{align*}
+
+\end{definition}
 
@@
   \item $V^\Gamma_v$ is the set of terms $\Gamma \vdash \cat C \isaCat$.
-  \item $V^\Gamma_v(\cat C, \cat D)$ is the set of objects $\Gamma
-    \pipe \alpha:\cat C \vdash a : \cat D$.
+  \item $V^\Gamma_v(\cat C, \cat D)$ is the set of ``objects'' $\Gamma
+    \pipe \alpha:\cat C \vdash a : \cat D$. The category structure is
+    given by substitution, with the variable $\alpha \vdash \alpha$
+    the identity.
   \item $V_h$ is the set of triples $(\cat C, \cat D, \Gamma \pipe \alpha:\cat C;\beta:\cat D \vdash R \isaSet)$
@@
     substitution in each component: $(a[a_1],b[a_n], t[\phi])$.
-  \item associativity and unit rules for composition follow by the
-    definition of substitution.
+  \item Associativity and unit rules for composition follow from the
+    definition of substitution by a straightforward induction.
   \end{enumerate}
 \end{definition}
+
+\subsection{Soundness}
+
+Let $$
 

### 2022-02-16 paper/lipics-v2021-sample-article.tex
get stuck...

@@
 %% \newtheorem{definition}{Definition}
+\newcommand{\sem}[1]{\llbracket{#1}\rrbracket}
 \newcommand{\pto}{\nrightarrow}
@@
   \item For any $\Gamma \vdash M : A$ and $\gamma \in \sem{\Gamma}_{\mathcal V}$, an element $\sem{M}_{\mathcal V} \in \sem{A}_{\mathcal V} \gamma$
-  \item categories...
-  \item functors...
-  \item profunctors...
-  \item trans contexts,
-  \item transformations
+  \item For any $\Gamma \vdash \cat C \isaCat$ and $\gamma \in \sem{\Gamma}_{\mathcal V}$, an object $\sem{\cat C}_{\mathcal V}\gamma$
+  \item For any $\Gamma\pipe\alpha:\cat C \vdash a :\cat D$ and $\gamma \in \sem{\Gamma}_{\mathcal V}$, a vertical morphism $\sem{\alpha.a}_{\mathcal V}\gamma \in \mathcal V_v(\sem{\cat C}_{\mathcal V}\gamma, \sem{\cat D}_{\mathcal V}\gamma,)$
+  \item For any $\Gamma\pipe\Xi \vdash R \isaSet$ and $\gamma \in
+    \sem{\Gamma}_{\mathcal V}$, a horizontal arrow $\sem{R}_{\mathcal
+      V}\gamma \in \mathcal V_h$ such that $d^\pm\sem{R}_{\mathcal
+      V}\gamma = \sem{d^\pm\Xi}_{\mathcal V}\gamma$. We also extend
+    this in the obvious way to transformation contexts
+    $\sem{\Phi}_{\mathcal V}\gamma$
+  \item For any $\Gamma\pipe\Phi \vdash t : R$ and $\gamma \in
+    \sem{\Gamma}_{\mathcal V}$, a 2-cell $\sem{t}_{\mathcal V}\gamma
+    \in V_h(\sem{\Phi}_{\mathcal V}\gamma;\sem{R}_{\mathcal V}\gamma)$
+    with $d^\pm\sem{t}_{\mathcal V}\gamma = \id$.
   \end{itemize}
 
+  %% \begin{align*}
+  %%   \sem{\cdot}_{\mathcal V} &= 1\\
+  %%   \sem{\Gamma, x:A}_{\mathcal V} &= \sum_{\gamma \in \sem{\Gamma}_{\mathcal V}} \sem{A}_{\mathcal V}\gamma\\
+  %% \end{align*}
+
+  %% with an 
+  %% \begin{align*}
+  %%   \sem{\Gamma, x:A}_{\mathcal V} &= \sem{\Gamma}_{\mathcal V} \times \sem{A}_{\mathcal V}\\
+  %%   \sem{\Gamma}_{\mathcal V} &= 1\\\\
+  %%   \sem{\Sort}_{\mathcal V} &= \mathcal V_0\\
+  %%   \sem{\varr{\cat C}{\cat D}}_{\mathcal V} &= \mathcal V_v(\sem{\cat C})
+  %% \end{align*}
+\end{definition}
+
+And we define a profunctor of interpretations of a signature in a
+virtual equipment as follows:
+\begin{definition}
   An interpretation of a signature $\Gamma$ in a virtual equipment
-  $\mathcal V$, $\textrm{Interp}(\Gamma,\mathcal V)$ is an element of
-  the set $\sem{\Gamma}_{\mathcal V}$ where we mutually define the set
-
-  \begin{align*}
-    \sem{\cdot}_{\mathcal V} &= 1\\
-    \sem{\Gamma, x:A}_{\mathcal V} &= \sum_{\gamma \in \sem{\Gamma}_{\mathcal V}} \sem{A}_{\mathcal V}\gamma\\
-  \end{align*}
-
-  with an 
-  \begin{align*}
-    \sem{\Gamma, x:A}_{\mathcal V} &= \sem{\Gamma}_{\mathcal V} \times \sem{A}_{\mathcal V}\\
-    \sem{\Gamma}_{\mathcal V} &= 1\\\\
-    \sem{\Sort}_{\mathcal V} &= \mathcal V_0\\
-    \sem{\varr{\cat C}{\cat D}}_{\mathcal V} &= \mathcal V_v(\sem{\cat C})
-  \end{align*}
-
+  $\mathcal V$ is an element of the set $\sem{\Gamma}_{\mathcal
+    V}$. Call this set $\textrm{Interp}(\Gamma;\mathcal V)$.
+
+  Note that this constitutes a profunctor contravariant in $\Gamma$
+  and covariant in $\mathcal V$.
 \end{definition}
 
-\subsection{Completeness}
-
+Next, we define the syntactic virtual equipment $\textrm{Syn}(\Gamma)$
+for a signature $\Gamma$.
 \begin{definition}
@@
   \item $V^\Gamma_v$ is the set of terms $\Gamma \vdash \cat C \isaCat$.
-  \item $V^\Gamma_v(\cat C, \cat D)$ is the set of ``objects'' $\Gamma
-    \pipe \alpha:\cat C \vdash a : \cat D$. The category structure is
-    given by substitution, with the variable $\alpha \vdash \alpha$
-    the identity.
-  \item $V_h$ is the set of triples $(\cat C, \cat D, \Gamma \pipe \alpha:\cat C;\beta:\cat D \vdash R \isaSet)$
-    where $d^\pm$ are the projections
+  \item $V^\Gamma_v(\cat C, \cat D)$ is the set of objects in context
+    $\Gamma \pipe \alpha:\cat C \vdash a : \cat D$. The category
+    structure is given by substitution, with the variable $\alpha
+    \vdash \alpha$ the identity.
+  \item $V_h$ is the set of triples $(\cat C, \cat D, \Gamma \pipe
+    \alpha:\cat C;\beta:\cat D \vdash R \isaSet)$ where $d^\pm$ are
+    the projections
   \item $V^\Gamma_h(\vec R; S)$ is the set of triples $(a,b,\Gamma
@@
 
-\subsection{Soundness}
-
-Let $$
+Then the soundness theorem says that for any $\Gamma$, the virtual
+double category $\textrm{Syn}(\Gamma)$ provides an isomorphism of profunctors:
+\[ \textrm{Interp}(\Gamma, \mathcal V \cong VE(\textrm{Syn}(\Gamma), \mathcal V) \]
+As we saw in a previous section, the backwards direction of this
+isomorphism should be a universal interpretation:
+
+\begin{definition}[Universal Interpretation]
+  For any signature $\Gamma$ there is a ``self-interpretation'' $\eta:
+  \textrm{Interp}(\Gamma, \textrm{Syn}(\Gamma))$ defined as the
+  identity substitution $\id \in \sem{\Gamma}_{\textrm{Syn}(\Gamma)}$
+
+\end{definition}
 

### 2022-02-27 paper/lipics-v2021-sample-article.tex
A missing connective?!? This should allow us to define slice/comma cats and potentially fibrations!

@@
 \[ \int_{\beta} P(-,\beta) \times Q(\beta,=) \]
+
+\subsection{Graph of a Profunctor}
+
+One final multiplicative we can define is the \emph{graph} of a
+profunctor. This is unique among the multiplicatives in that it
+defines a \emph{category} rather than a profunctor. It is a negative
+multiplicative but it looks like a positive ??. A functor into the
+category of elements $\exists (\alpha; \beta). R$ is equivalent to a
+morphism into the functor $R$.
+
+\begin{figure}
+  \begin{mathpar}
+    \inferrule*[right=Graph-Formation]
+    {\Gamma \vdash \alpha:\cat C; \beta:\cat D \vdash R \isaSet}
+    {\Gamma \vdash \exists (\alpha;\beta). R \isaCat}
+
+    \inferrule*[right=Graph-E+]
+    {\Gamma \vdash \gamma:\cat E \vdash \phi : \exists(\alpha:\cat C;\beta: \cat D).R}
+    {\Gamma \vdash \pi_+\phi : \cat C}
+
+    \inferrule*[right=Graph-E-]
+    {\Gamma \vdash \gamma:\cat E \vdash \phi : \exists(\alpha:\cat C;\beta: \cat D).R}
+    {\Gamma \vdash \pi_-\phi : \cat D}
+
+    \inferrule*[right=Graph-Ep]
+    {\Gamma \vdash \gamma:\cat E \vdash \phi : \exists(\alpha:\cat C;\beta: \cat D).R}
+    {\Gamma \vdash \gamma:\cat E \vdash \pi_p\phi : R[\pi_+\phi/\alpha,\pi_-\phi/\beta]}
+
+    \inferrule*[right=Graph-Intro]
+    {\Gamma\vdash \gamma:\cat E \vdash a : \cat C \and
+      \Gamma\vdash \gamma:\cat E \vdash b : \cat D \and
+      \Gamma\vdash \gamma:\cat E \vdash t : R[a/\alpha;b/\beta]}
+    {\Gamma\vdash \gamma:\cat E \vdash (a,b,t) : \exists (\alpha;\beta). R}
+
+    \inferrule*[right=Unit-Formation]
+    {\Gamma \vdash \cat C \isaCat\and
+     \Gamma \pipe d^-\Xi \vdash a_1 : \cat C\and \Gamma \pipe d^+ \Xi \vdash a_2 : \cat C}
+    {\Gamma \pipe \Xi \vdash \punitinXfromYtoZ {\cat C} {a_1} {a_2}}
+
+    \pi_+(a,b,t) = a \\
+    \pi_-(a,b,t) = b \\
+    \pi_p(a,b,t) = t\\
+
+    \inferrule
+    {\Gamma \vdash \gamma:\cat E \vdash \phi : \exists(\alpha:\cat C;\beta:\cat D). R}
+    {\phi = (\pi_+\phi, \pi_-\phi, \pi_p\phi)}
+  \end{mathpar}  
+  \caption{Graph of a Profunctor}
+\end{figure}
+
+Semantically, the graph of a profunctor is simply the category of
+elements of the profunctor. Taking the profunctor to be (restrictions
+of) the hom profunctor allows us to define the slice, arrow and comma
+categories.
 
@@
 
+\paragraph{As a tool for Formal Type Theory}
+
 %% \section{Typesetting instructions -- Summary}

### 2022-02-28 paper/lipics-v2021-sample-article.tex
fibrations??

@@
 One final multiplicative we can define is the \emph{graph} of a
-profunctor. This is unique among the multiplicatives in that it
-defines a \emph{category} rather than a profunctor. It is a negative
-multiplicative but it looks like a positive ??. A functor into the
-category of elements $\exists (\alpha; \beta). R$ is equivalent to a
-morphism into the functor $R$.
+profunctor. Given a profunctor $\alpha:\cat C;\beta:\cat D \vdash R$,
+the ``graph'' of $R$ is a category $\exists \alpha;\beta. R$. In
+ordinary category theory, this is usually defined as the category of
+elements of $R$ viewed as a functor $R : \cat C^o \times D \to \Set$.
+In detail, its objects are triples $(a:\cat C, r : R, b : \cat D)$ and
+morphism from $(a,r,b)$ to $(a', r', b')$ consists of a pair of
+morphisms $g : b \to b'$ and a backwards $f : a' \to a$ such that $r'
+= g\circ r \circ f$.
+%
+By picking $R = \punitinXfromYtoZ {\cat C} {\alpha}{\alpha'}$ we
+recover the \emph{arrow} category $\exists
+(\alpha;\alpha'). \punitinXfromYtoZ {\cat C}{\alpha}{\alpha'}$, by
+composing these with functors, we recover the notion of \emph{comma}
+category $\exists \alpha;\beta. \punitinXfromYtoZ {\cat
+  E}{(F\alpha)}{(G\beta)}$ and by picking either $F$ or $G$ to have
+domain the terminal category $1$, we recover the notion of a
+\emph{slice} category.
+
+This is unique among the multiplicatives in that it defines a
+\emph{category} rather than a profunctor. Also note that while similar
+to the tensor in that we use existential notation, the graph of a
+profunctor is a \emph{negative} multiplicative, analogous to a
+$\Sigma$-type. The universal property says that a functor from $\cat
+E$ into the graph $\exists (\alpha; \beta). R$ is equivalent to a
+transformation from $\cat E$ into the functor $R$, including functors
+from $\cat E$ to $\cat C$ and $\cat D$.
 
@@
 \end{figure}
-
-Semantically, the graph of a profunctor is simply the category of
-elements of the profunctor. Taking the profunctor to be (restrictions
-of) the hom profunctor allows us to define the slice, arrow and comma
-categories.
 
@@
 
-\subsection{Adjoint Functors}
+\subsection{Adjoint Functors and Representable Profunctors}
 
@@
 \begin{lemma}[(Co)-unit of a Representation]
-  If $G: \varr {\cat D}{\cat C}$ right-represents $R : \harr {\cat C}{\cat D}$, then there is a natural element
-
+  If $G: \varr {\cat D}{\cat C}$ right-represents $R : \harr {\cat
+    C}{\cat D}$, then there is a natural element, the co-unit:
   \[ \pendallXdotY \beta \harrapp R {G\beta} \beta \]
-
   Such that the half of the isomorphism 
-
   \[ \punitinXfromYtoZ {\cat C} {\alpha} {G \beta} \multimap \harrapp R {\alpha}{\beta} \]
-
   is defined by composition with the natural element.
@@
   Dually, any $F$ left-representing isomorphism with $R$ is determined
-  by a term of type $\pendallXdotY \alpha \harrapp R {\alpha}
+  by a ``unit'' of type $\pendallXdotY \alpha \harrapp R {\alpha}
   {F\alpha}$.
@@
   \end{align*}
+\end{proof}
+
+\subsection{(op)-Fibrations}
+
+Using the graph of a profunctor, we can give the following definition
+for when a functor is a fibration. This gives a very weak notion of
+fibration that is probably equivalent to a Street fibration (todo: is
+this an existing definition???).
+
+\begin{definition}{Street Fibration}
+  Let $F : \varr{\cat C}{\cat D}$, we say $F$ is a \emph{fibration}
+  when the induced functor
+  \[p : \exists {\alpha:\cat C;\alpha':\cat C} \punitinXfromYtoZ {\cat C}{\alpha}{\alpha'} \vdash (F\pi_-p,F_1\pi_pp, \pi_+p) : \exists {\beta:\cat D;\alpha':\cat C}.
+\punitinXfromYtoZ {\cat D} \beta {F \alpha'} \]
+  has a right adjoint inverse.
+
+  Dually, $F$ is an \emph{op-fibration} when the same induced functor
+  has a left adjoint inverse.
+\end{definition}
+
+We can get a strict notion of fibration if we allow ourselves to be a little ``evil'':
+
+\begin{definition}{Grothendieck Fibration}
+  Let $F : \varr{\cat C}{\cat D}$, we say $F$ is a \emph{strict
+  fibration} when the induced functor $F'$ has a strict right adjoint
+  inverse $R$, i.e. inverse $\lambda p. R(F'(p)) = \lambda p. p$. We
+  analogously define a strict op-fibration.  
+\end{definition}
+
+Alternatively, we can give a more traditional definition using slice
+categories, but the downside is the formulation is inevitably higher
+order.
+\begin{definition}{Street Fibration}
+  Let $F : \varr{\cat C}{\cat D}$, we say $F$ is a \emph{fibration}
+  when for every $a : \varr{1}{\cat C}$, the induced functor
+  \[p : \exists {\alpha:\cat C; o:\cat 1} \punitinXfromYtoZ {\cat C}{\alpha}{a()} \vdash (F\pi_-p,F_1\pi_pp, \pi_+p) : \exists {\beta:\cat D; o:\cat 1}. \punitinXfromYtoZ {\cat D} \beta {F(a())} \]
+  has a right adjoint inverse.
+
+  Dually, $F$ is an \emph{op-fibration} when the same induced functor
+  has a left adjoint inverse.
+\end{definition}
+
+\begin{lemma}{Fibrations compose}
+  If $F : \varr{\cat C}{\cat D}$ and $G : \varr{\cat D}{\cat E}$ are
+  (op)-fibrations then so is $G \circ F$.
+\end{lemma}
+\begin{proof}
+  By composition of adjoints.
 \end{proof}

### 2022-03-15 paper/paper.tex
fix the fibrations

--- 2022-02-28 paper/lipics-v2021-sample-article.tex
@@
 
-Using the graph of a profunctor, we can give the following definition
-for when a functor is a fibration. This gives a very weak notion of
-fibration that is probably equivalent to a Street fibration (todo: is
-this an existing definition???).
-
-\begin{definition}{Street Fibration}
-  Let $F : \varr{\cat C}{\cat D}$, we say $F$ is a \emph{fibration}
-  when the induced functor
-  \[p : \exists {\alpha:\cat C;\alpha':\cat C} \punitinXfromYtoZ {\cat C}{\alpha}{\alpha'} \vdash (F\pi_-p,F_1\pi_pp, \pi_+p) : \exists {\beta:\cat D;\alpha':\cat C}.
-\punitinXfromYtoZ {\cat D} \beta {F \alpha'} \]
-  has a right adjoint inverse.
-
-  Dually, $F$ is an \emph{op-fibration} when the same induced functor
-  has a left adjoint inverse.
-\end{definition}
-
-We can get a strict notion of fibration if we allow ourselves to be a little ``evil'':
-
-\begin{definition}{Grothendieck Fibration}
-  Let $F : \varr{\cat C}{\cat D}$, we say $F$ is a \emph{strict
-  fibration} when the induced functor $F'$ has a strict right adjoint
-  inverse $R$, i.e. inverse $\lambda p. R(F'(p)) = \lambda p. p$. We
-  analogously define a strict op-fibration.  
-\end{definition}
-
-Alternatively, we can give a more traditional definition using slice
-categories, but the downside is the formulation is inevitably higher
-order.
-\begin{definition}{Street Fibration}
-  Let $F : \varr{\cat C}{\cat D}$, we say $F$ is a \emph{fibration}
-  when for every $a : \varr{1}{\cat C}$, the induced functor
-  \[p : \exists {\alpha:\cat C; o:\cat 1} \punitinXfromYtoZ {\cat C}{\alpha}{a()} \vdash (F\pi_-p,F_1\pi_pp, \pi_+p) : \exists {\beta:\cat D; o:\cat 1}. \punitinXfromYtoZ {\cat D} \beta {F(a())} \]
-  has a right adjoint inverse.
-
-  Dually, $F$ is an \emph{op-fibration} when the same induced functor
-  has a left adjoint inverse.
-\end{definition}
+A (Street) fibration can be defined in any 2-category. A functor $p :
+\varr {\cat E}{\cat B}$ is a fibration when the morphism $p \to \cat
+B/p$ in the slice 2-category over $\cat B$ has a right adjoint.
+
+Since we can define all of the components of this slice 2-category we
+can internalize this definition in SCT.
+
+\begin{enumerate}
+\item A functor $\textrm{wp} : \exists (\alpha:\cat B; \beta:\cat E). \punitinXfromYtoZ {\cat B} \alpha {p\beta}$
+\item In the slice category: $p(\textrm{wp}(k)) = \pi_-k$
+\item With a universal morphism $\epsilon : \pforallXfromYtoZ {k} \punitinXfromYtoZ {} {(\pi_-k,k,\punitrefl{}{\pi_-k})} {k}$
+\item In the slice category: $\pi_k \epsilon = \id$ (how to express this nicely in SCT?)
+\end{enumerate}
+
+We can also weaken this to a Street fibration if we require the more sensible $p(\textrm{wp}(k)) \cong \pi_-k$ only.
 
@@
 \end{proof}
+
+\begin{lemma}{Fibrations are Stable under Pullback}
+
+\end{lemma}
 

### 2022-04-11 paper/paper.tex
comment out the internal category theory stuff

@@
 
-  We show that the type theory is sound for models called
-  \emph{virtual equipments}, a kind of double category which is known
-  to model not just ordinary categories but enriched, internal and
-  generalized multi-categories as well. This means that while proofs
-  and constructions in our theory read like typical category-theoretic
+  We show that the type theory is sound for models in \emph{virtual
+  equipments}, a kind of double category which is known to model not
+  just ordinary categories but enriched, internal and generalized
+  multi-categories as well. This means that while proofs and
+  constructions in our theory read like typical category-theoretic
   constructions, the proofs automatically apply to these generalized
@@
 %
-The solution to taming this complexity is to prove these general
+One approach to taming this complexity is to prove these general
 theorems at a higher level of generality, for instance, proving them
@@
 enriched and ``generalized multi-categories'', which include ordinary
-flavors of multi-categories like the cartesian multi-categories that
-model sequents in natural deduction, but even structures such as
-topological spaces.
+flavors of multi-categories that model intuitionistic and
+substructural type theories, but also a great deal more, including
+double categories, metric spaces and topological spaces.
 %
@@
 %
-First, the abstract arguments can be difficult to understand, and can
-seem quite hard to connect to the original argument for ordinary
-category theory.
+First, the abstract arguments can be difficult to understand compapred
+to the original argument for ordinary category theory.
 %
 Second, the algebraic structure of a virtual equipment is quite
-unwieldy to manipulate. The equivalent of a 2-cell is a kind of
-``multi-2-cell'' with many inputs and a single output, in
-diagrammatic form these are rendered as
+unwieldy to manipulate. The abstraction of a generalized natural
+transformation is a kind of ``multi-2-cell'' with many inputs and a
+single output. In diagrammatic form these are rendered as
 
@@
 transformation judgment
-\[ \Phi \vdash t : R[F(x);G(y)] \]
+\[ \Phi \vdash t : S[F(x);G(y)] \]
+where the composable string of input arrows is modeled by an ordered,
+linear context $\Phi$ and output $S$ and frames $F,G$ are combined in
+the output type.
+%
 We write natural transformations in a natural deduction style: a
@@
 spaces, metric spaces, etc.
+
+\subsection{Applications to Generalized Type Theory}
+
+While we mostly focus on the semantics in category theory, due to the
+tight connection between category theory and the semantics of type
+theory, there are direct applications to the study of ``abstract type
+theory'' as well. We demonstrate how we can model single-variable
+simple type theories with products and coproducts within SCT. This is
+very much in the spirit of logical frameworks \cite{TODO}, since we
+can faithfully model the terms of an object theory inside of our
+metatheory. Since we have models of our metatheory in virtual double
+categories, we establish sound semantics as well. Put succinctly,
+category theory itself is the logical framework for the study of type
+theories.
 
@@
 
-%% \subsection{Products}
-
-%% We add in parallel two different notions of product: products of sorts
-%% and products of spans, which have the universal property in their
-%% respective categories.
-
-%% \begin{figure}
-%%   \begin{mathpar}
-%%     \inferrule{}{\Gamma \vdash 1 : \Sort}\and
-%%     \inferrule*{}{\Gamma\pipe\alpha:C \vdash () : 1}\and
-%%     \inferrule*{\Gamma\pipe\alpha:C \vdash A : 1}{A = ()}
-%%     \inferrule{}{\Gamma\pipe \Xi \vdash 1 \isaSet}\and
-%%     \inferrule{}{\Gamma\pipe \Phi \vdash () :  1}\and
-%%     \inferrule{\Gamma\pipe \Phi \vdash M :  1}{M = ()}\\\\
-
-%%     \inferrule{\Gamma \vdash C : \Sort \and \Gamma \vdash D : \Sort}
-%%               {\Gamma \vdash C \times D : \Sort}\and
-
-%%     \inferrule{\forall i \in \{1,2\}.~ \Gamma \pipe \alpha:C \vdash A_i : D_i}{\Gamma \pipe \alpha:C \vdash (A_1,A_2) : D_1 \times D_2}
-
-%%     \inferrule{\Gamma \pipe \alpha:C \vdash A : D_1 \times D_2}{\Gamma \pipe \alpha:C \vdash \pi_i A : D_i}
-
-%%     \pi_i(A_1,A_2) = A_i\\
-%%     \inferrule{\Gamma\pipe \alpha:C \vdash A :  D_1\times D_2}{A = (\pi_1 A,\pi_2 A)}\\\\
-
-%%     \inferrule{\Gamma\pipe\Xi \vdash R \isaSet \and \Gamma\pipe\Xi \vdash S\isaSet}
-%%               {\Gamma\pipe\Xi \vdash R \times S\isaSet}\and
-
-%%     \inferrule{\forall i \in \{1,2\}.~ \Gamma \pipe\Phi \vdash M_i : R_i}{\Gamma \pipe\Phi \vdash (M_1,M_2) : R_1 \times R_2}
-
-%%     \inferrule{\Gamma \pipe\Phi\vdash M : R_1 \times R_2}{\Gamma \pipe\Phi \vdash \pi_i M : R_i}
-
-%%     \pi_i(M_1,M_2) = M_i\\
-%%     \inferrule{\Gamma\pipe\Phi \vdash M :  R_1\times R_2}{M = (\pi_1 M,\pi_2 M)}
-%%   \end{mathpar}
-%%   \caption{Products}
-%%   \label{fig:products}
-%% \end{figure}
-
-%% One additional connective we might desire is a ``parallel'' product of
-%% sets $R \times' S : \harr {C_1 \times C_2} {D_1 \times D_2}$ where $R
-%% : \harr {C_1} {D_1}$ and $R : \harr {C_2} {D_2}$. We don't need to
-%% define this as a primitive though because it is expressible in terms
-%% of the two kinds of products we have above:
-
-%% \[
-%% \inferrule
-%% {\Gamma \pipe \alpha : C_i; \beta : D_i \vdash R_i \isaSet}
-%% {\Gamma \pipe \alpha : C_1\times C_2; \beta : D_1 \times D_2
-%%   R_1[\pi_1\alpha/\alpha;\pi_1\beta/\beta] \times R_2[\pi_2\alpha/\alpha;\pi_2\beta/\beta]\isaSet}
-%% \]
-
-%% The empty products give us access now to several traditional notions
-%% of category theory that we couldn't before define, such as objects and
-%% presheaves of either variance:
-
-%% $$\textrm{Object}\,C = \varr 1 C$$
-%% $$\textrm{Presheaf}\,C = \harr 1 C$$
-%% $$\textrm{coPresheaf}\,C = \harr C 1$$
-
-%% However, we find that in practice we don't have much use for these
-%% concepts as most theorems about them easily generalize to more general
-%% theorems about arbitrary functors or profunctors.
+\subsection{Products}
+
+We add in parallel two different notions of product: products of sorts
+and products of spans, which have the universal property in their
+respective categories.
+
+\begin{figure}
+  \begin{mathpar}
+    \inferrule{}{\Gamma \vdash 1 : \Sort}\and
+    \inferrule*{}{\Gamma\pipe\alpha:C \vdash () : 1}\and
+    \inferrule*{\Gamma\pipe\alpha:C \vdash A : 1}{A = ()}
+    \inferrule{}{\Gamma\pipe \Xi \vdash 1 \isaSet}\and
+    \inferrule{}{\Gamma\pipe \Phi \vdash () :  1}\and
+    \inferrule{\Gamma\pipe \Phi \vdash M :  1}{M = ()}\\\\
+
+    \inferrule{\Gamma \vdash C : \Sort \and \Gamma \vdash D : \Sort}
+              {\Gamma \vdash C \times D : \Sort}\and
+
+    \inferrule{\forall i \in \{1,2\}.~ \Gamma \pipe \alpha:C \vdash A_i : D_i}{\Gamma \pipe \alpha:C \vdash (A_1,A_2) : D_1 \times D_2}
+
+    \inferrule{\Gamma \pipe \alpha:C \vdash A : D_1 \times D_2}{\Gamma \pipe \alpha:C \vdash \pi_i A : D_i}
+
+    \pi_i(A_1,A_2) = A_i\\
+    \inferrule{\Gamma\pipe \alpha:C \vdash A :  D_1\times D_2}{A = (\pi_1 A,\pi_2 A)}\\\\
+
+    \inferrule{\Gamma\pipe\Xi \vdash R \isaSet \and \Gamma\pipe\Xi \vdash S\isaSet}
+              {\Gamma\pipe\Xi \vdash R \times S\isaSet}\and
+
+    \inferrule{\forall i \in \{1,2\}.~ \Gamma \pipe\Phi \vdash M_i : R_i}{\Gamma \pipe\Phi \vdash (M_1,M_2) : R_1 \times R_2}
+
+    \inferrule{\Gamma \pipe\Phi\vdash M : R_1 \times R_2}{\Gamma \pipe\Phi \vdash \pi_i M : R_i}
+
+    \pi_i(M_1,M_2) = M_i\\
+    \inferrule{\Gamma\pipe\Phi \vdash M :  R_1\times R_2}{M = (\pi_1 M,\pi_2 M)}
+  \end{mathpar}
+  \caption{Products}
+  \label{fig:products}
+\end{figure}
+
+One additional connective we might desire is a ``parallel'' product of
+sets $R \times' S : \harr {C_1 \times C_2} {D_1 \times D_2}$ where $R
+: \harr {C_1} {D_1}$ and $R : \harr {C_2} {D_2}$. We don't need to
+define this as a primitive though because it is expressible in terms
+of the two kinds of products we have above:
+
+\[
+\inferrule
+{\Gamma \pipe \alpha : C_i; \beta : D_i \vdash R_i \isaSet}
+{\Gamma \pipe \alpha : C_1\times C_2; \beta : D_1 \times D_2
+  R_1[\pi_1\alpha/\alpha;\pi_1\beta/\beta] \times R_2[\pi_2\alpha/\alpha;\pi_2\beta/\beta]\isaSet}
+\]
+
+The empty products give us access now to several traditional notions
+of category theory that we couldn't before define, such as objects and
+presheaves of either variance:
+
+$$\textrm{Object}\,C = \varr 1 C$$
+$$\textrm{Presheaf}\,C = \harr 1 C$$
+$$\textrm{coPresheaf}\,C = \harr C 1$$
+
+However, we find that in practice we don't have much use for these
+concepts as most theorems about them easily generalize to more general
+theorems about arbitrary functors or profunctors.
 
@@
 \item In the slice category: $p(\textrm{wp}(k)) = \pi_-k$
-\item With a universal morphism $\epsilon : \pforallXfromYtoZ {k} \punitinXfromYtoZ {} {(\pi_-k,k,\punitrefl{}{\pi_-k})} {k}$
+\item With a universal morphism $\epsilon : \pendallXdotY {k} \punitinXfromYtoZ {} {(\pi_-k,k,\punitrefl{}{\pi_-k})} {k}$
 \item In the slice category: $\pi_k \epsilon = \id$ (how to express this nicely in SCT?)
@@
 \end{lemma}
-
-\section{Internal Category Theory}
-
-A downside of working synthetically is that since everything we define
-must be functorial/natural, we cannot prove any theorems about
-functions that are not assumed to be functorial or transformations
-that are not assumed ot be natural.
-%
-These kinds of constructions are surprisingly common in basic category theory.
-%
-For instance, when defining a left adjoint $F$ to a given right
-adjoint functor $U$, the functoriality of $F$ is often not assumed,
-rather we are given that each $F a$ has a certain universal mapping
-property and this can then be used to show the action of $F$ on
-objects extends to an action on morphisms.
-%
-This definition is quite important for applications of category theory
-to type theory: in type theory when a type constructor is defined it
-is given introduction, elimination and $\beta\eta$ rules, but no
-primitive syntax is given for a functorial action.
-%
-The fact that functoriality is determined from the universal mapping
-property of the type is then a key step in proving the type theory is
-sound and complete with respect to models which are expected to have
-certain functors on them.
-
-As another example, a common lemma states that an \emph{possibly
-unnatural} transformation is in fact natural if it is the inverse of a
-natural transformation.
-%
-Neither of these examples can be stated as synthetic theorems, as it
-is impossible to quantify over ``possibly unnatural transformations''
-or ``functions on objects''.
-%
-Inside the calculus, we cannot ``see'' that categories have an
-underlying set of objects.
-%
-Fortunately there is a way to use our calculus in a different manner
-to study these constructions: by defining categories \emph{internal
-to} SCT.
-%
-This reproduces classical constructions, since there is a model where
-a category $\cat C$ is interpreted simply as a set, functors as
-functions and profunctors $\harr {\cat C}{\cat D}$ as \emph{spans},
-i.e. sets $R$ with (source and target) projections $\cat C \leftarrow
-R \rightarrow \cat D$.
-%
-The proofs read like the classical proofs (with some care additional
-attention on the ordering of variable usage), but by virtue of being
-syntactic automatically apply to any notion of category internal to a
-virtual equipment, which includes classical internal categories,
-enriched categories and a broad range of generalized multicategory.
-
-We note that a synthetic approach to model this situation may be
-possible were we to extend our calculus with a \emph{modality} $|\cat
-C|$ that takes a category to the discrete category on the set of
-objects. We discuss this in \cref{sec:discussion}.
-
-Next, we introduce the notion of category internal to our syntactic
-virtual equipment and the \emph{monoids and modules} translation that
-shows that, with the inclusion of hom sets, our syntax is already
-suitable for doing syntactic category theory.
-
-The benefits of our syntax become clear when we define internal
-categories, functors, profunctors and transformations: the definitions
-read exactly as the classical definitions.
-\begin{definition}[Internal Category]
-  An \emph{(internal) category} consists of\footnote{the following is
-    easily interpreted as a $\Sigma$ type}:
-  \begin{enumerate}
-  \item A sort $C_0 \isaCat$
-  \item A span $C_1 : \harr{C_0}{ C_0}$
-  \item Identity arrows
-    $C_r : \pendallXdotY {\alpha :  C_0} {\harrapp{C_1} \alpha \alpha}$
-  \item Composition of arrows
-    \[ C_t : \pendallXdotY \alpha {\homrallXYtoZ \beta {\harrapp{C_1} \alpha \beta} {\homrallXYtoZ \gamma {\harrapp{C_1} \beta \gamma} {\harrapp{C_1} \alpha \gamma}}}\]
-
-    %% \[ C_t^\triangleright : \forall \alpha :  C_0. \forall \beta. ({C_1} \beta \alpha) \triangleright \forall \gamma. ({C_1} \gamma \beta) \triangleright {C_1} \gamma \alpha \]
-
-    %% or
-
-    %% \[ C_t^\triangleleft : \forall \alpha. \forall \beta. (\forall \gamma. {C_1} \alpha \gamma \triangleleft {C_1} \beta \gamma) \triangleleft {C_1} \alpha \beta
-    %% \]
-
-    %% or using tensors and covariant homs
-
-    %% \[ C_t^\odot : \forall \alpha. \forall \gamma. (\exists \beta. {C_1}\alpha\beta\odot {C_1}\beta\gamma)\triangleright {C_1}\alpha \gamma \]
-
-    %% or using tensors and contravariant homs
-
-    %% \[ C_t^\odot : \forall \gamma. \forall \alpha. {C_1}\alpha\gamma \triangleleft (\exists \beta. {C_1}\alpha\beta\odot {C_1}\beta\gamma)\]
-  \item Satisfying right unitality, i.e., for any $\Phi \vdash f :
-    \harrapp {C_1} A B$, we'd like $f; r(B) = f$. We can formalize this
-    using an equality of two transformations of type $\pendallXdotY
-    \alpha {\homrallXYtoZ \beta {\harrapp{C_1} \alpha \beta} {\harrapp{C_1} \alpha \beta}}$
-
-    \[ \pendlambdaXdotY \alpha {\homrlambdaXatYdotZ f \beta {\homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {C_t} \alpha} {f} \beta} {\pendappXtoY {C_r} \beta} \beta}}
-    = \pendlambdaXdotY \alpha {\homrlambdaXatYdotZ f \beta f} \]
-
-    These kinds of equalities between 2-cells are very common in
-    internal category theory, so to save space we will suppress the
-    sequence of $\lambda$s that start such equalities. So the above can simply be written as
-
-    \[ { {\homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {C_t} \alpha} {f} \beta} {\pendappXtoY {C_r} \beta} \beta}} = { f}\]
-
-  \item and left unitality (an equality of the same type as right unitality):
-
-    \[ { {\homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {C_t} \alpha} {\pendappXtoY {C_r} \alpha} \alpha} f \beta}} = { f}\]
-  \item and associativity, which is an equality at type
-    \[ \pendallXdotY \alpha {\homrallXYtoZ \beta {\harrapp{C_1} \alpha \beta} {\homrallXYtoZ \gamma {\harrapp{C_1} \beta \gamma} {\homrallXYtoZ \delta {\harrapp{C_1} \gamma \delta} {\harrapp{C_1} \alpha \delta}}}}\]
-
-    \[ \homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {C_t}\alpha} f \beta} {(\homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {C_t}\beta} g \gamma} h \delta)} \delta
-    =
-    \homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {C_t}\alpha} {(\homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {C_t}\alpha} f \beta} g \gamma)} \gamma} h \delta \]
-  \end{enumerate}
-\end{definition}
-
-We can define an internal functor to be a pair of an object function
-and a function on arrows that preserves identity and 
-\begin{definition}[Internal Functor]
-  A functor $F$ between internal categories $C$ and $D$ consists of
-  \begin{enumerate}
-  \item a map of objects $F_0 : \varr {C_0} {D_0}$
-  \item a map of arrows
-    $F_1 : \pendallXdotY {\alpha}
-    {\homrallXYtoZ \beta {\harrapp{C_1} \alpha \beta} {\harrapp{D_1} {F_0 \alpha}{F_0 \beta}}}$
-  \item That preserves identity. This is given by equality between the
-    following terms of type $\forall \alpha:{{C_0}}. D_1 {(F_0 \alpha)}{(F_0\alpha)}$:
-    \[ \homrappXtoYatZ {\pendappXtoY {F_1}\alpha} {\pendappXtoY {C_r} \alpha} \alpha = \pendappXtoY {D_r}{F_0\alpha} \]
-  \item And preserves composition.
-    This is given by an equality between the following terms of type
-
-    \[ \pendallXdotY \alpha
-    \homrallXYtoZ {\beta} {\harrapp{C_1} \alpha \beta} {\homrallXYtoZ \gamma {\harrapp{C_1} \beta \gamma} {\harrapp{D_1} {F_0 \alpha} {F_0 \gamma}}}\]
-
-    First, we can compose in $C$, and then map using $F$:
-    \[ \homrappXtoYatZ {\pendappXtoY{F_1} \alpha} {(\homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {C_t} \alpha} f \beta} g \gamma)}\gamma\]
-
-    or we can map the morphisms with $F$, and then compose them in $D$:
-    \[ \homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {D_t} {(F_0 \alpha)}} {(\homrappXtoYatZ {\pendappXtoY {F_1} \alpha} f \beta)} {(F_0 \beta)}} {(\homrappXtoYatZ {\pendappXtoY {F_1} \beta} g \gamma)} {{F_0} \gamma}\]
-  \end{enumerate}
-\end{definition}
-
-We can immediately define identity and composition of functors.
-
-\begin{definition}
-  For any $C : \textrm{Cat}$, we can define an identity functor
-  $\textrm{Id}_C : \textrm{Functor} C C$ by
-  \begin{enumerate}
-  \item $\textrm{Id}_0 = \lambda \alpha:C_0. \alpha$
-  \item $\textrm{Id}_1 = \pendlambdaXdotY \alpha
-    {\homrlambdaXatYdotZ f {\alpha'} f}$
-  \item Preservation of identity requires proving
-    \[
-    \lambda \alpha. \pendappXtoY {C_r}{\textrm{Id}_0\alpha} \]
-    is equal to
-    \[ \lambda \alpha.
-    \homrappXtoYatZ {\pendappXtoY {\textrm{Id}_1} \alpha} {(\pendappXtoY {C_r}\alpha)} \alpha\]
-    Both sides simply $\beta$ reduce to $ \lambda \alpha. \pendappXtoY {C_r} \alpha$
-    and the result follows by reflexivity.
-  \item Preservation of composition requires proving (under a $\lambda \alpha. \homrlambdaXatYdotZ f \beta \homrlambdaXatYdotZ g \gamma$)
-    \[
-    \homrappXtoYatZ {\pendappXtoY {{\textrm{Id}}_1} \alpha} {(\homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {C_t} \alpha} f \beta} g \gamma)}\gamma
-    = \homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {C_t} {{\textrm{Id}}_0 \alpha}} {(\homrappXtoYatZ {\pendappXtoY {{\textrm{Id}}_1} \alpha} f \beta)} {{\textrm{Id}}_0 \beta}} {(\homrappXtoYatZ {\pendappXtoY{{\textrm{Id}}_1} \beta} g \gamma)} {{\textrm{Id}}_0 \gamma}\]
-    Both sides $\beta$ reduce to
-
-    \[ \homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {C_t} \alpha} f \beta} g \gamma\]
-    so the result holds by reflexivity.
-  \end{enumerate}
-\end{definition}
-
-\begin{definition}
-  Given categories $C,D,E$ and functors $F : \textrm{Functor} C D$ and
-  $G : \textrm{Functor} D E$, we define the composition $G \circ F :
-  \textrm{Functor} C E$ as
-  \begin{enumerate}
-  \item $(G \circ F)_0 = \lambda \alpha. {G_0 ({F_0} \alpha)}$
-  \item $(G \circ F)_1 =
-    \pendlambdaXdotY \alpha {
-    \homrlambdaXatYdotZ f \beta
-    {\homrappXtoYatZ {\pendappXtoY {G_1} {F_0 \alpha}} {\homrappXtoYatZ {\pendappXtoY {F_1} \alpha} f \beta} {F_0 \beta}}}$
-  \item Preservation of identity requires proving
-    \[ \lambda \alpha. \pendappXtoY {E_r} {(G \circ F)_0\alpha} \]
-    is equal to
-    \[ \lambda \alpha.
-    \homrappXtoYatZ {\pendappXtoY {(G \circ F)_1} \alpha} {\pendappXtoY {C_r}\alpha} \alpha \]
-
-    Expanding definitions and $\beta$ reducing we need to show
-
-    \[ \pendlambdaXdotY\alpha {\pendappXtoY {E_r}{G_0 (F_0 \alpha)}} \]
-    is equal to
-    \[ \pendlambdaXdotY \alpha {\homrappXtoYatZ {\pendappXtoY {G_1} {F_0 \alpha}} {(\homrappXtoYatZ {\pendappXtoY {F_1} \alpha} {\pendappXtoY {C_r} \alpha} \alpha)} {F_0 \alpha}}\]
-
-    Proceeding from right to left applying transitivity:
-    \begin{align*}
-      &\pendlambdaXdotY \alpha {\homrappXtoYatZ {\pendappXtoY {G_1} {F_0 \alpha}} {(\homrappXtoYatZ {\pendappXtoY {F_1} \alpha} {\pendappXtoY {C_r} \alpha} \alpha)} {F_0 \alpha}}\\
-      &=\pendlambdaXdotY \alpha {\homrappXtoYatZ {\pendappXtoY {G_1} {F_0 \alpha}} {(\pendappXtoY{(\pendlambdaXdotY \alpha {\homrappXtoYatZ {\pendappXtoY {F_1} \alpha} {\pendappXtoY {C_r} \alpha} \alpha})}\alpha)}} {F_0 \alpha}\\
-      &=\pendlambdaXdotY \alpha {\homrappXtoYatZ {\pendappXtoY {G_1} {F_0 \alpha}} {(\pendappXtoY{(\pendlambdaXdotY \alpha {{\pendappXtoY {D_r} {F_0 \alpha}}})}\alpha)}} {F_0 \alpha}\tag{*}\\
-      &=\pendlambdaXdotY \alpha {\homrappXtoYatZ {\pendappXtoY {G_1} {F_0 \alpha}} {\pendappXtoY {D_r} {F_0 \alpha}}} {F_0 \alpha}\\
-      &= \pendlambdaXdotY \alpha
-      \pendappXtoY {(\pendlambdaXdotY \beta {{\homrappXtoYatZ {\pendappXtoY {G_1} \beta} {\pendappXtoY {D_r} \beta} \beta}})} {F_0 \alpha}\\
-      &= \pendlambdaXdotY \alpha
-      \pendappXtoY {(\pendlambdaXdotY \beta \pendappXtoY {E_r} {G_0 \beta})} {F_0 \alpha}\tag{*}\\
-      &= \pendlambdaXdotY \alpha \pendappXtoY {E_r} {G_0 (F_0 \beta)}\\
-    \end{align*}
-
-    Where the two marked lines follow from congruence using the
-    assumptions about $F,G$.
-
-  \item Preservation of Composition.
-
-    \[ \homrappXtoYatZ {\pendappXtoY{(G \circ F)_1} \alpha} {(\homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {C_t} \alpha} f \beta} g \gamma)}\gamma\]
-    equals
-
-    \[ \homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {E_t} {(G \circ F)_0 \alpha}} {(\homrappXtoYatZ {\pendappXtoY {(G\circ F)_1} \alpha} f \beta)} {(G\circ F)_0 \beta}} {(\homrappXtoYatZ {\pendappXtoY {(G\circ F)_1} \beta} g \gamma)} {{F_0} \gamma}\]
-
-    Reducing, we need to show
-    \[ \homrappXtoYatZ {\pendappXtoY{G_1} {F_0 \alpha}} {(\homrappXtoYatZ {\pendappXtoY {F_1} \alpha} {(\homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {C_t}\alpha} f {\beta}} g {\gamma})} \gamma)} {F_0 \gamma} \]
-
-    is equal to
-    \[ \homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {E_t} {G_0(F_0\alpha)}}
-      {({\homrappXtoYatZ {\pendappXtoY {G_1} {F_0 \alpha}} {(\homrappXtoYatZ {\pendappXtoY {F_1} \alpha} f \beta)} {F_0 \beta}})}
-      {G_0(F_0\beta)}}
-       {({\homrappXtoYatZ {\pendappXtoY {G_1} {F_0 \beta}} {(\homrappXtoYatZ {\pendappXtoY {F_1} \beta} g \gamma)} {F_0 \gamma}})}
-       {{F_0} \gamma}\]
-
-    Going from left to right:
-    \begin{align*}
-      &\homrappXtoYatZ {\pendappXtoY{G_1} {F_0 \alpha}} {(\homrappXtoYatZ {\pendappXtoY {F_1} \alpha} {(\homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {C_t}\alpha} f {\beta}} g {\gamma})} \gamma)} {F_0 \gamma}\\
-      &=
-      \homrappXtoYatZ
-      {\pendappXtoY{G_1} {F_0 \alpha}}
-      {(({\homrappXtoYatZ {\homrappXtoYatZ{\pendappXtoY{(\pendlambdaXdotY \alpha {\homrlambdaXatYdotZ f \beta {\homrlambdaXatYdotZ g \gamma {\homrappXtoYatZ {\pendappXtoY {F_1} \alpha} {(\homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {C_t}\alpha} f {\beta}} g {\gamma})} \gamma}}})}\alpha} f \beta} g \gamma}))}
-      {F_0 \gamma}\\
-      &=
-      \homrappXtoYatZ
-      {\pendappXtoY{G_1} {F_0 \alpha}}
-      {(({\homrappXtoYatZ {\homrappXtoYatZ{\pendappXtoY{(\pendlambdaXdotY \alpha {\homrlambdaXatYdotZ f \beta {\homrlambdaXatYdotZ g \gamma {\homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {D_t} {(F_0 \alpha)}} {(\homrappXtoYatZ {\pendappXtoY {F_1} \alpha} f \beta)} {(F_0 \beta)}} {(\homrappXtoYatZ {\pendappXtoY {F_1} \beta} g \gamma)} {{F_0} \gamma}}}})}\alpha} f \beta} g \gamma}))}
-      {F_0 \gamma}\tag{*}\\
-      &=
-      \homrappXtoYatZ
-      {\pendappXtoY{G_1} {F_0 \alpha}}
-      {\homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {D_t} {(F_0 \alpha)}} {(\homrappXtoYatZ {\pendappXtoY {F_1} \alpha} f \beta)} {(F_0 \beta)}} {(\homrappXtoYatZ {\pendappXtoY {F_1} \beta} g \gamma)} {{F_0} \gamma}}
-      {F_0 \gamma}\\
-      &=
-      {(({\homrappXtoYatZ {\homrappXtoYatZ{\pendappXtoY{(\pendlambdaXdotY {\alpha'} {\homrlambdaXatYdotZ {f'} {\beta'} {\homrlambdaXatYdotZ {g'} {\gamma'} {{\homrappXtoYatZ {\pendappXtoY {G_1} {\alpha'}} {(\homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {D_t}{\alpha'}} {f'} {\beta'}} {g'} {\gamma'})} {\gamma'}}}}})}{F_0\alpha}} {(\homrappXtoYatZ {\pendappXtoY {F_1} \alpha} f \beta)} {F_0\beta}} {(\homrappXtoYatZ {\pendappXtoY {F_1} \beta} g \gamma)} {F_0\gamma}}))}\\
-      &=
-      {(({\homrappXtoYatZ {\homrappXtoYatZ{\pendappXtoY{(\pendlambdaXdotY {\alpha'} {\homrlambdaXatYdotZ {f'} {\beta'} {\homrlambdaXatYdotZ {g'} {\gamma'} {{\homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {E_t} {(G_0 {\alpha'})}} {(\homrappXtoYatZ {\pendappXtoY {G_1} {\alpha'}} {f'} {\beta'})} {(G_0 \beta')}} {(\homrappXtoYatZ {\pendappXtoY {G_1} {\beta'}} {g'} {\gamma'})} {{G_0} \gamma'}}}}})}{F_0\alpha}} {(\homrappXtoYatZ {\pendappXtoY {F_1} \alpha} f \beta)} {F_0\beta}} {(\homrappXtoYatZ {\pendappXtoY {F_1} \beta} g \gamma)} {F_0\gamma}}))}\\
-      &= \homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {E_t} {G_0(F_0\alpha)}}
-      {({\homrappXtoYatZ {\pendappXtoY {G_1} {F_0 \alpha}} {(\homrappXtoYatZ {\pendappXtoY {F_1} \alpha} f \beta)} {F_0 \beta}})}
-      {G_0(F_0\beta)}}
-       {({\homrappXtoYatZ {\pendappXtoY {G_1} {F_0 \beta}} {(\homrappXtoYatZ {\pendappXtoY {F_1} \beta} g \gamma)} {F_0 \gamma}})}
-       {{F_0} \gamma}
-    \end{align*}
-  \end{enumerate}
-\end{definition}
-
-\begin{definition}[Internal Profunctor]
-  A profunctor $R$ between internal categories $\cat C$ and $\cat D$
-  consists of
-  \begin{enumerate}
-  \item A span $R_s : \harr {C_0} {D_0}$.
-  \item A ternary ``mixed composition'' action:
-    $R_t :\pendallXdotY \alpha {\homrallXYtoZ {\alpha'} {\harrapp {C_1} {\alpha}{\alpha'}} {\homrallXYtoZ {\beta'} {\harrapp {R_s} {\alpha'}{\beta'}} {\homrallXYtoZ {\beta} {\harrapp {D_1} {\beta'} {\beta}} {\harrapp {R_s} {\alpha}{\beta}}}}}$
-
-  \item Satisfying unitality (under a $\pendlambdaXdotY \alpha {\homrlambdaXatYdotZ f \beta {}}$):
-    \[
-    f = \homrappXtoYatZ {\homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {R_t} \alpha} {\pendappXtoY {C_r} \alpha} \alpha}f \beta} {\pendappXtoY {D_r} \beta} \beta
-    \]
-
-  \item And associativity (both with a large implicit
-    $\pendlambdaXdotY \alpha {\homrlambdaXatYdotZ g \beta \homrlambdaXatYdotZ {g'} {\beta'}
-    \homrlambdaXatYdotZ {r} {\beta''}\homrlambdaXatYdotZ {f'} {\alpha''}\homrlambdaXatYdotZ {f} {\alpha'}}$)
-
-    % f : a -> a'
-    % f' : a' -> a''
-    % r : a'' -> b''
-    % g' : b'' -> b'
-    % g : b' -> b
-    %%
-    %
-    % f o (f' o r o g') o g
-    %% and
-    % (f o f') o r o (g' o g)
-
-    \[ \homrappXtoYatZ {\homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {R_t} {\alpha}} {f} {\alpha'}}{({\homrappXtoYatZ {\homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {R_t} {\alpha'}} {f'} {\alpha''}}r {\beta''}} {g'} {\beta'}})} {\beta'}} {g} {\beta}\]
-    is equal to
-    \[ \homrappXtoYatZ
-       {\homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {R_t} {\alpha}} {(\homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {C_t} {\alpha}} {f} {\alpha'}} {f'} {\alpha''})} {\alpha''}} r {\beta''}}
-       {(\homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {D_t} {\beta''}} {g'} {\beta'}} g {\beta})} {\beta}\]
-  \end{enumerate}
-\end{definition}
-
-We can immediately define the Hom profunctor for any category $C$ as
-$(\text{Hom}_C R)_s = C_1$ with actions given by $C_t$.
-
-Next, we show that Profunctors can be restricted by functors.
-
-\begin{definition}[Restriction of Profunctors]
-  If $C, D, C', D' : \text{Cat}$, and $F : \text{Functor} C C', G : \text{Functor} D D'$ and
-  $R : \text{Prof} C' D'$, then we can define a profunctor $R[F;G] : \text{Prof} C D$ by
-  \begin{enumerate}
-  \item $R[F;G]_s = \lambda (\alpha:C,\beta:D). R_s (F_0 \alpha) (G_0 \beta)$
-  \item with composition action given by
-
-    \[ R[F;G]_t =
-    \pendlambdaXdotY \alpha {\homrlambdaXatYdotZ f {\alpha'} {\homrlambdaXatYdotZ r {\beta'} {\homrlambdaXatYdotZ g {\beta} {\homrappXtoYatZ {\homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {R_t} {F_0 \alpha}} {(\homrappXtoYatZ {\pendappXtoY {F_1} {\alpha}} f {\alpha'})} {F_0 \alpha'}} r {G_0 \beta'}} {(\homrappXtoYatZ {\pendappXtoY {G_1} {\beta'}} g \beta)} {G_0 \beta}}}} }
- \]
-  %% \item $\cat C'$ composition:
-  %%   \[ \lambda \alpha. \lambda^\triangleleft (\beta, (f : C_1 \alpha \alpha',\alpha',\alpha', r : R_s \alpha' \beta)). R_l(F_1 f, r) \]
-  %% \item $\cat D'$ composition:
-  %%   \[ \lambda \alpha. \lambda^\triangleleft ('\beta, (r : R_s \alpha \alpha',\alpha', \beta',h : D_1 \beta' \beta)). R_r(r, G_1 h) \]
-  \end{enumerate}
-  Identity and associativity properties should follow. TODO
-\end{definition}
-
-\begin{definition}[Homomorphism]
-  Given $R : \text{Prof} C D$ and $S : \text{Prof} C D$, a
-  homomorphism $m$ from $R$ to $S$ consists of
-  \begin{enumerate}
-  \item $m_f : \pendallXdotY \alpha {\homrallXYtoZ \beta {\harrapp R \alpha \beta}{\harrapp S \alpha \beta}}$
-    % m_f (R_t f r g) = S_t f (m_f r) g
-  \item Satisfying naturality (under a $\ldots$)
-
-    \[
-    \homrappXtoYatZ {\pendappXtoY{m_f} \alpha} {(\homrappXtoYatZ {\homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {R_t} \alpha} f {\alpha'}} r {\beta'}} g \beta)} \beta
-    =
-    \homrappXtoYatZ {\homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {S_t} \alpha} {f} {\alpha'}} {(\homrappXtoYatZ {\pendappXtoY {m_f}{\alpha'}} r {\beta'})} {\beta'}} g \beta
-    \]
-  \end{enumerate}
-\end{definition}
-
-\begin{definition}
-  An isomorphism between profunctors $R$ and $S$ is a mutually inverse
-  pair of homomorphisms.
-\end{definition}
-
-Additionally, we can define natural transformations between functors,
-but more generally, we can define a natural element of any
-endo-profunctor, where natural transformations from $F$ to $G$ are the
-same as natural elements of $\text{Hom}(F-,G=)$.
-
-\begin{definition}[Natural Element]
-  A natural element $\phi$ of a profunctor $R : \text{Prof}\,C C$ is a term
-  \[ \phi : \pendallXdotY \alpha {\harrapp R \alpha \alpha} \]
-  that is natural in that
-  \[
-  \homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {C_t} \alpha} {\pendappXtoY \phi \alpha} \alpha} f {\alpha'}
-  = \homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {C_t } \alpha} f {\alpha'}} {\pendappXtoY \phi {\alpha'}}{\alpha'}
-  \]
-\end{definition}
-
-\begin{definition}[Natural Transformation]
-  A natural transformation from $F : \text{Functor} \,C\, D$ and $G :
-  \text{Functor}\,C\,D$ is a natural element of $\text{Hom}(D)[F;G]$
-\end{definition}
-
-Composition of natural transformations is then given by $C_t$.
-
-Natural elements are closely related to homomorphisms: a natural
-element equivalent to a homomorphism out of a Hom profunctor. This is
-our first instance of a ``Yoneda'' principle that the behavior of some
-structure that varies naturally with the morphisms of a category is
-fully determined by its action on the identity morphisms.
-
-\begin{theorem}
-  Natural elements of $R : \text{Prof}\, C C$ are equivalent to
-  homomorphisms from $\text{Hom}_C$ to $R$.
-\end{theorem}
-\begin{enumerate}
-\item Given a natural element $\phi$ of $R$, we can define a
-  homomorphism by
-  \[ \pendlambdaXdotY \alpha {\homrlambdaXatYdotZ f {\alpha'} \homrappXtoYatZ {\homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {R_t} \alpha} {\pendappXtoY {C_r} \alpha} \alpha}{\pendappXtoY \phi \alpha} {\alpha}} f {\alpha'}} \]
-\item Given a homomorphism $m$ from $\text{Hom}(C)$ to $R$, define the
-  element as
-  \[ \pendlambdaXdotY \alpha {\homrappXtoYatZ {\pendappXtoY {m_f} \alpha} {\pendappXtoY {C_r}\alpha} \alpha}  \]
-
-  We should be able to show this is an isomorphism in the logic TODO!
-\end{enumerate}
-
-%% This last definition is probably unfamiliar to the reader that is not
-%% experienced with profunctors.
-%% %
-%% However, we can show that the traditional definition is equivalent to
-%% a natural transformation $\alpha : \cat C_0 \pipe \cdot \vdash
-%% t_{\alpha} : F(\alpha) \to_{\cat D} G(\alpha')$.
-
-%% \begin{definition}[Traditional Natural Transformation]
-%%   A traditional natural transformation from $\alpha : \cat C_0 \vdash
-%%   F(\alpha) : \cat D_0$ to $\alpha : \cat \vdash G(\alpha') : \cat
-%%   D_0$ consists of
-%%   \begin{enumerate}
-%%   \item A term $\cdot \vdash t_{\alpha} : F(\alpha) \to_{\cat D} G(\alpha)$
-%%   \item Such that
-%%     \[ \phi : \alpha \to_{\cat C} \alpha' \vdash t_{\alpha'} \circ F_1(\phi) \equiv G_1(\phi) \circ t_{\alpha} : F(\alpha) \to_{\cat D} G(\alpha')
-%%     \]
-%%   \end{enumerate}
-%% \end{definition}
-
-%% \begin{theorem}[Equivalent Presentations of Natural Transformations]
-%%   There is a bijection between traditional natural transformations $t
-%%   : F \to G$ and profunctor homomorphisms $h : \alpha \to_{\cat C}
-%%   \alpha' \vdash F(\alpha) \to_{\cat D} G(\alpha')$ given by
-%%   \begin{enumerate}
-%%   \item From a natural transformation $t$ we define a homomorphism
-%%     $h(t)$ by composing on the $F$ side. The choice of this side is
-%%     arbitrary and we could just as easily have used $G$ since they are
-%%     equal by naturality.
-%%     \[ \phi : \alpha \to_{\cat C} \alpha' \vdash t_{\alpha'} \circ F_1(\phi) : F(\alpha) \to_{\cat D} G_1(\alpha') \]
-%%   \item From a homomorphism $\phi.h$, we define a natural transformation $t(h)$ as
-%%     \[ \alpha \pipe \cdot \vdash h[{\id_{\alpha}}/\phi] \]
-%%   \end{enumerate}
-%% \end{theorem}
-%% \begin{proof}
-%%   First we show that the maps $h(\cdot),t(\cdot)$ actually produce
-%%   homomorphisms and natural transformations.
-%%   \begin{enumerate}
-%%   \item On the left, we need to show $h(t)_{\phi} \circ F_1(\psi) =
-%%     h(t)_{|phi \circ \psi}$. Expanding the definition, we just need
-%%     associativity and functoriality of $F_1$:
-%%     \begin{align*}
-%%       h(t)_{\phi} \circ F_1(\psi) &\equiv (t_{\alpha'} \circ F_1(\phi)) \circ F_1(\psi)\\
-%%       &\equiv t_{\alpha'} \circ F_1(\phi \circ \psi)\\
-%%       &\equiv h(t)_{\phi \circ \psi}
-%%     \end{align*}
-%%     on the right, we use naturality:
-%%     \begin{align*}
-%%       G_1(\theta) \circ h(t)_{\phi} &\equiv G_1(\theta) \circ (t_{\alpha'} \circ F_1(\phi))\\
-%%       &\equiv (t_{\alpha''} \circ F_1(\theta)) \circ F_1(\phi)\\
-%%       &\equiv t_{\alpha''} \circ F_1(\theta \circ \phi)\\
-%%       &\equiv t_{\theta \circ \phi}
-%%     \end{align*}
-%%   \item In the other direction, to prove naturality:
-%%     \begin{align*}
-%%       h[\id_\alpha/\phi] \circ F_1(\psi) &\equiv h[\id_\alpha \circ \psi/\phi]\\
-%%       &\equiv h[\psi/\phi]\\
-%%       &\equiv h[\psi \circ \id_{\alpha'}/\phi]\\
-%%       &\equiv G_1(\psi) \circ h[\id_{\alpha}/\phi]
-%%     \end{align*}
-%%   \end{enumerate}
-%%   Next we seek to prove that the functions are mutually inverse.
-%%   \begin{enumerate}
-%%   \item Round trip for homomorphisms:
-%%     \begin{align*}
-%%       h(t(h))_{\phi} &= t(h)_{\alpha} \circ F_1[\phi]\\
-%%       &= h[\id_{\alpha}] \circ F_1[\phi]\\
-%%       &= h[\id \circ \phi]\\
-%%       &= h[\phi]
-%%     \end{align*}
-%%   \item Round trip for natural transformations uses the fact that
-%%     functors preserve identity.
-%%     \begin{align*}
-%%       t(h(t))_{\alpha} &= h(t)_{\id_{\alpha}}\\
-%%       &= t_{\alpha} \circ F_1[\id_{\alpha}]\\
-%%       &= t_{\alpha} \circ \id_{F_0\alpha}\\
-%%       &= t_{\alpha}
-%%     \end{align*}
-%%   \end{enumerate}
-%% \end{proof}
-
-\subsection{Internal Category Translation}
-
-\subsection{Universal Properties}
-
-Since we have profunctors as a primitive notion, the most natural
-presentation of universal properties is given by the notion of a
-functor representing a profunctor.
-
-First, we have the ``algebraic'' notion of a functor representing a
-profunctor, i.e., when a profunctor $R : C\pto D$ is equivalent to
-$Hom_D(-,G=)$.
-\begin{definition}[Right Representability]
-  Let $C, D : \text{Cat}$ and $Q : \textrm{Prof}\,C D$ be a profunctor
-  between them. Then a \emph{right representation} of $Q$ consists of
-  \begin{enumerate}
-  \item A functor $G : \textrm{Functor}\, D C$
-  \item An isomorphism $I,I^-$ between $Q$ and $\text{Hom}(C)[Id;G]$
-  \end{enumerate}
-\end{definition}
-
-This is a nice definition for applications, but a lot of the structure
-(homomorphism, functoriality of G) is determined by other components
-of the structure. This can be simplified to the notion of a
-``universal morphism''.
-
-\begin{definition}[Universal Morphism]
-  A profunctor $Q : \textrm{Prof}\, C D$ is right representable when
-  there exists
-  \begin{enumerate}
-  \item A function on objects $G_0 : \varr {D_0} {C_0}$
-  \item A ``universal morphism'', i.e., a term $\epsilon :
-    \pendallXdotY \beta {\harrapp Q {G_0 \beta} \beta}$ (not
-    presupposed to be natural).
-  \item An ``introduction rule'', i.e., a term $I : \pendallXdotY
-    \alpha {\homrallXYtoZ \beta {\harrapp {Q_s}\alpha \beta} {\harrapp
-        {C_1} \alpha {G_0\beta}}}$ (not presupposed to satisfy the
-    homomorphism condition).
-  \item Such that composition with $\epsilon$ is a left and right
-    inverse to $I$. In more detail, we have two equalities.
-
-    \begin{enumerate}
-    \item First, the $\beta$ rule states that composing the universal
-      morphism with the introduction rule applied to a morphism $q : \harrapp {Q} \alpha {\beta}$ is
-      equal to the original $q$:
-
-      \[ q = \homrappXtoYatZ {\homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {Q_t} \alpha} {(\homrappXtoYatZ {\pendappXtoY I \alpha} {q} \beta)} {G_0 \beta}} {\pendappXtoY \epsilon \beta} {\beta}} {\pendappXtoY {D_r} \beta} {\beta} \]
-
-    \item Second, the $\eta$ rule says that every morphism $f :
-      \harrapp {C_1} \alpha {G_0 \beta}$ can be reconstructed using
-      $I$ after postcomposing with $\epsilon$:
-
-      \[ f = \homrappXtoYatZ {\pendappXtoY I \alpha} {({\homrappXtoYatZ {\homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {Q_t} \alpha} {f} {G_0 \beta}} {\pendappXtoY \epsilon \beta} {\beta}} {\pendappXtoY {D_r} \beta} {\beta}  })} {\beta}\]
-
-      As a special case, if $f = \pendappXtoY {C_r} {G_0 \beta}$, this shows that $\homrappXtoYatZ {\pendappXtoY I {G_0 \beta}} {\pendappXtoY \epsilon \beta} \beta = \pendappXtoY {C_r} {G_0 \beta}$
-    \end{enumerate}
-  \end{enumerate}
-\end{definition}
-
-These two notions turn out to be equivalent because the ``extra''
-information in the hom isomorphism is in fact fully determined by the
-data of a universal morphism: the functoriality of $G$ and the
-naturality of $I,I_{inv}$.
-%
-The proof can be broken up into 3 conceptual steps:
-\begin{enumerate}
-\item First,, the functoriality of $G$ is derivable from $I,\epsilon$.
-\item Next, $I_{inv}$ is a homomorphism automatically since composition
-  with a morphism is automatically natural, where $I_{inv}$ is defined as
-  \[\pendlambdaXdotY \alpha {\homrlambdaXatYdotZ f {\beta} \homrappXtoYatZ {\homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {Q_t} \alpha} f \beta}{\pendappXtoY \epsilon \beta} {\beta}} {\pendappXtoY {D_r }\beta} {\beta}} \]
-
-  \item Finally, $I$ is a homomorphism because the inverse of a
-    natural transformation is always natural.
-\end{enumerate}
-
-%% Part of this equivalence is the \emph{Yoneda Lemma}, which we state
-%% now in its internal, parameterized form.
-%% The usual non-parameterized Yoneda lemma can be formed if we have a
-%% unit sort.
-
-\begin{lemma}[Universal Constructions are functorial]
-  For any universal construction $(G_0,\epsilon,I)$ for $Q :
-  \text{Prof}\,C\,D$, we can extend $G_0$ to a functor by defining
-
-  \[ G_1 = \pendlambdaXdotY \beta {\homrlambdaXatYdotZ f {\beta'}
-    {\homrappXtoYatZ {\pendappXtoY I {G_0\beta}} {(\homrappXtoYatZ {\homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {Q_t} {G_0\beta}} {\pendappXtoY {C_r} {G_0\beta}} {G_0\beta}} {\pendappXtoY \epsilon {\beta}} \beta} {f} {\beta'})} {\beta'}}
-  }\]
-
-  Furthermore, $I,I^-1$ are homomorphisms with respect to the induced
-  profunctor structure.
-\end{lemma}
-\begin{proof}
-  \begin{enumerate}
-  \item To prove $G_1$ preserves identity we need to show
-    \[ {\homrappXtoYatZ {\pendappXtoY I {G_0\beta}} {(\homrappXtoYatZ {\homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {Q_t} {G_0\beta}} {\pendappXtoY {C_r} {G_0\beta}} {G_0\beta}} {\pendappXtoY \epsilon {\beta}} \beta} {\pendappXtoY {D_r} {\beta}} {\beta})} {\beta}} = \pendappXtoY{C_r} {G_0\beta}\]
-    which is precisely the $\eta$ principle.
-
-  \item Next, we show $I_{inv}$ is a homomorphism from $\text{Hom}(C)[\text{Id}(C);G]$ to $Q$.
-    All of this is under the binders
-    \[\lambdatrinary {\alpha : C} {f : \harrapp {C_s} \alpha {\alpha'}}
-    {\alpha' : C} {g : \harrapp{C_s} {\alpha'} {G_0 \beta'}}
-    {\beta' : D} {h : \harrapp {D_s} {\beta'} {\beta}}
-    {}\]
-
-    % given
-    % f : a -> a' in C
-    % g : a' -> G b'
-    % h : b' -> b
-
-    %% I_{inv} (f ;C (g ;C G_1 h)) = Q_t f (I_{inv} g) h
-
-    \begin{align*}
-      & \appunary
-      {I_{inv}}
-      \alpha
-      {(\appbinary {C_t} \alpha f {\alpha'} {(
-          \appbinary {C_t} {\alpha'} g {G_0\beta'}
-          {(\appunary {G_1} {\beta'} h \beta)} {G_0 \beta})} {G_0 \beta})}
-      {\beta}\\
-      &=
-      \apptrinary {Q_t} \alpha
-      {(\appbinary {C_t} \alpha f {\alpha'} {(
-          \appbinary {C_t} {\alpha'} g {G_0\beta'}
-          {(\appunary {G_1} {\beta'} h \beta)} {G_0 \beta})} {G_0 \beta})} \beta
-      {\pendappXtoY \epsilon \beta} \beta
-      {\pendappXtoY {D_r} \beta} {\beta}\\
-      &=
-      \apptrinary {Q_t} \alpha
-      {(\appbinary {C_t} \alpha f {\alpha'} {(
-          \appbinary {C_t} {\alpha'} g {G_0\beta'}
-          {(\appunary {G_1} {\beta'} h \beta)} {G_0 \beta})} {G_0 \beta})} \beta
-      {\pendappXtoY \epsilon \beta} \beta
-      {(\appbinary {D_t} \beta {\pendappXtoY {D_r} \beta} \beta {\pendappXtoY {D_r} \beta} \beta)} {\beta}\\
-      &=
-      \apptrinary {Q_t} \alpha
-      {(\appbinary {C_t} \alpha {(\appbinary {C_t} \alpha f {\alpha'} g {G_0 \beta'})} {G_0 \beta'}
-        {(\appunary {G_1} {\beta'} h \beta)} {G_0 \beta})} \beta
-      {\pendappXtoY \epsilon \beta} \beta
-      {(\appbinary {D_t} \beta {\pendappXtoY {D_r} \beta} \beta {\pendappXtoY {D_r} \beta} \beta)} {\beta}\\
-      &=
-      \apptrinary {Q_t} \alpha
-      {(\appbinary {C_t} \alpha f {\alpha'} g {G_0 \beta'})} {G_0 \beta'}
-      {(\apptrinary {Q_t} {G_0 \beta'}
-        {(\appunary {G_1} {\beta'} h {\beta})} {G_0 \beta}
-        {\pendappXtoY \epsilon \beta} \beta
-        {\pendappXtoY {D_r} \beta} \beta)} \beta
-      {\pendappXtoY {D_r} \beta} \beta
-      \\
-      &=
-      \apptrinary {Q_t} \alpha
-      {(\appbinary {C_t} \alpha f {\alpha'} g {G_0 \beta'})} {G_0 \beta'}
-      {(\appunary {I_{inv}} {G_0 \beta'} {(\appunary {G_1} {\beta'} h \beta)} \beta)} \beta
-      {\pendappXtoY {D_r} \beta} \beta
-      \\
-      &=
-      \apptrinary {Q_t} \alpha
-      {(\appbinary {C_t} \alpha f {\alpha'} g {G_0 \beta'})} {G_0 \beta'}
-      {(\appunary {I_{inv}} {G_0 \beta'} {(\appunary {I} {G_0\beta'} {(\apptrinary {Q_t} {G_0 \beta'} {\pendappXtoY {C_r} {G_0 \beta'}} {G_0 \beta'} {\pendappXtoY \epsilon {\beta'}} {\beta'} h \beta)} \beta)} \beta)} \beta
-      {\pendappXtoY {D_r} \beta} \beta
-      \\
-      &=
-      \apptrinary {Q_t} \alpha
-      {(\appbinary {C_t} \alpha f {\alpha'} g {G_0 \beta'})} {G_0 \beta'}
-      {(\apptrinary {Q_t} {G_0 \beta'} {\pendappXtoY {C_r} {G_0 \beta'}} {G_0 \beta'} {\pendappXtoY \epsilon {\beta'}} {\beta'} h \beta)} \beta
-      {\pendappXtoY {D_r} \beta} \beta
-      \\
-      &=
-      \apptrinary {Q_t} \alpha
-      {(\appbinary {C_t} \alpha {(\appbinary {C_t} \alpha f {\alpha'} g {G_0 \beta'})} {G_0 \beta'}
-        {\pendappXtoY {C_r} {G_0 \beta'}} {G_0 \beta'})} {G_0 \beta'}
-      {\pendappXtoY \epsilon {\beta'}} {\beta'}
-      {(\appbinary {D_t} {\beta'} h \beta {\pendappXtoY {D_r} \beta} \beta)} \beta\\
-      &=
-      \apptrinary {Q_t} \alpha
-      {(\appbinary {C_t} \alpha f {\alpha'} g {G_0 \beta'})} {G_0 \beta'}
-      {\pendappXtoY \epsilon {\beta'}} {\beta'}
-      {(\appbinary {D_t} {\beta'} {\pendappXtoY {D_r} {\beta'}} {\beta'} h \beta )} \beta\\
-      &=
-      \apptrinary {Q_t} \alpha
-      f {\alpha'}
-      {(\apptrinary {Q_t} {\alpha'} g {G_0 \beta'} {\pendappXtoY \epsilon {\beta'}} {\beta'}{\pendappXtoY {D_r} {\beta'}} {\beta'})} {\beta'}
-      h \beta\\
-      &= \homrappXtoYatZ {\homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {Q_t} \alpha} f {\alpha'}} {(\homrappXtoYatZ {\pendappXtoY {I_{inv}} {\alpha'}} {g} {\beta'})} {\beta'}} {h} {\beta}\\
-    \end{align*}
-
-  \item To prove $G_1$ preserves composition given $f : \harrapp {D_1} {\beta} {\beta'}$, $f' : \harrapp {D_1} {\beta'} {\beta''}$:
-    \begin{align*}
-      & \appunary {G_1} {\beta} {\appbinary {D_t} {\beta} f {\beta'} {f'} {\beta''}} {\beta'}\\
-      &= \appunary {I} {G_0 \beta} {(\apptrinary {Q_t} {G_0 \beta} {\pendappXtoY {C_r} {G_0 \beta}} {G_0 \beta} {\pendappXtoY \epsilon {\beta}} {\beta} {({\appbinary {D_t} {\beta} f {\beta'} {f'} {\beta''}})} {\beta''})} {\beta''}\\
-      &= \appunary {I} {G_0 \beta} {(\apptrinary {Q_t} {G_0 \beta} {(\appbinary {C_t} {G_0 \beta} {\pendappXtoY {C_r} {G_0 \beta}} {G_0 \beta} {\pendappXtoY {C_r} {G_0 \beta}} {G_0 \beta})} {G_0 \beta} {\pendappXtoY \epsilon {\beta}} {\beta} {({\appbinary {D_t} {\beta} f {\beta'} {f'} {\beta''}})} {\beta''})} {\beta''}\\
-      &=
-      \appunary {I} {G_0 \beta}
-      {(\apptrinary {Q_t} {G_0 \beta}
-        {\pendappXtoY {C_r} {G_0 \beta}} {G_0 \beta}
-        {(\apptrinary {Q_t} {G_0 \beta}
-          {\pendappXtoY {C_r} {G_0 \beta}} {G_0 \beta}
-          {\pendappXtoY \epsilon {\beta}} {\beta}
-          f {\beta'})} {\beta'}
-        {f'} {\beta''}
-        )}
-      {\beta''}\\
-      &=
-      \appbinary {C_t} {G_0 \beta}
-      {\pendappXtoY {C_r} {G_0 \beta}} {G_0 \beta}
-      {(\appbinary {C_t} {G_0 \beta}
-        {(\appunary {I} {G_0 \beta}
-        {(\apptrinary {Q_t} {G_0 \beta}
-          {\pendappXtoY {C_r} {G_0 \beta}} {G_0 \beta}
-          {\pendappXtoY \epsilon {\beta}} {\beta}
-          f {\beta'})} {\beta'})} {G_0 \beta'}
-        {(\appunary {G_1} {\beta'} {f'} {\beta''})} {G_0 \beta''})}
-      {G_0 \beta''}\\
-      &=
-      \appbinary {C_t} {G_0 \beta}
-        {(\appbinary {C_t} {G_0 \beta}
-          {\pendappXtoY {C_r} {G_0 \beta}} {G_0 \beta}
-          {(\appbinary {C_t} {G_0 \beta}
-            {(\appunary {I} {G_0 \beta} {\pendappXtoY \epsilon {\beta}} {\beta})} {G_0 \beta}
-            {(\appunary {G_1} \beta f {\beta'})} {G_0 \beta'})} {G_0 \beta'})} {G_0 \beta'}
-        {(\appunary {G_1} {\beta'} {f'} {\beta''})} {G_0 \beta''}\\
-      &=
-      \appbinary {C_t} {G_0 \beta}
-        {(\appbinary {C_t} {G_0 \beta}
-            {(\appunary {I} {G_0 \beta} {\pendappXtoY \epsilon {\beta}} {\beta})} {G_0 \beta}
-            {(\appunary {G_1} \beta f {\beta'})} {G_0 \beta'})} {G_0 \beta'}
-        {(\appunary {G_1} {\beta'} {f'} {\beta''})} {G_0 \beta''}\\
-      &=
-      \appbinary {C_t} {G_0 \beta}
-        {(\appbinary {C_t} {G_0 \beta}
-            {\pendappXtoY {C_r} {G_0 \beta}} {G_0 \beta}
-            {(\appunary {G_1} \beta f {\beta'})} {G_0 \beta'})} {G_0 \beta'}
-        {(\appunary {G_1} {\beta'} {f'} {\beta''})} {G_0 \beta''}\\
-      &=
-      \appbinary {C_t} {G_0 \beta}
-      {(\appunary {G_1} \beta f {\beta'})} {G_0 \beta'}
-        {(\appunary {G_1} {\beta'} {f'} {\beta''})} {G_0 \beta''}\\
-      %% &\homrappXtoYatZ {\pendappXtoY {G_1} \beta} {(\homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {D_t} {\beta}} {f} {\beta'}} {f'} {\beta''})} {\beta''}\\
-      %% &={\homrappXtoYatZ {\pendappXtoY I {G_0\beta}} {(\homrappXtoYatZ {\homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {Q_t} {G_0\beta}} {\pendappXtoY {C_r} {G_0\beta}} {G_0\beta}} {\pendappXtoY \epsilon {\beta}} \beta} {(\homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {D_t} {\beta}} {f} {\beta'}} {f'} {\beta''})} {\beta''})} {\beta''}}\\
-      %% &={\homrappXtoYatZ {\pendappXtoY I {G_0\beta}} {(\homrappXtoYatZ {\homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {Q_t} {G_0\beta}} {({\pendappXtoY {C_r} {G_0\beta}})} {G_0\beta}} {\pendappXtoY \epsilon {\beta}} \beta} {(\homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {D_t} {\beta}} {f} {\beta'}} {f'} {\beta''})} {\beta''})} {\beta''}}\\
-    \end{align*}
-    %% \[
-    %%   {\homrappXtoYatZ {\pendappXtoY I {G_0\beta}} {(\homrappXtoYatZ {\homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {Q_t} {G_0\beta}} {\pendappXtoY {C_r} {G_0\beta}} {G_0\beta}} {\pendappXtoY \epsilon {\beta}} \beta} {(\homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {D_t} {\beta}} {f} {\beta'}} {f'} {\beta''})} {\beta''})} {\beta''}}
-    %% \]
-  \end{enumerate}
-\end{proof}
-
-\begin{lemma}[Inverse of homomorphism is a homomorphism]
-  Let $\phi$ be a homomorphism from $R$ to $S : \text{Prof} C D$ and
-  let $\psi : \pendallXdotY \alpha {\homrallXYtoZ \beta {\harrapp S
-      \alpha \beta}{\harrapp R \alpha \beta}}$. If $\psi$ is a
-  two-sided inverse to $\phi_f$ then $\psi$ is a homomorphism.
-\end{lemma}
-\begin{proof}
-  We need to show the homomorphism condition for $\psi$:
-  \[ \homrappXtoYatZ {\pendappXtoY \psi \alpha} {(\homrappXtoYatZ {\homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {S_t} \alpha} f {\alpha'}} s {\beta'}} g {\beta})} {\beta}
-    = \homrappXtoYatZ {\homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {R_t} \alpha} f {\alpha'}} {(\homrappXtoYatZ {\pendappXtoY \psi {\alpha'}} s {\beta'})} {\beta'}} g \beta
-    \]
-
-    \begin{align*}
-      \homrappXtoYatZ {\pendappXtoY \psi \alpha} {(\homrappXtoYatZ {\homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {S_t} \alpha} f {\alpha'}} s {\beta'}} g {\beta})} {\beta}
-      &= \homrappXtoYatZ {\pendappXtoY \psi \alpha} {(\homrappXtoYatZ {\homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {S_t} \alpha} f {\alpha'}} {(\homrappXtoYatZ {\pendappXtoY {\phi_f} {\alpha'}} {(\homrappXtoYatZ {\pendappXtoY \psi {\alpha'}} s {\beta'})} {\beta'})} {\beta'}} g {\beta})} {\beta}\\
-      &= \homrappXtoYatZ {\pendappXtoY \psi \alpha}
-      {(\homrappXtoYatZ {\pendappXtoY {\phi_f} \alpha} {(\homrappXtoYatZ {\homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {R_t} \alpha} f {\alpha'}} {(\homrappXtoYatZ {\pendappXtoY \psi {\alpha'}} s {\beta'})} {\beta'}} g \beta)} \beta)} {\beta}\\
-      & = \homrappXtoYatZ {\homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {R_t} \alpha} f {\alpha'}} {(\homrappXtoYatZ {\pendappXtoY \psi {\alpha'}} s {\beta'})} {\beta'}} g \beta
-    \end{align*}
-\end{proof}
-
-%% \subsection{Self-Interpretation: Hom of Internal Profunctors}
-
-%% We can define a version of the hom $\homrallXYtoZ \alpha P Q$ on
-%% \emph{internal} profunctors using a combination of this connective,
-%% and an \emph{equalizer} to enforce a homomorphism condition.
-
-%% \begin{definition}[Internal Covariant Hom of Profunctors]
-%%   Let $P : \textrm{Prof}\,D\,E$ and $Q : \textrm{Prof}\,C\,E$. Define
-%%   the covariant Hom $P \triangleright Q : \textrm{Prof}\,C\,D$ by
-%%   \[ (P \triangleright Q)_1 = \lambda (\alpha;\beta).
-\equalizeVofWbyXeqYatZ \phi {\homrallXYtoZ \gamma {\harrapp {P_1} \alpha \gamma} {\harrapp {Q_1} \beta \gamma}} {\psi_{dom}} {\psi_{cod}} {\homrallXYtoZ {\gamma'} {\harrapp {P_1 }\beta {\gamma'}} {\homrallXYtoZ {\gamma} {\harrapp {E_1} {\gamma'} \gamma} {\harrapp {Q_1} \alpha \gamma}}}\]
-%%   where $\psi_{dom}$ and $\chi_{cod}$ are  as
-%%   \[ \psi_{dom} =\homrlambdaXatYdotZ p {\gamma'} {\homrlambdaXatYdotZ e \gamma
-%%     {\homrappXtoYatZ {\phi} {(\apptrinary {P_t} \beta {\pendappXtoY {D_r} \beta} \beta p {\gamma'} e \gamma)} \gamma}}
-%%   \]
-%%   \[
-%%   \psi_{cod} = \homrlambdaXatYdotZ p {\gamma'} {\homrlambdaXatYdotZ e \gamma
-%%     {\apptrinary {Q_t} \alpha {\pendappXtoY {C_r} \alpha} \alpha {(\homrappXtoYatZ \phi {p} {\gamma'})} {\gamma'} e \gamma}}
-%%   \]
-%% \end{definition}
-
-%% We can easily define a contravariant hom $P \triangleleft Q$ by a dual
-%% definition (TODO).
-
-%% We can also define a left Kan extensions using the contravariant
-%% cotensors.  Normally the left Kan extension is defined using tensors,
-%% but the definition using profunctors doesn't look right to me. Let's
-%% work backwards from the definition of colimit to see why.
-
-%% Let $\alpha : \cat C \vdash A : \cat D$ be a functor we want the
-%% colimit of. The UP of a colimit is that maps out of it are equivalent
-%% to cocones from $A$. Semi-formally this looks like
-%% \[ (\colim_{\alpha} A \to_{\cat D} \beta) \cong \forall \alpha:\cat C.~ A \to_{\cat D} \beta \]
-%% This looks a lot like a cotensor with a trivial domain:
-%% \[ (\colim_{\alpha} A \to_{\cat D} \beta) \cong \forall \alpha. (() \to_{1} \gamma) \triangleright (A \to_{\cat D} \beta) \]
-%% Suggesting that we define the left Kan extension profunctor as a
-%% contravariant cotensor as follows.
-%% \begin{definition}
-%%   Let $\cat C,\cat C',\cat D$ be cats and let $\alpha:\cat C \vdash A
-%%   : \cat D$ and $\alpha : \cat C \vdash J : \cat C'$ be functors.
-%%   Then the left Kan extension profunctor $\gamma^o:\cat C'; \beta:\cat
-%%   D \pipe (PLan_J A)[\gamma;\beta]$ is defined as
-%%   \[
-%%   \gamma^o:\cat C'; \beta:\cat D \pipe (PLan_J A)[\gamma;\beta]
-%%   := \forall \alpha : \cat C.~ (J \to_{\cat C'} \gamma) \triangleright (A \to_{\cat D} \beta)
-%%   \]
-%%   Then a left Kan extension of $A$ along $J$ is a functor $\gamma :
-%%   \cat C' \vdash Lan_J A : \cat D$ that represents $PLan_J A$ in that
-%%   \[ (Lan_J A \to_{\cat D} \beta) \cong (LPan_J A)[\gamma;\beta] \]
-%% \end{definition}
-
-%% \begin{definition}
-  %% Let $\cat C,\cat C',\cat D$ be cats and let $\alpha:\cat C \vdash
-  %% : \cat D$ and $\alpha : \cat C \vdash J : \cat C'$ be functors.
-  %% Then the left Kan extension profunctor $\beta^o:\cat D; \gamma:\cat
-  %% C' \pipe (PLan_J F)[\beta;\gamma]$ is defined as
-%%   \[
-%%   \beta^o:\cat D; \gamma:\cat C' \pipe (PLan_J A)[\beta;\gamma] :=
-%%   \exists \alpha:\cat C. (\beta \to_{\cat D} A[\alpha]) \odot (J[\alpha] \to_{\cat C'} \gamma)
-%%   \]
-%%   And a left Kan extension is a functor $\gamma :\cat C' \vdash Lan_J
-%%   A : \cat D$ that represents $PLan_J A$ in that
-%%   \[ (\beta \to_{\cat D} Lan_J A[\alpha]) \cong (PLan_J A)[\beta;\alpha] \]
-%% \end{definition}
-
-%% \begin{lemma}[Parameterized Yoneda Lemma]
-%%   Given categories $\cat C, \cat D$ and a profunctor $\alpha^o:\cat
-%%   C;\beta : \cat D \vdash \alpha \to_{Q} \beta$ and a function
-%%   $\beta : \cat D \vdash G[\beta] : \cat C$, then there is an
-%%   isomorphism between
-%%   \begin{enumerate}
-%%   \item Terms $\cdot \vdash t : G[\beta] \to_{Q} \beta$
-%%   \item left-homomorphisms $\phi : \alpha \to_{\cat C} G[\beta] \vdash
-%%     u : \alpha \to_{Q} \beta$ i.e. they satisfy homomorphism on the
-%%     left (the other side doesn't make sense if $G$ is not a functor):
-%%     $u[\phi \circ \psi] = u[\phi] \circ \psi$
-%%   \end{enumerate}
-%% \end{lemma}
-%% \begin{proof}
-%%   First, the constructions
-%%   \begin{enumerate}
-%%   \item Given $t$, we define $u(t)$ to be
-%%     \[\phi : \alpha \to_{\cat C} G[\beta] \vdash t \circ_{Q} \phi : \alpha \to_{Q} \beta\]
-
-%%     Which is a left-homomorphism because:
-%%     \[ (t \circ_{Q} \phi) \circ_{Q} \psi = t \circ_{Q} (\phi \circ_{\cat C} \psi)\]
-%%   \item Given $u$, we define $t(u)$ to be
-%%     \[ \cdot \vdash u[\id_{G[\beta]}/\phi] : G[\beta] \to_{Q} \beta \]
-%%   \end{enumerate}
-%%   Now we show they are inverse
-%%   \begin{enumerate}
-%%   \item First, the very easy case, $\cdot \vdash t(u(t)) : G[\beta] \to_{Q} \beta$:
-%%     \begin{align*}
-%%       t \circ_{Q} \id_{G[\beta]} &= t
-%%     \end{align*}
-%%     by the fact that $Q$ is a profunctor.
-%%   \item Next, $\phi : \alpha \to_{\cat C} G[\beta] \vdash u(t(u)) : \alpha \to_{Q} \beta$:
-%%     \begin{align*}
-%%       u[\id_{G[\beta]}/\phi] \circ_{Q} \phi &= u[\id \circ_{\cat C} \phi/\phi]\\
-%%       &= u[\phi/\phi]\\
-%%       &= u
-%%     \end{align*}
-%%     which uses the fact that $u$ is a left-homomorphism.
-%%   \end{enumerate}
-%% \end{proof}
-
-%% \begin{theorem}[Negative Types and Right Representables are Equivalent]
-%%   For any profunctor $\alpha^o: \cat C; \beta : \cat D \pipe \alpha
-%%   \to_{Q} \beta$, there is a bijection between
-%%   \begin{enumerate}
-%%   \item Right Represntables $\beta : \cat D \vdash G[\beta] : \cat C$ for $Q$.
-%%   \item Negative Types $\beta : \cat D \vdash N[\beta] : \cat C$
-%%     representing $Q$.
-%%   \end{enumerate}
-%% \end{theorem}
-%% \begin{proof}
-%%   \begin{enumerate}
-%%   \item Given a right representable $G[\beta]$, we set $N(G)[\beta] =
-%%     G[\beta]$, the introduction rule is the same and the elimination
-%%     rule is given as in the Yoneda lemma and the inverse property
-%%     follows by the Yoneda isomorphism and the fact that $I_G^{-1}$ is
-%%     an inverse for $I_G$.
-%%   \item Given a negative type $N[\beta]$ we set $G(N)[\beta] =
-%%     N[\beta]$, the introduction rule is the same and the inverse
-%%     introduction rule is given as in the Yoneda lemma.  We need to
-%%     show that $N[\beta]$ has the structure of a functor and that
-%%     $I_{N}$ and $I_{N}^{-1}$ are homomorphisms on the right with
-%%     respect to that structure.  Define $N_1$ as follows:
-%%     \begin{mathpar}
-%%       \inferrule
-%%           {\beta_1^o:\cat D;\beta_2:\cat D \vdash \psi \circ \epsilon : N[\beta_1/\beta] \to_{Q} \beta_2\\
-%%             \phi : N[\beta_1/\beta] \to_{Q} \beta_2 \vdash I_N[\phi] : N[\beta_1/\beta] \to_{\cat C} N[\beta_2/\beta]
-%%           }
-%%           {\beta_1^o:\cat D;\beta_2:\cat D\pipe \psi : \beta_1 \to_{\cat D} \beta_2 \vdash I_N[\psi \circ \epsilon_{\beta_1}/\phi] : N[\beta_1/\beta] \to_{\cat C} N[\beta_2/\beta]}
-%%     \end{mathpar}
-%%     We need to show functoriality.
-%%     For identity we have
-%%     \begin{align*}
-%%       I_N[\id_{\beta} \circ \epsilon_{\beta}/\phi] &= I_N[\epsilon_{\beta}]\\
-%%       &= I_N[\epsilon \circ \id_{N[\beta]}]\\
-%%       &= \id_{N[\beta]} \tag{$\eta$}
-%%     \end{align*}
-%%     For composition,
-%%     \begin{align*}
-%%       N_1[\theta] \circ N_1[\psi]
-%%       &= I_N[\theta \circ \epsilon] \circ I_N[\psi \circ \epsilon]\\
-%%       &= I_N[(\theta \circ \epsilon) \circ I_N[\psi \circ \epsilon]]\tag{left homomorphism}\\
-%%       &= I_N[\theta \circ (\epsilon \circ I_N[\psi \circ \epsilon])]\\
-%%       &= I_N[\theta \circ (\psi \circ \epsilon)] \tag{$\beta$}\\
-%%       &= I_N[(\theta \circ \psi) \circ \epsilon]\\
-%%       &= N_1[\theta \circ \psi]
-%%     \end{align*}
-%%     Next we show $I_N$ is a right homomorphism, the typing here is
-%%     $\phi : \alpha \to_{Q} \beta_1, \theta : \beta_1 \to_{\cat D}
-%%     \beta_2$.
-%%     \begin{align*}
-%%       N_1[\theta] \circ I_N[\phi] &= I_N[\theta \circ \epsilon] \circ I_N[\phi]\\
-%%       &= I_N[\theta \circ \epsilon \circ I_N[\phi]]\tag{left homomorphism}\\
-%%       &= I_N[\theta \circ \phi] \tag{$\beta$}
-%%     \end{align*}
-%%     and to show $I_N^{-1}$ is a right homomorphism under $\phi : \alpha \to_{\cat C}N[\beta_1], \theta : \beta_1 \to_{\cat D} \beta_2$
-%%     \begin{align*}
-%%       I_N^{-1}[N_1[\theta] \circ \phi] &= \epsilon \circ (N_1[\theta] \circ \phi) \tag{def}\\
-%%       &= \epsilon \circ (I_N[\theta \circ \epsilon] \circ \phi)\tag{def}\\
-%%       &= (\epsilon \circ I_N[\theta \circ \epsilon]) \circ \phi\tag{assoc}\\
-%%       &= (\theta \circ \epsilon) \circ \phi\tag{$\beta$}\\
-%%       &= \theta \circ (\epsilon \circ \phi)\tag{assoc}\\
-%%       &= \theta \circ I_N^{-1}[\phi] \tag{def}
-%%     \end{align*}
-%%   \end{enumerate}
-
-%%   Next we need to show that this is a \emph{bijection}.
-%%   \begin{enumerate}
-%%   \item Starting with a negative type $N[\beta]$, we recover the
-%%     original $\epsilon$ by the Yoneda lemma.
-%%   \item Starting with a right representable $G[\beta]$, we recover the
-%%     original $I_N^{-1}$ by the Yoneda lemma. We need to show that the
-%%     action of $G$ on arrows is the same as the one defined from $G$
-%%     viewed as a negative type, which we will call $G_1'$:
-%%     \begin{align*}
-%%       G_1'[\theta] &= I_G[\theta \circ \epsilon]\\
-%%       &= G_1[\theta] \circ I_G[\epsilon]\tag{right homomorphism}\\
-%%       &= G_1[\theta] \circ I_G[\epsilon \circ \id]\\
-%%       &= G_1[\theta] \circ \id\\
-%%       &= G_1[\theta]
-%%     \end{align*}
-%%   \end{enumerate}
-%% \end{proof}
-
-%% \subsection{Adjunctions by Universal Morphism}
-
-%% As an example of formal category theory in our syntax, let's prove the
-%% equivalence between two definitions of an adjunction: one by universal
-%% morphism and the other by natural hom isomorphism.
-%% %
-%% Though the definitions and proof will look like we are doing ordinary
-%% category theory, since it is syntactic it immediately applies to all
-%% models, including enriched and internal category theory.
-
-%% We start with the usual definition of an adjunction using
-%% hom-isomorphism and a profunctor.
-%% \begin{definition}
-%%   Let $C,D$ be categories and $c^o:C_0,d:D_0 \vdash H[c,d]$ a
-%%   profunctor between them. A \emph{left adjoint} to $H$ consists of
-%%   \begin{enumerate}
-%%   \item A functor $c:C \vdash F[c] : D$
-%%   \item A natural isomorphism $c,d \pipe h : H[c,d] \vdash
-%%     F_{ind}[h]: D_1[F_0[c],d]$. Call the inverse $g. F_{ind}^{-1}[g]$.
-%%   \end{enumerate}
-%% \end{definition}
-
-%% \begin{definition}
-%%   Let $c:C_0,d:D_0 \vdash H[c,d]$ be a profunctor.
-%%   A \emph{universal $H$-heteromorphism from $C$} consists of
-%%   \begin{enumerate}
-%%   \item For every $c:C_0$, an $H$-morphism $\eta_c : H(c,F_0[c])$ for
-%%     some function $c:C_0 \vdash F_0[c] : D_0$.
-%%   \item Such that every $h : H[c,d]$ factors through $\eta_c$ in that
-%%     there is a unique $h : H[c,d] \vdash F_{ind}[h] : D[F_0[c],d]$
-%%     such that $h = F_{ind}[h] \circ \eta_c$.  Uniqueness means
-%%     validity of the following rule:
-%%     \begin{mathpar}
-%%       \inferrule
-%%       {g:D[F_0[c],d] \vdash g \circ \eta = h}
-%%       {g:D[F_0[c],d] \vdash g = F_{ind}[h]}
-%%     \end{mathpar}
-%%     (Note that we do not require $F_{ind}$ to be natural in any sense)
-%%   \end{enumerate}
-%% \end{definition}
-
-%% It is easy to see that a left adjoint gives a universal heteromorphism
-%% by a Yoneda-like argument:
-%% \begin{construction}
-%%   Let $F$ be a left adjoint to $H$. Then we define a left universal
-%%   $H$-heteromorphism by
-%%   \begin{enumerate}
-%%   \item $c \vdash \eta_c = F_{ind}^{-1}[\id[F_0[c]]]: H[c,F_0[c]]$
-%%   \item Every $h : H[c,d]$ factorizes through $\eta$ because
-%%     \begin{align*}
-%%       F_{ind}[h] \circ F_{ind}^{-1}[\id] &= F_{ind}^{-1}[F_{ind}[h] \circ \id] \tag{naturality}\\
-%%       &= F_{ind}^{-1}[F_{ind}[h]] \tag{unit}\\
-%%       &= h \tag{isomorphism}
-%%     \end{align*}
-%%     This factorization is unique because if $g \circ F^{-1}[\id] = h$,
-%%     then
-%%     \begin{align*}
-%%       F_{ind}[h] &= F_{ind}[g \circ F^{-1}[\id]] \tag{congruence}\\
-%%       &= g \circ F_{ind}[F^{-1}[\id]] \tag{naturality}\\
-%%       &= g \circ \id \tag{iso}\\
-%%       &= g \tag{unit}
-%%     \end{align*}
-%%   \end{enumerate}
-%% \end{construction}
-
-%% But the utility of the universal morphism definition is that the
-%% universal morphism extracted from a left adjoint is enough to uniquely
-%% reconstruct the rest of the structure, which includes (1)
-%% functoriality of $F$ and (2) naturality of $F_{ind}$.
-%% \begin{construction}
-%%   Let $(F_0,\eta,F_{ind}[h])$ be a left universal $H$-morphism.
-%%   We extend this to a left adjoint as follows.
-%%   \begin{enumerate}
-%%   \item First, $F$ acts on morphisms by
-%%     \begin{mathpar}
-%%       \inferrule
-%%       {
-%%         \inferrule
-%%          {f : C[c,c'] \and \eta_{c'} : H[c',F_0[c']]}
-%%          {f : C[c,c'] \vdash \eta_{c'} \circ f : H[c,F_0[c']]}
-%%       }
-%%       {f : C[c,c'] \vdash F_{ind}[\eta_{c'} \circ f] : D[F_0[c],F_0[c']]}
-%%     \end{mathpar}
-%%     This preserves identity
-%%     \begin{mathpar}
-%%       \inferrule
-%%       {
-%%         \inferrule
-%%         {\id \circ \eta = \eta}
-%%         {\id = F_{ind}[\eta]}
-%%       }
-%%       {\id = F_{ind}[\eta \circ \id]}
-%%     \end{mathpar}
-%%     And composition
-%%     \begin{mathpar}
-%%       \inferrule
-%%       {F_{ind}[\eta \circ f] \circ F_{ind}[\eta \circ g] \circ \eta = \eta \circ f \circ g}
-%%       {F_{ind}[\eta \circ f] \circ F_{ind}[\eta \circ g] = F_{ind}[\eta \circ f \circ g]}
-%%     \end{mathpar}
-%%     Which follows by a couple uses of the factorization:
-%%     \begin{align*}
-%%       F_{ind}[\eta \circ f] \circ F_{ind}[\eta \circ g] \circ \eta
-%%       &= F_{ind}[\eta \circ f] \circ \eta \circ g \\
-%%       &= \eta \circ f \circ g
-%%     \end{align*}
-%%   \item Next, we show that $h : H[c,d] \vdash F_{ind}[h] :
-%%     D[F_0[c],d]$ is a natural transformation. As a first step, we show
-%%     that $\eta_c : H[c,F_0[c]]$ is natural in $c$ with the action of
-%%     $F$ on morphisms we have just defined.
-%%     \[ \eta \circ f = F[f] \circ \eta \]
-%%     expanding the definition, this is precisely the factorization
-%%     property:
-%%     \[ \eta \circ f = F_{ind}[\eta \circ f] \circ \eta \]
-
-%%     Next, to show that $F_{ind}$ is natural, we need to show that
-%%     (when it is well-defined):
-%%     \[F_{ind}[g \circ h \circ f] = g \circ F_{ind}[h] \circ F[f] \]
-%%     By the factorization property, the following sufficeth:
-%%     \begin{align*}
-%%      g \circ F_{ind}[h] \circ F[f] \circ \eta
-%%      &= g \circ F_{ind}[h] \circ \eta \circ f \tag{naturality of $\eta$}\\
-%%      &= g \circ h \circ f \tag{factorization}
-%%     \end{align*}
-%%   \item Finally, we define $F_{ind}^{-1}[g]$ as
-%%     \[ g : D[F_0[c],d] \vdash F_{ind}^{-1}[g] = g \circ \eta : H[c,d] \]
-%%     And we need to confirm that it is (1) a section of $F_{ind}$ (2) a
-%%     retraction of $F_{ind}$ and (3) natural.
-%%     \begin{enumerate}
-%%     \item For retraction we need to show $h \vdash F_{ind}[h] \circ
-%%       \eta = h$ which is precisely the factorization property.
-%%     \item For section, we need to show $g \vdash F_{ind}[g \circ \eta]
-%%       = g$. Immediately the uniquenes principle says its sufficient to
-%%       show $g \circ \eta = g \circ \eta$.
-%%     \item For naturality, we need to show
-%%       \[ F_{ind}^{-1}[k \circ g \circ F[f]] = k \circ F_{ind}^{-1}[g] \circ f \]
-%%       expanding the definition this is just:
-%%       \[ k \circ g \circ F[f] \circ \eta = k \circ g \circ \eta \circ f \]
-%%       which is just naturality of $\eta$.
-%%     \end{enumerate}
-%%   \end{enumerate}
-%% \end{construction}
-
-\subsection{Products and Units of Internal Categories}
-
-%% While we have now proved a beautiful internal theorem in great
-%% generality that has many useful instances in different models, we are
-%% hard pressed to actually \emph{instantiate} the theorem in the syntax
-%% itself because we can't really come up with any useful universal properties!
-%% The reason is that Judgment Theory itself is too bare-bones.
-
-%% For instance how would we say that a category $\cat C$ has a
-%% \emph{terminal object}? Well, it would have to be in the first place a
-%% \emph{single} object in $\cat C$, but so far our syntax only enables
-%% us to talk about functions and functors. Well an object of $\cat C$ is
-%% the same as a function from the unit sort, but so far our syntax only
-%% has base sorts.
-
-%% So we see that defining the terminal object, \emph{necessitates} some
-%% notion of terminal object in our ``metatheory'': Judgment Theory.
-%% This mysterious and pervasive aspect of higher category theory is
-%% called the \emph{microcosm principle} because it says that in order to
-%% discuss a property of a single object in a single category (the
-%% microcosm), we need the same sort of object at the meta-level (the
-%% macrocosm). The adage here is ``As above, so below''.
-
-%% That might all sound very spiritual, but we'll see that it's actually
-%% quite a nice heuristic for formalizing universal properties in
-%% internal category theory.
-%% %
-%% In the spirit of this heuristic, instead of first giving a bunch of
-%% new type constructions for Judgment Theory and then applying them to
-%% define universal properties, we will consider them in pairs, first
-%% above, and then below.
-
-First we consider the terminal sort and then the idea of a terminal
-object in an internal category.
-
-Next, the terminal sort trivially has a category structure using the
-trivial judgment as its hom set:
-
-\begin{definition}[Terminal Category]
-  The terminal category $\cat 1$ is defined as having
-  \begin{enumerate}
-  \item Its object sort as $1$
-  \item Its hom set as $\lambda \alpha:1;\beta:1. 1 : \harr 1 1$
-  \item Its identity arrow is given by $\lambda \alpha: 1. () : 1$
-  \item Composition is defined by $\phi : 1, \psi : 1 \vdash () : 1$
-  \item Associativity and unitality are trivial by \textsc{1-$\eta$}
-  \end{enumerate}
-\end{definition}
-
-\begin{definition}[Terminal Object Specification]
-  In judgment type theory with a terminal sort/judgment, we can define
-  for any category $\cat C$ a profunctor $\alpha^o:\cat C;\beta : \cat
-  1 \pipe 1$, which specifies a terminal object.
-\end{definition}
-
-Now let's instantiate the definition for the terminal object defined
-as a negative type and as a representable and see how it reproduces
-the standard type-theoretic and category-theoretic definitions.
-
-A representation of $\alpha^o:\cat C; \beta : \cat 1 \pipe 1$ consists
-of a functor $\beta : \cat 1 \vdash 1_{\cat C} : \cat C$, an introduction rule
-\[ \phi : 1 \vdash I_{1}[\phi] : \alpha \to_{\cat C} 1_{\cat C}[\beta] \]
-satisfying $I_{1}[\phi] \circ \gamma =I_{1}[\phi \circ \gamma]$ and an
-inverse $\psi : \alpha \to_{\cat C} 1_{\cat C}[\beta] \vdash
-I_{1}^{-1}[\psi] : 1$ which by \textsc{1-$\eta$} is equal to $()$ and
-therefore every $t : \alpha \to_{\cat C} 1_{\cat C}[\beta]$ is equal
-to $I_{1}[\phi]$. In light of this it would be appropriate to write
-$I_1[\phi]$ as $()$ and then the left-homomorphism property is $()
-\circ \gamma = ()$, which is precisely the definition of substitution.
-
-This alternate syntax makes even more sense in light of the negative
-type presentation. A negative type satisfying the terminal object
-specification consists of a functor $1_{\cat C}$ as above and an
-introduction rule as above and an elimination form $\vdash \epsilon :
-1$, which is trivial.
-
-First the product category
-\begin{definition}[Product category]
-  For any categories $\cat C_1$, $\cat C_2$ we can form the product
-  category whose sort of objects is $\cat C_1 \times \cat C_2$, whose
-  arrow judgment is $\alpha^o : \cat C_1 \times \cat C_2;\alpha \pipe
-  (\pi_1 \alpha \to_{\cat C_1} \pi_1 \alpha) \times (\pi_2 \alpha
-  \to_{\cat C_2} \pi_2 \alpha)$. Identity is given by $\alpha^o : \cat
-  C_1 \times \cat C_2;\alpha\pipe\cdot \vdash (\id_{\pi_1 \alpha},
-  \id_{\pi_2 \alpha}) : (\pi_1 \alpha \to_{\cat C_1} \pi_1 \alpha)
-  \times (\pi_2 \alpha \to_{\cat C_2} \pi_2 \alpha)$ and composition
-  by
-  \[
-  \phi %: (\pi_1 \alpha_0 \to_{\cat C_1} \pi_1 \alpha_1) \times (\pi_2 \alpha_0 \to_{\cat C_2} \pi_2 \alpha_1)
-  ,
-  \psi %: (\pi_1 \alpha_1 \to_{\cat C_1} \pi_1 \alpha_2) \times (\pi_2 \alpha_1 \to_{\cat C_2} \pi_2 \alpha_2)
-  \vdash ((\pi_1 \phi) \circ (\pi_1 \psi), (\pi_2 \psi) \circ (\pi_2 \phi)) :
-  (\pi_1 \alpha_0 \to_{\cat C_1} \pi_1 \alpha_2) \times (\pi_2 \alpha_0 \to_{\cat C_2} \pi_2 \alpha_2)  
-  \]
-  unitality, associativity follow from the same properties of $\cat C_1,\cat C_2$
-\end{definition}
-
-\begin{definition}[Product Functor Specification]
-  For any category $\cat C$, we can specify the product functor by
-  \begin{enumerate}
-  \item The span is $\alpha^o:\cat C; \beta : \cat C \times \cat C \pipe (\alpha \to_{\cat C} \pi_1 \beta) \times (\alpha \to_{\cat C} \pi_2 \beta)$
-  \item Left Composition is defined as
-    \[ \phi, \psi \vdash (\pi_1 \phi \circ \psi, \pi_2 \phi \circ \psi)\]
-  \item Left unitality:
-    \begin{align*}
-      (\pi_1 \phi \circ \id, \pi_2 \phi \circ \id)
-      &= (\pi_1 \phi, \pi_2 \phi)\\
-      &= \phi\tag{$\times-eta$}
-    \end{align*}
-  \item Left associativity:
-    \begin{align*}
-      (\phi \circ \psi) \circ \psi'
-      &= (\pi_1 \phi \circ \psi, \pi_2 \phi \circ \psi) \circ \psi'\\
-      &= (\pi_1 (\pi_1 \phi \circ \psi, \pi_2 \phi \circ \psi) \circ \psi',
-      \pi_2 (\pi_1 \phi \circ \psi, \pi_2 \phi \circ \psi) \circ \psi')\\
-      &= ((\pi_1 \phi \circ \psi) \circ \psi', (\pi_2 \phi \circ \psi) \circ \psi')\\
-      &= (\pi_1 \phi \circ (\psi \circ \psi'), \pi_2 \phi \circ (\psi \circ \psi'))\\
-      &= \phi \circ (\psi \circ \psi')
-    \end{align*}
-  \item Right composition is defined as
-    \[ \phi, \theta \vdash (\pi_1 \theta \circ \pi_1 \phi, \pi_2 \theta \circ \pi_2 \phi) \]
-  \item Right unitality:
-    \begin{align*}
-      (\pi_1 (\id,\id) \circ \pi_1 \phi, \pi_2 (\id,\id) \circ \pi_2 \phi)
-      &= (\id \circ \pi_1 \phi, \id \circ \pi_2 \phi) \tag{$\times-\beta$}\\
-      &= (\pi_1 \phi, \id \circ \pi_2 \phi)\\
-      &= \phi \tag{$\times-\eta$}
-    \end{align*}
-  \item Right associativity
-    \begin{align*}
-      \theta' \circ (\theta \circ \phi)
-      &= \theta' \circ (\pi_1 \theta \circ \pi_1 \phi, \pi_2 \theta \circ \pi_2 \phi)\\
-      &= (\pi_1 \theta' \circ \pi_1(\pi_1 \theta \circ \pi_1 \phi, \pi_2 \theta \circ \pi_2 \phi),
-      \pi_2 \theta' \circ \pi_2(\pi_1 \theta \circ \pi_1 \phi, \pi_2 \theta \circ \pi_2 \phi))\\
-      &= (\pi_1 \theta' \circ (\pi_1 \theta \circ \pi_1 \phi), \pi_2 \theta' \circ (\pi_2 \theta \circ \pi_2 \phi))\\
-      &= ((\pi_1 \theta' \circ \pi_1 \theta) \circ \pi_1 \phi, (\pi_2 \theta' \circ \pi_2 \theta) \circ \pi_2 \phi)\\
-      &= (\pi_1(\theta' \circ \theta) \circ \pi_1 \phi, \pi_2(\theta' \circ \theta) \circ \pi_2 \phi)\\
-      &= (\theta' \circ \theta) \circ \phi \\
-    \end{align*}
-  \end{enumerate}
-\end{definition}
-
-\subsection{Discrete and Codiscrete Categories}
-
-If we have terminal judgments, then we can define the
-\emph{codiscrete} category for any sort $\cat C$.
-\begin{definition}
-  The codiscrete category $\textrm{CoDisc} \cat C$ is defined as
-  \begin{enumerate}
-  \item Object sort is $\cat C$
-  \item Morphisms judgment is $1$
-  \item Identity and composition are given by the unique terms of
-    those types, making associativity and unitality trivial to verify.
-  \end{enumerate}
-\end{definition}
-
-If we have \emph{identity} objects, which we write as $\equiv_A$, then we
-can define discrete categories.
-\begin{definition}
-  The discrete category $\textrm{Disc} \cat C$ is defined as
-  \begin{enumerate}
-  \item Object sort is $\cat C$
-  \item Morphisms judgment is $\alpha;\alpha' \pipe \alpha \equiv \alpha'$
-  \item Identity and composition are defined as in the section on
-    synthetic category theory.
-  \end{enumerate}
-\end{definition}
-
-\subsection{Internal Tensor, Cotensor}
-
-We should be able to define internal \emph{tensor} using synthetic
-tensor and coequalizers and also internal \emph{cotensor} using
-synthetic cotensor and equalizers. These look exactly like the
-set-theoretic definitions.
-\begin{definition}
-  Let $R$ be an internal profunctor from $C$ to $D$ and $S$ be an
-  internal profunctor from $D$ to $E$.
-
-  Then we can define the internal tensor product $R \odot S$, a
-  profunctor from $C$ to $E$ as
-  \begin{enumerate}
-  \item The underlying span can be defined as a quotient of the synthetic tensor product
-    \[ (R \odot S)[\alpha;\gamma] = \exists \beta:\cat D. R \odot S/Rcomp = Scomp \]
-
-    Where R-comp and S-comp have as source the span
-    \[ \exists \beta:\cat D.~ R[\alpha;\beta] \odot (\exists \beta':\cat D. Hom_C(\beta,\beta') \odot S[\beta';\gamma]) \]
-    or the isomorphic
-    \[ \exists \beta':\cat D.~ (\exists \beta:\cat D. R[\alpha;\beta] \odot Hom_C(\beta,\beta')) \odot S[\beta';\gamma] \]
-    where R-comp composes the $C$-morphism with the $R$-morphism and
-    S-comp the $S$-morphism.
-  \item Composition is defined using composition in $R,S$.
-  \end{enumerate}
-\end{definition}
-Note that $R$-comp and $S$-comp are each constructed using the
-
-\begin{definition}
-  Let $R$ be an internal profunctor from $C$ to $D$ and $S$ be an
-  internal profunctor from $C$ to $E$. Then the internal contravariant
-  cotensor product $R \triangleright S$ is defined as
-  \[ (R \triangleright S)[\beta;\gamma] = \{ p : \forall \alpha:\cat C.~ R[\alpha;\beta] \triangleright S[\alpha;\gamma] 
-  \pipe R-comp = S-comp \}  \]
-  R-comp and S-comp have as codomain the span
-  \[ \forall \alpha:\cat C.~ R[\alpha;\beta] \triangleright \forall \alpha':\cat C.~ Hom_C(\alpha',\alpha) \triangleright S[\alpha';\gamma] \]
-  or the isomorphic
-  \[ \forall \alpha':\cat C.~ (\exists \alpha. Hom_C(\alpha',\alpha) \odot R[\alpha;\beta]) \triangleright S[\alpha';\gamma]
-  \]
-\end{definition}
-
-We can also construct the \emph{free profunctor} from a span using tensor:
-
-\begin{definition}
-  Let $\cat C$ and $\cat D$ be internal categories and let $R[\alpha;\beta]$ be a
-  span from $Ob(\cat C)$ to $Ob(\cat D)$. Then the free profunctor
-  extending $R$ to be an internal profunctor from $\cat C$ to $\cat D$ is
-  \[ F_{\cat C,\cat D}R[\alpha';\beta'] = \exists \alpha. Hom_C(\alpha',\alpha) \odot (\exists \beta. R[\alpha;\beta] \odot Hom_C(\beta,\beta')) \]
-  or the isomorphic
-  \[ F_{\cat C,\cat D}R[\alpha';\beta'] = \exists \beta. (\exists \alpha. Hom_C(\alpha',\alpha) \odot R[\alpha;\beta]) \odot Hom_C(\beta,\beta') \]
-\end{definition}
-
-And maybe we can even define the co-free profunctor from a span using cotensor?
-\begin{definition}
-  Let $\cat C$ and $\cat D$ be internal categories and let $R[\alpha;\beta]$ be a
-  span from $Ob(\cat C)$ to $Ob(\cat D)$. Then the co-free profunctor
-  extending $R$ to be an internal profunctor from $\cat C$ to $\cat D$ is
-  \[ G_{\cat C,\cat D}R[\alpha';\beta'] =
-  \forall \alpha. Hom_C(\alpha,\alpha') \triangleright (\forall \beta. R[\alpha;\beta] \triangleleft Hom_D(\beta',\beta)) \]
-  or the isomorphic
-  \[ G_{\cat C,\cat D}R[\alpha';\beta'] =
-  \forall \beta. 
-  (\forall \alpha. Hom_C(\alpha,\alpha') \triangleright R[\alpha;\beta]) \triangleleft Hom_D(\beta',\beta)) \]
-\end{definition}
 
@@
 
-\begin{definition}[Virtual Double Category]
-  A virtual double category $V$ consists of
+%% \begin{definition}[Virtual Double Category]
+%%   A virtual double category $V$ consists of
+%%   \begin{enumerate}
+%%     \item A set $V_0$ of objects.
+%%     \item For each $\cat C, \cat D \in V_0$, a set $V_v(\cat C, \cat
+%%       D)$ of vertical arrows, with composition and unit operations
+%%       satisfying the category axioms.
+%%     \item A set $V_h$ of horizontal arrows, with $d^+,d^- : V_h \to
+%%       V_0$.
+%%     \item For each composable string $\Phi$ of horizontal arrows
+%%       starting at $\cat C$ and ending at $\cat D$, and horizontal
+%%       arrow $R$, a set $V_2(\Phi; R)$ of 2-cells, with projection
+%%       functions $d^{\pm} : V_2(\Phi;R) \to V_v(d^\pm \Phi, d^\pm R)$.
+
+%%     \item For each $R \in V_h$, an identity arrow $\id_R \in
+%%       V_2(d^-R,R,d^+ R; R)$ with $d^\pm \id_R = \id_{d^\pm R}$.
+
+%%     \item For any $2$-cell $\alpha \in V_2(\Phi;R)$ and composable
+%%       string $\phi$ of 2-cells into $\Phi$, a composite $\alpha \circ
+%%       \phi$.
+
+%%     \item Composition should be unital in that:
+%%       \[ \id_R \circ [d^-\alpha, \alpha, d^+\alpha] = \alpha  \]
+%%       and
+%%       \[ \alpha \circ [\vec \id] = \alpha\]
+%%     \item And associative in that
+%%       \[ (\alpha \circ [\beta_1,\ldots]) \circ [\gamma_{11},\ldots,\gamma_{21},\ldots,] = \alpha \circ [(\beta_1 \circ \gamma_{11},\ldots), (\beta_2 \circ \gamma_{21},\ldots)]\]
+%%   \end{enumerate}
+%% \end{definition}
+
+%% \begin{definition}[Restriction]
+%%   Let $R \in V_h$ and $f \in V_v(\cat C, d^-R)$ and $g \in V_v(\cat D,
+%%   d^+R)$, then a restriction of $R$ along $f,g$ is a horizontal arrow
+%%   $R(f,g) \in V_h$ with $d^-R(f,g)= \cat C$ and $d^+R(f,g)= \cat D$
+%%   with a (op?)-cartesian 2-cell $\alpha \in V_2(R(f,g),R)$ with
+%%   $d^-\alpha = f$ and $d^+\alpha = g$.
+%% \end{definition}
+
+%% \begin{definition}[Unit]
+%%   A unit $U_{\cat C}\in V_h$ with $d^\pm U_{\cat C} = \cat C$ and an op-cartesian 2-cell
+%%   \[ r : V_h(\cat C, U_{\cat C}) \]
+%%   with $d^\pm r = \id_{\cat C}$.
+%% \end{definition}
+
+%% \begin{definition}[Tensor]
+%%   If $R,Q \in V_h$ with $d^+R = d^-Q$, a tensor is a horizontal arrow
+%%   $R \odot Q \in V_h$ with $d^+ (R\odot Q) = d^+ R$ and $d^-(R \odot
+%%   Q) = d^- Q$ with an op-cartesian 2-cell
+%%   \[ p : V_h(R,Q;R \odot Q) \]
+%% \end{definition}
+
+\subsection{Category of Models}
+
+Our goal is to establish that the syntax of SCT (with whichever type
+constructors we choose) presents the initial indexed virtual equipment
+(with the appropriate universal properties). To start we define
+appropriate categories of equipments.
+
+\begin{definition}{Indexed Virtual Equipment}
+  A VE-hyperdoctrine is a functor $\cat C^{op} \to VE$, and a morphism
+  of VE-hyperdoctrines is a natural transformation.
+\end{definition}
+
+\begin{definition}[Internal Logic]
+  Given any virtual equipment $\mathcal V$, we can construct the
+  ``internal logic'' by indexing $\mathcal V$ by $\Set$:
+
   \begin{enumerate}
-    \item A set $V_0$ of objects.
-    \item For each $\cat C, \cat D \in V_0$, a set $V_v(\cat C, \cat
-      D)$ of vertical arrows, with composition and unit operations
-      satisfying the category axioms.
-    \item A set $V_h$ of horizontal arrows, with $d^+,d^- : V_h \to
-      V_0$.
-    \item For each composable string $\Phi$ of horizontal arrows
-      starting at $\cat C$ and ending at $\cat D$, and horizontal
-      arrow $R$, a set $V_2(\Phi; R)$ of 2-cells, with projection
-      functions $d^{\pm} : V_2(\Phi;R) \to V_v(d^\pm \Phi, d^\pm R)$.
-
-    \item For each $R \in V_h$, an identity arrow $\id_R \in
-      V_2(d^-R,R,d^+ R; R)$ with $d^\pm \id_R = \id_{d^\pm R}$.
-
-    \item For any $2$-cell $\alpha \in V_2(\Phi;R)$ and composable
-      string $\phi$ of 2-cells into $\Phi$, a composite $\alpha \circ
-      \phi$.
-
-    \item Composition should be unital in that:
-      \[ \id_R \circ [d^-\alpha, \alpha, d^+\alpha] = \alpha  \]
-      and
-      \[ \alpha \circ [\vec \id] = \alpha\]
-    \item And associative in that
-      \[ (\alpha \circ [\beta_1,\ldots]) \circ [\gamma_{11},\ldots,\gamma_{21},\ldots,] = \alpha \circ [(\beta_1 \circ \gamma_{11},\ldots), (\beta_2 \circ \gamma_{21},\ldots)]\]
+  \item The indexing cateogry is $\Set$.
+  \item Given $\Gamma \in \Set$, the virtual equipment $\mathcal
+    V^\Gamma$ is the ``power'' of $\mathcal V$ by $\Gamma$:
+    \begin{enumerate}
+    \item Objects are functions $\Gamma \to \mathcal V_0$.
+    \item A vertical morphism from $O$ to $O'$ is a function $\Pi_{\gamma\in\Gamma} V_v(O\gamma,O'\gamma)$.
+    \item A horizontal morphism is a function $\Gamma \to \mathcal V_h$. This satisfies
+    \item A 2-cell from $R$ to $S$ is a family $\Pi_{\gamma\in\Gamma} \mathcal V_h(R\gamma,S\gamma)$.
+    \end{enumerate}
+    Compositions and restrictions are given pointwise.
+  \item Restriction by a function $\gamma : \Delta \to \Gamma$ is
+    given pointwise.
   \end{enumerate}
 \end{definition}
-
-\begin{definition}[Restriction]
-  Let $R \in V_h$ and $f \in V_v(\cat C, d^-R)$ and $g \in V_v(\cat D,
-  d^+R)$, then a restriction of $R$ along $f,g$ is a horizontal arrow
-  $R(f,g) \in V_h$ with $d^-R(f,g)= \cat C$ and $d^+R(f,g)= \cat D$
-  with a (op?)-cartesian 2-cell $\alpha \in V_2(R(f,g),R)$ with
-  $d^-\alpha = f$ and $d^+\alpha = g$.
-\end{definition}
-
-\begin{definition}[Unit]
-  A unit $U_{\cat C}\in V_h$ with $d^\pm U_{\cat C} = \cat C$ and an op-cartesian 2-cell
-  \[ r : V_h(\cat C, U_{\cat C}) \]
-  with $d^\pm r = \id_{\cat C}$.
-\end{definition}
-
-\begin{definition}[Tensor]
-  If $R,Q \in V_h$ with $d^+R = d^-Q$, a tensor is a horizontal arrow
-  $R \odot Q \in V_h$ with $d^+ (R\odot Q) = d^+ R$ and $d^-(R \odot
-  Q) = d^- Q$ with an op-cartesian 2-cell
-  \[ p : V_h(R,Q;R \odot Q) \]
-\end{definition}
-
-\begin{definition}[Hom]
-
-\end{definition}
-
-\subsection{Category of Signatures and Models}
-
-Our goal is to establish that the syntax of SCT (with whichever type
-constructors we choose) presents a free virtual double category (with
-the appropriate universal properties). To start we define appropriate
-categories of signatures and double categories.
-
-\begin{definition}
-  Define the category Sig to have as objects well-formed contexts
-  $\Gamma$ and morphisms $\gamma : \Delta \to \Gamma$ morphisms of
-  contexts, with the composition as substitution and identity as a
-  context morphism of all variable references.
-\end{definition}
-
-Second, there is a category of virtual double categories and functors.
-
-\begin{definition}
-  A morphism of virtual equipments with blah is a functor of virtual
-  double categories that strictly preserves the universal properties.
-\end{definition}
-
-Next, we define a profunctor between the category of signatures and
-the category of virtual equipments.
 
@@
 
-\paragraph{Semantic extensions}
-
-\paragraph{Extensions}
+\paragraph{Category Theoretic Extensions}
 
@@
 \paragraph{As a tool for Formal Type Theory}
+
+While we can model ``unary'' type theories using profunctors, to model
+theories like simple intuitionistic, linear or bunched type theories
+we need a way to talk about ``profunctors of arbitrarily many
+arguments''. This can be accomplished using \emph{multicategory}
+theory, which models the structure of the context with its structural
+rules as a monad of ``contexts and substitutions''. A suitable notion
+of monad on a virtual double category has been developed and was in
+fact the original motivation for the concept of a virtual double
+category. However while an abstract monad of contexts $T$ is enough to
+define what a $T$-multicategory internal to a virtual equipment is, it
+does not appear possible to define the multiplicative connectives as a
+universal property, i.e., to define when the multicategory is
+\emph{representable}.
+
+A recent development in the formalization of substructural type
+theories is to use monoidal connectives at the meta-level when
+defining multiplicative connectives\cite{woodsatkeyesop2022}. This
+suggest one approach to extending the theory to support the
+formalization of multiplicative universal properties would be, in
+adherence to the microcosm principle, to add linear judgments and
+multiplicative connectives to the metalevel.
 
@@
 
-\appendix
-
-TODO: complete substitution
-\begin{figure}
-  \begin{mathpar}
-      (\homrallXYtoZ \alpha R P)[\xi] = \homrallXYtoZ \alpha {R[d^+\xi;\alpha/\alpha]} {P[\alpha/\alpha;d^-\xi]}
-
-     (\homrlambdaXatYdotZ {x} {\alpha} M)[\phi]
-      = \homrlambdaXatYdotZ {x} {\alpha} M[\phi,x/x,\gamma/\gamma]
-
-     (\homllambdaXatYdotZ {x} {\alpha} M)[\phi]
-      = \homllambdaXatYdotZ {x} {\alpha} M[\phi,x/x,\gamma/\gamma]
-
-  \end{mathpar}
-  \caption{Substitution}
-
-\end{figure}
-
-\begin{figure}
-  \begin{mathpar}
-    \inferrule
-    {}
-    {\Gamma\pipe \cdot \vdash \Sort \isaSet}
-
-    \inferrule
-    {x:S \in \Gamma}
-    {\Gamma \pipe \cdot \vdash x : S}
-
-    \inferrule
-    {\Gamma\pipe\cdot \vdash C : \Sort}
-    {\Gamma \pipe \alpha:\cat C \vdash \alpha : \cat C}
-
-    \inferrule
-    {\Gamma \pipe \alpha:C,\beta:D \vdash R \isaSet}
-    {\Gamma\pipe \alpha:C, x:R, \beta: D \vdash x : R}
-
-    \inferrule{\forall x:A \in \Gamma. \Delta \vdash \gamma(x) : A}{\gamma : \Delta \to \Gamma}\\
-
-    \inferrule*%% [right=NullaryBoundaryMap]
-               {\Gamma \isadtctx}{\Gamma \vdash \cdot : \cdot \to \cdot}\and
-    \inferrule*{\Gamma \pipe \alpha':C' \vdash A : C}{\Gamma \vdash A/\alpha : \alpha':C' \to \alpha:C}
-
-    \inferrule*
-    {\Gamma \pipe \alpha':C' \vdash A : C\and
-      \Gamma \pipe \beta':D' \vdash B : D
-    }
-    {\Gamma \vdash (A/\alpha;B/\beta) : (\alpha':C';\beta':D') \to (\alpha:C;\beta':D')}\\
-
-    \inferrule*%% [right=NullCtxMap]
-               {\Gamma \isadtctx}{\Gamma \vdash \cdot : \cdot \to \cdot}
-
-    \inferrule*{\Gamma \pipe \alpha':C' \vdash A : C}{\Gamma \vdash A/\alpha : \alpha':C' \to \alpha:C}
-
-    \inferrule
-    {\Psi = \Psi_1 \jnctx \Psi_M \jnctx \Psi_2\and
-      \Gamma \vdash \phi_1 : \Psi_1 \to \Phi_1\and
-      \Gamma \pipe \Psi_M \vdash M : R[d^+\phi_1;d^-\phi_2] \and
-      \Gamma \vdash \phi_2 : \Psi_2 \to \Phi_2
-      }
-    {\Gamma \vdash \phi_1,M/x,\phi_2 : \Psi \to \Phi_1,x:R,\Phi_2}
-  \end{mathpar}
-  \caption{Substitutions in SCT}
-  \label{fig:basic-rules}
-\end{figure}
-
-\begin{figure}
-  \begin{mathpar}
-    \inferrule
-    {}
-    {\cdot : \cdot \to \cdot}
-
-    \inferrule
-    {\gamma : \Delta \to \Gamma \and \Delta \vdash M : A[\gamma]}
-    {\gamma,M/X : \Delta \to \Gamma,X:A}
-
-    \inferrule
-    {\Gamma \vdash \Phi \isavectx \and
-    \Gamma \vdash \Psi \isavectx \and
-    d^+\Phi = d^- \Psi}
-    {\Gamma \vdash \Phi \jnctx \Psi \isavectx}
-
-    \inferrule
-    {\Gamma \vdash \phi : \Phi' \to \Phi \and
-     \Gamma \vdash \psi : \Psi' \to \Psi\and
-     d^+\phi = d^- \psi}
-    {\Gamma \vdash \phi \jnctx \psi : \Phi' \jnctx \Psi' \to \Phi \jnctx \Psi}
-
-    \inferrule*[right=ObSubst]
-    {\alpha : \cat C \vdash B : \cat D\and
-    \beta : \cat D \vdash A : \cat E}
-    {\alpha : \cat C \vdash A[B/\beta] : \cat E}\\
-
-    \inferrule*[right=Everything Subst]{\gamma : \Delta \to \Gamma \and \Gamma J}{\Delta J}
-
-    \inferrule{\Gamma \vdash \xi : Z \to \Xi\and \Gamma \pipe \Xi \vdash R \isaSet}{\Gamma \pipe Z \vdash R[\xi] \isaSet}
-
-    \inferrule{\Gamma \vdash \phi : \Psi \to \Phi\and \Gamma \pipe \Phi \vdash M : R}{\Gamma \pipe \Psi \vdash M[\phi] : R[\underline \phi]}
-  \end{mathpar}
-  \begin{align*}
-    (\Phi \jnctx \Psi) \jnctx \Sigma &= \Phi \jnctx (\Psi \jnctx \Sigma)\\
-    \Phi \jnctx \alpha:C &= \Phi\\
-    \alpha:C \jnctx \Psi &= \Psi\\\\
-    d^\pm(A/\alpha) &= A/\alpha\\
-    d^+(\phi_1,M/x,\phi_2) &= d^+ \phi_2\\
-    d^-(\phi_1,M/x,\phi_2) &= d^- \phi_1\\\\
-    \underline \cdot &= \cdot\\
-    \underline {A/\alpha} &= A/\alpha\\
-    \underline {\phi_1,M/x,\phi_2} &= d^-\phi_1;d^+\phi_2
-  \end{align*}
-  \caption{Admissible Rules}
-\end{figure}
-
-\begin{figure}
-  \begin{mathpar}
-    \begin{array}{rcl}
-      x[\gamma] &=& \gamma(x)\\
-      J[\gamma][\delta] & = & J[\gamma[\delta]]\\
-      \gamma[\delta](x) &=& \gamma(x)[\delta]\\
-
-      A[\alpha/\alpha] &=& A\\
-      \alpha[A/\alpha] &=& A\\
-      A[B/\beta][C/\gamma] & = & A[B[C/\gamma]/\beta]\\
-      A[B/\beta][\gamma] &=& A[\gamma][B[\gamma]/\beta]\\\\
-
-      R[\cdot] &=& R\\
-      R[\alpha/\alpha] &=& R\\
-      R[\alpha/\alpha,\beta/\beta] &=& R\\\\
-      R[\xi][\xi'] & = & R[\xi[\xi']]\\
-      R[\xi][\gamma] &=& R[\gamma][\xi[\gamma]]\\\\
-
-      \cdot[\xi] &=& \cdot\\
-      (A/\alpha)[B/\beta] &=& A[B/\beta]/\alpha\\
-      (A/\alpha;B/\beta)[\xi] &=& A[d^-\xi]/\alpha,B[d^+\xi]/\beta\\
-      \cdot[\gamma] &=& \cdot\\
-      (A/\alpha)[\gamma] &=& A[\gamma]/\alpha\\
-      (A/\alpha;B/\beta)[\gamma] &=& A[\gamma]/\alpha;B[\gamma]/\beta\\\\
-
-      \cdot[\gamma] &=& \cdot\\
-      (A/\alpha)[\gamma] &=& A[\gamma]/\alpha\\
-      (\phi_1,M/x,\phi_2)[\gamma] &=& \phi_1[\gamma],M[\gamma]/x,\phi_2[\gamma]
-
-      TODO: cat substitution id/assoc
-    \end{array}
-  \end{mathpar}  
-  \caption{Algebraic Properties}
-  \label{fig:equations}
-\end{figure}
-
-\section{Additional Rules}
-
-The key property that the hom allows is that given
-any variable in context of a term $\Phi,\alpha,x,\beta,\Psi \vdash M :
-Q$ we can represent this term as a ``single-variable term'' by moving
-$\Phi$ and $\Psi$ to the right of the turnstile:
-
-\[ \alpha, x, \beta \vdash \lambda \Phi. \lambda \Psi M. : \Phi \triangleright Q \triangleleft \Psi \]
-
-This allows us to describe restricted positive eliminations that only
-work on single variable contexts. We give an admissible ``syntax
-sugar'' in Figure~\ref{fig:multi-hom}.
-
-Any $\Phi_l,\beta:C,\Phi_r \vdash M : R$ can be represented as a term
-$\beta:C \vdash M' : \Phi_r \triangleright_{\beta:C} R \triangleleft_{\beta:C} $
-
-Say we have a term in context
-\[ \alpha,x:P,\beta,y:Q,\gamma \vdash M : R \]
-Using $\triangleright\triangleleft$, we can produce an equivalent term
-of a different output type whose context is one of the single
-variables $\alpha,\beta$ or $\gamma$ in the context.
-
-We can make it $\alpha$ by using $\triangleleft$ twice:
-\[ \alpha \vdash \lambda^\triangleleft (x:P,\beta). \lambda^\triangleleft (y:Q, \gamma). M : \forall \beta. (\forall \gamma. R \triangleleft Q) \triangleleft P \]
-
-Or $\gamma$ by using $\triangleright$ twice:
-\[ \gamma \vdash\lambda^\triangleright (\beta, y:Q).  \lambda^\triangleright (\alpha, x:P). M : \forall \gamma. Q \triangleright \forall \alpha. P \triangleright R \]
-
-And we have two different ways to make the context $\beta$, either we use $\triangleright$ and then $\triangleleft$:
-\[ \beta \vdash\lambda^\triangleright (\alpha, x:P).  \lambda^\triangleleft (y:Q, \gamma). M : \forall \alpha. P \triangleright (\forall \gamma. R \triangleleft Q) \]
-or we do the opposite order:
-\[ \beta \vdash \lambda^\triangleleft (y:Q, \gamma). \lambda^\triangleright (\alpha, x:P). M : \forall \gamma. (\forall \alpha. P \triangleright R) \triangleleft Q \]
-
-In general we can iterate this process. That is given any $\Phi,\Psi$
-with $d^+\Phi = d^-\Psi = \beta:D$, we can take a term $\Phi \jnctx
-\Psi \vdash M : R$ and produce an equivalent term $\beta:D \vdash M :
-\Phi \triangleright R \triangleleft \Psi$. To formalize this, we show
-how we can create ``multi-hom'' jumbo connectives $\Phi \triangleright
-R$ and $R \triangleleft \Psi$, where the boundary conditions on
-$\Phi,\Psi$ are analogous to the ordinary case. These are given in
-Figure\ref{multi-hom}.
-
-\begin{figure}
-  \begin{align*}
-    P \triangleleft \gamma:E &= P\\
-    P \triangleleft (\Psi,x:R,\gamma:E) &= (\forall \gamma: E. P \triangleleft R) \triangleleft \Psi\\
-    %% \cdot \triangleright_{\alpha : C} P &= P\\
-    %% (x:R,\beta:D,\Phi_r)\triangleright_{\alpha:C} P  &= \Phi_R \triangleright_{\beta:D} \forall \beta. R \triangleright P
-  \end{align*}
-  \begin{mathpar}
-    \inferrule
-    {\Gamma \pipe d^+\Xi; \gamma: E \vdash P \isaSet\and
-     \Gamma \pipe d^-\Xi; \gamma:E \vdash \Psi \isavectx}
-    {\Gamma \pipe \Xi \vdash P \triangleleft \Psi}
-
-    \inferrule*[right=Covariant Multi-Hom introduction]
-    {\Gamma\pipe\Phi\jnctx \Psi \vdash M : P}
-    {\Gamma\pipe\Phi \vdash \lambda^\triangleleft \Psi. M : P \triangleleft \Psi}\and
-
-    \inferrule*[right=Covar Multi-Hom elim]
-    {~}
-    {\Gamma \pipe f:P\triangleleft \Psi,\Psi \vdash \textrm{app}\triangleleft_\Psi : P}
-
-    %%   (\lambda^\triangleleft \Psi. M)[\phi] = \lambda^\triangleleft \Psi. M[\phi,\Psi]\\
-    %%   \lambda^\triangleleft \gamma. M = M\\
-    %%   \lambda^\triangleleft (\Psi,x:R,\gamma:\cat E). M = \lambda^\triangleleft \Psi. \lambda^\triangleleft (x:R,\gamma:\cat E). M\\
-
-    %%   (\lambda^\triangleleft \gamma. M)[\phi] = M[\phi] = \lambda^\triangleleft \gamma. M[\phi]\\
-    %%   (\lambda^\triangleleft (\Psi,x:r,\gamma:\cat E). M)[\phi] =
-    %%   (\lambda^\triangleleft \Psi. \lambda^\triangleleft (x:R,\gamma:\cat E). M)[\phi]
-    %%   = \lambda^\triangleleft \Psi. (\lambda^\triangleleft (x:R,\gamma:\cat E). M)[\phi, \Psi]
-    %%   = \lambda^\triangleleft \Psi. \lambda^\triangleleft (x:R,\gamma:\cat E). M[\phi,\Psi,x:R,\gamma]
-
-    %%   \inferrule*[right=Covariant Multi-Hom elimination]
-    %%   {\Phi_1 \vdash M : P\triangleleft \Phi_2 \and \phi_2 : \Psi \vdash \Phi_2 \subst(\xi)}
-    %%   {\Phi_1,\Psi \vdash M\triangleleft \phi_2 : P[d^+\xi]}\\
-
-    %%   M \triangleleft (A/\gamma) = M[A/\gamma]\\
-    %%   M \triangleleft (\phi, N/x,A/\gamma) =  M \triangleleft (N, A) \triangleleft \phi
-
-    %%   (M \triangleleft \phi_2)[\phi_1,\psi] = M[\phi_1] \triangleleft [\phi_2[\psi]]\\
-    %%   TODO: prove
-
-    %%   TODO: (\lambda^\triangleleft \Phi. M) \triangleleft \phi = M[\phi]
-    %%   TODO: M = \lambda^\triangleleft \Phi. M\triangleleft \Phi
-
-    %%   TODO: contravariant by duality
-    %%   %% \inferrule*[right=Contra Multi-Hom formation]
-    %%   %% {\Phi \isavectx(\alpha:\cat C; d^-\Xi) \and P \prof(\alpha : \cat C; d^+\Xi)}
-    %%   %% {\Phi \triangleright P \prof(\xi)}
-
-    %%   %% \alpha \triangleright P = P\\
-    %%   %% \alpha:\cat C, x:R, \Phi \triangleright P = \forall \alpha. R \triangleright \Phi \triangleright P\\
-  \end{mathpar}
-  \begin{align*}
-    \lambda^\triangleleft \gamma:E. M &= M\\
-    \lambda^\triangleleft (\Psi,x:R,\gamma:E). M &= \lambda^\triangleleft \Psi. \lambda^\triangleleft (x:R,\gamma:E). M\\\\
-    \textrm{app}\triangleleft_{\gamma:E} &= f\\
-    \textrm{app}\triangleleft_{\Psi,x:R\gamma:E} &= \textrm{app}\triangleleft_\Psi \triangleleft (x,\gamma) \\
-  \end{align*}
-  \caption{Multi-Hom}
-  \label{multi-hom}
-\end{figure}
-
-\section{Unit and Tensor}
-we can define the
-\emph{unit} , which in set-theoretic settings we
-might call the \emph{identity/equality judgment} $A \equiv_{\cat C}
-A'$.
-%
-First the formation rule says given two objects $A,A'$ in the same
-category $\cat C$, we can form the hom-set from $A$ to $A'$ $A
-\to_{\cat C} A'$.
-%
-The introduction rule says that given any object $A$, we can define
-the identity arrow from $A$ to itself $\id(A) : A \to_{\cat C} A$.
-%
-Without knowing more about the category in question, these are the
-only arrows we know exist.
-%
-
-The elimination form is trickiest to define in a natural deduction style.
-%
-First, let's consider the universal property of the Hom. From
-Crutwell-Shulman, we want a 2-cell
-
-\[% https://tikzcd.yichuanshen.de/#N4Igdg9gJgpgziAXAbVABwnAlgFyxMJZARgBoAGAXVJADcBDAGwFcYkQBBEAX1PU1z5CKcqQBM1Ok1bsuvftjwEiY8ZIYs2iTjz4gMioUTLF10rSABOMAGaMekmFADm8IqBuWIAWyRkQOBBIqlKa7AASProeXr6IogFBiP6MWGAWUBDMAEaMbDQAFjD0UOyQ6SA0OPRY9trlbPIgnlHxVUkhqRXamTl5lSBFJWUE+QE1deCjDtxAA
-\begin{tikzcd}
-                    & A \arrow[ldd, no head, Rightarrow] \arrow[rdd, no head, Rightarrow] &   \\
-                    & refl                                                                &   \\
-A \arrow[rr, "Hom"] &                                                                     & A
-\end{tikzcd}\]
-
-That is universal/opcartesian in that for any 2-cell $t$:
-
-\[ % https://tikzcd.yichuanshen.de/#N4Igdg9gJgpgziAXAbVABwnAlgFyxMJZABgBpiBdUkANwEMAbAVxiRACEQBfU9TXfIRQBGclVqMWbbrxAZseAkQBMY6vWatEIAIIy+CwUQAsaiZrZ6eBgUpQA2Mxqnb2Acn1z+ioclXLxZy0QAGFPeVtfUwD1SWCAEXDvIxQAZlJhQLi2HA8ucRgoAHN4IlAAMwAnCABbJDIQHAgkYWsQKtqW6iakVRAGLDBgqDo4AAtCkGoJuig2SCHPDrrEPp7EdP7B4LgIAbnpmFmkMCYGBmoGOgAjGAYABWS7fphynCnzFxAACU625aQm3Wpi2i20I3Gk0Ox20C1Y-2qKwa6wArAjOogQet7OiVijus1EI5PsEAErcChcIA
-\begin{tikzcd}
-B \arrow[r] \arrow[rrdd] & {} \arrow[r, no head, dashed] & A \arrow[rr, "Hom"] &    & A \arrow[rr, no head, dashed] &  & B' \arrow[lldd] \\
-                         &                               &                     & t' &                               &  &                 \\
-                         &                               & C \arrow[rr, "R"]   &    & D                             &  &                
-\end{tikzcd}\]
-
-there exists a $t'$ which extends $t$ in that the following is equal
-to $t$:
-\[% https://tikzcd.yichuanshen.de/#N4Igdg9gJgpgziAXAbVABwnAlgFyxMJZABgBoAmAXVJADcBDAGwFcYkQAhEAX1PU1z5CKAIwVqdJq3Y8+IDNjwEi5cTQYs2iEAEFZ-RUKIAWNZM3s9vA4OUoAbGY3TtHAOT75ApcOSrjEs5aIADCngq2vqYB6lLBACLh3kYoAMykqYFx7Dge1l6GdsjpxFkW2lZyET5EZKWx5SBJhb5i9eYuTfnVKciO7UEy3clF6SJlnQBOMABmjDwSMFAA5vBEoDOTEAC2SGQgOBBIIvmbO8c0h0iqIIxYYMFQ9HAAFksgNG-0UOyQD55nXaIG5XRDpW73YJwCB3H6fGDfJBgZiMRg0Rj0ABGMEYAAURsJbrMcB8OsEABLnU5bIHg0GmCH-bRPV7veGI7R-NjU86IfaggCsPKBDNB9mFSAFlyOiEcZPYACUATSkAAOaXXdGQ9hQCDMTGMNjsn6cghGg70LDzU3-CWIdUHGXgu5MkC6-WG0lfE3gM2knCW62+21yQFIACcGr5dsjjuO+xdUJhWDhIG9SJRaItVt+frtIn5MrEjKTsK9CJNyNRl0DuZDGxViALUYdiZ1zzeqfTNu5ocbDtBImLbeZHbZaYrdd7Dd5Q6jDJHbr1BvN3eD5oDOZ7C24QA
-\begin{tikzcd}
-{} \arrow[dd] \arrow[r]  & {} \arrow[dd] \arrow[rr, no head, dashed] &                     & A \arrow[ldd, no head, Rightarrow] \arrow[rdd, no head, Rightarrow] \arrow[rrr, no head, dashed] &                               &  & {} \arrow[dd, no head, Rightarrow] \\
-                         &                                           &                     & refl                                                                                             &                               &  &                                    \\
-B \arrow[r] \arrow[rrdd] & {} \arrow[r, no head, dashed]             & A \arrow[rr, "Hom"] &                                                                                                  & A \arrow[rr, no head, dashed] &  & B' \arrow[lldd]                    \\
-                         &                                           &                     & t'                                                                                               &                               &  &                                    \\
-                         &                                           & C \arrow[rr, "R"]   &                                                                                                  & D                             &  &
-\end{tikzcd}\]
-
-Clearly, our introduction rule models this first 2-cell. Furthermore,
-the construction in the second condition is modeled faithfully by the
-following sequent rule in the figure. This is a J-like induction rule:
-to construct a term using an input of the Hom type, it is sufficient
-to consider the case in which the input is the identity morphism.
-%
-However, building in a substitution the usual way we would (pushing
-the substitution into the continuation $N$) doesn't work properly
-because we can't necessarily extract a substitution for the
-continuation from the substitution for the the whole term. A
-substitution for the whole term would be of the form
-\[\phi,D_1/\delta_1,M/f,D_2/\delta_2,\psi\]
-The the candidate for the continuation's substitution would be
-$\phi,?,\psi$ but there is not necessarily something to go in the
-middle here since $D_1,D_2$ do not necessarily have the same
-\emph{input variable}. However, if we assume the existence already of
-the cotensor products the problem can be simplified. Instead of
-allowing for a general elimination of a hom term anywhere in the
-context, we can instead make a remarkably restricted rule: we can only
-eliminate a Hom term when it is the \emph{sole} input in the context:
-
-\[ \inferrule*[right=Hom Elim Very Simple]{\alpha:\cat C \vdash N : Q[\alpha/\alpha_1,\alpha/\alpha_2]}{\alpha_1:\cat C, f : \alpha_1 \to_{\cat C} \alpha_2, \alpha_2:\cat C \vdash J(\alpha_1.\alpha_2.
+%% \appendix
+
+%% TODO: complete substitution
+%% \begin{figure}
+%%   \begin{mathpar}
+%%       (\homrallXYtoZ \alpha R P)[\xi] = \homrallXYtoZ \alpha {R[d^+\xi;\alpha/\alpha]} {P[\alpha/\alpha;d^-\xi]}
+
+%%      (\homrlambdaXatYdotZ {x} {\alpha} M)[\phi]
+%%       = \homrlambdaXatYdotZ {x} {\alpha} M[\phi,x/x,\gamma/\gamma]
+
+%%      (\homllambdaXatYdotZ {x} {\alpha} M)[\phi]
+%%       = \homllambdaXatYdotZ {x} {\alpha} M[\phi,x/x,\gamma/\gamma]
+
+%%   \end{mathpar}
+%%   \caption{Substitution}
+
+%% \end{figure}
+
+%% \begin{figure}
+%%   \begin{mathpar}
+%%     \inferrule
+%%     {}
+%%     {\Gamma\pipe \cdot \vdash \Sort \isaSet}
+
+%%     \inferrule
+%%     {x:S \in \Gamma}
+%%     {\Gamma \pipe \cdot \vdash x : S}
+
+%%     \inferrule
+%%     {\Gamma\pipe\cdot \vdash C : \Sort}
+%%     {\Gamma \pipe \alpha:\cat C \vdash \alpha : \cat C}
+
+%%     \inferrule
+%%     {\Gamma \pipe \alpha:C,\beta:D \vdash R \isaSet}
+%%     {\Gamma\pipe \alpha:C, x:R, \beta: D \vdash x : R}
+
+%%     \inferrule{\forall x:A \in \Gamma. \Delta \vdash \gamma(x) : A}{\gamma : \Delta \to \Gamma}\\
+
+%%     \inferrule*%% [right=NullaryBoundaryMap]
+%%                {\Gamma \isadtctx}{\Gamma \vdash \cdot : \cdot \to \cdot}\and
+%%     \inferrule*{\Gamma \pipe \alpha':C' \vdash A : C}{\Gamma \vdash A/\alpha : \alpha':C' \to \alpha:C}
+
+%%     \inferrule*
+%%     {\Gamma \pipe \alpha':C' \vdash A : C\and
+%%       \Gamma \pipe \beta':D' \vdash B : D
+%%     }
+%%     {\Gamma \vdash (A/\alpha;B/\beta) : (\alpha':C';\beta':D') \to (\alpha:C;\beta':D')}\\
+
+%%     \inferrule*%% [right=NullCtxMap]
+%%                {\Gamma \isadtctx}{\Gamma \vdash \cdot : \cdot \to \cdot}
+
+%%     \inferrule*{\Gamma \pipe \alpha':C' \vdash A : C}{\Gamma \vdash A/\alpha : \alpha':C' \to \alpha:C}
+
+%%     \inferrule
+%%     {\Psi = \Psi_1 \jnctx \Psi_M \jnctx \Psi_2\and
+%%       \Gamma \vdash \phi_1 : \Psi_1 \to \Phi_1\and
+%%       \Gamma \pipe \Psi_M \vdash M : R[d^+\phi_1;d^-\phi_2] \and
+%%       \Gamma \vdash \phi_2 : \Psi_2 \to \Phi_2
+%%       }
+%%     {\Gamma \vdash \phi_1,M/x,\phi_2 : \Psi \to \Phi_1,x:R,\Phi_2}
+%%   \end{mathpar}
+%%   \caption{Substitutions in SCT}
+%%   \label{fig:basic-rules}
+%% \end{figure}
+
+%% \begin{figure}
+%%   \begin{mathpar}
+%%     \inferrule
+%%     {}
+%%     {\cdot : \cdot \to \cdot}
+
+%%     \inferrule
+%%     {\gamma : \Delta \to \Gamma \and \Delta \vdash M : A[\gamma]}
+%%     {\gamma,M/X : \Delta \to \Gamma,X:A}
+
+%%     \inferrule
+%%     {\Gamma \vdash \Phi \isavectx \and
+%%     \Gamma \vdash \Psi \isavectx \and
+%%     d^+\Phi = d^- \Psi}
+%%     {\Gamma \vdash \Phi \jnctx \Psi \isavectx}
+
+%%     \inferrule
+%%     {\Gamma \vdash \phi : \Phi' \to \Phi \and
+%%      \Gamma \vdash \psi : \Psi' \to \Psi\and
+%%      d^+\phi = d^- \psi}
+%%     {\Gamma \vdash \phi \jnctx \psi : \Phi' \jnctx \Psi' \to \Phi \jnctx \Psi}
+
+%%     \inferrule*[right=ObSubst]
+%%     {\alpha : \cat C \vdash B : \cat D\and
+%%     \beta : \cat D \vdash A : \cat E}
+%%     {\alpha : \cat C \vdash A[B/\beta] : \cat E}\\
+
+%%     \inferrule*[right=Everything Subst]{\gamma : \Delta \to \Gamma \and \Gamma J}{\Delta J}
+
+%%     \inferrule{\Gamma \vdash \xi : Z \to \Xi\and \Gamma \pipe \Xi \vdash R \isaSet}{\Gamma \pipe Z \vdash R[\xi] \isaSet}
+
+%%     \inferrule{\Gamma \vdash \phi : \Psi \to \Phi\and \Gamma \pipe \Phi \vdash M : R}{\Gamma \pipe \Psi \vdash M[\phi] : R[\underline \phi]}
+%%   \end{mathpar}
+%%   \begin{align*}
+%%     (\Phi \jnctx \Psi) \jnctx \Sigma &= \Phi \jnctx (\Psi \jnctx \Sigma)\\
+%%     \Phi \jnctx \alpha:C &= \Phi\\
+%%     \alpha:C \jnctx \Psi &= \Psi\\\\
+%%     d^\pm(A/\alpha) &= A/\alpha\\
+%%     d^+(\phi_1,M/x,\phi_2) &= d^+ \phi_2\\
+%%     d^-(\phi_1,M/x,\phi_2) &= d^- \phi_1\\\\
+%%     \underline \cdot &= \cdot\\
+%%     \underline {A/\alpha} &= A/\alpha\\
+%%     \underline {\phi_1,M/x,\phi_2} &= d^-\phi_1;d^+\phi_2
+%%   \end{align*}
+%%   \caption{Admissible Rules}
+%% \end{figure}
+
+%% \begin{figure}
+%%   \begin{mathpar}
+%%     \begin{array}{rcl}
+%%       x[\gamma] &=& \gamma(x)\\
+%%       J[\gamma][\delta] & = & J[\gamma[\delta]]\\
+%%       \gamma[\delta](x) &=& \gamma(x)[\delta]\\
+
+%%       A[\alpha/\alpha] &=& A\\
+%%       \alpha[A/\alpha] &=& A\\
+%%       A[B/\beta][C/\gamma] & = & A[B[C/\gamma]/\beta]\\
+%%       A[B/\beta][\gamma] &=& A[\gamma][B[\gamma]/\beta]\\\\
+
+%%       R[\cdot] &=& R\\
+%%       R[\alpha/\alpha] &=& R\\
+%%       R[\alpha/\alpha,\beta/\beta] &=& R\\\\
+%%       R[\xi][\xi'] & = & R[\xi[\xi']]\\
+%%       R[\xi][\gamma] &=& R[\gamma][\xi[\gamma]]\\\\
+
+%%       \cdot[\xi] &=& \cdot\\
+%%       (A/\alpha)[B/\beta] &=& A[B/\beta]/\alpha\\
+%%       (A/\alpha;B/\beta)[\xi] &=& A[d^-\xi]/\alpha,B[d^+\xi]/\beta\\
+%%       \cdot[\gamma] &=& \cdot\\
+%%       (A/\alpha)[\gamma] &=& A[\gamma]/\alpha\\
+%%       (A/\alpha;B/\beta)[\gamma] &=& A[\gamma]/\alpha;B[\gamma]/\beta\\\\
+
+%%       \cdot[\gamma] &=& \cdot\\
+%%       (A/\alpha)[\gamma] &=& A[\gamma]/\alpha\\
+%%       (\phi_1,M/x,\phi_2)[\gamma] &=& \phi_1[\gamma],M[\gamma]/x,\phi_2[\gamma]
+
+%%       TODO: cat substitution id/assoc
+%%     \end{array}
+%%   \end{mathpar}  
+%%   \caption{Algebraic Properties}
+%%   \label{fig:equations}
+%% \end{figure}
+
+%% \section{Additional Rules}
+
+%% The key property that the hom allows is that given
+%% any variable in context of a term $\Phi,\alpha,x,\beta,\Psi \vdash M :
+%% Q$ we can represent this term as a ``single-variable term'' by moving
+%% $\Phi$ and $\Psi$ to the right of the turnstile:
+
+%% \[ \alpha, x, \beta \vdash \lambda \Phi. \lambda \Psi M. : \Phi \triangleright Q \triangleleft \Psi \]
+
+%% This allows us to describe restricted positive eliminations that only
+%% work on single variable contexts. We give an admissible ``syntax
+%% sugar'' in Figure~\ref{fig:multi-hom}.
+
+%% Any $\Phi_l,\beta:C,\Phi_r \vdash M : R$ can be represented as a term
+%% $\beta:C \vdash M' : \Phi_r \triangleright_{\beta:C} R \triangleleft_{\beta:C} $
+
+%% Say we have a term in context
+%% \[ \alpha,x:P,\beta,y:Q,\gamma \vdash M : R \]
+%% Using $\triangleright\triangleleft$, we can produce an equivalent term
+%% of a different output type whose context is one of the single
+%% variables $\alpha,\beta$ or $\gamma$ in the context.
+
+%% We can make it $\alpha$ by using $\triangleleft$ twice:
+%% \[ \alpha \vdash \lambda^\triangleleft (x:P,\beta). \lambda^\triangleleft (y:Q, \gamma). M : \forall \beta. (\forall \gamma. R \triangleleft Q) \triangleleft P \]
+
+%% Or $\gamma$ by using $\triangleright$ twice:
+%% \[ \gamma \vdash\lambda^\triangleright (\beta, y:Q).  \lambda^\triangleright (\alpha, x:P). M : \forall \gamma. Q \triangleright \forall \alpha. P \triangleright R \]
+
+%% And we have two different ways to make the context $\beta$, either we use $\triangleright$ and then $\triangleleft$:
+%% \[ \beta \vdash\lambda^\triangleright (\alpha, x:P).  \lambda^\triangleleft (y:Q, \gamma). M : \forall \alpha. P \triangleright (\forall \gamma. R \triangleleft Q) \]
+%% or we do the opposite order:
+%% \[ \beta \vdash \lambda^\triangleleft (y:Q, \gamma). \lambda^\triangleright (\alpha, x:P). M : \forall \gamma. (\forall \alpha. P \triangleright R) \triangleleft Q \]
+
+%% In general we can iterate this process. That is given any $\Phi,\Psi$
+%% with $d^+\Phi = d^-\Psi = \beta:D$, we can take a term $\Phi \jnctx
+%% \Psi \vdash M : R$ and produce an equivalent term $\beta:D \vdash M :
+%% \Phi \triangleright R \triangleleft \Psi$. To formalize this, we show
+%% how we can create ``multi-hom'' jumbo connectives $\Phi \triangleright
+%% R$ and $R \triangleleft \Psi$, where the boundary conditions on
+%% $\Phi,\Psi$ are analogous to the ordinary case. These are given in
+%% Figure\ref{multi-hom}.
+
+%% \begin{figure}
+%%   \begin{align*}
+%%     P \triangleleft \gamma:E &= P\\
+%%     P \triangleleft (\Psi,x:R,\gamma:E) &= (\forall \gamma: E. P \triangleleft R) \triangleleft \Psi\\
+%%     %% \cdot \triangleright_{\alpha : C} P &= P\\
+%%     %% (x:R,\beta:D,\Phi_r)\triangleright_{\alpha:C} P  &= \Phi_R \triangleright_{\beta:D} \forall \beta. R \triangleright P
+%%   \end{align*}
+%%   \begin{mathpar}
+%%     \inferrule
+%%     {\Gamma \pipe d^+\Xi; \gamma: E \vdash P \isaSet\and
+%%      \Gamma \pipe d^-\Xi; \gamma:E \vdash \Psi \isavectx}
+%%     {\Gamma \pipe \Xi \vdash P \triangleleft \Psi}
+
+%%     \inferrule*[right=Covariant Multi-Hom introduction]
+%%     {\Gamma\pipe\Phi\jnctx \Psi \vdash M : P}
+%%     {\Gamma\pipe\Phi \vdash \lambda^\triangleleft \Psi. M : P \triangleleft \Psi}\and
+
+%%     \inferrule*[right=Covar Multi-Hom elim]
+%%     {~}
+%%     {\Gamma \pipe f:P\triangleleft \Psi,\Psi \vdash \textrm{app}\triangleleft_\Psi : P}
+
+%%     %%   (\lambda^\triangleleft \Psi. M)[\phi] = \lambda^\triangleleft \Psi. M[\phi,\Psi]\\
+%%     %%   \lambda^\triangleleft \gamma. M = M\\
+%%     %%   \lambda^\triangleleft (\Psi,x:R,\gamma:\cat E). M = \lambda^\triangleleft \Psi. \lambda^\triangleleft (x:R,\gamma:\cat E). M\\
+
+%%     %%   (\lambda^\triangleleft \gamma. M)[\phi] = M[\phi] = \lambda^\triangleleft \gamma. M[\phi]\\
+%%     %%   (\lambda^\triangleleft (\Psi,x:r,\gamma:\cat E). M)[\phi] =
+%%     %%   (\lambda^\triangleleft \Psi. \lambda^\triangleleft (x:R,\gamma:\cat E). M)[\phi]
+%%     %%   = \lambda^\triangleleft \Psi. (\lambda^\triangleleft (x:R,\gamma:\cat E). M)[\phi, \Psi]
+%%     %%   = \lambda^\triangleleft \Psi. \lambda^\triangleleft (x:R,\gamma:\cat E). M[\phi,\Psi,x:R,\gamma]
+
+%%     %%   \inferrule*[right=Covariant Multi-Hom elimination]
+%%     %%   {\Phi_1 \vdash M : P\triangleleft \Phi_2 \and \phi_2 : \Psi \vdash \Phi_2 \subst(\xi)}
+%%     %%   {\Phi_1,\Psi \vdash M\triangleleft \phi_2 : P[d^+\xi]}\\
+
+%%     %%   M \triangleleft (A/\gamma) = M[A/\gamma]\\
+%%     %%   M \triangleleft (\phi, N/x,A/\gamma) =  M \triangleleft (N, A) \triangleleft \phi
+
+%%     %%   (M \triangleleft \phi_2)[\phi_1,\psi] = M[\phi_1] \triangleleft [\phi_2[\psi]]\\
+%%     %%   TODO: prove
+
+%%     %%   TODO: (\lambda^\triangleleft \Phi. M) \triangleleft \phi = M[\phi]
+%%     %%   TODO: M = \lambda^\triangleleft \Phi. M\triangleleft \Phi
+
+%%     %%   TODO: contravariant by duality
+%%     %%   %% \inferrule*[right=Contra Multi-Hom formation]
+%%     %%   %% {\Phi \isavectx(\alpha:\cat C; d^-\Xi) \and P \prof(\alpha : \cat C; d^+\Xi)}
+%%     %%   %% {\Phi \triangleright P \prof(\xi)}
+
+%%     %%   %% \alpha \triangleright P = P\\
+%%     %%   %% \alpha:\cat C, x:R, \Phi \triangleright P = \forall \alpha. R \triangleright \Phi \triangleright P\\
+%%   \end{mathpar}
+%%   \begin{align*}
+%%     \lambda^\triangleleft \gamma:E. M &= M\\
+%%     \lambda^\triangleleft (\Psi,x:R,\gamma:E). M &= \lambda^\triangleleft \Psi. \lambda^\triangleleft (x:R,\gamma:E). M\\\\
+%%     \textrm{app}\triangleleft_{\gamma:E} &= f\\
+%%     \textrm{app}\triangleleft_{\Psi,x:R\gamma:E} &= \textrm{app}\triangleleft_\Psi \triangleleft (x,\gamma) \\
+%%   \end{align*}
+%%   \caption{Multi-Hom}
+%%   \label{multi-hom}
+%% \end{figure}
+
+%% \section{Unit and Tensor}
+%% we can define the
+%% \emph{unit} , which in set-theoretic settings we
+%% might call the \emph{identity/equality judgment} $A \equiv_{\cat C}
+%% A'$.
+%% %
+%% First the formation rule says given two objects $A,A'$ in the same
+%% category $\cat C$, we can form the hom-set from $A$ to $A'$ $A
+%% \to_{\cat C} A'$.
+%% %
+%% The introduction rule says that given any object $A$, we can define
+%% the identity arrow from $A$ to itself $\id(A) : A \to_{\cat C} A$.
+%% %
+%% Without knowing more about the category in question, these are the
+%% only arrows we know exist.
+%% %
+
+%% The elimination form is trickiest to define in a natural deduction style.
+%% %
+%% First, let's consider the universal property of the Hom. From
+%% Crutwell-Shulman, we want a 2-cell
+
+%% \[% https://tikzcd.yichuanshen.de/#N4Igdg9gJgpgziAXAbVABwnAlgFyxMJZARgBoAGAXVJADcBDAGwFcYkQBBEAX1PU1z5CKcqQBM1Ok1bsuvftjwEiY8ZIYs2iTjz4gMioUTLF10rSABOMAGaMekmFADm8IqBuWIAWyRkQOBBIqlKa7AASProeXr6IogFBiP6MWGAWUBDMAEaMbDQAFjD0UOyQ6SA0OPRY9trlbPIgnlHxVUkhqRXamTl5lSBFJWUE+QE1deCjDtxAA
+%% \begin{tikzcd}
+%%                     & A \arrow[ldd, no head, Rightarrow] \arrow[rdd, no head, Rightarrow] &   \\
+%%                     & refl                                                                &   \\
+%% A \arrow[rr, "Hom"] &                                                                     & A
+%% \end{tikzcd}\]
+
+%% That is universal/opcartesian in that for any 2-cell $t$:
+
+%% \[ % https://tikzcd.yichuanshen.de/#N4Igdg9gJgpgziAXAbVABwnAlgFyxMJZABgBpiBdUkANwEMAbAVxiRACEQBfU9TXfIRQBGclVqMWbbrxAZseAkQBMY6vWatEIAIIy+CwUQAsaiZrZ6eBgUpQA2Mxqnb2Acn1z+ioclXLxZy0QAGFPeVtfUwD1SWCAEXDvIxQAZlJhQLi2HA8ucRgoAHN4IlAAMwAnCABbJDIQHAgkYWsQKtqW6iakVRAGLDBgqDo4AAtCkGoJuig2SCHPDrrEPp7EdP7B4LgIAbnpmFmkMCYGBmoGOgAjGAYABWS7fphynCnzFxAACU625aQm3Wpi2i20I3Gk0Ox20C1Y-2qKwa6wArAjOogQet7OiVijus1EI5PsEAErcChcIA
+%% \begin{tikzcd}
+%% B \arrow[r] \arrow[rrdd] & {} \arrow[r, no head, dashed] & A \arrow[rr, "Hom"] &    & A \arrow[rr, no head, dashed] &  & B' \arrow[lldd] \\
+%%                          &                               &                     & t' &                               &  &                 \\
+%%                          &                               & C \arrow[rr, "R"]   &    & D                             &  &                
+%% \end{tikzcd}\]
+
+%% there exists a $t'$ which extends $t$ in that the following is equal
+%% to $t$:
+%% \[% https://tikzcd.yichuanshen.de/#N4Igdg9gJgpgziAXAbVABwnAlgFyxMJZABgBoAmAXVJADcBDAGwFcYkQAhEAX1PU1z5CKAIwVqdJq3Y8+IDNjwEi5cTQYs2iEAEFZ-RUKIAWNZM3s9vA4OUoAbGY3TtHAOT75ApcOSrjEs5aIADCngq2vqYB6lLBACLh3kYoAMykqYFx7Dge1l6GdsjpxFkW2lZyET5EZKWx5SBJhb5i9eYuTfnVKciO7UEy3clF6SJlnQBOMABmjDwSMFAA5vBEoDOTEAC2SGQgOBBIIvmbO8c0h0iqIIxYYMFQ9HAAFksgNG-0UOyQD55nXaIG5XRDpW73YJwCB3H6fGDfJBgZiMRg0Rj0ABGMEYAAURsJbrMcB8OsEABLnU5bIHg0GmCH-bRPV7veGI7R-NjU86IfaggCsPKBDNB9mFSAFlyOiEcZPYACUATSkAAOaXXdGQ9hQCDMTGMNjsn6cghGg70LDzU3-CWIdUHGXgu5MkC6-WG0lfE3gM2knCW62+21yQFIACcGr5dsjjuO+xdUJhWDhIG9SJRaItVt+frtIn5MrEjKTsK9CJNyNRl0DuZDGxViALUYdiZ1zzeqfTNu5ocbDtBImLbeZHbZaYrdd7Dd5Q6jDJHbr1BvN3eD5oDOZ7C24QA
+%% \begin{tikzcd}
+%% {} \arrow[dd] \arrow[r]  & {} \arrow[dd] \arrow[rr, no head, dashed] &                     & A \arrow[ldd, no head, Rightarrow] \arrow[rdd, no head, Rightarrow] \arrow[rrr, no head, dashed] &                               &  & {} \arrow[dd, no head, Rightarrow] \\
+%%                          &                                           &                     & refl                                                                                             &                               &  &                                    \\
+%% B \arrow[r] \arrow[rrdd] & {} \arrow[r, no head, dashed]             & A \arrow[rr, "Hom"] &                                                                                                  & A \arrow[rr, no head, dashed] &  & B' \arrow[lldd]                    \\
+%%                          &                                           &                     & t'                                                                                               &                               &  &                                    \\
+%%                          &                                           & C \arrow[rr, "R"]   &                                                                                                  & D                             &  &
+%% \end{tikzcd}\]
+
+%% Clearly, our introduction rule models this first 2-cell. Furthermore,
+%% the construction in the second condition is modeled faithfully by the
+%% following sequent rule in the figure. This is a J-like induction rule:
+%% to construct a term using an input of the Hom type, it is sufficient
+%% to consider the case in which the input is the identity morphism.
+%% %
+%% However, building in a substitution the usual way we would (pushing
+%% the substitution into the continuation $N$) doesn't work properly
+%% because we can't necessarily extract a substitution for the
+%% continuation from the substitution for the the whole term. A
+%% substitution for the whole term would be of the form
+%% \[\phi,D_1/\delta_1,M/f,D_2/\delta_2,\psi\]
+%% The the candidate for the continuation's substitution would be
+%% $\phi,?,\psi$ but there is not necessarily something to go in the
+%% middle here since $D_1,D_2$ do not necessarily have the same
+%% \emph{input variable}. However, if we assume the existence already of
+%% the cotensor products the problem can be simplified. Instead of
+%% allowing for a general elimination of a hom term anywhere in the
+%% context, we can instead make a remarkably restricted rule: we can only
+%% eliminate a Hom term when it is the \emph{sole} input in the context:
+
+%% \[ \inferrule*[right=Hom Elim Very Simple]{\alpha:\cat C \vdash N : Q[\alpha/\alpha_1,\alpha/\alpha_2]}{\alpha_1:\cat C, f : \alpha_1 \to_{\cat C} \alpha_2, \alpha_2:\cat C \vdash J(\alpha_1.\alpha_2.
 Q, \alpha.
@@
 
-While this rule looks very impoverished at first sight, it has the
-virtue of easily supporting a built-in cut:
-
-\[ \inferrule*[right=Hom Elim VSimple Subst]{
-  Q \prof(\alpha_1:\cat C; \alpha_2:\cat C)\and
-  \Phi \vdash M : A_1 \to_{\cat C} A_2\and
-  \alpha:\cat C \vdash N : Q[\alpha/\alpha_1,\alpha/\alpha_2]}{\Phi \vdash J(\alpha_1.\alpha_2. Q, \alpha. N, A_1, M, A_2) : Q[A_1/\alpha_1, A_2/\alpha_2]}\]
-
-However, this rule on its own clearly is not enough to support our
-desired universal property shown above diagrammatically. To prove our
-completeness theorem, we will need to show that this rule is enough to
-give us the desired universal property \emph{assuming the presence of
-  left and right cotensors}.
-
-So first, we must show that the ``proper'' sequent rule
-
-\[ \inferrule*[right=Hom-Left (Sequent Style)]
-    {\Phi_1[\alpha/\alpha_1], \alpha:\cat C, \Phi_2[\alpha/\alpha_2]\vdash Q[\alpha/\alpha_1,\alpha/\alpha_2]}
-    {\Phi_1, \alpha_1 : \cat C, f : \alpha_1 \to_{\cat C} \alpha_2, \alpha_2 : \cat C, \Phi_2\vdash : Q}
-\]
-
-is admissible.
-
-Which we can sketch as follows:
-
-\[
-\inferrule*
-    {\inferrule*{\inferrule*{\Phi_1[\alpha/\alpha_1],\alpha:\cat C, \Phi_2[\alpha/\alpha_1] \vdash Q[\alpha/\alpha_1;\alpha/\alpha_2]}{\alpha:\cat C \vdash \Phi_1[\alpha/\alpha_1] \triangleright Q[\alpha/\alpha_1;\alpha/\alpha_2] \triangleleft \Phi_2[\alpha/\alpha_2]}}{\alpha_1 : \cat C, f : \alpha_1 \to_{\cat C} \alpha_2, \alpha_2 : \cat C\vdash \Phi_1 \triangleright Q \triangleleft \Phi_2}}
-    {\Phi_1, \alpha_1 : \cat C, f : \alpha_1 \to_{\cat C} \alpha_2, \alpha_2 : \cat C, \Phi_2\vdash Q}
-\]
-
-Which using our syntax sugar is given by the term:
-\begin{mathpar}
-  \inferrule*
-    {\inferrule*
-      {\inferrule*
-        {\Phi_1[\alpha/\alpha_1],\alpha:\cat C, \Phi_2[\alpha/\alpha_1] \vdash N : Q[\alpha/\alpha_1;\alpha/\alpha_2]}
-        {\alpha:\cat C \vdash \lambda^\triangleright \Phi_1[\alpha/\alpha_1].
+%% While this rule looks very impoverished at first sight, it has the
+%% virtue of easily supporting a built-in cut:
+
+%% \[ \inferrule*[right=Hom Elim VSimple Subst]{
+%%   Q \prof(\alpha_1:\cat C; \alpha_2:\cat C)\and
+%%   \Phi \vdash M : A_1 \to_{\cat C} A_2\and
+%%   \alpha:\cat C \vdash N : Q[\alpha/\alpha_1,\alpha/\alpha_2]}{\Phi \vdash J(\alpha_1.\alpha_2. Q, \alpha. N, A_1, M, A_2) : Q[A_1/\alpha_1, A_2/\alpha_2]}\]
+
+%% However, this rule on its own clearly is not enough to support our
+%% desired universal property shown above diagrammatically. To prove our
+%% completeness theorem, we will need to show that this rule is enough to
+%% give us the desired universal property \emph{assuming the presence of
+%%   left and right cotensors}.
+
+%% So first, we must show that the ``proper'' sequent rule
+
+%% \[ \inferrule*[right=Hom-Left (Sequent Style)]
+%%     {\Phi_1[\alpha/\alpha_1], \alpha:\cat C, \Phi_2[\alpha/\alpha_2]\vdash Q[\alpha/\alpha_1,\alpha/\alpha_2]}
+%%     {\Phi_1, \alpha_1 : \cat C, f : \alpha_1 \to_{\cat C} \alpha_2, \alpha_2 : \cat C, \Phi_2\vdash : Q}
+%% \]
+
+%% is admissible.
+
+%% Which we can sketch as follows:
+
+%% \[
+%% \inferrule*
+%%     {\inferrule*{\inferrule*{\Phi_1[\alpha/\alpha_1],\alpha:\cat C, \Phi_2[\alpha/\alpha_1] \vdash Q[\alpha/\alpha_1;\alpha/\alpha_2]}{\alpha:\cat C \vdash \Phi_1[\alpha/\alpha_1] \triangleright Q[\alpha/\alpha_1;\alpha/\alpha_2] \triangleleft \Phi_2[\alpha/\alpha_2]}}{\alpha_1 : \cat C, f : \alpha_1 \to_{\cat C} \alpha_2, \alpha_2 : \cat C\vdash \Phi_1 \triangleright Q \triangleleft \Phi_2}}
+%%     {\Phi_1, \alpha_1 : \cat C, f : \alpha_1 \to_{\cat C} \alpha_2, \alpha_2 : \cat C, \Phi_2\vdash Q}
+%% \]
+
+%% Which using our syntax sugar is given by the term:
+%% \begin{mathpar}
+%%   \inferrule*
+%%     {\inferrule*
+%%       {\inferrule*
+%%         {\Phi_1[\alpha/\alpha_1],\alpha:\cat C, \Phi_2[\alpha/\alpha_1] \vdash N : Q[\alpha/\alpha_1;\alpha/\alpha_2]}
+%%         {\alpha:\cat C \vdash \lambda^\triangleright \Phi_1[\alpha/\alpha_1].
 \lambda^\triangleleft \Phi_2[\alpha/\alpha_2].
 N : \Phi_1[\alpha/\alpha_1] \triangleright Q[\alpha/\alpha_1;\alpha/\alpha_2] \triangleleft \Phi_2[\alpha/\alpha_2]}}
-      {\alpha_1 : \cat C, f : \alpha_1 \to_{\cat C} \alpha_2, \alpha_2 : \cat C\vdash J(\alpha_1.\alpha_2.\Phi_1 \triangleright Q \triangleleft \Phi_2, \alpha.
+%%       {\alpha_1 : \cat C, f : \alpha_1 \to_{\cat C} \alpha_2, \alpha_2 : \cat C\vdash J(\alpha_1.\alpha_2.\Phi_1 \triangleright Q \triangleleft \Phi_2, \alpha.
 \lambda^\triangleright \Phi_1[\alpha/\alpha_1].
@@
 N, \alpha_1,f,\alpha_2) : \Phi_1 \triangleright Q \triangleleft \Phi_2}}
-    {\Phi_1, \alpha_1 : \cat C, f : \alpha_1 \to_{\cat C} \alpha_2, \alpha_2 : \cat C, \Phi_2\vdash
-      \Phi_1 \triangleright (J(\alpha_1.\alpha_2.\Phi_1 \triangleright Q \triangleleft \Phi_2, \alpha.
+%%     {\Phi_1, \alpha_1 : \cat C, f : \alpha_1 \to_{\cat C} \alpha_2, \alpha_2 : \cat C, \Phi_2\vdash
+%%       \Phi_1 \triangleright (J(\alpha_1.\alpha_2.\Phi_1 \triangleright Q \triangleleft \Phi_2, \alpha.
 \lambda^\triangleright \Phi_1[\alpha/\alpha_1].
@@
 N, \alpha_1,f,\alpha_2)) \triangleleft \Phi_2 : Q}
-\end{mathpar}
-
-We need to show $\beta$ and $\eta$ hold.
-
-First, $\beta$.
-
-\begin{align*}
-  &(\Phi_1 \triangleright (J_\to(\alpha_1.\alpha_2.\Phi_1 \triangleright Q \triangleleft \Phi_2, \alpha.
+%% \end{mathpar}
+
+%% We need to show $\beta$ and $\eta$ hold.
+
+%% First, $\beta$.
+
+%% \begin{align*}
+%%   &(\Phi_1 \triangleright (J_\to(\alpha_1.\alpha_2.\Phi_1 \triangleright Q \triangleleft \Phi_2, \alpha.
 \lambda^\triangleright \Phi_1[\alpha/\alpha_1].
@@
 N, \alpha_1,f,\alpha_2)) \triangleleft \Phi_2) [\Phi_1[A/\alpha_1],A/\alpha_1,\id(A),A/\alpha_2,\Phi_2[A/\alpha_2]] \\
-  &=
-  \Phi_1[A/\alpha_1] \triangleright (J_\to(\alpha_1.\alpha_2.\Phi_1 \triangleright Q \triangleleft \Phi_2, \alpha.
+%%   &=
+%%   \Phi_1[A/\alpha_1] \triangleright (J_\to(\alpha_1.\alpha_2.\Phi_1 \triangleright Q \triangleleft \Phi_2, \alpha.
 \lambda^\triangleright \Phi_1[\alpha/\alpha_1].
@@
 N, A,\id(A), A)) \triangleleft \Phi_2[A/\alpha_2]\\
-  &=
-  \Phi_1[A/\alpha_1] \triangleright (\lambda^\triangleright \Phi_1[A/\alpha_1]. \lambda^\triangleleft \Phi_2[A/\alpha_2]. N[A/\alpha]) \triangleleft \Phi_2[A/\alpha_2]\\
-  &=  (\lambda^\triangleleft \Phi_2[A/\alpha_2]. N[A/\alpha]) \triangleleft \Phi_2[A/\alpha_2]\\
-  &=  N[A/\alpha]
-\end{align*}
-
-\begin{figure}
-  \begin{mathpar}
-    \inferrule*[right=Full Unit Elim*]
-    {\Phi_1[\alpha/\alpha_1], \alpha:\cat C, \Phi_2[\alpha/\alpha_2]\vdash N : Q[\alpha/\alpha_1,\alpha/\alpha_2]}
-    {\Phi_1, \alpha_1 : \cat C, f : \alpha_1 \to_{\cat C} \alpha_2, \alpha_2 : \cat C, \Phi_2\vdash J_\to'(\Phi_1, \alpha_1 : \cat C, f : \alpha_1 \to_{\cat C} \alpha_2, \alpha_2 : \cat C, \Phi_2, Q, \alpha.
+%%   &=
+%%   \Phi_1[A/\alpha_1] \triangleright (\lambda^\triangleright \Phi_1[A/\alpha_1]. \lambda^\triangleleft \Phi_2[A/\alpha_2]. N[A/\alpha]) \triangleleft \Phi_2[A/\alpha_2]\\
+%%   &=  (\lambda^\triangleleft \Phi_2[A/\alpha_2]. N[A/\alpha]) \triangleleft \Phi_2[A/\alpha_2]\\
+%%   &=  N[A/\alpha]
+%% \end{align*}
+
+%% \begin{figure}
+%%   \begin{mathpar}
+%%     \inferrule*[right=Full Unit Elim*]
+%%     {\Phi_1[\alpha/\alpha_1], \alpha:\cat C, \Phi_2[\alpha/\alpha_2]\vdash N : Q[\alpha/\alpha_1,\alpha/\alpha_2]}
+%%     {\Phi_1, \alpha_1 : \cat C, f : \alpha_1 \to_{\cat C} \alpha_2, \alpha_2 : \cat C, \Phi_2\vdash J_\to'(\Phi_1, \alpha_1 : \cat C, f : \alpha_1 \to_{\cat C} \alpha_2, \alpha_2 : \cat C, \Phi_2, Q, \alpha.
 N) : Q}
 
-    J_\to'(\Phi_1, \alpha_1 : \cat C, f : \alpha_1 \to_{\cat C} \alpha_2, \alpha_2 : \cat C, \Phi_2, Q, \alpha. N) =
-    \Phi_1 \triangleright (J(\alpha_1.\alpha_2.\Phi_1 \triangleright Q \triangleleft \Phi_2, \alpha.
+%%     J_\to'(\Phi_1, \alpha_1 : \cat C, f : \alpha_1 \to_{\cat C} \alpha_2, \alpha_2 : \cat C, \Phi_2, Q, \alpha. N) =
+%%     \Phi_1 \triangleright (J(\alpha_1.\alpha_2.\Phi_1 \triangleright Q \triangleleft \Phi_2, \alpha.
 \lambda^\triangleright \Phi_1[\alpha/\alpha_1].
@@
 
-    \inferrule*[right=Full Unit $\beta$*]
-    {}
-    {J_\to'(\Phi_1, \alpha_1 : \cat C, f : \alpha_1 \to_{\cat C} \alpha_2, \alpha_2 : \cat C, \Phi_2, Q, \alpha.
+%%     \inferrule*[right=Full Unit $\beta$*]
+%%     {}
+%%     {J_\to'(\Phi_1, \alpha_1 : \cat C, f : \alpha_1 \to_{\cat C} \alpha_2, \alpha_2 : \cat C, \Phi_2, Q, \alpha.
 N)[\Phi_1[\alpha/\alpha_1],\alpha/\alpha_1, \id(\alpha)/f, \alpha/\alpha_2, \Phi_2[\alpha/\alpha_2]]
-    = N}
-
-    (\Phi_1 \triangleright J(\alpha_1.\alpha_2.\Phi_1 \triangleright Q \triangleleft \Phi_2, \alpha.
+%%     = N}
+
+%%     (\Phi_1 \triangleright J(\alpha_1.\alpha_2.\Phi_1 \triangleright Q \triangleleft \Phi_2, \alpha.
 \lambda^\triangleright \Phi_1[\alpha/\alpha_1].
@@
 N, \alpha_1,f,\alpha_2) \triangleleft \Phi_2)[\Phi_1[\alpha/\alpha_1],\alpha/\alpha_1, \id(\alpha)/f, \alpha/\alpha_2, \Phi_2[\alpha/\alpha_2]]\\
-    = \Phi_1[\alpha/\alpha_1] \triangleright J(\alpha_1.\alpha_2.\Phi_1 \triangleright Q \triangleleft \Phi_2, \alpha.
+%%     = \Phi_1[\alpha/\alpha_1] \triangleright J(\alpha_1.\alpha_2.\Phi_1 \triangleright Q \triangleleft \Phi_2, \alpha.
 \lambda^\triangleright \Phi_1[\alpha/\alpha_1].
@@
 N, \alpha,\id(\alpha),\alpha) \triangleleft \Phi_2[\alpha/\alpha_2]\\
-    = \Phi_1[\alpha/\alpha_1] \triangleright
-    (\lambda^\triangleright \Phi_1[\alpha/\alpha_1]. \lambda^\triangleleft \Phi_2[\alpha/\alpha_2]. N)
-    \triangleleft \Phi_2[\alpha/\alpha_2]\\
-    = N
-
-    \inferrule*[right=Unit$\eta$]
-    {\Phi_1, \alpha_1 : \cat C, f : \alpha_1 \to_{\cat C} \alpha_2, \alpha_2 : \cat C, \Phi_2 \vdash M : Q}
-    {M =
-      J_\to'(\Phi_1, \alpha_1 : \cat C, f : \alpha_1 \to_{\cat C} \alpha_2, \alpha_2 : \cat C, \Phi_2, Q, \alpha.
+%%     = \Phi_1[\alpha/\alpha_1] \triangleright
+%%     (\lambda^\triangleright \Phi_1[\alpha/\alpha_1]. \lambda^\triangleleft \Phi_2[\alpha/\alpha_2]. N)
+%%     \triangleleft \Phi_2[\alpha/\alpha_2]\\
+%%     = N
+
+%%     \inferrule*[right=Unit$\eta$]
+%%     {\Phi_1, \alpha_1 : \cat C, f : \alpha_1 \to_{\cat C} \alpha_2, \alpha_2 : \cat C, \Phi_2 \vdash M : Q}
+%%     {M =
+%%       J_\to'(\Phi_1, \alpha_1 : \cat C, f : \alpha_1 \to_{\cat C} \alpha_2, \alpha_2 : \cat C, \Phi_2, Q, \alpha.
 M[\Phi_1[\alpha/\alpha_1],\alpha/\alpha_1,\id(\alpha), \alpha/\alpha_2,\Phi_2[\alpha/\alpha_2]])}
 
-    M
-    =(\triangleleft\beta\triangleright\beta)
-    \Phi_1 \triangleright
-    (\lambda^\triangleright \Phi_1. \lambda^\triangleleft \Phi_2. M)
-    \triangleleft \Phi_2
-    = \Phi_1 \triangleright
-    J_\to(\alpha_1.\alpha_2.Q, \alpha.
+%%     M
+%%     =(\triangleleft\beta\triangleright\beta)
+%%     \Phi_1 \triangleright
+%%     (\lambda^\triangleright \Phi_1. \lambda^\triangleleft \Phi_2. M)
+%%     \triangleleft \Phi_2
+%%     = \Phi_1 \triangleright
+%%     J_\to(\alpha_1.\alpha_2.Q, \alpha.
 \lambda^\triangleright \Phi_1[\alpha/\alpha_1].
@@
 M[\alpha/\alpha_1, \id(\alpha)/f,\alpha/\alpha_2], \alpha_1, f, \alpha_2)
-    \triangleleft \Phi_2
-    =
-    J_\to'(\Phi_1, \alpha_1 : \cat C, f : \alpha_1 \to_{\cat C} \alpha_2, \alpha_2 : \cat C, \Phi_2, Q, \alpha. N)
-  \end{mathpar}  
-  \caption{Admissible Unit Universal Property Rules}
-\end{figure}
-
-First, the syntax for tensor is based on the coend formula that gives
-the definition in $Cat$, but using logical syntax $\exists$ rather
-than integrals because I don't like integral syntax and the syntax
-matches existential quantification in type theory.
-%
-Note that $\exists \beta. ~Q \odot R$ is the application of a single
-connective to $Q$ and $R$, and not separable into a quantifier
-$\exists$ connective and a product connective $\odot$.
-%
-Introduction, elimination and $\beta\eta$ are similar to a combination
-of existential quantification and the tensor product in linear logic.
-%
-The introduction rule is justified by the fact that semantically what
-we have is a transformation
-\[ \textrm{pair} : (\alpha : \cat C,Q,\beta:\cat D,R,\gamma:\cat E) \Rightarrow \exists \beta. R \odot Q \]
-And the introduction rule is just building in a substitution to this
-transformation.
-%
-
-For the elim rule we can pull a similar trick as with the Hom
-type. First, the rule where there's only one variable.
-\[ \inferrule*[right=Tensor elim Simple]
-  {\alpha : \cat C, q : Q, \beta : \cat D, r: R, \gamma : \cat E \vdash N : S}
-  {\alpha : \cat C, x : \exists \beta. Q \odot R, \gamma : \cat E \vdash J_\odot(x, q.\beta.r. N) : S}
-\]
-
-\begin{figure}
-\begin{mathpar}
-    \inferrule*[right=Full Tensor Elim*]
-    {\Phi_1, \alpha:\cat C, q:Q, \beta:\cat D, r:R, \gamma:\cat E \Phi_2\vdash N : S}
-    {\Phi_1, \alpha : \cat C, x : \exists \beta:\cat D. Q \odot R, \gamma : \cat E, \Phi_2
-      \vdash J_\odot'(\Phi_1, \alpha : \cat C, x : \exists \beta:\cat D. Q \odot R, \gamma : \cat E, \Phi_2; S;q.\beta.r.N) : S}
-
-    J_\odot'(\Phi_1, \alpha : \cat C, x : \exists \beta:\cat D. Q \odot R, \gamma : \cat E, \Phi_2;S; q.\beta.r.N)
-    = \Phi_1 \triangleright J_\odot(\alpha.\gamma.\Phi_1 \triangleright S \triangleleft \Phi_2, \alpha.q.\beta.r.\gamma.
+%%     \triangleleft \Phi_2
+%%     =
+%%     J_\to'(\Phi_1, \alpha_1 : \cat C, f : \alpha_1 \to_{\cat C} \alpha_2, \alpha_2 : \cat C, \Phi_2, Q, \alpha. N)
+%%   \end{mathpar}  
+%%   \caption{Admissible Unit Universal Property Rules}
+%% \end{figure}
+
+%% First, the syntax for tensor is based on the coend formula that gives
+%% the definition in $Cat$, but using logical syntax $\exists$ rather
+%% than integrals because I don't like integral syntax and the syntax
+%% matches existential quantification in type theory.
+%% %
+%% Note that $\exists \beta. ~Q \odot R$ is the application of a single
+%% connective to $Q$ and $R$, and not separable into a quantifier
+%% $\exists$ connective and a product connective $\odot$.
+%% %
+%% Introduction, elimination and $\beta\eta$ are similar to a combination
+%% of existential quantification and the tensor product in linear logic.
+%% %
+%% The introduction rule is justified by the fact that semantically what
+%% we have is a transformation
+%% \[ \textrm{pair} : (\alpha : \cat C,Q,\beta:\cat D,R,\gamma:\cat E) \Rightarrow \exists \beta. R \odot Q \]
+%% And the introduction rule is just building in a substitution to this
+%% transformation.
+%% %
+
+%% For the elim rule we can pull a similar trick as with the Hom
+%% type. First, the rule where there's only one variable.
+%% \[ \inferrule*[right=Tensor elim Simple]
+%%   {\alpha : \cat C, q : Q, \beta : \cat D, r: R, \gamma : \cat E \vdash N : S}
+%%   {\alpha : \cat C, x : \exists \beta. Q \odot R, \gamma : \cat E \vdash J_\odot(x, q.\beta.r. N) : S}
+%% \]
+
+%% \begin{figure}
+%% \begin{mathpar}
+%%     \inferrule*[right=Full Tensor Elim*]
+%%     {\Phi_1, \alpha:\cat C, q:Q, \beta:\cat D, r:R, \gamma:\cat E \Phi_2\vdash N : S}
+%%     {\Phi_1, \alpha : \cat C, x : \exists \beta:\cat D. Q \odot R, \gamma : \cat E, \Phi_2
+%%       \vdash J_\odot'(\Phi_1, \alpha : \cat C, x : \exists \beta:\cat D. Q \odot R, \gamma : \cat E, \Phi_2; S;q.\beta.r.N) : S}
+
+%%     J_\odot'(\Phi_1, \alpha : \cat C, x : \exists \beta:\cat D. Q \odot R, \gamma : \cat E, \Phi_2;S; q.\beta.r.N)
+%%     = \Phi_1 \triangleright J_\odot(\alpha.\gamma.\Phi_1 \triangleright S \triangleleft \Phi_2, \alpha.q.\beta.r.\gamma.
 \lambda^\triangleleft \Phi_1.
@@
 
-    J_\odot'(\Phi_1, \alpha : \cat C, x : \exists \beta:\cat D. Q \odot R, \gamma : \cat E, \Phi_2; q.\beta.r.N)[\Phi_1,\alpha/\alpha,(q,\beta,r)/x,\gamma/\gamma,\Phi_2] = N
-
-  \inferrule*[right=Full Tensor elim*]{ S \prof(\alpha:\cat C; \gamma : \cat E)\and
-    \Phi \vdash M : \exists \beta. Q[A/\alpha] \odot R[G/\gamma]\and
-    \alpha : \cat C, q : Q, \beta : \cat D, r: R, \gamma : \cat E \vdash N : S}
-  {\Phi \vdash J_\odot(\alpha. \gamma. S, \alpha.q.\beta.r.\gamma.N, A, M, G) : S[A/\alpha, G/\gamma]}
-
-  J_\odot(\alpha. \gamma. S, \alpha.q.\beta.r.\gamma.N, A, M, G)[\phi] =
-  J_\odot(\alpha. \gamma. S, \alpha.q.\beta.r.\gamma.N, A[d^-\phi], M[\phi], G[d^+\phi])
-
-  %% \inferrule*[right=Tensor elim Sequent]
-  %% {\Phi,\alpha : \cat C, q : Q, \beta : \cat D, r: R, \gamma : \cat E,\Psi \vdash N : S}
-  %% {\Phi, \alpha : \cat C, x : \exists \beta. Q \odot R, \gamma : \cat E,\Psi \vdash J_\odot(x;q.\beta.r. N) : S}
-
-  %% \inferrule*[right=Tensor elim Subst]
-  %% {\Phi,\alpha : \cat C, q : Q, \beta : \cat D, r: R, \gamma : \cat E,\Phi' \vdash N : S
-  %% \and
-  %% \phi : \Psi \to \Phi,\alpha : \cat C, q : Q, \beta : \cat D, r: R, \gamma : \cat E,\Phi'}
-  %% {\Psi  \vdash J_\odot(x;q.\beta.r. N)[\phi] : S}
-
-  \inferrule*[right=Tensor $\beta$]
-  {}{J_\odot(\alpha.\gamma.S, \alpha.q.\beta.r.\gamma.N,A,(M,B,N),G) = N[A/\alpha,M/q,B/\beta,N/r,G/\gamma]}
-
-  \inferrule*[right=Tensor$\eta$]
-  {\alpha : \cat C, x : \exists \beta. Q \odot R, \gamma : \cat E \vdash M : S}
-  {\alpha : \cat C, x : \exists \beta. Q \odot R, \gamma : \cat E \vdash M =
-    J(\alpha.\gamma.S, \alpha.q.\beta.r.\gamma. M[\alpha/\alpha,(q,\beta,r)/x,\gamma/\gamma])
-    : S}
-\end{mathpar}
-\caption{Tensor Admissible Universal Property Rules}
-\label{fig:tensor}
-\end{figure}
+%%     J_\odot'(\Phi_1, \alpha : \cat C, x : \exists \beta:\cat D. Q \odot R, \gamma : \cat E, \Phi_2; q.\beta.r.N)[\Phi_1,\alpha/\alpha,(q,\beta,r)/x,\gamma/\gamma,\Phi_2] = N
+
+%%   \inferrule*[right=Full Tensor elim*]{ S \prof(\alpha:\cat C; \gamma : \cat E)\and
+%%     \Phi \vdash M : \exists \beta. Q[A/\alpha] \odot R[G/\gamma]\and
+%%     \alpha : \cat C, q : Q, \beta : \cat D, r: R, \gamma : \cat E \vdash N : S}
+%%   {\Phi \vdash J_\odot(\alpha. \gamma. S, \alpha.q.\beta.r.\gamma.N, A, M, G) : S[A/\alpha, G/\gamma]}
+
+%%   J_\odot(\alpha. \gamma. S, \alpha.q.\beta.r.\gamma.N, A, M, G)[\phi] =
+%%   J_\odot(\alpha. \gamma. S, \alpha.q.\beta.r.\gamma.N, A[d^-\phi], M[\phi], G[d^+\phi])
+
+%%   %% \inferrule*[right=Tensor elim Sequent]
+%%   %% {\Phi,\alpha : \cat C, q : Q, \beta : \cat D, r: R, \gamma : \cat E,\Psi \vdash N : S}
+%%   %% {\Phi, \alpha : \cat C, x : \exists \beta. Q \odot R, \gamma : \cat E,\Psi \vdash J_\odot(x;q.\beta.r. N) : S}
+
+%%   %% \inferrule*[right=Tensor elim Subst]
+%%   %% {\Phi,\alpha : \cat C, q : Q, \beta : \cat D, r: R, \gamma : \cat E,\Phi' \vdash N : S
+%%   %% \and
+%%   %% \phi : \Psi \to \Phi,\alpha : \cat C, q : Q, \beta : \cat D, r: R, \gamma : \cat E,\Phi'}
+%%   %% {\Psi  \vdash J_\odot(x;q.\beta.r. N)[\phi] : S}
+
+%%   \inferrule*[right=Tensor $\beta$]
+%%   {}{J_\odot(\alpha.\gamma.S, \alpha.q.\beta.r.\gamma.N,A,(M,B,N),G) = N[A/\alpha,M/q,B/\beta,N/r,G/\gamma]}
+
+%%   \inferrule*[right=Tensor$\eta$]
+%%   {\alpha : \cat C, x : \exists \beta. Q \odot R, \gamma : \cat E \vdash M : S}
+%%   {\alpha : \cat C, x : \exists \beta. Q \odot R, \gamma : \cat E \vdash M =
+%%     J(\alpha.\gamma.S, \alpha.q.\beta.r.\gamma. M[\alpha/\alpha,(q,\beta,r)/x,\gamma/\gamma])
+%%     : S}
+%% \end{mathpar}
+%% \caption{Tensor Admissible Universal Property Rules}
+%% \label{fig:tensor}
+%% \end{figure}
+
+%% \section{Internal Category Theory}
+
+%% A downside of working synthetically is that since everything we define
+%% must be functorial/natural, we cannot prove any theorems about
+%% functions that are not assumed to be functorial or transformations
+%% that are not assumed ot be natural.
+%% %
+%% These kinds of constructions are surprisingly common in basic category theory.
+%% %
+%% For instance, when defining a left adjoint $F$ to a given right
+%% adjoint functor $U$, the functoriality of $F$ is often not assumed,
+%% rather we are given that each $F a$ has a certain universal mapping
+%% property and this can then be used to show the action of $F$ on
+%% objects extends to an action on morphisms.
+%% %
+%% This definition is quite important for applications of category theory
+%% to type theory: in type theory when a type constructor is defined it
+%% is given introduction, elimination and $\beta\eta$ rules, but no
+%% primitive syntax is given for a functorial action.
+%% %
+%% The fact that functoriality is determined from the universal mapping
+%% property of the type is then a key step in proving the type theory is
+%% sound and complete with respect to models which are expected to have
+%% certain functors on them.
+
+%% As another example, a common lemma states that an \emph{possibly
+%% unnatural} transformation is in fact natural if it is the inverse of a
+%% natural transformation.
+%% %
+%% Neither of these examples can be stated as synthetic theorems, as it
+%% is impossible to quantify over ``possibly unnatural transformations''
+%% or ``functions on objects''.
+%% %
+%% Inside the calculus, we cannot ``see'' that categories have an
+%% underlying set of objects.
+%% %
+%% Fortunately there is a way to use our calculus in a different manner
+%% to study these constructions: by defining categories \emph{internal
+%% to} SCT.
+%% %
+%% This reproduces classical constructions, since there is a model where
+%% a category $\cat C$ is interpreted simply as a set, functors as
+%% functions and profunctors $\harr {\cat C}{\cat D}$ as \emph{spans},
+%% i.e. sets $R$ with (source and target) projections $\cat C \leftarrow
+%% R \rightarrow \cat D$.
+%% %
+%% The proofs read like the classical proofs (with some care additional
+%% attention on the ordering of variable usage), but by virtue of being
+%% syntactic automatically apply to any notion of category internal to a
+%% virtual equipment, which includes classical internal categories,
+%% enriched categories and a broad range of generalized multicategory.
+
+%% We note that a synthetic approach to model this situation may be
+%% possible were we to extend our calculus with a \emph{modality} $|\cat
+%% C|$ that takes a category to the discrete category on the set of
+%% objects. We discuss this in \cref{sec:discussion}.
+
+%% Next, we introduce the notion of category internal to our syntactic
+%% virtual equipment and the \emph{monoids and modules} translation that
+%% shows that, with the inclusion of hom sets, our syntax is already
+%% suitable for doing syntactic category theory.
+
+%% The benefits of our syntax become clear when we define internal
+%% categories, functors, profunctors and transformations: the definitions
+%% read exactly as the classical definitions.
+%% \begin{definition}[Internal Category]
+%%   An \emph{(internal) category} consists of\footnote{the following is
+%%     easily interpreted as a $\Sigma$ type}:
+%%   \begin{enumerate}
+%%   \item A sort $C_0 \isaCat$
+%%   \item A span $C_1 : \harr{C_0}{ C_0}$
+%%   \item Identity arrows
+%%     $C_r : \pendallXdotY {\alpha :  C_0} {\harrapp{C_1} \alpha \alpha}$
+%%   \item Composition of arrows
+%%     \[ C_t : \pendallXdotY \alpha {\homrallXYtoZ \beta {\harrapp{C_1} \alpha \beta} {\homrallXYtoZ \gamma {\harrapp{C_1} \beta \gamma} {\harrapp{C_1} \alpha \gamma}}}\]
+
+%%     %% \[ C_t^\triangleright : \forall \alpha :  C_0. \forall \beta. ({C_1} \beta \alpha) \triangleright \forall \gamma. ({C_1} \gamma \beta) \triangleright {C_1} \gamma \alpha \]
+
+%%     %% or
+
+%%     %% \[ C_t^\triangleleft : \forall \alpha. \forall \beta. (\forall \gamma. {C_1} \alpha \gamma \triangleleft {C_1} \beta \gamma) \triangleleft {C_1} \alpha \beta
+%%     %% \]
+
+%%     %% or using tensors and covariant homs
+
+%%     %% \[ C_t^\odot : \forall \alpha. \forall \gamma. (\exists \beta. {C_1}\alpha\beta\odot {C_1}\beta\gamma)\triangleright {C_1}\alpha \gamma \]
+
+%%     %% or using tensors and contravariant homs
+
+%%     %% \[ C_t^\odot : \forall \gamma. \forall \alpha. {C_1}\alpha\gamma \triangleleft (\exists \beta. {C_1}\alpha\beta\odot {C_1}\beta\gamma)\]
+%%   \item Satisfying right unitality, i.e., for any $\Phi \vdash f :
+%%     \harrapp {C_1} A B$, we'd like $f; r(B) = f$. We can formalize this
+%%     using an equality of two transformations of type $\pendallXdotY
+%%     \alpha {\homrallXYtoZ \beta {\harrapp{C_1} \alpha \beta} {\harrapp{C_1} \alpha \beta}}$
+
+%%     \[ \pendlambdaXdotY \alpha {\homrlambdaXatYdotZ f \beta {\homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {C_t} \alpha} {f} \beta} {\pendappXtoY {C_r} \beta} \beta}}
+%%     = \pendlambdaXdotY \alpha {\homrlambdaXatYdotZ f \beta f} \]
+
+%%     These kinds of equalities between 2-cells are very common in
+%%     internal category theory, so to save space we will suppress the
+%%     sequence of $\lambda$s that start such equalities. So the above can simply be written as
+
+%%     \[ { {\homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {C_t} \alpha} {f} \beta} {\pendappXtoY {C_r} \beta} \beta}} = { f}\]
+
+%%   \item and left unitality (an equality of the same type as right unitality):
+
+%%     \[ { {\homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {C_t} \alpha} {\pendappXtoY {C_r} \alpha} \alpha} f \beta}} = { f}\]
+%%   \item and associativity, which is an equality at type
+%%     \[ \pendallXdotY \alpha {\homrallXYtoZ \beta {\harrapp{C_1} \alpha \beta} {\homrallXYtoZ \gamma {\harrapp{C_1} \beta \gamma} {\homrallXYtoZ \delta {\harrapp{C_1} \gamma \delta} {\harrapp{C_1} \alpha \delta}}}}\]
+
+%%     \[ \homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {C_t}\alpha} f \beta} {(\homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {C_t}\beta} g \gamma} h \delta)} \delta
+%%     =
+%%     \homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {C_t}\alpha} {(\homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {C_t}\alpha} f \beta} g \gamma)} \gamma} h \delta \]
+%%   \end{enumerate}
+%% \end{definition}
+
+%% We can define an internal functor to be a pair of an object function
+%% and a function on arrows that preserves identity and 
+%% \begin{definition}[Internal Functor]
+%%   A functor $F$ between internal categories $C$ and $D$ consists of
+%%   \begin{enumerate}
+%%   \item a map of objects $F_0 : \varr {C_0} {D_0}$
+%%   \item a map of arrows
+%%     $F_1 : \pendallXdotY {\alpha}
+%%     {\homrallXYtoZ \beta {\harrapp{C_1} \alpha \beta} {\harrapp{D_1} {F_0 \alpha}{F_0 \beta}}}$
+%%   \item That preserves identity. This is given by equality between the
+%%     following terms of type $\forall \alpha:{{C_0}}. D_1 {(F_0 \alpha)}{(F_0\alpha)}$:
+%%     \[ \homrappXtoYatZ {\pendappXtoY {F_1}\alpha} {\pendappXtoY {C_r} \alpha} \alpha = \pendappXtoY {D_r}{F_0\alpha} \]
+%%   \item And preserves composition.
+%%     This is given by an equality between the following terms of type
+
+%%     \[ \pendallXdotY \alpha
+%%     \homrallXYtoZ {\beta} {\harrapp{C_1} \alpha \beta} {\homrallXYtoZ \gamma {\harrapp{C_1} \beta \gamma} {\harrapp{D_1} {F_0 \alpha} {F_0 \gamma}}}\]
+
+%%     First, we can compose in $C$, and then map using $F$:
+%%     \[ \homrappXtoYatZ {\pendappXtoY{F_1} \alpha} {(\homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {C_t} \alpha} f \beta} g \gamma)}\gamma\]
+
+%%     or we can map the morphisms with $F$, and then compose them in $D$:
+%%     \[ \homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {D_t} {(F_0 \alpha)}} {(\homrappXtoYatZ {\pendappXtoY {F_1} \alpha} f \beta)} {(F_0 \beta)}} {(\homrappXtoYatZ {\pendappXtoY {F_1} \beta} g \gamma)} {{F_0} \gamma}\]
+%%   \end{enumerate}
+%% \end{definition}
+
+%% We can immediately define identity and composition of functors.
+
+%% \begin{definition}
+%%   For any $C : \textrm{Cat}$, we can define an identity functor
+%%   $\textrm{Id}_C : \textrm{Functor} C C$ by
+%%   \begin{enumerate}
+%%   \item $\textrm{Id}_0 = \lambda \alpha:C_0. \alpha$
+%%   \item $\textrm{Id}_1 = \pendlambdaXdotY \alpha
+%%     {\homrlambdaXatYdotZ f {\alpha'} f}$
+%%   \item Preservation of identity requires proving
+%%     \[
+%%     \lambda \alpha. \pendappXtoY {C_r}{\textrm{Id}_0\alpha} \]
+%%     is equal to
+%%     \[ \lambda \alpha.
+%%     \homrappXtoYatZ {\pendappXtoY {\textrm{Id}_1} \alpha} {(\pendappXtoY {C_r}\alpha)} \alpha\]
+%%     Both sides simply $\beta$ reduce to $ \lambda \alpha. \pendappXtoY {C_r} \alpha$
+%%     and the result follows by reflexivity.
+%%   \item Preservation of composition requires proving (under a $\lambda \alpha. \homrlambdaXatYdotZ f \beta \homrlambdaXatYdotZ g \gamma$)
+%%     \[
+%%     \homrappXtoYatZ {\pendappXtoY {{\textrm{Id}}_1} \alpha} {(\homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {C_t} \alpha} f \beta} g \gamma)}\gamma
+%%     = \homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {C_t} {{\textrm{Id}}_0 \alpha}} {(\homrappXtoYatZ {\pendappXtoY {{\textrm{Id}}_1} \alpha} f \beta)} {{\textrm{Id}}_0 \beta}} {(\homrappXtoYatZ {\pendappXtoY{{\textrm{Id}}_1} \beta} g \gamma)} {{\textrm{Id}}_0 \gamma}\]
+%%     Both sides $\beta$ reduce to
+
+%%     \[ \homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {C_t} \alpha} f \beta} g \gamma\]
+%%     so the result holds by reflexivity.
+%%   \end{enumerate}
+%% \end{definition}
+
+%% \begin{definition}
+%%   Given categories $C,D,E$ and functors $F : \textrm{Functor} C D$ and
+%%   $G : \textrm{Functor} D E$, we define the composition $G \circ F :
+%%   \textrm{Functor} C E$ as
+%%   \begin{enumerate}
+%%   \item $(G \circ F)_0 = \lambda \alpha. {G_0 ({F_0} \alpha)}$
+%%   \item $(G \circ F)_1 =
+%%     \pendlambdaXdotY \alpha {
+%%     \homrlambdaXatYdotZ f \beta
+%%     {\homrappXtoYatZ {\pendappXtoY {G_1} {F_0 \alpha}} {\homrappXtoYatZ {\pendappXtoY {F_1} \alpha} f \beta} {F_0 \beta}}}$
+%%   \item Preservation of identity requires proving
+%%     \[ \lambda \alpha. \pendappXtoY {E_r} {(G \circ F)_0\alpha} \]
+%%     is equal to
+%%     \[ \lambda \alpha.
+%%     \homrappXtoYatZ {\pendappXtoY {(G \circ F)_1} \alpha} {\pendappXtoY {C_r}\alpha} \alpha \]
+
+%%     Expanding definitions and $\beta$ reducing we need to show
+
+%%     \[ \pendlambdaXdotY\alpha {\pendappXtoY {E_r}{G_0 (F_0 \alpha)}} \]
+%%     is equal to
+%%     \[ \pendlambdaXdotY \alpha {\homrappXtoYatZ {\pendappXtoY {G_1} {F_0 \alpha}} {(\homrappXtoYatZ {\pendappXtoY {F_1} \alpha} {\pendappXtoY {C_r} \alpha} \alpha)} {F_0 \alpha}}\]
+
+%%     Proceeding from right to left applying transitivity:
+%%     \begin{align*}
+%%       &\pendlambdaXdotY \alpha {\homrappXtoYatZ {\pendappXtoY {G_1} {F_0 \alpha}} {(\homrappXtoYatZ {\pendappXtoY {F_1} \alpha} {\pendappXtoY {C_r} \alpha} \alpha)} {F_0 \alpha}}\\
+%%       &=\pendlambdaXdotY \alpha {\homrappXtoYatZ {\pendappXtoY {G_1} {F_0 \alpha}} {(\pendappXtoY{(\pendlambdaXdotY \alpha {\homrappXtoYatZ {\pendappXtoY {F_1} \alpha} {\pendappXtoY {C_r} \alpha} \alpha})}\alpha)}} {F_0 \alpha}\\
+%%       &=\pendlambdaXdotY \alpha {\homrappXtoYatZ {\pendappXtoY {G_1} {F_0 \alpha}} {(\pendappXtoY{(\pendlambdaXdotY \alpha {{\pendappXtoY {D_r} {F_0 \alpha}}})}\alpha)}} {F_0 \alpha}\tag{*}\\
+%%       &=\pendlambdaXdotY \alpha {\homrappXtoYatZ {\pendappXtoY {G_1} {F_0 \alpha}} {\pendappXtoY {D_r} {F_0 \alpha}}} {F_0 \alpha}\\
+%%       &= \pendlambdaXdotY \alpha
+%%       \pendappXtoY {(\pendlambdaXdotY \beta {{\homrappXtoYatZ {\pendappXtoY {G_1} \beta} {\pendappXtoY {D_r} \beta} \beta}})} {F_0 \alpha}\\
+%%       &= \pendlambdaXdotY \alpha
+%%       \pendappXtoY {(\pendlambdaXdotY \beta \pendappXtoY {E_r} {G_0 \beta})} {F_0 \alpha}\tag{*}\\
+%%       &= \pendlambdaXdotY \alpha \pendappXtoY {E_r} {G_0 (F_0 \beta)}\\
+%%     \end{align*}
+
+%%     Where the two marked lines follow from congruence using the
+%%     assumptions about $F,G$.
+
+%%   \item Preservation of Composition.
+
+%%     \[ \homrappXtoYatZ {\pendappXtoY{(G \circ F)_1} \alpha} {(\homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {C_t} \alpha} f \beta} g \gamma)}\gamma\]
+%%     equals
+
+%%     \[ \homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {E_t} {(G \circ F)_0 \alpha}} {(\homrappXtoYatZ {\pendappXtoY {(G\circ F)_1} \alpha} f \beta)} {(G\circ F)_0 \beta}} {(\homrappXtoYatZ {\pendappXtoY {(G\circ F)_1} \beta} g \gamma)} {{F_0} \gamma}\]
+
+%%     Reducing, we need to show
+%%     \[ \homrappXtoYatZ {\pendappXtoY{G_1} {F_0 \alpha}} {(\homrappXtoYatZ {\pendappXtoY {F_1} \alpha} {(\homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {C_t}\alpha} f {\beta}} g {\gamma})} \gamma)} {F_0 \gamma} \]
+
+%%     is equal to
+%%     \[ \homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {E_t} {G_0(F_0\alpha)}}
+%%       {({\homrappXtoYatZ {\pendappXtoY {G_1} {F_0 \alpha}} {(\homrappXtoYatZ {\pendappXtoY {F_1} \alpha} f \beta)} {F_0 \beta}})}
+%%       {G_0(F_0\beta)}}
+%%        {({\homrappXtoYatZ {\pendappXtoY {G_1} {F_0 \beta}} {(\homrappXtoYatZ {\pendappXtoY {F_1} \beta} g \gamma)} {F_0 \gamma}})}
+%%        {{F_0} \gamma}\]
+
+%%     Going from left to right:
+%%     \begin{align*}
+%%       &\homrappXtoYatZ {\pendappXtoY{G_1} {F_0 \alpha}} {(\homrappXtoYatZ {\pendappXtoY {F_1} \alpha} {(\homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {C_t}\alpha} f {\beta}} g {\gamma})} \gamma)} {F_0 \gamma}\\
+%%       &=
+%%       \homrappXtoYatZ
+%%       {\pendappXtoY{G_1} {F_0 \alpha}}
+%%       {(({\homrappXtoYatZ {\homrappXtoYatZ{\pendappXtoY{(\pendlambdaXdotY \alpha {\homrlambdaXatYdotZ f \beta {\homrlambdaXatYdotZ g \gamma {\homrappXtoYatZ {\pendappXtoY {F_1} \alpha} {(\homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {C_t}\alpha} f {\beta}} g {\gamma})} \gamma}}})}\alpha} f \beta} g \gamma}))}
+%%       {F_0 \gamma}\\
+%%       &=
+%%       \homrappXtoYatZ
+%%       {\pendappXtoY{G_1} {F_0 \alpha}}
+%%       {(({\homrappXtoYatZ {\homrappXtoYatZ{\pendappXtoY{(\pendlambdaXdotY \alpha {\homrlambdaXatYdotZ f \beta {\homrlambdaXatYdotZ g \gamma {\homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {D_t} {(F_0 \alpha)}} {(\homrappXtoYatZ {\pendappXtoY {F_1} \alpha} f \beta)} {(F_0 \beta)}} {(\homrappXtoYatZ {\pendappXtoY {F_1} \beta} g \gamma)} {{F_0} \gamma}}}})}\alpha} f \beta} g \gamma}))}
+%%       {F_0 \gamma}\tag{*}\\
+%%       &=
+%%       \homrappXtoYatZ
+%%       {\pendappXtoY{G_1} {F_0 \alpha}}
+%%       {\homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {D_t} {(F_0 \alpha)}} {(\homrappXtoYatZ {\pendappXtoY {F_1} \alpha} f \beta)} {(F_0 \beta)}} {(\homrappXtoYatZ {\pendappXtoY {F_1} \beta} g \gamma)} {{F_0} \gamma}}
+%%       {F_0 \gamma}\\
+%%       &=
+%%       {(({\homrappXtoYatZ {\homrappXtoYatZ{\pendappXtoY{(\pendlambdaXdotY {\alpha'} {\homrlambdaXatYdotZ {f'} {\beta'} {\homrlambdaXatYdotZ {g'} {\gamma'} {{\homrappXtoYatZ {\pendappXtoY {G_1} {\alpha'}} {(\homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {D_t}{\alpha'}} {f'} {\beta'}} {g'} {\gamma'})} {\gamma'}}}}})}{F_0\alpha}} {(\homrappXtoYatZ {\pendappXtoY {F_1} \alpha} f \beta)} {F_0\beta}} {(\homrappXtoYatZ {\pendappXtoY {F_1} \beta} g \gamma)} {F_0\gamma}}))}\\
+%%       &=
+%%       {(({\homrappXtoYatZ {\homrappXtoYatZ{\pendappXtoY{(\pendlambdaXdotY {\alpha'} {\homrlambdaXatYdotZ {f'} {\beta'} {\homrlambdaXatYdotZ {g'} {\gamma'} {{\homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {E_t} {(G_0 {\alpha'})}} {(\homrappXtoYatZ {\pendappXtoY {G_1} {\alpha'}} {f'} {\beta'})} {(G_0 \beta')}} {(\homrappXtoYatZ {\pendappXtoY {G_1} {\beta'}} {g'} {\gamma'})} {{G_0} \gamma'}}}}})}{F_0\alpha}} {(\homrappXtoYatZ {\pendappXtoY {F_1} \alpha} f \beta)} {F_0\beta}} {(\homrappXtoYatZ {\pendappXtoY {F_1} \beta} g \gamma)} {F_0\gamma}}))}\\
+%%       &= \homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {E_t} {G_0(F_0\alpha)}}
+%%       {({\homrappXtoYatZ {\pendappXtoY {G_1} {F_0 \alpha}} {(\homrappXtoYatZ {\pendappXtoY {F_1} \alpha} f \beta)} {F_0 \beta}})}
+%%       {G_0(F_0\beta)}}
+%%        {({\homrappXtoYatZ {\pendappXtoY {G_1} {F_0 \beta}} {(\homrappXtoYatZ {\pendappXtoY {F_1} \beta} g \gamma)} {F_0 \gamma}})}
+%%        {{F_0} \gamma}
+%%     \end{align*}
+%%   \end{enumerate}
+%% \end{definition}
+
+%% \begin{definition}[Internal Profunctor]
+%%   A profunctor $R$ between internal categories $\cat C$ and $\cat D$
+%%   consists of
+%%   \begin{enumerate}
+%%   \item A span $R_s : \harr {C_0} {D_0}$.
+%%   \item A ternary ``mixed composition'' action:
+%%     $R_t :\pendallXdotY \alpha {\homrallXYtoZ {\alpha'} {\harrapp {C_1} {\alpha}{\alpha'}} {\homrallXYtoZ {\beta'} {\harrapp {R_s} {\alpha'}{\beta'}} {\homrallXYtoZ {\beta} {\harrapp {D_1} {\beta'} {\beta}} {\harrapp {R_s} {\alpha}{\beta}}}}}$
+
+%%   \item Satisfying unitality (under a $\pendlambdaXdotY \alpha {\homrlambdaXatYdotZ f \beta {}}$):
+%%     \[
+%%     f = \homrappXtoYatZ {\homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {R_t} \alpha} {\pendappXtoY {C_r} \alpha} \alpha}f \beta} {\pendappXtoY {D_r} \beta} \beta
+%%     \]
+
+%%   \item And associativity (both with a large implicit
+%%     $\pendlambdaXdotY \alpha {\homrlambdaXatYdotZ g \beta \homrlambdaXatYdotZ {g'} {\beta'}
+%%     \homrlambdaXatYdotZ {r} {\beta''}\homrlambdaXatYdotZ {f'} {\alpha''}\homrlambdaXatYdotZ {f} {\alpha'}}$)
+
+%%     % f : a -> a'
+%%     % f' : a' -> a''
+%%     % r : a'' -> b''
+%%     % g' : b'' -> b'
+%%     % g : b' -> b
+%%     %%
+%%     %
+%%     % f o (f' o r o g') o g
+%%     %% and
+%%     % (f o f') o r o (g' o g)
+
+%%     \[ \homrappXtoYatZ {\homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {R_t} {\alpha}} {f} {\alpha'}}{({\homrappXtoYatZ {\homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {R_t} {\alpha'}} {f'} {\alpha''}}r {\beta''}} {g'} {\beta'}})} {\beta'}} {g} {\beta}\]
+%%     is equal to
+%%     \[ \homrappXtoYatZ
+%%        {\homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {R_t} {\alpha}} {(\homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {C_t} {\alpha}} {f} {\alpha'}} {f'} {\alpha''})} {\alpha''}} r {\beta''}}
+%%        {(\homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {D_t} {\beta''}} {g'} {\beta'}} g {\beta})} {\beta}\]
+%%   \end{enumerate}
+%% \end{definition}
+
+%% We can immediately define the Hom profunctor for any category $C$ as
+%% $(\text{Hom}_C R)_s = C_1$ with actions given by $C_t$.
+
+%% Next, we show that Profunctors can be restricted by functors.
+
+%% \begin{definition}[Restriction of Profunctors]
+%%   If $C, D, C', D' : \text{Cat}$, and $F : \text{Functor} C C', G : \text{Functor} D D'$ and
+%%   $R : \text{Prof} C' D'$, then we can define a profunctor $R[F;G] : \text{Prof} C D$ by
+%%   \begin{enumerate}
+%%   \item $R[F;G]_s = \lambda (\alpha:C,\beta:D). R_s (F_0 \alpha) (G_0 \beta)$
+%%   \item with composition action given by
+
+%%     \[ R[F;G]_t =
+%%     \pendlambdaXdotY \alpha {\homrlambdaXatYdotZ f {\alpha'} {\homrlambdaXatYdotZ r {\beta'} {\homrlambdaXatYdotZ g {\beta} {\homrappXtoYatZ {\homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {R_t} {F_0 \alpha}} {(\homrappXtoYatZ {\pendappXtoY {F_1} {\alpha}} f {\alpha'})} {F_0 \alpha'}} r {G_0 \beta'}} {(\homrappXtoYatZ {\pendappXtoY {G_1} {\beta'}} g \beta)} {G_0 \beta}}}} }
+%%  \]
+%%   %% \item $\cat C'$ composition:
+%%   %%   \[ \lambda \alpha. \lambda^\triangleleft (\beta, (f : C_1 \alpha \alpha',\alpha',\alpha', r : R_s \alpha' \beta)). R_l(F_1 f, r) \]
+%%   %% \item $\cat D'$ composition:
+%%   %%   \[ \lambda \alpha. \lambda^\triangleleft ('\beta, (r : R_s \alpha \alpha',\alpha', \beta',h : D_1 \beta' \beta)). R_r(r, G_1 h) \]
+%%   \end{enumerate}
+%%   Identity and associativity properties should follow. TODO
+%% \end{definition}
+
+%% \begin{definition}[Homomorphism]
+%%   Given $R : \text{Prof} C D$ and $S : \text{Prof} C D$, a
+%%   homomorphism $m$ from $R$ to $S$ consists of
+%%   \begin{enumerate}
+%%   \item $m_f : \pendallXdotY \alpha {\homrallXYtoZ \beta {\harrapp R \alpha \beta}{\harrapp S \alpha \beta}}$
+%%     % m_f (R_t f r g) = S_t f (m_f r) g
+%%   \item Satisfying naturality (under a $\ldots$)
+
+%%     \[
+%%     \homrappXtoYatZ {\pendappXtoY{m_f} \alpha} {(\homrappXtoYatZ {\homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {R_t} \alpha} f {\alpha'}} r {\beta'}} g \beta)} \beta
+%%     =
+%%     \homrappXtoYatZ {\homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {S_t} \alpha} {f} {\alpha'}} {(\homrappXtoYatZ {\pendappXtoY {m_f}{\alpha'}} r {\beta'})} {\beta'}} g \beta
+%%     \]
+%%   \end{enumerate}
+%% \end{definition}
+
+%% \begin{definition}
+%%   An isomorphism between profunctors $R$ and $S$ is a mutually inverse
+%%   pair of homomorphisms.
+%% \end{definition}
+
+%% Additionally, we can define natural transformations between functors,
+%% but more generally, we can define a natural element of any
+%% endo-profunctor, where natural transformations from $F$ to $G$ are the
+%% same as natural elements of $\text{Hom}(F-,G=)$.
+
+%% \begin{definition}[Natural Element]
+%%   A natural element $\phi$ of a profunctor $R : \text{Prof}\,C C$ is a term
+%%   \[ \phi : \pendallXdotY \alpha {\harrapp R \alpha \alpha} \]
+%%   that is natural in that
+%%   \[
+%%   \homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {C_t} \alpha} {\pendappXtoY \phi \alpha} \alpha} f {\alpha'}
+%%   = \homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {C_t } \alpha} f {\alpha'}} {\pendappXtoY \phi {\alpha'}}{\alpha'}
+%%   \]
+%% \end{definition}
+
+%% \begin{definition}[Natural Transformation]
+%%   A natural transformation from $F : \text{Functor} \,C\, D$ and $G :
+%%   \text{Functor}\,C\,D$ is a natural element of $\text{Hom}(D)[F;G]$
+%% \end{definition}
+
+%% Composition of natural transformations is then given by $C_t$.
+
+%% Natural elements are closely related to homomorphisms: a natural
+%% element equivalent to a homomorphism out of a Hom profunctor. This is
+%% our first instance of a ``Yoneda'' principle that the behavior of some
+%% structure that varies naturally with the morphisms of a category is
+%% fully determined by its action on the identity morphisms.
+
+%% \begin{theorem}
+%%   Natural elements of $R : \text{Prof}\, C C$ are equivalent to
+%%   homomorphisms from $\text{Hom}_C$ to $R$.
+%% \end{theorem}
+%% \begin{enumerate}
+%% \item Given a natural element $\phi$ of $R$, we can define a
+%%   homomorphism by
+%%   \[ \pendlambdaXdotY \alpha {\homrlambdaXatYdotZ f {\alpha'} \homrappXtoYatZ {\homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {R_t} \alpha} {\pendappXtoY {C_r} \alpha} \alpha}{\pendappXtoY \phi \alpha} {\alpha}} f {\alpha'}} \]
+%% \item Given a homomorphism $m$ from $\text{Hom}(C)$ to $R$, define the
+%%   element as
+%%   \[ \pendlambdaXdotY \alpha {\homrappXtoYatZ {\pendappXtoY {m_f} \alpha} {\pendappXtoY {C_r}\alpha} \alpha}  \]
+
+%%   We should be able to show this is an isomorphism in the logic TODO!
+%% \end{enumerate}
+
+%% %% This last definition is probably unfamiliar to the reader that is not
+%% %% experienced with profunctors.
+%% %% %
+%% %% However, we can show that the traditional definition is equivalent to
+%% %% a natural transformation $\alpha : \cat C_0 \pipe \cdot \vdash
+%% %% t_{\alpha} : F(\alpha) \to_{\cat D} G(\alpha')$.
+
+%% %% \begin{definition}[Traditional Natural Transformation]
+%% %%   A traditional natural transformation from $\alpha : \cat C_0 \vdash
+%% %%   F(\alpha) : \cat D_0$ to $\alpha : \cat \vdash G(\alpha') : \cat
+%% %%   D_0$ consists of
+%% %%   \begin{enumerate}
+%% %%   \item A term $\cdot \vdash t_{\alpha} : F(\alpha) \to_{\cat D} G(\alpha)$
+%% %%   \item Such that
+%% %%     \[ \phi : \alpha \to_{\cat C} \alpha' \vdash t_{\alpha'} \circ F_1(\phi) \equiv G_1(\phi) \circ t_{\alpha} : F(\alpha) \to_{\cat D} G(\alpha')
+%% %%     \]
+%% %%   \end{enumerate}
+%% %% \end{definition}
+
+%% %% \begin{theorem}[Equivalent Presentations of Natural Transformations]
+%% %%   There is a bijection between traditional natural transformations $t
+%% %%   : F \to G$ and profunctor homomorphisms $h : \alpha \to_{\cat C}
+%% %%   \alpha' \vdash F(\alpha) \to_{\cat D} G(\alpha')$ given by
+%% %%   \begin{enumerate}
+%% %%   \item From a natural transformation $t$ we define a homomorphism
+%% %%     $h(t)$ by composing on the $F$ side. The choice of this side is
+%% %%     arbitrary and we could just as easily have used $G$ since they are
+%% %%     equal by naturality.
+%% %%     \[ \phi : \alpha \to_{\cat C} \alpha' \vdash t_{\alpha'} \circ F_1(\phi) : F(\alpha) \to_{\cat D} G_1(\alpha') \]
+%% %%   \item From a homomorphism $\phi.h$, we define a natural transformation $t(h)$ as
+%% %%     \[ \alpha \pipe \cdot \vdash h[{\id_{\alpha}}/\phi] \]
+%% %%   \end{enumerate}
+%% %% \end{theorem}
+%% %% \begin{proof}
+%% %%   First we show that the maps $h(\cdot),t(\cdot)$ actually produce
+%% %%   homomorphisms and natural transformations.
+%% %%   \begin{enumerate}
+%% %%   \item On the left, we need to show $h(t)_{\phi} \circ F_1(\psi) =
+%% %%     h(t)_{|phi \circ \psi}$. Expanding the definition, we just need
+%% %%     associativity and functoriality of $F_1$:
+%% %%     \begin{align*}
+%% %%       h(t)_{\phi} \circ F_1(\psi) &\equiv (t_{\alpha'} \circ F_1(\phi)) \circ F_1(\psi)\\
+%% %%       &\equiv t_{\alpha'} \circ F_1(\phi \circ \psi)\\
+%% %%       &\equiv h(t)_{\phi \circ \psi}
+%% %%     \end{align*}
+%% %%     on the right, we use naturality:
+%% %%     \begin{align*}
+%% %%       G_1(\theta) \circ h(t)_{\phi} &\equiv G_1(\theta) \circ (t_{\alpha'} \circ F_1(\phi))\\
+%% %%       &\equiv (t_{\alpha''} \circ F_1(\theta)) \circ F_1(\phi)\\
+%% %%       &\equiv t_{\alpha''} \circ F_1(\theta \circ \phi)\\
+%% %%       &\equiv t_{\theta \circ \phi}
+%% %%     \end{align*}
+%% %%   \item In the other direction, to prove naturality:
+%% %%     \begin{align*}
+%% %%       h[\id_\alpha/\phi] \circ F_1(\psi) &\equiv h[\id_\alpha \circ \psi/\phi]\\
+%% %%       &\equiv h[\psi/\phi]\\
+%% %%       &\equiv h[\psi \circ \id_{\alpha'}/\phi]\\
+%% %%       &\equiv G_1(\psi) \circ h[\id_{\alpha}/\phi]
+%% %%     \end{align*}
+%% %%   \end{enumerate}
+%% %%   Next we seek to prove that the functions are mutually inverse.
+%% %%   \begin{enumerate}
+%% %%   \item Round trip for homomorphisms:
+%% %%     \begin{align*}
+%% %%       h(t(h))_{\phi} &= t(h)_{\alpha} \circ F_1[\phi]\\
+%% %%       &= h[\id_{\alpha}] \circ F_1[\phi]\\
+%% %%       &= h[\id \circ \phi]\\
+%% %%       &= h[\phi]
+%% %%     \end{align*}
+%% %%   \item Round trip for natural transformations uses the fact that
+%% %%     functors preserve identity.
+%% %%     \begin{align*}
+%% %%       t(h(t))_{\alpha} &= h(t)_{\id_{\alpha}}\\
+%% %%       &= t_{\alpha} \circ F_1[\id_{\alpha}]\\
+%% %%       &= t_{\alpha} \circ \id_{F_0\alpha}\\
+%% %%       &= t_{\alpha}
+%% %%     \end{align*}
+%% %%   \end{enumerate}
+%% %% \end{proof}
+
+%% \subsection{Internal Category Translation}
+
+%% \subsection{Universal Properties}
+
+%% Since we have profunctors as a primitive notion, the most natural
+%% presentation of universal properties is given by the notion of a
+%% functor representing a profunctor.
+
+%% First, we have the ``algebraic'' notion of a functor representing a
+%% profunctor, i.e., when a profunctor $R : C\pto D$ is equivalent to
+%% $Hom_D(-,G=)$.
+%% \begin{definition}[Right Representability]
+%%   Let $C, D : \text{Cat}$ and $Q : \textrm{Prof}\,C D$ be a profunctor
+%%   between them. Then a \emph{right representation} of $Q$ consists of
+%%   \begin{enumerate}
+%%   \item A functor $G : \textrm{Functor}\, D C$
+%%   \item An isomorphism $I,I^-$ between $Q$ and $\text{Hom}(C)[Id;G]$
+%%   \end{enumerate}
+%% \end{definition}
+
+%% This is a nice definition for applications, but a lot of the structure
+%% (homomorphism, functoriality of G) is determined by other components
+%% of the structure. This can be simplified to the notion of a
+%% ``universal morphism''.
+
+%% \begin{definition}[Universal Morphism]
+%%   A profunctor $Q : \textrm{Prof}\, C D$ is right representable when
+%%   there exists
+%%   \begin{enumerate}
+%%   \item A function on objects $G_0 : \varr {D_0} {C_0}$
+%%   \item A ``universal morphism'', i.e., a term $\epsilon :
+%%     \pendallXdotY \beta {\harrapp Q {G_0 \beta} \beta}$ (not
+%%     presupposed to be natural).
+%%   \item An ``introduction rule'', i.e., a term $I : \pendallXdotY
+%%     \alpha {\homrallXYtoZ \beta {\harrapp {Q_s}\alpha \beta} {\harrapp
+%%         {C_1} \alpha {G_0\beta}}}$ (not presupposed to satisfy the
+%%     homomorphism condition).
+%%   \item Such that composition with $\epsilon$ is a left and right
+%%     inverse to $I$. In more detail, we have two equalities.
+
+%%     \begin{enumerate}
+%%     \item First, the $\beta$ rule states that composing the universal
+%%       morphism with the introduction rule applied to a morphism $q : \harrapp {Q} \alpha {\beta}$ is
+%%       equal to the original $q$:
+
+%%       \[ q = \homrappXtoYatZ {\homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {Q_t} \alpha} {(\homrappXtoYatZ {\pendappXtoY I \alpha} {q} \beta)} {G_0 \beta}} {\pendappXtoY \epsilon \beta} {\beta}} {\pendappXtoY {D_r} \beta} {\beta} \]
+
+%%     \item Second, the $\eta$ rule says that every morphism $f :
+%%       \harrapp {C_1} \alpha {G_0 \beta}$ can be reconstructed using
+%%       $I$ after postcomposing with $\epsilon$:
+
+%%       \[ f = \homrappXtoYatZ {\pendappXtoY I \alpha} {({\homrappXtoYatZ {\homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {Q_t} \alpha} {f} {G_0 \beta}} {\pendappXtoY \epsilon \beta} {\beta}} {\pendappXtoY {D_r} \beta} {\beta}  })} {\beta}\]
+
+%%       As a special case, if $f = \pendappXtoY {C_r} {G_0 \beta}$, this shows that $\homrappXtoYatZ {\pendappXtoY I {G_0 \beta}} {\pendappXtoY \epsilon \beta} \beta = \pendappXtoY {C_r} {G_0 \beta}$
+%%     \end{enumerate}
+%%   \end{enumerate}
+%% \end{definition}
+
+%% These two notions turn out to be equivalent because the ``extra''
+%% information in the hom isomorphism is in fact fully determined by the
+%% data of a universal morphism: the functoriality of $G$ and the
+%% naturality of $I,I_{inv}$.
+%% %
+%% The proof can be broken up into 3 conceptual steps:
+%% \begin{enumerate}
+%% \item First,, the functoriality of $G$ is derivable from $I,\epsilon$.
+%% \item Next, $I_{inv}$ is a homomorphism automatically since composition
+%%   with a morphism is automatically natural, where $I_{inv}$ is defined as
+%%   \[\pendlambdaXdotY \alpha {\homrlambdaXatYdotZ f {\beta} \homrappXtoYatZ {\homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {Q_t} \alpha} f \beta}{\pendappXtoY \epsilon \beta} {\beta}} {\pendappXtoY {D_r }\beta} {\beta}} \]
+
+%%   \item Finally, $I$ is a homomorphism because the inverse of a
+%%     natural transformation is always natural.
+%% \end{enumerate}
+
+%% %% Part of this equivalence is the \emph{Yoneda Lemma}, which we state
+%% %% now in its internal, parameterized form.
+%% %% The usual non-parameterized Yoneda lemma can be formed if we have a
+%% %% unit sort.
+
+%% \begin{lemma}[Universal Constructions are functorial]
+%%   For any universal construction $(G_0,\epsilon,I)$ for $Q :
+%%   \text{Prof}\,C\,D$, we can extend $G_0$ to a functor by defining
+
+%%   \[ G_1 = \pendlambdaXdotY \beta {\homrlambdaXatYdotZ f {\beta'}
+%%     {\homrappXtoYatZ {\pendappXtoY I {G_0\beta}} {(\homrappXtoYatZ {\homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {Q_t} {G_0\beta}} {\pendappXtoY {C_r} {G_0\beta}} {G_0\beta}} {\pendappXtoY \epsilon {\beta}} \beta} {f} {\beta'})} {\beta'}}
+%%   }\]
+
+%%   Furthermore, $I,I^-1$ are homomorphisms with respect to the induced
+%%   profunctor structure.
+%% \end{lemma}
+%% \begin{proof}
+%%   \begin{enumerate}
+%%   \item To prove $G_1$ preserves identity we need to show
+%%     \[ {\homrappXtoYatZ {\pendappXtoY I {G_0\beta}} {(\homrappXtoYatZ {\homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {Q_t} {G_0\beta}} {\pendappXtoY {C_r} {G_0\beta}} {G_0\beta}} {\pendappXtoY \epsilon {\beta}} \beta} {\pendappXtoY {D_r} {\beta}} {\beta})} {\beta}} = \pendappXtoY{C_r} {G_0\beta}\]
+%%     which is precisely the $\eta$ principle.
+
+%%   \item Next, we show $I_{inv}$ is a homomorphism from $\text{Hom}(C)[\text{Id}(C);G]$ to $Q$.
+%%     All of this is under the binders
+%%     \[\lambdatrinary {\alpha : C} {f : \harrapp {C_s} \alpha {\alpha'}}
+%%     {\alpha' : C} {g : \harrapp{C_s} {\alpha'} {G_0 \beta'}}
+%%     {\beta' : D} {h : \harrapp {D_s} {\beta'} {\beta}}
+%%     {}\]
+
+%%     % given
+%%     % f : a -> a' in C
+%%     % g : a' -> G b'
+%%     % h : b' -> b
+
+%%     %% I_{inv} (f ;C (g ;C G_1 h)) = Q_t f (I_{inv} g) h
+
+%%     \begin{align*}
+%%       & \appunary
+%%       {I_{inv}}
+%%       \alpha
+%%       {(\appbinary {C_t} \alpha f {\alpha'} {(
+%%           \appbinary {C_t} {\alpha'} g {G_0\beta'}
+%%           {(\appunary {G_1} {\beta'} h \beta)} {G_0 \beta})} {G_0 \beta})}
+%%       {\beta}\\
+%%       &=
+%%       \apptrinary {Q_t} \alpha
+%%       {(\appbinary {C_t} \alpha f {\alpha'} {(
+%%           \appbinary {C_t} {\alpha'} g {G_0\beta'}
+%%           {(\appunary {G_1} {\beta'} h \beta)} {G_0 \beta})} {G_0 \beta})} \beta
+%%       {\pendappXtoY \epsilon \beta} \beta
+%%       {\pendappXtoY {D_r} \beta} {\beta}\\
+%%       &=
+%%       \apptrinary {Q_t} \alpha
+%%       {(\appbinary {C_t} \alpha f {\alpha'} {(
+%%           \appbinary {C_t} {\alpha'} g {G_0\beta'}
+%%           {(\appunary {G_1} {\beta'} h \beta)} {G_0 \beta})} {G_0 \beta})} \beta
+%%       {\pendappXtoY \epsilon \beta} \beta
+%%       {(\appbinary {D_t} \beta {\pendappXtoY {D_r} \beta} \beta {\pendappXtoY {D_r} \beta} \beta)} {\beta}\\
+%%       &=
+%%       \apptrinary {Q_t} \alpha
+%%       {(\appbinary {C_t} \alpha {(\appbinary {C_t} \alpha f {\alpha'} g {G_0 \beta'})} {G_0 \beta'}
+%%         {(\appunary {G_1} {\beta'} h \beta)} {G_0 \beta})} \beta
+%%       {\pendappXtoY \epsilon \beta} \beta
+%%       {(\appbinary {D_t} \beta {\pendappXtoY {D_r} \beta} \beta {\pendappXtoY {D_r} \beta} \beta)} {\beta}\\
+%%       &=
+%%       \apptrinary {Q_t} \alpha
+%%       {(\appbinary {C_t} \alpha f {\alpha'} g {G_0 \beta'})} {G_0 \beta'}
+%%       {(\apptrinary {Q_t} {G_0 \beta'}
+%%         {(\appunary {G_1} {\beta'} h {\beta})} {G_0 \beta}
+%%         {\pendappXtoY \epsilon \beta} \beta
+%%         {\pendappXtoY {D_r} \beta} \beta)} \beta
+%%       {\pendappXtoY {D_r} \beta} \beta
+%%       \\
+%%       &=
+%%       \apptrinary {Q_t} \alpha
+%%       {(\appbinary {C_t} \alpha f {\alpha'} g {G_0 \beta'})} {G_0 \beta'}
+%%       {(\appunary {I_{inv}} {G_0 \beta'} {(\appunary {G_1} {\beta'} h \beta)} \beta)} \beta
+%%       {\pendappXtoY {D_r} \beta} \beta
+%%       \\
+%%       &=
+%%       \apptrinary {Q_t} \alpha
+%%       {(\appbinary {C_t} \alpha f {\alpha'} g {G_0 \beta'})} {G_0 \beta'}
+%%       {(\appunary {I_{inv}} {G_0 \beta'} {(\appunary {I} {G_0\beta'} {(\apptrinary {Q_t} {G_0 \beta'} {\pendappXtoY {C_r} {G_0 \beta'}} {G_0 \beta'} {\pendappXtoY \epsilon {\beta'}} {\beta'} h \beta)} \beta)} \beta)} \beta
+%%       {\pendappXtoY {D_r} \beta} \beta
+%%       \\
+%%       &=
+%%       \apptrinary {Q_t} \alpha
+%%       {(\appbinary {C_t} \alpha f {\alpha'} g {G_0 \beta'})} {G_0 \beta'}
+%%       {(\apptrinary {Q_t} {G_0 \beta'} {\pendappXtoY {C_r} {G_0 \beta'}} {G_0 \beta'} {\pendappXtoY \epsilon {\beta'}} {\beta'} h \beta)} \beta
+%%       {\pendappXtoY {D_r} \beta} \beta
+%%       \\
+%%       &=
+%%       \apptrinary {Q_t} \alpha
+%%       {(\appbinary {C_t} \alpha {(\appbinary {C_t} \alpha f {\alpha'} g {G_0 \beta'})} {G_0 \beta'}
+%%         {\pendappXtoY {C_r} {G_0 \beta'}} {G_0 \beta'})} {G_0 \beta'}
+%%       {\pendappXtoY \epsilon {\beta'}} {\beta'}
+%%       {(\appbinary {D_t} {\beta'} h \beta {\pendappXtoY {D_r} \beta} \beta)} \beta\\
+%%       &=
+%%       \apptrinary {Q_t} \alpha
+%%       {(\appbinary {C_t} \alpha f {\alpha'} g {G_0 \beta'})} {G_0 \beta'}
+%%       {\pendappXtoY \epsilon {\beta'}} {\beta'}
+%%       {(\appbinary {D_t} {\beta'} {\pendappXtoY {D_r} {\beta'}} {\beta'} h \beta )} \beta\\
+%%       &=
+%%       \apptrinary {Q_t} \alpha
+%%       f {\alpha'}
+%%       {(\apptrinary {Q_t} {\alpha'} g {G_0 \beta'} {\pendappXtoY \epsilon {\beta'}} {\beta'}{\pendappXtoY {D_r} {\beta'}} {\beta'})} {\beta'}
+%%       h \beta\\
+%%       &= \homrappXtoYatZ {\homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {Q_t} \alpha} f {\alpha'}} {(\homrappXtoYatZ {\pendappXtoY {I_{inv}} {\alpha'}} {g} {\beta'})} {\beta'}} {h} {\beta}\\
+%%     \end{align*}
+
+%%   \item To prove $G_1$ preserves composition given $f : \harrapp {D_1} {\beta} {\beta'}$, $f' : \harrapp {D_1} {\beta'} {\beta''}$:
+%%     \begin{align*}
+%%       & \appunary {G_1} {\beta} {\appbinary {D_t} {\beta} f {\beta'} {f'} {\beta''}} {\beta'}\\
+%%       &= \appunary {I} {G_0 \beta} {(\apptrinary {Q_t} {G_0 \beta} {\pendappXtoY {C_r} {G_0 \beta}} {G_0 \beta} {\pendappXtoY \epsilon {\beta}} {\beta} {({\appbinary {D_t} {\beta} f {\beta'} {f'} {\beta''}})} {\beta''})} {\beta''}\\
+%%       &= \appunary {I} {G_0 \beta} {(\apptrinary {Q_t} {G_0 \beta} {(\appbinary {C_t} {G_0 \beta} {\pendappXtoY {C_r} {G_0 \beta}} {G_0 \beta} {\pendappXtoY {C_r} {G_0 \beta}} {G_0 \beta})} {G_0 \beta} {\pendappXtoY \epsilon {\beta}} {\beta} {({\appbinary {D_t} {\beta} f {\beta'} {f'} {\beta''}})} {\beta''})} {\beta''}\\
+%%       &=
+%%       \appunary {I} {G_0 \beta}
+%%       {(\apptrinary {Q_t} {G_0 \beta}
+%%         {\pendappXtoY {C_r} {G_0 \beta}} {G_0 \beta}
+%%         {(\apptrinary {Q_t} {G_0 \beta}
+%%           {\pendappXtoY {C_r} {G_0 \beta}} {G_0 \beta}
+%%           {\pendappXtoY \epsilon {\beta}} {\beta}
+%%           f {\beta'})} {\beta'}
+%%         {f'} {\beta''}
+%%         )}
+%%       {\beta''}\\
+%%       &=
+%%       \appbinary {C_t} {G_0 \beta}
+%%       {\pendappXtoY {C_r} {G_0 \beta}} {G_0 \beta}
+%%       {(\appbinary {C_t} {G_0 \beta}
+%%         {(\appunary {I} {G_0 \beta}
+%%         {(\apptrinary {Q_t} {G_0 \beta}
+%%           {\pendappXtoY {C_r} {G_0 \beta}} {G_0 \beta}
+%%           {\pendappXtoY \epsilon {\beta}} {\beta}
+%%           f {\beta'})} {\beta'})} {G_0 \beta'}
+%%         {(\appunary {G_1} {\beta'} {f'} {\beta''})} {G_0 \beta''})}
+%%       {G_0 \beta''}\\
+%%       &=
+%%       \appbinary {C_t} {G_0 \beta}
+%%         {(\appbinary {C_t} {G_0 \beta}
+%%           {\pendappXtoY {C_r} {G_0 \beta}} {G_0 \beta}
+%%           {(\appbinary {C_t} {G_0 \beta}
+%%             {(\appunary {I} {G_0 \beta} {\pendappXtoY \epsilon {\beta}} {\beta})} {G_0 \beta}
+%%             {(\appunary {G_1} \beta f {\beta'})} {G_0 \beta'})} {G_0 \beta'})} {G_0 \beta'}
+%%         {(\appunary {G_1} {\beta'} {f'} {\beta''})} {G_0 \beta''}\\
+%%       &=
+%%       \appbinary {C_t} {G_0 \beta}
+%%         {(\appbinary {C_t} {G_0 \beta}
+%%             {(\appunary {I} {G_0 \beta} {\pendappXtoY \epsilon {\beta}} {\beta})} {G_0 \beta}
+%%             {(\appunary {G_1} \beta f {\beta'})} {G_0 \beta'})} {G_0 \beta'}
+%%         {(\appunary {G_1} {\beta'} {f'} {\beta''})} {G_0 \beta''}\\
+%%       &=
+%%       \appbinary {C_t} {G_0 \beta}
+%%         {(\appbinary {C_t} {G_0 \beta}
+%%             {\pendappXtoY {C_r} {G_0 \beta}} {G_0 \beta}
+%%             {(\appunary {G_1} \beta f {\beta'})} {G_0 \beta'})} {G_0 \beta'}
+%%         {(\appunary {G_1} {\beta'} {f'} {\beta''})} {G_0 \beta''}\\
+%%       &=
+%%       \appbinary {C_t} {G_0 \beta}
+%%       {(\appunary {G_1} \beta f {\beta'})} {G_0 \beta'}
+%%         {(\appunary {G_1} {\beta'} {f'} {\beta''})} {G_0 \beta''}\\
+%%       %% &\homrappXtoYatZ {\pendappXtoY {G_1} \beta} {(\homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {D_t} {\beta}} {f} {\beta'}} {f'} {\beta''})} {\beta''}\\
+%%       %% &={\homrappXtoYatZ {\pendappXtoY I {G_0\beta}} {(\homrappXtoYatZ {\homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {Q_t} {G_0\beta}} {\pendappXtoY {C_r} {G_0\beta}} {G_0\beta}} {\pendappXtoY \epsilon {\beta}} \beta} {(\homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {D_t} {\beta}} {f} {\beta'}} {f'} {\beta''})} {\beta''})} {\beta''}}\\
+%%       %% &={\homrappXtoYatZ {\pendappXtoY I {G_0\beta}} {(\homrappXtoYatZ {\homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {Q_t} {G_0\beta}} {({\pendappXtoY {C_r} {G_0\beta}})} {G_0\beta}} {\pendappXtoY \epsilon {\beta}} \beta} {(\homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {D_t} {\beta}} {f} {\beta'}} {f'} {\beta''})} {\beta''})} {\beta''}}\\
+%%     \end{align*}
+%%     %% \[
+%%     %%   {\homrappXtoYatZ {\pendappXtoY I {G_0\beta}} {(\homrappXtoYatZ {\homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {Q_t} {G_0\beta}} {\pendappXtoY {C_r} {G_0\beta}} {G_0\beta}} {\pendappXtoY \epsilon {\beta}} \beta} {(\homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {D_t} {\beta}} {f} {\beta'}} {f'} {\beta''})} {\beta''})} {\beta''}}
+%%     %% \]
+%%   \end{enumerate}
+%% \end{proof}
+
+%% \begin{lemma}[Inverse of homomorphism is a homomorphism]
+%%   Let $\phi$ be a homomorphism from $R$ to $S : \text{Prof} C D$ and
+%%   let $\psi : \pendallXdotY \alpha {\homrallXYtoZ \beta {\harrapp S
+%%       \alpha \beta}{\harrapp R \alpha \beta}}$. If $\psi$ is a
+%%   two-sided inverse to $\phi_f$ then $\psi$ is a homomorphism.
+%% \end{lemma}
+%% \begin{proof}
+%%   We need to show the homomorphism condition for $\psi$:
+%%   \[ \homrappXtoYatZ {\pendappXtoY \psi \alpha} {(\homrappXtoYatZ {\homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {S_t} \alpha} f {\alpha'}} s {\beta'}} g {\beta})} {\beta}
+%%     = \homrappXtoYatZ {\homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {R_t} \alpha} f {\alpha'}} {(\homrappXtoYatZ {\pendappXtoY \psi {\alpha'}} s {\beta'})} {\beta'}} g \beta
+%%     \]
+
+%%     \begin{align*}
+%%       \homrappXtoYatZ {\pendappXtoY \psi \alpha} {(\homrappXtoYatZ {\homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {S_t} \alpha} f {\alpha'}} s {\beta'}} g {\beta})} {\beta}
+%%       &= \homrappXtoYatZ {\pendappXtoY \psi \alpha} {(\homrappXtoYatZ {\homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {S_t} \alpha} f {\alpha'}} {(\homrappXtoYatZ {\pendappXtoY {\phi_f} {\alpha'}} {(\homrappXtoYatZ {\pendappXtoY \psi {\alpha'}} s {\beta'})} {\beta'})} {\beta'}} g {\beta})} {\beta}\\
+%%       &= \homrappXtoYatZ {\pendappXtoY \psi \alpha}
+%%       {(\homrappXtoYatZ {\pendappXtoY {\phi_f} \alpha} {(\homrappXtoYatZ {\homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {R_t} \alpha} f {\alpha'}} {(\homrappXtoYatZ {\pendappXtoY \psi {\alpha'}} s {\beta'})} {\beta'}} g \beta)} \beta)} {\beta}\\
+%%       & = \homrappXtoYatZ {\homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {R_t} \alpha} f {\alpha'}} {(\homrappXtoYatZ {\pendappXtoY \psi {\alpha'}} s {\beta'})} {\beta'}} g \beta
+%%     \end{align*}
+%% \end{proof}
+
+%% %% \subsection{Self-Interpretation: Hom of Internal Profunctors}
+
+%% %% We can define a version of the hom $\homrallXYtoZ \alpha P Q$ on
+%% %% \emph{internal} profunctors using a combination of this connective,
+%% %% and an \emph{equalizer} to enforce a homomorphism condition.
+
+%% %% \begin{definition}[Internal Covariant Hom of Profunctors]
+%% %%   Let $P : \textrm{Prof}\,D\,E$ and $Q : \textrm{Prof}\,C\,E$. Define
+%% %%   the covariant Hom $P \triangleright Q : \textrm{Prof}\,C\,D$ by
+%% %%   \[ (P \triangleright Q)_1 = \lambda (\alpha;\beta).
+\equalizeVofWbyXeqYatZ \phi {\homrallXYtoZ \gamma {\harrapp {P_1} \alpha \gamma} {\harrapp {Q_1} \beta \gamma}} {\psi_{dom}} {\psi_{cod}} {\homrallXYtoZ {\gamma'} {\harrapp {P_1 }\beta {\gamma'}} {\homrallXYtoZ {\gamma} {\harrapp {E_1} {\gamma'} \gamma} {\harrapp {Q_1} \alpha \gamma}}}\]
+%% %%   where $\psi_{dom}$ and $\chi_{cod}$ are  as
+%% %%   \[ \psi_{dom} =\homrlambdaXatYdotZ p {\gamma'} {\homrlambdaXatYdotZ e \gamma
+%% %%     {\homrappXtoYatZ {\phi} {(\apptrinary {P_t} \beta {\pendappXtoY {D_r} \beta} \beta p {\gamma'} e \gamma)} \gamma}}
+%% %%   \]
+%% %%   \[
+%% %%   \psi_{cod} = \homrlambdaXatYdotZ p {\gamma'} {\homrlambdaXatYdotZ e \gamma
+%% %%     {\apptrinary {Q_t} \alpha {\pendappXtoY {C_r} \alpha} \alpha {(\homrappXtoYatZ \phi {p} {\gamma'})} {\gamma'} e \gamma}}
+%% %%   \]
+%% %% \end{definition}
+
+%% %% We can easily define a contravariant hom $P \triangleleft Q$ by a dual
+%% %% definition (TODO).
+
+%% %% We can also define a left Kan extensions using the contravariant
+%% %% cotensors.  Normally the left Kan extension is defined using tensors,
+%% %% but the definition using profunctors doesn't look right to me. Let's
+%% %% work backwards from the definition of colimit to see why.
+
+%% %% Let $\alpha : \cat C \vdash A : \cat D$ be a functor we want the
+%% %% colimit of. The UP of a colimit is that maps out of it are equivalent
+%% %% to cocones from $A$. Semi-formally this looks like
+%% %% \[ (\colim_{\alpha} A \to_{\cat D} \beta) \cong \forall \alpha:\cat C.~ A \to_{\cat D} \beta \]
+%% %% This looks a lot like a cotensor with a trivial domain:
+%% %% \[ (\colim_{\alpha} A \to_{\cat D} \beta) \cong \forall \alpha. (() \to_{1} \gamma) \triangleright (A \to_{\cat D} \beta) \]
+%% %% Suggesting that we define the left Kan extension profunctor as a
+%% %% contravariant cotensor as follows.
+%% %% \begin{definition}
+%% %%   Let $\cat C,\cat C',\cat D$ be cats and let $\alpha:\cat C \vdash A
+%% %%   : \cat D$ and $\alpha : \cat C \vdash J : \cat C'$ be functors.
+%% %%   Then the left Kan extension profunctor $\gamma^o:\cat C'; \beta:\cat
+%% %%   D \pipe (PLan_J A)[\gamma;\beta]$ is defined as
+%% %%   \[
+%% %%   \gamma^o:\cat C'; \beta:\cat D \pipe (PLan_J A)[\gamma;\beta]
+%% %%   := \forall \alpha : \cat C.~ (J \to_{\cat C'} \gamma) \triangleright (A \to_{\cat D} \beta)
+%% %%   \]
+%% %%   Then a left Kan extension of $A$ along $J$ is a functor $\gamma :
+%% %%   \cat C' \vdash Lan_J A : \cat D$ that represents $PLan_J A$ in that
+%% %%   \[ (Lan_J A \to_{\cat D} \beta) \cong (LPan_J A)[\gamma;\beta] \]
+%% %% \end{definition}
+
+%% %% \begin{definition}
+%%   %% Let $\cat C,\cat C',\cat D$ be cats and let $\alpha:\cat C \vdash
+%%   %% : \cat D$ and $\alpha : \cat C \vdash J : \cat C'$ be functors.
+%%   %% Then the left Kan extension profunctor $\beta^o:\cat D; \gamma:\cat
+%%   %% C' \pipe (PLan_J F)[\beta;\gamma]$ is defined as
+%% %%   \[
+%% %%   \beta^o:\cat D; \gamma:\cat C' \pipe (PLan_J A)[\beta;\gamma] :=
+%% %%   \exists \alpha:\cat C. (\beta \to_{\cat D} A[\alpha]) \odot (J[\alpha] \to_{\cat C'} \gamma)
+%% %%   \]
+%% %%   And a left Kan extension is a functor $\gamma :\cat C' \vdash Lan_J
+%% %%   A : \cat D$ that represents $PLan_J A$ in that
+%% %%   \[ (\beta \to_{\cat D} Lan_J A[\alpha]) \cong (PLan_J A)[\beta;\alpha] \]
+%% %% \end{definition}
+
+%% %% \begin{lemma}[Parameterized Yoneda Lemma]
+%% %%   Given categories $\cat C, \cat D$ and a profunctor $\alpha^o:\cat
+%% %%   C;\beta : \cat D \vdash \alpha \to_{Q} \beta$ and a function
+%% %%   $\beta : \cat D \vdash G[\beta] : \cat C$, then there is an
+%% %%   isomorphism between
+%% %%   \begin{enumerate}
+%% %%   \item Terms $\cdot \vdash t : G[\beta] \to_{Q} \beta$
+%% %%   \item left-homomorphisms $\phi : \alpha \to_{\cat C} G[\beta] \vdash
+%% %%     u : \alpha \to_{Q} \beta$ i.e. they satisfy homomorphism on the
+%% %%     left (the other side doesn't make sense if $G$ is not a functor):
+%% %%     $u[\phi \circ \psi] = u[\phi] \circ \psi$
+%% %%   \end{enumerate}
+%% %% \end{lemma}
+%% %% \begin{proof}
+%% %%   First, the constructions
+%% %%   \begin{enumerate}
+%% %%   \item Given $t$, we define $u(t)$ to be
+%% %%     \[\phi : \alpha \to_{\cat C} G[\beta] \vdash t \circ_{Q} \phi : \alpha \to_{Q} \beta\]
+
+%% %%     Which is a left-homomorphism because:
+%% %%     \[ (t \circ_{Q} \phi) \circ_{Q} \psi = t \circ_{Q} (\phi \circ_{\cat C} \psi)\]
+%% %%   \item Given $u$, we define $t(u)$ to be
+%% %%     \[ \cdot \vdash u[\id_{G[\beta]}/\phi] : G[\beta] \to_{Q} \beta \]
+%% %%   \end{enumerate}
+%% %%   Now we show they are inverse
+%% %%   \begin{enumerate}
+%% %%   \item First, the very easy case, $\cdot \vdash t(u(t)) : G[\beta] \to_{Q} \beta$:
+%% %%     \begin{align*}
+%% %%       t \circ_{Q} \id_{G[\beta]} &= t
+%% %%     \end{align*}
+%% %%     by the fact that $Q$ is a profunctor.
+%% %%   \item Next, $\phi : \alpha \to_{\cat C} G[\beta] \vdash u(t(u)) : \alpha \to_{Q} \beta$:
+%% %%     \begin{align*}
+%% %%       u[\id_{G[\beta]}/\phi] \circ_{Q} \phi &= u[\id \circ_{\cat C} \phi/\phi]\\
+%% %%       &= u[\phi/\phi]\\
+%% %%       &= u
+%% %%     \end{align*}
+%% %%     which uses the fact that $u$ is a left-homomorphism.
+%% %%   \end{enumerate}
+%% %% \end{proof}
+
+%% %% \begin{theorem}[Negative Types and Right Representables are Equivalent]
+%% %%   For any profunctor $\alpha^o: \cat C; \beta : \cat D \pipe \alpha
+%% %%   \to_{Q} \beta$, there is a bijection between
+%% %%   \begin{enumerate}
+%% %%   \item Right Represntables $\beta : \cat D \vdash G[\beta] : \cat C$ for $Q$.
+%% %%   \item Negative Types $\beta : \cat D \vdash N[\beta] : \cat C$
+%% %%     representing $Q$.
+%% %%   \end{enumerate}
+%% %% \end{theorem}
+%% %% \begin{proof}
+%% %%   \begin{enumerate}
+%% %%   \item Given a right representable $G[\beta]$, we set $N(G)[\beta] =
+%% %%     G[\beta]$, the introduction rule is the same and the elimination
+%% %%     rule is given as in the Yoneda lemma and the inverse property
+%% %%     follows by the Yoneda isomorphism and the fact that $I_G^{-1}$ is
+%% %%     an inverse for $I_G$.
+%% %%   \item Given a negative type $N[\beta]$ we set $G(N)[\beta] =
+%% %%     N[\beta]$, the introduction rule is the same and the inverse
+%% %%     introduction rule is given as in the Yoneda lemma.  We need to
+%% %%     show that $N[\beta]$ has the structure of a functor and that
+%% %%     $I_{N}$ and $I_{N}^{-1}$ are homomorphisms on the right with
+%% %%     respect to that structure.  Define $N_1$ as follows:
+%% %%     \begin{mathpar}
+%% %%       \inferrule
+%% %%           {\beta_1^o:\cat D;\beta_2:\cat D \vdash \psi \circ \epsilon : N[\beta_1/\beta] \to_{Q} \beta_2\\
+%% %%             \phi : N[\beta_1/\beta] \to_{Q} \beta_2 \vdash I_N[\phi] : N[\beta_1/\beta] \to_{\cat C} N[\beta_2/\beta]
+%% %%           }
+%% %%           {\beta_1^o:\cat D;\beta_2:\cat D\pipe \psi : \beta_1 \to_{\cat D} \beta_2 \vdash I_N[\psi \circ \epsilon_{\beta_1}/\phi] : N[\beta_1/\beta] \to_{\cat C} N[\beta_2/\beta]}
+%% %%     \end{mathpar}
+%% %%     We need to show functoriality.
+%% %%     For identity we have
+%% %%     \begin{align*}
+%% %%       I_N[\id_{\beta} \circ \epsilon_{\beta}/\phi] &= I_N[\epsilon_{\beta}]\\
+%% %%       &= I_N[\epsilon \circ \id_{N[\beta]}]\\
+%% %%       &= \id_{N[\beta]} \tag{$\eta$}
+%% %%     \end{align*}
+%% %%     For composition,
+%% %%     \begin{align*}
+%% %%       N_1[\theta] \circ N_1[\psi]
+%% %%       &= I_N[\theta \circ \epsilon] \circ I_N[\psi \circ \epsilon]\\
+%% %%       &= I_N[(\theta \circ \epsilon) \circ I_N[\psi \circ \epsilon]]\tag{left homomorphism}\\
+%% %%       &= I_N[\theta \circ (\epsilon \circ I_N[\psi \circ \epsilon])]\\
+%% %%       &= I_N[\theta \circ (\psi \circ \epsilon)] \tag{$\beta$}\\
+%% %%       &= I_N[(\theta \circ \psi) \circ \epsilon]\\
+%% %%       &= N_1[\theta \circ \psi]
+%% %%     \end{align*}
+%% %%     Next we show $I_N$ is a right homomorphism, the typing here is
+%% %%     $\phi : \alpha \to_{Q} \beta_1, \theta : \beta_1 \to_{\cat D}
+%% %%     \beta_2$.
+%% %%     \begin{align*}
+%% %%       N_1[\theta] \circ I_N[\phi] &= I_N[\theta \circ \epsilon] \circ I_N[\phi]\\
+%% %%       &= I_N[\theta \circ \epsilon \circ I_N[\phi]]\tag{left homomorphism}\\
+%% %%       &= I_N[\theta \circ \phi] \tag{$\beta$}
+%% %%     \end{align*}
+%% %%     and to show $I_N^{-1}$ is a right homomorphism under $\phi : \alpha \to_{\cat C}N[\beta_1], \theta : \beta_1 \to_{\cat D} \beta_2$
+%% %%     \begin{align*}
+%% %%       I_N^{-1}[N_1[\theta] \circ \phi] &= \epsilon \circ (N_1[\theta] \circ \phi) \tag{def}\\
+%% %%       &= \epsilon \circ (I_N[\theta \circ \epsilon] \circ \phi)\tag{def}\\
+%% %%       &= (\epsilon \circ I_N[\theta \circ \epsilon]) \circ \phi\tag{assoc}\\
+%% %%       &= (\theta \circ \epsilon) \circ \phi\tag{$\beta$}\\
+%% %%       &= \theta \circ (\epsilon \circ \phi)\tag{assoc}\\
+%% %%       &= \theta \circ I_N^{-1}[\phi] \tag{def}
+%% %%     \end{align*}
+%% %%   \end{enumerate}
+
+%% %%   Next we need to show that this is a \emph{bijection}.
+%% %%   \begin{enumerate}
+%% %%   \item Starting with a negative type $N[\beta]$, we recover the
+%% %%     original $\epsilon$ by the Yoneda lemma.
+%% %%   \item Starting with a right representable $G[\beta]$, we recover the
+%% %%     original $I_N^{-1}$ by the Yoneda lemma. We need to show that the
+%% %%     action of $G$ on arrows is the same as the one defined from $G$
+%% %%     viewed as a negative type, which we will call $G_1'$:
+%% %%     \begin{align*}
+%% %%       G_1'[\theta] &= I_G[\theta \circ \epsilon]\\
+%% %%       &= G_1[\theta] \circ I_G[\epsilon]\tag{right homomorphism}\\
+%% %%       &= G_1[\theta] \circ I_G[\epsilon \circ \id]\\
+%% %%       &= G_1[\theta] \circ \id\\
+%% %%       &= G_1[\theta]
+%% %%     \end{align*}
+%% %%   \end{enumerate}
+%% %% \end{proof}
+
+%% %% \subsection{Adjunctions by Universal Morphism}
+
+%% %% As an example of formal category theory in our syntax, let's prove the
+%% %% equivalence between two definitions of an adjunction: one by universal
+%% %% morphism and the other by natural hom isomorphism.
+%% %% %
+%% %% Though the definitions and proof will look like we are doing ordinary
+%% %% category theory, since it is syntactic it immediately applies to all
+%% %% models, including enriched and internal category theory.
+
+%% %% We start with the usual definition of an adjunction using
+%% %% hom-isomorphism and a profunctor.
+%% %% \begin{definition}
+%% %%   Let $C,D$ be categories and $c^o:C_0,d:D_0 \vdash H[c,d]$ a
+%% %%   profunctor between them. A \emph{left adjoint} to $H$ consists of
+%% %%   \begin{enumerate}
+%% %%   \item A functor $c:C \vdash F[c] : D$
+%% %%   \item A natural isomorphism $c,d \pipe h : H[c,d] \vdash
+%% %%     F_{ind}[h]: D_1[F_0[c],d]$. Call the inverse $g. F_{ind}^{-1}[g]$.
+%% %%   \end{enumerate}
+%% %% \end{definition}
+
+%% %% \begin{definition}
+%% %%   Let $c:C_0,d:D_0 \vdash H[c,d]$ be a profunctor.
+%% %%   A \emph{universal $H$-heteromorphism from $C$} consists of
+%% %%   \begin{enumerate}
+%% %%   \item For every $c:C_0$, an $H$-morphism $\eta_c : H(c,F_0[c])$ for
+%% %%     some function $c:C_0 \vdash F_0[c] : D_0$.
+%% %%   \item Such that every $h : H[c,d]$ factors through $\eta_c$ in that
+%% %%     there is a unique $h : H[c,d] \vdash F_{ind}[h] : D[F_0[c],d]$
+%% %%     such that $h = F_{ind}[h] \circ \eta_c$.  Uniqueness means
+%% %%     validity of the following rule:
+%% %%     \begin{mathpar}
+%% %%       \inferrule
+%% %%       {g:D[F_0[c],d] \vdash g \circ \eta = h}
+%% %%       {g:D[F_0[c],d] \vdash g = F_{ind}[h]}
+%% %%     \end{mathpar}
+%% %%     (Note that we do not require $F_{ind}$ to be natural in any sense)
+%% %%   \end{enumerate}
+%% %% \end{definition}
+
+%% %% It is easy to see that a left adjoint gives a universal heteromorphism
+%% %% by a Yoneda-like argument:
+%% %% \begin{construction}
+%% %%   Let $F$ be a left adjoint to $H$. Then we define a left universal
+%% %%   $H$-heteromorphism by
+%% %%   \begin{enumerate}
+%% %%   \item $c \vdash \eta_c = F_{ind}^{-1}[\id[F_0[c]]]: H[c,F_0[c]]$
+%% %%   \item Every $h : H[c,d]$ factorizes through $\eta$ because
+%% %%     \begin{align*}
+%% %%       F_{ind}[h] \circ F_{ind}^{-1}[\id] &= F_{ind}^{-1}[F_{ind}[h] \circ \id] \tag{naturality}\\
+%% %%       &= F_{ind}^{-1}[F_{ind}[h]] \tag{unit}\\
+%% %%       &= h \tag{isomorphism}
+%% %%     \end{align*}
+%% %%     This factorization is unique because if $g \circ F^{-1}[\id] = h$,
+%% %%     then
+%% %%     \begin{align*}
+%% %%       F_{ind}[h] &= F_{ind}[g \circ F^{-1}[\id]] \tag{congruence}\\
+%% %%       &= g \circ F_{ind}[F^{-1}[\id]] \tag{naturality}\\
+%% %%       &= g \circ \id \tag{iso}\\
+%% %%       &= g \tag{unit}
+%% %%     \end{align*}
+%% %%   \end{enumerate}
+%% %% \end{construction}
+
+%% %% But the utility of the universal morphism definition is that the
+%% %% universal morphism extracted from a left adjoint is enough to uniquely
+%% %% reconstruct the rest of the structure, which includes (1)
+%% %% functoriality of $F$ and (2) naturality of $F_{ind}$.
+%% %% \begin{construction}
+%% %%   Let $(F_0,\eta,F_{ind}[h])$ be a left universal $H$-morphism.
+%% %%   We extend this to a left adjoint as follows.
+%% %%   \begin{enumerate}
+%% %%   \item First, $F$ acts on morphisms by
+%% %%     \begin{mathpar}
+%% %%       \inferrule
+%% %%       {
+%% %%         \inferrule
+%% %%          {f : C[c,c'] \and \eta_{c'} : H[c',F_0[c']]}
+%% %%          {f : C[c,c'] \vdash \eta_{c'} \circ f : H[c,F_0[c']]}
+%% %%       }
+%% %%       {f : C[c,c'] \vdash F_{ind}[\eta_{c'} \circ f] : D[F_0[c],F_0[c']]}
+%% %%     \end{mathpar}
+%% %%     This preserves identity
+%% %%     \begin{mathpar}
+%% %%       \inferrule
+%% %%       {
+%% %%         \inferrule
+%% %%         {\id \circ \eta = \eta}
+%% %%         {\id = F_{ind}[\eta]}
+%% %%       }
+%% %%       {\id = F_{ind}[\eta \circ \id]}
+%% %%     \end{mathpar}
+%% %%     And composition
+%% %%     \begin{mathpar}
+%% %%       \inferrule
+%% %%       {F_{ind}[\eta \circ f] \circ F_{ind}[\eta \circ g] \circ \eta = \eta \circ f \circ g}
+%% %%       {F_{ind}[\eta \circ f] \circ F_{ind}[\eta \circ g] = F_{ind}[\eta \circ f \circ g]}
+%% %%     \end{mathpar}
+%% %%     Which follows by a couple uses of the factorization:
+%% %%     \begin{align*}
+%% %%       F_{ind}[\eta \circ f] \circ F_{ind}[\eta \circ g] \circ \eta
+%% %%       &= F_{ind}[\eta \circ f] \circ \eta \circ g \\
+%% %%       &= \eta \circ f \circ g
+%% %%     \end{align*}
+%% %%   \item Next, we show that $h : H[c,d] \vdash F_{ind}[h] :
+%% %%     D[F_0[c],d]$ is a natural transformation. As a first step, we show
+%% %%     that $\eta_c : H[c,F_0[c]]$ is natural in $c$ with the action of
+%% %%     $F$ on morphisms we have just defined.
+%% %%     \[ \eta \circ f = F[f] \circ \eta \]
+%% %%     expanding the definition, this is precisely the factorization
+%% %%     property:
+%% %%     \[ \eta \circ f = F_{ind}[\eta \circ f] \circ \eta \]
+
+%% %%     Next, to show that $F_{ind}$ is natural, we need to show that
+%% %%     (when it is well-defined):
+%% %%     \[F_{ind}[g \circ h \circ f] = g \circ F_{ind}[h] \circ F[f] \]
+%% %%     By the factorization property, the following sufficeth:
+%% %%     \begin{align*}
+%% %%      g \circ F_{ind}[h] \circ F[f] \circ \eta
+%% %%      &= g \circ F_{ind}[h] \circ \eta \circ f \tag{naturality of $\eta$}\\
+%% %%      &= g \circ h \circ f \tag{factorization}
+%% %%     \end{align*}
+%% %%   \item Finally, we define $F_{ind}^{-1}[g]$ as
+%% %%     \[ g : D[F_0[c],d] \vdash F_{ind}^{-1}[g] = g \circ \eta : H[c,d] \]
+%% %%     And we need to confirm that it is (1) a section of $F_{ind}$ (2) a
+%% %%     retraction of $F_{ind}$ and (3) natural.
+%% %%     \begin{enumerate}
+%% %%     \item For retraction we need to show $h \vdash F_{ind}[h] \circ
+%% %%       \eta = h$ which is precisely the factorization property.
+%% %%     \item For section, we need to show $g \vdash F_{ind}[g \circ \eta]
+%% %%       = g$. Immediately the uniquenes principle says its sufficient to
+%% %%       show $g \circ \eta = g \circ \eta$.
+%% %%     \item For naturality, we need to show
+%% %%       \[ F_{ind}^{-1}[k \circ g \circ F[f]] = k \circ F_{ind}^{-1}[g] \circ f \]
+%% %%       expanding the definition this is just:
+%% %%       \[ k \circ g \circ F[f] \circ \eta = k \circ g \circ \eta \circ f \]
+%% %%       which is just naturality of $\eta$.
+%% %%     \end{enumerate}
+%% %%   \end{enumerate}
+%% %% \end{construction}
+
+%% \subsection{Products and Units of Internal Categories}
+
+%% %% While we have now proved a beautiful internal theorem in great
+%% %% generality that has many useful instances in different models, we are
+%% %% hard pressed to actually \emph{instantiate} the theorem in the syntax
+%% %% itself because we can't really come up with any useful universal properties!
+%% %% The reason is that Judgment Theory itself is too bare-bones.
+
+%% %% For instance how would we say that a category $\cat C$ has a
+%% %% \emph{terminal object}? Well, it would have to be in the first place a
+%% %% \emph{single} object in $\cat C$, but so far our syntax only enables
+%% %% us to talk about functions and functors. Well an object of $\cat C$ is
+%% %% the same as a function from the unit sort, but so far our syntax only
+%% %% has base sorts.
+
+%% %% So we see that defining the terminal object, \emph{necessitates} some
+%% %% notion of terminal object in our ``metatheory'': Judgment Theory.
+%% %% This mysterious and pervasive aspect of higher category theory is
+%% %% called the \emph{microcosm principle} because it says that in order to
+%% %% discuss a property of a single object in a single category (the
+%% %% microcosm), we need the same sort of object at the meta-level (the
+%% %% macrocosm). The adage here is ``As above, so below''.
+
+%% %% That might all sound very spiritual, but we'll see that it's actually
+%% %% quite a nice heuristic for formalizing universal properties in
+%% %% internal category theory.
+%% %% %
+%% %% In the spirit of this heuristic, instead of first giving a bunch of
+%% %% new type constructions for Judgment Theory and then applying them to
+%% %% define universal properties, we will consider them in pairs, first
+%% %% above, and then below.
+
+%% First we consider the terminal sort and then the idea of a terminal
+%% object in an internal category.
+
+%% Next, the terminal sort trivially has a category structure using the
+%% trivial judgment as its hom set:
+
+%% \begin{definition}[Terminal Category]
+%%   The terminal category $\cat 1$ is defined as having
+%%   \begin{enumerate}
+%%   \item Its object sort as $1$
+%%   \item Its hom set as $\lambda \alpha:1;\beta:1. 1 : \harr 1 1$
+%%   \item Its identity arrow is given by $\lambda \alpha: 1. () : 1$
+%%   \item Composition is defined by $\phi : 1, \psi : 1 \vdash () : 1$
+%%   \item Associativity and unitality are trivial by \textsc{1-$\eta$}
+%%   \end{enumerate}
+%% \end{definition}
+
+%% \begin{definition}[Terminal Object Specification]
+%%   In judgment type theory with a terminal sort/judgment, we can define
+%%   for any category $\cat C$ a profunctor $\alpha^o:\cat C;\beta : \cat
+%%   1 \pipe 1$, which specifies a terminal object.
+%% \end{definition}
+
+%% Now let's instantiate the definition for the terminal object defined
+%% as a negative type and as a representable and see how it reproduces
+%% the standard type-theoretic and category-theoretic definitions.
+
+%% A representation of $\alpha^o:\cat C; \beta : \cat 1 \pipe 1$ consists
+%% of a functor $\beta : \cat 1 \vdash 1_{\cat C} : \cat C$, an introduction rule
+%% \[ \phi : 1 \vdash I_{1}[\phi] : \alpha \to_{\cat C} 1_{\cat C}[\beta] \]
+%% satisfying $I_{1}[\phi] \circ \gamma =I_{1}[\phi \circ \gamma]$ and an
+%% inverse $\psi : \alpha \to_{\cat C} 1_{\cat C}[\beta] \vdash
+%% I_{1}^{-1}[\psi] : 1$ which by \textsc{1-$\eta$} is equal to $()$ and
+%% therefore every $t : \alpha \to_{\cat C} 1_{\cat C}[\beta]$ is equal
+%% to $I_{1}[\phi]$. In light of this it would be appropriate to write
+%% $I_1[\phi]$ as $()$ and then the left-homomorphism property is $()
+%% \circ \gamma = ()$, which is precisely the definition of substitution.
+
+%% This alternate syntax makes even more sense in light of the negative
+%% type presentation. A negative type satisfying the terminal object
+%% specification consists of a functor $1_{\cat C}$ as above and an
+%% introduction rule as above and an elimination form $\vdash \epsilon :
+%% 1$, which is trivial.
+
+%% First the product category
+%% \begin{definition}[Product category]
+%%   For any categories $\cat C_1$, $\cat C_2$ we can form the product
+%%   category whose sort of objects is $\cat C_1 \times \cat C_2$, whose
+%%   arrow judgment is $\alpha^o : \cat C_1 \times \cat C_2;\alpha \pipe
+%%   (\pi_1 \alpha \to_{\cat C_1} \pi_1 \alpha) \times (\pi_2 \alpha
+%%   \to_{\cat C_2} \pi_2 \alpha)$. Identity is given by $\alpha^o : \cat
+%%   C_1 \times \cat C_2;\alpha\pipe\cdot \vdash (\id_{\pi_1 \alpha},
+%%   \id_{\pi_2 \alpha}) : (\pi_1 \alpha \to_{\cat C_1} \pi_1 \alpha)
+%%   \times (\pi_2 \alpha \to_{\cat C_2} \pi_2 \alpha)$ and composition
+%%   by
+%%   \[
+%%   \phi %: (\pi_1 \alpha_0 \to_{\cat C_1} \pi_1 \alpha_1) \times (\pi_2 \alpha_0 \to_{\cat C_2} \pi_2 \alpha_1)
+%%   ,
+%%   \psi %: (\pi_1 \alpha_1 \to_{\cat C_1} \pi_1 \alpha_2) \times (\pi_2 \alpha_1 \to_{\cat C_2} \pi_2 \alpha_2)
+%%   \vdash ((\pi_1 \phi) \circ (\pi_1 \psi), (\pi_2 \psi) \circ (\pi_2 \phi)) :
+%%   (\pi_1 \alpha_0 \to_{\cat C_1} \pi_1 \alpha_2) \times (\pi_2 \alpha_0 \to_{\cat C_2} \pi_2 \alpha_2)  
+%%   \]
+%%   unitality, associativity follow from the same properties of $\cat C_1,\cat C_2$
+%% \end{definition}
+
+%% \begin{definition}[Product Functor Specification]
+%%   For any category $\cat C$, we can specify the product functor by
+%%   \begin{enumerate}
+%%   \item The span is $\alpha^o:\cat C; \beta : \cat C \times \cat C \pipe (\alpha \to_{\cat C} \pi_1 \beta) \times (\alpha \to_{\cat C} \pi_2 \beta)$
+%%   \item Left Composition is defined as
+%%     \[ \phi, \psi \vdash (\pi_1 \phi \circ \psi, \pi_2 \phi \circ \psi)\]
+%%   \item Left unitality:
+%%     \begin{align*}
+%%       (\pi_1 \phi \circ \id, \pi_2 \phi \circ \id)
+%%       &= (\pi_1 \phi, \pi_2 \phi)\\
+%%       &= \phi\tag{$\times-eta$}
+%%     \end{align*}
+%%   \item Left associativity:
+%%     \begin{align*}
+%%       (\phi \circ \psi) \circ \psi'
+%%       &= (\pi_1 \phi \circ \psi, \pi_2 \phi \circ \psi) \circ \psi'\\
+%%       &= (\pi_1 (\pi_1 \phi \circ \psi, \pi_2 \phi \circ \psi) \circ \psi',
+%%       \pi_2 (\pi_1 \phi \circ \psi, \pi_2 \phi \circ \psi) \circ \psi')\\
+%%       &= ((\pi_1 \phi \circ \psi) \circ \psi', (\pi_2 \phi \circ \psi) \circ \psi')\\
+%%       &= (\pi_1 \phi \circ (\psi \circ \psi'), \pi_2 \phi \circ (\psi \circ \psi'))\\
+%%       &= \phi \circ (\psi \circ \psi')
+%%     \end{align*}
+%%   \item Right composition is defined as
+%%     \[ \phi, \theta \vdash (\pi_1 \theta \circ \pi_1 \phi, \pi_2 \theta \circ \pi_2 \phi) \]
+%%   \item Right unitality:
+%%     \begin{align*}
+%%       (\pi_1 (\id,\id) \circ \pi_1 \phi, \pi_2 (\id,\id) \circ \pi_2 \phi)
+%%       &= (\id \circ \pi_1 \phi, \id \circ \pi_2 \phi) \tag{$\times-\beta$}\\
+%%       &= (\pi_1 \phi, \id \circ \pi_2 \phi)\\
+%%       &= \phi \tag{$\times-\eta$}
+%%     \end{align*}
+%%   \item Right associativity
+%%     \begin{align*}
+%%       \theta' \circ (\theta \circ \phi)
+%%       &= \theta' \circ (\pi_1 \theta \circ \pi_1 \phi, \pi_2 \theta \circ \pi_2 \phi)\\
+%%       &= (\pi_1 \theta' \circ \pi_1(\pi_1 \theta \circ \pi_1 \phi, \pi_2 \theta \circ \pi_2 \phi),
+%%       \pi_2 \theta' \circ \pi_2(\pi_1 \theta \circ \pi_1 \phi, \pi_2 \theta \circ \pi_2 \phi))\\
+%%       &= (\pi_1 \theta' \circ (\pi_1 \theta \circ \pi_1 \phi), \pi_2 \theta' \circ (\pi_2 \theta \circ \pi_2 \phi))\\
+%%       &= ((\pi_1 \theta' \circ \pi_1 \theta) \circ \pi_1 \phi, (\pi_2 \theta' \circ \pi_2 \theta) \circ \pi_2 \phi)\\
+%%       &= (\pi_1(\theta' \circ \theta) \circ \pi_1 \phi, \pi_2(\theta' \circ \theta) \circ \pi_2 \phi)\\
+%%       &= (\theta' \circ \theta) \circ \phi \\
+%%     \end{align*}
+%%   \end{enumerate}
+%% \end{definition}
+
+%% \subsection{Discrete and Codiscrete Categories}
+
+%% If we have terminal judgments, then we can define the
+%% \emph{codiscrete} category for any sort $\cat C$.
+%% \begin{definition}
+%%   The codiscrete category $\textrm{CoDisc} \cat C$ is defined as
+%%   \begin{enumerate}
+%%   \item Object sort is $\cat C$
+%%   \item Morphisms judgment is $1$
+%%   \item Identity and composition are given by the unique terms of
+%%     those types, making associativity and unitality trivial to verify.
+%%   \end{enumerate}
+%% \end{definition}
+
+%% If we have \emph{identity} objects, which we write as $\equiv_A$, then we
+%% can define discrete categories.
+%% \begin{definition}
+%%   The discrete category $\textrm{Disc} \cat C$ is defined as
+%%   \begin{enumerate}
+%%   \item Object sort is $\cat C$
+%%   \item Morphisms judgment is $\alpha;\alpha' \pipe \alpha \equiv \alpha'$
+%%   \item Identity and composition are defined as in the section on
+%%     synthetic category theory.
+%%   \end{enumerate}
+%% \end{definition}
+
+%% \subsection{Internal Tensor, Cotensor}
+
+%% We should be able to define internal \emph{tensor} using synthetic
+%% tensor and coequalizers and also internal \emph{cotensor} using
+%% synthetic cotensor and equalizers. These look exactly like the
+%% set-theoretic definitions.
+%% \begin{definition}
+%%   Let $R$ be an internal profunctor from $C$ to $D$ and $S$ be an
+%%   internal profunctor from $D$ to $E$.
+
+%%   Then we can define the internal tensor product $R \odot S$, a
+%%   profunctor from $C$ to $E$ as
+%%   \begin{enumerate}
+%%   \item The underlying span can be defined as a quotient of the synthetic tensor product
+%%     \[ (R \odot S)[\alpha;\gamma] = \exists \beta:\cat D. R \odot S/Rcomp = Scomp \]
+
+%%     Where R-comp and S-comp have as source the span
+%%     \[ \exists \beta:\cat D.~ R[\alpha;\beta] \odot (\exists \beta':\cat D. Hom_C(\beta,\beta') \odot S[\beta';\gamma]) \]
+%%     or the isomorphic
+%%     \[ \exists \beta':\cat D.~ (\exists \beta:\cat D. R[\alpha;\beta] \odot Hom_C(\beta,\beta')) \odot S[\beta';\gamma] \]
+%%     where R-comp composes the $C$-morphism with the $R$-morphism and
+%%     S-comp the $S$-morphism.
+%%   \item Composition is defined using composition in $R,S$.
+%%   \end{enumerate}
+%% \end{definition}
+%% Note that $R$-comp and $S$-comp are each constructed using the
+
+%% \begin{definition}
+%%   Let $R$ be an internal profunctor from $C$ to $D$ and $S$ be an
+%%   internal profunctor from $C$ to $E$. Then the internal contravariant
+%%   cotensor product $R \triangleright S$ is defined as
+%%   \[ (R \triangleright S)[\beta;\gamma] = \{ p : \forall \alpha:\cat C.~ R[\alpha;\beta] \triangleright S[\alpha;\gamma] 
+%%   \pipe R-comp = S-comp \}  \]
+%%   R-comp and S-comp have as codomain the span
+%%   \[ \forall \alpha:\cat C.~ R[\alpha;\beta] \triangleright \forall \alpha':\cat C.~ Hom_C(\alpha',\alpha) \triangleright S[\alpha';\gamma] \]
+%%   or the isomorphic
+%%   \[ \forall \alpha':\cat C.~ (\exists \alpha. Hom_C(\alpha',\alpha) \odot R[\alpha;\beta]) \triangleright S[\alpha';\gamma]
+%%   \]
+%% \end{definition}
+
+%% We can also construct the \emph{free profunctor} from a span using tensor:
+
+%% \begin{definition}
+%%   Let $\cat C$ and $\cat D$ be internal categories and let $R[\alpha;\beta]$ be a
+%%   span from $Ob(\cat C)$ to $Ob(\cat D)$. Then the free profunctor
+%%   extending $R$ to be an internal profunctor from $\cat C$ to $\cat D$ is
+%%   \[ F_{\cat C,\cat D}R[\alpha';\beta'] = \exists \alpha. Hom_C(\alpha',\alpha) \odot (\exists \beta. R[\alpha;\beta] \odot Hom_C(\beta,\beta')) \]
+%%   or the isomorphic
+%%   \[ F_{\cat C,\cat D}R[\alpha';\beta'] = \exists \beta. (\exists \alpha. Hom_C(\alpha',\alpha) \odot R[\alpha;\beta]) \odot Hom_C(\beta,\beta') \]
+%% \end{definition}
+
+%% And maybe we can even define the co-free profunctor from a span using cotensor?
+%% \begin{definition}
+%%   Let $\cat C$ and $\cat D$ be internal categories and let $R[\alpha;\beta]$ be a
+%%   span from $Ob(\cat C)$ to $Ob(\cat D)$. Then the co-free profunctor
+%%   extending $R$ to be an internal profunctor from $\cat C$ to $\cat D$ is
+%%   \[ G_{\cat C,\cat D}R[\alpha';\beta'] =
+%%   \forall \alpha. Hom_C(\alpha,\alpha') \triangleright (\forall \beta. R[\alpha;\beta] \triangleleft Hom_D(\beta',\beta)) \]
+%%   or the isomorphic
+%%   \[ G_{\cat C,\cat D}R[\alpha';\beta'] =
+%%   \forall \beta. 
+%%   (\forall \alpha. Hom_C(\alpha,\alpha') \triangleright R[\alpha;\beta]) \triangleleft Hom_D(\beta',\beta)) \]
+%% \end{definition}
 

### 2022-04-15 paper/paper.tex
semantics work

@@
 
-In this section we establish our desired semantic soundness and
-completeness theorems for models in virtual equipments with the
-corresponding universal properties.
-%
-The first component of this is \emph{soundness}: every category,
-object, set and transformation denote an object, vertical arrow,
-horizontal arrow and 2-cell, respectively in any representable virtual
-equipment, such that all equations in the theory are satisfied.
-%
-We combine this with a set theoretic semantics for types $A$ and
-contexts $\Gamma$ to get a simple semantics of our logic over a
-virtual equipment.
-%
-The second component of this is the \emph{completeness} theorem:
-informally, the theory contains every category, object, set and
-transformation that \emph{must} exist in any virtual equipment, and
-further equations between syntactic forms hold if and only if they
-hold in every model.
-%
-While sounding more ambitious, the completeness theorem is
-typically simpler as it just amounts to showing that the syntax itself
-forms a virtual equipment.
+Next, we develop the basic model theory for SCT. Because SCT includes
+not just a syntax for category theory, but is also indexed by an
+ordinary context of dependent type theory $\Gamma$, a model of SCT is
+not given by a \emph{single} virtual double category, but instead by
+an \emph{indexed family} of VDCs, parameterized by an object of a
+category that interprets dependent type theory. We call such a
+structure as a \emph{VDC-hyperdoctrine}. We then show how to construct
+a VDC-hyperdoctrine indexed by the category of sets, which has the
+intuitive reading that any of the VDC judgments $\Gamma \vdash J$ is
+interpreted as simply a family of $J$-data parameterized by elements
+of the set $\Gamma$.
+
+%% The first component of this is \emph{soundness}: every category,
+%% object, set and transformation denote an object, vertical arrow,
+%% horizontal arrow and 2-cell, respectively in any representable virtual
+%% equipment, such that all equations in the theory are satisfied.
+%% %
+%% We combine this with a set theoretic semantics for types $A$ and
+%% contexts $\Gamma$ to get a simple semantics of our logic over a
+%% virtual equipment.
+%% %
+%% The second component of this is the \emph{completeness} theorem:
+%% informally, the theory contains every category, object, set and
+%% transformation that \emph{must} exist in any virtual equipment, and
+%% further equations between syntactic forms hold if and only if they
+%% hold in every model.
+%% %
+%% While sounding more ambitious, the completeness theorem is
+%% typically simpler as it just amounts to showing that the syntax itself
+%% forms a virtual equipment.
 
@@
 
-%% \begin{definition}[Virtual Double Category]
-%%   A virtual double category $V$ consists of
-%%   \begin{enumerate}
-%%     \item A set $V_0$ of objects.
-%%     \item For each $\cat C, \cat D \in V_0$, a set $V_v(\cat C, \cat
-%%       D)$ of vertical arrows, with composition and unit operations
-%%       satisfying the category axioms.
-%%     \item A set $V_h$ of horizontal arrows, with $d^+,d^- : V_h \to
-%%       V_0$.
-%%     \item For each composable string $\Phi$ of horizontal arrows
-%%       starting at $\cat C$ and ending at $\cat D$, and horizontal
-%%       arrow $R$, a set $V_2(\Phi; R)$ of 2-cells, with projection
-%%       functions $d^{\pm} : V_2(\Phi;R) \to V_v(d^\pm \Phi, d^\pm R)$.
-
-%%     \item For each $R \in V_h$, an identity arrow $\id_R \in
-%%       V_2(d^-R,R,d^+ R; R)$ with $d^\pm \id_R = \id_{d^\pm R}$.
-
-%%     \item For any $2$-cell $\alpha \in V_2(\Phi;R)$ and composable
-%%       string $\phi$ of 2-cells into $\Phi$, a composite $\alpha \circ
-%%       \phi$.
-
-%%     \item Composition should be unital in that:
-%%       \[ \id_R \circ [d^-\alpha, \alpha, d^+\alpha] = \alpha  \]
-%%       and
-%%       \[ \alpha \circ [\vec \id] = \alpha\]
-%%     \item And associative in that
-%%       \[ (\alpha \circ [\beta_1,\ldots]) \circ [\gamma_{11},\ldots,\gamma_{21},\ldots,] = \alpha \circ [(\beta_1 \circ \gamma_{11},\ldots), (\beta_2 \circ \gamma_{21},\ldots)]\]
-%%   \end{enumerate}
-%% \end{definition}
-
-%% \begin{definition}[Restriction]
-%%   Let $R \in V_h$ and $f \in V_v(\cat C, d^-R)$ and $g \in V_v(\cat D,
-%%   d^+R)$, then a restriction of $R$ along $f,g$ is a horizontal arrow
-%%   $R(f,g) \in V_h$ with $d^-R(f,g)= \cat C$ and $d^+R(f,g)= \cat D$
-%%   with a (op?)-cartesian 2-cell $\alpha \in V_2(R(f,g),R)$ with
-%%   $d^-\alpha = f$ and $d^+\alpha = g$.
-%% \end{definition}
+First, we recount the definition of a virtual double category.
+\begin{definition}[Virtual Double Category]
+  A (vertically strict) virtual equipment $\mathcal V$ consists of
+  \begin{enumerate}
+    \item A set $\mathcal V_0$ of objects.
+    \item For each $\cat C, \cat D \in V_0$, a set $V_v(\cat C, \cat
+      D)$ of vertical arrows, with composition and unit operations
+      satisfying the category axioms.
+    \item A set $V_h$ of horizontal arrows, with $d^+,d^- : V_h \to
+      V_0$.
+
+    \item For any vertical arrows $f \in V_v(\cat C, \cat D)$ and $g
+      \in V_v(\cat E, \cat F)$ and horizontal arrow $R \in V_h$ with
+      $d^+R = \cat F$ and $d^-R = \cat D$, a \emph{restriction}
+      $R(f,g)$ with $d^+R = \cat C$ and $d^-R = \cat E$.
+
+      This operation is furthermore associative and unital $R(f \circ
+      f',g \circ g') = R(f,g)(f',g')$ and $R(\id,\id) = R$.
+
+    \item For each composable string $\Phi$ (defined below) of horizontal arrows
+      starting at $\cat C$ and ending at $\cat D$, and horizontal
+      arrow $R$ with $d^-R = \cat C$ and $d^+R = \cat D$, a set
+      $V_2(\Phi; R)$ of 2-cells.
+
+    \item For any string of vertical arrows $\vec f$ into $\Phi$ and
+      2-cell $r \in \mathcal V_2(\Phi,S)$, a restriction operation
+      $r(\vec f) \in \mathcal V_2(\Phi(\vec f), S(d^-(\vec f),
+      d^+(\vec f)))$. Furthermore, this is associative and unital
+      $r(\vec id) = r$ and $r(\vec f \circ \vec g) = r(\vec {f \circ
+        g})$
+
+    \item For each $R \in V_h$, an identity arrow $\id_R \in
+      V_2(d^-R,R,d^+ R; R)$ with $d^\pm \id_R = \id_{d^\pm R}$.
+
+    \item For any $2$-cell $\alpha \in V_2(\Phi;R)$ and composable
+      string $\phi$ of 2-cells into $\Phi$, a composite $\alpha \circ
+      \phi$.
+
+    \item Composition should be unital in that:
+      \[ \id_R \circ [d^-\alpha, \alpha, d^+\alpha] = \alpha  \]
+      and
+      \[ \alpha \circ [\vec \id] = \alpha\]
+    \item And associative in that
+      \[ (\alpha \circ [\beta_1,\ldots]) \circ [\gamma_{11},\ldots,\gamma_{21},\ldots,] = \alpha \circ [(\beta_1 \circ \gamma_{11},\ldots), (\beta_2 \circ \gamma_{21},\ldots)]\]
+  \end{enumerate}
+
+  A composable string $\Phi$ of horizontal arrows is defined to be an
+  alternating sequence of objects and horizontal arrows, beginning and
+  ending with an object such that any substring $\cat C,R,\cat D$
+  satisfies $d^-R = \cat C$ and $d^+R = \cat D$.
+
+  A composable string of vertical arrows $\vec f$ into $\Phi$ is a
+  sequence of vertical arrows, one for each object in $\Phi$, with the
+  codomain of the arrow matching the object in $\Phi$. Then
+
+  A composable string of 2-cells from $\Phi$ to $\Psi$ is a sequence
+  $\vec r$ of 2-cells, one for each horizontal arrow in $\Psi$ such
+  that the codomains match the objects in
+\end{definition}
+
+Our main departure from\cite{crutwellshulman} is making restrictions
+part of the definition, and also requiring them to be strictly
+associative and unital. In the examples that we are most concerned
+with, however, these operations are strict and so we will not explore
+the possibility of coherence theorems for proving this equivalent to a
+weaker definition.
+
+\begin{definition}
+  A functor of virtual double categories $F : \mathcal V \to \mathcal
+  U$ is a mapping on objects, vertical arrows, horizontal arrows and
+  2-cells respecting typing and preserving composition, units, and
+  restriction.
+\end{definition}
 
@@
 
-\subsection{Category of Models}
-
-Our goal is to establish that the syntax of SCT (with whichever type
-constructors we choose) presents the initial indexed virtual equipment
-(with the appropriate universal properties). To start we define
-appropriate categories of equipments.
-
 \begin{definition}{Indexed Virtual Equipment}
-  A VE-hyperdoctrine is a functor $\cat C^{op} \to VE$, and a morphism
-  of VE-hyperdoctrines is a natural transformation.
+  A VE-hyperdoctrine is a pair of a category $C$ and a functor $\cat
+  C^{op} \to VE$. A morphism of VE-hyperdoctrines
+  \[(F,\phi) : (\cat C, \mathcal V : \cat C^{o} \to VE) \to (\cat D, \mathcal U : \cat D^{o} \to VE)\]
+  is a pair of a functor $F : \cat C \to \cat D$ and a natural
+  transformation $\phi : \mathcal V \Rightarrow\mathcal U \circ F^{o}$.
 \end{definition}
 
-\begin{definition}[Internal Logic]
-  Given any virtual equipment $\mathcal V$, we can construct the
-  ``internal logic'' by indexing $\mathcal V$ by $\Set$:
-
+\begin{definition}{Soundness}
+
+\end{definition}
+
+\begin{definition}{Completeness}
+  We define the syntactic VE-hyperdoctrine $\textrm{Syn}$ as follows:
   \begin{enumerate}
-  \item The indexing cateogry is $\Set$.
-  \item Given $\Gamma \in \Set$, the virtual equipment $\mathcal
-    V^\Gamma$ is the ``power'' of $\mathcal V$ by $\Gamma$:
+  \item The base category is the category of contexts $\Gamma$ and substitutions.
+  \item Define $\mathcal V_\Gamma$ as follows
     \begin{enumerate}
-    \item Objects are functions $\Gamma \to \mathcal V_0$.
-    \item A vertical morphism from $O$ to $O'$ is a function $\Pi_{\gamma\in\Gamma} V_v(O\gamma,O'\gamma)$.
-    \item A horizontal morphism is a function $\Gamma \to \mathcal V_h$. This satisfies
-    \item A 2-cell from $R$ to $S$ is a family $\Pi_{\gamma\in\Gamma} \mathcal V_h(R\gamma,S\gamma)$.
+    \item An object of $\mathcal V^\Gamma$ is a proof $\Gamma \vdash \cat C \isaCat$.
+    \item A vertical morphism $\mathcal V^\Gamma(\cat C, \cat D)$ is a
+      term $\Gamma\pipe \alpha:\cat C \vdash t : \cat D$ moduleo
+      $\alpha$-equivalence. Composition and identity are given by composition/variable.
+    \item A horizontal morphism $R \in \mathcal^\Gamma_h$ is a term
+      $\Gamma\pipe \alpha:\cat C;\beta:\cat D \vdash R$ with $d^-R =
+      \cat C$ and $d^+R = \cat D$.
+    \item Restriction of $R$ along $s,t$ is given by substitution
+      $R[s/\alpha;t/\beta]$. This is clearly associative and unital.
+    \item A 2-cell from $\Phi$ to $A$ is a term $\Phi \vdash t : B$
     \end{enumerate}
-    Compositions and restrictions are given pointwise.
-  \item Restriction by a function $\gamma : \Delta \to \Gamma$ is
-    given pointwise.
   \end{enumerate}
@@
 
-\begin{definition}
-  Given a well-formed signature $\Gamma$ and a virtual equipment
-  $\mathcal V$, we mutually define the following notions:
-
-  \begin{itemize}
-  \item The set $\sem{\Gamma}_{\mathcal V}$
-  \item For any $\Gamma \vdash A$ and $\gamma \in \sem{\Gamma}_{\mathcal V}$, a set $\sem{A}_{\mathcal V} \gamma$
-  \item For any $\Gamma \vdash M : A$ and $\gamma \in \sem{\Gamma}_{\mathcal V}$, an element $\sem{M}_{\mathcal V} \in \sem{A}_{\mathcal V} \gamma$
-  \item For any $\Gamma \vdash \cat C \isaCat$ and $\gamma \in \sem{\Gamma}_{\mathcal V}$, an object $\sem{\cat C}_{\mathcal V}\gamma$
-  \item For any $\Gamma\pipe\alpha:\cat C \vdash a :\cat D$ and $\gamma \in \sem{\Gamma}_{\mathcal V}$, a vertical morphism $\sem{\alpha.a}_{\mathcal V}\gamma \in \mathcal V_v(\sem{\cat C}_{\mathcal V}\gamma, \sem{\cat D}_{\mathcal V}\gamma,)$
-  \item For any $\Gamma\pipe\Xi \vdash R \isaSet$ and $\gamma \in
-    \sem{\Gamma}_{\mathcal V}$, a horizontal arrow $\sem{R}_{\mathcal
-      V}\gamma \in \mathcal V_h$ such that $d^\pm\sem{R}_{\mathcal
-      V}\gamma = \sem{d^\pm\Xi}_{\mathcal V}\gamma$. We also extend
-    this in the obvious way to transformation contexts
-    $\sem{\Phi}_{\mathcal V}\gamma$
-  \item For any $\Gamma\pipe\Phi \vdash t : R$ and $\gamma \in
-    \sem{\Gamma}_{\mathcal V}$, a 2-cell $\sem{t}_{\mathcal V}\gamma
-    \in V_h(\sem{\Phi}_{\mathcal V}\gamma;\sem{R}_{\mathcal V}\gamma)$
-    with $d^\pm\sem{t}_{\mathcal V}\gamma = \id$.
-  \end{itemize}
-
-  %% \begin{align*}
-  %%   \sem{\cdot}_{\mathcal V} &= 1\\
-  %%   \sem{\Gamma, x:A}_{\mathcal V} &= \sum_{\gamma \in \sem{\Gamma}_{\mathcal V}} \sem{A}_{\mathcal V}\gamma\\
-  %% \end{align*}
-
-  %% with an 
-  %% \begin{align*}
-  %%   \sem{\Gamma, x:A}_{\mathcal V} &= \sem{\Gamma}_{\mathcal V} \times \sem{A}_{\mathcal V}\\
-  %%   \sem{\Gamma}_{\mathcal V} &= 1\\\\
-  %%   \sem{\Sort}_{\mathcal V} &= \mathcal V_0\\
-  %%   \sem{\varr{\cat C}{\cat D}}_{\mathcal V} &= \mathcal V_v(\sem{\cat C})
-  %% \end{align*}
-\end{definition}
+%% \begin{definition}
+%%   Given a well-formed signature $\Gamma$ and a virtual equipment
+%%   $\mathcal V$, we mutually define the following notions:
+
+%%   \begin{itemize}
+%%   \item The set $\sem{\Gamma}_{\mathcal V}$
+%%   \item For any $\Gamma \vdash A$ and $\gamma \in \sem{\Gamma}_{\mathcal V}$, a set $\sem{A}_{\mathcal V} \gamma$
+%%   \item For any $\Gamma \vdash M : A$ and $\gamma \in \sem{\Gamma}_{\mathcal V}$, an element $\sem{M}_{\mathcal V} \in \sem{A}_{\mathcal V} \gamma$
+%%   \item For any $\Gamma \vdash \cat C \isaCat$ and $\gamma \in \sem{\Gamma}_{\mathcal V}$, an object $\sem{\cat C}_{\mathcal V}\gamma$
+%%   \item For any $\Gamma\pipe\alpha:\cat C \vdash a :\cat D$ and $\gamma \in \sem{\Gamma}_{\mathcal V}$, a vertical morphism $\sem{\alpha.a}_{\mathcal V}\gamma \in \mathcal V_v(\sem{\cat C}_{\mathcal V}\gamma, \sem{\cat D}_{\mathcal V}\gamma,)$
+%%   \item For any $\Gamma\pipe\Xi \vdash R \isaSet$ and $\gamma \in
+%%     \sem{\Gamma}_{\mathcal V}$, a horizontal arrow $\sem{R}_{\mathcal
+%%       V}\gamma \in \mathcal V_h$ such that $d^\pm\sem{R}_{\mathcal
+%%       V}\gamma = \sem{d^\pm\Xi}_{\mathcal V}\gamma$. We also extend
+%%     this in the obvious way to transformation contexts
+%%     $\sem{\Phi}_{\mathcal V}\gamma$
+%%   \item For any $\Gamma\pipe\Phi \vdash t : R$ and $\gamma \in
+%%     \sem{\Gamma}_{\mathcal V}$, a 2-cell $\sem{t}_{\mathcal V}\gamma
+%%     \in V_h(\sem{\Phi}_{\mathcal V}\gamma;\sem{R}_{\mathcal V}\gamma)$
+%%     with $d^\pm\sem{t}_{\mathcal V}\gamma = \id$.
+%%   \end{itemize}
+
+%%   %% \begin{align*}
+%%   %%   \sem{\cdot}_{\mathcal V} &= 1\\
+%%   %%   \sem{\Gamma, x:A}_{\mathcal V} &= \sum_{\gamma \in \sem{\Gamma}_{\mathcal V}} \sem{A}_{\mathcal V}\gamma\\
+%%   %% \end{align*}
+
+%%   %% with an 
+%%   %% \begin{align*}
+%%   %%   \sem{\Gamma, x:A}_{\mathcal V} &= \sem{\Gamma}_{\mathcal V} \times \sem{A}_{\mathcal V}\\
+%%   %%   \sem{\Gamma}_{\mathcal V} &= 1\\\\
+%%   %%   \sem{\Sort}_{\mathcal V} &= \mathcal V_0\\
+%%   %%   \sem{\varr{\cat C}{\cat D}}_{\mathcal V} &= \mathcal V_v(\sem{\cat C})
+%%   %% \end{align*}
+%% \end{definition}
 
@@
 
+\end{definition}
+
+\section{Example Models}
+
+Our main intended models are when the indexing category is $\Set$ and
+the hyperdoctrine is given by families of objects, arrows etc, of a
+fixed virtual equipment $\mathcal V$, indexed by the set
+$\sem\Gamma$. We can formalize this as the ``internal logic'' of the
+virtual equipment $\mathcal V$.
+
+\begin{definition}[Powers of Virtual Equipments]
+  Let $\mathcal V$ be a virtual equipment and let $X$ be a set.
+  Define a virtual equipment, the ``power'' $\mathcal V^X$ as follows.
+
+  \begin{enumerate}
+  \item Objects are functions $X \to \mathcal V_0$.
+  \item Vertical morphisms $\mathcal V^X_v(A,B)$ are functions
+    $\Pi_{x\inX} V_v(A x,B x)$. With pointwise composition, identity.
+  \item Horizontal morphisms $\mathcal V^X_h$ are functions $\Gamma \to \mathcal V_h$, with
+    $(d^\pm R)x = d^\pm (Rx)$.
+  \item Restriction of $R$ along $f,g$ is given by $(R(f,g))x = (R x)(f x,g x)$.
+  \item A 2-cell from $\Phi$ to $R$ is a family $\Pi_{x\in X} \mathcal
+    V^X_h(\Phix,R x)$ with composition given pointwise.
+  \end{enumerate}
+\end{definition}
+
+\begin{definition}[Internal Logic]
+  Given any virtual equipment $\mathcal V$, we can construct the
+  ``internal logic'' by indexing $\mathcal V$ by $\Set$:
+
+  \begin{enumerate}
+  \item The indexing cateogry is $\Set$.
+  \item The equipment is defined as $\mathcal U\Gamma = \mathcam V^\Gamma$.
+  \item Restriction by a function $\gamma : \Delta \to \Gamma$ is
+    given pointwise.
+  \end{enumerate}
 \end{definition}

### 2022-04-21 paper/paper.tex
category of elements, category with families? and start of GAT semantics

@@
 \newcommand{\equalizeVofWbyXeqYatZ}[5]{\{ {#1} : {#2} \pipe {#3} = {#4} : {#5} \}}
+
+\newcommand{\paramPresheaf}[1]{\mathcal P^{#1}}
+\newcommand{\pmPresheaf}{\paramPresheaf\pm}
+\newcommand{\posPresheaf}{\paramPresheaf+}
+\newcommand{\negPresheaf}{\paramPresheaf-}
+\newcommand{\negPresheafApp}[2]{#1\,#2}
+\newcommand{\posPresheafApp}[2]{#1\,#2}
 
@@
 
-\subsection{Graph of a Profunctor}
-
-One final multiplicative we can define is the \emph{graph} of a
-profunctor. Given a profunctor $\alpha:\cat C;\beta:\cat D \vdash R$,
-the ``graph'' of $R$ is a category $\exists \alpha;\beta. R$. In
-ordinary category theory, this is usually defined as the category of
-elements of $R$ viewed as a functor $R : \cat C^o \times D \to \Set$.
-In detail, its objects are triples $(a:\cat C, r : R, b : \cat D)$ and
-morphism from $(a,r,b)$ to $(a', r', b')$ consists of a pair of
-morphisms $g : b \to b'$ and a backwards $f : a' \to a$ such that $r'
-= g\circ r \circ f$.
+\subsection{Category of Elements}
+
+One final multiplicative we can define is the \emph{category of
+elements} of a profunctor. Given a profunctor $\alpha:\cat
+C;\beta:\cat D \vdash R$, the ``category of elements'' of $R$ is a
+category $\Sigma_{\alpha:\cat C;\beta:\cat D}. R$. In detail, its
+objects are triples $(a:\cat C, r : R, b : \cat D)$ and morphism from
+$(a,r,b)$ to $(a', r', b')$ consists of a pair of morphisms $g : b \to
+b'$ and a backwards $f : a' \to a$ such that $r' = g\circ r \circ f$.
 %
@@
   \caption{Graph of a Profunctor}
+\end{figure}
+
+\subsection{Positive and Negative Presheaf Categories}
+
+Though defining arbitrary functor categories type-theoretically would
+require a ``multi-argument functor'' judgment, we can define
+\emph{presheaf} categories by the universal property that a functor
+$\cat C \to \mathcal P {\cat D}$ is the same as a profunctor from
+$\cat D$ to $\cat C$. To distinguish the covariant from contravariant
+presheaves, we define them dually as $\negPresheaf {\cat C}$ and
+$\posPresheaf {\cat D}$
+
+\begin{figure}
+  \begin{mathpar}
+    \inferrule*[right=NegPresheaf-Formation]
+    {\Gamma \vdash \cat C \isaCat}
+    {\Gamma \vdash \negPresheaf {\cat C} \isaCat}
+
+    \inferrule*[right=NegPresheaf-E]
+    {\Gamma \pipe d^-\Xi \vdash a : \cat C \and
+     \Gamma \pipe d^+\Xi \vdash p : \negPresheaf {\cat C}
+    }
+    {\Gamma \pipe \Xi \vdash \negPresheafApp a p \isaSet}
+
+    \inferrule*[right=NegPresheaf-I]
+    {\Gamma \pipe \alpha:\cat C;\beta:\cat D \vdash R : \isaSet}
+    {\Gamma \pipe \beta:\cat D \vdash \lambda \alpha:\cat C. R : \negPresheaf {\cat C}}
+
+    \negPresheafApp a (\lambda \alpha. R) = R[a/\alpha]\\
+
+    \inferrule*[right=NegPresheaf-$\eta$]
+    {\Gamma \pipe \beta:\cat D \vdash p : \negPresheaf {\cat C}}
+    {p = \lambda \alpha. \negPresheafApp \alpha p}
+
+    \inferrule*[right=PosPresheaf-Formation]
+    {\Gamma \vdash \cat D \isaCat}
+    {\Gamma \vdash \posPresheaf {\cat D} \isaCat}
+
+    \inferrule*[right=PosPresheaf-E]
+    {\Gamma \pipe d^-\Xi \vdash p : \posPresheaf {\cat D}\and
+     \Gamma \pipe d^+\Xi \vdash a : \cat D
+    }
+    {\Gamma \pipe \Xi \vdash \posPresheafApp p a \isaSet}
+
+    \inferrule*[right=PosPresheaf-I]
+    {\Gamma \pipe \alpha:\cat C;\beta:\cat D \vdash R : \isaSet}
+    {\Gamma \pipe \alpha:\cat C \vdash \lambda \beta:\cat D. R : \posPresheaf {\cat D}}
+
+    \posPresheafApp {(\lambda \beta. R)} b = R[b/\beta]\\
+
+    \inferrule*[right=PosPresheaf-$\eta$]
+    {\Gamma \pipe \alpha:\cat C \vdash p : \posPresheaf {\cat D}}
+    {p = \lambda \beta. \posPresheafApp p \beta}
+  \end{mathpar}
+  \caption{Presheaf Categories}
 \end{figure}
@@
 
+\subsection{Yoneda Lemma and Representable Presheaves}
+
+Given any category $\cat C$, we can define the positive and negative
+Yoneda embeddings $Y^\pm : \cat C \to \pmPresheaf {\cat C}$ using the
+hom profunctor. Note that both of these functors are covariant in
+$\cat C$, so semantically, $\posPresheaf {\cat C}$ should be
+interpreted as $(\Set^{\cat C})^o$.
+
+\begin{definition}
+  Define the negative Yoneda embedding $Y^-[\alpha]$
+  \[ \alpha:\cat C \vdash Y^-[\alpha] = \lambda \alpha':\cat C. \punitinXfromYtoZ {\cat C}{\alpha'}{\alpha} : \negPresheaf {\cat C} \]
+
+  The positive Yoneda embedding is defined similarly $Y^+[\alpha]$.
+\end{definition}
+
+The Yoneda lemma says that the set of natural transformations from
+$Y^-[\alpha]$ to a presheaf $p$ is naturally isomorphic to the
+elements of the presheaf itself: $\negPresheafApp \alpha p$. In
+encoding this statement in our type theory, we find that we can use
+two different interpretations of ``natural transformations''.
+
+On the one hand, a natural transformation should be a morphism in the
+presheaf category. This interpretation of the Yoneda lemma would mean
+that we have a natural isomorphism of profunctors:
+\[ \alpha:\cat C; \pi:\negPresheaf {\cat C} \vdash \punitinXfromYtoZ{\negPresheaf {\cat C}} {Y^-[\alpha]} {\pi} \cong \negPresheafApp \alpha \pi \]
+Attempting to prove this we become stuck very quickly because we know
+nothing about the structure of morphisms in the presheaf category.
+
+On the other hand, a natural transformation of presheaves $\pi$ and
+$\rho$ can be defined as a family of functions from $\negPresheafApp
+\alpha \pi$ to $\negPresheafApp \alpha \rho$, or in other words as an
+\emph{end}. Specializing to the Yoneda lemma, we interpret ``a natural
+transformation from $Y^-[\alpha]$ to $\pi$'' as the following, which simplifies using $\beta$ reduction:
+\[ \alpha:\cat C; \pi:\negPresheaf {\cat C} \vdash \homrallXYtoZ {\alpha':\cat C} {\negPresheafApp {\alpha'} {Y^-[\alpha]}}{\negPresheafApp {\alpha'} \pi} = \homrallXYtoZ {\alpha':\cat C} {(\punitinXfromYtoZ {\cat C}{\alpha'}{\alpha})}{(\negPresheafApp {\alpha'} \pi)} \]
+And then the Yoneda lemma becomes:
+\[ \alpha:\cat C; \pi:\negPresheaf {\cat C} \vdash \homrallXYtoZ {\alpha':\cat C} {(\punitinXfromYtoZ {\cat C}{\alpha'}{\alpha})}{(\negPresheafApp {\alpha'} \pi)} \cong \negPresheafApp \alpha \pi\]
+Which is an instance of our earlier Lemma\,\ref{lem:yo-and-co-yo}
+which follows quite easily.
+
+Then we are left with the question of whether these two forms of the
+Yoneda lemma are equivalent, or slightly more generally are these two definitions of natural transformation of presheaves equivalent:
+\[ \rho:\negPresheaf{\cat C}; \pi:\negPresheaf {\cat C} \vdash \punitinXfromYtoZ{\negPresheaf {\cat C}} {\rho} {\pi} \]
+and
+\[ \rho:\negPresheaf{\cat C}; \pi:\negPresheaf {\cat C} \vdash \homrallXYtoZ {\alpha:\cat C} {\negPresheafApp \alpha \rho}{\negPresheafApp \alpha \pi} \]
+While we can show that there is a morphism from the hom set to the
+end, the inverse does not seem provable, and seems analogous to the
+\emph{univalence} axiom in homotopy type theory, as it tells us that
+the morphisms of the universe are given by a particular kind of
+function in our logic.
+
 \subsection{Adjoint Functors and Representable Profunctors}
@@
 
+\subsection{Formal Type Theory}
+
+\begin{definition}[Category with Families]
+  A category with families consists of
+  \begin{enumerate}
+  \item A category $\textrm{Ctx}\isaCat$
+  \item A presheaf $\textrm{Ty} : \harr {\textrm{Ctx}}{1}$
+  \item A presheaf $\textrm{Tm} : \harr {\Sigma_{\Gamma:\textrm{Ctx};*:1}\harrapp{\textrm{Ty}} \Gamma *} 1$
+  \item A terminal object in $\textrm{Ctx}$
+  \item An \emph{extension operation}
+    $\textrm{ext} : \Sigma_{\Gamma:\textrm{Ctx};*:1}\harrapp{\textrm{Ty}}\Gamma * \to \textrm{Ctx}$
+    right representing the profunctor
+    \[ \Delta:\textrm{Ctx};A:\Sigma_{\Gamma:\textrm{Ctx};*:1}\harrapp{\textrm{Ty}}{\Gamma}{*} \vdash \tensorexistsXwithYandZ {a}{(\punitinXfromYtoZ {} \Delta {\pi^-\pi^-a})}
+       {(\punitinXfromYtoZ {}{\pi^-a}{A})}\]
+       % {\Sigma_{\Gamma:\textrm{Ctx};*:1}\harrapp{\textrm{Ty}}{\Gamma}{*}}
+    % 
+  \end{enumerate}
+\end{definition}
+
 \section{Semantics}
@@
 category that interprets dependent type theory. We call such a
-structure as a \emph{VDC-hyperdoctrine}. We then show how to construct
-a VDC-hyperdoctrine indexed by the category of sets, which has the
-intuitive reading that any of the VDC judgments $\Gamma \vdash J$ is
-interpreted as simply a family of $J$-data parameterized by elements
-of the set $\Gamma$.
+structure a \emph{VDC-hyperdoctrine}. We then show how to construct a
+from any VDC, a VDC-hyperdoctrine indexed by the category of sets,
+which has the intuitive reading that any of the VDC judgments $\Gamma
+\vdash J$ is interpreted as simply a family of $J$-data parameterized
+by elements of the set $\Gamma$.
 
@@
 \begin{definition}{Indexed Virtual Equipment}
-  A VE-hyperdoctrine is a pair of a category $C$ and a functor $\cat
-  C^{op} \to VE$. A morphism of VE-hyperdoctrines
+  A VE-hyperdoctrine is a pair of a category with families $(C,
+  \textrm{Ty}, \textrm{Tm})$ and a functor $\cat C^{op} \to VE$. A
+  morphism of VE-hyperdoctrines
   \[(F,\phi) : (\cat C, \mathcal V : \cat C^{o} \to VE) \to (\cat D, \mathcal U : \cat D^{o} \to VE)\]
-  is a pair of a functor $F : \cat C \to \cat D$ and a natural
-  transformation $\phi : \mathcal V \Rightarrow\mathcal U \circ F^{o}$.
-\end{definition}
-
-\begin{definition}{Soundness}
-
-\end{definition}
-
-\begin{definition}{Completeness}
-  We define the syntactic VE-hyperdoctrine $\textrm{Syn}$ as follows:
-  \begin{enumerate}
-  \item The base category is the category of contexts $\Gamma$ and substitutions.
-  \item Define $\mathcal V_\Gamma$ as follows
-    \begin{enumerate}
-    \item An object of $\mathcal V^\Gamma$ is a proof $\Gamma \vdash \cat C \isaCat$.
-    \item A vertical morphism $\mathcal V^\Gamma(\cat C, \cat D)$ is a
-      term $\Gamma\pipe \alpha:\cat C \vdash t : \cat D$ moduleo
-      $\alpha$-equivalence. Composition and identity are given by composition/variable.
-    \item A horizontal morphism $R \in \mathcal^\Gamma_h$ is a term
-      $\Gamma\pipe \alpha:\cat C;\beta:\cat D \vdash R$ with $d^-R =
-      \cat C$ and $d^+R = \cat D$.
-    \item Restriction of $R$ along $s,t$ is given by substitution
-      $R[s/\alpha;t/\beta]$. This is clearly associative and unital.
-    \item A 2-cell from $\Phi$ to $A$ is a term $\Phi \vdash t : B$
-    \end{enumerate}
-  \end{enumerate}
+  is a pair of a functor of categories with families $F : \cat C \to
+  \cat D$ and a natural transformation $\phi : \mathcal V
+  \Rightarrow\mathcal U \circ F^{o}$.
 \end{definition}
@@
 
-And we define a profunctor of interpretations of a signature in a
-virtual equipment as follows:
-\begin{definition}
-  An interpretation of a signature $\Gamma$ in a virtual equipment
-  $\mathcal V$ is an element of the set $\sem{\Gamma}_{\mathcal
-    V}$. Call this set $\textrm{Interp}(\Gamma;\mathcal V)$.
-
-  Note that this constitutes a profunctor contravariant in $\Gamma$
-  and covariant in $\mathcal V$.
-\end{definition}
-
-Next, we define the syntactic virtual equipment $\textrm{Syn}(\Gamma)$
-for a signature $\Gamma$.
-\begin{definition}
-  For any $\Gamma \isadtctx$, we construct a virtual double category
-  $V^\Gamma$.
-  \begin{enumerate}
-  \item $V^\Gamma_v$ is the set of terms $\Gamma \vdash \cat C \isaCat$.
-  \item $V^\Gamma_v(\cat C, \cat D)$ is the set of objects in context
-    $\Gamma \pipe \alpha:\cat C \vdash a : \cat D$. The category
-    structure is given by substitution, with the variable $\alpha
-    \vdash \alpha$ the identity.
-  \item $V_h$ is the set of triples $(\cat C, \cat D, \Gamma \pipe
-    \alpha:\cat C;\beta:\cat D \vdash R \isaSet)$ where $d^\pm$ are
-    the projections
-  \item $V^\Gamma_h(\vec R; S)$ is the set of triples $(a,b,\Gamma
-    \pipe \Phi \vdash t : S[a;b])$ where $\Phi$ is the context
-    corresponding to $\vec R$ and $d^\pm$ are the obvious projections.
-
-  \item The identity morphisms are $(\cat C, \cat D, R)$ is $\Gamma\pipe\alpha:\cat C, x:R\beta:\cat D \vdash x:R$
-  \item Composition $(a,b,t) \circ [(a_1,),\ldots]$ is given by
-    substitution in each component: $(a[a_1],b[a_n], t[\phi])$.
-  \item Associativity and unit rules for composition follow from the
-    definition of substitution by a straightforward induction.
-  \end{enumerate}
-\end{definition}
-
-Then the soundness theorem says that for any $\Gamma$, the virtual
-double category $\textrm{Syn}(\Gamma)$ provides an isomorphism of profunctors:
-\[ \textrm{Interp}(\Gamma, \mathcal V \cong VE(\textrm{Syn}(\Gamma), \mathcal V) \]
-As we saw in a previous section, the backwards direction of this
-isomorphism should be a universal interpretation:
-
-\begin{definition}[Universal Interpretation]
-  For any signature $\Gamma$ there is a ``self-interpretation'' $\eta:
-  \textrm{Interp}(\Gamma, \textrm{Syn}(\Gamma))$ defined as the
-  identity substitution $\id \in \sem{\Gamma}_{\textrm{Syn}(\Gamma)}$
-
-\end{definition}
-
-\section{Example Models}
+\subsection{Generalized Algebraic Theory of VE Hyperdoctrines}
+
+To simplify our initiality proof, we work with an \emph{essentially
+algebraic} presentation of the notion of VE hyperdoctrine. As with all
+GATs, this has an initial model, which is a kind of ``combinator''
+syntax with explicit substitution for VE hyperdoctrines. Then we show
+that our type theoretic syntax is equivalent to the combinator syntax
+by a simple inductive ``simplification'' algorithm. We conjecture that
+a full normalization proof is possible, which would establish useful
+conservativity results for the calculus.
+
+The theory has sorts Ctx, Subst, Tm, Ty, Cat, Fun, Prof, ProfCtx and
+Trans.
+
+\subsection{Example Models}
 
@@
   \item Vertical morphisms $\mathcal V^X_v(A,B)$ are functions
-    $\Pi_{x\inX} V_v(A x,B x)$. With pointwise composition, identity.
+    $\Pi_{x\in X} V_v(A x,B x)$. With pointwise composition, identity.
   \item Horizontal morphisms $\mathcal V^X_h$ are functions $\Gamma \to \mathcal V_h$, with
@@
   \item A 2-cell from $\Phi$ to $R$ is a family $\Pi_{x\in X} \mathcal
-    V^X_h(\Phix,R x)$ with composition given pointwise.
+    V^X_h(\Phi x,R x)$ with composition given pointwise.
   \end{enumerate}
@@
   \item The indexing cateogry is $\Set$.
-  \item The equipment is defined as $\mathcal U\Gamma = \mathcam V^\Gamma$.
+  \item The equipment is defined as $\mathcal U\Gamma = \mathcal V^\Gamma$.
   \item Restriction by a function $\gamma : \Delta \to \Gamma$ is

### 2022-04-23 paper/paper.tex
overview of the algebraic semantics

@@
 
+Since many type theories are given semantics by free structured
+categories, we can use SCT as a logical framework in which to do
+``formal type theory''.
+
+\begin{definition}[Theory of unary Computational lambda calculus with booleans]
+  \begin{enumerate}
+  \item A category $C$
+  \item A terminal object $1$ of $C$
+  \item A monad $T$ on $C$
+  \item A type $\textrm{Bit} : \varr 1 C$
+  \item Morphisms $\textrm{y},\textrm{n} : \pendallXdotY * \punitinXfromYtoZ {1(*)} {\textrm{Bit}(*)}$
+  \item A natural transformation $\textrm{if} : \pendallXdotY * \homrallXYtoZ \alpha {(\punitinXfromYtoZ C {1(*)} \alpha) \times (\punitinXfromYtoZ C {1(*)} \alpha)} {\punitinXfromYtoZ C {\textrm{Bit}(*)} \alpha}$
+  \item Satisfying $\textrm{if} y = \pi_1$ and $\textrm{if} n = \pi_2$
+  \item A morphism $\textrm{print} : \pendallXdotY * \punitinXfromYtoZ {\textrm{Bit}(*)}{T (1(*))}$
+  \end{enumerate}
+\end{definition}
+
+Then we can ``program'' in this computational lambda calculus within
+our logical framework by constructing terms of type $\punitinXfromYtoZ
+C {A} {B}$ for any functors $A,B$ into $C$.
+
+Using a terminal category and the category of elements, we can define
+a formal category with families, an algebraic formulation of the
+judgmental structure of dependent type theory:
 \begin{definition}[Category with Families]
@@
 
-To simplify our initiality proof, we work with an \emph{essentially
-algebraic} presentation of the notion of VE hyperdoctrine. As with all
-GATs, this has an initial model, which is a kind of ``combinator''
-syntax with explicit substitution for VE hyperdoctrines. Then we show
-that our type theoretic syntax is equivalent to the combinator syntax
-by a simple inductive ``simplification'' algorithm. We conjecture that
-a full normalization proof is possible, which would establish useful
-conservativity results for the calculus.
-
-The theory has sorts Ctx, Subst, Tm, Ty, Cat, Fun, Prof, ProfCtx and
-Trans.
+To simplify our initiality proof, we use an \emph{essentially
+algebraic} presentation of our notion of model. To describe this, we
+give a ``combinator'' syntax for VE hyperdoctrines, where we have
+explicit sorts for contexts $\Gamma$ and $\Phi$ and explicit judgments
+for substitutions $\gamma : \Delta \to \Gamma$ and $\phi : \Psi
+\vdash_{f,g} \Phi$.
+%
+While the explicit description of the theory is quite long, we can
+describe it in simple modular pieces.
+%
+First, we use the standard theory of a category with families, which
+describes the algebraic structure of contexts $\Gamma$, types $\Gamma
+\vdash A$ and terms $\Gamma \vdash M : A$.
+%
+Second, we add to that a theory for a slight generalization of virtual
+double categories we call \emph{double categories with display
+pro-arrows}.
+%
+The difference is that in a double category with display pro-arrows,
+we have a true double category with strict vertical and horizontal
+composition, among whom are the ``display'' pro-arrows.
+%
+The display pro-arrows model the profunctors of the syntax, which
+don't have a strict composition, whereas the arbitrary pro-arrows
+model the contexts $\Phi$, which do have a strict composition, given
+by context concatenation $\Phi,\Psi$.
+%
+The difference is that in a virtual double category the contexts are
+assumed to be given inductively by formal compositions of the display
+pro-arrows, whereas in models of this algebraic theory, the contexts
+might include other exotic objects.
+%
+This difference is analogous to the difference between a category with
+families and a contextual category in models of dependent type theory,
+and we fully expect that there is a similar relationship between the
+models: virtual double categories should form a co-reflective
+subcategory characterized as those double categories with display
+pro-arrows where all pro-arrows are equal to some concatenation of
+displays.
+%
+Finally, we parameterize all of the constructions of the theory of
+double categories with pro-arrows by contexts $\Gamma$ and add in
+equations that substitution commutes with all double category
+operations.
+%
+This means the theory presents a category with families with an double
+category with display pro-arrows internal to the category of
+presheaves on the category of contexts, i.e., a hyperdoctrine of
+double categories with display pro-arrows.
+
+The main advantage of this algebraic approach is that the initiality
+theorem for the combinatory syntax is trivial. Instead, the work is to
+show that our type theoretic syntax, which restricts combinatory rules
+to make substitution an admissible rather than primitive operation, is
+equivalent to the combinatory syntax.
+
+\begin{figure}
+  \begin{mathpar}
+    \inferrule
+    {}
+    {\Gamma : \isaCtx}
+
+    \inferrule
+    {\Delta : \isaCtx \and \Gamma : \isaCtx}
+    {\gamma : \textrm{Subst}(\Delta,\Gamma)}
+
+    \inferrule
+    {\Gamma : \isaCtx}
+    {C : \textrm{Cat}(\Gamma)}
+
+    \inferrule
+    {\Gamma : \isaCtx \and C : \textrm{Cat}(\Gamma) \and D : \textrm{Cat}(\Gamma) \and
+    }
+    {F : \textrm{Functor}(\Gamma,C,D)}
+
+    \inferrule
+    {\Gamma : \isaCtx \and C : \textrm{Cat}(\Gamma) \and D : \textrm{Cat}(\Gamma) \and
+    }
+    {P : \textrm{Profunctor}(\Gamma,C,D)}
+
+    \inferrule
+    {\Gamma : \isaCtx \and C : \textrm{Cat}(\Gamma) \and D : \textrm{Cat}(\Gamma)
+    }
+    {P : \textrm{ProfCtx}(\Gamma,C,D)}
+
+    \inferrule
+    {\Gamma : \textrm{Ctx}\and
+      C  : \textrm{Cat}(\Gamma) \and D  : \textrm{Cat}(\Gamma) \and
+      C' : \textrm{Cat}(\Gamma) \and D' : \textrm{Cat}(\Gamma) \and
+      f : \textrm{Functor}(\Gamma,C,C') \and g : \textrm{Cat}(\Gamma,D,D') \and
+      \Phi : \textrm{ProfCtx}(\Gamma,C,D)\and
+      \Psi : \textrm{ProfCtx}(\Gamma,C',D')
+    }
+    {\phi : \textrm{ProfSubst}(\Gamma,C,D,C',D',f,g,\Phi,\Psi)}
+
+    \inferrule
+    {\Gamma : \textrm{Ctx}\and
+      C  : \textrm{Cat}(\Gamma) \and D  : \textrm{Cat}(\Gamma) \and
+      C' : \textrm{Cat}(\Gamma) \and D' : \textrm{Cat}(\Gamma) \and
+      f : \textrm{Functor}(\Gamma,C,C') \and g : \textrm{Cat}(\Gamma,D,D') \and
+      \Phi : \textrm{ProfCtx}(\Gamma,C,D)\and
+      P : \textrm{ProfCtx}(\Gamma,C',D')
+    }
+    {t : \textrm{Trans}(\Gamma,C,D,C',D',f,g,\Phi,P)}
+  \end{mathpar}
+  \caption{Sorts of Algebraic SCT}
+\end{figure}
+
+\begin{figure}
+  \begin{mathpar}
+    \cdots
+  \end{mathpar}
+  \caption{Algebraic Dependent Type Theory}
+\end{figure}
+
+\begin{figure}
+  \begin{mathpar}
+    \inferrule
+    {g : \textrm{Functor}(D,E) \and \textrm{Functor}(C,D)}
+    {g \circ_{\Gamma,C,D,E} f : \textrm{Functor}(C,E)}
+
+    (h \circ g) \circ f = h \circ (g \circ f)
+
+    \inferrule
+    {}
+    {\id_C : \textrm{Functor}(C,C)}
+
+    (h \circ \id) = h
+
+    \id \circ h = h
+
+    \inferrule
+    {P : \textrm{Profunctor}(\Gamma,C',D')
+    \and f : \textrm{Functor}(\Gamma,C,C')\and
+    \and g : \textrm{Functor}(\Gamma,D,D')}
+    {P \circ_{\Gamma,C,D,C',D'} (f,g) : \textrm{Profunctor}(\Gamma,C,D)}
+
+    P \circ (f,g) \circ (f',g') = P \circ (f \circ f', g \circ g')
+
+    P \circ (\id,\id) = P
+
+    \inferrule
+    {\psi : \textrm{ProfSubst}(f',g',\Phi,\Psi)\and \phi : \textrm{ProfSubst}(f,g,\Sigma,\Phi)
+    }
+    {\psi \circ \phi : \textrm{ProfSubst}(f' \circ f, g' \circ g, \Sigma, \Psi)}
+
+    \inferrule
+    {}
+    {restrict(f,g,\phi) : \textrm{ProfSubst}(\id,\id,\Psi,\Phi)}
+  \end{mathpar}
+  \caption{Algebraic Presentation of SCT}
+\end{figure}
 

### 2022-04-23 paper/paper.tex
sketch of the GAT

@@
 \newcommand{\posPresheafApp}[2]{#1\,#2}
+
+% Algebraic notation
+\newcommand{\algCtx}{\textrm{Ctx}}
+\newcommand{\algSubst}{\textrm{Subst}}
+\newcommand{\algTy}{\textrm{Type}}
+\newcommand{\algTm}{\textrm{Term}}
+\newcommand{\algCat}{\textrm{Cat}}
+\newcommand{\algVarr}{\textrm{Functor}}
+\newcommand{\algHCtx}{\textrm{ProfCtx}}
+\newcommand{\algHSubst}{\textrm{ProfSubst}}
+\newcommand{\algHarr}{\textrm{Prof}}
+\newcommand{\algTrans}{\textrm{Trans}}
 
@@
 
-To simplify our initiality proof, we use an \emph{essentially
-algebraic} presentation of our notion of model. To describe this, we
-give a ``combinator'' syntax for VE hyperdoctrines, where we have
-explicit sorts for contexts $\Gamma$ and $\Phi$ and explicit judgments
-for substitutions $\gamma : \Delta \to \Gamma$ and $\phi : \Psi
+To simplify our initiality proof, we use a \emph{generalized
+algebraic} presentation of our notion of
+model\cite{cartmell,sterling-eat}. To describe this, we give a
+``combinator'' syntax for VE hyperdoctrines, where we have explicit
+sorts for contexts $\Gamma$ and $\Phi$ and explicit judgments for
+substitutions $\gamma : \Delta \to \Gamma$ and $\phi : \Psi
 \vdash_{f,g} \Phi$.
@@
     {}
-    {\Gamma : \isaCtx}
+    {\Gamma : \algCtx}
 
     \inferrule
-    {\Delta : \isaCtx \and \Gamma : \isaCtx}
-    {\gamma : \textrm{Subst}(\Delta,\Gamma)}
+    {\Delta : \algCtx \and \Gamma : \algCtx}
+    {\gamma : \algSubst(\Delta,\Gamma)}
 
     \inferrule
-    {\Gamma : \isaCtx}
-    {C : \textrm{Cat}(\Gamma)}
+    {\Gamma : \algCtx}
+    {C : \algCat(\Gamma)}
 
     \inferrule
-    {\Gamma : \isaCtx \and C : \textrm{Cat}(\Gamma) \and D : \textrm{Cat}(\Gamma) \and
+    {\Gamma : \algCtx \and C : \algCat(\Gamma) \and D : \algCat(\Gamma) \and
     }
-    {F : \textrm{Functor}(\Gamma,C,D)}
+    {F : \algVarr(\Gamma,C,D)}
 
     \inferrule
-    {\Gamma : \isaCtx \and C : \textrm{Cat}(\Gamma) \and D : \textrm{Cat}(\Gamma) \and
+    {\Gamma : \algCtx \and C : \algCat(\Gamma) \and D : \algCat(\Gamma) \and
     }
@@
     \inferrule
-    {\Gamma : \isaCtx \and C : \textrm{Cat}(\Gamma) \and D : \textrm{Cat}(\Gamma)
+    {\Gamma : \algCtx \and C : \algCat(\Gamma) \and D : \algCat(\Gamma)
     }
@@
     {\Gamma : \textrm{Ctx}\and
-      C  : \textrm{Cat}(\Gamma) \and D  : \textrm{Cat}(\Gamma) \and
-      C' : \textrm{Cat}(\Gamma) \and D' : \textrm{Cat}(\Gamma) \and
-      f : \textrm{Functor}(\Gamma,C,C') \and g : \textrm{Cat}(\Gamma,D,D') \and
+      C  : \algCat(\Gamma) \and D  : \algCat(\Gamma) \and
+      C' : \algCat(\Gamma) \and D' : \algCat(\Gamma) \and
+      f : \algVarr(\Gamma,C,C') \and g : \algCat(\Gamma,D,D') \and
       \Phi : \textrm{ProfCtx}(\Gamma,C,D)\and
@@
 
+The algebraic theory of a double category with displays indexed by a
+category is given in Figure~\ref{fig:alg-dc-disp}.
+%
+First, we give the categories and functors the theory of a category
+indexed by a category.
+%
+Next, we give categories and \emph{profunctor contexts} the same
+theory, where the identity $\id_h$ is the ``empty'' context $\alpha:C$ and the composition $\Phi \circ_h$.
+%
+
 \begin{figure}
   \begin{mathpar}
+    % theory of an indexed category (vertical)
     \inferrule
-    {g : \textrm{Functor}(D,E) \and \textrm{Functor}(C,D)}
-    {g \circ_{\Gamma,C,D,E} f : \textrm{Functor}(C,E)}
-
-    (h \circ g) \circ f = h \circ (g \circ f)
+    {g : \algVarr(D,E) \and \algVarr(C,D)}
+    {g \circ_v f : \algVarr(C,E)}
+
+    (g \circ_v f) \circ_{vs} \gamma = (g \circ_{vs} \gamma) \circ_v (f \circ_s \gamma)
+
+    (h \circ_v g) \circ_v f = h \circ_v (g \circ_v f)
 
@@
     {}
-    {\id_C : \textrm{Functor}(C,C)}
-
-    (h \circ \id) = h
-
-    \id \circ h = h
+    {\id_h : \algVarr(C,C)}
+
+    \id_h \circ_{vs} \gamma = \id_h
+
+    (h \circ \id) = h\and \id \circ h = h
+
+    % theory of an indexed category (horizontal)
+
+    % embedding of displays
+
+    % strict equipment (horizontal)
+    % strict equipment  (displays)
+
+    % 2-cells: substitutions
+    % vertical category
+    % horizontal category
+    % interchange law
+
+    % 2-cells: transformations : Q: is any of this necessary or should
+    % we just conflate transformations with substitutions into a
+    % display?
+    % 
+    % action of trans on subst with assoc, unit
+    % identity trans with unit
+    % embedding of transformations into substitutions
 
     \inferrule
-    {P : \textrm{Profunctor}(\Gamma,C',D')
-    \and f : \textrm{Functor}(\Gamma,C,C')\and
-    \and g : \textrm{Functor}(\Gamma,D,D')}
-    {P \circ_{\Gamma,C,D,C',D'} (f,g) : \textrm{Profunctor}(\Gamma,C,D)}
-
-    P \circ (f,g) \circ (f',g') = P \circ (f \circ f', g \circ g')
-
-    P \circ (\id,\id) = P
+    {P : \algHarr(\Gamma,C',D')
+    \and f : \algVarr(\Gamma,C,C')\and
+    \and g : \algVarr(\Gamma,D,D')}
+    {P \circ_r (f,g) : \algHarr(\Gamma,C,D)}
+
+    (P \circ_r (f,g) \circ_s \gamma) = (P \circ_s \gamma) \circ_r (f \circ_s \gamma, g \circ_s \gamma)
+
+    P \circ_r (f,g) \circ_r (f',g') = P \circ_r (f \circ_r f', g \circ_r g')
+
+    P \circ_r (\id,\id) = P
 
     \inferrule
-    {\psi : \textrm{ProfSubst}(f',g',\Phi,\Psi)\and \phi : \textrm{ProfSubst}(f,g,\Sigma,\Phi)
+    {\psi : \algHSubst(f',g',\Phi,\Psi)\and \phi : \algHSubst(f,g,\Sigma,\Phi)
     }
-    {\psi \circ \phi : \textrm{ProfSubst}(f' \circ f, g' \circ g, \Sigma, \Psi)}
+    {\psi \circ \phi : \algHSubst(f' \circ f, g' \circ g, \Sigma, \Psi)}
 
@@
     {}
-    {restrict(f,g,\phi) : \textrm{ProfSubst}(\id,\id,\Psi,\Phi)}
+    {restrict(f,g,\phi) : \algHSubst(\id,\id,\Psi,\Phi)}
   \end{mathpar}
-  \caption{Algebraic Presentation of SCT}
+  \caption{Equipment with Displays}
+  \label{fig:alg-dc-disp}
+\end{figure}
+
+\begin{figure}
+  \begin{mathpar}
+    % Cat Multiplicatives
+    % Cat of Elements
+    % Presheaf cats
+
+    % Prof Multiplicatives
+    % Unit
+    % Tensor
+    % Hom
+
+    % Internalization
+    % Cats
+    % Functor
+    % Prof
+    % ProfCtx?
+    % Trans
+
+    % Cat Additives
+    % Product
+    % Terminal
+
+    % Prof Additives
+    % Product
+    % Terminal
+  \end{mathpar}
+  \caption{UMPs in an Equipment}
 \end{figure}
@@
 %% \begin{definition}
-%%   For any $C : \textrm{Cat}$, we can define an identity functor
-%%   $\textrm{Id}_C : \textrm{Functor} C C$ by
+%%   For any $C : \algCat$, we can define an identity functor
+%%   $\textrm{Id}_C : \algVarr C C$ by
 %%   \begin{enumerate}
@@
 %% \begin{definition}
-%%   Given categories $C,D,E$ and functors $F : \textrm{Functor} C D$ and
-%%   $G : \textrm{Functor} D E$, we define the composition $G \circ F :
-%%   \textrm{Functor} C E$ as
+%%   Given categories $C,D,E$ and functors $F : \algVarr C D$ and
+%%   $G : \algVarr D E$, we define the composition $G \circ F :
+%%   \algVarr C E$ as
 %%   \begin{enumerate}
@@
 %%   \begin{enumerate}
-%%   \item A functor $G : \textrm{Functor}\, D C$
+%%   \item A functor $G : \algVarr\, D C$
 %%   \item An isomorphism $I,I^-$ between $Q$ and $\text{Hom}(C)[Id;G]$

### 2022-04-25 paper/paper.tex
GAT progress: double cats w/ displays, restrictions, cat of elements
GAT Profunctor Multiplicatives
more GATs, start initiality, sketch internal DRA/PRA

@@
 %% \newtheorem{lemma}{Lemma}
-%% \newtheorem{construction}{Construction}
+\newtheorem{construction}{Construction}
 %% \newtheorem{definition}{Definition}
@@
 \newcommand{\algTrans}{\textrm{Trans}}
+\newcommand{\algElts}{\textrm{Elts}}
+\newcommand{\algEltsI}{\textrm{EltsI}}
+
+\newcommand{\algLHom}{\textrm{LHom}}
+\newcommand{\algLHomI}{\textrm{LHomI}}
+\newcommand{\algLHomE}{\textrm{LHomE}}
+\newcommand{\algRHom}{\textrm{RHom}}
+\newcommand{\algRHomI}{\textrm{RHomI}}
+\newcommand{\algRHomE}{\textrm{RHomE}}
+
+\newcommand{\algUnit}{\textrm{Unit}}
+\newcommand{\algUnitI}{\textrm{UnitI}}
+\newcommand{\algUnitE}{\textrm{UnitE}}
+
+\newcommand{\algTensor}{\textrm{Tensor}}
+\newcommand{\algTensorI}{\textrm{TensorI}}
+\newcommand{\algTensorE}{\textrm{TensorE}}
+
+\newcommand{\algCatTy}{\textrm{Cat}}
+\newcommand{\algCatQt}{\textrm{CatQt}}
+\newcommand{\algCatUnqt}{\textrm{CatUnQt}}
+
+\newcommand{\algTransTy}{\textrm{Trans}}
+\newcommand{\algTransQt}{\textrm{TransQt}}
+\newcommand{\algTransUnqt}{\textrm{TransUnQt}}
+
+\newcommand{\algVarrQt}{\textrm{FunctQt}}
+\newcommand{\algVarrUnqt}{\textrm{FunctUnQt}}
 
@@
 
+%% Can we do pras?
+%%
+%% A PRA mode theory is given by a free 2-category
+%% 1. Each mode m is a CwF whose context cat is [m]
+%% 2. Each 1-cell μ : m -> n is a functor [μ] : [n] -> [m] that has a DRA and is a PRA
+%% 3. Each 2-cell α : μ => ν is a 2-cell
+
+%% Let C, D be CwFs.
+%% A DRA is a pair of
+%% 1. A functor  L : C -> D
+%% 2. A morphism R : L^*(D_Ty) -> C_ty in ∫C, i.e., a morphism taking LΓ |- A to Γ |- RA
+%% 3. An iso     I : L^*(D_tm)?? =~ R^*(C_tm) in ∫(L^*(DTy)) a morphism taking LΓ |- M : A to Γ |- I(M) : RA
+
+%% Let C, D such that C has a terminal object.
+%% G : C -> D is a PRA if
+%% G / 1 : C -> D / G1 is a right adjoint
+
+%% What about fixed points?
+%% Say C has a terminal object
+%% fix : Hom(a,a) -> Hom(1, a)
+%% satisfying 
+%% fix[f] = f o fix[f]?
+%% We can write down fix, but not the equation I think
+
 \section{Semantics}
@@
 
+TODO: include all of the interactions with substitutions and restrictions.
+
 \begin{figure}
@@
     {}
-    {\id_h : \algVarr(C,C)}
-
-    \id_h \circ_{vs} \gamma = \id_h
-
-    (h \circ \id) = h\and \id \circ h = h
+    {\id_v : \algVarr(C,C)}
+
+    \id_v \circ_{vs} \gamma = \id_h
+
+    (f \circ_v \id_v) = f\and \id_v \circ_v f = f
 
@@
 
+    \inferrule
+    {\Psi : \algHCtx(D,E) \and \algHCtx(C,D)}
+    {\Psi \circ_h \Phi : \algHCtx(C,E)}
+
+    (\Psi \circ_h \Phi) \circ_{hs} \gamma = (\Psi \circ_{hs} \gamma) \circ_h (\Phi \circ_s \gamma)
+
+    (\Sigma \circ_h \Psi) \circ_h \Phi = \Sigma \circ_h (\Psi \circ_h \Phi)
+
+    \inferrule
+    {}
+    {\id_h : \algHCtx(C,C)}
+
+    \id_h \circ_{hs} \gamma = \id_h
+
+    (\Phi \circ_h \id_h) = \Phi\and \id_h \circ_h \Phi = \Phi
+
     % embedding of displays
 
-    % strict equipment (horizontal)
-    % strict equipment  (displays)
+    \inferrule{P : \algHarr(C,D)}{i(P) : \algHCtx(C,D)}
 
@@
     % interchange law
-
-    % 2-cells: transformations : Q: is any of this necessary or should
-    % we just conflate transformations with substitutions into a
-    % display?
-    % 
-    % action of trans on subst with assoc, unit
-    % identity trans with unit
-    % embedding of transformations into substitutions
-
+    \inferrule
+    {\psi : \algHSubst(\Phi, f',g',\Psi)\and \phi : \algHSubst(\Sigma,f,g,\Phi)
+    }
+    {\psi \circ_{2v} \phi : \algHSubst(\Sigma, f' \circ f, g' \circ g, \Psi)}
+
+    \inferrule
+    {}
+    {\id_{2v} : \algHSubst(\Phi,\id,\id,\Phi)}
+
+    (\phi \circ_{2v} \psi) \circ_{2v} \sigma = \phi \circ_{2v} (\psi \circ_{2v} \sigma)\\
+    \phi \circ_{2v} \id_{2v} = \phi \and \phi = \id_{2v} \circ_{2v} \phi\\
+
+    \inferrule
+    {\phi_1 : \algHSubst(\Psi_1,f,g,\Phi_1)\and \phi_2 : \algHSubst(\Psi_2,f,g,\Phi_2)
+    }
+    {\phi_1 \circ_{2h} \phi_2 : \algHSubst(\Psi_1 \circ_h \Psi_2, f, h, \Phi_1 \circ_h \Phi_2)}
+
+    \inferrule
+    {}
+    {\id_{2h} : \algHSubst(\id_h,f,f,\id_h)}
+
+    (\phi \circ_{2h} \psi) \circ_{2h} \sigma = \phi \circ_{2h} (\psi \circ_{2h} \sigma)\\
+    \phi \circ_{2h} \id_{2h} = \phi \and \id_{2h} \circ_{2h} \phi = \phi\\
+
+    (\phi_1 \circ_{2h} \phi_2) \circ_{2v} (\psi_1 \circ_{2h} \psi_2) = (\phi_1 \circ_{2v} \psi_1) \circ_{2h} (\phi_2 \circ_{2v} \psi_2)
+
+  \end{mathpar}
+  \caption{Double Category with Display Pro-arrows}
+  \label{fig:alg-dc-disp}
+\end{figure}
+
+\begin{figure}
+  \begin{mathpar}
+    % Restrictions
     \inferrule
@@
 
-    \inferrule
-    {\psi : \algHSubst(f',g',\Phi,\Psi)\and \phi : \algHSubst(f,g,\Sigma,\Phi)
+    % Prof Multiplicatives
+    % LHom
+    \inferrule
+    {P : \algHarr(C, D) \and Q : \algHarr(C,E)}
+    {\algLHom(P,Q) : \algHarr(D,E)}
+
+    \algLHom(P,Q) \circ_{hs} \gamma = \algLHom(P \circ_{hs} \gamma, Q \circ_{hs} \gamma)\and
+    \algLHom(P,Q) \circ_{r} (f,g) = \algLHom(P \circ_r (\id,f), Q \circ_r(\id,g))
+
+    \inferrule
+    {}
+    {\algLHomE(P,Q) : \algHSubst(i(P)\circ_h i(\algLHom(P,Q)),\id_v,\id_v,i(Q))}
+
+    \inferrule
+    {t : \algHSubst(i(P) \circ_h \Phi, \id,g, i(Q))}
+    {\algLHomI(\Phi,g,t,P,Q) : \algHSubst(\Phi,\id,g,i(\algLHom(P,Q)))}
+
+    \algLHomE \circ_{2v} (\id_v(P) \circ_h \algLHomI(t)) = t\and
+    \theta = \algLHomI(\algLHomE \circ_{2v} (\id_v(P) \circ_h \theta))\and
+
+    % RHom
+    \inferrule
+    {P : \algHarr(C, E) \and Q : \algHarr(D,E)}
+    {\algRHom(P,Q) : \algHarr(C, D)}
+
+    \algRHom(P,Q) \circ_{hs} \gamma = \algRHom(P \circ_{hs} \gamma, Q \circ_{hs} \gamma)\and
+    \algRHom(P,Q) \circ_{r} (f,g) = \algRHom(P \circ_r (f, \id), Q \circ_r(g,\id))
+
+    \inferrule
+    {}
+    {\algRHomE(P,Q) : \algHSubst(i(\algRHom(P,Q)) \id_h i(Q),\id_v,\id_v,i(P))}
+
+    \inferrule
+    {t : \algHSubst(\Phi \circ_h i(Q), f, \id, i(P))}
+    {\algRHomI(\Phi,f,t,P,Q) : \algHSubst(\Phi,\id,g,i(\algLHom(P,Q)))}
+
+    \algRHomE \circ_{2v} (\algRHomI(t) \circ_h \id(Q)) = t\and
+    \theta = \algRHomI(\algRHomE \circ_{2v} (\theta \circ_h \id(Q)))\and
+
+    % Unit
+    \inferrule
+    {C : \algCat}
+    {\algUnit(C) : \algHarr(C,C)}
+
+    \algUnit(C) \circ_{hs} \gamma = \algUnit(C \circ_{cs} \gamma)
+
+    \inferrule
+    {}
+    {\algUnitI(C) : \algSubst(\id_h(C),\id,\id,i(\algUnit(C)))}
+
+    \inferrule
+    {t : \algHSubst(\Phi \id_h \Psi,f,g,i(R))}
+    {\algUnitE(t) : \algHSubst((\Phi \circ_h i(\algUnit(C)) \circ_h \Psi), f, g, i(R))}
+
+    \algUnitE(t) \circ_{2v} (\id_{2v}(\Phi) \circ_h \algUnitI(C) \circ_h \id_{2v}(\Psi)) = t\and
+    s = \algUnitE(s \circ_{2v} (\id_{2v}(\Phi) \circ_h \algUnitI(C) \circ_h \id_{2v}(\Psi)))
+
+    % Tensor
+    \inferrule
+    {P : \algHarr(C,D) \and Q : \algHarr(D,E)}
+    {\algTensor(P,Q) : \algHarr(C,E)}
+
+    \algTensor(P,Q) \circ_{hs} \gamma = \algTensor(P \circ_{hs} \gamma, Q \circ_{hs} \gamma)\and
+    \algTensor(P,Q) \circ_{r} (f,g) = \algTensor(P \circ_r (f,\id), Q \circ_r (\id,g))
+
+    \algTensorI(P,Q) : \algHSubst(i(P) \circ_h i(Q),\id,\id,i(\algTensor(P,Q)))
+
+    \inferrule
+    {t : \algHSubst(\Phi \circ_h i(P) \circ_h i(Q) \circ_h \Psi, f,g,i(R))}
+    {\algTensorE(t) : \algHSubst(\Phi \circ_h i(\algTensor(P,Q)) \circ_h \Psi, f,g,i(R))}
+
+    \algTensorE(t) \circ_{2v} (\id_{2v}(\Phi) \circ_{2h} \algTensorI(P,Q) \circ_{2h} \id_{2v}(\Psi)) = t\and
+    s = \algTensorE(s \circ_{2v} (\id_{2v}(\Phi) \circ_{2h} \algTensorI(P,Q) \circ_{2h} \id_{2v}(\Psi)))
+
+    % Cat Multiplicatives
+    % Cat of Elements
+    \inferrule
+    {P : \algHarr(\Gamma,C,D)}
+    {\algElts(C,D,P) : \algCat(\Gamma)}
+
+    \inferrule
+    {}
+    {\pi^- : \algVarr(\algElts(C,D,P),C)}
+
+    \inferrule
+    {}
+    {\pi^+ : \algVarr(\algElts(C,D,P),D)}
+
+    \inferrule
+    {}
+    {\pi^p : \algHSubst(\id_h(\algElts(C,D,P)),\pi^-,\pi^+,P)}
+
+    \inferrule
+    {f : \algVarr(E,C) \and g : \algVarr(E,D)\and
+      t : \algHSubst(\id_h(E), f, g, P)
     }
-    {\psi \circ \phi : \algHSubst(f' \circ f, g' \circ g, \Sigma, \Psi)}
-
-    \inferrule
-    {}
-    {restrict(f,g,\phi) : \algHSubst(\id,\id,\Psi,\Phi)}
+    {\algEltsI(t,f,g) : \algVarr(E,\algElts(C,D,P))}
+
+    \pi^- \circ_v \algEltsI(t,f,g) = f\and
+    \pi^+ \circ_v \algEltsI(t,f,g) = g\and
+    \pi^p \circ_{2v} \algEltsI(t,f,g) = \alpha
+
+    \inferrule
+    {h : \algVarr(E,\algElts(C,D,P))}
+    {h = \algEltsI(\pi^-\circ_v h,\pi^+\circ_vh,\pi^p \circ_{2v} \id_{2h}(h))}
+
+    % Presheaf cats
+
+    % Cat Additives
+    % Product
+    % Terminal
+
+    % Prof Additives
+    % Product
+    % Terminal
   \end{mathpar}
-  \caption{Equipment with Displays}
-  \label{fig:alg-dc-disp}
+  \caption{Multiplicatives}
 \end{figure}
@@
   \begin{mathpar}
-    % Cat Multiplicatives
-    % Cat of Elements
-    % Presheaf cats
-
-    % Prof Multiplicatives
-    % Unit
-    % Tensor
-    % Hom
-
     % Internalization
     % Cats
+    \inferrule
+    {}
+    {\algCatTy : \algTy(1)}
+
+    \inferrule
+    {C : \algCat(\Gamma)}
+    {\algCatQt(C) : \algTm(\Gamma,\algCatTy \circ !)}
+
+    \algCatQt(C) \circ_{tms} \gamma = \algCatQt(C \circ_{cs} \gamma)\and
+
+    \inferrule
+    {M : \algTm(\Gamma,\algCatTy \circ !)}
+    {\algCatUnqt(M) : \algCat(\Gamma)}
+
+    \algCatUnqt(M) \circ_{cs} \gamma = \algCatUnqt(M \circ_{tms} \gamma)\and
+    \algCatUnqt(\algCatQt(C)) = C\and
+    M = \algCatQt(\algCatUnqt(M))
+
     % Functor
@@
     % Trans
-
-    % Cat Additives
-    % Product
-    % Terminal
-
-    % Prof Additives
-    % Product
-    % Terminal
+    \inferrule
+    {P : \algHarr(\Gamma,C,C)}
+    {\algTransTy(P) : \algTy(\Gamma)}
+
+    \inferrule
+    {t : \algHSubst(\id_h,\id_v,\id_v,i(P))}
+    {\algTransQt(t) : \algTm(\Gamma,\algTransTy(P))}
+
+    \algTransQt(t) \circ_{tms} \gamma = \algTransQt(t \circ_{2s} \gamma)
+
+    \inferrule
+    {}
+    {\algTransUnqt : \algHSubst(ext(\Gamma,\algTransTy(P)),\id_h,\id_v,\id_v,i(P \circ_{hs} \pi_1))}
+
+    \algTransUnqt \circ_{2s} (\gamma,\algTransQt(t)) = t\\
+
+    % M : Tm(G,∀ α. P α α)
+    % unquote : Subst(G,t : ∀ α:C. P α α; α:C, id,id, P α α)
+    % M = unquote()
+    M = \algTransQt(\algTransUnqt \circ_{2s} ext_s(\id,M))
   \end{mathpar}
-  \caption{UMPs in an Equipment}
+  \caption{Internalization}
 \end{figure}
+
+\subsection{Equivalence between Natural Deduction and Combinatory Syntaxes}
+
+Now we aim to prove that our natural deduction syntax of SCT is sound
+and complete for the algebraic notion of hyperdoctines of double
+categories with displays.
+%
+In category-theoretic terms, we want to show that the syntax of SCT
+presents an \emph{initial object} in the category of models.
+%
+This is proved in two steps: first, we show that SCT is a model
+
+\begin{construction}{Completeness}
+  SCT presents a hyperdoctrine of double categories with displays.
+\end{construction}
+\begin{proof}
+  First, the sorts.
+  \begin{align*}
+    \eta(\algCtx) &= \{ \Gamma | \Gamma \isadtctx \}\\
+    \eta(\algSubst(\Delta,\Gamma)) &= \{ \gamma | \eta(\Delta) \vdash \gamma : \eta(\Gamma) \}\\
+    \eta(\algTy(\Gamma)) &= \{ A |  \eta(\Gamma) \vdash A \isaTy \}\\
+    \eta(\algTm(\Gamma,A)) &= \{ M | \eta(\Gamma) \vdash M : \eta(A) \}\\
+    \eta(\algCat(\Gamma)) &= \{ \cat C | \eta(\Gamma) \vdash \cat C \isaCat \}\\
+    \eta(\algVarr(\Gamma,C,D)) &= \{ a | \eta(\Gamma)|\alpha:\eta(C) \vdash a : \eta(D) \}\\
+    \eta(\algHarr(\Gamma,C,D)) &= \{ R | \eta(\Gamma)|\alpha:\eta(C);\beta:\eta(D) \vdash R \isaSet \}\\
+    \eta(\algHCtx(\Gamma,C,D)) &= \{ \Phi | \eta(\Gamma) \vdash \Phi \isavectx \wedge d^-(\Phi) = \eta(C) \and d^+(\Phi) = \eta(D) \}\\
+    \eta(\algHSubst(\Gamma, \Psi, f, g, \Phi)) &= \{ \phi | \eta(\gamma)\pipe \eta(\Psi) \vdash \phi : \eta(\Phi) \wedge d^-\phi = \eta(f) \wedge d^+\phi = \eta(g) \}\\
+    1 &= \cdot\\
+    ! &= \cdot\\
+    ext(\Gamma,A) = \Gamma, x:A \\
+    \cdots\\
+    \Phi \circ_h \Psi &= \Phi \jnctx \Psi\\
+    \id_h(\cat C) &= \alpha:\cat C\\
+    i(\cat C, \cat D, R) &= \alpha:\cat C, x: R, \beta:\cat D\\
+    \phi \circ_{2v} \psi &= \phi[\psi]\\
+    \id_{2v}(\Phi) &= \id_\Phi\\
+    \id_{2h}(\alpha:\cat C \vdash a : \cat D) &= a/\beta\\
+    \phi \circ_{2h} \psi &= \phi \jnctx \psi
+  \end{align*}
+\end{proof}
+\begin{corollary}
+  There is a unique homomorphism of HDCDs
+  \[ \eta : Comb \to SCT \]
+\end{corollary}
+
+\begin{construction}{Soundness}
+  We construct a homomorphism of HDCDs
+  \[ i : SCT \to Comb \]
+  such that the composite
+  \[ \eta \circ i : SCT \to SCT \]
+  is equal to the identity.
+\end{construction}
+\begin{proof}
+  \begin{align*}
+    i(\cdot) &= 1\\
+    i(\Gamma,x:A)  &= ext(i(\Gamma),ext(A))\\
+    \cdots
+    i(\Gamma \pipe \alpha:\cat C \vdash \alpha:\cat C) &= \id_v\\
+    i(\Gamma \pipe \alpha:\cat C \vdash M a :\cat C) &= (\algVarrUnqt \circ_{vs} (\id, (i(M)))) \circ_v i(a)\\
+    % Presheaf intro
+    % elements intro
+    i(\Gamma \pipe \alpha:\cat C \vdash (a,t,b) : \Sigma_{\alpha:\cat C;\beta:\cat D} R) &= \algEltsI(i(a),i(t),i(b))\\
+    %
+  \end{align*}
+\end{proof}
+
+\begin{corollary}
+  SCT is initial in the category of HDCDs.
+\end{corollary}
+\begin{proof}
+  Since Comb is initial, the composite $i \circ \eta$ is the identity,
+  and as above the composite $\eta \circ i$ so SCT is isomorphic to
+  Comb, and therefore it is also initial.
+\end{proof}
 

### 2022-04-25 paper/paper.tex
intro tweaks

@@
 
-More experienced users of category theory in programming language
-theory find that for modeling computational effects such as general
-recursion or indexed-state, or modeling logics for parametricity,
-gradual typing, or step-indexed logical relations involves not just
-ordinary category theory, but more general notions such as enriched,
-internal or multi-categories. However, model construction theorems and
-especially soundness and completeness theorems in these areas will
-require their own, even more complex versions of the same common
-lemmas of category theory, in addition to the domain-specific results
-such as existence of fixed points.
+More experienced practitioners of category theory in programming
+language theory find that for modeling computational effects such as
+general recursion or indexed-state, or modeling logics for
+parametricity, gradual typing, or step-indexed logical relations
+involves not just ordinary category theory, but more general notions
+such as enriched, internal or multi-categories. However, model
+construction theorems and especially soundness and completeness
+theorems in these areas will require their own, even more complex
+versions of the same common lemmas of category theory, in addition to
+the domain-specific results such as existence of fixed points.
 
@@
 %
-First, the abstract arguments can be difficult to understand compapred
-to the original argument for ordinary category theory.
+First, the algebraic arguments can be difficult to understand compared
+to the original argument for ordinary category theory, which is an
+obstacle to wider adoption of this powerful approach.
 %
@@
 
-We solve both of these issues by developing an alternative \emph{type
-theoretic syntax} for the constructions in a virtual equipment.
+We address both of these issues by developing an alternative
+\emph{type theoretic syntax} for virtual equipments.
 %
@@
 linear context $\Phi$ and output $S$ and frames $F,G$ are combined in
-the output type.
+the output type using the restriction operation of an equipment.
 %

### 2022-04-25 paper/paper.tex
A straightforward presentation of the models. Eureka!

@@
 
-First, we recount the definition of a virtual double category.
-\begin{definition}[Virtual Double Category]
-  A (vertically strict) virtual equipment $\mathcal V$ consists of
+%% 1. A strict double category is a monoid in Span(Cat)
+%% 2. A strict double category internal to a topos T is a monoid in Span(T-Cat)
+%% 3. A double category with display pro-arrows is a strict double category with a span of profunctors that is representable
+%%    C0 <- C1 -> C0
+%% 4. UMPs are given on top of that structure
+
+We build up the notion of a model in several stages.
+
+\begin{definition}
+  A strict double category is a monoid in the double category $Span(Cat)$.
+\end{definition}
+
+This data consists of a span $C_0 \leftarrow C_1 \rightarrow C_0$
+where $C_0$ is the vertical category and $C_1$ is the category whose
+objects are horizontal morphisms and morphisms are squares, with the
+two projections giving the left and right ``frame'' for the horizontal
+morphisms/squares. The unit gives the horizontal identity morphisms
+and the monoid multiplication is the horizontal composition.
+
+\begin{definition}
+  A double category with displays (DCD) consists of
   \begin{enumerate}
-    \item A set $\mathcal V_0$ of objects.
-    \item For each $\cat C, \cat D \in V_0$, a set $V_v(\cat C, \cat
-      D)$ of vertical arrows, with composition and unit operations
-      satisfying the category axioms.
-    \item A set $V_h$ of horizontal arrows, with $d^+,d^- : V_h \to
-      V_0$.
-
-    \item For any vertical arrows $f \in V_v(\cat C, \cat D)$ and $g
-      \in V_v(\cat E, \cat F)$ and horizontal arrow $R \in V_h$ with
-      $d^+R = \cat F$ and $d^-R = \cat D$, a \emph{restriction}
-      $R(f,g)$ with $d^+R = \cat C$ and $d^-R = \cat E$.
-
-      This operation is furthermore associative and unital $R(f \circ
-      f',g \circ g') = R(f,g)(f',g')$ and $R(\id,\id) = R$.
-
-    \item For each composable string $\Phi$ (defined below) of horizontal arrows
-      starting at $\cat C$ and ending at $\cat D$, and horizontal
-      arrow $R$ with $d^-R = \cat C$ and $d^+R = \cat D$, a set
-      $V_2(\Phi; R)$ of 2-cells.
-
-    \item For any string of vertical arrows $\vec f$ into $\Phi$ and
-      2-cell $r \in \mathcal V_2(\Phi,S)$, a restriction operation
-      $r(\vec f) \in \mathcal V_2(\Phi(\vec f), S(d^-(\vec f),
-      d^+(\vec f)))$. Furthermore, this is associative and unital
-      $r(\vec id) = r$ and $r(\vec f \circ \vec g) = r(\vec {f \circ
-        g})$
-
-    \item For each $R \in V_h$, an identity arrow $\id_R \in
-      V_2(d^-R,R,d^+ R; R)$ with $d^\pm \id_R = \id_{d^\pm R}$.
-
-    \item For any $2$-cell $\alpha \in V_2(\Phi;R)$ and composable
-      string $\phi$ of 2-cells into $\Phi$, a composite $\alpha \circ
-      \phi$.
-
-    \item Composition should be unital in that:
-      \[ \id_R \circ [d^-\alpha, \alpha, d^+\alpha] = \alpha  \]
-      and
-      \[ \alpha \circ [\vec \id] = \alpha\]
-    \item And associative in that
-      \[ (\alpha \circ [\beta_1,\ldots]) \circ [\gamma_{11},\ldots,\gamma_{21},\ldots,] = \alpha \circ [(\beta_1 \circ \gamma_{11},\ldots), (\beta_2 \circ \gamma_{21},\ldots)]\]
+  \item A strict double category $C$
+  \item A category $C_d$ of (horizontal) ``display morphisms'' and ``display squares''.
+  \item A span of profunctors:
+    % https://q.uiver.app/?q=WzAsOCxbMiwyLCJDXzEiXSxbMCwyLCJDXzAiXSxbNCwyLCJDXzAiXSxbMiwwLCJDX2QiXSxbMCwwLCJDXzAiXSxbNCwwLCJDXzAiXSxbMSwxLCJcXExlZnRhcnJvdyJdLFszLDEsIlxcUmlnaHRhcnJvdyJdLFszLDAsIkNfXFx2ZGFzaCIsMix7InN0eWxlIjp7ImJvZHkiOnsibmFtZSI6ImJhcnJlZCJ9fX1dLFs1LDIsIlxcdGV4dHJte0hvbX0iLDAseyJzdHlsZSI6eyJib2R5Ijp7Im5hbWUiOiJiYXJyZWQifX19XSxbNCwxLCJcXHRleHRybXtIb219IiwyLHsic3R5bGUiOnsiYm9keSI6eyJuYW1lIjoiYmFycmVkIn19fV0sWzMsNCwicyIsMl0sWzMsNSwidCJdLFswLDIsInQiLDJdLFswLDEsInMiXV0=
+\[\begin{tikzcd}
+	{C_0} && {C_d} && {C_0} \\
+	& \Leftarrow && \Rightarrow \\
+	{C_0} && {C_1} && {C_0}
+	\arrow["{C_\vdash}"', "\shortmid"{marking}, from=1-3, to=3-3]
+	\arrow["{\textrm{Hom}}", "\shortmid"{marking}, from=1-5, to=3-5]
+	\arrow["{\textrm{Hom}}"', "\shortmid"{marking}, from=1-1, to=3-1]
+	\arrow["s"', from=1-3, to=1-1]
+	\arrow["t", from=1-3, to=1-5]
+	\arrow["t"', from=3-3, to=3-5]
+	\arrow["s", from=3-3, to=3-1]
+\end{tikzcd}\]
+\item A fully faithful functor $i : C_d \to C_1$ that respects source and target in that the following commute:
+% https://q.uiver.app/?q=WzAsNixbMiwyLCJDXzEiXSxbMCwyLCJDXzAiXSxbNCwyLCJDXzAiXSxbMiwwLCJDX2QiXSxbMCwwLCJDXzAiXSxbNCwwLCJDXzAiXSxbMywwLCJpIiwyXSxbNSwyLCJcXGlkIl0sWzQsMSwiXFxpZCIsMl0sWzMsNCwicyIsMl0sWzMsNSwidCJdLFswLDIsInQiLDJdLFswLDEsInMiXV0=
+\[\begin{tikzcd}
+	{C_0} && {C_d} && {C_0} \\
+	\\
+	{C_0} && {C_1} && {C_0}
+	\arrow["i"', from=1-3, to=3-3]
+	\arrow["\id", from=1-5, to=3-5]
+	\arrow["\id"', from=1-1, to=3-1]
+	\arrow["s"', from=1-3, to=1-1]
+	\arrow["t", from=1-3, to=1-5]
+	\arrow["t"', from=3-3, to=3-5]
+	\arrow["s", from=3-3, to=3-1]
+\end{tikzcd}\]
+and $i$ represents $C_\vdash$ in that there is a natural isomorphism:
+    \[ e : C_\vdash(\Phi, R) \cong C_1(\Phi, i(R)) \]
+    that commutes with the natural transformations $s,t$.
   \end{enumerate}
-
-  A composable string $\Phi$ of horizontal arrows is defined to be an
-  alternating sequence of objects and horizontal arrows, beginning and
-  ending with an object such that any substring $\cat C,R,\cat D$
-  satisfies $d^-R = \cat C$ and $d^+R = \cat D$.
-
-  A composable string of vertical arrows $\vec f$ into $\Phi$ is a
-  sequence of vertical arrows, one for each object in $\Phi$, with the
-  codomain of the arrow matching the object in $\Phi$. Then
-
-  A composable string of 2-cells from $\Phi$ to $\Psi$ is a sequence
-  $\vec r$ of 2-cells, one for each horizontal arrow in $\Psi$ such
-  that the codomains match the objects in
 \end{definition}
 
-Our main departure from\cite{crutwellshulman} is making restrictions
-part of the definition, and also requiring them to be strictly
-associative and unital. In the examples that we are most concerned
-with, however, these operations are strict and so we will not explore
-the possibility of coherence theorems for proving this equivalent to a
-weaker definition.
+In terms of the syntax of SCT, the category $C_0$ models the category
+and functor judgments, the category $C_1$ models the profunctor
+\emph{contexts} and substitutions and the category $C_d$ models the
+profunctor judgment. The profunctor $C_\vdash$ models the
+transformation judgment and the functor $i$ models the construction of
+a ``singleton'' context $\alpha:\cat C,x:R,\beta:\cat D$.
+
+A DCD is a more general notion that a VDC, but they are closely
+related, in that VDCs form a coreflective subcategory of DCDs. This is
+analogous to the situation in models of dependent type theory, where
+DCDs are analogous to Categories with Families \cite{cwfs,natmodels}
+and VDCs are analogous to the more ``syntactic'' notion of a
+Contextual Category/C-system \cite{ctxcat,csystem}. First, any DCD
+induces a VDC by taking the display proarrows as the proarrows and
+interpreting 2-cells using $i$ and $C_\vdash$.  Second, any VDC
+presents a DCD whose double category has the original VDC pro-arrows
+as the display morphisms and \emph{composable strings} of proarrows as
+horizontal morphisms, which gives a strict double category.
+%
+Taking as morphisms the strict morphisms of DCDs/VDCs, these
+constructions form a coreflective embedding of VDCs into DCDs.
+
+The principle advantage of DCDs over VDCs is that the presentation of
+their algebraic structure is considerably simplified and is
+straightforward to describe as a Generalized Algebraic Theory in the
+sense of \citet{CartmellGAT}. This is because we don't \emph{require}
+that our profunctor contexts are lists, but instead merely the fact
+that their concatenation forms a strict category. Then the
+associativity/identity equations for a DCD are simply those of a
+double category.
+
+Next, each of the connectives of SCT are modeled by different
+universal properties that are profunctor isomorphisms involving
+principally $C_\vdash$.
+TODO: write them down.
 
 \begin{definition}
-  A functor of virtual double categories $F : \mathcal V \to \mathcal
-  U$ is a mapping on objects, vertical arrows, horizontal arrows and
-  2-cells respecting typing and preserving composition, units, and
-  restriction.
+  An \emph{equipment with displays} (ED) is a DCD with all of the
+  universal properties in Figure\ref{eq-umps}.
 \end{definition}
+
+Finally, to model the logic over SCT provided by the dependently typed
+context $\Gamma$, we use a category with families $Ctx$, and interpret
+all of the prior constructions in this section to be internal to
+presheaves on $Ctx$. This models the syntactic aspect that all double
+category structure is parameterized by a context $\Gamma$ and stable
+under substitution. Since the category of presheaves is a topos, all
+definitions given in set-theoretic language can be interpreted
+internally.
+
+\begin{definition}
+  A Category with Families is...
+\end{definition}
+
+\begin{definition}
+  A hyperdoctrine of equipment with displays (HED) is a pair of a
+  category with families $Ctx$ modeling $\Sigma,\Pi,\textrm{Id}$ and a
+  DCD internal to $\Set^{Ctx^o}$ with categories, functors, displays
+  and unary cells representable as types.
+\end{definition}
+
+Again this notion is straightforwardly, if laboriously, interpretable
+as a GAT.
 

### 2022-04-26 paper/paper.tex
Intro/abstract rewrite

@@
 \usepackage{cleveref}
+\usepackage{natbib}
 
@@
   theoretic proofs about the relationship between unit, tensor and
-  (ordered) function types.
-
-  We show that the type theory is sound for models in \emph{virtual
-  equipments}, a kind of double category which is known to model not
-  just ordinary categories but enriched, internal and generalized
-  multi-categories as well. This means that while proofs and
-  constructions in our theory read like typical category-theoretic
-  constructions, the proofs automatically apply to these generalized
-  categories. Our type theory simplifies arguments over the algebraic
-  description of virtual equipments by modeling the algebraic
-  operation of restriction of profunctors along functors as instead
-  the strict syntactic substitution.
+  (ordered) function types. We demonstrate that this type theory can
+  be used as a kind of logical framework in which we can define common
+  categorical structures such as categories with families that serve
+  as models of type theories.
+
+  Finally, we give a sound and complete notion of model of our type
+  theory in a structure we call a hyperdoctrine of pro-arrow
+  equipments with displays, a generalization of prior notions of
+  virtual equipments that have been shown to model enriched and
+  internal category theory. This means that while the proofs in our
+  type theory look like standard set-based arguments, the
+  substructural syntactic restrictions ensure that the proofs carry
+  over to enriched and internal settings as well.
 \end{abstract}
@@
 
-Category theory is a branch of mathematics that formalizes the
-algebraic structure of function composition. The high level of
-abstraction allows for succinct proofs relevant to structures across a
-broad variety of mathematical domains. Category theory is of
-particular interest in the theory of programming languages and proof
-assistants, where it is commonly used to aid in the design of typed
-languages, proof assistants and functional programming abstractions.
-
+Category theory is a branch of mathematics that gives a systematic
+theory of typed, ``higher-dimensional'' algebraic structures.
+%
+Originally developed for applications to homological algebra, it was
+quickly discovered that these algebraic structures were common in
+logic and computer science, and that formal systems like type theories
+and programming languages can be modeled by a sound and complete
+notion of semantic model given by some notion of structured
+category\cite{lawvere,goguen,lambek,moggi}.
+%
+Examples include the simply typed lambda calculus \cite{lambekscott},
+computational lambda calculus \cite{moggi}, linear logic \cite{girard}
+and dependent type theory\cite{not-sure}.
+%
+These seminal works have led to a fruitful field of research applying
+such categorical logic techniques to the design of new type theories
+based directly on category-theoretic semantics.
+%
+The slogan is that the syntax of a type theory should present an
+initial object in its category of models, a category-theoretic
+reformulation and generalization of the logical notion of soundness
+and completeness.
+
+% TODO: do we have citations for this?
+While this research program has been quite successful,
+category-theoretic techniques in computer science are sometimes
+criticized for being overly complex, and there is difficulty in
+acquisition of competence in category theory by computer scientists
+who do not have a strong background in ``pure'' mathemantics.
+%
+While much of this can be explained by lack of, or lack of awareness
+of, accessible material for computer scientists, in some ways category
+theory is objectively complex.
+%
+In a traditional set-theoretic formulation, notions such as adjoint
+functors and Kan extensions produce a proliferation of ``naturality''
+and ``functoriality'' side-conditions that must be discharged.
+%
+This is compounded when we move from ordinary category theory to
+enriched and internal category theory, where constructions must be
+additionally proven continuous, monotone, etc, in addition to natural
+or functorial.
+%
+These generalizations have long been recognized as being important in
+programming language applications: domain-,metric- and
+step-index-enriched categories have been used to model general
+recursion and internal categories have been used to model
+parametricity and gradual typing.
+
+% TODO: Yoneda example
 Novices in category theory are often overwhelmed by the number of
@@
 holds.
+
+Fortunately, the tools of category theory itself can be employed to
+simplify these constructions, specifically the tools of \emph{higher}
+category theory.
+%
+Consider the analogous situation in differential calculus. When a
+physicist writes down a function, they do not expand out the
+$\epsilon-\delta$ definition of continuity and prove that it is
+continuous. Rather, they are aware of certain \emph{syntactic
+principles}, most importantly that composition of continuous functions
+is continuous and that previously defined functions are known to be
+continuous.
+%
+Similar principles apply to category theory itself: a composition of
+functors is functorial, a composition of natural transformations is
+natual, etc.
+%
+In the case of category theory, these principles can be formalized
+using category-like algebraic structures such as 2-categories,
+bicategories, (virtual) double categories and pro-arrow
+equipments\cite{}.
+%
+This approach has the advantage of avoiding the need to prove
+the side conditions of naturality or enrichment.
+%
+The downside is that these algebraic structures are quite complicated,
+with a 2-dimensional diagrammatic language that can be quite beautiful
+and elegant, but is also seems very far removed from the traditional
+formulation of category theory that while complex is at least
+described in terms of a concrete notion of sets or types.
+
+In this work we show that these algebraic formulations of category
+theory can be given a type-theoretic syntax that allows us to make
+``na\"ive'' set theoretic constructions for categories, but by
+adhering to a certain syntactic discipline provide correct arguments
+for not just ordinary category theory but enriched and internal
+categories as well.
+%
+The type theory, which we call SCT, is a ``synthetic'' theory of
+categories, functors, profunctors and natural transformations.
+%
+This provides a formulation of category theory in which profunctors
+play a primary role, most recognizable in ordinary category theory in
+the prominence of ``hom isomorphism'' formulations of universal
+properties.
+%
+The semantics of SCT is given by a slight generalization of virtual
+double categories \cite{} that we call double categories with
+displays.
+%
+Additionally, to define signatures and parameterize theorems like the
+Yoneda lemma, SCT includes an ordinary dependent type theory over it
+which includes usual $\Pi,\Sigma,\Id$ types.
+%
+In an implementation of SCT, this dependent type theory might be the
+host language itself.
+
+In this work, we focus on the syntax and semantics of SCT, but the
+ultimate goal is for this theory to serve as the foundation for a
+framework in which we can more easily study and program with new type
+theories. This falls in the tradition of \emph{logical frameworks}
+such as the LF system. To provide evidence for the feasibility of this
+approach, we show that we can model structures used in recent work on
+modal dependent type theory. This shows that an implementation of SCT
+would provide at least a toy prototype implementation/type checker for
+modal dependent type theory. This can be summarized in the slogan
+\begin{quote}
+  category theory as a logical framework
+\end{quote}
+that is, that a syntax for category theory can be used as a logical
+framework for type theories with semantics in structured categories.
+%
+If implemented in a proof assistant this would also provide with it
+the soundness of the syntax in any model.
+%
+The emphasis on profunctors in SCT fits nicely with the
+\emph{judgmental} approach of Per-Martin L\"of and suggests the
+following slogan
+\begin{quote}
+  profunctors as judgments, natural transformations as derivations
+\end{quote}
+because we observe that the ``derivation tree'' style of writing
+judgments corresponds closely to the type structure of profunctors and
+natural transformations in SCT, and it should be simple to implement a
+program that converts formal natural transformations into derivation
+tree syntax.
+
+\subsection{Category theory with Profunctors}
+
+\subsection{Double Categories and Pro-arrow Equipments}
 

### 2022-04-27 paper/paper.tex
some more intro

@@
 
+This work is merely a first step in this direction, as there are still
+three major components of category theory: opposite categories,
+functor categories and ``universes'' such as the category of sets or
+categories that are not nicely axiomatizable in the current system.
+%
+Nontheless, a great deal of category theory can be carried out in this
+more limited setting, and we can still take advantage of duality at a
+meta level.
+
 \subsection{Category theory with Profunctors}
 
+First, we review the 4 basic notions of our theory of categories in
+the traditional formulation. A category $\cat C$ is a set of objects
+$\cat C_0$ with a span of arrows $\cat C_0 \leftarrow \cat C_1
+\rightarrow \cat C_0$ that pick out the source and target of each
+arrow, such that there are composition and identity arrows with
+appropriate laws. A functor $F : \cat C \to \cat D$ consists of a
+function on objects $F_0 : \cat C_0 \to \cat D_0$ and a function on
+morphisms $F_1 : \cat C_1 \to \cat D_1$ that is compatible with source
+and target and preserves identity and composition.
+
+A profunctor $R$ from $\cat C$ to $\cat D$ is a span $\cat C_0
+\leftarrow R \rightarrow \cat D_0$ that is compatible with
+\emph{left}-composition by $\cat C$ and \emph{right}-composition by
+$\cat D$. If we think of categories as generalizations of (ordered)
+sets then functors generalize (monotone) functions and profunctors
+generalize (monotone) \emph{relations}. We think of an element $r \in
+R$ with $s(r) = c$ and $t(r) = d$ as a ``proof'' that $c R d$ or some
+kind of ``heteromorphism'' from $c$ to $d$. The compatibility with
+composition says that these relations are down-closed in $C$ and
+up-closed in $D$, or that heteromorphisms can be composed with
+homomorphisms.
+%
+
 \subsection{Double Categories and Pro-arrow Equipments}
 
-More experienced practitioners of category theory in programming
-language theory find that for modeling computational effects such as
-general recursion or indexed-state, or modeling logics for
-parametricity, gradual typing, or step-indexed logical relations
-involves not just ordinary category theory, but more general notions
-such as enriched, internal or multi-categories. However, model
-construction theorems and especially soundness and completeness
-theorems in these areas will require their own, even more complex
-versions of the same common lemmas of category theory, in addition to
-the domain-specific results such as existence of fixed points.
-
-At this level of generality, the proofs can be difficult to
-understand, as they often involve some combination of the original
-category theoretic proof with some additional side-conditions that
-everythinng is in fact appropriately step-indexed, continuous etc.
-%
-One approach to taming this complexity is to prove these general
-theorems at a higher level of generality, for instance, proving them
-in some abstract 2-category of enriched/internal categories, functors
-and transformations.
-%
-For proofs in generalized category theory specifically, the notion of
-a \emph{virtual equipment} has been shown to be occupy an algebraic
-sweet-spot of being complex enough to express most basic constructions
-(functors, profunctors, natural transformations, adjunctions, kan
-extensions, etc) but also general enough to encompass internal,
-enriched and ``generalized multi-categories'', which include ordinary
-flavors of multi-categories that model intuitionistic and
-substructural type theories, but also a great deal more, including
-double categories, metric spaces and topological spaces.
-%
-However, there are two difficulties with this approach.
-%
-First, the algebraic arguments can be difficult to understand compared
-to the original argument for ordinary category theory, which is an
-obstacle to wider adoption of this powerful approach.
-%
-Second, the algebraic structure of a virtual equipment is quite
-unwieldy to manipulate. The abstraction of a generalized natural
-transformation is a kind of ``multi-2-cell'' with many inputs and a
-single output. In diagrammatic form these are rendered as
+We can abstract the algebraic structure of categories, functors,
+profunctors, and natural transformations as a \emph{double category}.
+%
+A double category is a generalization of categories where we have a
+collection of objects $C_0$, and \emph{two separate} notions of
+morphism, which we call vertical $C_v$ and horizontal $C_h$ over the
+same objects $C_0$ which each have their own identity and composition,
+and finally a notion of 2-cell that is most easily described
 
@@
 
-\subsection{Category Theory with Profunctors}
-
-The syntax of Judgment Theory gives not just a syntax for the familiar
-trio of category, functor and natural transformation, but additionally
-a concept that is not usually highlighted in introductory category
-theory texts: the profunctor. A simple intuition is that profunctors
-are to functors as relations are to functions. Indeed, the syntax for
-judgment theory has a straightforward semantics in the category of
-sets and relations.
-
-The reason profunctors are not commonly introduced is that they are
-seen as an auxiliary notion: a profunctor $R : C \pto D$ is
-typically defined as a functor $C^{\textrm{op}}\times D \to
-\textrm{Set}$. However this definition requires two concepts that are
-somewhat problematic: the opposite category $C^\textrm{op}$ and the
-category of (small) sets $\textrm{Set}$. However, these notions do not
-necessarily exist in all of our intended applications. First, we are
-interested in semantics where are objects are \emph{multicategories}
-where morphisms have multiple inputs but only a single output. In this
-case the opposite of a multicategory is not a multicategory at all but
-instead some kind of ``co-multicategory'' with multiple outputs but
-one input. Second, having a universe object like $\textrm{Set}$ is also
-
-\subsection{Applications to Enriched, Internal and Multicategory Theory}
-
-The syntax of judgment theory is an internal language for
-\emph{virtual equipments}. These are a kind of joint generalization of
-double categories and multicategories.
-
-Notably, the definition of a category in \cref{sec:internal} can be
-applied as an extremely general notion of category, including internal
-categories and enriched categories. This general notion is often
-called a ``generalized multi-category'' and encompasses an astonishing
-range of mathematical structures, including categories, topological
-spaces, metric spaces, etc.
-
-\subsection{Applications to Generalized Type Theory}
-
-While we mostly focus on the semantics in category theory, due to the
-tight connection between category theory and the semantics of type
-theory, there are direct applications to the study of ``abstract type
-theory'' as well. We demonstrate how we can model single-variable
-simple type theories with products and coproducts within SCT. This is
-very much in the spirit of logical frameworks \cite{TODO}, since we
-can faithfully model the terms of an object theory inside of our
-metatheory. Since we have models of our metatheory in virtual double
-categories, we establish sound semantics as well. Put succinctly,
-category theory itself is the logical framework for the study of type
-theories.
+%% \subsection{Category Theory with Profunctors}
+
+%% The syntax of Judgment Theory gives not just a syntax for the familiar
+%% trio of category, functor and natural transformation, but additionally
+%% a concept that is not usually highlighted in introductory category
+%% theory texts: the profunctor. A simple intuition is that profunctors
+%% are to functors as relations are to functions. Indeed, the syntax for
+%% judgment theory has a straightforward semantics in the category of
+%% sets and relations.
+
+%% The reason profunctors are not commonly introduced is that they are
+%% seen as an auxiliary notion: a profunctor $R : C \pto D$ is
+%% typically defined as a functor $C^{\textrm{op}}\times D \to
+%% \textrm{Set}$. However this definition requires two concepts that are
+%% somewhat problematic: the opposite category $C^\textrm{op}$ and the
+%% category of (small) sets $\textrm{Set}$. However, these notions do not
+%% necessarily exist in all of our intended applications. First, we are
+%% interested in semantics where are objects are \emph{multicategories}
+%% where morphisms have multiple inputs but only a single output. In this
+%% case the opposite of a multicategory is not a multicategory at all but
+%% instead some kind of ``co-multicategory'' with multiple outputs but
+%% one input. Second, having a universe object like $\textrm{Set}$ is also
+
+%% \subsection{Applications to Enriched, Internal and Multicategory Theory}
+
+%% The syntax of judgment theory is an internal language for
+%% \emph{virtual equipments}. These are a kind of joint generalization of
+%% double categories and multicategories.
+
+%% Notably, the definition of a category in \cref{sec:internal} can be
+%% applied as an extremely general notion of category, including internal
+%% categories and enriched categories. This general notion is often
+%% called a ``generalized multi-category'' and encompasses an astonishing
+%% range of mathematical structures, including categories, topological
+%% spaces, metric spaces, etc.
+
+%% \subsection{Applications to Generalized Type Theory}
+
+%% While we mostly focus on the semantics in category theory, due to the
+%% tight connection between category theory and the semantics of type
+%% theory, there are direct applications to the study of ``abstract type
+%% theory'' as well. We demonstrate how we can model single-variable
+%% simple type theories with products and coproducts within SCT. This is
+%% very much in the spirit of logical frameworks \cite{TODO}, since we
+%% can faithfully model the terms of an object theory inside of our
+%% metatheory. Since we have models of our metatheory in virtual double
+%% categories, we establish sound semantics as well. Put succinctly,
+%% category theory itself is the logical framework for the study of type
+%% theories.
 
@@
 
-To demonstrate the utility of SCT, we proceed as follows
 \begin{enumerate}
@@
   Theory.
-\item In \cref{sec:synthetic} we show how common theorems of category
-  theory such as the (co)-Yoneda lemma, equivalence of different
-  formulations of adjunctions, and preservation of limits can all be
-  cleanly reproduced in SCT.
-\item In \cref{sec:internal}, we show how we can use SCT to define
-  \emph{internal categories}, which allows us to prove additional
-  theorems that are not formalizable as synthetic theorems.
+
+\item In \cref{sec:examples} we show how to use our syntax for
+  synthetic category theory. Much of this is just rendering well-known
+  2-categorical/double categorical constructions in our type theoretic
+  syntax to demonstrate the ``naive''/synthetic character of proofs
+  and constructions in SCT. We additionally give some examples of how
+  this can be applied to model a recent formulation of modal dependent
+  type theory.
+
+%% \item In \cref{sec:synthetic} we show how common theorems of category
+%%   theory such as the (co)-Yoneda lemma, equivalence of different
+%%   formulations of adjunctions, and preservation of limits can all be
+%%   cleanly reproduced in SCT.
+%% \item In \cref{sec:internal}, we show how we can use SCT to define
+%%   \emph{internal categories}, which allows us to prove additional
+%%   theorems that are not formalizable as synthetic theorems.
+
 \item In \cref{sec:semantics}, we show that the syntax of SCT is sound
@@
   corresponding universal constructions.
+
 \item In \cref{sec:discussion}, we discuss related type theories and
@@
 \end{enumerate}
+
+We have aimed that the syntax and examples sections should be
+accessible to anyone familiar with basic category theory and
+willingness to understand profunctors and (co-)ends.
+%
+The semantics section is by necessity more advanced and may be skipped
+or skimmed by readers who are willing to trust the mathematical
+foundations of the system.
 
@@
 
-Next, we develop the basic model theory for SCT. Because SCT includes
-not just a syntax for category theory, but is also indexed by an
-ordinary context of dependent type theory $\Gamma$, a model of SCT is
-not given by a \emph{single} virtual double category, but instead by
-an \emph{indexed family} of VDCs, parameterized by an object of a
+Next, we develop the basics of the model theory for SCT. Because SCT
+includes not just a syntax for category theory, but is also indexed by
+an ordinary context of dependent type theory $\Gamma$, a model of SCT
+is not given by a \emph{single} virtual double category, but instead
+by an \emph{indexed family} of VDCs, parameterized by an object of a
 category that interprets dependent type theory. We call such a
@@
 
+\paragraph{Calculi for Presheaves/Profunctors}
+
+David Jaz Myers String Diagrams
+
+Coend calculus
+
+Cartesian Bicategories.
+
+Double-cat-like syntax: Gradual Type Theory, System P of parametricity, cubical stuff?
+
 \paragraph{Directed Type Theory}
 
-Cartesian Bicategories.
-
 Directed Type theory
@@
 Fibrational approaches to logical relations
-
-Double-cat-like syntax: Gradual Type Theory, System P of parametricity, cubical stuff?
-
-String Diagrams?
 
@@
 An orthogonal extension to these would be to consider adding a
-\emph{discretization} modality $|\cat C|$ that gives the discrete
+\emph{discretization} modality $\Delta\cat C$ that gives the discrete
 category on the underlying set of objects of $\cat C$. This would give
@@
 
-\paragraph{As a tool for Formal Type Theory}
-
-While we can model ``unary'' type theories using profunctors, to model
-theories like simple intuitionistic, linear or bunched type theories
-we need a way to talk about ``profunctors of arbitrarily many
-arguments''. This can be accomplished using \emph{multicategory}
-theory, which models the structure of the context with its structural
-rules as a monad of ``contexts and substitutions''. A suitable notion
-of monad on a virtual double category has been developed and was in
-fact the original motivation for the concept of a virtual double
-category. However while an abstract monad of contexts $T$ is enough to
-define what a $T$-multicategory internal to a virtual equipment is, it
-does not appear possible to define the multiplicative connectives as a
-universal property, i.e., to define when the multicategory is
-\emph{representable}.
-
-A recent development in the formalization of substructural type
-theories is to use monoidal connectives at the meta-level when
-defining multiplicative connectives\cite{woodsatkeyesop2022}. This
-suggest one approach to extending the theory to support the
-formalization of multiplicative universal properties would be, in
-adherence to the microcosm principle, to add linear judgments and
-multiplicative connectives to the metalevel.
+A third approach would be to extend the calculus to model general
+monoidal closed structures and more generally multicategories, which
+provides semantics of modal substructural type theories.
+%
+This can be modeled by extending the ``universe'' $\Set$ to be a
+monoidal closed category, which induces a monoidal closed structure on
+profunctors by the Day convolution product \cite{...}.
+%
+Similar developments along these lines have already been done for
+coend calculus \citep{coend-calculus} and in formalization of
+substructural type theories \citep{atkey-woods}.
+
+%% A recent development in the formalization of substructural type
+%% theories is to use monoidal connectives at the meta-level when
+%% defining multiplicative connectives\cite{woodsatkeyesop2022}. This
+%% suggest one approach to extending the theory to support the
+%% formalization of multiplicative universal properties would be, in
+%% adherence to the microcosm principle, to add linear judgments and
+%% multiplicative connectives to the metalevel.
+
+%% \paragraph{As a tool for Formal Type Theory}
+
+%% While we can model ``unary'' type theories using profunctors, to model
+%% theories like simple intuitionistic, linear or bunched type theories
+%% we need a way to talk about ``profunctors of many arguments''. This
+%% can be accomplished using \emph{multicategory} theory, which models
+%% the structure of the context with its structural rules as a monad of
+%% ``contexts and substitutions''. A suitable notion of monad on a
+%% virtual double category has been developed and was in fact the
+%% original motivation for the concept of a virtual double
+%% category. However while an abstract monad of contexts $T$ is enough to
+%% define what a $T$-multicategory internal to a virtual equipment is, it
+%% does not appear possible to define the multiplicative connectives as a
+%% universal property, i.e., to define when the multicategory is
+%% \emph{representable}.
+
+\paragraph{Semantics}
+
+Our definition of a double category with displays is suggestive of a
+simple generalization. A double category is a monoid object in the
+double category of $Span(Cat)$, and a double category with displays
+furthermore uses certain diagrams involving profunctors, which also
+form a double category $Prof$. This suggests that the algebraic
+structure of a double category with displays is naturally modeled in
+some kind of \emph{triple category}, modeling the three types of
+morphisms: functors, profunctors and spans of functors.
 

### 2022-04-28 paper/paper.tex
The interesting part of the completeness proof

@@
 
-    \inferrule{\Gamma \vdash \cat C : \isaCat}{\Gamma \vdash \alpha: \cat C \boundary}
-
-    \inferrule{\Gamma \vdash \cat C : \isaCat\and
-      \Gamma\vdash \cat D : \isaCat}{\Gamma \vdash \alpha: \cat C;\beta:\cat D \boundary}\\
-
-    \inferrule*
-        {\Gamma \pipe \alpha':\cat C' \vdash A : \cat C}
-        {\Gamma \vdash a/\alpha : \alpha':\cat C' \to \alpha:\cat C}
-
-    \inferrule*
-    {\Gamma \pipe d^-\Xi \vdash a : \cat C\and
-     \Gamma \pipe d^+\Xi \vdash b : \cat D
-    }
-    {\Gamma \vdash (a/\alpha;b/\beta) : \Xi \to (\alpha:\cat C;\beta:\cat D)}\\
-
     \inferrule
@@
     {\Gamma \vdash \Phi \isavectx\and
-    \Gamma \vdash \Psi \isavectx \and
-    \Gamma \pipe d^+\Phi; d^- \Psi \vdash R \isaSet}
-    {\Gamma \vdash \Phi, x:R, \Psi \isavectx}
+    \Gamma \pipe d^+\Phi; \beta:\cat D \vdash R \isaSet}
+    {\Gamma \vdash \Phi, x:R, \beta:\cat D \isavectx}
 
@@
 
-    \inferrule
-    {\Psi = \Psi_1 \jnctx \Psi_M \jnctx \Psi_2\and
-      \Gamma \vdash \phi_1 : \Psi_1 \to \Phi_1\and
-      \Gamma \pipe \Psi_s \vdash s : R[d^+\phi_1;d^-\phi_2] \and
+    % horizontal identity
+    \inferrule
+    {\Gamma\pipe \beta:\cat C \vdash a : \cat C}
+    {\Gamma \pipe \beta:\cat C \vdash a/\alpha : (\alpha : \cat C)}
+
+    \inferrule
+    {\Psi = \Psi_1 \jnctx \Psi_2\and
+      \Gamma\pipe\Psi_1 \vdash \phi : \Phi\and
+      \Gamma \pipe \Psi_2 \vdash t : R[d^+\phi;b/\beta] \and
       \Gamma \vdash \phi_2 : \Psi_2 \to \Phi_2
       }
-    {\Gamma \vdash \phi_1,t/x,\phi_2 : \Psi \to \Phi_1,x:R,\Phi_2}
+    {\Gamma\pipe\Psi \vdash \phi,t/x,b/\beta :: (\Phi,x:R,\beta:\cat D)}
   \end{mathpar}
@@
 
-%% \appendix
-
-%% TODO: complete substitution
+\appendix
+
+\section{Completeness}
+
 %% \begin{figure}
@@
 %%   \caption{Substitution}
-
-%% \end{figure}
-
-%% \begin{figure}
-%%   \begin{mathpar}
-%%     \inferrule
-%%     {}
-%%     {\Gamma\pipe \cdot \vdash \Sort \isaSet}
-
-%%     \inferrule
-%%     {x:S \in \Gamma}
-%%     {\Gamma \pipe \cdot \vdash x : S}
-
-%%     \inferrule
-%%     {\Gamma\pipe\cdot \vdash C : \Sort}
-%%     {\Gamma \pipe \alpha:\cat C \vdash \alpha : \cat C}
-
-%%     \inferrule
-%%     {\Gamma \pipe \alpha:C,\beta:D \vdash R \isaSet}
-%%     {\Gamma\pipe \alpha:C, x:R, \beta: D \vdash x : R}
-
-%%     \inferrule{\forall x:A \in \Gamma. \Delta \vdash \gamma(x) : A}{\gamma : \Delta \to \Gamma}\\
-
-%%     \inferrule*%% [right=NullaryBoundaryMap]
-%%                {\Gamma \isadtctx}{\Gamma \vdash \cdot : \cdot \to \cdot}\and
-%%     \inferrule*{\Gamma \pipe \alpha':C' \vdash A : C}{\Gamma \vdash A/\alpha : \alpha':C' \to \alpha:C}
-
-%%     \inferrule*
-%%     {\Gamma \pipe \alpha':C' \vdash A : C\and
-%%       \Gamma \pipe \beta':D' \vdash B : D
+%% \end{figure
 %% }
-%%     {\Gamma \vdash (A/\alpha;B/\beta) : (\alpha':C';\beta':D') \to (\alpha:C;\beta':D')}\\
-
-%%     \inferrule*%% [right=NullCtxMap]
-%%                {\Gamma \isadtctx}{\Gamma \vdash \cdot : \cdot \to \cdot}
-
-%%     \inferrule*{\Gamma \pipe \alpha':C' \vdash A : C}{\Gamma \vdash A/\alpha : \alpha':C' \to \alpha:C}
-
-%%     \inferrule
-%%     {\Psi = \Psi_1 \jnctx \Psi_M \jnctx \Psi_2\and
-%%       \Gamma \vdash \phi_1 : \Psi_1 \to \Phi_1\and
-%%       \Gamma \pipe \Psi_M \vdash M : R[d^+\phi_1;d^-\phi_2] \and
-%%       \Gamma \vdash \phi_2 : \Psi_2 \to \Phi_2
-%%       }
-%%     {\Gamma \vdash \phi_1,M/x,\phi_2 : \Psi \to \Phi_1,x:R,\Phi_2}
-%%   \end{mathpar}
-%%   \caption{Substitutions in SCT}
-%%   \label{fig:basic-rules}
-%% \end{figure}
-
-%% \begin{figure}
-%%   \begin{mathpar}
-%%     \inferrule
-%%     {}
-%%     {\cdot : \cdot \to \cdot}
-
-%%     \inferrule
-%%     {\gamma : \Delta \to \Gamma \and \Delta \vdash M : A[\gamma]}
-%%     {\gamma,M/X : \Delta \to \Gamma,X:A}
-
-%%     \inferrule
-%%     {\Gamma \vdash \Phi \isavectx \and
-%%     \Gamma \vdash \Psi \isavectx \and
-%%     d^+\Phi = d^- \Psi}
-%%     {\Gamma \vdash \Phi \jnctx \Psi \isavectx}
-
-%%     \inferrule
-%%     {\Gamma \vdash \phi : \Phi' \to \Phi \and
-%%      \Gamma \vdash \psi : \Psi' \to \Psi\and
-%%      d^+\phi = d^- \psi}
-%%     {\Gamma \vdash \phi \jnctx \psi : \Phi' \jnctx \Psi' \to \Phi \jnctx \Psi}
-
-%%     \inferrule*[right=ObSubst]
-%%     {\alpha : \cat C \vdash B : \cat D\and
-%%     \beta : \cat D \vdash A : \cat E}
-%%     {\alpha : \cat C \vdash A[B/\beta] : \cat E}\\
-
-%%     \inferrule*[right=Everything Subst]{\gamma : \Delta \to \Gamma \and \Gamma J}{\Delta J}
-
-%%     \inferrule{\Gamma \vdash \xi : Z \to \Xi\and \Gamma \pipe \Xi \vdash R \isaSet}{\Gamma \pipe Z \vdash R[\xi] \isaSet}
-
-%%     \inferrule{\Gamma \vdash \phi : \Psi \to \Phi\and \Gamma \pipe \Phi \vdash M : R}{\Gamma \pipe \Psi \vdash M[\phi] : R[\underline \phi]}
-%%   \end{mathpar}
-%%   \begin{align*}
-%%     (\Phi \jnctx \Psi) \jnctx \Sigma &= \Phi \jnctx (\Psi \jnctx \Sigma)\\
-%%     \Phi \jnctx \alpha:C &= \Phi\\
-%%     \alpha:C \jnctx \Psi &= \Psi\\\\
-%%     d^\pm(A/\alpha) &= A/\alpha\\
-%%     d^+(\phi_1,M/x,\phi_2) &= d^+ \phi_2\\
-%%     d^-(\phi_1,M/x,\phi_2) &= d^- \phi_1\\\\
-%%     \underline \cdot &= \cdot\\
-%%     \underline {A/\alpha} &= A/\alpha\\
-%%     \underline {\phi_1,M/x,\phi_2} &= d^-\phi_1;d^+\phi_2
-%%   \end{align*}
-%%   \caption{Admissible Rules}
-%% \end{figure}
+
+\begin{definition}
+  We define composition of transformation contexts $\Phi \jnctx \Psi$ by induction on $\Psi$
+  \begin{align*}
+    \Phi \jnctx \beta:\cat D &= \Phi    \\
+    \Phi \jnctx \Psi,x:R,\beta:\cat D &= (\Phi\jnctx\Psi),x:R,\beta:\cat D
+  \end{align*}
+\end{definition}
+
+It follows by an inductive argument that this definition is
+associative and unital.
+
+\begin{definition}
+  We define the horizontal composition of transformation substitutions
+  $\phi \jnctx \psi$ whenever $d^+\phi = d^-\psi$ by induction on
+  $\psi$:
+  \begin{align*}
+    \phi \jnctx a/\alpha &= \phi\\
+    \phi \jnctx (\psi,t/x,a/\alpha) &= (\phi \jnctx \psi),t/x,a/\alpha
+  \end{align*}
+\end{definition}
+
+It follows by an inductive argument that this operation is associative
+and unital.
+
+\begin{lemma}[Inversion]
+  \begin{enumerate}
+  \item If $\Phi \vdash \psi :: (\alpha:\cat C)$ then $\Phi =
+    \beta:\cat D$ for some $\cat D$ and $\psi = a/\alpha$ where
+    $\beta:\cat D \vdash a : \cat C$.
+  \item If $\Phi \vdash \psi :: \Psi_1 \jnctx \Psi_2$, then there
+    exists unique $\Phi_1,\Phi_2,\psi_1,\psi_2$ such that $\Phi =
+    \Phi_1\jnctx\Phi_2$ and $\Phi_1\vdash\psi_1 :: \Psi_1$ and
+    $\Phi_2\vdash \psi_2 : \Psi_2$ and $\psi = \psi_1\jnctx\psi_2$.
+  \end{enumerate}
+\end{lemma}
+
+\begin{definition}
+  We define the vertical composition of transformation substitutions
+  $\phi[\psi]$ inductively on $\phi$.
+  \begin{align*}
+    (a/\alpha)[b/\beta] &= a[b/\beta]/\alpha\\
+    (\phi_1,t/x,a/\alpha)[\psi_1\jnctx\psi_2] &= \phi_1[\psi_1], t[\psi_2],a[d^+\psi_2]
+  \end{align*}
+  This covers all cases by \cref{inversion}.
+
+  We define the vertical identity $\id_\Phi$ by induction on $\Phi$
+  \begin{align*}
+    \id_{\alpha:\cat C} &= \alpha/\alpha\\
+    \id_{\Phi,x:R,\alpha:\cat C} &= \id_{\Phi},x/x,\alpha/\alpha
+  \end{align*}
+\end{definition}
+By an inductive argument, this operation is associative and unital.
+
+\begin{lemma}{Interchange}
+  We prove the interchange law
+  $$(\phi\jnctx\psi)[\phi'\jnctx\psi'] = \phi[\phi'] \jnctx \psi[\psi']$$
+  by induction on $\psi$.
+
+  \begin{enumerate}
+  \item When $\psi=a/\alpha$
+      \begin{align*}
+    (\phi\jnctx a/\alpha)[\phi'\jnctx b/\beta]
+        &= \phi[\phi']\\
+        &= \phi[\phi'] \jnctx (a[b/\beta]/\alpha)\\
+        &= \phi[\phi'] \jnctx (a/\alpha)[b/\beta]
+  \end{align*}
+    \item When $\psi=\psi,t/x,a/\alpha$
+      \begin{align*}
+        (\phi\jnctx (\psi,t/x,a/\alpha))[\phi' \jnctx (\psi'_1\jnctx\psi'_2)]
+        &= ((\phi\jnctx \psi),t/x,a/\alpha)[(\phi' \jnctx \psi'_1)\jnctx\psi'_2]\\
+        &= (\phi\jnctx \psi)[\phi' \jnctx \psi'_1],t[\psi'_2]/x,a[d^+\psi'_2]/\alpha\\
+        &= (\phi[\psi'] \jnctx \psi[\psi']),t[\psi'_2]/x,a[d^+\psi'_2]/\alpha\\
+        &= \phi[\psi'] \jnctx (\psi[\psi'],t[\psi'_2]/x,a[d^+\psi'_2]/\alpha)\\
+      \end{align*}
+  \end{enumerate}
+\end{lemma}
+
+\begin{figure}
+  \begin{mathpar}
+    % vertical arrows
+    \inferrule*[Right=VComp*]
+    {\alpha:\cat C \vdash b : \cat D \and
+     \beta:\cat D \vdash c : \cat E}
+    {\alpha:\cat C \vdash c[b/\beta] : \cat E}\\
+
+    \beta[b/\beta] = b\and
+    c[\beta/\beta] = c\and
+    c[b/\beta][a/\alpha] = c[b[a/\alpha]/\beta]\\
+
+    % horizontal arrows
+    \inferrule*[Right=HComp*]
+    {\Gamma \vdash \Phi \isavectx\\\\
+      \Gamma\vdash\Psi\isavectx\\\\
+      d^+\Phi = d^-\Psi}
+    {\Gamma \vdash \Phi \jnctx \Psi \isavectx}
+
+    d^+(\Phi \jnctx \Psi) = d^+\Psi\and
+    d^-(\Phi \jnctx \Psi) = d^-\Phi\\
+
+    \Phi \jnctx (d^+\Phi) = \Phi \and
+    (d^-\Phi) \jnctx \Phi = \Phi \and
+    (\Phi \jnctx \Psi) \jnctx \Sigma = \Phi \jnctx (\Psi \jnctx \Sigma)\\
+
+    % 2-cells
+    %% vertical composition
+    \inferrule*[Right=V2Comp*]
+    {\Sigma \vdash \psi :: \Psi \and
+     \Psi \vdash \phi :: \Phi}
+    {\Sigma \vdash \phi[\psi] :: \Phi}
+
+    \inferrule*[Right=V2Id*]
+    {}
+    {\Phi \vdash \id_\Phi :: \Phi}\\
+
+    \id_\Phi[\phi] = \phi\and
+    \phi[\id_\Psi] = \phi\and
+    \phi[\psi][\sigma] = \phi[\psi[\sigma]]\\
+
+    %% Horizontal composition
+    \inferrule*[Right=H2Comp*]
+    {\Phi' \vdash \phi :: \Phi \\\\
+     \Psi' \vdash \psi :: \Psi \\\\
+     d^+\phi = d^-\psi 
+    }
+    {\Phi' \jnctx \Psi' \vdash \phi \jnctx \psi :: \Phi \jnctx \Psi}
+
+    d^-(\phi\jnctx\psi) = d^-\phi\and
+    d^+(\phi\jnctx\psi) = d^+\phi\and
+
+    \phi\jnctx d^+\phi = \phi\and
+    d^-\phi \jnctx \phi = \phi\and
+    \phi\jnctx(\psi\jnctx\sigma) = (\phi\jnctx\psi)\jnctx\sigma
+
+    %% Interchange
+    \inferrule*[Right=Interchange]
+    {\Phi'' \vdash \phi' :: \Phi' \and
+     \Psi'' \vdash \psi' :: \Psi' \and
+     d^+\phi' = d^-\psi'\\\\
+     \Phi' \vdash \phi :: \Phi \and
+     \Psi' \vdash \psi :: \Psi \and
+     d^+\phi = d^-\psi
+    }
+    {(\phi\jnctx\psi)[\phi'\jnctx\psi'] = \phi[\phi'] \jnctx \psi[\psi']}\\
+
+    % Interaction with displays
+    \inferrule
+    {\Psi \vdash \phi :: \Phi\and
+     \Phi \vdash t : R[a/\alpha;b/\beta]}
+    {\Psi\vdash t[\phi] : R[a[d^-\phi]/\alpha;b[d^+\phi]/\beta]}
+
+    t[\id_\Phi] = t\and
+    t[\phi][\psi] = t[\phi[\psi]]
+  \end{mathpar}
+  \caption{Syntactic Double Category with Displays Structure}
+\end{figure}
+
+\begin{figure}
+  \begin{mathpar}
+    \inferrule
+    {}
+    {\cdot : \Gamma \to \cdot}
+
+    \inferrule
+    {\gamma : \Delta \to \Gamma \and \Delta \vdash M : A[\gamma]}
+    {\gamma,M/X : \Delta \to \Gamma,X:A}
+
+    \inferrule
+    {\Gamma \vdash \Phi \isavectx \and
+    \Gamma \vdash \Psi \isavectx \and
+    d^+\Phi = d^- \Psi}
+    {\Gamma \vdash \Phi \jnctx \Psi \isavectx}
+
+    \inferrule
+    {\Gamma \vdash \phi : \Phi' \to \Phi \and
+     \Gamma \vdash \psi : \Psi' \to \Psi\and
+     d^+\phi = d^- \psi}
+    {\Gamma \vdash \phi \jnctx \psi : \Phi' \jnctx \Psi' \to \Phi \jnctx \Psi}
+
+    \inferrule*[right=ObSubst]
+    {\alpha : \cat C \vdash B : \cat D\and
+    \beta : \cat D \vdash A : \cat E}
+    {\alpha : \cat C \vdash A[B/\beta] : \cat E}\\
+
+    \inferrule*[right=Everything Subst]{\gamma : \Delta \to \Gamma \and \Gamma J}{\Delta J}
+
+    \inferrule{\Gamma \vdash \xi : Z \to \Xi\and \Gamma \pipe \Xi \vdash R \isaSet}{\Gamma \pipe Z \vdash R[\xi] \isaSet}
+
+    \inferrule{\Gamma \vdash \phi : \Psi \to \Phi\and \Gamma \pipe \Phi \vdash M : R}{\Gamma \pipe \Psi \vdash M[\phi] : R[\underline \phi]}
+  \end{mathpar}
+  \begin{align*}
+    (\Phi \jnctx \Psi) \jnctx \Sigma &= \Phi \jnctx (\Psi \jnctx \Sigma)\\
+    \Phi \jnctx \alpha:C &= \Phi\\
+    \alpha:C \jnctx \Psi &= \Psi\\\\
+    d^\pm(A/\alpha) &= A/\alpha\\
+    d^+(\phi_1,M/x,\phi_2) &= d^+ \phi_2\\
+    d^-(\phi_1,M/x,\phi_2) &= d^- \phi_1\\\\
+    \underline \cdot &= \cdot\\
+    \underline {A/\alpha} &= A/\alpha\\
+    \underline {\phi_1,M/x,\phi_2} &= d^-\phi_1;d^+\phi_2
+  \end{align*}
+  \caption{Admissible Rules}
+\end{figure}
 

### 2022-04-28 paper/paper.tex
missing last step of interchange
complete draft of the semantics section

@@
 transformation judgment and the functor $i$ models the construction of
-a ``singleton'' context $\alpha:\cat C,x:R,\beta:\cat D$.
+a ``singleton'' context $\alpha:\cat C,x:R,\beta:\cat D$. The fact
+that $i$ is fully faithful corresponds to the fact that we don't
+provide in the syntax a notion of morphism of profunctors separate
+from the notion of substitution of singleton contexts.
 
@@
 
+The final notion of composition that we require for our structure is
+that display morphisms have \emph{restrictions} along vertical arrows.
+
+\begin{definition}
+  A DCD is an Equipment with displays when
+  \begin{enumerate}
+  \item The functor $(s,t) : C_d \to C_0$ is a split fibration, with
+    restriction operation written as $(f,g)^*R$.
+  \item The profunctor $C_\vdash$ respects restrictions in that
+    \[ C_\vdash^{f',g'}(\Phi, (f,g)^*R) \cong C_\vdash^{f\circ f',g\circ g'}(\Phi, R) \]
+  \end{enumerate}
+\end{definition}
+Specializing this property give us that $C_\vdash^{f,g}(\Phi,R) \cong
+C_\vdash^{\id,\id}(\Phi,(f,g)^*R)$, which justifies that we only need
+to give syntax for terms $\Phi \vdash t : R$ where $\Phi$ and $R$ have
+the same covariant/contravariant dependencies.
+
+We can then define the remaining connectives by UMPs in an equipment
+with displays, adding in the extra strictness criterion that these
+connectives appropriately commute with restriction.
+
+\begin{figure}
+  \begin{mathpar}
+    C_\vdash^{f,g}(\Phi \circ_{C} \Psi, R) \cong C_\vdash^{f,g}(\Phi \circ_C i(U_C) \circ_c \Psi, R) \\
+    C_\vdash^{f,g}(\Phi \circ_{C} iP \circ_D iQ \circ_E \Psi, R) \cong C_\vdash^{f,g}(\Phi \circ_C i(P \odot Q) \circ_E \Psi, R) \and (f,g)^*(P \odot Q) = (f,\id)^*P \odot (\id,g)^*Q\\
+    C_\vdash^{\id_C,g}(P \circ_c \Phi, R) \cong C_\vdash^{\id_D,g}(\Phi, P \triangleright Q) \and (f,g)^*(P \triangleright Q) = (\id,f)^*P \triangleright (\id,g)^*Q\\
+    C_\vdash^{f, \id_C}(\Phi \circ_c R, P) \cong C_\vdash^{f, \id_D}(\Phi, P \triangleleft Q)\and (f,g)^*(P \triangleleft Q) = (f,\id)^*P \triangleleft (g,\id)^*Q\\
+    C_\vdash^{f,g}(\Phi,1) \cong 1 \and (f,g)^*1 = 1\\
+    C_\vdash^{f,g}(\Phi,P \times Q) \cong C_\vdash^{f,g}(\Phi,P) \times C_\vdash^{f,g}(\Phi, Q)\and (f,g)^*(P \times Q) = (f,g)^*P \times (f,g)^*Q\\
+    C_0(C,\smallint R) \cong C_\vdash(\id_C, R)\\
+  \end{mathpar}
+  \caption{UMPs in an Equipment with Displays}
+\end{figure}
+
 Next, each of the connectives of SCT are modeled by different
 universal properties that are profunctor isomorphisms involving
-principally $C_\vdash$.
-TODO: write them down.
+principally $C_\vdash$. We define these in Figure\ref{fig:eq-umps}
 
@@
 \begin{definition}
-  A Category with Families is...
+  A Category with Families consists of
+  \begin{enumerate}
+  \item A category $\textrm{Ctx}$
+  \item with a terminal object $1 \in \textrm{Ctx}$
+  \item A presheaf $\textrm{Ty} : \Set^\textrm{Ctx}$ of types in context.
+  \item A presheaf $\textrm{Tm} : \Set^{\int \textrm{Tm}}$ of terms of
+    a given type.
+  \item With a context extension functor $\textrm{ext} : \int
+    \textrm{Ty} \to \textrm{Ctx}$ with universal property:
+    \[ \textrm{Ctx}(\Delta,\textrm{ext}(A)) \cong (\gamma : \textrm{Ctx}(\Delta,\textrm{ctx}(A))) \times \textrm{Tm}(A[\gamma])  \]
+  \end{enumerate}
 \end{definition}
 
+This can be extended in a standard way to model $\Pi,\Sigma, \Id$
+types \citet{cwf-stuff}.
+%
+The final connectives we need are the ``universes'' that internalize
+the double category structure as types in the CwF.
+
 \begin{definition}
-  A hyperdoctrine of equipment with displays (HED) is a pair of a
-  category with families $Ctx$ modeling $\Sigma,\Pi,\textrm{Id}$ and a
-  DCD internal to $\Set^{Ctx^o}$ with categories, functors, displays
-  and unary cells representable as types.
+  Let $(\textrm{Ctx},\textrm{Ty},\textrm{Tm},\textrm{ext})$ be a
+  category with families and let $\mathcal V$ be a DCD internal to
+  $\Set^\textrm{Ctx}$. Then
+  \begin{enumerate}
+  \item $\mathcal V$ objects are representable as types if there is a
+    type given by a section $\textrm{Ob} \in \textrm{Ty}(1)$ with a
+    natural isomorphism
+    \[ \textrm{Tm}(!_\Gamma^*\textrm{Ob}) \cong \textrm{Ob}(\mathcal V)_\Gamma \]
+  \item $\mathcal V$ vertical arrows are representable as types if
+    there is a type
+    \[ \textrm{Tm}(((\gamma,\lceil a\rceil),\lceil b\rceil)^*\textrm{Functor} \cong
+    \textrm{VArr}(\mathcal V_{\Delta})(a,b)
+    \]
+    and similarly horizontal displays are representable as types if
+    \[ \textrm{Tm}(((\gamma,\lceil a\rceil),\lceil b\rceil)^*\textrm{Profunctor} \cong
+    \textrm{Disp}(\mathcal V_{\Delta})(a,b)
+    \]
+  \item Closed $\mathcal V$ transformations are representable as types if
+    \[ \textrm{Tm}() \]
+  \end{enumerate}
 \end{definition}
 
-Again this notion is straightforwardly, if laboriously, interpretable
+We then arrive at our final notion of model by collecting all of this
+algebraic structure into one definition.
+\begin{definition}
+  A hyperdoctrine of equipments with displays (HED) is a pair of a
+  category with families $Ctx$ modeling $\Sigma,\Pi,\Id$ and a DCD
+  internal to $\Set^{Ctx^o}$ with categories, functors, displays and
+  unary cells representable as types.
+\end{definition}
+
+This notion of HED is straightforwardly, if laboriously, interpretable
 as a GAT.
-
-%% \begin{definition}[Unit]
-%%   A unit $U_{\cat C}\in V_h$ with $d^\pm U_{\cat C} = \cat C$ and an op-cartesian 2-cell
-%%   \[ r : V_h(\cat C, U_{\cat C}) \]
-%%   with $d^\pm r = \id_{\cat C}$.
-%% \end{definition}
-
-%% \begin{definition}[Tensor]
-%%   If $R,Q \in V_h$ with $d^+R = d^-Q$, a tensor is a horizontal arrow
-%%   $R \odot Q \in V_h$ with $d^+ (R\odot Q) = d^+ R$ and $d^-(R \odot
-%%   Q) = d^- Q$ with an op-cartesian 2-cell
-%%   \[ p : V_h(R,Q;R \odot Q) \]
-%% \end{definition}
-
-\begin{definition}{Indexed Virtual Equipment}
-  A VE-hyperdoctrine is a pair of a category with families $(C,
-  \textrm{Ty}, \textrm{Tm})$ and a functor $\cat C^{op} \to VE$. A
-  morphism of VE-hyperdoctrines
-  \[(F,\phi) : (\cat C, \mathcal V : \cat C^{o} \to VE) \to (\cat D, \mathcal U : \cat D^{o} \to VE)\]
-  is a pair of a functor of categories with families $F : \cat C \to
-  \cat D$ and a natural transformation $\phi : \mathcal V
-  \Rightarrow\mathcal U \circ F^{o}$.
-\end{definition}
 
@@
 
-\subsection{Generalized Algebraic Theory of VE Hyperdoctrines}
-
-To simplify our initiality proof, we use a \emph{generalized
-algebraic} presentation of our notion of
-model\cite{cartmell,sterling-eat}. To describe this, we give a
-``combinator'' syntax for VE hyperdoctrines, where we have explicit
-sorts for contexts $\Gamma$ and $\Phi$ and explicit judgments for
-substitutions $\gamma : \Delta \to \Gamma$ and $\phi : \Psi
-\vdash_{f,g} \Phi$.
-%
-While the explicit description of the theory is quite long, we can
-describe it in simple modular pieces.
-%
-First, we use the standard theory of a category with families, which
-describes the algebraic structure of contexts $\Gamma$, types $\Gamma
-\vdash A$ and terms $\Gamma \vdash M : A$.
-%
-Second, we add to that a theory for a slight generalization of virtual
-double categories we call \emph{double categories with display
-pro-arrows}.
-%
-The difference is that in a double category with display pro-arrows,
-we have a true double category with strict vertical and horizontal
-composition, among whom are the ``display'' pro-arrows.
-%
-The display pro-arrows model the profunctors of the syntax, which
-don't have a strict composition, whereas the arbitrary pro-arrows
-model the contexts $\Phi$, which do have a strict composition, given
-by context concatenation $\Phi,\Psi$.
-%
-The difference is that in a virtual double category the contexts are
-assumed to be given inductively by formal compositions of the display
-pro-arrows, whereas in models of this algebraic theory, the contexts
-might include other exotic objects.
-%
-This difference is analogous to the difference between a category with
-families and a contextual category in models of dependent type theory,
-and we fully expect that there is a similar relationship between the
-models: virtual double categories should form a co-reflective
-subcategory characterized as those double categories with display
-pro-arrows where all pro-arrows are equal to some concatenation of
-displays.
-%
-Finally, we parameterize all of the constructions of the theory of
-double categories with pro-arrows by contexts $\Gamma$ and add in
-equations that substitution commutes with all double category
-operations.
-%
-This means the theory presents a category with families with an double
-category with display pro-arrows internal to the category of
-presheaves on the category of contexts, i.e., a hyperdoctrine of
-double categories with display pro-arrows.
-
-The main advantage of this algebraic approach is that the initiality
-theorem for the combinatory syntax is trivial. Instead, the work is to
-show that our type theoretic syntax, which restricts combinatory rules
-to make substitution an admissible rather than primitive operation, is
-equivalent to the combinatory syntax.
-
-\begin{figure}
-  \begin{mathpar}
-    \inferrule
-    {}
-    {\Gamma : \algCtx}
-
-    \inferrule
-    {\Delta : \algCtx \and \Gamma : \algCtx}
-    {\gamma : \algSubst(\Delta,\Gamma)}
-
-    \inferrule
-    {\Gamma : \algCtx}
-    {C : \algCat(\Gamma)}
-
-    \inferrule
-    {\Gamma : \algCtx \and C : \algCat(\Gamma) \and D : \algCat(\Gamma) \and
-    }
-    {F : \algVarr(\Gamma,C,D)}
-
-    \inferrule
-    {\Gamma : \algCtx \and C : \algCat(\Gamma) \and D : \algCat(\Gamma) \and
-    }
-    {P : \textrm{Profunctor}(\Gamma,C,D)}
-
-    \inferrule
-    {\Gamma : \algCtx \and C : \algCat(\Gamma) \and D : \algCat(\Gamma)
-    }
-    {P : \textrm{ProfCtx}(\Gamma,C,D)}
-
-    \inferrule
-    {\Gamma : \textrm{Ctx}\and
-      C  : \algCat(\Gamma) \and D  : \algCat(\Gamma) \and
-      C' : \algCat(\Gamma) \and D' : \algCat(\Gamma) \and
-      f : \algVarr(\Gamma,C,C') \and g : \algCat(\Gamma,D,D') \and
-      \Phi : \textrm{ProfCtx}(\Gamma,C,D)\and
-      \Psi : \textrm{ProfCtx}(\Gamma,C',D')
-    }
-    {\phi : \textrm{ProfSubst}(\Gamma,C,D,C',D',f,g,\Phi,\Psi)}
-
-    \inferrule
-    {\Gamma : \textrm{Ctx}\and
-      C  : \algCat(\Gamma) \and D  : \algCat(\Gamma) \and
-      C' : \algCat(\Gamma) \and D' : \algCat(\Gamma) \and
-      f : \algVarr(\Gamma,C,C') \and g : \algCat(\Gamma,D,D') \and
-      \Phi : \textrm{ProfCtx}(\Gamma,C,D)\and
-      P : \textrm{ProfCtx}(\Gamma,C',D')
-    }
-    {t : \textrm{Trans}(\Gamma,C,D,C',D',f,g,\Phi,P)}
-  \end{mathpar}
-  \caption{Sorts of Algebraic SCT}
-\end{figure}
-
-\begin{figure}
-  \begin{mathpar}
-    \cdots
-  \end{mathpar}
-  \caption{Algebraic Dependent Type Theory}
-\end{figure}
-
-The algebraic theory of a double category with displays indexed by a
-category is given in Figure~\ref{fig:alg-dc-disp}.
-%
-First, we give the categories and functors the theory of a category
-indexed by a category.
-%
-Next, we give categories and \emph{profunctor contexts} the same
-theory, where the identity $\id_h$ is the ``empty'' context $\alpha:C$ and the composition $\Phi \circ_h$.
-%
-
-TODO: include all of the interactions with substitutions and restrictions.
-
-\begin{figure}
-  \begin{mathpar}
-    % theory of an indexed category (vertical)
-    \inferrule
-    {g : \algVarr(D,E) \and \algVarr(C,D)}
-    {g \circ_v f : \algVarr(C,E)}
-
-    (g \circ_v f) \circ_{vs} \gamma = (g \circ_{vs} \gamma) \circ_v (f \circ_s \gamma)
-
-    (h \circ_v g) \circ_v f = h \circ_v (g \circ_v f)
-
-    \inferrule
-    {}
-    {\id_v : \algVarr(C,C)}
-
-    \id_v \circ_{vs} \gamma = \id_h
-
-    (f \circ_v \id_v) = f\and \id_v \circ_v f = f
-
-    % theory of an indexed category (horizontal)
-
-    \inferrule
-    {\Psi : \algHCtx(D,E) \and \algHCtx(C,D)}
-    {\Psi \circ_h \Phi : \algHCtx(C,E)}
-
-    (\Psi \circ_h \Phi) \circ_{hs} \gamma = (\Psi \circ_{hs} \gamma) \circ_h (\Phi \circ_s \gamma)
-
-    (\Sigma \circ_h \Psi) \circ_h \Phi = \Sigma \circ_h (\Psi \circ_h \Phi)
-
-    \inferrule
-    {}
-    {\id_h : \algHCtx(C,C)}
-
-    \id_h \circ_{hs} \gamma = \id_h
-
-    (\Phi \circ_h \id_h) = \Phi\and \id_h \circ_h \Phi = \Phi
-
-    % embedding of displays
-
-    \inferrule{P : \algHarr(C,D)}{i(P) : \algHCtx(C,D)}
-
-    % 2-cells: substitutions
-    % vertical category
-    % horizontal category
-    % interchange law
-    \inferrule
-    {\psi : \algHSubst(\Phi, f',g',\Psi)\and \phi : \algHSubst(\Sigma,f,g,\Phi)
-    }
-    {\psi \circ_{2v} \phi : \algHSubst(\Sigma, f' \circ f, g' \circ g, \Psi)}
-
-    \inferrule
-    {}
-    {\id_{2v} : \algHSubst(\Phi,\id,\id,\Phi)}
-
-    (\phi \circ_{2v} \psi) \circ_{2v} \sigma = \phi \circ_{2v} (\psi \circ_{2v} \sigma)\\
-    \phi \circ_{2v} \id_{2v} = \phi \and \phi = \id_{2v} \circ_{2v} \phi\\
-
-    \inferrule
-    {\phi_1 : \algHSubst(\Psi_1,f,g,\Phi_1)\and \phi_2 : \algHSubst(\Psi_2,f,g,\Phi_2)
-    }
-    {\phi_1 \circ_{2h} \phi_2 : \algHSubst(\Psi_1 \circ_h \Psi_2, f, h, \Phi_1 \circ_h \Phi_2)}
-
-    \inferrule
-    {}
-    {\id_{2h} : \algHSubst(\id_h,f,f,\id_h)}
-
-    (\phi \circ_{2h} \psi) \circ_{2h} \sigma = \phi \circ_{2h} (\psi \circ_{2h} \sigma)\\
-    \phi \circ_{2h} \id_{2h} = \phi \and \id_{2h} \circ_{2h} \phi = \phi\\
-
-    (\phi_1 \circ_{2h} \phi_2) \circ_{2v} (\psi_1 \circ_{2h} \psi_2) = (\phi_1 \circ_{2v} \psi_1) \circ_{2h} (\phi_2 \circ_{2v} \psi_2)
-
-  \end{mathpar}
-  \caption{Double Category with Display Pro-arrows}
-  \label{fig:alg-dc-disp}
-\end{figure}
-
-\begin{figure}
-  \begin{mathpar}
-    % Restrictions
-    \inferrule
-    {P : \algHarr(\Gamma,C',D')
-    \and f : \algVarr(\Gamma,C,C')\and
-    \and g : \algVarr(\Gamma,D,D')}
-    {P \circ_r (f,g) : \algHarr(\Gamma,C,D)}
-
-    (P \circ_r (f,g) \circ_s \gamma) = (P \circ_s \gamma) \circ_r (f \circ_s \gamma, g \circ_s \gamma)
-
-    P \circ_r (f,g) \circ_r (f',g') = P \circ_r (f \circ_r f', g \circ_r g')
-
-    P \circ_r (\id,\id) = P
-
-    % Prof Multiplicatives
-    % LHom
-    \inferrule
-    {P : \algHarr(C, D) \and Q : \algHarr(C,E)}
-    {\algLHom(P,Q) : \algHarr(D,E)}
-
-    \algLHom(P,Q) \circ_{hs} \gamma = \algLHom(P \circ_{hs} \gamma, Q \circ_{hs} \gamma)\and
-    \algLHom(P,Q) \circ_{r} (f,g) = \algLHom(P \circ_r (\id,f), Q \circ_r(\id,g))
-
-    \inferrule
-    {}
-    {\algLHomE(P,Q) : \algHSubst(i(P)\circ_h i(\algLHom(P,Q)),\id_v,\id_v,i(Q))}
-
-    \inferrule
-    {t : \algHSubst(i(P) \circ_h \Phi, \id,g, i(Q))}
-    {\algLHomI(\Phi,g,t,P,Q) : \algHSubst(\Phi,\id,g,i(\algLHom(P,Q)))}
-
-    \algLHomE \circ_{2v} (\id_v(P) \circ_h \algLHomI(t)) = t\and
-    \theta = \algLHomI(\algLHomE \circ_{2v} (\id_v(P) \circ_h \theta))\and
-
-    % RHom
-    \inferrule
-    {P : \algHarr(C, E) \and Q : \algHarr(D,E)}
-    {\algRHom(P,Q) : \algHarr(C, D)}
-
-    \algRHom(P,Q) \circ_{hs} \gamma = \algRHom(P \circ_{hs} \gamma, Q \circ_{hs} \gamma)\and
-    \algRHom(P,Q) \circ_{r} (f,g) = \algRHom(P \circ_r (f, \id), Q \circ_r(g,\id))
-
-    \inferrule
-    {}
-    {\algRHomE(P,Q) : \algHSubst(i(\algRHom(P,Q)) \id_h i(Q),\id_v,\id_v,i(P))}
-
-    \inferrule
-    {t : \algHSubst(\Phi \circ_h i(Q), f, \id, i(P))}
-    {\algRHomI(\Phi,f,t,P,Q) : \algHSubst(\Phi,\id,g,i(\algLHom(P,Q)))}
-
-    \algRHomE \circ_{2v} (\algRHomI(t) \circ_h \id(Q)) = t\and
-    \theta = \algRHomI(\algRHomE \circ_{2v} (\theta \circ_h \id(Q)))\and
-
-    % Unit
-    \inferrule
-    {C : \algCat}
-    {\algUnit(C) : \algHarr(C,C)}
-
-    \algUnit(C) \circ_{hs} \gamma = \algUnit(C \circ_{cs} \gamma)
-
-    \inferrule
-    {}
-    {\algUnitI(C) : \algSubst(\id_h(C),\id,\id,i(\algUnit(C)))}
-
-    \inferrule
-    {t : \algHSubst(\Phi \id_h \Psi,f,g,i(R))}
-    {\algUnitE(t) : \algHSubst((\Phi \circ_h i(\algUnit(C)) \circ_h \Psi), f, g, i(R))}
-
-    \algUnitE(t) \circ_{2v} (\id_{2v}(\Phi) \circ_h \algUnitI(C) \circ_h \id_{2v}(\Psi)) = t\and
-    s = \algUnitE(s \circ_{2v} (\id_{2v}(\Phi) \circ_h \algUnitI(C) \circ_h \id_{2v}(\Psi)))
-
-    % Tensor
-    \inferrule
-    {P : \algHarr(C,D) \and Q : \algHarr(D,E)}
-    {\algTensor(P,Q) : \algHarr(C,E)}
-
-    \algTensor(P,Q) \circ_{hs} \gamma = \algTensor(P \circ_{hs} \gamma, Q \circ_{hs} \gamma)\and
-    \algTensor(P,Q) \circ_{r} (f,g) = \algTensor(P \circ_r (f,\id), Q \circ_r (\id,g))
-
-    \algTensorI(P,Q) : \algHSubst(i(P) \circ_h i(Q),\id,\id,i(\algTensor(P,Q)))
-
-    \inferrule
-    {t : \algHSubst(\Phi \circ_h i(P) \circ_h i(Q) \circ_h \Psi, f,g,i(R))}
-    {\algTensorE(t) : \algHSubst(\Phi \circ_h i(\algTensor(P,Q)) \circ_h \Psi, f,g,i(R))}
-
-    \algTensorE(t) \circ_{2v} (\id_{2v}(\Phi) \circ_{2h} \algTensorI(P,Q) \circ_{2h} \id_{2v}(\Psi)) = t\and
-    s = \algTensorE(s \circ_{2v} (\id_{2v}(\Phi) \circ_{2h} \algTensorI(P,Q) \circ_{2h} \id_{2v}(\Psi)))
-
-    % Cat Multiplicatives
-    % Cat of Elements
-    \inferrule
-    {P : \algHarr(\Gamma,C,D)}
-    {\algElts(C,D,P) : \algCat(\Gamma)}
-
-    \inferrule
-    {}
-    {\pi^- : \algVarr(\algElts(C,D,P),C)}
-
-    \inferrule
-    {}
-    {\pi^+ : \algVarr(\algElts(C,D,P),D)}
-
-    \inferrule
-    {}
-    {\pi^p : \algHSubst(\id_h(\algElts(C,D,P)),\pi^-,\pi^+,P)}
-
-    \inferrule
-    {f : \algVarr(E,C) \and g : \algVarr(E,D)\and
-      t : \algHSubst(\id_h(E), f, g, P)
-    }
-    {\algEltsI(t,f,g) : \algVarr(E,\algElts(C,D,P))}
-
-    \pi^- \circ_v \algEltsI(t,f,g) = f\and
-    \pi^+ \circ_v \algEltsI(t,f,g) = g\and
-    \pi^p \circ_{2v} \algEltsI(t,f,g) = \alpha
-
-    \inferrule
-    {h : \algVarr(E,\algElts(C,D,P))}
-    {h = \algEltsI(\pi^-\circ_v h,\pi^+\circ_vh,\pi^p \circ_{2v} \id_{2h}(h))}
-
-    % Presheaf cats
-
-    % Cat Additives
-    % Product
-    % Terminal
-
-    % Prof Additives
-    % Product
-    % Terminal
-  \end{mathpar}
-  \caption{Multiplicatives}
-\end{figure}
-
-\begin{figure}
-  \begin{mathpar}
-    % Internalization
-    % Cats
-    \inferrule
-    {}
-    {\algCatTy : \algTy(1)}
-
-    \inferrule
-    {C : \algCat(\Gamma)}
-    {\algCatQt(C) : \algTm(\Gamma,\algCatTy \circ !)}
-
-    \algCatQt(C) \circ_{tms} \gamma = \algCatQt(C \circ_{cs} \gamma)\and
-
-    \inferrule
-    {M : \algTm(\Gamma,\algCatTy \circ !)}
-    {\algCatUnqt(M) : \algCat(\Gamma)}
-
-    \algCatUnqt(M) \circ_{cs} \gamma = \algCatUnqt(M \circ_{tms} \gamma)\and
-    \algCatUnqt(\algCatQt(C)) = C\and
-    M = \algCatQt(\algCatUnqt(M))
-
-    % Functor
-    % Prof
-    % ProfCtx?
-    % Trans
-    \inferrule
-    {P : \algHarr(\Gamma,C,C)}
-    {\algTransTy(P) : \algTy(\Gamma)}
-
-    \inferrule
-    {t : \algHSubst(\id_h,\id_v,\id_v,i(P))}
-    {\algTransQt(t) : \algTm(\Gamma,\algTransTy(P))}
-
-    \algTransQt(t) \circ_{tms} \gamma = \algTransQt(t \circ_{2s} \gamma)
-
-    \inferrule
-    {}
-    {\algTransUnqt : \algHSubst(ext(\Gamma,\algTransTy(P)),\id_h,\id_v,\id_v,i(P \circ_{hs} \pi_1))}
-
-    \algTransUnqt \circ_{2s} (\gamma,\algTransQt(t)) = t\\
-
-    % M : Tm(G,∀ α. P α α)
-    % unquote : Subst(G,t : ∀ α:C. P α α; α:C, id,id, P α α)
-    % M = unquote()
-    M = \algTransQt(\algTransUnqt \circ_{2s} ext_s(\id,M))
-  \end{mathpar}
-  \caption{Internalization}
-\end{figure}
-
-\subsection{Equivalence between Natural Deduction and Combinatory Syntaxes}
-
-Now we aim to prove that our natural deduction syntax of SCT is sound
-and complete for the algebraic notion of hyperdoctines of double
-categories with displays.
-%
-In category-theoretic terms, we want to show that the syntax of SCT
-presents an \emph{initial object} in the category of models.
-%
-This is proved in two steps: first, we show that SCT is a model
+\subsection{Initiality}
+
+In this section we give a sketch of an initiality proof of the syntax
+of SCT in HEDs. Note that we consider all the following sets up to
+$\alpha$-equivalence of names.
 
 \begin{construction}{Completeness}
-  SCT presents a hyperdoctrine of double categories with displays.
+  SCT presents a HED with
+  \begin{enumerate}
+  \item Category with Families provided by the contexts,
+    substitutions, types and terms.
+  \item Objects of $C_0^\Gamma$ provided by the terms $\Gamma \vdash \cat C \isaCat$
+  \item Morphisms of $C_0^\Gamma$ provided by terms $\Gamma
+    \alpha:\cat C \vdash b \cat D$ with source $\cat C$ and target
+    $\cat D$. With identity given by a variable and composition by identity.
+  \item Objects of $C_1^\Gamma$ are provided by the profunctor
+    contexts $\Gamma \vdash \Phi \isavectx$ with morphisms given by
+    transformation substitutions $\Gamma \pipe \Phi \vdash \psi :
+    \Psi$.
+  \item Horizontal composition of contexts is defined by gluing along
+    a variable $\Phi\jnctx \Psi$ and the horizontal identity is given
+    by the single category variable $\alpha:\cat C$.
+  \item Objects of $C_\vdash^\Gamma$ are provided by the profunctors
+    $\Gamma\alpha:\cat C;\beta:\cat D \vdash R$, and a morphism from
+    $\alpha:\cat C;\beta:\cat D \vdash R$ to $\alpha':\cat
+    C';\beta':\cat D' \vdash R'$ is a triple $(a,t,b)$ of functor
+    terms $\alpha:\cat C \vdash a : \cat C'$ and $\beta:\cat D\vdash b
+    : \cat D'$ and a term $\alpha:\cat C,x:R,\beta:\cat D \vdash t :
+    R'[a/\alpha';b/\beta']$
+
+    Composition $(a,t,b) \circ (a',t',b') = (a[a'],t[t'],b[b'])$ is
+    given by component-wise single-place substitution with the triple
+    of all variables the identity.
+  \item The inclusion functor $i^\Gamma : C_d \to C_1$ is defined on
+    types as the singleton context $i(R) = \alpha:\cat
+    C,x:R,\beta:\cat D$ and on morphisms as the singleton
+    substitution.
+  \item The restriction operation $R \circ (a,b)$ is defined as the
+    substitution $R[a;b]$.
+  \item The remaining universal properties are satisfied using the
+    analogous type theoretic construction.
+  \end{enumerate}
 \end{construction}
+
+The associativity, unit and interchange laws of a category are proven
+in the \cref{appendix}.
+
+\begin{theorem}{(Sketch) Soundness/Initiality}
+  Given any HED $\mathcal V$, there is a unique homomorphism of HEDs $\sem\cdot$
+  from $\textrm{Syn}(\textrm{SCT})$ to $\mathcal V$.
+\end{theorem}
 \begin{proof}
-  First, the sorts.
-  \begin{align*}
-    \eta(\algCtx) &= \{ \Gamma | \Gamma \isadtctx \}\\
-    \eta(\algSubst(\Delta,\Gamma)) &= \{ \gamma | \eta(\Delta) \vdash \gamma : \eta(\Gamma) \}\\
-    \eta(\algTy(\Gamma)) &= \{ A |  \eta(\Gamma) \vdash A \isaTy \}\\
-    \eta(\algTm(\Gamma,A)) &= \{ M | \eta(\Gamma) \vdash M : \eta(A) \}\\
-    \eta(\algCat(\Gamma)) &= \{ \cat C | \eta(\Gamma) \vdash \cat C \isaCat \}\\
-    \eta(\algVarr(\Gamma,C,D)) &= \{ a | \eta(\Gamma)|\alpha:\eta(C) \vdash a : \eta(D) \}\\
-    \eta(\algHarr(\Gamma,C,D)) &= \{ R | \eta(\Gamma)|\alpha:\eta(C);\beta:\eta(D) \vdash R \isaSet \}\\
-    \eta(\algHCtx(\Gamma,C,D)) &= \{ \Phi | \eta(\Gamma) \vdash \Phi \isavectx \wedge d^-(\Phi) = \eta(C) \and d^+(\Phi) = \eta(D) \}\\
-    \eta(\algHSubst(\Gamma, \Psi, f, g, \Phi)) &= \{ \phi | \eta(\gamma)\pipe \eta(\Psi) \vdash \phi : \eta(\Phi) \wedge d^-\phi = \eta(f) \wedge d^+\phi = \eta(g) \}\\
-    1 &= \cdot\\
-    ! &= \cdot\\
-    ext(\Gamma,A) = \Gamma, x:A \\
-    \cdots\\
-    \Phi \circ_h \Psi &= \Phi \jnctx \Psi\\
-    \id_h(\cat C) &= \alpha:\cat C\\
-    i(\cat C, \cat D, R) &= \alpha:\cat C, x: R, \beta:\cat D\\
-    \phi \circ_{2v} \psi &= \phi[\psi]\\
-    \id_{2v}(\Phi) &= \id_\Phi\\
-    \id_{2h}(\alpha:\cat C \vdash a : \cat D) &= a/\beta\\
-    \phi \circ_{2h} \psi &= \phi \jnctx \psi
-  \end{align*}
+  Roughly the converse of the completeness theorem. For instance,
+  profunctor contexts are interpreted using horizontal identity and
+  composition as well as the $i$ functor. Preservation of the
+  restriction operation requires that restriction commutes with the
+  display connectives.
 \end{proof}
-\begin{corollary}
-  There is a unique homomorphism of HDCDs
-  \[ \eta : Comb \to SCT \]
-\end{corollary}
-
-\begin{construction}{Soundness}
-  We construct a homomorphism of HDCDs
-  \[ i : SCT \to Comb \]
-  such that the composite
-  \[ \eta \circ i : SCT \to SCT \]
-  is equal to the identity.
-\end{construction}
-\begin{proof}
-  \begin{align*}
-    i(\cdot) &= 1\\
-    i(\Gamma,x:A)  &= ext(i(\Gamma),ext(A))\\
-    \cdots
-    i(\Gamma \pipe \alpha:\cat C \vdash \alpha:\cat C) &= \id_v\\
-    i(\Gamma \pipe \alpha:\cat C \vdash M a :\cat C) &= (\algVarrUnqt \circ_{vs} (\id, (i(M)))) \circ_v i(a)\\
-    % Presheaf intro
-    % elements intro
-    i(\Gamma \pipe \alpha:\cat C \vdash (a,t,b) : \Sigma_{\alpha:\cat C;\beta:\cat D} R) &= \algEltsI(i(a),i(t),i(b))\\
-    %
-  \end{align*}
-\end{proof}
-
-\begin{corollary}
-  SCT is initial in the category of HDCDs.
-\end{corollary}
-\begin{proof}
-  Since Comb is initial, the composite $i \circ \eta$ is the identity,
-  and as above the composite $\eta \circ i$ so SCT is isomorphic to
-  Comb, and therefore it is also initial.
-\end{proof}
+
+Details of the interpretation of profunctor contexts and substitutions
+are given in the appendix\cref{appendix}.
 
@@
 
-Our main intended models are when the indexing category is $\Set$ and
-the hyperdoctrine is given by families of objects, arrows etc, of a
-fixed virtual equipment $\mathcal V$, indexed by the set
-$\sem\Gamma$. We can formalize this as the ``internal logic'' of the
-virtual equipment $\mathcal V$.
-
+We now describe how to interpret the syntax in any virtual equipment
+(that can model the connectives). First, from any virtual equipment we
+can construct a DCD:
+
+\begin{definition}
+  Given a virtual equipment $\mathcal V$, define a DCD by taking the
+  horizontal morphisms of $\mathcal V$ as the display morphisms and
+  the horizontal morphisms as the ``composable strings'' of horizontal
+  morphisms. Then we define $C_\vdash$ using the notion of
+  ``multi-2-cell''.
+\end{definition}
+
+Next, we can upgrade any DCD in $\Set$ to a DCD internal to
+$\Set^\Set$ by taking the presheaf of objects, vertical arrows, etc as
+families of objects, vertical arrows etc indexed by the indexing set
+$\Gamma$.
 \begin{definition}[Powers of Virtual Equipments]
@@
   \item Objects are functions $X \to \mathcal V_0$.
-  \item Vertical morphisms $\mathcal V^X_v(A,B)$ are functions
-    $\Pi_{x\in X} V_v(A x,B x)$. With pointwise composition, identity.
-  \item Horizontal morphisms $\mathcal V^X_h$ are functions $\Gamma \to \mathcal V_h$, with
-    $(d^\pm R)x = d^\pm (Rx)$.
-  \item Restriction of $R$ along $f,g$ is given by $(R(f,g))x = (R x)(f x,g x)$.
+  \item Vertical morphisms $\mathcal V^X_0(A,B)$ are functions
+    $\Pi_{x\in X} V_0(A x,B x)$. With pointwise composition, identity.
+  \item Horizontal morphisms $\mathcal V^X_1$ are functions $\Gamma
+    \to \mathcal V_1$.
   \item A 2-cell from $\Phi$ to $R$ is a family $\Pi_{x\in X} \mathcal
-    V^X_h(\Phi x,R x)$ with composition given pointwise.
+    V^X_h(\Phi x,R x)$.
   \end{enumerate}
+  All operations are simply lifted pointwise.
 \end{definition}
 
+This defines the ``natural'' interpretation of SCT in a virtual
+equipment once we see that the powering construction is functorial in
+$\Gamma$.
 \begin{definition}[Internal Logic]
@@
   \begin{enumerate}
-  \item The indexing cateogry is $\Set$.
+  \item The indexing cateogry is $\Set$, which is a category with families.
   \item The equipment is defined as $\mathcal U\Gamma = \mathcal V^\Gamma$.
@@
     given pointwise.
+  \item The representing object for objects is simply
+    $\textrm{Ob}_\Gamma = \textrm{Ob}(\mathcal V)^\Gamma$ and
+    similarly for the other universe objects.
   \end{enumerate}
 \end{definition}
-
-%% \section{Conservativity of Homs, Tensors}
-
-%% The soundness theorem is not entirely sufficient for our goals,
-%% however.
-%% %
-%% In particular, we are interested in applying the internal category
-%% constructions to models that only have units, and do not necessarily
-%% admit left/right homs or tensor products of horizontal arrows.
-%% %
-%% To do so, we now prove a \emph{conservativity} result for the calculus
-%% with homs and tensors over a simpler calculus whose only profunctor
-%% constructor is the unit.
-%% %
-%% We must take care in defining the more restricted calculus, however.
-%% %
-%% If we simply remove homs and tensors from the calculus, then there is
-%% no way to internalize the transformations as types, and so no way to
-%% actually define internal categories!
-%% %
-%% For example, if we remove the hom sets, there is no way to abstract
-%% over the composition transformation:
-%% \[\alpha_1:\cat C, f : \harrapp {\cat C_1}{\alpha_1}{\alpha_2}, \alpha_2:\cat C, g : \harrapp{\cat C_1}{\alpha_2}{\alpha_3}, \alpha_3:\cat C \vdash \textrm{compose} : \harrapp{\cat C_1}{\alpha_1}{\alpha_3}
-%% \]
-%% %
-%% So in addition to removing the hom sets and tensors, we add a
-%% ``jumbo'' combined form of the hom set and end type, described in
-%% \cref{fig:jumbo}.
-%% %
-%% This type directly internalizes a transformation judgment of arbitrary
-%% large context, as opposed to the end type which internalized only
-%% transformations from an ``empty'' context.
-%% %
-
-%% \begin{figure}
-%%   \begin{mathpar}
-%%     \inferrule*[right=JumboHomForm]
-%%     {\Gamma \vdash \Phi \isavectx \and \Gamma \pipe \underline \Phi \vdash R \isaSet}
-%%     {\Gamma \vdash \Phi \triangleright R}
-
-%%     \inferrule*[right=JumboHomI]
-%%     {\Gamma \pipe \Phi \vdash s : R}
-%%     {\Gamma \vdash \lambda \Phi. s : \Phi \triangleright R}
-
-%%     \inferrule*[right=JumboHomE]
-%%     {\Gamma \vdash M : \Phi \triangleright R\and
-%%       \Gamma \vdash \phi : \Psi \to \Phi}
-%%     {\Gamma \pipe \Psi \vdash M \triangleright \phi : R[\underline \phi]}
-
-%%     \inferrule*[right=Jumbo$\beta$]
-%%     {}
-%%     {\lambda \Phi. M \triangleright (\Phi) = M}
-
-%%     \inferrule*[right=Jumbo$\eta$]
-%%     {\Gamma \vdash M : \Phi \triangleright R}
-%%     {M = \lambda \Phi. M \triangleright \Phi}
-%%   \end{mathpar}
-%%   \caption{Jumbo End/Homs}
-%%   \label{fig:jumbo}
-%% \end{figure}
-
-%% %
-%% To demonstrate this, we prove a \emph{conservativity} result for the
-%% calculus with homs and tensors over a calculus containing only units,
-%% and a ``jumbo'' internalization
-
-A model of our theory is given by a virtual double category with
-restrictions, units, tensors and homs. We informally describe the
-semantics of each judgment now.
-
-A context $\Gamma \isadtctx$ denotes a set.
-%
-A set with empty boundary $\Gamma \pipe \cdot \vdash S \isaSet$ denotes a
-set for every $\gamma:\Gamma$.
-%
-The type $\Gamma \pipe \cdot \vdash \Sort \isaSet$ denotes the set of
-objects of the virtual equipment (for any $\gamma:\Gamma$).
-%
-A term $\Gamma \pipe \alpha:C \vdash A : D$ denotes a vertical arrow
-from $C \gamma$ to $D \gamma$ for every $\gamma:\Gamma$.
-%
-A term $\Gamma \pipe \alpha:C; \beta : D \vdash R \isaSet$ denotes for
-each $\gamma:\Gamma$ a horizontal arrow from $C\gamma$ to $D\gamma$.
-%
-A term $\Gamma \pipe \alpha:C \vdash R \isaSet$ denotes for each
-$\gamma$, a horizontal arrow from $C\gamma$ to itself.
-%
-A non-$\cdot$ context $\Gamma \vdash \Phi\isavectx$ denotes a
-``composable string'' of horizontal arrows from $d^-\Phi \gamma$ to
-$d^+ \Phi \gamma$.
-%
-A term $\Gamma \pipe \cdot \vdash M : S$ denotes a function from
-$\gamma :\Gamma$ to $S \gamma$ (section of the bundle blah blah).
-%
-A term $\Gamma \pipe \Phi \vdash M : R$ for non-empty $\Phi$ denotes a
-``globular'' 2-cell in the virtual equipment for every
-$\gamma:\Gamma$.
-
-Next, substitutions model the forms of composition in the system.
-Substitution $A[B/\alpha]$ denotes composition of vertical arrows
-(with variable being the identity). Substitution $R[A/\alpha;B/\beta]$
-denotes restriction of a horizontal arrow $R$ along the vertical
-arrows $A,B$. And substitution $M[\phi]$ denotes composition of
-2-cells, where some restrictions are of course involved, and again a
-variable is the identity (globular) 2-cell.
-
-Next, we would also like to prove some kind of \emph{completeness}
-theorem for our calculus, at least the part relevant to virtual
-equipments. One way to do this would be to prove a kind of
-parameterized completeness theorem, where we treat the $\Gamma$ like a
-signature, and show that the syntax is complete for every fixed
-$\Gamma$. This amounts to proving that for any fixed $\Gamma$, the
-judgments above present a virtual equipment with a trivial model
-(i.e., each judgment is interpreted as itself). This shows that the
-syntax is closed under all constructions that are present in every
-model.
-
-Here is the outline of how we expect we should be able to prove
-this. For any fixed $\Gamma$, the objects of our double category are
-the terms $\Gamma \vdash C : \Sort$. A vertical arrow from $C$ to $D$
-is a term $\Gamma\pipe \alpha:C \vdash A : D$. A horizontal arrow from
-$C$ to $D$ is a term $\Gamma \pipe \alpha:C;\beta:D \vdash R
-\isaSet$. From any composable string of such $R$s we can construct a
-context $\Gamma \vdash \Phi \isavectx$ and then a 2-cell from $\Phi$ to
-$R$ along $A,B$ is given by a term $\Gamma \pipe \Phi \vdash
-R[A/\alpha;B/\beta]$. Compositions and restriction are modeled by the
-appropriate notion of substitution and units/tensors/homs are modeled
-by the corresponding type formers. There are a lot of details to be
-verified, especially the equations.
 
@@
         &= (\phi\jnctx \psi)[\phi' \jnctx \psi'_1],t[\psi'_2]/x,a[d^+\psi'_2]/\alpha\\
-        &= (\phi[\psi'] \jnctx \psi[\psi']),t[\psi'_2]/x,a[d^+\psi'_2]/\alpha\\
-        &= \phi[\psi'] \jnctx (\psi[\psi'],t[\psi'_2]/x,a[d^+\psi'_2]/\alpha)\\
+        &= (\phi[\psi'] \jnctx \psi[\psi'_1]),t[\psi'_2]/x,a[d^+\psi'_2]/\alpha\\
+        &= \phi[\psi'] \jnctx (\psi[\psi'_1],t[\psi'_2]/x,a[d^+\psi'_2]/\alpha)\\
+        &= \phi[\psi'] \jnctx (\psi,t/x,a/\alpha)[\psi'_1\jnctx \psi'_2]
       \end{align*}
   \end{enumerate}
+  Again, this covers all cases by \cref{inversion}.
 \end{lemma}
@@
 
-%% \begin{figure}
-%%   \begin{mathpar}
-%%     \begin{array}{rcl}
-%%       x[\gamma] &=& \gamma(x)\\
-%%       J[\gamma][\delta] & = & J[\gamma[\delta]]\\
-%%       \gamma[\delta](x) &=& \gamma(x)[\delta]\\
-
-%%       A[\alpha/\alpha] &=& A\\
-%%       \alpha[A/\alpha] &=& A\\
-%%       A[B/\beta][C/\gamma] & = & A[B[C/\gamma]/\beta]\\
-%%       A[B/\beta][\gamma] &=& A[\gamma][B[\gamma]/\beta]\\\\
-
-%%       R[\cdot] &=& R\\
-%%       R[\alpha/\alpha] &=& R\\
-%%       R[\alpha/\alpha,\beta/\beta] &=& R\\\\
-%%       R[\xi][\xi'] & = & R[\xi[\xi']]\\
-%%       R[\xi][\gamma] &=& R[\gamma][\xi[\gamma]]\\\\
-
-%%       \cdot[\xi] &=& \cdot\\
-%%       (A/\alpha)[B/\beta] &=& A[B/\beta]/\alpha\\
-%%       (A/\alpha;B/\beta)[\xi] &=& A[d^-\xi]/\alpha,B[d^+\xi]/\beta\\
-%%       \cdot[\gamma] &=& \cdot\\
-%%       (A/\alpha)[\gamma] &=& A[\gamma]/\alpha\\
-%%       (A/\alpha;B/\beta)[\gamma] &=& A[\gamma]/\alpha;B[\gamma]/\beta\\\\
-
-%%       \cdot[\gamma] &=& \cdot\\
-%%       (A/\alpha)[\gamma] &=& A[\gamma]/\alpha\\
-%%       (\phi_1,M/x,\phi_2)[\gamma] &=& \phi_1[\gamma],M[\gamma]/x,\phi_2[\gamma]
-
-%%       TODO: cat substitution id/assoc
-%%     \end{array}
-%%   \end{mathpar}  
-%%   \caption{Algebraic Properties}
-%%   \label{fig:equations}
-%% \end{figure}
+\section{Soundness}
+
+We next show how to uniquely interpret the profunctor contexts and
+substitutions in an arbitrary HED.
+
+\begin{proof}
+  We sketch only the interpretation of profunctor contexts in $C_1$,
+  as the rest is a straightforward translation from type theoretic to
+  categorical syntax.  We define the translation of a context
+  $\sem{\Phi}$ by induction on $\Phi$:
+
+  \begin{align*}
+    \sem{\alpha:\cat C} &= \sem{\id^h(\cat C)}\\
+    &= \id^h(\sem{\cat C})\\\\
+    \sem{\Phi , x:R, \beta:\cat C}\\
+    &= \sem{\Phi \circ_h i(R)}\\
+    &= \sem{\Phi} \sem{\Phi} \circ_h i(\sem{R})
+  \end{align*}
+  As these derivations show every step was uniquely determined by the
+  homomorphism property. Preservation of identity follows by
+  definition. Preservation of composition $\sem{\Phi \jnctx \Psi} =
+  \sem{\Phi} \circ_h \sem{\Psi}$ follows by induction on $\Psi$:
+  \begin{align*}
+    \sem{\Phi \jnctx (\alpha:\cat C)} &= \sem{\Phi} = \sem{\Phi} \circ_h \sem{\id(C)}\\
+    \sem{\Phi \jnctx (\Psi,x:R,\alpha:\cat C)}
+    &= \sem{(\Phi \jnctx \Psi),x:R,\alpha:\cat C} \\
+    &= \sem{(\Phi \jnctx \Psi)} \circ_h i(\sem{R})\\
+    &= (\sem{\Phi} \circ_h \sem{\Psi}) \circ_h i(\sem{R})\\
+    &= \sem{\Phi} \circ_h (\sem{\Psi} \circ_h i(\sem{R}))\\
+    &= \sem{\Phi} \circ_h \sem{\Psi, x:R, \alpha:\cat C}
+  \end{align*}
+  The interpretation of transformation substitutions is analogous.
+\end{proof}
 

### 2022-05-19 paper/paper.tex
some changes to the semantics section

@@
 \begin{definition}
-  A strict double category is a monoid in the double category $Span(Cat)$.
+  A strict double category is a monoid in the bicategory $Span(Cat)$.
 \end{definition}
@@
   \begin{enumerate}
-  \item The functor $(s,t) : C_d \to C_0$ is a split fibration, with
-    restriction operation written as $(f,g)^*R$.
+  \item The functor $(s,t) : C_d \to C_0 \times C_0$ is a split
+    fibration, with restriction operation written as $(f,g)^*R$.
   \item The profunctor $C_\vdash$ respects restrictions in that
@@
   \item A category $\textrm{Ctx}$
-  \item with a terminal object $1 \in \textrm{Ctx}$
-  \item A presheaf $\textrm{Ty} : \Set^\textrm{Ctx}$ of types in context.
-  \item A presheaf $\textrm{Tm} : \Set^{\int \textrm{Tm}}$ of terms of
+  \item with a terminal object $1 : \textrm{Ctx}$
+  \item A presheaf $\textrm{Ty} : \widehat {\textrm{Ctx}}$ of types in context.
+  \item A presheaf $\textrm{Tm} : \widehat{\int\textrm{Ty}}$ of terms of
     a given type.
@@
   Let $(\textrm{Ctx},\textrm{Ty},\textrm{Tm},\textrm{ext})$ be a
-  category with families and let $\mathcal V$ be a DCD internal to
+  category with families and let $\mathcal V = (C,C_d)$ be a DCD internal to
   $\Set^\textrm{Ctx}$. Then
@@
     natural isomorphism
-    \[ \textrm{Tm}(!_\Gamma^*\textrm{Ob}) \cong \textrm{Ob}(\mathcal V)_\Gamma \]
+    \[ \textrm{Tm}(!_\Gamma^*\textrm{Ob}) \cong C_{0}(\Gamma) \]
   \item $\mathcal V$ vertical arrows are representable as types if
     there is a type
-    \[ \textrm{Tm}(((\gamma,\lceil a\rceil),\lceil b\rceil)^*\textrm{Functor} \cong
-    \textrm{VArr}(\mathcal V_{\Delta})(a,b)
+    \[ \textrm{Tm}(Functor)(\Gamma)(a,b) \cong C_0(\Gamma)(a,b)
     \]
     and similarly horizontal displays are representable as types if
-    \[ \textrm{Tm}(((\gamma,\lceil a\rceil),\lceil b\rceil)^*\textrm{Profunctor} \cong
-    \textrm{Disp}(\mathcal V_{\Delta})(a,b)
+    \[ \textrm{Tm}(Prof)(\Gamma)(a,b) \cong \{ R \in C_d(\Gamma) \,|\, s(R) = a \wedge t(R) = b \}
     \]
   \item Closed $\mathcal V$ transformations are representable as types if
-    \[ \textrm{Tm}() \]
+    \[ \textrm{Tm}(\textrm{End})(\Gamma)(a,R) \cong C_\vdash^{\id,\id}(r(a),R) \]
   \end{enumerate}

### 2022-05-25 paper/paper.tex
stop abusing the word "simple"

@@
 {\cat C} a {a'}$. This semantically is a ``restriction'' of a
-profunctor along two functors, and modeling this operation simply as a
+profunctor along two functors, and modeling this operation as a
 substitution considerably simplifies reasoning using profunctors.
@@
 \begin{figure}
+  % TODO: formalize the one-place substitution
   \begin{mathpar}
@@
 %
-The basic idea is to simply reverse the orientation of all \emph{set}
+The basic idea is to reverse the orientation of all \emph{set}
 expressions by swapping which variable is covariant and which is
@@
 
-We can define this as a simple translation $\cdot^* : SCT \to SCT$
-defined on syntactic forms of SCT.
+We can define this as a translation $\cdot^* : SCT \to SCT$ defined on
+syntactic forms of SCT.
 
@@
 which has the intuitive reading that any of the VDC judgments $\Gamma
-\vdash J$ is interpreted as simply a family of $J$-data parameterized
+\vdash J$ is interpreted as a family of $J$-data parameterized
 by elements of the set $\Gamma$.
@@
 that their concatenation forms a strict category. Then the
-associativity/identity equations for a DCD are simply those of a
+associativity/identity equations for a DCD are precisely those of a
 double category.
@@
 Next, we can upgrade any DCD in $\Set$ to a DCD internal to
-$\Set^\Set$ by taking the presheaf of objects, vertical arrows, etc as
+$\Set^{\Set^o}$ by taking the presheaf of objects, vertical arrows, etc as
 families of objects, vertical arrows etc indexed by the indexing set
-$\Gamma$.
+$\Gamma$ as follows:
 \begin{definition}[Powers of Virtual Equipments]
@@
   \end{enumerate}
-  All operations are simply lifted pointwise.
+  All operations are lifted pointwise.
 \end{definition}

### 2022-05-25 paper/paper.tex
work on related work

@@
     given pointwise.
-  \item The representing object for objects is simply
-    $\textrm{Ob}_\Gamma = \textrm{Ob}(\mathcal V)^\Gamma$ and
-    similarly for the other universe objects.
+  \item The representing object for objects is $\textrm{Ob}_\Gamma =
+    \textrm{Ob}(\mathcal V)^\Gamma$ and similarly for the other
+    universe objects.
   \end{enumerate}
@@
 
-\paragraph{Calculi for Presheaves/Profunctors}
-
-David Jaz Myers String Diagrams
-
-Coend calculus
-
-Cartesian Bicategories.
-
-Double-cat-like syntax: Gradual Type Theory, System P of parametricity, cubical stuff?
-
-\paragraph{Directed Type Theory}
+\paragraph{Calculi for 2-dimensional category theory}
+
+\citet{djm-string-diagrams} provides a string diagram calculus for
+double categories and pro-arrow equipments, generalizing string
+diagrams for monoidal categories \citet{joyal?}. These are an
+alternative approach to type theoretic calculi, with the string
+diagrams typically making tensor products simpler to work with, while
+a type theoretic calculus like SCT makes the closed structure
+$\homrallXYtoZ \alpha P Q$ simpler to work with by using bound
+variables.
+
+One alternative to proarrow equipments is to axiomatize the bicategory
+of categories and profunctors rather than the double category of
+categories functors and profunctors. One abstract structure along
+these lines is a \emph{Cartesian bicategory}\footnote{which is not the
+same as a bicategory with products/finite limits, for which a syntax
+was developed \citet{fioresaville}} is given in
+\citet{carboni-walters,carboni-walters-someonelese}. Some preliminary
+work on a proof system for Cartesian bicateogires was presented at
+\citet{jonasfreycartesianbicategories}. This bears strong similarities
+to our calculus SCT, except that (1) their profunctors may have 0, 1
+or more covariant or contravariant variables but (2) they do not have
+a term syntax like our natural transformation syntax and (3) they do
+not have functors and restriction of profunctors by functors (since
+this structure is not present in a Cartesian
+bicategory). Semantically, Cartesian bicategories are somewhat less
+general than pro-arrow equipments because in general the set of
+functors between two categories can only be recovered from the set of
+profunctors if the codomain category is Cauchy complete, in which case
+functors are equivalent to adjoint pairs of profunctors.
+
+Our type-theoretic syntax for virtual double categories is intended to
+be a formal notation for the manipulation of profunctors. As such it
+has similarities to informal syntax for manipulating profunctors
+involving ends and coends, such as the coend calculus of
+\citet{coend-calculus}.
+
+Finally, some variations on double categories have been used to model
+the structure of certain program logics. \citet{newlicata20} present a
+logic for \emph{vertically thin} pro-arrow equipments, i.e., pro-arrow
+equipments where the functor and natural transformation judgments are
+posetal. This calculus is quite different from ours in that the
+horizontal morphisms are terms of a programming language, and so they
+are not presented as having co-variant and contravariant variables,
+and the equipment structure is modeled by explicit coercions rather
+than by syntactic substitution of vertical arrows in horizontal
+arrows. Another similar calculus is System P \citep{systemP} which is
+an internal language of \emph{reflexive graph categories}, which are
+like double categories without horizontal composition.
+
+Double-cat-like syntax:
+Gradual Type Theory, System P of parametricity, cubical stuff?
+
+\subsection{Directed Type Theory}
 
@@
 
-\paragraph{Category Theoretic Extensions}
+\subsection{Category Theoretic Extensions}
 
@@
 
-\paragraph{Semantics}
+\subsection{Semantics}
 

### 2022-05-25 paper/paper.tex
some "category theories"

@@
 categories, we can use SCT as a logical framework in which to do
-``formal type theory''.
+``formal type theory'' by axiomatizing the structure of the categories
+involved. The theories we can define are limited because we do not
+have access to opposite categories, but we can define some modal and
+monoidal structures.
+
+\begin{definition}
+  An \emph{SCT-signature} is a well-formed context $\Gamma \isadtctx$.
+\end{definition}
+
+\begin{lemma}
+  Any \emph{unary mode theory} can be defined as an SCT-signature as follows.
+  \begin{enumerate}
+  \item For each mode $m$ add an assumption $m : \Cat$
+  \item For each 1-morphism $f : m \to n$ add an assumption $R_f :
+    \harr m n$ and assumptions $F_f : \varr m n$ with and $G_f : \varr
+    n m$ with assumptions that they represent $R_f$.
+  \item For each 2-morphism $\alpha : f \Rightarrow g$ add an assumption $\alpha : R_f \multimap R_g$
+  \item Add each equality as an assumption using the identity type Id.
+  \end{enumerate}
+\end{lemma}
+
+This translation illustrates some ways the notion of SCT-signature is
+more flexible than a mode theory. For instance, in an SCT-signature we
+can add a profunctor assumption $R_f : \harr m n$ without saying
+whether it is representable as $F_f$ or $G_f$, and we can add more
+complex 2-morphisms.
+
+\begin{definition}
+  The \emph{theory of a strict monoidal category} consists of
+  \begin{enumerate}
+  \item A category $C : \Cat$
+  \item A unit $I : \varr 1 C$
+  \item A tensor product $\otimes : \varr {(C \times C)} C$
+  \item Such that $\alpha:(C \times C) \times C \vdash (\pi_2\pi_1\alpha \otimes \pi_2\pi_1 \alpha) \otimes \pi_2 \alpha = \pi_2\pi_1\alpha \otimes (\pi_2\pi_1 \alpha \otimes \pi_2 \alpha)$
+    and $\alpha \otimes I() = \alpha = I() \otimes \alpha$
+  \end{enumerate}
+\end{definition}
+
+\begin{definition}
+  The theory of a cartesian strict monoidal category is an extension
+  of the theory of a strict monoidal category with representability
+  data
+  \[ \punitinXfromYtoZ {C} {\Gamma}{\Delta \otimes \Sigma} \cong (\punitinXfromYtoZ C \Gamma \Delta)\times (\punitinXfromYtoZ C \Gamma \Sigma) \]
+  and
+  \[ \punitinXfromYtoZ {C} {\Gamma} I \cong 1 \]
+\end{definition}
+
+\begin{definition}
+  The theory of a simple type theory is an extension of the theory of
+  a strict cartesian monoidal category with a
+  \begin{enumerate}
+  \item A category $\text{Ty}$ of types
+  \item A profunctor $\vdash : \harr C {\textrm{Ty}}$
+  \item A functor $\text{var} : \varr{\textrm{Ty}} C$ representing $\vdash$ :
+    \[ \Gamma \vdash A \cong \punitinXfromYtoZ {} \Gamma {\text{var} A} \]
+  \item As assumption that $\text{var}$ is fully faithful:
+     \[ \punitinXfromYtoZ {} {\text{var} A} {\text{var} B} \cong \punitinXfromYtoZ {} A B \]
+  \end{enumerate}
+\end{definition}
+
+\begin{definition}
+  The theory of CBPV is an extension of the theory of a simple type
+  theory $(\text{VCtx},\times,1,\text{VTy},\vdash_V,\text{var}_V)$ with
+  \begin{enumerate}
+  \item A category $\text{CCtx}$
+  \item An object $I : \varr 1 {\text{CCtx}}$
+  \item An action of $\text{VCtx}$ on $\text{CCtx}$, i.e.,
+    \begin{enumerate}
+    \item A functor $\oslash : \text{VCtx}\times \text{CCtx} \to \text{CCtx}$
+    \item Such that $1 \oslash \underline\Delta = \underline\Delta$
+    \item And $(\Gamma \times \Delta) \oslash \underline\Sigma = \Gamma \oslash \Delta \oslash \underline\Sigma$
+    \end{enumerate}
+  \item A category $\text{CTy}$ with a profunctor $\vdash_C : \harr
+    {\text{CCtx}}{\text{CTy}}$ and a fully faithful functor $\text{var}_C$ that represents it
+    \[ \underline\Delta \vdash_C \underline B \cong \punitinXfromYtoZ {} {\underline\Delta}{\text{var}_C\underline B}\]
+  \end{enumerate}
+  Furthermore we can add in the universal properties of any
+  non-contravariant CBPV connectives. For instance:
+  \begin{enumerate}
+  \item $F : \varr {VTy} {CTy}$ satisfying
+    \[ \Gamma \oslash \text{var}_C F A \vdash_C \underline B \cong \Gamma \times \text{var}_V A \oslash I \vdash_C \underline B  \]
+    \[ \Gamma \vdash_V U \underline B \cong \Gamma \oslash I \vdash_C \underline B  \]
+  \end{enumerate}
+\end{definition}
 

### 2022-05-27 paper/paper.tex

@@
 
+In this section we establish the most significant portion of the
+completeness proof:
+\begin{construction}[Double category completeness]
+  For any $\Gamma$, define a strict double category as follows
+  \begin{enumerate}
+  \item Objects are categories $\Gamma \vdash \cat C \isaCat$
+  \item Vertical morphisms from $\cat C$ to $\cat D$ are functor/open
+    object terms $\Gamma \pipe\alpha:\cat C \vdash B : \cat D$ modulo
+    $\alpha$-equivalence.
+  \item Horizontal morphisms from $\cat C$ to $\cat D$ are profunctor
+    contexts $\Gamma\vdash \Phi \isavectx$ with $d^-\Phi = \cat C$ and
+    $d^+\Phi = \cat D$ modulo $\alpha$-equivalence.
+  \item Squares from $\Psi$ to $\Phi$ along $A$ and $B$ are
+    substitutions $\Gamma \vdash \phi : \Psi \to \Phi$ with $d^-\phi =
+    A$ and $d^+\phi = B$ modulo $\alpha$-equivalence.
+  \end{enumerate}
+\end{construction}
+\begin{proof}
+  The details are in the linked lemmas. Here we give an overview of
+  the relevant structure and properties.
+  \begin{enumerate}
+  \item Vertical composition is given by substitution of object terms
+    and is associative and unital with the variable as the identity
+    (Lemma \ref{lem:vertical-cat}).
+  \item Horizontal composition is given by the gluing of contexts
+    $\Phi\jnctx \Psi$ and is associative and unital with the single
+    category variable as the identity (Lemma
+    \ref{lem:horizontal-cat}).
+  \item Vertical composition of substitutions is given by substitution
+    $\phi[\psi]$ and is associative and unital with the identity
+    substitution given by a substitution of variables for themselves
+    (Lemma \ref{lem:vertical-sq-cat}).
+  \item Horizontal composition of substitutions is given by gluing of
+    substitutions $\phi \jnctx \psi$ and is associative and unital
+    with the identity substitution given by a single functor/open
+    object substitution (Lemma \ref{lem:horizontal-sq-cat}).
+  \item The process of constructing horizontal identity substitutions
+    is functorial in that the horizontal identity of a vertical
+    identity arrow is the vertical identity
+    (Lemma~\ref{lem:hor-vert-id-id}) and the vertical composition of
+    horizontal identities is a horizontal identity
+    (Lemma~\ref{lemma:lem-hor-sq-id}).
+  \item The process of horizontal composition of substitutions is
+    functorial in that the horizontal composition of vertical identity
+    squares is the identity (Lemma~\ref{lem:lem-vert-sq-id}) and the
+    interchange law holds (Lemma~\ref{lem:interchange}).
+  \end{enumerate}
+\end{proof}
+
+\begin{lemma}[Vertical category]
+  \label{lem:vertical-cat}
+  Substitution of object terms $A[B/\beta]$ is associative with
+  identity given by the variable $\alpha:\cat C \vdash \alpha:\cat C$.
+\end{lemma}
+
 \begin{definition}
+  \label{lem:horizontal-cat}
   We define composition of transformation contexts $\Phi \jnctx \Psi$ by induction on $\Psi$
@@
   \end{align*}
+  By induction, this is seen to be associative
+  \[ (\Phi \jnctx \Psi) \jnctx \Sigma = \Phi \jnctx (\Psi \jnctx \Sigma) \]
+  and unital with identity for $\cat C$ given by the single category
+  variable context $\alpha:\cat C$
+  \[ \alpha:\cat C \jnctx \Phi = \Phi = \Phi \jnctx \beta:\cat D\]
+  when $d^-\Phi = \alpha:\cat C$ and $d^+\Phi = \beta:\cat D$.
 \end{definition}
 
-It follows by an inductive argument that this definition is
-associative and unital.
-
 \begin{definition}
+  \label{lem:horizontal-sq-cat}
   We define the horizontal composition of transformation substitutions
@@
   \end{align*}
+  By induction, this is seen to be associative
+  \[ \phi \jnctx (\psi \jnctx \sigma) = (\phi \jnctx \psi) \jnctx \sigma \]
+  and unital with identity given by a single object substitution
+  $a/\alpha : \alpha:\cat C \to \beta:\cat D$
+  \[ a/\alpha \jnctx \phi = \phi = \phi \jnctx b/\beta \]
+  when $d^-\phi = a/\alpha$ and $d^+\phi = b/\beta$.
 \end{definition}
 
-It follows by an inductive argument that this operation is associative
-and unital.
-
+To define vertical composition of squares, we need the following lemma.
 \begin{lemma}[Inversion]
@@
 \begin{definition}
+  \label{lem:vertical-sq-cat}
   We define the vertical composition of transformation substitutions
@@
   \end{align*}
+
+  By induction this is seen to be associative:
+  \[ \phi[\psi][\sigma] = \phi[\psi[\sigma]] \]
+  and unital
+  \[ \id_{\Phi}[\phi] = \phi = \phi[\id_{\Psi}] \]
 \end{definition}
-By an inductive argument, this operation is associative and unital.
+
+\begin{lemma}
+  \label{lem:sq-vid-hid}
+  For any $\cat C$, $\alpha/\alpha : \alpha:\cat C \to \alpha:\cat C$
+  is both a horizontal and vertical identity.
+\end{lemma}
+
+\begin{lemma}
+  \label{lem:sq-vcomp-hid}
+  For any $\beta:\cat D \vdash a : \cat C$ and $\alpha:\cat C \vdash a': \cat C'$
+  the substitution
+  \[ (a'/\alpha')[a/\alpha] : \beta:\cat D \to \alpha':\cat C' \]
+  satisfies
+  \[ (a'/\alpha')[a/\alpha] = a'[a/\alpha]/\alpha' \]
+\end{lemma}
+
+\begin{lemma}
+  \label{lem:sq-vid-hcomp}
+  For any $\Phi$ and $\Psi$ with $\Phi\jnctx \Psi$ well-defined,
+  \[ \id_\Phi \jnctx \id_\Psi = \id_{\Phi\jnctx\Psi} \]
+  follows by induction.
+\end{lemma}
 
 \begin{lemma}{Interchange}
+  \label{lem:sq-vcomp-hcomp}
   We prove the interchange law

### 2022-08-30 paper/paper.tex
lncs format (very wide margins?)

@@
-\documentclass[a4paper,UKenglish,cleveref, autoref, thm-restate]{lipics-v2021}
+\documentclass{llncs}
 
@@
 \usepackage{amsmath}
-\usepackage{amsthm}
+%% \usepackage{amsthm}
 \usepackage{mathtools}
@@
 \DeclareMathOperator*{\colim}{colim}
-
-%This is a template for producing LIPIcs articles. 
-%See lipics-v2021-authors-guidelines.pdf for further information.
-%for A4 paper format use option "a4paper", for US-letter use option "letterpaper"
-%for british hyphenation rules use option "UKenglish", for american hyphenation rules use option "USenglish"
-%for section-numbered lemmas etc., use "numberwithinsect"
-%for enabling cleveref support, use "cleveref"
-%for enabling autoref support, use "autoref"
-%for anonymousing the authors (e.g. for double-blind review), add "anonymous"
-%for enabling thm-restate support, use "thm-restate"
-%for enabling a two-column layout for the author/affilation part (only applicable for > 6 authors), use "authorcolumns"
-%for producing a PDF according the PDF/A standard, add "pdfa"
-
-%\pdfoutput=1 %uncomment to ensure pdflatex processing (mandatatory e.g. to submit to arXiv)
-%\hideLIPIcs  %uncomment to remove references to LIPIcs series (logo, DOI, ...), e.g. when preparing a pre-final version to be uploaded to arXiv or another public repository
-
-%\graphicspath{{./graphics/}}%helpful if your graphic files are in another directory
-
-\bibliographystyle{plainurl}% the mandatory bibstyle
-
-\title{A Simple Type Theory for Category Theory} %TODO Please add
-
-%\titlerunning{Dummy short title} %TODO optional, please use if title is longer than one line
-
-\author{Max S. New}
-       {University of Michigan Computer Science \& Engineering, Ann Arbor, Michigan, USA \and \url{http://www.maxsnew.com} }
-       {maxsnew@umich.edu}
-       {}
-       {}
-       %% {https://orcid.org/0000-0002-1825-0097}
-       %% {(Optional) author-specific funding acknowledgements}%TODO mandatory, please use full name
-       %% ; only 1 author per \author macro; first two parameters are mandatory, other parameters ca
-       %% n be empty.
-Please provide at least the name of the affiliation and the country.
-The full address is optional.
-Use additional curly braces to indicate the correct name splitting when the last name consists of multiple name parts.
-
-\author{Daniel R. Licata}{Wesleyan University, Middletown, Connecticut, USA}
-       {dlicata@wesleyan.edu}{}{}
-
-\authorrunning{M.S. New and D.R. Licata} %TODO mandatory. First: Use abbreviated first/middle names. Second (only in severe cases): Use first author plus 'et al.'
-
-\Copyright{Max S. New and Daniel R. Licata} %TODO mandatory, please use full first names. LIPIcs license is "CC-BY";  http://creativecommons.org/licenses/by/3.0/
-
-\ccsdesc[100]{\textcolor{red}{Replace ccsdesc macro with valid one}} %TODO mandatory: Please choose ACM 2012 classifications from https://dl.acm.org/ccs/ccs_flat.cfm
-
-\keywords{Dummy keyword} %TODO mandatory; please add comma-separated list of keywords
-
-\category{} %optional, e.g. invited paper
-
-\relatedversion{} %optional, e.g. full version hosted on arXiv, HAL, or other respository/website
-%\relatedversiondetails[linktext={opt.
-text shown instead of the URL}, cite=DBLP:books/mk/GrayR93]{Classification (e.g.
-Full Version, Extended Version, Previous Version}{URL to related version} %linktext and cite are optional
-
-%\supplement{}%optional, e.g. related research data, source code, ... hosted on a repository like zenodo, figshare, GitHub, ...
-%\supplementdetails[linktext={opt.
-text shown instead of the URL}, cite=DBLP:books/mk/GrayR93, subcategory={Description, Subcategory}, swhid={Software Heritage Identifier}]{General Classification (e.g.
-Software, Dataset, Model, ...)}{URL to related version} %linktext, cite, and subcategory are optional
-
-%\funding{(Optional) general funding statement \dots}%optional, to capture a funding statement, which applies to all authors.
-Please enter author specific funding statements as fifth argument of the \author macro.
-
-%% \acknowledgements{I want to thank \dots}%optional
-
-%\nolinenumbers %uncomment to disable line numbering
-
-%Editor-only macros:: begin (do not touch as author)%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\EventEditors{John Q. Open and Joan R. Access}
-\EventNoEds{2}
-\EventLongTitle{42nd Conference on Very Important Topics (CVIT 2016)}
-\EventShortTitle{CVIT 2016}
-\EventAcronym{CVIT}
-\EventYear{2016}
-\EventDate{December 24--27, 2016}
-\EventLocation{Little Whinging, United Kingdom}
-\EventLogo{}
-\SeriesVolume{42}
-\ArticleNo{23}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 
+\title{A Simple Type Theory for Category Theory}
+\author{Max S. New\inst{1,2} and Daniel R. Licata\inst{2}}
+\institute{University of Michigan \and Wesleyan University}
+
 \maketitle
@@
 %
+
+%% Embracing the analogy ``profunctors as judgments'' we can view
+%% elements of a profunctor $r \in R$ with $s(r) = A$ and $t(r) = B$ in
+%% type-theoretic notation
+%% \[ \inferrule*[right={r}]{~}{A \vdash_R B}\]
+%% Noting that every category $\cat C$ has a hom profunctor $Hom_C : \cat
+%% C \pto \cat D$, we get for each category a judgment
+%% \[ \inferrule*{}{A \vdash_{Hom_C} A'} \]
+%% and the category operations provide us with rules for use of the assumption:
+%% \[ \inferrule*[right=id]{}{A \vdash_{Hom_C} A} \]
+%% and cut
+%% \[ \inferrule*[right=cut]{A \vdash_{Hom_C} A' \and A' \vdash_{Hom_C} A''}{A \vdash_{Hom_C} A''} \]
+%% satisfying unit and associativity equations.
+%% %
+%% An arbitrary profunctor $R$ supports a kind of ``bi-cut''
+%% rule\footnote{alternatively, this can be defined as 2 cuts: left and
+%% right, with an associativity law}:
+%% \[ \inferrule*[right=bicut]{A \vdash_{Hom_C} A' \and A' \vdash_R B' \and B' \vdash_{Hom_D} B}{A \vdash_R B} \]
+%% That satisfies similar bi-unit and bi-associativity laws.
 

### 2022-09-01 paper/paper.tex
a table comparing with higher order logic

@@
 
-%TODO mandatory: add short abstract of the document
 \begin{abstract}
@@
   transformations. By restricting the type theory to a form of ordered
-  logic, we guarantee that all mappings between categories are
-  functorial and all transformations are natural by construction, with
-  no separate proof necessary. Important category theoretic proofs
-  such as the Yoneda lemma and Co-yoneda lemma become simple type
-  theoretic proofs about the relationship between unit, tensor and
-  (ordered) function types. We demonstrate that this type theory can
-  be used as a kind of logical framework in which we can define common
-  categorical structures such as categories with families that serve
-  as models of type theories.
+  higher-order logic, we guarantee that all mappings between categories
+  are functorial and all transformations are natural by construction,
+  with no separate proof necessary. Important category theoretic
+  proofs such as the Yoneda lemma and Co-yoneda lemma become simple
+  type theoretic proofs about the relationship between unit, tensor
+  and (ordered) function types.
 
@@
 recursion and internal categories have been used to model
-parametricity and gradual typing.
+parametricity and gradual
+typing\cite{order-enriched-cats,topos-of-trees,reflexive-graphs,double-cats-gradual-typing}.
 
@@
 the Yoneda lemma, often the first ``big'' proof in an introductory
-category theory text. ... Here the reader is left to themselves to
-show that the ``evaluation at identity'' morphism and its inverse are
-actually natural transformations. They will then work through a
-diagram chase to determine that this naturality condition indeed
-holds.
+category theory text. The Yoneda lemma states that for any functor $P
+: C^o \to \Set$ and object $A \in C$, that the set of natural
+transformations $C(-,A) \to P$ is in bijection with the elements
+$P(A)$. Furthermore, this bijection is natural in both $A$ and $P$. In
+the authors' experience teaching category theory the sheer amount of
+naturality side conditions in this result can be very intimidating to
+learners, while experienced practitioners will observe that these
+naturality conditions are tedious to carry out but routine. In our
+type theory, none of these naturality conditions need to be explicitly
+proven. Instead, all terms expressible in the type theory are
+automatically natural by construction.
 
@@
 functors is functorial, a composition of natural transformations is
-natual, etc.
+natural, etc.
 %
 In the case of category theory, these principles can be formalized
-using category-like algebraic structures such as 2-categories,
-bicategories, (virtual) double categories and pro-arrow
-equipments\cite{}.
+using algebraic structures such as 2-categories, bicategories,
+(virtual) double categories and pro-arrow equipments\cite{}.
 %
@@
 foundations of the system.
+
+In Figure~\ref{fig:hol-cats} we give a table summarizing the
+relationship between the judgments and connectives of higher-order
+logic with our category logic. Since category logic is an ordered
+variant of higher-order logic, some notions splinter into
+multiple. For instance, covariant and contravariant presheaf
+categories generalize the power set. Further, because we only have
+binary relations rather than relations of arbitrary arity, we have
+only restricted forms of universal and existential quantification
+which combined with implications and conjunctions.
+
+\begin{figure}
+  \begin{center}
+    \begin{tabular}{ |c|c| } 
+      \hline
+      Higher-Order Logic & Category Logic \\
+      \hline
+      Set $X$ & Category $\mathbb{C}$\\
+      $X \times Y$ & $\mathbb C \times \mathbb D$ \\ 
+      1 & $1$ \\
+      $\mathcal P X$ & $\mathcal P^+ X, \mathcal P^- X$ \\
+      $\{ (x,y) \in X \times Y | R(x,y) \}$ & $\sum_{\alpha:C;\beta:D} R$\\
+      Function $f(x:X) : Y$ & Functor $F(\alpha:\cat C): \cat B$\\
+      Relation $R(x,y)$ & Profunctor $R(\alpha:\cat C; \beta: \cat D)$\\
+      $R \wedge Q$ & $R \times Q$\\
+      $\top$ & $1$\\
+      $\forall x. P \Rightarrow Q$ & $\homrallXYtoZ {\alpha:\cat C} P Q, \homlallXYtoZ {\alpha:\cat C} P Q$ \\
+      $\exists x. P \wedge Q$ & $\tensorexistsXwithYandZ {\alpha:\cat C} P Q$\\
+      $x =_X y$ & $\punitinXfromYtoZ {\cat C} \alpha \beta$\\
+      Proof $R_1 \wedge \cdots \Rightarrow Q$ & Natural Transformation $x_1: R_1,\ldots \vdash t : Q$\\
+      \hline
+    \end{tabular}
+  \end{center}
+  \caption{Analogy between Higher-Order Logic and Category Logic
+    Judgments and Connectives}
+\end{figure}
 
@@
 
-\subsection{Category of Elements}
-
-One final multiplicative we can define is the \emph{category of
-elements} of a profunctor. Given a profunctor $\alpha:\cat
-C;\beta:\cat D \vdash R$, the ``category of elements'' of $R$ is a
-category $\Sigma_{\alpha:\cat C;\beta:\cat D}. R$. In detail, its
-objects are triples $(a:\cat C, r : R, b : \cat D)$ and morphism from
-$(a,r,b)$ to $(a', r', b')$ consists of a pair of morphisms $g : b \to
-b'$ and a backwards $f : a' \to a$ such that $r' = g\circ r \circ f$.
+\subsection{Graph of a profunctor}
+
+One final multiplicative we can define is the \emph{graph} of a
+profunctor. Given a profunctor $\alpha:\cat C;\beta:\cat D \vdash R$,
+the ``graph'' of $R$ is a category $\Sigma_{\alpha:\cat C;\beta:\cat
+  D}. R$. In detail, its objects are triples $(a:\cat C, r : R, b :
+\cat D)$ and morphism from $(a,r,b)$ to $(a', r', b')$ consists of a
+pair of morphisms $g : b \to b'$ and a backwards $f : a' \to a$ such
+that $r' = g\circ r \circ f$.
+%
+Intuitively this generalizes the ``graph'' of a relation $R : A\times
+B \to \Omega$, where we think of the graph as the corresponding subset
+$gr(R) \subseteq A\times B$
 %
@@
 composing these with functors, we recover the notion of \emph{comma}
-category $\exists \alpha;\beta. \punitinXfromYtoZ {\cat
+category $\Sigma \alpha;\beta. \punitinXfromYtoZ {\cat
   E}{(F\alpha)}{(G\beta)}$ and by picking either $F$ or $G$ to have
@@
 This is unique among the multiplicatives in that it defines a
-\emph{category} rather than a profunctor. Also note that while similar
-to the tensor in that we use existential notation, the graph of a
-profunctor is a \emph{negative} multiplicative, analogous to a
-$\Sigma$-type. The universal property says that a functor from $\cat
-E$ into the graph $\exists (\alpha; \beta). R$ is equivalent to a
+\emph{category} rather than a profunctor. The graph of a profunctor is
+a \emph{negative} multiplicative, a kind of $\Sigma$-type for
+profunctors. The universal property says that a functor from $\cat E$
+into the graph $\Sigma (\alpha; \beta). R$ is equivalent to a
 transformation from $\cat E$ into the functor $R$, including functors
 from $\cat E$ to $\cat C$ and $\cat D$.
+%
+The graph of a profunctor is a joint generalization of several common
+constructions in category theory: the arrow category, slice
+categories, comma categories and the covariant and the covariant and
+contravariant category of elements.
 
@@
 %% %% Please use bibtex, 
-
-\bibliography{lipics-v2021-sample-article}
+\subsubsection*{Acknowledgements}
+Grants stuff...? The authors would like to thank Emily Riehl, Mike
+Shulman,...for helpful feedback on this work
+
+\bibliography{cats}
 

### 2022-09-02 paper/paper.tex
work on shortened synthetic cats section

@@
 %% \newtheorem{definition}{Definition}
+\newcommand{\ohol}{OHol}
 \newcommand{\sem}[1]{\llbracket{#1}\rrbracket}
@@
 \newcommand{\hombinary}[6]{\pendallXdotY {#1} {\homrallXYtoZ {#3} {#2} \homrallXYtoZ {#5} {#4} {#6}}}
+\newcommand{\homtrinary}[8]{\pendallXdotY {#1} {\homrallXYtoZ {#3} {#2} \homrallXYtoZ {#5} {#4} \homrallXYtoZ {#7} {#6} {#8}}}
+
+\newcommand{\homlunary}[4]{\pendallXdotY {#1} {\homrallXYtoZ {#3} {#2} {#4}}}
+\newcommand{\homlbinary}[6]{\pendallXdotY {#1} {\homrallXYtoZ {#3} {#2} \homrallXYtoZ {#5} {#4} {#6}}}
+\newcommand{\homltrinary}[8]{\pendallXdotY {#1} {\homrallXYtoZ {#3} {#2} \homrallXYtoZ {#5} {#4} \homrallXYtoZ {#7} {#6} {#8}}}
 
@@
 \newcommand{\hlambda}[2]{\lambda^{P}{#1};{#2}.}
+\newcommand{\ran}[1]{\textrm{Ran}_{#1}\,}
 
@@
 \newcommand{\negPresheafApp}[2]{#1\,#2}
+\newcommand{\negPresheafAppPtoX}[2]{#2\,#1}
 \newcommand{\posPresheafApp}[2]{#1\,#2}
@@
 
-\title{A Simple Type Theory for Category Theory}
+\title{An Ordered Logic for Category Theory}
 \author{Max S. New\inst{1,2} and Daniel R. Licata\inst{2}}
@@
   category, functor, profunctor and a generalized form of natural
-  transformations. By restricting the type theory to a form of ordered
-  higher-order logic, we guarantee that all mappings between categories
-  are functorial and all transformations are natural by construction,
-  with no separate proof necessary. Important category theoretic
-  proofs such as the Yoneda lemma and Co-yoneda lemma become simple
-  type theoretic proofs about the relationship between unit, tensor
-  and (ordered) function types.
-
-  Finally, we give a sound and complete notion of model of our type
-  theory in a structure we call a hyperdoctrine of pro-arrow
-  equipments with displays, a generalization of prior notions of
-  virtual equipments that have been shown to model enriched and
-  internal category theory. This means that while the proofs in our
-  type theory look like standard set-based arguments, the
-  substructural syntactic restrictions ensure that the proofs carry
-  over to enriched and internal settings as well.
+  transformations. By restricting the type theory to an ordered
+  linear restriction of higher-order logic, we guarantee that all
+  mappings between categories are functorial and all transformations
+  are natural by construction, with no separate proof
+  necessary. Important category theoretic proofs such as the Yoneda
+  lemma and Co-yoneda lemma become simple type theoretic proofs about
+  the relationship between unit, tensor and (ordered) function types,
+  and can be seen to be ordered refinements of theorems in
+  intuitionistic higher-order logic.
+
+  We give a sound and complete notion of model of our type theory in a
+  structure we call a hyperdoctrine of pro-arrow equipments with
+  displays, which are known to model enriched and internal category
+  theory. This means that while the proofs in our type theory look
+  like standard set-based arguments, the substructural syntactic
+  restrictions ensure that the proofs carry over to enriched and
+  internal settings as well.
 \end{abstract}
@@
 meta level.
+
+\subsection{Ordered Logic of Functions and (Constructive) Relations}
+
+Our logic is based on four basic notions: categories, functors,
+profunctors and (multi-argument) natural transformations.
+%
+These generalize the basic notions already present in formulations of
+higher-order logic: sorts, functions, relations and proofs.
+%
+In fact, one way to view our logic is that it is an \emph{ordered},
+\emph{constructive} higher-order logic
+
+While profunctors are not typically employed in introductory category
+theory texts,
 
@@
 which combined with implications and conjunctions.
-
-\begin{figure}
-  \begin{center}
-    \begin{tabular}{ |c|c| } 
-      \hline
-      Higher-Order Logic & Category Logic \\
-      \hline
-      Set $X$ & Category $\mathbb{C}$\\
-      $X \times Y$ & $\mathbb C \times \mathbb D$ \\ 
-      1 & $1$ \\
-      $\mathcal P X$ & $\mathcal P^+ X, \mathcal P^- X$ \\
-      $\{ (x,y) \in X \times Y | R(x,y) \}$ & $\sum_{\alpha:C;\beta:D} R$\\
-      Function $f(x:X) : Y$ & Functor $F(\alpha:\cat C): \cat B$\\
-      Relation $R(x,y)$ & Profunctor $R(\alpha:\cat C; \beta: \cat D)$\\
-      $R \wedge Q$ & $R \times Q$\\
-      $\top$ & $1$\\
-      $\forall x. P \Rightarrow Q$ & $\homrallXYtoZ {\alpha:\cat C} P Q, \homlallXYtoZ {\alpha:\cat C} P Q$ \\
-      $\exists x. P \wedge Q$ & $\tensorexistsXwithYandZ {\alpha:\cat C} P Q$\\
-      $x =_X y$ & $\punitinXfromYtoZ {\cat C} \alpha \beta$\\
-      Proof $R_1 \wedge \cdots \Rightarrow Q$ & Natural Transformation $x_1: R_1,\ldots \vdash t : Q$\\
-      \hline
-    \end{tabular}
-  \end{center}
-  \caption{Analogy between Higher-Order Logic and Category Logic
-    Judgments and Connectives}
-\end{figure}
 
@@
 
-\section{Synthetic Category Theory}
-
-Next, we look at synthetic category theory constructions in SCT.
-
-\subsection{Profunctor Isomorphisms}
-
-We start our study of synthetic category theory constructions with
-some isomorphisms that show the interaction between the 5
-multiplicative connectives: right/left hom, end, unit and tensor.
-
-First, we fix some terminology
+\section{Synthetic Category Theory in Ordered Higher-order Logic}
+
+To demonstrate that \ohol{} provides a setting in which to carry out
+synthetic category theory, we demonstrate some basic definitions and
+theorems in category theory in \ohol{}. We emphasize throughout the
+interpretations as tautologies in higher order-logic, as well as the
+relation to ordered linear logic.
+
+First, we fix some terminology: TODO, isomorphism of profunctors, anything else we'll need?
+%% \begin{definition}{Isomorphism of Profunctors}
+
+%% \end{definition}
+
+We can witness the fact that all constructions are (pro-)functorial by
+using the hom type $\punitinXfromYtoZ {\cat C}{\alpha}{\alpha'}$.
 \begin{lemma}
-  Let $P, Q : \harr {\cat C} {\cat D}$.
-  A homomorphism from $P$ to $Q$ is a term of type
-  \[ \homunary {\alpha} {\harrapp P \alpha \beta} {\beta} {\harrapp Q \alpha \beta} \]
-  Given homomorphisms $M : P \multimap Q$, $N : Q \multimap R$ we can define their composite as
-  \[ (N \circ M) = \lambdaunary {\alpha} {x:\harrapp P \alpha \beta} \beta {\appunary \alpha N \beta {\appunary \alpha M \beta x}} \]
-  This notion of composition is associative and unital using the
-  $\beta\eta$ equations for the end/right hom connectives, with unit
-  $\lambdaunary {\alpha} x \beta x$.
-
-  An isomorphism between $P, Q : \harr {\cat C}{\cat D}$ is then a
-  homomorphism $P \multimap Q$ with an inverse $Q \multimap P$.
-\end{lemma}
-
-Our first synthetic theorem is that we have that the unit, i.e., the
-hom set, is the unit for tensor and hom.
-%
-The theorem for hom is the synthetic \emph{Yoneda lemma} and the
-theorem for tensor is the synthetic \emph{Co-Yoneda lemma}.
-\begin{lemma}[{Yoneda and Co-Yoneda}]
-  \label{lem:yo-and-co-yo}
-  Let $P : \harr {\cat C} {\cat D}$.
-
-  For any $\gamma : \cat E \vdash b : \cat D$, the following are naturally isomorphic to $\harrapp P \alpha b$:
   \begin{enumerate}
-  \item $\homrallXYtoZ {\beta} {(\punitinXfromYtoZ {D} {\beta} {b})} {\harrapp P {\alpha} {\beta}}$
-  \item $\tensorexistsXwithYandZ {\beta} {\harrapp P \alpha {\beta}} {(\punitinXfromYtoZ {\cat D} {\beta} {b})}$
-  \end{enumerate}
-
-  For any $\gamma : \cat E \vdash a : \cat C$, the following are naturally isomorphic to $\harrapp P a \beta$:
-  \begin{enumerate}
-  \item $\homlallXYtoZ {{\alpha}} {(\punitinXfromYtoZ {C} {\alpha} {a})} {\harrapp P {{\alpha}} {\beta}}$
-  \item $\tensorexistsXwithYandZ {\alpha} {(\punitinXfromYtoZ {C} {a} {\alpha})} {\harrapp P {\alpha}{\beta}}$
+  \item Identity: $\pendallXdotY {\alpha} {\punitinXfromYtoZ {\cat C}{\alpha}{\alpha'}}$
+  \item Composition: $\hombinary {\alpha_1} {\punitinXfromYtoZ {}{\alpha_1}{\alpha_2}} {\alpha_2} {\punitinXfromYtoZ {}{\alpha_1}{\alpha_2}} {\alpha_3} {\punitinXfromYtoZ {}{\alpha_1}{\alpha_3}}$
+  \item Functoriality: for any $F : \varr {\cat C}{\cat D}$, $\homunary {\alpha_1} {\punitinXfromYtoZ {}{\alpha_1}{\alpha_2}} {\alpha_2} {\punitinXfromYtoZ {} {F(\alpha_1)}{F(\alpha_2)}}$.
+   \item Profunctoriality: for any $R : \harr {\cat C}{\cat D}$ $\homtrinary {\alpha_1} {\punitinXfromYtoZ{} {\alpha_1}{\alpha_2}} {\alpha_2} {R{\alpha_2}{\beta_2}} {\beta_2} {\punitinXfromYtoZ{} {\beta_2} {\beta_1}} {\beta_1} {R\alpha_1\beta_1}$
   \end{enumerate}
 \end{lemma}
-\begin{proof}
-  TODO: what's better, build the substitution in or prove the
-  substitution can be built in afterwards? The one building in the
-  substitution obscures the connection to the normal tensor/hom/unit
-  relationship, but the one with the built in subst is obviously more
-  general.
-\end{proof}
-
-Next, we have various associativities between the homs, tensor and end
-constructions. Semantically, these are related to so-called
+Reflexivity and transitivity generalize the corresponding properties
+of equality, with the lack of symmetry being a key feature of the
+generalization.
+%
+In addition, we can prove that the (pro)-functoriality axioms commute
+with the composition proof by the $\eta$ principle for the unit.
+%
+(Pro-)Functoriality generalizes the statement that all functions and
+relations respect equality.
+
+Naturality is more complex to state, and it is a statement about the
+\emph{proofs} so it has no analogue in ordinary higher-order
+logic. The following version is stated for any \emph{profunctor}, with
+the usual case of naturality arising when $R \alpha \beta =
+\punitinXfromYtoZ {\cat C}{F \alpha}{G \beta}$.
+\begin{lemma}[Naturality]
+  For any $t : \pendallXdotY {\alpha:\cat C}{\harrapp R \alpha
+    \alpha}$, by composing with profunctoriality, we can construct
+  terms
+  \[ \alpha_1, f : \punitinXfromYtoZ {\cat C} {\alpha_1} {\alpha_2}, \alpha_2 \vdash \textrm{lcomp}(f,\pendappXtoY t {\alpha_2}) \quad \textrm{and} \quad \textrm{rcomp}(\pendappXtoY t {\alpha_1}, f) : \harrapp R {\alpha_1}{\alpha_2}\]
+  These are both equal to $\punitelimtoYwithkontZ f {t}$.
+\end{lemma}
+
+Next, we turn to some of the central theorems of category theory, the
+Yoneda and Co-Yoneda lemmas. Despite being ulimately quite elementary,
+these are notoriously abstract. In \ohol{}, we can view these as
+ordered generalizations of some very simple tautologies about
+equality. For instance, the Yoneda lemma (and its proof) generalizes
+the equivalence between the formulae $\forall x. \forall y. x = y
+\Rightarrow P y$ and $P x$.
+\begin{lemma}
+  Let $\alpha^o : \cat C$ and $\pi : \negPresheaf {\cat C}$. Then
+  \begin{enumerate}
+  \item (Yoneda) The profunctor $\homrallXYtoZ {\alpha'} {\punitinXfromYtoZ {\cat C} {\alpha'}{\alpha}} {\negPresheafAppPtoX \pi {\alpha'}}$ is isomorphic to $\negPresheafAppPtoX \pi \alpha$
+  \item (Co-Yoneda) The profunctor $\tensorexistsXwithYandZ {\alpha'}{\punitinXfromYtoZ{}{\alpha}{\alpha'}} {\negPresheafAppPtoX \pi \alpha}$ is isomorphic to $\negPresheafAppPtoX \pi \alpha$
+  \end{enumerate}
+\end{lemma}
+
+Next, we have the ``Fubini'' theorems. These correspond to quantifier
+re-ordering and associativity of conjunction and implication in
+higher-order logic.
 \emph{Fubini} theorems about interchange of (co)-ends.
@@
   \label{lem:fubini}
-  The following isomorphisms hold for any $P : \harr {\cat C}{\cat D},
-  P' : \harr {\cat C}{\cat D} Q :\harr {\cat D}{\cat E}, R : \harr
-  {\cat E} {\cat F}, S : \harr {\cat C} {\cat E}$. where $\alpha:\cat
-  C, \beta:\cat D, \gamma : \cat E, \delta: \cat F$
+  The following isomorphisms hold when the corresponding profunctors
+  are well typed.
   \begin{enumerate}
@@
 \end{lemma}
-\begin{proof}
-  TODO
-\end{proof}
-\subsection{Composition, Functoriality, Profunctoriality, Naturality}
-
-Hom sets don't provide primitive syntax for composition of morphisms
-in a category. As our first synthetic construction, we can
-\emph{define} composition using the elimination form for the unit.
-%
-\begin{definition}[Composition of Morphisms]
-  We define composition of morphisms as the following term of type $\Pi_{\cat C:\Cat}\hombinary {\alpha_1} {\punitinXfromYtoZ {\cat C}{\alpha_1} {\alpha_2}} {\alpha_2} {\punitinXfromYtoZ {\cat C}{\alpha_2} {\alpha_3}} {\alpha_3} {\punitinXfromYtoZ {\cat C}{\alpha_1} {\alpha_3}}$:
-  \[
-  \lambda \cat C.
-\punitelimtoYwithkontZ {\hlambda {\alpha_1}{\alpha_2} \homrallXYtoZ {\alpha_3} {\punitinXfromYtoZ {\cat C} {\alpha_2} {\alpha_3}} {\punitinXfromYtoZ {\cat C} {\alpha_1} {\alpha_3}}} {\lambdaunary {\alpha} {f : \punitinXfromYtoZ {\cat C}{\alpha} {\alpha_3}} {\alpha_3} f}
-  \]
-  We call this term $\textrm{comp}$
-\end{definition}
-This has the same flavor as a directed type theory: to construct the
-composite it is sufficient to define the action on the identity
-morphism, which is of course the identity for composition.
-%
-Note that we gave a slight bias to the definition to follow by
-induction on the \emph{first} function, whereas we could make the
-following alternative definition, naturally of type
-$\pendallXdotY {\alpha_2} \homrallXYtoZ {\alpha_3} {\punitinXfromYtoZ {\cat C}{\alpha_2} {\alpha_3}}
-{(\homlallXYtoZ {{\alpha_1}} {\punitinXfromYtoZ {\cat C}{\alpha_1} {\alpha_2}}{\punitinXfromYtoZ {\cat C}{\alpha_1} {\alpha_3}})}$:
-\[
-  \punitelimtoYwithkontZ {\hlambda {\alpha_2}{\alpha_3} \homlallXYtoZ {{\alpha_1}} {\punitinXfromYtoZ {\cat C}{\alpha_1} {\alpha_2}}{\punitinXfromYtoZ {\cat C}{\alpha_1} {\alpha_3}} } {\lambdaunary {\alpha} {g : \punitinXfromYtoZ {\cat C}{\alpha_1}{\alpha}} {\alpha} g}
-\]
-
-\begin{lemma}[Equivalence of Composition Definitions]
-  TODO
+Note that the last item here is different from the others in that it
+is an equivalence of sets rather than profunctors.
+
+Next, we turn to basic universal properties. While there are many
+equivalent definitions of a universal property in category theory, the
+one most natural in \ohol{} is the notion of a functor representing a
+profunctor, a mild generalization of the more common notion of
+representable presheaf. We say a functor $F : \varr {\cat C}{\cat D}$
+left represents a profunctor $R : \harr {\cat C}{\cat D}$ when
+$\punitinXfromYtoZ {}{\beta}{F \alpha} \cong \harrapp R \alpha
+\beta$. This is an ordered generalization of the notion of a relation
+$R$ being the \emph{graph} of a function $F$. Due to the asymmetry,
+there is a dual notion of a functor $G : \varr {\cat D}{\cat C}$
+right-representing a profunctor $\punitinXfromYtoZ {\alpha}{G\beta}
+\cong \harrapp R \alpha \beta$. Then $F$ and $G$ are adjoint functors
+when $F$ left represents a profunctor that $G$ right-represents. This
+generalizes the formulation that a function $g$ is an inverse to $f$
+if its corresponding relation is the converse of the relation
+corresponding to $f$.
+
+The definition we gave for adjoint functors is the one in terms of
+``hom-isomorphism''. The equivalent in terms of unit and co-unit is a
+simple corollary of the Yoneda lemma, because for instance the unit
+can be extracted from the transformation $\homunary \alpha
+{\punitinXfromYtoZ {}{F \alpha}{\beta}} \beta {\punitinXfromYtoZ
+  {}{\alpha}{G\beta}}$. The formulation in terms of unit and co-unit
+generalizes the formulation of inverses that each composition is equal
+to the identity.
+%% \begin{lemma}[Formulations of Representability]
+%%   The following are equivalent
+%%   \begin{enumerate}
+%%   \item An isomorphism between $\punitinXfromYtoZ {}{F \alpha}{\beta}$ and $\harrapp R \alpha \beta$
+%%   \item A term $\pendallXdotY \alpha {\harrapp R \alpha {F \alpha}}$
+%%     such that the induced morphism from $\punitinXfromYtoZ {}{F
+%%       \alpha}{\beta}$ to $\harrapp R \alpha \beta$ is an isomorphism.
+%%   \end{enumerate}
+%% \end{lemma}
+
+Next, we turn to limits, or more generally, Kan extensions.
+%
+Given functors $J : \varr {\cat C}{\cat C'}$ and $F : \varr{\cat
+  C}{\cat D}$, a right Kan extension is functor $\ran J F : \varr{\cat
+  C'}{\cat D}$ such that the composition $\ran J F \circ J$ has a
+universal natural transformation to $F$.
+\begin{lemma}
+  A right Kan extension of $F$ along $J$ is a functor $\ran J F$ that
+  right-represents the profunctor $\homrallXYtoZ {\alpha} {\punitinXfromYtoZ {} {\alpha'}{J \alpha}}{\punitinXfromYtoZ {} \beta {F \alpha}}$
 \end{lemma}
-
-It should follow by $\beta\eta$ that this is associative and unital.
-
-Next, let's prove synthetically that every functor $F : \varr{\cat
-  C}{\cat D}$ is functorial in this notion of composition.
-\begin{definition}
-  For any $F : \varr {\cat C} {\cat D}$, we can define a morphism $F_1
-  : \homunary {\alpha_1} {\punitinXfromYtoZ {\cat C} {\alpha_1}
-    {\alpha_2}} {\alpha_2} {\punitinXfromYtoZ {\cat D} {F \alpha_1}
-    {F \alpha_2}}$ as
-  \[ F_1 = \punitelimtoYwithkontZ {\hlambda {\alpha_1}{\alpha_2} \punitinXfromYtoZ {\cat D} {F \alpha_1}{F \alpha_2}}
-  {\pendlambdaXdotY \alpha \punitrefl {\cat D} {F \alpha}}
-  \]
-  Note that
-  \[ A[\id_\alpha/\alpha] = \id_{A} \]
-  \[ A[f;g] = A[f];A[g] \]
-\end{definition}
-Note the similarity to the proof that all functions are congruences in
-MLTT/HoTT.
-
-And similarly, every profunctor $R : \harr {\cat C}{\cat D}$ is
-\emph{pro-functorial}.
-\begin{definition}
-  For any $R : \varr {\cat C} {\cat D}$, we can define left and right
-  actions of composition:
-  \[ \homunary {\alpha_1} {\tensorexistsXwithYandZ {\alpha_2} {(\punitinXfromYtoZ {\cat C} {\alpha_1}{\alpha_2})} {\tensorexistsXwithYandZ {\alpha_3} {(\harrapp R {\alpha_2} {\alpha_3})}}{(\punitinXfromYtoZ {\cat C} {\alpha_3}{\alpha_4})}} {\alpha_4} {\harrapp R {\alpha_1}{\alpha_4}}
-  \]
-  Furthermore we can show that they are functorial in that the induced functions
-  TODO
-  are equal
-\end{definition}
-
-Next we can define natural transformations as specific kinds of terms
-whose type is $\pendallXdotY {\alpha} \punitinXfromYtoZ {\cat D} {F
-  \alpha}{G \alpha}$. In fact we can generalize this to any
-endo-profunctor $P$:
-\begin{definition}[Natural element]
-  A natural element of an endo-profunctor $P : \harr {\cat C}{\cat C}$
-  is a term of type $\pendallXdotY {\alpha} \harrapp P \alpha \alpha$.
-
-  A natural transformation is the special case where $P =
-  \punitinXfromYtoZ {\cat D} {F \alpha}{G \alpha}$ for functors $F, G
-  : \varr {\cat C}{\cat D}$.
-
-  This is natural in the sense that the incuded terms
-  \[ \homunary {\alpha_1}{\punitinXfromYtoZ {} {\alpha_1}{\alpha_2}} {\alpha_2} {\harrapp P {\alpha_1}{\alpha_2}}  \]
-  and
-  \[ \pendallXdotY {\alpha_2} \homlallXYtoZ {\alpha_1} {\punitinXfromYtoZ {} {\alpha_1}{\alpha_2}} {\harrapp P {\alpha_1}{\alpha_2}}\]
-
-  are equal under the associativity equivalence between these two
-  types.
-\end{definition}
-
-\subsection{Yoneda Lemma and Representable Presheaves}
-
-Given any category $\cat C$, we can define the positive and negative
-Yoneda embeddings $Y^\pm : \cat C \to \pmPresheaf {\cat C}$ using the
-hom profunctor. Note that both of these functors are covariant in
-$\cat C$, so semantically, $\posPresheaf {\cat C}$ should be
-interpreted as $(\Set^{\cat C})^o$.
-
-\begin{definition}
-  Define the negative Yoneda embedding $Y^-[\alpha]$
-  \[ \alpha:\cat C \vdash Y^-[\alpha] = \lambda \alpha':\cat C. \punitinXfromYtoZ {\cat C}{\alpha'}{\alpha} : \negPresheaf {\cat C} \]
-
-  The positive Yoneda embedding is defined similarly $Y^+[\alpha]$.
-\end{definition}
-
-The Yoneda lemma says that the set of natural transformations from
-$Y^-[\alpha]$ to a presheaf $p$ is naturally isomorphic to the
-elements of the presheaf itself: $\negPresheafApp \alpha p$. In
-encoding this statement in our type theory, we find that we can use
-two different interpretations of ``natural transformations''.
-
-On the one hand, a natural transformation should be a morphism in the
-presheaf category. This interpretation of the Yoneda lemma would mean
-that we have a natural isomorphism of profunctors:
-\[ \alpha:\cat C; \pi:\negPresheaf {\cat C} \vdash \punitinXfromYtoZ{\negPresheaf {\cat C}} {Y^-[\alpha]} {\pi} \cong \negPresheafApp \alpha \pi \]
-Attempting to prove this we become stuck very quickly because we know
-nothing about the structure of morphisms in the presheaf category.
-
-On the other hand, a natural transformation of presheaves $\pi$ and
-$\rho$ can be defined as a family of functions from $\negPresheafApp
-\alpha \pi$ to $\negPresheafApp \alpha \rho$, or in other words as an
-\emph{end}. Specializing to the Yoneda lemma, we interpret ``a natural
-transformation from $Y^-[\alpha]$ to $\pi$'' as the following, which simplifies using $\beta$ reduction:
-\[ \alpha:\cat C; \pi:\negPresheaf {\cat C} \vdash \homrallXYtoZ {\alpha':\cat C} {\negPresheafApp {\alpha'} {Y^-[\alpha]}}{\negPresheafApp {\alpha'} \pi} = \homrallXYtoZ {\alpha':\cat C} {(\punitinXfromYtoZ {\cat C}{\alpha'}{\alpha})}{(\negPresheafApp {\alpha'} \pi)} \]
-And then the Yoneda lemma becomes:
-\[ \alpha:\cat C; \pi:\negPresheaf {\cat C} \vdash \homrallXYtoZ {\alpha':\cat C} {(\punitinXfromYtoZ {\cat C}{\alpha'}{\alpha})}{(\negPresheafApp {\alpha'} \pi)} \cong \negPresheafApp \alpha \pi\]
-Which is an instance of our earlier Lemma\,\ref{lem:yo-and-co-yo}
-which follows quite easily.
-
-Then we are left with the question of whether these two forms of the
-Yoneda lemma are equivalent, or slightly more generally are these two definitions of natural transformation of presheaves equivalent:
-\[ \rho:\negPresheaf{\cat C}; \pi:\negPresheaf {\cat C} \vdash \punitinXfromYtoZ{\negPresheaf {\cat C}} {\rho} {\pi} \]
-and
-\[ \rho:\negPresheaf{\cat C}; \pi:\negPresheaf {\cat C} \vdash \homrallXYtoZ {\alpha:\cat C} {\negPresheafApp \alpha \rho}{\negPresheafApp \alpha \pi} \]
-While we can show that there is a morphism from the hom set to the
-end, the inverse does not seem provable, and seems analogous to the
-\emph{univalence} axiom in homotopy type theory, as it tells us that
-the morphisms of the universe are given by a particular kind of
-function in our logic.
-
-\subsection{Adjoint Functors and Representable Profunctors}
-
-Next, we demonstrate that the two common definitions of adjoint
-functors are equivalent: isomorphism of hom-sets and units and
-co-units.
-%
-The proof goes through most easily, and utilizes duality most
-effectively, if we factorize the construction through an intermediate
-notion of when a functor \emph{represents} a profunctor.
-
-\begin{definition}[Functor Represents a Profunctor]
-  A functor $G : \varr{\cat D}{\cat C}$ right-represents a profunctor $R :
-  \harr {\cat C}{\cat D}$ when there is an isomorphism of profunctors:
-  \[ \punitinXfromYtoZ {\cat C} {\alpha} {G \beta} \cong \harrapp R {\alpha}{\beta}\]
-
-  We say a functor $F : \varr{\cat C}{\cat D}$ \emph{left-represents} a profunctor $R : \harr {\cat C}{\cat D}$ when there is an isomorphism of profunctors
-  \[ \punitinXfromYtoZ {\cat D} {F \alpha} {\beta} \cong \harrapp R {\alpha}{\beta}\]
-\end{definition}
-
-This is a parameterized generalization of the notion of an object
-representing a functor $R : C^{op} \to Set$.
-These notions are precisely dual. $F$ left-represents a profunctor $R$
-precisely if it right-represents it in the opposite SCT.
-
-Note that trivially any functor $G$ right-represents
-$\hlambda\alpha\beta \punitinXfromYtoZ {\cat D}{\alpha} {G \beta}$ and
-dually, any functor $F$ left-represents
-$\hlambda\alpha\beta\punitinXfromYtoZ {\cat C}{F\alpha}{\beta}$.
-
-Then we can show that any right-representation isomorphism is
-determined by a co-unit:
-\begin{lemma}[(Co)-unit of a Representation]
-  If $G: \varr {\cat D}{\cat C}$ right-represents $R : \harr {\cat
-    C}{\cat D}$, then there is a natural element, the co-unit:
-  \[ \pendallXdotY \beta \harrapp R {G\beta} \beta \]
-  Such that the half of the isomorphism 
-  \[ \punitinXfromYtoZ {\cat C} {\alpha} {G \beta} \multimap \harrapp R {\alpha}{\beta} \]
-  is defined by composition with the natural element.
-
-  Dually, any $F$ left-representing isomorphism with $R$ is determined
-  by a ``unit'' of type $\pendallXdotY \alpha \harrapp R {\alpha}
-  {F\alpha}$.
-\end{lemma}
-\begin{proof}
-  A homomorphism $\punitinXfromYtoZ {\cat C} {\alpha} {G \beta} \multimap \harrapp R {\alpha}{\beta}$ is by definition a term of type
-  \[ \pendallXdotY \alpha \homrallXYtoZ {\beta} {(\punitinXfromYtoZ {\cat C} \alpha {G \beta})} {\harrapp R \alpha \beta} \]
-  By \cref{lem:fubini}, this type is isomorphic to 
-  \[ \pendallXdotY \beta \homlallXYtoZ {\alpha} {(\punitinXfromYtoZ {\cat C} \alpha {G \beta})} {\harrapp R \alpha \beta} \]
-  By \cref{lem:yo-and-co-yo}, this type is isomorphic to
-  \[ \pendallXdotY \beta {{\harrapp R {(G \beta)} \beta}} \]
-
-  Therefore this half of the isomorphism is determined by a natural
-  element. The construction in the proof of \cref{lem:yo-and-co-yo} is
-  indeed by composition (i.e., unit elimination).
-
-  The $F$ case is precisely dual.
-\end{proof}
-
-Combining these two, we get the unit-counit definition for adjoint
-functors
-
-\begin{definition}[Adjunction by unit/co-unit]
-  We say $G : \varr {\cat D}{\cat C}$ is right adjoint to $F : \varr
-  {\cat C}{\cat D}$ if there is a unit
-  \[ \eta : \pendallXdotY \alpha \punitinXfromYtoZ \alpha {G(F(\alpha))}\]
-  and a co-unit
-  \[ \epsilon : \pendallXdotY \beta \punitinXfromYtoZ {F(G(\beta))} \beta \]
-
-  satisfying the triangle identities...
-  \[ \id_{R[\beta]} = R[\epsilon_{\beta}] \circ \eta_{R[\beta]} \]
-  \[ \id_{L[\alpha]} = \epsilon_{L[\alpha]} \circ L[\eta_{\alpha}] \]
-\end{definition}
-
-\begin{theorem}[Adjunctions]
-  $G$ is right adjoint to $F$ if and only if there is a natural isomorphism
-  $\punitinXfromYtoZ {\cat D} {(F \alpha)} \beta \cong \punitinXfromYtoZ {\cat D} \alpha {G \beta}$
+Viewing this in higher-order logic, this says that $\beta = (\ran J
+F)\alpha'$ if and only if $\forall \alpha. \alpha' = J \alpha
+\Rightarrow \beta = F \alpha$. If $J$ is surjective this says that
+$\ran J F$ composed with $J$ is equal to $F$, but if $J$ is not
+surjective then this is impossible to satisfy for ordinary
+functions. Kan extensions generalize limits in that the limit of a
+diagram $F : \varr {\cat C}{\cat D}$ is the same as a right Kan
+extension along the unique $! : \varr{\cat C}{1}$.
+
+Then we can prove the famous theorem that right adjoints preserve
+right Kan extensions, and so in particular, limits.
+\begin{theorem}
+  If $\ran J F$ is a right Kan extension and $G : \varr{\cat D}{\cat
+    E}$ has a left adjoint $L$, then $G \circ (\ran J F)$ is the right
+  Kan extension of $G \circ F$ along $J$.
 \end{theorem}
 \begin{proof}
-  First, by \cref{lem:co-unit}, the forward direction of the
-  isomorphism is equivalent to providing a term
-  \[ \eta : \pendallXdotY \alpha \punitinXfromYtoZ \alpha {G(F(\alpha)) } \]
-
-  Second, the backward direction of the isomorphism is equivalent to
-  providing a term
-  \[ \epsilon : \pendallXdotY \beta \punitinXfromYtoZ {F(G(\beta))} \beta \]
-
-  And the triangle laws state exactly that the induced functions are
-  isomorphisms.
-\end{proof}
-
-\subsection{Kan Extensions/Lifts}
-
-As a further example of the synthetic approach to universal
-properties, we consider right and left \emph{Kan extensions}, which
-are generalizations of limits and colimits, respectively.
-
-The right Kan extension of $F$ along $J$ is the ``best'' (maximal) way to
-fill in the following square:
-\[% https://tikzcd.yichuanshen.de/#N4Igdg9gJgpgziAXAbVABwnAlgFyxMJZABgBpiBdUkANwEMAbAVxiRAGEQBfU9TXfIRQBGclVqMWbACLdeIDNjwEiZYePrNWiDgHJu4mFADm8IqABmAJwgBbJGRA4ISAEzVNUnQCk5lm-aIjs5IohJabABiBlxAA
-\begin{tikzcd}
-\cat C \arrow[d, "J"] \arrow[r, "F"] & \cat D \\
-\cat C'                              &  
-\end{tikzcd}\]
-I.e. a functor $\textrm{Ran}_J F : \cat C' \to \cat D$ with a
-universal natural transformation $\textrm{Ran}_J F \circ J \Rightarrow
-F$. A left kan extension $\textrm{Lan}_J F$ is the same but with a
-natural transformation $F \Rightarrow \textrm{Lan}_J F \circ J $.
-
-We can express the concept in SCT as representing a certain profunctor
-constructed using the profunctor hom:
-\begin{definition}[Kan Extensions]
-  Given functors $F : \varr {\cat C}{\cat D}$ and $J : \varr {\cat
-    C}{\cat C'}$, a right Kan extension is a functor $\textrm{Ran}_J F
-  : \varr {\cat C'}{\cat D} $that right-represents the profunctor
-  \[ \hlambda {\beta:\cat D}{\alpha': \cat C} \homrallXYtoZ {\alpha}{(\punitinXfromYtoZ {\cat C'}{\alpha'}{J\alpha})} {\punitinXfromYtoZ {\cat D}{\beta}{F \alpha}} \]
-
-  A left Kan extensions is a functor $\textrm{Lan}_J F : \varr {\cat
-    C'}{\cat D}$ is a functor satisfying the dual property: it
-  \emph{left}-represents the profunctor
-  \[\hlambda {\alpha':\cat C}{\beta:\cat D} \homlallXYtoZ {\alpha}{(\punitinXfromYtoZ {\cat C'}{J\alpha}{\alpha'})} {\punitinXfromYtoZ {\cat D}{F \alpha}{\beta}}
-  \]
-\end{definition}
-
-The limit of a functor $F : \varr{\cat C}{\cat D}$ is a special case
-of a Kan extension where $J : \varr{\cat C}{1}$ is the unique functor
-to the terminal category. Similarly, the colimit of $F$ is the left
-Kan extension with respect to the same $J$.
-%
-Specializing to this case, the representability for limits means:
-\[ \punitinXfromYtoZ {\cat D}{\beta}{\textrm{lim} F} \cong \homrallXYtoZ {\alpha}{(\punitinXfromYtoZ {\cat 1}{\alpha'}{()})}{\punitinXfromYtoZ {\cat D}{\beta}{F \alpha}} \]
-And if we view a morphism in $\cat 1$ as being trivial, then this
-essentially states that mapping $\beta$ to $\textrm{lim} F$ is the
-same thing as constructing a cone: for all $\alpha :\cat C$, a
-morphism from $\beta$ to $F \alpha$, naturally in $\alpha$.
-
-The usual formulation of Kan extensions is in terms of a universal
-morphism.
-%
-We can reproduce this formulation using our established theory of
-equivalences for homs of profunctors.
-\begin{lemma}
-  One half of the representability isomorphism for a right Kan
-  extension $\textrm{Ran}_J F$ is determined by a natural
-  transformation
-  \[ \pendallXdotY \alpha \punitinXfromYtoZ {\cat D}{(\textrm{Ran}_JF)(J\alpha)}{F\alpha} \]
-
-  Dually, one half of the representability isomorphism for a left Kan
-  extension $\textrm{Lan}_J F$ is determined by a natural
-  transformation
-  \[ \pendallXdotY \alpha \punitinXfromYtoZ {\cat D}{F\alpha}{(\textrm{Lan}_JF)(J\alpha)} \]
-\end{lemma}
-\begin{proof}
-  By \cref{lem:unit-repr}, one half of representability is determined by a natural element
-  \[ \pendallXdotY {\alpha'} \homrallXYtoZ {\alpha}{(\punitinXfromYtoZ {\cat C'}{\alpha'}{J\alpha})} {\punitinXfromYtoZ {\cat D}{(\textrm{Ran}_J F)\alpha'}{F \alpha}} \]
-  which, by \cref{lem:fubini} is isomorphic to
-  \[
-  \pendallXdotY {\alpha} \homlallXYtoZ {\alpha'}{(\punitinXfromYtoZ {\cat C'}{\alpha'}{J\alpha})}{\punitinXfromYtoZ {\cat D}{(\textrm{Ran}_J F)\alpha'}{F \alpha}}
-  \]
-  which, by \cref{lem:yo-and-co-yo}, is isomorphic to 
-  \[ \pendallXdotY \alpha \punitinXfromYtoZ {\cat D}{(\textrm{Ran}_JF)(J\alpha)}{F\alpha} \]
-  by taking $P = \hlambda {\alpha'}{\alpha} {\punitinXfromYtoZ {\cat D}{(\textrm{Ran}_J F)\alpha'}{F \alpha}}$.
-
-  The proof for left Kan extensions follows by \cref{lem:duality}.
-\end{proof}
-
-With the definition based on representability, we can easily prove the
-generalization of the famous theorem, that right adjoints preserve not
-just limits but all right Kan extensions
-
-\begin{theorem}[Right/Left Adjoints preserve Right/Left Kan Extensions]
-  Let $K : \varr{\cat C}{\cat D}$ and $J : \varr{\cat C}{\cat C'}$.
-
-  If the right Kan extension $\textrm{Ran}_J K$ exists, and if $G :
-  \varr {\cat D}{\cat D'}$ has a left adjoint $F : \varr{\cat D'}{\cat
-    D}$, then the right Kan extension $\textrm{Ran}_J G\circ K$ exists
-  and is given by $G \circ \textrm{Ran}_J K$.
-
-  Similarly, if the left Kan extension $\textrm{Lan}_J K$ exists and
-  if $F : \varr{\cat D}{\cat D'}$ has a right adjoint $G : \varr{\cat
-    D'}{\cat D}$, then the left Kan extension $\textrm{Lan}_J F\circ
-  K$ exists and is given by $F \circ \textrm{Ran}_J K$.
-\end{theorem}
-\begin{proof}
-  By the composition of the following natural isomorphisms:
   \begin{align*}
-    \punitinXfromYtoZ {\cat D'} {\beta'} {G ((\textrm{Ran}_J K) \alpha')} &\cong \punitinXfromYtoZ {\cat D}{F \beta'} {(\textrm{Ran}_J K) \alpha'} \tag{Adjunction}\\
-    &\cong \homrallXYtoZ {\alpha}{(\punitinXfromYtoZ {\cat C'}{\alpha'}{J\alpha})} {\punitinXfromYtoZ {\cat D}{F \beta'}{K \alpha}} \tag{Ran assumption, \cref{lem:natural-restriction}}\\
-    &\cong \homrallXYtoZ {\alpha}{(\punitinXfromYtoZ {\cat C'}{\alpha'}{J\alpha})} {\punitinXfromYtoZ {\cat D}{\beta'}{G(K \alpha)}} \tag{Adjunction}\\
+    \punitinXfromYtoZ {} {\gamma} {G (\ran J F \alpha')}
+    &\cong \punitinXfromYtoZ {}{L \gamma} {\ran J F \alpha'}\\
+    &\cong \homrallXYtoZ {\alpha}{(\punitinXfromYtoZ {}{\alpha'}{J\alpha})} {\punitinXfromYtoZ {}{L \gamma}{F \alpha}}\\
+    &\cong \homrallXYtoZ {\alpha}{(\punitinXfromYtoZ {}{\alpha'}{J\alpha})} {\punitinXfromYtoZ {}{\gamma}{G (F \alpha)}}
   \end{align*}  
@@
 
-\subsection{(op)-Fibrations}
-
-A (Street) fibration can be defined in any 2-category. A functor $p :
-\varr {\cat E}{\cat B}$ is a fibration when the morphism $p \to \cat
-B/p$ in the slice 2-category over $\cat B$ has a right adjoint.
-
-Since we can define all of the components of this slice 2-category we
-can internalize this definition in SCT.
-
-\begin{enumerate}
-\item A functor $\textrm{wp} : \exists (\alpha:\cat B; \beta:\cat E). \punitinXfromYtoZ {\cat B} \alpha {p\beta}$
-\item In the slice category: $p(\textrm{wp}(k)) = \pi_-k$
-\item With a universal morphism $\epsilon : \pendallXdotY {k} \punitinXfromYtoZ {} {(\pi_-k,k,\punitrefl{}{\pi_-k})} {k}$
-\item In the slice category: $\pi_k \epsilon = \id$ (how to express this nicely in SCT?)
-\end{enumerate}
-
-We can also weaken this to a Street fibration if we require the more sensible $p(\textrm{wp}(k)) \cong \pi_-k$ only.
-
-\begin{lemma}{Fibrations compose}
-  If $F : \varr{\cat C}{\cat D}$ and $G : \varr{\cat D}{\cat E}$ are
-  (op)-fibrations then so is $G \circ F$.
-\end{lemma}
-\begin{proof}
-  By composition of adjoints.
-\end{proof}
-
-\begin{lemma}{Fibrations are Stable under Pullback}
-
-\end{lemma}
+\subsection{Ordered Theories}
+
+Just as we can define theories in first-order logic or higher-order
+logic by assuming base sorts, function and relation symbols and
+axioms, we can define ordered, or categorical theories in \ohol{}
+easily by using the two-level type theory structure.
+%
+We can define an \emph{ordered theory} to be any context $\Gamma$ and
+a translation of theories to be a substitution.
+%
+Then we can define theories for cartesian categories, monoidal
+categories, bicartesian categories, strong monads, (strong)
+adjunctions.
+%
+However the utility for applications to programming language semantics
+is limited due to the lack of opposite categories.
 
 \subsection{Duality Principle}
-
 One common source of category theoretic arguments is the duality
@@
 This has the effect of reversing the order of the context $\Phi$.
-
+%
 We can define this as a translation $\cdot^* : SCT \to SCT$ defined on
-syntactic forms of SCT.
-
+syntactic forms of SCT in Figure~\ref{fig:duality}.
 \begin{figure}
@@
     \end{align*}
-
   \caption{Duality Involution}
+  \label{fig:duality}
 \end{figure}
 
-TODO: example of applying the principle in our calculus.
-
-\subsection{Formal Type Theory}
-
-Since many type theories are given semantics by free structured
-categories, we can use SCT as a logical framework in which to do
-``formal type theory'' by axiomatizing the structure of the categories
-involved. The theories we can define are limited because we do not
-have access to opposite categories, but we can define some modal and
-monoidal structures.
-
-\begin{definition}
-  An \emph{SCT-signature} is a well-formed context $\Gamma \isadtctx$.
-\end{definition}
-
-\begin{lemma}
-  Any \emph{unary mode theory} can be defined as an SCT-signature as follows.
-  \begin{enumerate}
-  \item For each mode $m$ add an assumption $m : \Cat$
-  \item For each 1-morphism $f : m \to n$ add an assumption $R_f :
-    \harr m n$ and assumptions $F_f : \varr m n$ with and $G_f : \varr
-    n m$ with assumptions that they represent $R_f$.
-  \item For each 2-morphism $\alpha : f \Rightarrow g$ add an assumption $\alpha : R_f \multimap R_g$
-  \item Add each equality as an assumption using the identity type Id.
-  \end{enumerate}
-\end{lemma}
-
-This translation illustrates some ways the notion of SCT-signature is
-more flexible than a mode theory. For instance, in an SCT-signature we
-can add a profunctor assumption $R_f : \harr m n$ without saying
-whether it is representable as $F_f$ or $G_f$, and we can add more
-complex 2-morphisms.
-
-\begin{definition}
-  The \emph{theory of a strict monoidal category} consists of
-  \begin{enumerate}
-  \item A category $C : \Cat$
-  \item A unit $I : \varr 1 C$
-  \item A tensor product $\otimes : \varr {(C \times C)} C$
-  \item Such that $\alpha:(C \times C) \times C \vdash (\pi_2\pi_1\alpha \otimes \pi_2\pi_1 \alpha) \otimes \pi_2 \alpha = \pi_2\pi_1\alpha \otimes (\pi_2\pi_1 \alpha \otimes \pi_2 \alpha)$
-    and $\alpha \otimes I() = \alpha = I() \otimes \alpha$
-  \end{enumerate}
-\end{definition}
-
-\begin{definition}
-  The theory of a cartesian strict monoidal category is an extension
-  of the theory of a strict monoidal category with representability
-  data
-  \[ \punitinXfromYtoZ {C} {\Gamma}{\Delta \otimes \Sigma} \cong (\punitinXfromYtoZ C \Gamma \Delta)\times (\punitinXfromYtoZ C \Gamma \Sigma) \]
-  and
-  \[ \punitinXfromYtoZ {C} {\Gamma} I \cong 1 \]
-\end{definition}
-
-\begin{definition}
-  The theory of a simple type theory is an extension of the theory of
-  a strict cartesian monoidal category with a
-  \begin{enumerate}
-  \item A category $\text{Ty}$ of types
-  \item A profunctor $\vdash : \harr C {\textrm{Ty}}$
-  \item A functor $\text{var} : \varr{\textrm{Ty}} C$ representing $\vdash$ :
-    \[ \Gamma \vdash A \cong \punitinXfromYtoZ {} \Gamma {\text{var} A} \]
-  \item As assumption that $\text{var}$ is fully faithful:
-     \[ \punitinXfromYtoZ {} {\text{var} A} {\text{var} B} \cong \punitinXfromYtoZ {} A B \]
-  \end{enumerate}
-\end{definition}
-
-\begin{definition}
-  The theory of CBPV is an extension of the theory of a simple type
-  theory $(\text{VCtx},\times,1,\text{VTy},\vdash_V,\text{var}_V)$ with
-  \begin{enumerate}
-  \item A category $\text{CCtx}$
-  \item An object $I : \varr 1 {\text{CCtx}}$
-  \item An action of $\text{VCtx}$ on $\text{CCtx}$, i.e.,
-    \begin{enumerate}
-    \item A functor $\oslash : \text{VCtx}\times \text{CCtx} \to \text{CCtx}$
-    \item Such that $1 \oslash \underline\Delta = \underline\Delta$
-    \item And $(\Gamma \times \Delta) \oslash \underline\Sigma = \Gamma \oslash \Delta \oslash \underline\Sigma$
-    \end{enumerate}
-  \item A category $\text{CTy}$ with a profunctor $\vdash_C : \harr
-    {\text{CCtx}}{\text{CTy}}$ and a fully faithful functor $\text{var}_C$ that represents it
-    \[ \underline\Delta \vdash_C \underline B \cong \punitinXfromYtoZ {} {\underline\Delta}{\text{var}_C\underline B}\]
-  \end{enumerate}
-  Furthermore we can add in the universal properties of any
-  non-contravariant CBPV connectives. For instance:
-  \begin{enumerate}
-  \item $F : \varr {VTy} {CTy}$ satisfying
-    \[ \Gamma \oslash \text{var}_C F A \vdash_C \underline B \cong \Gamma \times \text{var}_V A \oslash I \vdash_C \underline B  \]
-    \[ \Gamma \vdash_V U \underline B \cong \Gamma \oslash I \vdash_C \underline B  \]
-  \end{enumerate}
-\end{definition}
-
-\begin{definition}[Theory of unary Computational lambda calculus with booleans]
-  \begin{enumerate}
-  \item A category $C$
-  \item A terminal object $1$ of $C$
-  \item A monad $T$ on $C$
-  \item A type $\textrm{Bit} : \varr 1 C$
-  \item Morphisms $\textrm{y},\textrm{n} : \pendallXdotY * \punitinXfromYtoZ {1(*)} {\textrm{Bit}(*)}$
-  \item A natural transformation $\textrm{if} : \pendallXdotY * \homrallXYtoZ \alpha {(\punitinXfromYtoZ C {1(*)} \alpha) \times (\punitinXfromYtoZ C {1(*)} \alpha)} {\punitinXfromYtoZ C {\textrm{Bit}(*)} \alpha}$
-  \item Satisfying $\textrm{if} y = \pi_1$ and $\textrm{if} n = \pi_2$
-  \item A morphism $\textrm{print} : \pendallXdotY * \punitinXfromYtoZ {\textrm{Bit}(*)}{T (1(*))}$
-  \end{enumerate}
-\end{definition}
-
-Then we can ``program'' in this computational lambda calculus within
-our logical framework by constructing terms of type $\punitinXfromYtoZ
-C {A} {B}$ for any functors $A,B$ into $C$.
-
-Using a terminal category and the category of elements, we can define
-a formal category with families, an algebraic formulation of the
-judgmental structure of dependent type theory:
-\begin{definition}[Category with Families]
-  A category with families consists of
-  \begin{enumerate}
-  \item A category $\textrm{Ctx}\isaCat$
-  \item A presheaf $\textrm{Ty} : \harr {\textrm{Ctx}}{1}$
-  \item A presheaf $\textrm{Tm} : \harr {\Sigma_{\Gamma:\textrm{Ctx};*:1}\harrapp{\textrm{Ty}} \Gamma *} 1$
-  \item A terminal object in $\textrm{Ctx}$
-  \item An \emph{extension operation}
-    $\textrm{ext} : \Sigma_{\Gamma:\textrm{Ctx};*:1}\harrapp{\textrm{Ty}}\Gamma * \to \textrm{Ctx}$
-    right representing the profunctor
-    \[ \Delta:\textrm{Ctx};A:\Sigma_{\Gamma:\textrm{Ctx};*:1}\harrapp{\textrm{Ty}}{\Gamma}{*} \vdash \tensorexistsXwithYandZ {a}{(\punitinXfromYtoZ {} \Delta {\pi^-\pi^-a})}
-       {(\punitinXfromYtoZ {}{\pi^-a}{A})}\]
-       % {\Sigma_{\Gamma:\textrm{Ctx};*:1}\harrapp{\textrm{Ty}}{\Gamma}{*}}
-    % 
-  \end{enumerate}
-\end{definition}
-
-%% Can we do pras?
-%%
-%% A PRA mode theory is given by a free 2-category
-%% 1. Each mode m is a CwF whose context cat is [m]
-%% 2. Each 1-cell μ : m -> n is a functor [μ] : [n] -> [m] that has a DRA and is a PRA
-%% 3. Each 2-cell α : μ => ν is a 2-cell
-
-%% Let C, D be CwFs.
-%% A DRA is a pair of
-%% 1. A functor  L : C -> D
-%% 2. A morphism R : L^*(D_Ty) -> C_ty in ∫C, i.e., a morphism taking LΓ |- A to Γ |- RA
-%% 3. An iso     I : L^*(D_tm)?? =~ R^*(C_tm) in ∫(L^*(DTy)) a morphism taking LΓ |- M : A to Γ |- I(M) : RA
-
-%% Let C, D such that C has a terminal object.
-%% G : C -> D is a PRA if
-%% G / 1 : C -> D / G1 is a right adjoint
-
-%% What about fixed points?
-%% Say C has a terminal object
-%% fix : Hom(a,a) -> Hom(1, a)
-%% satisfying 
-%% fix[f] = f o fix[f]?
-%% We can write down fix, but not the equation I think
+\section{Semantics}
+Next, we develop the basics of the model theory for \ohol{}. Being a
+2-level type theory, \ohol{} has models that are a kind of
+\emph{hyperdoctrine}. First, a model must include a model of
+intuitionistic type theory to interpret the type-level context
+$\Gamma$ and corresponding type/term judgments. Next, there should be
+a functor from the category of semantic contexts and substitutions to
+a category of models of ordered higher-order logic.
+
+\subsection{Soundness and Completeness}
+
+\subsection{Concrete Models}
+
+Because SCT
+includes not just a syntax for category theory, but is also indexed by
+an ordinary context of dependent type theory $\Gamma$, a model of SCT
+is not given by a \emph{single} virtual double category, but instead
+by an \emph{indexed family} of VDCs, parameterized by an object of a
+category that interprets dependent type theory. We call such a
+structure a \emph{VDC-hyperdoctrine}. We then show how to construct a
+from any VDC, a VDC-hyperdoctrine indexed by the category of sets,
+which has the intuitive reading that any of the VDC judgments $\Gamma
+\vdash J$ is interpreted as a family of $J$-data parameterized
+by elements of the set $\Gamma$.
 
@@
 \end{proof}
+
+\section{Extended Synthetic Category Theory}
+
+Next, we look at synthetic category theory constructions in SCT.
+
+\subsection{Profunctor Isomorphisms}
+
+We start our study of synthetic category theory constructions with
+some isomorphisms that show the interaction between the 5
+multiplicative connectives: right/left hom, end, unit and tensor.
+
+First, we fix some terminology
+\begin{lemma}
+  Let $P, Q : \harr {\cat C} {\cat D}$.
+  A homomorphism from $P$ to $Q$ is a term of type
+  \[ \homunary {\alpha} {\harrapp P \alpha \beta} {\beta} {\harrapp Q \alpha \beta} \]
+  Given homomorphisms $M : P \multimap Q$, $N : Q \multimap R$ we can define their composite as
+  \[ (N \circ M) = \lambdaunary {\alpha} {x:\harrapp P \alpha \beta} \beta {\appunary \alpha N \beta {\appunary \alpha M \beta x}} \]
+  This notion of composition is associative and unital using the
+  $\beta\eta$ equations for the end/right hom connectives, with unit
+  $\lambdaunary {\alpha} x \beta x$.
+
+  An isomorphism between $P, Q : \harr {\cat C}{\cat D}$ is then a
+  homomorphism $P \multimap Q$ with an inverse $Q \multimap P$.
+\end{lemma}
+
+Our first synthetic theorem is that we have that the unit, i.e., the
+hom set, is the unit for tensor and hom.
+%
+The theorem for hom is the synthetic \emph{Yoneda lemma} and the
+theorem for tensor is the synthetic \emph{Co-Yoneda lemma}.
+\begin{lemma}[{Yoneda and Co-Yoneda}]
+  \label{lem:yo-and-co-yo}
+  Let $P : \harr {\cat C} {\cat D}$.
+
+  For any $\gamma : \cat E \vdash b : \cat D$, the following are naturally isomorphic to $\harrapp P \alpha b$:
+  \begin{enumerate}
+  \item $\homrallXYtoZ {\beta} {(\punitinXfromYtoZ {D} {\beta} {b})} {\harrapp P {\alpha} {\beta}}$
+  \item $\tensorexistsXwithYandZ {\beta} {\harrapp P \alpha {\beta}} {(\punitinXfromYtoZ {\cat D} {\beta} {b})}$
+  \end{enumerate}
+
+  For any $\gamma : \cat E \vdash a : \cat C$, the following are naturally isomorphic to $\harrapp P a \beta$:
+  \begin{enumerate}
+  \item $\homlallXYtoZ {{\alpha}} {(\punitinXfromYtoZ {C} {\alpha} {a})} {\harrapp P {{\alpha}} {\beta}}$
+  \item $\tensorexistsXwithYandZ {\alpha} {(\punitinXfromYtoZ {C} {a} {\alpha})} {\harrapp P {\alpha}{\beta}}$
+  \end{enumerate}
+\end{lemma}
+\begin{proof}
+  TODO: what's better, build the substitution in or prove the
+  substitution can be built in afterwards? The one building in the
+  substitution obscures the connection to the normal tensor/hom/unit
+  relationship, but the one with the built in subst is obviously more
+  general.
+\end{proof}
+
+Next, we have various associativities between the homs, tensor and end
+constructions. Semantically, these are related to so-called
+\emph{Fubini} theorems about interchange of (co)-ends.
+\begin{lemma}[{Fubini}]
+  \label{lem:fubini}
+  The following isomorphisms hold for any $P : \harr {\cat C}{\cat D},
+  P' : \harr {\cat C}{\cat D} Q :\harr {\cat D}{\cat E}, R : \harr
+  {\cat E} {\cat F}, S : \harr {\cat C} {\cat E}$. where $\alpha:\cat
+  C, \beta:\cat D, \gamma : \cat E, \delta: \cat F$
+  \begin{enumerate}
+  \item $\tensorexistsXwithYandZ {\beta} {\harrapp P \alpha \beta} {(\tensorexistsXwithYandZ \gamma {\harrapp Q \beta \gamma} {\harrapp R \gamma \delta})} \cong \tensorexistsXwithYandZ {\gamma} {(\tensorexistsXwithYandZ \beta {\harrapp P \alpha \beta} {\harrapp Q \beta \gamma})} {\harrapp R \gamma \delta}$
+  \item $\homrallXYtoZ {\gamma} {(\tensorexistsXwithYandZ \beta {\harrapp P \alpha \beta} {\harrapp Q \beta \gamma})} {\harrapp S \alpha \gamma} \cong \homrallXYtoZ \beta {\harrapp P \alpha \beta} {\homrallXYtoZ \gamma {\harrapp Q \beta \gamma} {\harrapp S \alpha \gamma}}$
+  \item $\homlallXYtoZ {\alpha} {(\tensorexistsXwithYandZ \beta {\harrapp P \alpha \beta} {\harrapp Q \beta \gamma})} {\harrapp S \alpha \gamma} \cong \homlallXYtoZ \beta {\harrapp Q \beta \gamma} {\homlallXYtoZ \alpha {\harrapp P \alpha \beta} {\harrapp S \alpha \gamma}}$
+  \item $\homrallXYtoZ \gamma {\harrapp Q \beta \gamma} {(\homlallXYtoZ \alpha {\harrapp P \alpha \beta} {\harrapp S \alpha \gamma})} \cong \homlallXYtoZ \alpha {\harrapp P \alpha \beta} {(\homrallXYtoZ \gamma {\harrapp Q \beta \gamma} {\harrapp S \alpha \gamma})}$
+  \item $\pendallXdotY \alpha \homrallXYtoZ \beta P {P'} \cong \pendallXdotY \beta \homlallXYtoZ \alpha P {P'}$
+  \end{enumerate}  
+\end{lemma}
+\begin{proof}
+  TODO
+\end{proof}
+\subsection{Composition, Functoriality, Profunctoriality, Naturality}
+
+Hom sets don't provide primitive syntax for composition of morphisms
+in a category. As our first synthetic construction, we can
+\emph{define} composition using the elimination form for the unit.
+%
+\begin{definition}[Composition of Morphisms]
+  We define composition of morphisms as the following term of type $\Pi_{\cat C:\Cat}\hombinary {\alpha_1} {\punitinXfromYtoZ {\cat C}{\alpha_1} {\alpha_2}} {\alpha_2} {\punitinXfromYtoZ {\cat C}{\alpha_2} {\alpha_3}} {\alpha_3} {\punitinXfromYtoZ {\cat C}{\alpha_1} {\alpha_3}}$:
+  \[
+  \lambda \cat C.
+\punitelimtoYwithkontZ {\hlambda {\alpha_1}{\alpha_2} \homrallXYtoZ {\alpha_3} {\punitinXfromYtoZ {\cat C} {\alpha_2} {\alpha_3}} {\punitinXfromYtoZ {\cat C} {\alpha_1} {\alpha_3}}} {\lambdaunary {\alpha} {f : \punitinXfromYtoZ {\cat C}{\alpha} {\alpha_3}} {\alpha_3} f}
+  \]
+  We call this term $\textrm{comp}$
+\end{definition}
+This has the same flavor as a directed type theory: to construct the
+composite it is sufficient to define the action on the identity
+morphism, which is of course the identity for composition.
+%
+Note that we gave a slight bias to the definition to follow by
+induction on the \emph{first} function, whereas we could make the
+following alternative definition, naturally of type
+$\pendallXdotY {\alpha_2} \homrallXYtoZ {\alpha_3} {\punitinXfromYtoZ {\cat C}{\alpha_2} {\alpha_3}}
+{(\homlallXYtoZ {{\alpha_1}} {\punitinXfromYtoZ {\cat C}{\alpha_1} {\alpha_2}}{\punitinXfromYtoZ {\cat C}{\alpha_1} {\alpha_3}})}$:
+\[
+  \punitelimtoYwithkontZ {\hlambda {\alpha_2}{\alpha_3} \homlallXYtoZ {{\alpha_1}} {\punitinXfromYtoZ {\cat C}{\alpha_1} {\alpha_2}}{\punitinXfromYtoZ {\cat C}{\alpha_1} {\alpha_3}} } {\lambdaunary {\alpha} {g : \punitinXfromYtoZ {\cat C}{\alpha_1}{\alpha}} {\alpha} g}
+\]
+
+\begin{lemma}[Equivalence of Composition Definitions]
+  TODO
+\end{lemma}
+
+It should follow by $\beta\eta$ that this is associative and unital.
+
+Next, let's prove synthetically that every functor $F : \varr{\cat
+  C}{\cat D}$ is functorial in this notion of composition.
+\begin{definition}
+  For any $F : \varr {\cat C} {\cat D}$, we can define a morphism $F_1
+  : \homunary {\alpha_1} {\punitinXfromYtoZ {\cat C} {\alpha_1}
+    {\alpha_2}} {\alpha_2} {\punitinXfromYtoZ {\cat D} {F \alpha_1}
+    {F \alpha_2}}$ as
+  \[ F_1 = \punitelimtoYwithkontZ {\hlambda {\alpha_1}{\alpha_2} \punitinXfromYtoZ {\cat D} {F \alpha_1}{F \alpha_2}}
+  {\pendlambdaXdotY \alpha \punitrefl {\cat D} {F \alpha}}
+  \]
+  Note that
+  \[ A[\id_\alpha/\alpha] = \id_{A} \]
+  \[ A[f;g] = A[f];A[g] \]
+\end{definition}
+Note the similarity to the proof that all functions are congruences in
+MLTT/HoTT.
+
+And similarly, every profunctor $R : \harr {\cat C}{\cat D}$ is
+\emph{pro-functorial}.
+\begin{definition}
+  For any $R : \varr {\cat C} {\cat D}$, we can define left and right
+  actions of composition:
+  \[ \homunary {\alpha_1} {\tensorexistsXwithYandZ {\alpha_2} {(\punitinXfromYtoZ {\cat C} {\alpha_1}{\alpha_2})} {\tensorexistsXwithYandZ {\alpha_3} {(\harrapp R {\alpha_2} {\alpha_3})}}{(\punitinXfromYtoZ {\cat C} {\alpha_3}{\alpha_4})}} {\alpha_4} {\harrapp R {\alpha_1}{\alpha_4}}
+  \]
+  Furthermore we can show that they are functorial in that the induced functions
+  TODO
+  are equal
+\end{definition}
+
+Next we can define natural transformations as specific kinds of terms
+whose type is $\pendallXdotY {\alpha} \punitinXfromYtoZ {\cat D} {F
+  \alpha}{G \alpha}$. In fact we can generalize this to any
+endo-profunctor $P$:
+\begin{definition}[Natural element]
+  A natural element of an endo-profunctor $P : \harr {\cat C}{\cat C}$
+  is a term of type $\pendallXdotY {\alpha} \harrapp P \alpha \alpha$.
+
+  A natural transformation is the special case where $P =
+  \punitinXfromYtoZ {\cat D} {F \alpha}{G \alpha}$ for functors $F, G
+  : \varr {\cat C}{\cat D}$.
+
+  This is natural in the sense that the incuded terms
+  \[ \homunary {\alpha_1}{\punitinXfromYtoZ {} {\alpha_1}{\alpha_2}} {\alpha_2} {\harrapp P {\alpha_1}{\alpha_2}}  \]
+  and
+  \[ \pendallXdotY {\alpha_2} \homlallXYtoZ {\alpha_1} {\punitinXfromYtoZ {} {\alpha_1}{\alpha_2}} {\harrapp P {\alpha_1}{\alpha_2}}\]
+
+  are equal under the associativity equivalence between these two
+  types.
+\end{definition}
+
+\subsection{Yoneda Lemma and Representable Presheaves}
+
+Given any category $\cat C$, we can define the positive and negative
+Yoneda embeddings $Y^\pm : \cat C \to \pmPresheaf {\cat C}$ using the
+hom profunctor. Note that both of these functors are covariant in
+$\cat C$, so semantically, $\posPresheaf {\cat C}$ should be
+interpreted as $(\Set^{\cat C})^o$.
+
+\begin{definition}
+  Define the negative Yoneda embedding $Y^-[\alpha]$
+  \[ \alpha:\cat C \vdash Y^-[\alpha] = \lambda \alpha':\cat C. \punitinXfromYtoZ {\cat C}{\alpha'}{\alpha} : \negPresheaf {\cat C} \]
+
+  The positive Yoneda embedding is defined similarly $Y^+[\alpha]$.
+\end{definition}
+
+The Yoneda lemma says that the set of natural transformations from
+$Y^-[\alpha]$ to a presheaf $p$ is naturally isomorphic to the
+elements of the presheaf itself: $\negPresheafApp \alpha p$. In
+encoding this statement in our type theory, we find that we can use
+two different interpretations of ``natural transformations''.
+
+On the one hand, a natural transformation should be a morphism in the
+presheaf category. This interpretation of the Yoneda lemma would mean
+that we have a natural isomorphism of profunctors:
+\[ \alpha:\cat C; \pi:\negPresheaf {\cat C} \vdash \punitinXfromYtoZ{\negPresheaf {\cat C}} {Y^-[\alpha]} {\pi} \cong \negPresheafApp \alpha \pi \]
+Attempting to prove this we become stuck very quickly because we know
+nothing about the structure of morphisms in the presheaf category.
+
+On the other hand, a natural transformation of presheaves $\pi$ and
+$\rho$ can be defined as a family of functions from $\negPresheafApp
+\alpha \pi$ to $\negPresheafApp \alpha \rho$, or in other words as an
+\emph{end}. Specializing to the Yoneda lemma, we interpret ``a natural
+transformation from $Y^-[\alpha]$ to $\pi$'' as the following, which simplifies using $\beta$ reduction:
+\[ \alpha:\cat C; \pi:\negPresheaf {\cat C} \vdash \homrallXYtoZ {\alpha':\cat C} {\negPresheafApp {\alpha'} {Y^-[\alpha]}}{\negPresheafApp {\alpha'} \pi} = \homrallXYtoZ {\alpha':\cat C} {(\punitinXfromYtoZ {\cat C}{\alpha'}{\alpha})}{(\negPresheafApp {\alpha'} \pi)} \]
+And then the Yoneda lemma becomes:
+\[ \alpha:\cat C; \pi:\negPresheaf {\cat C} \vdash \homrallXYtoZ {\alpha':\cat C} {(\punitinXfromYtoZ {\cat C}{\alpha'}{\alpha})}{(\negPresheafApp {\alpha'} \pi)} \cong \negPresheafApp \alpha \pi\]
+Which is an instance of our earlier Lemma\,\ref{lem:yo-and-co-yo}
+which follows quite easily.
+
+Then we are left with the question of whether these two forms of the
+Yoneda lemma are equivalent, or slightly more generally are these two definitions of natural transformation of presheaves equivalent:
+\[ \rho:\negPresheaf{\cat C}; \pi:\negPresheaf {\cat C} \vdash \punitinXfromYtoZ{\negPresheaf {\cat C}} {\rho} {\pi} \]
+and
+\[ \rho:\negPresheaf{\cat C}; \pi:\negPresheaf {\cat C} \vdash \homrallXYtoZ {\alpha:\cat C} {\negPresheafApp \alpha \rho}{\negPresheafApp \alpha \pi} \]
+While we can show that there is a morphism from the hom set to the
+end, the inverse does not seem provable, and seems analogous to the
+\emph{univalence} axiom in homotopy type theory, as it tells us that
+the morphisms of the universe are given by a particular kind of
+function in our logic.
+
+\subsection{Adjoint Functors and Representable Profunctors}
+
+Next, we demonstrate that the two common definitions of adjoint
+functors are equivalent: isomorphism of hom-sets and units and
+co-units.
+%
+The proof goes through most easily, and utilizes duality most
+effectively, if we factorize the construction through an intermediate
+notion of when a functor \emph{represents} a profunctor.
+
+\begin{definition}[Functor Represents a Profunctor]
+  A functor $G : \varr{\cat D}{\cat C}$ right-represents a profunctor $R :
+  \harr {\cat C}{\cat D}$ when there is an isomorphism of profunctors:
+  \[ \punitinXfromYtoZ {\cat C} {\alpha} {G \beta} \cong \harrapp R {\alpha}{\beta}\]
+
+  We say a functor $F : \varr{\cat C}{\cat D}$ \emph{left-represents} a profunctor $R : \harr {\cat C}{\cat D}$ when there is an isomorphism of profunctors
+  \[ \punitinXfromYtoZ {\cat D} {F \alpha} {\beta} \cong \harrapp R {\alpha}{\beta}\]
+\end{definition}
+
+This is a parameterized generalization of the notion of an object
+representing a functor $R : C^{op} \to Set$.
+These notions are precisely dual. $F$ left-represents a profunctor $R$
+precisely if it right-represents it in the opposite SCT.
+
+Note that trivially any functor $G$ right-represents
+$\hlambda\alpha\beta \punitinXfromYtoZ {\cat D}{\alpha} {G \beta}$ and
+dually, any functor $F$ left-represents
+$\hlambda\alpha\beta\punitinXfromYtoZ {\cat C}{F\alpha}{\beta}$.
+
+Then we can show that any right-representation isomorphism is
+determined by a co-unit:
+\begin{lemma}[(Co)-unit of a Representation]
+  If $G: \varr {\cat D}{\cat C}$ right-represents $R : \harr {\cat
+    C}{\cat D}$, then there is a natural element, the co-unit:
+  \[ \pendallXdotY \beta \harrapp R {G\beta} \beta \]
+  Such that the half of the isomorphism 
+  \[ \punitinXfromYtoZ {\cat C} {\alpha} {G \beta} \multimap \harrapp R {\alpha}{\beta} \]
+  is defined by composition with the natural element.
+
+  Dually, any $F$ left-representing isomorphism with $R$ is determined
+  by a ``unit'' of type $\pendallXdotY \alpha \harrapp R {\alpha}
+  {F\alpha}$.
+\end{lemma}
+\begin{proof}
+  A homomorphism $\punitinXfromYtoZ {\cat C} {\alpha} {G \beta} \multimap \harrapp R {\alpha}{\beta}$ is by definition a term of type
+  \[ \pendallXdotY \alpha \homrallXYtoZ {\beta} {(\punitinXfromYtoZ {\cat C} \alpha {G \beta})} {\harrapp R \alpha \beta} \]
+  By \cref{lem:fubini}, this type is isomorphic to 
+  \[ \pendallXdotY \beta \homlallXYtoZ {\alpha} {(\punitinXfromYtoZ {\cat C} \alpha {G \beta})} {\harrapp R \alpha \beta} \]
+  By \cref{lem:yo-and-co-yo}, this type is isomorphic to
+  \[ \pendallXdotY \beta {{\harrapp R {(G \beta)} \beta}} \]
+
+  Therefore this half of the isomorphism is determined by a natural
+  element. The construction in the proof of \cref{lem:yo-and-co-yo} is
+  indeed by composition (i.e., unit elimination).
+
+  The $F$ case is precisely dual.
+\end{proof}
+
+Combining these two, we get the unit-counit definition for adjoint
+functors
+
+\begin{definition}[Adjunction by unit/co-unit]
+  We say $G : \varr {\cat D}{\cat C}$ is right adjoint to $F : \varr
+  {\cat C}{\cat D}$ if there is a unit
+  \[ \eta : \pendallXdotY \alpha \punitinXfromYtoZ \alpha {G(F(\alpha))}\]
+  and a co-unit
+  \[ \epsilon : \pendallXdotY \beta \punitinXfromYtoZ {F(G(\beta))} \beta \]
+
+  satisfying the triangle identities...
+  \[ \id_{R[\beta]} = R[\epsilon_{\beta}] \circ \eta_{R[\beta]} \]
+  \[ \id_{L[\alpha]} = \epsilon_{L[\alpha]} \circ L[\eta_{\alpha}] \]
+\end{definition}
+
+\begin{theorem}[Adjunctions]
+  $G$ is right adjoint to $F$ if and only if there is a natural isomorphism
+  $\punitinXfromYtoZ {\cat D} {(F \alpha)} \beta \cong \punitinXfromYtoZ {\cat D} \alpha {G \beta}$
+\end{theorem}
+\begin{proof}
+  First, by \cref{lem:co-unit}, the forward direction of the
+  isomorphism is equivalent to providing a term
+  \[ \eta : \pendallXdotY \alpha \punitinXfromYtoZ \alpha {G(F(\alpha)) } \]
+
+  Second, the backward direction of the isomorphism is equivalent to
+  providing a term
+  \[ \epsilon : \pendallXdotY \beta \punitinXfromYtoZ {F(G(\beta))} \beta \]
+
+  And the triangle laws state exactly that the induced functions are
+  isomorphisms.
+\end{proof}
+
+\subsection{Kan Extensions/Lifts}
+
+As a further example of the synthetic approach to universal
+properties, we consider right and left \emph{Kan extensions}, which
+are generalizations of limits and colimits, respectively.
+
+The right Kan extension of $F$ along $J$ is the ``best'' (maximal) way to
+fill in the following square:
+\[% https://tikzcd.yichuanshen.de/#N4Igdg9gJgpgziAXAbVABwnAlgFyxMJZABgBpiBdUkANwEMAbAVxiRAGEQBfU9TXfIRQBGclVqMWbACLdeIDNjwEiZYePrNWiDgHJu4mFADm8IqABmAJwgBbJGRA4ISAEzVNUnQCk5lm-aIjs5IohJabABiBlxAA
+\begin{tikzcd}
+\cat C \arrow[d, "J"] \arrow[r, "F"] & \cat D \\
+\cat C'                              &  
+\end{tikzcd}\]
+I.e. a functor $\textrm{Ran}_J F : \cat C' \to \cat D$ with a
+universal natural transformation $\textrm{Ran}_J F \circ J \Rightarrow
+F$. A left kan extension $\textrm{Lan}_J F$ is the same but with a
+natural transformation $F \Rightarrow \textrm{Lan}_J F \circ J $.
+
+We can express the concept in SCT as representing a certain profunctor
+constructed using the profunctor hom:
+\begin{definition}[Kan Extensions]
+  Given functors $F : \varr {\cat C}{\cat D}$ and $J : \varr {\cat
+    C}{\cat C'}$, a right Kan extension is a functor $\textrm{Ran}_J F
+  : \varr {\cat C'}{\cat D} $that right-represents the profunctor
+  \[ \hlambda {\beta:\cat D}{\alpha': \cat C} \homrallXYtoZ {\alpha}{(\punitinXfromYtoZ {\cat C'}{\alpha'}{J\alpha})} {\punitinXfromYtoZ {\cat D}{\beta}{F \alpha}} \]
+
+  A left Kan extensions is a functor $\textrm{Lan}_J F : \varr {\cat
+    C'}{\cat D}$ is a functor satisfying the dual property: it
+  \emph{left}-represents the profunctor
+  \[\hlambda {\alpha':\cat C}{\beta:\cat D} \homlallXYtoZ {\alpha}{(\punitinXfromYtoZ {\cat C'}{J\alpha}{\alpha'})} {\punitinXfromYtoZ {\cat D}{F \alpha}{\beta}}
+  \]
+\end{definition}
+
+The limit of a functor $F : \varr{\cat C}{\cat D}$ is a special case
+of a Kan extension where $J : \varr{\cat C}{1}$ is the unique functor
+to the terminal category. Similarly, the colimit of $F$ is the left
+Kan extension with respect to the same $J$.
+%
+Specializing to this case, the representability for limits means:
+\[ \punitinXfromYtoZ {\cat D}{\beta}{\textrm{lim} F} \cong \homrallXYtoZ {\alpha}{(\punitinXfromYtoZ {\cat 1}{\alpha'}{()})}{\punitinXfromYtoZ {\cat D}{\beta}{F \alpha}} \]
+And if we view a morphism in $\cat 1$ as being trivial, then this
+essentially states that mapping $\beta$ to $\textrm{lim} F$ is the
+same thing as constructing a cone: for all $\alpha :\cat C$, a
+morphism from $\beta$ to $F \alpha$, naturally in $\alpha$.
+
+The usual formulation of Kan extensions is in terms of a universal
+morphism.
+%
+We can reproduce this formulation using our established theory of
+equivalences for homs of profunctors.
+\begin{lemma}
+  One half of the representability isomorphism for a right Kan
+  extension $\textrm{Ran}_J F$ is determined by a natural
+  transformation
+  \[ \pendallXdotY \alpha \punitinXfromYtoZ {\cat D}{(\textrm{Ran}_JF)(J\alpha)}{F\alpha} \]
+
+  Dually, one half of the representability isomorphism for a left Kan
+  extension $\textrm{Lan}_J F$ is determined by a natural
+  transformation
+  \[ \pendallXdotY \alpha \punitinXfromYtoZ {\cat D}{F\alpha}{(\textrm{Lan}_JF)(J\alpha)} \]
+\end{lemma}
+\begin{proof}
+  By \cref{lem:unit-repr}, one half of representability is determined by a natural element
+  \[ \pendallXdotY {\alpha'} \homrallXYtoZ {\alpha}{(\punitinXfromYtoZ {\cat C'}{\alpha'}{J\alpha})} {\punitinXfromYtoZ {\cat D}{(\textrm{Ran}_J F)\alpha'}{F \alpha}} \]
+  which, by \cref{lem:fubini} is isomorphic to
+  \[
+  \pendallXdotY {\alpha} \homlallXYtoZ {\alpha'}{(\punitinXfromYtoZ {\cat C'}{\alpha'}{J\alpha})}{\punitinXfromYtoZ {\cat D}{(\textrm{Ran}_J F)\alpha'}{F \alpha}}
+  \]
+  which, by \cref{lem:yo-and-co-yo}, is isomorphic to 
+  \[ \pendallXdotY \alpha \punitinXfromYtoZ {\cat D}{(\textrm{Ran}_JF)(J\alpha)}{F\alpha} \]
+  by taking $P = \hlambda {\alpha'}{\alpha} {\punitinXfromYtoZ {\cat D}{(\textrm{Ran}_J F)\alpha'}{F \alpha}}$.
+
+  The proof for left Kan extensions follows by \cref{lem:duality}.
+\end{proof}
+
+With the definition based on representability, we can easily prove the
+generalization of the famous theorem, that right adjoints preserve not
+just limits but all right Kan extensions
+
+\begin{theorem}[Right/Left Adjoints preserve Right/Left Kan Extensions]
+  Let $K : \varr{\cat C}{\cat D}$ and $J : \varr{\cat C}{\cat C'}$.
+
+  If the right Kan extension $\textrm{Ran}_J K$ exists, and if $G :
+  \varr {\cat D}{\cat D'}$ has a left adjoint $F : \varr{\cat D'}{\cat
+    D}$, then the right Kan extension $\textrm{Ran}_J G\circ K$ exists
+  and is given by $G \circ \textrm{Ran}_J K$.
+
+  Similarly, if the left Kan extension $\textrm{Lan}_J K$ exists and
+  if $F : \varr{\cat D}{\cat D'}$ has a right adjoint $G : \varr{\cat
+    D'}{\cat D}$, then the left Kan extension $\textrm{Lan}_J F\circ
+  K$ exists and is given by $F \circ \textrm{Ran}_J K$.
+\end{theorem}
+\begin{proof}
+  By the composition of the following natural isomorphisms:
+  \begin{align*}
+    \punitinXfromYtoZ {\cat D'} {\beta'} {G ((\textrm{Ran}_J K) \alpha')} &\cong \punitinXfromYtoZ {\cat D}{F \beta'} {(\textrm{Ran}_J K) \alpha'} \tag{Adjunction}\\
+    &\cong \homrallXYtoZ {\alpha}{(\punitinXfromYtoZ {\cat C'}{\alpha'}{J\alpha})} {\punitinXfromYtoZ {\cat D}{F \beta'}{K \alpha}} \tag{Ran assumption, \cref{lem:natural-restriction}}\\
+    &\cong \homrallXYtoZ {\alpha}{(\punitinXfromYtoZ {\cat C'}{\alpha'}{J\alpha})} {\punitinXfromYtoZ {\cat D}{\beta'}{G(K \alpha)}} \tag{Adjunction}\\
+  \end{align*}
+\end{proof}
+
+\subsection{(op)-Fibrations}
+
+A (Street) fibration can be defined in any 2-category. A functor $p :
+\varr {\cat E}{\cat B}$ is a fibration when the morphism $p \to \cat
+B/p$ in the slice 2-category over $\cat B$ has a right adjoint.
+
+Since we can define all of the components of this slice 2-category we
+can internalize this definition in SCT.
+
+\begin{enumerate}
+\item A functor $\textrm{wp} : \exists (\alpha:\cat B; \beta:\cat E). \punitinXfromYtoZ {\cat B} \alpha {p\beta}$
+\item In the slice category: $p(\textrm{wp}(k)) = \pi_-k$
+\item With a universal morphism $\epsilon : \pendallXdotY {k} \punitinXfromYtoZ {} {(\pi_-k,k,\punitrefl{}{\pi_-k})} {k}$
+\item In the slice category: $\pi_k \epsilon = \id$ (how to express this nicely in SCT?)
+\end{enumerate}
+
+We can also weaken this to a Street fibration if we require the more sensible $p(\textrm{wp}(k)) \cong \pi_-k$ only.
+
+\begin{lemma}{Fibrations compose}
+  If $F : \varr{\cat C}{\cat D}$ and $G : \varr{\cat D}{\cat E}$ are
+  (op)-fibrations then so is $G \circ F$.
+\end{lemma}
+\begin{proof}
+  By composition of adjoints.
+\end{proof}
+
+\begin{lemma}{Fibrations are Stable under Pullback}
+
+\end{lemma}
+
+\subsection{Duality Principle}
+
+TODO: example of applying the principle in our calculus.
+
+\subsection{Formal Type Theory}
+
+Since many type theories are given semantics by free structured
+categories, we can use SCT as a logical framework in which to do
+``formal type theory'' by axiomatizing the structure of the categories
+involved. The theories we can define are limited because we do not
+have access to opposite categories, but we can define some modal and
+monoidal structures.
+
+\begin{definition}
+  An \emph{SCT-signature} is a well-formed context $\Gamma \isadtctx$.
+\end{definition}
+
+\begin{lemma}
+  Any \emph{unary mode theory} can be defined as an SCT-signature as follows.
+  \begin{enumerate}
+  \item For each mode $m$ add an assumption $m : \Cat$
+  \item For each 1-morphism $f : m \to n$ add an assumption $R_f :
+    \harr m n$ and assumptions $F_f : \varr m n$ with and $G_f : \varr
+    n m$ with assumptions that they represent $R_f$.
+  \item For each 2-morphism $\alpha : f \Rightarrow g$ add an assumption $\alpha : R_f \multimap R_g$
+  \item Add each equality as an assumption using the identity type Id.
+  \end{enumerate}
+\end{lemma}
+
+This translation illustrates some ways the notion of SCT-signature is
+more flexible than a mode theory. For instance, in an SCT-signature we
+can add a profunctor assumption $R_f : \harr m n$ without saying
+whether it is representable as $F_f$ or $G_f$, and we can add more
+complex 2-morphisms.
+
+\begin{definition}
+  The \emph{theory of a strict monoidal category} consists of
+  \begin{enumerate}
+  \item A category $C : \Cat$
+  \item A unit $I : \varr 1 C$
+  \item A tensor product $\otimes : \varr {(C \times C)} C$
+  \item Such that $\alpha:(C \times C) \times C \vdash (\pi_2\pi_1\alpha \otimes \pi_2\pi_1 \alpha) \otimes \pi_2 \alpha = \pi_2\pi_1\alpha \otimes (\pi_2\pi_1 \alpha \otimes \pi_2 \alpha)$
+    and $\alpha \otimes I() = \alpha = I() \otimes \alpha$
+  \end{enumerate}
+\end{definition}
+
+\begin{definition}
+  The theory of a cartesian strict monoidal category is an extension
+  of the theory of a strict monoidal category with representability
+  data
+  \[ \punitinXfromYtoZ {C} {\Gamma}{\Delta \otimes \Sigma} \cong (\punitinXfromYtoZ C \Gamma \Delta)\times (\punitinXfromYtoZ C \Gamma \Sigma) \]
+  and
+  \[ \punitinXfromYtoZ {C} {\Gamma} I \cong 1 \]
+\end{definition}
+
+\begin{definition}
+  The theory of a simple type theory is an extension of the theory of
+  a strict cartesian monoidal category with a
+  \begin{enumerate}
+  \item A category $\text{Ty}$ of types
+  \item A profunctor $\vdash : \harr C {\textrm{Ty}}$
+  \item A functor $\text{var} : \varr{\textrm{Ty}} C$ representing $\vdash$ :
+    \[ \Gamma \vdash A \cong \punitinXfromYtoZ {} \Gamma {\text{var} A} \]
+  \item As assumption that $\text{var}$ is fully faithful:
+     \[ \punitinXfromYtoZ {} {\text{var} A} {\text{var} B} \cong \punitinXfromYtoZ {} A B \]
+  \end{enumerate}
+\end{definition}
+
+\begin{definition}
+  The theory of CBPV is an extension of the theory of a simple type
+  theory $(\text{VCtx},\times,1,\text{VTy},\vdash_V,\text{var}_V)$ with
+  \begin{enumerate}
+  \item A category $\text{CCtx}$
+  \item An object $I : \varr 1 {\text{CCtx}}$
+  \item An action of $\text{VCtx}$ on $\text{CCtx}$, i.e.,
+    \begin{enumerate}
+    \item A functor $\oslash : \text{VCtx}\times \text{CCtx} \to \text{CCtx}$
+    \item Such that $1 \oslash \underline\Delta = \underline\Delta$
+    \item And $(\Gamma \times \Delta) \oslash \underline\Sigma = \Gamma \oslash \Delta \oslash \underline\Sigma$
+    \end{enumerate}
+  \item A category $\text{CTy}$ with a profunctor $\vdash_C : \harr
+    {\text{CCtx}}{\text{CTy}}$ and a fully faithful functor $\text{var}_C$ that represents it
+    \[ \underline\Delta \vdash_C \underline B \cong \punitinXfromYtoZ {} {\underline\Delta}{\text{var}_C\underline B}\]
+  \end{enumerate}
+  Furthermore we can add in the universal properties of any
+  non-contravariant CBPV connectives. For instance:
+  \begin{enumerate}
+  \item $F : \varr {VTy} {CTy}$ satisfying
+    \[ \Gamma \oslash \text{var}_C F A \vdash_C \underline B \cong \Gamma \times \text{var}_V A \oslash I \vdash_C \underline B  \]
+    \[ \Gamma \vdash_V U \underline B \cong \Gamma \oslash I \vdash_C \underline B  \]
+  \end{enumerate}
+\end{definition}
+
+\begin{definition}[Theory of unary Computational lambda calculus with booleans]
+  \begin{enumerate}
+  \item A category $C$
+  \item A terminal object $1$ of $C$
+  \item A monad $T$ on $C$
+  \item A type $\textrm{Bit} : \varr 1 C$
+  \item Morphisms $\textrm{y},\textrm{n} : \pendallXdotY * \punitinXfromYtoZ {1(*)} {\textrm{Bit}(*)}$
+  \item A natural transformation $\textrm{if} : \pendallXdotY * \homrallXYtoZ \alpha {(\punitinXfromYtoZ C {1(*)} \alpha) \times (\punitinXfromYtoZ C {1(*)} \alpha)} {\punitinXfromYtoZ C {\textrm{Bit}(*)} \alpha}$
+  \item Satisfying $\textrm{if} y = \pi_1$ and $\textrm{if} n = \pi_2$
+  \item A morphism $\textrm{print} : \pendallXdotY * \punitinXfromYtoZ {\textrm{Bit}(*)}{T (1(*))}$
+  \end{enumerate}
+\end{definition}
+
+Then we can ``program'' in this computational lambda calculus within
+our logical framework by constructing terms of type $\punitinXfromYtoZ
+C {A} {B}$ for any functors $A,B$ into $C$.
+
+Using a terminal category and the category of elements, we can define
+a formal category with families, an algebraic formulation of the
+judgmental structure of dependent type theory:
+\begin{definition}[Category with Families]
+  A category with families consists of
+  \begin{enumerate}
+  \item A category $\textrm{Ctx}\isaCat$
+  \item A presheaf $\textrm{Ty} : \harr {\textrm{Ctx}}{1}$
+  \item A presheaf $\textrm{Tm} : \harr {\Sigma_{\Gamma:\textrm{Ctx};*:1}\harrapp{\textrm{Ty}} \Gamma *} 1$
+  \item A terminal object in $\textrm{Ctx}$
+  \item An \emph{extension operation}
+    $\textrm{ext} : \Sigma_{\Gamma:\textrm{Ctx};*:1}\harrapp{\textrm{Ty}}\Gamma * \to \textrm{Ctx}$
+    right representing the profunctor
+    \[ \Delta:\textrm{Ctx};A:\Sigma_{\Gamma:\textrm{Ctx};*:1}\harrapp{\textrm{Ty}}{\Gamma}{*} \vdash \tensorexistsXwithYandZ {a}{(\punitinXfromYtoZ {} \Delta {\pi^-\pi^-a})}
+       {(\punitinXfromYtoZ {}{\pi^-a}{A})}\]
+       % {\Sigma_{\Gamma:\textrm{Ctx};*:1}\harrapp{\textrm{Ty}}{\Gamma}{*}}
+    % 
+  \end{enumerate}
+\end{definition}
+
+%% Can we do pras?
+%%
+%% A PRA mode theory is given by a free 2-category
+%% 1. Each mode m is a CwF whose context cat is [m]
+%% 2. Each 1-cell μ : m -> n is a functor [μ] : [n] -> [m] that has a DRA and is a PRA
+%% 3. Each 2-cell α : μ => ν is a 2-cell
+
+%% Let C, D be CwFs.
+%% A DRA is a pair of
+%% 1. A functor  L : C -> D
+%% 2. A morphism R : L^*(D_Ty) -> C_ty in ∫C, i.e., a morphism taking LΓ |- A to Γ |- RA
+%% 3. An iso     I : L^*(D_tm)?? =~ R^*(C_tm) in ∫(L^*(DTy)) a morphism taking LΓ |- M : A to Γ |- I(M) : RA
+
+%% Let C, D such that C has a terminal object.
+%% G : C -> D is a PRA if
+%% G / 1 : C -> D / G1 is a right adjoint
+
+%% What about fixed points?
+%% Say C has a terminal object
+%% fix : Hom(a,a) -> Hom(1, a)
+%% satisfying 
+%% fix[f] = f o fix[f]?
+%% We can write down fix, but not the equation I think
 

### 2022-09-07 paper/paper.tex
some work on the semantics section

@@
 
-\usepackage{tikz-cd}
-\usepackage{amssymb}
-\usepackage{amsmath}
+\usepackage{tikz-cd} \usepackage{amssymb} \usepackage{amsmath}
 %% \usepackage{amsthm}
@@
   category, functor, profunctor and a generalized form of natural
-  transformations. By restricting the type theory to an ordered
-  linear restriction of higher-order logic, we guarantee that all
-  mappings between categories are functorial and all transformations
-  are natural by construction, with no separate proof
-  necessary. Important category theoretic proofs such as the Yoneda
-  lemma and Co-yoneda lemma become simple type theoretic proofs about
-  the relationship between unit, tensor and (ordered) function types,
-  and can be seen to be ordered refinements of theorems in
-  intuitionistic higher-order logic.
-
-  We give a sound and complete notion of model of our type theory in a
-  structure we call a hyperdoctrine of pro-arrow equipments with
-  displays, which are known to model enriched and internal category
-  theory. This means that while the proofs in our type theory look
-  like standard set-based arguments, the substructural syntactic
-  restrictions ensure that the proofs carry over to enriched and
-  internal settings as well.
+  transformations. By restricting the type theory to an ordered linear
+  restriction of predicate logic, we guarantee that all functions
+  between categories are functorial, all relations are profunctorial,
+  and all transformations are natural by construction, with no
+  separate proof necessary. Important category theoretic proofs such
+  as the Yoneda lemma and Co-yoneda lemma become simple type theoretic
+  proofs about the relationship between unit, tensor and (ordered)
+  function types, and can be seen to be ordered refinements of
+  theorems in predicate logic.
+
+  We develop a sound and complete notion of categorical model and give
+  a model construction that shows that the theorems are valid in
+  enriched and internal categories. This means that while the proofs
+  in our type theory look like standard set-based arguments, the
+  substructural syntactic restrictions ensure that the proofs carry
+  over to enriched and internal settings as well.
 \end{abstract}
@@
 Originally developed for applications to homological algebra, it was
-quickly discovered that these algebraic structures were common in
-logic and computer science, and that formal systems like type theories
-and programming languages can be modeled by a sound and complete
-notion of semantic model given by some notion of structured
-category\cite{lawvere,goguen,lambek,moggi}.
+quickly discovered that these categorical structures were common in
+logic and computer science, and that formal systems like logics, type
+theories and programming languages can be modeled by sound and
+complete semantic models given by notions of structured
+category\cite{lawvere,goguen,lambek,moggi}, a phenomenon sometimes
+referred to as the Curry-Howard-Lambek correspondence
+\cite{curry-howard-lambek}.
 %
@@
 %
-These seminal works have led to a fruitful field of research applying
-such categorical logic techniques to the design of new type theories
-based directly on category-theoretic semantics.
+These seminal works have led to a fruitful field of research known as
+categorical logic, where
+new type theories are designed based directly on category-theoretic
+semantics.
 %
@@
 acquisition of competence in category theory by computer scientists
-who do not have a strong background in ``pure'' mathemantics.
+who do not have a strong background in pure mathemantics.
 %
@@
 %
-Consider the analogous situation in differential calculus. When a
-physicist writes down a function, they do not expand out the
+Consider the analogous situation in differential calculus. When an
+adept analyst writes down a function, they do not expand out the
 $\epsilon-\delta$ definition of continuity and prove that it is
@@
 
-In this work we show that these algebraic formulations of category
-theory can be given a type-theoretic syntax that allows us to make
-``na\"ive'' set theoretic constructions for categories, but by
-adhering to a certain syntactic discipline provide correct arguments
-for not just ordinary category theory but enriched and internal
-categories as well.
-%
-The type theory, which we call SCT, is a ``synthetic'' theory of
-categories, functors, profunctors and natural transformations.
-%
-This provides a formulation of category theory in which profunctors
-play a primary role, most recognizable in ordinary category theory in
-the prominence of ``hom isomorphism'' formulations of universal
-properties.
-%
-The semantics of SCT is given by a slight generalization of virtual
-double categories \cite{} that we call double categories with
-displays.
-%
-Additionally, to define signatures and parameterize theorems like the
-Yoneda lemma, SCT includes an ordinary dependent type theory over it
-which includes usual $\Pi,\Sigma,\Id$ types.
-%
-In an implementation of SCT, this dependent type theory might be the
-host language itself.
-
-In this work, we focus on the syntax and semantics of SCT, but the
-ultimate goal is for this theory to serve as the foundation for a
-framework in which we can more easily study and program with new type
-theories. This falls in the tradition of \emph{logical frameworks}
-such as the LF system. To provide evidence for the feasibility of this
-approach, we show that we can model structures used in recent work on
-modal dependent type theory. This shows that an implementation of SCT
-would provide at least a toy prototype implementation/type checker for
-modal dependent type theory. This can be summarized in the slogan
-\begin{quote}
-  category theory as a logical framework
-\end{quote}
-that is, that a syntax for category theory can be used as a logical
-framework for type theories with semantics in structured categories.
-%
-If implemented in a proof assistant this would also provide with it
-the soundness of the syntax in any model.
-%
-The emphasis on profunctors in SCT fits nicely with the
-\emph{judgmental} approach of Per-Martin L\"of and suggests the
-following slogan
-\begin{quote}
-  profunctors as judgments, natural transformations as derivations
-\end{quote}
-because we observe that the ``derivation tree'' style of writing
-judgments corresponds closely to the type structure of profunctors and
-natural transformations in SCT, and it should be simple to implement a
-program that converts formal natural transformations into derivation
-tree syntax.
-
-This work is merely a first step in this direction, as there are still
-three major components of category theory: opposite categories,
-functor categories and ``universes'' such as the category of sets or
-categories that are not nicely axiomatizable in the current system.
-%
-Nontheless, a great deal of category theory can be carried out in this
-more limited setting, and we can still take advantage of duality at a
-meta level.
-
-\subsection{Ordered Logic of Functions and (Constructive) Relations}
-
-Our logic is based on four basic notions: categories, functors,
-profunctors and (multi-argument) natural transformations.
-%
-These generalize the basic notions already present in formulations of
-higher-order logic: sorts, functions, relations and proofs.
-%
-In fact, one way to view our logic is that it is an \emph{ordered},
-\emph{constructive} higher-order logic
-
-While profunctors are not typically employed in introductory category
-theory texts,
-
-\subsection{Category theory with Profunctors}
-
-First, we review the 4 basic notions of our theory of categories in
-the traditional formulation. A category $\cat C$ is a set of objects
-$\cat C_0$ with a span of arrows $\cat C_0 \leftarrow \cat C_1
-\rightarrow \cat C_0$ that pick out the source and target of each
-arrow, such that there are composition and identity arrows with
-appropriate laws. A functor $F : \cat C \to \cat D$ consists of a
-function on objects $F_0 : \cat C_0 \to \cat D_0$ and a function on
-morphisms $F_1 : \cat C_1 \to \cat D_1$ that is compatible with source
-and target and preserves identity and composition.
-
-A profunctor $R$ from $\cat C$ to $\cat D$ is a span $\cat C_0
-\leftarrow R \rightarrow \cat D_0$ that is compatible with
-\emph{left}-composition by $\cat C$ and \emph{right}-composition by
-$\cat D$. If we think of categories as generalizations of (ordered)
-sets then functors generalize (monotone) functions and profunctors
-generalize (monotone) \emph{relations}. We think of an element $r \in
-R$ with $s(r) = c$ and $t(r) = d$ as a ``proof'' that $c R d$ or some
-kind of ``heteromorphism'' from $c$ to $d$. The compatibility with
-composition says that these relations are down-closed in $C$ and
-up-closed in $D$, or that heteromorphisms can be composed with
-homomorphisms.
-%
-
-%% Embracing the analogy ``profunctors as judgments'' we can view
-%% elements of a profunctor $r \in R$ with $s(r) = A$ and $t(r) = B$ in
-%% type-theoretic notation
-%% \[ \inferrule*[right={r}]{~}{A \vdash_R B}\]
-%% Noting that every category $\cat C$ has a hom profunctor $Hom_C : \cat
-%% C \pto \cat D$, we get for each category a judgment
-%% \[ \inferrule*{}{A \vdash_{Hom_C} A'} \]
-%% and the category operations provide us with rules for use of the assumption:
-%% \[ \inferrule*[right=id]{}{A \vdash_{Hom_C} A} \]
-%% and cut
-%% \[ \inferrule*[right=cut]{A \vdash_{Hom_C} A' \and A' \vdash_{Hom_C} A''}{A \vdash_{Hom_C} A''} \]
-%% satisfying unit and associativity equations.
-%% %
-%% An arbitrary profunctor $R$ supports a kind of ``bi-cut''
-%% rule\footnote{alternatively, this can be defined as 2 cuts: left and
-%% right, with an associativity law}:
-%% \[ \inferrule*[right=bicut]{A \vdash_{Hom_C} A' \and A' \vdash_R B' \and B' \vdash_{Hom_D} B}{A \vdash_R B} \]
-%% That satisfies similar bi-unit and bi-associativity laws.
-
-\subsection{Double Categories and Pro-arrow Equipments}
-
-We can abstract the algebraic structure of categories, functors,
-profunctors, and natural transformations as a \emph{double category}.
-%
-A double category is a generalization of categories where we have a
-collection of objects $C_0$, and \emph{two separate} notions of
-morphism, which we call vertical $C_v$ and horizontal $C_h$ over the
-same objects $C_0$ which each have their own identity and composition,
-and finally a notion of 2-cell that is most easily described
-
-\[ % https://tikzcd.yichuanshen.de/#N4Igdg9gJgpgziAXAbVABwnAlgFyxMJZABgBpiBdUkANwEMAbAVxiRAGEB9YkAX1PSZc+QigCM5KrUYs2XMXwEgM2PASIAmSdXrNWiDpw2LBqkUQDM26XrYAdOwygQcCfqeHqUAFmu7ZBlzAYADUYrwmykJqoiSkYlL++iCRKp6xvgk6Msl8UjBQAObwRKAAZgBOEAC2SBIgOBBIWjYBIABKnAruIJU1zdSNSFatyZ3GPX21iCNDiL6jbJ2Ek1XTZA1NiPVJS9yRU0gbcwCs2bYGAGIHa0hnm0gAbOdtAMo3-fODW8+LBgDieV4QA
-\begin{tikzcd}
-C_0 \arrow[r, "R_0"] \arrow[d, "F"] & C_1 \arrow[r, "R_1"] & C_2 \arrow[r, "R_2"] & \ldots \arrow[r, "R_n"] & C_{n+1} \arrow[d, "G"] \\
-{} \arrow[rrrr, "S"]                &                      &                      &                         & {}                    
-\end{tikzcd} \]
-
-This 2-cell has one output pro-arrow, many input pro-arrows which must
-match the source from one to the target object of the next and a left
-and right ``frame'' arrow. These can be composed in a manner similar
-to multi-categories, where you can compose this with many input
-2-cells, one for each input pro-arrow.
-%
-These then have complex associativity and unit rules and side
-conditions that say the framing arrows compose alongside the 2-cells.
-
-We address both of these issues by developing an alternative
-\emph{type theoretic syntax} for virtual equipments.
-%
-In this syntax the ``multi-2-cell'' above instead becomes a natural
-transformation judgment
-\[ \Phi \vdash t : S[F(x);G(y)] \]
-where the composable string of input arrows is modeled by an ordered,
-linear context $\Phi$ and output $S$ and frames $F,G$ are combined in
-the output type using the restriction operation of an equipment.
-%
-We write natural transformations in a natural deduction style: a
-transformation is written as a ``parameterized'' element of the output
-profunctor. Furthermore the restriction operators for vertical arrows
-into horizontal arrows is naturally modeled by a syntactic
-substitution.
-%
-This interpretation of restriction as substitution both matches the
-semantic intuition for the restriction operation as well as greatly
-simplifies reasoning about composition of transformations.
-
-%% Experienced category theorists have no trouble with this, as they are
-%% familiar with a great many principles for constructing transformations
-%% that are guaranteed to be natural. Consider an analogy with
-%% calculus. When a mathematician writes a formula such as
-
-%% $$\sin(f(x) - g(x))$$
-
-%% They do not need to spend time with an explicit $\epsilon,\delta$
-%% proof that the function is continuous or furthermore smooth. Instead,
-%% they follow syntactic principles that compositions of continuous
-%% functions are continuous.
-
-%% There are similar syntactic notions for guaranteeing that
-%% transformations are natural.
-
-%% \subsection{Category Theory with Profunctors}
-
-%% The syntax of Judgment Theory gives not just a syntax for the familiar
-%% trio of category, functor and natural transformation, but additionally
-%% a concept that is not usually highlighted in introductory category
-%% theory texts: the profunctor. A simple intuition is that profunctors
-%% are to functors as relations are to functions. Indeed, the syntax for
-%% judgment theory has a straightforward semantics in the category of
-%% sets and relations.
-
-%% The reason profunctors are not commonly introduced is that they are
-%% seen as an auxiliary notion: a profunctor $R : C \pto D$ is
-%% typically defined as a functor $C^{\textrm{op}}\times D \to
-%% \textrm{Set}$. However this definition requires two concepts that are
-%% somewhat problematic: the opposite category $C^\textrm{op}$ and the
-%% category of (small) sets $\textrm{Set}$. However, these notions do not
-%% necessarily exist in all of our intended applications. First, we are
-%% interested in semantics where are objects are \emph{multicategories}
-%% where morphisms have multiple inputs but only a single output. In this
-%% case the opposite of a multicategory is not a multicategory at all but
-%% instead some kind of ``co-multicategory'' with multiple outputs but
-%% one input. Second, having a universe object like $\textrm{Set}$ is also
-
-%% \subsection{Applications to Enriched, Internal and Multicategory Theory}
-
-%% The syntax of judgment theory is an internal language for
-%% \emph{virtual equipments}. These are a kind of joint generalization of
-%% double categories and multicategories.
-
-%% Notably, the definition of a category in \cref{sec:internal} can be
-%% applied as an extremely general notion of category, including internal
-%% categories and enriched categories. This general notion is often
-%% called a ``generalized multi-category'' and encompasses an astonishing
-%% range of mathematical structures, including categories, topological
-%% spaces, metric spaces, etc.
-
-%% \subsection{Applications to Generalized Type Theory}
-
-%% While we mostly focus on the semantics in category theory, due to the
-%% tight connection between category theory and the semantics of type
-%% theory, there are direct applications to the study of ``abstract type
-%% theory'' as well. We demonstrate how we can model single-variable
-%% simple type theories with products and coproducts within SCT. This is
-%% very much in the spirit of logical frameworks \cite{TODO}, since we
-%% can faithfully model the terms of an object theory inside of our
-%% metatheory. Since we have models of our metatheory in virtual double
-%% categories, we establish sound semantics as well. Put succinctly,
-%% category theory itself is the logical framework for the study of type
-%% theories.
+In this work, we apply the technique of categorical logic to these
+algebraic structures to design a formal logic, which we call
+\emph{ordered predicate logic}, that can be soundly interpreted in
+virtual equipments.
+%
+In ordered predicate logic, the sorts are interpreted as categories,
+functions as functors, relations as profunctors and implications of
+relations as natural transformations.
+%
+The type theoretic syntax allows us to write all of our constructions
+as if they were ordinary functors and profunctors in set-based
+category theory, but the proofs are valid in an arbitrary virtual
+equipments and so are valid in enriched and internal category theory
+as well.
+%
+This is analogous to the situation in intuitionistic type theory,
+where the syntax is similar to that of informal set theory, but by
+restricting to intuitionistic reasoning, the constructions can all be
+interpreted in models where types are spaces and terms are continuous
+functions ``for free''.
+
+Ordered predicate logic is a kind of \emph{ordered},
+\emph{constructive} variant of predicate logic over a unary type
+theory.
+%
+It is \emph{ordered} in that our notion of predicate is that of a
+binary relation $R(\alpha;\beta)$ where the $\alpha$ is thought of as
+a contravariant dependence and the $\beta$ has a covariant dependence.
+%
+Proofs in the predicate logic $R_1, R_2,\ldots \vdash S$ are
+restricted as in ordered linear logic \cite{lambek-calculus, ordered-linear-logic} in
+that each input must be used exactly once and in the order they were
+declared.
+%
+This order discipline ensures that every proof in the logic can be
+interpreted as a \emph{natural} transformation.
+%
+The logic is \emph{constructive} in that we care about the identity of
+proofs: they are natural transformations and there may be more than
+one natural transformation of the same type.
+%
+If we ignore the identity of proofs, ordered predicate logic can be
+viewed as a logic of preordered sets rather than categories.
+
+The restriction to unary type theory and binary relations is fairly
+strict: we cannot discuss important concepts such as functor
+categories, opposite categories and arbitrary (co)-ends of
+profunctors. However, the restriction is natural in that it
+corresponds to a well-understood notion of model (virtual equipments)
+and so we leave the extension to future work.
+%
+We show that a great deal of fundamental results in category theory
+are already expressible in this fragment.
+
+Due to the restriction to unary functors and binary profunctors, it is
+difficult to state some theorems internally to the theory. For
+instance, since we cannot express functor categories, we cannot
+express any theorem that quantifies over functors purely in ordered
+predicate logic, and instead only as a metatheorem. To more easily
+formalize these metatheorems, we add a third layer to the type theory:
+an ordinary dependent type theory in the style of Martin-L\"of type
+theory. Then all of our ordered predicate logic judgments will also be
+indexed by a context from this dependent type theory, and the type
+theory includes types for categories, functors, profunctors and
+natural transformations that allow us to formalize theorems the logic
+is too restrictive to encode.
+
+The dependent type theory layer also has an additional purpose: since
+we can quantify over the structures in ordered predicate logic, we can
+additionally encode \emph{signatures} for ordered predicate logic as
+simply contexts of the dependent type theory. This allows for a clean
+formalization of many categorical structures such as monoidal
+categories, or strong monads.
+
+While we emphasize the applications to enriched and internal category
+theory in this work, there is potential for more direct application to
+programming language semantics.
+%
+Ordinary predicate logic has been used to develop logics in which
+logical relations proofs can be developed such as Abadi-Plotkin logic
+for parametricity and LSLR for step-indexed logical relations proofs
+\cite{abadi-plotkin,lslr}.
+%
+Many \emph{binary} logical relations are based on ordered structures,
+such as logical relations for approximation in non-deterministic or
+partial languages or logical relations for graduality in gradual
+typing.
+%
+We conjecture that some extension of the ordered predicate logic
+defined here will be useful for describing these relations as well.
+
+\subsection{Basics of Profunctors}
+
+While we assume the reader has some background knowledge of category
+theory, we briefly define profunctors, which are not included in many
+introductory texts. First, a category $\cat C$ has a collection of
+objects and morphisms with identity and composition, and a functor $F
+: \cat C \to \cat D$ is a function on objects and a function on
+morphisms that respects source and target. It will be useful for
+intuition to think of a category as a generalization of a preordered
+set, which has a set of elements and a binary \emph{relation} on its
+objects satisfying reflexivity and transitivity. It is fruitful to
+think of a category as a \emph{proof-relevant preorder}, where
+morphisms are thought of as proofs of ordering and the reflexivity and
+transitivity constructions must additionally satisfy equations that
+make them act like identity and composition. Then a functor is a
+constructive version of a monotone function.
+
+Given categories $\mathcal C$ and $\mathcal D$, a profunctor $R$ from
+$\mathcal C$ to $\mathcal D$, written $R : \cat C \pto \cat D$ is a
+functor $R : {\cat C}^{op} \times {\cat D} \to \Set$. We can think of
+each set $R(c,d)$ as a set of ``heteromorphisms'' from $c$, an object
+of $\cat C$ to $d$, an object of $\cat D$. Functoriality then defines
+a notion of composition of heteromorphisms with homomorphisms in $\cat
+C$ and $\cat D$. Thinking of categories as constructive preorders, a
+profunctor is a constructive relation that is downward-closed in $\cat
+C$ and upward-closed in $\cat D$, a kind of ``hetero-ordering''.
+%
+Given profunctors $R, S : \cat C \pto \cat D$, a homomorphism from
+$R$ to $S$ is simply a natural transformation, which in the preordered
+setting is simply an implication of relations.
+
+The simplest example of a profunctor is for any category $\cat C$,
+$\text{Hom}_{\cat C} : \cat C \pto \cat C$ is a profunctor. On
+preorders this is just the ordering itself. The hom profunctor is the
+unit for a composition of profunctors $R \odot S$ which is defined as
+a \emph{co-end}. The heteromorphisms of $(R \odot S)(c,e)$ are triples
+$(d \in \cat D, f \in R(c,d), g \in S(d,e))$ modulo the equivalence
+relation generated by $(d,f,g) ~ (d', f', g')$ if there exists $h \in
+D(d,d')$ such that $h\circ f = f'$ and $g'\circ h = g$. The
+composition of profunctors is a generalization of the composition of
+relations, and the hom profunctor is analogous in the same way to the
+diagonal, or equality relation in that it is the identity for this
+composition.
+For any profunctor $R : \cat C \pto \cat D$ and functors $F : \cat
+C' \to \cat C$ and $G : \cat D' \to \cat D$, by composition we can
+define a profunctor $R(F,G)$, the ``restriction'' of $R$ along $F$ and
+$G$.
+
+Profunctors are not explicitly defined in most introductory texts on
+category theory but arise implicitly in formulations of universal
+properties such a representable functors and adjunctions.
 
@@
   syntax to demonstrate the ``naive''/synthetic character of proofs
-  and constructions in SCT. We additionally give some examples of how
-  this can be applied to model a recent formulation of modal dependent
-  type theory.
+  and constructions in SCT.
 
@@
 
-\item In \cref{sec:semantics}, we show that the syntax of SCT is sound
-  for models in an arbitrary virtual double category with
-  corresponding universal constructions.
+\item In \cref{sec:semantics}, we develop some basic model theory for
+  ordered predicate logic, including a sound and complete notion of
+  categorical model and sound interpretation in virtual equipments
+  modeling ordinary, enriched and internal category theory.
 
@@
 
-We have aimed that the syntax and examples sections should be
-accessible to anyone familiar with basic category theory and
-willingness to understand profunctors and (co-)ends.
-%
-The semantics section is by necessity more advanced and may be skipped
-or skimmed by readers who are willing to trust the mathematical
-foundations of the system.
+\section{Syntax of SCT}
 
@@
 relationship between the judgments and connectives of higher-order
-logic with our category logic. Since category logic is an ordered
-variant of higher-order logic, some notions splinter into
-multiple. For instance, covariant and contravariant presheaf
+predicate logic with our ordered variant. Due to the incorporation of
+variance, some unorderd concepts generalize to multiple different
+ordered notions. For instance, covariant and contravariant presheaf
 categories generalize the power set. Further, because we only have
@@
 only restricted forms of universal and existential quantification
-which combined with implications and conjunctions.
-
-\section{Syntax of SCT}
+which come combined with implications and conjunctions.
 
@@
     \begin{align*}
-      (\harr P Q)^* &= \harr {P^*} {Q^*}\\
+      (\harr {\cat C} {\cat D})^* &= \harr {\cat D^*} {\cat C^*}\\
       (\alpha:\cat C)^* &= \alpha: \cat C^*\\
@@
       (\tensorintroatXwithYandZ b s t)^* &= \tensorintroatXwithYandZ {b^*} {t^*} {s^*}\\
-      (\punitinXfromYtoZ {\cat C} {a_1}{a_2})^* &= \punitinXfromYtoZ {\cat C^*} {a_1^*}{a_2^*}
+      (\punitinXfromYtoZ {\cat C} {a_1}{a_2})^* &= \punitinXfromYtoZ {\cat C^*} {a_2^*}{a_1^*}
     \end{align*}
@@
 \section{Semantics}
-Next, we develop the basics of the model theory for \ohol{}. Being a
-2-level type theory, \ohol{} has models that are a kind of
-\emph{hyperdoctrine}. First, a model must include a model of
-intuitionistic type theory to interpret the type-level context
-$\Gamma$ and corresponding type/term judgments. Next, there should be
-a functor from the category of semantic contexts and substitutions to
-a category of models of ordered higher-order logic.
-
-\subsection{Soundness and Completeness}
+
+Next, we develop the basics of the model theory for \ohol{}.
+%
+First, we demonstrate that a hyperdoctrine of virtual double
+categories with split restrictions provides a sound and complete
+notion of categorical model.
+%
+Then we instantiate this general notion of model to show that the
+\ohol{} can be interpreted in ordinary category theory as well as
+enriched and internal category theory, and more trivially, in terms of
+sets and relations or preordered sets and ordered relations.
+
+\subsection{Categorical Models}
+
+First, we can model the judgmental structure of the unary type theory
+and predicate logic in certain\emph{virtual double categories}
+\cite{leinster,crutwell-shulman}. We briefly recount the structure
+present in a virtual double category, but see \citet{crutwell-shulman}
+for a precise definition.
+The following definition is
+equivalent to the formulation as a ``normalized $T$-monoid'' in
+\citet{crutwell-shulman}, and we direct the reader there for the full
+description.
+%
+
+\begin{definition}[Virtual Double Category]
+  A virtual double category $\mathcal V$ consists of
+  \begin{enumerate}
+  \item A category $V_o$ of ``objects and vertical arrows''
+  \item A set ${\mathcal V}_h$ of ``horizontal arrows'' with source
+    and target functions $s,t : {\mathcal V}_h \to {\mathcal V_o}^2$
+  \item Sets of 2-cells of the following form:
+% https://q.uiver.app/?q=WzAsNixbMCwwLCJDXzAiXSxbMCwyLCJEXzAiXSxbMiwwLCJDX3tufSJdLFsxLDAsIlxcY2RvdHMiXSxbMiwyLCJEXzEiXSxbMSwxLCJcXHBoaSJdLFszLDIsIlJfbiIsMCx7InN0eWxlIjp7ImJvZHkiOnsibmFtZSI6ImJhcnJlZCJ9fX1dLFswLDEsIkYiLDJdLFsyLDQsIkciXSxbMSw0LCJTIiwyLHsic3R5bGUiOnsiYm9keSI6eyJuYW1lIjoiYmFycmVkIn19fV0sWzAsMywiUl8wIiwwLHsic3R5bGUiOnsiYm9keSI6eyJuYW1lIjoiYmFycmVkIn19fV1d
+\[\begin{tikzcd}
+	{C_0} & \cdots & {C_{n}} \\
+	& \phi \\
+	{D_0} && {D_1}
+	\arrow["{R_n}", "\shortmid"{marking}, from=1-2, to=1-3]
+	\arrow["F"', from=1-1, to=3-1]
+	\arrow["G", from=1-3, to=3-3]
+	\arrow["S"', "\shortmid"{marking}, from=3-1, to=3-3]
+	\arrow["{R_0}", "\shortmid"{marking}, from=1-1, to=1-2]
+\end{tikzcd}\]
+  \end{enumerate}
+  We say a virtual double category has \emph{split restrictions} when
+  for any horizontal arrow $R : C \pto D$ and vertical arrows $f : C'
+  \to C$ and $g : D' \to D$ there is a chosen profunctor $R(f,g) : C'
+  \pto D'$ with a cartesian 2-cell to $R$ framed by $f,g$ and these
+  chosen cartesian lifts are functorial in $f,g$.  A functor of
+  virtual double categories preserves split restrictions if it
+  preserves these specified cartesian lifts.  
+\end{definition}
+
+Each component of this definition has a direct correspondence to a
+syntactic structure in \ohol{}. The category $\mathcal V_o$ models the
+category and functor judgments. The set $\mathcal V_h$ models the
+profunctor judgment. A composable string $R_0 \cdots R_n$ models the
+profunctor contexts. The 2-cells correspond to the natural
+transformation judgment where we have taken the restriction $S(F,G)$
+of the codomain.
+
+To model the dependent type theory and fact that the \ohol{} judgments
+are all under a dependent type theory context, we use a variation on
+Lawvere's notion of \emph{hyperdoctrine} for modeling predicate logic:
+\begin{definition}[Hyperdoctrine of Virtual Double Categories with Restrictions]
+  A hyperdoctrine of virtual double categories with restrictions is a
+  pair of a category with families $\mathcal C$ and a functor $L :
+  \mathcal C^o \to \text{vDbl}_r$ to the category of virtual double
+  categories with split restrictions and restriction-preserving
+  functors.
+\end{definition}
+Categories with families $\mathcal C$ model dependent type theory
+\citep{cwf} and for each semantic context $\Gamma$, $L(\Gamma)$ models
+the \ohol{} judgments in context $\Gamma$, with the functoriality
+modeling the fact that all of these judgments admit a well-behaved
+action of substitution.
+
+\begin{construction}[Soundness and Completeness]
+  For any choice of connectives, the syntax of parameterized \ohol{}
+  defines a hyperdoctrine of VDCrs that is an initial object in the
+  category of hyperdoctrines of VDCrs with chosen universal
+  constructions and functors preserving those universal constructions.
+\end{construction}
+\begin{proof}
+  Most of the details of the proof are straightforward. In the
+  appendix we demonstrate the syntactic elements of the proof which
+  are that the syntax itself is a hyperdoctrine of VDCrs, i.e., that
+  the notions of substitution are appropriately functorial.
+\end{proof}
 
@@
 
-Because SCT
-includes not just a syntax for category theory, but is also indexed by
-an ordinary context of dependent type theory $\Gamma$, a model of SCT
-is not given by a \emph{single} virtual double category, but instead
-by an \emph{indexed family} of VDCs, parameterized by an object of a
-category that interprets dependent type theory. We call such a
-structure a \emph{VDC-hyperdoctrine}. We then show how to construct a
-from any VDC, a VDC-hyperdoctrine indexed by the category of sets,
-which has the intuitive reading that any of the VDC judgments $\Gamma
-\vdash J$ is interpreted as a family of $J$-data parameterized
-by elements of the set $\Gamma$.
+Now that we have a category-theoretic notion of model, we give some
+model construction theorems that can be used to justify our intuitive
+notion of semantics in (enriched, internal) category theory.
+
+First, we can extend any model of \ohol{} to a model of \ohol{}
+parameterized by dependent type theory as follows:
+\begin{construction}
+  Given a virtual double category with restrictions $\mathcal V$, we
+  can construct a functor $\mathcal V^{-}: \Set \to \text{vDbl}_r$ by
+  defining the objects of $\mathcal V^\Gamma$ to be functions $\Gamma
+  \to \mathcal V_o$, and similarly for morphisms and 2-cells with all
+  operations given pointwise.
+\end{construction}
+% NOTE: this should be some kind of general theorem right?
+% Sketch:
+% 1. VDCrs are a finite limit theory
+% 2. A hyperdoctrine L : C -> VDCr is the same as a VDCr in Pr(C)
+% 3. The Yoneda embedding preserves finite limits
+
+TODO: find a reference that proves things have all the UMPs we want
+lol.
 
@@
 
-%% \section{Typesetting instructions -- Summary}
-%% \label{sec:typesetting-summary}
-
-%% LIPIcs is a series of open access high-quality conference proceedings across all fields in informatics established in cooperation with Schloss Dagstuhl. 
-%% In order to do justice to the high scientific quality of the conferences that publish their proceedings in the LIPIcs series, which is ensured by the thorough review process of the respective events, we believe that LIPIcs proceedings must have an attractive and consistent layout matching the standard of the series.
-%% Moreover, the quality of the metadata, the typesetting and the layout must also meet the requirements of other external parties such as indexing service, DOI registry, funding agencies, among others.
-The guidelines contained in this document serve as the baseline for the authors, editors, and the publisher to create documents that meet as many different requirements as possible.
-
-%% Please comply with the following instructions when preparing your article for a LIPIcs proceedings volume. 
-%% \paragraph*{Minimum requirements}
-
-%% \begin{itemize}
-%% \item Use pdflatex and an up-to-date \LaTeX{} system.
-%% \item Use further \LaTeX{} packages and custom made macros carefully and only if required.
-%% \item Use the provided sectioning macros: \verb+\section+, \verb+\subsection+, \verb+\subsubsection+, \linebreak \verb+\paragraph+, \verb+\paragraph*+, and \verb+\subparagraph*+.
-%% \item Provide suitable graphics of at least 300dpi (preferably in PDF format).
-%% \item Use BibTeX and keep the standard style (\verb+plainurl+) for the bibliography.
-%% \item Please try to keep the warnings log as small as possible. Avoid overfull \verb+\hboxes+ and any kind of warnings/errors with the referenced BibTeX entries.
-%% \item Use a spellchecker to correct typos.
-%% \end{itemize}
-
-%% \paragraph*{Mandatory metadata macros}
-%% Please set the values of the metadata macros carefully since the information parsed from these macros will be passed to publication servers, catalogues and search engines.
-%% Avoid placing macros inside the metadata macros. The following metadata macros/environments are mandatory:
-%% \begin{itemize}
-%% \item \verb+\title+ and, in case of long titles, \verb+\titlerunning+.
-%% \item \verb+\author+, one for each author, even if two or more authors have the same affiliation.
-%% \item \verb+\authorrunning+ and \verb+\Copyright+ (concatenated author names)\\
-%% The \verb+\author+ macros and the \verb+\Copyright+ macro should contain full author names (especially with regard to the first name), while \verb+\authorrunning+ should contain abbreviated first names.
-%% \item \verb+\ccsdesc+ (ACM classification, see \url{https://www.acm.org/publications/class-2012}).
-%% \item \verb+\keywords+ (a comma-separated list of keywords).
-%% \item \verb+\relatedversion+ (if there is a related version, typically the ``full version''); please make sure to provide a persistent URL, e.\,g., at arXiv.
-%% \item \verb+\begin{abstract}...\end{abstract}+ .
-%% \end{itemize}
-
-%% \paragraph*{Please do not \ldots} %Do not override the \texttt{\seriesstyle}-defaults}
-%% Generally speaking, please do not override the \texttt{lipics-v2021}-style defaults.
-To be more specific, a short checklist also used by Dagstuhl Publishing during the final typesetting is given below.
-%% In case of \textbf{non-compliance} with these rules Dagstuhl Publishing will remove the corresponding parts of \LaTeX{} code and \textbf{replace it with the \texttt{lipics-v2021} defaults}.
-In serious cases, we may reject the LaTeX-source and expect the corresponding author to revise the relevant parts.
-%% \begin{itemize}
-%% \item Do not use a different main font. (For example, the \texttt{times} package is forbidden.)
-%% \item Do not alter the spacing of the \texttt{lipics-v2021.cls} style file.
-%% \item Do not use \verb+enumitem+ and \verb+paralist+. (The \texttt{enumerate} package is preloaded, so you can use
-%%  \verb+\begin{enumerate}[(a)]+ or the like.)
-%% \item Do not use ``self-made'' sectioning commands (e.\,g., \verb+\noindent{\bf My+ \verb+Paragraph}+).
-%% \item Do not hide large text blocks using comments or \verb+\iffalse+ $\ldots$ \verb+\fi+ constructions. 
-%% \item Do not use conditional structures to include/exclude content. Instead, please provide only the content that should be published -- in one file -- and nothing else.
-%% \item Do not wrap figures and tables with text. In particular, the package \texttt{wrapfig} is not supported.
-%% \item Do not change the bibliography style. In particular, do not use author-year citations. (The
-%% \texttt{natbib} package is not supported.)
-%% \end{itemize}
-
-%% \enlargethispage{\baselineskip}
-
-%% This is only a summary containing the most relevant details.
-Please read the complete document ``LIPIcs: Instructions for Authors and the \texttt{lipics-v2021} Class'' for all details and don't hesitate to contact Dagstuhl Publishing (\url{mailto:publishing@dagstuhl.de}) in case of questions or comments:
-%% \href{http://drops.dagstuhl.de/styles/lipics-v2021/lipics-v2021-authors/lipics-v2021-authors-guidelines.pdf}{\texttt{http://drops.dagstuhl.de/styles/lipics-v2021/\newline lipics-v2021-authors/lipics-v2021-authors-guidelines.pdf}}
-
-%% \section{Lorem ipsum dolor sit amet}
-
-%% Lorem ipsum dolor sit amet, consectetur adipiscing elit \cite{DBLP:journals/cacm/Knuth74}.
-Praesent convallis orci arcu, eu mollis dolor.
-Aliquam eleifend suscipit lacinia.
-Maecenas quam mi, porta ut lacinia sed, convallis ac dui.
-Lorem ipsum dolor sit amet, consectetur adipiscing elit.
-Suspendisse potenti.
-Donec eget odio et magna ullamcorper vehicula ut vitae libero.
-Maecenas lectus nulla, auctor nec varius ac, ultricies et turpis.
-Pellentesque id ante erat.
-In hac habitasse platea dictumst.
-Curabitur a scelerisque odio.
-Pellentesque elit risus, posuere quis elementum at, pellentesque ut diam.
-Quisque aliquam libero id mi imperdiet quis convallis turpis eleifend.
-
-%% \begin{lemma}[Lorem ipsum]
-%% \label{lemma:lorem}
-%% Vestibulum sodales dolor et dui cursus iaculis.
-Nullam ullamcorper purus vel turpis lobortis eu tempus lorem semper.
-Proin facilisis gravida rutrum.
-Etiam sed sollicitudin lorem.
-Proin pellentesque risus at elit hendrerit pharetra.
-Integer at turpis varius libero rhoncus fermentum vitae vitae metus.
-%% \end{lemma}
-
-%% \begin{proof}
-%% Cras purus lorem, pulvinar et fermentum sagittis, suscipit quis magna.
-
-%% \proofsubparagraph*{Just some paragraph within the proof.}
-%% Nam liber tempor cum soluta nobis eleifend option congue nihil imperdiet doming id quod mazim placerat facer possim assum.
-Lorem ipsum dolor sit amet, consectetuer adipiscing elit, sed diam nonummy nibh euismod tincidunt ut laoreet dolore magna aliquam erat volutpat.
-%% \begin{claim}
-%% content...
-%% \end{claim}
-%% \begin{claimproof}
-%% content...
-%%     \begin{enumerate}
-%%         \item abc abc abc \claimqedhere{}
-%%     \end{enumerate}
-%% \end{claimproof}
-
-%% \end{proof}
-
-%% \begin{corollary}[Curabitur pulvinar, \cite{DBLP:books/mk/GrayR93}]
-%% \label{lemma:curabitur}
-%% Nam liber tempor cum soluta nobis eleifend option congue nihil imperdiet doming id quod mazim placerat facer possim assum.
-Lorem ipsum dolor sit amet, consectetuer adipiscing elit, sed diam nonummy nibh euismod tincidunt ut laoreet dolore magna aliquam erat volutpat.
-%% \end{corollary}
-
-%% \begin{proposition}\label{prop1}
-%% This is a proposition
-%% \end{proposition}
-
-%% \autoref{prop1} and \cref{prop1} \ldots
-
-%% \subsection{Curabitur dictum felis id sapien}
-
-%% Curabitur dictum \cref{lemma:curabitur} felis id sapien \autoref{lemma:curabitur} mollis ut venenatis tortor feugiat.
-Curabitur sed velit diam.
-Integer aliquam, nunc ac egestas lacinia, nibh est vehicula nibh, ac auctor velit tellus non arcu.
-Vestibulum lacinia ipsum vitae nisi ultrices eget gravida turpis laoreet.
-Duis rutrum dapibus ornare.
-Nulla vehicula vulputate iaculis.
-Proin a consequat neque.
-Donec ut rutrum urna.
-Morbi scelerisque turpis sed elit sagittis eu scelerisque quam condimentum.
-Pellentesque habitant morbi tristique senectus et netus et malesuada fames ac turpis egestas.
-Aenean nec faucibus leo.
-Cras ut nisl odio, non tincidunt lorem.
-Integer purus ligula, venenatis et convallis lacinia, scelerisque at erat.
-Fusce risus libero, convallis at fermentum in, dignissim sed sem.
-Ut dapibus orci vitae nisl viverra nec adipiscing tortor condimentum \cite{DBLP:journals/cacm/Dijkstra68a}.
-Donec non suscipit lorem.
-Nam sit amet enim vitae nisl accumsan pretium.
-
-%% \begin{lstlisting}[caption={Useless code.},label=list:8-6,captionpos=t,float,abovecaptionskip=-\medskipamount]
-%% for i:=maxint to 0 do 
-%% begin 
-%%     j:=square(root(i));
-%% end;
-%% \end{lstlisting}
-
-%% \subsection{Proin ac fermentum augue}
-
-%% Proin ac fermentum augue.
-Nullam bibendum enim sollicitudin tellus egestas lacinia euismod orci mollis.
-Nulla facilisi.
-Vivamus volutpat venenatis sapien, vitae feugiat arcu fringilla ac.
-Mauris sapien tortor, sagittis eget auctor at, vulputate pharetra magna.
-Sed congue, dui nec vulputate convallis, sem nunc adipiscing dui, vel venenatis mauris sem in dui.
-Praesent a pretium quam.
-Mauris non mauris sit amet eros rutrum aliquam id ut sapien.
-Nulla aliquet fringilla sagittis.
-Pellentesque eu metus posuere nunc tincidunt dignissim in tempor dolor.
-Nulla cursus aliquet enim.
-Cras sapien risus, accumsan eu cursus ut, commodo vel velit.
-Praesent aliquet consectetur ligula, vitae iaculis ligula interdum vel.
-Integer faucibus faucibus felis.
-
-%% \begin{itemize}
-%% \item Ut vitae diam augue. 
-%% \item Integer lacus ante, pellentesque sed sollicitudin et, pulvinar adipiscing sem. 
-%% \item Maecenas facilisis, leo quis tincidunt egestas, magna ipsum condimentum orci, vitae facilisis nibh turpis et elit. 
-%% \end{itemize}
-
-%% \begin{remark}
-%% content...
-%% \end{remark}
-
-%% \section{Pellentesque quis tortor}
-
-%% Nec urna malesuada sollicitudin.
-Nulla facilisi.
-Vivamus aliquam tempus ligula eget ornare.
-Praesent eget magna ut turpis mattis cursus.
-Aliquam vel condimentum orci.
-Nunc congue, libero in gravida convallis \cite{DBLP:conf/focs/HopcroftPV75}, orci nibh sodales quam, id egestas felis mi nec nisi.
-Suspendisse tincidunt, est ac vestibulum posuere, justo odio bibendum urna, rutrum bibendum dolor sem nec tellus.
-
-%% \begin{lemma} [Quisque blandit tempus nunc]
-%% Sed interdum nisl pretium non.
-Mauris sodales consequat risus vel consectetur.
-Aliquam erat volutpat.
-Nunc sed sapien ligula.
-Proin faucibus sapien luctus nisl feugiat convallis faucibus elit cursus.
-Nunc vestibulum nunc ac massa pretium pharetra.
-Nulla facilisis turpis id augue venenatis blandit.
-Cum sociis natoque penatibus et magnis dis parturient montes, nascetur ridiculus mus.
-%% \end{lemma}
-
-%% Fusce eu leo nisi. Cras eget orci neque, eleifend dapibus felis. Duis et leo dui. Nam vulputate, velit et laoreet porttitor, quam arcu facilisis dui, sed malesuada risus massa sit amet neque.
-
-%% \section{Morbi eros magna}
-
-%% Morbi eros magna, vestibulum non posuere non, porta eu quam.
-Maecenas vitae orci risus, eget imperdiet mauris.
-Donec massa mauris, pellentesque vel lobortis eu, molestie ac turpis.
-Sed condimentum convallis dolor, a dignissim est ultrices eu.
-Donec consectetur volutpat eros, et ornare dui ultricies id.
-Vivamus eu augue eget dolor euismod ultrices et sit amet nisi.
-Vivamus malesuada leo ac leo ullamcorper tempor.
-Donec justo mi, tempor vitae aliquet non, faucibus eu lacus.
-Donec dictum gravida neque, non porta turpis imperdiet eget.
-Curabitur quis euismod ligula.
-
 %% %%
@@
 \subsubsection*{Acknowledgements}
-Grants stuff...? The authors would like to thank Emily Riehl, Mike
-Shulman,...for helpful feedback on this work
+Grants stuff...? The authors would like to thank David Jaz Myers, Emily Riehl, Mike
+Shulman, Dominic Verity,...for helpful feedback on this work
 

### 2022-09-07 paper/paper.tex
related work

@@
 
-\paragraph{Calculi for 2-dimensional category theory}
+\paragraph{Calculi for formal category theory}
+
+In their book on $\infty$-category theory, Riehl and Verity use the
+formal language of virtual equipments to prove results valid for
+$\infty$-categories without concrete manipulation of model
+categories. They formalize this language using Makkai's FOLDS
+framework of first-order logic with dependent sorts. The approach is
+similar in spirit to our \ohol{}, and has the same models. The main
+difference is that their theory is built on top of first-order logic,
+whereas ours is built on a custom-made variant of higher-order logic
+over a (unary) simple type theory. For this reason, statements in
+their theory are quite complex as they have to be encoded into
+first-order logic, whereas we argue our type-theoretic approach
+formalizes informal arguments much more directly, as shown in
+Section~\ref{sec:examples}.
+%
+An informal syntax for manipulating profunctors involving ends and
+coends, called the coend calculus is closely related, an extension of
+\ohol{} to treat profunctors of many variables of different variances
+should provide a formal treatment for the coend calculus
+\citet{coend-calculus}.
+
+Another approach to formal category theory is to develop a
+\emph{directed dependent type theory}, a type theory where the
+identity type is interpreted as morphisms in a (possibly
+$\infty$-)category. Research on directed dependent type theories is
+ongoing \cite{riehl-shulman,north}. Directed dependent type theories
+should generalize \ohol{} in the same way that dependent type theory
+generalizes higher order logic: internally there should be a type that
+models the category of sets, and terms of this type should model the
+profunctors of \ohol{}. As a foundation for a domain-specific proof
+assistant for category theory, \ohol{} may provide a foundation for a
+proof assistant in the style of Isabelle/HOL \cite{isabelle-hol},
+whereas directed type theory would provide a foundation for a proof
+assistant in the style of Agda or Coq \cite{agda,coq}.
 
@@
 
-Our type-theoretic syntax for virtual double categories is intended to
-be a formal notation for the manipulation of profunctors. As such it
-has similarities to informal syntax for manipulating profunctors
-involving ends and coends, such as the coend calculus of
-\citet{coend-calculus}.
-
 Finally, some variations on double categories have been used to model
@@
 like double categories without horizontal composition.
-
-Double-cat-like syntax:
-Gradual Type Theory, System P of parametricity, cubical stuff?
-
-\subsection{Directed Type Theory}
-
-Directed Type theory
-
-Fibrational approaches to logical relations
 

### 2022-09-07 paper/paper.tex
talk about extensionality

@@
 \end{figure}
+
+\subsection{Extensionality Axioms}
+
+Finally, we discuss some non-theorems that would be desirable to add
+as axioms to the theory of \ohol{}.
+%
+Chief among these are principles about the hom type $\punitinXfromYtoZ
+{\cat C} \alpha \beta$ when $\cat C$ is a connective.
+%
+Since the hom type is defined as a kind of unit type, rather than
+defined inductively, we are not guaranteed that the hom type for the
+product category $\cat C \times \cat D$ or the presheaf categories are
+what we expect.
+%
+For the product type, we expect there should be an isomorphism
+$\punitinXfromYtoZ {\cat C \times \cat D} \alpha \beta \cong
+(\punitinXfromYtoZ {\cat C} {\pi_1 \alpha} {\pi_1 \beta}) \times
+(\punitinXfromYtoZ {\cat D} {\pi_2 \alpha} {\pi_2 \beta})$ saying an
+arrow in a product category is a pair of arrows.
+%
+Similarly, the hom type in the presheaf category should be given by
+the end formula $\punitinXfromYtoZ {\negPresheaf {\cat C}} \alpha
+\beta \cong \homlallXYtoZ {\gamma}{\negPresheafApp \alpha
+  \gamma}{\negPresheafApp \beta \gamma}$. This would allow us to
+relate the Yoneda Lemma \ref{lem:yoneda} to the morphisms of
+presheaves, and prove that the Yoneda embedding is fully faithful.
+%
+We call these extensionality principles because they are analogous to
+principles such as function extensionality but for the directed hom
+type rather than equality.
+
+The presheaf extensionality principle is independent from the existing
+axioms of \ohol{}: it is true in the ordinary category theoretic
+model, but it is false in the set theoretic model. That is, in the set
+theoretic model, the presheaf category is simply the powerset, the hom
+type is \emph{equality} and the function profunctor is interpreted as
+a universally quantified \emph{implication}, but equality in the
+powerset should be \emph{bi-implication}, so the theorem is false. We
+expect the product extensionality is independent as well, but a
+counter-model would likely be quite exotic as we expect it to be
+provable in a system with functor categories, in analogy with the
+proof for the identity type in MLTT.
+%
+Note that these axioms can easily be added as assumptions in the
+dependently typed context, though this will break the constructive
+character of the system. We expect techniques from type theory might
+be applied to incorporate these as computationally meaningful
+principles rather than assumed as axioms \cite{obs-tt,xtt,cubical-type-theory}.
 

### 2022-09-21 paper/paper.tex
intro tweaks, start working on shortening the syntax section

@@
 %% \newtheorem{definition}{Definition}
+\newcommand{\vett}{VETT}
 \newcommand{\ohol}{OHol}
@@
 \newcommand{\negPresheafApp}[2]{#1\,#2}
-\newcommand{\negPresheafAppPtoX}[2]{#2\,#1}
+\newcommand{\negPresheafAppPtoX}[2]{#2\in #1}
+\newcommand{\posPresheafAppPtoX}[2]{#1 \ni #2}
 \newcommand{\posPresheafApp}[2]{#1\,#2}
+
+\newcommand{\graphProf}[3]{\sum_{{#1};{#2}} #3}
 
@@
 
-\title{An Ordered Logic for Category Theory}
+\title{A Constructive Ordered Predicate Logic for Category Theory}
 \author{Max S. New\inst{1,2} and Daniel R. Licata\inst{2}}
@@
 
-  We develop a sound and complete notion of categorical model and give
-  a model construction that shows that the theorems are valid in
-  enriched and internal categories. This means that while the proofs
-  in our type theory look like standard set-based arguments, the
-  substructural syntactic restrictions ensure that the proofs carry
-  over to enriched and internal settings as well.
+  We develop a sound and complete notion of categorical model for the
+  type theory based on virtual equipments, which are known to model
+  both internal and enriched category theory. This means that while
+  the proofs in our type theory look like standard set-based
+  arguments, the syntactic discipline ensure that all proofs and
+  constructions carry over to enriched and internal settings as well.
 \end{abstract}
@@
 Originally developed for applications to homological algebra, it was
-quickly discovered that these categorical structures were common in
-logic and computer science, and that formal systems like logics, type
-theories and programming languages can be modeled by sound and
-complete semantic models given by notions of structured
-category\cite{lawvere,goguen,lambek,moggi}, a phenomenon sometimes
-referred to as the Curry-Howard-Lambek correspondence
-\cite{curry-howard-lambek}.
+quickly discovered that (among many other applications) these
+categorical structures were common in logic and computer science, and
+that formal systems like logics, type theories and programming
+languages can be modeled by sound and complete semantic models given
+by notions of structured category\cite{lawvere,goguen,lambek,moggi}, a
+phenomenon sometimes referred to as the Curry-Howard-Lambek
+nncorrespondence \cite{curry-howard-lambek}.
 %
@@
 These seminal works have led to a fruitful field of research known as
-categorical logic, where
-new type theories are designed based directly on category-theoretic
-semantics.
+categorical logic, where new type theories are designed based directly
+on category-theoretic semantics.
 %
@@
 parametricity and gradual
-typing\cite{order-enriched-cats,topos-of-trees,reflexive-graphs,double-cats-gradual-typing}.
+typing\cite{order-enriched-cats,topos-of-trees,reflexive-graphs,double-cats-gradual-typing},
+to name a few.
 
@@
 algebraic structures to design a formal logic, which we call
-\emph{ordered predicate logic}, that can be soundly interpreted in
-virtual equipments.
-%
-In ordered predicate logic, the sorts are interpreted as categories,
-functions as functors, relations as profunctors and implications of
-relations as natural transformations.
+\emph{virtual equipment type theory}, that can be soundly interpreted
+in virtual equipments.
+%
+In \vett{}, the sorts are interpreted as categories, functions as
+functors, relations as profunctors and implications of relations as
+profunctor homomorphisms, a generalization of natural transformations.
 %
@@
 
-Ordered predicate logic is a kind of \emph{ordered},
-\emph{constructive} variant of predicate logic over a unary type
-theory.
+\vett{} is a kind of \emph{ordered}, \emph{constructive} variant of
+predicate logic over a unary type theory.
 %
@@
 binary relation $R(\alpha;\beta)$ where the $\alpha$ is thought of as
-a contravariant dependence and the $\beta$ has a covariant dependence.
+a contravariant dependence and the $\beta$ has a covariant dependence,
+and the rules of the logic distinguish between these two variances.
 %
@@
 %
-We show that a great deal of fundamental results in category theory
-are already expressible in this fragment.
+It is well known that a great deal of fundamental results and
+constructions in category theory such as Yoneda and co-Yoneda lemmas,
+adjoint functors and Kan extensions are already expressible in virtual
+equipments, for instance virtual equipments are used for this purpose
+in a text on $\infty$-category theory to avoid explicit manipulation
+of model categories \cite{elements-of-oo-category-theory}.
+%
+We demonstrate some of these standard constructions
+in \ref{sec:synthetic} to show that \vett{} gives a syntax in which
+these formal constructions look like standard set-theoretic arguments.
 
@@
 natural transformations that allow us to formalize theorems the logic
-is too restrictive to encode.
+is too restrictive to encode. This is analogous to the notion of a
+\emph{2-level type theory}: we add a dependent type theory for
+reasoning about the terms of our virtual equipment type theory.
 
 The dependent type theory layer also has an additional purpose: since
-we can quantify over the structures in ordered predicate logic, we can
-additionally encode \emph{signatures} for ordered predicate logic as
-simply contexts of the dependent type theory. This allows for a clean
-formalization of many categorical structures such as monoidal
-categories, or strong monads.
+we can quantify over the structures in \vett{}, we can additionally
+encode \emph{signatures} for \vett{} as simply contexts of the
+dependent type theory. This allows for a clean formalization of many
+categorical structures such as monoidal categories, or strong monads.
 
@@
 logical relations proofs can be developed such as Abadi-Plotkin logic
-for parametricity and LSLR for step-indexed logical relations proofs
-\cite{abadi-plotkin,lslr}.
-%
-Many \emph{binary} logical relations are based on ordered structures,
-such as logical relations for approximation in non-deterministic or
-partial languages or logical relations for graduality in gradual
-typing.
-%
-We conjecture that some extension of the ordered predicate logic
-defined here will be useful for describing these relations as well.
+for parametricity and LSLR and Iris for step-indexed logical relations
+proofs \cite{abadi-plotkin,lslr,iris}.
+%
+We conjecture that \vett{} might similarly serve as the foundation for
+a logic of \emph{ordered} structures, which abound in applications:
+rewriting and approximation relations can both be modeled as orderings
+and logical relations involving these structures are proven to respect
+orderings: operational logical relations must be downward-closed and
+approximation relations should satisfy transitivity.
+%
+Just as LSLR and Iris release the user from the syntactic burden of
+explicit step-indexing, \vett{} may be used to release the user from
+the syntactic burden of proving downward-closure or transitivity
+side-conditions.
 
@@
 
-The syntactic forms of Simple Category Theory (SCT) are given in
+\begin{figure}
+  \begin{center}
+    \begin{tabular}{ |c|c| } 
+      \hline
+      Higher-Order Logic & Virtual Equipment Type Theory \\
+      \hline
+      Set $X$ & Category $\mathbb{C}$\\
+      $X \times Y$ & $\mathbb C \times \mathbb D$ \\ 
+      1 & $1$ \\
+      $\mathcal P X$ & $\mathcal P^+ X, \mathcal P^- X$ \\
+      $\{ (x,y) \in X \times Y | R(x,y) \}$ & $\sum_{\alpha:C;\beta:D} R$\\
+      Function $f(x:X) : Y$ & Functor/Object $\alpha:\cat C \vdash A : \cat D$\\
+      Relation $R(x,y)$ & Profunctor/Set $\alpha:\cat C; \beta:\cat D \vdash R$\\
+      $R \wedge Q$ & $R \times Q$\\
+      $\top$ & $1$\\
+      $\forall x. P \Rightarrow Q$ & $\homrallXYtoZ {\alpha:\cat C} P Q, \homlallXYtoZ {\alpha:\cat C} P Q$ \\
+      $\exists x. P \wedge Q$ & $\tensorexistsXwithYandZ {\alpha:\cat C} P Q$\\
+      $x =_X y$ & $\punitinXfromYtoZ {\cat C} \alpha \beta$\\
+      Proof $\forall \overrightarrow \alpha. R_1(\alpha_1,\alpha_2) \wedge \cdots \Rightarrow Q$ & Nat. Trans./Element $\alpha_1,x_1: R_1(\alpha_1,\alpha_2),\ldots \vdash t : Q$\\
+      \hline
+    \end{tabular}
+  \end{center}
+  \caption{Analogy between Higher-Order Logic and \vett{} Judgments and Connectives}
+\end{figure}
+
+The syntactic forms of \vett{} are given in
 \cref{fig:syntax}.
 %
-First, we have the types and terms of the logic over our category
-theory. Types include the usual dependent sum, product and identity
-types of dependent type theory, as well as several forms that allow us
-to abstract over the category theoretic constructions: a type $\Cat$
-of categories, a type $\varr{\cat C}{\cat D}$ for functors from $\cat
-C$ to $\cat D$, a type $\harr{\cat C}{\cat D}$ for profunctors
-contravariant in $\cat C$ and covariant in $\cat D$, and finally a
-type $\pendallXdotY {\alpha:{\cat C}} R$ that, when combined with
-other constructions on profunctors, allows for abstraction over
-arbitrary natural transformations.
-%
-Terms $L,M,N$ describe the elements of our logic's types, and include
-the usual constructions as well as forms for defining categories,
-functors, profunctors and natural transformations.
-%
-Next, we have a syntax for categories, which are written using a
-blackboard font $\cat C, \cat D, \cat E$. Categories include terms $M$
-of the category type (allowing for abstraction), in addition to
-primitives for product and unit categories.
-%
-Next, we have a syntax for objects of a category, which, because they
-are typed with respect to an input variable, provide a syntax for
-functors.
-%
-The objects we can define are object variables, application of a
-functor $M a$, and pairs of objects (for the product category) and the
-unique object of the unit category.
-%
-Next, $P,Q,R$ stand for \emph{sets}, but similar to objects, they are
-always typed with respect to a contravariant and covariant variable,
-and so denote profunctors.
-%
-These include first the multiplicatives: the hom set
-$\punitinXfromYtoZ {\cat C} a b$ representing the morphisms from $a$
-to $b$ in $\cat C$, a tensor product $\tensorexistsXwithYandZ
-{\beta:\cat D} P Q$, and two kinds of ``ordered function types''
-$\homrallXYtoZ \beta P Q$ and $\homlallXYtoZ \alpha R S$. 
-%
-We also include application of a profunctor term $M$ to two objects,
-and cartesian product and unit sets.
-%
-Next, we have transformations $s,t,u$, which are typed with respect
-to (0 or more) input and (one) output set.
-
-Finally, we have three forms of contexts, type contexts $\Gamma$, set
-contexts $\Xi$ and transformation contexts $\Phi$. A type context
-$\Gamma$ is the familiar notion of context from dependent type
-theory. We use uppercase $X,Y,Z$ for term variables to distinguish
-them from set variables. Every set expression $P$ must depend on an
-object of one category covariantly and an object of one category
-contravariantly, however these are allowed to be the same
-object. These two scenarios are captured by the notion of set context
-$\Xi$. The set context $\alpha:\cat C; \beta: \cat D$ gives the common
-case where the contravariant variable $\alpha$ is different from the
-covariant variable $\beta$, and the set context $\alpha: \cat C$ alone
-gives the case where the profunctor depends on the same variable both
-covariantly and contravariantly. Note that we write object variables
-as greek letters $\alpha, \beta$. Transformation contexts $\Phi$
-describe the inputs to a natural transformation. The input variables
-of a natural transformation $x: R$ stand for elements of sets, but as
-just described, sets always have covariant and contravariant
-dependencies. For this reason, transformation contexts include
-category variables $\alpha: \cat C$ interleaved with the set variables
-$x : R$, a typical context $\Phi$ would look like $\alpha_1:\cat C_1,
-x_1:R_1, \alpha_2: \cat C_2, x_2 : R_2, \alpha_3: \cat C_3$, where
-$R_1$ is parameterized by $\alpha_1$ contravariantly and $\alpha_2$
-covariantly and $R_2$ is similarly parameterized by
-$\alpha_2;\alpha_3$. Note that we consider transformation contexts up
-to associativity in that $(\Phi, x:R, \Psi), y : S, \Sigma$ should be
-considered equal to $\Phi, x:R, (\Psi, y: S, \Sigma)$.  In this
-calculus there is no single ``empty'' transformation context, but
-rather the empty context is one that consists of a single category
-variable $\alpha:\cat C$ and no set variables. Each form of context
-has a corresponding form of substitution, denoted using the lower-case
-form of the same greek letters.
+First, we have categories, which are analogous to sorts in a
+first-order theory. We have $M$ a base sort, product and unit sorts,
+as well as the graph of a profunctor and the negative and positive
+presheaf categories.
+%
+Next, objects $a,b,c$ are the syntax for the \emph{functors} between
+categories. We call them objects rather than functors, because in
+type-theoretic style, a functor is viewed as a ``generalized object''
+parameterized by an input variable $\alpha : \cat C$.
+%
+Next, sets $P,Q,R$ are the syntax for \emph{sets}. These sets denote
+\emph{profunctors}, i.e., a categorification of relations. Similar to
+functors, rather than writing profunctors as functions ${\cat
+  C}^o\times {\cat D} \to \Set$, we write them as sets with a
+contravariant variable $\alpha:\cat C$ and a covariant variable
+$\beta:\cat D$. The sets we can define are the Hom-set, the tensor and
+internal hom, as well as products of sets, profunctors applied to two
+objects and elements of positive and negative presheaves. Finally we
+have elements of sets, which correspond to natural transformations of
+multiple inputs, where again we view natural transformations valued in
+a profunctor as generalized elements of profunctors.
+
+After these forms we have types and terms, which are for the most part
+the same as in dependent type theory, except that we have universe
+objects for classifying categories, functors, profunctors and natural
+transformations.
+%
+Finally we have several forms of context which are used in the theory:
+contexts $\Gamma$ of term variables with their types, contexts $\Xi$
+of object variables with their categories, and contexts $\Phi$ of
+element variables with their sets, alternating with those sets' object
+variables.
+
+%% %
+%% First, we have the types and terms of the logic over our category
+%% theory. Types include the usual dependent sum, product and identity
+%% types of dependent type theory, as well as several forms that allow us
+%% to abstract over the category theoretic constructions: a type $\Cat$
+%% of categories, a type $\varr{\cat C}{\cat D}$ for functors from $\cat
+%% C$ to $\cat D$, a type $\harr{\cat C}{\cat D}$ for profunctors
+%% contravariant in $\cat C$ and covariant in $\cat D$, and finally a
+%% type $\pendallXdotY {\alpha:{\cat C}} R$ that, when combined with
+%% other constructions on profunctors, allows for abstraction over
+%% arbitrary natural transformations.
+%% %
+%% Terms $L,M,N$ describe the elements of our logic's types, and include
+%% the usual constructions as well as forms for defining categories,
+%% functors, profunctors and natural transformations.
+%% %
+%% Next, we have a syntax for categories, which are written using a
+%% blackboard font $\cat C, \cat D, \cat E$. Categories include terms $M$
+%% of the category type (allowing for abstraction), in addition to
+%% primitives for product and unit categories.
+%% %
+%% Next, we have a syntax for objects of a category, which, because they
+%% are typed with respect to an input variable, provide a syntax for
+%% functors.
+%% %
+%% The objects we can define are object variables, application of a
+%% functor $M a$, and pairs of objects (for the product category) and the
+%% unique object of the unit category.
+%% %
+%% Next, $P,Q,R$ stand for \emph{sets}, but similar to objects, they are
+%% always typed with respect to a contravariant and covariant variable,
+%% and so denote profunctors.
+%% %
+%% These include first the multiplicatives: the hom set
+%% $\punitinXfromYtoZ {\cat C} a b$ representing the morphisms from $a$
+%% to $b$ in $\cat C$, a tensor product $\tensorexistsXwithYandZ
+%% {\beta:\cat D} P Q$, and two kinds of ``ordered function types''
+%% $\homrallXYtoZ \beta P Q$ and $\homlallXYtoZ \alpha R S$. 
+%% %
+%% We also include application of a profunctor term $M$ to two objects,
+%% and cartesian product and unit sets.
+%% %
+%% Next, we have transformations $s,t,u$, which are typed with respect
+%% to (0 or more) input and (one) output set.
+
+%% Finally, we have three forms of contexts, type contexts $\Gamma$, set
+%% contexts $\Xi$ and transformation contexts $\Phi$. A type context
+%% $\Gamma$ is the familiar notion of context from dependent type
+%% theory. We use uppercase $X,Y,Z$ for term variables to distinguish
+%% them from set variables. Every set expression $P$ must depend on an
+%% object of one category covariantly and an object of one category
+%% contravariantly, however these are allowed to be the same
+%% object. These two scenarios are captured by the notion of set context
+%% $\Xi$. The set context $\alpha:\cat C; \beta: \cat D$ gives the common
+%% case where the contravariant variable $\alpha$ is different from the
+%% covariant variable $\beta$, and the set context $\alpha: \cat C$ alone
+%% gives the case where the profunctor depends on the same variable both
+%% covariantly and contravariantly. Note that we write object variables
+%% as greek letters $\alpha, \beta$. Transformation contexts $\Phi$
+%% describe the inputs to a natural transformation. The input variables
+%% of a natural transformation $x: R$ stand for elements of sets, but as
+%% just described, sets always have covariant and contravariant
+%% dependencies. For this reason, transformation contexts include
+%% category variables $\alpha: \cat C$ interleaved with the set variables
+%% $x : R$, a typical context $\Phi$ would look like $\alpha_1:\cat C_1,
+%% x_1:R_1, \alpha_2: \cat C_2, x_2 : R_2, \alpha_3: \cat C_3$, where
+%% $R_1$ is parameterized by $\alpha_1$ contravariantly and $\alpha_2$
+%% covariantly and $R_2$ is similarly parameterized by
+%% $\alpha_2;\alpha_3$. Note that we consider transformation contexts up
+%% to associativity in that $(\Phi, x:R, \Psi), y : S, \Sigma$ should be
+%% considered equal to $\Phi, x:R, (\Psi, y: S, \Sigma)$.  In this
+%% calculus there is no single ``empty'' transformation context, but
+%% rather the empty context is one that consists of a single category
+%% variable $\alpha:\cat C$ and no set variables. Each form of context
+%% has a corresponding form of substitution, denoted using the lower-case
+%% form of the same greek letters.
 
@@
     \begin{array}{rccl}
+      \textrm{Categories} &\cat C, \cat D, \cat E & ::= & M \pipe \cat C \times \cat D \pipe \cat 1
+      \pipe \graphProf \alpha \beta P \pipe \negPresheaf {\cat C} \pipe \posPresheaf {\cat C}
+      \\
+      \textrm{Objects} & a, b, c & ::= & \alpha \pipe M a \pipe (a,b) \pipe () \ldots\\
+      \textrm{Sets} & P,Q,R & ::= & \punitinXfromYtoZ {\cat C} a b \pipe \tensorexistsXwithYandZ {\beta} P Q \pipe \homrallXYtoZ \beta P Q \pipe \homlallXYtoZ \alpha R S \\
+      &&& 1 \pipe P \times Q\\
+      &&& \pipe \harrapp M a b \pipe \negPresheafAppPtoX a b \pipe \posPresheafAppPtoX a b\\
+      \textrm{Elements} & s,t,u & ::= & \ldots \\
+      \\
       \textrm{Type} & A,B,C & ::= &
@@
       &&&\lceil \cat C \rceil \pipe \vlambda {\alpha:\cat C} a \pipe \hlambda {\alpha:\cat C}{\beta: \cat D} R \ldots \\
-      \textrm{Categories} &\cat C, \cat D, \cat E & ::= & M \pipe \cat C \times \cat D \pipe \cat 1\\
-      \textrm{Objects} & a, b, c & ::= & \alpha \pipe M a \pipe (a,b) \pipe () \ldots\\
-      \textrm{Sets} & P,Q,R & ::= & \punitinXfromYtoZ {\cat C} a b \pipe \tensorexistsXwithYandZ {\beta} P Q \pipe \homrallXYtoZ \beta P Q \pipe \homlallXYtoZ \alpha R S \\
-      &&& M(a,b) \pipe 1 \pipe P \times Q\\
-      \textrm{Transformations} & s,t,u & ::= & \ldots \\
-      \\
       \textrm{Type Context} & \Gamma, \Delta & ::= & \cdot \pipe \Gamma, X : A\\
       \textrm{Category Context} & \Xi, Z & ::= & \alpha:\cat C \pipe \alpha:\cat C;\beta: \cat D\\
-      \textrm{Set Context} & \Phi,\Psi & ::= & \alpha : \cat C \pipe \Phi , x : P , \Psi\\
-      \textrm{Term Substitution} & \gamma, \delta & ::= & \cdot \pipe \gamma, M/X\\
-      \textrm{Object Substitution} & \xi, \zeta & ::= & a/\alpha \pipe a/\alpha;b/\beta\\
-      \textrm{Transformation Substitution} & \phi, \psi & ::= & a/\alpha \pipe \phi,s/x, \psi      
+      \textrm{Set Context} & \Phi,\Psi & ::= & \alpha : \cat C \pipe \Phi , x : P,\beta:\cat D\\
     \end{array}
   \end{mathpar}
-  \caption{Simple Category Theory Syntactic Forms}
+  \caption{\vett{} Syntactic Forms}
   \label{fig:syntax}
@@
   \begin{mathpar}
-    \inferrule*[right=TyCtxForm]
-        {}
-        {\Gamma \isadtctx}
-
     \inferrule*[right=Cat-form]
@@
 
-    \inferrule*[right=SetCtxForm]
-    {\Gamma \isadtctx}
-    {\Gamma \vdash \Xi \boundary}
-
     \inferrule*[right=SetForm]
@@
     {\Gamma \pipe \Xi \vdash S \isaSet}
-
-    \inferrule*[right=TransCtxForm]
-    {\Gamma \isadtctx}
-     {\Gamma \vdash \Phi \isavectx}
 
@@
     {\Gamma \pipe \Phi \vdash s : R}
-
-    \inferrule
-    {\Delta \isadtctx \and \Gamma \isadtctx}
-    {\gamma : \Delta \to \Gamma}
-
-    \inferrule*[right=Boundary Map formation]
-    {\Gamma \vdash Z \boundary \and \Gamma \vdash \Xi \boundary}
-    {\Gamma \vdash \xi : Z \to \Xi}
-
-    \inferrule*[right=Cat Subst formation]
-    {\Gamma \vdash \Psi \isavectx \and \Gamma \vdash \Phi \isavectx}
-    {\Gamma \vdash \phi : \Psi \to \Phi}
   \end{mathpar}
-  \caption{Formation Rules}
+  \caption{Selected Formation Rules}
   \label{fig:formation}
@@
 \appendix
+
+\section{Syntactic Details}
+
+\begin{figure}
+  \begin{mathpar}
+    \begin{array}{rccl}
+      \textrm{Term Substitution} & \gamma, \delta & ::= & \cdot \pipe \gamma, M/X\\
+      \textrm{Object Substitution} & \xi, \zeta & ::= & a/\alpha \pipe a/\alpha;b/\beta\\
+      \textrm{Transformation Substitution} & \phi, \psi & ::= & a/\alpha \pipe \phi,s/x, \psi      
+    \end{array}
+
+      \inferrule*[right=TyCtxForm]
+        {}
+        {\Gamma \isadtctx}
+
+    \inferrule*[right=SetCtxForm]
+    {\Gamma \isadtctx}
+    {\Gamma \vdash \Xi \boundary}
+
+    \inferrule*[right=TransCtxForm]
+    {\Gamma \isadtctx}
+     {\Gamma \vdash \Phi \isavectx}
+    \inferrule
+    {\Delta \isadtctx \and \Gamma \isadtctx}
+    {\gamma : \Delta \to \Gamma}
+
+    \inferrule*[right=Boundary Map formation]
+    {\Gamma \vdash Z \boundary \and \Gamma \vdash \Xi \boundary}
+    {\Gamma \vdash \xi : Z \to \Xi}
+
+    \inferrule*[right=Cat Subst formation]
+    {\Gamma \vdash \Psi \isavectx \and \Gamma \vdash \Phi \isavectx}
+    {\Gamma \vdash \phi : \Psi \to \Phi}
+
+  \end{mathpar}
+  \caption{Substitution}
+
+\end{figure}
 

### 2022-10-03 paper/paper.tex
progress on syntax section

@@
 \newcommand{\isaCat}{\,\,\textrm{Cat}}
+\newcommand{\isSmall}{\,\,\textrm{Small}}
 \newcommand{\Set}{\textrm{Set}}
@@
 \newcommand{\harrapp}[3]{#1(#2;#3)}
+
+\newcommand{\defaultObCtx}{\Gamma\pipe\alpha:\cat C}
 
@@
 %
-The downside is that these algebraic structures are quite complicated,
-with a 2-dimensional diagrammatic language that can be quite beautiful
-and elegant, but is also seems very far removed from the traditional
-formulation of category theory that while complex is at least
-described in terms of a concrete notion of sets or types.
+A downside is that these algebraic structures are quite complicated,
+and practitioners typically employ either an algebraic combinator
+syntax or a 2-dimensional diagrammatic language that can be quite
+beautiful and elegant, but is also seems very far removed from the
+traditional formulation of category theory in terms of sets and
+functions.
 
@@
 functions ``for free''.
+%
+The syntax provides an alternative to algebraic and
+string-diagrammatic syntaxes for working with virtual equipments,
+similar to how the lambda calculus provides an alternative to
+categorical combinators and string diagram calculi for cartesian
+closed categories.
 
@@
 
-\section{Syntax of SCT}
+\section{Syntax of VETT}
 
@@
 %% have rules for constructing an ``empty'' context, which is just a
+
 In typical natural-deduction style, all syntactic forms typed in
@@
 \cat D$ in other objects. We can also substitute objects into
-\emph{sets}. For instance, if we have a hom set in a category
-$\punitinXfromYtoZ {\cat C} \alpha {\alpha'}$, and we can define
-objects $a$ and $a'$ in the category $\cat C$, then we can substitute
-those in to form the hom set between $a$ and $a'$: $\punitinXfromYtoZ
-{\cat C} a {a'}$. This semantically is a ``restriction'' of a
-profunctor along two functors, and modeling this operation as a
-substitution considerably simplifies reasoning using profunctors.
+\emph{sets}, that is, if we have a set $P$ parameterized by a
+contravariant variable $\alpha : \cat C$ and a covariant variable
+$\beta: \cat D$, then we can substitute objects $a : \cat C$ and $b :
+\cat D$ for these variables $P[a/\alpha;b\beta]$. This generalizes the
+ordinary precomposition of a relation by a function.  Semantically
+this is the ``restriction'' of a profunctor along two functors, which
+is just composition of functors if a profunctor is viewed as a functor
+to $\Set$. Modeling this operation as a substitution considerably
+simplifies reasoning using profunctors.
 %
@@
 category variables with objects.
-
-\cref{fig:contexts} details the typing rules for contexts and the
-typing for variables and the admissible actions of substitution (the
-definition of substitution is straightforward and found in the
-appendix). the rules for term and object variables are as usual.
-%
-An object substitution into a single object context $\alpha:\cat C$
-consists of an object $a : \cat C$ parameterized by an object of
-another category.
-%
-An object substitution into a pair of co- and contra-variant variables
-$\alpha:\cat C;\beta:\cat D$ is a pair of an object substituted for
-$\alpha$ and one for $\beta$, each typed with respect to the
-appropriate variable of an object context $\Xi$.
-%
-Note that this allows for $\Xi$ to consist of either two variables or
-a single variable.
-%
-When
-constructing a transformation context $\Phi, x : R, \Psi$, we need to
-ensure that the contravariant dependence of $R$ matches the covariant
-dependence of $\Phi$ and vice-versa for $\Psi$. We define this with
-auxiliary metafunctions $d^+,d^-$ which extract the covariant and
-contravariant object variable of a transformation context.
-%
-The rules for transformation variables and substitutions define the
-use of variables to be \emph{linear} (each variables is used exactly
-once) and \emph{ordered}, they occur in the term in the same order
-that they appear in the context.
-%
-The rule for a transformation variables says a variable $x$ can only
-be used when it is the sole variable in the context.
-%
-A transformation subsitution into an ``empty'' transformation context
-$\alpha: \cat C$ is the same as an object substitution into
-$\alpha:\cat C$.
-%
-The rule for constructing a transformation substitution into a
-non-empty context $\Phi_1,x:R,\Phi_2$ says we can construct a
-substitution by constructing substitutions $\phi_1,\phi_2$ for
-$\Phi_1$ and $\Phi_2$ as well as a transformation $s$ into an
-instantiation of $R$ where the input contexts come from splitting the
-input context $\Psi$ into three components, and ensuring that the
-instantiation of $R$ agrees with how $\phi_1,\phi_2$ instantiate the
-variables. This uses two new metafunctions. First, $\Phi \jnctx \Psi$
-is an operation for concatenating two contexts when $d^+(\Phi) =
-d^-(\Psi)$. Next, we extend the boundary maps $d^\pm$ to substitutions
-to extract their object substitutions.
+%
+The full syntactic details of substitution are included in the appendix.
+
+\subsection{Category Connectives}
+
+We introduce the connectives for defining categories in
+Figure~\ref{fig:category-connectives}. We show the formation,
+introduction and elimination rules, but additionally there are rules
+for $\beta\eta$ equality and substitution, which are included in the
+appendix.
+%
+First we introduce the additives: the unit category $1$ and product
+category $\cat C \times \cat D$ with their usual introduction and
+elimination rules.
+%
+Next, we introduce the \emph{graph of a profunctor} $\graphProf \alpha
+\beta P$. Just as a relation $R : A \times B \to \Set$ can be viewed
+as a subset $\{ (a,b) \in A\times B | R(a,b)\}$, any profunctor $P :
+C_-^o\times C_+ \to \Set$ can be viewed as a category with a functor
+to $C_- \times C_+$ (no op), specifically a two-sided discrete
+fibration. In set-based category theory, the objects are triples
+$(a_-, a_+, s : P(a_-,a_+))$ and morphisms from $(a_-,a_+, s)$ to
+$(a_-',a_+', s')$ are pairs of morphisms $f_- : a_- \to a_-'$ and $f_+
+: a_+ \to a_+'$ such that $P(\id,f_+)(s) = P(f_-,\id)(s')$.
+%
+With various choices of $P$, this connective can be used to define the
+arrow category, slice category and comma category constructions.
+%
+In our syntax we define it as the universal category $\cat C$ equipped
+with functors to $\cat C_-$ and $\cat C_+$ and a morphism to $P$.
+
+Lastly, we define the \emph{negative} and \emph{positive} presheaf
+categories $\negPresheaf \cat C$ and $\posPresheaf \cat D$. These are
+given a syntax suggestive of the fact that they generalize the notion
+of a powerset, and so can be thought of as ``power categories''.
+%
+Note that we include a restriction that the input category is
+\emph{small}, which is an inductively defined by saying all base
+categories are small, the unit is small, product of small categories
+is small and the graph of a profunctor over small categories is
+small. Notably, the presheaf categories themselves are not small.
+%
+The negative presheaf category is defined by its universal property
+that a functor into it $\cat D \to \negPresheaf \cat C$ is equivalent
+to a profunctor $\cat C^o \times \cat D \to \Set$.
+%
+The introduction rule constructs an object of the negative presheaf
+category from such a profunctor and the elimination rule inverts
+it. We use the notation $\negPresheafApp a p$ for the elements of the
+induced profunctor.
+%
+The positive presheaf category is then the dual. In ordinary
+set-theoretic category theory the negative presheaf category is the
+usual presheaf category $\Set^{\cat C^o}$, and the positive presheaf
+category is the opposite of the dual presheaf category $(\Set^{\cat
+  D})^o$.
 
@@
   \begin{mathpar}
-    \inferrule{}{\cdot \isadtctx}
+    \inferrule{}{\Gamma \vdash 1 \isaCat}\and
+    \inferrule*{}{\Gamma\pipe\alpha:C \vdash () : 1}\and
 
     \inferrule
-    {\Gamma \isadtctx \and \Gamma \vdash A \isaTy}
-    {\Gamma , X : A \isadtctx}\\
-
+        {\Gamma \vdash \cat C_1 \isaCat \and \cat C_2 \isaCat}
+        {\Gamma \vdash {\cat C_1} \times {\cat C_2} \isaCat}
     \inferrule
-    {X : A \in \Gamma}
-    {\Gamma \vdash X : A}
+    {\Gamma \pipe \alpha:\cat C \vdash a_1 : \cat C_1 \and \Gamma \pipe \alpha:\cat C \vdash a_2 : \cat C_2}
+    {\Gamma \pipe\alpha:\cat C \vdash (a_1,a_2) : {\cat C_1} \times {\cat C_2}}
 
     \inferrule
-    {}
-    {\alpha: \cat C \vdash \alpha: \cat C}
+    {\Gamma \pipe\alpha:\cat C \vdash a : {\cat C_1} \times {\cat C_2}}
+    {\Gamma\pipe\alpha:\cat C\vdash \pi_i a : \cat C_i}
 
     \inferrule
-    {\Gamma \vdash \cat C : \isaCat}
-    {\Gamma \vdash \alpha: \cat C \isavectx}
+    {\Gamma \pipe \alpha:\cat C;\beta:\cat D \vdash P \isaSet}
+    {\Gamma \vdash \graphProf{\alpha}{\beta} P \isaCat}
 
     \inferrule
-    {\Gamma \vdash \Phi \isavectx\and
-    \Gamma \pipe d^+\Phi; \beta:\cat D \vdash R \isaSet}
-    {\Gamma \vdash \Phi, x:R, \beta:\cat D \isavectx}
+    {\defaultObCtx \vdash a : \graphProf{\alpha:\cat C_-}{\beta} P}
+    {\defaultObCtx \vdash \pi_- a : \cat C_-}
 
     \inferrule
-    {}
-    {\Gamma \pipe \alpha:\cat C, x : R, \beta: \cat D \vdash x : R}
-
-    % horizontal identity
+    {\defaultObCtx \vdash a : \graphProf{\alpha}{\beta:C_+} P}
+    {\defaultObCtx \vdash \pi_+ a : \cat C_+}
+
     \inferrule
-    {\Gamma\pipe \beta:\cat C \vdash a : \cat C}
-    {\Gamma \pipe \beta:\cat C \vdash a/\alpha : (\alpha : \cat C)}
+    {\defaultObCtx \vdash a : \graphProf{\alpha}{\beta} P}
+    {\defaultObCtx \vdash \pi_e a : P[\pi_-a/\alpha;\pi_+a/\beta]}
 
     \inferrule
-    {\Psi = \Psi_1 \jnctx \Psi_2\and
-      \Gamma\pipe\Psi_1 \vdash \phi : \Phi\and
-      \Gamma \pipe \Psi_2 \vdash t : R[d^+\phi;b/\beta] \and
-      \Gamma \vdash \phi_2 : \Psi_2 \to \Phi_2
+    {\defaultObCtx \vdash a_- : C_-\and
+      \defaultObCtx \vdash a_+ : C_+\and
+      \defaultObCtx \vdash s : P[a_-/\alpha;a_+/\beta]}
+    {\defaultObCtx \vdash (a_-,a_+,s) : \graphProf{\alpha:\cat C_-}{\beta:\cat C_+} P}
+
+    \inferrule
+    {\Gamma \vdash \cat C \isaCat \and \cat C \isSmall}
+    {\Gamma \vdash \negPresheaf {\cat C} \isaCat}
+
+    \inferrule
+    {\Gamma \pipe d^-\Xi \vdash a : \cat C \and
+     \Gamma \pipe d^+\Xi \vdash p : \negPresheaf {\cat C}
     }
-    {\Gamma\pipe\Psi \vdash \phi,t/x,b/\beta :: (\Phi,x:R,\beta:\cat D)}
+    {\Gamma \pipe \Xi \vdash \negPresheafAppPtoX p a \isaSet}
+
+    \inferrule
+    {\Gamma \pipe \alpha:\cat C;\beta:\cat D \vdash R : \isaSet}
+    {\Gamma \pipe \beta:\cat D \vdash \lambda \alpha:\cat C. R : \negPresheaf {\cat C}}
+
+    \inferrule
+    {\Gamma \vdash \cat D \isaCat \and \cat D \isSmall}
+    {\Gamma \vdash \posPresheaf {\cat D} \isaCat}
+
+    \inferrule
+    {\Gamma \pipe d^-\Xi \vdash p : \posPresheaf {\cat D}\and
+     \Gamma \pipe d^+\Xi \vdash a : \cat D
+    }
+    {\Gamma \pipe \Xi \vdash \posPresheafAppPtoX p a \isaSet}
+
+    \inferrule
+    {\Gamma \pipe \alpha:\cat C;\beta:\cat D \vdash R : \isaSet}
+    {\Gamma \pipe \alpha:\cat C \vdash \lambda \beta:\cat D. R : \posPresheaf {\cat D}}
   \end{mathpar}
-  \begin{align*}
-    \underline{\alpha:C} &= \alpha:C\\
-    \underline{\Phi,x:R,\Psi} &= d^-\Phi; d^+{\Psi}
-  \end{align*}
-  \begin{align*}
-    d^{\pm}(\alpha:\cat C) &= \alpha : \cat C\\
-    d^{-}(\Phi,x:R,\Psi) &= d^-\Phi\\
-    d^{+}(\Phi,x:R,\Psi) &= d^-\Psi
-  \end{align*}
-  \begin{align*}
-    d^{\pm}(a/\alpha) &= a/\alpha\\
-    d^{-}(\phi,t/x,\psi) &= d^-\phi\\
-    d^{+}(\phi,t/x,\psi) &= d^+\psi\\
-  \end{align*}
-  \caption{Basic Rules of Judgment Theory}
-  \label{fig:contexts}
+  \caption{Category Conectives}
 \end{figure}
 
-Now that we have established a judgmental structure, in typical type
-theoretic fashion, our type constructs follow naturally as
-internalizing the judgments. Though the positive multiplicatives are
-more familiar, we start by describing the negative multiplicatives,
-since they are better behaved, and are useful to have already defined
-when discussing the positives.
-
-\subsection{Types}
+\subsection{Set Connectives}
+
+Next, we cover the connectives for the sets, i.e., constructions on
+profunctors in Figure~\ref{fig:set-connectives}.
+
+\begin{figure}
+  TODO
+  \caption{Set Connectives}
+\end{figure}
+
+\subsection{Type Connectives}
+
+Finally, we briefly describe the connectives for the
+``meta-logic''. These include the standard rules for Martin-L\"of type
+theory with Pi/Sigma and extensional identity types, whose rules we
+elide as they are standard \cite{emltt}. We use extensional identity
+types so that the description of models is simpler, but there is no
+obstacle to including models with intensional identity types.
+%
+The types we include are what we call \emph{universes} for the object
+categorical logic: types of small categories and locally small
+categories, functors, profunctors and natural transformations.
+
+\begin{figure}
+
+  \caption{Type Connectives}
+\end{figure}
 
@@
 \end{figure}
+
+\cref{fig:contexts} details the typing rules for contexts and the
+typing for variables and the admissible actions of substitution (the
+definition of substitution is straightforward and found in the
+appendix). the rules for term and object variables are as usual.
+%
+An object substitution into a single object context $\alpha:\cat C$
+consists of an object $a : \cat C$ parameterized by an object of
+another category.
+%
+An object substitution into a pair of co- and contra-variant variables
+$\alpha:\cat C;\beta:\cat D$ is a pair of an object substituted for
+$\alpha$ and one for $\beta$, each typed with respect to the
+appropriate variable of an object context $\Xi$.
+%
+Note that this allows for $\Xi$ to consist of either two variables or
+a single variable.
+%
+When
+constructing a transformation context $\Phi, x : R, \Psi$, we need to
+ensure that the contravariant dependence of $R$ matches the covariant
+dependence of $\Phi$ and vice-versa for $\Psi$. We define this with
+auxiliary metafunctions $d^+,d^-$ which extract the covariant and
+contravariant object variable of a transformation context.
+%
+The rules for transformation variables and substitutions define the
+use of variables to be \emph{linear} (each variables is used exactly
+once) and \emph{ordered}, they occur in the term in the same order
+that they appear in the context.
+%
+The rule for a transformation variables says a variable $x$ can only
+be used when it is the sole variable in the context.
+%
+A transformation subsitution into an ``empty'' transformation context
+$\alpha: \cat C$ is the same as an object substitution into
+$\alpha:\cat C$.
+%
+The rule for constructing a transformation substitution into a
+non-empty context $\Phi_1,x:R,\Phi_2$ says we can construct a
+substitution by constructing substitutions $\phi_1,\phi_2$ for
+$\Phi_1$ and $\Phi_2$ as well as a transformation $s$ into an
+instantiation of $R$ where the input contexts come from splitting the
+input context $\Psi$ into three components, and ensuring that the
+instantiation of $R$ agrees with how $\phi_1,\phi_2$ instantiate the
+variables. This uses two new metafunctions. First, $\Phi \jnctx \Psi$
+is an operation for concatenating two contexts when $d^+(\Phi) =
+d^-(\Psi)$. Next, we extend the boundary maps $d^\pm$ to substitutions
+to extract their object substitutions.
+
+\begin{figure}
+  \begin{mathpar}
+    \inferrule{}{\cdot \isadtctx}
+
+    \inferrule
+    {\Gamma \isadtctx \and \Gamma \vdash A \isaTy}
+    {\Gamma , X : A \isadtctx}\\
+
+    \inferrule
+    {X : A \in \Gamma}
+    {\Gamma \vdash X : A}
+
+    \inferrule
+    {}
+    {\alpha: \cat C \vdash \alpha: \cat C}
+
+    \inferrule
+    {\Gamma \vdash \cat C : \isaCat}
+    {\Gamma \vdash \alpha: \cat C \isavectx}
+
+    \inferrule
+    {\Gamma \vdash \Phi \isavectx\and
+    \Gamma \pipe d^+\Phi; \beta:\cat D \vdash R \isaSet}
+    {\Gamma \vdash \Phi, x:R, \beta:\cat D \isavectx}
+
+    \inferrule
+    {}
+    {\Gamma \pipe \alpha:\cat C, x : R, \beta: \cat D \vdash x : R}
+
+    % horizontal identity
+    \inferrule
+    {\Gamma\pipe \beta:\cat C \vdash a : \cat C}
+    {\Gamma \pipe \beta:\cat C \vdash a/\alpha : (\alpha : \cat C)}
+
+    \inferrule
+    {\Psi = \Psi_1 \jnctx \Psi_2\and
+      \Gamma\pipe\Psi_1 \vdash \phi : \Phi\and
+      \Gamma \pipe \Psi_2 \vdash t : R[d^+\phi;b/\beta] \and
+      \Gamma \vdash \phi_2 : \Psi_2 \to \Phi_2
+      }
+    {\Gamma\pipe\Psi \vdash \phi,t/x,b/\beta :: (\Phi,x:R,\beta:\cat D)}
+  \end{mathpar}
+  \begin{align*}
+    \underline{\alpha:C} &= \alpha:C\\
+    \underline{\Phi,x:R,\Psi} &= d^-\Phi; d^+{\Psi}
+  \end{align*}
+  \begin{align*}
+    d^{\pm}(\alpha:\cat C) &= \alpha : \cat C\\
+    d^{-}(\Phi,x:R,\Psi) &= d^-\Phi\\
+    d^{+}(\Phi,x:R,\Psi) &= d^-\Psi
+  \end{align*}
+  \begin{align*}
+    d^{\pm}(a/\alpha) &= a/\alpha\\
+    d^{-}(\phi,t/x,\psi) &= d^-\phi\\
+    d^{+}(\phi,t/x,\psi) &= d^+\psi\\
+  \end{align*}
+  \caption{Basic Rules of Judgment Theory}
+  \label{fig:contexts}
+\end{figure}
+
+Now that we have established a judgmental structure, in typical type
+theoretic fashion, our type constructs follow naturally as
+internalizing the judgments. Though the positive multiplicatives are
+more familiar, we start by describing the negative multiplicatives,
+since they are better behaved, and are useful to have already defined
+when discussing the positives.
 

### 2022-10-03 paper/paper.tex
progress on syntax re-org

@@
 \newcommand{\lmod}{\text{LMod}}
-\newcommand{\varr}[2]{\text{Fun}{#1}{#2}}
-\newcommand{\harr}[2]{\text{Rel}{#1}{#2}}
+
+\newcommand{\smallCats}{\text{SmallCat}}
+\newcommand{\Cats}{\text{Cat}}
+\newcommand{\varr}[2]{\text{Fun}\,{#1}\,{#2}}
+\newcommand{\harr}[2]{\text{Prof}\,{#1}\,{#2}}
 \newcommand{\harrapp}[3]{#1(#2;#3)}
@@
 \newcommand{\punitelimtoYwithkontZ}[2]{\textrm{ind}_{\to}(#1,#2)}
+\newcommand{\punitelimkontZatABC}[4]{\textrm{ind}_{\to}(#1,#2, #3, #4)}
 
@@
 \newcommand{\tensorelimfromWXtoYwithkontZ}[4]{\textrm{ind}_{\odot}(#1,#2,#3,#4)}
-\newcommand{\tensorelimVatWtoXtoYwithkontZ}[5]{\textrm{ind}_{\to}(#4,#2,#1,#3,#5)}
+\newcommand{\tensorelimVatWtoXtoYwithkontZ}[5]{\textrm{ind}_{\odot}(#4,#2,#1,#3,#5)}
+\newcommand{\tensorelimWfromXtoYkontZ}[4]{\textrm{ind}_{\odot}(#4;#1,#2,#3)}
 
@@
 \begin{figure}
-  TODO
+  \begin{mathpar}
+    % unit,
+
+    \inferrule
+    {\Gamma \pipe d^-\Xi; \beta:D\vdash P \isaSet\\\\
+     \Gamma \pipe \beta:D;d^+\Xi \vdash Q \isaSet}
+    {\Gamma \pipe \Xi \vdash \tensorexistsXwithYandZ {\beta:D} P Q}
+
+    \inferrule
+    {\Gamma \pipe \alpha:\cat C; d^-\Xi\vdash R \\\\ \Gamma \pipe \alpha : \cat C; d^+\Xi\vdash P}
+    {\Gamma \pipe \Xi \vdash \homlallXYtoZ \alpha R P}
+
+  \inferrule
+  {\Gamma \pipe d^+\Xi; \alpha : \cat C \vdash  R \\\\ \Gamma \pipe  d^-\Xi; \alpha : \cat C \vdash P}
+  {\Gamma \pipe \Xi \vdash \homrallXYtoZ \alpha R P}\\
+
+  \inferrule
+    {\Gamma \pipe d^-\Xi \vdash a_1 : \cat C\\\\ \Gamma \pipe d^+ \Xi \vdash a_2 : \cat C}
+    {\Gamma \pipe \Xi \vdash \punitinXfromYtoZ {\cat C} {a_1} {a_2}}
+
+    \inferrule{}{\Gamma\pipe \Xi \vdash 1 \isaSet}
+
+    \inferrule{\Gamma\pipe\Xi \vdash R \isaSet \\\\ \Gamma\pipe\Xi \vdash S\isaSet}
+              {\Gamma\pipe\Xi \vdash R \times S\isaSet}\\
+
+    \inferrule
+    {\Gamma\pipe\beta: \cat D \vdash a : \cat C}
+    {\Gamma \pipe\cdot \vdash \punitrefl a : \punitinXfromYtoZ {\cat C} a a}
+
+%%     \inferrule*[right=Unit Elim]
+%%     {\Gamma\vdash Q : \harr {\cat C} {\cat C}\and
+%%      \Gamma\vdash t : \pendallXdotY {\alpha:\cat C} {\harrapp Q \alpha \alpha}}
+%%     {\Gamma \vdash \punitelimtoYwithkontZ Q N :
+%%       \pendallXdotY {\alpha_1} {\homrallXYtoZ {\alpha_2} {\punitinXfromYtoZ {\cat C} {\alpha_1}{\alpha_2}} {\harrapp Q {\alpha_1}{\alpha_2}}}}
+
+    \inferrule
+    {\Gamma\pipe \alpha:\cat C\vdash t : P[\alpha/\alpha^-;\alpha/\alpha^+]\\\\
+     \Gamma\pipe\Phi \vdash s : \punitinXfromYtoZ {\cat C} A B}
+    {\Gamma \vdash \punitelimkontZatABC {\alpha. t} A s B : P[A/\alpha^-;B/\alpha^+]}
+
+  %% \inferrule*[right=Tensor intro function]
+  %% {}
+  %% {\alpha : \cat C, q : Q, \beta : \cat D, r : R, \gamma : \cat E \vdash (q, \beta,r) : \exists \beta : \cat D. Q \odot R}
+
+  %% \inferrule*[right=Tensor introduction]
+  %% {\Gamma \vdash P : \harr C D \and
+  %%   \Gamma \vdash Q : \harr D E}
+  %% {\Gamma\pipe\cdot\vdash \tensorintroofXandY P Q : \pendallXdotY {\alpha} \homrallXYtoZ \beta {\harrapp P \alpha \beta} {\homrallXYtoZ \gamma {\harrapp Q \beta \gamma} {(\tensorexistsXwithYandZ \beta {\harrapp P \alpha \beta}{\harrapp Q \beta \gamma})}}}
+
+  \inferrule
+  {\Gamma \pipe d^+\Psi_s \vdash b : \cat D\\\\
+   \Gamma \pipe \Psi_s \vdash s : P[b/\beta] \\\\
+   \Gamma \pipe \Psi_t \vdash t : Q[b/\beta]}
+  {\Gamma\pipe\Psi_s \jnctx \Psi_t\vdash \tensorintroatXwithYandZ b s t : \tensorexistsXwithYandZ {\beta : \cat D} P Q}
+
+  \inferrule
+  {
+   \Gamma\pipe \alpha:\cat C, x:P, \beta:\cat D, y:Q, \gamma:\cat E \vdash t : R\\\\
+   \Gamma \pipe\Phi \vdash s : \tensorexistsXwithYandZ {\beta} {P[a/\alpha]}{Q[c/\gamma]}}
+  {\Gamma \vdash \tensorelimWfromXtoYkontZ a s b {\alpha.x.\beta.y.\gamma. t} : R[a/\alpha;c/\gamma]}
+
+  %% \tensorintroatXwithYandZ B M N [\phi,B'/\beta',\psi] = \tensorintroatXwithYandZ {B[B'\beta]} {M[\phi]} {N[\psi]}
+
+  %% \inferrule*[right=Tensor simpl]
+  %% {\Gamma\pipe\cdot \vdash R : \harr {C} {E}\and
+  %%  \Gamma\pipe\cdot \vdash N : \pendallXdotY {\alpha:C} {\homrallXYtoZ {\beta:D} P {\homrallXYtoZ {\gamma:E} Q {\harrapp R \alpha\gamma}}}}
+  %% {\Gamma\pipe\alpha:C,p:\tensorexistsXwithYandZ {\beta:D} P Q,\gamma:E \vdash\tensorelimVatWtoXtoYwithkontZ p \alpha \gamma R N : \harrapp R A G}
+
+%%   \inferrule*[right=Tensor-elim]
+%%   {\Gamma \pipe \vdash P : \harr {\cat C} {\cat D}\and\Gamma \vdash Q : \harr {\cat D}{\cat E}\and
+%%    \Gamma \vdash R : \harr {\cat C}{\cat E}\and
+%%    \Gamma \vdash N : \hombinary \alpha {\harrapp P \alpha \beta} \beta {\harrapp Q \beta \gamma} \gamma {\harrapp R \alpha \gamma}}
+%%   {\Gamma \vdash \tensorelimfromWXtoYwithkontZ P Q R N : \homunary \alpha {(\tensorexistsXwithYandZ \beta {\harrapp P \alpha \beta} {\harrapp Q \beta \gamma})} {\gamma} {\harrapp R \alpha \gamma}}
+  \inferrule
+  {\Gamma \pipe \Phi, x : R, \alpha : \cat C \vdash t : P}
+  {\Gamma \pipe \Phi \vdash
+   \homrlambdaXatYdotZ {x : R} {\alpha : \cat C} t : \homrallXYtoZ \alpha R P}
+
+      %% \inferrule*[right=Covar Hom elim simple]
+      %% {}
+      %% {\alpha: \cat C, f:\forall \gamma. P \triangleleft R, \beta : \cat D, x : R, \gamma : \cat E \vdash f\triangleleft (x, \gamma) : P}
+
+  \inferrule
+  {\Gamma\pipe\Phi_f \vdash s : \homrallXYtoZ {\alpha:\cat C} R P \\\\
+    d^+\Phi_a \vdash a : \cat C\\\\
+    \Phi_a \vdash t : R[a/\alpha]
+  }
+  {\Gamma\pipe \Phi_f \jnctx \Phi_a \vdash \homrappXtoYatZ s t a : P[a/\alpha]}\\
+
+  \inferrule
+  {\Gamma\pipe\alpha:\cat C, x : R, \Phi \vdash t : P}
+  {\Gamma\pipe\Phi \vdash \homllambdaXatYdotZ {x : R} {\alpha : \cat C} t : \homlallXYtoZ \alpha R P}
+
+  \inferrule
+  {\Gamma \pipe d^-\Phi_a \vdash a : \cat C\\\\
+   \Gamma \pipe \Phi_a \vdash s : R[a/\alpha] \\\\
+   \Gamma \pipe \Phi_f \vdash t : \homlallXYtoZ {\alpha:\cat C} R P}
+  {\Gamma \pipe \Phi_a \jnctx \Phi_f \vdash \homlappXtoYatZ t s a : P[a/\alpha]}\\
+  % tensor,
+  % hom,
+  % cartesian products,
+  %
+
+  \inferrule{}{\Gamma\pipe \Phi \vdash () :  1}
+
+    \inferrule{\forall i \in \{1,2\}.~ \Gamma \pipe\Phi \vdash M_i : R_i}{\Gamma \pipe\Phi \vdash (M_1,M_2) : R_1 \times R_2}
+
+    \inferrule{\Gamma \pipe\Phi\vdash M : R_1 \times R_2}{\Gamma \pipe\Phi \vdash \pi_i M : R_i}
+  \end{mathpar}
   \caption{Set Connectives}
@@
 \begin{figure}
-
-  \caption{Type Connectives}
-\end{figure}
-
-Finally, we describe the \emph{end}, $\pendallXdotY {\alpha} P$, which
-when combined with the left or right hom types described in the next
-section allows for us to internalize arbitrary transformation sequents
-$\Phi \vdash P$ as a type.
-%
-This type is well formed when $\alpha$ is the covariant \emph{and}
-contravariant variable of $P$.
-%
-An element of the type is a polymorphic element of $P$, varying in
-$\alpha$.
-%
-The introduction form is naturally a $\lambda$ abstracting over
-$\alpha:\cat C$ and the elimination form instantiates $\alpha$ with an
-arbitrary object $a : \cat C$.
-%
-We also include the necessary $\beta\eta$ equalities.
-
-\begin{figure}
   \begin{mathpar}
+    \inferrule
+    {}
+    {\Gamma\vdash \smallCats}
+
+    \inferrule
+    {}
+    {\Gamma\vdash \Cats}
+
+    \inferrule
+    {\Gamma \vdash \cat C \isaCat \and\Gamma \vdash \cat D \isaCat}
+    {\Gamma\pipe\cdot \vdash \varr{\cat C}{\cat D} \isaSet}
+
+    \inferrule
+    {\Gamma \vdash \cat C \isaCat \and\Gamma \vdash \cat D \isaCat}
+    {\Gamma\pipe\cdot \vdash \harr{\cat C}{\cat D} \isaSet}
+
     \inferrule
@@
     {\Gamma \vdash \pendallXdotY {\alpha:{\cat C}} P \isaTy}
-
-    \inferrule
-    {\Gamma \pipe \alpha:\cat C\vdash t : P}
-    {\Gamma \vdash \pendlambdaXdotY \alpha t : \pendallXdotY \alpha P}
-
-    \inferrule
-    {\Gamma \vdash M : \pendallXdotY \alpha P\and
-     \Gamma \pipe \beta:\cat D \vdash a : \cat C}
-    {\Gamma \pipe \beta:\cat D \vdash \pendappXtoY M a : P[a/\alpha]}
-
-    \pendappXtoY {(\pendlambdaXdotY \alpha t)} a = t[a/\alpha]
-
-    \inferrule*[right=2Cell$\eta$]
-    {\Gamma \vdash M : \pendallXdotY \alpha P }
-    {\Gamma \vdash M = \pendlambdaXdotY \alpha {\pendappXtoY M \alpha} : \pendallXdotY \alpha P}
-  \end{mathpar}
-  \caption{End}
-  \label{fig:end}
-\end{figure}
-
-Semantically, this type corresponds to taking the end of a profunctor
-$P : {\cat C}^{op} \times {\cat C} \to \Set$, typically written
-\[ \int^{\alpha:\cat C} P \alpha \alpha \]
-
-\subsection{Negative Multiplicatives}
-
-Since our calculus of natural transformations is ordered linear, we
-have two distinct ``function sets'', internalizing concatenation on
-the left or right of the context. The typing rules for these are given
-in \cref{fig:cotensor}. The two types are the \emph{right hom}
-$\homrallXYtoZ \alpha R P$ and the \emph{left hom} $\homlallXYtoZ
-\alpha R P$. As a type, we can think of these both as a kind of
-polymorphic function type in that in addition to parameterizing an
-element of the set $P$ by an element of the set $R$, they also
-quantify over an object $\alpha$ that is in $P$ and $R$.
-
-These behave like function types, but their typing is somewhat
-involved. First consider the right hom $\homrallXYtoZ \alpha R
-P$. This acts very much like a polymorphic function type $\forall
-\alpha. R \to P$, but with restrictions on the object
-variables. First, the object variable $\alpha$ must occur covariantly
-in $P$ and contravariantly in $R$. Next, the contravariant variable of
-the type is the contravariant variable of $P$ and the covariant
-variable of the type is the \emph{contravariant} variable of $R$, with
-the flip in variance intuitively justified since $R$ occurs in the
-negative position of the function. The placement of the $\forall
-\alpha$ on the right side of the triangle is intended to be suggestive
-of the fact that $\alpha$ occurs \emph{covariantly} in the input and
-output sets $R,P$ and also that the corresponding $\lambda$ introduces
-variables on the right.
-%
-As you would expect from a polymorphic function type, the introduction
-form is a $\lambda$ that abstracts over both a transformation variable
-and an object variable, which are concatenated on the right of the
-context, aligning with the condition that the contravariant variable
-of $R$ match the covariant variable of the original context.
-%
-The elimination rule is a combined instantiation and application form,
-$\homrappXtoYatZ s t a$ which instantiates the object $\alpha$ to $a$
-and passes an element $t : R[a/\alpha]$ as input, yielding an element
-$P[a/\alpha]$ as output. Finally, we have the expected $\beta$ and
-$\eta$ reasoning principles that we would have for a function type.
-
-The left hom $\homlallXYtoZ \alpha R P$ dually abstracts over context
-extension on the left. This time, the variable $\alpha$ occurs
-\emph{contravariantly} in the input and output sets, and the covariant
-dependence of the function set matches the covariant dependence of the
-ouput set, while the contravariant dependence is the covariant
-dependence of the input. In our notation, we place the $\forall\alpha$
-on the left to be suggestive of the fact that it is contravariant in
-$R$ and $P$. Next, the introduction form is again a polymorphic
-$\lambda$, but this time we write the object variable $\alpha$ first,
-and then the set variable $x$, to match the structure of the
-context. The elimination form is similarly an application, but to
-agree with the ordered use of variables, this time we place the
-argument to the left and the function to the right.
-
-\begin{figure}
-  % TODO: formalize the one-place substitution
-  \begin{mathpar}
-      \inferrule*[right=Covariant Hom formation]
-      {\Gamma \pipe d^+\Xi; \alpha : \cat C \vdash  R \and \Gamma \pipe  d^-\Xi; \alpha : \cat C \vdash P}
-      {\Gamma \pipe \Xi \vdash \homrallXYtoZ \alpha R P}
-
-      \inferrule*[right=Covariant Hom introduction]
-      {\Gamma \pipe \Phi, x : R, \alpha : \cat C \vdash t : P}
-      {\Gamma \pipe \Phi \vdash
-        \homrlambdaXatYdotZ {x : R} {\alpha : \cat C} t : \homrallXYtoZ \alpha R P}
-
-      %% \inferrule*[right=Covar Hom elim simple]
-      %% {}
-      %% {\alpha: \cat C, f:\forall \gamma. P \triangleleft R, \beta : \cat D, x : R, \gamma : \cat E \vdash f\triangleleft (x, \gamma) : P}
-
-      \inferrule*[right=Covariant Hom elimination]
-      {\Phi = \Phi_f \jnctx \Phi_a\and
-       \Gamma\pipe\Phi_f \vdash s : \homrallXYtoZ {\alpha:\cat C} R P \and
-       d^+\Phi_a \vdash a : \cat C\and
-       \Phi_a \vdash t : R[a/\alpha]
-        }
-      {\Gamma\pipe \Phi \vdash \homrappXtoYatZ s t a : P[a/\alpha]}\\
-
-      \homrappXtoYatZ {(\homrlambdaXatYdotZ x \alpha s)} t a = s[t/x,a/\alpha] (\beta)
-      %% (\lambda^\triangleleft (x, \gamma). M)\triangleleft (x, \gamma) = M (\beta)
-
-      \inferrule
-      {\Gamma\pipe\Phi \vdash s : \homrallXYtoZ \alpha R P}
-      {s = \homrlambdaXatYdotZ {x}{\alpha} \homrappXtoYatZ s x \alpha}
-
-      %% \inferrule
-      %% {\Gamma\pipe\Phi, x:R, \alpha : \cat C \vdash \homrappXtoYatZ s x \alpha = \homrappXtoYatZ t x \alpha : P}
-      %% {\Gamma\pipe\Phi \vdash s = t : \homrallXYtoZ \alpha R P}
-
-      \inferrule*[right=Contra Hom formation]
-      {\Gamma \pipe \alpha:\cat C; d^-\Xi\vdash R \and \Gamma \pipe \alpha : \cat C; d^+\Xi\vdash P}
-      {\Gamma \pipe \Xi \vdash \homlallXYtoZ \alpha R P}
-
-      \inferrule*[right=Contra Hom introduction]
-      {\Gamma\pipe\alpha:\cat C, x : R, \Phi \vdash t : P}
-      {\Gamma\pipe\Phi \vdash \homllambdaXatYdotZ {x : R} {\alpha : \cat C} t : \homlallXYtoZ \alpha R P}
-
-      \inferrule*[right=Contra Hom elimination]
-      {\Phi = \Phi_a \jnctx \Phi_f\and
-        \Gamma \pipe d^-\Phi_a \vdash a : \cat C\and
-       \Gamma \pipe \Phi_a \vdash s : R[a/\alpha] \and
-       \Gamma \pipe \Phi_f \vdash t : \homlallXYtoZ {\alpha:\cat C} R P}
-      {\Gamma \pipe \Phi \vdash \homlappXtoYatZ t s a : P[a/\alpha]}\\
-
-      \inferrule*[Right=$\beta$]
-                 {}
-                 {\homlappXtoYatZ {(\homllambdaXatYdotZ x \alpha t)} s a = t[a/\alpha,s/x]}
-
-      \inferrule*[right=$\eta$]
-      {\Gamma \pipe \Phi \vdash t : \homlallXYtoZ \alpha R P}
-      {\Gamma \pipe \Phi \vdash t = \homllambdaXatYdotZ x \alpha \homlappXtoYatZ t x \alpha : \homlallXYtoZ \alpha R P}
-
-      %% \inferrule*[right=$\eta$]
-      %% {\Gamma \pipe \alpha:\cat C, x:R, \Phi \vdash \homlappXtoYatZ M x \alpha = \homlappXtoYatZ N x \alpha : P}
-      %% {\Gamma \pipe \Phi \vdash M = N : \homlallXYtoZ \alpha R P}
-  \end{mathpar}
-  \caption{Hom of Relations/Profunctors}
-  \label{fig:hom}
-\end{figure}
-
-In category theory, these profunctors can be constructed using an
-\emph{end} and the hom of functor categories\cite{somecats}. Given two
-profunctors $P : C^o \times E \to \Set$ and $R : D^o \times E \to
-Set$, their hom is given as
-\[ \int_{e \in E} R(-,e) \Rightarrow P(e,=) \]
-
-%% The syntax is given in Figure~\ref{fig:cotensor}.
-%% %
-%% First, the formation rule for covariant cotensor formation requires
-%% that the covariant variables for $R,P$ are the same, and the
-%% contravariant variables match the boundary of the overall
-%% judgment. Note that we use $d^+\Xi,d^-\Xi$ to allow for the case that
-%% the boundary is a single variable $\Xi = \alpha:\cat C$.
-%% %
-%% Next, the substitution rule is straightforward.
-%% %
-%% Then we have the elimination form, first written in a simple form with
-%% no substitution built in.
-%% %
-%% This says that given a variable of the cotensor judgment, and another
-%% variable of judgment $R$, we can apply these to get a term of judgment
-%% $P$. As a diagram this is written as follows:
-
-%% \[ % https://tikzcd.yichuanshen.de/#N4Igdg9gJgpgziAXAbVABwnAlgFyxMJZABgBpiBdUkANwEMAbAVxiRAGEQBfU9TXfIRQBGclVqMWbACLdeIDNjwEiAJjHV6zVohABROXyWCiZYeK1TdnHkYEqU685sk793cTCgBzeEVAAZgBOEAC2SGQgOBBIohLabAA6iQEQQYwMAATJ3nShoXQAdJkACtmJOEFYdGDeDDD1ATiZAEqGIMFhsdTRSOrxViBtth0h4YgAzD0xiAAsLgm6Je2d45G9k9QMWGBuUBBMAEb1INQAFjB0UGyQu6dRdFgMNwSsI6t900jzIBdXL3ctjs9gdjqweo9nrpbm8KFwgA
-%% \begin{tikzcd}
-%% C \arrow[r, "\forall \gamma. P \triangleleft R"] \arrow[d, no head, Rightarrow] & D \arrow[r, "R"] & E \arrow[d, no head, Rightarrow] \\
-%% C \arrow[rr, "P"]                                                               &                  & E                               
-%% \end{tikzcd}\]
-
-%% The universal property of this judgment is that this is the universal
-%% 2-cell of this type, in that any 2-cell
-
-%% \[ % https://tikzcd.yichuanshen.de/#N4Igdg9gJgpgziAXAbVABwnAlgFyxMJZABgBpiBdUkANwEMAbAVxiRAGEByEAX1PUy58hFACZyVWoxZsAIr34gM2PASIBmCdXrNWiEAFEFAlcKJkAjJJ0z97Y0sGqRyTVe3S9hh8qFqUFlpSumy8kjBQAObwRKAAZgBOEAC2SIEgOBBI4sG2IABKDokpSJoZWYgALB4h+gAKRUmpiGTlpdQMWGBecBCdUCDUABYwdAOIYEwMDNQ4dFgMbJDdjSWIOZlI1SAjY0sErB1dXlAQTABGDIcZ84v6y6x88U1IrZuIAKxPIMXNH7MVdKdFb6KB0OAjAbDUbjcAHMI8IA
-%% \begin{tikzcd}
-%% C' \arrow[d] \arrow[r] & {} \arrow[r, no head, dashed] & D \arrow[r, "R"] & E \arrow[d, no head, Rightarrow] \\
-%% C \arrow[rrr, "P"]     &                               &                  & E                               
-%% \end{tikzcd}\]
-
-%% Exhibits a unique factorization:
-
-%% \[ % https://tikzcd.yichuanshen.de/#N4Igdg9gJgpgziAXAbVABwnAlgFyxMJZABgBoBGAXVJADcBDAGwFcYkQBhEAX1PU1z5CKAEwVqdJq3YARHnxAZseAkQDM4mgxZtEIAKLz+yoUTIiJ26Xq69jg1Sg0WtU3QaOKBK4SVLFLN3YOAHJPJQdfcn9AnXZw71NRGNc4vTk7LxNHZA0A1OsPbgkYKABzeCJQADMAJwgAWyRokBwIJDFJNJAAJU86xqQNVvbEABYC9wAFfvqmxDIRoZpGLDB3KAhmACNGNhoACxh6KHZIdZAaHHosRjOCNkyB+c62pAmQI5P7i5W1ja2u32rRudz050eCmeSAArFdRgA2J5zJAI+FIADsfwueig9DgR1Oh2Op3BD1mg0QWKWiBaqxxIE2Oz2l0+JJ+wOutw5FPmcJpxGRlMWb1pk3YAB0JdUILUmIwAARSsr0BoNegAOgVUyVEpwtSw9DAZT2e2qOAVfSF82pooAHOK9FaoSjEA6aZ16QDmcCvqTwOSrqCecVuEA
-%% \begin{tikzcd}
-%% C' \arrow[r] \arrow[d]                                                           & {} \arrow[r, no head, dashed] & D \arrow[d, no head, Rightarrow] \arrow[r, "R"] & E \arrow[d, no head, Rightarrow] \\
-%% C \arrow[d, no head, Rightarrow] \arrow[rr, "\forall \gamma.
-P \triangleleft R"] &                               & D \arrow[r, "R"]                                & E \arrow[d, no head, Rightarrow] \\
-%% C \arrow[rrr, "P"]                                                               &                               &                                                 & E                               
-%% \end{tikzcd}\]
-
-%% We give elimination rules and introduction rules in a style similar to
-%% $\lambda$-calculus, with corresponding $\beta\eta$ rules.
-
-These hom types are useful both for category-theoretic applications
-(Kan extensions, the Yoneda lemma), but also for purely syntactic
-reasons.  The universal property of these function types allows us, by
-iterated hom to express any sequent $\Phi \vdash R$ as an equivalent
-one with an ``empty'' context, i.e., a single object variable $d^\pm
-\Phi$. Then using the end type defined in the previous section, we can
-internalize that judgment as a type in our ambient dependent type
-theory.
-
-As an example, any sequent $\Gamma \pipe \alpha, x: R, \beta, y : S, \gamma \vdash
-P$ is equivalent to either of
-\[ \Gamma \vdash \pendallXdotY\alpha \homrallXYtoZ \beta R \homrallXYtoZ \gamma S P 
-\qquad\qquad\qquad \Gamma \vdash \pendallXdotY \gamma \homlallXYtoZ \beta S {\homlallXYtoZ \alpha R P}\]
-This property is useful in giving a simple syntax to positive types.
-
-\subsection{Unit and Tensor of Profunctors}
-
-While the hom sets of the previous section work quite smoothly since
-they are negative connectives in a natural deduction system, the positive
-connectives, the unit and tensor, are trickier to define.
-
-First, the multiplicative unit is the set $\punitinXfromYtoZ {\cat C}
-{a_1} {a_2}$, semantically, the hom set of the category $\cat C$. The
-rules for this type act like an identity type in a directed type
-theory. The introduction rule defines an identity morphism of an
-object in any category. The elimination rule is then the appropriate
-analogue of the $J$ rule of the identity type. Given any profunctor
-$Q$ from $\cat C$ to itself, we can construct an element $\harrapp Q
-{\alpha_1}{\alpha_2}$ for any objects in $\cat C$ if we specify a
-natural element $\harrapp Q \alpha \alpha$ for any $\alpha$ in $\cat
-C$. The reason this is sufficient, semantically, is that all
-profunctors must be functorial in their two indices. In fact, as we
-will see in the section on synthetic category theory, this induction
-principle is how we can show internally that profunctors are
-functorial in their indices. We choose to express this elimination
-form using the right hom type as its syntax more closely matches the
-usual syntax of function types, but this choice is arbitrary, and we
-would get an equivalent calculus had we used the left hom, or added
-both.
-
-Next, we continue with the tensor product $\tensorexistsXwithYandZ
-\beta P Q$. Adjoint to the right/left hom sets, this provides a form
-of \emph{existentially quantified pair}. The formation rule states the
-set is well-formed when $\beta$ is the covariant variable of $P$ and
-the contravariant variable of $Q$, and, its contravariant dependence
-is the same as $P$, while its covariant dependence is the same as $Q$.
-%
-The introduction rule is then a tuple $\tensorintroatXwithYandZ b s t$
-with an instantiation $b$ of the quantified variable $\beta$. Note
-that the use of $d^+\Psi_s$ in the rule could be replaced by
-$d^-\Psi_t$ since they must be equal by the side condition $\Phi =
-\jnctx \Psi_s \Psi_t$.
-%
-The elimination rule says to construct a transformation taking a pair
-as input it is sufficient to construct one that takes in a $P$ and $Q$
-for any $\beta$, directly corresponding to the adjoint relationship
-between the tensor and the right hom.
-%
-Finally, we have the expected $\beta\eta$ equations.
-
-\begin{figure}
-  \begin{mathpar}
-    \inferrule*[right=Unit-Formation]
-    {\Gamma \vdash \cat C \isaCat\and
-     \Gamma \pipe d^-\Xi \vdash a_1 : \cat C\and \Gamma \pipe d^+ \Xi \vdash a_2 : \cat C}
-    {\Gamma \pipe \Xi \vdash \punitinXfromYtoZ {\cat C} {a_1} {a_2}}
-
-    %% (A \to A')[\xi] = A[d^-\xi] \to A'[d^+\xi]
-
-    \inferrule*[right=Unit-Introduction]
-    {\Gamma\pipe\beta: \cat D \vdash a : \cat C}
-    {\Gamma \pipe\cdot \vdash \punitrefl a : \punitinXfromYtoZ {\cat C} a a}
-
-    \inferrule*[right=Unit Elim]{
-      \Gamma\vdash Q : \harr {\cat C} {\cat C}\and
-      \Gamma\vdash t : \pendallXdotY {\alpha:\cat C} {\harrapp Q \alpha \alpha}}
-    {\Gamma \vdash \punitelimtoYwithkontZ Q N :
-      \pendallXdotY {\alpha_1} {\homrallXYtoZ {\alpha_2} {\punitinXfromYtoZ {\cat C} {\alpha_1}{\alpha_2}} {\harrapp Q {\alpha_1}{\alpha_2}}}
-      }
-
-    %% J_\to(\alpha_1.\alpha_2. Q, \alpha. N, A_1, M, A_2)[\phi] = J_\to(\alpha_1.\alpha_2. Q, \alpha. N, A_1[d^-\phi], M[\phi], A_2[d^+\phi])
-
-    %% \inferrule*[right=Unit$\beta$]{\beta:\cat D \vdash A : \cat C\and
-    %%   \alpha:\cat C \vdash N : Q[\alpha/\alpha_1,\alpha/\alpha_2]
-    %% }{\beta : \cat D \vdash J_\to(\alpha_1.\alpha_2. Q, \alpha. N, A, \id(A), A) = N[A/\alpha] : Q[A/\alpha_1,A/\alpha_2]}
-    %%
-
-    \appunary {\punitelimtoYwithkontZ Q N} a {\punitrefl a} a = \pendappXtoY N a
-
-    \inferrule*[right=Unit$\eta$]
-    {\Gamma \vdash M : \pendallXdotY {\alpha_1} \homrallXYtoZ {\alpha_2} {\punitinXfromYtoZ {\cat C} {\alpha_1} {\alpha_2}} {\harrapp Q {\alpha_1} {\alpha_2}}}
-    {M = {\punitelimtoYwithkontZ Q {\appunary M \alpha {\punitrefl \alpha} \alpha}}}
-    %% {}
-    %% {\alpha_1 : \cat C, f : \alpha_1 \to_{\cat C} \alpha_2, \alpha_2:\cat C \vdash M : \harrapp Q {\alpha_1} {\alpha_2}}
-    %% {\alpha_1 : \cat C, f : \alpha_1 \to_{\cat C} \alpha_2, \alpha_2:\cat C \vdash M =
-    %%   \homrappXtoYatZ {\pendappXtoY {\punitelimtoYwithkontZ Q {\pendlambdaXdotY \alpha M[\alpha/\alpha_1,\pendappXtoY {\punitrefl C} \alpha/f,\alpha/\alpha_2]}} {\alpha_1}} f {\alpha_2}
-    %% : \harrapp Q {\alpha_1}{\alpha_2}}
-
-    \inferrule*[right=Tensor formation]
-    {\Gamma \pipe d^-\Xi; \beta:D\vdash P \isaSet\and
-     \Gamma \pipe \beta:D;d^+\Xi \vdash Q \isaSet}
-    {\Gamma \pipe \Xi \vdash \tensorexistsXwithYandZ {\beta:D} P Q}
-
-  %% \inferrule*[right=Tensor intro function]
-  %% {}
-  %% {\alpha : \cat C, q : Q, \beta : \cat D, r : R, \gamma : \cat E \vdash (q, \beta,r) : \exists \beta : \cat D. Q \odot R}
-
-  %% \inferrule*[right=Tensor introduction]
-  %% {\Gamma \vdash P : \harr C D \and
-  %%   \Gamma \vdash Q : \harr D E}
-  %% {\Gamma\pipe\cdot\vdash \tensorintroofXandY P Q : \pendallXdotY {\alpha} \homrallXYtoZ \beta {\harrapp P \alpha \beta} {\homrallXYtoZ \gamma {\harrapp Q \beta \gamma} {(\tensorexistsXwithYandZ \beta {\harrapp P \alpha \beta}{\harrapp Q \beta \gamma})}}}
-
-  \inferrule*[right=Tensor introduction]
-  {\Phi = \Psi_s \jnctx \Psi_t \and
-   \Gamma \pipe d^+\Psi_s \vdash b : \cat D\and
-   \Gamma \pipe \Psi_s \vdash s : P[b/\beta] \and
-   \Gamma \pipe \Psi_t \vdash t : Q[b/\beta]}
-  {\Gamma\pipe\Phi\vdash \tensorintroatXwithYandZ b s t : \tensorexistsXwithYandZ {\beta : \cat D} P Q}
-
-  %% \tensorintroatXwithYandZ B M N [\phi,B'/\beta',\psi] = \tensorintroatXwithYandZ {B[B'\beta]} {M[\phi]} {N[\psi]}
-
-  %% \inferrule*[right=Tensor simpl]
-  %% {\Gamma\pipe\cdot \vdash R : \harr {C} {E}\and
-  %%  \Gamma\pipe\cdot \vdash N : \pendallXdotY {\alpha:C} {\homrallXYtoZ {\beta:D} P {\homrallXYtoZ {\gamma:E} Q {\harrapp R \alpha\gamma}}}}
-  %% {\Gamma\pipe\alpha:C,p:\tensorexistsXwithYandZ {\beta:D} P Q,\gamma:E \vdash\tensorelimVatWtoXtoYwithkontZ p \alpha \gamma R N : \harrapp R A G}
-
-  \inferrule*[right=Tensor-elim]
-  {\Gamma \pipe \vdash P : \harr {\cat C} {\cat D}\and\Gamma \vdash Q : \harr {\cat D}{\cat E}\and
-   \Gamma \vdash R : \harr {\cat C}{\cat E}\and
-   \Gamma \vdash N : \hombinary \alpha {\harrapp P \alpha \beta} \beta {\harrapp Q \beta \gamma} \gamma {\harrapp R \alpha \gamma}}
-  {\Gamma \vdash \tensorelimfromWXtoYwithkontZ P Q R N : \homunary \alpha {(\tensorexistsXwithYandZ \beta {\harrapp P \alpha \beta} {\harrapp Q \beta \gamma})} {\gamma} {\harrapp R \alpha \gamma}}
-
-  %% \inferrule*[right=Tensor elim]
-  %% {\Gamma\pipe\cdot \vdash R : \harr {C} {E}\and
-  %%  \Gamma\pipe\Phi \vdash M : \tensorexistsXwithYandZ {\beta:D} {P[A/\alpha;]} {Q[;G/\gamma]}\and
-  %%  \Gamma\pipe\cdot \vdash N : \pendallXdotY {\alpha:C} {\homrallXYtoZ {\beta:D} P {\homrallXYtoZ {\gamma:E} Q {\harrapp R \alpha\gamma}}}}
-  %% {\Gamma\pipe\Phi \vdash\tensorelimVatWtoXtoYwithkontZ M A G R N : \harrapp R A G}
-
-  %% J_\odot(\alpha. \gamma. S, \alpha.q.\beta.r.\gamma.N, A, M, G)[\phi] =
-  %% J_\odot(\alpha. \gamma. S, \alpha.q.\beta.r.\gamma.N, A[d^-\phi], M[\phi], G[d^+\phi])
-
-  %% \inferrule*[right=Tensor elim Sequent]
-  %% {\Phi,\alpha : \cat C, q : Q, \beta : \cat D, r: R, \gamma : \cat E,\Psi \vdash N : S}
-  %% {\Phi, \alpha : \cat C, x : \exists \beta. Q \odot R, \gamma : \cat E,\Psi \vdash J_\odot(x;q.\beta.r. N) : S}
-
-  %% \inferrule*[right=Tensor elim Subst]
-  %% {\Phi,\alpha : \cat C, q : Q, \beta : \cat D, r: R, \gamma : \cat E,\Phi' \vdash N : S
-  %% \and
-  %% \phi : \Psi \to \Phi,\alpha : \cat C, q : Q, \beta : \cat D, r: R, \gamma : \cat E,\Phi'}
-  %% {\Psi  \vdash J_\odot(x;q.\beta.r. N)[\phi] : S}
-
-  \inferrule*[right=Tensor $\beta$]
-  {}
-  {\appunary {\tensorelimfromWXtoYwithkontZ P Q R N} a {\tensorintroatXwithYandZ b s t} c = \appbinary N a s b t c}
-  %% {\homrappXtoYatZ {\pendappXtoY{\tensorelimfromWXtoYwithkontZ P Q R N} A} {(\homrappXtoYatZ{\homrappXtoYatZ{\pendappXtoY {\tensorintroofXandY P Q} A} f B} g G)} G =
-  %%   \homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY N A} f B} g G}
-
-  %% \inferrule*[right=Tensor $\beta$]
-  %% {}{J_\odot(\alpha.\gamma.S, \alpha.q.\beta.r.\gamma.N,A,(M,B,N),G) = N[A/\alpha,M/q,B/\beta,N/r,G/\gamma]}
-
-  \inferrule*[right=Tensor$\eta$]
-  {\Gamma \vdash N : \homunary \alpha {(\tensorexistsXwithYandZ \beta {\harrapp P \alpha \beta}{\harrapp Q \beta \gamma})} \gamma {\harrapp R \alpha \gamma}}
-  {N = \tensorelimfromWXtoYwithkontZ P Q R {\lambdabinary \alpha x \beta y \gamma \appunary N \alpha {\tensorintroatXwithYandZ \beta x y} \gamma}}
-
-  %% \inferrule*[right=Tensor$\eta$]
-  %% {\alpha : \cat C, x : \tensorexistsXwithYandZ \beta P Q, \gamma : \cat E \vdash M : R}
-  %% {\alpha : \cat C, x : \exists \beta. Q \odot R, \gamma : \cat E \vdash M =
-  %%   \homrappXtoYatZ {\pendappXtoY{\tensorelimfromWXtoYwithkontZ P Q R
-  %%       {\pendlambdaXdotY \alpha {\homrlambdaXatYdotZ p \beta {\homrlambdaXatYdotZ q \gamma {M[\homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {\tensorintroofXandY P Q} \alpha} p \beta} q \gamma/x]}}}}
-  %%     } \alpha} x \gamma : S}
-  \end{mathpar}  
-  \caption{Unit, Tensor of Profunctors}
-\end{figure}
-
-%% Next, we can internalize the $,\alpha,$ as a positive connective as a
-%% kind of ``existentially quantified tensor product'', abstracting the
-%% composition of profunctors.
-
-%% The ordinary formula for this is that for profunctors $R : C^o \times
-%% D \to \Set$ and $S : D^o \times E \to \Set$, the composite is given by
-%% a coend:
-
-%% \[ \int^{d \in D} R(-,d) \otimes S(d,=) \]
-
-%% We will instead use slightly more logical notation, writing the coend as
-%% an existential quantifier, justified by the analogy between the two
-%% and the structure of the introduction/elimination forms in our syntax
-%% in particular.
-%% \begin{figure}
-%% \begin{mathpar}
-%% \end{mathpar}
-%% \caption{Tensor Primitive Rules}
-%% \label{fig:tensor}
-%% \end{figure}
-
-Semantically, the tensor product is the composition of profunctors,
-which is typically written as a coend
-\[ \int_{\beta} P(-,\beta) \times Q(\beta,=) \]
-
-\subsection{Graph of a profunctor}
-
-One final multiplicative we can define is the \emph{graph} of a
-profunctor. Given a profunctor $\alpha:\cat C;\beta:\cat D \vdash R$,
-the ``graph'' of $R$ is a category $\Sigma_{\alpha:\cat C;\beta:\cat
-  D}. R$. In detail, its objects are triples $(a:\cat C, r : R, b :
-\cat D)$ and morphism from $(a,r,b)$ to $(a', r', b')$ consists of a
-pair of morphisms $g : b \to b'$ and a backwards $f : a' \to a$ such
-that $r' = g\circ r \circ f$.
-%
-Intuitively this generalizes the ``graph'' of a relation $R : A\times
-B \to \Omega$, where we think of the graph as the corresponding subset
-$gr(R) \subseteq A\times B$
-%
-By picking $R = \punitinXfromYtoZ {\cat C} {\alpha}{\alpha'}$ we
-recover the \emph{arrow} category $\exists
-(\alpha;\alpha'). \punitinXfromYtoZ {\cat C}{\alpha}{\alpha'}$, by
-composing these with functors, we recover the notion of \emph{comma}
-category $\Sigma \alpha;\beta. \punitinXfromYtoZ {\cat
-  E}{(F\alpha)}{(G\beta)}$ and by picking either $F$ or $G$ to have
-domain the terminal category $1$, we recover the notion of a
-\emph{slice} category.
-
-This is unique among the multiplicatives in that it defines a
-\emph{category} rather than a profunctor. The graph of a profunctor is
-a \emph{negative} multiplicative, a kind of $\Sigma$-type for
-profunctors. The universal property says that a functor from $\cat E$
-into the graph $\Sigma (\alpha; \beta). R$ is equivalent to a
-transformation from $\cat E$ into the functor $R$, including functors
-from $\cat E$ to $\cat C$ and $\cat D$.
-%
-The graph of a profunctor is a joint generalization of several common
-constructions in category theory: the arrow category, slice
-categories, comma categories and the covariant and the covariant and
-contravariant category of elements.
-
-\begin{figure}
-  \begin{mathpar}
-    \inferrule*[right=Graph-Formation]
-    {\Gamma \vdash \alpha:\cat C; \beta:\cat D \vdash R \isaSet}
-    {\Gamma \vdash \exists (\alpha;\beta). R \isaCat}
-
-    \inferrule*[right=Graph-E+]
-    {\Gamma \vdash \gamma:\cat E \vdash \phi : \exists(\alpha:\cat C;\beta: \cat D).R}
-    {\Gamma \vdash \pi_+\phi : \cat C}
-
-    \inferrule*[right=Graph-E-]
-    {\Gamma \vdash \gamma:\cat E \vdash \phi : \exists(\alpha:\cat C;\beta: \cat D).R}
-    {\Gamma \vdash \pi_-\phi : \cat D}
-
-    \inferrule*[right=Graph-Ep]
-    {\Gamma \vdash \gamma:\cat E \vdash \phi : \exists(\alpha:\cat C;\beta: \cat D).R}
-    {\Gamma \vdash \gamma:\cat E \vdash \pi_p\phi : R[\pi_+\phi/\alpha,\pi_-\phi/\beta]}
-
-    \inferrule*[right=Graph-Intro]
-    {\Gamma\vdash \gamma:\cat E \vdash a : \cat C \and
-      \Gamma\vdash \gamma:\cat E \vdash b : \cat D \and
-      \Gamma\vdash \gamma:\cat E \vdash t : R[a/\alpha;b/\beta]}
-    {\Gamma\vdash \gamma:\cat E \vdash (a,b,t) : \exists (\alpha;\beta). R}
-
-    \inferrule*[right=Unit-Formation]
-    {\Gamma \vdash \cat C \isaCat\and
-     \Gamma \pipe d^-\Xi \vdash a_1 : \cat C\and \Gamma \pipe d^+ \Xi \vdash a_2 : \cat C}
-    {\Gamma \pipe \Xi \vdash \punitinXfromYtoZ {\cat C} {a_1} {a_2}}
-
-    \pi_+(a,b,t) = a \\
-    \pi_-(a,b,t) = b \\
-    \pi_p(a,b,t) = t\\
-
-    \inferrule
-    {\Gamma \vdash \gamma:\cat E \vdash \phi : \exists(\alpha:\cat C;\beta:\cat D). R}
-    {\phi = (\pi_+\phi, \pi_-\phi, \pi_p\phi)}
-  \end{mathpar}  
-  \caption{Graph of a Profunctor}
-\end{figure}
-
-\subsection{Positive and Negative Presheaf Categories}
-
-Though defining arbitrary functor categories type-theoretically would
-require a ``multi-argument functor'' judgment, we can define
-\emph{presheaf} categories by the universal property that a functor
-$\cat C \to \mathcal P {\cat D}$ is the same as a profunctor from
-$\cat D$ to $\cat C$. To distinguish the covariant from contravariant
-presheaves, we define them dually as $\negPresheaf {\cat C}$ and
-$\posPresheaf {\cat D}$
-
-\begin{figure}
-  \begin{mathpar}
-    \inferrule*[right=NegPresheaf-Formation]
-    {\Gamma \vdash \cat C \isaCat}
-    {\Gamma \vdash \negPresheaf {\cat C} \isaCat}
-
-    \inferrule*[right=NegPresheaf-E]
-    {\Gamma \pipe d^-\Xi \vdash a : \cat C \and
-     \Gamma \pipe d^+\Xi \vdash p : \negPresheaf {\cat C}
-    }
-    {\Gamma \pipe \Xi \vdash \negPresheafApp a p \isaSet}
-
-    \inferrule*[right=NegPresheaf-I]
-    {\Gamma \pipe \alpha:\cat C;\beta:\cat D \vdash R : \isaSet}
-    {\Gamma \pipe \beta:\cat D \vdash \lambda \alpha:\cat C. R : \negPresheaf {\cat C}}
-
-    \negPresheafApp a (\lambda \alpha. R) = R[a/\alpha]\\
-
-    \inferrule*[right=NegPresheaf-$\eta$]
-    {\Gamma \pipe \beta:\cat D \vdash p : \negPresheaf {\cat C}}
-    {p = \lambda \alpha. \negPresheafApp \alpha p}
-
-    \inferrule*[right=PosPresheaf-Formation]
-    {\Gamma \vdash \cat D \isaCat}
-    {\Gamma \vdash \posPresheaf {\cat D} \isaCat}
-
-    \inferrule*[right=PosPresheaf-E]
-    {\Gamma \pipe d^-\Xi \vdash p : \posPresheaf {\cat D}\and
-     \Gamma \pipe d^+\Xi \vdash a : \cat D
-    }
-    {\Gamma \pipe \Xi \vdash \posPresheafApp p a \isaSet}
-
-    \inferrule*[right=PosPresheaf-I]
-    {\Gamma \pipe \alpha:\cat C;\beta:\cat D \vdash R : \isaSet}
-    {\Gamma \pipe \alpha:\cat C \vdash \lambda \beta:\cat D. R : \posPresheaf {\cat D}}
-
-    \posPresheafApp {(\lambda \beta. R)} b = R[b/\beta]\\
-
-    \inferrule*[right=PosPresheaf-$\eta$]
-    {\Gamma \pipe \alpha:\cat C \vdash p : \posPresheaf {\cat D}}
-    {p = \lambda \beta. \posPresheafApp p \beta}
-  \end{mathpar}
-  \caption{Presheaf Categories}
-\end{figure}
-
-\subsection{Dependent Types}
-
-Next, we cover the basic dependent type theory structure: $\Pi,\Sigma$
-and identity types.
-
-\begin{figure}
-  \begin{mathpar}
-  \inferrule
-  {\Gamma \pipe\cdot \vdash S \isaSet\and \Gamma \pipe\cdot \vdash T \isaSet}
-  {\Gamma\pipe\cdot \vdash \prod_{x:S} T \isaSet}
-
-  \inferrule
-  {\Gamma \pipe \cdot \vdash M : \prod_{x:S} T\and
-    \Gamma \pipe \cdot \vdash N : S}
-  {\Gamma \pipe \cdot \vdash M N : T[N/x]}
-
-  \inferrule
-  {\Gamma,x:S \pipe \cdot \vdash M : T}
-  {\Gamma \pipe \cdot \vdash \lambda x:S. M : \prod_{x:S} T}
-
-  \inferrule
-  {\Gamma \pipe\cdot \vdash S \isaSet\and \Gamma \pipe\cdot \vdash T \isaSet}
-  {\Gamma\pipe\cdot \vdash \sum_{x:S} T \isaSet}
-
-  \inferrule
-  {~}
-  {\Gamma\pipe\cdot \vdash \top \isaSet}
-
-  \inferrule
-  {\Gamma \pipe \cdot \vdash M : S\and \Gamma \pipe \cdot \vdash N : S}
-  {\Gamma \pipe \cdot \vdash \textrm{Id}_S M N \isaSet}
-
-  \inferrule
-  {\Gamma \pipe \cdot \vdash M : S}
-  {\Gamma \pipe \cdot \vdash \textrm{refl}(M) : \textrm{Id}_S M M}
-
-  \inferrule
-  {\Gamma,x:S,y:S,p:\textrm{Id}_S x y \pipe \cdot \vdash C \isaSet\and
-   \Gamma \pipe \cdot \vdash M_k : \prod_{x:S} C[x/y,\textrm{refl}(x)/p]}
-  {\Gamma \pipe \cdot \vdash \textrm{J}(x,y,p.C,M) : \prod_{x:S,y:S,p:\textrm{Id}_S x y} C}
-  \end{mathpar}
-  \caption{Ordinary Dependent Type Structure}
-\end{figure}
-
-\subsection{Internalizing Virtual Equipment Structure}
-
-\begin{figure}
-  \begin{mathpar}
-    \inferrule
-    {\Gamma \vdash \cat C : \Sort \and\Gamma \vdash \cat D : \Sort}
-    {\Gamma\pipe\cdot \vdash \varr{\cat C}{\cat D} \isaSet}
 
@@
     {\Gamma\pipe \alpha:\cat C \vdash M A : \cat E}
-
-    (\lambda\alpha:C. A) B = A[B/\alpha]
-
-    \inferrule*[right=VArr $\eta$]
-    {\Gamma \pipe\cdot \vdash M : \varr C D}
-    {\Gamma \pipe\cdot \vdash M = (\lambda \alpha:C. M \alpha) : \varr C D}
-    \\
-
-    \inferrule
-    {\Gamma \vdash \cat C : \Sort \and\Gamma \vdash \cat D : \Sort}
-    {\Gamma\pipe\cdot \vdash \harr{\cat C}{\cat D} \isaSet}
 
@@
 
-    (\lambda \alpha:C;\beta:D. R)A;B = R[A/\alpha;B/\beta]
-
-    \inferrule*[right=HArr $\eta$]
-    {\Gamma \pipe \cdot \vdash M : \harr C D}
-    {\Gamma \pipe \cdot \vdash M = (\lambda \alpha;\beta. M \alpha;\beta) : \harr C D}\\
-
+    \inferrule
+    {\Gamma \pipe \alpha:\cat C\vdash t : P}
+    {\Gamma \vdash \pendlambdaXdotY \alpha t : \pendallXdotY \alpha P}
+
+    \inferrule
+    {\Gamma \vdash M : \pendallXdotY \alpha P\and
+     \Gamma \pipe \beta:\cat D \vdash a : \cat C}
+    {\Gamma \pipe \beta:\cat D \vdash \pendappXtoY M a : P[a/\alpha]}
   \end{mathpar}
-  \caption{Internalizing VE Structure}
+  \caption{Type Connectives}
 \end{figure}
-
-%% \subsection{Equalizers}
-
-%% To encode the concept of internal hom of profunctors, kan extensions
-%% and (co-)limits, we need to define the \emph{equalizer} of horizontal
-%% arrows. Specifically, the universal property that we need is an
-%% equalizer in the category whose objects are horizontal arrows (for
-%% fixed endpoints) and whose arrows are the 2-cells.
-
-%% \begin{figure}
-%%   \begin{mathpar}
-%%     \inferrule*[right=Equalizer-formation]
-%%     {\Gamma \pipe \cdot \vdash P : \harr\, C\,D\and
-%%      \Gamma \pipe \cdot \vdash Q : \harr\, C \, D\\\\
-%%      \Gamma \pipe \cdot \vdash \phi : \homunary \alpha {\harrapp P \alpha \beta} \beta {\harrapp Q \alpha \beta}\and
-%%      \Gamma \pipe \cdot \vdash \psi : \homunary \alpha {\harrapp P \alpha \beta} \beta {\harrapp Q \alpha \beta}\and}
-%%     {\Gamma \pipe \cdot \vdash \textrm{Equalizer}(C,D,P,\phi,\psi) : \harr\, C\, D}\and
-
-%%     \inferrule
-%%     {\Gamma \pipe \Phi \vdash M : \harrapp{\textrm{Equalizer}(C,D,P,\phi,\psi)} A B}
-%%     {\Gamma \pipe \Phi \vdash \textrm{forget}(M) : \harrapp P A B}
-
-%%     \inferrule
-%%     {\Gamma \pipe \Phi \vdash M : \harrapp P A B \and \appunary \phi A M B = \appunary \psi A M B}
-%%     {\Gamma \pipe \Phi \vdash \textrm{equalizing}_{\phi,\psi}(M) : \harrapp{\textrm{Equalizer}(C,D,P,\phi,\psi)} A B}
-
-%%     TODO: obvious equations
-%%   \end{mathpar}
-%%   \caption{Equalizers (Logical Framework style)}
-%% \end{figure}
-
-%% \begin{figure}
-%%   \begin{mathpar}
-%%     \inferrule*[right=Equalizer-formation]
-%%     {\Gamma \pipe \Xi \vdash P \isaSet\and\Gamma \pipe \Xi \vdash Q \isaSet\and
-%%       \Gamma\pipe d^-\Xi, p : P, d^+ \Xi \vdash \phi : Q\and
-%%       \Gamma\pipe d^-\Xi, p : P, d^+ \Xi \vdash \psi : Q }
-%%     {\Gamma \pipe \Xi \vdash \equalizeVofWbyXeqYatZ p P \phi \psi Q : \harr\, C\, D}\and
-
-%%     \inferrule*[right=Equalizer-formation]
-%%     {\Gamma \pipe \alpha:C;\beta:D \vdash P \isaSet\and\Gamma \pipe \alpha:C;\beta:D \vdash Q \isaSet\and
-%%       \Gamma \pipe d^-\Xi \vdash A : C\and
-%%       \Gamma \pipe d^+\Xi \vdash B : D\\\\
-%%       \Gamma\pipe \cdot \vdash \phi : \homunary \alpha P \beta Q\and
-%%       \Gamma\pipe \cdot \vdash \psi : \homunary \alpha P \beta Q
-%%     }
-%%     {\Gamma \pipe \Xi \vdash \equalizeVofWbyXeqYatZ p {P[A/\alpha;B/\beta]} {\appunary \phi A p B} {\appunary \psi A p B} {Q[A/\alpha;B/\beta]} : \harr\, C\, D}\and
-
-%%     \inferrule
-%%     {\Gamma \pipe \Phi \vdash M : \equalizeVofWbyXeqYatZ p P \phi \psi Q}
-%%     {\Gamma \pipe \Phi \vdash \pi_{eq}\,M : P}
-
-%%     \inferrule
-%%     {\Gamma \pipe \Phi \vdash M : \harrapp P A B \and \appunary \phi A M B = \appunary \psi A M B}
-%%     {\Gamma \pipe \Phi \vdash \textrm{equalizing}_{\phi,\psi}(M) : \harrapp{\textrm{Equalizer}(C,D,P,\phi,\psi)} A B}
-
-%%     TODO: equations
-%%   \end{mathpar}
-%%   \caption{Equalizers (Natural Deduction Style)}
-%% \end{figure}
-
-\subsection{Products}
-
-We add in parallel two different notions of product: products of sorts
-and products of spans, which have the universal property in their
-respective categories.
-
-\begin{figure}
-  \begin{mathpar}
-    \inferrule{}{\Gamma \vdash 1 : \Sort}\and
-    \inferrule*{}{\Gamma\pipe\alpha:C \vdash () : 1}\and
-    \inferrule*{\Gamma\pipe\alpha:C \vdash A : 1}{A = ()}
-    \inferrule{}{\Gamma\pipe \Xi \vdash 1 \isaSet}\and
-    \inferrule{}{\Gamma\pipe \Phi \vdash () :  1}\and
-    \inferrule{\Gamma\pipe \Phi \vdash M :  1}{M = ()}\\\\
-
-    \inferrule{\Gamma \vdash C : \Sort \and \Gamma \vdash D : \Sort}
-              {\Gamma \vdash C \times D : \Sort}\and
-
-    \inferrule{\forall i \in \{1,2\}.~ \Gamma \pipe \alpha:C \vdash A_i : D_i}{\Gamma \pipe \alpha:C \vdash (A_1,A_2) : D_1 \times D_2}
-
-    \inferrule{\Gamma \pipe \alpha:C \vdash A : D_1 \times D_2}{\Gamma \pipe \alpha:C \vdash \pi_i A : D_i}
-
-    \pi_i(A_1,A_2) = A_i\\
-    \inferrule{\Gamma\pipe \alpha:C \vdash A :  D_1\times D_2}{A = (\pi_1 A,\pi_2 A)}\\\\
-
-    \inferrule{\Gamma\pipe\Xi \vdash R \isaSet \and \Gamma\pipe\Xi \vdash S\isaSet}
-              {\Gamma\pipe\Xi \vdash R \times S\isaSet}\and
-
-    \inferrule{\forall i \in \{1,2\}.~ \Gamma \pipe\Phi \vdash M_i : R_i}{\Gamma \pipe\Phi \vdash (M_1,M_2) : R_1 \times R_2}
-
-    \inferrule{\Gamma \pipe\Phi\vdash M : R_1 \times R_2}{\Gamma \pipe\Phi \vdash \pi_i M : R_i}
-
-    \pi_i(M_1,M_2) = M_i\\
-    \inferrule{\Gamma\pipe\Phi \vdash M :  R_1\times R_2}{M = (\pi_1 M,\pi_2 M)}
-  \end{mathpar}
-  \caption{Products}
-  \label{fig:products}
-\end{figure}
-
-One additional connective we might desire is a ``parallel'' product of
-sets $R \times' S : \harr {C_1 \times C_2} {D_1 \times D_2}$ where $R
-: \harr {C_1} {D_1}$ and $R : \harr {C_2} {D_2}$. We don't need to
-define this as a primitive though because it is expressible in terms
-of the two kinds of products we have above:
-
-\[
-\inferrule
-{\Gamma \pipe \alpha : C_i; \beta : D_i \vdash R_i \isaSet}
-{\Gamma \pipe \alpha : C_1\times C_2; \beta : D_1 \times D_2
-  R_1[\pi_1\alpha/\alpha;\pi_1\beta/\beta] \times R_2[\pi_2\alpha/\alpha;\pi_2\beta/\beta]\isaSet}
-\]
-
-The empty products give us access now to several traditional notions
-of category theory that we couldn't before define, such as objects and
-presheaves of either variance:
-
-$$\textrm{Object}\,C = \varr 1 C$$
-$$\textrm{Presheaf}\,C = \harr 1 C$$
-$$\textrm{coPresheaf}\,C = \harr C 1$$
-
-However, we find that in practice we don't have much use for these
-concepts as most theorems about them easily generalize to more general
-theorems about arbitrary functors or profunctors.
-
-%% %% To define internal ends and coends, we need to go beyond the usual
-%% %% simple types and add equalizers (a form of subsets) and co-equalizers
-%% %% (a form of quotients) of judgments. Here's a sketch of the syntax for
-%% %% each. These make type checking very difficult as they result in
-%% %% obligations for proving equality of certain terms, which may or may
-%% %% not be decidable.
-
-%% %% \begin{mathpar}
-%% %%   \inferrule*[right=Equalizer Formation]
-%% %%   {\alpha;\beta\,|\, R \and
-%% %%   x : R \vdash t : S x:R \vdash u : S}
-%% %%   {\alpha;\beta\,|\,\{ x : R \,|\, t = u \}}
-
-%% %%   \inferrule*[right=Equalizer Intro]
-%% %%   {\Phi \vdash r : R \and t[r/x] = u[r/x]}
-%% %%   {\Phi \vdash \textrm{assert}(x.t = u)(r) : \{ x : R \,|\, t = u \}}
-
-%% %%   \inferrule*[right=Equalizer Elim]
-%% %%   {\Phi \vdash r : \{ x : R \,|\, t = u \}}
-%% %%   {\Phi \vdash \textrm{forget}(r) : R}
-
-%% %%   t[\textrm{forget}(r)/x] = u[\textrm{forget(r)}/x]
-
-%% %%   \textrm{assert}(x.t = u)(\textrm{forget}(r)) = r
-%% %%   \textrm{forget}(\textrm{assert}(x.t = u)(r)) = r
-%% %% \end{mathpar}
-
-%% %% \begin{mathpar}
-%% %%   \inferrule*[right=Co-equalizer Formation]
-%% %%   {\alpha;\beta\,|\, R \and
-%% %%   x : S \vdash t : R x:S \vdash u : R}
-%% %%   {\alpha;\beta\,|\,R/(\forall x:S. t = u)}
-
-%% %%   \inferrule*[right=Coequalizer Intro]
-%% %%   {\Phi \vdash r : R}
-%% %%   {\Phi \vdash \textrm{equate}(x:S.t=u)(r) : R/(\forall x.t=u)}
-
-%% %%   \inferrule*[right=Coequalizer Elim]
-%% %%   {\Psi \vdash r : R/(x.t = u) \and
-%% %%     \Phi,y:R,\Phi' \vdash s : S \and
-%% %%     s[t/y] = s[u/y]}
-%% %%   {\Phi,\Psi,\Phi' \vdash \textrm{lett} \textrm{equate}(y) = r.~s : S}
-
-%% %%   \textrm{equate}(x:S.t = u)(t[\gamma]) = \textrm{equate}(x:S.t = u)(u[\gamma])
-
-%% %%   \cdots
-%% %% \end{mathpar}
-
-%% %% \subsection{Positive Multiplicatives}
-
-%% %% \subsection{Negative Multiplicatives}
-
-%% %% \subsection{Additives}
 
@@
 
+\section{Discussion}
+
+\paragraph{Calculi for formal category theory}
+
+In their book on $\infty$-category theory, Riehl and Verity use the
+formal language of virtual equipments to prove results valid for
+$\infty$-categories without concrete manipulation of model
+categories. They formalize this language using Makkai's FOLDS
+framework of first-order logic with dependent sorts. The approach is
+similar in spirit to our \ohol{}, and has the same models. The main
+difference is that their theory is built on top of first-order logic,
+whereas ours is built on a custom-made variant of higher-order logic
+over a (unary) simple type theory. For this reason, statements in
+their theory are quite complex as they have to be encoded into
+first-order logic, whereas we argue our type-theoretic approach
+formalizes informal arguments much more directly, as shown in
+Section~\ref{sec:examples}.
+%
+An informal syntax for manipulating profunctors involving ends and
+coends, called the coend calculus is closely related, an extension of
+\ohol{} to treat profunctors of many variables of different variances
+should provide a formal treatment for the coend calculus
+\citet{coend-calculus}.
+
+Another approach to formal category theory is to develop a
+\emph{directed dependent type theory}, a type theory where the
+identity type is interpreted as morphisms in a (possibly
+$\infty$-)category. Research on directed dependent type theories is
+ongoing \cite{riehl-shulman,north}. Directed dependent type theories
+should generalize \ohol{} in the same way that dependent type theory
+generalizes higher order logic: internally there should be a type that
+models the category of sets, and terms of this type should model the
+profunctors of \ohol{}.  As a foundation for a domain-specific proof
+assistant for category theory, \ohol{} may provide a foundation for a
+proof assistant in the style of Isabelle/HOL \cite{isabelle-hol},
+whereas directed type theory would provide a foundation for a proof
+assistant in the style of Agda or Coq \cite{agda,coq}.
+
+\citet{djm-string-diagrams} provides a string diagram calculus for
+double categories and pro-arrow equipments, generalizing string
+diagrams for monoidal categories \citet{joyal?}. These are an
+alternative approach to type theoretic calculi, with the string
+diagrams typically making tensor products simpler to work with, while
+a type theoretic calculus like SCT makes the closed structure
+$\homrallXYtoZ \alpha P Q$ simpler to work with by using bound
+variables.
+
+One alternative to proarrow equipments is to axiomatize the bicategory
+of categories, profunctors and natural transformations rather than the
+virtual double category of categories, functors, profunctors and
+natural transformations. An abstract structure along
+these lines is a \emph{Cartesian bicategory}\footnote{which is not the
+same as a bicategory with products/finite limits, for which a syntax
+was developed \citet{fioresaville}} is given in
+\citet{carboni-walters,carboni-walters-someonelese}. 
+Some preliminary work on a proof system for Cartesian bicateogires was
+presented at \citet{jonasfreycartesianbicategories}. This bears strong
+similarities to our calculus VETT, except that (1) their profunctors
+may have 0, 1 or more covariant or contravariant variables but (2)
+they do not have a term syntax like our natural transformation syntax
+and (3) they do not have functors and restriction of profunctors by
+functors (since this structure is not present in a Cartesian
+bicategory). Semantically, Cartesian bicategories are somewhat less
+general than virtual equipments because in general the set of
+functors between two categories can only be recovered from the set of
+profunctors if the codomain category is Cauchy complete, in which case
+functors are equivalent to adjoint pairs of profunctors.
+
+Finally, some variations on double categories have been used to model
+the structure of certain program logics. \citet{newlicata20} present a
+logic for \emph{vertically thin} pro-arrow equipments, i.e., pro-arrow
+equipments where the functor and natural transformation judgments are
+posetal. This calculus is quite different from ours in that the
+horizontal morphisms are terms of a programming language, and so they
+are not presented as having co-variant and contravariant variables,
+and the equipment structure is modeled by explicit coercions rather
+than by syntactic substitution of vertical arrows in horizontal
+arrows. Another similar calculus is System P \citep{systemP} which is
+an internal language of \emph{reflexive graph categories}, which are
+like double categories without horizontal composition.
+
+\subsection{Category Theoretic Extensions}
+
+While we have demonstrated that some common constructions in category
+theory can be expressed using the simple combination of homs, tensors
+and units, there are several that are out of reach for SCT due to its
+restrictive judgmental structure. These require extensions to the
+judgmental structure that would move beyond virtual equipments.
+
+First, while we can add products to the calculus, adding in a
+\emph{functor category} ${\cat C}^{\cat D}$ would necessitate adding a
+syntax for functors of multiple arguments, to allow for the lambda rule:
+\[ \inferrule {\alpha:\cat C, \beta:\cat D \vdash e : \cat E} {\alpha:\cat C \vdash \lambda \beta:\cat D. e : \cat E} \]
+This would then require also some extension to profunctors for
+allowing 0 or more covariant/contravariant dependencies as well.
+%
+Semantically, this would correspond to generalizing from virtual
+equipments to some kind of (possibly cartesian) \emph{monoidal closed}
+virtual equipments.
+
+In addition, we could consider extending our semantics with a notion
+of \emph{opposite category}, with the property that a covariant
+dependence of a profunctor on a variable in a category should be
+equivalent to a contravariant dependence on a variable in the opposite
+category.
+%
+This could possibly be done by introducing a modal distinction
+$\alpha:^{+}\cat C$ for covariant dependence and $\alpha:^{-}\cat C$
+for contravariant dependence, with the opposite category being a
+modality that internalizes flipping the variance as a type
+constructor.
+%
+With these two extensions, possibly, the category contexts for
+functors and profunctors could be unified as one notion of a sequence
+of object variables with variances.
+%
+With this unification, it would then be simpler to incorporate a third
+extension: a universe object, corresponding to an object
+$\textrm{Set}$, which would allow us to internalize the profunctor
+judgment in the functor judgment, modeling that in classical category
+theory, profunctors are the same as certain functors to the category
+of Sets.
+%
+While all of these extensions seem plausible, they all move away to
+the simple ordered type theory structure that allowed us to define
+units, tensor and hom profunctors so easily. It is unclear how to
+define these types in this extended structure. As a simple example, if
+profunctors have many variables of mixed variance, rather than exactly
+one of each variance, it is unclear what the appropriate
+generalization of the transformation context $\Phi$ should be.
+
+An orthogonal extension to these would be to consider adding a
+\emph{discretization} modality $\Delta\cat C$ that gives the discrete
+category on the underlying set of objects of $\cat C$. This would give
+us the ability to formalize concepts like ``function on objects'' or
+``unnatural transformation'' without the need to define internal
+categories. This could be modeled by adding a modal distinction
+$\alpha:^{\#}\cat C$ for ``crisp'' dependence on an object of $\cat
+C$, i.e., dependence merely on the object of $\cat C$ with no
+functoriality/naturality conditions. This would then be internalized
+by the discretization modality.
+
+A third approach would be to extend the calculus to model general
+monoidal closed structures and more generally multicategories, which
+provides semantics of modal substructural type theories.
+%
+This can be modeled by extending the ``universe'' $\Set$ to be a
+monoidal closed category, which induces a monoidal closed structure on
+profunctors by the Day convolution product \cite{...}.
+%
+Similar developments along these lines have already been done for
+coend calculus \citep{coend-calculus} and in formalization of
+substructural type theories \citep{atkey-woods}.
+
+%% A recent development in the formalization of substructural type
+%% theories is to use monoidal connectives at the meta-level when
+%% defining multiplicative connectives\cite{woodsatkeyesop2022}. This
+%% suggest one approach to extending the theory to support the
+%% formalization of multiplicative universal properties would be, in
+%% adherence to the microcosm principle, to add linear judgments and
+%% multiplicative connectives to the metalevel.
+
+%% \paragraph{As a tool for Formal Type Theory}
+
+%% While we can model ``unary'' type theories using profunctors, to model
+%% theories like simple intuitionistic, linear or bunched type theories
+%% we need a way to talk about ``profunctors of many arguments''. This
+%% can be accomplished using \emph{multicategory} theory, which models
+%% the structure of the context with its structural rules as a monad of
+%% ``contexts and substitutions''. A suitable notion of monad on a
+%% virtual double category has been developed and was in fact the
+%% original motivation for the concept of a virtual double
+%% category. However while an abstract monad of contexts $T$ is enough to
+%% define what a $T$-multicategory internal to a virtual equipment is, it
+%% does not appear possible to define the multiplicative connectives as a
+%% universal property, i.e., to define when the multicategory is
+%% \emph{representable}.
+
+\subsection{Semantics}
+
+Our definition of a double category with displays is suggestive of a
+simple generalization. A double category is a monoid object in the
+double category of $Span(Cat)$, and a double category with displays
+furthermore uses certain diagrams involving profunctors, which also
+form a double category $Prof$. This suggests that the algebraic
+structure of a double category with displays is naturally modeled in
+some kind of \emph{triple category}, modeling the three types of
+morphisms: functors, profunctors and spans of functors.
+
+%% %%
+%% %% Bibliography
+%% %%
+
+%% %% Please use bibtex, 
+\subsubsection*{Acknowledgements}
+Grants stuff...? The authors would like to thank David Jaz Myers, Emily Riehl, Mike
+Shulman, Dominic Verity,...for helpful feedback on this work
+
+\bibliography{cats}
+
+\appendix
+
+\section{Syntactic Details}
+
+\begin{figure}
+  \begin{mathpar}
+    \begin{array}{rccl}
+      \textrm{Term Substitution} & \gamma, \delta & ::= & \cdot \pipe \gamma, M/X\\
+      \textrm{Object Substitution} & \xi, \zeta & ::= & a/\alpha \pipe a/\alpha;b/\beta\\
+      \textrm{Transformation Substitution} & \phi, \psi & ::= & a/\alpha \pipe \phi,s/x, \psi      
+    \end{array}
+
+      \inferrule*[right=TyCtxForm]
+        {}
+        {\Gamma \isadtctx}
+
+    \inferrule*[right=SetCtxForm]
+    {\Gamma \isadtctx}
+    {\Gamma \vdash \Xi \boundary}
+
+    \inferrule*[right=TransCtxForm]
+    {\Gamma \isadtctx}
+     {\Gamma \vdash \Phi \isavectx}
+    \inferrule
+    {\Delta \isadtctx \and \Gamma \isadtctx}
+    {\gamma : \Delta \to \Gamma}
+
+    \inferrule*[right=Boundary Map formation]
+    {\Gamma \vdash Z \boundary \and \Gamma \vdash \Xi \boundary}
+    {\Gamma \vdash \xi : Z \to \Xi}
+
+    \inferrule*[right=Cat Subst formation]
+    {\Gamma \vdash \Psi \isavectx \and \Gamma \vdash \Phi \isavectx}
+    {\Gamma \vdash \phi : \Psi \to \Phi}
+
+  \end{mathpar}
+  \caption{Substitution}
+
+\end{figure}
+
+\cref{fig:contexts} details the typing rules for contexts and the
+typing for variables and the admissible actions of substitution (the
+definition of substitution is straightforward and found in the
+appendix). the rules for term and object variables are as usual.
+%
+An object substitution into a single object context $\alpha:\cat C$
+consists of an object $a : \cat C$ parameterized by an object of
+another category.
+%
+An object substitution into a pair of co- and contra-variant variables
+$\alpha:\cat C;\beta:\cat D$ is a pair of an object substituted for
+$\alpha$ and one for $\beta$, each typed with respect to the
+appropriate variable of an object context $\Xi$.
+%
+Note that this allows for $\Xi$ to consist of either two variables or
+a single variable.
+%
+When
+constructing a transformation context $\Phi, x : R, \Psi$, we need to
+ensure that the contravariant dependence of $R$ matches the covariant
+dependence of $\Phi$ and vice-versa for $\Psi$. We define this with
+auxiliary metafunctions $d^+,d^-$ which extract the covariant and
+contravariant object variable of a transformation context.
+%
+The rules for transformation variables and substitutions define the
+use of variables to be \emph{linear} (each variables is used exactly
+once) and \emph{ordered}, they occur in the term in the same order
+that they appear in the context.
+%
+The rule for a transformation variables says a variable $x$ can only
+be used when it is the sole variable in the context.
+%
+A transformation subsitution into an ``empty'' transformation context
+$\alpha: \cat C$ is the same as an object substitution into
+$\alpha:\cat C$.
+%
+The rule for constructing a transformation substitution into a
+non-empty context $\Phi_1,x:R,\Phi_2$ says we can construct a
+substitution by constructing substitutions $\phi_1,\phi_2$ for
+$\Phi_1$ and $\Phi_2$ as well as a transformation $s$ into an
+instantiation of $R$ where the input contexts come from splitting the
+input context $\Psi$ into three components, and ensuring that the
+instantiation of $R$ agrees with how $\phi_1,\phi_2$ instantiate the
+variables. This uses two new metafunctions. First, $\Phi \jnctx \Psi$
+is an operation for concatenating two contexts when $d^+(\Phi) =
+d^-(\Psi)$. Next, we extend the boundary maps $d^\pm$ to substitutions
+to extract their object substitutions.
+
+\begin{figure}
+  \begin{mathpar}
+    \inferrule{}{\cdot \isadtctx}
+
+    \inferrule
+    {\Gamma \isadtctx \and \Gamma \vdash A \isaTy}
+    {\Gamma , X : A \isadtctx}\\
+
+    \inferrule
+    {X : A \in \Gamma}
+    {\Gamma \vdash X : A}
+
+    \inferrule
+    {}
+    {\alpha: \cat C \vdash \alpha: \cat C}
+
+    \inferrule
+    {\Gamma \vdash \cat C : \isaCat}
+    {\Gamma \vdash \alpha: \cat C \isavectx}
+
+    \inferrule
+    {\Gamma \vdash \Phi \isavectx\and
+    \Gamma \pipe d^+\Phi; \beta:\cat D \vdash R \isaSet}
+    {\Gamma \vdash \Phi, x:R, \beta:\cat D \isavectx}
+
+    \inferrule
+    {}
+    {\Gamma \pipe \alpha:\cat C, x : R, \beta: \cat D \vdash x : R}
+
+    % horizontal identity
+    \inferrule
+    {\Gamma\pipe \beta:\cat C \vdash a : \cat C}
+    {\Gamma \pipe \beta:\cat C \vdash a/\alpha : (\alpha : \cat C)}
+
+    \inferrule
+    {\Psi = \Psi_1 \jnctx \Psi_2\and
+      \Gamma\pipe\Psi_1 \vdash \phi : \Phi\and
+      \Gamma \pipe \Psi_2 \vdash t : R[d^+\phi;b/\beta] \and
+      \Gamma \vdash \phi_2 : \Psi_2 \to \Phi_2
+      }
+    {\Gamma\pipe\Psi \vdash \phi,t/x,b/\beta :: (\Phi,x:R,\beta:\cat D)}
+  \end{mathpar}
+  \begin{align*}
+    \underline{\alpha:C} &= \alpha:C\\
+    \underline{\Phi,x:R,\Psi} &= d^-\Phi; d^+{\Psi}
+  \end{align*}
+  \begin{align*}
+    d^{\pm}(\alpha:\cat C) &= \alpha : \cat C\\
+    d^{-}(\Phi,x:R,\Psi) &= d^-\Phi\\
+    d^{+}(\Phi,x:R,\Psi) &= d^-\Psi
+  \end{align*}
+  \begin{align*}
+    d^{\pm}(a/\alpha) &= a/\alpha\\
+    d^{-}(\phi,t/x,\psi) &= d^-\phi\\
+    d^{+}(\phi,t/x,\psi) &= d^+\psi\\
+  \end{align*}
+  \caption{Basic Rules of Judgment Theory}
+  \label{fig:contexts}
+\end{figure}
+
+Now that we have established a judgmental structure, in typical type
+theoretic fashion, our type constructs follow naturally as
+internalizing the judgments. Though the positive multiplicatives are
+more familiar, we start by describing the negative multiplicatives,
+since they are better behaved, and are useful to have already defined
+when discussing the positives.
+
+\subsection{Negative Multiplicatives}
+
+Since our calculus of natural transformations is ordered linear, we
+have two distinct ``function sets'', internalizing concatenation on
+the left or right of the context. The typing rules for these are given
+in \cref{fig:cotensor}. The two types are the \emph{right hom}
+$\homrallXYtoZ \alpha R P$ and the \emph{left hom} $\homlallXYtoZ
+\alpha R P$. As a type, we can think of these both as a kind of
+polymorphic function type in that in addition to parameterizing an
+element of the set $P$ by an element of the set $R$, they also
+quantify over an object $\alpha$ that is in $P$ and $R$.
+
+These behave like function types, but their typing is somewhat
+involved. First consider the right hom $\homrallXYtoZ \alpha R
+P$. This acts very much like a polymorphic function type $\forall
+\alpha. R \to P$, but with restrictions on the object
+variables. First, the object variable $\alpha$ must occur covariantly
+in $P$ and contravariantly in $R$. Next, the contravariant variable of
+the type is the contravariant variable of $P$ and the covariant
+variable of the type is the \emph{contravariant} variable of $R$, with
+the flip in variance intuitively justified since $R$ occurs in the
+negative position of the function. The placement of the $\forall
+\alpha$ on the right side of the triangle is intended to be suggestive
+of the fact that $\alpha$ occurs \emph{covariantly} in the input and
+output sets $R,P$ and also that the corresponding $\lambda$ introduces
+variables on the right.
+%
+As you would expect from a polymorphic function type, the introduction
+form is a $\lambda$ that abstracts over both a transformation variable
+and an object variable, which are concatenated on the right of the
+context, aligning with the condition that the contravariant variable
+of $R$ match the covariant variable of the original context.
+%
+The elimination rule is a combined instantiation and application form,
+$\homrappXtoYatZ s t a$ which instantiates the object $\alpha$ to $a$
+and passes an element $t : R[a/\alpha]$ as input, yielding an element
+$P[a/\alpha]$ as output. Finally, we have the expected $\beta$ and
+$\eta$ reasoning principles that we would have for a function type.
+
+The left hom $\homlallXYtoZ \alpha R P$ dually abstracts over context
+extension on the left. This time, the variable $\alpha$ occurs
+\emph{contravariantly} in the input and output sets, and the covariant
+dependence of the function set matches the covariant dependence of the
+ouput set, while the contravariant dependence is the covariant
+dependence of the input. In our notation, we place the $\forall\alpha$
+on the left to be suggestive of the fact that it is contravariant in
+$R$ and $P$. Next, the introduction form is again a polymorphic
+$\lambda$, but this time we write the object variable $\alpha$ first,
+and then the set variable $x$, to match the structure of the
+context. The elimination form is similarly an application, but to
+agree with the ordered use of variables, this time we place the
+argument to the left and the function to the right.
+
+\begin{figure}
+  % TODO: formalize the one-place substitution
+  \begin{mathpar}
+      \inferrule*[right=Covariant Hom formation]
+      {\Gamma \pipe d^+\Xi; \alpha : \cat C \vdash  R \and \Gamma \pipe  d^-\Xi; \alpha : \cat C \vdash P}
+      {\Gamma \pipe \Xi \vdash \homrallXYtoZ \alpha R P}
+
+      \inferrule*[right=Covariant Hom introduction]
+      {\Gamma \pipe \Phi, x : R, \alpha : \cat C \vdash t : P}
+      {\Gamma \pipe \Phi \vdash
+        \homrlambdaXatYdotZ {x : R} {\alpha : \cat C} t : \homrallXYtoZ \alpha R P}
+
+      %% \inferrule*[right=Covar Hom elim simple]
+      %% {}
+      %% {\alpha: \cat C, f:\forall \gamma. P \triangleleft R, \beta : \cat D, x : R, \gamma : \cat E \vdash f\triangleleft (x, \gamma) : P}
+
+      \inferrule*[right=Covariant Hom elimination]
+      {\Phi = \Phi_f \jnctx \Phi_a\and
+       \Gamma\pipe\Phi_f \vdash s : \homrallXYtoZ {\alpha:\cat C} R P \and
+       d^+\Phi_a \vdash a : \cat C\and
+       \Phi_a \vdash t : R[a/\alpha]
+        }
+      {\Gamma\pipe \Phi \vdash \homrappXtoYatZ s t a : P[a/\alpha]}\\
+
+      \homrappXtoYatZ {(\homrlambdaXatYdotZ x \alpha s)} t a = s[t/x,a/\alpha] (\beta)
+      %% (\lambda^\triangleleft (x, \gamma). M)\triangleleft (x, \gamma) = M (\beta)
+
+      \inferrule
+      {\Gamma\pipe\Phi \vdash s : \homrallXYtoZ \alpha R P}
+      {s = \homrlambdaXatYdotZ {x}{\alpha} \homrappXtoYatZ s x \alpha}
+
+      %% \inferrule
+      %% {\Gamma\pipe\Phi, x:R, \alpha : \cat C \vdash \homrappXtoYatZ s x \alpha = \homrappXtoYatZ t x \alpha : P}
+      %% {\Gamma\pipe\Phi \vdash s = t : \homrallXYtoZ \alpha R P}
+
+      \inferrule*[right=Contra Hom formation]
+      {\Gamma \pipe \alpha:\cat C; d^-\Xi\vdash R \and \Gamma \pipe \alpha : \cat C; d^+\Xi\vdash P}
+      {\Gamma \pipe \Xi \vdash \homlallXYtoZ \alpha R P}
+
+      \inferrule*[right=Contra Hom introduction]
+      {\Gamma\pipe\alpha:\cat C, x : R, \Phi \vdash t : P}
+      {\Gamma\pipe\Phi \vdash \homllambdaXatYdotZ {x : R} {\alpha : \cat C} t : \homlallXYtoZ \alpha R P}
+
+      \inferrule*[right=Contra Hom elimination]
+      {\Phi = \Phi_a \jnctx \Phi_f\and
+        \Gamma \pipe d^-\Phi_a \vdash a : \cat C\and
+       \Gamma \pipe \Phi_a \vdash s : R[a/\alpha] \and
+       \Gamma \pipe \Phi_f \vdash t : \homlallXYtoZ {\alpha:\cat C} R P}
+      {\Gamma \pipe \Phi \vdash \homlappXtoYatZ t s a : P[a/\alpha]}\\
+
+      \inferrule*[Right=$\beta$]
+                 {}
+                 {\homlappXtoYatZ {(\homllambdaXatYdotZ x \alpha t)} s a = t[a/\alpha,s/x]}
+
+      \inferrule*[right=$\eta$]
+      {\Gamma \pipe \Phi \vdash t : \homlallXYtoZ \alpha R P}
+      {\Gamma \pipe \Phi \vdash t = \homllambdaXatYdotZ x \alpha \homlappXtoYatZ t x \alpha : \homlallXYtoZ \alpha R P}
+
+      %% \inferrule*[right=$\eta$]
+      %% {\Gamma \pipe \alpha:\cat C, x:R, \Phi \vdash \homlappXtoYatZ M x \alpha = \homlappXtoYatZ N x \alpha : P}
+      %% {\Gamma \pipe \Phi \vdash M = N : \homlallXYtoZ \alpha R P}
+  \end{mathpar}
+  \caption{Hom of Relations/Profunctors}
+  \label{fig:hom}
+\end{figure}
+
+In category theory, these profunctors can be constructed using an
+\emph{end} and the hom of functor categories\cite{somecats}. Given two
+profunctors $P : C^o \times E \to \Set$ and $R : D^o \times E \to
+Set$, their hom is given as
+\[ \int_{e \in E} R(-,e) \Rightarrow P(e,=) \]
+
+%% The syntax is given in Figure~\ref{fig:cotensor}.
+%% %
+%% First, the formation rule for covariant cotensor formation requires
+%% that the covariant variables for $R,P$ are the same, and the
+%% contravariant variables match the boundary of the overall
+%% judgment. Note that we use $d^+\Xi,d^-\Xi$ to allow for the case that
+%% the boundary is a single variable $\Xi = \alpha:\cat C$.
+%% %
+%% Next, the substitution rule is straightforward.
+%% %
+%% Then we have the elimination form, first written in a simple form with
+%% no substitution built in.
+%% %
+%% This says that given a variable of the cotensor judgment, and another
+%% variable of judgment $R$, we can apply these to get a term of judgment
+%% $P$. As a diagram this is written as follows:
+
+%% \[ % https://tikzcd.yichuanshen.de/#N4Igdg9gJgpgziAXAbVABwnAlgFyxMJZABgBpiBdUkANwEMAbAVxiRAGEQBfU9TXfIRQBGclVqMWbACLdeIDNjwEiAJjHV6zVohABROXyWCiZYeK1TdnHkYEqU685sk793cTCgBzeEVAAZgBOEAC2SGQgOBBIohLabAA6iQEQQYwMAATJ3nShoXQAdJkACtmJOEFYdGDeDDD1ATiZAEqGIMFhsdTRSOrxViBtth0h4YgAzD0xiAAsLgm6Je2d45G9k9QMWGBuUBBMAEb1INQAFjB0UGyQu6dRdFgMNwSsI6t900jzIBdXL3ctjs9gdjqweo9nrpbm8KFwgA
+%% \begin{tikzcd}
+%% C \arrow[r, "\forall \gamma. P \triangleleft R"] \arrow[d, no head, Rightarrow] & D \arrow[r, "R"] & E \arrow[d, no head, Rightarrow] \\
+%% C \arrow[rr, "P"]                                                               &                  & E                               
+%% \end{tikzcd}\]
+
+%% The universal property of this judgment is that this is the universal
+%% 2-cell of this type, in that any 2-cell
+
+%% \[ % https://tikzcd.yichuanshen.de/#N4Igdg9gJgpgziAXAbVABwnAlgFyxMJZABgBpiBdUkANwEMAbAVxiRAGEByEAX1PUy58hFACZyVWoxZsAIr34gM2PASIBmCdXrNWiEAFEFAlcKJkAjJJ0z97Y0sGqRyTVe3S9hh8qFqUFlpSumy8kjBQAObwRKAAZgBOEAC2SIEgOBBI4sG2IABKDokpSJoZWYgALB4h+gAKRUmpiGTlpdQMWGBecBCdUCDUABYwdAOIYEwMDNQ4dFgMbJDdjSWIOZlI1SAjY0sErB1dXlAQTABGDIcZ84v6y6x88U1IrZuIAKxPIMXNH7MVdKdFb6KB0OAjAbDUbjcAHMI8IA
+%% \begin{tikzcd}
+%% C' \arrow[d] \arrow[r] & {} \arrow[r, no head, dashed] & D \arrow[r, "R"] & E \arrow[d, no head, Rightarrow] \\
+%% C \arrow[rrr, "P"]     &                               &                  & E                               
+%% \end{tikzcd}\]
+
+%% Exhibits a unique factorization:
+
+%% \[ % https://tikzcd.yichuanshen.de/#N4Igdg9gJgpgziAXAbVABwnAlgFyxMJZABgBoBGAXVJADcBDAGwFcYkQBhEAX1PU1z5CKAEwVqdJq3YARHnxAZseAkQDM4mgxZtEIAKLz+yoUTIiJ26Xq69jg1Sg0WtU3QaOKBK4SVLFLN3YOAHJPJQdfcn9AnXZw71NRGNc4vTk7LxNHZA0A1OsPbgkYKABzeCJQADMAJwgAWyRokBwIJDFJNJAAJU86xqQNVvbEABYC9wAFfvqmxDIRoZpGLDB3KAhmACNGNhoACxh6KHZIdZAaHHosRjOCNkyB+c62pAmQI5P7i5W1ja2u32rRudz050eCmeSAArFdRgA2J5zJAI+FIADsfwueig9DgR1Oh2Op3BD1mg0QWKWiBaqxxIE2Oz2l0+JJ+wOutw5FPmcJpxGRlMWb1pk3YAB0JdUILUmIwAARSsr0BoNegAOgVUyVEpwtSw9DAZT2e2qOAVfSF82pooAHOK9FaoSjEA6aZ16QDmcCvqTwOSrqCecVuEA
+%% \begin{tikzcd}
+%% C' \arrow[r] \arrow[d]                                                           & {} \arrow[r, no head, dashed] & D \arrow[d, no head, Rightarrow] \arrow[r, "R"] & E \arrow[d, no head, Rightarrow] \\
+%% C \arrow[d, no head, Rightarrow] \arrow[rr, "\forall \gamma.
+P \triangleleft R"] &                               & D \arrow[r, "R"]                                & E \arrow[d, no head, Rightarrow] \\
+%% C \arrow[rrr, "P"]                                                               &                               &                                                 & E                               
+%% \end{tikzcd}\]
+
+%% We give elimination rules and introduction rules in a style similar to
+%% $\lambda$-calculus, with corresponding $\beta\eta$ rules.
+
+These hom types are useful both for category-theoretic applications
+(Kan extensions, the Yoneda lemma), but also for purely syntactic
+reasons.  The universal property of these function types allows us, by
+iterated hom to express any sequent $\Phi \vdash R$ as an equivalent
+one with an ``empty'' context, i.e., a single object variable $d^\pm
+\Phi$. Then using the end type defined in the previous section, we can
+internalize that judgment as a type in our ambient dependent type
+theory.
+
+As an example, any sequent $\Gamma \pipe \alpha, x: R, \beta, y : S, \gamma \vdash
+P$ is equivalent to either of
+\[ \Gamma \vdash \pendallXdotY\alpha \homrallXYtoZ \beta R \homrallXYtoZ \gamma S P 
+\qquad\qquad\qquad \Gamma \vdash \pendallXdotY \gamma \homlallXYtoZ \beta S {\homlallXYtoZ \alpha R P}\]
+This property is useful in giving a simple syntax to positive types.
+
+\subsection{Unit and Tensor of Profunctors}
+
+While the hom sets of the previous section work quite smoothly since
+they are negative connectives in a natural deduction system, the positive
+connectives, the unit and tensor, are trickier to define.
+
+First, the multiplicative unit is the set $\punitinXfromYtoZ {\cat C}
+{a_1} {a_2}$, semantically, the hom set of the category $\cat C$. The
+rules for this type act like an identity type in a directed type
+theory. The introduction rule defines an identity morphism of an
+object in any category. The elimination rule is then the appropriate
+analogue of the $J$ rule of the identity type. Given any profunctor
+$Q$ from $\cat C$ to itself, we can construct an element $\harrapp Q
+{\alpha_1}{\alpha_2}$ for any objects in $\cat C$ if we specify a
+natural element $\harrapp Q \alpha \alpha$ for any $\alpha$ in $\cat
+C$. The reason this is sufficient, semantically, is that all
+profunctors must be functorial in their two indices. In fact, as we
+will see in the section on synthetic category theory, this induction
+principle is how we can show internally that profunctors are
+functorial in their indices. We choose to express this elimination
+form using the right hom type as its syntax more closely matches the
+usual syntax of function types, but this choice is arbitrary, and we
+would get an equivalent calculus had we used the left hom, or added
+both.
+
+Next, we continue with the tensor product $\tensorexistsXwithYandZ
+\beta P Q$. Adjoint to the right/left hom sets, this provides a form
+of \emph{existentially quantified pair}. The formation rule states the
+set is well-formed when $\beta$ is the covariant variable of $P$ and
+the contravariant variable of $Q$, and, its contravariant dependence
+is the same as $P$, while its covariant dependence is the same as $Q$.
+%
+The introduction rule is then a tuple $\tensorintroatXwithYandZ b s t$
+with an instantiation $b$ of the quantified variable $\beta$. Note
+that the use of $d^+\Psi_s$ in the rule could be replaced by
+$d^-\Psi_t$ since they must be equal by the side condition $\Phi =
+\jnctx \Psi_s \Psi_t$.
+%
+The elimination rule says to construct a transformation taking a pair
+as input it is sufficient to construct one that takes in a $P$ and $Q$
+for any $\beta$, directly corresponding to the adjoint relationship
+between the tensor and the right hom.
+%
+Finally, we have the expected $\beta\eta$ equations.
+
+\begin{figure}
+  \begin{mathpar}
+    \inferrule*[right=Unit-Formation]
+    {\Gamma \vdash \cat C \isaCat\and
+     \Gamma \pipe d^-\Xi \vdash a_1 : \cat C\and \Gamma \pipe d^+ \Xi \vdash a_2 : \cat C}
+    {\Gamma \pipe \Xi \vdash \punitinXfromYtoZ {\cat C} {a_1} {a_2}}
+
+    %% (A \to A')[\xi] = A[d^-\xi] \to A'[d^+\xi]
+
+    \inferrule*[right=Unit-Introduction]
+    {\Gamma\pipe\beta: \cat D \vdash a : \cat C}
+    {\Gamma \pipe\cdot \vdash \punitrefl a : \punitinXfromYtoZ {\cat C} a a}
+
+    \inferrule*[right=Unit Elim]{
+      \Gamma\vdash Q : \harr {\cat C} {\cat C}\and
+      \Gamma\vdash t : \pendallXdotY {\alpha:\cat C} {\harrapp Q \alpha \alpha}}
+    {\Gamma \vdash \punitelimtoYwithkontZ Q N :
+      \pendallXdotY {\alpha_1} {\homrallXYtoZ {\alpha_2} {\punitinXfromYtoZ {\cat C} {\alpha_1}{\alpha_2}} {\harrapp Q {\alpha_1}{\alpha_2}}}
+      }
+
+    %% J_\to(\alpha_1.\alpha_2. Q, \alpha. N, A_1, M, A_2)[\phi] = J_\to(\alpha_1.\alpha_2. Q, \alpha. N, A_1[d^-\phi], M[\phi], A_2[d^+\phi])
+
+    %% \inferrule*[right=Unit$\beta$]{\beta:\cat D \vdash A : \cat C\and
+    %%   \alpha:\cat C \vdash N : Q[\alpha/\alpha_1,\alpha/\alpha_2]
+    %% }{\beta : \cat D \vdash J_\to(\alpha_1.\alpha_2. Q, \alpha. N, A, \id(A), A) = N[A/\alpha] : Q[A/\alpha_1,A/\alpha_2]}
+    %%
+
+    \appunary {\punitelimtoYwithkontZ Q N} a {\punitrefl a} a = \pendappXtoY N a
+
+    \inferrule*[right=Unit$\eta$]
+    {\Gamma \vdash M : \pendallXdotY {\alpha_1} \homrallXYtoZ {\alpha_2} {\punitinXfromYtoZ {\cat C} {\alpha_1} {\alpha_2}} {\harrapp Q {\alpha_1} {\alpha_2}}}
+    {M = {\punitelimtoYwithkontZ Q {\appunary M \alpha {\punitrefl \alpha} \alpha}}}
+    %% {}
+    %% {\alpha_1 : \cat C, f : \alpha_1 \to_{\cat C} \alpha_2, \alpha_2:\cat C \vdash M : \harrapp Q {\alpha_1} {\alpha_2}}
+    %% {\alpha_1 : \cat C, f : \alpha_1 \to_{\cat C} \alpha_2, \alpha_2:\cat C \vdash M =
+    %%   \homrappXtoYatZ {\pendappXtoY {\punitelimtoYwithkontZ Q {\pendlambdaXdotY \alpha M[\alpha/\alpha_1,\pendappXtoY {\punitrefl C} \alpha/f,\alpha/\alpha_2]}} {\alpha_1}} f {\alpha_2}
+    %% : \harrapp Q {\alpha_1}{\alpha_2}}
+
+    \inferrule*[right=Tensor formation]
+    {\Gamma \pipe d^-\Xi; \beta:D\vdash P \isaSet\and
+     \Gamma \pipe \beta:D;d^+\Xi \vdash Q \isaSet}
+    {\Gamma \pipe \Xi \vdash \tensorexistsXwithYandZ {\beta:D} P Q}
+
+  %% \inferrule*[right=Tensor intro function]
+  %% {}
+  %% {\alpha : \cat C, q : Q, \beta : \cat D, r : R, \gamma : \cat E \vdash (q, \beta,r) : \exists \beta : \cat D. Q \odot R}
+
+  %% \inferrule*[right=Tensor introduction]
+  %% {\Gamma \vdash P : \harr C D \and
+  %%   \Gamma \vdash Q : \harr D E}
+  %% {\Gamma\pipe\cdot\vdash \tensorintroofXandY P Q : \pendallXdotY {\alpha} \homrallXYtoZ \beta {\harrapp P \alpha \beta} {\homrallXYtoZ \gamma {\harrapp Q \beta \gamma} {(\tensorexistsXwithYandZ \beta {\harrapp P \alpha \beta}{\harrapp Q \beta \gamma})}}}
+
+  \inferrule*[right=Tensor introduction]
+  {\Phi = \Psi_s \jnctx \Psi_t \and
+   \Gamma \pipe d^+\Psi_s \vdash b : \cat D\and
+   \Gamma \pipe \Psi_s \vdash s : P[b/\beta] \and
+   \Gamma \pipe \Psi_t \vdash t : Q[b/\beta]}
+  {\Gamma\pipe\Phi\vdash \tensorintroatXwithYandZ b s t : \tensorexistsXwithYandZ {\beta : \cat D} P Q}
+
+  %% \tensorintroatXwithYandZ B M N [\phi,B'/\beta',\psi] = \tensorintroatXwithYandZ {B[B'\beta]} {M[\phi]} {N[\psi]}
+
+  %% \inferrule*[right=Tensor simpl]
+  %% {\Gamma\pipe\cdot \vdash R : \harr {C} {E}\and
+  %%  \Gamma\pipe\cdot \vdash N : \pendallXdotY {\alpha:C} {\homrallXYtoZ {\beta:D} P {\homrallXYtoZ {\gamma:E} Q {\harrapp R \alpha\gamma}}}}
+  %% {\Gamma\pipe\alpha:C,p:\tensorexistsXwithYandZ {\beta:D} P Q,\gamma:E \vdash\tensorelimVatWtoXtoYwithkontZ p \alpha \gamma R N : \harrapp R A G}
+
+  \inferrule*[right=Tensor-elim]
+  {\Gamma \pipe \vdash P : \harr {\cat C} {\cat D}\and\Gamma \vdash Q : \harr {\cat D}{\cat E}\and
+   \Gamma \vdash R : \harr {\cat C}{\cat E}\and
+   \Gamma \vdash N : \hombinary \alpha {\harrapp P \alpha \beta} \beta {\harrapp Q \beta \gamma} \gamma {\harrapp R \alpha \gamma}}
+  {\Gamma \vdash \tensorelimfromWXtoYwithkontZ P Q R N : \homunary \alpha {(\tensorexistsXwithYandZ \beta {\harrapp P \alpha \beta} {\harrapp Q \beta \gamma})} {\gamma} {\harrapp R \alpha \gamma}}
+
+  %% \inferrule*[right=Tensor elim]
+  %% {\Gamma\pipe\cdot \vdash R : \harr {C} {E}\and
+  %%  \Gamma\pipe\Phi \vdash M : \tensorexistsXwithYandZ {\beta:D} {P[A/\alpha;]} {Q[;G/\gamma]}\and
+  %%  \Gamma\pipe\cdot \vdash N : \pendallXdotY {\alpha:C} {\homrallXYtoZ {\beta:D} P {\homrallXYtoZ {\gamma:E} Q {\harrapp R \alpha\gamma}}}}
+  %% {\Gamma\pipe\Phi \vdash\tensorelimVatWtoXtoYwithkontZ M A G R N : \harrapp R A G}
+
+  %% J_\odot(\alpha. \gamma. S, \alpha.q.\beta.r.\gamma.N, A, M, G)[\phi] =
+  %% J_\odot(\alpha. \gamma. S, \alpha.q.\beta.r.\gamma.N, A[d^-\phi], M[\phi], G[d^+\phi])
+
+  %% \inferrule*[right=Tensor elim Sequent]
+  %% {\Phi,\alpha : \cat C, q : Q, \beta : \cat D, r: R, \gamma : \cat E,\Psi \vdash N : S}
+  %% {\Phi, \alpha : \cat C, x : \exists \beta. Q \odot R, \gamma : \cat E,\Psi \vdash J_\odot(x;q.\beta.r. N) : S}
+
+  %% \inferrule*[right=Tensor elim Subst]
+  %% {\Phi,\alpha : \cat C, q : Q, \beta : \cat D, r: R, \gamma : \cat E,\Phi' \vdash N : S
+  %% \and
+  %% \phi : \Psi \to \Phi,\alpha : \cat C, q : Q, \beta : \cat D, r: R, \gamma : \cat E,\Phi'}
+  %% {\Psi  \vdash J_\odot(x;q.\beta.r. N)[\phi] : S}
+
+  \inferrule*[right=Tensor $\beta$]
+  {}
+  {\appunary {\tensorelimfromWXtoYwithkontZ P Q R N} a {\tensorintroatXwithYandZ b s t} c = \appbinary N a s b t c}
+  %% {\homrappXtoYatZ {\pendappXtoY{\tensorelimfromWXtoYwithkontZ P Q R N} A} {(\homrappXtoYatZ{\homrappXtoYatZ{\pendappXtoY {\tensorintroofXandY P Q} A} f B} g G)} G =
+  %%   \homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY N A} f B} g G}
+
+  %% \inferrule*[right=Tensor $\beta$]
+  %% {}{J_\odot(\alpha.\gamma.S, \alpha.q.\beta.r.\gamma.N,A,(M,B,N),G) = N[A/\alpha,M/q,B/\beta,N/r,G/\gamma]}
+
+  \inferrule*[right=Tensor$\eta$]
+  {\Gamma \vdash N : \homunary \alpha {(\tensorexistsXwithYandZ \beta {\harrapp P \alpha \beta}{\harrapp Q \beta \gamma})} \gamma {\harrapp R \alpha \gamma}}
+  {N = \tensorelimfromWXtoYwithkontZ P Q R {\lambdabinary \alpha x \beta y \gamma \appunary N \alpha {\tensorintroatXwithYandZ \beta x y} \gamma}}
+
+  %% \inferrule*[right=Tensor$\eta$]
+  %% {\alpha : \cat C, x : \tensorexistsXwithYandZ \beta P Q, \gamma : \cat E \vdash M : R}
+  %% {\alpha : \cat C, x : \exists \beta. Q \odot R, \gamma : \cat E \vdash M =
+  %%   \homrappXtoYatZ {\pendappXtoY{\tensorelimfromWXtoYwithkontZ P Q R
+  %%       {\pendlambdaXdotY \alpha {\homrlambdaXatYdotZ p \beta {\homrlambdaXatYdotZ q \gamma {M[\homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {\tensorintroofXandY P Q} \alpha} p \beta} q \gamma/x]}}}}
+  %%     } \alpha} x \gamma : S}
+  \end{mathpar}  
+  \caption{Unit, Tensor of Profunctors}
+\end{figure}
+
+%% Next, we can internalize the $,\alpha,$ as a positive connective as a
+%% kind of ``existentially quantified tensor product'', abstracting the
+%% composition of profunctors.
+
+%% The ordinary formula for this is that for profunctors $R : C^o \times
+%% D \to \Set$ and $S : D^o \times E \to \Set$, the composite is given by
+%% a coend:
+
+%% \[ \int^{d \in D} R(-,d) \otimes S(d,=) \]
+
+%% We will instead use slightly more logical notation, writing the coend as
+%% an existential quantifier, justified by the analogy between the two
+%% and the structure of the introduction/elimination forms in our syntax
+%% in particular.
+%% \begin{figure}
+%% \begin{mathpar}
+%% \end{mathpar}
+%% \caption{Tensor Primitive Rules}
+%% \label{fig:tensor}
+%% \end{figure}
+
+Semantically, the tensor product is the composition of profunctors,
+which is typically written as a coend
+\[ \int_{\beta} P(-,\beta) \times Q(\beta,=) \]
+
+\subsection{Graph of a profunctor}
+
+One final multiplicative we can define is the \emph{graph} of a
+profunctor. Given a profunctor $\alpha:\cat C;\beta:\cat D \vdash R$,
+the ``graph'' of $R$ is a category $\Sigma_{\alpha:\cat C;\beta:\cat
+  D}. R$. In detail, its objects are triples $(a:\cat C, r : R, b :
+\cat D)$ and morphism from $(a,r,b)$ to $(a', r', b')$ consists of a
+pair of morphisms $g : b \to b'$ and a backwards $f : a' \to a$ such
+that $r' = g\circ r \circ f$.
+%
+Intuitively this generalizes the ``graph'' of a relation $R : A\times
+B \to \Omega$, where we think of the graph as the corresponding subset
+$gr(R) \subseteq A\times B$
+%
+By picking $R = \punitinXfromYtoZ {\cat C} {\alpha}{\alpha'}$ we
+recover the \emph{arrow} category $\exists
+(\alpha;\alpha'). \punitinXfromYtoZ {\cat C}{\alpha}{\alpha'}$, by
+composing these with functors, we recover the notion of \emph{comma}
+category $\Sigma \alpha;\beta. \punitinXfromYtoZ {\cat
+  E}{(F\alpha)}{(G\beta)}$ and by picking either $F$ or $G$ to have
+domain the terminal category $1$, we recover the notion of a
+\emph{slice} category.
+
+This is unique among the multiplicatives in that it defines a
+\emph{category} rather than a profunctor. The graph of a profunctor is
+a \emph{negative} multiplicative, a kind of $\Sigma$-type for
+profunctors. The universal property says that a functor from $\cat E$
+into the graph $\Sigma (\alpha; \beta). R$ is equivalent to a
+transformation from $\cat E$ into the functor $R$, including functors
+from $\cat E$ to $\cat C$ and $\cat D$.
+%
+The graph of a profunctor is a joint generalization of several common
+constructions in category theory: the arrow category, slice
+categories, comma categories and the covariant and the covariant and
+contravariant category of elements.
+
+\begin{figure}
+  \begin{mathpar}
+    \inferrule*[right=Graph-Formation]
+    {\Gamma \vdash \alpha:\cat C; \beta:\cat D \vdash R \isaSet}
+    {\Gamma \vdash \exists (\alpha;\beta). R \isaCat}
+
+    \inferrule*[right=Graph-E+]
+    {\Gamma \vdash \gamma:\cat E \vdash \phi : \exists(\alpha:\cat C;\beta: \cat D).R}
+    {\Gamma \vdash \pi_+\phi : \cat C}
+
+    \inferrule*[right=Graph-E-]
+    {\Gamma \vdash \gamma:\cat E \vdash \phi : \exists(\alpha:\cat C;\beta: \cat D).R}
+    {\Gamma \vdash \pi_-\phi : \cat D}
+
+    \inferrule*[right=Graph-Ep]
+    {\Gamma \vdash \gamma:\cat E \vdash \phi : \exists(\alpha:\cat C;\beta: \cat D).R}
+    {\Gamma \vdash \gamma:\cat E \vdash \pi_p\phi : R[\pi_+\phi/\alpha,\pi_-\phi/\beta]}
+
+    \inferrule*[right=Graph-Intro]
+    {\Gamma\vdash \gamma:\cat E \vdash a : \cat C \and
+      \Gamma\vdash \gamma:\cat E \vdash b : \cat D \and
+      \Gamma\vdash \gamma:\cat E \vdash t : R[a/\alpha;b/\beta]}
+    {\Gamma\vdash \gamma:\cat E \vdash (a,b,t) : \exists (\alpha;\beta). R}
+
+    \inferrule*[right=Unit-Formation]
+    {\Gamma \vdash \cat C \isaCat\and
+     \Gamma \pipe d^-\Xi \vdash a_1 : \cat C\and \Gamma \pipe d^+ \Xi \vdash a_2 : \cat C}
+    {\Gamma \pipe \Xi \vdash \punitinXfromYtoZ {\cat C} {a_1} {a_2}}
+
+    \pi_+(a,b,t) = a \\
+    \pi_-(a,b,t) = b \\
+    \pi_p(a,b,t) = t\\
+
+    \inferrule
+    {\Gamma \vdash \gamma:\cat E \vdash \phi : \exists(\alpha:\cat C;\beta:\cat D). R}
+    {\phi = (\pi_+\phi, \pi_-\phi, \pi_p\phi)}
+  \end{mathpar}  
+  \caption{Graph of a Profunctor}
+\end{figure}
+
+\subsection{Positive and Negative Presheaf Categories}
+
+Though defining arbitrary functor categories type-theoretically would
+require a ``multi-argument functor'' judgment, we can define
+\emph{presheaf} categories by the universal property that a functor
+$\cat C \to \mathcal P {\cat D}$ is the same as a profunctor from
+$\cat D$ to $\cat C$. To distinguish the covariant from contravariant
+presheaves, we define them dually as $\negPresheaf {\cat C}$ and
+$\posPresheaf {\cat D}$
+
+\begin{figure}
+  \begin{mathpar}
+    \inferrule*[right=NegPresheaf-Formation]
+    {\Gamma \vdash \cat C \isaCat}
+    {\Gamma \vdash \negPresheaf {\cat C} \isaCat}
+
+    \inferrule*[right=NegPresheaf-E]
+    {\Gamma \pipe d^-\Xi \vdash a : \cat C \and
+     \Gamma \pipe d^+\Xi \vdash p : \negPresheaf {\cat C}
+    }
+    {\Gamma \pipe \Xi \vdash \negPresheafApp a p \isaSet}
+
+    \inferrule*[right=NegPresheaf-I]
+    {\Gamma \pipe \alpha:\cat C;\beta:\cat D \vdash R : \isaSet}
+    {\Gamma \pipe \beta:\cat D \vdash \lambda \alpha:\cat C. R : \negPresheaf {\cat C}}
+
+    \negPresheafApp a (\lambda \alpha. R) = R[a/\alpha]\\
+
+    \inferrule*[right=NegPresheaf-$\eta$]
+    {\Gamma \pipe \beta:\cat D \vdash p : \negPresheaf {\cat C}}
+    {p = \lambda \alpha. \negPresheafApp \alpha p}
+
+    \inferrule*[right=PosPresheaf-Formation]
+    {\Gamma \vdash \cat D \isaCat}
+    {\Gamma \vdash \posPresheaf {\cat D} \isaCat}
+
+    \inferrule*[right=PosPresheaf-E]
+    {\Gamma \pipe d^-\Xi \vdash p : \posPresheaf {\cat D}\and
+     \Gamma \pipe d^+\Xi \vdash a : \cat D
+    }
+    {\Gamma \pipe \Xi \vdash \posPresheafApp p a \isaSet}
+
+    \inferrule*[right=PosPresheaf-I]
+    {\Gamma \pipe \alpha:\cat C;\beta:\cat D \vdash R : \isaSet}
+    {\Gamma \pipe \alpha:\cat C \vdash \lambda \beta:\cat D. R : \posPresheaf {\cat D}}
+
+    \posPresheafApp {(\lambda \beta. R)} b = R[b/\beta]\\
+
+    \inferrule*[right=PosPresheaf-$\eta$]
+    {\Gamma \pipe \alpha:\cat C \vdash p : \posPresheaf {\cat D}}
+    {p = \lambda \beta. \posPresheafApp p \beta}
+  \end{mathpar}
+  \caption{Presheaf Categories}
+\end{figure}
+
+\subsection{Dependent Types}
+
+Next, we cover the basic dependent type theory structure: $\Pi,\Sigma$
+and identity types.
+
+\begin{figure}
+  \begin{mathpar}
+  \inferrule
+  {\Gamma \pipe\cdot \vdash S \isaSet\and \Gamma \pipe\cdot \vdash T \isaSet}
+  {\Gamma\pipe\cdot \vdash \prod_{x:S} T \isaSet}
+
+  \inferrule
+  {\Gamma \pipe \cdot \vdash M : \prod_{x:S} T\and
+    \Gamma \pipe \cdot \vdash N : S}
+  {\Gamma \pipe \cdot \vdash M N : T[N/x]}
+
+  \inferrule
+  {\Gamma,x:S \pipe \cdot \vdash M : T}
+  {\Gamma \pipe \cdot \vdash \lambda x:S. M : \prod_{x:S} T}
+
+  \inferrule
+  {\Gamma \pipe\cdot \vdash S \isaSet\and \Gamma \pipe\cdot \vdash T \isaSet}
+  {\Gamma\pipe\cdot \vdash \sum_{x:S} T \isaSet}
+
+  \inferrule
+  {~}
+  {\Gamma\pipe\cdot \vdash \top \isaSet}
+
+  \inferrule
+  {\Gamma \pipe \cdot \vdash M : S\and \Gamma \pipe \cdot \vdash N : S}
+  {\Gamma \pipe \cdot \vdash \textrm{Id}_S M N \isaSet}
+
+  \inferrule
+  {\Gamma \pipe \cdot \vdash M : S}
+  {\Gamma \pipe \cdot \vdash \textrm{refl}(M) : \textrm{Id}_S M M}
+
+  \inferrule
+  {\Gamma,x:S,y:S,p:\textrm{Id}_S x y \pipe \cdot \vdash C \isaSet\and
+   \Gamma \pipe \cdot \vdash M_k : \prod_{x:S} C[x/y,\textrm{refl}(x)/p]}
+  {\Gamma \pipe \cdot \vdash \textrm{J}(x,y,p.C,M) : \prod_{x:S,y:S,p:\textrm{Id}_S x y} C}
+  \end{mathpar}
+  \caption{Ordinary Dependent Type Structure}
+\end{figure}
+
+\begin{figure}
+  \begin{mathpar}
+    \inferrule
+    {}
+    {\Gamma\vdash \smallCats}
+
+    \inferrule
+    {}
+    {\Gamma\vdash \Cats}
+
+    \inferrule
+    {\Gamma \vdash \cat C : \Sort \and\Gamma \vdash \cat D : \Sort}
+    {\Gamma\pipe\cdot \vdash \varr{\cat C}{\cat D} \isaSet}
+
+    \inferrule
+    {\Gamma \vdash \cat C : \Sort \and\Gamma \vdash \cat D : \Sort}
+    {\Gamma\pipe\cdot \vdash \harr{\cat C}{\cat D} \isaSet}
+
+    \inferrule
+    {\Gamma \pipe \alpha:\cat C \vdash A : \cat D}
+    {\Gamma \vdash \lambda \alpha:\cat C. A : \varr{\cat C}{\cat D}}
+
+    \inferrule
+    {\Gamma \pipe \alpha:\cat C \vdash A : \cat D\and
+      \Gamma\pipe\cdot \vdash M : \varr{\cat D}{\cat E}}
+    {\Gamma\pipe \alpha:\cat C \vdash M A : \cat E}
+
+    \inferrule
+    {\Gamma \pipe \alpha:\cat C;\beta:\cat D \vdash R \,\isaSet}
+    {\Gamma \vdash \lambda \alpha:\cat C;\beta:\cat D. R : \harr{\cat C}{\cat D}}
+
+    \inferrule
+    {\Gamma \vdash M : \harr{\cat C}{\cat D}\and
+     \Gamma \pipe d^- \Xi \vdash A : \cat C\and
+     \Gamma \pipe d^+ \Xi \vdash B : \cat C}
+    {\Gamma \pipe \Xi \vdash M A; B \isaSet}
+
+    \inferrule
+    {\Gamma \pipe \alpha:\cat C\vdash P \isaSet}
+    {\Gamma \vdash \pendallXdotY {\alpha:{\cat C}} P \isaTy}
+
+    \inferrule
+    {\Gamma \pipe \alpha:\cat C\vdash t : P}
+    {\Gamma \vdash \pendlambdaXdotY \alpha t : \pendallXdotY \alpha P}
+
+    \inferrule
+    {\Gamma \vdash M : \pendallXdotY \alpha P\and
+     \Gamma \pipe \beta:\cat D \vdash a : \cat C}
+    {\Gamma \pipe \beta:\cat D \vdash \pendappXtoY M a : P[a/\alpha]}
+  \end{mathpar}
+  \caption{Type Connectives}
+\end{figure}
+\begin{figure}
+  \begin{mathpar}
+    \inferrule
+    {\Gamma \pipe \alpha:\cat C\vdash P \isaSet}
+    {\Gamma \vdash \pendallXdotY {\alpha:{\cat C}} P \isaTy}
+
+    \inferrule
+    {\Gamma \pipe \alpha:\cat C\vdash t : P}
+    {\Gamma \vdash \pendlambdaXdotY \alpha t : \pendallXdotY \alpha P}
+
+    \inferrule
+    {\Gamma \vdash M : \pendallXdotY \alpha P\and
+     \Gamma \pipe \beta:\cat D \vdash a : \cat C}
+    {\Gamma \pipe \beta:\cat D \vdash \pendappXtoY M a : P[a/\alpha]}
+
+    \pendappXtoY {(\pendlambdaXdotY \alpha t)} a = t[a/\alpha]
+
+    \inferrule*[right=2Cell$\eta$]
+    {\Gamma \vdash M : \pendallXdotY \alpha P }
+    {\Gamma \vdash M = \pendlambdaXdotY \alpha {\pendappXtoY M \alpha} : \pendallXdotY \alpha P}
+  \end{mathpar}
+  \caption{End}
+  \label{fig:end}
+\end{figure}
 \section{Semantics}
@@
 \end{definition}
-
-\section{Discussion}
-
-\paragraph{Calculi for formal category theory}
-
-In their book on $\infty$-category theory, Riehl and Verity use the
-formal language of virtual equipments to prove results valid for
-$\infty$-categories without concrete manipulation of model
-categories. They formalize this language using Makkai's FOLDS
-framework of first-order logic with dependent sorts. The approach is
-similar in spirit to our \ohol{}, and has the same models. The main
-difference is that their theory is built on top of first-order logic,
-whereas ours is built on a custom-made variant of higher-order logic
-over a (unary) simple type theory. For this reason, statements in
-their theory are quite complex as they have to be encoded into
-first-order logic, whereas we argue our type-theoretic approach
-formalizes informal arguments much more directly, as shown in
-Section~\ref{sec:examples}.
-%
-An informal syntax for manipulating profunctors involving ends and
-coends, called the coend calculus is closely related, an extension of
-\ohol{} to treat profunctors of many variables of different variances
-should provide a formal treatment for the coend calculus
-\citet{coend-calculus}.
-
-Another approach to formal category theory is to develop a
-\emph{directed dependent type theory}, a type theory where the
-identity type is interpreted as morphisms in a (possibly
-$\infty$-)category. Research on directed dependent type theories is
-ongoing \cite{riehl-shulman,north}. Directed dependent type theories
-should generalize \ohol{} in the same way that dependent type theory
-generalizes higher order logic: internally there should be a type that
-models the category of sets, and terms of this type should model the
-profunctors of \ohol{}.  As a foundation for a domain-specific proof
-assistant for category theory, \ohol{} may provide a foundation for a
-proof assistant in the style of Isabelle/HOL \cite{isabelle-hol},
-whereas directed type theory would provide a foundation for a proof
-assistant in the style of Agda or Coq \cite{agda,coq}.
-
-\citet{djm-string-diagrams} provides a string diagram calculus for
-double categories and pro-arrow equipments, generalizing string
-diagrams for monoidal categories \citet{joyal?}. These are an
-alternative approach to type theoretic calculi, with the string
-diagrams typically making tensor products simpler to work with, while
-a type theoretic calculus like SCT makes the closed structure
-$\homrallXYtoZ \alpha P Q$ simpler to work with by using bound
-variables.
-
-One alternative to proarrow equipments is to axiomatize the bicategory
-of categories and profunctors rather than the double category of
-categories functors and profunctors. One abstract structure along
-these lines is a \emph{Cartesian bicategory}\footnote{which is not the
-same as a bicategory with products/finite limits, for which a syntax
-was developed \citet{fioresaville}} is given in
-\citet{carboni-walters,carboni-walters-someonelese}. Some preliminary
-work on a proof system for Cartesian bicateogires was presented at
-\citet{jonasfreycartesianbicategories}. This bears strong similarities
-to our calculus SCT, except that (1) their profunctors may have 0, 1
-or more covariant or contravariant variables but (2) they do not have
-a term syntax like our natural transformation syntax and (3) they do
-not have functors and restriction of profunctors by functors (since
-this structure is not present in a Cartesian
-bicategory). Semantically, Cartesian bicategories are somewhat less
-general than pro-arrow equipments because in general the set of
-functors between two categories can only be recovered from the set of
-profunctors if the codomain category is Cauchy complete, in which case
-functors are equivalent to adjoint pairs of profunctors.
-
-Finally, some variations on double categories have been used to model
-the structure of certain program logics. \citet{newlicata20} present a
-logic for \emph{vertically thin} pro-arrow equipments, i.e., pro-arrow
-equipments where the functor and natural transformation judgments are
-posetal. This calculus is quite different from ours in that the
-horizontal morphisms are terms of a programming language, and so they
-are not presented as having co-variant and contravariant variables,
-and the equipment structure is modeled by explicit coercions rather
-than by syntactic substitution of vertical arrows in horizontal
-arrows. Another similar calculus is System P \citep{systemP} which is
-an internal language of \emph{reflexive graph categories}, which are
-like double categories without horizontal composition.
-
-\subsection{Category Theoretic Extensions}
-
-While we have demonstrated that many common constructions in category
-theory can be expressed using the simple combination of homs, tensors
-and units, there are several that are out of reach for SCT due to its
-restrictive judgmental structure.
-
-First, while we can add products to the calculus, adding in a
-\emph{functor category} ${\cat C}^{\cat D}$ would necessitate adding a
-syntax for functors of multiple arguments, to allow for the lambda rule:
-\[ \inferrule {\alpha:\cat C, \beta:\cat D \vdash e : \cat E} {\alpha:\cat C \vdash \lambda \beta:\cat D. e : \cat E} \]
-This would then require also some extension to profunctors for
-allowing 0 or more covariant/contravariant dependencies as well.
-%
-In addition, we could consider extending our semantics with a notion
-of \emph{opposite category}, with the property that a covariant
-dependence of a profunctor on a variable in a category should be
-equivalent to a contravariant dependence on a variable in the opposite
-category.
-%
-This could possibly be done by introducing a modal distinction
-$\alpha:^{+}\cat C$ for covariant dependence and $\alpha:^{-}\cat C$
-for contravariant dependence, with the opposite category being a
-modality that internalizes flipping the variance as a type
-constructor.
-%
-With these two extensions, possibly, the category contexts for
-functors and profunctors could be unified as one notion of a sequence
-of object variables with variances.
-%
-With this unification, it would then be simpler to incorporate a third
-extension: a universe object, corresponding to an object
-$\textrm{Set}$, which would allow us to internalize the profunctor
-judgment in the functor judgment, modeling that in classical category
-theory, profunctors are the same as certain functors to the category
-of Sets.
-%
-While all of these extensions seem plausible, they all move away to
-the simple ordered type theory structure that allowed us to define
-units, tensor and hom profunctors so easily. It is unclear how to
-define these types in this extended structure. As a simple example, if
-profunctors have many variables of mixed variance, rather than exactly
-one of each variance, it is unclear what the appropriate
-generalization of the transformation context $\Phi$ should be.
-
-An orthogonal extension to these would be to consider adding a
-\emph{discretization} modality $\Delta\cat C$ that gives the discrete
-category on the underlying set of objects of $\cat C$. This would give
-us the ability to formalize concepts like ``function on objects'' or
-``unnatural transformation'' without the need to define internal
-categories. This could be modeled by adding a modal distinction
-$\alpha:^{\#}\cat C$ for ``crisp'' dependence on an object of $\cat
-C$, i.e., dependence merely on the object of $\cat C$ with no
-functoriality/naturality conditions. This would then be internalized
-by the discretization modality.
-
-A third approach would be to extend the calculus to model general
-monoidal closed structures and more generally multicategories, which
-provides semantics of modal substructural type theories.
-%
-This can be modeled by extending the ``universe'' $\Set$ to be a
-monoidal closed category, which induces a monoidal closed structure on
-profunctors by the Day convolution product \cite{...}.
-%
-Similar developments along these lines have already been done for
-coend calculus \citep{coend-calculus} and in formalization of
-substructural type theories \citep{atkey-woods}.
-
-%% A recent development in the formalization of substructural type
-%% theories is to use monoidal connectives at the meta-level when
-%% defining multiplicative connectives\cite{woodsatkeyesop2022}. This
-%% suggest one approach to extending the theory to support the
-%% formalization of multiplicative universal properties would be, in
-%% adherence to the microcosm principle, to add linear judgments and
-%% multiplicative connectives to the metalevel.
-
-%% \paragraph{As a tool for Formal Type Theory}
-
-%% While we can model ``unary'' type theories using profunctors, to model
-%% theories like simple intuitionistic, linear or bunched type theories
-%% we need a way to talk about ``profunctors of many arguments''. This
-%% can be accomplished using \emph{multicategory} theory, which models
-%% the structure of the context with its structural rules as a monad of
-%% ``contexts and substitutions''. A suitable notion of monad on a
-%% virtual double category has been developed and was in fact the
-%% original motivation for the concept of a virtual double
-%% category. However while an abstract monad of contexts $T$ is enough to
-%% define what a $T$-multicategory internal to a virtual equipment is, it
-%% does not appear possible to define the multiplicative connectives as a
-%% universal property, i.e., to define when the multicategory is
-%% \emph{representable}.
-
-\subsection{Semantics}
-
-Our definition of a double category with displays is suggestive of a
-simple generalization. A double category is a monoid object in the
-double category of $Span(Cat)$, and a double category with displays
-furthermore uses certain diagrams involving profunctors, which also
-form a double category $Prof$. This suggests that the algebraic
-structure of a double category with displays is naturally modeled in
-some kind of \emph{triple category}, modeling the three types of
-morphisms: functors, profunctors and spans of functors.
-
-%% %%
-%% %% Bibliography
-%% %%
-
-%% %% Please use bibtex, 
-\subsubsection*{Acknowledgements}
-Grants stuff...? The authors would like to thank David Jaz Myers, Emily Riehl, Mike
-Shulman, Dominic Verity,...for helpful feedback on this work
-
-\bibliography{cats}
-
-\appendix
-
-\section{Syntactic Details}
-
-\begin{figure}
-  \begin{mathpar}
-    \begin{array}{rccl}
-      \textrm{Term Substitution} & \gamma, \delta & ::= & \cdot \pipe \gamma, M/X\\
-      \textrm{Object Substitution} & \xi, \zeta & ::= & a/\alpha \pipe a/\alpha;b/\beta\\
-      \textrm{Transformation Substitution} & \phi, \psi & ::= & a/\alpha \pipe \phi,s/x, \psi      
-    \end{array}
-
-      \inferrule*[right=TyCtxForm]
-        {}
-        {\Gamma \isadtctx}
-
-    \inferrule*[right=SetCtxForm]
-    {\Gamma \isadtctx}
-    {\Gamma \vdash \Xi \boundary}
-
-    \inferrule*[right=TransCtxForm]
-    {\Gamma \isadtctx}
-     {\Gamma \vdash \Phi \isavectx}
-    \inferrule
-    {\Delta \isadtctx \and \Gamma \isadtctx}
-    {\gamma : \Delta \to \Gamma}
-
-    \inferrule*[right=Boundary Map formation]
-    {\Gamma \vdash Z \boundary \and \Gamma \vdash \Xi \boundary}
-    {\Gamma \vdash \xi : Z \to \Xi}
-
-    \inferrule*[right=Cat Subst formation]
-    {\Gamma \vdash \Psi \isavectx \and \Gamma \vdash \Phi \isavectx}
-    {\Gamma \vdash \phi : \Psi \to \Phi}
-
-  \end{mathpar}
-  \caption{Substitution}
-
-\end{figure}
-
-\cref{fig:contexts} details the typing rules for contexts and the
-typing for variables and the admissible actions of substitution (the
-definition of substitution is straightforward and found in the
-appendix). the rules for term and object variables are as usual.
-%
-An object substitution into a single object context $\alpha:\cat C$
-consists of an object $a : \cat C$ parameterized by an object of
-another category.
-%
-An object substitution into a pair of co- and contra-variant variables
-$\alpha:\cat C;\beta:\cat D$ is a pair of an object substituted for
-$\alpha$ and one for $\beta$, each typed with respect to the
-appropriate variable of an object context $\Xi$.
-%
-Note that this allows for $\Xi$ to consist of either two variables or
-a single variable.
-%
-When
-constructing a transformation context $\Phi, x : R, \Psi$, we need to
-ensure that the contravariant dependence of $R$ matches the covariant
-dependence of $\Phi$ and vice-versa for $\Psi$. We define this with
-auxiliary metafunctions $d^+,d^-$ which extract the covariant and
-contravariant object variable of a transformation context.
-%
-The rules for transformation variables and substitutions define the
-use of variables to be \emph{linear} (each variables is used exactly
-once) and \emph{ordered}, they occur in the term in the same order
-that they appear in the context.
-%
-The rule for a transformation variables says a variable $x$ can only
-be used when it is the sole variable in the context.
-%
-A transformation subsitution into an ``empty'' transformation context
-$\alpha: \cat C$ is the same as an object substitution into
-$\alpha:\cat C$.
-%
-The rule for constructing a transformation substitution into a
-non-empty context $\Phi_1,x:R,\Phi_2$ says we can construct a
-substitution by constructing substitutions $\phi_1,\phi_2$ for
-$\Phi_1$ and $\Phi_2$ as well as a transformation $s$ into an
-instantiation of $R$ where the input contexts come from splitting the
-input context $\Psi$ into three components, and ensuring that the
-instantiation of $R$ agrees with how $\phi_1,\phi_2$ instantiate the
-variables. This uses two new metafunctions. First, $\Phi \jnctx \Psi$
-is an operation for concatenating two contexts when $d^+(\Phi) =
-d^-(\Psi)$. Next, we extend the boundary maps $d^\pm$ to substitutions
-to extract their object substitutions.
-
-\begin{figure}
-  \begin{mathpar}
-    \inferrule{}{\cdot \isadtctx}
-
-    \inferrule
-    {\Gamma \isadtctx \and \Gamma \vdash A \isaTy}
-    {\Gamma , X : A \isadtctx}\\
-
-    \inferrule
-    {X : A \in \Gamma}
-    {\Gamma \vdash X : A}
-
-    \inferrule
-    {}
-    {\alpha: \cat C \vdash \alpha: \cat C}
-
-    \inferrule
-    {\Gamma \vdash \cat C : \isaCat}
-    {\Gamma \vdash \alpha: \cat C \isavectx}
-
-    \inferrule
-    {\Gamma \vdash \Phi \isavectx\and
-    \Gamma \pipe d^+\Phi; \beta:\cat D \vdash R \isaSet}
-    {\Gamma \vdash \Phi, x:R, \beta:\cat D \isavectx}
-
-    \inferrule
-    {}
-    {\Gamma \pipe \alpha:\cat C, x : R, \beta: \cat D \vdash x : R}
-
-    % horizontal identity
-    \inferrule
-    {\Gamma\pipe \beta:\cat C \vdash a : \cat C}
-    {\Gamma \pipe \beta:\cat C \vdash a/\alpha : (\alpha : \cat C)}
-
-    \inferrule
-    {\Psi = \Psi_1 \jnctx \Psi_2\and
-      \Gamma\pipe\Psi_1 \vdash \phi : \Phi\and
-      \Gamma \pipe \Psi_2 \vdash t : R[d^+\phi;b/\beta] \and
-      \Gamma \vdash \phi_2 : \Psi_2 \to \Phi_2
-      }
-    {\Gamma\pipe\Psi \vdash \phi,t/x,b/\beta :: (\Phi,x:R,\beta:\cat D)}
-  \end{mathpar}
-  \begin{align*}
-    \underline{\alpha:C} &= \alpha:C\\
-    \underline{\Phi,x:R,\Psi} &= d^-\Phi; d^+{\Psi}
-  \end{align*}
-  \begin{align*}
-    d^{\pm}(\alpha:\cat C) &= \alpha : \cat C\\
-    d^{-}(\Phi,x:R,\Psi) &= d^-\Phi\\
-    d^{+}(\Phi,x:R,\Psi) &= d^-\Psi
-  \end{align*}
-  \begin{align*}
-    d^{\pm}(a/\alpha) &= a/\alpha\\
-    d^{-}(\phi,t/x,\psi) &= d^-\phi\\
-    d^{+}(\phi,t/x,\psi) &= d^+\psi\\
-  \end{align*}
-  \caption{Basic Rules of Judgment Theory}
-  \label{fig:contexts}
-\end{figure}
-
-Now that we have established a judgmental structure, in typical type
-theoretic fashion, our type constructs follow naturally as
-internalizing the judgments. Though the positive multiplicatives are
-more familiar, we start by describing the negative multiplicatives,
-since they are better behaved, and are useful to have already defined
-when discussing the positives.
 

### 2022-10-04 paper/paper.tex
prose for the profunctor connectives

@@
 \newcommand{\tensorelimWfromXtoYkontZ}[4]{\textrm{ind}_{\odot}(#4;#1,#2,#3)}
+\newcommand{\tensorelimWkontZ}[2]{\textrm{ind}_{\odot}(#2;#1)}
 
@@
 profunctors in Figure~\ref{fig:set-connectives}.
+%
+The top of the figure includes the formation rules for tensor products
+$\tensorexistsXwithYandZ \beta P Q$, the covariant $\homrallXYtoZ
+\alpha R P$ and contravariant homs $\homlallXYtoZ {\alpha} R P$ and
+the unit type $\punitinXfromYtoZ {\cat C} {a_1}{a_2}$.
+%
+First, the unit set is our syntax for the profunctor of morphisms of a
+category $\cat C$.
+%
+Its introduction and elimination rules are like those of an
+\emph{ordered} variant of equality: the introduction rule is the
+identity morphism (reflexivity) and the elimination rule is an
+induction principle that is a variant of the J rule of equality
+elimination: we can use a term of $s : \punitinXfromYtoZ {\cat C} a b$
+by specifying the behavior when $s$ is of the form $\punitrefl \alpha$
+in the form of a continuation $\alpha. t$. Again we elide all
+$\beta\eta$ rules, which are standard. Next, the tensor product of
+sets is a kind of combined existential quantifier and monoidal
+product, which we combine into a single notation
+$\tensorexistsXwithYandZ \beta P Q$, where $\beta$ is the covariant
+variable of $P$ and the contravariant variable of $Q$. Then the
+covariant variable of the tensor product is the covariant variable of
+$Q$ and the contravariant variable similarly comes from $P$.
+%
+In ordinary category theory, this is the \emph{composition} of
+profunctors, and is defined by a coend of a product. We require that
+the variable $\beta$ quantifies over a small category $\cat D$, as in
+general this composite doesn't exist for large categories.
+%
+The introduction and elimination are like those for a combined tensor
+product and existential type: the introduction rule is a pair of
+terms, with an appropriate instantiation of $\beta$, and the
+elimination rule says to use a term of a tensor product, it is
+sufficient to specify the behavior on two elements typed with an
+arbitrary middle object $\beta$.
+%
+Note that in the unit elimination, the continuation is typed with a
+context consisting of only a single object $\alpha$, in contrast with
+normal rules for a unit type in natural deduction for ordered linear
+logic such as \cite{polakawpfenning}, which would split the context
+$\Phi$ into $\Phi_l \jnctx \Phi_m \jnctx \Phi_r$ where $\Phi_m$ is
+used by the scrutinee and $\Phi_l \jnctx \Phi_r$ is used by the
+continuation. However, unlike in ordered linear logic, the $\Phi_l
+\jnctx \Phi_r$ is not, necessarily well formed because there is no
+reason in general that $\Phi_m$ has contravariant and covariant
+dependence on the same category.
+%
+This is analogous to a similar situation in dependent type theory for
+identity types: TODO: Dan please help with this explanation!!
+
+Next, we introduce the contravariant ($\homlallXYtoZ \alpha R P$) and
+covariant ($\homrallXYtoZ \alpha R P$) homs of sets, which are
+different because we are in an orderd logic.
+%
+These are a kind of universally quantified
+function type, where the universally quantified variable must occur
+with the same variance in domain and codomain. In the contravariant
+case, it occurs as the contravariant variable in both, and vice-versa
+for the covariant case.
+%
+To highlight this, the notation for the contravariant dependence puts
+the quantified variable on the \emph{left} of the triangle, as
+contravariant variables occur to the left of the covariant variable,
+and similarly the covariant hom has the quantified variable on the
+right.
+%
+Then the
+covariant variable of the contravariant hom set is the covariant
+variable of the codomain and, and the contravariant variable of the
+hom set is the \emph{covariant} variable of the domain, as the two
+contravariances cancel. The covariant hom is dual.
+%
+Semantically, in ordinary category theory these are known as the
+\emph{hom} of profunctors and are constructed using a combination of
+an end and a function set.
+%
+The two connectives have similar introduction and elimination rules in
+the form of $\lambda$ terms abstracting over both the object of the
+category and the element of the set, and appropriate application
+forms.
+%
+To keep with our invariant that the variable occurrences occur left to
+right in the term syntax in a manner matching the context, we write
+the covariant application in the usual order $\homrappXtoYatZ s t a$
+where the function is on the left and the argument is on the right,
+and the contravariant application in the flipped order.
+
+Finally, we have the cartesian unit and product sets, which are
+analogous to the normal unit and product of types. The most notable
+point to emphasize is that in the formation rule for the product, the
+two subformulae should have the same covariant and contravariant
+dependence. So as with linear logic, the cartesian unit and product
+complicate the syntactic formulation of linearity: the unit
+constructor ``uses'' all variables and the product ``uses'' all
+variables in both cases.
 
@@
     \inferrule
-    {\Gamma \pipe d^-\Xi; \beta:D\vdash P \isaSet\\\\
+    {\cat D \isSmall\\\\
+      \Gamma \pipe d^-\Xi; \beta:D\vdash P \isaSet\\\\
      \Gamma \pipe \beta:D;d^+\Xi \vdash Q \isaSet}
@@
     \inferrule
-    {\Gamma \pipe \alpha:\cat C; d^-\Xi\vdash R \\\\ \Gamma \pipe \alpha : \cat C; d^+\Xi\vdash P}
-    {\Gamma \pipe \Xi \vdash \homlallXYtoZ \alpha R P}
-
-  \inferrule
-  {\Gamma \pipe d^+\Xi; \alpha : \cat C \vdash  R \\\\ \Gamma \pipe  d^-\Xi; \alpha : \cat C \vdash P}
-  {\Gamma \pipe \Xi \vdash \homrallXYtoZ \alpha R P}\\
+    {d^-\Xi \isSmall\\\\
+      \Gamma \pipe \alpha:\cat C; d^-\Xi\vdash R \\\\ \Gamma \pipe \alpha : \cat C; d^+\Xi\vdash P}
+    {\Gamma \pipe \Xi \vdash \homlallXYtoZ {\alpha: \cat C} R P}
+
+  \inferrule
+  {d^+\Xi \isSmall\\\\ \Gamma \pipe d^+\Xi; \alpha : \cat C \vdash  R \\\\ \Gamma \pipe  d^-\Xi; \alpha : \cat C \vdash P}
+  {\Gamma \pipe \Xi \vdash \homrallXYtoZ {\alpha:\cat C} R P}\\
 
@@
   {
-   \Gamma\pipe \alpha:\cat C, x:P, \beta:\cat D, y:Q, \gamma:\cat E \vdash t : R\\\\
-   \Gamma \pipe\Phi \vdash s : \tensorexistsXwithYandZ {\beta} {P[a/\alpha]}{Q[c/\gamma]}}
-  {\Gamma \vdash \tensorelimWfromXtoYkontZ a s b {\alpha.x.\beta.y.\gamma. t} : R[a/\alpha;c/\gamma]}
+   \Gamma\pipe \Phi_m \jnctx x:P, \beta:\cat D, y:Q \jnctx \Phi_r \vdash t : R\\\\
+   \Gamma \pipe\Phi_m \vdash s : \tensorexistsXwithYandZ {\beta} P Q}
+  {\Gamma\pipe\Phi_l\jnctx \Phi_m \jnctx \Phi_r \vdash \tensorelimWkontZ s {x.\beta.y. t} : R}
 
@@
 %
-First, we demonstrate that a hyperdoctrine of virtual double
+First, we demonstrate that a hyperdoctrine of virtual equipments
 categories with split restrictions provides a sound and complete

### 2022-10-06 paper/paper.tex
very briefly describe the models from Shulman's enriched indexed categories

@@
 \newcommand{\Set}{\textrm{Set}}
+\newcommand{\MonCat}{\textrm{MonCat}}
+\newcommand{\SymMonCat}{\textrm{SymMonCat}}
 \newcommand{\vtkmnd}{\mathbb{K}\text{Mod} (\vcat{V},T)}
@@
 and the contravariant application in the flipped order.
+%
+We also write the instantiating object as a superscript to
+de-emphasize it, as in practice it can often be inferred.
 
@@
 categories, functors, profunctors and natural transformations.
+%
+The rule for the types of small categories and categories are very
+similar: any definable category defines an element of type $\Cats$,
+and any element of that type can be reflected back into a
+category. The only difference for $\smallCats$ is that the categories
+involved additionally satisfy $\cat C \isSmall$. Again we elide the
+standard $\beta\eta$ principles.
+%
+Since every small category $\cat C \isSmall$ is a category $\cat C
+\isaCat$, there is a definable inclusion function from $\smallCats$ to
+$\Cat$ and the $\beta\eta$ properties ensure that this is a
+monomorphism.
+
+Next, we have the types of all functors and profunctors between any
+two fixed categories.
+%
+The introduction and elimination forms are those for unary and binary
+function types respectively.
+%
+Finally we include a type $\pendallXdotY {\alpha:\cat C} P$ which we
+call the set of ``natural elements'' of $P$.
+%
+The name comes from the case that $P$ is of the form
+$\punitinXfromYtoZ {}{F(\alpha)}{G(\alpha)}$ in which case the type
+$\pendallXdotY {\alpha:\cat C} \punitinXfromYtoZ
+{}{F(\alpha)}{G(\alpha)}$ can be interpreted as the set of all natural
+transformations from $F$ to $G$.
+%
+More generally this is modeled as an end, and matching our notation
+for hom types we use a universal quantifier to notate this.
 
@@
     {\Gamma \vdash \cat C \isaCat \and\Gamma \vdash \cat D \isaCat}
-    {\Gamma\pipe\cdot \vdash \varr{\cat C}{\cat D} \isaSet}
+    {\Gamma\pipe\cdot \vdash \varr{\cat C}{\cat D} \isaTy}
 
@@
     {\Gamma \vdash \cat C \isaCat \and\Gamma \vdash \cat D \isaCat}
-    {\Gamma\pipe\cdot \vdash \harr{\cat C}{\cat D} \isaSet}
+    {\Gamma\pipe\cdot \vdash \harr{\cat C}{\cat D} \isaTy}
 
@@
 
-\section{Synthetic Category Theory in Ordered Higher-order Logic}
+\section{Synthetic Category Theory in VETT}
 
@@
 
-\subsection{Categorical Models}
+\subsection{Notion of Model}
 
 First, we can model the judgmental structure of the unary type theory
-and predicate logic in certain\emph{virtual double categories}
-\cite{leinster,crutwell-shulman}. We briefly recount the structure
-present in a virtual double category, but see \citet{crutwell-shulman}
-for a precise definition.
-The following definition is
-equivalent to the formulation as a ``normalized $T$-monoid'' in
-\citet{crutwell-shulman}, and we direct the reader there for the full
-description.
-%
-
-\begin{definition}[Virtual Double Category]
-  A virtual double category $\mathcal V$ consists of
+and predicate logic in certain\emph{virtual equipments}
+\cite{leinster,crutwell-shulman}, i.e., virtual double categories with
+restrictions. We briefly recount the structure present in a virtual
+double category, but see \citet{crutwell-shulman} for a precise
+definition of the composition rules for 2-cells and functor of virtual
+double categories.
+%
+Our notion here differs from \citet{crutwell-shulman} in several ways
+to more closely match our judgmental structure: (1) we require that
+restrictions are given as coherent structure rather than mere
+existence (2) we include a notion of ``small'' object and (3) we don't
+yet require the existence of unit horizontal arrows, as these are
+modeled by a connective rather than judgmental structure.
+
+\begin{definition}[Virtual Equipment]
+  A virtual equipment $\mathcal V$ consists of
   \begin{enumerate}
   \item A category $V_o$ of ``objects and vertical arrows''
+  \item A chosen subset of the objects $V_s \subseteq V_o$ of ``small objects''
   \item A set ${\mathcal V}_h$ of ``horizontal arrows'' with source
@@
 	\arrow["{R_n}", "\shortmid"{marking}, from=1-2, to=1-3]
-	\arrow["F"', from=1-1, to=3-1]
-	\arrow["G", from=1-3, to=3-3]
+	\arrow["f"', from=1-1, to=3-1]
+	\arrow["g", from=1-3, to=3-3]
 	\arrow["S"', "\shortmid"{marking}, from=3-1, to=3-3]
@@
 \end{tikzcd}\]
+    With appropriate ``multi-categorical'' notions of identity and
+    composition.
+
+    We say that the 2-cell $\phi$ has $S$ as codomain, the sequence
+    $R_0 \ldots R_n$ as domain and call $f$ and $g$ the left and right
+    ``frames'', or that $\phi$ is framed by $f$ and $g$.
   \end{enumerate}
@@
   for any horizontal arrow $R : C \pto D$ and vertical arrows $f : C'
-  \to C$ and $g : D' \to D$ there is a chosen profunctor $R(f,g) : C'
-  \pto D'$ with a cartesian 2-cell to $R$ framed by $f,g$ and these
-  chosen cartesian lifts are functorial in $f,g$.  A functor of
-  virtual double categories preserves split restrictions if it
-  preserves these specified cartesian lifts.  
+  \to C$ and $g : D' \to D$ there is a chosen horizontal arrow $R(f,g)
+  : C' \pto D'$ with a cartesian 2-cell to $R$ framed by $f,g$ and
+  these chosen cartesian lifts are functorial in $f,g$.
+
+  A functor of virtual equipments is a functor of the virtual double
+  categories that additionally preserves the restrictions and
+  smallness of objects.
 \end{definition}
 
+In the presence of restrictions, every 2-cell can be represented as a
+``globular'' 2-cell where the left and right frame are identities. For
+example the 2-cell $\phi$ above can be represented as one with the
+same domain but whose codomain is $S(f,g)$. This property is crucial
+for the completeness of our semantics as we only include a syntax for these globular terms 
+%
 Each component of this definition has a direct correspondence to a
-syntactic structure in \ohol{}. The category $\mathcal V_o$ models the
-category and functor judgments. The set $\mathcal V_h$ models the
-profunctor judgment. A composable string $R_0 \cdots R_n$ models the
-profunctor contexts. The 2-cells correspond to the natural
-transformation judgment where we have taken the restriction $S(F,G)$
-of the codomain.
-
-To model the dependent type theory and fact that the \ohol{} judgments
-are all under a dependent type theory context, we use a variation on
-Lawvere's notion of \emph{hyperdoctrine} for modeling predicate logic:
-\begin{definition}[Hyperdoctrine of Virtual Double Categories with Restrictions]
-  A hyperdoctrine of virtual double categories with restrictions is a
-  pair of a category with families $\mathcal C$ and a functor $L :
-  \mathcal C^o \to \text{vDbl}_r$ to the category of virtual double
-  categories with split restrictions and restriction-preserving
-  functors.
+syntactic structure in \ohol{}. The objects of $\mathcal V_o$ models
+the category judgment and the morphisms model the functor
+judgment. The set $\mathcal V_h$ models the profunctor judgment. A
+composable string $R_0 \cdots R_n$ models the profunctor contexts. The
+2-cells correspond to the natural transformation judgment where we
+have taken the restriction $S(F,G)$ of the codomain.
+
+To model the dependent type theory and fact that the categorical
+judgments are all typed under a context $\Gamma$ with an action of
+substitution, we use a variation on Lawvere's notion of
+\emph{hyperdoctrine} for modeling predicate logic:
+\begin{definition}[Hyperdoctrine of Virtual Equipments]
+  A hyperdoctrine of virtual equipments is a pair of a category with
+  families $\mathcal C$ and a functor $V^{(-)} : \mathcal C^o \to
+  \text{vEq}$ to the category of virtual equipments and functors.
 \end{definition}
 Categories with families $\mathcal C$ model dependent type theory
-\citep{cwf} and for each semantic context $\Gamma$, $L(\Gamma)$ models
-the \ohol{} judgments in context $\Gamma$, with the functoriality
+\citep{cwf} and for each semantic context $\Gamma$, $V^\Gamma$ models
+the VETT judgments in context $\Gamma$, with the functoriality
 modeling the fact that all of these judgments admit a well-behaved
@@
 
-\subsection{Concrete Models}
+\subsection{Examples of Models}
 
@@
 model construction theorems that can be used to justify our intuitive
-notion of semantics in (enriched, internal) category theory.
+notion of semantics in (enriched, internal, indexed) category theory.
+
+First, we can extend any virtual equipment to a hyperdoctrine of
+virtual equipments
 
@@
 \begin{construction}
-  Given a virtual double category with restrictions $\mathcal V$, we
+  Given a virtual equipment $\mathcal V$, we
   can construct a functor $\mathcal V^{-}: \Set \to \text{vDbl}_r$ by
@@
 
-TODO: find a reference that proves things have all the UMPs we want
-lol.
+%% Semantic Qs:
+%
+%% (1) The virtual equipment of locally small categories is a model of everything.
+%% (2) Are Shulman's enriched indexed categories strict enough?
+
+Then to define a model of VETT with a collection of connectives is is
+sufficient to construct a virtual equipment with the universal
+properties corresponding to those connectives.
+%
+\begin{construction}
+  Fix a cardinal $\kappa$. The virtual equipment $\Cat_{\kappa}$ is
+  defined to have as objects locally $\kappa$-small categories, small
+  objects as $\kappa$-small categories, vertical morphisms as functors,
+  horizontal arrows as functors $\cat C^o \times \cat D \to \kappa Set$ and
+  2-cells as morphisms of profunctors. Restriction of profunctors is
+  given by composition, which is strictly associative and unital.
+  $\Cat_{U}$ has objects satisfying the universal properties of all
+  connectives in VETT.
+\end{construction}
+
+More generally, categories internal to, enriched in and/or indexed by
+sufficiently nice categories define a virtual equipment that model
+most of the connectives of our type theory. We highlight one example
+from the literature that is highly general: Shulman's enriched indexed
+categories. Shulman's construction defines a virtual double categories
+of large and small $\mathcal V$-categories for any pseudofunctor
+$\mathcal V : S^o \to \MonCat$ where $S$ is a category with finite
+products. He gives examples that show that this internal, enriched and
+indexed categories for suitable choices of $\mathcal V$.
+%
+This is slightly weaker then what we require: to have
+\emph{split} restrictions, we need that $\mathcal V$ be a
+\emph{strict} functor, not merely a pseudo-functor.
+%
+This is analogous to the situation for dependent type theory where
+substitution is strictly associative but in models is typically
+modeled by pullback which is only associative up to unique
+isomorphism. In fact, one of the main examples is only a
+pseudo-functor for precisely this reason that the action on morphisms
+is given by pullback.
+%
+We expect this difference can be handled in the same way: by a general
+strictification theorem or by moving to a univalent setting.
+
+\begin{construction}[(\citet{shulman-enriched-indexed})]
+  Given any functor $\mathcal V : S^o \to \SymMonCat$ such that $S$
+  and $\mathcal V$ have sufficiently well-behaved (indexed)
+  $\kappa$-products, then there is a virtual equipment $\mathcal
+  V-\textrm{Cat}$ whose objects are locally $\kappa$-small $\mathcal
+  V$-categories, small objects are $\kappa$-small $\mathcal
+  V$-categories etc. This virtual equipment has been shown in
+  loc. cit. to have objects satisfying all of the universal properties
+  used in models of VETT except the graph of a profunctor.
+\end{construction}
 

### 2022-10-07 paper/paper.tex
clarity edits to intro

@@
   category, functor, profunctor and a generalized form of natural
-  transformations. By restricting the type theory to an ordered linear
-  restriction of predicate logic, we guarantee that all functions
+  transformations. The type theory imposes an ordered linear
+  restriction on standard predicate logic, which guarantees that all functions
   between categories are functorial, all relations are profunctorial,
   and all transformations are natural by construction, with no
-  separate proof necessary. Important category theoretic proofs such
+  separate proofs necessary. Important category theoretic proofs such
   as the Yoneda lemma and Co-yoneda lemma become simple type theoretic
@@
 that formal systems like logics, type theories and programming
-languages can be modeled by sound and complete semantic models given
-by notions of structured category\cite{lawvere,goguen,lambek,moggi}, a
+languages can be soundly and completely modeled
+by notions of structured categories\cite{lawvere,goguen,lambek,moggi}, a
 phenomenon sometimes referred to as the Curry-Howard-Lambek
@@
 These generalizations have long been recognized as being important in
-programming language applications: domain-,metric- and
+programming language applications: domain-, metric- and
 step-index-enriched categories have been used to model general
@@
 naturality and functoriality side conditions that abound in otherwise
-simple proofs in category theory. For instance consider the proof of
-the Yoneda lemma, often the first ``big'' proof in an introductory
-category theory text. The Yoneda lemma states that for any functor $P
-: C^o \to \Set$ and object $A \in C$, that the set of natural
-transformations $C(-,A) \to P$ is in bijection with the elements
-$P(A)$. Furthermore, this bijection is natural in both $A$ and $P$. In
-the authors' experience teaching category theory the sheer amount of
-naturality side conditions in this result can be very intimidating to
-learners, while experienced practitioners will observe that these
-naturality conditions are tedious to carry out but routine. In our
-type theory, none of these naturality conditions need to be explicitly
-proven. Instead, all terms expressible in the type theory are
-automatically natural by construction.
+simple proofs in category theory. For instance consider the proof of the
+Yoneda lemma, often the first ``big'' proof in an introductory category
+theory text. The Yoneda lemma states that for any category $C$, functor
+$P : C^{o} \to \Set$ (where $C^{o}$ is the ``opposite'' of $C$, i.e. 
+$C$ with the direction of morphisms reversed) and object $A \in C$, the
+set of natural transformations $C(-,A) \to P$ is in bijection with the
+elements $P(A)$. Furthermore, this bijection is natural in both $A$ and
+$P$. In the authors' experience teaching category theory the sheer
+amount of naturality side conditions in this result can be very
+intimidating to learners, while experienced practitioners will observe
+that these naturality conditions are tedious to carry out but
+routine. In the type theory we introduce in this paper, none of these
+naturality conditions need to be explicitly proven. Instead, all terms
+expressible in the type theory are automatically natural by
+construction.
 
@@
 adept analyst writes down a function, they do not expand out the
-$\epsilon-\delta$ definition of continuity and prove that it is
+$\epsilon\mathord{-}\delta$ definition of continuity and prove that it is
 continuous. Rather, they are aware of certain \emph{syntactic
@@
 %
-In \vett{}, the sorts are interpreted as categories, functions as
-functors, relations as profunctors and implications of relations as
-profunctor homomorphisms, a generalization of natural transformations.
+In the ``standard model'' of \vett{}, the sorts are interpreted as
+categories, functions as functors, relations as profunctors and
+implications of relations as profunctor homomorphisms, a generalization
+of natural transformations.
 %
@@
 as if they were ordinary functors and profunctors in set-based
-category theory, but the proofs are valid in an arbitrary virtual
-equipments and so are valid in enriched and internal category theory
-as well.
-%
-This is analogous to the situation in intuitionistic type theory,
-where the syntax is similar to that of informal set theory, but by
-restricting to intuitionistic reasoning, the constructions can all be
-interpreted in models where types are spaces and terms are continuous
-functions ``for free''.
+category theory.
+%
+\vett also can be modeled in an arbitrary virtual equipments, and so the
+constructions are valid in enriched and internal category theory as
+well.
+%%
+%% DRL: I know what you mean, but I think conflating set theory and
+%% type theory will put some people off... 
+%% 
+%% This is analogous to the situation in intuitionistic type theory,
+%% where the syntax is similar to that of informal set theory, but by
+%% restricting to intuitionistic reasoning, the constructions can all be
+%% interpreted in models where types are spaces and terms are continuous
+%% functions ``for free''.
 %
@@
 
-\vett{} is a kind of \emph{ordered}, \emph{constructive} variant of
+\vett{} is an \emph{ordered}, \emph{constructive} variant of
 predicate logic over a unary type theory.
@@
 corresponds to a well-understood notion of model (virtual equipments)
-and so we leave the extension to future work.
+and so we leave these extensions to future work.
 %
@@
 
-Due to the restriction to unary functors and binary profunctors, it is
-difficult to state some theorems internally to the theory. For
-instance, since we cannot express functor categories, we cannot
-express any theorem that quantifies over functors purely in ordered
-predicate logic, and instead only as a metatheorem. To more easily
-formalize these metatheorems, we add a third layer to the type theory:
-an ordinary dependent type theory in the style of Martin-L\"of type
-theory. Then all of our ordered predicate logic judgments will also be
-indexed by a context from this dependent type theory, and the type
-theory includes types for categories, functors, profunctors and
-natural transformations that allow us to formalize theorems the logic
-is too restrictive to encode. This is analogous to the notion of a
-\emph{2-level type theory}: we add a dependent type theory for
-reasoning about the terms of our virtual equipment type theory.
-
+However, we can work around these unary/binary restrictions to some
+extent by viewing the type theory as a domain-specific language embedded
+in a metalanguage.  For example, while we cannot talk about functor
+categories inside the ordered part of the type theory, we can state a
+theorem that quantifies over functors using the meta-language's
+universal quantifier.  To more easily formalize these metatheorems, we
+add a third layer to the type theory: an ordinary dependent type theory
+in the style of Martin-L\"of type theory. Then all of our ordered
+predicate logic judgments will also be indexed by a context from this
+dependent type theory, and the type theory includes types for
+categories, functors, profunctors and natural transformations that allow
+us to formalize theorems the logic is too restrictive to encode. This is
+analogous to the notion of a \emph{2-level type
+  theory}\cite{voevodsky12,altenkirchXX}: we add a dependent type theory
+for reasoning about the terms of our virtual equipment type theory.
 The dependent type theory layer also has an additional purpose: since
@@
 
-\subsection{Basics of Profunctors}
+\paragraph{Basics of Profunctors.}
 
@@
 theory, we briefly define profunctors, which are not included in many
-introductory texts. First, a category $\cat C$ has a collection of
-objects and morphisms with identity and composition, and a functor $F
-: \cat C \to \cat D$ is a function on objects and a function on
-morphisms that respects source and target. It will be useful for
+introductory texts.  Profunctors arise implicitly in formulations of
+universal properties such as representable functors and adjunctions.
+
+First, a category $\cat C$ has a collection of objects and morphisms
+with identity and composition, and a functor $F : \cat C \to \cat D$ is
+a function on objects and a function on morphisms that respects source
+and target and preserves identity and composition. It will be useful for
 intuition to think of a category as a generalization of a preordered
 set, which has a set of elements and a binary \emph{relation} on its
-objects satisfying reflexivity and transitivity. It is fruitful to
-think of a category as a \emph{proof-relevant preorder}, where
-morphisms are thought of as proofs of ordering and the reflexivity and
-transitivity constructions must additionally satisfy equations that
-make them act like identity and composition. Then a functor is a
-constructive version of a monotone function.
+objects satisfying reflexivity and transitivity. It is fruitful to think
+of a category as a \emph{proof-relevant preorder}, where morphisms are
+the proofs of ordering, and the reflexivity and transitivity
+constructions must additionally satisfy equations that make them act
+like identity and composition. Then a functor is a proof-relevant
+version of a monotone function.
 
@@
 functor $R : {\cat C}^{op} \times {\cat D} \to \Set$. We can think of
-each set $R(c,d)$ as a set of ``heteromorphisms'' from $c$, an object
-of $\cat C$ to $d$, an object of $\cat D$. Functoriality then defines
-a notion of composition of heteromorphisms with homomorphisms in $\cat
-C$ and $\cat D$. Thinking of categories as constructive preorders, a
-profunctor is a constructive relation that is downward-closed in $\cat
-C$ and upward-closed in $\cat D$, a kind of ``hetero-ordering''.
+each set $R(c,d)$ as a set of ``heteromorphisms'' from $c$ (an object of
+$\cat C$) to $d$ (an object of $\cat D$).  These ``heteromorphisms''
+differ from the usual ``homomorphisms'' in a single category $\cat C$
+because they relate two objects from two different categories.
+Functoriality then defines a notion of composition of heteromorphisms
+with homomorphisms in $\cat C$ and $\cat D$. Thinking of categories as
+proof-relevant preorders, a profunctor is a proof-relevant relation that
+is downward-closed in $\cat C$ and upward-closed in $\cat D$, a kind of
+``hetero-ordering''.
 %
@@
 
-The simplest example of a profunctor is for any category $\cat C$,
-$\text{Hom}_{\cat C} : \cat C \pto \cat C$ is a profunctor. On
-preorders this is just the ordering itself. The hom profunctor is the
-unit for a composition of profunctors $R \odot S$ which is defined as
-a \emph{co-end}. The heteromorphisms of $(R \odot S)(c,e)$ are triples
-$(d \in \cat D, f \in R(c,d), g \in S(d,e))$ modulo the equivalence
-relation generated by $(d,f,g) ~ (d', f', g')$ if there exists $h \in
-D(d,d')$ such that $h\circ f = f'$ and $g'\circ h = g$. The
-composition of profunctors is a generalization of the composition of
-relations, and the hom profunctor is analogous in the same way to the
-diagonal, or equality relation in that it is the identity for this
-composition.
-For any profunctor $R : \cat C \pto \cat D$ and functors $F : \cat
-C' \to \cat C$ and $G : \cat D' \to \cat D$, by composition we can
-define a profunctor $R(F,G)$, the ``restriction'' of $R$ along $F$ and
-$G$.
-
-Profunctors are not explicitly defined in most introductory texts on
-category theory but arise implicitly in formulations of universal
-properties such a representable functors and adjunctions.
-
-\subsection{Paper Overview}
+In addition to their usefulness for formalizing category theory, one
+reason we take profunctors to be a basic concept of \vett{} is that they
+allow us to give a \emph{universal property} for the type of ``morphisms
+in a category ${\cat C}$''.  This is analogous to how the $J$ elimination
+rule for the identity type in Martin-L\"of type theory gives a universal
+property for morphisms in a groupoid (the special case of a category
+where all morphisms are
+invertible)~\cite{hofmannstreicher,awodey,voevodsky}. The reason
+profunctors are useful for this purpose is that, for any category $\cat
+C$, $\text{Hom}_{\cat C} : \cat C \pto \cat C$ is a profunctor. On
+preorders this is just the preorder's ordering relation
+itself. Moreover, the hom profunctor is the unit for a composition of
+profunctors $R \odot S$ which is defined as a \emph{co-end}. The
+heteromorphisms of $(R \odot S)(c,e)$ are triples $(d \in \cat D, f \in
+R(c,d), g \in S(d,e))$ modulo the equivalence relation generated by
+$(d,f,g) ~ (d', f', g')$ if there exists $h \in D(d,d')$ such that
+$h\circ f = f'$ and $g'\circ h = g$. The composition of profunctors is a
+generalization of the composition of relations, and just as the equality
+relation is the identity for the composition of relations, the hom
+profunctor is the identity for this composition.  The unit law for the
+hom profunctor can be seen as a ``morphism induction'' principle,
+analogous to the ``path induction'' used in homotopy type theory (though
+in this paper we consider only ordinary 1-dimensional categories, not
+higher generalizations).
+
+%% orphaned --- move later? 
+%% For any profunctor $R : \cat C \pto \cat D$ and functors $F : \cat C'
+%% \to \cat C$ and $G : \cat D' \to \cat D$, by composition we can define a
+%% profunctor $R(F,G)$, the ``restriction'' of $R$ along $F$ and $G$.
+
+\paragraph{Paper Overview.}
 
 \begin{enumerate}
-\item In \cref{sec:syntax} we introduce the syntax of Simple Category
-  Theory.
+\item In \cref{sec:syntax} we introduce the syntax of \vett{}.
 
@@
 \item In \cref{sec:semantics}, we develop some basic model theory for
-  ordered predicate logic, including a sound and complete notion of
-  categorical model and sound interpretation in virtual equipments
-  modeling ordinary, enriched and internal category theory.
+  \vett{}, including a sound and complete notion of categorical model
+  and sound interpretation in virtual equipments modeling ordinary,
+  enriched and internal category theory.
 

### 2022-10-07 paper/paper.tex
full draft of semantics section

@@
 
-\begin{construction}[Soundness and Completeness]
-  For any choice of connectives, the syntax of parameterized \ohol{}
-  defines a hyperdoctrine of VDCrs that is an initial object in the
-  category of hyperdoctrines of VDCrs with chosen universal
-  constructions and functors preserving those universal constructions.
+An HVE is then precisely the structure corresponding to the
+\emph{judgments} and actions of substitution in VETT.
+\begin{construction}[Syntactic Model]
+  \label{cons:syn-model}
+  The syntax of VETT with any choice of which subset of connectives
+  are included presents an HVE.
 \end{construction}
 \begin{proof}
-  Most of the details of the proof are straightforward. In the
-  appendix we demonstrate the syntactic elements of the proof which
-  are that the syntax itself is a hyperdoctrine of VDCrs, i.e., that
-  the notions of substitution are appropriately functorial.
+  Straightforward, defining the category of families using the
+  dependent type structure and the virtual equipment structure having
+  ($\alpha$-equivalence classes of) syntactic categories as objects,
+  functors/sets as vertical/horizontal arrows and interpreting
+  compositions/restrictions as substitutions.  The biggest gap is in
+  the definition of the 2-cells.
+
+  A 2-cell from $(\alpha_1:\cat C_1;\alpha_2:\cat C_2\vdash
+  R_1),(\alpha_2:\cat C_2;\alpha_3:\cat C_3),\ldots$ to $(\beta_1:\cat
+  D_1;\beta_2\vdash S)$ with frames $\alpha_1:\cat C_1\vdash b_1 :
+  \cat D_1$ and $\alpha_n:\cat C_n\vdash b_2 : \cat D_2$ given by a
+  term $x_1:R_1,x_2:R_2\ldots \vdash s :
+  S[b_1/\beta_1;b_2/\beta_2]$. Composition is then defined by
+  substitution.
+\end{proof}
+
+Then the \emph{connectives} of VETT each precisely correspond to a
+universal construction of structures in an HVE. The
+$\Pi,\Sigma,\text{Id}$ types correspond to their standard semantics in
+a CwF, products of categories interpreted as products in the vertical
+category, and products of sets as products in the category of
+pro-arrows and 2-cells. The units, tensor and covariant and
+contravariant homs are modeled by the universal properties of the same
+names, as described for instance in \citet{shulmanFrBicats}. The graph
+of a profunctor is modeled by \emph{tabulators}
+\citet{grandis-pare}. Finally, the covariant and contravariant
+presheaf categories can be described as a weakening of the definition
+of a Yoneda equipment from \citet{yosegi-boxes} to virtual double
+categories. More detailed descriptions of these universal properties
+are included in the appendix.
+%
+Then we can package up the soundness and completeness of this notion
+of categorical model in the following modular initiality theorem.
+
+\begin{theorem}[Initiality]
+  The syntax of VETT with almost any choice of connectives presents an
+  HVE that is initial in the category of HVEs with the chosen
+  instances of the universal properties and functors that preserve
+  such chosen instances. The only exception is that if units are
+  included, then contravariant and covariant hom sets must be included
+  as well.
+\end{theorem}
+\begin{proof}
+  It is straightforward to extend the construction
+  \ref{cons:syn-model} for any connective modularly, with the
+  exception that the unit relies on the presence of hom sets in order
+  to satisfy the ``distributivity'' requirement that its elimination
+  can occur in any context. Then we can construct the unique morphism
+  to any HVE by straightforward induction on syntax.
 \end{proof}
@@
   V$-categories, small objects are $\kappa$-small $\mathcal
-  V$-categories etc. This virtual equipment has been shown in
-  loc. cit. to have objects satisfying all of the universal properties
-  used in models of VETT except the graph of a profunctor.
+  V$-categories etc. This virtual equipment has objects satisfying all
+  of the universal properties needed for a model of VETT.
 \end{construction}

### 2022-10-07 paper/paper.tex
cleanup

@@
 notion of semantics in (enriched, internal, indexed) category theory.
-
-First, we can extend any virtual equipment to a hyperdoctrine of
-virtual equipments
-
-First, we can extend any model of \ohol{} to a model of \ohol{}
-parameterized by dependent type theory as follows:
+%
+First, we can extend any virtual equipment to an HVE as follows:
 \begin{construction}
@@
 More generally, categories internal to, enriched in and/or indexed by
-sufficiently nice categories define a virtual equipment that model
-most of the connectives of our type theory. We highlight one example
-from the literature that is highly general: Shulman's enriched indexed
-categories. Shulman's construction defines a virtual double categories
-of large and small $\mathcal V$-categories for any pseudofunctor
-$\mathcal V : S^o \to \MonCat$ where $S$ is a category with finite
-products. He gives examples that show that this internal, enriched and
-indexed categories for suitable choices of $\mathcal V$.
+sufficiently nice categories define a virtual equipment that model the
+connectives of VETT. We highlight one example from the literature that
+is highly general: Shulman's enriched indexed categories. Shulman's
+construction defines a virtual double categories of large and small
+$\mathcal V$-categories for any pseudofunctor $\mathcal V : S^o \to
+\MonCat$ where $S$ is a category with finite products. He gives
+examples that show that this subsumes ordinary internal, enriched and
+indexed categories for suitable choices of $\mathcal V$, as well as
+more general categories that can be thought of as both indexed and
+enriched.
 %

### 2022-10-07 paper/paper.tex
init the bib

@@
 computational lambda calculus \cite{moggi}, linear logic \cite{girard}
-and dependent type theory\cite{not-sure}.
+and dependent type theory\cite{cartmell1986}.
 %

### 2022-10-08 paper/paper.tex
first pass over syntax section

--- 2022-10-07 paper/paper.tex
@@
       $\top$ & $1$\\
-      $\forall x. P \Rightarrow Q$ & $\homrallXYtoZ {\alpha:\cat C} P Q, \homlallXYtoZ {\alpha:\cat C} P Q$ \\
+      $\forall x. P \Rightarrow Q$ & $\homrallXYtoZ {\alpha:\cat C} P Q \text{ and } \homlallXYtoZ {\alpha:\cat C} P Q$ \\
       $\exists x. P \wedge Q$ & $\tensorexistsXwithYandZ {\alpha:\cat C} P Q$\\
       $x =_X y$ & $\punitinXfromYtoZ {\cat C} \alpha \beta$\\
-      Proof $\forall \overrightarrow \alpha. R_1(\alpha_1,\alpha_2) \wedge \cdots \Rightarrow Q$ & Nat. Trans./Element $\alpha_1,x_1: R_1(\alpha_1,\alpha_2),\ldots \vdash t : Q$\\
+      Proof $\forall \overrightarrow \alpha. R_1 \wedge \cdots \Rightarrow Q$ & Nat. Trans./Element $\alpha_1,x_1: R_1(\alpha_1,\alpha_2),\ldots \vdash t : Q$\\
       \hline
@@
 The syntactic forms of \vett{} are given in
-\cref{fig:syntax}.
+Figure~\ref{fig:syntax}.
 %
@@
 
-After these forms we have types and terms, which are for the most part
-the same as in dependent type theory, except that we have universe
-objects for classifying categories, functors, profunctors and natural
-transformations.
+After these forms we have types and terms, which represent the
+meta-language that we use to talk about categories/profunctors/natural
+transformations.  The types include the usual $\Pi$ and $\Sigma$ and
+identity types of dependent type theory, along with universes of
+categories, functors, profunctors and natural transformations.
 %
@@
 element variables with their sets, alternating with those sets' object
-variables.
+variables.  A typical $\Phi$ will look like
+\[
+\alpha_1 : \cat C_1, x_1:R_1(\alpha_1,\alpha_2), \alpha_2 : \cat C_2,
+x_2:R_2(\alpha_2,\alpha_3), \alpha_3:\cat C_3, \ldots
+\]
+and represents the composition of the relations $R_1, R_2, R_3, \ldots,
+R_n$.  We write $d^-(\Phi)$ for the first category variable in $\Phi$
+(which we regard as the negative/contravariant position), $d^+(\Phi)$
+for the last category variable in $\Phi$ (which we regard as the
+positive/covariant position), and $\Phi_1 \jnctx \Phi_2$ for the append
+of two category contexts where the last variable in $\Phi_1$ is equal to
+the first variable in $\Phi_2$.  Formal inductive definitions are in the
+appendix, but intuitively we have: 
+\[
+\begin{array}{ccl}
+d^-(\alpha_1 : \cat C_1, x_1:R_1(\alpha_1,\alpha_2), \ldots, x_n:R_n(\alpha_n,\alpha_n), \alpha_{n+1} : \cat C_{n+1}) & = & \alpha_1 : \cat C_1\\
+d^+(\alpha_1 : \cat C_1, x_1:R_1(\alpha_1,\alpha_2), \ldots, x_n:R_n(\alpha_n,\alpha_n), \alpha_{n+1} : \cat C_{n+1}) & = & \alpha_{n+1}:\cat C_{n+1}\\
+(\Phi_1,\beta:\cat D) \jnctx (\beta:\cat D, \Phi_2) & = & \Phi_1, \beta : \cat D, \Phi_2
+\end{array}
+\]
 
@@
 
-Next, \cref{fig:formation} shows the \emph{formation} rules for these
-various syntactic forms. All other forms are typed with respect to a
-well-formed type context $\Gamma$. Objects are typed additionally with
-an input object variable $\alpha : \cat C$ and an output category
-$\cat D$; in the semantics these objects are modeled as functors. A
-set $S$ is typed with respect to a set context $\Xi$ to describe its
-covariant/contravariant dependence on some input objects. These are
-semantically modeled as profunctors. A transformation $s$ has a
-context of transformation variable and a single set output. To be well
-formed, the context and set must be parameterized by the same
-contravariant and covariant object variables. To ensure this, we use a
-coercion operation $\underline \Phi$ that erases everything in the
-context but the left-most and right-most object variables.
-
-\begin{figure}
-  \begin{mathpar}
-    \inferrule*[right=Cat-form]
-    {\Gamma \isadtctx}
-    {\Gamma\vdash \cat C \isaCat}
-
-    \inferrule*[right=Ob-form]
-    {\Gamma \vdash \cat C\isaCat\and \Gamma \vdash \cat D \isaCat}
-    {\Gamma \pipe \alpha : \cat C \vdash a : \cat D}\\
-
-    \inferrule*[right=SetForm]
-    {\Gamma \isadtctx \and \Gamma \vdash \Xi \boundary}
-    {\Gamma \pipe \Xi \vdash S \isaSet}
-
-    \inferrule*[right=TransFormation]
-    {\Gamma \vdash \Phi \isavectx\and
-     \Gamma\pipe \underline \Phi \vdash R \isaSet}
-    {\Gamma \pipe \Phi \vdash s : R}
-  \end{mathpar}
-  \caption{Selected Formation Rules}
-  \label{fig:formation}
-\end{figure}
+The \emph{formation conditions} or \emph{presuppositions} of a judgement
+are the conditions under which a judgement ``type checks''.  We have
+\begin{itemize}
+\item ${\Gamma\vdash \cat C \isaCat}$ presupposes ${\Gamma \isadtctx}$
+\item ${\Gamma \pipe \alpha : \cat C \vdash a : \cat D}$ presupposes
+       $\Gamma \vdash \cat C\isaCat$ and $\Gamma \vdash \cat D \isaCat$.
+\item ${\Gamma \pipe \Xi \vdash S \isaSet}$ presupposes
+      $\Gamma \isadtctx$ and $\Gamma \vdash \Xi \boundary$
+\item ${\Gamma \pipe \Phi \vdash s : R}$ presupposes
+     ${\Gamma \vdash \Phi \isavectx}$
+     and $\Gamma\pipe \underline \Phi \vdash R \isaSet$
+\end{itemize}
+
+%% Next, \cref{fig:formation} shows the \emph{formation} rules for these
+%% various syntactic forms.
+
+All other forms are typed with respect to a well-formed type context
+$\Gamma$. Objects are typed additionally with an input object variable
+$\alpha : \cat C$ and an output category $\cat D$; in the semantics
+these objects are modeled as functors. A set $S$ is typed with respect
+to a set context $\Xi$ to describe its covariant/contravariant
+dependence on some input objects. These are semantically modeled as
+profunctors. A transformation $s$ has a context of transformation
+variable and a single set output. To be well formed, the context and set
+must be parameterized by the same contravariant and covariant object
+variables. To ensure this, we use a coercion operation $\underline \Phi$
+that erases everything in the context but the left-most and right-most
+object variables.
+
+%% \begin{figure}
+%%   \begin{mathpar}
+%%     \inferrule*[right=Cat-form]
+%%     {\Gamma \isadtctx}
+%%     {\Gamma\vdash \cat C \isaCat}
+
+%%     \inferrule*[right=Ob-form]
+%%     {\Gamma \vdash \cat C\isaCat\and \Gamma \vdash \cat D \isaCat}
+%%     {\Gamma \pipe \alpha : \cat C \vdash a : \cat D}\\
+
+%%     \inferrule*[right=SetForm]
+%%     {\Gamma \isadtctx \and \Gamma \vdash \Xi \boundary}
+%%     {\Gamma \pipe \Xi \vdash S \isaSet}
+
+%%     \inferrule*[right=TransFormation]
+%%     {\Gamma \vdash \Phi \isavectx\and
+%%      \Gamma\pipe \underline \Phi \vdash R \isaSet}
+%%     {\Gamma \pipe \Phi \vdash s : R}
+%%   \end{mathpar}
+%%   \caption{Selected Formation Rules}
+%%   \label{fig:formation}
+%% \end{figure}
 
@@
 context admit an action of substitution. For terms, this is the
-ordinary type theoretic substitution. Naturally objects $\alpha:\cat C
+ordinary type theoretic substitution. Objects $\alpha:\cat C
 \vdash a : \cat D$ can be substituted for object variables $\beta :
@@
 $\beta: \cat D$, then we can substitute objects $a : \cat C$ and $b :
-\cat D$ for these variables $P[a/\alpha;b\beta]$. This generalizes the
+\cat D$ for these variables $P[a/\alpha;b/\beta]$. This generalizes the
 ordinary precomposition of a relation by a function.  Semantically
@@
 
+Recall that the basic entities of our type theory are categories,
+objects/functors, sets/profunctors, and elements/natural
+transformations.  In this section we discuss some connectives for
+constructing categories, which are specified by introduction and
+elimination rules (the $\beta\eta$ equality and substitution rules are
+included in the appendix).  The introduction and elimination rules make
+use of functors, profunctors, and natural transformations.
+%
 We introduce the connectives for defining categories in
-Figure~\ref{fig:category-connectives}. We show the formation,
-introduction and elimination rules, but additionally there are rules
-for $\beta\eta$ equality and substitution, which are included in the
-appendix.
+Figure~\ref{fig:category-connectives}. 
 %
 First we introduce the additives: the unit category $1$ and product
-category $\cat C \times \cat D$ with their usual introduction and
-elimination rules.
+category $\cat C \times \cat D$ have the usual introduction and
+elimination rules defining functors to/from them.  
 %
@@
   \begin{mathpar}
-    \inferrule{}{\Gamma \vdash 1 \isaCat}\and
-    \inferrule*{}{\Gamma\pipe\alpha:C \vdash () : 1}\and
+    \inferrule{ ~ }{\Gamma \vdash 1 \isaCat}
+    \quad
+    \inferrule*{ ~ }{\Gamma\pipe\alpha:C \vdash () : 1}\and
 
@@
         {\Gamma \vdash {\cat C_1} \times {\cat C_2} \isaCat}
+    \quad    
     \inferrule
@@
     {\Gamma \pipe\alpha:\cat C \vdash (a_1,a_2) : {\cat C_1} \times {\cat C_2}}
-
+    \quad
     \inferrule
@@
     {\Gamma \vdash \graphProf{\alpha}{\beta} P \isaCat}
-
-    \inferrule
-    {\defaultObCtx \vdash a : \graphProf{\alpha:\cat C_-}{\beta} P}
-    {\defaultObCtx \vdash \pi_- a : \cat C_-}
-
-    \inferrule
-    {\defaultObCtx \vdash a : \graphProf{\alpha}{\beta:C_+} P}
-    {\defaultObCtx \vdash \pi_+ a : \cat C_+}
-
-    \inferrule
-    {\defaultObCtx \vdash a : \graphProf{\alpha}{\beta} P}
-    {\defaultObCtx \vdash \pi_e a : P[\pi_-a/\alpha;\pi_+a/\beta]}
-
+    \quad
     \inferrule
@@
     \inferrule
+    {\defaultObCtx \vdash a : \graphProf{\alpha:\cat C_-}{\beta} P}
+    {\defaultObCtx \vdash \pi_- a : \cat C_-}
+    \quad
+    \inferrule
+    {\defaultObCtx \vdash a : \graphProf{\alpha}{\beta:C_+} P}
+    {\defaultObCtx \vdash \pi_+ a : \cat C_+}
+    \quad
+    \inferrule
+    {\defaultObCtx \vdash a : \graphProf{\alpha}{\beta} P}
+    {\defaultObCtx \vdash \pi_e a : P[\pi_-a/\alpha;\pi_+a/\beta]}
+
+    \inferrule
     {\Gamma \vdash \cat C \isaCat \and \cat C \isSmall}
     {\Gamma \vdash \negPresheaf {\cat C} \isaCat}
-
+    \quad
     \inferrule
@@
     {\Gamma \pipe \Xi \vdash \negPresheafAppPtoX p a \isaSet}
-
+    \quad
     \inferrule
@@
     {\Gamma \vdash \posPresheaf {\cat D} \isaCat}
-
+    \quad
     \inferrule
@@
     {\Gamma \pipe \Xi \vdash \posPresheafAppPtoX p a \isaSet}
-
+    \quad
     \inferrule
@@
   \caption{Category Conectives}
+  \label{fig:category-connectives}
 \end{figure}
@@
 
-Next, we cover the connectives for the sets, i.e., constructions on
-profunctors in Figure~\ref{fig:set-connectives}.
+Next, in Figure~\ref{fig:set-connectives}, we cover the connectives for
+the sets, i.e., constructions on profunctors.  In these constructions,
+the sets/profunctors act like types, and they are given 
+introduction and elimination rules for elements/natural
+transformations.  
 %
@@
 \alpha R P$ and contravariant homs $\homlallXYtoZ {\alpha} R P$ and
-the unit type $\punitinXfromYtoZ {\cat C} {a_1}{a_2}$.
+the unit type $\punitinXfromYtoZ {\cat C} {a_1}{a_2}$, and
+the cartesian unit and product types $1$ and $R \times S$.  
+%
+The remainder of the figure gives the introduction and elimination rules
+for these sets (the $\beta/eta$-rules are in the appendix).  
 %
@@
 by specifying the behavior when $s$ is of the form $\punitrefl \alpha$
-in the form of a continuation $\alpha. t$. Again we elide all
-$\beta\eta$ rules, which are standard. Next, the tensor product of
+in the form of a continuation $\alpha. t$. Next, the tensor product of
 sets is a kind of combined existential quantifier and monoidal
@@
 arbitrary middle object $\beta$.
-%
-Note that in the unit elimination, the continuation is typed with a
-context consisting of only a single object $\alpha$, in contrast with
+
+FIXME: Note that in the unit elimination, the continuation is typed with
+a context consisting of only a single object $\alpha$, in contrast with
 normal rules for a unit type in natural deduction for ordered linear
 logic such as \cite{polakawpfenning}, which would split the context
-$\Phi$ into $\Phi_l \jnctx \Phi_m \jnctx \Phi_r$ where $\Phi_m$ is
-used by the scrutinee and $\Phi_l \jnctx \Phi_r$ is used by the
+$\Phi$ into $\Phi_l \jnctx \Phi_m \jnctx \Phi_r$ where $\Phi_m$ is used
+by the scrutinee and $\Phi_l \jnctx \Phi_r$ is used by the
 continuation. However, unlike in ordered linear logic, the $\Phi_l
-\jnctx \Phi_r$ is not, necessarily well formed because there is no
-reason in general that $\Phi_m$ has contravariant and covariant
-dependence on the same category.
+\jnctx \Phi_r$ is not necessarily well formed because there is no reason
+in general that $\Phi_m$ has contravariant and covariant dependence on
+the same category.
 %
@@
 covariant ($\homrallXYtoZ \alpha R P$) homs of sets, which are
-different because we are in an orderd logic.
+different from each other because we are in an ordered logic.
 %
@@
 
-Finally, we briefly describe the connectives for the
-``meta-logic''. These include the standard rules for Martin-L\"of type
-theory with Pi/Sigma and extensional identity types, whose rules we
-elide as they are standard \cite{emltt}. We use extensional identity
-types so that the description of models is simpler, but there is no
-obstacle to including models with intensional identity types.
+Finally, we briefly describe the connectives for the ``meta-logic'',
+which extends Martin-L\"of type theory with $Pi$/$Sigma$ and extensional
+identity types (with their standard rules)~\cite{emltt}. We use
+extensional identity types so that the description of models is simpler,
+but intensional identity types could be used instead.
 %
@@
 %
-The rule for the types of small categories and categories are very
+The rule for the types of small categories and (large) categories are very
 similar: any definable category defines an element of type $\Cats$,
@@
 involved additionally satisfy $\cat C \isSmall$. Again we elide the
-standard $\beta\eta$ principles.
+$\beta\eta$ principles, which state that the introduction and
+elimination rules are a bijection between judgement forms.  
 %
@@
     {\Gamma\vdash \Cats}
+
+    FIXME missing intro/elim rules for smallcat and cat
 
@@
 
-An HVE is then precisely the structure corresponding to the
-\emph{judgments} and actions of substitution in VETT.
-\begin{construction}[Syntactic Model]
-  \label{cons:syn-model}
-  The syntax of VETT with any choice of which subset of connectives
-  are included presents an HVE.
+\begin{construction}[Soundness and Completeness]
+  For any choice of connectives, the syntax of parameterized \ohol{}
+  defines a hyperdoctrine of VDCrs that is an initial object in the
+  category of hyperdoctrines of VDCrs with chosen universal
+  constructions and functors preserving those universal constructions.
 \end{construction}
 \begin{proof}
-  Straightforward, defining the category of families using the
-  dependent type structure and the virtual equipment structure having
-  ($\alpha$-equivalence classes of) syntactic categories as objects,
-  functors/sets as vertical/horizontal arrows and interpreting
-  compositions/restrictions as substitutions.  The biggest gap is in
-  the definition of the 2-cells.
-
-  A 2-cell from $(\alpha_1:\cat C_1;\alpha_2:\cat C_2\vdash
-  R_1),(\alpha_2:\cat C_2;\alpha_3:\cat C_3),\ldots$ to $(\beta_1:\cat
-  D_1;\beta_2\vdash S)$ with frames $\alpha_1:\cat C_1\vdash b_1 :
-  \cat D_1$ and $\alpha_n:\cat C_n\vdash b_2 : \cat D_2$ given by a
-  term $x_1:R_1,x_2:R_2\ldots \vdash s :
-  S[b_1/\beta_1;b_2/\beta_2]$. Composition is then defined by
-  substitution.
-\end{proof}
-
-Then the \emph{connectives} of VETT each precisely correspond to a
-universal construction of structures in an HVE. The
-$\Pi,\Sigma,\text{Id}$ types correspond to their standard semantics in
-a CwF, products of categories interpreted as products in the vertical
-category, and products of sets as products in the category of
-pro-arrows and 2-cells. The units, tensor and covariant and
-contravariant homs are modeled by the universal properties of the same
-names, as described for instance in \citet{shulmanFrBicats}. The graph
-of a profunctor is modeled by \emph{tabulators}
-\citet{grandis-pare}. Finally, the covariant and contravariant
-presheaf categories can be described as a weakening of the definition
-of a Yoneda equipment from \citet{yosegi-boxes} to virtual double
-categories. More detailed descriptions of these universal properties
-are included in the appendix.
-%
-Then we can package up the soundness and completeness of this notion
-of categorical model in the following modular initiality theorem.
-
-\begin{theorem}[Initiality]
-  The syntax of VETT with almost any choice of connectives presents an
-  HVE that is initial in the category of HVEs with the chosen
-  instances of the universal properties and functors that preserve
-  such chosen instances. The only exception is that if units are
-  included, then contravariant and covariant hom sets must be included
-  as well.
-\end{theorem}
-\begin{proof}
-  It is straightforward to extend the construction
-  \ref{cons:syn-model} for any connective modularly, with the
-  exception that the unit relies on the presence of hom sets in order
-  to satisfy the ``distributivity'' requirement that its elimination
-  can occur in any context. Then we can construct the unique morphism
-  to any HVE by straightforward induction on syntax.
+  Most of the details of the proof are straightforward. In the
+  appendix we demonstrate the syntactic elements of the proof which
+  are that the syntax itself is a hyperdoctrine of VDCrs, i.e., that
+  the notions of substitution are appropriately functorial.
 \end{proof}
@@
   V$-categories, small objects are $\kappa$-small $\mathcal
-  V$-categories etc. This virtual equipment has objects satisfying all
-  of the universal properties needed for a model of VETT.
+  V$-categories etc. This virtual equipment has been shown in
+  loc. cit. to have objects satisfying all of the universal properties
+  used in models of VETT except the graph of a profunctor.
 \end{construction}

### 2022-10-10 paper/paper.tex
intro/elim for cats and small cats
reformat figures

@@
       1 & $1$ \\
-      $\mathcal P X$ & $\mathcal P^+ X, \mathcal P^- X$ \\
+      $\mathcal P X$ & $\mathcal P^+ X$ \text{ and } $\mathcal P^- X$ \\
       $\{ (x,y) \in X \times Y | R(x,y) \}$ & $\sum_{\alpha:C;\beta:D} R$\\
@@
   \caption{Analogy between Higher-Order Logic and \vett{} Judgments and Connectives}
+\label{fig:hol-cats}
 \end{figure}
@@
 categories, functors, profunctors and natural transformations.
-%
-Finally we have several forms of context which are used in the theory:
-contexts $\Gamma$ of term variables with their types, contexts $\Xi$
-of object variables with their categories, and contexts $\Phi$ of
+
+Finally we have several forms of context which are used in the theory.
+The contexts $\Gamma$ of term variables with their types are as usual.
+The category contexts $\Xi$ contain object variables with their
+categories; the two forms of category context are $\alpha : \cat C$,
+containing one variable that can be used both contravariantly and
+covariantly, and $alpha : \cat C ; \beta : \cat D$, containing a
+contravariant variable $\alpha$ and covariant variable $\beta$.
+Finally, we have contexts $\Phi$ of
 element variables with their sets, alternating with those sets' object
-variables.  A typical $\Phi$ will look like
+variables.  A typical $\Phi$ has the shape
 \[
 \alpha_1 : \cat C_1, x_1:R_1(\alpha_1,\alpha_2), \alpha_2 : \cat C_2,
-x_2:R_2(\alpha_2,\alpha_3), \alpha_3:\cat C_3, \ldots
+x_2:R_2(\alpha_2,\alpha_3), \ldots, R_n(\alpha_n,\alpha_{n+1}), \alpha_{n+1} : C_{n+1}
 \]
@@
 
-\begin{figure}
+\begin{figure}[t]
   \begin{mathpar}
@@
       &&& \Cat \pipe \varr{\cat C}{\cat D}\pipe \harr {\cat C} {\cat D}\pipe \pendallXdotY {\alpha:{\cat C}} R\\
-      \textrm{Term} & L,M,N & ::= & X \pipe \lambda X:M. N \pipe M\,N \pipe (M,N) \pipe \pi_1 M \pipe \pi_2 M\\
-      &&&\lceil \cat C \rceil \pipe \vlambda {\alpha:\cat C} a \pipe \hlambda {\alpha:\cat C}{\beta: \cat D} R \ldots \\
+      \textrm{Term} & L,M,N & ::= & X \pipe \lambda X:M. N \pipe M\,N \pipe (M,N) \pipe \pi_1 M \pipe \pi_2 M \mid \text{refl}_{M} \\
+      &&&\lceil \cat C \rceil \pipe \vlambda {\alpha:\cat C} a \pipe \hlambda {\alpha:\cat C}{\beta: \cat D} R \mid FIXME \ldots \\
       \textrm{Type Context} & \Gamma, \Delta & ::= & \cdot \pipe \Gamma, X : A\\
@@
 
-The \emph{formation conditions} or \emph{presuppositions} of a judgement
-are the conditions under which a judgement ``type checks''.  We have
+Next, we overview our basic judgement forms along with the
+\emph{formation conditions} or \emph{presuppositions} (the conditions
+under which a judgement ``type checks'').  We have
 \begin{itemize}
-\item ${\Gamma\vdash \cat C \isaCat}$ presupposes ${\Gamma \isadtctx}$
-\item ${\Gamma \pipe \alpha : \cat C \vdash a : \cat D}$ presupposes
+\item Categories: ${\Gamma\vdash \cat C \isaCat}$ presupposes ${\Gamma \isadtctx}$.
+
+\item Objects/functors: ${\Gamma \pipe \alpha : \cat C \vdash a : \cat D}$ presupposes
   $\Gamma \vdash \cat C\isaCat$ and $\Gamma \vdash \cat D \isaCat$.
-\item ${\Gamma \pipe \Xi \vdash S \isaSet}$ presupposes
+Objects are typed with an input object variable
+$\alpha : \cat C$ and an output category $\cat D$; in the semantics
+these objects are modeled as functors.
+
+\item Sets/profunctors: ${\Gamma \pipe \Xi \vdash S \isaSet}$ presupposes
       $\Gamma \isadtctx$ and $\Gamma \vdash \Xi \boundary$
-\item ${\Gamma \pipe \Phi \vdash s : R}$ presupposes
-     ${\Gamma \vdash \Phi \isavectx}$
-     and $\Gamma\pipe \underline \Phi \vdash R \isaSet$
-\end{itemize}
-
-%% Next, \cref{fig:formation} shows the \emph{formation} rules for these
-%% various syntactic forms.
-
-All other forms are typed with respect to a well-formed type context
-$\Gamma$. Objects are typed additionally with an input object variable
-$\alpha : \cat C$ and an output category $\cat D$; in the semantics
-these objects are modeled as functors. A set $S$ is typed with respect
+A set $S$ is typed with respect
 to a set context $\Xi$ to describe its covariant/contravariant
 dependence on some input objects. These are semantically modeled as
-profunctors. A transformation $s$ has a context of transformation
+profunctors.
+
+\item Elements/natural transformations: ${\Gamma \pipe \Phi \vdash s : R}$ presupposes
+     ${\Gamma \vdash \Phi \isavectx}$
+  and $\Gamma\pipe \underline \Phi \vdash R \isaSet$.
+  A transformation $s$ has a context of transformation
 variable and a single set output. To be well formed, the context and set
@@
 that erases everything in the context but the left-most and right-most
-object variables.
+object variables, i.e. $\underline{\alpha : \cat C} = \alpha : \cat C$ and
+$\underline{\Phi} = d^-(\Phi);d^+(\Phi)$ otherwise.
+
+\item Meta-language types and terms: $\Gamma \vdash A \isaTy$ presupposes $\Gamma \isadtctx$ and $\Gamma
+  \vdash M : A$ presupposes $\Gamma \isadtctx$ and $\Gamma \vdash A
+  \isaTy$.
+\end{itemize}
 
@@
 
-In typical natural-deduction style, all syntactic forms typed in
-context admit an action of substitution. For terms, this is the
-ordinary type theoretic substitution. Objects $\alpha:\cat C
-\vdash a : \cat D$ can be substituted for object variables $\beta :
-\cat D$ in other objects. We can also substitute objects into
-\emph{sets}, that is, if we have a set $P$ parameterized by a
-contravariant variable $\alpha : \cat C$ and a covariant variable
-$\beta: \cat D$, then we can substitute objects $a : \cat C$ and $b :
-\cat D$ for these variables $P[a/\alpha;b/\beta]$. This generalizes the
-ordinary precomposition of a relation by a function.  Semantically
-this is the ``restriction'' of a profunctor along two functors, which
-is just composition of functors if a profunctor is viewed as a functor
-to $\Set$. Modeling this operation as a substitution considerably
-simplifies reasoning using profunctors.
-%
-Finally we have the action of substitution on natural
-transformations. A natural transformation is indexed by set variables
-$x:R$, whose sets are each parameterized by a contravariant and a
-covariant category variable $\alpha;\beta$. We can think of natural
-transformations as \emph{polymorphic} in the categories involved, and
-so when we make a transformation substitution, we substitute not just
-the set variables, but we also \emph{instantiate} the polymorphic
-category variables with objects.
+All syntactic forms typed in context admit an action of
+substitution. For (meta-language) types and terms, this is as
+usual. Objects $\alpha:\cat C \vdash a : \cat D$ can be substituted for
+object variables $\beta : \cat D$ in other objects. We can also
+substitute objects into \emph{sets}, that is, if we have a set $P$
+parameterized by a contravariant variable $\alpha : \cat C$ and a
+covariant variable $\beta: \cat D$, then we can substitute objects $a :
+\cat C$ and $b : \cat D$ for these variables $P[a/\alpha;b/\beta]$. This
+generalizes the ordinary precomposition of a relation by a function.
+Semantically this is the ``restriction'' of a profunctor along two
+functors, which is just composition of functors if a profunctor is
+viewed as a functor to $\Set$. Modeling this operation as a substitution
+considerably simplifies reasoning using profunctors.
+%
+Finally we have the action of substitution on elements/natural transformations.
+First, we can substitute elements/natural transformations for the set
+variables in elements, denoting the composition of natural
+transformations.  Second, an element is also parameterized
+by a contravariant and a covariant category variable $\alpha;\beta$. We
+can think of natural transformations as \emph{polymorphic} in the
+categories involved, and so when we make a transformation substitution,
+we also \emph{instantiate} the polymorphic category variables with
+objects.
 %
@@
 
-Recall that the basic entities of our type theory are categories,
-objects/functors, sets/profunctors, and elements/natural
-transformations.  In this section we discuss some connectives for
-constructing categories, which are specified by introduction and
-elimination rules (the $\beta\eta$ equality and substitution rules are
-included in the appendix).  The introduction and elimination rules make
-use of functors, profunctors, and natural transformations.
-%
-We introduce the connectives for defining categories in
-Figure~\ref{fig:category-connectives}. 
+In this section we discuss some connectives for constructing categories,
+which are specified by introduction and elimination rules in
+Figure~\ref{fig:category-connectives} (the $\beta\eta$ equality and
+substitution rules are included in the appendix).  The introduction and
+elimination rules make use of functors, profunctors, and natural
+transformations.
 %
@@
 
-\begin{figure}
-  \begin{mathpar}
+\begin{figure}[t]
+  \[
+  \begin{array}{l}
+    \text{Unit:} \quad
     \inferrule{ ~ }{\Gamma \vdash 1 \isaCat}
     \quad
-    \inferrule*{ ~ }{\Gamma\pipe\alpha:C \vdash () : 1}\and
-
+    \inferrule*{ ~ }{\Gamma\pipe\alpha:C \vdash () : 1}\\\\
+
+    \text{Product:} \quad
     \inferrule
@@
     {\Gamma\pipe\alpha:\cat C\vdash \pi_i a : \cat C_i}
-
+    \\\\
+
+    \text{Graph of a profuctor:} \\
     \inferrule
@@
     {\defaultObCtx \vdash (a_-,a_+,s) : \graphProf{\alpha:\cat C_-}{\beta:\cat C_+} P}
-
+    \\
     \inferrule
@@
     {\defaultObCtx \vdash \pi_e a : P[\pi_-a/\alpha;\pi_+a/\beta]}
-
+    \\\\
+    \text{Negative Presehaf:} \\
     \inferrule
@@
     {\Gamma \pipe \beta:\cat D \vdash \lambda \alpha:\cat C. R : \negPresheaf {\cat C}}
-
+    \\
+    \\
+    \text{Positive Presehaf:} \\
     \inferrule
@@
     {\Gamma \pipe \alpha:\cat C \vdash \lambda \beta:\cat D. R : \posPresheaf {\cat D}}
-  \end{mathpar}
+  \end{array}
+  \]
   \caption{Category Conectives}
@@
 Next, in Figure~\ref{fig:set-connectives}, we cover the connectives for
-the sets, i.e., constructions on profunctors.  In these constructions,
-the sets/profunctors act like types, and they are given 
-introduction and elimination rules for elements/natural
-transformations.  
+the sets/profunctors.  In these constructions, the sets/profunctors act
+like types, and they are given introduction and elimination rules for
+elements/natural transformations.
 %
@@
 \begin{figure}
-  \begin{mathpar}
     % unit,
-
-    \inferrule
-    {\cat D \isSmall\\\\
-      \Gamma \pipe d^-\Xi; \beta:D\vdash P \isaSet\\\\
-     \Gamma \pipe \beta:D;d^+\Xi \vdash Q \isaSet}
-    {\Gamma \pipe \Xi \vdash \tensorexistsXwithYandZ {\beta:D} P Q}
-
-    \inferrule
-    {d^-\Xi \isSmall\\\\
-      \Gamma \pipe \alpha:\cat C; d^-\Xi\vdash R \\\\ \Gamma \pipe \alpha : \cat C; d^+\Xi\vdash P}
-    {\Gamma \pipe \Xi \vdash \homlallXYtoZ {\alpha: \cat C} R P}
-
-  \inferrule
-  {d^+\Xi \isSmall\\\\ \Gamma \pipe d^+\Xi; \alpha : \cat C \vdash  R \\\\ \Gamma \pipe  d^-\Xi; \alpha : \cat C \vdash P}
-  {\Gamma \pipe \Xi \vdash \homrallXYtoZ {\alpha:\cat C} R P}\\
-
+  \[
+  \begin{array}{l}
+
+    \text{Unit/morphism set:}\\
     \inferrule
@@
   {\Gamma \pipe \Xi \vdash \punitinXfromYtoZ {\cat C} {a_1} {a_2}}
-
-    \inferrule{}{\Gamma\pipe \Xi \vdash 1 \isaSet}
-
-    \inferrule{\Gamma\pipe\Xi \vdash R \isaSet \\\\ \Gamma\pipe\Xi \vdash S\isaSet}
-              {\Gamma\pipe\Xi \vdash R \times S\isaSet}\\
-
+  \quad
     \inferrule
@@
     {\Gamma \pipe\cdot \vdash \punitrefl a : \punitinXfromYtoZ {\cat C} a a}
-
+    \quad
 %%     \inferrule*[right=Unit Elim]
@@
 %%       \pendallXdotY {\alpha_1} {\homrallXYtoZ {\alpha_2} {\punitinXfromYtoZ {\cat C} {\alpha_1}{\alpha_2}} {\harrapp Q {\alpha_1}{\alpha_2}}}}
-
     \inferrule
@@
     {\Gamma \vdash \punitelimkontZatABC {\alpha. t} A s B : P[A/\alpha^-;B/\alpha^+]}
-
+    \\
   %% \inferrule*[right=Tensor intro function]
@@
 
+    \\
+    \text{Tensor product:}\\
+    \inferrule
+    {\cat D \isSmall\\\\
+      \Gamma \pipe d^-\Xi; \beta:D\vdash P \isaSet\\\\
+     \Gamma \pipe \beta:D;d^+\Xi \vdash Q \isaSet}
+    {\Gamma \pipe \Xi \vdash \tensorexistsXwithYandZ {\beta:D} P Q}
+    \quad
   \inferrule
@@
   {\Gamma\pipe\Psi_s \jnctx \Psi_t\vdash \tensorintroatXwithYandZ b s t : \tensorexistsXwithYandZ {\beta : \cat D} P Q}
-
+  \quad
   \inferrule
@@
   {\Gamma\pipe\Phi_l\jnctx \Phi_m \jnctx \Phi_r \vdash \tensorelimWkontZ s {x.\beta.y. t} : R}
-
+  \\
   %% \tensorintroatXwithYandZ B M N [\phi,B'/\beta',\psi] = \tensorintroatXwithYandZ {B[B'\beta]} {M[\phi]} {N[\psi]}
@@
 %%   {\Gamma \vdash \tensorelimfromWXtoYwithkontZ P Q R N : \homunary \alpha {(\tensorexistsXwithYandZ \beta {\harrapp P \alpha \beta} {\harrapp Q \beta \gamma})} {\gamma} {\harrapp R \alpha \gamma}}
+
+  \\
+  \text{Right hom:} \quad
+  \inferrule
+  {d^+\Xi \isSmall\\\\ \Gamma \pipe d^+\Xi; \alpha : \cat C \vdash  R \\\\ \Gamma \pipe  d^-\Xi; \alpha : \cat C \vdash P}
+  {\Gamma \pipe \Xi \vdash \homrallXYtoZ {\alpha:\cat C} R P}
+  \quad
   \inferrule
@@
    \homrlambdaXatYdotZ {x : R} {\alpha : \cat C} t : \homrallXYtoZ \alpha R P}
-
-      %% \inferrule*[right=Covar Hom elim simple]
-      %% {}
-      %% {\alpha: \cat C, f:\forall \gamma. P \triangleleft R, \beta : \cat D, x : R, \gamma : \cat E \vdash f\triangleleft (x, \gamma) : P}
-
+  \quad
   \inferrule
@@
   {\Gamma\pipe \Phi_f \jnctx \Phi_a \vdash \homrappXtoYatZ s t a : P[a/\alpha]}\\
-
+  \quad
+
+      %% \inferrule*[right=Covar Hom elim simple]
+      %% {}
+      %% {\alpha: \cat C, f:\forall \gamma. P \triangleleft R, \beta : \cat D, x : R, \gamma : \cat E \vdash f\triangleleft (x, \gamma) : P}
+
+  \\ 
+  \text{Left hom:} \quad
+  \inferrule
+    {d^-\Xi \isSmall\\\\
+      \Gamma \pipe \alpha:\cat C; d^-\Xi\vdash R \\\\ \Gamma \pipe \alpha : \cat C; d^+\Xi\vdash P}
+    {\Gamma \pipe \Xi \vdash \homlallXYtoZ {\alpha: \cat C} R P}
+  \quad
   \inferrule
@@
   {\Gamma\pipe\Phi \vdash \homllambdaXatYdotZ {x : R} {\alpha : \cat C} t : \homlallXYtoZ \alpha R P}
-
+  \quad
   \inferrule
@@
 
-  \inferrule{}{\Gamma\pipe \Phi \vdash () :  1}
-
+  \\
+  \text{Cartesian unit and products:} \quad
+      \inferrule{~}{\Gamma\pipe \Xi \vdash 1 \isaSet}
+      \quad
+      \inferrule{~}{\Gamma\pipe \Phi \vdash () :  1}
+      \\ 
+    \inferrule{\Gamma\pipe\Xi \vdash R \isaSet \\\\ \Gamma\pipe\Xi \vdash S\isaSet}
+              {\Gamma\pipe\Xi \vdash R \times S\isaSet}
+    \quad
     \inferrule{\forall i \in \{1,2\}.~ \Gamma \pipe\Phi \vdash M_i : R_i}{\Gamma \pipe\Phi \vdash (M_1,M_2) : R_1 \times R_2}
-
+    \quad
     \inferrule{\Gamma \pipe\Phi\vdash M : R_1 \times R_2}{\Gamma \pipe\Phi \vdash \pi_i M : R_i}
-  \end{mathpar}
+  \end{array}
+  \]
   \caption{Set Connectives}
+  \label{fig:set-connectives}
 \end{figure}

### 2022-10-10 paper/paper.tex
more edits to set connectives section

@@
 \newcommand{\algVarrUnqt}{\textrm{FunctUnQt}}
+
+\newcommand{\quoth}[1]{\lceil{} {#1}\rceil{}}
+\newcommand{\unquoth}[1]{\lfloor{} {#1}\rfloor{}}
 
@@
 Next, in Figure~\ref{fig:set-connectives}, we cover the connectives for
-the sets/profunctors.  In these constructions, the sets/profunctors act
-like types, and they are given introduction and elimination rules for
-elements/natural transformations.
-%
-The top of the figure includes the formation rules for tensor products
-$\tensorexistsXwithYandZ \beta P Q$, the covariant $\homrallXYtoZ
-\alpha R P$ and contravariant homs $\homlallXYtoZ {\alpha} R P$ and
-the unit type $\punitinXfromYtoZ {\cat C} {a_1}{a_2}$, and
-the cartesian unit and product types $1$ and $R \times S$.  
-%
-The remainder of the figure gives the introduction and elimination rules
-for these sets (the $\beta/eta$-rules are in the appendix).  
-%
+the sets/profunctors, which classify elements/natural transformations
+(the $\beta/eta$-rules are in the appendix).
+%% %
+%% The top of the figure includes the formation rules for tensor products
+%% $\tensorexistsXwithYandZ \beta P Q$, the covariant $\homrallXYtoZ
+%% \alpha R P$ and contravariant homs $\homlallXYtoZ {\alpha} R P$ and
+%% the unit type $\punitinXfromYtoZ {\cat C} {a_1}{a_2}$, and
+%% the cartesian unit and product types $1$ and $R \times S$.  
+%% %
+%% The remainder of the figure gives the introduction and elimination rules
+%% for these sets .  
+%% %
 First, the unit set is our syntax for the profunctor of morphisms of a
-category $\cat C$.
-%
-Its introduction and elimination rules are like those of an
-\emph{ordered} variant of equality: the introduction rule is the
-identity morphism (reflexivity) and the elimination rule is an
-induction principle that is a variant of the J rule of equality
-elimination: we can use a term of $s : \punitinXfromYtoZ {\cat C} a b$
-by specifying the behavior when $s$ is of the form $\punitrefl \alpha$
-in the form of a continuation $\alpha. t$. Next, the tensor product of
-sets is a kind of combined existential quantifier and monoidal
-product, which we combine into a single notation
+category $\cat C$, which is well-formed when given two objects of the
+same category.  
+%
+Its introduction and elimination rules are analogous to the usual rules
+for equality in Martin-L\"of type theory.  The introduction rule is the
+identity morphism (reflexivity) and the elimination rule is an induction
+principle: we can use a term of $s : \punitinXfromYtoZ {\cat C} A B$ by
+specifying the behavior when $s$ is of the form $\punitrefl \alpha$ in
+the form of a continuation $\alpha. t$.  Like the $J$ elimination rule
+for equality in Martin-L\"of type theory, $P$ must be ``fully general'',
+i.e. well-typed for variables $\alpha$ and $\beta$.  Those familiar with
+linear logic~\cite{polakawpfenning} might expect a more general rule,
+with the continuation $t$ typed in a context $\Phi_l \jnctx \Phi_r$ and
+$\Phi_l \jnctx \Phi \jnctx \Phi_r$ as the context of the conclusion of
+the rule.  Because of dependency, this is not necessarily well-formed in
+cases where the endpoints $A$ and $B$ of $\punitinXfromYtoZ{\cat C} A B$
+are not fully general (i.e. not distinct variables).  However, the
+instances of this more general rule that do type check are derivable
+from our more restricted rule using right/left-hom types.
+
+The tensor product of sets is a kind of combined existential quantifier
+and monoidal product, which we combine into a single notation
 $\tensorexistsXwithYandZ \beta P Q$, where $\beta$ is the covariant
@@
 
-FIXME: Note that in the unit elimination, the continuation is typed with
-a context consisting of only a single object $\alpha$, in contrast with
-normal rules for a unit type in natural deduction for ordered linear
-logic such as \cite{polakawpfenning}, which would split the context
-$\Phi$ into $\Phi_l \jnctx \Phi_m \jnctx \Phi_r$ where $\Phi_m$ is used
-by the scrutinee and $\Phi_l \jnctx \Phi_r$ is used by the
-continuation. However, unlike in ordered linear logic, the $\Phi_l
-\jnctx \Phi_r$ is not necessarily well formed because there is no reason
-in general that $\Phi_m$ has contravariant and covariant dependence on
-the same category.
-%
-This is analogous to a similar situation in dependent type theory for
-identity types: TODO: Dan please help with this explanation!!
-
 Next, we introduce the contravariant ($\homlallXYtoZ \alpha R P$) and
@@
 two subformulae should have the same covariant and contravariant
-dependence. So as with linear logic, the cartesian unit and product
-complicate the syntactic formulation of linearity: the unit
-constructor ``uses'' all variables and the product ``uses'' all
+dependence. So, as with linear logic, some constructions can
+syntactically use a variable more than once and still be ``linear'': the
+unit constructor ``uses'' all variables and the product ``uses'' all
 variables in both cases.
@@
   {\Gamma \pipe d^-\Xi \vdash a_1 : \cat C\\\\ \Gamma \pipe d^+ \Xi \vdash a_2 : \cat C}
-  {\Gamma \pipe \Xi \vdash \punitinXfromYtoZ {\cat C} {a_1} {a_2}}
+  {\Gamma \pipe \Xi \vdash \punitinXfromYtoZ {\cat C} {a_1} {a_2} \isaSet}
   \quad
@@
     {\Gamma\pipe\beta: \cat D \vdash a : \cat C}
-    {\Gamma \pipe\cdot \vdash \punitrefl a : \punitinXfromYtoZ {\cat C} a a}
+    {\Gamma \pipe \beta : \cat D \vdash \punitrefl a : \punitinXfromYtoZ {\cat C} a a}
     \quad
@@
     \inferrule
-    {\Gamma\pipe \alpha:\cat C\vdash t : P[\alpha/\alpha^-;\alpha/\alpha^+]\\\\
+    {\Gamma\pipe \alpha:\cat C; \beta:\cat C \vdash P \isaSet \\\\
+     \Gamma\pipe \alpha:\cat C\vdash t : P[\alpha/\alpha;\alpha/\beta]\\\\
      \Gamma\pipe\Phi \vdash s : \punitinXfromYtoZ {\cat C} A B}
-    {\Gamma \vdash \punitelimkontZatABC {\alpha. t} A s B : P[A/\alpha^-;B/\alpha^+]}
+    {\Gamma \pipe \Phi \vdash \punitelimkontZatABC {\alpha. t} A s B : P[A/\alpha;B/\beta]}
     \\
@@
      \Gamma \pipe \beta:D;d^+\Xi \vdash Q \isaSet}
-    {\Gamma \pipe \Xi \vdash \tensorexistsXwithYandZ {\beta:D} P Q}
+    {\Gamma \pipe \Xi \vdash \tensorexistsXwithYandZ {\beta:D} P Q \isaSet}
     \quad
@@
   \inferrule
-  {d^+\Xi \isSmall\\\\ \Gamma \pipe d^+\Xi; \alpha : \cat C \vdash  R \\\\ \Gamma \pipe  d^-\Xi; \alpha : \cat C \vdash P}
-  {\Gamma \pipe \Xi \vdash \homrallXYtoZ {\alpha:\cat C} R P}
+  {d^+\Xi \isSmall\\\\ \Gamma \pipe d^+\Xi; \alpha : \cat C \vdash  R
+    \isaSet \\\\ \Gamma \pipe  d^-\Xi; \alpha : \cat C \vdash P \isaSet}
+  {\Gamma \pipe \Xi \vdash \homrallXYtoZ {\alpha:\cat C} R P \isaSet}
   \quad
@@
     {d^-\Xi \isSmall\\\\
-      \Gamma \pipe \alpha:\cat C; d^-\Xi\vdash R \\\\ \Gamma \pipe \alpha : \cat C; d^+\Xi\vdash P}
-    {\Gamma \pipe \Xi \vdash \homlallXYtoZ {\alpha: \cat C} R P}
+      \Gamma \pipe \alpha:\cat C; d^-\Xi\vdash R \isaSet \\\\ \Gamma
+      \pipe \alpha : \cat C; d^+\Xi\vdash P \isaSet }
+    {\Gamma \pipe \Xi \vdash \homlallXYtoZ {\alpha: \cat C} R P \isaSet}
   \quad
@@
     \inferrule
-    {}
+    {~}
     {\Gamma\vdash \smallCats}
@@
     \inferrule
-    {}
+    {~}
     {\Gamma\vdash \Cats}
-
-    FIXME missing intro/elim rules for smallcat and cat
 
@@
     {\Gamma \vdash \pendallXdotY {\alpha:{\cat C}} P \isaTy}
+
+    \inferrule
+    {\Gamma \vdash \cat C \isSmall}
+    {\Gamma \vdash \quoth {\cat C} : \smallCats}
+
+    \inferrule
+    {\Gamma \vdash M : \smallCats}
+    {\Gamma \vdash \unquoth M \isSmall}
+
+    \inferrule
+    {\Gamma \vdash \cat C \isaCat}
+    {\Gamma \vdash \quoth {\cat C} : \Cat}
+
+    \inferrule
+    {\Gamma \vdash M : \Cat}
+    {\Gamma \vdash \unquoth M \isaCat}
 

### 2022-10-10 paper/paper.tex
edit type section

@@
 Finally, we briefly describe the connectives for the ``meta-logic'',
-which extends Martin-L\"of type theory with $Pi$/$Sigma$ and extensional
+which extends Martin-L\"of type theory with $\Pi$/$\Sigma$ and extensional
 identity types (with their standard rules)~\cite{emltt}. We use
@@
 %
-The rule for the types of small categories and (large) categories are very
-similar: any definable category defines an element of type $\Cats$,
-and any element of that type can be reflected back into a
-category. The only difference for $\smallCats$ is that the categories
-involved additionally satisfy $\cat C \isSmall$. Again we elide the
-$\beta\eta$ principles, which state that the introduction and
-elimination rules are a bijection between judgement forms.  
+The rule for the types of small categories and (large) categories are
+very similar: any definable category defines an element of type $\Cats$,
+and any element of that type can be reflected back into a category. The
+only difference for $\smallCats$ is that the categories involved
+additionally satisfy $\cat C \isSmall$. Again we elide the $\beta\eta$
+principles, which state that $\quoth{-}$ and $\unquoth{-}$ are mutually
+inverse.
 %
@@
 The introduction and elimination forms are those for unary and binary
-function types respectively.
-%
+function types respectively, where metalanguage terms of type
+$\varr{\cat C}{\cat D}$ can be used to construct an object/functor,
+while metalanguage terms of type $\harr{\cat C}{\cat D}$ can be used to
+construct a set/profunctor.
+
 Finally we include a type $\pendallXdotY {\alpha:\cat C} P$ which we
@@
 %
-More generally this is modeled as an end, and matching our notation
-for hom types we use a universal quantifier to notate this.
+More generally this is modeled as an end, and we notate it with a
+universal quantifier (just as we do for the quantifiers in left/right
+hom types).  Syntactically, $\pendallXdotY{\alpha}{P}$ is a
+meta-language type that represents elements/natural transformations with
+exactly one free variable.
 
 \begin{figure}
-  \begin{mathpar}
+  \[
+  \begin{array}{l}
     \inferrule
@@
     {\Gamma\vdash \smallCats}
-
+    \quad
+    \inferrule
+    {\Gamma \vdash \cat C \isSmall}
+    {\Gamma \vdash \quoth {\cat C} : \smallCats}
+    \quad
+    \inferrule
+    {\Gamma \vdash M : \smallCats}
+    {\Gamma \vdash \unquoth M \isSmall}
+    \\\\
     \inferrule
@@
     {\Gamma\vdash \Cats}
+    \quad
+    \inferrule
+    {\Gamma \vdash \cat C \isaCat}
+    {\Gamma \vdash \quoth {\cat C} : \Cat}
+    \quad
+    \inferrule
+    {\Gamma \vdash M : \Cat}
+    {\Gamma \vdash \unquoth M \isaCat}
+    \\ \\
 
@@
     {\Gamma\pipe\cdot \vdash \varr{\cat C}{\cat D} \isaTy}
-
-    \inferrule
-    {\Gamma \vdash \cat C \isaCat \and\Gamma \vdash \cat D \isaCat}
-    {\Gamma\pipe\cdot \vdash \harr{\cat C}{\cat D} \isaTy}
-
-    \inferrule
-    {\Gamma \pipe \alpha:\cat C\vdash P \isaSet}
-    {\Gamma \vdash \pendallXdotY {\alpha:{\cat C}} P \isaTy}
-
-    \inferrule
-    {\Gamma \vdash \cat C \isSmall}
-    {\Gamma \vdash \quoth {\cat C} : \smallCats}
-
-    \inferrule
-    {\Gamma \vdash M : \smallCats}
-    {\Gamma \vdash \unquoth M \isSmall}
-
-    \inferrule
-    {\Gamma \vdash \cat C \isaCat}
-    {\Gamma \vdash \quoth {\cat C} : \Cat}
-
-    \inferrule
-    {\Gamma \vdash M : \Cat}
-    {\Gamma \vdash \unquoth M \isaCat}
-
+    \quad
     \inferrule
@@
     {\Gamma \vdash \lambda \alpha:\cat C. A : \varr{\cat C}{\cat D}}
-
+    \quad
     \inferrule
@@
     {\Gamma\pipe \alpha:\cat C \vdash M A : \cat E}
-
+    \\ \\
+
+    \inferrule
+    {\Gamma \vdash \cat C \isaCat \and\Gamma \vdash \cat D \isaCat}
+    {\Gamma\pipe\cdot \vdash \harr{\cat C}{\cat D} \isaTy}
+    \quad
     \inferrule
@@
     {\Gamma \vdash \lambda \alpha:\cat C;\beta:\cat D. R : \harr{\cat C}{\cat D}}
-
-    \inferrule
-    {\Gamma \vdash M : \harr{\cat C}{\cat D}\and
-     \Gamma \pipe d^- \Xi \vdash A : \cat C\and
+    \quad
+    \inferrule
+    {\Gamma \vdash M : \harr{\cat C}{\cat D}\\\\
+     \Gamma \pipe d^- \Xi \vdash A : \cat C\\\\
      \Gamma \pipe d^+ \Xi \vdash B : \cat C}
     {\Gamma \pipe \Xi \vdash M A; B \isaSet}
-
+    \\ \\
+
+    \inferrule
+    {\Gamma \pipe \alpha:\cat C\vdash P \isaSet}
+    {\Gamma \vdash \pendallXdotY {\alpha:{\cat C}} P \isaTy}
+    \quad
     \inferrule
@@
     {\Gamma \vdash \pendlambdaXdotY \alpha t : \pendallXdotY \alpha P}
-
+    \quad
     \inferrule
@@
     {\Gamma \pipe \beta:\cat D \vdash \pendappXtoY M a : P[a/\alpha]}
-  \end{mathpar}
+  \end{array}
+  \]
   \caption{Type Connectives}

### 2022-10-10 paper/paper.tex
new draft of examples section, more emphasis on showing the syntax

@@
 \newcommand{\vett}{VETT}
-\newcommand{\ohol}{OHol}
+\newcommand{\ohol}{\vett}
 \newcommand{\sem}[1]{\llbracket{#1}\rrbracket}
@@
 \newcommand{\ran}[1]{\textrm{Ran}_{#1}\,}
+\newcommand{\weightedLimitDW}[2]{\textrm{lim}^{#2}{#1}\,}
 
@@
 
-\section{Synthetic Category Theory in VETT}
-
-To demonstrate that \ohol{} provides a setting in which to carry out
-synthetic category theory, we demonstrate some basic definitions and
-theorems in category theory in \ohol{}. We emphasize throughout the
-interpretations as tautologies in higher order-logic, as well as the
-relation to ordered linear logic.
-
-First, we fix some terminology: TODO, isomorphism of profunctors, anything else we'll need?
-%% \begin{definition}{Isomorphism of Profunctors}
-
-%% \end{definition}
-
-We can witness the fact that all constructions are (pro-)functorial by
-using the hom type $\punitinXfromYtoZ {\cat C}{\alpha}{\alpha'}$.
-\begin{lemma}
+\section{Formal Category Theory in VETT}
+
+To demonstrate what formal category theory in VETT looks like, we
+demonstrate some basic definitions and theorems. We make no claims to
+the originality of the results of this section, as it is well known
+that much category theory can be formalized in virtual equipments, we
+only show these to demonstrate that the VETT syntax does give a more
+familiar syntax to these constructions, while still avoiding the need
+for explicit naturality and functoriality side conditions.
+
+First, we using the elimination for the unit set, we can see that all
+constructions are (pro-)functorial:
+\begin{construction}
+  For any small category $\cat C$, we can construct natural elements
   \begin{enumerate}
@@
   \item Functoriality: for any $F : \varr {\cat C}{\cat D}$, $\homunary {\alpha_1} {\punitinXfromYtoZ {}{\alpha_1}{\alpha_2}} {\alpha_2} {\punitinXfromYtoZ {} {F(\alpha_1)}{F(\alpha_2)}}$.
-   \item Profunctoriality: for any $R : \harr {\cat C}{\cat D}$ $\homtrinary {\alpha_1} {\punitinXfromYtoZ{} {\alpha_1}{\alpha_2}} {\alpha_2} {R{\alpha_2}{\beta_2}} {\beta_2} {\punitinXfromYtoZ{} {\beta_2} {\beta_1}} {\beta_1} {R\alpha_1\beta_1}$
+   \item Profunctoriality: for any $R : \harr {\cat C}{\cat D}$ if
+     $\cat D$ is small then $\homtrinary {\alpha_1}
+     {\punitinXfromYtoZ{} {\alpha_1}{\alpha_2}} {\alpha_2}
+     {R{\alpha_2}{\beta_2}} {\beta_2} {\punitinXfromYtoZ{} {\beta_2}
+       {\beta_1}} {\beta_1} {R\alpha_1\beta_1}$
   \end{enumerate}
-\end{lemma}
+\end{construction}
 Reflexivity and transitivity generalize the corresponding properties
@@
 relations respect equality.
-
+%
 Naturality is more complex to state, and it is a statement about the
@@
     \alpha}$, by composing with profunctoriality, we can construct
-  terms
-  \[ \alpha_1, f : \punitinXfromYtoZ {\cat C} {\alpha_1} {\alpha_2}, \alpha_2 \vdash \textrm{lcomp}(f,\pendappXtoY t {\alpha_2}) \quad \textrm{and} \quad \textrm{rcomp}(\pendappXtoY t {\alpha_1}, f) : \harrapp R {\alpha_1}{\alpha_2}\]
-  These are both equal to $\punitelimtoYwithkontZ f {t}$.
+  terms $\alpha_1, f : \punitinXfromYtoZ {\cat C} {\alpha_1}
+  {\alpha_2}, \alpha_2 \vdash \textrm{lcomp}(f,\pendappXtoY t
+  {\alpha_2})$ and $\textrm{rcomp}(\pendappXtoY t {\alpha_1}, f) :
+  \harrapp R {\alpha_1}{\alpha_2}$ that are both equal to
+  $\punitelimtoYwithkontZ f {t}$.
 \end{lemma}
@@
 \end{lemma}
-
-Next, we have the ``Fubini'' theorems. These correspond to quantifier
-re-ordering and associativity of conjunction and implication in
-higher-order logic.
-\emph{Fubini} theorems about interchange of (co)-ends.
+The proofs both follow from the unit elimination rule, and we can view
+the unit elimination rule as essentially making the Yoneda lemma a
+primitive axiom, as the two cases of showing (1) is an isomorphism are
+precisely the $\beta$ and $\eta$ rules for the unit.
+
+Next, we have the ``Fubini'' theorems, which are analogous to simple
+theorems relating tensor and hom types in ordered logic, and whose proofs are 
 \begin{lemma}[{Fubini}]
@@
 \end{lemma}
-Note that the last item here is different from the others in that it
-is an equivalence of sets rather than profunctors.
-
-Next, we turn to basic universal properties. While there are many
-equivalent definitions of a universal property in category theory, the
-one most natural in \ohol{} is the notion of a functor representing a
-profunctor, a mild generalization of the more common notion of
-representable presheaf. We say a functor $F : \varr {\cat C}{\cat D}$
-left represents a profunctor $R : \harr {\cat C}{\cat D}$ when
-$\punitinXfromYtoZ {}{\beta}{F \alpha} \cong \harrapp R \alpha
-\beta$. This is an ordered generalization of the notion of a relation
-$R$ being the \emph{graph} of a function $F$. Due to the asymmetry,
-there is a dual notion of a functor $G : \varr {\cat D}{\cat C}$
-right-representing a profunctor $\punitinXfromYtoZ {\alpha}{G\beta}
-\cong \harrapp R \alpha \beta$. Then $F$ and $G$ are adjoint functors
-when $F$ left represents a profunctor that $G$ right-represents. This
-generalizes the formulation that a function $g$ is an inverse to $f$
-if its corresponding relation is the converse of the relation
-corresponding to $f$.
-
-The definition we gave for adjoint functors is the one in terms of
-``hom-isomorphism''. The equivalent in terms of unit and co-unit is a
-simple corollary of the Yoneda lemma, because for instance the unit
-can be extracted from the transformation $\homunary \alpha
-{\punitinXfromYtoZ {}{F \alpha}{\beta}} \beta {\punitinXfromYtoZ
-  {}{\alpha}{G\beta}}$. The formulation in terms of unit and co-unit
-generalizes the formulation of inverses that each composition is equal
-to the identity.
-%% \begin{lemma}[Formulations of Representability]
-%%   The following are equivalent
-%%   \begin{enumerate}
-%%   \item An isomorphism between $\punitinXfromYtoZ {}{F \alpha}{\beta}$ and $\harrapp R \alpha \beta$
-%%   \item A term $\pendallXdotY \alpha {\harrapp R \alpha {F \alpha}}$
-%%     such that the induced morphism from $\punitinXfromYtoZ {}{F
-%%       \alpha}{\beta}$ to $\harrapp R \alpha \beta$ is an isomorphism.
-%%   \end{enumerate}
-%% \end{lemma}
-
-Next, we turn to limits, or more generally, Kan extensions.
-%
-Given functors $J : \varr {\cat C}{\cat C'}$ and $F : \varr{\cat
-  C}{\cat D}$, a right Kan extension is functor $\ran J F : \varr{\cat
-  C'}{\cat D}$ such that the composition $\ran J F \circ J$ has a
-universal natural transformation to $F$.
+\begin{proof}
+  Straightforward, for instance the forward direction of (1) is given
+  by $\lambdaunary \alpha {x} \delta {\tensorelimWkontZ x
+    {p.\beta.y. \tensorelimWkontZ y
+      {q.\gamma.r. \tensorintroatXwithYandZ {\gamma}
+        {\tensorintroatXwithYandZ \beta p q} {r}}}}$
+\end{proof}
+
+Next, we turn to basic universal properties.
 \begin{lemma}
-  A right Kan extension of $F$ along $J$ is a functor $\ran J F$ that
-  right-represents the profunctor $\homrallXYtoZ {\alpha} {\punitinXfromYtoZ {} {\alpha'}{J \alpha}}{\punitinXfromYtoZ {} \beta {F \alpha}}$
+  Let $R : \varr {\cat D}{\cat C}$ and $L : \varr {\cat
+    C}{\cat D}$. Then the following data are in bijection:
+  \begin{enumerate}
+  \item An isomorphism of profunctors $(\punitinXfromYtoZ {\cat D} {L \alpha}{\beta}) \cong (\punitinXfromYtoZ {\cat C} \alpha {R \beta})$
+  \item A unit $\eta : \pendallXdotY \alpha {\punitinXfromYtoZ {\cat C} \alpha
+    {R(L\alpha)}}$ and co-unit $\varepsilon: \pendallXdotY \beta
+    {\punitinXfromYtoZ {\cat D} {L(R(\beta))}{\beta}}$ satisfying triangle
+    identities.
+  \end{enumerate}
 \end{lemma}
-Viewing this in higher-order logic, this says that $\beta = (\ran J
-F)\alpha'$ if and only if $\forall \alpha. \alpha' = J \alpha
-\Rightarrow \beta = F \alpha$. If $J$ is surjective this says that
-$\ran J F$ composed with $J$ is equal to $F$, but if $J$ is not
-surjective then this is impossible to satisfy for ordinary
-functions. Kan extensions generalize limits in that the limit of a
-diagram $F : \varr {\cat C}{\cat D}$ is the same as a right Kan
-extension along the unique $! : \varr{\cat C}{1}$.
-
-Then we can prove the famous theorem that right adjoints preserve
-right Kan extensions, and so in particular, limits.
+\begin{proof}
+  Given the forward homomorphism $\text{lr}$, we can construct $\eta =
+  \pendlambdaXdotY \alpha \appunary {\text{lr}} \alpha {\punitrefl
+    \alpha} {L \alpha}$ and given the unit we can reconstruct the
+  forward homomorphism using
+  $\text{comp}$ and $\text{fctor}$ as defined in
+  Construction\ref{construction:synthetic-composition} as
+  $\appbinary {\text{comp}} \alpha
+  {\pendappXtoY \eta \alpha} {R(L\alpha)} {(\appunary
+    {\text{fctor}(R)}{L \alpha} {f} {\beta})} {R \beta}$.
+\end{proof}
+
+Next, we turn to weighted limits, which generalize ordinary limits and
+Kan extensions.
+\begin{definition}
+  For a functor $D : \varr {\cat J}{\cat C}$ and a profunctor $W :
+  \harr{\cat K}{\cat J}$, the limit of $D$ weighted by $W$ is (if it
+  exists) a functor $\weightedLimitDW D W : \varr{\cat K}{\cat C}$
+  with an isomorphism $\punitinXfromYtoZ {\cat C} {\alpha} {(\weightedLimitDW D W) k} \cong \homrallXYtoZ j {W k j} {(\punitinXfromYtoZ {\cat C} \alpha {D j})}$
+\end{definition}
+This generalizes the usual definition that a morphism into a limit is
+a cone over the diagram $(\punitinXfromYtoZ {\cat C} \alpha {D j})$ to
+be parameterized by a weight $W k j$.
+Then we can prove the famous theorem that right adjoints preserve (weighted) limits:
 \begin{theorem}
-  If $\ran J F$ is a right Kan extension and $G : \varr{\cat D}{\cat
-    E}$ has a left adjoint $L$, then $G \circ (\ran J F)$ is the right
-  Kan extension of $G \circ F$ along $J$.
+  If $\weightedLimitDW D W$ exists and is a limit and $R : \varr{\cat
+    C}{\cat C'}$ has a left adjoint $L$, then $\lambda
+  \kappa. R((\weightedLimitDW D W) \kappa)$ is the limit of $\lambda
+  j. R(D j)$ weighted by $W$.
 \end{theorem}
@@
   \begin{align*}
-    \punitinXfromYtoZ {} {\gamma} {G (\ran J F \alpha')}
-    &\cong \punitinXfromYtoZ {}{L \gamma} {\ran J F \alpha'}\\
-    &\cong \homrallXYtoZ {\alpha}{(\punitinXfromYtoZ {}{\alpha'}{J\alpha})} {\punitinXfromYtoZ {}{L \gamma}{F \alpha}}\\
-    &\cong \homrallXYtoZ {\alpha}{(\punitinXfromYtoZ {}{\alpha'}{J\alpha})} {\punitinXfromYtoZ {}{\gamma}{G (F \alpha)}}
+    \punitinXfromYtoZ {} {\gamma} {R ((\weightedLimitDW D W) \kappa)}
+    \cong \punitinXfromYtoZ {}{L \gamma} {(\weightedLimitDW D W) \kappa}
+    \cong \homrallXYtoZ {j}{W k j} {\punitinXfromYtoZ {}{L \gamma}{D j}}
+    \cong \homrallXYtoZ {j}{W k j} {\punitinXfromYtoZ {}{\gamma}{R(D j)}}\\
   \end{align*}  
 \end{proof}
-
-\subsection{Ordered Theories}
-
-Just as we can define theories in first-order logic or higher-order
-logic by assuming base sorts, function and relation symbols and
-axioms, we can define ordered, or categorical theories in \ohol{}
-easily by using the two-level type theory structure.
-%
-We can define an \emph{ordered theory} to be any context $\Gamma$ and
-a translation of theories to be a substitution.
-%
-Then we can define theories for cartesian categories, monoidal
-categories, bicartesian categories, strong monads, (strong)
-adjunctions.
-%
-However the utility for applications to programming language semantics
-is limited due to the lack of opposite categories.
-
-\subsection{Duality Principle}
-One common source of category theoretic arguments is the duality
-principle: since the opposite of any category is also a category, any
-statement that quantifies over all categories is equally true of their
-opposites. Furthermore in a formal language of categories, any
-statement can be dualized to a statement where domains and codomains
-are swapped and composition is flipped.
-%
-A similar duality principle holds in SCT.
-%
-The basic idea is to reverse the orientation of all \emph{set}
-expressions by swapping which variable is covariant and which is
-contravariant.
-%
-This has the effect of reversing the order of the context $\Phi$.
-%
-We can define this as a translation $\cdot^* : SCT \to SCT$ defined on
-syntactic forms of SCT in Figure~\ref{fig:duality}.
-\begin{figure}
-    \begin{align*}
-      (\harr {\cat C} {\cat D})^* &= \harr {\cat D^*} {\cat C^*}\\
-      (\alpha:\cat C)^* &= \alpha: \cat C^*\\
-      (\alpha:\cat C; \beta:\cat D)^* &= \beta:\cat D^*; \alpha:\cat C^*\\
-      (\Phi,x:R,\Psi)^* &= \Psi^*,x:R^*, \Phi^*\\
-      (\homrallXYtoZ \alpha P Q)^* &= \homlallXYtoZ \alpha {P^*} {Q^*}\\
-      (\homlallXYtoZ \alpha P Q)^* &= \homrallXYtoZ \alpha {P^*} {Q^*}\\
-      (\homrlambdaXatYdotZ x \alpha s)^* &= \homllambdaXatYdotZ x \alpha s^*\\
-      (\homrappXtoYatZ s t a)^* &= \homlappXtoYatZ {s^*} {t^*} {a^*}\\
-      (\tensorexistsXwithYandZ \beta P Q)^* &=
-      \tensorexistsXwithYandZ \beta {Q^*}{P^*}\\
-      (\tensorintroatXwithYandZ b s t)^* &= \tensorintroatXwithYandZ {b^*} {t^*} {s^*}\\
-      (\punitinXfromYtoZ {\cat C} {a_1}{a_2})^* &= \punitinXfromYtoZ {\cat C^*} {a_2^*}{a_1^*}
-    \end{align*}
-  \caption{Duality Involution}
-  \label{fig:duality}
-\end{figure}
+This is a high level proof in terms of isomorphisms that may be
+written in VETT. The first two steps are the instantiation of
+assumptions (adjointness, weighted limits). The last step uses the
+fact that a natural isomorphisms lift to natural isomorphism of homs
+of profunctors. We show the construction of this isomorphism as follows, giving an example that naturality is simply 
+\begin{construction}
+  For any $\phi: \homunary \alpha {R' \alpha \beta} \beta {R \alpha
+    \beta}$ and $\psi : \homunary \gamma {S \gamma \beta} \beta {S'
+    \gamma \beta}$ we can construct a natural transformation $\phi
+  \triangleright \psi : \hombinary \gamma {(\homrallXYtoZ \beta {R
+      \alpha \beta}{S \gamma \beta})} {\alpha} {R'\alpha\beta} \beta {S'\gamma\beta}$ as
+  $\lambdabinary {\gamma} {f} {\alpha} r \beta {\appunary \psi \gamma {(\homrappXtoYatZ f {(\appunary \phi \alpha r \beta)} \beta)} \beta}$.
+Furthermore if $\phi$ and $\psi$ have inverses, then $\phi^{-1} \triangleright \psi^{-1}$ is the inverse of $\phi \triangleright \psi$.
+\end{construction}
+
+%% \subsection{Ordered Theories}
+
+%% Just as we can define theories in first-order logic or higher-order
+%% logic by assuming base sorts, function and relation symbols and
+%% axioms, we can define ordered, or categorical theories in \ohol{}
+%% easily by using the two-level type theory structure.
+%% %
+%% We can define an \emph{ordered theory} to be any context $\Gamma$ and
+%% a translation of theories to be a substitution.
+%% %
+%% Then we can define theories for cartesian categories, monoidal
+%% categories, bicartesian categories, strong monads, (strong)
+%% adjunctions.
+%% %
+%% However the utility for applications to programming language semantics
+%% is limited due to the lack of opposite categories.
+
+%% \subsection{Duality Principle}
+%% One common source of category theoretic arguments is the duality
+%% principle: since the opposite of any category is also a category, any
+%% statement that quantifies over all categories is equally true of their
+%% opposites. Furthermore in a formal language of categories, any
+%% statement can be dualized to a statement where domains and codomains
+%% are swapped and composition is flipped.
+%% %
+%% A similar duality principle holds in SCT.
+%% %
+%% The basic idea is to reverse the orientation of all \emph{set}
+%% expressions by swapping which variable is covariant and which is
+%% contravariant.
+%% %
+%% This has the effect of reversing the order of the context $\Phi$.
+%% %
+%% We can define this as a translation $\cdot^* : SCT \to SCT$ defined on
+%% syntactic forms of SCT in Figure~\ref{fig:duality}.
+%% \begin{figure}
+%%     \begin{align*}
+%%       (\harr {\cat C} {\cat D})^* &= \harr {\cat D^*} {\cat C^*}\\
+%%       (\alpha:\cat C)^* &= \alpha: \cat C^*\\
+%%       (\alpha:\cat C; \beta:\cat D)^* &= \beta:\cat D^*; \alpha:\cat C^*\\
+%%       (\Phi,x:R,\Psi)^* &= \Psi^*,x:R^*, \Phi^*\\
+%%       (\homrallXYtoZ \alpha P Q)^* &= \homlallXYtoZ \alpha {P^*} {Q^*}\\
+%%       (\homlallXYtoZ \alpha P Q)^* &= \homrallXYtoZ \alpha {P^*} {Q^*}\\
+%%       (\homrlambdaXatYdotZ x \alpha s)^* &= \homllambdaXatYdotZ x \alpha s^*\\
+%%       (\homrappXtoYatZ s t a)^* &= \homlappXtoYatZ {s^*} {t^*} {a^*}\\
+%%       (\tensorexistsXwithYandZ \beta P Q)^* &=
+%%       \tensorexistsXwithYandZ \beta {Q^*}{P^*}\\
+%%       (\tensorintroatXwithYandZ b s t)^* &= \tensorintroatXwithYandZ {b^*} {t^*} {s^*}\\
+%%       (\punitinXfromYtoZ {\cat C} {a_1}{a_2})^* &= \punitinXfromYtoZ {\cat C^*} {a_2^*}{a_1^*}
+%%     \end{align*}
+%%   \caption{Duality Involution}
+%%   \label{fig:duality}
+%% \end{figure}
 

### 2022-10-11 paper/paper.tex
get the bib compiling
pass over the intro (not any shorter uh oh)

@@
 \usepackage{cleveref}
-\usepackage{natbib}
 
@@
 
+\newcommand{\citet}[1]{\cite{#1}}
+\newcommand{\citep}[1]{(\cite{#1})}
+
 \title{A Constructive Ordered Predicate Logic for Category Theory}
@@
 
-Category theory is a branch of mathematics that gives a systematic
-theory of typed, ``higher-dimensional'' algebraic structures.
+Category theory is a branch of mathematics that studies
+higher-dimensional typed algebraic structures .
 %
@@
 categorical structures were common in logic and computer science, and
-that formal systems like logics, type theories and programming
-languages can be soundly and completely modeled
-by notions of structured categories\cite{lawvere,goguen,lambek,moggi}, a
-phenomenon sometimes referred to as the Curry-Howard-Lambek
-nncorrespondence \cite{curry-howard-lambek}.
+formal systems like logics, type theories and programming languages
+typically have sound and complete models given by notions of
+structured categories\cite{lawvere,goguen,lambek,moggi}, a phenomenon
+sometimes referred to as the Curry-Howard-Lambek nncorrespondence
+\cite{curry-howard-lambek}.
 %
@@
 In a traditional set-theoretic formulation, notions such as adjoint
-functors and Kan extensions produce a proliferation of ``naturality''
-and ``functoriality'' side-conditions that must be discharged.
+functors and limits produce a proliferation of ``naturality'' and
+``functoriality'' side-conditions that must be discharged.
 %
@@
 programming language applications: domain-, metric- and
-step-index-enriched categories have been used to model general
-recursion and internal categories have been used to model
+step-index-enriched categories have been used to model recursive
+programming languages and internal categories have been used to model
 parametricity and gradual
-typing\cite{order-enriched-cats,topos-of-trees,reflexive-graphs,double-cats-gradual-typing},
-to name a few.
-
-% TODO: Yoneda example
-Novices in category theory are often overwhelmed by the number of
-naturality and functoriality side conditions that abound in otherwise
-simple proofs in category theory. For instance consider the proof of the
-Yoneda lemma, often the first ``big'' proof in an introductory category
-theory text. The Yoneda lemma states that for any category $C$, functor
-$P : C^{o} \to \Set$ (where $C^{o}$ is the ``opposite'' of $C$, i.e. 
-$C$ with the direction of morphisms reversed) and object $A \in C$, the
-set of natural transformations $C(-,A) \to P$ is in bijection with the
-elements $P(A)$. Furthermore, this bijection is natural in both $A$ and
-$P$. In the authors' experience teaching category theory the sheer
-amount of naturality side conditions in this result can be very
-intimidating to learners, while experienced practitioners will observe
-that these naturality conditions are tedious to carry out but
-routine. In the type theory we introduce in this paper, none of these
-naturality conditions need to be explicitly proven. Instead, all terms
-expressible in the type theory are automatically natural by
-construction.
+typing\cite{order-enriched-cats,topos-of-trees,reflexive-graphs,double-cats-gradual-typing}.
+
+For instance, when constructing an adjoint pair of functors between
+two categories, a na\"ive approach would define all of the data of the
+action on objects, action on arrows, prove the functoriality of such
+actions, as well as construct two families of transformations, prove
+they are natural and then finally proving a pair of equalities
+relating compositions of natural transformations.
+%
+Furthermore, when moving into the enriched or internal setting, the
+proofs can have further continuity, monotonicity, or other
+side-conditions.
+%
+Carrying out the proofs explicitly is quite tedious and many newcomers
+to are left with the impression that category theory is full of long,
+but ultimately trivial constructions.
 
@@
 adept analyst writes down a function, they do not expand out the
-$\epsilon\mathord{-}\delta$ definition of continuity and prove that it is
-continuous. Rather, they are aware of certain \emph{syntactic
-principles}, most importantly that composition of continuous functions
-is continuous and that previously defined functions are known to be
-continuous.
-%
-Similar principles apply to category theory itself: a composition of
-functors is functorial, a composition of natural transformations is
-natural, etc.
+$\epsilon\mathord{-}\delta$ definition of continuity for a function
+and proceed from first principles. Rather, they are aware of certain
+\emph{syntactic principles} for defining functions that are continuous
+by construction, most importantly that composition of continuous
+functions is continuous and that previously defined functions are
+known to be continuous.
+%
+Similar principles apply to category theory itself: functors and
+natural transformations are closed under composition and whiskering
+operations, and experienced category theorists rely on these syntactic
+principles to eliminate the tedium of explicit proofs.
 %
@@
 using algebraic structures such as 2-categories, bicategories,
-(virtual) double categories and pro-arrow equipments\cite{}.
-%
-This approach has the advantage of avoiding the need to prove
-the side conditions of naturality or enrichment.
+(virtual) double categories and pro-arrow equipments\cite{TODO}, an
+approach known as \emph{formal category theory}.
+%
+In these structures, rather than defining notions of category, functor
+and natural transformation from first principles, they are axiomatized
+in a manner similar to how a category axiomatizes a notion of space
+and homomorphism.
+%
+One benefit is that the formal approaches also apply to enriched and
+internal settings, so \emph{formal} category theory proofs apply to
+these settings.
 %
@@
 functions.
+
+In this work, we apply the techniques of categorical logic to define a
+more familiar logical syntax for carrying out constructions and proofs
+in formal category theory.
+%
+We call the resulting theory \emph{virtual equipment type theory}
+(VETT) as (hyperdoctrines of) virtual equipments provide a sound and
+complete notion of model for the theory.
+%
+In VETT, functors and natural transformations are defined using
+ordinary term syntax, using familiar notations and $\beta\eta$
+reasoning principles for $\lambda$-functions, bound variables, tuples,
+etc.
+%
+By adhering to a \emph{syntactic discipline}, the logic guarantees
+that all functor terms are automatically functorial, all and natural
+transformation terms are natural.
+%
+More specifically, the syntax for transformations is a kind of
+\emph{indexed, ordered linear} $lambda$ calculus, where the indexing
+ensures that transformations are correctly natural and the ordering
+and linearity ensure that the proofs are valid in a large class of
+enriched and internal categories: for instance enrichment in a
+non-symmetric monoidal category.
+%
+VETT provides an alternative alternative to algebraic and
+string-diagram syntaxes for working with virtual equipments,
+similar to how the lambda calculus provides an alternative to
+categorical combinators and string diagram calculi for cartesian
+closed categories.
 
@@
 %
-The syntax provides an alternative to algebraic and
-string-diagrammatic syntaxes for working with virtual equipments,
-similar to how the lambda calculus provides an alternative to
-categorical combinators and string diagram calculi for cartesian
-closed categories.
-
-\vett{} is an \emph{ordered}, \emph{constructive} variant of
-predicate logic over a unary type theory.
-%
-It is \emph{ordered} in that our notion of predicate is that of a
-binary relation $R(\alpha;\beta)$ where the $\alpha$ is thought of as
-a contravariant dependence and the $\beta$ has a covariant dependence,
-and the rules of the logic distinguish between these two variances.
-%
-Proofs in the predicate logic $R_1, R_2,\ldots \vdash S$ are
-restricted as in ordered linear logic \cite{lambek-calculus, ordered-linear-logic} in
-that each input must be used exactly once and in the order they were
-declared.
-%
-This order discipline ensures that every proof in the logic can be
-interpreted as a \emph{natural} transformation.
-%
-The logic is \emph{constructive} in that we care about the identity of
-proofs: they are natural transformations and there may be more than
-one natural transformation of the same type.
-%
-If we ignore the identity of proofs, ordered predicate logic can be
-viewed as a logic of preordered sets rather than categories.
+
+The syntax of \vett{} is a kind of indexed, ordered linear,
+proof-relevant variant of predicate logic over a unary type theory.
+%
+That is just as a predicate logic has a notion of type, term,
+relation and implication, \vett{} is based on four analogous category
+theoretic concepts: categories, functors, profunctors and natural
+transformations of profunctors.
+%
+So categories are simply treated like types, and functors are simply
+terms whose type is a category and whose free variables have
+categories as types as well. The functor syntax is a unary type
+theory, meaning functor terms always have precisely one free variable
+representing the input object.
+%
+The analogue of a relation is a \emph{profunctor} (defined below),
+which are written as a kind of set with free category variables. In
+keeping with our models, like the restriction to unary functors, we
+restrict to profunctors with two free variables.
+%
+The logic is proof-relevant in that the implications of relations are
+generalized to natural transformations of profunctors, and we use a
+$\lambda$-calculus notation to describe these ``proof terms''.
+%
+We can strenghten this analogy into a formal construction: any
+construction in \vett{} can be erased to a corresponding construction
+or proof in predicate logic, as sets, functions and relations define a
+(somewhat degenerate) virtual equipment.
+%
+Our syntax highlights this analogy in that some of our constructions
+are generalizations of set comprehension or power-sets and we use
+similar notation.
 
@@
 strict: we cannot discuss important concepts such as functor
-categories, opposite categories and arbitrary (co)-ends of
-profunctors. However, the restriction is natural in that it
-corresponds to a well-understood notion of model (virtual equipments)
-and so we leave these extensions to future work.
+categories or opposite categories. However, the restriction is natural
+in that it corresponds to a well-understood notion of model (virtual
+equipments).
 %
 It is well known that a great deal of fundamental results and
-constructions in category theory such as Yoneda and co-Yoneda lemmas,
-adjoint functors and Kan extensions are already expressible in virtual
-equipments, for instance virtual equipments are used for this purpose
-in a text on $\infty$-category theory to avoid explicit manipulation
-of model categories \cite{elements-of-oo-category-theory}.
-%
-We demonstrate some of these standard constructions
-in \ref{sec:synthetic} to show that \vett{} gives a syntax in which
-these formal constructions look like standard set-theoretic arguments.
-
+constructions in category theory, and they have been used for this
+purpose in the formal category theory literature
+\cite{elements-of-oo-category-theory, enriched-indexed-categories}.
+%
 However, we can work around these unary/binary restrictions to some
-extent by viewing the type theory as a domain-specific language embedded
-in a metalanguage.  For example, while we cannot talk about functor
-categories inside the ordered part of the type theory, we can state a
-theorem that quantifies over functors using the meta-language's
-universal quantifier.  To more easily formalize these metatheorems, we
-add a third layer to the type theory: an ordinary dependent type theory
-in the style of Martin-L\"of type theory. Then all of our ordered
-predicate logic judgments will also be indexed by a context from this
-dependent type theory, and the type theory includes types for
-categories, functors, profunctors and natural transformations that allow
-us to formalize theorems the logic is too restrictive to encode. This is
-analogous to the notion of a \emph{2-level type
-  theory}\cite{voevodsky12,altenkirchXX}: we add a dependent type theory
-for reasoning about the terms of our virtual equipment type theory.
-The dependent type theory layer also has an additional purpose: since
-we can quantify over the structures in \vett{}, we can additionally
-encode \emph{signatures} for \vett{} as simply contexts of the
-dependent type theory. This allows for a clean formalization of many
-categorical structures such as monoidal categories, or strong monads.
+extent by viewing the type theory as a domain-specific language
+embedded in a metalanguage.  For example, while we cannot talk about
+functor categories inside the ordered part of the type theory, we can
+state a theorem that quantifies over functors using the
+meta-language's universal quantifier.  To more easily formalize these
+metatheorems, we add a third layer to the type theory: an extensional
+dependent type theory in the style of Martin-L\"of type theory. Then
+all of our ordered predicate logic judgments will also be indexed by a
+context from this dependent type theory, and the type theory includes
+universe types for categories, functors, profunctors and natural
+transformations that allow us to formalize theorems the logic is too
+restrictive to encode, analogous to the use of 2-level type theory in
+homotopy type theory \cite{voevodsky12,altenkirchXX}.
+
+%% This is analogous to the notion of a
+%% \emph{2-level type theory}: we add a
+%% dependent type theory for reasoning about the terms of our virtual
+%% equipment type theory.  The dependent type theory layer also has an
+%% additional purpose: since we can quantify over the structures in
+%% \vett{}, we can additionally encode \emph{signatures} for \vett{} as
+%% simply contexts of the dependent type theory. This allows for a clean
+%% formalization of many categorical structures such as monoidal
+%% categories, or strong monads.
 
@@
 side-conditions.
+%
+Additionally, \vett{} may serve as the basis of a future domain
+specific proof assistant for category theoretic proofs.
 
@@
 
-\item In \cref{sec:examples} we show how to use our syntax for
-  synthetic category theory. Much of this is just rendering well-known
-  2-categorical/double categorical constructions in our type theoretic
-  syntax to demonstrate the ``naive''/synthetic character of proofs
-  and constructions in SCT.
-
-%% \item In \cref{sec:synthetic} we show how common theorems of category
-%%   theory such as the (co)-Yoneda lemma, equivalence of different
-%%   formulations of adjunctions, and preservation of limits can all be
-%%   cleanly reproduced in SCT.
-%% \item In \cref{sec:internal}, we show how we can use SCT to define
-%%   \emph{internal categories}, which allows us to prove additional
-%%   theorems that are not formalizable as synthetic theorems.
-
-\item In \cref{sec:semantics}, we develop some basic model theory for
+\item In \cref{sec:examples} we demonstrate how to use our syntax for
+  formal category theory
+
+\item In \cref{sec:semantics}, we develop some model theory for
   \vett{}, including a sound and complete notion of categorical model
@@
   \item Sets of 2-cells of the following form:
-% https://q.uiver.app/?q=WzAsNixbMCwwLCJDXzAiXSxbMCwyLCJEXzAiXSxbMiwwLCJDX3tufSJdLFsxLDAsIlxcY2RvdHMiXSxbMiwyLCJEXzEiXSxbMSwxLCJcXHBoaSJdLFszLDIsIlJfbiIsMCx7InN0eWxlIjp7ImJvZHkiOnsibmFtZSI6ImJhcnJlZCJ9fX1dLFswLDEsIkYiLDJdLFsyLDQsIkciXSxbMSw0LCJTIiwyLHsic3R5bGUiOnsiYm9keSI6eyJuYW1lIjoiYmFycmVkIn19fV0sWzAsMywiUl8wIiwwLHsic3R5bGUiOnsiYm9keSI6eyJuYW1lIjoiYmFycmVkIn19fV1d
-\[\begin{tikzcd}
+    With appropriate ``multi-categorical'' notions of identity and
+    composition.
+\[\begin{tikzcd}% https://q.uiver.app/?q=WzAsNixbMCwwLCJDXzAiXSxbMCwyLCJEXzAiXSxbMiwwLCJDX3tufSJdLFsxLDAsIlxcY2RvdHMiXSxbMiwyLCJEXzEiXSxbMSwxLCJcXHBoaSJdLFszLDIsIlJfbiIsMCx7InN0eWxlIjp7ImJvZHkiOnsibmFtZSI6ImJhcnJlZCJ9fX1dLFswLDEsIkYiLDJdLFsyLDQsIkciXSxbMSw0LCJTIiwyLHsic3R5bGUiOnsiYm9keSI6eyJuYW1lIjoiYmFycmVkIn19fV0sWzAsMywiUl8wIiwwLHsic3R5bGUiOnsiYm9keSI6eyJuYW1lIjoiYmFycmVkIn19fV1d
 	{C_0} & \cdots & {C_{n}} \\
@@
 \end{tikzcd}\]
-    With appropriate ``multi-categorical'' notions of identity and
-    composition.
-
     We say that the 2-cell $\phi$ has $S$ as codomain, the sequence
@@
 Grants stuff...? The authors would like to thank David Jaz Myers, Emily Riehl, Mike
-Shulman, Dominic Verity,...for helpful feedback on this work
-
+Shulman, Dominic Verity for helpful feedback on this work.
+
+\bibliographystyle{splncs04}
 \bibliography{cats}

### 2022-10-12 paper/paper.tex
missing period
edits to related work

@@
 \item In \cref{sec:examples} we demonstrate how to use our syntax for
-  formal category theory
+  formal category theory.
 
@@
 
-\section{Discussion}
-
-\paragraph{Calculi for formal category theory}
-
-In their book on $\infty$-category theory, Riehl and Verity use the
-formal language of virtual equipments to prove results valid for
-$\infty$-categories without concrete manipulation of model
-categories. They formalize this language using Makkai's FOLDS
-framework of first-order logic with dependent sorts. The approach is
-similar in spirit to our \ohol{}, and has the same models. The main
-difference is that their theory is built on top of first-order logic,
-whereas ours is built on a custom-made variant of higher-order logic
-over a (unary) simple type theory. For this reason, statements in
-their theory are quite complex as they have to be encoded into
-first-order logic, whereas we argue our type-theoretic approach
-formalizes informal arguments much more directly, as shown in
-Section~\ref{sec:examples}.
-%
-An informal syntax for manipulating profunctors involving ends and
-coends, called the coend calculus is closely related, an extension of
-\ohol{} to treat profunctors of many variables of different variances
-should provide a formal treatment for the coend calculus
-\citet{coend-calculus}.
-
-Another approach to formal category theory is to develop a
-\emph{directed dependent type theory}, a type theory where the
-identity type is interpreted as morphisms in a (possibly
-$\infty$-)category. Research on directed dependent type theories is
-ongoing \cite{riehl-shulman,north}. Directed dependent type theories
-should generalize \ohol{} in the same way that dependent type theory
-generalizes higher order logic: internally there should be a type that
-models the category of sets, and terms of this type should model the
-profunctors of \ohol{}.  As a foundation for a domain-specific proof
-assistant for category theory, \ohol{} may provide a foundation for a
-proof assistant in the style of Isabelle/HOL \cite{isabelle-hol},
-whereas directed type theory would provide a foundation for a proof
-assistant in the style of Agda or Coq \cite{agda,coq}.
+\section{Related and Future Work}
+
+We now comapre \vett{} with other calculi for formal category theory.
+
+Riehl and Verity~\cite{rvXXelements} use a formal language of virtual
+equipments to prove results valid for $\infty$-categories without
+concrete manipulation of model categories. They formalize this language
+as a theory in Makkai's framework of first-order logic with dependent
+sorts (FOLDS).  While this previous work has the same models as \vett{},
+we believe that the syntax we propose in this paper formalizes informal
+arguments more directly, as shown in Section~\ref{sec:examples}.  This
+is because FOLDS approach approach is entirely relational, whereas we
+formalize concepts like restriction of a profunctor or composition of
+natural transformations as functional operations (substitution).  In
+particular, this means that our calculus requires only vertically
+degenerate squares (elements/natural transformations) as a
+``user-facing'' notion, with general squares occuring only in the
+admissible substitution operations.
+
+%% The main difference is that their theory is built on top of
+%% first-order logic, whereas ours is built on a custom-made variant of
+%% higher-order logic over a (unary) simple type theory. For this reason,
+%% statements in their theory are quite complex as they have to be encoded
+%% into first-order logic, whereas we argue our type-theoretic approach
+%% %
+
+The coend calculus~\citet{coend-calculus} is an informal syntax for
+manipulating profunctors involving ends and coends; an extension of
+\vett{} to treat profunctors of many variables of different variances
+should would provide a formal treatment of it.
 
@@
 of categories, profunctors and natural transformations rather than the
-virtual double category of categories, functors, profunctors and
-natural transformations. An abstract structure along
-these lines is a \emph{Cartesian bicategory}\footnote{which is not the
-same as a bicategory with products/finite limits, for which a syntax
-was developed \citet{fioresaville}} is given in
-\citet{carboni-walters,carboni-walters-someonelese}. 
-Some preliminary work on a proof system for Cartesian bicateogires was
-presented at \citet{jonasfreycartesianbicategories}. This bears strong
-similarities to our calculus VETT, except that (1) their profunctors
-may have 0, 1 or more covariant or contravariant variables but (2)
-they do not have a term syntax like our natural transformation syntax
-and (3) they do not have functors and restriction of profunctors by
-functors (since this structure is not present in a Cartesian
-bicategory). Semantically, Cartesian bicategories are somewhat less
-general than virtual equipments because in general the set of
-functors between two categories can only be recovered from the set of
-profunctors if the codomain category is Cauchy complete, in which case
-functors are equivalent to adjoint pairs of profunctors.
+virtual double category of categories, functors, profunctors and natural
+transformations. An abstract structure along these lines is a
+\emph{Cartesian
+  bicategory}~\citet{carboni-walters,carboni-walters-someonelese}.  (In
+a terminological collision, \citet{fioresaville} give a syntax for a a
+bicategory with products/finite limits, but this is a different notion.)
+\citet{jonasfreycartesianbicategories} describe some preliminary work on
+a proof system for Cartesian bicateogires.  This calculus is more
+general than \vett{} in that their profunctors may have 0, 1 or more
+covariant or contravariant variables, but they do not have a term syntax
+for the natural transformations/elements, and Cartesian bicategories do
+not have functors and restriction of profunctors by
+functors. Semantically, Cartesian bicategories are somewhat less general
+than virtual equipments because in general the set of functors between
+two categories can only be recovered from the set of profunctors if the
+codomain category is Cauchy complete, in which case functors are
+equivalent to adjoint pairs of profunctors.
+
+Our work in this paper fits broadly into a line of work on
+\emph{directed dependent type theories}, a type theory where the
+identity type is interpreted as morphisms in a (possibly
+$\infty$-)category.  In directed type theories based on a bisimplicial
+model~\cite{riehl-shulman,weinberger,otherweinbegerpaper,licataweaver},
+morphism types are defined using an interval object, like in cubical
+type theory~\cite{bch,cchm,afh,abcfhl}, and universal properties like
+``morphism induction'' are an internally definable property of certain
+types.  Other type theories~\cite{north,ahrensnorth} define morphism
+types via an induction principle, corresponding to the lifting
+properties of certain kinds of fibrations of categories.  While these
+previous works can express some constructions on $\Cat$ that are not
+expressible in \vett{}, because \vett{} is more restricted, it has more
+models, e.g. in internal and enriched categories, so the theorems that
+are provable in \vett{} apply in more settings.
+
+Our use of dependent type theory as a metalanguage, with new judgements
+for the virtual equipment structure dependent on the meta-language
+context, is similar to the indexed type theories developed by
+\citet{isaev}.
+
+%% Directed dependent type theories
+%% should generalize \ohol{} in the same way that dependent type theory
+%% generalizes higher order logic: internally there should be a type that
+%% models the category of sets, and terms of this type should model the
+%% profunctors of \ohol{}.  As a foundation for a domain-specific proof
+%% assistant for category theory, \ohol{} may provide a foundation for a
+%% proof assistant in the style of Isabelle/HOL \cite{isabelle-hol},
+%% whereas directed type theory would provide a foundation for a proof
+%% assistant in the style of Agda or Coq \cite{agda,coq}.
 
@@
 horizontal morphisms are terms of a programming language, and so they
-are not presented as having co-variant and contravariant variables,
-and the equipment structure is modeled by explicit coercions rather
-than by syntactic substitution of vertical arrows in horizontal
-arrows. Another similar calculus is System P \citep{systemP} which is
-an internal language of \emph{reflexive graph categories}, which are
-like double categories without horizontal composition.
+are not presented as relations with co-variant and contravariant
+variables, and the equipment structure is modeled by explicit coercions
+rather than by syntactic substitution of vertical arrows in horizontal
+arrows. Another similar calculus is System P \citep{systemP} which is an
+internal language of \emph{reflexive graph categories}, which are like
+double categories without horizontal composition.
+
+%% DRL note: I would compress the rest of this to a couple of sentences
+%% as the first thing to remove for space.
 

### 2022-10-12 paper/paper.tex
edits to sec 3

@@
 To demonstrate what formal category theory in VETT looks like, we
-demonstrate some basic definitions and theorems. We make no claims to
-the originality of the results of this section, as it is well known
+demonstrate some basic definitions and theorems.  While it is well known
 that much category theory can be formalized in virtual equipments, we
-only show these to demonstrate that the VETT syntax does give a more
+show these examples to demonstrate how the VETT syntax gives a more
 familiar syntax to these constructions, while still avoiding the need
-for explicit naturality and functoriality side conditions.
+for explicit naturality and functoriality side conditions.  We have
+formalized the syntax of \vett{} in Agda, using Agda's rewriting
+mechanism to implement substitution and $\beta$-reduction, and checked
+some of the results in this section (e.g. Lemma~\ref{lem:yoneda} and
+Lemma~\ref{lem:fubini-short} and the maps in
+Lemma~\ref{lem:adjunction}).\footnote{\url{https://github.com/maxsnew/virtual-equipments/blob/master/agda/Examples.agda}}
 
@@
    \item Profunctoriality: for any $R : \harr {\cat C}{\cat D}$ if
-     $\cat D$ is small then $\homtrinary {\alpha_1}
+     $\cat D$ is small then \\ $\homtrinary {\alpha_1}
      {\punitinXfromYtoZ{} {\alpha_1}{\alpha_2}} {\alpha_2}
@@
 \Rightarrow P y$ and $P x$.
-\begin{lemma}
+\begin{lemma} \label{lem:yoneda}
   Let $\alpha^o : \cat C$ and $\pi : \negPresheaf {\cat C}$. Then
@@
 \end{lemma}
-The proofs both follow from the unit elimination rule, and we can view
+The proofs both follow from the unit elimination rule: we can view
 the unit elimination rule as essentially making the Yoneda lemma a
@@
 \begin{lemma}[{Fubini}]
-  \label{lem:fubini}
+  \label{lem:fubini-short}
   The following isomorphisms hold when the corresponding profunctors
@@
 
-Next, we turn to basic universal properties.
-\begin{lemma}
+Next, we can prove that two definitions of an adjunction are equivalent:
+\begin{lemma} \label{lem:adjunction}
   Let $R : \varr {\cat D}{\cat C}$ and $L : \varr {\cat
@@
 fact that a natural isomorphisms lift to natural isomorphism of homs
-of profunctors. We show the construction of this isomorphism as follows, giving an example that naturality is simply 
+of profunctors. We show the construction of this isomorphism as follows,
+giving an example that naturality need not be proved explicitly.
 \begin{construction}

### 2022-10-13 paper/paper.tex
hide the acks for dbl blind
shorten the discussion section

@@
 \vett{} to treat profunctors of many variables of different variances
-should would provide a formal treatment of it.
+may provide a formal treatment of it.
 
@@
 previous works can express some constructions on $\Cat$ that are not
-expressible in \vett{}, because \vett{} is more restricted, it has more
-models, e.g. in internal and enriched categories, so the theorems that
-are provable in \vett{} apply in more settings.
+expressible in \vett{}, because \vett{} is more restricted, \vett{}
+contrariwise has more models, for instance categories enriched in
+non-cartesian monoidal categories, so the theorems that are provable
+in \vett{} apply in more settings.
 
@@
 
-\subsection{Category Theoretic Extensions}
-
-While we have demonstrated that some common constructions in category
-theory can be expressed using the simple combination of homs, tensors
-and units, there are several that are out of reach for SCT due to its
-restrictive judgmental structure. These require extensions to the
-judgmental structure that would move beyond virtual equipments.
-
-First, while we can add products to the calculus, adding in a
-\emph{functor category} ${\cat C}^{\cat D}$ would necessitate adding a
-syntax for functors of multiple arguments, to allow for the lambda rule:
-\[ \inferrule {\alpha:\cat C, \beta:\cat D \vdash e : \cat E} {\alpha:\cat C \vdash \lambda \beta:\cat D. e : \cat E} \]
-This would then require also some extension to profunctors for
-allowing 0 or more covariant/contravariant dependencies as well.
-%
-Semantically, this would correspond to generalizing from virtual
-equipments to some kind of (possibly cartesian) \emph{monoidal closed}
-virtual equipments.
-
-In addition, we could consider extending our semantics with a notion
-of \emph{opposite category}, with the property that a covariant
-dependence of a profunctor on a variable in a category should be
-equivalent to a contravariant dependence on a variable in the opposite
-category.
-%
-This could possibly be done by introducing a modal distinction
-$\alpha:^{+}\cat C$ for covariant dependence and $\alpha:^{-}\cat C$
-for contravariant dependence, with the opposite category being a
-modality that internalizes flipping the variance as a type
-constructor.
-%
-With these two extensions, possibly, the category contexts for
-functors and profunctors could be unified as one notion of a sequence
-of object variables with variances.
-%
-With this unification, it would then be simpler to incorporate a third
-extension: a universe object, corresponding to an object
-$\textrm{Set}$, which would allow us to internalize the profunctor
-judgment in the functor judgment, modeling that in classical category
-theory, profunctors are the same as certain functors to the category
-of Sets.
-%
-While all of these extensions seem plausible, they all move away to
-the simple ordered type theory structure that allowed us to define
-units, tensor and hom profunctors so easily. It is unclear how to
-define these types in this extended structure. As a simple example, if
-profunctors have many variables of mixed variance, rather than exactly
-one of each variance, it is unclear what the appropriate
-generalization of the transformation context $\Phi$ should be.
-
-An orthogonal extension to these would be to consider adding a
-\emph{discretization} modality $\Delta\cat C$ that gives the discrete
-category on the underlying set of objects of $\cat C$. This would give
-us the ability to formalize concepts like ``function on objects'' or
-``unnatural transformation'' without the need to define internal
-categories. This could be modeled by adding a modal distinction
-$\alpha:^{\#}\cat C$ for ``crisp'' dependence on an object of $\cat
-C$, i.e., dependence merely on the object of $\cat C$ with no
-functoriality/naturality conditions. This would then be internalized
-by the discretization modality.
-
-A third approach would be to extend the calculus to model general
-monoidal closed structures and more generally multicategories, which
-provides semantics of modal substructural type theories.
-%
-This can be modeled by extending the ``universe'' $\Set$ to be a
-monoidal closed category, which induces a monoidal closed structure on
-profunctors by the Day convolution product \cite{...}.
-%
-Similar developments along these lines have already been done for
-coend calculus \citep{coend-calculus} and in formalization of
-substructural type theories \citep{atkey-woods}.
+\subsection{Extending VETT}
+
+While many constructions in category theory can be formalized in
+virtual equipments, and therefore \vett{}, others require more
+structure. Future extensions to \vett{} could incorporate functor
+categories by generalizing the unary type theory of functors to
+functors of many variables, in which case ordinary $\lambda$ calculus
+can be used to define functor categories as function types. Similarly,
+ideas from modal type theory/graded linear logic could be used for
+defining opposite categories
+\cite{mike-contravariance-enrichment,semiring-stuf??}.
+
+%% While we have demonstrated that some common constructions in category
+%% theory can be expressed using the simple combination of homs, tensors
+%% and units, there are several that are out of reach for SCT due to its
+%% restrictive judgmental structure. These require extensions to the
+%% judgmental structure that would move beyond virtual equipments.
+
+%% First, while we can add products to the calculus, adding in a
+%% \emph{functor category} ${\cat C}^{\cat D}$ would necessitate adding a
+%% syntax for functors of multiple arguments, to allow for the lambda rule:
+%% \[ \inferrule {\alpha:\cat C, \beta:\cat D \vdash e : \cat E} {\alpha:\cat C \vdash \lambda \beta:\cat D. e : \cat E} \]
+%% This would then require also some extension to profunctors for
+%% allowing 0 or more covariant/contravariant dependencies as well.
+%% %
+%% Semantically, this would correspond to generalizing from virtual
+%% equipments to some kind of (possibly cartesian) \emph{monoidal closed}
+%% virtual equipments.
+
+%% In addition, we could consider extending our semantics with a notion
+%% of \emph{opposite category}, with the property that a covariant
+%% dependence of a profunctor on a variable in a category should be
+%% equivalent to a contravariant dependence on a variable in the opposite
+%% category.
+%% %
+%% This could possibly be done by introducing a modal distinction
+%% $\alpha:^{+}\cat C$ for covariant dependence and $\alpha:^{-}\cat C$
+%% for contravariant dependence, with the opposite category being a
+%% modality that internalizes flipping the variance as a type
+%% constructor.
+%% %
+%% With these two extensions, possibly, the category contexts for
+%% functors and profunctors could be unified as one notion of a sequence
+%% of object variables with variances.
+%% %
+%% With this unification, it would then be simpler to incorporate a third
+%% extension: a universe object, corresponding to an object
+%% $\textrm{Set}$, which would allow us to internalize the profunctor
+%% judgment in the functor judgment, modeling that in classical category
+%% theory, profunctors are the same as certain functors to the category
+%% of Sets.
+%% %
+%% While all of these extensions seem plausible, they all move away to
+%% the simple ordered type theory structure that allowed us to define
+%% units, tensor and hom profunctors so easily. It is unclear how to
+%% define these types in this extended structure. As a simple example, if
+%% profunctors have many variables of mixed variance, rather than exactly
+%% one of each variance, it is unclear what the appropriate
+%% generalization of the transformation context $\Phi$ should be.
+
+%% An orthogonal extension to these would be to consider adding a
+%% \emph{discretization} modality $\Delta\cat C$ that gives the discrete
+%% category on the underlying set of objects of $\cat C$. This would give
+%% us the ability to formalize concepts like ``function on objects'' or
+%% ``unnatural transformation'' without the need to define internal
+%% categories. This could be modeled by adding a modal distinction
+%% $\alpha:^{\#}\cat C$ for ``crisp'' dependence on an object of $\cat
+%% C$, i.e., dependence merely on the object of $\cat C$ with no
+%% functoriality/naturality conditions. This would then be internalized
+%% by the discretization modality.
+
+%% A third approach would be to extend the calculus to model general
+%% monoidal closed structures and more generally multicategories, which
+%% provides semantics of modal substructural type theories.
+%% %
+%% This can be modeled by extending the ``universe'' $\Set$ to be a
+%% monoidal closed category, which induces a monoidal closed structure on
+%% profunctors by the Day convolution product \cite{...}.
+%% %
+%% Similar developments along these lines have already been done for
+%% coend calculus \citep{coend-calculus} and in formalization of
+%% substructural type theories \citep{atkey-woods}.
 
@@
 
-\subsection{Semantics}
-
-Our definition of a double category with displays is suggestive of a
-simple generalization. A double category is a monoid object in the
-double category of $Span(Cat)$, and a double category with displays
-furthermore uses certain diagrams involving profunctors, which also
-form a double category $Prof$. This suggests that the algebraic
-structure of a double category with displays is naturally modeled in
-some kind of \emph{triple category}, modeling the three types of
-morphisms: functors, profunctors and spans of functors.
+%% \subsubsection*{Acknowledgements}
+%% Grants stuff...? The authors would like to thank David Jaz Myers, Emily Riehl, Mike
+%% Shulman, Dominic Verity for helpful feedback on this work.
 
@@
 %% %% Please use bibtex,
-\subsubsection*{Acknowledgements}
-Grants stuff...? The authors would like to thank David Jaz Myers, Emily Riehl, Mike
-Shulman, Dominic Verity for helpful feedback on this work.
 

### 2022-10-13 paper/paper.tex
tighten up intro

--- 2022-10-12 paper/paper.tex
@@
 \begin{abstract}
-  We present a domain-specific type theory for constructions and
-  proofs in category theory. The type theory axiomatizes notions of
-  category, functor, profunctor and a generalized form of natural
-  transformations. The type theory imposes an ordered linear
-  restriction on standard predicate logic, which guarantees that all functions
+  We present a domain-specific type theory for constructions and proofs
+  in category theory. The type theory axiomatizes notions of category,
+  functor, profunctor and a generalized form of natural
+  transformations. The type theory imposes an ordered linear restriction
+  on standard predicate logic, which guarantees that all functions
   between categories are functorial, all relations are profunctorial,
-  and all transformations are natural by construction, with no
-  separate proofs necessary. Important category theoretic proofs such
-  as the Yoneda lemma and Co-yoneda lemma become simple type theoretic
-  proofs about the relationship between unit, tensor and (ordered)
-  function types, and can be seen to be ordered refinements of
-  theorems in predicate logic.
-
-  We develop a sound and complete notion of categorical model for the
-  type theory based on virtual equipments, which are known to model
-  both internal and enriched category theory. This means that while
-  the proofs in our type theory look like standard set-based
-  arguments, the syntactic discipline ensure that all proofs and
-  constructions carry over to enriched and internal settings as well.
+  and all transformations are natural by construction, with no separate
+  proofs necessary. Important category theoretic proofs such as the
+  Yoneda lemma and Co-yoneda lemma become simple type theoretic proofs
+  about the relationship between unit, tensor and (ordered) function
+  types, and can be seen to be ordered refinements of theorems in
+  predicate logic.  The type theory is sound and complete for a
+  categorical model in \emph{virtual equipments}, which model both
+  internal and enriched category theory. While the proofs in our type
+  theory look like standard set-based arguments, the syntactic
+  discipline ensure that all proofs and constructions carry over to
+  enriched and internal settings as well.
 \end{abstract}
@@
 Category theory is a branch of mathematics that studies
-higher-dimensional typed algebraic structures .
+higher-dimensional typed algebraic structures.
 %
 Originally developed for applications to homological algebra, it was
-quickly discovered that (among many other applications) these
-categorical structures were common in logic and computer science, and
-formal systems like logics, type theories and programming languages
-typically have sound and complete models given by notions of
-structured categories\cite{lawvere,goguen,lambek,moggi}, a phenomenon
-sometimes referred to as the Curry-Howard-Lambek nncorrespondence
-\cite{curry-howard-lambek}.
-%
-Examples include the simply typed lambda calculus \cite{lambekscott},
-computational lambda calculus \cite{moggi}, linear logic \cite{girard}
-and dependent type theory\cite{cartmell1986}.
-%
-These seminal works have led to a fruitful field of research known as
-categorical logic, where new type theories are designed based directly
-on category-theoretic semantics.
-%
-The slogan is that the syntax of a type theory should present an
-initial object in its category of models, a category-theoretic
-reformulation and generalization of the logical notion of soundness
-and completeness.
+quickly discovered that categorical structures were common in logic and
+computer science.  Formal systems like logics, type theories and
+programming languages typically have sound and complete models given by
+notions of structured categories,\cite{lawvere,goguen,lambek,moggi}.
+%
+This Curry-Howard-Lambek correspondence~\cite{curry-howard-lambek}
+applies to simply typed lambda calculus~\cite{lambekscott},
+computational lambda calculus~\cite{moggi}, linear logic~\cite{girard}
+dependent type theory\cite{cartmell1986}, and many other type theories
+designed based on category-theoretic semantics.
+%
+The syntax of a type theory should present an initial object in its
+category of models, a category-theoretic reformulation of logical
+soundness and completeness.
 
@@
 category-theoretic techniques in computer science are sometimes
-criticized for being overly complex, and there is difficulty in
-acquisition of competence in category theory by computer scientists
-who do not have a strong background in pure mathemantics.
-%
-While much of this can be explained by lack of, or lack of awareness
-of, accessible material for computer scientists, in some ways category
-theory is objectively complex.
+criticized for being overly complex and hard to learn, especially for
+computer scientists who do not have a strong background in pure
+mathemantics.
+%
+While in some ways this could be remedied by better introductory
+materials, in some ways category theory is objectively complex.
 %
@@
 %
-This is compounded when we move from ordinary category theory to
+For example, when constructing an adjoint pair of
+functors between two categories, a na\"ive approach would define all of
+the data of the action on objects, action on arrows, prove the
+functoriality of such actions, as well as construct two families of
+transformations, prove they are natural and then finally proving a pair
+of equalities relating compositions of natural transformations.
+%
+Carrying out these proofs explicitly is quite tedious and many newcomers
+to are left with the impression that category theory is full of long,
+but ultimately trivial constructions.
+%
+This complexity is compounded when we move from ordinary category theory to
 enriched and internal category theory, where constructions must be
@@
 %
-These generalizations have long been recognized as being important in
-programming language applications: domain-, metric- and
+Moreover, these generalizations are often what is needed for programming
+language applications; for example, domain- and metric- and
 step-index-enriched categories have been used to model recursive
@@
 
-For instance, when constructing an adjoint pair of functors between
-two categories, a na\"ive approach would define all of the data of the
-action on objects, action on arrows, prove the functoriality of such
-actions, as well as construct two families of transformations, prove
-they are natural and then finally proving a pair of equalities
-relating compositions of natural transformations.
-%
-Furthermore, when moving into the enriched or internal setting, the
-proofs can have further continuity, monotonicity, or other
-side-conditions.
-%
-Carrying out the proofs explicitly is quite tedious and many newcomers
-to are left with the impression that category theory is full of long,
-but ultimately trivial constructions.
-
 Fortunately, the tools of category theory itself can be employed to
-simplify these constructions, specifically the tools of \emph{higher}
+simplify this complexity, specifically the tools of \emph{higher}
 category theory.
 %
-Consider the analogous situation in differential calculus. When an
-adept analyst writes down a function, they do not expand out the
-$\epsilon\mathord{-}\delta$ definition of continuity for a function
-and proceed from first principles. Rather, they are aware of certain
-\emph{syntactic principles} for defining functions that are continuous
-by construction, most importantly that composition of continuous
-functions is continuous and that previously defined functions are
-known to be continuous.
+As an analogy in differential calculus, when an adept analyst writes
+down a function, they do not expand out the $\epsilon\mathord{-}\delta$
+definition of continuity for a function and proceed from first
+principles, but rather use certain \emph{syntactic principles} for
+defining functions that are continuous by construction --- e.g. that
+composition of continuous functions is continuous and that previously
+defined functions are known to be continuous.
 %
@@
 %
-One benefit is that the formal approaches also apply to enriched and
-internal settings, so \emph{formal} category theory proofs apply to
-these settings.
-%
-A downside is that these algebraic structures are quite complicated,
-and practitioners typically employ either an algebraic combinator
-syntax or a 2-dimensional diagrammatic language that can be quite
-beautiful and elegant, but is also seems very far removed from the
-traditional formulation of category theory in terms of sets and
-functions.
+The formal approaches also apply to enriched and internal settings, so
+\emph{formal} category theory proofs apply to these settings.
+%
+A downside is that these algebraic structures are quite complicated, and
+practitioners typically employ either an algebraic combinator syntax or
+a 2-dimensional diagrammatic language that can be quite beautiful and
+elegant, but is also somewhat removed from the traditional formulation
+of category theory in terms of sets and functions.
 
@@
 %
-In VETT, functors and natural transformations are defined using
-ordinary term syntax, using familiar notations and $\beta\eta$
-reasoning principles for $\lambda$-functions, bound variables, tuples,
-etc.
+VETT provides syntax for categories, functors, profunctors, and natural
+transformations, which are defined using familiar term syntax and
+$\beta\eta$ reasoning principles for $\lambda$-functions, bound
+variables, tuples, etc.  
 %
@@
 \emph{indexed, ordered linear} $lambda$ calculus, where the indexing
-ensures that transformations are correctly natural and the ordering
-and linearity ensure that the proofs are valid in a large class of
-enriched and internal categories: for instance enrichment in a
-non-symmetric monoidal category.
+ensures that transformations are correctly natural and the ordering and
+linearity ensure that the proofs are valid in a large class of enriched
+and internal categories, such as enrichment in a non-symmetric monoidal
+category.
 %
@@
 
-In this work, we apply the technique of categorical logic to these
-algebraic structures to design a formal logic, which we call
-\emph{virtual equipment type theory}, that can be soundly interpreted
-in virtual equipments.
-%
-In the ``standard model'' of \vett{}, the sorts are interpreted as
-categories, functions as functors, relations as profunctors and
-implications of relations as profunctor homomorphisms, a generalization
-of natural transformations.
-%
-The type theoretic syntax allows us to write all of our constructions
-as if they were ordinary functors and profunctors in set-based
-category theory.
-%
-\vett also can be modeled in an arbitrary virtual equipments, and so the
-constructions are valid in enriched and internal category theory as
-well.
+%% The type theoretic syntax allows us to write all of our 
+%% constructions are written as if they were
+%% ordinary functors and profunctors in set-based category theory.
+
 %%
@@
 
-The syntax of \vett{} is a kind of indexed, ordered linear,
-proof-relevant variant of predicate logic over a unary type theory.
-%
-That is just as a predicate logic has a notion of type, term,
-relation and implication, \vett{} is based on four analogous category
-theoretic concepts: categories, functors, profunctors and natural
-transformations of profunctors.
-%
-So categories are simply treated like types, and functors are simply
-terms whose type is a category and whose free variables have
-categories as types as well. The functor syntax is a unary type
-theory, meaning functor terms always have precisely one free variable
-representing the input object.
-%
-The analogue of a relation is a \emph{profunctor} (defined below),
-which are written as a kind of set with free category variables. In
-keeping with our models, like the restriction to unary functors, we
-restrict to profunctors with two free variables.
+The syntax of \vett{} is an indexed, ordered linear, proof-relevant
+variant of predicate logic over a unary type theory.
+%
+Just as a predicate logic has a notion of type, term, relation and
+implication, \vett{} is based on four analogous category theoretic
+concepts: categories, functors, profunctors and natural transformations
+of profunctors.
+%
+Categories are treated like types, and the unary functors we consider in
+this paper are each represented by a term whose type is a category and
+whose one free variable ranges over a category.
+%
+The analogue of a relation is a \emph{profunctor} (defined below), which
+are written as a kind of set with free category variables.  Like the
+restriction to unary functors, we restrict to profunctors with two free
+variables.
 %
@@
 (somewhat degenerate) virtual equipment.
-%
-Our syntax highlights this analogy in that some of our constructions
-are generalizations of set comprehension or power-sets and we use
-similar notation.
-
-The restriction to unary type theory and binary relations is fairly
-strict: we cannot discuss important concepts such as functor
-categories or opposite categories. However, the restriction is natural
-in that it corresponds to a well-understood notion of model (virtual
-equipments).
-%
-It is well known that a great deal of fundamental results and
-constructions in category theory, and they have been used for this
-purpose in the formal category theory literature
-\cite{elements-of-oo-category-theory, enriched-indexed-categories}.
-%
-However, we can work around these unary/binary restrictions to some
-extent by viewing the type theory as a domain-specific language
-embedded in a metalanguage.  For example, while we cannot talk about
-functor categories inside the ordered part of the type theory, we can
-state a theorem that quantifies over functors using the
-meta-language's universal quantifier.  To more easily formalize these
-metatheorems, we add a third layer to the type theory: an extensional
-dependent type theory in the style of Martin-L\"of type theory. Then
-all of our ordered predicate logic judgments will also be indexed by a
-context from this dependent type theory, and the type theory includes
-universe types for categories, functors, profunctors and natural
-transformations that allow us to formalize theorems the logic is too
-restrictive to encode, analogous to the use of 2-level type theory in
-homotopy type theory \cite{voevodsky12,altenkirchXX}.
+%% %
+%% Our syntax highlights this analogy in that some of our constructions
+%% are generalizations of set comprehension or power-sets and we use
+%% similar notation.
+
+While the restricted syntax developed in this paper does not express
+important concepts such as functor categories or opposite categories,
+the restriction is natural in that it corresponds exactly to a
+well-understood notion of model, virtual equipments, which can express a
+great deal of fundamental results and constructions in category
+theory~\cite{elements-of-oo-category-theory,
+  enriched-indexed-categories}.
+%
+Moveover, we can work around these unary/binary restrictions to some
+extent by viewing the type theory as a domain-specific language embedded
+in a metalanguage.  For example, while we cannot talk about functor
+categories, we can state a theorem that quantifies over functors using
+the meta-language's ``external'' universal quantifier (which does not
+have automatic functoriality/naturality properties).  To support this,
+\vett includes a third layer, an extensional dependent type theory in
+the style of Martin-L\"of type theory. All of our ordered predicate
+logic judgments are also be indexed by a context from this dependent
+type theory, and the type theory includes universe types for categories,
+functors, profunctors and natural transformations.  This allow us to
+formalize theorems the logic is too restrictive to encode, analogous to
+2-level~\cite{voevodsky12,altenkirchXX} or indexed type
+theories~\cite{isaev,linearlf,vakar,krishnaswamibenton}.
 
@@
 %
-Ordinary predicate logic has been used to develop logics in which
-logical relations proofs can be developed such as Abadi-Plotkin logic
-for parametricity and LSLR and Iris for step-indexed logical relations
-proofs \cite{abadi-plotkin,lslr,iris}.
+Ordinary predicate logic is the foundation for logical presentations of
+logical relations, such as Abadi-Plotkin logic for parametricity and
+LSLR and Iris for step-indexed logical relations proofs
+\cite{abadi-plotkin,lslr,iris}.
 %
@@
 %
-Additionally, \vett{} may serve as the basis of a future domain
-specific proof assistant for category theoretic proofs.
+Additionally, \vett{} may serve as the basis of a future domain specific
+proof assistant for category theoretic proofs.  To pilot-test this, we
+have formalized the syntax of \vett{} in Agda 2.6.2.2, using the rewrite
+mechanism to make \vett's substitution and $\beta$-reduction rules
+definitional
+equalities.\footnote{\url{https://github.com/maxsnew/virtual-equipments/blob/master/agda/STC.agda}}
+We have used this lightweight implementation to check a number of
+examples.
 
 \paragraph{Basics of Profunctors.}
-
 While we assume the reader has some background knowledge of category
 theory, we briefly define profunctors, which are not included in many
-introductory texts.  Profunctors arise implicitly in formulations of
-universal properties such as representable functors and adjunctions.
-
-First, a category $\cat C$ has a collection of objects and morphisms
-with identity and composition, and a functor $F : \cat C \to \cat D$ is
-a function on objects and a function on morphisms that respects source
-and target and preserves identity and composition. It will be useful for
-intuition to think of a category as a generalization of a preordered
-set, which has a set of elements and a binary \emph{relation} on its
-objects satisfying reflexivity and transitivity. It is fruitful to think
-of a category as a \emph{proof-relevant preorder}, where morphisms are
-the proofs of ordering, and the reflexivity and transitivity
-constructions must additionally satisfy equations that make them act
-like identity and composition. Then a functor is a proof-relevant
-version of a monotone function.
-
-Given categories $\mathcal C$ and $\mathcal D$, a profunctor $R$ from
-$\mathcal C$ to $\mathcal D$, written $R : \cat C \pto \cat D$ is a
-functor $R : {\cat C}^{op} \times {\cat D} \to \Set$. We can think of
-each set $R(c,d)$ as a set of ``heteromorphisms'' from $c$ (an object of
-$\cat C$) to $d$ (an object of $\cat D$).  These ``heteromorphisms''
-differ from the usual ``homomorphisms'' in a single category $\cat C$
-because they relate two objects from two different categories.
-Functoriality then defines a notion of composition of heteromorphisms
-with homomorphisms in $\cat C$ and $\cat D$. Thinking of categories as
-proof-relevant preorders, a profunctor is a proof-relevant relation that
-is downward-closed in $\cat C$ and upward-closed in $\cat D$, a kind of
-``hetero-ordering''.
-%
+introductory texts.  Recall that a category $\cat C$ has a collection of
+objects and morphisms with identity and composition, and a functor $F :
+\cat C \to \cat D$ is a function on objects and a function on morphisms
+that preserves identity and composition.  A category can be thought of
+as a generalization of a preordered set, which has a set of elements and
+a binary \emph{relation} on its objects satisfying reflexivity and
+transitivity.  A category is then a \emph{proof-relevant preorder},
+where morphisms are the proofs of ordering, and the reflexivity and
+transitivity proofs must identity and unit equations. A functor is then
+a \emph{proof-relevant monotone function}.  Given categories $\mathcal
+C$ and $\mathcal D$, a profunctor $R$ from $\mathcal C$ to $\mathcal D$,
+written $R : \cat C \pto \cat D$ is a functor $R : {\cat C}^{op} \times
+{\cat D} \to \Set$.  Because a profunctor outputs a $\Set$ rather than a
+proposition, it is itself a \emph{proof-relevant relation}.  Thinking of
+categories as proof-relevant preorders, functoriality says that the
+profunctor is downward-closed in $\cat C$ and upward-closed in $\cat D$.
 Given profunctors $R, S : \cat C \pto \cat D$, a homomorphism from
@@
 
-In addition to their usefulness for formalizing category theory, one
-reason we take profunctors to be a basic concept of \vett{} is that they
+%% We can think of each set $R(c,d)$ as a set of
+%% ``heteromorphisms'' from $c$ (an object of $\cat C$) to $d$ (an object
+%% of $\cat D$).  These ``heteromorphisms'' differ from the usual
+%% ``homomorphisms'' in a single category $\cat C$ because they relate two
+%% objects from two different categories.  Functoriality then defines a
+%% notion of composition of heteromorphisms with homomorphisms in $\cat C$
+%% and $\cat D$.
+
+Profunctors are very useful for formalizing category theory, but an
+additional reason we make them a basic concept of \vett{} is that they
 allow us to give a \emph{universal property} for the type of ``morphisms
-in a category ${\cat C}$''.  This is analogous to how the $J$ elimination
-rule for the identity type in Martin-L\"of type theory gives a universal
-property for morphisms in a groupoid (the special case of a category
-where all morphisms are
-invertible)~\cite{hofmannstreicher,awodey,voevodsky}. The reason
+in a category ${\cat C}$''.  This is analogous to how the $J$
+elimination rule for the identity type in Martin-L\"of type theory gives
+a universal property for morphisms in a groupoid (the special case of a
+category where all morphisms are
+invertible)~\cite{hofmannstreicher,awodeywarren,voevodsky}. The reason
 profunctors are useful for this purpose is that, for any category $\cat
@@
 itself. Moreover, the hom profunctor is the unit for a composition of
-profunctors $R \odot S$ which is defined as a \emph{co-end}. The
-heteromorphisms of $(R \odot S)(c,e)$ are triples $(d \in \cat D, f \in
-R(c,d), g \in S(d,e))$ modulo the equivalence relation generated by
-$(d,f,g) ~ (d', f', g')$ if there exists $h \in D(d,d')$ such that
-$h\circ f = f'$ and $g'\circ h = g$. The composition of profunctors is a
-generalization of the composition of relations, and just as the equality
-relation is the identity for the composition of relations, the hom
-profunctor is the identity for this composition.  The unit law for the
-hom profunctor can be seen as a ``morphism induction'' principle,
-analogous to the ``path induction'' used in homotopy type theory (though
-in this paper we consider only ordinary 1-dimensional categories, not
-higher generalizations).
+profunctors $R \odot S$ which is defined as a \emph{co-end}.
+%% The
+%% heteromorphisms of $(R \odot S)(c,e)$ are triples $(d \in \cat D, f \in
+%% R(c,d), g \in S(d,e))$ modulo the equivalence relation generated by
+%% $(d,f,g) ~ (d', f', g')$ if there exists $h \in D(d,d')$ such that
+%% $h\circ f = f'$ and $g'\circ h = g$.
+The composition of profunctors is a generalization of the composition of
+relations, and just as the equality relation is the identity for the
+composition of relations, the hom profunctor is the identity for this
+composition.  The unit law for the hom profunctor can be seen as a
+``morphism induction'' principle, analogous to the ``path induction''
+used in homotopy type theory (though in this paper we consider only
+ordinary 1-dimensional categories, not higher generalizations).
 
@@
 %% profunctor $R(F,G)$, the ``restriction'' of $R$ along $F$ and $G$.
-
-\paragraph{Paper Overview.}
-
-\begin{enumerate}
-\item In \cref{sec:syntax} we introduce the syntax of \vett{}.
-
-\item In \cref{sec:examples} we demonstrate how to use our syntax for
-  formal category theory.
-
-\item In \cref{sec:semantics}, we develop some model theory for
-  \vett{}, including a sound and complete notion of categorical model
+The remainder of this paper is organized as follows.  
+In \cref{sec:syntax} we introduce the syntax of \vett{}.  In
+\cref{sec:examples} we demonstrate how to use our syntax for formal
+category theory.  In \cref{sec:semantics}, we develop some model theory
+for \vett{}, including a sound and complete notion of categorical model
 and sound interpretation in virtual equipments modeling ordinary,
-  enriched and internal category theory.
-
-\item In \cref{sec:discussion}, we discuss related type theories and
-  potential extensions to model multi-variable and higher-order
-  categorical constructions.
-\end{enumerate}
+enriched and internal category theory.  In \cref{sec:discussion}, we
+discuss related type theories and potential extensions.
 
 \section{Syntax of VETT}
+\label{sec:syntax}
 
@@
 for equality in Martin-L\"of type theory, $P$ must be ``fully general'',
-i.e. well-typed for variables $\alpha$ and $\beta$.  Those familiar with
-linear logic~\cite{polakawpfenning} might expect a more general rule,
-with the continuation $t$ typed in a context $\Phi_l \jnctx \Phi_r$ and
-$\Phi_l \jnctx \Phi \jnctx \Phi_r$ as the context of the conclusion of
-the rule.  Because of dependency, this is not necessarily well-formed in
-cases where the endpoints $A$ and $B$ of $\punitinXfromYtoZ{\cat C} A B$
-are not fully general (i.e. not distinct variables).  However, the
-instances of this more general rule that do type check are derivable
-from our more restricted rule using right/left-hom types.
+i.e. well-typed for variables $\alpha$ and $\beta$; this is because for
+distinct variables $\alpha$ and $\beta$, $\punitinXfromYtoZ{\cat
+  C}{\alpha}{\beta}$ denotes the unit, which has a universal property,
+but $\punitinXfromYtoZ{\cat C} A B$ denotes a restriction of a unit,
+which in general does not.  Those familiar with linear
+logic~\cite{polakawpfenning} might expect a more general rule, with the
+continuation $t$ typed in a context $\Phi_l \jnctx \Phi_r$ and $\Phi_l
+\jnctx \Phi \jnctx \Phi_r$ as the context of the conclusion of the rule.
+Because of dependency, this is not necessarily well-formed in cases
+where the endpoints $A$ and $B$ of $\punitinXfromYtoZ{\cat C} A B$ are
+not fully general (i.e. not distinct variables).  However, the instances
+of this more general rule that do type check are derivable from our more
+restricted rule using right/left-hom types.
 
@@
   {
-   \Gamma\pipe \Phi_m \jnctx x:P, \beta:\cat D, y:Q \jnctx \Phi_r \vdash t : R\\\\
+   \Gamma\pipe \Phi_l \jnctx x:P, \beta:\cat D, y:Q \jnctx \Phi_r \vdash t : R\\\\
    \Gamma \pipe\Phi_m \vdash s : \tensorexistsXwithYandZ {\beta} P Q}
@@
 \section{Formal Category Theory in VETT}
+\label{sec:examples}
 
@@
 \section{Semantics}
+\label{sec:semantics}
 
@@
 
-Our use of dependent type theory as a metalanguage, with new judgements
-for the virtual equipment structure dependent on the meta-language
-context, is similar to the indexed type theories developed by
-\citet{isaev}.
-
 %% Directed dependent type theories
@@
 
+%% \subsubsection*{Acknowledgements}
+%% Grants stuff...? The authors would like to thank David Jaz Myers, Emily Riehl, Mike
+%% Shulman, Dominic Verity for helpful feedback on this work.
+
 %% %%

### 2022-10-13 paper/paper.tex

@@
 %% \emph{representable}.
-
-\subsection{Semantics}
-
-Our definition of a double category with displays is suggestive of a
-simple generalization. A double category is a monoid object in the
-double category of $Span(Cat)$, and a double category with displays
-furthermore uses certain diagrams involving profunctors, which also
-form a double category $Prof$. This suggests that the algebraic
-structure of a double category with displays is naturally modeled in
-some kind of \emph{triple category}, modeling the three types of
-morphisms: functors, profunctors and spans of functors.
 

### 2022-10-13 paper/paper.tex
make it fit with really small inference rules

@@
 Finally we have several forms of context which are used in the theory.
-The contexts $\Gamma$ of term variables with their types are as usual.
-The category contexts $\Xi$ contain object variables with their
-categories; the two forms of category context are $\alpha : \cat C$,
-containing one variable that can be used both contravariantly and
-covariantly, and $alpha : \cat C ; \beta : \cat D$, containing a
-contravariant variable $\alpha$ and covariant variable $\beta$.
-Finally, we have contexts $\Phi$ of
-element variables with their sets, alternating with those sets' object
-variables.  A typical $\Phi$ has the shape
+The contexts $\Gamma$ of term variables with their types are as usual;
+we write ${\Gamma \isadtctx}$ to indicate that a context is well-formed.
+The category contexts $\Gamma \vdash \Xi \boundary$ contain object
+variables with their categories; the two forms of category context are
+$\alpha : \cat C$, containing one variable that can be used both
+contravariantly and covariantly, and $\alpha : \cat C ; \beta : \cat D$,
+containing a contravariant variable $\alpha$ and covariant variable
+$\beta$.  Finally, we have contexts $\Phi$ of element variables with
+their sets, alternating with those sets' object variables.  A typical
+$\Phi$ has the shape
 \[
@@
       \textrm{Term} & L,M,N & ::= & X \pipe \lambda X:M. N \pipe M\,N \pipe (M,N) \pipe \pi_1 M \pipe \pi_2 M \mid \text{refl}_{M} \\
-      &&&\lceil \cat C \rceil \pipe \vlambda {\alpha:\cat C} a \pipe \hlambda {\alpha:\cat C}{\beta: \cat D} R \mid FIXME \ldots \\
+      &&&\lceil \cat C \rceil \pipe \vlambda {\alpha:\cat C} a \pipe \hlambda {\alpha:\cat C}{\beta: \cat D} R \mid \ldots \\
       \textrm{Type Context} & \Gamma, \Delta & ::= & \cdot \pipe \Gamma, X : A\\
@@
 
-Next, we overview our basic judgement forms along with the
-\emph{formation conditions} or \emph{presuppositions} (the conditions
-under which a judgement ``type checks'').  We have
+Next, we overview our basic judgement forms.  We have
 \begin{itemize}
-\item Categories: ${\Gamma\vdash \cat C \isaCat}$ presupposes ${\Gamma \isadtctx}$.
-
-\item Objects/functors: ${\Gamma \pipe \alpha : \cat C \vdash a : \cat D}$ presupposes
-  $\Gamma \vdash \cat C\isaCat$ and $\Gamma \vdash \cat D \isaCat$.
-Objects are typed with an input object variable
-$\alpha : \cat C$ and an output category $\cat D$; in the semantics
-these objects are modeled as functors.
-
-\item Sets/profunctors: ${\Gamma \pipe \Xi \vdash S \isaSet}$ presupposes
-      $\Gamma \isadtctx$ and $\Gamma \vdash \Xi \boundary$
-A set $S$ is typed with respect
-to a set context $\Xi$ to describe its covariant/contravariant
-dependence on some input objects. These are semantically modeled as
-profunctors.
-
-\item Elements/natural transformations: ${\Gamma \pipe \Phi \vdash s : R}$ presupposes
-     ${\Gamma \vdash \Phi \isavectx}$
-  and $\Gamma\pipe \underline \Phi \vdash R \isaSet$.
-  A transformation $s$ has a context of transformation
-variable and a single set output. To be well formed, the context and set
-must be parameterized by the same contravariant and covariant object
-variables. To ensure this, we use a coercion operation $\underline \Phi$
-that erases everything in the context but the left-most and right-most
-object variables, i.e. $\underline{\alpha : \cat C} = \alpha : \cat C$ and
-$\underline{\Phi} = d^-(\Phi);d^+(\Phi)$ otherwise.
-
-\item Meta-language types and terms: $\Gamma \vdash A \isaTy$ presupposes $\Gamma \isadtctx$ and $\Gamma
-  \vdash M : A$ presupposes $\Gamma \isadtctx$ and $\Gamma \vdash A
-  \isaTy$.
+\item Categories: ${\Gamma\vdash \cat C \isaCat}$, where ${\Gamma \isadtctx}$.
+
+\item Objects/functors: ${\Gamma \pipe \alpha : \cat C \vdash a : \cat
+  D}$, where $\Gamma \vdash \cat C\isaCat$ and $\Gamma \vdash \cat D
+  \isaCat$.  Objects are typed with an input object variable $\alpha :
+  \cat C$ and an output category $\cat D$; in the semantics, objects are
+  modeled as functors $\cat C \to \cat D$.
+
+\item Sets/profunctors: ${\Gamma \pipe \Xi \vdash S \isaSet}$, where
+  $\Gamma \vdash \Xi \boundary$. A set $S$ is typed with respect to a set
+  context $\Xi$ to describe its covariant/contravariant dependence on
+  some input objects. Sets are semantically modeled as profunctors.
+
+\item Elements/natural transformations: ${\Gamma \pipe \Phi \vdash s :
+  R}$, where ${\Gamma \vdash \Phi \isavectx}$ and $\Gamma\pipe
+  \underline \Phi \vdash R \isaSet$.  A transformation $s$ has a context
+  of transformation variable and a single set output. To be well formed,
+  the context and set must be parameterized by the same contravariant
+  and covariant object variables. To ensure this, we use a coercion
+  operation $\underline \Phi$ that erases everything in the context but
+  the left-most and right-most object variables, i.e. $\underline{\alpha
+    : \cat C} = \alpha : \cat C$ and $\underline{\Phi} =
+  d^-(\Phi);d^+(\Phi)$ otherwise.
+
+\item Meta-language types and terms: $\Gamma \vdash A \isaTy$ and
+  $\Gamma \vdash M : A$ as in standard dependent type theory.
 \end{itemize}
@@
 \begin{figure}[t]
+  \begin{scriptsize}
   \[
@@
 
-    \text{Graph of a profuctor:} \\
+    \text{Graph of a profunctor:} \\
     \inferrule
@@
   \]
+  \end{scriptsize}
   \caption{Category Conectives}
@@
 two subformulae should have the same covariant and contravariant
-dependence. So, as with linear logic, some constructions can
-syntactically use a variable more than once and still be ``linear'': the
-unit constructor ``uses'' all variables and the product ``uses'' all
-variables in both cases.
-
-\begin{figure}
+dependence (so, as with linear logic, some constructions can
+syntactically use a variable more than once and still be ``linear'').  
+%% the
+%% unit constructor ``uses'' all variables and the product ``uses'' all
+%% variables in both cases.
+
+\begin{figure}[t]
   % unit,
+  \begin{scriptsize}
   \[
@@
   \]
+  \end{scriptsize}
   \caption{Set Connectives}
@@
 %
-The types we include are what we call \emph{universes} for the object
-categorical logic: types of small categories and locally small
-categories, functors, profunctors and natural transformations.
+The types we include are \emph{universes} for the object categorical
+logic: types of small categories and locally small categories, functors,
+profunctors and natural transformations.
 %
@@
 \begin{figure}
+  \begin{scriptsize}
   \[
@@
     {\Gamma \vdash \unquoth M \isSmall}
-    \\\\
+    \quad
     \inferrule
@@
   \]
+  \end{scriptsize}
   \caption{Type Connectives}
@@
 \begin{lemma} \label{lem:adjunction}
-  Let $R : \varr {\cat D}{\cat C}$ and $L : \varr {\cat
-    C}{\cat D}$. Then the following data are in bijection:
+  For $R : \varr {\cat D}{\cat C}$ and $L : \varr {\cat
+    C}{\cat D}$, the following are in bijection:
   \begin{enumerate}
@@
 
-Next, we turn to weighted limits, which generalize ordinary limits and
+\noindent We can define weighted limits, which include ordinary limits and
 Kan extensions.
@@
 be parameterized by a weight $W k j$.
-Then we can prove the famous theorem that right adjoints preserve (weighted) limits:
+Then we can prove the well-known theorem that right adjoints preserve (weighted) limits:
 \begin{theorem}
@@
 \begin{proof}
-  \begin{align*}
+\[
     \punitinXfromYtoZ {} {\gamma} {R ((\weightedLimitDW D W) \kappa)}
@@
     \cong \homrallXYtoZ {j}{W k j} {\punitinXfromYtoZ {}{\gamma}{R(D j)}}\\
-  \end{align*}  
+\]
+This is a high level proof in terms of isomorphisms that may be written
+in VETT. The first two steps are the instantiation of assumptions
+(adjointness, weighted limits). The last step uses the fact that a
+natural isomorphisms lift to natural isomorphism of homs of profunctors.
+The construction of this isomorphism illustrates how naturality need not
+be proved explicitly in \vett{}.  For any $\phi: \homunary \alpha {R'
+  \alpha \beta} \beta {R \alpha \beta}$ and $\psi : \homunary \gamma {S
+  \gamma \beta} \beta {S' \gamma \beta}$ we can construct a natural
+transformation $\phi \triangleright \psi : \hombinary \gamma
+{(\homrallXYtoZ \beta {R \alpha \beta}{S \gamma \beta})} {\alpha}
+{R'\alpha\beta} \beta {S'\gamma\beta}$ as \\ $\lambdabinary {\gamma} {f}
+{\alpha} r \beta {\appunary \psi \gamma {(\homrappXtoYatZ f {(\appunary
+      \phi \alpha r \beta)} \beta)} \beta}$. Furthermore if $\phi$ and
+$\psi$ have inverses, then $\phi^{-1} \triangleright \psi^{-1}$ is the
+inverse of $\phi \triangleright \psi$.
 \end{proof}
-This is a high level proof in terms of isomorphisms that may be
-written in VETT. The first two steps are the instantiation of
-assumptions (adjointness, weighted limits). The last step uses the
-fact that a natural isomorphisms lift to natural isomorphism of homs
-of profunctors. We show the construction of this isomorphism as follows,
-giving an example that naturality need not be proved explicitly.
-\begin{construction}
-  For any $\phi: \homunary \alpha {R' \alpha \beta} \beta {R \alpha
-    \beta}$ and $\psi : \homunary \gamma {S \gamma \beta} \beta {S'
-    \gamma \beta}$ we can construct a natural transformation $\phi
-  \triangleright \psi : \hombinary \gamma {(\homrallXYtoZ \beta {R
-      \alpha \beta}{S \gamma \beta})} {\alpha} {R'\alpha\beta} \beta {S'\gamma\beta}$ as
-  $\lambdabinary {\gamma} {f} {\alpha} r \beta {\appunary \psi \gamma {(\homrappXtoYatZ f {(\appunary \phi \alpha r \beta)} \beta)} \beta}$.
-Furthermore if $\phi$ and $\psi$ have inverses, then $\phi^{-1} \triangleright \psi^{-1}$ is the inverse of $\phi \triangleright \psi$.
-\end{construction}
 
@@
 
-\subsection{Extensionality Axioms}
-
-Finally, we discuss some non-theorems that would be desirable to add
-as axioms to the theory of \ohol{}.
-%
-Chief among these are principles about the hom type $\punitinXfromYtoZ
-{\cat C} \alpha \beta$ when $\cat C$ is a connective.
-%
-Since the hom type is defined as a kind of unit type, rather than
-defined inductively, we are not guaranteed that the hom type for the
-product category $\cat C \times \cat D$ or the presheaf categories are
-what we expect.
-%
-For the product type, we expect there should be an isomorphism
-$\punitinXfromYtoZ {\cat C \times \cat D} \alpha \beta \cong
-(\punitinXfromYtoZ {\cat C} {\pi_1 \alpha} {\pi_1 \beta}) \times
-(\punitinXfromYtoZ {\cat D} {\pi_2 \alpha} {\pi_2 \beta})$ saying an
-arrow in a product category is a pair of arrows.
-%
-Similarly, the hom type in the presheaf category should be given by
-the end formula $\punitinXfromYtoZ {\negPresheaf {\cat C}} \alpha
-\beta \cong \homlallXYtoZ {\gamma}{\negPresheafApp \alpha
-  \gamma}{\negPresheafApp \beta \gamma}$. This would allow us to
-relate the Yoneda Lemma \ref{lem:yoneda} to the morphisms of
-presheaves, and prove that the Yoneda embedding is fully faithful.
-%
-We call these extensionality principles because they are analogous to
-principles such as function extensionality but for the directed hom
-type rather than equality.
-
-The presheaf extensionality principle is independent from the existing
-axioms of \ohol{}: it is true in the ordinary category theoretic
-model, but it is false in the set theoretic model. That is, in the set
-theoretic model, the presheaf category is simply the powerset, the hom
-type is \emph{equality} and the function profunctor is interpreted as
-a universally quantified \emph{implication}, but equality in the
-powerset should be \emph{bi-implication}, so the theorem is false. We
-expect the product extensionality is independent as well, but a
-counter-model would likely be quite exotic as we expect it to be
-provable in a system with functor categories, in analogy with the
-proof for the identity type in MLTT.
-%
-Note that these axioms can easily be added as assumptions in the
-dependently typed context, though this will break the constructive
-character of the system. We expect techniques from type theory might
-be applied to incorporate these as computationally meaningful
-principles rather than assumed as axioms \cite{obs-tt,xtt,cubical-type-theory}.
+%% MOVE TO APPENDIX examples section?
+
+%% \subsection{Extensionality Axioms}
+
+%% Finally, we discuss some non-theorems that would be desirable to add
+%% as axioms to the theory of \ohol{}.
+%% %
+%% Chief among these are principles about the hom type $\punitinXfromYtoZ
+%% {\cat C} \alpha \beta$ when $\cat C$ is a connective.
+%% %
+%% Since the hom type is defined as a kind of unit type, rather than
+%% defined inductively, we are not guaranteed that the hom type for the
+%% product category $\cat C \times \cat D$ or the presheaf categories are
+%% what we expect.
+%% %
+%% For the product type, we expect there should be an isomorphism
+%% $\punitinXfromYtoZ {\cat C \times \cat D} \alpha \beta \cong
+%% (\punitinXfromYtoZ {\cat C} {\pi_1 \alpha} {\pi_1 \beta}) \times
+%% (\punitinXfromYtoZ {\cat D} {\pi_2 \alpha} {\pi_2 \beta})$ saying an
+%% arrow in a product category is a pair of arrows.
+%% %
+%% Similarly, the hom type in the presheaf category should be given by
+%% the end formula $\punitinXfromYtoZ {\negPresheaf {\cat C}} \alpha
+%% \beta \cong \homlallXYtoZ {\gamma}{\negPresheafApp \alpha
+%%   \gamma}{\negPresheafApp \beta \gamma}$. This would allow us to
+%% relate the Yoneda Lemma \ref{lem:yoneda} to the morphisms of
+%% presheaves, and prove that the Yoneda embedding is fully faithful.
+%% %
+%% We call these extensionality principles because they are analogous to
+%% principles such as function extensionality but for the directed hom
+%% type rather than equality.
+
+%% The presheaf extensionality principle is independent from the existing
+%% axioms of \ohol{}: it is true in the ordinary category theoretic
+%% model, but it is false in the set theoretic model. That is, in the set
+%% theoretic model, the presheaf category is simply the powerset, the hom
+%% type is \emph{equality} and the function profunctor is interpreted as
+%% a universally quantified \emph{implication}, but equality in the
+%% powerset should be \emph{bi-implication}, so the theorem is false. We
+%% expect the product extensionality is independent as well, but a
+%% counter-model would likely be quite exotic as we expect it to be
+%% provable in a system with functor categories, in analogy with the
+%% proof for the identity type in MLTT.
+%% %
+%% Note that these axioms can easily be added as assumptions in the
+%% dependently typed context, though this will break the constructive
+%% character of the system. We expect techniques from type theory might
+%% be applied to incorporate these as computationally meaningful
+%% principles rather than assumed as axioms \cite{obs-tt,xtt,cubical-type-theory}.
 
@@
 
-\subsection{Notion of Model}
+%% \subsection{Notion of Model}
 
@@
 \begin{proof}
-  Straightforward, defining the category of families using the
-  dependent type structure and the virtual equipment structure having
+  Straightforward, defining the category of families using the dependent
+  type structure and the virtual equipment structure having
   ($\alpha$-equivalence classes of) syntactic categories as objects,
   functors/sets as vertical/horizontal arrows and interpreting
-  compositions/restrictions as substitutions.  The biggest gap is in
-  the definition of the 2-cells.
-
-  A 2-cell from $(\alpha_1:\cat C_1;\alpha_2:\cat C_2\vdash
-  R_1),(\alpha_2:\cat C_2;\alpha_3:\cat C_3),\ldots$ to $(\beta_1:\cat
-  D_1;\beta_2\vdash S)$ with frames $\alpha_1:\cat C_1\vdash b_1 :
-  \cat D_1$ and $\alpha_n:\cat C_n\vdash b_2 : \cat D_2$ given by a
-  term $x_1:R_1,x_2:R_2\ldots \vdash s :
-  S[b_1/\beta_1;b_2/\beta_2]$. Composition is then defined by
-  substitution.
+  compositions/restrictions as substitutions.  The biggest gap is in the
+  definition of the 2-cells.  A 2-cell from\\ $(\alpha_1:\cat
+  C_1\mathord{;}\alpha_2:\cat C_2\vdash R_1),(\alpha_2:\cat
+  C_2;\alpha_3:\cat C_3 \vdash R_2),\ldots$ to $(\beta_1:\cat
+  D_1;\beta_2\vdash S)$ with frames $\alpha_1:\cat C_1\vdash b_1 : \cat
+  D_1$ and $\alpha_n:\cat C_n\vdash b_2 : \cat D_2$ given by a term
+  $x_1:R_1,x_2:R_2\ldots \vdash s :
+  S[b_1/\beta_1;b_2/\beta_2]$. Composition is defined by substitution.
 \end{proof}
@@
 
-\subsection{Examples of Models}
+%% \subsection{Examples of Models}
 
@@
 
-Then to define a model of VETT with a collection of connectives is is
-sufficient to construct a virtual equipment with the universal
-properties corresponding to those connectives.
+Then to define a model of VETT with a collection of connectives it is
+sufficient to construct a virtual equipment with the corresponding
+universal properties.
 %
@@
 is highly general: Shulman's enriched indexed categories. Shulman's
-construction defines a virtual double categories of large and small
+construction defines a virtual double category of large and small
 $\mathcal V$-categories for any pseudofunctor $\mathcal V : S^o \to
@@
 modeled by pullback which is only associative up to unique
-isomorphism. In fact, one of the main examples is only a
-pseudo-functor for precisely this reason that the action on morphisms
-is given by pullback.
-%
-We expect this difference can be handled in the same way: by a general
-strictification theorem or by moving to a univalent setting.
+isomorphism.
+%% In fact, one of the main examples is only a
+%% pseudo-functor for precisely this reason that the action on morphisms
+%% is given by pullback.
+%
+We leave strictification theorems to future work.
 
@@
 not have functors and restriction of profunctors by
-functors. Semantically, Cartesian bicategories are somewhat less general
+functors. Semantically, Cartesian bicategories are less general
 than virtual equipments because in general the set of functors between
@@
 
-\subsection{Extending VETT}
-
-While many constructions in category theory can be formalized in
-virtual equipments, and therefore \vett{}, others require more
-structure. Future extensions to \vett{} could incorporate functor
-categories by generalizing the unary type theory of functors to
-functors of many variables, in which case ordinary $\lambda$ calculus
-can be used to define functor categories as function types. Similarly,
-ideas from modal type theory/graded linear logic could be used for
-defining opposite categories
-\cite{mike-contravariance-enrichment,semiring-stuf??}.
+In future work, \vett{} could incorporate functor categories by
+generalizing the unary type theory of functors to functors of many
+variables, in which case ordinary $\lambda$ calculus can be used to
+define functor categories as function types. Similarly, ideas from modal
+type theory/graded linear logic could be used for defining opposite
+categories \cite{mike-contravariance-enrichment,semiring-stuf??}.
+
+%% while many constructions in category theory can be formalized in
+%% virtual equipments, and therefore \vett{}, others require more
+%% structure. Future extensions to
 
@@
 
-%% \subsubsection*{Acknowledgements}
-%% Grants stuff...? The authors would like to thank David Jaz Myers, Emily Riehl, Mike
-%% Shulman, Dominic Verity for helpful feedback on this work.
+\textbf{Acknowledgements.}  This material is based on research sponsored
+by the National Science Foundation under agreement number CCF-1909517
+and the United States Air Force Research Laboratory under agreement
+number FA9550-21-0009 (Tristan Nguyen, program manager).
+%% The U.S. Government is authorized to reproduce and distribute
+%% reprints for Governmental purposes notwithstanding any copyright
+%% notation thereon.  The views and conclusions contained herein are
+%% those of the authors and should not be interpreted as necessarily
+%% representing the official policies or endorsements, either expressed
+%% or implied, of the National Science Foundation, the United States Air
+%% Force Research Laboratory, or the U.S. Government.
+The authors would like to thank David Jaz Myers, Emily Riehl, Mike
+Shulman, Dominic Verity for helpful feedback on this work.
 

### 2022-10-13 paper/paper.tex
\beta\eta rules

@@
 \newcommand{\negPresheaf}{\paramPresheaf-}
-\newcommand{\negPresheafApp}[2]{#1\,#2}
 \newcommand{\negPresheafAppPtoX}[2]{#2\in #1}
 \newcommand{\posPresheafAppPtoX}[2]{#1 \ni #2}
-\newcommand{\posPresheafApp}[2]{#1\,#2}
+\newcommand{\negPresheafApp}{\negPresheafAppPtoX}
+\newcommand{\posPresheafApp}{\posPresheafAppPtoX}
 
@@
 %
-We call the resulting theory \emph{virtual equipment type theory}
-(VETT) as (hyperdoctrines of) virtual equipments provide a sound and
-complete notion of model for the theory.
+We call the resulting theory \emph{virtual equipment type theory} (VETT)
+as (hyperdoctrines of) \emph{virtual
+  equipments}~\cite{virtualequipments}, a particular semantic model of
+formal category theory, provide a sound and complete notion of model for
+the theory.
 %
@@
 formalize theorems the logic is too restrictive to encode, analogous to
-2-level~\cite{voevodsky12,altenkirchXX} or indexed type
-theories~\cite{isaev,linearlf,vakar,krishnaswamibenton}.
+2-level~\cite{voevodsky13hts,altenkirch+16strict} or indexed type
+theories~\cite{isaev21indexed,cervesatopfenning02llf,vakar15linear,krishnaswami+15linear}.
 
@@
 category where all morphisms are
-invertible)~\cite{hofmannstreicher,awodeywarren,voevodsky}. The reason
+invertible)~\cite{hofmann98groupoid,awodeywarren09identity,voevodsky11wollic}. The reason
 profunctors are useful for this purpose is that, for any category $\cat
@@
 The remainder of this paper is organized as follows.  
-In \cref{sec:syntax} we introduce the syntax of \vett{}.  In
-\cref{sec:examples} we demonstrate how to use our syntax for formal
-category theory.  In \cref{sec:semantics}, we develop some model theory
+In Section~\ref{sec:syntax} we introduce the syntax of \vett{}.  In
+Section~\ref{sec:examples} we demonstrate how to use our syntax for formal
+category theory.  In Section~\ref{sec:semantics}, we develop some model theory
 for \vett{}, including a sound and complete notion of categorical model
 and sound interpretation in virtual equipments modeling ordinary,
-enriched and internal category theory.  In \cref{sec:discussion}, we
+enriched and internal category theory.  In Section~\ref{sec:discussion}, we
 discuss related type theories and potential extensions.
@@
 two subformulae should have the same covariant and contravariant
-dependence (so, as with linear logic, some constructions can
+dependence (as with linear logic, some constructions can
 syntactically use a variable more than once and still be ``linear'').  
@@
      \Gamma \pipe d^+ \Xi \vdash B : \cat C}
-    {\Gamma \pipe \Xi \vdash M A; B \isaSet}
+    {\Gamma \pipe \Xi \vdash M A\, B \isaSet}
     \\ \\
@@
 for explicit naturality and functoriality side conditions.  We have
-formalized the syntax of \vett{} in Agda, using Agda's rewriting
-mechanism to implement substitution and $\beta$-reduction, and checked
-some of the results in this section (e.g. Lemma~\ref{lem:yoneda} and
-Lemma~\ref{lem:fubini-short} and the maps in
-Lemma~\ref{lem:adjunction}).\footnote{\url{https://github.com/maxsnew/virtual-equipments/blob/master/agda/Examples.agda}}
+checked some of the results in this section (e.g. Lemma~\ref{lem:yoneda}
+and Lemma~\ref{lem:fubini-short} and the maps in
+Lemma~\ref{lem:adjunction}) in
+Agda.\footnote{\url{https://github.com/maxsnew/virtual-equipments/blob/master/agda/Examples.agda}}
 
@@
 ordered generalizations of some very simple tautologies about
-equality. For instance, the Yoneda lemma (and its proof) generalizes
+equality. For instance, the Yoneda lemma generalizes
 the equivalence between the formulae $\forall x. \forall y. x = y
@@
 \end{lemma}
-The proofs both follow from the unit elimination rule: we can view
-the unit elimination rule as essentially making the Yoneda lemma a
-primitive axiom, as the two cases of showing (1) is an isomorphism are
-precisely the $\beta$ and $\eta$ rules for the unit.
+The proofs both follow from the unit elimination rule, which is
+essential the Yoneda lemma---the two cases of showing (1) is an
+isomorphism are precisely the $\beta$ and $\eta$ rules for the unit.
 
@@
     and target functions $s,t : {\mathcal V}_h \to {\mathcal V_o}^2$
-  \item Sets of 2-cells of the following form:
-    With appropriate ``multi-categorical'' notions of identity and
-    composition.
+  \item Sets of 2-cells of the following form, 
+    with appropriate ``multi-categorical'' notions of identity and
+    composition:
 \[\begin{tikzcd}% https://q.uiver.app/?q=WzAsNixbMCwwLCJDXzAiXSxbMCwyLCJEXzAiXSxbMiwwLCJDX3tufSJdLFsxLDAsIlxcY2RvdHMiXSxbMiwyLCJEXzEiXSxbMSwxLCJcXHBoaSJdLFszLDIsIlJfbiIsMCx7InN0eWxlIjp7ImJvZHkiOnsibmFtZSI6ImJhcnJlZCJ9fX1dLFswLDEsIkYiLDJdLFsyLDQsIkciXSxbMSw0LCJTIiwyLHsic3R5bGUiOnsiYm9keSI6eyJuYW1lIjoiYmFycmVkIn19fV0sWzAsMywiUl8wIiwwLHsic3R5bGUiOnsiYm9keSI6eyJuYW1lIjoiYmFycmVkIn19fV1d
@@
 \section{Related and Future Work}
+\label{sec:discussion}
 
@@
 $\infty$-)category.  In directed type theories based on a bisimplicial
-model~\cite{riehl-shulman,weinberger,otherweinbegerpaper,licataweaver},
+model~\cite{riehlshulman17directed,buchholtzweinberger21fibered,weinberger22thesis,WeaverLicata20},
 morphism types are defined using an interval object, like in cubical
-type theory~\cite{bch,cchm,afh,abcfhl}, and universal properties like
-``morphism induction'' are an internally definable property of certain
-types.  Other type theories~\cite{north,ahrensnorth} define morphism
-types via an induction principle, corresponding to the lifting
-properties of certain kinds of fibrations of categories.  While these
-previous works can express some constructions on $\Cat$ that are not
-expressible in \vett{}, because \vett{} is more restricted, \vett{}
-contrariwise has more models, for instance categories enriched in
-non-cartesian monoidal categories, so the theorems that are provable
-in \vett{} apply in more settings.
+type
+theory~\cite{bch18,cchm18cubical,angiuli+18cartesian-csl,abcfhl17cartesian},
+and universal properties like ``morphism induction'' are an internally
+definable property of certain types.  Other type
+theories~\cite{north18,ahrens+22semantics-2dtt} define morphism types via an induction
+principle, corresponding to the lifting properties of certain kinds of
+fibrations of categories.  While these previous works can express some
+constructions on $\Cat$ that are not expressible in \vett{}, because
+\vett{} is more restricted, \vett{} contrariwise has more models, for
+instance categories enriched in non-cartesian monoidal categories, so
+the theorems that are provable in \vett{} apply in more settings.
 
@@
 
-\section{Syntactic Details}
+\section{\vett{} Syntactic Metatheory}
+\label{sec:appendix:syntax}
+
+\subsection{Substitution}
+
+In Figure~\ref{fig:subst} we give full definitions for well-formedness
+of contexts and substitutions.
 
@@
     {\Gamma \vdash \phi : \Psi \to \Phi}
-
   \end{mathpar}
-  \caption{Substitution}
-
-\end{figure}
-
-\cref{fig:contexts} details the typing rules for contexts and the
-typing for variables and the admissible actions of substitution (the
-definition of substitution is straightforward and found in the
-appendix). the rules for term and object variables are as usual.
-%
-An object substitution into a single object context $\alpha:\cat C$
-consists of an object $a : \cat C$ parameterized by an object of
-another category.
-%
-An object substitution into a pair of co- and contra-variant variables
-$\alpha:\cat C;\beta:\cat D$ is a pair of an object substituted for
-$\alpha$ and one for $\beta$, each typed with respect to the
-appropriate variable of an object context $\Xi$.
-%
-Note that this allows for $\Xi$ to consist of either two variables or
-a single variable.
-%
-When
-constructing a transformation context $\Phi, x : R, \Psi$, we need to
-ensure that the contravariant dependence of $R$ matches the covariant
-dependence of $\Phi$ and vice-versa for $\Psi$. We define this with
-auxiliary metafunctions $d^+,d^-$ which extract the covariant and
-contravariant object variable of a transformation context.
-%
-The rules for transformation variables and substitutions define the
-use of variables to be \emph{linear} (each variables is used exactly
-once) and \emph{ordered}, they occur in the term in the same order
-that they appear in the context.
-%
-The rule for a transformation variables says a variable $x$ can only
-be used when it is the sole variable in the context.
-%
-A transformation subsitution into an ``empty'' transformation context
-$\alpha: \cat C$ is the same as an object substitution into
-$\alpha:\cat C$.
-%
-The rule for constructing a transformation substitution into a
-non-empty context $\Phi_1,x:R,\Phi_2$ says we can construct a
-substitution by constructing substitutions $\phi_1,\phi_2$ for
-$\Phi_1$ and $\Phi_2$ as well as a transformation $s$ into an
-instantiation of $R$ where the input contexts come from splitting the
-input context $\Psi$ into three components, and ensuring that the
-instantiation of $R$ agrees with how $\phi_1,\phi_2$ instantiate the
-variables. This uses two new metafunctions. First, $\Phi \jnctx \Psi$
-is an operation for concatenating two contexts when $d^+(\Phi) =
-d^-(\Psi)$. Next, we extend the boundary maps $d^\pm$ to substitutions
-to extract their object substitutions.
-
-\begin{figure}
   \begin{mathpar}
@@
   \end{mathpar}
+
+  \caption{Substitution}
+\end{figure}
+
+\begin{figure}
   \begin{align*}
@@
   \end{align*}
+
+  \caption{Operations on Contexts}
+
+\end{figure}
+
+\begin{figure}
+  \begin{mathpar}
+    \inferrule
+    {}
+    {\cdot : \Gamma \to \cdot}
+
+    \inferrule
+    {\gamma : \Delta \to \Gamma \and \Delta \vdash M : A[\gamma]}
+    {\gamma,M/X : \Delta \to \Gamma,X:A}
+
+    \inferrule
+    {\Gamma \vdash \Phi \isavectx \and
+    \Gamma \vdash \Psi \isavectx \and
+    d^+\Phi = d^- \Psi}
+    {\Gamma \vdash \Phi \jnctx \Psi \isavectx}
+
+    \inferrule
+    {\Gamma \vdash \phi : \Phi' \to \Phi \and
+     \Gamma \vdash \psi : \Psi' \to \Psi\and
+     d^+\phi = d^- \psi}
+    {\Gamma \vdash \phi \jnctx \psi : \Phi' \jnctx \Psi' \to \Phi \jnctx \Psi}
+
+    \inferrule*[right=ObSubst]
+    {\alpha : \cat C \vdash B : \cat D\and
+    \beta : \cat D \vdash A : \cat E}
+    {\alpha : \cat C \vdash A[B/\beta] : \cat E}\\
+
+    \inferrule*[right=Everything Subst]{\gamma : \Delta \to \Gamma \and \Gamma J}{\Delta J}
+
+    \inferrule{\Gamma \vdash \xi : Z \to \Xi\and \Gamma \pipe \Xi \vdash R \isaSet}{\Gamma \pipe Z \vdash R[\xi] \isaSet}
+
+    \inferrule{\Gamma \vdash \phi : \Psi \to \Phi\and \Gamma \pipe \Phi \vdash M : R}{\Gamma \pipe \Psi \vdash M[\phi] : R[\underline \phi]}
+  \end{mathpar}
+  \begin{align*}
+    (\Phi \jnctx \Psi) \jnctx \Sigma &= \Phi \jnctx (\Psi \jnctx \Sigma)\\
+    \Phi \jnctx \alpha:C &= \Phi\\
+    \alpha:C \jnctx \Psi &= \Psi\\\\
+    d^\pm(A/\alpha) &= A/\alpha\\
+    d^+(\phi_1,M/x,\phi_2) &= d^+ \phi_2\\
+    d^-(\phi_1,M/x,\phi_2) &= d^- \phi_1\\\\
+    \underline \cdot &= \cdot\\
+    \underline {A/\alpha} &= A/\alpha\\
+    \underline {\phi_1,M/x,\phi_2} &= d^-\phi_1;d^+\phi_2
+  \end{align*}
+  \caption{Admissible Rules}
+\end{figure}
+
+\cref{fig:contexts} details the typing rules for contexts and the
+typing for variables and the admissible actions of substitution (the
+definition of substitution is straightforward and found in the
+appendix). the rules for term and object variables are as usual.
+%
+An object substitution into a single object context $\alpha:\cat C$
+consists of an object $a : \cat C$ parameterized by an object of
+another category.
+%
+An object substitution into a pair of co- and contra-variant variables
+$\alpha:\cat C;\beta:\cat D$ is a pair of an object substituted for
+$\alpha$ and one for $\beta$, each typed with respect to the
+appropriate variable of an object context $\Xi$.
+%
+Note that this allows for $\Xi$ to consist of either two variables or
+a single variable.
+%
+When
+constructing a transformation context $\Phi, x : R, \Psi$, we need to
+ensure that the contravariant dependence of $R$ matches the covariant
+dependence of $\Phi$ and vice-versa for $\Psi$. We define this with
+auxiliary metafunctions $d^+,d^-$ which extract the covariant and
+contravariant object variable of a transformation context.
+%
+The rules for transformation variables and substitutions define the
+use of variables to be \emph{linear} (each variables is used exactly
+once) and \emph{ordered}, they occur in the term in the same order
+that they appear in the context.
+%
+The rule for a transformation variables says a variable $x$ can only
+be used when it is the sole variable in the context.
+%
+A transformation subsitution into an ``empty'' transformation context
+$\alpha: \cat C$ is the same as an object substitution into
+$\alpha:\cat C$.
+%
+The rule for constructing a transformation substitution into a
+non-empty context $\Phi_1,x:R,\Phi_2$ says we can construct a
+substitution by constructing substitutions $\phi_1,\phi_2$ for
+$\Phi_1$ and $\Phi_2$ as well as a transformation $s$ into an
+instantiation of $R$ where the input contexts come from splitting the
+input context $\Psi$ into three components, and ensuring that the
+instantiation of $R$ agrees with how $\phi_1,\phi_2$ instantiate the
+variables. This uses two new metafunctions. First, $\Phi \jnctx \Psi$
+is an operation for concatenating two contexts when $d^+(\Phi) =
+d^-(\Psi)$. Next, we extend the boundary maps $d^\pm$ to substitutions
+to extract their object substitutions.
+
+\begin{figure}
+
   \caption{Basic Rules of Judgment Theory}
@@
 when discussing the positives.
+
+\subsection{Equational Theory}
+
+Next we present the $\beta\eta$ rules that generate the equational
+theory of the terms. In keeping with the extensional style of the type
+theory, we do not present explicit transitivity, congruence, or
+transport rules, but rather consider these as inherent to the notion
+of equality. This can be formalized by modeling the terms of our type
+theory as a quotient inductive inductive type
+\cite{altenkirch-kaposi-popl17}.
+
+\begin{figure}
+  \begin{mathpar}
+    % Type connectives: small and large cats, functors, profs, nat transformations, and if we want also Pi/Sigma/Id
+    \inferrule*[right=SmallCat$\beta$]
+    {~}
+    {\unquoth {\quoth {\cat C}}  = \cat C}
+
+    \inferrule*[right=SmallCat$\eta$]
+    {\Gamma \vdash M : \smallCats}
+    {\Gamma \vdash \quoth {\unquoth {M}}  = M : \smallCats}
+
+    \inferrule*[right=Cat$\beta$]
+    {~}
+    {\unquoth {\quoth {\cat C}}  = \cat C}
+
+    \inferrule*[right=Cat$\eta$]
+    {\Gamma \vdash M : \Cats}
+    {\Gamma \vdash \quoth {\unquoth {M}}  = M : \Cats}
+
+    \\
+    \inferrule*[right=Fctor$\beta$]
+    {~}
+    {(\lambda \alpha. b)\,a = b[a/\alpha]}
+
+    \inferrule*[right=Fctor$\eta$]
+    {\Gamma \vdash M : \varr {\cat C}{\cat D}}
+    {\Gamma \vdash M = \lambda \alpha. M\,\alpha : \varr {\cat C}{\cat D}}
+
+    \inferrule*[right=Prof$\beta$]
+    {~}
+    {(\lambda \alpha \beta. P)\,a\,b = P[a/\alpha;b/\beta]}
+
+    \inferrule*[right=Prof$\eta$]
+    {\Gamma \vdash M : \harr {\cat C}{\cat D}}
+    {\Gamma \vdash M = \lambda \alpha\,\beta. M\,\alpha\,\beta : \harr {\cat C}{\cat D}}
+
+    \inferrule*[right=NatElt$\beta$]
+    {~}
+    {(\pendlambdaXdotY \alpha s) a = s[a/\alpha]}
+
+    \inferrule*[right=NatElt$\eta$]
+    {\Gamma \vdash M : \pendallXdotY \alpha P}
+    {\Gamma \vdash M = \pendlambdaXdotY \alpha {M\,a} : \pendallXdotY\alpha P}
+
+    \\\\\\\\
+
+    % Object connectives: 0,2-products, tabulators, +/- presheafs
+
+    \inferrule*[right=NegPresheaf$\beta$]
+    {~}
+    {\negPresheafApp a (\lambda \alpha. R) = R[a/\alpha]}
+
+    \inferrule*[right=NegPresheaf$\eta$]
+    {\Gamma \pipe \beta:\cat D \vdash p : \negPresheaf {\cat C}}
+    {p = \lambda \alpha. \negPresheafApp \alpha p}
+
+    \inferrule*[right=NegPresheaf$\beta$]
+    {~}
+    {\posPresheafApp {(\lambda \beta. R)} b = R[b/\beta]}
+
+    \inferrule*[right=PosPresheaf$\eta$]
+    {\Gamma \pipe \alpha:\cat C \vdash p : \posPresheaf {\cat D}}
+    {p = \lambda \beta. \posPresheafApp p \beta}
+
+    \inferrule*[right=Graph$\beta-$]
+    {~}
+    {\pi_- (a_-,a_+,s) = a_-}
+
+    \inferrule*[right=Graph$\beta+$]
+    {~}
+    {\pi_+ (a_-,a_+,s) = a_+}
+
+    \inferrule*[right=Graph$\beta e$]
+    {~}
+    {\pi_e (a_-,a_+,s) = s}
+
+    \inferrule*[right=Graph$\eta$]
+    {\Gamma \pipe \alpha:\cat C \vdash b : \graphProf{\beta_-}{\beta_+} P}
+    {\Gamma \pipe \alpha:\cat C \vdash b = (\pi_-b, \pi_+b, \pi_e b) : \graphProf{\beta_-}{\beta_+} P}
+
+  \inferrule*[right=${1}\eta$]
+  {\Gamma \pipe \alpha:\cat C \vdash a : 1}
+  {\Gamma \pipe \alpha:\cat C \vdash a = () : 1}\\
+
+  \inferrule*[right=${\times}\beta$]
+  {~}
+  {\pi_i(a_1,a_2) = a_i}
+
+  \inferrule*[right=${\times}\eta$]
+  {\Gamma \pipe \alpha:\cat C \vdash a : \cat D_1 \times \cat D_2}
+  {\Gamma \pipe \alpha:\cat C \vdash a = (\pi_1a, \pi_2a) : \cat D_1 \times \cat D_2}
+  \end{mathpar}
+  \caption{$\beta\eta$ Equality for type and object connectives}
+\end{figure}
+
+\begin{figure}
+  \begin{mathpar}
+  % Set connectives: unit, tensor, +/- hom, 0,2-products
+  \inferrule*[right=CovHom$\beta$]
+    {~}
+    {\homrappXtoYatZ {(\homrlambdaXatYdotZ x \alpha s)} t a = s[t/x,a/\alpha]}
+    \inferrule*[right=CovHom$\eta$]
+               {\Gamma\pipe\Phi \vdash s : \homrallXYtoZ \alpha R P}
+               {\Gamma\pipe\Phi \vdash s = \homrlambdaXatYdotZ {x}{\alpha} \homrappXtoYatZ s x \alpha}
+
+    \inferrule*[right=ConHom$\beta$]
+                 {~}
+                 {\homlappXtoYatZ {(\homllambdaXatYdotZ x \alpha t)} s a = t[a/\alpha,s/x]}
+      \inferrule*[right=ConHom$\eta$]
+      {\Gamma \pipe \Phi \vdash t : \homlallXYtoZ \alpha R P}
+      {\Gamma \pipe \Phi \vdash t = \homllambdaXatYdotZ x \alpha \homlappXtoYatZ t x \alpha : \homlallXYtoZ \alpha R P}
+
+    \inferrule*[right=Unit$\beta$]
+    {~}
+    {(\punitelimkontZatABC {\alpha. t} a {\punitrefl a} a) = t[a/\alpha]}
+    \inferrule*[right=Unit$\eta$]
+    {\Gamma\pipe\Phi \vdash s : \punitinXfromYtoZ {\cat C}{a}{b}}
+    {\Gamma\pipe\Phi\vdash s = \punitelimkontZatABC {\alpha. \punitrefl \alpha} a s b: \punitinXfromYtoZ {\cat C}{a}{b}}
+
+    \inferrule*[right=Tensor$\beta$]
+    {~}
+    {\tensorelimWkontZ {\tensorintroatXwithYandZ b s t} {x.\beta.y. r} = t[s/x;b/\beta;t/y]}
+    \inferrule*[right=Tensor$\eta$]
+    {\Gamma\pipe\Phi \vdash s : \tensorexistsXwithYandZ \beta P Q}
+    {\Gamma\pipe\Phi\vdash s = \tensorelimWkontZ s {x.\beta.y. \tensorintroatXwithYandZ \beta x y}: \tensorexistsXwithYandZ \beta P Q}
+
+  \inferrule*[right=$1\eta$]
+  {\Phi \vdash s : 1}
+  {\Phi \vdash s = () : 1}\\
+
+  \inferrule*[right=${\times}\beta$]
+  {~}
+  {\pi_i(s_1,s_2) = s_i}
+
+  \inferrule*[right=${\times}\eta$]
+  {\Phi \vdash s : P \times Q}
+  {\Phi \vdash s = (\pi_1 s, \pi_2 s) : P \times Q}
+  \end{mathpar}
+  \caption{$\beta\eta$ Equality for set connectives}
+  \label{fig:betaeta-sets}
+\end{figure}
+
+\subsection{Identity Elimination}
+
+As a final syntactic meta-theorem, we prove that the more general identity elimination principle
+\[ \]
+is admissible in the presence of let and right hom types.
+
+TODO
 
@@
 \end{figure}
+
+\section{Extended Formal Category Theory Examples}
+\label{sec:appendix:examples}
+
+Next, we provide some futher detailed examples of the formal category
+theory constructions and theorems from Section~\ref{sec:examples}.
+
 \section{Semantics}
+\label{sec:appendix:semantics}
+
+In this section, we provide the full definitions of the universal
+properties in a virtual equipment corresponding to each connective in
+\vett{}, and the most critical details of the soundness and
+completeness proof.
+
+TODO
 
@@
   \end{enumerate}
-\end{construction}
-
-The associativity, unit and interchange laws of a category are proven
-in the \cref{appendix}.
-
-\begin{theorem}{(Sketch) Soundness/Initiality}
-  Given any HED $\mathcal V$, there is a unique homomorphism of HEDs $\sem\cdot$
-  from $\textrm{Syn}(\textrm{SCT})$ to $\mathcal V$.
-\end{theorem}
-\begin{proof}
-  Roughly the converse of the completeness theorem. For instance,
-  profunctor contexts are interpreted using horizontal identity and
-  composition as well as the $i$ functor. Preservation of the
-  restriction operation requires that restriction commutes with the
-  display connectives.
-\end{proof}
-
-Details of the interpretation of profunctor contexts and substitutions
-are given in the appendix\cref{appendix}.
-
-\subsection{Example Models}
-
-We now describe how to interpret the syntax in any virtual equipment
-(that can model the connectives). First, from any virtual equipment we
-can construct a DCD:
-
-\begin{definition}
-  Given a virtual equipment $\mathcal V$, define a DCD by taking the
-  horizontal morphisms of $\mathcal V$ as the display morphisms and
-  the horizontal morphisms as the ``composable strings'' of horizontal
-  morphisms. Then we define $C_\vdash$ using the notion of
-  ``multi-2-cell''.
-\end{definition}
-
-Next, we can upgrade any DCD in $\Set$ to a DCD internal to
-$\Set^{\Set^o}$ by taking the presheaf of objects, vertical arrows, etc as
-families of objects, vertical arrows etc indexed by the indexing set
-$\Gamma$ as follows:
-\begin{definition}[Powers of Virtual Equipments]
-  Let $\mathcal V$ be a virtual equipment and let $X$ be a set.
-  Define a virtual equipment, the ``power'' $\mathcal V^X$ as follows.
-
-  \begin{enumerate}
-  \item Objects are functions $X \to \mathcal V_0$.
-  \item Vertical morphisms $\mathcal V^X_0(A,B)$ are functions
-    $\Pi_{x\in X} V_0(A x,B x)$. With pointwise composition, identity.
-  \item Horizontal morphisms $\mathcal V^X_1$ are functions $\Gamma
-    \to \mathcal V_1$.
-  \item A 2-cell from $\Phi$ to $R$ is a family $\Pi_{x\in X} \mathcal
-    V^X_h(\Phi x,R x)$.
-  \end{enumerate}
-  All operations are lifted pointwise.
-\end{definition}
-
-This defines the ``natural'' interpretation of SCT in a virtual
-equipment once we see that the powering construction is functorial in
-$\Gamma$.
-\begin{definition}[Internal Logic]
-  Given any virtual equipment $\mathcal V$, we can construct the
-  ``internal logic'' by indexing $\mathcal V$ by $\Set$:
-
-  \begin{enumerate}
-  \item The indexing cateogry is $\Set$, which is a category with families.
-  \item The equipment is defined as $\mathcal U\Gamma = \mathcal V^\Gamma$.
-  \item Restriction by a function $\gamma : \Delta \to \Gamma$ is
-    given pointwise.
-  \item The representing object for objects is $\textrm{Ob}_\Gamma =
-    \textrm{Ob}(\mathcal V)^\Gamma$ and similarly for the other
-    universe objects.
-  \end{enumerate}
-\end{definition}
-
-\section{Completeness}
+\end{construction}\section{Completeness}
 
@@
   \caption{Syntactic Double Category with Displays Structure}
-\end{figure}
-
-\begin{figure}
-  \begin{mathpar}
-    \inferrule
-    {}
-    {\cdot : \Gamma \to \cdot}
-
-    \inferrule
-    {\gamma : \Delta \to \Gamma \and \Delta \vdash M : A[\gamma]}
-    {\gamma,M/X : \Delta \to \Gamma,X:A}
-
-    \inferrule
-    {\Gamma \vdash \Phi \isavectx \and
-    \Gamma \vdash \Psi \isavectx \and
-    d^+\Phi = d^- \Psi}
-    {\Gamma \vdash \Phi \jnctx \Psi \isavectx}
-
-    \inferrule
-    {\Gamma \vdash \phi : \Phi' \to \Phi \and
-     \Gamma \vdash \psi : \Psi' \to \Psi\and
-     d^+\phi = d^- \psi}
-    {\Gamma \vdash \phi \jnctx \psi : \Phi' \jnctx \Psi' \to \Phi \jnctx \Psi}
-
-    \inferrule*[right=ObSubst]
-    {\alpha : \cat C \vdash B : \cat D\and
-    \beta : \cat D \vdash A : \cat E}
-    {\alpha : \cat C \vdash A[B/\beta] : \cat E}\\
-
-    \inferrule*[right=Everything Subst]{\gamma : \Delta \to \Gamma \and \Gamma J}{\Delta J}
-
-    \inferrule{\Gamma \vdash \xi : Z \to \Xi\and \Gamma \pipe \Xi \vdash R \isaSet}{\Gamma \pipe Z \vdash R[\xi] \isaSet}
-
-    \inferrule{\Gamma \vdash \phi : \Psi \to \Phi\and \Gamma \pipe \Phi \vdash M : R}{\Gamma \pipe \Psi \vdash M[\phi] : R[\underline \phi]}
-  \end{mathpar}
-  \begin{align*}
-    (\Phi \jnctx \Psi) \jnctx \Sigma &= \Phi \jnctx (\Psi \jnctx \Sigma)\\
-    \Phi \jnctx \alpha:C &= \Phi\\
-    \alpha:C \jnctx \Psi &= \Psi\\\\
-    d^\pm(A/\alpha) &= A/\alpha\\
-    d^+(\phi_1,M/x,\phi_2) &= d^+ \phi_2\\
-    d^-(\phi_1,M/x,\phi_2) &= d^- \phi_1\\\\
-    \underline \cdot &= \cdot\\
-    \underline {A/\alpha} &= A/\alpha\\
-    \underline {\phi_1,M/x,\phi_2} &= d^-\phi_1;d^+\phi_2
-  \end{align*}
-  \caption{Admissible Rules}
 \end{figure}

### 2022-10-13 paper/paper.tex
add some cites

@@
 $\infty$-)category.  In directed type theories based on a bisimplicial
-model~\cite{riehl-shulman,weinberger,otherweinbegerpaper,licataweaver},
+model~\cite{riehlshulman17directed,buchholtzweinberger21fibered,weinberger22thesis,WeaverLicata20},
 morphism types are defined using an interval object, like in cubical
-type theory~\cite{bch,cchm,afh,abcfhl}, and universal properties like
-``morphism induction'' are an internally definable property of certain
-types.  Other type theories~\cite{north,ahrensnorth} define morphism
-types via an induction principle, corresponding to the lifting
-properties of certain kinds of fibrations of categories.  While these
-previous works can express some constructions on $\Cat$ that are not
-expressible in \vett{}, because \vett{} is more restricted, \vett{}
-contrariwise has more models, for instance categories enriched in
-non-cartesian monoidal categories, so the theorems that are provable
-in \vett{} apply in more settings.
+type
+theory~\cite{bch18,cchm18cubical,angiuli+18cartesian-csl,abcfhl17cartesian},
+and universal properties like ``morphism induction'' are an internally
+definable property of certain types.  Other type
+theories~\cite{north,ahrensnorth} define morphism types via an induction
+principle, corresponding to the lifting properties of certain kinds of
+fibrations of categories.  While these previous works can express some
+constructions on $\Cat$ that are not expressible in \vett{}, because
+\vett{} is more restricted, \vett{} contrariwise has more models, for
+instance categories enriched in non-cartesian monoidal categories, so
+the theorems that are provable in \vett{} apply in more settings.
 

### 2022-10-13 paper/paper.tex
more cites

@@
 definable property of certain types.  Other type
-theories~\cite{north,ahrensnorth} define morphism types via an induction
+theories~\cite{north18,ahrens+22semantics-2dtt} define morphism types via an induction
 principle, corresponding to the lifting properties of certain kinds of

### 2022-10-13 paper/paper.tex
more cites

@@
 2-level~\cite{voevodsky12,altenkirchXX} or indexed type
-theories~\cite{isaev,linearlf,vakar,krishnaswamibenton}.
+theories~\cite{isaev21indexed,cervesatopfenning02llf,vakar15linear,krishnaswami+15linear}.
 

### 2022-10-13 paper/paper.tex
work on substitutions in the appendix

--- 2022-10-13 paper/paper.tex
@@
   \end{align*}
+  \caption{Operations on Contexts}
+\end{figure}
+
+In Figure~\ref{fig:admissible}, we describe the admissible typing
+rules for the actions of substitution.
+
+\begin{definition}[Substitution Actions]
+  For any $\Gamma \pipe \alpha: \cat C \vdash a : \cat D$ and $\Gamma
+  \pipe \beta:\cat D \vdash b \cat E$, we define $b[a/\beta]$ by
+  recursion on $b$:
+
+  \begin{align*}
+    \beta[a/\beta] &= a\\
+    (M\,b)[a/\beta] &= M\,(b[a/\beta])\\
+    (\lambda \beta)    
+    (\pi_\pm b)[a/\beta] &= \pi_\pm b[a/\beta]\\
+    (b_-,b_+,s)[a/\beta] &= (b_-[a/\beta],b_+[a/\beta], s[a/\beta])\\
+  \end{align*}
+
+  Simultaneously by recursion on $s$ we define $s[\phi]$
+  \begin{align*}
+    x[a/\alpha,t,b/\beta] &= t\\
+    \pendappXtoY M b[a/\alpha] &= \pendappXtoY M {b[a/\alpha]}\\
+    {\punitelimkontZatABC {\alpha. t}{b_1} s {b_2}}[\phi] &=
+    {\punitelimkontZatABC {\alpha. t}{b_1[d^-\phi]} {s[\phi]} {b_2[d^+\phi]}}\\
+    (\punitrefl b)[a/\alpha] &= \punitrefl {b[a/\alpha]}\\
+    {\tensorelimWkontZ s {x.\beta.y.r}[\phi_l\jnctx\phi_m\jnctx\phi_r]} &=
+    \tensorelimWkontZ {s[\phi_m]} {x.\beta.y. r[\phi_l,x/x,\beta/\beta \jnctx \beta/\beta,y/y,\phi_r]}\\
+    {\tensorintroatXwithYandZ b s t}[\phi_s \jnctx \phi_t] &= \tensorintroatXwithYandZ {d^+\phi_s} {s[\phi_s]} {t[\phi_t]}\\
+    {(\homrappXtoYatZ s t a)}[\phi_f\jnctx \phi_a] &=
+    {\homrappXtoYatZ {s[\phi_f]} {t[\phi_a]} {a[d^+\phi_f]}}\\
+    {(\homrlambdaXatYdotZ {x}{\alpha} s)}[\phi] &=
+    {\homrlambdaXatYdotZ {x}{\alpha} {s[\phi,x/x,\alpha/\alpha]}}\\
+    {(\homlappXtoYatZ t s a)}[\phi_a\jnctx \phi_f] &=
+    {\homlappXtoYatZ {t[\phi_a]} {s[\phi_f]} {a[d^-\phi_f]}}\\
+    {(\homllambdaXatYdotZ {x}{\alpha} s)}[\phi] &=
+    {\homllambdaXatYdotZ {x}{\alpha} {s[\alpha/\alpha,x/x,\phi]}}\\
+    (\pi_i s)[\phi] &= \pi_i s[\phi]\\
+    (s_1,s_2)[\phi] &= (s_1[\phi], s_2[\phi])\\
+    ()[\phi] &= ()
+  \end{align*}
+
+  And finally, we define $P[\xi]$ by recursion on $P$:
+  \begin{align*}
+    (M\,a\,b)[\xi] &= M\,(a[d^-\xi])\,(b[d^+\xi])\\
+    (\punitinXfromYtoZ {\cat C} a b)[\xi] &= \punitinXfromYtoZ {\cat C} {a[\xi]}{b[\xi]}\\
+    (\tensorexistsXwithYandZ \beta P Q)[\xi] &= \tensorexistsXwithYandZ \beta {P[d^-\xi;\beta/\beta]}{P[\beta/\beta;d^+\xi]}\\
+    (\homrallXYtoZ \alpha R P)[\xi] &= \homrallXYtoZ \alpha {R[d^+\xi;\alpha/\alpha]} {P[d^-\xi;\alpha/\alpha]}\\
+    (\homlallXYtoZ \alpha P Q)[\xi] &= \homlallXYtoZ \alpha {P[\alpha/\alpha;d^-\xi]} {Q[\alpha/\alpha;d^+\xi]}\\
+    1[\xi] &= 1\\
+    (P_1 \times P_2)[\xi] &= P_1[\xi] \times P_2[\xi]
+  \end{align*}
+\end{definition}
+
+%% Theorems about substitution?
+
+%% \begin{definition}[Substitution of Objects in Sets]
+%%   3
+%% \end{definition}
+%% \begin{definition}[Substitution of Elements in Elements]
+%%   3
+%% \end{definition}
+
+\begin{figure}
+  \begin{mathpar}
+    \inferrule
+    {}
+    {\cdot : \Gamma \to \cdot}
+
+    \inferrule
+    {\gamma : \Delta \to \Gamma \and \Delta \vdash M : A[\gamma]}
+    {\gamma,M/X : \Delta \to \Gamma,X:A}
+
+    \inferrule
+    {\Gamma \vdash \Phi \isavectx \and
+    \Gamma \vdash \Psi \isavectx \and
+    d^+\Phi = d^- \Psi}
+    {\Gamma \vdash \Phi \jnctx \Psi \isavectx}
+
+    \inferrule
+    {\Gamma \vdash \phi : \Phi' \to \Phi \and
+     \Gamma \vdash \psi : \Psi' \to \Psi\and
+     d^+\phi = d^- \psi}
+    {\Gamma \vdash \phi \jnctx \psi : \Phi' \jnctx \Psi' \to \Phi \jnctx \Psi}
+
+    \inferrule*[right=ObSubst]
+    {\alpha : \cat C \vdash B : \cat D\and
+    \beta : \cat D \vdash A : \cat E}
+    {\alpha : \cat C \vdash A[B/\beta] : \cat E}\\
+
+    \inferrule*[right=Everything Subst]{\gamma : \Delta \to \Gamma \and \Gamma J}{\Delta J}
+
+    \inferrule*[right=BoundarySubst]{\Gamma \vdash \xi : Z \to \Xi\and \Gamma \pipe \Xi \vdash R \isaSet}{\Gamma \pipe Z \vdash R[\xi] \isaSet}
+
+    \inferrule*[right=TransSubst]{\Gamma \vdash \phi : \Psi \to \Phi\and \Gamma \pipe \Phi \vdash s : R}{\Gamma \pipe \Psi \vdash s[\phi] : R[\underline \phi]}
+  \end{mathpar}
+  \begin{align*}
+    (\Phi \jnctx \Psi) \jnctx \Sigma &= \Phi \jnctx (\Psi \jnctx \Sigma)\\
+    \Phi \jnctx \alpha:C &= \Phi\\
+    \alpha:C \jnctx \Psi &= \Psi\\\\
+    d^\pm(A/\alpha) &= A/\alpha\\
+    d^+(\phi_1,M/x,\phi_2) &= d^+ \phi_2\\
+    d^-(\phi_1,M/x,\phi_2) &= d^- \phi_1\\\\
+    \underline \cdot &= \cdot\\
+    \underline {A/\alpha} &= A/\alpha\\
+    \underline {\phi_1,M/x,\phi_2} &= d^-\phi_1;d^+\phi_2
+  \end{align*}
+  \caption{Admissible Rules}
+\end{figure}
+
+\cref{fig:contexts} details the typing rules for contexts and the
+typing for variables and the admissible actions of substitution (the
+definition of substitution is straightforward and found in the
+appendix). the rules for term and object variables are as usual.
+%
+An object substitution into a single object context $\alpha:\cat C$
+consists of an object $a : \cat C$ parameterized by an object of
+another category.
+%
+An object substitution into a pair of co- and contra-variant variables
+$\alpha:\cat C;\beta:\cat D$ is a pair of an object substituted for
+$\alpha$ and one for $\beta$, each typed with respect to the
+appropriate variable of an object context $\Xi$.
+%
+Note that this allows for $\Xi$ to consist of either two variables or
+a single variable.
+%
+When
+constructing a transformation context $\Phi, x : R, \Psi$, we need to
+ensure that the contravariant dependence of $R$ matches the covariant
+dependence of $\Phi$ and vice-versa for $\Psi$. We define this with
+auxiliary metafunctions $d^+,d^-$ which extract the covariant and
+contravariant object variable of a transformation context.
+%
+The rules for transformation variables and substitutions define the
+use of variables to be \emph{linear} (each variables is used exactly
+once) and \emph{ordered}, they occur in the term in the same order
+that they appear in the context.
+%
+The rule for a transformation variables says a variable $x$ can only
+be used when it is the sole variable in the context.
+%
+A transformation subsitution into an ``empty'' transformation context
+$\alpha: \cat C$ is the same as an object substitution into
+$\alpha:\cat C$.
+%
+The rule for constructing a transformation substitution into a
+non-empty context $\Phi_1,x:R,\Phi_2$ says we can construct a
+substitution by constructing substitutions $\phi_1,\phi_2$ for
+$\Phi_1$ and $\Phi_2$ as well as a transformation $s$ into an
+instantiation of $R$ where the input contexts come from splitting the
+input context $\Psi$ into three components, and ensuring that the
+instantiation of $R$ agrees with how $\phi_1,\phi_2$ instantiate the
+variables. This uses two new metafunctions. First, $\Phi \jnctx \Psi$
+is an operation for concatenating two contexts when $d^+(\Phi) =
+d^-(\Psi)$. Next, we extend the boundary maps $d^\pm$ to substitutions
+to extract their object substitutions.
+
+\begin{figure}
+
   \caption{Basic Rules of Judgment Theory}

### 2022-10-14 paper/paper.tex
appendix work

--- 2022-10-13 paper/paper.tex
@@
 
-In Figure~\ref{fig:subst} we give full definitions for well-formedness
-of contexts and substitutions.
+In Figure~\ref{fig:ctx} we include the formation rules and definitions
+of the three kinds of contexts that are used in \vett{}.
+%
+In Figure~\ref{fig:substitutions} we give definitions for
+well-typedness of the corresponding three kinds of substitutions.
+
+\begin{figure}
+  \begin{mathpar}
+    \inferrule*[right=TyCtxForm]
+    {}
+    {\Gamma \isadtctx}
+
+    \inferrule*[right=MtTyCtx]{}{\cdot \isadtctx}
+
+    \inferrule*[right=TyCtxExt]
+    {\Gamma \isadtctx \and \Gamma \vdash A \isaTy}
+    {\Gamma , X : A \isadtctx}
+
+    \inferrule*[right=BoundaryForm]
+    {\Gamma \isadtctx}
+    {\Gamma \vdash \Xi \boundary}
+
+    \inferrule*[right=BoundarySingle]
+    {\Gamma \vdash \cat C \isaCat}
+    {\Gamma \vdash \alpha: \cat C \boundary}
+
+    \inferrule*[right=BoundaryDbl]
+    {\Gamma \vdash \cat C \isaCat \and \Gamma\vdash\cat D \isaCat}
+    {\Gamma \vdash \alpha: \cat C; \beta: \cat D \isaboundary}
+
+    \inferrule*[right=SetCtxForm]
+    {\Gamma \isadtctx}
+    {\Gamma \vdash \Phi \isavectx}
+
+    \inferrule*[right=SetCtxMt]
+    {\Gamma \vdash \cat C \isaCat}
+    {\Gamma \vdash \alpha: \cat C \isavectx}
+
+    \inferrule*[right=SetCtxExt]
+    {\Gamma \vdash \Phi \isavectx\and
+    \Gamma \pipe d^+\Phi; \beta:\cat D \vdash R \isaSet}
+    {\Gamma \vdash \Phi, x:R, \beta:\cat D \isavectx}
+  \end{mathpar}
+  \caption{Contexts}
+  \label{fig:ctx}
+\end{figure}
 
@@
 
-      \inferrule*[right=TyCtxForm]
+    \inferrule*[right=TermSubstFormation]
+    {\Delta \isadtctx \and \Gamma \isadtctx}
+    {\Delta \vdash \gamma :: \Gamma}
+
+    \inferrule*[right=TermSubstMt]
     {}
-        {\Gamma \isadtctx}
-
-    \inferrule*[right=SetCtxForm]
-    {\Gamma \isadtctx}
-    {\Gamma \vdash \Xi \boundary}
-
-    \inferrule*[right=TransCtxForm]
-    {\Gamma \isadtctx}
-     {\Gamma \vdash \Phi \isavectx}
-    \inferrule
-    {\Delta \isadtctx \and \Gamma \isadtctx}
-    {\gamma : \Delta \to \Gamma}
-
-    \inferrule*[right=Boundary Map formation]
+    {\cdot : \Gamma \to \cdot}
+
+    \inferrule*[right=TermSubstExt]
+    {\gamma : \Delta \to \Gamma \and \Delta \vdash M : A[\gamma]}
+    {\gamma,M/X : \Delta \to \Gamma,X:A}
+
+    \inferrule*[right=BoundarySubstFormation]
     {\Gamma \vdash Z \boundary \and \Gamma \vdash \Xi \boundary}
-    {\Gamma \vdash \xi : Z \to \Xi}
-
-    \inferrule*[right=Cat Subst formation]
+    {\Gamma\pipe Z \vdash \xi :: \Xi}
+
+    \inferrule*[right=BoundarySubstSingle]
+    {\Gamma\pipe \alpha:\cat C \vdash b : \cat D}
+    {\Gamma \pipe \alpha:\cat C \vdash b/\beta :: \beta:\cat D}
+
+    \inferrule*[right=BoundarySubstDbl]
+    {\Gamma\pipe d^-\Xi \vdash a : \cat C\and
+     \Gamma\pipe d^+\Xi \vdash b : \cat D}
+    {\Gamma \pipe \Xi \vdash a/\alpha;b/\beta :: (\alpha : \cat C; \beta:\cat D)}
+
+    \inferrule*[right=ElementSubstFormation]
     {\Gamma \vdash \Psi \isavectx \and \Gamma \vdash \Phi \isavectx}
-    {\Gamma \vdash \phi : \Psi \to \Phi}
-  \end{mathpar}
-  \begin{mathpar}
-    \inferrule{}{\cdot \isadtctx}
-
-    \inferrule
-    {\Gamma \isadtctx \and \Gamma \vdash A \isaTy}
-    {\Gamma , X : A \isadtctx}\\
-
-    \inferrule
-    {X : A \in \Gamma}
-    {\Gamma \vdash X : A}
-
-    \inferrule
-    {}
-    {\alpha: \cat C \vdash \alpha: \cat C}
-
-    \inferrule
-    {\Gamma \vdash \cat C : \isaCat}
-    {\Gamma \vdash \alpha: \cat C \isavectx}
-
-    \inferrule
-    {\Gamma \vdash \Phi \isavectx\and
-    \Gamma \pipe d^+\Phi; \beta:\cat D \vdash R \isaSet}
-    {\Gamma \vdash \Phi, x:R, \beta:\cat D \isavectx}
-
-    \inferrule
-    {}
-    {\Gamma \pipe \alpha:\cat C, x : R, \beta: \cat D \vdash x : R}
+    {\Gamma\pipe \Psi \vdash \phi :: \Phi}
 
     % horizontal identity
-    \inferrule
+    \inferrule*[right=ElementSubstMt]
     {\Gamma\pipe \beta:\cat C \vdash a : \cat C}
@@
 
-    \inferrule
-    {\Psi = \Psi_1 \jnctx \Psi_2\and
-      \Gamma\pipe\Psi_1 \vdash \phi : \Phi\and
-      \Gamma \pipe \Psi_2 \vdash t : R[d^+\phi;b/\beta] \and
-      \Gamma \vdash \phi_2 : \Psi_2 \to \Phi_2
-      }
+    \inferrule*[right=ElementSubstExt]
+    {\Gamma\pipe\Psi \vdash \phi :: \Phi \and
+      d^+\Psi = d^-\Psi'
+     \Gamma\pipe\Psi' \vdash t : R[d^+\phi;b/\beta]\and
+     \Gamma\pipe d^+\Psi'\vdash b : \cat D}
     {\Gamma\pipe\Psi \vdash \phi,t/x,b/\beta :: (\Phi,x:R,\beta:\cat D)}
   \end{mathpar}
-
   \caption{Substitution}
+  \label{fig:substitions}
 \end{figure}
 
-\begin{figure}
-  \begin{align*}
-    \underline{\alpha:C} &= \alpha:C\\
-    \underline{\Phi,x:R,\Psi} &= d^-\Phi; d^+{\Psi}
-  \end{align*}
+These definitions involve several operations $d^\pm\Phi$, $\underline
+\Phi$ and $\Phi\jnctx \Psi$ on contexts (and their functorial lift to
+substitutions) that we now define.
+
+\begin{definition}
+  We define operations $d^\pm$ that project out the covariant and
+  contravariant boundary of a set context.
+  This can be typed with the admissible rule
+  \begin{mathpar}
+    \inferrule*[right=(*)]
+               {\Gamma \vdash \Phi \vectx}
+               {\Gamma \vdash d^\pm\Phi \boundary}
+  \end{mathpar}
+  and is defined as
   \begin{align*}
@@
   \end{align*}
+  This operation extends to the substitutions with admissible rule
+  \begin{mathpar}
+    \inferrule*[right=(*)]
+    {\Gamma\pipe\Psi \vdash \phi :: \Phi}
+    {\Gamma\pipe d^\pm \Psi \vdash d^\pm\phi :: d^\pm \Phi}
+  \end{mathpar}
+  defined as
   \begin{align*}
@@
   \end{align*}
-  \caption{Operations on Contexts}
-\end{figure}
-
-In Figure~\ref{fig:admissible}, we describe the admissible typing
-rules for the actions of substitution.
-
+\end{definition}
+
+\begin{definition}
+  We define the operation of restricting a set context to both sides
+  of its boundary with admissible typing
+  \begin{mathpar}
+    \inferrule*[right=(*)]
+    {\Gamma \vdash \Phi \isavectx}
+    {\Gamma \vdash \underline\Phi \boundary}
+  \end{mathpar}
+  and definition
+  \begin{align*}
+    \underline{\alpha:C} &= \alpha:C\\
+    \underline{\Phi,x:R,\Psi} &= d^-\Phi; d^+{\Psi}
+  \end{align*}
+  The extension to substitutions has admissible typing
+  \begin{mathpar}
+    \inferrule*[right=(*)]
+    {\Gamma \pipe \Psi \vdash \phi :: \Phi}
+    {\Gamma\pipe\underline \vdash \Phi \boundary}
+  \end{mathpar}
+  and definition
+  \begin{align*}
+    \underline{\alpha:C} &= \alpha:C\\
+    \underline{\Phi,x:R,\Psi} &= d^-\Phi; d^+{\Psi}
+  \end{align*}
+\end{definition}
+
+% RETURN
+Finally, we define the operation of ``horizontal composition'' of set
+contexts $\Phi \jnctx \Psi$ and its functorial lift $\phi \jnctx
+\psi$. Note that the horizontal composition is only well-defined when $$
+
+Next, we define the actions of substitutions on terms. We elide the
+obvious action of term substitutions $\gamma$ and include only the
+more unusual substructural substitutions.
 \begin{definition}[Substitution Actions]
@@
 
+\begin{lemma}[Properties of Substitution]
+  Substitution is associative and unital?
+\end{lemma}
+
 %% Theorems about substitution?
-
-%% \begin{definition}[Substitution of Objects in Sets]
-%%   3
-%% \end{definition}
-%% \begin{definition}[Substitution of Elements in Elements]
-%%   3
-%% \end{definition}
 
@@
   \begin{mathpar}
-    \inferrule
-    {}
-    {\cdot : \Gamma \to \cdot}
-
-    \inferrule
-    {\gamma : \Delta \to \Gamma \and \Delta \vdash M : A[\gamma]}
-    {\gamma,M/X : \Delta \to \Gamma,X:A}
 

### 2022-10-14 paper/paper.tex
fix unicode issue/bugs
fix terminology skew with how contexts are named in Fig 1

--- 2022-10-14 paper/paper.tex
@@
 
-\paragraph{Basics of Profunctors.}
-While we assume the reader has some background knowledge of category
-theory, we briefly define profunctors, which are not included in many
-introductory texts.  Recall that a category $\cat C$ has a collection of
-objects and morphisms with identity and composition, and a functor $F :
-\cat C \to \cat D$ is a function on objects and a function on morphisms
-that preserves identity and composition.  A category can be thought of
-as a generalization of a preordered set, which has a set of elements and
-a binary \emph{relation} on its objects satisfying reflexivity and
-transitivity.  A category is then a \emph{proof-relevant preorder},
-where morphisms are the proofs of ordering, and the reflexivity and
-transitivity proofs must identity and unit equations. A functor is then
-a \emph{proof-relevant monotone function}.  Given categories $\mathcal
-C$ and $\mathcal D$, a profunctor $R$ from $\mathcal C$ to $\mathcal D$,
-written $R : \cat C \pto \cat D$ is a functor $R : {\cat C}^{op} \times
-{\cat D} \to \Set$.  Because a profunctor outputs a $\Set$ rather than a
-proposition, it is itself a \emph{proof-relevant relation}.  Thinking of
-categories as proof-relevant preorders, functoriality says that the
-profunctor is downward-closed in $\cat C$ and upward-closed in $\cat D$.
-Given profunctors $R, S : \cat C \pto \cat D$, a homomorphism from
-$R$ to $S$ is simply a natural transformation, which in the preordered
-setting is simply an implication of relations.
+\textbf{Basics of Profunctors.}  While we assume the reader has some
+background knowledge of category theory, we briefly define profunctors,
+which are not included in many introductory texts.  Recall that a
+category $\cat C$ has a collection of objects and morphisms with
+identity and composition, and a functor $F : \cat C \to \cat D$ is a
+function on objects and a function on morphisms that preserves identity
+and composition.  A category can be thought of as a generalization of a
+preordered set, which has a set of elements and a binary \emph{relation}
+on its objects satisfying reflexivity and transitivity.  A category is
+then a \emph{proof-relevant preorder}, where morphisms are the proofs of
+ordering, and the reflexivity and transitivity proofs must identity and
+unit equations. A functor is then a \emph{proof-relevant monotone
+  function}.  Given categories $\mathcal C$ and $\mathcal D$, a
+profunctor $R$ from $\mathcal C$ to $\mathcal D$, written $R : \cat C
+\pto \cat D$ is a functor $R : {\cat C}^{op} \times {\cat D} \to \Set$.
+Because a profunctor outputs a $\Set$ rather than a proposition, it is
+itself a \emph{proof-relevant relation}.  Thinking of categories as
+proof-relevant preorders, functoriality says that the profunctor is
+downward-closed in $\cat C$ and upward-closed in $\cat D$.  Given
+profunctors $R, S : \cat C \pto \cat D$, a homomorphism from $R$ to $S$
+is simply a natural transformation, which in the preordered setting is
+simply an implication of relations.
 
@@
 we write ${\Gamma \isadtctx}$ to indicate that a context is well-formed.
-The category contexts $\Gamma \vdash \Xi \boundary$ contain object
-variables with their categories; the two forms of category context are
-$\alpha : \cat C$, containing one variable that can be used both
-contravariantly and covariantly, and $\alpha : \cat C ; \beta : \cat D$,
-containing a contravariant variable $\alpha$ and covariant variable
-$\beta$.  Finally, we have contexts $\Phi$ of element variables with
-their sets, alternating with those sets' object variables.  A typical
-$\Phi$ has the shape
+We name the remaining contexts after the judgements that they are used
+by.  The set contexts $\Gamma \vdash \Xi \boundary$, which will be used
+to type-check sets, contain object variables with their categories. The
+two forms of set context are $\alpha : \cat C$, containing one variable
+that can be used both contravariantly and covariantly, and $\alpha :
+\cat C ; \beta : \cat D$, containing a contravariant variable $\alpha$
+and covariant variable $\beta$.  Finally, the transformation contexts
+$\Gamma \vdash \Phi \isavectx$ contain element variables with their sets, alternating with those
+sets' object variables with their categories.  A typical $\Phi$ has the
+shape
 \[
 \alpha_1 : \cat C_1, x_1:R_1(\alpha_1,\alpha_2), \alpha_2 : \cat C_2,
-x_2:R_2(\alpha_2,\alpha_3), \ldots, R_n(\alpha_n,\alpha_{n+1}), \alpha_{n+1} : C_{n+1}
+x_2:R_2(\alpha_2,\alpha_3), \ldots, R_n(\alpha_n,\alpha_{n+1}),
+\alpha_{n+1} : \cat C_{n+1}
 \]
-and represents the composition of the relations $R_1, R_2, R_3, \ldots,
+and represents the composition of the ``relations'' $R_1, R_2, R_3, \ldots,
 R_n$.  We write $d^-(\Phi)$ for the first category variable in $\Phi$
@@
       \textrm{Type Context} & \Gamma, \Delta & ::= & \cdot \pipe \Gamma, X : A\\
-      \textrm{Category Context} & \Xi, Z & ::= & \alpha:\cat C \pipe \alpha:\cat C;\beta: \cat D\\
-      \textrm{Set Context} & \Phi,\Psi & ::= & \alpha : \cat C \pipe \Phi , x : P,\beta:\cat D\\
+      \textrm{Set Context} & \Xi, Z & ::= & \alpha:\cat C \pipe \alpha:\cat C;\beta: \cat D\\
+      \textrm{Transformation Context} & \Phi,\Psi & ::= & \alpha : \cat C \pipe \Phi , x : P,\beta:\cat D\\
     \end{array}
@@
   \underline \Phi \vdash R \isaSet$.  A transformation $s$ has a context
-  of transformation variable and a single set output. To be well formed,
-  the context and set must be parameterized by the same contravariant
-  and covariant object variables. To ensure this, we use a coercion
-  operation $\underline \Phi$ that erases everything in the context but
-  the left-most and right-most object variables, i.e. $\underline{\alpha
-    : \cat C} = \alpha : \cat C$ and $\underline{\Phi} =
-  d^-(\Phi);d^+(\Phi)$ otherwise.
+  $\Phi$ of transformation variable and a single set output $R$. To be
+  well formed, the context and set must be parameterized by the same
+  contravariant and covariant object variables. To ensure this, we use a
+  coercion operation $\underline \Phi$ from transformation contexts to
+  set contexts that erases everything in the context but the left-most
+  and right-most object variables ($\underline{\alpha : \cat C} = \alpha
+  : \cat C$ and $\underline{\Phi} = d^-(\Phi);d^+(\Phi)$).
 
@@
 where the endpoints $A$ and $B$ of $\punitinXfromYtoZ{\cat C} A B$ are
-not fully general (i.e. not distinct variables).  However, the instances
-of this more general rule that do type check are derivable from our more
-restricted rule using right/left-hom types.
+not distinct variables.  However, the instances of this more general
+rule that do type check are derivable from our more restricted rule
+using right/left-hom types.
 

### 2022-10-14 paper/paper.tex

--- 2022-10-14 paper/paper.tex
@@
 \usepackage{cleveref}
+\usepackage[utf8]{inputenc}
 
@@
     {\Gamma \vdash \cat C \isaCat \and \Gamma\vdash\cat D \isaCat}
-    {\Gamma \vdash \alpha: \cat C; \beta: \cat D \isaboundary}
+    {\Gamma \vdash \alpha: \cat C; \beta: \cat D \boundary}
 
@@
     \inferrule*[right=(*)]
-               {\Gamma \vdash \Phi \vectx}
+               {\Gamma \vdash \Phi \isavectx}
                {\Gamma \vdash d^\pm\Phi \boundary}
@@
 contexts $\Phi \jnctx \Psi$ and its functorial lift $\phi \jnctx
-\psi$. Note that the horizontal composition is only well-defined when $$
+\psi$. Note that the horizontal composition is only well-defined when
 

### 2022-10-14 paper/paper.tex
progress on appendix, a few fixes to the semantics sections

@@
 First, we demonstrate that a hyperdoctrine of virtual equipments
-categories with split restrictions provides a sound and complete
-notion of categorical model.
+provides a sound and complete notion of categorical model.
 %
@@
 \ohol{} can be interpreted in ordinary category theory as well as
-enriched and internal category theory, and more trivially, in terms of
-sets and relations or preordered sets and ordered relations.
+enriched, internal and indexed notions.
 
@@
 First, we can model the judgmental structure of the unary type theory
-and predicate logic in certain\emph{virtual equipments}
+and predicate logic in certain \emph{virtual equipments}
 \cite{leinster,crutwell-shulman}, i.e., virtual double categories with
@@
 substitution, we use a variation on Lawvere's notion of
-\emph{hyperdoctrine} for modeling predicate logic:
+\emph{hyperdoctrine} for modeling predicate logic\cite{Lawvere69}:
 \begin{definition}[Hyperdoctrine of Virtual Equipments]
@@
 $\Pi,\Sigma,\text{Id}$ types correspond to their standard semantics in
-a CwF, products of categories interpreted as products in the vertical
-category, and products of sets as products in the category of
-pro-arrows and 2-cells. The units, tensor and covariant and
-contravariant homs are modeled by the universal properties of the same
-names, as described for instance in \citet{shulmanFrBicats}. The graph
-of a profunctor is modeled by \emph{tabulators}
-\citet{grandis-pare}. Finally, the covariant and contravariant
-presheaf categories can be described as a weakening of the definition
-of a Yoneda equipment from \citet{yosegi-boxes} to virtual double
-categories. More detailed descriptions of these universal properties
-are included in the appendix.
+a CwF and the types of categories, functors etc as representing
+objects for the virtual equipment structures. Products of categories
+are interpreted as products in the vertical category, and products of
+sets as products in the category of pro-arrows and 2-cells. The units,
+tensor and covariant and contravariant homs are modeled by the
+universal properties of the same names, as described for instance in
+\citet{shulmanFrBicats}. The graph of a profunctor is modeled by
+\emph{tabulators} \citet{grandis-pare}. Finally, the covariant and
+contravariant presheaf categories can be described as a weakening of
+the definition of a Yoneda equipment from \citet{yosegi-boxes} to
+virtual double categories. More detailed descriptions of these
+universal properties are included in the appendix.
 %
@@
 
-\subsection{Substitution}
+\subsection{Contexts and Substitutions}
 
@@
     {}
-    {\cdot : \Gamma \to \cdot}
+    {\Delta \vdash \cdot :: \cdot}
 
     \inferrule*[right=TermSubstExt]
-    {\gamma : \Delta \to \Gamma \and \Delta \vdash M : A[\gamma]}
-    {\gamma,M/X : \Delta \to \Gamma,X:A}
+    {\Delta \vdash \gamma :: \Gamma \and \Delta \vdash M : A[\gamma]}
+    {\Delta \vdash \gamma,M/X :: \Gamma,X:A}
 
@@
   \end{align*}
+  Several rules assume the substitution is in a particular form, such
+  as the tensor elimination which expects an input context
+  $\phi_s\jnctx\phi_t$. The fact that the context can be uniquely
+  decomposed in a well-typed way follows from an inversion principle (Lemma~\ref{lem:inversion})
+  for well-typed substitutions.
 
@@
 
+\begin{lemma}[Inversion]
+  \label{lem:inversion}
+  \begin{enumerate}
+  \item If $\Phi \vdash \psi :: (\alpha:\cat C)$ then $\Phi =
+    \beta:\cat D$ for some $\cat D$ and $\psi = a/\alpha$ where
+    $\beta:\cat D \vdash a : \cat C$.
+  \item If $\Phi \vdash \psi :: \Psi_1 \jnctx \Psi_2$, then there
+    exists unique $\Phi_1,\Phi_2,\psi_1,\psi_2$ such that $\Phi =
+    \Phi_1\jnctx\Phi_2$ and $\Phi_1\vdash\psi_1 :: \Psi_1$ and
+    $\Phi_2\vdash \psi_2 : \Psi_2$ and $\psi = \psi_1\jnctx\psi_2$.
+  \end{enumerate}
+\end{lemma}
+
+\begin{lemma}[Vertical category]
+  \label{lem:vertical-cat}
+  Substitution of object terms $A[B/\beta]$ is associative with
+  identity given by the variable $\alpha:\cat C \vdash \alpha:\cat C$.
+\end{lemma}
+
+To define vertical composition of squares, we need the following lemma.
+
+\begin{lemma}[Horizontal Category of Contexts]
+  Horizontal composition of contexts is associative (when defined)
+  \[ (\Phi \jnctx \Psi) \jnctx \Sigma = \Phi \jnctx (\Psi \jnctx \Sigma) \]
+  and unital with identity for $\cat C$ given by the single category
+  variable context $\alpha:\cat C$:
+  \[ \alpha:\cat C \jnctx \Phi = \Phi = \Phi \jnctx \beta:\cat D\]
+  when $d^-\Phi = \alpha:\cat C$ and $d^+\Phi = \beta:\cat D$.
+
+  These properties extend to the horizontal composition of element substitutions:
+  \[ \phi \jnctx (\psi \jnctx \sigma) = (\phi \jnctx \psi) \jnctx \sigma \]
+  where the identity is the single variable substitution:
+  \[ a/\alpha \jnctx \phi = \phi = \phi \jnctx b/\beta \]
+  when $d^-\phi = a/\alpha$ and $d^+\phi = b/\beta$.
+\end{lemma}
+
 \begin{lemma}[Properties of Substitution]
-  Substitution is associative and unital?
+  TODO
 \end{lemma}
@@
 %% Theorems about substitution?
-
-\begin{figure}
-  \begin{mathpar}
-
-    \inferrule
-    {\Gamma \vdash \Phi \isavectx \and
-    \Gamma \vdash \Psi \isavectx \and
-    d^+\Phi = d^- \Psi}
-    {\Gamma \vdash \Phi \jnctx \Psi \isavectx}
-
-    \inferrule
-    {\Gamma \vdash \phi : \Phi' \to \Phi \and
-     \Gamma \vdash \psi : \Psi' \to \Psi\and
-     d^+\phi = d^- \psi}
-    {\Gamma \vdash \phi \jnctx \psi : \Phi' \jnctx \Psi' \to \Phi \jnctx \Psi}
-
-    \inferrule*[right=ObSubst]
-    {\alpha : \cat C \vdash B : \cat D\and
-    \beta : \cat D \vdash A : \cat E}
-    {\alpha : \cat C \vdash A[B/\beta] : \cat E}\\
-
-    \inferrule*[right=Everything Subst]{\gamma : \Delta \to \Gamma \and \Gamma J}{\Delta J}
-
-    \inferrule*[right=BoundarySubst]{\Gamma \vdash \xi : Z \to \Xi\and \Gamma \pipe \Xi \vdash R \isaSet}{\Gamma \pipe Z \vdash R[\xi] \isaSet}
-
-    \inferrule*[right=TransSubst]{\Gamma \vdash \phi : \Psi \to \Phi\and \Gamma \pipe \Phi \vdash s : R}{\Gamma \pipe \Psi \vdash s[\phi] : R[\underline \phi]}
-  \end{mathpar}
-  \begin{align*}
-    (\Phi \jnctx \Psi) \jnctx \Sigma &= \Phi \jnctx (\Psi \jnctx \Sigma)\\
-    \Phi \jnctx \alpha:C &= \Phi\\
-    \alpha:C \jnctx \Psi &= \Psi\\\\
-    d^\pm(A/\alpha) &= A/\alpha\\
-    d^+(\phi_1,M/x,\phi_2) &= d^+ \phi_2\\
-    d^-(\phi_1,M/x,\phi_2) &= d^- \phi_1\\\\
-    \underline \cdot &= \cdot\\
-    \underline {A/\alpha} &= A/\alpha\\
-    \underline {\phi_1,M/x,\phi_2} &= d^-\phi_1;d^+\phi_2
-  \end{align*}
-  \caption{Admissible Rules}
-\end{figure}
-
-\cref{fig:contexts} details the typing rules for contexts and the
-typing for variables and the admissible actions of substitution (the
-definition of substitution is straightforward and found in the
-appendix). the rules for term and object variables are as usual.
-%
-An object substitution into a single object context $\alpha:\cat C$
-consists of an object $a : \cat C$ parameterized by an object of
-another category.
-%
-An object substitution into a pair of co- and contra-variant variables
-$\alpha:\cat C;\beta:\cat D$ is a pair of an object substituted for
-$\alpha$ and one for $\beta$, each typed with respect to the
-appropriate variable of an object context $\Xi$.
-%
-Note that this allows for $\Xi$ to consist of either two variables or
-a single variable.
-%
-When
-constructing a transformation context $\Phi, x : R, \Psi$, we need to
-ensure that the contravariant dependence of $R$ matches the covariant
-dependence of $\Phi$ and vice-versa for $\Psi$. We define this with
-auxiliary metafunctions $d^+,d^-$ which extract the covariant and
-contravariant object variable of a transformation context.
-%
-The rules for transformation variables and substitutions define the
-use of variables to be \emph{linear} (each variables is used exactly
-once) and \emph{ordered}, they occur in the term in the same order
-that they appear in the context.
-%
-The rule for a transformation variables says a variable $x$ can only
-be used when it is the sole variable in the context.
-%
-A transformation subsitution into an ``empty'' transformation context
-$\alpha: \cat C$ is the same as an object substitution into
-$\alpha:\cat C$.
-%
-The rule for constructing a transformation substitution into a
-non-empty context $\Phi_1,x:R,\Phi_2$ says we can construct a
-substitution by constructing substitutions $\phi_1,\phi_2$ for
-$\Phi_1$ and $\Phi_2$ as well as a transformation $s$ into an
-instantiation of $R$ where the input contexts come from splitting the
-input context $\Psi$ into three components, and ensuring that the
-instantiation of $R$ agrees with how $\phi_1,\phi_2$ instantiate the
-variables. This uses two new metafunctions. First, $\Phi \jnctx \Psi$
-is an operation for concatenating two contexts when $d^+(\Phi) =
-d^-(\Psi)$. Next, we extend the boundary maps $d^\pm$ to substitutions
-to extract their object substitutions.
-
-\begin{figure}
-
-  \caption{Basic Rules of Judgment Theory}
-  \label{fig:contexts}
-\end{figure}
-
-Now that we have established a judgmental structure, in typical type
-theoretic fashion, our type constructs follow naturally as
-internalizing the judgments. Though the positive multiplicatives are
-more familiar, we start by describing the negative multiplicatives,
-since they are better behaved, and are useful to have already defined
-when discussing the positives.
 
@@
 TODO
-
-\subsection{Negative Multiplicatives}
-
-Since our calculus of natural transformations is ordered linear, we
-have two distinct ``function sets'', internalizing concatenation on
-the left or right of the context. The typing rules for these are given
-in \cref{fig:cotensor}. The two types are the \emph{right hom}
-$\homrallXYtoZ \alpha R P$ and the \emph{left hom} $\homlallXYtoZ
-\alpha R P$. As a type, we can think of these both as a kind of
-polymorphic function type in that in addition to parameterizing an
-element of the set $P$ by an element of the set $R$, they also
-quantify over an object $\alpha$ that is in $P$ and $R$.
-
-These behave like function types, but their typing is somewhat
-involved. First consider the right hom $\homrallXYtoZ \alpha R
-P$. This acts very much like a polymorphic function type $\forall
-\alpha. R \to P$, but with restrictions on the object
-variables. First, the object variable $\alpha$ must occur covariantly
-in $P$ and contravariantly in $R$. Next, the contravariant variable of
-the type is the contravariant variable of $P$ and the covariant
-variable of the type is the \emph{contravariant} variable of $R$, with
-the flip in variance intuitively justified since $R$ occurs in the
-negative position of the function. The placement of the $\forall
-\alpha$ on the right side of the triangle is intended to be suggestive
-of the fact that $\alpha$ occurs \emph{covariantly} in the input and
-output sets $R,P$ and also that the corresponding $\lambda$ introduces
-variables on the right.
-%
-As you would expect from a polymorphic function type, the introduction
-form is a $\lambda$ that abstracts over both a transformation variable
-and an object variable, which are concatenated on the right of the
-context, aligning with the condition that the contravariant variable
-of $R$ match the covariant variable of the original context.
-%
-The elimination rule is a combined instantiation and application form,
-$\homrappXtoYatZ s t a$ which instantiates the object $\alpha$ to $a$
-and passes an element $t : R[a/\alpha]$ as input, yielding an element
-$P[a/\alpha]$ as output. Finally, we have the expected $\beta$ and
-$\eta$ reasoning principles that we would have for a function type.
-
-The left hom $\homlallXYtoZ \alpha R P$ dually abstracts over context
-extension on the left. This time, the variable $\alpha$ occurs
-\emph{contravariantly} in the input and output sets, and the covariant
-dependence of the function set matches the covariant dependence of the
-ouput set, while the contravariant dependence is the covariant
-dependence of the input. In our notation, we place the $\forall\alpha$
-on the left to be suggestive of the fact that it is contravariant in
-$R$ and $P$. Next, the introduction form is again a polymorphic
-$\lambda$, but this time we write the object variable $\alpha$ first,
-and then the set variable $x$, to match the structure of the
-context. The elimination form is similarly an application, but to
-agree with the ordered use of variables, this time we place the
-argument to the left and the function to the right.
-
-\begin{figure}
-  % TODO: formalize the one-place substitution
-  \begin{mathpar}
-      \inferrule*[right=Covariant Hom formation]
-      {\Gamma \pipe d^+\Xi; \alpha : \cat C \vdash  R \and \Gamma \pipe  d^-\Xi; \alpha : \cat C \vdash P}
-      {\Gamma \pipe \Xi \vdash \homrallXYtoZ \alpha R P}
-
-      \inferrule*[right=Covariant Hom introduction]
-      {\Gamma \pipe \Phi, x : R, \alpha : \cat C \vdash t : P}
-      {\Gamma \pipe \Phi \vdash
-        \homrlambdaXatYdotZ {x : R} {\alpha : \cat C} t : \homrallXYtoZ \alpha R P}
-
-      %% \inferrule*[right=Covar Hom elim simple]
-      %% {}
-      %% {\alpha: \cat C, f:\forall \gamma. P \triangleleft R, \beta : \cat D, x : R, \gamma : \cat E \vdash f\triangleleft (x, \gamma) : P}
-
-      \inferrule*[right=Covariant Hom elimination]
-      {\Phi = \Phi_f \jnctx \Phi_a\and
-       \Gamma\pipe\Phi_f \vdash s : \homrallXYtoZ {\alpha:\cat C} R P \and
-       d^+\Phi_a \vdash a : \cat C\and
-       \Phi_a \vdash t : R[a/\alpha]
-        }
-      {\Gamma\pipe \Phi \vdash \homrappXtoYatZ s t a : P[a/\alpha]}\\
-
-      \homrappXtoYatZ {(\homrlambdaXatYdotZ x \alpha s)} t a = s[t/x,a/\alpha] (\beta)
-      %% (\lambda^\triangleleft (x, \gamma). M)\triangleleft (x, \gamma) = M (\beta)
-
-      \inferrule
-      {\Gamma\pipe\Phi \vdash s : \homrallXYtoZ \alpha R P}
-      {s = \homrlambdaXatYdotZ {x}{\alpha} \homrappXtoYatZ s x \alpha}
-
-      %% \inferrule
-      %% {\Gamma\pipe\Phi, x:R, \alpha : \cat C \vdash \homrappXtoYatZ s x \alpha = \homrappXtoYatZ t x \alpha : P}
-      %% {\Gamma\pipe\Phi \vdash s = t : \homrallXYtoZ \alpha R P}
-
-      \inferrule*[right=Contra Hom formation]
-      {\Gamma \pipe \alpha:\cat C; d^-\Xi\vdash R \and \Gamma \pipe \alpha : \cat C; d^+\Xi\vdash P}
-      {\Gamma \pipe \Xi \vdash \homlallXYtoZ \alpha R P}
-
-      \inferrule*[right=Contra Hom introduction]
-      {\Gamma\pipe\alpha:\cat C, x : R, \Phi \vdash t : P}
-      {\Gamma\pipe\Phi \vdash \homllambdaXatYdotZ {x : R} {\alpha : \cat C} t : \homlallXYtoZ \alpha R P}
-
-      \inferrule*[right=Contra Hom elimination]
-      {\Phi = \Phi_a \jnctx \Phi_f\and
-        \Gamma \pipe d^-\Phi_a \vdash a : \cat C\and
-       \Gamma \pipe \Phi_a \vdash s : R[a/\alpha] \and
-       \Gamma \pipe \Phi_f \vdash t : \homlallXYtoZ {\alpha:\cat C} R P}
-      {\Gamma \pipe \Phi \vdash \homlappXtoYatZ t s a : P[a/\alpha]}\\
-
-      \inferrule*[Right=$\beta$]
-                 {}
-                 {\homlappXtoYatZ {(\homllambdaXatYdotZ x \alpha t)} s a = t[a/\alpha,s/x]}
-
-      \inferrule*[right=$\eta$]
-      {\Gamma \pipe \Phi \vdash t : \homlallXYtoZ \alpha R P}
-      {\Gamma \pipe \Phi \vdash t = \homllambdaXatYdotZ x \alpha \homlappXtoYatZ t x \alpha : \homlallXYtoZ \alpha R P}
-
-      %% \inferrule*[right=$\eta$]
-      %% {\Gamma \pipe \alpha:\cat C, x:R, \Phi \vdash \homlappXtoYatZ M x \alpha = \homlappXtoYatZ N x \alpha : P}
-      %% {\Gamma \pipe \Phi \vdash M = N : \homlallXYtoZ \alpha R P}
-  \end{mathpar}
-  \caption{Hom of Relations/Profunctors}
-  \label{fig:hom}
-\end{figure}
-
-In category theory, these profunctors can be constructed using an
-\emph{end} and the hom of functor categories\cite{somecats}. Given two
-profunctors $P : C^o \times E \to \Set$ and $R : D^o \times E \to
-Set$, their hom is given as
-\[ \int_{e \in E} R(-,e) \Rightarrow P(e,=) \]
-
-%% The syntax is given in Figure~\ref{fig:cotensor}.
-%% %
-%% First, the formation rule for covariant cotensor formation requires
-%% that the covariant variables for $R,P$ are the same, and the
-%% contravariant variables match the boundary of the overall
-%% judgment. Note that we use $d^+\Xi,d^-\Xi$ to allow for the case that
-%% the boundary is a single variable $\Xi = \alpha:\cat C$.
-%% %
-%% Next, the substitution rule is straightforward.
-%% %
-%% Then we have the elimination form, first written in a simple form with
-%% no substitution built in.
-%% %
-%% This says that given a variable of the cotensor judgment, and another
-%% variable of judgment $R$, we can apply these to get a term of judgment
-%% $P$. As a diagram this is written as follows:
-
-%% \[ % https://tikzcd.yichuanshen.de/#N4Igdg9gJgpgziAXAbVABwnAlgFyxMJZABgBpiBdUkANwEMAbAVxiRAGEQBfU9TXfIRQBGclVqMWbACLdeIDNjwEiAJjHV6zVohABROXyWCiZYeK1TdnHkYEqU685sk793cTCgBzeEVAAZgBOEAC2SGQgOBBIohLabAA6iQEQQYwMAATJ3nShoXQAdJkACtmJOEFYdGDeDDD1ATiZAEqGIMFhsdTRSOrxViBtth0h4YgAzD0xiAAsLgm6Je2d45G9k9QMWGBuUBBMAEb1INQAFjB0UGyQu6dRdFgMNwSsI6t900jzIBdXL3ctjs9gdjqweo9nrpbm8KFwgA
-%% \begin{tikzcd}
-%% C \arrow[r, "\forall \gamma. P \triangleleft R"] \arrow[d, no head, Rightarrow] & D \arrow[r, "R"] & E \arrow[d, no head, Rightarrow] \\
-%% C \arrow[rr, "P"]                                                               &                  & E                               
-%% \end{tikzcd}\]
-
-%% The universal property of this judgment is that this is the universal
-%% 2-cell of this type, in that any 2-cell
-
-%% \[ % https://tikzcd.yichuanshen.de/#N4Igdg9gJgpgziAXAbVABwnAlgFyxMJZABgBpiBdUkANwEMAbAVxiRAGEByEAX1PUy58hFACZyVWoxZsAIr34gM2PASIBmCdXrNWiEAFEFAlcKJkAjJJ0z97Y0sGqRyTVe3S9hh8qFqUFlpSumy8kjBQAObwRKAAZgBOEAC2SIEgOBBI4sG2IABKDokpSJoZWYgALB4h+gAKRUmpiGTlpdQMWGBecBCdUCDUABYwdAOIYEwMDNQ4dFgMbJDdjSWIOZlI1SAjY0sErB1dXlAQTABGDIcZ84v6y6x88U1IrZuIAKxPIMXNH7MVdKdFb6KB0OAjAbDUbjcAHMI8IA
-%% \begin{tikzcd}
-%% C' \arrow[d] \arrow[r] & {} \arrow[r, no head, dashed] & D \arrow[r, "R"] & E \arrow[d, no head, Rightarrow] \\
-%% C \arrow[rrr, "P"]     &                               &                  & E                               
-%% \end{tikzcd}\]
-
-%% Exhibits a unique factorization:
-
-%% \[ % https://tikzcd.yichuanshen.de/#N4Igdg9gJgpgziAXAbVABwnAlgFyxMJZABgBoBGAXVJADcBDAGwFcYkQBhEAX1PU1z5CKAEwVqdJq3YARHnxAZseAkQDM4mgxZtEIAKLz+yoUTIiJ26Xq69jg1Sg0WtU3QaOKBK4SVLFLN3YOAHJPJQdfcn9AnXZw71NRGNc4vTk7LxNHZA0A1OsPbgkYKABzeCJQADMAJwgAWyRokBwIJDFJNJAAJU86xqQNVvbEABYC9wAFfvqmxDIRoZpGLDB3KAhmACNGNhoACxh6KHZIdZAaHHosRjOCNkyB+c62pAmQI5P7i5W1ja2u32rRudz050eCmeSAArFdRgA2J5zJAI+FIADsfwueig9DgR1Oh2Op3BD1mg0QWKWiBaqxxIE2Oz2l0+JJ+wOutw5FPmcJpxGRlMWb1pk3YAB0JdUILUmIwAARSsr0BoNegAOgVUyVEpwtSw9DAZT2e2qOAVfSF82pooAHOK9FaoSjEA6aZ16QDmcCvqTwOSrqCecVuEA
-%% \begin{tikzcd}
-%% C' \arrow[r] \arrow[d]                                                           & {} \arrow[r, no head, dashed] & D \arrow[d, no head, Rightarrow] \arrow[r, "R"] & E \arrow[d, no head, Rightarrow] \\
-%% C \arrow[d, no head, Rightarrow] \arrow[rr, "\forall \gamma.
-P \triangleleft R"] &                               & D \arrow[r, "R"]                                & E \arrow[d, no head, Rightarrow] \\
-%% C \arrow[rrr, "P"]                                                               &                               &                                                 & E                               
-%% \end{tikzcd}\]
-
-%% We give elimination rules and introduction rules in a style similar to
-%% $\lambda$-calculus, with corresponding $\beta\eta$ rules.
-
-These hom types are useful both for category-theoretic applications
-(Kan extensions, the Yoneda lemma), but also for purely syntactic
-reasons.  The universal property of these function types allows us, by
-iterated hom to express any sequent $\Phi \vdash R$ as an equivalent
-one with an ``empty'' context, i.e., a single object variable $d^\pm
-\Phi$. Then using the end type defined in the previous section, we can
-internalize that judgment as a type in our ambient dependent type
-theory.
-
 As an example, any sequent $\Gamma \pipe \alpha, x: R, \beta, y : S, \gamma \vdash
@@
 This property is useful in giving a simple syntax to positive types.
-
-\subsection{Unit and Tensor of Profunctors}
-
-While the hom sets of the previous section work quite smoothly since
-they are negative connectives in a natural deduction system, the positive
-connectives, the unit and tensor, are trickier to define.
 
@@
 
-Next, we continue with the tensor product $\tensorexistsXwithYandZ
-\beta P Q$. Adjoint to the right/left hom sets, this provides a form
-of \emph{existentially quantified pair}. The formation rule states the
-set is well-formed when $\beta$ is the covariant variable of $P$ and
-the contravariant variable of $Q$, and, its contravariant dependence
-is the same as $P$, while its covariant dependence is the same as $Q$.
-%
-The introduction rule is then a tuple $\tensorintroatXwithYandZ b s t$
-with an instantiation $b$ of the quantified variable $\beta$. Note
-that the use of $d^+\Psi_s$ in the rule could be replaced by
-$d^-\Psi_t$ since they must be equal by the side condition $\Phi =
-\jnctx \Psi_s \Psi_t$.
-%
-The elimination rule says to construct a transformation taking a pair
-as input it is sufficient to construct one that takes in a $P$ and $Q$
-for any $\beta$, directly corresponding to the adjoint relationship
-between the tensor and the right hom.
-%
-Finally, we have the expected $\beta\eta$ equations.
-
-\begin{figure}
-  \begin{mathpar}
-    \inferrule*[right=Unit-Formation]
-    {\Gamma \vdash \cat C \isaCat\and
-     \Gamma \pipe d^-\Xi \vdash a_1 : \cat C\and \Gamma \pipe d^+ \Xi \vdash a_2 : \cat C}
-    {\Gamma \pipe \Xi \vdash \punitinXfromYtoZ {\cat C} {a_1} {a_2}}
-
-    %% (A \to A')[\xi] = A[d^-\xi] \to A'[d^+\xi]
-
-    \inferrule*[right=Unit-Introduction]
-    {\Gamma\pipe\beta: \cat D \vdash a : \cat C}
-    {\Gamma \pipe\cdot \vdash \punitrefl a : \punitinXfromYtoZ {\cat C} a a}
-
-    \inferrule*[right=Unit Elim]{
-      \Gamma\vdash Q : \harr {\cat C} {\cat C}\and
-      \Gamma\vdash t : \pendallXdotY {\alpha:\cat C} {\harrapp Q \alpha \alpha}}
-    {\Gamma \vdash \punitelimtoYwithkontZ Q N :
-      \pendallXdotY {\alpha_1} {\homrallXYtoZ {\alpha_2} {\punitinXfromYtoZ {\cat C} {\alpha_1}{\alpha_2}} {\harrapp Q {\alpha_1}{\alpha_2}}}
-      }
-
-    %% J_\to(\alpha_1.\alpha_2. Q, \alpha. N, A_1, M, A_2)[\phi] = J_\to(\alpha_1.\alpha_2. Q, \alpha. N, A_1[d^-\phi], M[\phi], A_2[d^+\phi])
-
-    %% \inferrule*[right=Unit$\beta$]{\beta:\cat D \vdash A : \cat C\and
-    %%   \alpha:\cat C \vdash N : Q[\alpha/\alpha_1,\alpha/\alpha_2]
-    %% }{\beta : \cat D \vdash J_\to(\alpha_1.\alpha_2. Q, \alpha. N, A, \id(A), A) = N[A/\alpha] : Q[A/\alpha_1,A/\alpha_2]}
-    %%
-
-    \appunary {\punitelimtoYwithkontZ Q N} a {\punitrefl a} a = \pendappXtoY N a
-
-    \inferrule*[right=Unit$\eta$]
-    {\Gamma \vdash M : \pendallXdotY {\alpha_1} \homrallXYtoZ {\alpha_2} {\punitinXfromYtoZ {\cat C} {\alpha_1} {\alpha_2}} {\harrapp Q {\alpha_1} {\alpha_2}}}
-    {M = {\punitelimtoYwithkontZ Q {\appunary M \alpha {\punitrefl \alpha} \alpha}}}
-    %% {}
-    %% {\alpha_1 : \cat C, f : \alpha_1 \to_{\cat C} \alpha_2, \alpha_2:\cat C \vdash M : \harrapp Q {\alpha_1} {\alpha_2}}
-    %% {\alpha_1 : \cat C, f : \alpha_1 \to_{\cat C} \alpha_2, \alpha_2:\cat C \vdash M =
-    %%   \homrappXtoYatZ {\pendappXtoY {\punitelimtoYwithkontZ Q {\pendlambdaXdotY \alpha M[\alpha/\alpha_1,\pendappXtoY {\punitrefl C} \alpha/f,\alpha/\alpha_2]}} {\alpha_1}} f {\alpha_2}
-    %% : \harrapp Q {\alpha_1}{\alpha_2}}
-
-    \inferrule*[right=Tensor formation]
-    {\Gamma \pipe d^-\Xi; \beta:D\vdash P \isaSet\and
-     \Gamma \pipe \beta:D;d^+\Xi \vdash Q \isaSet}
-    {\Gamma \pipe \Xi \vdash \tensorexistsXwithYandZ {\beta:D} P Q}
-
-  %% \inferrule*[right=Tensor intro function]
-  %% {}
-  %% {\alpha : \cat C, q : Q, \beta : \cat D, r : R, \gamma : \cat E \vdash (q, \beta,r) : \exists \beta : \cat D. Q \odot R}
-
-  %% \inferrule*[right=Tensor introduction]
-  %% {\Gamma \vdash P : \harr C D \and
-  %%   \Gamma \vdash Q : \harr D E}
-  %% {\Gamma\pipe\cdot\vdash \tensorintroofXandY P Q : \pendallXdotY {\alpha} \homrallXYtoZ \beta {\harrapp P \alpha \beta} {\homrallXYtoZ \gamma {\harrapp Q \beta \gamma} {(\tensorexistsXwithYandZ \beta {\harrapp P \alpha \beta}{\harrapp Q \beta \gamma})}}}
-
-  \inferrule*[right=Tensor introduction]
-  {\Phi = \Psi_s \jnctx \Psi_t \and
-   \Gamma \pipe d^+\Psi_s \vdash b : \cat D\and
-   \Gamma \pipe \Psi_s \vdash s : P[b/\beta] \and
-   \Gamma \pipe \Psi_t \vdash t : Q[b/\beta]}
-  {\Gamma\pipe\Phi\vdash \tensorintroatXwithYandZ b s t : \tensorexistsXwithYandZ {\beta : \cat D} P Q}
-
-  %% \tensorintroatXwithYandZ B M N [\phi,B'/\beta',\psi] = \tensorintroatXwithYandZ {B[B'\beta]} {M[\phi]} {N[\psi]}
-
-  %% \inferrule*[right=Tensor simpl]
-  %% {\Gamma\pipe\cdot \vdash R : \harr {C} {E}\and
-  %%  \Gamma\pipe\cdot \vdash N : \pendallXdotY {\alpha:C} {\homrallXYtoZ {\beta:D} P {\homrallXYtoZ {\gamma:E} Q {\harrapp R \alpha\gamma}}}}
-  %% {\Gamma\pipe\alpha:C,p:\tensorexistsXwithYandZ {\beta:D} P Q,\gamma:E \vdash\tensorelimVatWtoXtoYwithkontZ p \alpha \gamma R N : \harrapp R A G}
-
-  \inferrule*[right=Tensor-elim]
-  {\Gamma \pipe \vdash P : \harr {\cat C} {\cat D}\and\Gamma \vdash Q : \harr {\cat D}{\cat E}\and
-   \Gamma \vdash R : \harr {\cat C}{\cat E}\and
-   \Gamma \vdash N : \hombinary \alpha {\harrapp P \alpha \beta} \beta {\harrapp Q \beta \gamma} \gamma {\harrapp R \alpha \gamma}}
-  {\Gamma \vdash \tensorelimfromWXtoYwithkontZ P Q R N : \homunary \alpha {(\tensorexistsXwithYandZ \beta {\harrapp P \alpha \beta} {\harrapp Q \beta \gamma})} {\gamma} {\harrapp R \alpha \gamma}}
-
-  %% \inferrule*[right=Tensor elim]
-  %% {\Gamma\pipe\cdot \vdash R : \harr {C} {E}\and
-  %%  \Gamma\pipe\Phi \vdash M : \tensorexistsXwithYandZ {\beta:D} {P[A/\alpha;]} {Q[;G/\gamma]}\and
-  %%  \Gamma\pipe\cdot \vdash N : \pendallXdotY {\alpha:C} {\homrallXYtoZ {\beta:D} P {\homrallXYtoZ {\gamma:E} Q {\harrapp R \alpha\gamma}}}}
-  %% {\Gamma\pipe\Phi \vdash\tensorelimVatWtoXtoYwithkontZ M A G R N : \harrapp R A G}
-
-  %% J_\odot(\alpha. \gamma. S, \alpha.q.\beta.r.\gamma.N, A, M, G)[\phi] =
-  %% J_\odot(\alpha. \gamma. S, \alpha.q.\beta.r.\gamma.N, A[d^-\phi], M[\phi], G[d^+\phi])
-
-  %% \inferrule*[right=Tensor elim Sequent]
-  %% {\Phi,\alpha : \cat C, q : Q, \beta : \cat D, r: R, \gamma : \cat E,\Psi \vdash N : S}
-  %% {\Phi, \alpha : \cat C, x : \exists \beta. Q \odot R, \gamma : \cat E,\Psi \vdash J_\odot(x;q.\beta.r. N) : S}
-
-  %% \inferrule*[right=Tensor elim Subst]
-  %% {\Phi,\alpha : \cat C, q : Q, \beta : \cat D, r: R, \gamma : \cat E,\Phi' \vdash N : S
-  %% \and
-  %% \phi : \Psi \to \Phi,\alpha : \cat C, q : Q, \beta : \cat D, r: R, \gamma : \cat E,\Phi'}
-  %% {\Psi  \vdash J_\odot(x;q.\beta.r. N)[\phi] : S}
-
-  \inferrule*[right=Tensor $\beta$]
-  {}
-  {\appunary {\tensorelimfromWXtoYwithkontZ P Q R N} a {\tensorintroatXwithYandZ b s t} c = \appbinary N a s b t c}
-  %% {\homrappXtoYatZ {\pendappXtoY{\tensorelimfromWXtoYwithkontZ P Q R N} A} {(\homrappXtoYatZ{\homrappXtoYatZ{\pendappXtoY {\tensorintroofXandY P Q} A} f B} g G)} G =
-  %%   \homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY N A} f B} g G}
-
-  %% \inferrule*[right=Tensor $\beta$]
-  %% {}{J_\odot(\alpha.\gamma.S, \alpha.q.\beta.r.\gamma.N,A,(M,B,N),G) = N[A/\alpha,M/q,B/\beta,N/r,G/\gamma]}
-
-  \inferrule*[right=Tensor$\eta$]
-  {\Gamma \vdash N : \homunary \alpha {(\tensorexistsXwithYandZ \beta {\harrapp P \alpha \beta}{\harrapp Q \beta \gamma})} \gamma {\harrapp R \alpha \gamma}}
-  {N = \tensorelimfromWXtoYwithkontZ P Q R {\lambdabinary \alpha x \beta y \gamma \appunary N \alpha {\tensorintroatXwithYandZ \beta x y} \gamma}}
-
-  %% \inferrule*[right=Tensor$\eta$]
-  %% {\alpha : \cat C, x : \tensorexistsXwithYandZ \beta P Q, \gamma : \cat E \vdash M : R}
-  %% {\alpha : \cat C, x : \exists \beta. Q \odot R, \gamma : \cat E \vdash M =
-  %%   \homrappXtoYatZ {\pendappXtoY{\tensorelimfromWXtoYwithkontZ P Q R
-  %%       {\pendlambdaXdotY \alpha {\homrlambdaXatYdotZ p \beta {\homrlambdaXatYdotZ q \gamma {M[\homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {\tensorintroofXandY P Q} \alpha} p \beta} q \gamma/x]}}}}
-  %%     } \alpha} x \gamma : S}
-  \end{mathpar}  
-  \caption{Unit, Tensor of Profunctors}
-\end{figure}
-
-%% Next, we can internalize the $,\alpha,$ as a positive connective as a
-%% kind of ``existentially quantified tensor product'', abstracting the
-%% composition of profunctors.
-
-%% The ordinary formula for this is that for profunctors $R : C^o \times
-%% D \to \Set$ and $S : D^o \times E \to \Set$, the composite is given by
-%% a coend:
-
-%% \[ \int^{d \in D} R(-,d) \otimes S(d,=) \]
-
-%% We will instead use slightly more logical notation, writing the coend as
-%% an existential quantifier, justified by the analogy between the two
-%% and the structure of the introduction/elimination forms in our syntax
-%% in particular.
+%% \subsection{Dependent Types}
+
+%% Next, we cover the basic dependent type theory structure: $\Pi,\Sigma$
+%% and identity types.
+
 %% \begin{figure}
 %%   \begin{mathpar}
+%%   \inferrule
+%%   {\Gamma \pipe\cdot \vdash S \isaSet\and \Gamma \pipe\cdot \vdash T \isaSet}
+%%   {\Gamma\pipe\cdot \vdash \prod_{x:S} T \isaSet}
+
+%%   \inferrule
+%%   {\Gamma \pipe \cdot \vdash M : \prod_{x:S} T\and
+%%     \Gamma \pipe \cdot \vdash N : S}
+%%   {\Gamma \pipe \cdot \vdash M N : T[N/x]}
+
+%%   \inferrule
+%%   {\Gamma,x:S \pipe \cdot \vdash M : T}
+%%   {\Gamma \pipe \cdot \vdash \lambda x:S. M : \prod_{x:S} T}
+
+%%   \inferrule
+%%   {\Gamma \pipe\cdot \vdash S \isaSet\and \Gamma \pipe\cdot \vdash T \isaSet}
+%%   {\Gamma\pipe\cdot \vdash \sum_{x:S} T \isaSet}
+
+%%   \inferrule
+%%   {~}
+%%   {\Gamma\pipe\cdot \vdash \top \isaSet}
+
+%%   \inferrule
+%%   {\Gamma \pipe \cdot \vdash M : S\and \Gamma \pipe \cdot \vdash N : S}
+%%   {\Gamma \pipe \cdot \vdash \textrm{Id}_S M N \isaSet}
+
+%%   \inferrule
+%%   {\Gamma \pipe \cdot \vdash M : S}
+%%   {\Gamma \pipe \cdot \vdash \textrm{refl}(M) : \textrm{Id}_S M M}
+
+%%   \inferrule
+%%   {\Gamma,x:S,y:S,p:\textrm{Id}_S x y \pipe \cdot \vdash C \isaSet\and
+%%    \Gamma \pipe \cdot \vdash M_k : \prod_{x:S} C[x/y,\textrm{refl}(x)/p]}
+%%   {\Gamma \pipe \cdot \vdash \textrm{J}(x,y,p.C,M) : \prod_{x:S,y:S,p:\textrm{Id}_S x y} C}
 %%   \end{mathpar}
-%% \caption{Tensor Primitive Rules}
-%% \label{fig:tensor}
+%%   \caption{Ordinary Dependent Type Structure}
 %% \end{figure}
-
-Semantically, the tensor product is the composition of profunctors,
-which is typically written as a coend
-\[ \int_{\beta} P(-,\beta) \times Q(\beta,=) \]
-
-\subsection{Graph of a profunctor}
-
-One final multiplicative we can define is the \emph{graph} of a
-profunctor. Given a profunctor $\alpha:\cat C;\beta:\cat D \vdash R$,
-the ``graph'' of $R$ is a category $\Sigma_{\alpha:\cat C;\beta:\cat
-  D}. R$. In detail, its objects are triples $(a:\cat C, r : R, b :
-\cat D)$ and morphism from $(a,r,b)$ to $(a', r', b')$ consists of a
-pair of morphisms $g : b \to b'$ and a backwards $f : a' \to a$ such
-that $r' = g\circ r \circ f$.
-%
-Intuitively this generalizes the ``graph'' of a relation $R : A\times
-B \to \Omega$, where we think of the graph as the corresponding subset
-$gr(R) \subseteq A\times B$
-%
-By picking $R = \punitinXfromYtoZ {\cat C} {\alpha}{\alpha'}$ we
-recover the \emph{arrow} category $\exists
-(\alpha;\alpha'). \punitinXfromYtoZ {\cat C}{\alpha}{\alpha'}$, by
-composing these with functors, we recover the notion of \emph{comma}
-category $\Sigma \alpha;\beta. \punitinXfromYtoZ {\cat
-  E}{(F\alpha)}{(G\beta)}$ and by picking either $F$ or $G$ to have
-domain the terminal category $1$, we recover the notion of a
-\emph{slice} category.
-
-This is unique among the multiplicatives in that it defines a
-\emph{category} rather than a profunctor. The graph of a profunctor is
-a \emph{negative} multiplicative, a kind of $\Sigma$-type for
-profunctors. The universal property says that a functor from $\cat E$
-into the graph $\Sigma (\alpha; \beta). R$ is equivalent to a
-transformation from $\cat E$ into the functor $R$, including functors
-from $\cat E$ to $\cat C$ and $\cat D$.
-%
-The graph of a profunctor is a joint generalization of several common
-constructions in category theory: the arrow category, slice
-categories, comma categories and the covariant and the covariant and
-contravariant category of elements.
-
-\begin{figure}
-  \begin{mathpar}
-    \inferrule*[right=Graph-Formation]
-    {\Gamma \vdash \alpha:\cat C; \beta:\cat D \vdash R \isaSet}
-    {\Gamma \vdash \exists (\alpha;\beta). R \isaCat}
-
-    \inferrule*[right=Graph-E+]
-    {\Gamma \vdash \gamma:\cat E \vdash \phi : \exists(\alpha:\cat C;\beta: \cat D).R}
-    {\Gamma \vdash \pi_+\phi : \cat C}
-
-    \inferrule*[right=Graph-E-]
-    {\Gamma \vdash \gamma:\cat E \vdash \phi : \exists(\alpha:\cat C;\beta: \cat D).R}
-    {\Gamma \vdash \pi_-\phi : \cat D}
-
-    \inferrule*[right=Graph-Ep]
-    {\Gamma \vdash \gamma:\cat E \vdash \phi : \exists(\alpha:\cat C;\beta: \cat D).R}
-    {\Gamma \vdash \gamma:\cat E \vdash \pi_p\phi : R[\pi_+\phi/\alpha,\pi_-\phi/\beta]}
-
-    \inferrule*[right=Graph-Intro]
-    {\Gamma\vdash \gamma:\cat E \vdash a : \cat C \and
-      \Gamma\vdash \gamma:\cat E \vdash b : \cat D \and
-      \Gamma\vdash \gamma:\cat E \vdash t : R[a/\alpha;b/\beta]}
-    {\Gamma\vdash \gamma:\cat E \vdash (a,b,t) : \exists (\alpha;\beta). R}
-
-    \inferrule*[right=Unit-Formation]
-    {\Gamma \vdash \cat C \isaCat\and
-     \Gamma \pipe d^-\Xi \vdash a_1 : \cat C\and \Gamma \pipe d^+ \Xi \vdash a_2 : \cat C}
-    {\Gamma \pipe \Xi \vdash \punitinXfromYtoZ {\cat C} {a_1} {a_2}}
-
-    \pi_+(a,b,t) = a \\
-    \pi_-(a,b,t) = b \\
-    \pi_p(a,b,t) = t\\
-
-    \inferrule
-    {\Gamma \vdash \gamma:\cat E \vdash \phi : \exists(\alpha:\cat C;\beta:\cat D). R}
-    {\phi = (\pi_+\phi, \pi_-\phi, \pi_p\phi)}
-  \end{mathpar}  
-  \caption{Graph of a Profunctor}
-\end{figure}
-
-\subsection{Positive and Negative Presheaf Categories}
-
-Though defining arbitrary functor categories type-theoretically would
-require a ``multi-argument functor'' judgment, we can define
-\emph{presheaf} categories by the universal property that a functor
-$\cat C \to \mathcal P {\cat D}$ is the same as a profunctor from
-$\cat D$ to $\cat C$. To distinguish the covariant from contravariant
-presheaves, we define them dually as $\negPresheaf {\cat C}$ and
-$\posPresheaf {\cat D}$
-
-\begin{figure}
-  \begin{mathpar}
-    \inferrule*[right=NegPresheaf-Formation]
-    {\Gamma \vdash \cat C \isaCat}
-    {\Gamma \vdash \negPresheaf {\cat C} \isaCat}
-
-    \inferrule*[right=NegPresheaf-E]
-    {\Gamma \pipe d^-\Xi \vdash a : \cat C \and
-     \Gamma \pipe d^+\Xi \vdash p : \negPresheaf {\cat C}
-    }
-    {\Gamma \pipe \Xi \vdash \negPresheafApp a p \isaSet}
-
-    \inferrule*[right=NegPresheaf-I]
-    {\Gamma \pipe \alpha:\cat C;\beta:\cat D \vdash R : \isaSet}
-    {\Gamma \pipe \beta:\cat D \vdash \lambda \alpha:\cat C. R : \negPresheaf {\cat C}}
-
-    \negPresheafApp a (\lambda \alpha. R) = R[a/\alpha]\\
-
-    \inferrule*[right=NegPresheaf-$\eta$]
-    {\Gamma \pipe \beta:\cat D \vdash p : \negPresheaf {\cat C}}
-    {p = \lambda \alpha. \negPresheafApp \alpha p}
-
-    \inferrule*[right=PosPresheaf-Formation]
-    {\Gamma \vdash \cat D \isaCat}
-    {\Gamma \vdash \posPresheaf {\cat D} \isaCat}
-
-    \inferrule*[right=PosPresheaf-E]
-    {\Gamma \pipe d^-\Xi \vdash p : \posPresheaf {\cat D}\and
-     \Gamma \pipe d^+\Xi \vdash a : \cat D
-    }
-    {\Gamma \pipe \Xi \vdash \posPresheafApp p a \isaSet}
-
-    \inferrule*[right=PosPresheaf-I]
-    {\Gamma \pipe \alpha:\cat C;\beta:\cat D \vdash R : \isaSet}
-    {\Gamma \pipe \alpha:\cat C \vdash \lambda \beta:\cat D. R : \posPresheaf {\cat D}}
-
-    \posPresheafApp {(\lambda \beta. R)} b = R[b/\beta]\\
-
-    \inferrule*[right=PosPresheaf-$\eta$]
-    {\Gamma \pipe \alpha:\cat C \vdash p : \posPresheaf {\cat D}}
-    {p = \lambda \beta. \posPresheafApp p \beta}
-  \end{mathpar}
-  \caption{Presheaf Categories}
-\end{figure}
-
-\subsection{Dependent Types}
-
-Next, we cover the basic dependent type theory structure: $\Pi,\Sigma$
-and identity types.
-
-\begin{figure}
-  \begin{mathpar}
-  \inferrule
-  {\Gamma \pipe\cdot \vdash S \isaSet\and \Gamma \pipe\cdot \vdash T \isaSet}
-  {\Gamma\pipe\cdot \vdash \prod_{x:S} T \isaSet}
-
-  \inferrule
-  {\Gamma \pipe \cdot \vdash M : \prod_{x:S} T\and
-    \Gamma \pipe \cdot \vdash N : S}
-  {\Gamma \pipe \cdot \vdash M N : T[N/x]}
-
-  \inferrule
-  {\Gamma,x:S \pipe \cdot \vdash M : T}
-  {\Gamma \pipe \cdot \vdash \lambda x:S. M : \prod_{x:S} T}
-
-  \inferrule
-  {\Gamma \pipe\cdot \vdash S \isaSet\and \Gamma \pipe\cdot \vdash T \isaSet}
-  {\Gamma\pipe\cdot \vdash \sum_{x:S} T \isaSet}
-
-  \inferrule
-  {~}
-  {\Gamma\pipe\cdot \vdash \top \isaSet}
-
-  \inferrule
-  {\Gamma \pipe \cdot \vdash M : S\and \Gamma \pipe \cdot \vdash N : S}
-  {\Gamma \pipe \cdot \vdash \textrm{Id}_S M N \isaSet}
-
-  \inferrule
-  {\Gamma \pipe \cdot \vdash M : S}
-  {\Gamma \pipe \cdot \vdash \textrm{refl}(M) : \textrm{Id}_S M M}
-
-  \inferrule
-  {\Gamma,x:S,y:S,p:\textrm{Id}_S x y \pipe \cdot \vdash C \isaSet\and
-   \Gamma \pipe \cdot \vdash M_k : \prod_{x:S} C[x/y,\textrm{refl}(x)/p]}
-  {\Gamma \pipe \cdot \vdash \textrm{J}(x,y,p.C,M) : \prod_{x:S,y:S,p:\textrm{Id}_S x y} C}
-  \end{mathpar}
-  \caption{Ordinary Dependent Type Structure}
-\end{figure}
-
-\begin{figure}
-  \begin{mathpar}
-    \inferrule
-    {}
-    {\Gamma\vdash \smallCats}
-
-    \inferrule
-    {}
-    {\Gamma\vdash \Cats}
-
-    \inferrule
-    {\Gamma \vdash \cat C : \Sort \and\Gamma \vdash \cat D : \Sort}
-    {\Gamma\pipe\cdot \vdash \varr{\cat C}{\cat D} \isaSet}
-
-    \inferrule
-    {\Gamma \vdash \cat C : \Sort \and\Gamma \vdash \cat D : \Sort}
-    {\Gamma\pipe\cdot \vdash \harr{\cat C}{\cat D} \isaSet}
-
-    \inferrule
-    {\Gamma \pipe \alpha:\cat C \vdash A : \cat D}
-    {\Gamma \vdash \lambda \alpha:\cat C. A : \varr{\cat C}{\cat D}}
-
-    \inferrule
-    {\Gamma \pipe \alpha:\cat C \vdash A : \cat D\and
-      \Gamma\pipe\cdot \vdash M : \varr{\cat D}{\cat E}}
-    {\Gamma\pipe \alpha:\cat C \vdash M A : \cat E}
-
-    \inferrule
-    {\Gamma \pipe \alpha:\cat C;\beta:\cat D \vdash R \,\isaSet}
-    {\Gamma \vdash \lambda \alpha:\cat C;\beta:\cat D. R : \harr{\cat C}{\cat D}}
-
-    \inferrule
-    {\Gamma \vdash M : \harr{\cat C}{\cat D}\and
-     \Gamma \pipe d^- \Xi \vdash A : \cat C\and
-     \Gamma \pipe d^+ \Xi \vdash B : \cat C}
-    {\Gamma \pipe \Xi \vdash M A; B \isaSet}
-
-    \inferrule
-    {\Gamma \pipe \alpha:\cat C\vdash P \isaSet}
-    {\Gamma \vdash \pendallXdotY {\alpha:{\cat C}} P \isaTy}
-
-    \inferrule
-    {\Gamma \pipe \alpha:\cat C\vdash t : P}
-    {\Gamma \vdash \pendlambdaXdotY \alpha t : \pendallXdotY \alpha P}
-
-    \inferrule
-    {\Gamma \vdash M : \pendallXdotY \alpha P\and
-     \Gamma \pipe \beta:\cat D \vdash a : \cat C}
-    {\Gamma \pipe \beta:\cat D \vdash \pendappXtoY M a : P[a/\alpha]}
-  \end{mathpar}
-  \caption{Type Connectives}
-\end{figure}
-\begin{figure}
-  \begin{mathpar}
-    \inferrule
-    {\Gamma \pipe \alpha:\cat C\vdash P \isaSet}
-    {\Gamma \vdash \pendallXdotY {\alpha:{\cat C}} P \isaTy}
-
-    \inferrule
-    {\Gamma \pipe \alpha:\cat C\vdash t : P}
-    {\Gamma \vdash \pendlambdaXdotY \alpha t : \pendallXdotY \alpha P}
-
-    \inferrule
-    {\Gamma \vdash M : \pendallXdotY \alpha P\and
-     \Gamma \pipe \beta:\cat D \vdash a : \cat C}
-    {\Gamma \pipe \beta:\cat D \vdash \pendappXtoY M a : P[a/\alpha]}
-
-    \pendappXtoY {(\pendlambdaXdotY \alpha t)} a = t[a/\alpha]
-
-    \inferrule*[right=2Cell$\eta$]
-    {\Gamma \vdash M : \pendallXdotY \alpha P }
-    {\Gamma \vdash M = \pendlambdaXdotY \alpha {\pendappXtoY M \alpha} : \pendallXdotY \alpha P}
-  \end{mathpar}
-  \caption{End}
-  \label{fig:end}
-\end{figure}
 
@@
   \end{enumerate}
-\end{construction}\section{Completeness}
+\end{construction}
+
+\section{Completeness}
 
@@
 
-\begin{lemma}[Vertical category]
-  \label{lem:vertical-cat}
-  Substitution of object terms $A[B/\beta]$ is associative with
-  identity given by the variable $\alpha:\cat C \vdash \alpha:\cat C$.
-\end{lemma}
-
-\begin{definition}
-  \label{lem:horizontal-cat}
-  We define composition of transformation contexts $\Phi \jnctx \Psi$ by induction on $\Psi$
-  \begin{align*}
-    \Phi \jnctx \beta:\cat D &= \Phi    \\
-    \Phi \jnctx \Psi,x:R,\beta:\cat D &= (\Phi\jnctx\Psi),x:R,\beta:\cat D
-  \end{align*}
-  By induction, this is seen to be associative
-  \[ (\Phi \jnctx \Psi) \jnctx \Sigma = \Phi \jnctx (\Psi \jnctx \Sigma) \]
-  and unital with identity for $\cat C$ given by the single category
-  variable context $\alpha:\cat C$
-  \[ \alpha:\cat C \jnctx \Phi = \Phi = \Phi \jnctx \beta:\cat D\]
-  when $d^-\Phi = \alpha:\cat C$ and $d^+\Phi = \beta:\cat D$.
-\end{definition}
-
-\begin{definition}
-  \label{lem:horizontal-sq-cat}
-  We define the horizontal composition of transformation substitutions
-  $\phi \jnctx \psi$ whenever $d^+\phi = d^-\psi$ by induction on
-  $\psi$:
-  \begin{align*}
-    \phi \jnctx a/\alpha &= \phi\\
-    \phi \jnctx (\psi,t/x,a/\alpha) &= (\phi \jnctx \psi),t/x,a/\alpha
-  \end{align*}
-  By induction, this is seen to be associative
-  \[ \phi \jnctx (\psi \jnctx \sigma) = (\phi \jnctx \psi) \jnctx \sigma \]
-  and unital with identity given by a single object substitution
-  $a/\alpha : \alpha:\cat C \to \beta:\cat D$
-  \[ a/\alpha \jnctx \phi = \phi = \phi \jnctx b/\beta \]
-  when $d^-\phi = a/\alpha$ and $d^+\phi = b/\beta$.
-\end{definition}
-
-To define vertical composition of squares, we need the following lemma.
-\begin{lemma}[Inversion]
-  \begin{enumerate}
-  \item If $\Phi \vdash \psi :: (\alpha:\cat C)$ then $\Phi =
-    \beta:\cat D$ for some $\cat D$ and $\psi = a/\alpha$ where
-    $\beta:\cat D \vdash a : \cat C$.
-  \item If $\Phi \vdash \psi :: \Psi_1 \jnctx \Psi_2$, then there
-    exists unique $\Phi_1,\Phi_2,\psi_1,\psi_2$ such that $\Phi =
-    \Phi_1\jnctx\Phi_2$ and $\Phi_1\vdash\psi_1 :: \Psi_1$ and
-    $\Phi_2\vdash \psi_2 : \Psi_2$ and $\psi = \psi_1\jnctx\psi_2$.
-  \end{enumerate}
-\end{lemma}
-
 \begin{definition}
@@
 
-With the definition based on representability, we can easily prove the
-generalization of the famous theorem, that right adjoints preserve not
-just limits but all right Kan extensions
-
-\begin{theorem}[Right/Left Adjoints preserve Right/Left Kan Extensions]
-  Let $K : \varr{\cat C}{\cat D}$ and $J : \varr{\cat C}{\cat C'}$.
-
-  If the right Kan extension $\textrm{Ran}_J K$ exists, and if $G :
-  \varr {\cat D}{\cat D'}$ has a left adjoint $F : \varr{\cat D'}{\cat
-    D}$, then the right Kan extension $\textrm{Ran}_J G\circ K$ exists
-  and is given by $G \circ \textrm{Ran}_J K$.
-
-  Similarly, if the left Kan extension $\textrm{Lan}_J K$ exists and
-  if $F : \varr{\cat D}{\cat D'}$ has a right adjoint $G : \varr{\cat
-    D'}{\cat D}$, then the left Kan extension $\textrm{Lan}_J F\circ
-  K$ exists and is given by $F \circ \textrm{Ran}_J K$.
-\end{theorem}
-\begin{proof}
-  By the composition of the following natural isomorphisms:
-  \begin{align*}
-    \punitinXfromYtoZ {\cat D'} {\beta'} {G ((\textrm{Ran}_J K) \alpha')} &\cong \punitinXfromYtoZ {\cat D}{F \beta'} {(\textrm{Ran}_J K) \alpha'} \tag{Adjunction}\\
-    &\cong \homrallXYtoZ {\alpha}{(\punitinXfromYtoZ {\cat C'}{\alpha'}{J\alpha})} {\punitinXfromYtoZ {\cat D}{F \beta'}{K \alpha}} \tag{Ran assumption, \cref{lem:natural-restriction}}\\
-    &\cong \homrallXYtoZ {\alpha}{(\punitinXfromYtoZ {\cat C'}{\alpha'}{J\alpha})} {\punitinXfromYtoZ {\cat D}{\beta'}{G(K \alpha)}} \tag{Adjunction}\\
-  \end{align*}
-\end{proof}
-
-\subsection{(op)-Fibrations}
-
-A (Street) fibration can be defined in any 2-category. A functor $p :
-\varr {\cat E}{\cat B}$ is a fibration when the morphism $p \to \cat
-B/p$ in the slice 2-category over $\cat B$ has a right adjoint.
-
-Since we can define all of the components of this slice 2-category we
-can internalize this definition in SCT.
-
-\begin{enumerate}
-\item A functor $\textrm{wp} : \exists (\alpha:\cat B; \beta:\cat E). \punitinXfromYtoZ {\cat B} \alpha {p\beta}$
-\item In the slice category: $p(\textrm{wp}(k)) = \pi_-k$
-\item With a universal morphism $\epsilon : \pendallXdotY {k} \punitinXfromYtoZ {} {(\pi_-k,k,\punitrefl{}{\pi_-k})} {k}$
-\item In the slice category: $\pi_k \epsilon = \id$ (how to express this nicely in SCT?)
-\end{enumerate}
-
-We can also weaken this to a Street fibration if we require the more sensible $p(\textrm{wp}(k)) \cong \pi_-k$ only.
-
-\begin{lemma}{Fibrations compose}
-  If $F : \varr{\cat C}{\cat D}$ and $G : \varr{\cat D}{\cat E}$ are
-  (op)-fibrations then so is $G \circ F$.
-\end{lemma}
-\begin{proof}
-  By composition of adjoints.
-\end{proof}
-
-\begin{lemma}{Fibrations are Stable under Pullback}
-
-\end{lemma}
-
-\subsection{Duality Principle}
-
-TODO: example of applying the principle in our calculus.
-
-\subsection{Formal Type Theory}
-
-Since many type theories are given semantics by free structured
-categories, we can use SCT as a logical framework in which to do
-``formal type theory'' by axiomatizing the structure of the categories
-involved. The theories we can define are limited because we do not
-have access to opposite categories, but we can define some modal and
-monoidal structures.
-
-\begin{definition}
-  An \emph{SCT-signature} is a well-formed context $\Gamma \isadtctx$.
-\end{definition}
-
-\begin{lemma}
-  Any \emph{unary mode theory} can be defined as an SCT-signature as follows.
-  \begin{enumerate}
-  \item For each mode $m$ add an assumption $m : \Cat$
-  \item For each 1-morphism $f : m \to n$ add an assumption $R_f :
-    \harr m n$ and assumptions $F_f : \varr m n$ with and $G_f : \varr
-    n m$ with assumptions that they represent $R_f$.
-  \item For each 2-morphism $\alpha : f \Rightarrow g$ add an assumption $\alpha : R_f \multimap R_g$
-  \item Add each equality as an assumption using the identity type Id.
-  \end{enumerate}
-\end{lemma}
-
-This translation illustrates some ways the notion of SCT-signature is
-more flexible than a mode theory. For instance, in an SCT-signature we
-can add a profunctor assumption $R_f : \harr m n$ without saying
-whether it is representable as $F_f$ or $G_f$, and we can add more
-complex 2-morphisms.
-
-\begin{definition}
-  The \emph{theory of a strict monoidal category} consists of
-  \begin{enumerate}
-  \item A category $C : \Cat$
-  \item A unit $I : \varr 1 C$
-  \item A tensor product $\otimes : \varr {(C \times C)} C$
-  \item Such that $\alpha:(C \times C) \times C \vdash (\pi_2\pi_1\alpha \otimes \pi_2\pi_1 \alpha) \otimes \pi_2 \alpha = \pi_2\pi_1\alpha \otimes (\pi_2\pi_1 \alpha \otimes \pi_2 \alpha)$
-    and $\alpha \otimes I() = \alpha = I() \otimes \alpha$
-  \end{enumerate}
-\end{definition}
-
-\begin{definition}
-  The theory of a cartesian strict monoidal category is an extension
-  of the theory of a strict monoidal category with representability
-  data
-  \[ \punitinXfromYtoZ {C} {\Gamma}{\Delta \otimes \Sigma} \cong (\punitinXfromYtoZ C \Gamma \Delta)\times (\punitinXfromYtoZ C \Gamma \Sigma) \]
-  and
-  \[ \punitinXfromYtoZ {C} {\Gamma} I \cong 1 \]
-\end{definition}
-
-\begin{definition}
-  The theory of a simple type theory is an extension of the theory of
-  a strict cartesian monoidal category with a
-  \begin{enumerate}
-  \item A category $\text{Ty}$ of types
-  \item A profunctor $\vdash : \harr C {\textrm{Ty}}$
-  \item A functor $\text{var} : \varr{\textrm{Ty}} C$ representing $\vdash$ :
-    \[ \Gamma \vdash A \cong \punitinXfromYtoZ {} \Gamma {\text{var} A} \]
-  \item As assumption that $\text{var}$ is fully faithful:
-     \[ \punitinXfromYtoZ {} {\text{var} A} {\text{var} B} \cong \punitinXfromYtoZ {} A B \]
-  \end{enumerate}
-\end{definition}
-
-\begin{definition}
-  The theory of CBPV is an extension of the theory of a simple type
-  theory $(\text{VCtx},\times,1,\text{VTy},\vdash_V,\text{var}_V)$ with
-  \begin{enumerate}
-  \item A category $\text{CCtx}$
-  \item An object $I : \varr 1 {\text{CCtx}}$
-  \item An action of $\text{VCtx}$ on $\text{CCtx}$, i.e.,
-    \begin{enumerate}
-    \item A functor $\oslash : \text{VCtx}\times \text{CCtx} \to \text{CCtx}$
-    \item Such that $1 \oslash \underline\Delta = \underline\Delta$
-    \item And $(\Gamma \times \Delta) \oslash \underline\Sigma = \Gamma \oslash \Delta \oslash \underline\Sigma$
-    \end{enumerate}
-  \item A category $\text{CTy}$ with a profunctor $\vdash_C : \harr
-    {\text{CCtx}}{\text{CTy}}$ and a fully faithful functor $\text{var}_C$ that represents it
-    \[ \underline\Delta \vdash_C \underline B \cong \punitinXfromYtoZ {} {\underline\Delta}{\text{var}_C\underline B}\]
-  \end{enumerate}
-  Furthermore we can add in the universal properties of any
-  non-contravariant CBPV connectives. For instance:
-  \begin{enumerate}
-  \item $F : \varr {VTy} {CTy}$ satisfying
-    \[ \Gamma \oslash \text{var}_C F A \vdash_C \underline B \cong \Gamma \times \text{var}_V A \oslash I \vdash_C \underline B  \]
-    \[ \Gamma \vdash_V U \underline B \cong \Gamma \oslash I \vdash_C \underline B  \]
-  \end{enumerate}
-\end{definition}
-
-\begin{definition}[Theory of unary Computational lambda calculus with booleans]
-  \begin{enumerate}
-  \item A category $C$
-  \item A terminal object $1$ of $C$
-  \item A monad $T$ on $C$
-  \item A type $\textrm{Bit} : \varr 1 C$
-  \item Morphisms $\textrm{y},\textrm{n} : \pendallXdotY * \punitinXfromYtoZ {1(*)} {\textrm{Bit}(*)}$
-  \item A natural transformation $\textrm{if} : \pendallXdotY * \homrallXYtoZ \alpha {(\punitinXfromYtoZ C {1(*)} \alpha) \times (\punitinXfromYtoZ C {1(*)} \alpha)} {\punitinXfromYtoZ C {\textrm{Bit}(*)} \alpha}$
-  \item Satisfying $\textrm{if} y = \pi_1$ and $\textrm{if} n = \pi_2$
-  \item A morphism $\textrm{print} : \pendallXdotY * \punitinXfromYtoZ {\textrm{Bit}(*)}{T (1(*))}$
-  \end{enumerate}
-\end{definition}
-
-Then we can ``program'' in this computational lambda calculus within
-our logical framework by constructing terms of type $\punitinXfromYtoZ
-C {A} {B}$ for any functors $A,B$ into $C$.
-
-Using a terminal category and the category of elements, we can define
-a formal category with families, an algebraic formulation of the
-judgmental structure of dependent type theory:
-\begin{definition}[Category with Families]
-  A category with families consists of
-  \begin{enumerate}
-  \item A category $\textrm{Ctx}\isaCat$
-  \item A presheaf $\textrm{Ty} : \harr {\textrm{Ctx}}{1}$
-  \item A presheaf $\textrm{Tm} : \harr {\Sigma_{\Gamma:\textrm{Ctx};*:1}\harrapp{\textrm{Ty}} \Gamma *} 1$
-  \item A terminal object in $\textrm{Ctx}$
-  \item An \emph{extension operation}
-    $\textrm{ext} : \Sigma_{\Gamma:\textrm{Ctx};*:1}\harrapp{\textrm{Ty}}\Gamma * \to \textrm{Ctx}$
-    right representing the profunctor
-    \[ \Delta:\textrm{Ctx};A:\Sigma_{\Gamma:\textrm{Ctx};*:1}\harrapp{\textrm{Ty}}{\Gamma}{*} \vdash \tensorexistsXwithYandZ {a}{(\punitinXfromYtoZ {} \Delta {\pi^-\pi^-a})}
-       {(\punitinXfromYtoZ {}{\pi^-a}{A})}\]
-       % {\Sigma_{\Gamma:\textrm{Ctx};*:1}\harrapp{\textrm{Ty}}{\Gamma}{*}}
-    % 
-  \end{enumerate}
-\end{definition}
-
-%% Can we do pras?
-%%
-%% A PRA mode theory is given by a free 2-category
-%% 1. Each mode m is a CwF whose context cat is [m]
-%% 2. Each 1-cell μ : m -> n is a functor [μ] : [n] -> [m] that has a DRA and is a PRA
-%% 3. Each 2-cell α : μ => ν is a 2-cell
-
-%% Let C, D be CwFs.
-%% A DRA is a pair of
-%% 1. A functor  L : C -> D
-%% 2. A morphism R : L^*(D_Ty) -> C_ty in ∫C, i.e., a morphism taking LΓ |- A to Γ |- RA
-%% 3. An iso     I : L^*(D_tm)?? =~ R^*(C_tm) in ∫(L^*(DTy)) a morphism taking LΓ |- M : A to Γ |- I(M) : RA
-
-%% Let C, D such that C has a terminal object.
-%% G : C -> D is a PRA if
-%% G / 1 : C -> D / G1 is a right adjoint
-
-%% What about fixed points?
-%% Say C has a terminal object
-%% fix : Hom(a,a) -> Hom(1, a)
-%% satisfying 
-%% fix[f] = f o fix[f]?
-%% We can write down fix, but not the equation I think
-
-%% \section{Additional Rules}
-
-%% The key property that the hom allows is that given
-%% any variable in context of a term $\Phi,\alpha,x,\beta,\Psi \vdash M :
-%% Q$ we can represent this term as a ``single-variable term'' by moving
-%% $\Phi$ and $\Psi$ to the right of the turnstile:
-
-%% \[ \alpha, x, \beta \vdash \lambda \Phi. \lambda \Psi M. : \Phi \triangleright Q \triangleleft \Psi \]
-
-%% This allows us to describe restricted positive eliminations that only
-%% work on single variable contexts. We give an admissible ``syntax
-%% sugar'' in Figure~\ref{fig:multi-hom}.
-
-%% Any $\Phi_l,\beta:C,\Phi_r \vdash M : R$ can be represented as a term
-%% $\beta:C \vdash M' : \Phi_r \triangleright_{\beta:C} R \triangleleft_{\beta:C} $
-
-%% Say we have a term in context
-%% \[ \alpha,x:P,\beta,y:Q,\gamma \vdash M : R \]
-%% Using $\triangleright\triangleleft$, we can produce an equivalent term
-%% of a different output type whose context is one of the single
-%% variables $\alpha,\beta$ or $\gamma$ in the context.
-
-%% We can make it $\alpha$ by using $\triangleleft$ twice:
-%% \[ \alpha \vdash \lambda^\triangleleft (x:P,\beta). \lambda^\triangleleft (y:Q, \gamma). M : \forall \beta. (\forall \gamma. R \triangleleft Q) \triangleleft P \]
-
-%% Or $\gamma$ by using $\triangleright$ twice:
-%% \[ \gamma \vdash\lambda^\triangleright (\beta, y:Q).  \lambda^\triangleright (\alpha, x:P). M : \forall \gamma. Q \triangleright \forall \alpha. P \triangleright R \]
-
-%% And we have two different ways to make the context $\beta$, either we use $\triangleright$ and then $\triangleleft$:
-%% \[ \beta \vdash\lambda^\triangleright (\alpha, x:P).  \lambda^\triangleleft (y:Q, \gamma). M : \forall \alpha. P \triangleright (\forall \gamma. R \triangleleft Q) \]
-%% or we do the opposite order:
-%% \[ \beta \vdash \lambda^\triangleleft (y:Q, \gamma). \lambda^\triangleright (\alpha, x:P). M : \forall \gamma. (\forall \alpha. P \triangleright R) \triangleleft Q \]
-
-%% In general we can iterate this process. That is given any $\Phi,\Psi$
-%% with $d^+\Phi = d^-\Psi = \beta:D$, we can take a term $\Phi \jnctx
-%% \Psi \vdash M : R$ and produce an equivalent term $\beta:D \vdash M :
-%% \Phi \triangleright R \triangleleft \Psi$. To formalize this, we show
-%% how we can create ``multi-hom'' jumbo connectives $\Phi \triangleright
-%% R$ and $R \triangleleft \Psi$, where the boundary conditions on
-%% $\Phi,\Psi$ are analogous to the ordinary case. These are given in
-%% Figure\ref{multi-hom}.
-
-%% \begin{figure}
-%%   \begin{align*}
-%%     P \triangleleft \gamma:E &= P\\
-%%     P \triangleleft (\Psi,x:R,\gamma:E) &= (\forall \gamma: E. P \triangleleft R) \triangleleft \Psi\\
-%%     %% \cdot \triangleright_{\alpha : C} P &= P\\
-%%     %% (x:R,\beta:D,\Phi_r)\triangleright_{\alpha:C} P  &= \Phi_R \triangleright_{\beta:D} \forall \beta. R \triangleright P
-%%   \end{align*}
-%%   \begin{mathpar}
-%%     \inferrule
-%%     {\Gamma \pipe d^+\Xi; \gamma: E \vdash P \isaSet\and
-%%      \Gamma \pipe d^-\Xi; \gamma:E \vdash \Psi \isavectx}
-%%     {\Gamma \pipe \Xi \vdash P \triangleleft \Psi}
-
-%%     \inferrule*[right=Covariant Multi-Hom introduction]
-%%     {\Gamma\pipe\Phi\jnctx \Psi \vdash M : P}
-%%     {\Gamma\pipe\Phi \vdash \lambda^\triangleleft \Psi. M : P \triangleleft \Psi}\and
-
-%%     \inferrule*[right=Covar Multi-Hom elim]
-%%     {~}
-%%     {\Gamma \pipe f:P\triangleleft \Psi,\Psi \vdash \textrm{app}\triangleleft_\Psi : P}
-
-%%     %%   (\lambda^\triangleleft \Psi. M)[\phi] = \lambda^\triangleleft \Psi. M[\phi,\Psi]\\
-%%     %%   \lambda^\triangleleft \gamma. M = M\\
-%%     %%   \lambda^\triangleleft (\Psi,x:R,\gamma:\cat E). M = \lambda^\triangleleft \Psi. \lambda^\triangleleft (x:R,\gamma:\cat E). M\\
-
-%%     %%   (\lambda^\triangleleft \gamma. M)[\phi] = M[\phi] = \lambda^\triangleleft \gamma. M[\phi]\\
-%%     %%   (\lambda^\triangleleft (\Psi,x:r,\gamma:\cat E). M)[\phi] =
-%%     %%   (\lambda^\triangleleft \Psi. \lambda^\triangleleft (x:R,\gamma:\cat E). M)[\phi]
-%%     %%   = \lambda^\triangleleft \Psi. (\lambda^\triangleleft (x:R,\gamma:\cat E). M)[\phi, \Psi]
-%%     %%   = \lambda^\triangleleft \Psi. \lambda^\triangleleft (x:R,\gamma:\cat E). M[\phi,\Psi,x:R,\gamma]
-
-%%     %%   \inferrule*[right=Covariant Multi-Hom elimination]
-%%     %%   {\Phi_1 \vdash M : P\triangleleft \Phi_2 \and \phi_2 : \Psi \vdash \Phi_2 \subst(\xi)}
-%%     %%   {\Phi_1,\Psi \vdash M\triangleleft \phi_2 : P[d^+\xi]}\\
-
-%%     %%   M \triangleleft (A/\gamma) = M[A/\gamma]\\
-%%     %%   M \triangleleft (\phi, N/x,A/\gamma) =  M \triangleleft (N, A) \triangleleft \phi
-
-%%     %%   (M \triangleleft \phi_2)[\phi_1,\psi] = M[\phi_1] \triangleleft [\phi_2[\psi]]\\
-%%     %%   TODO: prove
-
-%%     %%   TODO: (\lambda^\triangleleft \Phi. M) \triangleleft \phi = M[\phi]
-%%     %%   TODO: M = \lambda^\triangleleft \Phi. M\triangleleft \Phi
-
-%%     %%   TODO: contravariant by duality
-%%     %%   %% \inferrule*[right=Contra Multi-Hom formation]
-%%     %%   %% {\Phi \isavectx(\alpha:\cat C; d^-\Xi) \and P \prof(\alpha : \cat C; d^+\Xi)}
-%%     %%   %% {\Phi \triangleright P \prof(\xi)}
-
-%%     %%   %% \alpha \triangleright P = P\\
-%%     %%   %% \alpha:\cat C, x:R, \Phi \triangleright P = \forall \alpha. R \triangleright \Phi \triangleright P\\
-%%   \end{mathpar}
-%%   \begin{align*}
-%%     \lambda^\triangleleft \gamma:E. M &= M\\
-%%     \lambda^\triangleleft (\Psi,x:R,\gamma:E). M &= \lambda^\triangleleft \Psi. \lambda^\triangleleft (x:R,\gamma:E). M\\\\
-%%     \textrm{app}\triangleleft_{\gamma:E} &= f\\
-%%     \textrm{app}\triangleleft_{\Psi,x:R\gamma:E} &= \textrm{app}\triangleleft_\Psi \triangleleft (x,\gamma) \\
-%%   \end{align*}
-%%   \caption{Multi-Hom}
-%%   \label{multi-hom}
-%% \end{figure}
-
-%% \section{Unit and Tensor}
-%% we can define the
-%% \emph{unit} , which in set-theoretic settings we
-%% might call the \emph{identity/equality judgment} $A \equiv_{\cat C}
-%% A'$.
-%% %
-%% First the formation rule says given two objects $A,A'$ in the same
-%% category $\cat C$, we can form the hom-set from $A$ to $A'$ $A
-%% \to_{\cat C} A'$.
-%% %
-%% The introduction rule says that given any object $A$, we can define
-%% the identity arrow from $A$ to itself $\id(A) : A \to_{\cat C} A$.
-%% %
-%% Without knowing more about the category in question, these are the
-%% only arrows we know exist.
-%% %
-
-%% The elimination form is trickiest to define in a natural deduction style.
-%% %
-%% First, let's consider the universal property of the Hom. From
-%% Crutwell-Shulman, we want a 2-cell
-
-%% \[% https://tikzcd.yichuanshen.de/#N4Igdg9gJgpgziAXAbVABwnAlgFyxMJZARgBoAGAXVJADcBDAGwFcYkQBBEAX1PU1z5CKcqQBM1Ok1bsuvftjwEiY8ZIYs2iTjz4gMioUTLF10rSABOMAGaMekmFADm8IqBuWIAWyRkQOBBIqlKa7AASProeXr6IogFBiP6MWGAWUBDMAEaMbDQAFjD0UOyQ6SA0OPRY9trlbPIgnlHxVUkhqRXamTl5lSBFJWUE+QE1deCjDtxAA
-%% \begin{tikzcd}
-%%                     & A \arrow[ldd, no head, Rightarrow] \arrow[rdd, no head, Rightarrow] &   \\
-%%                     & refl                                                                &   \\
-%% A \arrow[rr, "Hom"] &                                                                     & A
-%% \end{tikzcd}\]
-
-%% That is universal/opcartesian in that for any 2-cell $t$:
-
-%% \[ % https://tikzcd.yichuanshen.de/#N4Igdg9gJgpgziAXAbVABwnAlgFyxMJZABgBpiBdUkANwEMAbAVxiRACEQBfU9TXfIRQBGclVqMWbbrxAZseAkQBMY6vWatEIAIIy+CwUQAsaiZrZ6eBgUpQA2Mxqnb2Acn1z+ioclXLxZy0QAGFPeVtfUwD1SWCAEXDvIxQAZlJhQLi2HA8ucRgoAHN4IlAAMwAnCABbJDIQHAgkYWsQKtqW6iakVRAGLDBgqDo4AAtCkGoJuig2SCHPDrrEPp7EdP7B4LgIAbnpmFmkMCYGBmoGOgAjGAYABWS7fphynCnzFxAACU625aQm3Wpi2i20I3Gk0Ox20C1Y-2qKwa6wArAjOogQet7OiVijus1EI5PsEAErcChcIA
-%% \begin{tikzcd}
-%% B \arrow[r] \arrow[rrdd] & {} \arrow[r, no head, dashed] & A \arrow[rr, "Hom"] &    & A \arrow[rr, no head, dashed] &  & B' \arrow[lldd] \\
-%%                          &                               &                     & t' &                               &  &                 \\
-%%                          &                               & C \arrow[rr, "R"]   &    & D                             &  &                
-%% \end{tikzcd}\]
-
-%% there exists a $t'$ which extends $t$ in that the following is equal
-%% to $t$:
-%% \[% https://tikzcd.yichuanshen.de/#N4Igdg9gJgpgziAXAbVABwnAlgFyxMJZABgBoAmAXVJADcBDAGwFcYkQAhEAX1PU1z5CKAIwVqdJq3Y8+IDNjwEi5cTQYs2iEAEFZ-RUKIAWNZM3s9vA4OUoAbGY3TtHAOT75ApcOSrjEs5aIADCngq2vqYB6lLBACLh3kYoAMykqYFx7Dge1l6GdsjpxFkW2lZyET5EZKWx5SBJhb5i9eYuTfnVKciO7UEy3clF6SJlnQBOMABmjDwSMFAA5vBEoDOTEAC2SGQgOBBIIvmbO8c0h0iqIIxYYMFQ9HAAFksgNG-0UOyQD55nXaIG5XRDpW73YJwCB3H6fGDfJBgZiMRg0Rj0ABGMEYAAURsJbrMcB8OsEABLnU5bIHg0GmCH-bRPV7veGI7R-NjU86IfaggCsPKBDNB9mFSAFlyOiEcZPYACUATSkAAOaXXdGQ9hQCDMTGMNjsn6cghGg70LDzU3-CWIdUHGXgu5MkC6-WG0lfE3gM2knCW62+21yQFIACcGr5dsjjuO+xdUJhWDhIG9SJRaItVt+frtIn5MrEjKTsK9CJNyNRl0DuZDGxViALUYdiZ1zzeqfTNu5ocbDtBImLbeZHbZaYrdd7Dd5Q6jDJHbr1BvN3eD5oDOZ7C24QA
-%% \begin{tikzcd}
-%% {} \arrow[dd] \arrow[r]  & {} \arrow[dd] \arrow[rr, no head, dashed] &                     & A \arrow[ldd, no head, Rightarrow] \arrow[rdd, no head, Rightarrow] \arrow[rrr, no head, dashed] &                               &  & {} \arrow[dd, no head, Rightarrow] \\
-%%                          &                                           &                     & refl                                                                                             &                               &  &                                    \\
-%% B \arrow[r] \arrow[rrdd] & {} \arrow[r, no head, dashed]             & A \arrow[rr, "Hom"] &                                                                                                  & A \arrow[rr, no head, dashed] &  & B' \arrow[lldd]                    \\
-%%                          &                                           &                     & t'                                                                                               &                               &  &                                    \\
-%%                          &                                           & C \arrow[rr, "R"]   &                                                                                                  & D                             &  &
-%% \end{tikzcd}\]
-
-%% Clearly, our introduction rule models this first 2-cell. Furthermore,
-%% the construction in the second condition is modeled faithfully by the
-%% following sequent rule in the figure. This is a J-like induction rule:
-%% to construct a term using an input of the Hom type, it is sufficient
-%% to consider the case in which the input is the identity morphism.
-%% %
-%% However, building in a substitution the usual way we would (pushing
-%% the substitution into the continuation $N$) doesn't work properly
-%% because we can't necessarily extract a substitution for the
-%% continuation from the substitution for the the whole term. A
-%% substitution for the whole term would be of the form
-%% \[\phi,D_1/\delta_1,M/f,D_2/\delta_2,\psi\]
-%% The the candidate for the continuation's substitution would be
-%% $\phi,?,\psi$ but there is not necessarily something to go in the
-%% middle here since $D_1,D_2$ do not necessarily have the same
-%% \emph{input variable}. However, if we assume the existence already of
-%% the cotensor products the problem can be simplified. Instead of
-%% allowing for a general elimination of a hom term anywhere in the
-%% context, we can instead make a remarkably restricted rule: we can only
-%% eliminate a Hom term when it is the \emph{sole} input in the context:
-
-%% \[ \inferrule*[right=Hom Elim Very Simple]{\alpha:\cat C \vdash N : Q[\alpha/\alpha_1,\alpha/\alpha_2]}{\alpha_1:\cat C, f : \alpha_1 \to_{\cat C} \alpha_2, \alpha_2:\cat C \vdash J(\alpha_1.\alpha_2.
-Q, \alpha.
-N, \alpha_1, f, \alpha_2) : Q}\]
-
-%% While this rule looks very impoverished at first sight, it has the
-%% virtue of easily supporting a built-in cut:
-
-%% \[ \inferrule*[right=Hom Elim VSimple Subst]{
-%%   Q \prof(\alpha_1:\cat C; \alpha_2:\cat C)\and
-%%   \Phi \vdash M : A_1 \to_{\cat C} A_2\and
-%%   \alpha:\cat C \vdash N : Q[\alpha/\alpha_1,\alpha/\alpha_2]}{\Phi \vdash J(\alpha_1.\alpha_2. Q, \alpha. N, A_1, M, A_2) : Q[A_1/\alpha_1, A_2/\alpha_2]}\]
-
-%% However, this rule on its own clearly is not enough to support our
-%% desired universal property shown above diagrammatically. To prove our
-%% completeness theorem, we will need to show that this rule is enough to
-%% give us the desired universal property \emph{assuming the presence of
-%%   left and right cotensors}.
-
-%% So first, we must show that the ``proper'' sequent rule
-
-%% \[ \inferrule*[right=Hom-Left (Sequent Style)]
-%%     {\Phi_1[\alpha/\alpha_1], \alpha:\cat C, \Phi_2[\alpha/\alpha_2]\vdash Q[\alpha/\alpha_1,\alpha/\alpha_2]}
-%%     {\Phi_1, \alpha_1 : \cat C, f : \alpha_1 \to_{\cat C} \alpha_2, \alpha_2 : \cat C, \Phi_2\vdash : Q}
-%% \]
-
-%% is admissible.
-
-%% Which we can sketch as follows:
-
-%% \[
-%% \inferrule*
-%%     {\inferrule*{\inferrule*{\Phi_1[\alpha/\alpha_1],\alpha:\cat C, \Phi_2[\alpha/\alpha_1] \vdash Q[\alpha/\alpha_1;\alpha/\alpha_2]}{\alpha:\cat C \vdash \Phi_1[\alpha/\alpha_1] \triangleright Q[\alpha/\alpha_1;\alpha/\alpha_2] \triangleleft \Phi_2[\alpha/\alpha_2]}}{\alpha_1 : \cat C, f : \alpha_1 \to_{\cat C} \alpha_2, \alpha_2 : \cat C\vdash \Phi_1 \triangleright Q \triangleleft \Phi_2}}
-%%     {\Phi_1, \alpha_1 : \cat C, f : \alpha_1 \to_{\cat C} \alpha_2, \alpha_2 : \cat C, \Phi_2\vdash Q}
-%% \]
-
-%% Which using our syntax sugar is given by the term:
-%% \begin{mathpar}
-%%   \inferrule*
-%%     {\inferrule*
-%%       {\inferrule*
-%%         {\Phi_1[\alpha/\alpha_1],\alpha:\cat C, \Phi_2[\alpha/\alpha_1] \vdash N : Q[\alpha/\alpha_1;\alpha/\alpha_2]}
-%%         {\alpha:\cat C \vdash \lambda^\triangleright \Phi_1[\alpha/\alpha_1].
-\lambda^\triangleleft \Phi_2[\alpha/\alpha_2].
-N : \Phi_1[\alpha/\alpha_1] \triangleright Q[\alpha/\alpha_1;\alpha/\alpha_2] \triangleleft \Phi_2[\alpha/\alpha_2]}}
-%%       {\alpha_1 : \cat C, f : \alpha_1 \to_{\cat C} \alpha_2, \alpha_2 : \cat C\vdash J(\alpha_1.\alpha_2.\Phi_1 \triangleright Q \triangleleft \Phi_2, \alpha.
-\lambda^\triangleright \Phi_1[\alpha/\alpha_1].
-\lambda^\triangleleft \Phi_2[\alpha/\alpha_2].
-N, \alpha_1,f,\alpha_2) : \Phi_1 \triangleright Q \triangleleft \Phi_2}}
-%%     {\Phi_1, \alpha_1 : \cat C, f : \alpha_1 \to_{\cat C} \alpha_2, \alpha_2 : \cat C, \Phi_2\vdash
-%%       \Phi_1 \triangleright (J(\alpha_1.\alpha_2.\Phi_1 \triangleright Q \triangleleft \Phi_2, \alpha.
-\lambda^\triangleright \Phi_1[\alpha/\alpha_1].
-\lambda^\triangleleft \Phi_2[\alpha/\alpha_2].
-N, \alpha_1,f,\alpha_2)) \triangleleft \Phi_2 : Q}
-%% \end{mathpar}
-
-%% We need to show $\beta$ and $\eta$ hold.
-
-%% First, $\beta$.
-
-%% \begin{align*}
-%%   &(\Phi_1 \triangleright (J_\to(\alpha_1.\alpha_2.\Phi_1 \triangleright Q \triangleleft \Phi_2, \alpha.
-\lambda^\triangleright \Phi_1[\alpha/\alpha_1].
-\lambda^\triangleleft \Phi_2[\alpha/\alpha_2].
-N, \alpha_1,f,\alpha_2)) \triangleleft \Phi_2) [\Phi_1[A/\alpha_1],A/\alpha_1,\id(A),A/\alpha_2,\Phi_2[A/\alpha_2]] \\
-%%   &=
-%%   \Phi_1[A/\alpha_1] \triangleright (J_\to(\alpha_1.\alpha_2.\Phi_1 \triangleright Q \triangleleft \Phi_2, \alpha.
-\lambda^\triangleright \Phi_1[\alpha/\alpha_1].
-\lambda^\triangleleft \Phi_2[\alpha/\alpha_2].
-N, A,\id(A), A)) \triangleleft \Phi_2[A/\alpha_2]\\
-%%   &=
-%%   \Phi_1[A/\alpha_1] \triangleright (\lambda^\triangleright \Phi_1[A/\alpha_1]. \lambda^\triangleleft \Phi_2[A/\alpha_2]. N[A/\alpha]) \triangleleft \Phi_2[A/\alpha_2]\\
-%%   &=  (\lambda^\triangleleft \Phi_2[A/\alpha_2]. N[A/\alpha]) \triangleleft \Phi_2[A/\alpha_2]\\
-%%   &=  N[A/\alpha]
-%% \end{align*}
-
-%% \begin{figure}
-%%   \begin{mathpar}
-%%     \inferrule*[right=Full Unit Elim*]
-%%     {\Phi_1[\alpha/\alpha_1], \alpha:\cat C, \Phi_2[\alpha/\alpha_2]\vdash N : Q[\alpha/\alpha_1,\alpha/\alpha_2]}
-%%     {\Phi_1, \alpha_1 : \cat C, f : \alpha_1 \to_{\cat C} \alpha_2, \alpha_2 : \cat C, \Phi_2\vdash J_\to'(\Phi_1, \alpha_1 : \cat C, f : \alpha_1 \to_{\cat C} \alpha_2, \alpha_2 : \cat C, \Phi_2, Q, \alpha.
-N) : Q}
-
-%%     J_\to'(\Phi_1, \alpha_1 : \cat C, f : \alpha_1 \to_{\cat C} \alpha_2, \alpha_2 : \cat C, \Phi_2, Q, \alpha. N) =
-%%     \Phi_1 \triangleright (J(\alpha_1.\alpha_2.\Phi_1 \triangleright Q \triangleleft \Phi_2, \alpha.
-\lambda^\triangleright \Phi_1[\alpha/\alpha_1].
-\lambda^\triangleleft \Phi_2[\alpha/\alpha_2].
-N, \alpha_1,f,\alpha_2)) \triangleleft \Phi_2
-
-%%     \inferrule*[right=Full Unit $\beta$*]
-%%     {}
-%%     {J_\to'(\Phi_1, \alpha_1 : \cat C, f : \alpha_1 \to_{\cat C} \alpha_2, \alpha_2 : \cat C, \Phi_2, Q, \alpha.
-N)[\Phi_1[\alpha/\alpha_1],\alpha/\alpha_1, \id(\alpha)/f, \alpha/\alpha_2, \Phi_2[\alpha/\alpha_2]]
-%%     = N}
-
-%%     (\Phi_1 \triangleright J(\alpha_1.\alpha_2.\Phi_1 \triangleright Q \triangleleft \Phi_2, \alpha.
-\lambda^\triangleright \Phi_1[\alpha/\alpha_1].
-\lambda^\triangleleft \Phi_2[\alpha/\alpha_2].
-N, \alpha_1,f,\alpha_2) \triangleleft \Phi_2)[\Phi_1[\alpha/\alpha_1],\alpha/\alpha_1, \id(\alpha)/f, \alpha/\alpha_2, \Phi_2[\alpha/\alpha_2]]\\
-%%     = \Phi_1[\alpha/\alpha_1] \triangleright J(\alpha_1.\alpha_2.\Phi_1 \triangleright Q \triangleleft \Phi_2, \alpha.
-\lambda^\triangleright \Phi_1[\alpha/\alpha_1].
-\lambda^\triangleleft \Phi_2[\alpha/\alpha_2].
-N, \alpha,\id(\alpha),\alpha) \triangleleft \Phi_2[\alpha/\alpha_2]\\
-%%     = \Phi_1[\alpha/\alpha_1] \triangleright
-%%     (\lambda^\triangleright \Phi_1[\alpha/\alpha_1]. \lambda^\triangleleft \Phi_2[\alpha/\alpha_2]. N)
-%%     \triangleleft \Phi_2[\alpha/\alpha_2]\\
-%%     = N
-
-%%     \inferrule*[right=Unit$\eta$]
-%%     {\Phi_1, \alpha_1 : \cat C, f : \alpha_1 \to_{\cat C} \alpha_2, \alpha_2 : \cat C, \Phi_2 \vdash M : Q}
-%%     {M =
-%%       J_\to'(\Phi_1, \alpha_1 : \cat C, f : \alpha_1 \to_{\cat C} \alpha_2, \alpha_2 : \cat C, \Phi_2, Q, \alpha.
-M[\Phi_1[\alpha/\alpha_1],\alpha/\alpha_1,\id(\alpha), \alpha/\alpha_2,\Phi_2[\alpha/\alpha_2]])}
-
-%%     M
-%%     =(\triangleleft\beta\triangleright\beta)
-%%     \Phi_1 \triangleright
-%%     (\lambda^\triangleright \Phi_1. \lambda^\triangleleft \Phi_2. M)
-%%     \triangleleft \Phi_2
-%%     = \Phi_1 \triangleright
-%%     J_\to(\alpha_1.\alpha_2.Q, \alpha.
-\lambda^\triangleright \Phi_1[\alpha/\alpha_1].
-\lambda^\triangleleft \Phi_2[\alpha/\alpha_2].
-M[\alpha/\alpha_1, \id(\alpha)/f,\alpha/\alpha_2], \alpha_1, f, \alpha_2)
-%%     \triangleleft \Phi_2
-%%     =
-%%     J_\to'(\Phi_1, \alpha_1 : \cat C, f : \alpha_1 \to_{\cat C} \alpha_2, \alpha_2 : \cat C, \Phi_2, Q, \alpha. N)
-%%   \end{mathpar}  
-%%   \caption{Admissible Unit Universal Property Rules}
-%% \end{figure}
-
-%% First, the syntax for tensor is based on the coend formula that gives
-%% the definition in $Cat$, but using logical syntax $\exists$ rather
-%% than integrals because I don't like integral syntax and the syntax
-%% matches existential quantification in type theory.
-%% %
-%% Note that $\exists \beta. ~Q \odot R$ is the application of a single
-%% connective to $Q$ and $R$, and not separable into a quantifier
-%% $\exists$ connective and a product connective $\odot$.
-%% %
-%% Introduction, elimination and $\beta\eta$ are similar to a combination
-%% of existential quantification and the tensor product in linear logic.
-%% %
-%% The introduction rule is justified by the fact that semantically what
-%% we have is a transformation
-%% \[ \textrm{pair} : (\alpha : \cat C,Q,\beta:\cat D,R,\gamma:\cat E) \Rightarrow \exists \beta. R \odot Q \]
-%% And the introduction rule is just building in a substitution to this
-%% transformation.
-%% %
-
-%% For the elim rule we can pull a similar trick as with the Hom
-%% type. First, the rule where there's only one variable.
-%% \[ \inferrule*[right=Tensor elim Simple]
-%%   {\alpha : \cat C, q : Q, \beta : \cat D, r: R, \gamma : \cat E \vdash N : S}
-%%   {\alpha : \cat C, x : \exists \beta. Q \odot R, \gamma : \cat E \vdash J_\odot(x, q.\beta.r. N) : S}
-%% \]
-
-%% \begin{figure}
-%% \begin{mathpar}
-%%     \inferrule*[right=Full Tensor Elim*]
-%%     {\Phi_1, \alpha:\cat C, q:Q, \beta:\cat D, r:R, \gamma:\cat E \Phi_2\vdash N : S}
-%%     {\Phi_1, \alpha : \cat C, x : \exists \beta:\cat D. Q \odot R, \gamma : \cat E, \Phi_2
-%%       \vdash J_\odot'(\Phi_1, \alpha : \cat C, x : \exists \beta:\cat D. Q \odot R, \gamma : \cat E, \Phi_2; S;q.\beta.r.N) : S}
-
-%%     J_\odot'(\Phi_1, \alpha : \cat C, x : \exists \beta:\cat D. Q \odot R, \gamma : \cat E, \Phi_2;S; q.\beta.r.N)
-%%     = \Phi_1 \triangleright J_\odot(\alpha.\gamma.\Phi_1 \triangleright S \triangleleft \Phi_2, \alpha.q.\beta.r.\gamma.
-\lambda^\triangleleft \Phi_1.
-\lambda^\triangleright\Phi_2.
-N, \alpha, x, \gamma) \triangleleft \Phi_2
-
-%%     J_\odot'(\Phi_1, \alpha : \cat C, x : \exists \beta:\cat D. Q \odot R, \gamma : \cat E, \Phi_2; q.\beta.r.N)[\Phi_1,\alpha/\alpha,(q,\beta,r)/x,\gamma/\gamma,\Phi_2] = N
-
-%%   \inferrule*[right=Full Tensor elim*]{ S \prof(\alpha:\cat C; \gamma : \cat E)\and
-%%     \Phi \vdash M : \exists \beta. Q[A/\alpha] \odot R[G/\gamma]\and
-%%     \alpha : \cat C, q : Q, \beta : \cat D, r: R, \gamma : \cat E \vdash N : S}
-%%   {\Phi \vdash J_\odot(\alpha. \gamma. S, \alpha.q.\beta.r.\gamma.N, A, M, G) : S[A/\alpha, G/\gamma]}
-
-%%   J_\odot(\alpha. \gamma. S, \alpha.q.\beta.r.\gamma.N, A, M, G)[\phi] =
-%%   J_\odot(\alpha. \gamma. S, \alpha.q.\beta.r.\gamma.N, A[d^-\phi], M[\phi], G[d^+\phi])
-
-%%   %% \inferrule*[right=Tensor elim Sequent]
-%%   %% {\Phi,\alpha : \cat C, q : Q, \beta : \cat D, r: R, \gamma : \cat E,\Psi \vdash N : S}
-%%   %% {\Phi, \alpha : \cat C, x : \exists \beta. Q \odot R, \gamma : \cat E,\Psi \vdash J_\odot(x;q.\beta.r. N) : S}
-
-%%   %% \inferrule*[right=Tensor elim Subst]
-%%   %% {\Phi,\alpha : \cat C, q : Q, \beta : \cat D, r: R, \gamma : \cat E,\Phi' \vdash N : S
-%%   %% \and
-%%   %% \phi : \Psi \to \Phi,\alpha : \cat C, q : Q, \beta : \cat D, r: R, \gamma : \cat E,\Phi'}
-%%   %% {\Psi  \vdash J_\odot(x;q.\beta.r. N)[\phi] : S}
-
-%%   \inferrule*[right=Tensor $\beta$]
-%%   {}{J_\odot(\alpha.\gamma.S, \alpha.q.\beta.r.\gamma.N,A,(M,B,N),G) = N[A/\alpha,M/q,B/\beta,N/r,G/\gamma]}
-
-%%   \inferrule*[right=Tensor$\eta$]
-%%   {\alpha : \cat C, x : \exists \beta. Q \odot R, \gamma : \cat E \vdash M : S}
-%%   {\alpha : \cat C, x : \exists \beta. Q \odot R, \gamma : \cat E \vdash M =
-%%     J(\alpha.\gamma.S, \alpha.q.\beta.r.\gamma. M[\alpha/\alpha,(q,\beta,r)/x,\gamma/\gamma])
-%%     : S}
-%% \end{mathpar}
-%% \caption{Tensor Admissible Universal Property Rules}
-%% \label{fig:tensor}
-%% \end{figure}
-
-%% \section{Internal Category Theory}
-
-%% A downside of working synthetically is that since everything we define
-%% must be functorial/natural, we cannot prove any theorems about
-%% functions that are not assumed to be functorial or transformations
-%% that are not assumed ot be natural.
-%% %
-%% These kinds of constructions are surprisingly common in basic category theory.
-%% %
-%% For instance, when defining a left adjoint $F$ to a given right
-%% adjoint functor $U$, the functoriality of $F$ is often not assumed,
-%% rather we are given that each $F a$ has a certain universal mapping
-%% property and this can then be used to show the action of $F$ on
-%% objects extends to an action on morphisms.
-%% %
-%% This definition is quite important for applications of category theory
-%% to type theory: in type theory when a type constructor is defined it
-%% is given introduction, elimination and $\beta\eta$ rules, but no
-%% primitive syntax is given for a functorial action.
-%% %
-%% The fact that functoriality is determined from the universal mapping
-%% property of the type is then a key step in proving the type theory is
-%% sound and complete with respect to models which are expected to have
-%% certain functors on them.
-
-%% As another example, a common lemma states that an \emph{possibly
-%% unnatural} transformation is in fact natural if it is the inverse of a
-%% natural transformation.
-%% %
-%% Neither of these examples can be stated as synthetic theorems, as it
-%% is impossible to quantify over ``possibly unnatural transformations''
-%% or ``functions on objects''.
-%% %
-%% Inside the calculus, we cannot ``see'' that categories have an
-%% underlying set of objects.
-%% %
-%% Fortunately there is a way to use our calculus in a different manner
-%% to study these constructions: by defining categories \emph{internal
-%% to} SCT.
-%% %
-%% This reproduces classical constructions, since there is a model where
-%% a category $\cat C$ is interpreted simply as a set, functors as
-%% functions and profunctors $\harr {\cat C}{\cat D}$ as \emph{spans},
-%% i.e. sets $R$ with (source and target) projections $\cat C \leftarrow
-%% R \rightarrow \cat D$.
-%% %
-%% The proofs read like the classical proofs (with some care additional
-%% attention on the ordering of variable usage), but by virtue of being
-%% syntactic automatically apply to any notion of category internal to a
-%% virtual equipment, which includes classical internal categories,
-%% enriched categories and a broad range of generalized multicategory.
-
-%% We note that a synthetic approach to model this situation may be
-%% possible were we to extend our calculus with a \emph{modality} $|\cat
-%% C|$ that takes a category to the discrete category on the set of
-%% objects. We discuss this in \cref{sec:discussion}.
-
-%% Next, we introduce the notion of category internal to our syntactic
-%% virtual equipment and the \emph{monoids and modules} translation that
-%% shows that, with the inclusion of hom sets, our syntax is already
-%% suitable for doing syntactic category theory.
-
-%% The benefits of our syntax become clear when we define internal
-%% categories, functors, profunctors and transformations: the definitions
-%% read exactly as the classical definitions.
-%% \begin{definition}[Internal Category]
-%%   An \emph{(internal) category} consists of\footnote{the following is
-%%     easily interpreted as a $\Sigma$ type}:
-%%   \begin{enumerate}
-%%   \item A sort $C_0 \isaCat$
-%%   \item A span $C_1 : \harr{C_0}{ C_0}$
-%%   \item Identity arrows
-%%     $C_r : \pendallXdotY {\alpha :  C_0} {\harrapp{C_1} \alpha \alpha}$
-%%   \item Composition of arrows
-%%     \[ C_t : \pendallXdotY \alpha {\homrallXYtoZ \beta {\harrapp{C_1} \alpha \beta} {\homrallXYtoZ \gamma {\harrapp{C_1} \beta \gamma} {\harrapp{C_1} \alpha \gamma}}}\]
-
-%%     %% \[ C_t^\triangleright : \forall \alpha :  C_0. \forall \beta. ({C_1} \beta \alpha) \triangleright \forall \gamma. ({C_1} \gamma \beta) \triangleright {C_1} \gamma \alpha \]
-
-%%     %% or
-
-%%     %% \[ C_t^\triangleleft : \forall \alpha. \forall \beta. (\forall \gamma. {C_1} \alpha \gamma \triangleleft {C_1} \beta \gamma) \triangleleft {C_1} \alpha \beta
-%%     %% \]
-
-%%     %% or using tensors and covariant homs
-
-%%     %% \[ C_t^\odot : \forall \alpha. \forall \gamma. (\exists \beta. {C_1}\alpha\beta\odot {C_1}\beta\gamma)\triangleright {C_1}\alpha \gamma \]
-
-%%     %% or using tensors and contravariant homs
-
-%%     %% \[ C_t^\odot : \forall \gamma. \forall \alpha. {C_1}\alpha\gamma \triangleleft (\exists \beta. {C_1}\alpha\beta\odot {C_1}\beta\gamma)\]
-%%   \item Satisfying right unitality, i.e., for any $\Phi \vdash f :
-%%     \harrapp {C_1} A B$, we'd like $f; r(B) = f$. We can formalize this
-%%     using an equality of two transformations of type $\pendallXdotY
-%%     \alpha {\homrallXYtoZ \beta {\harrapp{C_1} \alpha \beta} {\harrapp{C_1} \alpha \beta}}$
-
-%%     \[ \pendlambdaXdotY \alpha {\homrlambdaXatYdotZ f \beta {\homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {C_t} \alpha} {f} \beta} {\pendappXtoY {C_r} \beta} \beta}}
-%%     = \pendlambdaXdotY \alpha {\homrlambdaXatYdotZ f \beta f} \]
-
-%%     These kinds of equalities between 2-cells are very common in
-%%     internal category theory, so to save space we will suppress the
-%%     sequence of $\lambda$s that start such equalities. So the above can simply be written as
-
-%%     \[ { {\homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {C_t} \alpha} {f} \beta} {\pendappXtoY {C_r} \beta} \beta}} = { f}\]
-
-%%   \item and left unitality (an equality of the same type as right unitality):
-
-%%     \[ { {\homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {C_t} \alpha} {\pendappXtoY {C_r} \alpha} \alpha} f \beta}} = { f}\]
-%%   \item and associativity, which is an equality at type
-%%     \[ \pendallXdotY \alpha {\homrallXYtoZ \beta {\harrapp{C_1} \alpha \beta} {\homrallXYtoZ \gamma {\harrapp{C_1} \beta \gamma} {\homrallXYtoZ \delta {\harrapp{C_1} \gamma \delta} {\harrapp{C_1} \alpha \delta}}}}\]
-
-%%     \[ \homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {C_t}\alpha} f \beta} {(\homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {C_t}\beta} g \gamma} h \delta)} \delta
-%%     =
-%%     \homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {C_t}\alpha} {(\homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {C_t}\alpha} f \beta} g \gamma)} \gamma} h \delta \]
-%%   \end{enumerate}
-%% \end{definition}
-
-%% We can define an internal functor to be a pair of an object function
-%% and a function on arrows that preserves identity and 
-%% \begin{definition}[Internal Functor]
-%%   A functor $F$ between internal categories $C$ and $D$ consists of
-%%   \begin{enumerate}
-%%   \item a map of objects $F_0 : \varr {C_0} {D_0}$
-%%   \item a map of arrows
-%%     $F_1 : \pendallXdotY {\alpha}
-%%     {\homrallXYtoZ \beta {\harrapp{C_1} \alpha \beta} {\harrapp{D_1} {F_0 \alpha}{F_0 \beta}}}$
-%%   \item That preserves identity. This is given by equality between the
-%%     following terms of type $\forall \alpha:{{C_0}}. D_1 {(F_0 \alpha)}{(F_0\alpha)}$:
-%%     \[ \homrappXtoYatZ {\pendappXtoY {F_1}\alpha} {\pendappXtoY {C_r} \alpha} \alpha = \pendappXtoY {D_r}{F_0\alpha} \]
-%%   \item And preserves composition.
-%%     This is given by an equality between the following terms of type
-
-%%     \[ \pendallXdotY \alpha
-%%     \homrallXYtoZ {\beta} {\harrapp{C_1} \alpha \beta} {\homrallXYtoZ \gamma {\harrapp{C_1} \beta \gamma} {\harrapp{D_1} {F_0 \alpha} {F_0 \gamma}}}\]
-
-%%     First, we can compose in $C$, and then map using $F$:
-%%     \[ \homrappXtoYatZ {\pendappXtoY{F_1} \alpha} {(\homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {C_t} \alpha} f \beta} g \gamma)}\gamma\]
-
-%%     or we can map the morphisms with $F$, and then compose them in $D$:
-%%     \[ \homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {D_t} {(F_0 \alpha)}} {(\homrappXtoYatZ {\pendappXtoY {F_1} \alpha} f \beta)} {(F_0 \beta)}} {(\homrappXtoYatZ {\pendappXtoY {F_1} \beta} g \gamma)} {{F_0} \gamma}\]
-%%   \end{enumerate}
-%% \end{definition}
-
-%% We can immediately define identity and composition of functors.
-
-%% \begin{definition}
-%%   For any $C : \algCat$, we can define an identity functor
-%%   $\textrm{Id}_C : \algVarr C C$ by
-%%   \begin{enumerate}
-%%   \item $\textrm{Id}_0 = \lambda \alpha:C_0. \alpha$
-%%   \item $\textrm{Id}_1 = \pendlambdaXdotY \alpha
-%%     {\homrlambdaXatYdotZ f {\alpha'} f}$
-%%   \item Preservation of identity requires proving
-%%     \[
-%%     \lambda \alpha. \pendappXtoY {C_r}{\textrm{Id}_0\alpha} \]
-%%     is equal to
-%%     \[ \lambda \alpha.
-%%     \homrappXtoYatZ {\pendappXtoY {\textrm{Id}_1} \alpha} {(\pendappXtoY {C_r}\alpha)} \alpha\]
-%%     Both sides simply $\beta$ reduce to $ \lambda \alpha. \pendappXtoY {C_r} \alpha$
-%%     and the result follows by reflexivity.
-%%   \item Preservation of composition requires proving (under a $\lambda \alpha. \homrlambdaXatYdotZ f \beta \homrlambdaXatYdotZ g \gamma$)
-%%     \[
-%%     \homrappXtoYatZ {\pendappXtoY {{\textrm{Id}}_1} \alpha} {(\homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {C_t} \alpha} f \beta} g \gamma)}\gamma
-%%     = \homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {C_t} {{\textrm{Id}}_0 \alpha}} {(\homrappXtoYatZ {\pendappXtoY {{\textrm{Id}}_1} \alpha} f \beta)} {{\textrm{Id}}_0 \beta}} {(\homrappXtoYatZ {\pendappXtoY{{\textrm{Id}}_1} \beta} g \gamma)} {{\textrm{Id}}_0 \gamma}\]
-%%     Both sides $\beta$ reduce to
-
-%%     \[ \homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {C_t} \alpha} f \beta} g \gamma\]
-%%     so the result holds by reflexivity.
-%%   \end{enumerate}
-%% \end{definition}
-
-%% \begin{definition}
-%%   Given categories $C,D,E$ and functors $F : \algVarr C D$ and
-%%   $G : \algVarr D E$, we define the composition $G \circ F :
-%%   \algVarr C E$ as
-%%   \begin{enumerate}
-%%   \item $(G \circ F)_0 = \lambda \alpha. {G_0 ({F_0} \alpha)}$
-%%   \item $(G \circ F)_1 =
-%%     \pendlambdaXdotY \alpha {
-%%     \homrlambdaXatYdotZ f \beta
-%%     {\homrappXtoYatZ {\pendappXtoY {G_1} {F_0 \alpha}} {\homrappXtoYatZ {\pendappXtoY {F_1} \alpha} f \beta} {F_0 \beta}}}$
-%%   \item Preservation of identity requires proving
-%%     \[ \lambda \alpha. \pendappXtoY {E_r} {(G \circ F)_0\alpha} \]
-%%     is equal to
-%%     \[ \lambda \alpha.
-%%     \homrappXtoYatZ {\pendappXtoY {(G \circ F)_1} \alpha} {\pendappXtoY {C_r}\alpha} \alpha \]
-
-%%     Expanding definitions and $\beta$ reducing we need to show
-
-%%     \[ \pendlambdaXdotY\alpha {\pendappXtoY {E_r}{G_0 (F_0 \alpha)}} \]
-%%     is equal to
-%%     \[ \pendlambdaXdotY \alpha {\homrappXtoYatZ {\pendappXtoY {G_1} {F_0 \alpha}} {(\homrappXtoYatZ {\pendappXtoY {F_1} \alpha} {\pendappXtoY {C_r} \alpha} \alpha)} {F_0 \alpha}}\]
-
-%%     Proceeding from right to left applying transitivity:
-%%     \begin{align*}
-%%       &\pendlambdaXdotY \alpha {\homrappXtoYatZ {\pendappXtoY {G_1} {F_0 \alpha}} {(\homrappXtoYatZ {\pendappXtoY {F_1} \alpha} {\pendappXtoY {C_r} \alpha} \alpha)} {F_0 \alpha}}\\
-%%       &=\pendlambdaXdotY \alpha {\homrappXtoYatZ {\pendappXtoY {G_1} {F_0 \alpha}} {(\pendappXtoY{(\pendlambdaXdotY \alpha {\homrappXtoYatZ {\pendappXtoY {F_1} \alpha} {\pendappXtoY {C_r} \alpha} \alpha})}\alpha)}} {F_0 \alpha}\\
-%%       &=\pendlambdaXdotY \alpha {\homrappXtoYatZ {\pendappXtoY {G_1} {F_0 \alpha}} {(\pendappXtoY{(\pendlambdaXdotY \alpha {{\pendappXtoY {D_r} {F_0 \alpha}}})}\alpha)}} {F_0 \alpha}\tag{*}\\
-%%       &=\pendlambdaXdotY \alpha {\homrappXtoYatZ {\pendappXtoY {G_1} {F_0 \alpha}} {\pendappXtoY {D_r} {F_0 \alpha}}} {F_0 \alpha}\\
-%%       &= \pendlambdaXdotY \alpha
-%%       \pendappXtoY {(\pendlambdaXdotY \beta {{\homrappXtoYatZ {\pendappXtoY {G_1} \beta} {\pendappXtoY {D_r} \beta} \beta}})} {F_0 \alpha}\\
-%%       &= \pendlambdaXdotY \alpha
-%%       \pendappXtoY {(\pendlambdaXdotY \beta \pendappXtoY {E_r} {G_0 \beta})} {F_0 \alpha}\tag{*}\\
-%%       &= \pendlambdaXdotY \alpha \pendappXtoY {E_r} {G_0 (F_0 \beta)}\\
-%%     \end{align*}
-
-%%     Where the two marked lines follow from congruence using the
-%%     assumptions about $F,G$.
-
-%%   \item Preservation of Composition.
-
-%%     \[ \homrappXtoYatZ {\pendappXtoY{(G \circ F)_1} \alpha} {(\homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {C_t} \alpha} f \beta} g \gamma)}\gamma\]
-%%     equals
-
-%%     \[ \homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {E_t} {(G \circ F)_0 \alpha}} {(\homrappXtoYatZ {\pendappXtoY {(G\circ F)_1} \alpha} f \beta)} {(G\circ F)_0 \beta}} {(\homrappXtoYatZ {\pendappXtoY {(G\circ F)_1} \beta} g \gamma)} {{F_0} \gamma}\]
-
-%%     Reducing, we need to show
-%%     \[ \homrappXtoYatZ {\pendappXtoY{G_1} {F_0 \alpha}} {(\homrappXtoYatZ {\pendappXtoY {F_1} \alpha} {(\homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {C_t}\alpha} f {\beta}} g {\gamma})} \gamma)} {F_0 \gamma} \]
-
-%%     is equal to
-%%     \[ \homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {E_t} {G_0(F_0\alpha)}}
-%%       {({\homrappXtoYatZ {\pendappXtoY {G_1} {F_0 \alpha}} {(\homrappXtoYatZ {\pendappXtoY {F_1} \alpha} f \beta)} {F_0 \beta}})}
-%%       {G_0(F_0\beta)}}
-%%        {({\homrappXtoYatZ {\pendappXtoY {G_1} {F_0 \beta}} {(\homrappXtoYatZ {\pendappXtoY {F_1} \beta} g \gamma)} {F_0 \gamma}})}
-%%        {{F_0} \gamma}\]
-
-%%     Going from left to right:
-%%     \begin{align*}
-%%       &\homrappXtoYatZ {\pendappXtoY{G_1} {F_0 \alpha}} {(\homrappXtoYatZ {\pendappXtoY {F_1} \alpha} {(\homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {C_t}\alpha} f {\beta}} g {\gamma})} \gamma)} {F_0 \gamma}\\
-%%       &=
-%%       \homrappXtoYatZ
-%%       {\pendappXtoY{G_1} {F_0 \alpha}}
-%%       {(({\homrappXtoYatZ {\homrappXtoYatZ{\pendappXtoY{(\pendlambdaXdotY \alpha {\homrlambdaXatYdotZ f \beta {\homrlambdaXatYdotZ g \gamma {\homrappXtoYatZ {\pendappXtoY {F_1} \alpha} {(\homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {C_t}\alpha} f {\beta}} g {\gamma})} \gamma}}})}\alpha} f \beta} g \gamma}))}
-%%       {F_0 \gamma}\\
-%%       &=
-%%       \homrappXtoYatZ
-%%       {\pendappXtoY{G_1} {F_0 \alpha}}
-%%       {(({\homrappXtoYatZ {\homrappXtoYatZ{\pendappXtoY{(\pendlambdaXdotY \alpha {\homrlambdaXatYdotZ f \beta {\homrlambdaXatYdotZ g \gamma {\homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {D_t} {(F_0 \alpha)}} {(\homrappXtoYatZ {\pendappXtoY {F_1} \alpha} f \beta)} {(F_0 \beta)}} {(\homrappXtoYatZ {\pendappXtoY {F_1} \beta} g \gamma)} {{F_0} \gamma}}}})}\alpha} f \beta} g \gamma}))}
-%%       {F_0 \gamma}\tag{*}\\
-%%       &=
-%%       \homrappXtoYatZ
-%%       {\pendappXtoY{G_1} {F_0 \alpha}}
-%%       {\homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {D_t} {(F_0 \alpha)}} {(\homrappXtoYatZ {\pendappXtoY {F_1} \alpha} f \beta)} {(F_0 \beta)}} {(\homrappXtoYatZ {\pendappXtoY {F_1} \beta} g \gamma)} {{F_0} \gamma}}
-%%       {F_0 \gamma}\\
-%%       &=
-%%       {(({\homrappXtoYatZ {\homrappXtoYatZ{\pendappXtoY{(\pendlambdaXdotY {\alpha'} {\homrlambdaXatYdotZ {f'} {\beta'} {\homrlambdaXatYdotZ {g'} {\gamma'} {{\homrappXtoYatZ {\pendappXtoY {G_1} {\alpha'}} {(\homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {D_t}{\alpha'}} {f'} {\beta'}} {g'} {\gamma'})} {\gamma'}}}}})}{F_0\alpha}} {(\homrappXtoYatZ {\pendappXtoY {F_1} \alpha} f \beta)} {F_0\beta}} {(\homrappXtoYatZ {\pendappXtoY {F_1} \beta} g \gamma)} {F_0\gamma}}))}\\
-%%       &=
-%%       {(({\homrappXtoYatZ {\homrappXtoYatZ{\pendappXtoY{(\pendlambdaXdotY {\alpha'} {\homrlambdaXatYdotZ {f'} {\beta'} {\homrlambdaXatYdotZ {g'} {\gamma'} {{\homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {E_t} {(G_0 {\alpha'})}} {(\homrappXtoYatZ {\pendappXtoY {G_1} {\alpha'}} {f'} {\beta'})} {(G_0 \beta')}} {(\homrappXtoYatZ {\pendappXtoY {G_1} {\beta'}} {g'} {\gamma'})} {{G_0} \gamma'}}}}})}{F_0\alpha}} {(\homrappXtoYatZ {\pendappXtoY {F_1} \alpha} f \beta)} {F_0\beta}} {(\homrappXtoYatZ {\pendappXtoY {F_1} \beta} g \gamma)} {F_0\gamma}}))}\\
-%%       &= \homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {E_t} {G_0(F_0\alpha)}}
-%%       {({\homrappXtoYatZ {\pendappXtoY {G_1} {F_0 \alpha}} {(\homrappXtoYatZ {\pendappXtoY {F_1} \alpha} f \beta)} {F_0 \beta}})}
-%%       {G_0(F_0\beta)}}
-%%        {({\homrappXtoYatZ {\pendappXtoY {G_1} {F_0 \beta}} {(\homrappXtoYatZ {\pendappXtoY {F_1} \beta} g \gamma)} {F_0 \gamma}})}
-%%        {{F_0} \gamma}
-%%     \end{align*}
-%%   \end{enumerate}
-%% \end{definition}
-
-%% \begin{definition}[Internal Profunctor]
-%%   A profunctor $R$ between internal categories $\cat C$ and $\cat D$
-%%   consists of
-%%   \begin{enumerate}
-%%   \item A span $R_s : \harr {C_0} {D_0}$.
-%%   \item A ternary ``mixed composition'' action:
-%%     $R_t :\pendallXdotY \alpha {\homrallXYtoZ {\alpha'} {\harrapp {C_1} {\alpha}{\alpha'}} {\homrallXYtoZ {\beta'} {\harrapp {R_s} {\alpha'}{\beta'}} {\homrallXYtoZ {\beta} {\harrapp {D_1} {\beta'} {\beta}} {\harrapp {R_s} {\alpha}{\beta}}}}}$
-
-%%   \item Satisfying unitality (under a $\pendlambdaXdotY \alpha {\homrlambdaXatYdotZ f \beta {}}$):
-%%     \[
-%%     f = \homrappXtoYatZ {\homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {R_t} \alpha} {\pendappXtoY {C_r} \alpha} \alpha}f \beta} {\pendappXtoY {D_r} \beta} \beta
-%%     \]
-
-%%   \item And associativity (both with a large implicit
-%%     $\pendlambdaXdotY \alpha {\homrlambdaXatYdotZ g \beta \homrlambdaXatYdotZ {g'} {\beta'}
-%%     \homrlambdaXatYdotZ {r} {\beta''}\homrlambdaXatYdotZ {f'} {\alpha''}\homrlambdaXatYdotZ {f} {\alpha'}}$)
-
-%%     % f : a -> a'
-%%     % f' : a' -> a''
-%%     % r : a'' -> b''
-%%     % g' : b'' -> b'
-%%     % g : b' -> b
-%%     %%
-%%     %
-%%     % f o (f' o r o g') o g
-%%     %% and
-%%     % (f o f') o r o (g' o g)
-
-%%     \[ \homrappXtoYatZ {\homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {R_t} {\alpha}} {f} {\alpha'}}{({\homrappXtoYatZ {\homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {R_t} {\alpha'}} {f'} {\alpha''}}r {\beta''}} {g'} {\beta'}})} {\beta'}} {g} {\beta}\]
-%%     is equal to
-%%     \[ \homrappXtoYatZ
-%%        {\homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {R_t} {\alpha}} {(\homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {C_t} {\alpha}} {f} {\alpha'}} {f'} {\alpha''})} {\alpha''}} r {\beta''}}
-%%        {(\homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {D_t} {\beta''}} {g'} {\beta'}} g {\beta})} {\beta}\]
-%%   \end{enumerate}
-%% \end{definition}
-
-%% We can immediately define the Hom profunctor for any category $C$ as
-%% $(\text{Hom}_C R)_s = C_1$ with actions given by $C_t$.
-
-%% Next, we show that Profunctors can be restricted by functors.
-
-%% \begin{definition}[Restriction of Profunctors]
-%%   If $C, D, C', D' : \text{Cat}$, and $F : \text{Functor} C C', G : \text{Functor} D D'$ and
-%%   $R : \text{Prof} C' D'$, then we can define a profunctor $R[F;G] : \text{Prof} C D$ by
-%%   \begin{enumerate}
-%%   \item $R[F;G]_s = \lambda (\alpha:C,\beta:D). R_s (F_0 \alpha) (G_0 \beta)$
-%%   \item with composition action given by
-
-%%     \[ R[F;G]_t =
-%%     \pendlambdaXdotY \alpha {\homrlambdaXatYdotZ f {\alpha'} {\homrlambdaXatYdotZ r {\beta'} {\homrlambdaXatYdotZ g {\beta} {\homrappXtoYatZ {\homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {R_t} {F_0 \alpha}} {(\homrappXtoYatZ {\pendappXtoY {F_1} {\alpha}} f {\alpha'})} {F_0 \alpha'}} r {G_0 \beta'}} {(\homrappXtoYatZ {\pendappXtoY {G_1} {\beta'}} g \beta)} {G_0 \beta}}}} }
-%%  \]
-%%   %% \item $\cat C'$ composition:
-%%   %%   \[ \lambda \alpha. \lambda^\triangleleft (\beta, (f : C_1 \alpha \alpha',\alpha',\alpha', r : R_s \alpha' \beta)). R_l(F_1 f, r) \]
-%%   %% \item $\cat D'$ composition:
-%%   %%   \[ \lambda \alpha. \lambda^\triangleleft ('\beta, (r : R_s \alpha \alpha',\alpha', \beta',h : D_1 \beta' \beta)). R_r(r, G_1 h) \]
-%%   \end{enumerate}
-%%   Identity and associativity properties should follow. TODO
-%% \end{definition}
-
-%% \begin{definition}[Homomorphism]
-%%   Given $R : \text{Prof} C D$ and $S : \text{Prof} C D$, a
-%%   homomorphism $m$ from $R$ to $S$ consists of
-%%   \begin{enumerate}
-%%   \item $m_f : \pendallXdotY \alpha {\homrallXYtoZ \beta {\harrapp R \alpha \beta}{\harrapp S \alpha \beta}}$
-%%     % m_f (R_t f r g) = S_t f (m_f r) g
-%%   \item Satisfying naturality (under a $\ldots$)
-
-%%     \[
-%%     \homrappXtoYatZ {\pendappXtoY{m_f} \alpha} {(\homrappXtoYatZ {\homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {R_t} \alpha} f {\alpha'}} r {\beta'}} g \beta)} \beta
-%%     =
-%%     \homrappXtoYatZ {\homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {S_t} \alpha} {f} {\alpha'}} {(\homrappXtoYatZ {\pendappXtoY {m_f}{\alpha'}} r {\beta'})} {\beta'}} g \beta
-%%     \]
-%%   \end{enumerate}
-%% \end{definition}
-
-%% \begin{definition}
-%%   An isomorphism between profunctors $R$ and $S$ is a mutually inverse
-%%   pair of homomorphisms.
-%% \end{definition}
-
-%% Additionally, we can define natural transformations between functors,
-%% but more generally, we can define a natural element of any
-%% endo-profunctor, where natural transformations from $F$ to $G$ are the
-%% same as natural elements of $\text{Hom}(F-,G=)$.
-
-%% \begin{definition}[Natural Element]
-%%   A natural element $\phi$ of a profunctor $R : \text{Prof}\,C C$ is a term
-%%   \[ \phi : \pendallXdotY \alpha {\harrapp R \alpha \alpha} \]
-%%   that is natural in that
-%%   \[
-%%   \homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {C_t} \alpha} {\pendappXtoY \phi \alpha} \alpha} f {\alpha'}
-%%   = \homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {C_t } \alpha} f {\alpha'}} {\pendappXtoY \phi {\alpha'}}{\alpha'}
-%%   \]
-%% \end{definition}
-
-%% \begin{definition}[Natural Transformation]
-%%   A natural transformation from $F : \text{Functor} \,C\, D$ and $G :
-%%   \text{Functor}\,C\,D$ is a natural element of $\text{Hom}(D)[F;G]$
-%% \end{definition}
-
-%% Composition of natural transformations is then given by $C_t$.
-
-%% Natural elements are closely related to homomorphisms: a natural
-%% element equivalent to a homomorphism out of a Hom profunctor. This is
-%% our first instance of a ``Yoneda'' principle that the behavior of some
-%% structure that varies naturally with the morphisms of a category is
-%% fully determined by its action on the identity morphisms.
-
-%% \begin{theorem}
-%%   Natural elements of $R : \text{Prof}\, C C$ are equivalent to
-%%   homomorphisms from $\text{Hom}_C$ to $R$.
-%% \end{theorem}
-%% \begin{enumerate}
-%% \item Given a natural element $\phi$ of $R$, we can define a
-%%   homomorphism by
-%%   \[ \pendlambdaXdotY \alpha {\homrlambdaXatYdotZ f {\alpha'} \homrappXtoYatZ {\homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {R_t} \alpha} {\pendappXtoY {C_r} \alpha} \alpha}{\pendappXtoY \phi \alpha} {\alpha}} f {\alpha'}} \]
-%% \item Given a homomorphism $m$ from $\text{Hom}(C)$ to $R$, define the
-%%   element as
-%%   \[ \pendlambdaXdotY \alpha {\homrappXtoYatZ {\pendappXtoY {m_f} \alpha} {\pendappXtoY {C_r}\alpha} \alpha}  \]
-
-%%   We should be able to show this is an isomorphism in the logic TODO!
-%% \end{enumerate}
-
-%% %% This last definition is probably unfamiliar to the reader that is not
-%% %% experienced with profunctors.
-%% %% %
-%% %% However, we can show that the traditional definition is equivalent to
-%% %% a natural transformation $\alpha : \cat C_0 \pipe \cdot \vdash
-%% %% t_{\alpha} : F(\alpha) \to_{\cat D} G(\alpha')$.
-
-%% %% \begin{definition}[Traditional Natural Transformation]
-%% %%   A traditional natural transformation from $\alpha : \cat C_0 \vdash
-%% %%   F(\alpha) : \cat D_0$ to $\alpha : \cat \vdash G(\alpha') : \cat
-%% %%   D_0$ consists of
-%% %%   \begin{enumerate}
-%% %%   \item A term $\cdot \vdash t_{\alpha} : F(\alpha) \to_{\cat D} G(\alpha)$
-%% %%   \item Such that
-%% %%     \[ \phi : \alpha \to_{\cat C} \alpha' \vdash t_{\alpha'} \circ F_1(\phi) \equiv G_1(\phi) \circ t_{\alpha} : F(\alpha) \to_{\cat D} G(\alpha')
-%% %%     \]
-%% %%   \end{enumerate}
-%% %% \end{definition}
-
-%% %% \begin{theorem}[Equivalent Presentations of Natural Transformations]
-%% %%   There is a bijection between traditional natural transformations $t
-%% %%   : F \to G$ and profunctor homomorphisms $h : \alpha \to_{\cat C}
-%% %%   \alpha' \vdash F(\alpha) \to_{\cat D} G(\alpha')$ given by
-%% %%   \begin{enumerate}
-%% %%   \item From a natural transformation $t$ we define a homomorphism
-%% %%     $h(t)$ by composing on the $F$ side. The choice of this side is
-%% %%     arbitrary and we could just as easily have used $G$ since they are
-%% %%     equal by naturality.
-%% %%     \[ \phi : \alpha \to_{\cat C} \alpha' \vdash t_{\alpha'} \circ F_1(\phi) : F(\alpha) \to_{\cat D} G_1(\alpha') \]
-%% %%   \item From a homomorphism $\phi.h$, we define a natural transformation $t(h)$ as
-%% %%     \[ \alpha \pipe \cdot \vdash h[{\id_{\alpha}}/\phi] \]
-%% %%   \end{enumerate}
-%% %% \end{theorem}
-%% %% \begin{proof}
-%% %%   First we show that the maps $h(\cdot),t(\cdot)$ actually produce
-%% %%   homomorphisms and natural transformations.
-%% %%   \begin{enumerate}
-%% %%   \item On the left, we need to show $h(t)_{\phi} \circ F_1(\psi) =
-%% %%     h(t)_{|phi \circ \psi}$. Expanding the definition, we just need
-%% %%     associativity and functoriality of $F_1$:
-%% %%     \begin{align*}
-%% %%       h(t)_{\phi} \circ F_1(\psi) &\equiv (t_{\alpha'} \circ F_1(\phi)) \circ F_1(\psi)\\
-%% %%       &\equiv t_{\alpha'} \circ F_1(\phi \circ \psi)\\
-%% %%       &\equiv h(t)_{\phi \circ \psi}
-%% %%     \end{align*}
-%% %%     on the right, we use naturality:
-%% %%     \begin{align*}
-%% %%       G_1(\theta) \circ h(t)_{\phi} &\equiv G_1(\theta) \circ (t_{\alpha'} \circ F_1(\phi))\\
-%% %%       &\equiv (t_{\alpha''} \circ F_1(\theta)) \circ F_1(\phi)\\
-%% %%       &\equiv t_{\alpha''} \circ F_1(\theta \circ \phi)\\
-%% %%       &\equiv t_{\theta \circ \phi}
-%% %%     \end{align*}
-%% %%   \item In the other direction, to prove naturality:
-%% %%     \begin{align*}
-%% %%       h[\id_\alpha/\phi] \circ F_1(\psi) &\equiv h[\id_\alpha \circ \psi/\phi]\\
-%% %%       &\equiv h[\psi/\phi]\\
-%% %%       &\equiv h[\psi \circ \id_{\alpha'}/\phi]\\
-%% %%       &\equiv G_1(\psi) \circ h[\id_{\alpha}/\phi]
-%% %%     \end{align*}
-%% %%   \end{enumerate}
-%% %%   Next we seek to prove that the functions are mutually inverse.
-%% %%   \begin{enumerate}
-%% %%   \item Round trip for homomorphisms:
-%% %%     \begin{align*}
-%% %%       h(t(h))_{\phi} &= t(h)_{\alpha} \circ F_1[\phi]\\
-%% %%       &= h[\id_{\alpha}] \circ F_1[\phi]\\
-%% %%       &= h[\id \circ \phi]\\
-%% %%       &= h[\phi]
-%% %%     \end{align*}
-%% %%   \item Round trip for natural transformations uses the fact that
-%% %%     functors preserve identity.
-%% %%     \begin{align*}
-%% %%       t(h(t))_{\alpha} &= h(t)_{\id_{\alpha}}\\
-%% %%       &= t_{\alpha} \circ F_1[\id_{\alpha}]\\
-%% %%       &= t_{\alpha} \circ \id_{F_0\alpha}\\
-%% %%       &= t_{\alpha}
-%% %%     \end{align*}
-%% %%   \end{enumerate}
-%% %% \end{proof}
-
-%% \subsection{Internal Category Translation}
-
-%% \subsection{Universal Properties}
-
-%% Since we have profunctors as a primitive notion, the most natural
-%% presentation of universal properties is given by the notion of a
-%% functor representing a profunctor.
-
-%% First, we have the ``algebraic'' notion of a functor representing a
-%% profunctor, i.e., when a profunctor $R : C\pto D$ is equivalent to
-%% $Hom_D(-,G=)$.
-%% \begin{definition}[Right Representability]
-%%   Let $C, D : \text{Cat}$ and $Q : \textrm{Prof}\,C D$ be a profunctor
-%%   between them. Then a \emph{right representation} of $Q$ consists of
-%%   \begin{enumerate}
-%%   \item A functor $G : \algVarr\, D C$
-%%   \item An isomorphism $I,I^-$ between $Q$ and $\text{Hom}(C)[Id;G]$
-%%   \end{enumerate}
-%% \end{definition}
-
-%% This is a nice definition for applications, but a lot of the structure
-%% (homomorphism, functoriality of G) is determined by other components
-%% of the structure. This can be simplified to the notion of a
-%% ``universal morphism''.
-
-%% \begin{definition}[Universal Morphism]
-%%   A profunctor $Q : \textrm{Prof}\, C D$ is right representable when
-%%   there exists
-%%   \begin{enumerate}
-%%   \item A function on objects $G_0 : \varr {D_0} {C_0}$
-%%   \item A ``universal morphism'', i.e., a term $\epsilon :
-%%     \pendallXdotY \beta {\harrapp Q {G_0 \beta} \beta}$ (not
-%%     presupposed to be natural).
-%%   \item An ``introduction rule'', i.e., a term $I : \pendallXdotY
-%%     \alpha {\homrallXYtoZ \beta {\harrapp {Q_s}\alpha \beta} {\harrapp
-%%         {C_1} \alpha {G_0\beta}}}$ (not presupposed to satisfy the
-%%     homomorphism condition).
-%%   \item Such that composition with $\epsilon$ is a left and right
-%%     inverse to $I$. In more detail, we have two equalities.
-
-%%     \begin{enumerate}
-%%     \item First, the $\beta$ rule states that composing the universal
-%%       morphism with the introduction rule applied to a morphism $q : \harrapp {Q} \alpha {\beta}$ is
-%%       equal to the original $q$:
-
-%%       \[ q = \homrappXtoYatZ {\homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {Q_t} \alpha} {(\homrappXtoYatZ {\pendappXtoY I \alpha} {q} \beta)} {G_0 \beta}} {\pendappXtoY \epsilon \beta} {\beta}} {\pendappXtoY {D_r} \beta} {\beta} \]
-
-%%     \item Second, the $\eta$ rule says that every morphism $f :
-%%       \harrapp {C_1} \alpha {G_0 \beta}$ can be reconstructed using
-%%       $I$ after postcomposing with $\epsilon$:
-
-%%       \[ f = \homrappXtoYatZ {\pendappXtoY I \alpha} {({\homrappXtoYatZ {\homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {Q_t} \alpha} {f} {G_0 \beta}} {\pendappXtoY \epsilon \beta} {\beta}} {\pendappXtoY {D_r} \beta} {\beta}  })} {\beta}\]
-
-%%       As a special case, if $f = \pendappXtoY {C_r} {G_0 \beta}$, this shows that $\homrappXtoYatZ {\pendappXtoY I {G_0 \beta}} {\pendappXtoY \epsilon \beta} \beta = \pendappXtoY {C_r} {G_0 \beta}$
-%%     \end{enumerate}
-%%   \end{enumerate}
-%% \end{definition}
-
-%% These two notions turn out to be equivalent because the ``extra''
-%% information in the hom isomorphism is in fact fully determined by the
-%% data of a universal morphism: the functoriality of $G$ and the
-%% naturality of $I,I_{inv}$.
-%% %
-%% The proof can be broken up into 3 conceptual steps:
-%% \begin{enumerate}
-%% \item First,, the functoriality of $G$ is derivable from $I,\epsilon$.
-%% \item Next, $I_{inv}$ is a homomorphism automatically since composition
-%%   with a morphism is automatically natural, where $I_{inv}$ is defined as
-%%   \[\pendlambdaXdotY \alpha {\homrlambdaXatYdotZ f {\beta} \homrappXtoYatZ {\homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {Q_t} \alpha} f \beta}{\pendappXtoY \epsilon \beta} {\beta}} {\pendappXtoY {D_r }\beta} {\beta}} \]
-
-%%   \item Finally, $I$ is a homomorphism because the inverse of a
-%%     natural transformation is always natural.
-%% \end{enumerate}
-
-%% %% Part of this equivalence is the \emph{Yoneda Lemma}, which we state
-%% %% now in its internal, parameterized form.
-%% %% The usual non-parameterized Yoneda lemma can be formed if we have a
-%% %% unit sort.
-
-%% \begin{lemma}[Universal Constructions are functorial]
-%%   For any universal construction $(G_0,\epsilon,I)$ for $Q :
-%%   \text{Prof}\,C\,D$, we can extend $G_0$ to a functor by defining
-
-%%   \[ G_1 = \pendlambdaXdotY \beta {\homrlambdaXatYdotZ f {\beta'}
-%%     {\homrappXtoYatZ {\pendappXtoY I {G_0\beta}} {(\homrappXtoYatZ {\homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {Q_t} {G_0\beta}} {\pendappXtoY {C_r} {G_0\beta}} {G_0\beta}} {\pendappXtoY \epsilon {\beta}} \beta} {f} {\beta'})} {\beta'}}
-%%   }\]
-
-%%   Furthermore, $I,I^-1$ are homomorphisms with respect to the induced
-%%   profunctor structure.
-%% \end{lemma}
-%% \begin{proof}
-%%   \begin{enumerate}
-%%   \item To prove $G_1$ preserves identity we need to show
-%%     \[ {\homrappXtoYatZ {\pendappXtoY I {G_0\beta}} {(\homrappXtoYatZ {\homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {Q_t} {G_0\beta}} {\pendappXtoY {C_r} {G_0\beta}} {G_0\beta}} {\pendappXtoY \epsilon {\beta}} \beta} {\pendappXtoY {D_r} {\beta}} {\beta})} {\beta}} = \pendappXtoY{C_r} {G_0\beta}\]
-%%     which is precisely the $\eta$ principle.
-
-%%   \item Next, we show $I_{inv}$ is a homomorphism from $\text{Hom}(C)[\text{Id}(C);G]$ to $Q$.
-%%     All of this is under the binders
-%%     \[\lambdatrinary {\alpha : C} {f : \harrapp {C_s} \alpha {\alpha'}}
-%%     {\alpha' : C} {g : \harrapp{C_s} {\alpha'} {G_0 \beta'}}
-%%     {\beta' : D} {h : \harrapp {D_s} {\beta'} {\beta}}
-%%     {}\]
-
-%%     % given
-%%     % f : a -> a' in C
-%%     % g : a' -> G b'
-%%     % h : b' -> b
-
-%%     %% I_{inv} (f ;C (g ;C G_1 h)) = Q_t f (I_{inv} g) h
-
-%%     \begin{align*}
-%%       & \appunary
-%%       {I_{inv}}
-%%       \alpha
-%%       {(\appbinary {C_t} \alpha f {\alpha'} {(
-%%           \appbinary {C_t} {\alpha'} g {G_0\beta'}
-%%           {(\appunary {G_1} {\beta'} h \beta)} {G_0 \beta})} {G_0 \beta})}
-%%       {\beta}\\
-%%       &=
-%%       \apptrinary {Q_t} \alpha
-%%       {(\appbinary {C_t} \alpha f {\alpha'} {(
-%%           \appbinary {C_t} {\alpha'} g {G_0\beta'}
-%%           {(\appunary {G_1} {\beta'} h \beta)} {G_0 \beta})} {G_0 \beta})} \beta
-%%       {\pendappXtoY \epsilon \beta} \beta
-%%       {\pendappXtoY {D_r} \beta} {\beta}\\
-%%       &=
-%%       \apptrinary {Q_t} \alpha
-%%       {(\appbinary {C_t} \alpha f {\alpha'} {(
-%%           \appbinary {C_t} {\alpha'} g {G_0\beta'}
-%%           {(\appunary {G_1} {\beta'} h \beta)} {G_0 \beta})} {G_0 \beta})} \beta
-%%       {\pendappXtoY \epsilon \beta} \beta
-%%       {(\appbinary {D_t} \beta {\pendappXtoY {D_r} \beta} \beta {\pendappXtoY {D_r} \beta} \beta)} {\beta}\\
-%%       &=
-%%       \apptrinary {Q_t} \alpha
-%%       {(\appbinary {C_t} \alpha {(\appbinary {C_t} \alpha f {\alpha'} g {G_0 \beta'})} {G_0 \beta'}
-%%         {(\appunary {G_1} {\beta'} h \beta)} {G_0 \beta})} \beta
-%%       {\pendappXtoY \epsilon \beta} \beta
-%%       {(\appbinary {D_t} \beta {\pendappXtoY {D_r} \beta} \beta {\pendappXtoY {D_r} \beta} \beta)} {\beta}\\
-%%       &=
-%%       \apptrinary {Q_t} \alpha
-%%       {(\appbinary {C_t} \alpha f {\alpha'} g {G_0 \beta'})} {G_0 \beta'}
-%%       {(\apptrinary {Q_t} {G_0 \beta'}
-%%         {(\appunary {G_1} {\beta'} h {\beta})} {G_0 \beta}
-%%         {\pendappXtoY \epsilon \beta} \beta
-%%         {\pendappXtoY {D_r} \beta} \beta)} \beta
-%%       {\pendappXtoY {D_r} \beta} \beta
-%%       \\
-%%       &=
-%%       \apptrinary {Q_t} \alpha
-%%       {(\appbinary {C_t} \alpha f {\alpha'} g {G_0 \beta'})} {G_0 \beta'}
-%%       {(\appunary {I_{inv}} {G_0 \beta'} {(\appunary {G_1} {\beta'} h \beta)} \beta)} \beta
-%%       {\pendappXtoY {D_r} \beta} \beta
-%%       \\
-%%       &=
-%%       \apptrinary {Q_t} \alpha
-%%       {(\appbinary {C_t} \alpha f {\alpha'} g {G_0 \beta'})} {G_0 \beta'}
-%%       {(\appunary {I_{inv}} {G_0 \beta'} {(\appunary {I} {G_0\beta'} {(\apptrinary {Q_t} {G_0 \beta'} {\pendappXtoY {C_r} {G_0 \beta'}} {G_0 \beta'} {\pendappXtoY \epsilon {\beta'}} {\beta'} h \beta)} \beta)} \beta)} \beta
-%%       {\pendappXtoY {D_r} \beta} \beta
-%%       \\
-%%       &=
-%%       \apptrinary {Q_t} \alpha
-%%       {(\appbinary {C_t} \alpha f {\alpha'} g {G_0 \beta'})} {G_0 \beta'}
-%%       {(\apptrinary {Q_t} {G_0 \beta'} {\pendappXtoY {C_r} {G_0 \beta'}} {G_0 \beta'} {\pendappXtoY \epsilon {\beta'}} {\beta'} h \beta)} \beta
-%%       {\pendappXtoY {D_r} \beta} \beta
-%%       \\
-%%       &=
-%%       \apptrinary {Q_t} \alpha
-%%       {(\appbinary {C_t} \alpha {(\appbinary {C_t} \alpha f {\alpha'} g {G_0 \beta'})} {G_0 \beta'}
-%%         {\pendappXtoY {C_r} {G_0 \beta'}} {G_0 \beta'})} {G_0 \beta'}
-%%       {\pendappXtoY \epsilon {\beta'}} {\beta'}
-%%       {(\appbinary {D_t} {\beta'} h \beta {\pendappXtoY {D_r} \beta} \beta)} \beta\\
-%%       &=
-%%       \apptrinary {Q_t} \alpha
-%%       {(\appbinary {C_t} \alpha f {\alpha'} g {G_0 \beta'})} {G_0 \beta'}
-%%       {\pendappXtoY \epsilon {\beta'}} {\beta'}
-%%       {(\appbinary {D_t} {\beta'} {\pendappXtoY {D_r} {\beta'}} {\beta'} h \beta )} \beta\\
-%%       &=
-%%       \apptrinary {Q_t} \alpha
-%%       f {\alpha'}
-%%       {(\apptrinary {Q_t} {\alpha'} g {G_0 \beta'} {\pendappXtoY \epsilon {\beta'}} {\beta'}{\pendappXtoY {D_r} {\beta'}} {\beta'})} {\beta'}
-%%       h \beta\\
-%%       &= \homrappXtoYatZ {\homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {Q_t} \alpha} f {\alpha'}} {(\homrappXtoYatZ {\pendappXtoY {I_{inv}} {\alpha'}} {g} {\beta'})} {\beta'}} {h} {\beta}\\
-%%     \end{align*}
-
-%%   \item To prove $G_1$ preserves composition given $f : \harrapp {D_1} {\beta} {\beta'}$, $f' : \harrapp {D_1} {\beta'} {\beta''}$:
-%%     \begin{align*}
-%%       & \appunary {G_1} {\beta} {\appbinary {D_t} {\beta} f {\beta'} {f'} {\beta''}} {\beta'}\\
-%%       &= \appunary {I} {G_0 \beta} {(\apptrinary {Q_t} {G_0 \beta} {\pendappXtoY {C_r} {G_0 \beta}} {G_0 \beta} {\pendappXtoY \epsilon {\beta}} {\beta} {({\appbinary {D_t} {\beta} f {\beta'} {f'} {\beta''}})} {\beta''})} {\beta''}\\
-%%       &= \appunary {I} {G_0 \beta} {(\apptrinary {Q_t} {G_0 \beta} {(\appbinary {C_t} {G_0 \beta} {\pendappXtoY {C_r} {G_0 \beta}} {G_0 \beta} {\pendappXtoY {C_r} {G_0 \beta}} {G_0 \beta})} {G_0 \beta} {\pendappXtoY \epsilon {\beta}} {\beta} {({\appbinary {D_t} {\beta} f {\beta'} {f'} {\beta''}})} {\beta''})} {\beta''}\\
-%%       &=
-%%       \appunary {I} {G_0 \beta}
-%%       {(\apptrinary {Q_t} {G_0 \beta}
-%%         {\pendappXtoY {C_r} {G_0 \beta}} {G_0 \beta}
-%%         {(\apptrinary {Q_t} {G_0 \beta}
-%%           {\pendappXtoY {C_r} {G_0 \beta}} {G_0 \beta}
-%%           {\pendappXtoY \epsilon {\beta}} {\beta}
-%%           f {\beta'})} {\beta'}
-%%         {f'} {\beta''}
-%%         )}
-%%       {\beta''}\\
-%%       &=
-%%       \appbinary {C_t} {G_0 \beta}
-%%       {\pendappXtoY {C_r} {G_0 \beta}} {G_0 \beta}
-%%       {(\appbinary {C_t} {G_0 \beta}
-%%         {(\appunary {I} {G_0 \beta}
-%%         {(\apptrinary {Q_t} {G_0 \beta}
-%%           {\pendappXtoY {C_r} {G_0 \beta}} {G_0 \beta}
-%%           {\pendappXtoY \epsilon {\beta}} {\beta}
-%%           f {\beta'})} {\beta'})} {G_0 \beta'}
-%%         {(\appunary {G_1} {\beta'} {f'} {\beta''})} {G_0 \beta''})}
-%%       {G_0 \beta''}\\
-%%       &=
-%%       \appbinary {C_t} {G_0 \beta}
-%%         {(\appbinary {C_t} {G_0 \beta}
-%%           {\pendappXtoY {C_r} {G_0 \beta}} {G_0 \beta}
-%%           {(\appbinary {C_t} {G_0 \beta}
-%%             {(\appunary {I} {G_0 \beta} {\pendappXtoY \epsilon {\beta}} {\beta})} {G_0 \beta}
-%%             {(\appunary {G_1} \beta f {\beta'})} {G_0 \beta'})} {G_0 \beta'})} {G_0 \beta'}
-%%         {(\appunary {G_1} {\beta'} {f'} {\beta''})} {G_0 \beta''}\\
-%%       &=
-%%       \appbinary {C_t} {G_0 \beta}
-%%         {(\appbinary {C_t} {G_0 \beta}
-%%             {(\appunary {I} {G_0 \beta} {\pendappXtoY \epsilon {\beta}} {\beta})} {G_0 \beta}
-%%             {(\appunary {G_1} \beta f {\beta'})} {G_0 \beta'})} {G_0 \beta'}
-%%         {(\appunary {G_1} {\beta'} {f'} {\beta''})} {G_0 \beta''}\\
-%%       &=
-%%       \appbinary {C_t} {G_0 \beta}
-%%         {(\appbinary {C_t} {G_0 \beta}
-%%             {\pendappXtoY {C_r} {G_0 \beta}} {G_0 \beta}
-%%             {(\appunary {G_1} \beta f {\beta'})} {G_0 \beta'})} {G_0 \beta'}
-%%         {(\appunary {G_1} {\beta'} {f'} {\beta''})} {G_0 \beta''}\\
-%%       &=
-%%       \appbinary {C_t} {G_0 \beta}
-%%       {(\appunary {G_1} \beta f {\beta'})} {G_0 \beta'}
-%%         {(\appunary {G_1} {\beta'} {f'} {\beta''})} {G_0 \beta''}\\
-%%       %% &\homrappXtoYatZ {\pendappXtoY {G_1} \beta} {(\homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {D_t} {\beta}} {f} {\beta'}} {f'} {\beta''})} {\beta''}\\
-%%       %% &={\homrappXtoYatZ {\pendappXtoY I {G_0\beta}} {(\homrappXtoYatZ {\homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {Q_t} {G_0\beta}} {\pendappXtoY {C_r} {G_0\beta}} {G_0\beta}} {\pendappXtoY \epsilon {\beta}} \beta} {(\homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {D_t} {\beta}} {f} {\beta'}} {f'} {\beta''})} {\beta''})} {\beta''}}\\
-%%       %% &={\homrappXtoYatZ {\pendappXtoY I {G_0\beta}} {(\homrappXtoYatZ {\homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {Q_t} {G_0\beta}} {({\pendappXtoY {C_r} {G_0\beta}})} {G_0\beta}} {\pendappXtoY \epsilon {\beta}} \beta} {(\homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {D_t} {\beta}} {f} {\beta'}} {f'} {\beta''})} {\beta''})} {\beta''}}\\
-%%     \end{align*}
-%%     %% \[
-%%     %%   {\homrappXtoYatZ {\pendappXtoY I {G_0\beta}} {(\homrappXtoYatZ {\homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {Q_t} {G_0\beta}} {\pendappXtoY {C_r} {G_0\beta}} {G_0\beta}} {\pendappXtoY \epsilon {\beta}} \beta} {(\homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {D_t} {\beta}} {f} {\beta'}} {f'} {\beta''})} {\beta''})} {\beta''}}
-%%     %% \]
-%%   \end{enumerate}
-%% \end{proof}
-
-%% \begin{lemma}[Inverse of homomorphism is a homomorphism]
-%%   Let $\phi$ be a homomorphism from $R$ to $S : \text{Prof} C D$ and
-%%   let $\psi : \pendallXdotY \alpha {\homrallXYtoZ \beta {\harrapp S
-%%       \alpha \beta}{\harrapp R \alpha \beta}}$. If $\psi$ is a
-%%   two-sided inverse to $\phi_f$ then $\psi$ is a homomorphism.
-%% \end{lemma}
-%% \begin{proof}
-%%   We need to show the homomorphism condition for $\psi$:
-%%   \[ \homrappXtoYatZ {\pendappXtoY \psi \alpha} {(\homrappXtoYatZ {\homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {S_t} \alpha} f {\alpha'}} s {\beta'}} g {\beta})} {\beta}
-%%     = \homrappXtoYatZ {\homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {R_t} \alpha} f {\alpha'}} {(\homrappXtoYatZ {\pendappXtoY \psi {\alpha'}} s {\beta'})} {\beta'}} g \beta
-%%     \]
-
-%%     \begin{align*}
-%%       \homrappXtoYatZ {\pendappXtoY \psi \alpha} {(\homrappXtoYatZ {\homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {S_t} \alpha} f {\alpha'}} s {\beta'}} g {\beta})} {\beta}
-%%       &= \homrappXtoYatZ {\pendappXtoY \psi \alpha} {(\homrappXtoYatZ {\homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {S_t} \alpha} f {\alpha'}} {(\homrappXtoYatZ {\pendappXtoY {\phi_f} {\alpha'}} {(\homrappXtoYatZ {\pendappXtoY \psi {\alpha'}} s {\beta'})} {\beta'})} {\beta'}} g {\beta})} {\beta}\\
-%%       &= \homrappXtoYatZ {\pendappXtoY \psi \alpha}
-%%       {(\homrappXtoYatZ {\pendappXtoY {\phi_f} \alpha} {(\homrappXtoYatZ {\homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {R_t} \alpha} f {\alpha'}} {(\homrappXtoYatZ {\pendappXtoY \psi {\alpha'}} s {\beta'})} {\beta'}} g \beta)} \beta)} {\beta}\\
-%%       & = \homrappXtoYatZ {\homrappXtoYatZ {\homrappXtoYatZ {\pendappXtoY {R_t} \alpha} f {\alpha'}} {(\homrappXtoYatZ {\pendappXtoY \psi {\alpha'}} s {\beta'})} {\beta'}} g \beta
-%%     \end{align*}
-%% \end{proof}
-
-%% %% \subsection{Self-Interpretation: Hom of Internal Profunctors}
-
-%% %% We can define a version of the hom $\homrallXYtoZ \alpha P Q$ on
-%% %% \emph{internal} profunctors using a combination of this connective,
-%% %% and an \emph{equalizer} to enforce a homomorphism condition.
-
-%% %% \begin{definition}[Internal Covariant Hom of Profunctors]
-%% %%   Let $P : \textrm{Prof}\,D\,E$ and $Q : \textrm{Prof}\,C\,E$. Define
-%% %%   the covariant Hom $P \triangleright Q : \textrm{Prof}\,C\,D$ by
-%% %%   \[ (P \triangleright Q)_1 = \lambda (\alpha;\beta).
-\equalizeVofWbyXeqYatZ \phi {\homrallXYtoZ \gamma {\harrapp {P_1} \alpha \gamma} {\harrapp {Q_1} \beta \gamma}} {\psi_{dom}} {\psi_{cod}} {\homrallXYtoZ {\gamma'} {\harrapp {P_1 }\beta {\gamma'}} {\homrallXYtoZ {\gamma} {\harrapp {E_1} {\gamma'} \gamma} {\harrapp {Q_1} \alpha \gamma}}}\]
-%% %%   where $\psi_{dom}$ and $\chi_{cod}$ are  as
-%% %%   \[ \psi_{dom} =\homrlambdaXatYdotZ p {\gamma'} {\homrlambdaXatYdotZ e \gamma
-%% %%     {\homrappXtoYatZ {\phi} {(\apptrinary {P_t} \beta {\pendappXtoY {D_r} \beta} \beta p {\gamma'} e \gamma)} \gamma}}
-%% %%   \]
-%% %%   \[
-%% %%   \psi_{cod} = \homrlambdaXatYdotZ p {\gamma'} {\homrlambdaXatYdotZ e \gamma
-%% %%     {\apptrinary {Q_t} \alpha {\pendappXtoY {C_r} \alpha} \alpha {(\homrappXtoYatZ \phi {p} {\gamma'})} {\gamma'} e \gamma}}
-%% %%   \]
-%% %% \end{definition}
-
-%% %% We can easily define a contravariant hom $P \triangleleft Q$ by a dual
-%% %% definition (TODO).
-
-%% %% We can also define a left Kan extensions using the contravariant
-%% %% cotensors.  Normally the left Kan extension is defined using tensors,
-%% %% but the definition using profunctors doesn't look right to me. Let's
-%% %% work backwards from the definition of colimit to see why.
-
-%% %% Let $\alpha : \cat C \vdash A : \cat D$ be a functor we want the
-%% %% colimit of. The UP of a colimit is that maps out of it are equivalent
-%% %% to cocones from $A$. Semi-formally this looks like
-%% %% \[ (\colim_{\alpha} A \to_{\cat D} \beta) \cong \forall \alpha:\cat C.~ A \to_{\cat D} \beta \]
-%% %% This looks a lot like a cotensor with a trivial domain:
-%% %% \[ (\colim_{\alpha} A \to_{\cat D} \beta) \cong \forall \alpha. (() \to_{1} \gamma) \triangleright (A \to_{\cat D} \beta) \]
-%% %% Suggesting that we define the left Kan extension profunctor as a
-%% %% contravariant cotensor as follows.
-%% %% \begin{definition}
-%% %%   Let $\cat C,\cat C',\cat D$ be cats and let $\alpha:\cat C \vdash A
-%% %%   : \cat D$ and $\alpha : \cat C \vdash J : \cat C'$ be functors.
-%% %%   Then the left Kan extension profunctor $\gamma^o:\cat C'; \beta:\cat
-%% %%   D \pipe (PLan_J A)[\gamma;\beta]$ is defined as
-%% %%   \[
-%% %%   \gamma^o:\cat C'; \beta:\cat D \pipe (PLan_J A)[\gamma;\beta]
-%% %%   := \forall \alpha : \cat C.~ (J \to_{\cat C'} \gamma) \triangleright (A \to_{\cat D} \beta)
-%% %%   \]
-%% %%   Then a left Kan extension of $A$ along $J$ is a functor $\gamma :
-%% %%   \cat C' \vdash Lan_J A : \cat D$ that represents $PLan_J A$ in that
-%% %%   \[ (Lan_J A \to_{\cat D} \beta) \cong (LPan_J A)[\gamma;\beta] \]
-%% %% \end{definition}
-
-%% %% \begin{definition}
-%%   %% Let $\cat C,\cat C',\cat D$ be cats and let $\alpha:\cat C \vdash
-%%   %% : \cat D$ and $\alpha : \cat C \vdash J : \cat C'$ be functors.
-%%   %% Then the left Kan extension profunctor $\beta^o:\cat D; \gamma:\cat
-%%   %% C' \pipe (PLan_J F)[\beta;\gamma]$ is defined as
-%% %%   \[
-%% %%   \beta^o:\cat D; \gamma:\cat C' \pipe (PLan_J A)[\beta;\gamma] :=
-%% %%   \exists \alpha:\cat C. (\beta \to_{\cat D} A[\alpha]) \odot (J[\alpha] \to_{\cat C'} \gamma)
-%% %%   \]
-%% %%   And a left Kan extension is a functor $\gamma :\cat C' \vdash Lan_J
-%% %%   A : \cat D$ that represents $PLan_J A$ in that
-%% %%   \[ (\beta \to_{\cat D} Lan_J A[\alpha]) \cong (PLan_J A)[\beta;\alpha] \]
-%% %% \end{definition}
-
-%% %% \begin{lemma}[Parameterized Yoneda Lemma]
-%% %%   Given categories $\cat C, \cat D$ and a profunctor $\alpha^o:\cat
-%% %%   C;\beta : \cat D \vdash \alpha \to_{Q} \beta$ and a function
-%% %%   $\beta : \cat D \vdash G[\beta] : \cat C$, then there is an
-%% %%   isomorphism between
-%% %%   \begin{enumerate}
-%% %%   \item Terms $\cdot \vdash t : G[\beta] \to_{Q} \beta$
-%% %%   \item left-homomorphisms $\phi : \alpha \to_{\cat C} G[\beta] \vdash
-%% %%     u : \alpha \to_{Q} \beta$ i.e. they satisfy homomorphism on the
-%% %%     left (the other side doesn't make sense if $G$ is not a functor):
-%% %%     $u[\phi \circ \psi] = u[\phi] \circ \psi$
-%% %%   \end{enumerate}
-%% %% \end{lemma}
-%% %% \begin{proof}
-%% %%   First, the constructions
-%% %%   \begin{enumerate}
-%% %%   \item Given $t$, we define $u(t)$ to be
-%% %%     \[\phi : \alpha \to_{\cat C} G[\beta] \vdash t \circ_{Q} \phi : \alpha \to_{Q} \beta\]
-
-%% %%     Which is a left-homomorphism because:
-%% %%     \[ (t \circ_{Q} \phi) \circ_{Q} \psi = t \circ_{Q} (\phi \circ_{\cat C} \psi)\]
-%% %%   \item Given $u$, we define $t(u)$ to be
-%% %%     \[ \cdot \vdash u[\id_{G[\beta]}/\phi] : G[\beta] \to_{Q} \beta \]
-%% %%   \end{enumerate}
-%% %%   Now we show they are inverse
-%% %%   \begin{enumerate}
-%% %%   \item First, the very easy case, $\cdot \vdash t(u(t)) : G[\beta] \to_{Q} \beta$:
-%% %%     \begin{align*}
-%% %%       t \circ_{Q} \id_{G[\beta]} &= t
-%% %%     \end{align*}
-%% %%     by the fact that $Q$ is a profunctor.
-%% %%   \item Next, $\phi : \alpha \to_{\cat C} G[\beta] \vdash u(t(u)) : \alpha \to_{Q} \beta$:
-%% %%     \begin{align*}
-%% %%       u[\id_{G[\beta]}/\phi] \circ_{Q} \phi &= u[\id \circ_{\cat C} \phi/\phi]\\
-%% %%       &= u[\phi/\phi]\\
-%% %%       &= u
-%% %%     \end{align*}
-%% %%     which uses the fact that $u$ is a left-homomorphism.
-%% %%   \end{enumerate}
-%% %% \end{proof}
-
-%% %% \begin{theorem}[Negative Types and Right Representables are Equivalent]
-%% %%   For any profunctor $\alpha^o: \cat C; \beta : \cat D \pipe \alpha
-%% %%   \to_{Q} \beta$, there is a bijection between
-%% %%   \begin{enumerate}
-%% %%   \item Right Represntables $\beta : \cat D \vdash G[\beta] : \cat C$ for $Q$.
-%% %%   \item Negative Types $\beta : \cat D \vdash N[\beta] : \cat C$
-%% %%     representing $Q$.
-%% %%   \end{enumerate}
-%% %% \end{theorem}
-%% %% \begin{proof}
-%% %%   \begin{enumerate}
-%% %%   \item Given a right representable $G[\beta]$, we set $N(G)[\beta] =
-%% %%     G[\beta]$, the introduction rule is the same and the elimination
-%% %%     rule is given as in the Yoneda lemma and the inverse property
-%% %%     follows by the Yoneda isomorphism and the fact that $I_G^{-1}$ is
-%% %%     an inverse for $I_G$.
-%% %%   \item Given a negative type $N[\beta]$ we set $G(N)[\beta] =
-%% %%     N[\beta]$, the introduction rule is the same and the inverse
-%% %%     introduction rule is given as in the Yoneda lemma.  We need to
-%% %%     show that $N[\beta]$ has the structure of a functor and that
-%% %%     $I_{N}$ and $I_{N}^{-1}$ are homomorphisms on the right with
-%% %%     respect to that structure.  Define $N_1$ as follows:
-%% %%     \begin{mathpar}
-%% %%       \inferrule
-%% %%           {\beta_1^o:\cat D;\beta_2:\cat D \vdash \psi \circ \epsilon : N[\beta_1/\beta] \to_{Q} \beta_2\\
-%% %%             \phi : N[\beta_1/\beta] \to_{Q} \beta_2 \vdash I_N[\phi] : N[\beta_1/\beta] \to_{\cat C} N[\beta_2/\beta]
-%% %%           }
-%% %%           {\beta_1^o:\cat D;\beta_2:\cat D\pipe \psi : \beta_1 \to_{\cat D} \beta_2 \vdash I_N[\psi \circ \epsilon_{\beta_1}/\phi] : N[\beta_1/\beta] \to_{\cat C} N[\beta_2/\beta]}
-%% %%     \end{mathpar}
-%% %%     We need to show functoriality.
-%% %%     For identity we have
-%% %%     \begin{align*}
-%% %%       I_N[\id_{\beta} \circ \epsilon_{\beta}/\phi] &= I_N[\epsilon_{\beta}]\\
-%% %%       &= I_N[\epsilon \circ \id_{N[\beta]}]\\
-%% %%       &= \id_{N[\beta]} \tag{$\eta$}
-%% %%     \end{align*}
-%% %%     For composition,
-%% %%     \begin{align*}
-%% %%       N_1[\theta] \circ N_1[\psi]
-%% %%       &= I_N[\theta \circ \epsilon] \circ I_N[\psi \circ \epsilon]\\
-%% %%       &= I_N[(\theta \circ \epsilon) \circ I_N[\psi \circ \epsilon]]\tag{left homomorphism}\\
-%% %%       &= I_N[\theta \circ (\epsilon \circ I_N[\psi \circ \epsilon])]\\
-%% %%       &= I_N[\theta \circ (\psi \circ \epsilon)] \tag{$\beta$}\\
-%% %%       &= I_N[(\theta \circ \psi) \circ \epsilon]\\
-%% %%       &= N_1[\theta \circ \psi]
-%% %%     \end{align*}
-%% %%     Next we show $I_N$ is a right homomorphism, the typing here is
-%% %%     $\phi : \alpha \to_{Q} \beta_1, \theta : \beta_1 \to_{\cat D}
-%% %%     \beta_2$.
-%% %%     \begin{align*}
-%% %%       N_1[\theta] \circ I_N[\phi] &= I_N[\theta \circ \epsilon] \circ I_N[\phi]\\
-%% %%       &= I_N[\theta \circ \epsilon \circ I_N[\phi]]\tag{left homomorphism}\\
-%% %%       &= I_N[\theta \circ \phi] \tag{$\beta$}
-%% %%     \end{align*}
-%% %%     and to show $I_N^{-1}$ is a right homomorphism under $\phi : \alpha \to_{\cat C}N[\beta_1], \theta : \beta_1 \to_{\cat D} \beta_2$
-%% %%     \begin{align*}
-%% %%       I_N^{-1}[N_1[\theta] \circ \phi] &= \epsilon \circ (N_1[\theta] \circ \phi) \tag{def}\\
-%% %%       &= \epsilon \circ (I_N[\theta \circ \epsilon] \circ \phi)\tag{def}\\
-%% %%       &= (\epsilon \circ I_N[\theta \circ \epsilon]) \circ \phi\tag{assoc}\\
-%% %%       &= (\theta \circ \epsilon) \circ \phi\tag{$\beta$}\\
-%% %%       &= \theta \circ (\epsilon \circ \phi)\tag{assoc}\\
-%% %%       &= \theta \circ I_N^{-1}[\phi] \tag{def}
-%% %%     \end{align*}
-%% %%   \end{enumerate}
-
-%% %%   Next we need to show that this is a \emph{bijection}.
-%% %%   \begin{enumerate}
-%% %%   \item Starting with a negative type $N[\beta]$, we recover the
-%% %%     original $\epsilon$ by the Yoneda lemma.
-%% %%   \item Starting with a right representable $G[\beta]$, we recover the
-%% %%     original $I_N^{-1}$ by the Yoneda lemma. We need to show that the
-%% %%     action of $G$ on arrows is the same as the one defined from $G$
-%% %%     viewed as a negative type, which we will call $G_1'$:
-%% %%     \begin{align*}
-%% %%       G_1'[\theta] &= I_G[\theta \circ \epsilon]\\
-%% %%       &= G_1[\theta] \circ I_G[\epsilon]\tag{right homomorphism}\\
-%% %%       &= G_1[\theta] \circ I_G[\epsilon \circ \id]\\
-%% %%       &= G_1[\theta] \circ \id\\
-%% %%       &= G_1[\theta]
-%% %%     \end{align*}
-%% %%   \end{enumerate}
-%% %% \end{proof}
-
-%% %% \subsection{Adjunctions by Universal Morphism}
-
-%% %% As an example of formal category theory in our syntax, let's prove the
-%% %% equivalence between two definitions of an adjunction: one by universal
-%% %% morphism and the other by natural hom isomorphism.
-%% %% %
-%% %% Though the definitions and proof will look like we are doing ordinary
-%% %% category theory, since it is syntactic it immediately applies to all
-%% %% models, including enriched and internal category theory.
-
-%% %% We start with the usual definition of an adjunction using
-%% %% hom-isomorphism and a profunctor.
-%% %% \begin{definition}
-%% %%   Let $C,D$ be categories and $c^o:C_0,d:D_0 \vdash H[c,d]$ a
-%% %%   profunctor between them. A \emph{left adjoint} to $H$ consists of
-%% %%   \begin{enumerate}
-%% %%   \item A functor $c:C \vdash F[c] : D$
-%% %%   \item A natural isomorphism $c,d \pipe h : H[c,d] \vdash
-%% %%     F_{ind}[h]: D_1[F_0[c],d]$. Call the inverse $g. F_{ind}^{-1}[g]$.
-%% %%   \end{enumerate}
-%% %% \end{definition}
-
-%% %% \begin{definition}
-%% %%   Let $c:C_0,d:D_0 \vdash H[c,d]$ be a profunctor.
-%% %%   A \emph{universal $H$-heteromorphism from $C$} consists of
-%% %%   \begin{enumerate}
-%% %%   \item For every $c:C_0$, an $H$-morphism $\eta_c : H(c,F_0[c])$ for
-%% %%     some function $c:C_0 \vdash F_0[c] : D_0$.
-%% %%   \item Such that every $h : H[c,d]$ factors through $\eta_c$ in that
-%% %%     there is a unique $h : H[c,d] \vdash F_{ind}[h] : D[F_0[c],d]$
-%% %%     such that $h = F_{ind}[h] \circ \eta_c$.  Uniqueness means
-%% %%     validity of the following rule:
-%% %%     \begin{mathpar}
-%% %%       \inferrule
-%% %%       {g:D[F_0[c],d] \vdash g \circ \eta = h}
-%% %%       {g:D[F_0[c],d] \vdash g = F_{ind}[h]}
-%% %%     \end{mathpar}
-%% %%     (Note that we do not require $F_{ind}$ to be natural in any sense)
-%% %%   \end{enumerate}
-%% %% \end{definition}
-
-%% %% It is easy to see that a left adjoint gives a universal heteromorphism
-%% %% by a Yoneda-like argument:
-%% %% \begin{construction}
-%% %%   Let $F$ be a left adjoint to $H$. Then we define a left universal
-%% %%   $H$-heteromorphism by
-%% %%   \begin{enumerate}
-%% %%   \item $c \vdash \eta_c = F_{ind}^{-1}[\id[F_0[c]]]: H[c,F_0[c]]$
-%% %%   \item Every $h : H[c,d]$ factorizes through $\eta$ because
-%% %%     \begin{align*}
-%% %%       F_{ind}[h] \circ F_{ind}^{-1}[\id] &= F_{ind}^{-1}[F_{ind}[h] \circ \id] \tag{naturality}\\
-%% %%       &= F_{ind}^{-1}[F_{ind}[h]] \tag{unit}\\
-%% %%       &= h \tag{isomorphism}
-%% %%     \end{align*}
-%% %%     This factorization is unique because if $g \circ F^{-1}[\id] = h$,
-%% %%     then
-%% %%     \begin{align*}
-%% %%       F_{ind}[h] &= F_{ind}[g \circ F^{-1}[\id]] \tag{congruence}\\
-%% %%       &= g \circ F_{ind}[F^{-1}[\id]] \tag{naturality}\\
-%% %%       &= g \circ \id \tag{iso}\\
-%% %%       &= g \tag{unit}
-%% %%     \end{align*}
-%% %%   \end{enumerate}
-%% %% \end{construction}
-
-%% %% But the utility of the universal morphism definition is that the
-%% %% universal morphism extracted from a left adjoint is enough to uniquely
-%% %% reconstruct the rest of the structure, which includes (1)
-%% %% functoriality of $F$ and (2) naturality of $F_{ind}$.
-%% %% \begin{construction}
-%% %%   Let $(F_0,\eta,F_{ind}[h])$ be a left universal $H$-morphism.
-%% %%   We extend this to a left adjoint as follows.
-%% %%   \begin{enumerate}
-%% %%   \item First, $F$ acts on morphisms by
-%% %%     \begin{mathpar}
-%% %%       \inferrule
-%% %%       {
-%% %%         \inferrule
-%% %%          {f : C[c,c'] \and \eta_{c'} : H[c',F_0[c']]}
-%% %%          {f : C[c,c'] \vdash \eta_{c'} \circ f : H[c,F_0[c']]}
-%% %%       }
-%% %%       {f : C[c,c'] \vdash F_{ind}[\eta_{c'} \circ f] : D[F_0[c],F_0[c']]}
-%% %%     \end{mathpar}
-%% %%     This preserves identity
-%% %%     \begin{mathpar}
-%% %%       \inferrule
-%% %%       {
-%% %%         \inferrule
-%% %%         {\id \circ \eta = \eta}
-%% %%         {\id = F_{ind}[\eta]}
-%% %%       }
-%% %%       {\id = F_{ind}[\eta \circ \id]}
-%% %%     \end{mathpar}
-%% %%     And composition
-%% %%     \begin{mathpar}
-%% %%       \inferrule
-%% %%       {F_{ind}[\eta \circ f] \circ F_{ind}[\eta \circ g] \circ \eta = \eta \circ f \circ g}
-%% %%       {F_{ind}[\eta \circ f] \circ F_{ind}[\eta \circ g] = F_{ind}[\eta \circ f \circ g]}
-%% %%     \end{mathpar}
-%% %%     Which follows by a couple uses of the factorization:
-%% %%     \begin{align*}
-%% %%       F_{ind}[\eta \circ f] \circ F_{ind}[\eta \circ g] \circ \eta
-%% %%       &= F_{ind}[\eta \circ f] \circ \eta \circ g \\
-%% %%       &= \eta \circ f \circ g
-%% %%     \end{align*}
-%% %%   \item Next, we show that $h : H[c,d] \vdash F_{ind}[h] :
-%% %%     D[F_0[c],d]$ is a natural transformation. As a first step, we show
-%% %%     that $\eta_c : H[c,F_0[c]]$ is natural in $c$ with the action of
-%% %%     $F$ on morphisms we have just defined.
-%% %%     \[ \eta \circ f = F[f] \circ \eta \]
-%% %%     expanding the definition, this is precisely the factorization
-%% %%     property:
-%% %%     \[ \eta \circ f = F_{ind}[\eta \circ f] \circ \eta \]
-
-%% %%     Next, to show that $F_{ind}$ is natural, we need to show that
-%% %%     (when it is well-defined):
-%% %%     \[F_{ind}[g \circ h \circ f] = g \circ F_{ind}[h] \circ F[f] \]
-%% %%     By the factorization property, the following sufficeth:
-%% %%     \begin{align*}
-%% %%      g \circ F_{ind}[h] \circ F[f] \circ \eta
-%% %%      &= g \circ F_{ind}[h] \circ \eta \circ f \tag{naturality of $\eta$}\\
-%% %%      &= g \circ h \circ f \tag{factorization}
-%% %%     \end{align*}
-%% %%   \item Finally, we define $F_{ind}^{-1}[g]$ as
-%% %%     \[ g : D[F_0[c],d] \vdash F_{ind}^{-1}[g] = g \circ \eta : H[c,d] \]
-%% %%     And we need to confirm that it is (1) a section of $F_{ind}$ (2) a
-%% %%     retraction of $F_{ind}$ and (3) natural.
-%% %%     \begin{enumerate}
-%% %%     \item For retraction we need to show $h \vdash F_{ind}[h] \circ
-%% %%       \eta = h$ which is precisely the factorization property.
-%% %%     \item For section, we need to show $g \vdash F_{ind}[g \circ \eta]
-%% %%       = g$. Immediately the uniquenes principle says its sufficient to
-%% %%       show $g \circ \eta = g \circ \eta$.
-%% %%     \item For naturality, we need to show
-%% %%       \[ F_{ind}^{-1}[k \circ g \circ F[f]] = k \circ F_{ind}^{-1}[g] \circ f \]
-%% %%       expanding the definition this is just:
-%% %%       \[ k \circ g \circ F[f] \circ \eta = k \circ g \circ \eta \circ f \]
-%% %%       which is just naturality of $\eta$.
-%% %%     \end{enumerate}
-%% %%   \end{enumerate}
-%% %% \end{construction}
-
-%% \subsection{Products and Units of Internal Categories}
-
-%% %% While we have now proved a beautiful internal theorem in great
-%% %% generality that has many useful instances in different models, we are
-%% %% hard pressed to actually \emph{instantiate} the theorem in the syntax
-%% %% itself because we can't really come up with any useful universal properties!
-%% %% The reason is that Judgment Theory itself is too bare-bones.
-
-%% %% For instance how would we say that a category $\cat C$ has a
-%% %% \emph{terminal object}? Well, it would have to be in the first place a
-%% %% \emph{single} object in $\cat C$, but so far our syntax only enables
-%% %% us to talk about functions and functors. Well an object of $\cat C$ is
-%% %% the same as a function from the unit sort, but so far our syntax only
-%% %% has base sorts.
-
-%% %% So we see that defining the terminal object, \emph{necessitates} some
-%% %% notion of terminal object in our ``metatheory'': Judgment Theory.
-%% %% This mysterious and pervasive aspect of higher category theory is
-%% %% called the \emph{microcosm principle} because it says that in order to
-%% %% discuss a property of a single object in a single category (the
-%% %% microcosm), we need the same sort of object at the meta-level (the
-%% %% macrocosm). The adage here is ``As above, so below''.
-
-%% %% That might all sound very spiritual, but we'll see that it's actually
-%% %% quite a nice heuristic for formalizing universal properties in
-%% %% internal category theory.
-%% %% %
-%% %% In the spirit of this heuristic, instead of first giving a bunch of
-%% %% new type constructions for Judgment Theory and then applying them to
-%% %% define universal properties, we will consider them in pairs, first
-%% %% above, and then below.
-
-%% First we consider the terminal sort and then the idea of a terminal
-%% object in an internal category.
-
-%% Next, the terminal sort trivially has a category structure using the
-%% trivial judgment as its hom set:
-
-%% \begin{definition}[Terminal Category]
-%%   The terminal category $\cat 1$ is defined as having
-%%   \begin{enumerate}
-%%   \item Its object sort as $1$
-%%   \item Its hom set as $\lambda \alpha:1;\beta:1. 1 : \harr 1 1$
-%%   \item Its identity arrow is given by $\lambda \alpha: 1. () : 1$
-%%   \item Composition is defined by $\phi : 1, \psi : 1 \vdash () : 1$
-%%   \item Associativity and unitality are trivial by \textsc{1-$\eta$}
-%%   \end{enumerate}
-%% \end{definition}
-
-%% \begin{definition}[Terminal Object Specification]
-%%   In judgment type theory with a terminal sort/judgment, we can define
-%%   for any category $\cat C$ a profunctor $\alpha^o:\cat C;\beta : \cat
-%%   1 \pipe 1$, which specifies a terminal object.
-%% \end{definition}
-
-%% Now let's instantiate the definition for the terminal object defined
-%% as a negative type and as a representable and see how it reproduces
-%% the standard type-theoretic and category-theoretic definitions.
-
-%% A representation of $\alpha^o:\cat C; \beta : \cat 1 \pipe 1$ consists
-%% of a functor $\beta : \cat 1 \vdash 1_{\cat C} : \cat C$, an introduction rule
-%% \[ \phi : 1 \vdash I_{1}[\phi] : \alpha \to_{\cat C} 1_{\cat C}[\beta] \]
-%% satisfying $I_{1}[\phi] \circ \gamma =I_{1}[\phi \circ \gamma]$ and an
-%% inverse $\psi : \alpha \to_{\cat C} 1_{\cat C}[\beta] \vdash
-%% I_{1}^{-1}[\psi] : 1$ which by \textsc{1-$\eta$} is equal to $()$ and
-%% therefore every $t : \alpha \to_{\cat C} 1_{\cat C}[\beta]$ is equal
-%% to $I_{1}[\phi]$. In light of this it would be appropriate to write
-%% $I_1[\phi]$ as $()$ and then the left-homomorphism property is $()
-%% \circ \gamma = ()$, which is precisely the definition of substitution.
-
-%% This alternate syntax makes even more sense in light of the negative
-%% type presentation. A negative type satisfying the terminal object
-%% specification consists of a functor $1_{\cat C}$ as above and an
-%% introduction rule as above and an elimination form $\vdash \epsilon :
-%% 1$, which is trivial.
-
-%% First the product category
-%% \begin{definition}[Product category]
-%%   For any categories $\cat C_1$, $\cat C_2$ we can form the product
-%%   category whose sort of objects is $\cat C_1 \times \cat C_2$, whose
-%%   arrow judgment is $\alpha^o : \cat C_1 \times \cat C_2;\alpha \pipe
-%%   (\pi_1 \alpha \to_{\cat C_1} \pi_1 \alpha) \times (\pi_2 \alpha
-%%   \to_{\cat C_2} \pi_2 \alpha)$. Identity is given by $\alpha^o : \cat
-%%   C_1 \times \cat C_2;\alpha\pipe\cdot \vdash (\id_{\pi_1 \alpha},
-%%   \id_{\pi_2 \alpha}) : (\pi_1 \alpha \to_{\cat C_1} \pi_1 \alpha)
-%%   \times (\pi_2 \alpha \to_{\cat C_2} \pi_2 \alpha)$ and composition
-%%   by
-%%   \[
-%%   \phi %: (\pi_1 \alpha_0 \to_{\cat C_1} \pi_1 \alpha_1) \times (\pi_2 \alpha_0 \to_{\cat C_2} \pi_2 \alpha_1)
-%%   ,
-%%   \psi %: (\pi_1 \alpha_1 \to_{\cat C_1} \pi_1 \alpha_2) \times (\pi_2 \alpha_1 \to_{\cat C_2} \pi_2 \alpha_2)
-%%   \vdash ((\pi_1 \phi) \circ (\pi_1 \psi), (\pi_2 \psi) \circ (\pi_2 \phi)) :
-%%   (\pi_1 \alpha_0 \to_{\cat C_1} \pi_1 \alpha_2) \times (\pi_2 \alpha_0 \to_{\cat C_2} \pi_2 \alpha_2)  
-%%   \]
-%%   unitality, associativity follow from the same properties of $\cat C_1,\cat C_2$
-%% \end{definition}
-
-%% \begin{definition}[Product Functor Specification]
-%%   For any category $\cat C$, we can specify the product functor by
-%%   \begin{enumerate}
-%%   \item The span is $\alpha^o:\cat C; \beta : \cat C \times \cat C \pipe (\alpha \to_{\cat C} \pi_1 \beta) \times (\alpha \to_{\cat C} \pi_2 \beta)$
-%%   \item Left Composition is defined as
-%%     \[ \phi, \psi \vdash (\pi_1 \phi \circ \psi, \pi_2 \phi \circ \psi)\]
-%%   \item Left unitality:
-%%     \begin{align*}
-%%       (\pi_1 \phi \circ \id, \pi_2 \phi \circ \id)
-%%       &= (\pi_1 \phi, \pi_2 \phi)\\
-%%       &= \phi\tag{$\times-eta$}
-%%     \end{align*}
-%%   \item Left associativity:
-%%     \begin{align*}
-%%       (\phi \circ \psi) \circ \psi'
-%%       &= (\pi_1 \phi \circ \psi, \pi_2 \phi \circ \psi) \circ \psi'\\
-%%       &= (\pi_1 (\pi_1 \phi \circ \psi, \pi_2 \phi \circ \psi) \circ \psi',
-%%       \pi_2 (\pi_1 \phi \circ \psi, \pi_2 \phi \circ \psi) \circ \psi')\\
-%%       &= ((\pi_1 \phi \circ \psi) \circ \psi', (\pi_2 \phi \circ \psi) \circ \psi')\\
-%%       &= (\pi_1 \phi \circ (\psi \circ \psi'), \pi_2 \phi \circ (\psi \circ \psi'))\\
-%%       &= \phi \circ (\psi \circ \psi')
-%%     \end{align*}
-%%   \item Right composition is defined as
-%%     \[ \phi, \theta \vdash (\pi_1 \theta \circ \pi_1 \phi, \pi_2 \theta \circ \pi_2 \phi) \]
-%%   \item Right unitality:
-%%     \begin{align*}
-%%       (\pi_1 (\id,\id) \circ \pi_1 \phi, \pi_2 (\id,\id) \circ \pi_2 \phi)
-%%       &= (\id \circ \pi_1 \phi, \id \circ \pi_2 \phi) \tag{$\times-\beta$}\\
-%%       &= (\pi_1 \phi, \id \circ \pi_2 \phi)\\
-%%       &= \phi \tag{$\times-\eta$}
-%%     \end{align*}
-%%   \item Right associativity
-%%     \begin{align*}
-%%       \theta' \circ (\theta \circ \phi)
-%%       &= \theta' \circ (\pi_1 \theta \circ \pi_1 \phi, \pi_2 \theta \circ \pi_2 \phi)\\
-%%       &= (\pi_1 \theta' \circ \pi_1(\pi_1 \theta \circ \pi_1 \phi, \pi_2 \theta \circ \pi_2 \phi),
-%%       \pi_2 \theta' \circ \pi_2(\pi_1 \theta \circ \pi_1 \phi, \pi_2 \theta \circ \pi_2 \phi))\\
-%%       &= (\pi_1 \theta' \circ (\pi_1 \theta \circ \pi_1 \phi), \pi_2 \theta' \circ (\pi_2 \theta \circ \pi_2 \phi))\\
-%%       &= ((\pi_1 \theta' \circ \pi_1 \theta) \circ \pi_1 \phi, (\pi_2 \theta' \circ \pi_2 \theta) \circ \pi_2 \phi)\\
-%%       &= (\pi_1(\theta' \circ \theta) \circ \pi_1 \phi, \pi_2(\theta' \circ \theta) \circ \pi_2 \phi)\\
-%%       &= (\theta' \circ \theta) \circ \phi \\
-%%     \end{align*}
-%%   \end{enumerate}
-%% \end{definition}
-
-%% \subsection{Discrete and Codiscrete Categories}
-
-%% If we have terminal judgments, then we can define the
-%% \emph{codiscrete} category for any sort $\cat C$.
-%% \begin{definition}
-%%   The codiscrete category $\textrm{CoDisc} \cat C$ is defined as
-%%   \begin{enumerate}
-%%   \item Object sort is $\cat C$
-%%   \item Morphisms judgment is $1$
-%%   \item Identity and composition are given by the unique terms of
-%%     those types, making associativity and unitality trivial to verify.
-%%   \end{enumerate}
-%% \end{definition}
-
-%% If we have \emph{identity} objects, which we write as $\equiv_A$, then we
-%% can define discrete categories.
-%% \begin{definition}
-%%   The discrete category $\textrm{Disc} \cat C$ is defined as
-%%   \begin{enumerate}
-%%   \item Object sort is $\cat C$
-%%   \item Morphisms judgment is $\alpha;\alpha' \pipe \alpha \equiv \alpha'$
-%%   \item Identity and composition are defined as in the section on
-%%     synthetic category theory.
-%%   \end{enumerate}
-%% \end{definition}
-
-%% \subsection{Internal Tensor, Cotensor}
-
-%% We should be able to define internal \emph{tensor} using synthetic
-%% tensor and coequalizers and also internal \emph{cotensor} using
-%% synthetic cotensor and equalizers. These look exactly like the
-%% set-theoretic definitions.
-%% \begin{definition}
-%%   Let $R$ be an internal profunctor from $C$ to $D$ and $S$ be an
-%%   internal profunctor from $D$ to $E$.
-
-%%   Then we can define the internal tensor product $R \odot S$, a
-%%   profunctor from $C$ to $E$ as
-%%   \begin{enumerate}
-%%   \item The underlying span can be defined as a quotient of the synthetic tensor product
-%%     \[ (R \odot S)[\alpha;\gamma] = \exists \beta:\cat D. R \odot S/Rcomp = Scomp \]
-
-%%     Where R-comp and S-comp have as source the span
-%%     \[ \exists \beta:\cat D.~ R[\alpha;\beta] \odot (\exists \beta':\cat D. Hom_C(\beta,\beta') \odot S[\beta';\gamma]) \]
-%%     or the isomorphic
-%%     \[ \exists \beta':\cat D.~ (\exists \beta:\cat D. R[\alpha;\beta] \odot Hom_C(\beta,\beta')) \odot S[\beta';\gamma] \]
-%%     where R-comp composes the $C$-morphism with the $R$-morphism and
-%%     S-comp the $S$-morphism.
-%%   \item Composition is defined using composition in $R,S$.
-%%   \end{enumerate}
-%% \end{definition}
-%% Note that $R$-comp and $S$-comp are each constructed using the
-
-%% \begin{definition}
-%%   Let $R$ be an internal profunctor from $C$ to $D$ and $S$ be an
-%%   internal profunctor from $C$ to $E$. Then the internal contravariant
-%%   cotensor product $R \triangleright S$ is defined as
-%%   \[ (R \triangleright S)[\beta;\gamma] = \{ p : \forall \alpha:\cat C.~ R[\alpha;\beta] \triangleright S[\alpha;\gamma] 
-%%   \pipe R-comp = S-comp \}  \]
-%%   R-comp and S-comp have as codomain the span
-%%   \[ \forall \alpha:\cat C.~ R[\alpha;\beta] \triangleright \forall \alpha':\cat C.~ Hom_C(\alpha',\alpha) \triangleright S[\alpha';\gamma] \]
-%%   or the isomorphic
-%%   \[ \forall \alpha':\cat C.~ (\exists \alpha. Hom_C(\alpha',\alpha) \odot R[\alpha;\beta]) \triangleright S[\alpha';\gamma]
-%%   \]
-%% \end{definition}
-
-%% We can also construct the \emph{free profunctor} from a span using tensor:
-
-%% \begin{definition}
-%%   Let $\cat C$ and $\cat D$ be internal categories and let $R[\alpha;\beta]$ be a
-%%   span from $Ob(\cat C)$ to $Ob(\cat D)$. Then the free profunctor
-%%   extending $R$ to be an internal profunctor from $\cat C$ to $\cat D$ is
-%%   \[ F_{\cat C,\cat D}R[\alpha';\beta'] = \exists \alpha. Hom_C(\alpha',\alpha) \odot (\exists \beta. R[\alpha;\beta] \odot Hom_C(\beta,\beta')) \]
-%%   or the isomorphic
-%%   \[ F_{\cat C,\cat D}R[\alpha';\beta'] = \exists \beta. (\exists \alpha. Hom_C(\alpha',\alpha) \odot R[\alpha;\beta]) \odot Hom_C(\beta,\beta') \]
-%% \end{definition}
-
-%% And maybe we can even define the co-free profunctor from a span using cotensor?
-%% \begin{definition}
-%%   Let $\cat C$ and $\cat D$ be internal categories and let $R[\alpha;\beta]$ be a
-%%   span from $Ob(\cat C)$ to $Ob(\cat D)$. Then the co-free profunctor
-%%   extending $R$ to be an internal profunctor from $\cat C$ to $\cat D$ is
-%%   \[ G_{\cat C,\cat D}R[\alpha';\beta'] =
-%%   \forall \alpha. Hom_C(\alpha,\alpha') \triangleright (\forall \beta. R[\alpha;\beta] \triangleleft Hom_D(\beta',\beta)) \]
-%%   or the isomorphic
-%%   \[ G_{\cat C,\cat D}R[\alpha';\beta'] =
-%%   \forall \beta. 
-%%   (\forall \alpha. Hom_C(\alpha,\alpha') \triangleright R[\alpha;\beta]) \triangleleft Hom_D(\beta',\beta)) \]
-%% \end{definition}
-
 \end{document}

### 2022-10-14 paper/paper.tex
variable rules and syntax figure

--- 2022-10-14 paper/paper.tex
@@
     \begin{array}{rccl}
-      \textrm{Categories} &\cat C, \cat D, \cat E & ::= & M \pipe \cat C \times \cat D \pipe \cat 1
+      \textrm{Categories} &\cat C, \cat D, \cat E & ::= & \unquoth M \pipe \cat C \times \cat D \pipe \cat 1
       \pipe \graphProf \alpha \beta P \pipe \negPresheaf {\cat C} \pipe \posPresheaf {\cat C}
       \\
-      \textrm{Objects} & a, b, c & ::= & \alpha \pipe M a \pipe (a,b) \pipe () \ldots\\
-      \textrm{Sets} & P,Q,R & ::= & \punitinXfromYtoZ {\cat C} a b \pipe \tensorexistsXwithYandZ {\beta} P Q \pipe \homrallXYtoZ \beta P Q \pipe \homlallXYtoZ \alpha R S \\
-      &&& 1 \pipe P \times Q\\
+      \textrm{Objects} & A, B, C, a, b, c & ::= & \alpha \pipe M a \pipe (a,b)
+      \pipe () \pipe \pi_i a \pipe (a_-, a_+, s) \pipe \pi_- a \pipe
+      \pi_+ a \pipe \lambda \alpha:\cat C.R \\
+      \textrm{Sets} & P,Q,R & ::= & \punitinXfromYtoZ {\cat C} a b \pipe \tensorexistsXwithYandZ {\beta} P Q \pipe \homrallXYtoZ \beta P Q \pipe \homlallXYtoZ \alpha R S 
+      \pipe 1 \pipe P \times Q\\
       &&& \pipe \harrapp M a b \pipe \negPresheafAppPtoX a b \pipe \posPresheafAppPtoX a b\\
-      \textrm{Elements} & s,t,u & ::= & \ldots \\
-      \\
+      \textrm{Elements} & s,t,u & ::= & 
+    x \pipe 
+    {\punitelimkontZatABC {\alpha. t}{b_1} s {b_2}} \pipe
+    \punitrefl b \pipe 
+    {\tensorelimWkontZ s {x.\beta.y.r}} \pipe 
+    {\tensorintroatXwithYandZ b s t} \pipe
+    {\homrappXtoYatZ s t a} \\
+    &&&
+    \pipe {\homrlambdaXatYdotZ {x}{\alpha} s}\pipe 
+    {\homlappXtoYatZ t s a} \pipe 
+    {\homllambdaXatYdotZ {x}{\alpha} s} \pipe
+    \pi_i s \pipe 
+    (s_1,s_2) \pipe
+    () \pipe \pi_e a \pipe \pendappXtoY M b \\ 
       \textrm{Type} & A,B,C & ::= &
-      \sum_{X:A} B \pipe \prod_{X:A} B \pipe \Id A M N \\
-      &&& \Cat \pipe \varr{\cat C}{\cat D}\pipe \harr {\cat C} {\cat D}\pipe \pendallXdotY {\alpha:{\cat C}} R\\
-      \textrm{Term} & L,M,N & ::= & X \pipe \lambda X:M. N \pipe M\,N \pipe (M,N) \pipe \pi_1 M \pipe \pi_2 M \mid \text{refl}_{M} \\
-      &&&\lceil \cat C \rceil \pipe \vlambda {\alpha:\cat C} a \pipe \hlambda {\alpha:\cat C}{\beta: \cat D} R \mid \ldots \\
+%      \sum_{X:A} B \pipe \prod_{X:A} B \pipe \Id A M N \\
+      \ldots \pipe \smallCats \pipe \Cat \pipe \varr{\cat C}{\cat D}\pipe \harr {\cat C} {\cat D}\pipe \pendallXdotY {\alpha:{\cat C}} R\\
+      \textrm{Term} & L,M,N & ::= & 
+      %% & X \pipe \lambda X:M. N \pipe M\,N \pipe (M,N) \pipe \pi_1 M \pipe \pi_2 M \mid \text{refl}_{M} \\
+      \ldots \pipe \lceil \cat C \rceil \pipe \lambda {\alpha:\cat C}. a \pipe \lambda ({\alpha:\cat C};{\beta: \cat D}). R \mid \lambda \alpha.t \\
       \textrm{Type Context} & \Gamma, \Delta & ::= & \cdot \pipe \Gamma, X : A\\
       \textrm{Set Context} & \Xi, Z & ::= & \alpha:\cat C \pipe \alpha:\cat C;\beta: \cat D\\
-      \textrm{Transformation Context} & \Phi,\Psi & ::= & \alpha : \cat C \pipe \Phi , x : P,\beta:\cat D\\
+      \textrm{Trans. Context} & \Phi,\Psi & ::= & \alpha : \cat C \pipe \Phi , x : P,\beta:\cat D\\
     \end{array}
@@
 
-All syntactic forms typed in context admit an action of
-substitution. For (meta-language) types and terms, this is as
-usual. Objects $\alpha:\cat C \vdash a : \cat D$ can be substituted for
-object variables $\beta : \cat D$ in other objects. We can also
-substitute objects into \emph{sets}, that is, if we have a set $P$
-parameterized by a contravariant variable $\alpha : \cat C$ and a
-covariant variable $\beta: \cat D$, then we can substitute objects $a :
-\cat C$ and $b : \cat D$ for these variables $P[a/\alpha;b/\beta]$. This
-generalizes the ordinary precomposition of a relation by a function.
-Semantically this is the ``restriction'' of a profunctor along two
-functors, which is just composition of functors if a profunctor is
-viewed as a functor to $\Set$. Modeling this operation as a substitution
-considerably simplifies reasoning using profunctors.
+The variable rules for objects and elements are 
+\[
+\inferrule*{}{ \Gamma \pipe \alpha : \cat C \vdash \alpha : \cat C}
+\qquad
+\inferrule*{}{ \Gamma \pipe \alpha : \cat C, x : R, \beta : \cat D \vdash x : R}
+\]
+All syntactic forms typed in context admit an action of substitution.
+For types and terms, this is as usual. Objects $\alpha:\cat C \vdash a :
+\cat D$ can be substituted for object variables $\beta : \cat D$ in
+other objects. We can also substitute objects into \emph{sets}, that is,
+if we have a set $P$ parameterized by a contravariant variable $\alpha :
+\cat C$ and a covariant variable $\beta: \cat D$, then we can substitute
+objects $a : \cat C$ and $b : \cat D$ for these variables
+$P[a/\alpha;b/\beta]$. This generalizes the ordinary precomposition of a
+relation by a function.  Semantically this is the ``restriction'' of a
+profunctor along two functors, which is just composition of functors if
+a profunctor is viewed as a functor to $\Set$. Modeling this operation
+as a substitution considerably simplifies reasoning using profunctors.
 %
@@
     \quad
-    \inferrule{\forall i \in \{1,2\}.~ \Gamma \pipe\Phi \vdash M_i : R_i}{\Gamma \pipe\Phi \vdash (M_1,M_2) : R_1 \times R_2}
+    \inferrule{\forall i \in \{1,2\}.~ \Gamma \pipe\Phi \vdash s_i : R_i}{\Gamma \pipe\Phi \vdash (s_1,s_2) : R_1 \times R_2}
     \quad
-    \inferrule{\Gamma \pipe\Phi\vdash M : R_1 \times R_2}{\Gamma \pipe\Phi \vdash \pi_i M : R_i}
+    \inferrule{\Gamma \pipe\Phi\vdash s : R_1 \times R_2}{\Gamma \pipe\Phi \vdash \pi_i s : R_i}
   \end{array}
@@
     {\Gamma \vdash \cat C \isaCat \and\Gamma \vdash \cat D \isaCat}
-    {\Gamma\pipe\cdot \vdash \varr{\cat C}{\cat D} \isaTy}
+    {\Gamma \vdash \varr{\cat C}{\cat D} \isaTy}
     \quad
@@
     {\Gamma \pipe \alpha:\cat C \vdash A : \cat D\and
-      \Gamma\pipe\cdot \vdash M : \varr{\cat D}{\cat E}}
+      \Gamma \vdash M : \varr{\cat D}{\cat E}}
     {\Gamma\pipe \alpha:\cat C \vdash M A : \cat E}
@@
     {\Gamma \vdash \cat C \isaCat \and\Gamma \vdash \cat D \isaCat}
-    {\Gamma\pipe\cdot \vdash \harr{\cat C}{\cat D} \isaTy}
+    {\Gamma \vdash \harr{\cat C}{\cat D} \isaTy}
     \quad

### 2022-10-14 paper/paper.tex
missing cite

@@
 
-\title{A Constructive Ordered Predicate Logic for Category Theory}
+\title{A Constructive Ordered Predicate Logic \\ for Category Theory}
 \author{Max S. New\inst{1,2} and Daniel R. Licata\inst{2}}
@@
 category where all morphisms are
-invertible)~\cite{hofmann98groupoid,awodeywarren09identity,voevodsky11wollic}. The reason
+invertible)~\cite{hofmann98groupoid,awodeywarren09identity,voevodsky06homotopy}. The reason
 profunctors are useful for this purpose is that, for any category $\cat
@@
 meta-language that we use to talk about categories/profunctors/natural
-transformations.  The types include the usual $\Pi$ and $\Sigma$ and
-identity types of dependent type theory, along with universes of
+transformations.  In addition to standard dependent type theory with
+$\Pi$ and $\Sigma$ and identity types, we have universes of
 categories, functors, profunctors and natural transformations.

### 2022-10-14 paper/paper.tex
spell check main paper

@@
 computer scientists who do not have a strong background in pure
-mathemantics.
+mathematics.
 %
@@
 %
-We can strenghten this analogy into a formal construction: any
+We can strengthen this analogy into a formal construction: any
 construction in \vett{} can be erased to a corresponding construction
@@
 %
-Moveover, we can work around these unary/binary restrictions to some
+Moreover, we can work around these unary/binary restrictions to some
 extent by viewing the type theory as a domain-specific language embedded
@@
 predicate logic with our ordered variant. Due to the incorporation of
-variance, some unorderd concepts generalize to multiple different
+variance, some unordered concepts generalize to multiple different
 ordered notions. For instance, covariant and contravariant presheaf
@@
 Next, we turn to some of the central theorems of category theory, the
-Yoneda and Co-Yoneda lemmas. Despite being ulimately quite elementary,
+Yoneda and Co-Yoneda lemmas. Despite being ultimately quite elementary,
 these are notoriously abstract. In \ohol{}, we can view these as
@@
 
-We now comapre \vett{} with other calculi for formal category theory.
+We now compare \vett{} with other calculi for formal category theory.
 
@@
 degenerate squares (elements/natural transformations) as a
-``user-facing'' notion, with general squares occuring only in the
+``user-facing'' notion, with general squares occurring only in the
 admissible substitution operations.
@@
 
-\textbf{Acknowledgements.}  This material is based on research sponsored
+\textbf{Acknowledgments.}  This material is based on research sponsored
 by the National Science Foundation under agreement number CCF-1909517

### 2022-10-14 paper/paper.tex
admissibility of general unit elim

@@
 
-% RETURN
 Finally, we define the operation of ``horizontal composition'' of set
 contexts $\Phi \jnctx \Psi$ and its functorial lift $\phi \jnctx
-\psi$. Note that the horizontal composition is only well-defined when
+\psi$.
+\begin{definition}
+  We define horizontal composition of transformation contexts with the
+  admissible typing rule
+  \[\inferrule*[right=(*)]
+  {\Gamma\vdash \Phi\isavectx\and \Gamma\vdash \Psi\isavectx\and d^+\Phi=d^-\Psi}
+  {\Gamma\vdash \Phi \jnctx \Psi \isavectx }\]
+  as follows
+  \begin{align*}
+    \Phi \jnctx \alpha:\cat C &= \Phi\\
+    \Phi \jnctx (\Psi,x:R,\alpha:\cat C) &= (\Phi\jnctx \Psi), x:R, \alpha:\cat C
+  \end{align*}
+
+  And we extend this to an operation on substitutions with the
+  admissible rule
+  \[\inferrule*[right=(*)]
+  {\Gamma\pipe\Psi \vdash \phi : \Phi \and
+   \Gamma\pipe\Psi' \vdash \phi' : \Phi'\and
+   d^+\phi = d^-\phi'
+  }
+  {\Gamma\pipe\Psi \jnctx \Psi' \vdash \phi \jnctx \phi' :: \Phi \jnctx \Phi' }
+  \]
+  Defined as follows
+  \begin{align*}
+    \phi \jnctx a/\alpha &= \phi\\
+    \phi\jnctx (\psi, s/x,a/\alpha) &= (\phi\jnctx\psi), s/x, a/\alpha
+  \end{align*}  
+\end{definition}
+
+\begin{lemma}[Horizontal Category of Contexts/Substitutions]
+  Horizontal composition of contexts is associative (when defined)
+  \[ (\Phi \jnctx \Psi) \jnctx \Sigma = \Phi \jnctx (\Psi \jnctx \Sigma) \]
+  and unital with identity for $\cat C$ given by the single category
+  variable context $\alpha:\cat C$:
+  \[ \alpha:\cat C \jnctx \Phi = \Phi = \Phi \jnctx \beta:\cat D\]
+  when $d^-\Phi = \alpha:\cat C$ and $d^+\Phi = \beta:\cat D$.
+
+  These properties extend to the horizontal composition of element substitutions:
+  \[ \phi \jnctx (\psi \jnctx \sigma) = (\phi \jnctx \psi) \jnctx \sigma \]
+  where the identity is the single variable substitution:
+  \[ a/\alpha \jnctx \phi = \phi = \phi \jnctx b/\beta \]
+  when $d^-\phi = a/\alpha$ and $d^+\phi = b/\beta$.
+\end{lemma}
 
@@
 
-\begin{lemma}[Vertical category]
-  \label{lem:vertical-cat}
-  Substitution of object terms $A[B/\beta]$ is associative with
-  identity given by the variable $\alpha:\cat C \vdash \alpha:\cat C$.
-\end{lemma}
-
-To define vertical composition of squares, we need the following lemma.
-
-\begin{lemma}[Horizontal Category of Contexts]
-  Horizontal composition of contexts is associative (when defined)
-  \[ (\Phi \jnctx \Psi) \jnctx \Sigma = \Phi \jnctx (\Psi \jnctx \Sigma) \]
-  and unital with identity for $\cat C$ given by the single category
-  variable context $\alpha:\cat C$:
-  \[ \alpha:\cat C \jnctx \Phi = \Phi = \Phi \jnctx \beta:\cat D\]
-  when $d^-\Phi = \alpha:\cat C$ and $d^+\Phi = \beta:\cat D$.
-
-  These properties extend to the horizontal composition of element substitutions:
-  \[ \phi \jnctx (\psi \jnctx \sigma) = (\phi \jnctx \psi) \jnctx \sigma \]
-  where the identity is the single variable substitution:
-  \[ a/\alpha \jnctx \phi = \phi = \phi \jnctx b/\beta \]
-  when $d^-\phi = a/\alpha$ and $d^+\phi = b/\beta$.
-\end{lemma}
-
-\begin{lemma}[Properties of Substitution]
-  TODO
-\end{lemma}
+%% \begin{lemma}[Vertical category]
+%%   \label{lem:vertical-cat}
+%%   Substitution of object terms $A[B/\beta]$ is associative with
+%%   identity given by the variable $\alpha:\cat C \vdash \alpha:\cat C$.
+%% \end{lemma}
+
+%% To define vertical composition of squares, we need the following lemma.
 
 %% Theorems about substitution?
+
+% TODO: more properties of substitution??
 
@@
 
-\subsection{Identity Elimination}
-
-As a final syntactic meta-theorem, we prove that the more general identity elimination principle
-\[ \]
-is admissible in the presence of let and right hom types.
-
-TODO
-As an example, any sequent $\Gamma \pipe \alpha, x: R, \beta, y : S, \gamma \vdash
-P$ is equivalent to either of
-\[ \Gamma \vdash \pendallXdotY\alpha \homrallXYtoZ \beta R \homrallXYtoZ \gamma S P 
-\qquad\qquad\qquad \Gamma \vdash \pendallXdotY \gamma \homlallXYtoZ \beta S {\homlallXYtoZ \alpha R P}\]
-This property is useful in giving a simple syntax to positive types.
-
-First, the multiplicative unit is the set $\punitinXfromYtoZ {\cat C}
-{a_1} {a_2}$, semantically, the hom set of the category $\cat C$. The
-rules for this type act like an identity type in a directed type
-theory. The introduction rule defines an identity morphism of an
-object in any category. The elimination rule is then the appropriate
-analogue of the $J$ rule of the identity type. Given any profunctor
-$Q$ from $\cat C$ to itself, we can construct an element $\harrapp Q
-{\alpha_1}{\alpha_2}$ for any objects in $\cat C$ if we specify a
-natural element $\harrapp Q \alpha \alpha$ for any $\alpha$ in $\cat
-C$. The reason this is sufficient, semantically, is that all
-profunctors must be functorial in their two indices. In fact, as we
-will see in the section on synthetic category theory, this induction
-principle is how we can show internally that profunctors are
-functorial in their indices. We choose to express this elimination
-form using the right hom type as its syntax more closely matches the
-usual syntax of function types, but this choice is arbitrary, and we
-would get an equivalent calculus had we used the left hom, or added
-both.
-
-%% \subsection{Dependent Types}
-
-%% Next, we cover the basic dependent type theory structure: $\Pi,\Sigma$
-%% and identity types.
-
-%% \begin{figure}
-%%   \begin{mathpar}
-%%   \inferrule
-%%   {\Gamma \pipe\cdot \vdash S \isaSet\and \Gamma \pipe\cdot \vdash T \isaSet}
-%%   {\Gamma\pipe\cdot \vdash \prod_{x:S} T \isaSet}
-
-%%   \inferrule
-%%   {\Gamma \pipe \cdot \vdash M : \prod_{x:S} T\and
-%%     \Gamma \pipe \cdot \vdash N : S}
-%%   {\Gamma \pipe \cdot \vdash M N : T[N/x]}
-
-%%   \inferrule
-%%   {\Gamma,x:S \pipe \cdot \vdash M : T}
-%%   {\Gamma \pipe \cdot \vdash \lambda x:S. M : \prod_{x:S} T}
-
-%%   \inferrule
-%%   {\Gamma \pipe\cdot \vdash S \isaSet\and \Gamma \pipe\cdot \vdash T \isaSet}
-%%   {\Gamma\pipe\cdot \vdash \sum_{x:S} T \isaSet}
-
-%%   \inferrule
-%%   {~}
-%%   {\Gamma\pipe\cdot \vdash \top \isaSet}
-
-%%   \inferrule
-%%   {\Gamma \pipe \cdot \vdash M : S\and \Gamma \pipe \cdot \vdash N : S}
-%%   {\Gamma \pipe \cdot \vdash \textrm{Id}_S M N \isaSet}
-
-%%   \inferrule
-%%   {\Gamma \pipe \cdot \vdash M : S}
-%%   {\Gamma \pipe \cdot \vdash \textrm{refl}(M) : \textrm{Id}_S M M}
-
-%%   \inferrule
-%%   {\Gamma,x:S,y:S,p:\textrm{Id}_S x y \pipe \cdot \vdash C \isaSet\and
-%%    \Gamma \pipe \cdot \vdash M_k : \prod_{x:S} C[x/y,\textrm{refl}(x)/p]}
-%%   {\Gamma \pipe \cdot \vdash \textrm{J}(x,y,p.C,M) : \prod_{x:S,y:S,p:\textrm{Id}_S x y} C}
-%%   \end{mathpar}
-%%   \caption{Ordinary Dependent Type Structure}
-%% \end{figure}
+\subsection{Generalized Unit Elimination}
+
+The unit elimination rule presented in Section~\ref{sec:syntax} is
+more restrictive than universal property of a unit in a virtual double
+category that we use in the semantics. So in order for our calculus to
+be complete for virtual equipments with units, we need to show that
+the more general unit elimination principle is admissible and
+satisfies the correct $\beta\eta$ rules.
+
+The more general rules are as follows
+\begin{mathpar}
+  \inferrule*[Right=UnitElim]
+  {\Phi[\alpha/\alpha_-]\jnctx \Psi[\alpha/\alpha_+] \vdash t : P[\alpha/\alpha_-;\alpha/\alpha_+]}
+  {\Phi,x:\punitinXfromYtoZ {}{\alpha_-} {\alpha_+}, \Psi \vdash \textrm{ind}_{\to}^{\Phi;\Psi}(\alpha.t; x) : P}
+
+  (\textrm{ind}_{\to}^{\Phi;\Psi}(\alpha.t; x))[\phi,\punitrefl \alpha,\psi] = t[\phi\jnctx\psi]
+
+  \inferrule
+  {\Phi,x:\punitinXfromYtoZ {}{\alpha_-} {\alpha_+}, \Psi \vdash s : P}
+  {\Phi,x:\punitinXfromYtoZ {}{\alpha_-} {\alpha_+}, \Psi \vdash s = \textrm{ind}_{\to}^{\Phi;\Psi}(\alpha. s[\punitrefl \alpha/x]; x){}}
+\end{mathpar}
+
+The rule is more general because it allows the elimination of an input
+of the unit type with non-trivial contexts $\Phi, \Psi$ surrounding
+it, whereas the rule presented earlier would only allow this
+elimination if $x$ were the only variable. We did not include this
+more general rule because it does not seem possible to define it with
+an admissible action of substitution, and would require a notion of
+explicit substitution.
+
+In the presence of hom types, we can prove this more general
+elimination is admissible, because the judgment
+\[ \Phi,x:\punitinXfromYtoZ {}{\alpha_-} {\alpha_+}, \Psi \vdash s : P \]
+is in natural bijection with the judgment
+\[ \alpha_-,x:\punitinXfromYtoZ {}{\alpha_-} {\alpha_+}, \alpha_+ \vdash s : \Phi \triangleright P \triangleleft \Psi \]
+where $\Psi \triangleright P \triangleleft \Phi$ (note the reversal of
+order) is a type constructed by recursion on $\Phi$ and $\Psi$ using
+uses the hom types.
+
+\begin{definition}[Generalized Unit Elimination]
+  We define $\textrm{ind}_{\to}^{\Phi;\Psi}(\alpha.t; x)$ by induction
+  on $\Phi/\Psi$.
+  \begin{align*}
+    \textrm{ind}_{\to}^{\alpha_-;\alpha_+}(\alpha.t; x) &= \punitelimkontZatABC{\alpha.t}{\alpha_-}{x}{\alpha_+}\\
+    \textrm{ind}_{\to}^{\alpha_-;\Psi,y,\beta}(\alpha.t; x) &=
+    \homrappXtoYatZ
+        {(\textrm{ind}_{\to}^{\alpha_-;\Psi}(\alpha.\homrlambdaXatYdotZ y \beta t; x))} y \beta \\
+    \textrm{ind}_{\to}^{\beta, y, \Phi;\Psi}(\alpha.t; x) &=
+    \homlappXtoYatZ
+        {(\textrm{ind}_{\to}^{\alpha_-;\Psi}(\alpha.\homllambdaXatYdotZ y \beta t; x))} y \beta \\
+  \end{align*}
+\end{definition}
+
+\begin{lemma}[Generalized Unit Elim $\beta\eta$]
+  The admissible generalizaed unit elimination satisfies the described $\beta\eta$ equations.
+\end{lemma}
+\begin{proof}
+  By induction on $\Phi/\Psi$
+  First, $\beta$
+  \begin{itemize}
+  \item If $\Phi = \Psi = \alpha$
+    \begin{align*}
+    (\punitelimkontZatABC{\alpha.t}{\alpha_-}{x}{\alpha_+})[\alpha,\punitrefl \alpha, \alpha] &=
+    \punitelimkontZatABC{\alpha.t}{\alpha}{\punitrefl}{\alpha}\\
+    &= t \tag{Unit$\beta$}
+  \end{align*}
+  \item If $\Phi = \alpha$ and $\Psi = \Psi, y, \beta$
+    \begin{align*}
+    (\homrappXtoYatZ
+      {(\textrm{ind}_{\to}^{\alpha_-;\Psi}(\alpha.\homrlambdaXatYdotZ y \beta t; x))} y \beta)[\phi,\punitrefl \alpha, \psi, s/y, b/\beta]
+      &= (\homrappXtoYatZ
+        {(\textrm{ind}_{\to}^{\alpha_-;\Psi}(\alpha.\homrlambdaXatYdotZ y \beta t; x)[[\phi,\punitrefl \alpha, \psi]])} s b)\tag{Definition}\\
+      &= (\homrappXtoYatZ {\homrlambdaXatYdotZ y \beta t[\phi,\psi]} s b) \tag{Induction}\\
+      &= t[\phi,\psi,s/y,b/\beta]\tag{Hom$\beta$}
+    \end{align*}
+  \item $\Phi = \beta, y, \Phi$ case is similar to previous.
+  \end{itemize}
+
+  Next $\eta$.
+  \begin{itemize}
+  \item $\Phi = \alpha_- $ and $\Psi = \alpha_+$: immediate by unit $\eta$.
+  \item $\Phi = \alpha_-$ and $\Psi = \Psi, y, \beta$:
+    \begin{align*}
+      s &= \homrappXtoYatZ {\homrlambdaXatYdotZ y \beta s} y \beta\tag{Hom $\beta$}\\
+      &= \homrappXtoYatZ {(\textrm{ind}_{\to}^{\alpha_-;\Psi}(\alpha.\homrlambdaXatYdotZ y \beta {s[\punitrefl \alpha/x]}; x))} y \beta \tag{Induction}\\
+      &= \textrm{ind}_{\to}^{\Phi;\Psi,y,\beta}(\alpha. s[\punitrefl \alpha/x]; x)\tag{Definition}
+    \end{align*}
+  \end{itemize}
+\end{proof}
 

### 2022-10-14 paper/paper.tex
semantics section of appendix

@@
 \newcommand{\SymMonCat}{\textrm{SymMonCat}}
+\newcommand{\Syn}{\textrm{Syn}}
 \newcommand{\vtkmnd}{\mathbb{K}\text{Mod} (\vcat{V},T)}
@@
 
-In this section, we provide the full definitions of the universal
+In this section, we provide the full descriptions of the universal
 properties in a virtual equipment corresponding to each connective in
-\vett{}, and the most critical details of the soundness and
-completeness proof.
-
-TODO
-
-Next, we develop the basics of the model theory for SCT. Because SCT
-includes not just a syntax for category theory, but is also indexed by
-an ordinary context of dependent type theory $\Gamma$, a model of SCT
-is not given by a \emph{single} virtual double category, but instead
-by an \emph{indexed family} of VDCs, parameterized by an object of a
-category that interprets dependent type theory. We call such a
-structure a \emph{VDC-hyperdoctrine}. We then show how to construct a
-from any VDC, a VDC-hyperdoctrine indexed by the category of sets,
-which has the intuitive reading that any of the VDC judgments $\Gamma
-\vdash J$ is interpreted as a family of $J$-data parameterized
-by elements of the set $\Gamma$.
+\vett{}.
+
+\begin{definition}[Universal Properties for Category Connectives]
+  Let $\mathcal V$ be a virtual equipment.
+  \begin{enumerate}
+  \item Let $C$ be a small object, then a contravariant presheaf object $\mathcal P^- C$ is an object with natural isomorphism
+    $\mathcal V_o(A, \mathcal P^- C) \cong \{ R \in V_h \,|\, s(R) = C \wedge t(R) = A \}$
+  \item Let $C$ be a small object, then a covariant presheaf object $\mathcal P^+ C$ is an object with natural isomorphism
+    $\mathcal V_o(A, \mathcal P^+ C) \cong \{ R \in V_h \,|\, t(R) = C \wedge s(R) = A \}$
+  \item Let $R$ be a horizontal arrow, then a tabulator $\int R$ is an object with natural isomorphism
+    $\mathcal V_o(A, \int R) \cong \sum_{f : {\mathcal V}_o(A, s(R))}\sum_{g : {\mathcal V}_o(A, t(R))}{\mathcal V}_2(\cdot;f;g;R)$
+  \item A nullary product is an object $1$ with natural isomorphism $\mathcal V_o(A,1) \cong 1$
+  \item A binary product of $B$ and $C$ is an object $B \times C$ with natural isomorphism $\mathcal V_o(A,B \times C) \cong {\mathcal V}_o(A,B)  \times {\mathcal V}_o(A,C) $
+  \end{enumerate}
+\end{definition}
+
+\renewcommand{\vec}{\overrightarrow}
+\begin{definition}[Universal Properties for Set Connectives]
+  Let $\mathcal V$ be a virtual equipment.
+  \begin{enumerate}
+  \item A unit $U_C$ for an object $C$ is a horizontal arrow $U_C$ with $s(U_C) = t(U_c) = C$ with natural isomorphism
+    $\mathcal V_2(\vec P, U_c, \vec Q; f; g; R) \cong \mathcal V_2(\vec P, \vec Q; f; g; R)$
+  \item A tensor of horizontal arrows $P$ and $Q$ where $t(P) = s(Q)$ is a horizontal arrow $P \odot Q$ with $s(P\odot Q) = s(P)$ and $t(P \odot Q) = t(Q)$ with natural isomorphism
+    $\mathcal V_2(\vec R, P \odot Q,\vec S; f; g; T) \cong \mathcal V_2(\vec R, P, Q,\vec S; f; g; T)$.
+  \item A covariant hom of $P$ and $Q$ where $t(P) = t(Q)$ is a horizontal arrow $P \triangleright Q$ with $s(P\triangleright Q) = s(Q)$ and $t(P \triangleright Q) = s(P)$ with natural isomorphism
+    $\mathcal V_2(\vec R; f; \id; P \triangleright Q) \cong \mathcal V_2(\vec R, P; f; \id; Q)$
+  \item A contravariant hom of $P$ and $Q$ where $s(P) = s(Q)$ is a horizontal arrow $P \triangleleft Q$ with $s(P\triangleleft Q) = t(Q)$ and $t(P \triangleleft Q) = t(P)$ with natural isomorphism
+    $\mathcal V_2(\vec R; \id; g; P \triangleleft Q) \cong \mathcal V_2(Q, \vec R; \id; g; P)$
+  \item A nullary product for an object $C$ is a horizontal arrow $1_C$ with $s(1_C) = t(1_C) = C$ with natural isomorphism
+    $\mathcal V_2(\vec P; f; g; 1_C) \cong 1$
+  \item A binary product of horizontal arrows $P$ and $Q$ where $s(P)= s(Q)$ and $t(P) = t(Q)$ is a horizontal arrow $P \times Q$ with $s(P \times Q) = s(P)$ and $t(P \times Q) = t(P)$ with natural isomorphism
+    $\mathcal V_2(\vec R; f; g; P \times Q) \cong \mathcal V_2(\vec R; f; g; P) \times \mathcal V_2(\vec R; f; g; Q)$
+  \end{enumerate}
+
+  We require that in our models, units exist for all objects, tensors
+  and homs overs small objects exist and all finite products exist. We
+  additionally require that the choice of tensors, homs and products
+  commute strictly with restrictions in that
+  \begin{enumerate}
+  \item $(P \odot Q)(f,g) = (P(f,\id) \odot Q(\id, g))$
+  \item $(P \triangleright Q)(f,g) = (P(g, \id) \triangleright Q(f,\id))$
+  \item $(P \triangleleft Q)(f,g) = (P(\id, g) \triangleleft Q(\id, f))$
+  \item $1(f,g) = 1$
+  \item $(P \times Q)(f,g) = (P(f, g) \times Q(f, g))$
+  \end{enumerate}
+  Note that these equations necessarily hold up to isomorphism, even
+  if we do not require them to commute strictly.
+\end{definition}
 
@@
 
-\subsection{Notion of Model}
-
-%% 1. A strict double category is a monoid in Span(Cat)
-%% 2. A strict double category internal to a topos T is a monoid in Span(T-Cat)
-%% 3. A double category with display pro-arrows is a strict double category with a span of profunctors that is representable
-%%    C0 <- C1 -> C0
-%% 4. UMPs are given on top of that structure
-
-We build up the notion of a model in several stages.
-
-\begin{definition}
-  A strict double category is a monoid in the bicategory $Span(Cat)$.
-\end{definition}
-
-This data consists of a span $C_0 \leftarrow C_1 \rightarrow C_0$
-where $C_0$ is the vertical category and $C_1$ is the category whose
-objects are horizontal morphisms and morphisms are squares, with the
-two projections giving the left and right ``frame'' for the horizontal
-morphisms/squares. The unit gives the horizontal identity morphisms
-and the monoid multiplication is the horizontal composition.
-
-\begin{definition}
-  A double category with displays (DCD) consists of
+\subsection{Completeness}
+
+Next we describe the syntactic properties of substitution that are
+needed in order to prove the \emph{completeness} theorem, that is,
+that the syntax of \vett{} presents a hyperdoctrine of virtual
+equipments.
+
+\begin{definition}[Syntactic Virtual Equipment]
+  Fix a context $\Gamma$. Define a virtual equipment $\Syn^\Gamma$ as
+  follows:
   \begin{enumerate}
-  \item A strict double category $C$
-  \item A category $C_d$ of (horizontal) ``display morphisms'' and ``display squares''.
-  \item A span of profunctors:
-    % https://q.uiver.app/?q=WzAsOCxbMiwyLCJDXzEiXSxbMCwyLCJDXzAiXSxbNCwyLCJDXzAiXSxbMiwwLCJDX2QiXSxbMCwwLCJDXzAiXSxbNCwwLCJDXzAiXSxbMSwxLCJcXExlZnRhcnJvdyJdLFszLDEsIlxcUmlnaHRhcnJvdyJdLFszLDAsIkNfXFx2ZGFzaCIsMix7InN0eWxlIjp7ImJvZHkiOnsibmFtZSI6ImJhcnJlZCJ9fX1dLFs1LDIsIlxcdGV4dHJte0hvbX0iLDAseyJzdHlsZSI6eyJib2R5Ijp7Im5hbWUiOiJiYXJyZWQifX19XSxbNCwxLCJcXHRleHRybXtIb219IiwyLHsic3R5bGUiOnsiYm9keSI6eyJuYW1lIjoiYmFycmVkIn19fV0sWzMsNCwicyIsMl0sWzMsNSwidCJdLFswLDIsInQiLDJdLFswLDEsInMiXV0=
-\[\begin{tikzcd}
-	{C_0} && {C_d} && {C_0} \\
-	& \Leftarrow && \Rightarrow \\
-	{C_0} && {C_1} && {C_0}
-	\arrow["{C_\vdash}"', "\shortmid"{marking}, from=1-3, to=3-3]
-	\arrow["{\textrm{Hom}}", "\shortmid"{marking}, from=1-5, to=3-5]
-	\arrow["{\textrm{Hom}}"', "\shortmid"{marking}, from=1-1, to=3-1]
-	\arrow["s"', from=1-3, to=1-1]
-	\arrow["t", from=1-3, to=1-5]
-	\arrow["t"', from=3-3, to=3-5]
-	\arrow["s", from=3-3, to=3-1]
-\end{tikzcd}\]
-\item A fully faithful functor $i : C_d \to C_1$ that respects source and target in that the following commute:
-% https://q.uiver.app/?q=WzAsNixbMiwyLCJDXzEiXSxbMCwyLCJDXzAiXSxbNCwyLCJDXzAiXSxbMiwwLCJDX2QiXSxbMCwwLCJDXzAiXSxbNCwwLCJDXzAiXSxbMywwLCJpIiwyXSxbNSwyLCJcXGlkIl0sWzQsMSwiXFxpZCIsMl0sWzMsNCwicyIsMl0sWzMsNSwidCJdLFswLDIsInQiLDJdLFswLDEsInMiXV0=
-\[\begin{tikzcd}
-	{C_0} && {C_d} && {C_0} \\
-	\\
-	{C_0} && {C_1} && {C_0}
-	\arrow["i"', from=1-3, to=3-3]
-	\arrow["\id", from=1-5, to=3-5]
-	\arrow["\id"', from=1-1, to=3-1]
-	\arrow["s"', from=1-3, to=1-1]
-	\arrow["t", from=1-3, to=1-5]
-	\arrow["t"', from=3-3, to=3-5]
-	\arrow["s", from=3-3, to=3-1]
-\end{tikzcd}\]
-and $i$ represents $C_\vdash$ in that there is a natural isomorphism:
-    \[ e : C_\vdash(\Phi, R) \cong C_1(\Phi, i(R)) \]
-    that commutes with the natural transformations $s,t$.
+  \item The vertical category $\Syn^\Gamma_o$ has categories $\Gamma
+    \vdash \cat C \isaCat$ as objects, small categories as small
+    objects and as arrows from $\cat C$ to $\cat D$ objects
+    $\alpha:\cat C \vdash b : \cat D$ modulo renaming of the input
+    variable. Composition is given by substitution and identity is the
+    variable.
+  \item The horizontal arrows are the sets $\alpha:\cat C; \beta:\cat
+    D \vdash R$ up to renaming with source $\cat C$ and target $\cat
+    D$.
+  \item Note that composable strings $\vec R$ of horizontal arrows are
+    in bijection with contexts $\Phi$. Then we can define a 2-cell
+    $\Syn^\Gamma_2(\Phi, a, b, S)$ to be an element
+    $\Gamma\pipe\Phi\vdash s : S[a/\alpha;b/\beta]$.
+
+    Composition is defined by substitution $t[\phi]$ as substitutions
+    are in bijection with the ``sequences of 2-cells'' used in the
+    definition of a virtual equipment. Associativity says that
+    $t[\phi][\psi] = t[\phi[\psi]]$ where the composition $\phi[\psi]$
+    is defined below and corresponds exactly to the associativity rule
+    in a virtual equipment. The unit is the variable, and they are
+    unital as $x[s/x] = s$ and $s[\vec {x/x}] = s$.
+
+  \item Restriction along vertical arrows is given by substitution
+    $R(a,b) = R[a/\alpha;b/\beta]$. This is strictly associative and
+    unital, and the cartesian cell from $R$ to $R[a/\alpha;b/\beta]$
+    is just the identity $x:R[a/\alpha;b/\beta] \vdash x :
+    R[a/\alpha;b/\beta]$.
   \end{enumerate}
 \end{definition}
-
-In terms of the syntax of SCT, the category $C_0$ models the category
-and functor judgments, the category $C_1$ models the profunctor
-\emph{contexts} and substitutions and the category $C_d$ models the
-profunctor judgment. The profunctor $C_\vdash$ models the
-transformation judgment and the functor $i$ models the construction of
-a ``singleton'' context $\alpha:\cat C,x:R,\beta:\cat D$. The fact
-that $i$ is fully faithful corresponds to the fact that we don't
-provide in the syntax a notion of morphism of profunctors separate
-from the notion of substitution of singleton contexts.
-
-A DCD is a more general notion that a VDC, but they are closely
-related, in that VDCs form a coreflective subcategory of DCDs. This is
-analogous to the situation in models of dependent type theory, where
-DCDs are analogous to Categories with Families \cite{cwfs,natmodels}
-and VDCs are analogous to the more ``syntactic'' notion of a
-Contextual Category/C-system \cite{ctxcat,csystem}. First, any DCD
-induces a VDC by taking the display proarrows as the proarrows and
-interpreting 2-cells using $i$ and $C_\vdash$.  Second, any VDC
-presents a DCD whose double category has the original VDC pro-arrows
-as the display morphisms and \emph{composable strings} of proarrows as
-horizontal morphisms, which gives a strict double category.
-%
-Taking as morphisms the strict morphisms of DCDs/VDCs, these
-constructions form a coreflective embedding of VDCs into DCDs.
-
-The principle advantage of DCDs over VDCs is that the presentation of
-their algebraic structure is considerably simplified and is
-straightforward to describe as a Generalized Algebraic Theory in the
-sense of \citet{CartmellGAT}. This is because we don't \emph{require}
-that our profunctor contexts are lists, but instead merely the fact
-that their concatenation forms a strict category. Then the
-associativity/identity equations for a DCD are precisely those of a
-double category.
-
-The final notion of composition that we require for our structure is
-that display morphisms have \emph{restrictions} along vertical arrows.
-
-\begin{definition}
-  A DCD is an Equipment with displays when
-  \begin{enumerate}
-  \item The functor $(s,t) : C_d \to C_0 \times C_0$ is a split
-    fibration, with restriction operation written as $(f,g)^*R$.
-  \item The profunctor $C_\vdash$ respects restrictions in that
-    \[ C_\vdash^{f',g'}(\Phi, (f,g)^*R) \cong C_\vdash^{f\circ f',g\circ g'}(\Phi, R) \]
-  \end{enumerate}
-\end{definition}
-Specializing this property give us that $C_\vdash^{f,g}(\Phi,R) \cong
-C_\vdash^{\id,\id}(\Phi,(f,g)^*R)$, which justifies that we only need
-to give syntax for terms $\Phi \vdash t : R$ where $\Phi$ and $R$ have
-the same covariant/contravariant dependencies.
-
-We can then define the remaining connectives by UMPs in an equipment
-with displays, adding in the extra strictness criterion that these
-connectives appropriately commute with restriction.
-
-\begin{figure}
-  \begin{mathpar}
-    C_\vdash^{f,g}(\Phi \circ_{C} \Psi, R) \cong C_\vdash^{f,g}(\Phi \circ_C i(U_C) \circ_c \Psi, R) \\
-    C_\vdash^{f,g}(\Phi \circ_{C} iP \circ_D iQ \circ_E \Psi, R) \cong C_\vdash^{f,g}(\Phi \circ_C i(P \odot Q) \circ_E \Psi, R) \and (f,g)^*(P \odot Q) = (f,\id)^*P \odot (\id,g)^*Q\\
-    C_\vdash^{\id_C,g}(P \circ_c \Phi, R) \cong C_\vdash^{\id_D,g}(\Phi, P \triangleright Q) \and (f,g)^*(P \triangleright Q) = (\id,f)^*P \triangleright (\id,g)^*Q\\
-    C_\vdash^{f, \id_C}(\Phi \circ_c R, P) \cong C_\vdash^{f, \id_D}(\Phi, P \triangleleft Q)\and (f,g)^*(P \triangleleft Q) = (f,\id)^*P \triangleleft (g,\id)^*Q\\
-    C_\vdash^{f,g}(\Phi,1) \cong 1 \and (f,g)^*1 = 1\\
-    C_\vdash^{f,g}(\Phi,P \times Q) \cong C_\vdash^{f,g}(\Phi,P) \times C_\vdash^{f,g}(\Phi, Q)\and (f,g)^*(P \times Q) = (f,g)^*P \times (f,g)^*Q\\
-    C_0(C,\smallint R) \cong C_\vdash(\id_C, R)\\
-  \end{mathpar}
-  \caption{UMPs in an Equipment with Displays}
-\end{figure}
-
-Next, each of the connectives of SCT are modeled by different
-universal properties that are profunctor isomorphisms involving
-principally $C_\vdash$. We define these in Figure\ref{fig:eq-umps}
-
-\begin{definition}
-  An \emph{equipment with displays} (ED) is a DCD with all of the
-  universal properties in Figure\ref{eq-umps}.
-\end{definition}
-
-Finally, to model the logic over SCT provided by the dependently typed
-context $\Gamma$, we use a category with families $Ctx$, and interpret
-all of the prior constructions in this section to be internal to
-presheaves on $Ctx$. This models the syntactic aspect that all double
-category structure is parameterized by a context $\Gamma$ and stable
-under substitution. Since the category of presheaves is a topos, all
-definitions given in set-theoretic language can be interpreted
-internally.
-
-\begin{definition}
-  A Category with Families consists of
-  \begin{enumerate}
-  \item A category $\textrm{Ctx}$
-  \item with a terminal object $1 : \textrm{Ctx}$
-  \item A presheaf $\textrm{Ty} : \widehat {\textrm{Ctx}}$ of types in context.
-  \item A presheaf $\textrm{Tm} : \widehat{\int\textrm{Ty}}$ of terms of
-    a given type.
-  \item With a context extension functor $\textrm{ext} : \int
-    \textrm{Ty} \to \textrm{Ctx}$ with universal property:
-    \[ \textrm{Ctx}(\Delta,\textrm{ext}(A)) \cong (\gamma : \textrm{Ctx}(\Delta,\textrm{ctx}(A))) \times \textrm{Tm}(A[\gamma])  \]
-  \end{enumerate}
-\end{definition}
-
-This can be extended in a standard way to model $\Pi,\Sigma, \Id$
-types \citet{cwf-stuff}.
-%
-The final connectives we need are the ``universes'' that internalize
-the double category structure as types in the CwF.
-
-\begin{definition}
-  Let $(\textrm{Ctx},\textrm{Ty},\textrm{Tm},\textrm{ext})$ be a
-  category with families and let $\mathcal V = (C,C_d)$ be a DCD internal to
-  $\Set^\textrm{Ctx}$. Then
-  \begin{enumerate}
-  \item $\mathcal V$ objects are representable as types if there is a
-    type given by a section $\textrm{Ob} \in \textrm{Ty}(1)$ with a
-    natural isomorphism
-    \[ \textrm{Tm}(!_\Gamma^*\textrm{Ob}) \cong C_{0}(\Gamma) \]
-  \item $\mathcal V$ vertical arrows are representable as types if
-    there is a type
-    \[ \textrm{Tm}(Functor)(\Gamma)(a,b) \cong C_0(\Gamma)(a,b)
-    \]
-    and similarly horizontal displays are representable as types if
-    \[ \textrm{Tm}(Prof)(\Gamma)(a,b) \cong \{ R \in C_d(\Gamma) \,|\, s(R) = a \wedge t(R) = b \}
-    \]
-  \item Closed $\mathcal V$ transformations are representable as types if
-    \[ \textrm{Tm}(\textrm{End})(\Gamma)(a,R) \cong C_\vdash^{\id,\id}(r(a),R) \]
-  \end{enumerate}
-\end{definition}
-
-We then arrive at our final notion of model by collecting all of this
-algebraic structure into one definition.
-\begin{definition}
-  A hyperdoctrine of equipments with displays (HED) is a pair of a
-  category with families $Ctx$ modeling $\Sigma,\Pi,\Id$ and a DCD
-  internal to $\Set^{Ctx^o}$ with categories, functors, displays and
-  unary cells representable as types.
-\end{definition}
-
-This notion of HED is straightforwardly, if laboriously, interpretable
-as a GAT.
-
-%% \begin{definition}
-%%   Given a well-formed signature $\Gamma$ and a virtual equipment
-%%   $\mathcal V$, we mutually define the following notions:
-
-%%   \begin{itemize}
-%%   \item The set $\sem{\Gamma}_{\mathcal V}$
-%%   \item For any $\Gamma \vdash A$ and $\gamma \in \sem{\Gamma}_{\mathcal V}$, a set $\sem{A}_{\mathcal V} \gamma$
-%%   \item For any $\Gamma \vdash M : A$ and $\gamma \in \sem{\Gamma}_{\mathcal V}$, an element $\sem{M}_{\mathcal V} \in \sem{A}_{\mathcal V} \gamma$
-%%   \item For any $\Gamma \vdash \cat C \isaCat$ and $\gamma \in \sem{\Gamma}_{\mathcal V}$, an object $\sem{\cat C}_{\mathcal V}\gamma$
-%%   \item For any $\Gamma\pipe\alpha:\cat C \vdash a :\cat D$ and $\gamma \in \sem{\Gamma}_{\mathcal V}$, a vertical morphism $\sem{\alpha.a}_{\mathcal V}\gamma \in \mathcal V_v(\sem{\cat C}_{\mathcal V}\gamma, \sem{\cat D}_{\mathcal V}\gamma,)$
-%%   \item For any $\Gamma\pipe\Xi \vdash R \isaSet$ and $\gamma \in
-%%     \sem{\Gamma}_{\mathcal V}$, a horizontal arrow $\sem{R}_{\mathcal
-%%       V}\gamma \in \mathcal V_h$ such that $d^\pm\sem{R}_{\mathcal
-%%       V}\gamma = \sem{d^\pm\Xi}_{\mathcal V}\gamma$. We also extend
-%%     this in the obvious way to transformation contexts
-%%     $\sem{\Phi}_{\mathcal V}\gamma$
-%%   \item For any $\Gamma\pipe\Phi \vdash t : R$ and $\gamma \in
-%%     \sem{\Gamma}_{\mathcal V}$, a 2-cell $\sem{t}_{\mathcal V}\gamma
-%%     \in V_h(\sem{\Phi}_{\mathcal V}\gamma;\sem{R}_{\mathcal V}\gamma)$
-%%     with $d^\pm\sem{t}_{\mathcal V}\gamma = \id$.
-%%   \end{itemize}
-
-%%   %% \begin{align*}
-%%   %%   \sem{\cdot}_{\mathcal V} &= 1\\
-%%   %%   \sem{\Gamma, x:A}_{\mathcal V} &= \sum_{\gamma \in \sem{\Gamma}_{\mathcal V}} \sem{A}_{\mathcal V}\gamma\\
-%%   %% \end{align*}
-
-%%   %% with an 
-%%   %% \begin{align*}
-%%   %%   \sem{\Gamma, x:A}_{\mathcal V} &= \sem{\Gamma}_{\mathcal V} \times \sem{A}_{\mathcal V}\\
-%%   %%   \sem{\Gamma}_{\mathcal V} &= 1\\\\
-%%   %%   \sem{\Sort}_{\mathcal V} &= \mathcal V_0\\
-%%   %%   \sem{\varr{\cat C}{\cat D}}_{\mathcal V} &= \mathcal V_v(\sem{\cat C})
-%%   %% \end{align*}
-%% \end{definition}
-
-\subsection{Initiality}
-
-In this section we give a sketch of an initiality proof of the syntax
-of SCT in HEDs. Note that we consider all the following sets up to
-$\alpha$-equivalence of names.
-
-\begin{construction}{Completeness}
-  SCT presents a HED with
-  \begin{enumerate}
-  \item Category with Families provided by the contexts,
-    substitutions, types and terms.
-  \item Objects of $C_0^\Gamma$ provided by the terms $\Gamma \vdash \cat C \isaCat$
-  \item Morphisms of $C_0^\Gamma$ provided by terms $\Gamma
-    \alpha:\cat C \vdash b \cat D$ with source $\cat C$ and target
-    $\cat D$. With identity given by a variable and composition by identity.
-  \item Objects of $C_1^\Gamma$ are provided by the profunctor
-    contexts $\Gamma \vdash \Phi \isavectx$ with morphisms given by
-    transformation substitutions $\Gamma \pipe \Phi \vdash \psi :
-    \Psi$.
-  \item Horizontal composition of contexts is defined by gluing along
-    a variable $\Phi\jnctx \Psi$ and the horizontal identity is given
-    by the single category variable $\alpha:\cat C$.
-  \item Objects of $C_\vdash^\Gamma$ are provided by the profunctors
-    $\Gamma\alpha:\cat C;\beta:\cat D \vdash R$, and a morphism from
-    $\alpha:\cat C;\beta:\cat D \vdash R$ to $\alpha':\cat
-    C';\beta':\cat D' \vdash R'$ is a triple $(a,t,b)$ of functor
-    terms $\alpha:\cat C \vdash a : \cat C'$ and $\beta:\cat D\vdash b
-    : \cat D'$ and a term $\alpha:\cat C,x:R,\beta:\cat D \vdash t :
-    R'[a/\alpha';b/\beta']$
-
-    Composition $(a,t,b) \circ (a',t',b') = (a[a'],t[t'],b[b'])$ is
-    given by component-wise single-place substitution with the triple
-    of all variables the identity.
-  \item The inclusion functor $i^\Gamma : C_d \to C_1$ is defined on
-    types as the singleton context $i(R) = \alpha:\cat
-    C,x:R,\beta:\cat D$ and on morphisms as the singleton
-    substitution.
-  \item The restriction operation $R \circ (a,b)$ is defined as the
-    substitution $R[a;b]$.
-  \item The remaining universal properties are satisfied using the
-    analogous type theoretic construction.
-  \end{enumerate}
-\end{construction}
-
-\section{Completeness}
-
-%% \begin{figure}
-%%   \begin{mathpar}
-%%       (\homrallXYtoZ \alpha R P)[\xi] = \homrallXYtoZ \alpha {R[d^+\xi;\alpha/\alpha]} {P[\alpha/\alpha;d^-\xi]}
-
-%%      (\homrlambdaXatYdotZ {x} {\alpha} M)[\phi]
-%%       = \homrlambdaXatYdotZ {x} {\alpha} M[\phi,x/x,\gamma/\gamma]
-
-%%      (\homllambdaXatYdotZ {x} {\alpha} M)[\phi]
-%%       = \homllambdaXatYdotZ {x} {\alpha} M[\phi,x/x,\gamma/\gamma]
-
-%%   \end{mathpar}
-%%   \caption{Substitution}
-%% \end{figure
-%% }
-
-In this section we establish the most significant portion of the
-completeness proof:
-\begin{construction}[Double category completeness]
-  For any $\Gamma$, define a strict double category as follows
-  \begin{enumerate}
-  \item Objects are categories $\Gamma \vdash \cat C \isaCat$
-  \item Vertical morphisms from $\cat C$ to $\cat D$ are functor/open
-    object terms $\Gamma \pipe\alpha:\cat C \vdash B : \cat D$ modulo
-    $\alpha$-equivalence.
-  \item Horizontal morphisms from $\cat C$ to $\cat D$ are profunctor
-    contexts $\Gamma\vdash \Phi \isavectx$ with $d^-\Phi = \cat C$ and
-    $d^+\Phi = \cat D$ modulo $\alpha$-equivalence.
-  \item Squares from $\Psi$ to $\Phi$ along $A$ and $B$ are
-    substitutions $\Gamma \vdash \phi : \Psi \to \Phi$ with $d^-\phi =
-    A$ and $d^+\phi = B$ modulo $\alpha$-equivalence.
-  \end{enumerate}
-\end{construction}
-\begin{proof}
-  The details are in the linked lemmas. Here we give an overview of
-  the relevant structure and properties.
-  \begin{enumerate}
-  \item Vertical composition is given by substitution of object terms
-    and is associative and unital with the variable as the identity
-    (Lemma \ref{lem:vertical-cat}).
-  \item Horizontal composition is given by the gluing of contexts
-    $\Phi\jnctx \Psi$ and is associative and unital with the single
-    category variable as the identity (Lemma
-    \ref{lem:horizontal-cat}).
-  \item Vertical composition of substitutions is given by substitution
-    $\phi[\psi]$ and is associative and unital with the identity
-    substitution given by a substitution of variables for themselves
-    (Lemma \ref{lem:vertical-sq-cat}).
-  \item Horizontal composition of substitutions is given by gluing of
-    substitutions $\phi \jnctx \psi$ and is associative and unital
-    with the identity substitution given by a single functor/open
-    object substitution (Lemma \ref{lem:horizontal-sq-cat}).
-  \item The process of constructing horizontal identity substitutions
-    is functorial in that the horizontal identity of a vertical
-    identity arrow is the vertical identity
-    (Lemma~\ref{lem:hor-vert-id-id}) and the vertical composition of
-    horizontal identities is a horizontal identity
-    (Lemma~\ref{lemma:lem-hor-sq-id}).
-  \item The process of horizontal composition of substitutions is
-    functorial in that the horizontal composition of vertical identity
-    squares is the identity (Lemma~\ref{lem:lem-vert-sq-id}) and the
-    interchange law holds (Lemma~\ref{lem:interchange}).
-  \end{enumerate}
-\end{proof}
 
@@
 \end{definition}
-
-\begin{lemma}
-  \label{lem:sq-vid-hid}
-  For any $\cat C$, $\alpha/\alpha : \alpha:\cat C \to \alpha:\cat C$
-  is both a horizontal and vertical identity.
-\end{lemma}
-
-\begin{lemma}
-  \label{lem:sq-vcomp-hid}
-  For any $\beta:\cat D \vdash a : \cat C$ and $\alpha:\cat C \vdash a': \cat C'$
-  the substitution
-  \[ (a'/\alpha')[a/\alpha] : \beta:\cat D \to \alpha':\cat C' \]
-  satisfies
-  \[ (a'/\alpha')[a/\alpha] = a'[a/\alpha]/\alpha' \]
-\end{lemma}
-
-\begin{lemma}
-  \label{lem:sq-vid-hcomp}
-  For any $\Phi$ and $\Psi$ with $\Phi\jnctx \Psi$ well-defined,
-  \[ \id_\Phi \jnctx \id_\Psi = \id_{\Phi\jnctx\Psi} \]
-  follows by induction.
-\end{lemma}
-
-\begin{lemma}{Interchange}
-  \label{lem:sq-vcomp-hcomp}
-  We prove the interchange law
-  $$(\phi\jnctx\psi)[\phi'\jnctx\psi'] = \phi[\phi'] \jnctx \psi[\psi']$$
-  by induction on $\psi$.
-
-  \begin{enumerate}
-  \item When $\psi=a/\alpha$
-      \begin{align*}
-    (\phi\jnctx a/\alpha)[\phi'\jnctx b/\beta]
-        &= \phi[\phi']\\
-        &= \phi[\phi'] \jnctx (a[b/\beta]/\alpha)\\
-        &= \phi[\phi'] \jnctx (a/\alpha)[b/\beta]
-  \end{align*}
-    \item When $\psi=\psi,t/x,a/\alpha$
-      \begin{align*}
-        (\phi\jnctx (\psi,t/x,a/\alpha))[\phi' \jnctx (\psi'_1\jnctx\psi'_2)]
-        &= ((\phi\jnctx \psi),t/x,a/\alpha)[(\phi' \jnctx \psi'_1)\jnctx\psi'_2]\\
-        &= (\phi\jnctx \psi)[\phi' \jnctx \psi'_1],t[\psi'_2]/x,a[d^+\psi'_2]/\alpha\\
-        &= (\phi[\psi'] \jnctx \psi[\psi'_1]),t[\psi'_2]/x,a[d^+\psi'_2]/\alpha\\
-        &= \phi[\psi'] \jnctx (\psi[\psi'_1],t[\psi'_2]/x,a[d^+\psi'_2]/\alpha)\\
-        &= \phi[\psi'] \jnctx (\psi,t/x,a/\alpha)[\psi'_1\jnctx \psi'_2]
-      \end{align*}
-  \end{enumerate}
-  Again, this covers all cases by \cref{inversion}.
-\end{lemma}
-
-\begin{figure}
-  \begin{mathpar}
-    % vertical arrows
-    \inferrule*[Right=VComp*]
-    {\alpha:\cat C \vdash b : \cat D \and
-     \beta:\cat D \vdash c : \cat E}
-    {\alpha:\cat C \vdash c[b/\beta] : \cat E}\\
-
-    \beta[b/\beta] = b\and
-    c[\beta/\beta] = c\and
-    c[b/\beta][a/\alpha] = c[b[a/\alpha]/\beta]\\
-
-    % horizontal arrows
-    \inferrule*[Right=HComp*]
-    {\Gamma \vdash \Phi \isavectx\\\\
-      \Gamma\vdash\Psi\isavectx\\\\
-      d^+\Phi = d^-\Psi}
-    {\Gamma \vdash \Phi \jnctx \Psi \isavectx}
-
-    d^+(\Phi \jnctx \Psi) = d^+\Psi\and
-    d^-(\Phi \jnctx \Psi) = d^-\Phi\\
-
-    \Phi \jnctx (d^+\Phi) = \Phi \and
-    (d^-\Phi) \jnctx \Phi = \Phi \and
-    (\Phi \jnctx \Psi) \jnctx \Sigma = \Phi \jnctx (\Psi \jnctx \Sigma)\\
-
-    % 2-cells
-    %% vertical composition
-    \inferrule*[Right=V2Comp*]
-    {\Sigma \vdash \psi :: \Psi \and
-     \Psi \vdash \phi :: \Phi}
-    {\Sigma \vdash \phi[\psi] :: \Phi}
-
-    \inferrule*[Right=V2Id*]
-    {}
-    {\Phi \vdash \id_\Phi :: \Phi}\\
-
-    \id_\Phi[\phi] = \phi\and
-    \phi[\id_\Psi] = \phi\and
-    \phi[\psi][\sigma] = \phi[\psi[\sigma]]\\
-
-    %% Horizontal composition
-    \inferrule*[Right=H2Comp*]
-    {\Phi' \vdash \phi :: \Phi \\\\
-     \Psi' \vdash \psi :: \Psi \\\\
-     d^+\phi = d^-\psi 
-    }
-    {\Phi' \jnctx \Psi' \vdash \phi \jnctx \psi :: \Phi \jnctx \Psi}
-
-    d^-(\phi\jnctx\psi) = d^-\phi\and
-    d^+(\phi\jnctx\psi) = d^+\phi\and
-
-    \phi\jnctx d^+\phi = \phi\and
-    d^-\phi \jnctx \phi = \phi\and
-    \phi\jnctx(\psi\jnctx\sigma) = (\phi\jnctx\psi)\jnctx\sigma
-
-    %% Interchange
-    \inferrule*[Right=Interchange]
-    {\Phi'' \vdash \phi' :: \Phi' \and
-     \Psi'' \vdash \psi' :: \Psi' \and
-     d^+\phi' = d^-\psi'\\\\
-     \Phi' \vdash \phi :: \Phi \and
-     \Psi' \vdash \psi :: \Psi \and
-     d^+\phi = d^-\psi
-    }
-    {(\phi\jnctx\psi)[\phi'\jnctx\psi'] = \phi[\phi'] \jnctx \psi[\psi']}\\
-
-    % Interaction with displays
-    \inferrule
-    {\Psi \vdash \phi :: \Phi\and
-     \Phi \vdash t : R[a/\alpha;b/\beta]}
-    {\Psi\vdash t[\phi] : R[a[d^-\phi]/\alpha;b[d^+\phi]/\beta]}
-
-    t[\id_\Phi] = t\and
-    t[\phi][\psi] = t[\phi[\psi]]
-  \end{mathpar}
-  \caption{Syntactic Double Category with Displays Structure}
-\end{figure}
-
-\section{Soundness}
-
-We next show how to uniquely interpret the profunctor contexts and
-substitutions in an arbitrary HED.
-
-\begin{proof}
-  We sketch only the interpretation of profunctor contexts in $C_1$,
-  as the rest is a straightforward translation from type theoretic to
-  categorical syntax.  We define the translation of a context
-  $\sem{\Phi}$ by induction on $\Phi$:
-
-  \begin{align*}
-    \sem{\alpha:\cat C} &= \sem{\id^h(\cat C)}\\
-    &= \id^h(\sem{\cat C})\\\\
-    \sem{\Phi , x:R, \beta:\cat C}\\
-    &= \sem{\Phi \circ_h i(R)}\\
-    &= \sem{\Phi} \sem{\Phi} \circ_h i(\sem{R})
-  \end{align*}
-  As these derivations show every step was uniquely determined by the
-  homomorphism property. Preservation of identity follows by
-  definition. Preservation of composition $\sem{\Phi \jnctx \Psi} =
-  \sem{\Phi} \circ_h \sem{\Psi}$ follows by induction on $\Psi$:
-  \begin{align*}
-    \sem{\Phi \jnctx (\alpha:\cat C)} &= \sem{\Phi} = \sem{\Phi} \circ_h \sem{\id(C)}\\
-    \sem{\Phi \jnctx (\Psi,x:R,\alpha:\cat C)}
-    &= \sem{(\Phi \jnctx \Psi),x:R,\alpha:\cat C} \\
-    &= \sem{(\Phi \jnctx \Psi)} \circ_h i(\sem{R})\\
-    &= (\sem{\Phi} \circ_h \sem{\Psi}) \circ_h i(\sem{R})\\
-    &= \sem{\Phi} \circ_h (\sem{\Psi} \circ_h i(\sem{R}))\\
-    &= \sem{\Phi} \circ_h \sem{\Psi, x:R, \alpha:\cat C}
-  \end{align*}
-  The interpretation of transformation substitutions is analogous.
-\end{proof}
 

### 2022-10-14 paper/paper.tex
bib

@@
 programming languages typically have sound and complete models given by
-notions of structured categories,\cite{lawvere,goguen,lambek,moggi}.
-%
-This Curry-Howard-Lambek correspondence~\cite{curry-howard-lambek}
-applies to simply typed lambda calculus~\cite{lambekscott},
+notions of structured categories,\cite{Lawvere69,LambekScott,moggi}.
+%
+This Curry-Howard-Lambek correspondence~
+applies to simply typed lambda calculus~\cite{LambekScott},
 computational lambda calculus~\cite{moggi}, linear logic~\cite{girard}
@@
 
-% TODO: do we have citations for this?
 While this research program has been quite successful,
@@
 parametricity and gradual
-typing\cite{order-enriched-cats,topos-of-trees,reflexive-graphs,double-cats-gradual-typing}.
+typing\cite{wand197913,topos-of-trees,reflexive-graphs,double-cats-gradual-typing}.
 
@@
 using algebraic structures such as 2-categories, bicategories,
-(virtual) double categories and pro-arrow equipments\cite{TODO}, an
+(virtual) double categories and pro-arrow equipments\cite{bicategories,proarrow-equipments,LEINSTER2002391,Cruttwell2010}, an
 approach known as \emph{formal category theory}.
@@
 as (hyperdoctrines of) \emph{virtual
-  equipments}~\cite{virtualequipments}, a particular semantic model of
+  equipments}~\cite{LEINSTER2002391,Cruttwell2010}, a particular semantic model of
 formal category theory, provide a sound and complete notion of model for
@@
 great deal of fundamental results and constructions in category
-theory~\cite{elements-of-oo-category-theory,
-  enriched-indexed-categories}.
+theory~\cite{riehl_verity_2022,Shulman13}.
 %
@@
 which in general does not.  Those familiar with linear
-logic~\cite{polakawpfenning} might expect a more general rule, with the
+logic~\cite{polakow-pfenning} might expect a more general rule, with the
 continuation $t$ typed in a context $\Phi_l \jnctx \Phi_r$ and $\Phi_l
@@
 which extends Martin-L\"of type theory with $\Pi$/$\Sigma$ and extensional
-identity types (with their standard rules)~\cite{emltt}. We use
+identity types (with their standard rules). We use
 extensional identity types so that the description of models is simpler,
@@
 for explicit naturality and functoriality side conditions.  We have
-checked some of the results in this section (e.g. Lemma~\ref{lem:yoneda}
+mechanized some of the results in this section (e.g. Lemma~\ref{lem:yoneda}
 and Lemma~\ref{lem:fubini-short} and the maps in
@@
 and predicate logic in certain \emph{virtual equipments}
-\cite{leinster,crutwell-shulman}, i.e., virtual double categories with
+\cite{LEINSTER2002391,Cruttwell2010}, i.e., virtual double categories with
 restrictions. We briefly recount the structure present in a virtual
-double category, but see \citet{crutwell-shulman} for a precise
+double category, but see \citet{Cruttwell2010} for a precise
 definition of the composition rules for 2-cells and functor of virtual
@@
 %
-Our notion here differs from \citet{crutwell-shulman} in several ways
+Our notion here differs from \citet{Cruttwell2010} in several ways
 to more closely match our judgmental structure: (1) we require that
@@
 Categories with families $\mathcal C$ model dependent type theory
-\citep{cwf} and for each semantic context $\Gamma$, $V^\Gamma$ models
+\citep{dybjer-cwf} and for each semantic context $\Gamma$, $V^\Gamma$ models
 the VETT judgments in context $\Gamma$, with the functoriality
@@
 universal properties of the same names, as described for instance in
-\citet{shulmanFrBicats}. The graph of a profunctor is modeled by
-\emph{tabulators} \citet{grandis-pare}. Finally, the covariant and
+\citet{Shulman08}. The graph of a profunctor is modeled by
+\emph{tabulators} \citet{grandis-pare99}. Finally, the covariant and
 contravariant presheaf categories can be described as a weakening of
-the definition of a Yoneda equipment from \citet{yosegi-boxes} to
+the definition of a Yoneda equipment from \citet{diliberti-loregian} to
 virtual double categories. More detailed descriptions of these
@@
 
-\begin{construction}[(\citet{shulman-enriched-indexed})]
+\begin{construction}[(\citet{Shulman13})]
   Given any functor $\mathcal V : S^o \to \SymMonCat$ such that $S$
@@
 
-Riehl and Verity~\cite{rvXXelements} use a formal language of virtual
+Riehl and Verity~\cite{riehl_verity_2022} use a formal language of virtual
 equipments to prove results valid for $\infty$-categories without
@@
 
-The coend calculus~\citet{coend-calculus} is an informal syntax for
+The coend calculus~\citet{loregian_2021} is an informal syntax for
 manipulating profunctors involving ends and coends; an extension of
@@
 
-\citet{djm-string-diagrams} provides a string diagram calculus for
+\citet{jazmyers-strings} provides a string diagram calculus for
 double categories and pro-arrow equipments, generalizing string
-diagrams for monoidal categories \citet{joyal?}. These are an
+diagrams for monoidal categories. These are an
 alternative approach to type theoretic calculi, with the string
@@
 \emph{Cartesian
-  bicategory}~\citet{carboni-walters,carboni-walters-someonelese}.  (In
-a terminological collision, \citet{fioresaville} give a syntax for a a
-bicategory with products/finite limits, but this is a different notion.)
-\citet{jonasfreycartesianbicategories} describe some preliminary work on
+  bicategory}~\citet{CARBONI198711}.
+\citet{frey} describe some preliminary work on
 a proof system for Cartesian bicateogires.  This calculus is more
@@
 Finally, some variations on double categories have been used to model
-the structure of certain program logics. \citet{newlicata20} present a
+the structure of certain program logics. \citet{double-cats-gradual-typing} present a
 logic for \emph{vertically thin} pro-arrow equipments, i.e., pro-arrow
@@
 rather than by syntactic substitution of vertical arrows in horizontal
-arrows. Another similar calculus is System P \citep{systemP} which is an
+arrows. Another similar calculus is System P \citep{dunphyreddy} which is an
 internal language of \emph{reflexive graph categories}, which are like
@@
 variables, in which case ordinary $\lambda$ calculus can be used to
-define functor categories as function types. Similarly, ideas from modal
-type theory/graded linear logic could be used for defining opposite
-categories \cite{mike-contravariance-enrichment,semiring-stuf??}.
+define functor categories as function types. Similarly, ideas from
+coeffects and enriched category theory could be used for defining
+opposite categories \cite{Shulman18,coeffect}.
 
@@
 theory as a quotient inductive inductive type
-\cite{altenkirch-kaposi-popl17}.
+\cite{qit}.
 

### 2022-10-14 paper/paper.tex
fix refs

@@
 \begin{construction}
+  \label{construction:synthetic-composition}
   For any small category $\cat C$, we can construct natural elements
@@
   \caption{Substitution}
-  \label{fig:substitions}
+  \label{fig:substitutions}
 \end{figure}
@@
   \end{align*}
-  This covers all cases by \cref{inversion}.
+  This covers all cases by \cref{lem:inversion}.
 
@@
 
-\begin{theorem}[Adjunctions]
-  $G$ is right adjoint to $F$ if and only if there is a natural isomorphism
-  $\punitinXfromYtoZ {\cat D} {(F \alpha)} \beta \cong \punitinXfromYtoZ {\cat D} \alpha {G \beta}$
-\end{theorem}
-\begin{proof}
-  First, by \cref{lem:co-unit}, the forward direction of the
-  isomorphism is equivalent to providing a term
-  \[ \eta : \pendallXdotY \alpha \punitinXfromYtoZ \alpha {G(F(\alpha)) } \]
-
-  Second, the backward direction of the isomorphism is equivalent to
-  providing a term
-  \[ \epsilon : \pendallXdotY \beta \punitinXfromYtoZ {F(G(\beta))} \beta \]
-
-  And the triangle laws state exactly that the induced functions are
-  isomorphisms.
-\end{proof}
-
-\subsection{Kan Extensions/Lifts}
-
-As a further example of the synthetic approach to universal
-properties, we consider right and left \emph{Kan extensions}, which
-are generalizations of limits and colimits, respectively.
-
-The right Kan extension of $F$ along $J$ is the ``best'' (maximal) way to
-fill in the following square:
-\[% https://tikzcd.yichuanshen.de/#N4Igdg9gJgpgziAXAbVABwnAlgFyxMJZABgBpiBdUkANwEMAbAVxiRAGEQBfU9TXfIRQBGclVqMWbACLdeIDNjwEiZYePrNWiDgHJu4mFADm8IqABmAJwgBbJGRA4ISAEzVNUnQCk5lm-aIjs5IohJabABiBlxAA
-\begin{tikzcd}
-\cat C \arrow[d, "J"] \arrow[r, "F"] & \cat D \\
-\cat C'                              &  
-\end{tikzcd}\]
-I.e. a functor $\textrm{Ran}_J F : \cat C' \to \cat D$ with a
-universal natural transformation $\textrm{Ran}_J F \circ J \Rightarrow
-F$. A left kan extension $\textrm{Lan}_J F$ is the same but with a
-natural transformation $F \Rightarrow \textrm{Lan}_J F \circ J $.
-
-We can express the concept in SCT as representing a certain profunctor
-constructed using the profunctor hom:
-\begin{definition}[Kan Extensions]
-  Given functors $F : \varr {\cat C}{\cat D}$ and $J : \varr {\cat
-    C}{\cat C'}$, a right Kan extension is a functor $\textrm{Ran}_J F
-  : \varr {\cat C'}{\cat D} $that right-represents the profunctor
-  \[ \hlambda {\beta:\cat D}{\alpha': \cat C} \homrallXYtoZ {\alpha}{(\punitinXfromYtoZ {\cat C'}{\alpha'}{J\alpha})} {\punitinXfromYtoZ {\cat D}{\beta}{F \alpha}} \]
-
-  A left Kan extensions is a functor $\textrm{Lan}_J F : \varr {\cat
-    C'}{\cat D}$ is a functor satisfying the dual property: it
-  \emph{left}-represents the profunctor
-  \[\hlambda {\alpha':\cat C}{\beta:\cat D} \homlallXYtoZ {\alpha}{(\punitinXfromYtoZ {\cat C'}{J\alpha}{\alpha'})} {\punitinXfromYtoZ {\cat D}{F \alpha}{\beta}}
-  \]
-\end{definition}
-
-The limit of a functor $F : \varr{\cat C}{\cat D}$ is a special case
-of a Kan extension where $J : \varr{\cat C}{1}$ is the unique functor
-to the terminal category. Similarly, the colimit of $F$ is the left
-Kan extension with respect to the same $J$.
-%
-Specializing to this case, the representability for limits means:
-\[ \punitinXfromYtoZ {\cat D}{\beta}{\textrm{lim} F} \cong \homrallXYtoZ {\alpha}{(\punitinXfromYtoZ {\cat 1}{\alpha'}{()})}{\punitinXfromYtoZ {\cat D}{\beta}{F \alpha}} \]
-And if we view a morphism in $\cat 1$ as being trivial, then this
-essentially states that mapping $\beta$ to $\textrm{lim} F$ is the
-same thing as constructing a cone: for all $\alpha :\cat C$, a
-morphism from $\beta$ to $F \alpha$, naturally in $\alpha$.
-
-The usual formulation of Kan extensions is in terms of a universal
-morphism.
-%
-We can reproduce this formulation using our established theory of
-equivalences for homs of profunctors.
-\begin{lemma}
-  One half of the representability isomorphism for a right Kan
-  extension $\textrm{Ran}_J F$ is determined by a natural
-  transformation
-  \[ \pendallXdotY \alpha \punitinXfromYtoZ {\cat D}{(\textrm{Ran}_JF)(J\alpha)}{F\alpha} \]
-
-  Dually, one half of the representability isomorphism for a left Kan
-  extension $\textrm{Lan}_J F$ is determined by a natural
-  transformation
-  \[ \pendallXdotY \alpha \punitinXfromYtoZ {\cat D}{F\alpha}{(\textrm{Lan}_JF)(J\alpha)} \]
-\end{lemma}
-\begin{proof}
-  By \cref{lem:unit-repr}, one half of representability is determined by a natural element
-  \[ \pendallXdotY {\alpha'} \homrallXYtoZ {\alpha}{(\punitinXfromYtoZ {\cat C'}{\alpha'}{J\alpha})} {\punitinXfromYtoZ {\cat D}{(\textrm{Ran}_J F)\alpha'}{F \alpha}} \]
-  which, by \cref{lem:fubini} is isomorphic to
-  \[
-  \pendallXdotY {\alpha} \homlallXYtoZ {\alpha'}{(\punitinXfromYtoZ {\cat C'}{\alpha'}{J\alpha})}{\punitinXfromYtoZ {\cat D}{(\textrm{Ran}_J F)\alpha'}{F \alpha}}
-  \]
-  which, by \cref{lem:yo-and-co-yo}, is isomorphic to 
-  \[ \pendallXdotY \alpha \punitinXfromYtoZ {\cat D}{(\textrm{Ran}_JF)(J\alpha)}{F\alpha} \]
-  by taking $P = \hlambda {\alpha'}{\alpha} {\punitinXfromYtoZ {\cat D}{(\textrm{Ran}_J F)\alpha'}{F \alpha}}$.
-
-  The proof for left Kan extensions follows by \cref{lem:duality}.
-\end{proof}
+%% \begin{theorem}[Adjunctions]
+%%   $G$ is right adjoint to $F$ if and only if there is a natural isomorphism
+%%   $\punitinXfromYtoZ {\cat D} {(F \alpha)} \beta \cong \punitinXfromYtoZ {\cat D} \alpha {G \beta}$
+%% \end{theorem}
+%% \begin{proof}
+%%   First, by \cref{lem:co-unit}, the forward direction of the
+%%   isomorphism is equivalent to providing a term
+%%   \[ \eta : \pendallXdotY \alpha \punitinXfromYtoZ \alpha {G(F(\alpha)) } \]
+
+%%   Second, the backward direction of the isomorphism is equivalent to
+%%   providing a term
+%%   \[ \epsilon : \pendallXdotY \beta \punitinXfromYtoZ {F(G(\beta))} \beta \]
+
+%%   And the triangle laws state exactly that the induced functions are
+%%   isomorphisms.
+%% \end{proof}
+
+%% \subsection{Kan Extensions/Lifts}
+
+%% As a further example of the synthetic approach to universal
+%% properties, we consider right and left \emph{Kan extensions}, which
+%% are generalizations of limits and colimits, respectively.
+
+%% The right Kan extension of $F$ along $J$ is the ``best'' (maximal) way to
+%% fill in the following square:
+%% \[% https://tikzcd.yichuanshen.de/#N4Igdg9gJgpgziAXAbVABwnAlgFyxMJZABgBpiBdUkANwEMAbAVxiRAGEQBfU9TXfIRQBGclVqMWbACLdeIDNjwEiZYePrNWiDgHJu4mFADm8IqABmAJwgBbJGRA4ISAEzVNUnQCk5lm-aIjs5IohJabABiBlxAA
+%% \begin{tikzcd}
+%% \cat C \arrow[d, "J"] \arrow[r, "F"] & \cat D \\
+%% \cat C'                              &  
+%% \end{tikzcd}\]
+%% I.e. a functor $\textrm{Ran}_J F : \cat C' \to \cat D$ with a
+%% universal natural transformation $\textrm{Ran}_J F \circ J \Rightarrow
+%% F$. A left kan extension $\textrm{Lan}_J F$ is the same but with a
+%% natural transformation $F \Rightarrow \textrm{Lan}_J F \circ J $.
+
+%% We can express the concept in SCT as representing a certain profunctor
+%% constructed using the profunctor hom:
+%% \begin{definition}[Kan Extensions]
+%%   Given functors $F : \varr {\cat C}{\cat D}$ and $J : \varr {\cat
+%%     C}{\cat C'}$, a right Kan extension is a functor $\textrm{Ran}_J F
+%%   : \varr {\cat C'}{\cat D} $that right-represents the profunctor
+%%   \[ \hlambda {\beta:\cat D}{\alpha': \cat C} \homrallXYtoZ {\alpha}{(\punitinXfromYtoZ {\cat C'}{\alpha'}{J\alpha})} {\punitinXfromYtoZ {\cat D}{\beta}{F \alpha}} \]
+
+%%   A left Kan extensions is a functor $\textrm{Lan}_J F : \varr {\cat
+%%     C'}{\cat D}$ is a functor satisfying the dual property: it
+%%   \emph{left}-represents the profunctor
+%%   \[\hlambda {\alpha':\cat C}{\beta:\cat D} \homlallXYtoZ {\alpha}{(\punitinXfromYtoZ {\cat C'}{J\alpha}{\alpha'})} {\punitinXfromYtoZ {\cat D}{F \alpha}{\beta}}
+%%   \]
+%% \end{definition}
+
+%% The limit of a functor $F : \varr{\cat C}{\cat D}$ is a special case
+%% of a Kan extension where $J : \varr{\cat C}{1}$ is the unique functor
+%% to the terminal category. Similarly, the colimit of $F$ is the left
+%% Kan extension with respect to the same $J$.
+%% %
+%% Specializing to this case, the representability for limits means:
+%% \[ \punitinXfromYtoZ {\cat D}{\beta}{\textrm{lim} F} \cong \homrallXYtoZ {\alpha}{(\punitinXfromYtoZ {\cat 1}{\alpha'}{()})}{\punitinXfromYtoZ {\cat D}{\beta}{F \alpha}} \]
+%% And if we view a morphism in $\cat 1$ as being trivial, then this
+%% essentially states that mapping $\beta$ to $\textrm{lim} F$ is the
+%% same thing as constructing a cone: for all $\alpha :\cat C$, a
+%% morphism from $\beta$ to $F \alpha$, naturally in $\alpha$.
+
+%% The usual formulation of Kan extensions is in terms of a universal
+%% morphism.
+%% %
+%% We can reproduce this formulation using our established theory of
+%% equivalences for homs of profunctors.
+%% \begin{lemma}
+%%   One half of the representability isomorphism for a right Kan
+%%   extension $\textrm{Ran}_J F$ is determined by a natural
+%%   transformation
+%%   \[ \pendallXdotY \alpha \punitinXfromYtoZ {\cat D}{(\textrm{Ran}_JF)(J\alpha)}{F\alpha} \]
+
+%%   Dually, one half of the representability isomorphism for a left Kan
+%%   extension $\textrm{Lan}_J F$ is determined by a natural
+%%   transformation
+%%   \[ \pendallXdotY \alpha \punitinXfromYtoZ {\cat D}{F\alpha}{(\textrm{Lan}_JF)(J\alpha)} \]
+%% \end{lemma}
+%% \begin{proof}
+%%   By \cref{lem:unit-repr}, one half of representability is determined by a natural element
+%%   \[ \pendallXdotY {\alpha'} \homrallXYtoZ {\alpha}{(\punitinXfromYtoZ {\cat C'}{\alpha'}{J\alpha})} {\punitinXfromYtoZ {\cat D}{(\textrm{Ran}_J F)\alpha'}{F \alpha}} \]
+%%   which, by \cref{lem:fubini} is isomorphic to
+%%   \[
+%%   \pendallXdotY {\alpha} \homlallXYtoZ {\alpha'}{(\punitinXfromYtoZ {\cat C'}{\alpha'}{J\alpha})}{\punitinXfromYtoZ {\cat D}{(\textrm{Ran}_J F)\alpha'}{F \alpha}}
+%%   \]
+%%   which, by \cref{lem:yo-and-co-yo}, is isomorphic to 
+%%   \[ \pendallXdotY \alpha \punitinXfromYtoZ {\cat D}{(\textrm{Ran}_JF)(J\alpha)}{F\alpha} \]
+%%   by taking $P = \hlambda {\alpha'}{\alpha} {\punitinXfromYtoZ {\cat D}{(\textrm{Ran}_J F)\alpha'}{F \alpha}}$.
+
+%%   The proof for left Kan extensions follows by \cref{lem:duality}.
+%% \end{proof}
 

### 2022-10-14 paper/paper.tex
detailed proof of Yoneda

@@
   \begin{enumerate}
-  \item Identity: $\pendallXdotY {\alpha} {\punitinXfromYtoZ {\cat C}{\alpha}{\alpha'}}$
+  \item Identity: $\pendallXdotY {\alpha} {\punitinXfromYtoZ {\cat C}{\alpha}{\alpha}}$
   \item Composition: $\hombinary {\alpha_1} {\punitinXfromYtoZ {}{\alpha_1}{\alpha_2}} {\alpha_2} {\punitinXfromYtoZ {}{\alpha_1}{\alpha_2}} {\alpha_3} {\punitinXfromYtoZ {}{\alpha_1}{\alpha_3}}$
@@
 Next, we turn to some of the central theorems of category theory, the
-Yoneda and Co-Yoneda lemmas. Despite being ultimately quite elementary,
-these are notoriously abstract. In \ohol{}, we can view these as
-ordered generalizations of some very simple tautologies about
-equality. For instance, the Yoneda lemma generalizes
-the equivalence between the formulae $\forall x. \forall y. x = y
-\Rightarrow P y$ and $P x$.
+Yoneda and Co-Yoneda lemmas. Despite being ultimately quite
+elementary, these are notoriously abstract. In \ohol{}, we view
+these as ordered generalizations of some very simple tautologies about
+equality. For instance, the Yoneda lemma generalizes the equivalence
+between the formulae $\forall x. \forall y. x = y \Rightarrow P y$ and
+$P x$.
 \begin{lemma} \label{lem:yoneda}
-  Let $\alpha^o : \cat C$ and $\pi : \negPresheaf {\cat C}$. Then
+  Let $\alpha : \cat C$ and $\pi : \posPresheaf {\cat C}$. Then
   \begin{enumerate}
-  \item (Yoneda) The profunctor $\homrallXYtoZ {\alpha'} {\punitinXfromYtoZ {\cat C} {\alpha'}{\alpha}} {\negPresheafAppPtoX \pi {\alpha'}}$ is isomorphic to $\negPresheafAppPtoX \pi \alpha$
-  \item (Co-Yoneda) The profunctor $\tensorexistsXwithYandZ {\alpha'}{\punitinXfromYtoZ{}{\alpha}{\alpha'}} {\negPresheafAppPtoX \pi \alpha}$ is isomorphic to $\negPresheafAppPtoX \pi \alpha$
+  \item (Yoneda) The profunctor 
+    $\homrallXYtoZ {\alpha'} {\punitinXfromYtoZ {\cat C} {\alpha}{\alpha'}} {\posPresheafAppPtoX \pi {\alpha'}}$ is isomorphic to $\posPresheafAppPtoX \pi \alpha$
+  \item (Co-Yoneda) The profunctor $\tensorexistsXwithYandZ {\alpha'}{\posPresheafAppPtoX \pi \alpha'} {\punitinXfromYtoZ{}{\alpha'}{\alpha}}$ is isomorphic to $\posPresheafAppPtoX \pi \alpha$
   \end{enumerate}
@@
 The proofs both follow from the unit elimination rule, which is
-essential the Yoneda lemma---the two cases of showing (1) is an
+essentialy the Yoneda lemma---the two cases of showing (1) is an
 isomorphism are precisely the $\beta$ and $\eta$ rules for the unit.
@@
     {(\punitelimkontZatABC {\alpha. t} a {\punitrefl a} a) = t[a/\alpha]}
+
     \inferrule*[right=Unit$\eta$]
-    {\Gamma\pipe\Phi \vdash s : \punitinXfromYtoZ {\cat C}{a}{b}}
-    {\Gamma\pipe\Phi\vdash s = \punitelimkontZatABC {\alpha. \punitrefl \alpha} a s b: \punitinXfromYtoZ {\cat C}{a}{b}}
+    {\Gamma\pipe \alpha_1:\cat C,z:\punitinXfromYtoZ {\cat C}{\alpha_1}{\alpha_2},\alpha_2:\cat C \vdash s : R}
+    {\Gamma\pipe\alpha_1:\cat C,z:\punitinXfromYtoZ {\cat C}{\alpha_1}{\alpha_2},\alpha_2:\cat C\vdash s = \punitelimkontZatABC {\alpha.
+s[\alpha/\alpha_1;\punitrefl{\alpha}/z,\alpha/\alpha_2]} {\alpha_1} z {\alpha_2}: R}
 
@@
     {\tensorelimWkontZ {\tensorintroatXwithYandZ b s t} {x.\beta.y. r} = t[s/x;b/\beta;t/y]}
+
     \inferrule*[right=Tensor$\eta$]
-    {\Gamma\pipe\Phi \vdash s : \tensorexistsXwithYandZ \beta P Q}
-    {\Gamma\pipe\Phi\vdash s = \tensorelimWkontZ s {x.\beta.y. \tensorintroatXwithYandZ \beta x y}: \tensorexistsXwithYandZ \beta P Q}
+    {\Gamma\pipe\Phi,z:\tensorexistsXwithYandZ \beta P Q, \Psi \vdash s : R}
+    {\Gamma\pipe\Phi,z:\tensorexistsXwithYandZ \beta P Q, \Psi \vdash s = \tensorelimWkontZ s {x.\beta.y. s[\tensorintroatXwithYandZ \beta x y/z]}: \tensorexistsXwithYandZ \beta P Q}
 
@@
   {\Phi,x:\punitinXfromYtoZ {}{\alpha_-} {\alpha_+}, \Psi \vdash s : P}
-  {\Phi,x:\punitinXfromYtoZ {}{\alpha_-} {\alpha_+}, \Psi \vdash s = \textrm{ind}_{\to}^{\Phi;\Psi}(\alpha. s[\punitrefl \alpha/x]; x){}}
+  {\Phi,x:\punitinXfromYtoZ {}{\alpha_-} {\alpha_+}, \Psi \vdash s = \textrm{ind}_{\to}^{\Phi;\Psi}(\alpha. s[\alpha/\alpha_-,\punitrefl \alpha/x,\alpha/\alpha_+]; x) : P}
 \end{mathpar}
@@
   \begin{itemize}
-  \item $\Phi = \alpha_- $ and $\Psi = \alpha_+$: immediate by unit $\eta$.
+  \item $\Phi = \alpha_- $ and $\Psi = \alpha_+$:
+    \[ \textrm{ind}_{\to}^{\alpha_-;\alpha_+}(\alpha. s[\alpha/\alpha_-,\punitrefl \alpha/x,\alpha/\alpha_+]; x)
+    = \punitelimkontZatABC {\alpha. s[\alpha/\alpha_-,\punitrefl \alpha/x,\alpha/\alpha_+]} {\alpha_1}{x}{\alpha_2}
+    \]
+    Which is equal to $s$ by the primitive unit $\eta$.
   \item $\Phi = \alpha_-$ and $\Psi = \Psi, y, \beta$:
@@
 theory constructions and theorems from Section~\ref{sec:examples}.
+
+\begin{definition}[Synthetic Composition/Functoriality]
+  We provide the definitions of the terms in \cref{construction:synthetic-composition}
+  \begin{enumerate}
+  \item Identity $\id = \pendlambdaXdotY \alpha \punitrefl \alpha$
+  \item Composition $\textrm{comp} = \lambdaunary {\alpha_1} f {\alpha_2} \punitelimkontZatABC {\alpha. \homrallXYtoZ {\alpha_3} g g} {\alpha_1} f {\alpha_2}$
+  \item Functoriality $\textrm{fctor}(F) = \lambdaunary {\alpha_1} f {\alpha_2} \punitelimkontZatABC {\alpha. \punitrefl{F\,\alpha}} {\alpha_1} f {\alpha_2}$
+  \item Profunctoriality $\textrm{prof}(R) = \lambdaunary {\alpha_1} f {\alpha_2} {\punitelimkontZatABC {\alpha.
+\homrlambdaXatYdotZ r {\beta_1} \homrlambdaXatYdotZ g {\beta_2} {\homlappXtoYatZ {(\punitelimkontZatABC {\beta.
+\homllambdaXatYdotZ r \alpha r}{\beta_1}{g}{\beta_2})} {r} {\alpha}}}{\alpha_1} f {\alpha_2}}$
+
+    We can also define left and right composition for profunctors by
+    applying the profunctorial action to a reflexivity on one side or
+    the other:
+
+    \[ \textrm{lcomp}(R) = \lambdabinary {\alpha_1} f {\alpha_2} r {\beta} {\apptrinary {\textrm{prof}(R)} {\alpha_1} f {\alpha_2} r {\beta} {\punitrefl \beta} {\beta}}\]
+
+    \[ \textrm{rcomp}(R) = \lambdabinary {\alpha} r {\beta_1} g {\beta_2} {\apptrinary {\textrm{prof}(R)}  {\alpha} {\punitrefl \alpha} {\alpha} r {\beta_1} g {\beta_2}}\]
+  \end{enumerate}
+  Associativity and unit folllow by $\beta\eta$ for unit and homs.
+\end{definition}
+
+\begin{lemma}[Naturality]
+  For any $t : \pendallXdotY {\alpha:\cat C}{\harrapp R \alpha \alpha}$,
+  \[ \lambdaunary {\alpha_1} f {\alpha_2} {\appbinary {\textrm{lcomp}(R)}{\alpha_1} f {\alpha_2} {\pendappXtoY t {\alpha_2}} {\alpha_2}}
+  =  \lambdaunary {\alpha_1} f {\alpha_2} {\appbinary {\textrm{rcomp}(R)}{\alpha_1}  {\pendappXtoY t {\alpha_1}} {\alpha_1} f {\alpha_2}}\]
+\end{lemma}
+\begin{proof}
+  Expanding the definitions and applying $\beta$ reductions, both are equal to $\lambdaunary {\alpha_1} f {\alpha_2} \punitelimkontZatABC {\alpha. \pendappXtoY t \alpha} {\alpha_1} f {\alpha_2}$
+\end{proof}
+
+\begin{lemma}[Yoneda, Co-Yoneda]
+  Let $\alpha^o : \cat C$ and $\pi : \posPresheaf {\cat C}$. Then
+  \item (Yoneda) The profunctor $\homrallXYtoZ {\alpha'} {\punitinXfromYtoZ {\cat C} {\alpha'}{\alpha}} {\negPresheafAppPtoX \pi {\alpha'}}$ is isomorphic to $\negPresheafAppPtoX \pi \alpha$
+  \item (Co-Yoneda) The profunctor $\tensorexistsXwithYandZ {\alpha'}{\punitinXfromYtoZ{}{\alpha}{\alpha'}} {\negPresheafAppPtoX \pi \alpha}$ is isomorphic to $\negPresheafAppPtoX \pi \alpha$
+\end{lemma}
+\begin{proof}
+  We show Yoneda in detail.
+  \begin{itemize}
+  \item The left-to-right homomorphism is defined as
+    \[ M = \lambdaunary {\alpha} {\phi} \pi {\homrappXtoYatZ {\phi} {\punitrefl {\alpha}} \alpha} \]
+  \item The right-to-left homomorphism is defined as
+    \[ N = \lambdabinary \pi x \alpha f \alpha' {\homlappXtoYatZ {\punitelimkontZatABC {\alpha. \homllambdaXatYdotZ x \pi x} {\alpha'} f {\alpha}} x \pi}\]
+  \item First, right-to-left-to-right:
+    \begin{align*}
+      \lambdaunary \pi x \alpha {\appunary M {\alpha} {(\appunary N \pi x \alpha)} \pi}
+      &= \lambdaunary \pi x \alpha {\appunary M {\alpha} {({\homrlambdaXatYdotZ f {\alpha'} {{\homlappXtoYatZ {\punitelimkontZatABC {\alpha.
+\homllambdaXatYdotZ x \pi x} {\alpha'} f {\alpha}} x \pi}}})} \pi}\\
+      &= \lambdaunary \pi x \alpha {\homrappXtoYatZ {({\homrlambdaXatYdotZ f {\alpha'} {{\homlappXtoYatZ {\punitelimkontZatABC {\alpha.
+\homllambdaXatYdotZ x \pi x} {\alpha'} f {\alpha}} x \pi}}})} {\punitrefl \alpha} \alpha}\\
+      &= \lambdaunary \pi x \alpha {{\homlappXtoYatZ {\punitelimkontZatABC {\alpha. \homllambdaXatYdotZ x \pi x} {\alpha'} {\punitrefl \alpha} {\alpha}} x \pi}}\tag{covhom$\beta$}\\
+      &= \lambdaunary \pi x \alpha {{\homlappXtoYatZ {(\homllambdaXatYdotZ x \pi x)} x \pi}}\tag{unit$\beta$}\\
+      &= \lambdaunary \pi x \alpha {x}\tag{contrahom$\beta$}\\
+    \end{align*}
+  \item Left-to-right-to-left
+    \begin{align*}
+      &\lambdaunary {\alpha} {\phi} \pi \appunary N {\pi} {({\appunary M \alpha \phi \pi})} {\alpha}\\
+      &= \lambdaunary {\alpha} {\phi} \pi \homrlambdaXatYdotZ {f}{\alpha'} {\homlappXtoYatZ {\punitelimkontZatABC {\alpha.
+\homllambdaXatYdotZ x \pi x} {\alpha'} f {\alpha}} {({\appunary M \alpha \phi \pi})} \pi}\\
+      &= \lambdaunary {\alpha} {\phi} \pi \homrlambdaXatYdotZ {f}{\alpha'} {\homlappXtoYatZ {\punitelimkontZatABC {\alpha.
+\homllambdaXatYdotZ x \pi x} {\alpha'} f {\alpha}} {(\appunary \phi \alpha {\punitrefl\alpha}{\alpha})} \pi}\\
+      &= \lambdaunary {\alpha} {\phi} \pi \homrlambdaXatYdotZ {f}{\alpha'} \punitelimkontZatABC {\alpha.
+({\homlappXtoYatZ {\punitelimkontZatABC {\alpha.
+\homllambdaXatYdotZ x \pi x} {\alpha} {\punitrefl \alpha} {\alpha}} {(\appunary \phi \alpha {\punitrefl\alpha}{\alpha})} \pi})} {\alpha'}{f}{\alpha}\tag{unit $\eta$}
+      \\ 
+      &= \lambdaunary {\alpha} {\phi} \pi \homrlambdaXatYdotZ {f}{\alpha'} \punitelimkontZatABC {\alpha.
+({\homlappXtoYatZ {\homllambdaXatYdotZ x \pi x} {(\appunary \phi \alpha {\punitrefl\alpha}{\alpha})} \pi})} {\alpha'}{f}{\alpha}\tag{unit $\beta$} \\
+      &= \lambdaunary {\alpha} {\phi} \pi \homrlambdaXatYdotZ {f}{\alpha'} \punitelimkontZatABC {\alpha.
+({{(\appunary \phi \alpha {\punitrefl\alpha}{\alpha})}})} {\alpha'}{f}{\alpha}\tag{contrahom $\beta$} \\
+      &= \lambdaunary {\alpha} {\phi} \pi \homrlambdaXatYdotZ {f}{\alpha'} {{(\appunary \phi {\alpha'} f {\alpha})}}\tag{unit $\eta$} \\ 
+      &= \lambdaunary {\alpha} {\phi} \pi \phi \tag{covhom $\eta$} \\ 
+    \end{align*}
+  \end{itemize}
+\end{proof}
 

### 2022-10-14 paper/paper.tex
fix overflow
fubini examples done

@@
 
-\title{A Constructive Ordered Predicate Logic \\ for Category Theory}
+\title{A Formal Logic for Formal Category Theory}
 \author{Max S. New\inst{1,2} and Daniel R. Licata\inst{2}}
@@
   \item $\tensorexistsXwithYandZ {\beta} {\harrapp P \alpha \beta} {(\tensorexistsXwithYandZ \gamma {\harrapp Q \beta \gamma} {\harrapp R \gamma \delta})} \cong \tensorexistsXwithYandZ {\gamma} {(\tensorexistsXwithYandZ \beta {\harrapp P \alpha \beta} {\harrapp Q \beta \gamma})} {\harrapp R \gamma \delta}$
-  \item $\homrallXYtoZ {\gamma} {(\tensorexistsXwithYandZ \beta {\harrapp P \alpha \beta} {\harrapp Q \beta \gamma})} {\harrapp S \alpha \gamma} \cong \homrallXYtoZ \beta {\harrapp P \alpha \beta} {\homrallXYtoZ \gamma {\harrapp Q \beta \gamma} {\harrapp S \alpha \gamma}}$
-  \item $\homlallXYtoZ {\alpha} {(\tensorexistsXwithYandZ \beta {\harrapp P \alpha \beta} {\harrapp Q \beta \gamma})} {\harrapp S \alpha \gamma} \cong \homlallXYtoZ \beta {\harrapp Q \beta \gamma} {\homlallXYtoZ \alpha {\harrapp P \alpha \beta} {\harrapp S \alpha \gamma}}$
-  \item $\homrallXYtoZ \gamma {\harrapp Q \beta \gamma} {(\homlallXYtoZ \alpha {\harrapp P \alpha \beta} {\harrapp S \alpha \gamma})} \cong \homlallXYtoZ \alpha {\harrapp P \alpha \beta} {(\homrallXYtoZ \gamma {\harrapp Q \beta \gamma} {\harrapp S \alpha \gamma})}$
-  \item $\pendallXdotY \alpha \homrallXYtoZ \beta P {P'} \cong \pendallXdotY \beta \homlallXYtoZ \alpha P {P'}$
+  \item $\homrallXYtoZ {\gamma} {(\tensorexistsXwithYandZ \beta {\harrapp P \delta \beta} {\harrapp Q \beta \gamma})} {\harrapp S \alpha \gamma} \cong \homrallXYtoZ \beta {\harrapp P \delta \beta} {\homrallXYtoZ \gamma {\harrapp Q \beta \gamma} {\harrapp S \alpha \gamma}}$
+  \item $\homlallXYtoZ {\gamma} {(\tensorexistsXwithYandZ \beta {\harrapp P \gamma \beta} {\harrapp Q \beta \alpha})} {\harrapp S \gamma \delta} \cong \homlallXYtoZ \beta {\harrapp Q \beta \alpha} {\homlallXYtoZ \gamma {\harrapp P \gamma \beta} {\harrapp S \gamma \delta}}$
+  \item $\homrallXYtoZ \gamma {\harrapp Q \delta \gamma} {(\homlallXYtoZ \beta {\harrapp P \beta \alpha} {\harrapp S \beta \gamma})} \cong \homlallXYtoZ \beta {\harrapp P \beta \alpha} {(\homrallXYtoZ \gamma {\harrapp Q \delta \gamma} {\harrapp S \beta \gamma})}$
+  \item $\pendallXdotY \alpha \homrallXYtoZ \beta {\harrapp P \alpha \beta} {\harrapp Q \alpha \beta} \cong \pendallXdotY \beta \homlallXYtoZ \alpha {\harrapp P \alpha \beta} {\harrapp Q \alpha \beta}$
   \end{enumerate}  
@@
 
-Next, we provide some futher detailed examples of the formal category
-theory constructions and theorems from Section~\ref{sec:examples}.
+Next, we provide some futher details for some of the examples of the
+formal category theory constructions and theorems from
+Section~\ref{sec:examples}.
 
 \begin{definition}[Synthetic Composition/Functoriality]
-  We provide the definitions of the terms in \cref{construction:synthetic-composition}
+  We provide the definitions of the terms in Construction~\ref{construction:synthetic-composition}
   \begin{enumerate}
@@
   \item Functoriality $\textrm{fctor}(F) = \lambdaunary {\alpha_1} f {\alpha_2} \punitelimkontZatABC {\alpha. \punitrefl{F\,\alpha}} {\alpha_1} f {\alpha_2}$
-  \item Profunctoriality $\textrm{prof}(R) = \lambdaunary {\alpha_1} f {\alpha_2} {\punitelimkontZatABC {\alpha.
+  \item Profunctoriality
+    \[\textrm{prof}(R) = \lambdaunary {\alpha_1} f {\alpha_2} {\punitelimkontZatABC {\alpha.
 \homrlambdaXatYdotZ r {\beta_1} \homrlambdaXatYdotZ g {\beta_2} {\homlappXtoYatZ {(\punitelimkontZatABC {\beta.
-\homllambdaXatYdotZ r \alpha r}{\beta_1}{g}{\beta_2})} {r} {\alpha}}}{\alpha_1} f {\alpha_2}}$
+\homllambdaXatYdotZ r \alpha r}{\beta_1}{g}{\beta_2})} {r} {\alpha}}}{\alpha_1} f {\alpha_2}}\]
 
@@
 
+\begin{lemma}[Fubini]
+  We show two of the Fubini cases in detail:
+
+  \begin{enumerate}
+  \item $\homlallXYtoZ {\gamma} {(\tensorexistsXwithYandZ \beta {\harrapp P \gamma \beta} {\harrapp Q \beta \alpha})} {\harrapp S \gamma \delta} \cong \homlallXYtoZ \beta {\harrapp Q \beta \alpha} {\homlallXYtoZ \gamma {\harrapp P \gamma \beta} {\harrapp S \gamma \delta}}$
+  \item $\pendallXdotY \alpha \homrallXYtoZ \beta {\harrapp P \alpha \beta} {\harrapp Q \alpha \beta} \cong \pendallXdotY \beta \homlallXYtoZ \alpha {\harrapp P \alpha \beta} {\harrapp Q \alpha \beta}$
+  \end{enumerate}
+\end{lemma}
+\begin{proof}
+  \begin{enumerate}
+  \item This is a form of Currying isomorphism, as the $\lambda$ term
+    makes clear:
+    \begin{itemize}
+    \item Left to Right
+      \[ \lambdaunary \alpha h \delta \homllambdaXatYdotZ q \beta \homllambdaXatYdotZ p \gamma \homlappXtoYatZ h {(p,\beta,q)} \gamma  \]
+    \item Right to Left
+      \[ \lambdaunary {\alpha} k \delta \homllambdaXatYdotZ w \gamma \tensorelimWkontZ {w} {p.\beta.q. \homlappXtoYatZ {\homlappXtoYatZ k p \gamma} q \beta } \]
+    \item Left to Right to Left
+      \begin{align*}
+        &\lambdaunary \alpha h \delta \homllambdaXatYdotZ w \gamma {\tensorelimWkontZ {w} {p.\beta.q.
+(\homlappXtoYatZ {\homlappXtoYatZ {(\homllambdaXatYdotZ q \beta \homllambdaXatYdotZ p \gamma {\homlappXtoYatZ h {(p,\beta,q)} \gamma})} q \beta }  p \gamma) }}\\
+        &=\lambdaunary \alpha h \delta \homllambdaXatYdotZ w \gamma {\tensorelimWkontZ {w} {p.\beta.q.
+(\homlappXtoYatZ {{(\homllambdaXatYdotZ p \gamma {\homlappXtoYatZ h {(p,\beta,q)} \gamma})} }  p \gamma) }}\tag{contrahom$\beta$}\\
+        &=\lambdaunary \alpha h \delta \homllambdaXatYdotZ w \gamma {\tensorelimWkontZ {w} {p.\beta.q. {{({\homlappXtoYatZ h {(p,\beta,q)} \gamma})} } }}\tag{contrahom$\beta$}\\
+        &=\lambdaunary \alpha h \delta \homllambdaXatYdotZ w \gamma \homlappXtoYatZ h w \gamma\tag{tensor$\eta$}\\
+        &=\lambdaunary \alpha h h\tag{contrahom$\eta$}
+      \end{align*}
+    \item Right to Left to Right
+      \begin{align*}
+        &\lambdaunary \alpha k \delta \homllambdaXatYdotZ q \beta \homllambdaXatYdotZ p \gamma \homlappXtoYatZ {(\homllambdaXatYdotZ w \gamma {\tensorelimWkontZ {w} {p.\beta.q.
+\homlappXtoYatZ {\homlappXtoYatZ k q \beta}  p \gamma }})} {(p,\beta,q)} \gamma\\
+        &=\lambdaunary \alpha k \delta \homllambdaXatYdotZ q \beta \homllambdaXatYdotZ p \gamma { {\tensorelimWkontZ { {(p,\beta,q)}} {p.\beta.q.
+\homlappXtoYatZ {\homlappXtoYatZ k q \beta } p \gamma }}}\\
+        &=\lambdaunary \alpha k \delta \homllambdaXatYdotZ q \beta \homllambdaXatYdotZ p \gamma {\homlappXtoYatZ {\homlappXtoYatZ k q \beta } p \gamma}\\
+        &=\lambdaunary \alpha k \delta \homllambdaXatYdotZ q \beta { {\homlappXtoYatZ k q \beta}}\\
+        &=\lambdaunary \alpha k \delta k\\
+      \end{align*}
+    \end{itemize}
+  \item This isomorphism relates left and right homs. Unlike the
+    previous cases, the isomorphism is of types, not sets/profunctors.
+    \begin{itemize}
+    \item Left to right
+      \[\lambda X. \pendlambdaXdotY \beta \homllambdaXatYdotZ p \alpha {\appunary X \alpha p \beta}\]
+    \item Right to left
+      \[ \lambda Y. \pendlambdaXdotY \alpha \homrlambdaXatYdotZ p \beta {\homlappXtoYatZ {\pendappXtoY Y \beta} p \alpha} \]
+    \item Left to right to left
+      \begin{align*}
+        &\lambda X.
+\pendlambdaXdotY \alpha \homrlambdaXatYdotZ p \beta {\homlappXtoYatZ {\pendappXtoY {(\pendlambdaXdotY \beta \homllambdaXatYdotZ p \alpha {\appunary X \alpha p \beta})} \beta} p \alpha}\\
+        &= \lambda X. \pendlambdaXdotY \alpha \homrlambdaXatYdotZ p \beta {\homlappXtoYatZ { {(\homllambdaXatYdotZ p \alpha {\appunary X \alpha p \beta})}} p \alpha}\tag{nat.elt.$\beta$}\\
+        &= \lambda X. \pendlambdaXdotY \alpha \homrlambdaXatYdotZ p \beta {\appunary X \alpha p \beta}\tag{contrahom$\beta$}\\
+        &= \lambda X. \pendlambdaXdotY \alpha {\pendappXtoY X \alpha}\tag{contrahom$\eta$}\\
+        &= \lambda X. X\tag{nat.elt.$\eta$}\\
+      \end{align*}
+    \item The other case is similar.
+    \end{itemize}
+  \end{enumerate}
+\end{proof}
+
 \section{Semantics}
@@
 
-\section{Extended Synthetic Category Theory}
-
-Next, we look at synthetic category theory constructions in SCT.
-
-\subsection{Profunctor Isomorphisms}
-
-We start our study of synthetic category theory constructions with
-some isomorphisms that show the interaction between the 5
-multiplicative connectives: right/left hom, end, unit and tensor.
-
-First, we fix some terminology
-\begin{lemma}
-  Let $P, Q : \harr {\cat C} {\cat D}$.
-  A homomorphism from $P$ to $Q$ is a term of type
-  \[ \homunary {\alpha} {\harrapp P \alpha \beta} {\beta} {\harrapp Q \alpha \beta} \]
-  Given homomorphisms $M : P \multimap Q$, $N : Q \multimap R$ we can define their composite as
-  \[ (N \circ M) = \lambdaunary {\alpha} {x:\harrapp P \alpha \beta} \beta {\appunary \alpha N \beta {\appunary \alpha M \beta x}} \]
-  This notion of composition is associative and unital using the
-  $\beta\eta$ equations for the end/right hom connectives, with unit
-  $\lambdaunary {\alpha} x \beta x$.
-
-  An isomorphism between $P, Q : \harr {\cat C}{\cat D}$ is then a
-  homomorphism $P \multimap Q$ with an inverse $Q \multimap P$.
-\end{lemma}
-
-Our first synthetic theorem is that we have that the unit, i.e., the
-hom set, is the unit for tensor and hom.
-%
-The theorem for hom is the synthetic \emph{Yoneda lemma} and the
-theorem for tensor is the synthetic \emph{Co-Yoneda lemma}.
-\begin{lemma}[{Yoneda and Co-Yoneda}]
-  \label{lem:yo-and-co-yo}
-  Let $P : \harr {\cat C} {\cat D}$.
-
-  For any $\gamma : \cat E \vdash b : \cat D$, the following are naturally isomorphic to $\harrapp P \alpha b$:
-  \begin{enumerate}
-  \item $\homrallXYtoZ {\beta} {(\punitinXfromYtoZ {D} {\beta} {b})} {\harrapp P {\alpha} {\beta}}$
-  \item $\tensorexistsXwithYandZ {\beta} {\harrapp P \alpha {\beta}} {(\punitinXfromYtoZ {\cat D} {\beta} {b})}$
-  \end{enumerate}
-
-  For any $\gamma : \cat E \vdash a : \cat C$, the following are naturally isomorphic to $\harrapp P a \beta$:
-  \begin{enumerate}
-  \item $\homlallXYtoZ {{\alpha}} {(\punitinXfromYtoZ {C} {\alpha} {a})} {\harrapp P {{\alpha}} {\beta}}$
-  \item $\tensorexistsXwithYandZ {\alpha} {(\punitinXfromYtoZ {C} {a} {\alpha})} {\harrapp P {\alpha}{\beta}}$
-  \end{enumerate}
-\end{lemma}
-\begin{proof}
-  TODO: what's better, build the substitution in or prove the
-  substitution can be built in afterwards? The one building in the
-  substitution obscures the connection to the normal tensor/hom/unit
-  relationship, but the one with the built in subst is obviously more
-  general.
-\end{proof}
-
-Next, we have various associativities between the homs, tensor and end
-constructions. Semantically, these are related to so-called
-\emph{Fubini} theorems about interchange of (co)-ends.
-\begin{lemma}[{Fubini}]
-  \label{lem:fubini}
-  The following isomorphisms hold for any $P : \harr {\cat C}{\cat D},
-  P' : \harr {\cat C}{\cat D} Q :\harr {\cat D}{\cat E}, R : \harr
-  {\cat E} {\cat F}, S : \harr {\cat C} {\cat E}$. where $\alpha:\cat
-  C, \beta:\cat D, \gamma : \cat E, \delta: \cat F$
-  \begin{enumerate}
-  \item $\tensorexistsXwithYandZ {\beta} {\harrapp P \alpha \beta} {(\tensorexistsXwithYandZ \gamma {\harrapp Q \beta \gamma} {\harrapp R \gamma \delta})} \cong \tensorexistsXwithYandZ {\gamma} {(\tensorexistsXwithYandZ \beta {\harrapp P \alpha \beta} {\harrapp Q \beta \gamma})} {\harrapp R \gamma \delta}$
-  \item $\homrallXYtoZ {\gamma} {(\tensorexistsXwithYandZ \beta {\harrapp P \alpha \beta} {\harrapp Q \beta \gamma})} {\harrapp S \alpha \gamma} \cong \homrallXYtoZ \beta {\harrapp P \alpha \beta} {\homrallXYtoZ \gamma {\harrapp Q \beta \gamma} {\harrapp S \alpha \gamma}}$
-  \item $\homlallXYtoZ {\alpha} {(\tensorexistsXwithYandZ \beta {\harrapp P \alpha \beta} {\harrapp Q \beta \gamma})} {\harrapp S \alpha \gamma} \cong \homlallXYtoZ \beta {\harrapp Q \beta \gamma} {\homlallXYtoZ \alpha {\harrapp P \alpha \beta} {\harrapp S \alpha \gamma}}$
-  \item $\homrallXYtoZ \gamma {\harrapp Q \beta \gamma} {(\homlallXYtoZ \alpha {\harrapp P \alpha \beta} {\harrapp S \alpha \gamma})} \cong \homlallXYtoZ \alpha {\harrapp P \alpha \beta} {(\homrallXYtoZ \gamma {\harrapp Q \beta \gamma} {\harrapp S \alpha \gamma})}$
-  \item $\pendallXdotY \alpha \homrallXYtoZ \beta P {P'} \cong \pendallXdotY \beta \homlallXYtoZ \alpha P {P'}$
-  \end{enumerate}  
-\end{lemma}
-\begin{proof}
-  TODO
-\end{proof}
-\subsection{Composition, Functoriality, Profunctoriality, Naturality}
-
-Hom sets don't provide primitive syntax for composition of morphisms
-in a category. As our first synthetic construction, we can
-\emph{define} composition using the elimination form for the unit.
-%
-\begin{definition}[Composition of Morphisms]
-  We define composition of morphisms as the following term of type $\Pi_{\cat C:\Cat}\hombinary {\alpha_1} {\punitinXfromYtoZ {\cat C}{\alpha_1} {\alpha_2}} {\alpha_2} {\punitinXfromYtoZ {\cat C}{\alpha_2} {\alpha_3}} {\alpha_3} {\punitinXfromYtoZ {\cat C}{\alpha_1} {\alpha_3}}$:
-  \[
-  \lambda \cat C.
-\punitelimtoYwithkontZ {\hlambda {\alpha_1}{\alpha_2} \homrallXYtoZ {\alpha_3} {\punitinXfromYtoZ {\cat C} {\alpha_2} {\alpha_3}} {\punitinXfromYtoZ {\cat C} {\alpha_1} {\alpha_3}}} {\lambdaunary {\alpha} {f : \punitinXfromYtoZ {\cat C}{\alpha} {\alpha_3}} {\alpha_3} f}
-  \]
-  We call this term $\textrm{comp}$
-\end{definition}
-This has the same flavor as a directed type theory: to construct the
-composite it is sufficient to define the action on the identity
-morphism, which is of course the identity for composition.
-%
-Note that we gave a slight bias to the definition to follow by
-induction on the \emph{first} function, whereas we could make the
-following alternative definition, naturally of type
-$\pendallXdotY {\alpha_2} \homrallXYtoZ {\alpha_3} {\punitinXfromYtoZ {\cat C}{\alpha_2} {\alpha_3}}
-{(\homlallXYtoZ {{\alpha_1}} {\punitinXfromYtoZ {\cat C}{\alpha_1} {\alpha_2}}{\punitinXfromYtoZ {\cat C}{\alpha_1} {\alpha_3}})}$:
-\[
-  \punitelimtoYwithkontZ {\hlambda {\alpha_2}{\alpha_3} \homlallXYtoZ {{\alpha_1}} {\punitinXfromYtoZ {\cat C}{\alpha_1} {\alpha_2}}{\punitinXfromYtoZ {\cat C}{\alpha_1} {\alpha_3}} } {\lambdaunary {\alpha} {g : \punitinXfromYtoZ {\cat C}{\alpha_1}{\alpha}} {\alpha} g}
-\]
-
-\begin{lemma}[Equivalence of Composition Definitions]
-  TODO
-\end{lemma}
-
-It should follow by $\beta\eta$ that this is associative and unital.
-
-Next, let's prove synthetically that every functor $F : \varr{\cat
-  C}{\cat D}$ is functorial in this notion of composition.
-\begin{definition}
-  For any $F : \varr {\cat C} {\cat D}$, we can define a morphism $F_1
-  : \homunary {\alpha_1} {\punitinXfromYtoZ {\cat C} {\alpha_1}
-    {\alpha_2}} {\alpha_2} {\punitinXfromYtoZ {\cat D} {F \alpha_1}
-    {F \alpha_2}}$ as
-  \[ F_1 = \punitelimtoYwithkontZ {\hlambda {\alpha_1}{\alpha_2} \punitinXfromYtoZ {\cat D} {F \alpha_1}{F \alpha_2}}
-  {\pendlambdaXdotY \alpha \punitrefl {\cat D} {F \alpha}}
-  \]
-  Note that
-  \[ A[\id_\alpha/\alpha] = \id_{A} \]
-  \[ A[f;g] = A[f];A[g] \]
-\end{definition}
-Note the similarity to the proof that all functions are congruences in
-MLTT/HoTT.
-
-And similarly, every profunctor $R : \harr {\cat C}{\cat D}$ is
-\emph{pro-functorial}.
-\begin{definition}
-  For any $R : \varr {\cat C} {\cat D}$, we can define left and right
-  actions of composition:
-  \[ \homunary {\alpha_1} {\tensorexistsXwithYandZ {\alpha_2} {(\punitinXfromYtoZ {\cat C} {\alpha_1}{\alpha_2})} {\tensorexistsXwithYandZ {\alpha_3} {(\harrapp R {\alpha_2} {\alpha_3})}}{(\punitinXfromYtoZ {\cat C} {\alpha_3}{\alpha_4})}} {\alpha_4} {\harrapp R {\alpha_1}{\alpha_4}}
-  \]
-  Furthermore we can show that they are functorial in that the induced functions
-  TODO
-  are equal
-\end{definition}
-
-Next we can define natural transformations as specific kinds of terms
-whose type is $\pendallXdotY {\alpha} \punitinXfromYtoZ {\cat D} {F
-  \alpha}{G \alpha}$. In fact we can generalize this to any
-endo-profunctor $P$:
-\begin{definition}[Natural element]
-  A natural element of an endo-profunctor $P : \harr {\cat C}{\cat C}$
-  is a term of type $\pendallXdotY {\alpha} \harrapp P \alpha \alpha$.
-
-  A natural transformation is the special case where $P =
-  \punitinXfromYtoZ {\cat D} {F \alpha}{G \alpha}$ for functors $F, G
-  : \varr {\cat C}{\cat D}$.
-
-  This is natural in the sense that the incuded terms
-  \[ \homunary {\alpha_1}{\punitinXfromYtoZ {} {\alpha_1}{\alpha_2}} {\alpha_2} {\harrapp P {\alpha_1}{\alpha_2}}  \]
-  and
-  \[ \pendallXdotY {\alpha_2} \homlallXYtoZ {\alpha_1} {\punitinXfromYtoZ {} {\alpha_1}{\alpha_2}} {\harrapp P {\alpha_1}{\alpha_2}}\]
-
-  are equal under the associativity equivalence between these two
-  types.
-\end{definition}
-
-\subsection{Yoneda Lemma and Representable Presheaves}
-
-Given any category $\cat C$, we can define the positive and negative
-Yoneda embeddings $Y^\pm : \cat C \to \pmPresheaf {\cat C}$ using the
-hom profunctor. Note that both of these functors are covariant in
-$\cat C$, so semantically, $\posPresheaf {\cat C}$ should be
-interpreted as $(\Set^{\cat C})^o$.
-
-\begin{definition}
-  Define the negative Yoneda embedding $Y^-[\alpha]$
-  \[ \alpha:\cat C \vdash Y^-[\alpha] = \lambda \alpha':\cat C. \punitinXfromYtoZ {\cat C}{\alpha'}{\alpha} : \negPresheaf {\cat C} \]
-
-  The positive Yoneda embedding is defined similarly $Y^+[\alpha]$.
-\end{definition}
-
-The Yoneda lemma says that the set of natural transformations from
-$Y^-[\alpha]$ to a presheaf $p$ is naturally isomorphic to the
-elements of the presheaf itself: $\negPresheafApp \alpha p$. In
-encoding this statement in our type theory, we find that we can use
-two different interpretations of ``natural transformations''.
-
-On the one hand, a natural transformation should be a morphism in the
-presheaf category. This interpretation of the Yoneda lemma would mean
-that we have a natural isomorphism of profunctors:
-\[ \alpha:\cat C; \pi:\negPresheaf {\cat C} \vdash \punitinXfromYtoZ{\negPresheaf {\cat C}} {Y^-[\alpha]} {\pi} \cong \negPresheafApp \alpha \pi \]
-Attempting to prove this we become stuck very quickly because we know
-nothing about the structure of morphisms in the presheaf category.
-
-On the other hand, a natural transformation of presheaves $\pi$ and
-$\rho$ can be defined as a family of functions from $\negPresheafApp
-\alpha \pi$ to $\negPresheafApp \alpha \rho$, or in other words as an
-\emph{end}. Specializing to the Yoneda lemma, we interpret ``a natural
-transformation from $Y^-[\alpha]$ to $\pi$'' as the following, which simplifies using $\beta$ reduction:
-\[ \alpha:\cat C; \pi:\negPresheaf {\cat C} \vdash \homrallXYtoZ {\alpha':\cat C} {\negPresheafApp {\alpha'} {Y^-[\alpha]}}{\negPresheafApp {\alpha'} \pi} = \homrallXYtoZ {\alpha':\cat C} {(\punitinXfromYtoZ {\cat C}{\alpha'}{\alpha})}{(\negPresheafApp {\alpha'} \pi)} \]
-And then the Yoneda lemma becomes:
-\[ \alpha:\cat C; \pi:\negPresheaf {\cat C} \vdash \homrallXYtoZ {\alpha':\cat C} {(\punitinXfromYtoZ {\cat C}{\alpha'}{\alpha})}{(\negPresheafApp {\alpha'} \pi)} \cong \negPresheafApp \alpha \pi\]
-Which is an instance of our earlier Lemma\,\ref{lem:yo-and-co-yo}
-which follows quite easily.
-
-Then we are left with the question of whether these two forms of the
-Yoneda lemma are equivalent, or slightly more generally are these two definitions of natural transformation of presheaves equivalent:
-\[ \rho:\negPresheaf{\cat C}; \pi:\negPresheaf {\cat C} \vdash \punitinXfromYtoZ{\negPresheaf {\cat C}} {\rho} {\pi} \]
-and
-\[ \rho:\negPresheaf{\cat C}; \pi:\negPresheaf {\cat C} \vdash \homrallXYtoZ {\alpha:\cat C} {\negPresheafApp \alpha \rho}{\negPresheafApp \alpha \pi} \]
-While we can show that there is a morphism from the hom set to the
-end, the inverse does not seem provable, and seems analogous to the
-\emph{univalence} axiom in homotopy type theory, as it tells us that
-the morphisms of the universe are given by a particular kind of
-function in our logic.
-
-\subsection{Adjoint Functors and Representable Profunctors}
-
-Next, we demonstrate that the two common definitions of adjoint
-functors are equivalent: isomorphism of hom-sets and units and
-co-units.
-%
-The proof goes through most easily, and utilizes duality most
-effectively, if we factorize the construction through an intermediate
-notion of when a functor \emph{represents} a profunctor.
-
-\begin{definition}[Functor Represents a Profunctor]
-  A functor $G : \varr{\cat D}{\cat C}$ right-represents a profunctor $R :
-  \harr {\cat C}{\cat D}$ when there is an isomorphism of profunctors:
-  \[ \punitinXfromYtoZ {\cat C} {\alpha} {G \beta} \cong \harrapp R {\alpha}{\beta}\]
-
-  We say a functor $F : \varr{\cat C}{\cat D}$ \emph{left-represents} a profunctor $R : \harr {\cat C}{\cat D}$ when there is an isomorphism of profunctors
-  \[ \punitinXfromYtoZ {\cat D} {F \alpha} {\beta} \cong \harrapp R {\alpha}{\beta}\]
-\end{definition}
-
-This is a parameterized generalization of the notion of an object
-representing a functor $R : C^{op} \to Set$.
-These notions are precisely dual. $F$ left-represents a profunctor $R$
-precisely if it right-represents it in the opposite SCT.
-
-Note that trivially any functor $G$ right-represents
-$\hlambda\alpha\beta \punitinXfromYtoZ {\cat D}{\alpha} {G \beta}$ and
-dually, any functor $F$ left-represents
-$\hlambda\alpha\beta\punitinXfromYtoZ {\cat C}{F\alpha}{\beta}$.
-
-Then we can show that any right-representation isomorphism is
-determined by a co-unit:
-\begin{lemma}[(Co)-unit of a Representation]
-  If $G: \varr {\cat D}{\cat C}$ right-represents $R : \harr {\cat
-    C}{\cat D}$, then there is a natural element, the co-unit:
-  \[ \pendallXdotY \beta \harrapp R {G\beta} \beta \]
-  Such that the half of the isomorphism 
-  \[ \punitinXfromYtoZ {\cat C} {\alpha} {G \beta} \multimap \harrapp R {\alpha}{\beta} \]
-  is defined by composition with the natural element.
-
-  Dually, any $F$ left-representing isomorphism with $R$ is determined
-  by a ``unit'' of type $\pendallXdotY \alpha \harrapp R {\alpha}
-  {F\alpha}$.
-\end{lemma}
-\begin{proof}
-  A homomorphism $\punitinXfromYtoZ {\cat C} {\alpha} {G \beta} \multimap \harrapp R {\alpha}{\beta}$ is by definition a term of type
-  \[ \pendallXdotY \alpha \homrallXYtoZ {\beta} {(\punitinXfromYtoZ {\cat C} \alpha {G \beta})} {\harrapp R \alpha \beta} \]
-  By \cref{lem:fubini}, this type is isomorphic to 
-  \[ \pendallXdotY \beta \homlallXYtoZ {\alpha} {(\punitinXfromYtoZ {\cat C} \alpha {G \beta})} {\harrapp R \alpha \beta} \]
-  By \cref{lem:yo-and-co-yo}, this type is isomorphic to
-  \[ \pendallXdotY \beta {{\harrapp R {(G \beta)} \beta}} \]
-
-  Therefore this half of the isomorphism is determined by a natural
-  element. The construction in the proof of \cref{lem:yo-and-co-yo} is
-  indeed by composition (i.e., unit elimination).
-
-  The $F$ case is precisely dual.
-\end{proof}
-
-Combining these two, we get the unit-counit definition for adjoint
-functors
-
-\begin{definition}[Adjunction by unit/co-unit]
-  We say $G : \varr {\cat D}{\cat C}$ is right adjoint to $F : \varr
-  {\cat C}{\cat D}$ if there is a unit
-  \[ \eta : \pendallXdotY \alpha \punitinXfromYtoZ \alpha {G(F(\alpha))}\]
-  and a co-unit
-  \[ \epsilon : \pendallXdotY \beta \punitinXfromYtoZ {F(G(\beta))} \beta \]
-
-  satisfying the triangle identities...
-  \[ \id_{R[\beta]} = R[\epsilon_{\beta}] \circ \eta_{R[\beta]} \]
-  \[ \id_{L[\alpha]} = \epsilon_{L[\alpha]} \circ L[\eta_{\alpha}] \]
-\end{definition}
-
-%% \begin{theorem}[Adjunctions]
-%%   $G$ is right adjoint to $F$ if and only if there is a natural isomorphism
-%%   $\punitinXfromYtoZ {\cat D} {(F \alpha)} \beta \cong \punitinXfromYtoZ {\cat D} \alpha {G \beta}$
-%% \end{theorem}
-%% \begin{proof}
-%%   First, by \cref{lem:co-unit}, the forward direction of the
-%%   isomorphism is equivalent to providing a term
-%%   \[ \eta : \pendallXdotY \alpha \punitinXfromYtoZ \alpha {G(F(\alpha)) } \]
-
-%%   Second, the backward direction of the isomorphism is equivalent to
-%%   providing a term
-%%   \[ \epsilon : \pendallXdotY \beta \punitinXfromYtoZ {F(G(\beta))} \beta \]
-
-%%   And the triangle laws state exactly that the induced functions are
-%%   isomorphisms.
-%% \end{proof}
-
-%% \subsection{Kan Extensions/Lifts}
-
-%% As a further example of the synthetic approach to universal
-%% properties, we consider right and left \emph{Kan extensions}, which
-%% are generalizations of limits and colimits, respectively.
-
-%% The right Kan extension of $F$ along $J$ is the ``best'' (maximal) way to
-%% fill in the following square:
-%% \[% https://tikzcd.yichuanshen.de/#N4Igdg9gJgpgziAXAbVABwnAlgFyxMJZABgBpiBdUkANwEMAbAVxiRAGEQBfU9TXfIRQBGclVqMWbACLdeIDNjwEiZYePrNWiDgHJu4mFADm8IqABmAJwgBbJGRA4ISAEzVNUnQCk5lm-aIjs5IohJabABiBlxAA
-%% \begin{tikzcd}
-%% \cat C \arrow[d, "J"] \arrow[r, "F"] & \cat D \\
-%% \cat C'                              &  
-%% \end{tikzcd}\]
-%% I.e. a functor $\textrm{Ran}_J F : \cat C' \to \cat D$ with a
-%% universal natural transformation $\textrm{Ran}_J F \circ J \Rightarrow
-%% F$. A left kan extension $\textrm{Lan}_J F$ is the same but with a
-%% natural transformation $F \Rightarrow \textrm{Lan}_J F \circ J $.
-
-%% We can express the concept in SCT as representing a certain profunctor
-%% constructed using the profunctor hom:
-%% \begin{definition}[Kan Extensions]
-%%   Given functors $F : \varr {\cat C}{\cat D}$ and $J : \varr {\cat
-%%     C}{\cat C'}$, a right Kan extension is a functor $\textrm{Ran}_J F
-%%   : \varr {\cat C'}{\cat D} $that right-represents the profunctor
-%%   \[ \hlambda {\beta:\cat D}{\alpha': \cat C} \homrallXYtoZ {\alpha}{(\punitinXfromYtoZ {\cat C'}{\alpha'}{J\alpha})} {\punitinXfromYtoZ {\cat D}{\beta}{F \alpha}} \]
-
-%%   A left Kan extensions is a functor $\textrm{Lan}_J F : \varr {\cat
-%%     C'}{\cat D}$ is a functor satisfying the dual property: it
-%%   \emph{left}-represents the profunctor
-%%   \[\hlambda {\alpha':\cat C}{\beta:\cat D} \homlallXYtoZ {\alpha}{(\punitinXfromYtoZ {\cat C'}{J\alpha}{\alpha'})} {\punitinXfromYtoZ {\cat D}{F \alpha}{\beta}}
-%%   \]
-%% \end{definition}
-
-%% The limit of a functor $F : \varr{\cat C}{\cat D}$ is a special case
-%% of a Kan extension where $J : \varr{\cat C}{1}$ is the unique functor
-%% to the terminal category. Similarly, the colimit of $F$ is the left
-%% Kan extension with respect to the same $J$.
-%% %
-%% Specializing to this case, the representability for limits means:
-%% \[ \punitinXfromYtoZ {\cat D}{\beta}{\textrm{lim} F} \cong \homrallXYtoZ {\alpha}{(\punitinXfromYtoZ {\cat 1}{\alpha'}{()})}{\punitinXfromYtoZ {\cat D}{\beta}{F \alpha}} \]
-%% And if we view a morphism in $\cat 1$ as being trivial, then this
-%% essentially states that mapping $\beta$ to $\textrm{lim} F$ is the
-%% same thing as constructing a cone: for all $\alpha :\cat C$, a
-%% morphism from $\beta$ to $F \alpha$, naturally in $\alpha$.
-
-%% The usual formulation of Kan extensions is in terms of a universal
-%% morphism.
-%% %
-%% We can reproduce this formulation using our established theory of
-%% equivalences for homs of profunctors.
-%% \begin{lemma}
-%%   One half of the representability isomorphism for a right Kan
-%%   extension $\textrm{Ran}_J F$ is determined by a natural
-%%   transformation
-%%   \[ \pendallXdotY \alpha \punitinXfromYtoZ {\cat D}{(\textrm{Ran}_JF)(J\alpha)}{F\alpha} \]
-
-%%   Dually, one half of the representability isomorphism for a left Kan
-%%   extension $\textrm{Lan}_J F$ is determined by a natural
-%%   transformation
-%%   \[ \pendallXdotY \alpha \punitinXfromYtoZ {\cat D}{F\alpha}{(\textrm{Lan}_JF)(J\alpha)} \]
-%% \end{lemma}
-%% \begin{proof}
-%%   By \cref{lem:unit-repr}, one half of representability is determined by a natural element
-%%   \[ \pendallXdotY {\alpha'} \homrallXYtoZ {\alpha}{(\punitinXfromYtoZ {\cat C'}{\alpha'}{J\alpha})} {\punitinXfromYtoZ {\cat D}{(\textrm{Ran}_J F)\alpha'}{F \alpha}} \]
-%%   which, by \cref{lem:fubini} is isomorphic to
-%%   \[
-%%   \pendallXdotY {\alpha} \homlallXYtoZ {\alpha'}{(\punitinXfromYtoZ {\cat C'}{\alpha'}{J\alpha})}{\punitinXfromYtoZ {\cat D}{(\textrm{Ran}_J F)\alpha'}{F \alpha}}
-%%   \]
-%%   which, by \cref{lem:yo-and-co-yo}, is isomorphic to 
-%%   \[ \pendallXdotY \alpha \punitinXfromYtoZ {\cat D}{(\textrm{Ran}_JF)(J\alpha)}{F\alpha} \]
-%%   by taking $P = \hlambda {\alpha'}{\alpha} {\punitinXfromYtoZ {\cat D}{(\textrm{Ran}_J F)\alpha'}{F \alpha}}$.
-
-%%   The proof for left Kan extensions follows by \cref{lem:duality}.
-%% \end{proof}
-
 \end{document}

### 2022-10-14 paper/paper.tex
short Adjoint example

@@
 
+\begin{lemma}[Equivalent Definitions of Adjoints]
+  We show that given a morphism of profunctors
+  \[ \homunary \alpha {\punitinXfromYtoZ{} {L\alpha} \beta} \beta {\punitinXfromYtoZ {} \alpha {R \beta}}\]
+  we can extract a unit natural transformation $\eta : \pendallXdotY
+  \alpha \punitinXfromYtoZ {}{\alpha}{R(L\alpha)}$ and vice-versa.
+\end{lemma}
+\begin{proof}
+  The construction is exactly the ordinary proof but formalized in \vett{} syntax.
+  Given the morphism of profunctors $M$, we define the unit by evaluating at the identity:
+  \[ \pendallXdotY \alpha \appunary M \alpha {\punitrefl \alpha} \alpha \]
+  and given the unit $\eta$, we can define a morphism of profunctors
+  by composing the unit with the functorial lift of the input:
+  \[ \homunary \alpha {f} \beta \textrm{comp}(\pendappXtoY \eta \alpha, \textrm{fctor}(R)(f))\]
+  That this is an isomorphism follows by a similar argument to the
+  proof of the Yoneda lemma.
+\end{proof}
+
 \section{Semantics}

### 2022-10-14 paper/paper.tex
typos in intro

@@
 programming languages typically have sound and complete models given by
-notions of structured categories,\cite{Lawvere69,LambekScott,moggi}.
+notions of structured categories~\cite{Lawvere69,LambekScott,moggi}.
 %
@@
 computational lambda calculus~\cite{moggi}, linear logic~\cite{girard}
-dependent type theory\cite{cartmell1986}, and many other type theories
+dependent type theory~\cite{cartmell1986}, and many other type theories
 designed based on category-theoretic semantics.
@@
 %
-Moreover, these generalizations are often what is needed for programming
+However, these generalizations are often exactly what is needed for programming
 language applications; for example, domain- and metric- and
@@
 parametricity and gradual
-typing\cite{wand197913,topos-of-trees,reflexive-graphs,double-cats-gradual-typing}.
+typing~\cite{wand197913,topos-of-trees,reflexive-graphs,double-cats-gradual-typing}.
 
@@
 using algebraic structures such as 2-categories, bicategories,
-(virtual) double categories and pro-arrow equipments\cite{bicategories,proarrow-equipments,LEINSTER2002391,Cruttwell2010}, an
+(virtual) double categories and pro-arrow equipments~\cite{bicategories,proarrow-equipments,LEINSTER2002391,Cruttwell2010}, an
 approach known as \emph{formal category theory}.
@@
 %
-The formal approaches also apply to enriched and internal settings, so
-\emph{formal} category theory proofs apply to these settings.
+Proofs in formal category theory apply to enriched and internal
+settings, which are instances of the formal axioms.
 %
@@
 By adhering to a \emph{syntactic discipline}, the logic guarantees
-that all functor terms are automatically functorial, all and natural
+that all functor terms are automatically functorial, and all natural
 transformation terms are natural.
@@
 %
-VETT provides an alternative alternative to algebraic and
+VETT provides an alternative to algebraic and
 string-diagram syntaxes for working with virtual equipments,
@@
 The analogue of a relation is a \emph{profunctor} (defined below), which
-are written as a kind of set with free category variables.  Like the
+is written like a set with free category variables.  Like the
 restriction to unary functors, we restrict to profunctors with two free
@@
 %
-We can strengthen this analogy into a formal construction: any
-construction in \vett{} can be erased to a corresponding construction
-or proof in predicate logic, as sets, functions and relations define a 
-(somewhat degenerate) virtual equipment.
+This analogy to predicate logic can be made formal: any construction in
+\vett{} can be erased to a corresponding construction or proof in
+predicate logic, as sets, functions, relations, and implication of
+relations define a (somewhat degenerate) virtual equipment.
 %% %
@@
 While the restricted syntax developed in this paper does not express
-important concepts such as functor categories or opposite categories,
-the restriction is natural in that it corresponds exactly to a
-well-understood notion of model, virtual equipments, which can express a
+some important concepts such as functor categories or opposite
+categories, the restriction is natural in that it corresponds exactly to
+virtual equipments, a well-understood notion of model that can express a
 great deal of fundamental results and constructions in category
@@
 have automatic functoriality/naturality properties).  To support this,
-\vett includes a third layer, an extensional dependent type theory in
+\vett{} includes a third layer, an extensional dependent type theory in
 the style of Martin-L\"of type theory. All of our ordered predicate
-logic judgments are also be indexed by a context from this dependent
+logic judgments are also indexed by a context from this dependent
 type theory, and the type theory includes universe types for categories,
 functors, profunctors and natural transformations.  This allow us to
-formalize theorems the logic is too restrictive to encode, analogous to
+formalize theorems the object logic is too restrictive to encode, analogous to
 2-level~\cite{voevodsky13hts,altenkirch+16strict} or indexed type
@@
 %
-Ordinary predicate logic is the foundation for logical presentations of
+Ordinary predicate logic is the foundation for proof-theoretic presentations of
 logical relations, such as Abadi-Plotkin logic for parametricity and
@@
 and logical relations involving these structures are proven to respect
-orderings: operational logical relations must be downward-closed and
+orderings; operational logical relations must be downward-closed and
 approximation relations should satisfy transitivity.
@@
 then a \emph{proof-relevant preorder}, where morphisms are the proofs of
-ordering, and the reflexivity and transitivity proofs must identity and
+ordering, and the reflexivity and transitivity proofs must satisfy identity and
 unit equations. A functor is then a \emph{proof-relevant monotone
@@
 profunctors $R, S : \cat C \pto \cat D$, a homomorphism from $R$ to $S$
-is simply a natural transformation, which in the preordered setting is
+is a natural transformation, which in the preordered setting is
 simply an implication of relations.
@@
 %% profunctor $R(F,G)$, the ``restriction'' of $R$ along $F$ and $G$.
-The remainder of this paper is organized as follows.  
-In Section~\ref{sec:syntax} we introduce the syntax of \vett{}.  In
-Section~\ref{sec:examples} we demonstrate how to use our syntax for formal
-category theory.  In Section~\ref{sec:semantics}, we develop some model theory
-for \vett{}, including a sound and complete notion of categorical model
-and sound interpretation in virtual equipments modeling ordinary,
-enriched and internal category theory.  In Section~\ref{sec:discussion}, we
-discuss related type theories and potential extensions.
+\textbf{Outline.}  In Section~\ref{sec:syntax} we introduce the syntax
+of \vett{}.  In Section~\ref{sec:examples} we demonstrate how to use our
+syntax for formal category theory.  In Section~\ref{sec:semantics}, we
+develop some model theory for \vett{}, including a sound and complete
+notion of categorical model and sound interpretation in virtual
+equipments modeling ordinary, enriched and internal category theory.  In
+Section~\ref{sec:discussion}, we discuss related type theories and
+potential extensions.
 

### 2022-10-14 paper/paper.tex
finish proofreading pass over main body

@@
 R_n$.  We write $d^-(\Phi)$ for the first category variable in $\Phi$
-(which we regard as the negative/contravariant position), $d^+(\Phi)$
+(which we regard as the negative or contravariant position), $d^+(\Phi)$
 for the last category variable in $\Phi$ (which we regard as the
-positive/covariant position), and $\Phi_1 \jnctx \Phi_2$ for the append
-of two category contexts where the last variable in $\Phi_1$ is equal to
+positive or covariant position), and $\Phi_1 \jnctx \Phi_2$ for the append
+of two transformation contexts where the last variable in $\Phi_1$ is equal to
 the first variable in $\Phi_2$.  Formal inductive definitions are in the
-appendix, but intuitively we have: 
+appendix, but intuitively: 
 \[
@@
       \\
-      \textrm{Objects} & A, B, C, a, b, c & ::= & \alpha \pipe M a \pipe (a,b)
+      \textrm{Objects} & a, b, c & ::= & \alpha \pipe M a \pipe (a,b)
       \pipe () \pipe \pi_i a \pipe (a_-, a_+, s) \pipe \pi_- a \pipe
@@
 \]
-All syntactic forms typed in context admit an action of substitution.
-For types and terms, this is as usual. Objects $\alpha:\cat C \vdash a :
-\cat D$ can be substituted for object variables $\beta : \cat D$ in
-other objects. We can also substitute objects into \emph{sets}, that is,
-if we have a set $P$ parameterized by a contravariant variable $\alpha :
-\cat C$ and a covariant variable $\beta: \cat D$, then we can substitute
-objects $a : \cat C$ and $b : \cat D$ for these variables
-$P[a/\alpha;b/\beta]$. This generalizes the ordinary precomposition of a
-relation by a function.  Semantically this is the ``restriction'' of a
-profunctor along two functors, which is just composition of functors if
-a profunctor is viewed as a functor to $\Set$. Modeling this operation
-as a substitution considerably simplifies reasoning using profunctors.
+As in linear logic, the latter rule applies only when the context
+contains a single set $R$.  All syntactic forms typed in context admit
+an action of substitution.  For types and terms, this is as
+usual. Objects $\alpha:\cat C \vdash a : \cat D$ can be substituted for
+object variables $\beta : \cat D$ in other objects. We can also
+substitute objects into \emph{sets}, that is, if we have a set $P$
+parameterized by a contravariant variable $\alpha : \cat C$ and a
+covariant variable $\beta: \cat D$, then we can substitute objects $a :
+\cat C$ and $b : \cat D$ for these variables $P[a/\alpha;b/\beta]$. This
+generalizes the ordinary precomposition of a relation by a function.
+Semantically this is the ``restriction'' of a profunctor along two
+functors, which is just composition of functors if a profunctor is
+viewed as a functor to $\Set$. Modeling this operation as a substitution
+considerably simplifies reasoning using profunctors.
 %
@@
 In our syntax we define it as the universal category $\cat C$ equipped
-with functors to $\cat C_-$ and $\cat C_+$ and a morphism to $P$.
+with functors to $\cat C_-$ and $\cat C_+$ and a natural transformation to $P$.
 
@@
 the sets/profunctors, which classify elements/natural transformations
-(the $\beta/eta$-rules are in the appendix).
+(the $\beta/\eta$-rules are in the appendix).
 %% %
@@
 Its introduction and elimination rules are analogous to the usual rules
-for equality in Martin-L\"of type theory.  The introduction rule is the
-identity morphism (reflexivity) and the elimination rule is an induction
-principle: we can use a term of $s : \punitinXfromYtoZ {\cat C} A B$ by
-specifying the behavior when $s$ is of the form $\punitrefl \alpha$ in
-the form of a continuation $\alpha. t$.  Like the $J$ elimination rule
-for equality in Martin-L\"of type theory, $P$ must be ``fully general'',
-i.e. well-typed for variables $\alpha$ and $\beta$; this is because for
-distinct variables $\alpha$ and $\beta$, $\punitinXfromYtoZ{\cat
-  C}{\alpha}{\beta}$ denotes the unit, which has a universal property,
-but $\punitinXfromYtoZ{\cat C} A B$ denotes a restriction of a unit,
-which in general does not.  Those familiar with linear
-logic~\cite{polakow-pfenning} might expect a more general rule, with the
-continuation $t$ typed in a context $\Phi_l \jnctx \Phi_r$ and $\Phi_l
-\jnctx \Phi \jnctx \Phi_r$ as the context of the conclusion of the rule.
-Because of dependency, this is not necessarily well-formed in cases
-where the endpoints $A$ and $B$ of $\punitinXfromYtoZ{\cat C} A B$ are
-not distinct variables.  However, the instances of this more general
-rule that do type check are derivable from our more restricted rule
-using right/left-hom types.
+for equality in intensional Martin-L\"of type theory.  The introduction
+rule is the identity morphism (reflexivity) and the elimination rule is
+an induction principle: we can use a term of $s : \punitinXfromYtoZ
+{\cat C} a b$ by specifying the behavior when $s$ is of the form
+$\punitrefl \alpha$ in the form of a continuation $\alpha. t$.  Like the
+$J$ elimination rule for equality in Martin-L\"of type theory, $P$ must
+be ``fully general'', i.e. well-typed for variables $\alpha$ and
+$\beta$; this is because for distinct variables $\alpha$ and $\beta$,
+$\punitinXfromYtoZ{\cat C}{\alpha}{\beta}$ denotes the unit, which has a
+universal property, but $\punitinXfromYtoZ{\cat C} a b$ denotes a
+restriction of a unit, which in general does not.  Those familiar with
+linear logic as in e.g.~\cite{polakow-pfenning} might expect a more
+general rule, with the continuation $t$ typed in a context $\Phi_l
+\jnctx \Phi_r$, and $\Phi_l \jnctx \Phi \jnctx \Phi_r$ as the context of
+the conclusion of the rule.  Because of dependency, this is not
+necessarily well-formed in cases where the endpoints $a$ and $b$ of
+$\punitinXfromYtoZ{\cat C} a b$ are not distinct variables.  However,
+the instances of this more general rule that do type check are derivable
+from our more restricted rule using right/left-hom types.
 
@@
      \Gamma\pipe \alpha:\cat C\vdash t : P[\alpha/\alpha;\alpha/\beta]\\\\
-     \Gamma\pipe\Phi \vdash s : \punitinXfromYtoZ {\cat C} A B}
-    {\Gamma \pipe \Phi \vdash \punitelimkontZatABC {\alpha. t} A s B : P[A/\alpha;B/\beta]}
+     \Gamma\pipe\Phi \vdash s : \punitinXfromYtoZ {\cat C} a b}
+    {\Gamma \pipe \Phi \vdash \punitelimkontZatABC {\alpha. t} A s B : P[a/\alpha;b/\beta]}
     \\
@@
   \pendlambdaXdotY \alpha \appunary {\text{lr}} \alpha {\punitrefl
-    \alpha} {L \alpha}$ and given the unit we can reconstruct the
+    \alpha} {L \alpha}$.  Given the unit we can reconstruct the
   forward homomorphism using
-  $\text{comp}$ and $\text{fctor}$ as defined in
-  Construction\ref{construction:synthetic-composition} as
+  $\text{comp}$ (composition) and $\text{fctor}$ (functoriality) from
+  Construction~\ref{construction:synthetic-composition} as\\
   $\appbinary {\text{comp}} \alpha
@@
 
-\noindent We can define weighted limits, which include ordinary limits and
-Kan extensions.
+We can define weighted limits, which as special cases include ordinary
+limits and Kan extensions.
 \begin{definition}
@@
 same domain but whose codomain is $S(f,g)$. This property is crucial
-for the completeness of our semantics as we only include a syntax for these globular terms 
+for the completeness of our semantics as we only include a syntax for these globular terms.
 %
@@
 \begin{definition}[Hyperdoctrine of Virtual Equipments]
-  A hyperdoctrine of virtual equipments is a pair of a category with
+  A hyperdoctrine of virtual equipments (HVE) is a pair of a category with
   families $\mathcal C$ and a functor $V^{(-)} : \mathcal C^o \to
@@
 \end{definition}
-Categories with families $\mathcal C$ model dependent type theory
-\citep{dybjer-cwf} and for each semantic context $\Gamma$, $V^\Gamma$ models
-the VETT judgments in context $\Gamma$, with the functoriality
-modeling the fact that all of these judgments admit a well-behaved
-action of substitution.
+Categories with families $\mathcal C$ model dependent type
+theory~\cite{dybjer-cwf} and for each semantic context $\Gamma$,
+$V^\Gamma$ models the VETT judgments in context $\Gamma$, with the
+functoriality modeling the fact that all of these judgments admit a
+well-behaved action of substitution.
 
@@
   C_2;\alpha_3:\cat C_3 \vdash R_2),\ldots$ to $(\beta_1:\cat
-  D_1;\beta_2\vdash S)$ with frames $\alpha_1:\cat C_1\vdash b_1 : \cat
-  D_1$ and $\alpha_n:\cat C_n\vdash b_2 : \cat D_2$ given by a term
+  D_1;\beta_2 : \cat D_2\vdash S)$ with frames $\alpha_1:\cat C_1\vdash b_1 : \cat
+  D_1$ and $\alpha_n:\cat C_n\vdash b_2 : \cat D_2$ is given by a term
   $x_1:R_1,x_2:R_2\ldots \vdash s :
@@
 connectives of VETT. We highlight one example from the literature that
-is highly general: Shulman's enriched indexed categories. Shulman's
-construction defines a virtual double category of large and small
-$\mathcal V$-categories for any pseudofunctor $\mathcal V : S^o \to
-\MonCat$ where $S$ is a category with finite products. He gives
-examples that show that this subsumes ordinary internal, enriched and
-indexed categories for suitable choices of $\mathcal V$, as well as
-more general categories that can be thought of as both indexed and
-enriched.
+is highly general: Shulman's enriched indexed
+categories~\cite{Shulman13}. Shulman's construction defines a virtual
+double category of large and small $\mathcal V$-categories for any
+pseudofunctor $\mathcal V : S^o \to \MonCat$ where $S$ is a category
+with finite products. He gives examples that show that this subsumes
+ordinary internal, enriched and indexed categories for suitable choices
+of $\mathcal V$, as well as more general categories that can be thought
+of as both indexed and enriched.
 %
@@
 %
-This is analogous to the situation for dependent type theory where
-substitution is strictly associative but in models is typically
-modeled by pullback which is only associative up to unique
-isomorphism.
+This is analogous to the situation for dependent type theory, where
+syntactic substitution is strictly associative, but semantic
+substitution is typically given by pullback, which is only associative
+up to unique isomorphism.
 %% In fact, one of the main examples is only a
@@
 
-\citet{jazmyers-strings} provides a string diagram calculus for
+Myers~\citet{jazmyers-strings} provides a string diagram calculus for
 double categories and pro-arrow equipments, generalizing string
@@
   bicategory}~\citet{CARBONI198711}.
-\citet{frey} describe some preliminary work on
+Frey~\citet{frey} describes some preliminary work on
 a proof system for Cartesian bicateogires.  This calculus is more
@@
 Finally, some variations on double categories have been used to model
-the structure of certain program logics. \citet{double-cats-gradual-typing} present a
-logic for \emph{vertically thin} pro-arrow equipments, i.e., pro-arrow
-equipments where the functor and natural transformation judgments are
-posetal. This calculus is quite different from ours in that the
-horizontal morphisms are terms of a programming language, and so they
-are not presented as relations with co-variant and contravariant
-variables, and the equipment structure is modeled by explicit coercions
-rather than by syntactic substitution of vertical arrows in horizontal
-arrows. Another similar calculus is System P \citep{dunphyreddy} which is an
-internal language of \emph{reflexive graph categories}, which are like
-double categories without horizontal composition.
+the structure of certain program
+logics. GTT~\cite{double-cats-gradual-typing} is a logic for
+\emph{vertically thin} pro-arrow equipments, i.e., pro-arrow equipments
+where the functor and natural transformation judgments are posetal. This
+calculus is quite different from ours in that the horizontal morphisms
+are terms of a programming language, and so they are not presented as
+relations with co-variant and contravariant variables, and the equipment
+structure is modeled by explicit coercions rather than by syntactic
+substitution of vertical arrows in horizontal arrows. Another similar
+calculus is System P~\cite{dunphyreddy} which is an internal language
+of \emph{reflexive graph categories}, which are like double categories
+without horizontal composition.
 
@@
 variables, in which case ordinary $\lambda$ calculus can be used to
-define functor categories as function types. Similarly, ideas from
+define functor categories as function types, and incorporate
+multi-variable profunctors as in~\cite{frey}.  Similarly, ideas from
 coeffects and enriched category theory could be used for defining

### 2022-10-14 paper/paper.tex
typos in A.1

@@
   \end{mathpar}
-  and is defined as
+  This is defined as
   \begin{align*}
@@
 
+  Note that $d^\pm\Phi$ will always be a set context with a single
+  variable $\alpha : \cat C$---we exploit the fact that we have these
+  singleton set contexts to avoid introducing a separate syntactic class
+  of category contexts $\alpha : \cat C$ and substitutions between them.
+
 \begin{definition}
@@
   \begin{align*}
-    \underline{\alpha:C} &= \alpha:C\\
+    \underline{\alpha:\cat C} &= \alpha:\cat C\\
     \underline{\Phi,x:R,\Psi} &= d^-\Phi; d^+{\Psi}
@@
     {\Gamma \pipe \Psi \vdash \phi :: \Phi}
-    {\Gamma\pipe\underline \vdash \Phi \boundary}
+    {\Gamma\pipe\underline \Psi \vdash \underline \phi : \underline \Phi}
   \end{mathpar}
@@
   \begin{align*}
-    \underline{\alpha:C} &= \alpha:C\\
-    \underline{\Phi,x:R,\Psi} &= d^-\Phi; d^+{\Psi}
+    \underline{b/\beta} &= b/\beta\\
+    \underline{\phi} &= d^-\phi; d^+{\phi} \text{ otherwise }
   \end{align*}
@@
   For any $\Gamma \pipe \alpha: \cat C \vdash a : \cat D$ and $\Gamma
-  \pipe \beta:\cat D \vdash b \cat E$, we define $b[a/\beta]$ by
+  \pipe \beta:\cat D \vdash b : \cat E$, we define $b[a/\beta]$ by
   recursion on $b$:
@@
     (M\,b)[a/\beta] &= M\,(b[a/\beta])\\
-    (\lambda \beta)    
+    (b_1,b_2)[a/\beta] &= (b_1[a/\beta],b_2[a/\beta])\\
+    (\pi_i b)[a/\beta] &= \pi_i b[a/\beta]\\
+    () [a/\beta] &= () \\
     (\pi_\pm b)[a/\beta] &= \pi_\pm b[a/\beta]\\
     (b_-,b_+,s)[a/\beta] &= (b_-[a/\beta],b_+[a/\beta], s[a/\beta])\\
+    (\lambda \alpha.R)[a/\beta] &=\lambda \alpha.R[a/\beta] %% FIXME: not sure if this is formally legal... split cases on positive and negative
   \end{align*}
 
-  Simultaneously by recursion on $s$ we define $s[\phi]$
+  Simultaneously by recursion on an element $s$ we define $s[\phi]$
   \begin{align*}
-    x[a/\alpha,t,b/\beta] &= t\\
+    x[a/\alpha,t/x,b/\beta] &= t\\
     \pendappXtoY M b[a/\alpha] &= \pendappXtoY M {b[a/\alpha]}\\
@@
     \tensorelimWkontZ {s[\phi_m]} {x.\beta.y. r[\phi_l,x/x,\beta/\beta \jnctx \beta/\beta,y/y,\phi_r]}\\
-    {\tensorintroatXwithYandZ b s t}[\phi_s \jnctx \phi_t] &= \tensorintroatXwithYandZ {d^+\phi_s} {s[\phi_s]} {t[\phi_t]}\\
+    {\tensorintroatXwithYandZ b s t}[\phi_s \jnctx \phi_t] &= \tensorintroatXwithYandZ {b[d^+\phi_s]} {s[\phi_s]} {t[\phi_t]}\\
     {(\homrappXtoYatZ s t a)}[\phi_f\jnctx \phi_a] &=
@@
     (M\,a\,b)[\xi] &= M\,(a[d^-\xi])\,(b[d^+\xi])\\
-    (\punitinXfromYtoZ {\cat C} a b)[\xi] &= \punitinXfromYtoZ {\cat C} {a[\xi]}{b[\xi]}\\
+    (\punitinXfromYtoZ {\cat C} a b)[\xi] &= \punitinXfromYtoZ {\cat C} {a[d^- \xi]}{b[d^+ \xi]}\\
     (\tensorexistsXwithYandZ \beta P Q)[\xi] &= \tensorexistsXwithYandZ \beta {P[d^-\xi;\beta/\beta]}{P[\beta/\beta;d^+\xi]}\\

### 2022-10-14 paper/paper.tex
state substitution principles' types in A.1

@@
   For any $\Gamma \pipe \alpha: \cat C \vdash a : \cat D$ and $\Gamma
-  \pipe \beta:\cat D \vdash b : \cat E$, we define $b[a/\beta]$ by
-  recursion on $b$:
+  \pipe \beta:\cat D \vdash b : \cat E$, we define
+  $\Gamma \pipe \alpha : \cat C \vdash b[a/\beta] : \cat E$ by recursion on $b$:
 
@@
 
-  Simultaneously by recursion on an element $s$ we define $s[\phi]$
+  Simultaneously,
+  for
+  $\Gamma \pipe \Psi \vdash \phi : \Phi$
+  and 
+  $\Gamma \pipe \Phi \vdash s : R$ we define\\
+  $\Gamma \pipe \Psi \vdash s[\phi] : R [ \underline \phi ]$ by recursion on $s$:
   \begin{align*}
@@
 
-  And finally, we define $P[\xi]$ by recursion on $P$:
+  And finally, for $\Gamma \pipe \Xi' \vdash \xi : \Xi$ and $\Gamma \pipe \Xi \vdash P \isaSet$,
+  we define $\Gamma \pipe \Xi' \vdash P[\xi] \isaSet$ by recursion on $P$:
   \begin{align*}

### 2022-10-14 paper/paper.tex
A.2
A.3

@@
 theory as a quotient inductive inductive type
-\cite{qit}.
+\cite{qit}.  We elide the types on the $\beta$ rules, as they can be
+inferred from the shape of the term, but include them for clarity on the
+$\eta$ rules.
 
@@
     {~}
-    {\negPresheafApp a (\lambda \alpha. R) = R[a/\alpha]}
+    {\negPresheafApp a {(\lambda \alpha. R)} = R[a/\alpha]}
 
@@
 elimination if $x$ were the only variable. We did not include this
-more general rule because it does not seem possible to define it with
-an admissible action of substitution, and would require a notion of
-explicit substitution.
-
+more general rule as a basic inference rule because 
+it requires an additional explicit substitution for the context
+$\Phi,x:\punitinXfromYtoZ {}{\alpha_-} {\alpha_+}, \Psi$,
+which would require making the substitutions part of the basic syntax.
 In the presence of hom types, we can prove this more general
@@
 order) is a type constructed by recursion on $\Phi$ and $\Psi$ using
-uses the hom types.
+uses the hom types.  The function applications for the hom types then
+provide a more lightweight way to incorporate the explicit substitution
+into the definition of the type theory.
 
@@
     \begin{align*}
-      s &= \homrappXtoYatZ {\homrlambdaXatYdotZ y \beta s} y \beta\tag{Hom $\beta$}\\
+      s &= \homrappXtoYatZ {\homrlambdaXatYdotZ y \beta s} y \beta\tag{Hom $\eta$}\\
       &= \homrappXtoYatZ {(\textrm{ind}_{\to}^{\alpha_-;\Psi}(\alpha.\homrlambdaXatYdotZ y \beta {s[\punitrefl \alpha/x]}; x))} y \beta \tag{Induction}\\

### 2022-10-14 paper/paper.tex
last part

@@
 
-\section{\vett{} Syntactic Metatheory}
+\section{Details of \vett{} Syntax and Syntactic Metatheory}
 \label{sec:appendix:syntax}
@@
 
-\section{Extended Formal Category Theory Examples}
+\section{Details of Formal Category Theory Examples}
 \label{sec:appendix:examples}
@@
 
-\section{Semantics}
+\section{Details of Semantics}
 \label{sec:appendix:semantics}
@@
     variable.
-  \item The horizontal arrows are the sets $\alpha:\cat C; \beta:\cat
-    D \vdash R$ up to renaming with source $\cat C$ and target $\cat
-    D$.
+  \item The horizontal arrows are the sets $\alpha:\cat C; \beta:\cat D
+    \vdash R$ (up to renaming $\alpha$ and $\beta$) with source $\cat C$
+    and target $\cat D$.
   \item Note that composable strings $\vec R$ of horizontal arrows are

### 2022-10-14 paper/paper.tex
spell check

@@
 The proofs both follow from the unit elimination rule, which is
-essentialy the Yoneda lemma---the two cases of showing (1) is an
+essentially the Yoneda lemma---the two cases of showing (1) is an
 isomorphism are precisely the $\beta$ and $\eta$ rules for the unit.
@@
 covariant or contravariant variables, but they do not have a term syntax
-for the natural transformations/elements, and Cartesian bicategories do
+or equational theory for the natural transformations/elements, and Cartesian bicategories do
 not have functors and restriction of profunctors by
@@
 \begin{lemma}[Generalized Unit Elim $\beta\eta$]
-  The admissible generalizaed unit elimination satisfies the described $\beta\eta$ equations.
+  The admissible generalized unit elimination satisfies the described $\beta\eta$ equations.
 \end{lemma}
@@
 
-Next, we provide some futher details for some of the examples of the
+Next, we provide some further details for some of the examples of the
 formal category theory constructions and theorems from
@@
   \end{enumerate}
-  Associativity and unit folllow by $\beta\eta$ for unit and homs.
+  Associativity and unit follow by $\beta\eta$ for unit and homs.
 \end{definition}

### 2022-12-08 paper/reviews/response.txt
reformatted response

We thank the reviewers for their very detailed and helpful feedback.  We
will integrate them into the final version of the paper --- including
all of the clarification comments not mentioned below.

We respond to the most pertinent questions first. And each response
includes "Reviewer X" so you can search for the responses to your
specific questions.

We agree with Review 3 that the most important advantages of our
approach are for enriched category theory (indeed, this was our
motivation for starting this work), and can revise the introduction to
emphasize this aspect more.  However, we do feel there are some
advantages even for ordinary category theory: while a synthetic theory
cannot eliminate all side conditions, it can help structure when the
user really needs to prove them.  One of the examples the reviewer
mentions, respect for quotients, is a classic example where an
"analytic" approach is unpleasant in practice ("setoid hell"), and
improved by synthetic approaches in NuPRL/Observational Type
Theory/HoTT/Cubical type theory.  While an adept category theorist may
know how to manage these obligations and e.g. pick a definition that
is more automatically natural, our experience explaining category
theory to programming languages researchers as background for
understanding some of our past papers suggests that the support of a
type theory may be useful for explaining how experts think about these
issues. We agree with the reviewer that the form the "domain specific
proof assistant" could take is as a library design for an existing
proof assistant, i.e., an embedded DSL.

Review 3 asks what the relevant CwF structures are in Constructions 4
and 5.  Our type theory distinguishes between the virtual equipment
layer (categories, functors, profunctors, natural transformations) and
the "meta" layer, a dependent type theory.  The semantics interprets
this in a hyperdoctrine of virtual equipments, with the CwF structure
for the meta-layer on the base of the hyperdoctrine.  In the "standard"
model, the base is just the category of sets Set, while the fibers of
the hyperdoctrine are Cat.  So there is no need for a CwF structure on
Cat.

Review 3 asks if every CwF gives rise to a HVE. Certainly construction
4 can be adapted to any elementary topos with sufficient universes
since the category-theoretic constructions needed are all
intuitionistically valid. So this should correspond to a Category with
Families extended with certain universal properties, for instance,
extensional identity types.

Review 3 suggests some additional references and examples (Palmgren,
  Weber, Di Liberti and Loregian). Thank you; we will include them!

Review 1 asks about the limitations of the present formalism, and
Review 2 similarly asked about why we restrict to unary
functors/binary profunctors. There are several useful concepts we
cannot formalize because our judgments are limited to covariant
functors of a single argument: categories of functors and opposite
categories, for instance. Next steps for the current work would be to
extend the judgmental structure to support these. Some preliminary
work along these lines is in Reference [21]. However, there is a
tradeoff: by adding more structure in the type theory, we support
fewer models. We chose these restrictions as a starting point because
they correspond to a well-understood notion of model (virtual
equipments), future extensions would likely require the equipment to
come with a monoidal closed structure. In particular, we expect that
categories enriched in a non-symmetric monoidal closed category will
not be a model of these extensions.

Review 1 asks about the definition of d(Ξ) and the presheaf
elimination rules. d-(x:C) and d+(x:C) are both x:C; we will add this
definition. The negative presheaf elimination rule is the closure
under substitution of a : C ; p : P- C |- a in p Set i.e. evaluation
is a map C^op x (Set^{C^op}) -> Set

Review 1 asks about the tensor product of profunctors. The tensor
product of profunctors is indeed a colimit --- e.g. it can be defined
using a coend.  Syntactically, beta is just a bound variable (and can
of course be alpha-converted).  The fact that maps out of the tensor
respect the quotienting is implicit/synthetic, and comes from the fact
that all open terms ... beta : C ... |- t : S respect this
quotienting.  This is similar to how quotient types are handled in
NuPRL/Observational type theory/HoTT/Cubical type theory.

Review 1 had several questions about Section 4. We will rephrase
Theorem 2 with less shorthand. The formal meaning of "almost any" is
to emphasize that the initiality theorem is modular in that most
connectives are independent of each other. The exception is that to
get the right universal property for hom and tensor of profunctors, we
need the exponential of profunctors. vDbl_r and Cat_U should, as the
reviewer guessed, be vEq and Cat_κ. The purpose of Construction 3 is
to reduce the construction of a model of VETT to just constructing a
virtual equipment, since any virtual equipment can be "upgraded" to a
hyperdoctrine of equipments using this construction.

Review 2 asks if one can generate explicit proofs of functoriality
/naturality from proofs in the type theory.  Yes, "running" the
(constructive) proof of soundness (Theorem 2) would do exactly this, by
compiling our "synthetic" proofs to "analytic" ones.

Review 2 asks about "P ▹∀α Q" and the difference with "∀α.P".

P ▹∀α Q is the 'hom' of profunctors, the right adjoint to - (x)∃α P.  This
type binds both an object variable α and a transformation variable x:P.

In the type of composition 
   ∀α1. (α1 → α2) |>∀α2 (α2 → α3 |>∀α3 (α1 → α3))
α2 occurs covariantly in (α1 → α2)
and covariantly in (α2 → α3 |>∀α3 (α1 → α3)) because the
contravariance of |> and → cancel out.

∀α.P is the 'end' over a profunctor with 1 free variable (Figure 5,
bottom row).  This type is used to bind the "last" object variable in a
singleton context. The type for functoriality that the reviewer gives is
isomorphic to the type we gave by Lemma 3, part 5.

Review 2 asks about our comments on the Yoneda lemma. We had a typo in
this formula, thanks --- we meant ∀x. [ (∀y.(x=y) ⇒ Py) iff P x].

Review 2 asks about the proof of initiality theorem. The initiality
theory could in principle be mechanized in Agda, though we have not
yet done so.

### 2023-01-27 paper/paper.tex
final version?

@@
 \begin{document}
+
+\newif\ifextended
+\extendedfalse
+\newcommand{\theappendix}{\ifextended the appendix \else the extended version \cite{extended-version}\fi}
+
 %% \newtheorem{theorem}{Theorem}
@@
 \newcommand{\ohol}{\vett}
+\newcommand{\fVDC}{\textrm{fVDCs}}
+\newcommand{\VMJ}{\textrm{VM}_J}
 \newcommand{\sem}[1]{\llbracket{#1}\rrbracket}
@@
 \newcommand{\tensorintroatXwithYandZ}[3]{(#2,#1,#3)}
-\newcommand{\tensorelimXtoYwithkontZ}[3]{\textrm{ind}_{\to}(#1,#2,#3)}
-\newcommand{\tensorelimfromWXtoYwithkontZ}[4]{\textrm{ind}_{\odot}(#1,#2,#3,#4)}
-\newcommand{\tensorelimVatWtoXtoYwithkontZ}[5]{\textrm{ind}_{\odot}(#4,#2,#1,#3,#5)}
-\newcommand{\tensorelimWfromXtoYkontZ}[4]{\textrm{ind}_{\odot}(#4;#1,#2,#3)}
 \newcommand{\tensorelimWkontZ}[2]{\textrm{ind}_{\odot}(#2;#1)}
@@
 \title{A Formal Logic for Formal Category Theory}
-\author{Max S. New\inst{1,2} and Daniel R. Licata\inst{2}}
+\author{Max S. New\inst{1,2}\orcidID{0000-0001-8141-195X} and Daniel R. Licata\inst{2}}
 \institute{University of Michigan \and Wesleyan University}
@@
   and all transformations are natural by construction, with no separate
-  proofs necessary. Important category theoretic proofs such as the
-  Yoneda lemma and Co-yoneda lemma become simple type theoretic proofs
+  proofs necessary. Important category-theoretic proofs such as the
+  Yoneda lemma and Co-yoneda lemma become simple type-theoretic proofs
   about the relationship between unit, tensor and (ordered) function
@@
 computational lambda calculus~\cite{moggi}, linear logic~\cite{girard}
-dependent type theory~\cite{cartmell1986}, and many other type theories
+dependent type theory~\cite{cartmell1986,seely_1984}, and many other type theories
 designed based on category-theoretic semantics.
@@
 While this research program has been quite successful,
-category-theoretic techniques in computer science are sometimes
-criticized for being overly complex and hard to learn, especially for
-computer scientists who do not have a strong background in pure
-mathematics.
-%
-While in some ways this could be remedied by better introductory
-materials, in some ways category theory is objectively complex.
+category-theoretic notions can be overwhelming for beginners.
 %
@@
 %
-Carrying out these proofs explicitly is quite tedious and many newcomers
-to are left with the impression that category theory is full of long,
-but ultimately trivial constructions.
-%
-This complexity is compounded when we move from ordinary category theory to
-enriched and internal category theory, where constructions must be
-additionally proven continuous, monotone, etc, in addition to natural
-or functorial.
+Carrying out these proofs explicitly is quite tedious and many
+newcomers are left with the impression that category theory is full of
+long, but ultimately trivial constructions.
+%
+This complexity is compounded when moving from ordinary category
+theory to enriched and internal category theory, where constructions
+must be additionally proven continuous, monotone, etc, in addition to
+natural or functorial.
 %
 However, these generalizations are often exactly what is needed for programming
-language applications; for example, domain- and metric- and
+language applications; for example, domain-, metric- and
 step-index-enriched categories have been used to model recursive
@@
 defining functions that are continuous by construction --- e.g. that
-composition of continuous functions is continuous and that previously
-defined functions are known to be continuous.
+composition of continuous functions is continuous.
 %
@@
 In the case of category theory, these principles can be formalized
-using algebraic structures such as 2-categories, bicategories,
-(virtual) double categories and pro-arrow equipments~\cite{bicategories,proarrow-equipments,LEINSTER2002391,Cruttwell2010}, an
-approach known as \emph{formal category theory}.
+using algebraic structures such as 2-categories, bicategories, Yoneda
+structures, (virtual) double categories, pro-arrow
+equipments~\cite{bicategories,proarrow-equipments,STREET1978350,LEINSTER2002391,Cruttwell2010},
+an approach known as \emph{formal category theory}.
 %
@@
 %
-A downside is that these algebraic structures are quite complicated, and
-practitioners typically employ either an algebraic combinator syntax or
-a 2-dimensional diagrammatic language that can be quite beautiful and
+A downside is that these algebraic structures are quite complicated,
+and practitioners typically employ either an algebraic combinator
+syntax (formalized in \citet{CURIEN1986188}) or a
+2-dimensional diagrammatic language that can be quite beautiful and
 elegant, but is also somewhat removed from the traditional formulation
@@
 
-%% The type theoretic syntax allows us to write all of our 
+%% The type-theoretic syntax allows us to write all of our 
 %% constructions are written as if they were
@@
 Just as a predicate logic has a notion of type, term, relation and
-implication, \vett{} is based on four analogous category theoretic
+implication, \vett{} is based on four analogous category-theoretic
 concepts: categories, functors, profunctors and natural transformations
@@
 %
-The analogue of a relation is a \emph{profunctor} (defined below), which
+The analog of a relation is a \emph{profunctor} (defined below), which
 is written like a set with free category variables.  Like the
@@
 formalize theorems the object logic is too restrictive to encode, analogous to
-2-level~\cite{voevodsky13hts,altenkirch+16strict} or indexed type
+2-level~\cite{voevodsky13hts,altenkirch+16strict,PALMGREN2019102715} or indexed type
 theories~\cite{isaev21indexed,cervesatopfenning02llf,vakar15linear,krishnaswami+15linear}.
@@
 Additionally, \vett{} may serve as the basis of a future domain specific
-proof assistant for category theoretic proofs.  To pilot-test this, we
+proof assistant for category-theoretic proofs.  To pilot-test this, we
 have formalized the syntax of \vett{} in Agda 2.6.2.2, using the rewrite
@@
 profunctor $R$ from $\mathcal C$ to $\mathcal D$, written $R : \cat C
-\pto \cat D$ is a functor $R : {\cat C}^{op} \times {\cat D} \to \Set$.
+\pto \cat D$ is a functor $R : {\cat C}^{o} \times {\cat D} \to \Set$\footnote{${\cat C}^o$ is the notation we use for the opposite category of $\cat C$}.
 Because a profunctor outputs a $\Set$ rather than a proposition, it is
@@
 The contexts $\Gamma$ of term variables with their types are as usual;
-we write ${\Gamma \isadtctx}$ to indicate that a context is well-formed.
+we write ``${\Gamma \isadtctx}$'' to indicate that a context is well-formed.
 We name the remaining contexts after the judgements that they are used
-by.  The set contexts $\Gamma \vdash \Xi \boundary$, which will be used
+by.  The set contexts $\Xi$, which will be used
 to type-check sets, contain object variables with their categories. The
@@
 and covariant variable $\beta$.  Finally, the transformation contexts
-$\Gamma \vdash \Phi \isavectx$ contain element variables with their sets, alternating with those
+$\Phi$ contain element variables with their sets, alternating with those
 sets' object variables with their categories.  A typical $\Phi$ has the
@@
 \]
-and represents the composition of the ``relations'' $R_1, R_2, R_3, \ldots,
-R_n$.  We write $d^-(\Phi)$ for the first category variable in $\Phi$
-(which we regard as the negative or contravariant position), $d^+(\Phi)$
-for the last category variable in $\Phi$ (which we regard as the
-positive or covariant position), and $\Phi_1 \jnctx \Phi_2$ for the append
-of two transformation contexts where the last variable in $\Phi_1$ is equal to
-the first variable in $\Phi_2$.  Formal inductive definitions are in the
-appendix, but intuitively: 
+and represents the composition of the ``relations'' $R_1, R_2, R_3,
+\ldots, R_n$.  We write $d^-(\Phi)$ for the first category variable in
+$\Phi$ (which we regard as the negative or contravariant position),
+$d^+(\Phi)$ for the last category variable in $\Phi$ (which we regard
+as the positive or covariant position) and use the notation $d^\pm\Xi$
+with the same meaning. We write $\Phi_1 \jnctx \Phi_2$ for the append
+of two transformation contexts, which is only well-formed when the
+last variable in $\Phi_1$ is equal to the first variable in
+$\Phi_2$. Formal inductive definitions are in the appendix, but
+intuitively:
 \[
@@
 %% types of dependent type theory, as well as several forms that allow us
-%% to abstract over the category theoretic constructions: a type $\Cat$
+%% to abstract over the category-theoretic constructions: a type $\Cat$
 %% of categories, a type $\varr{\cat C}{\cat D}$ for functors from $\cat
@@
     \punitrefl b \pipe 
-    {\tensorelimWkontZ s {x.\beta.y.r}} \pipe 
+    {\tensorelimWkontZ s {x,\beta,y.r}} \pipe 
     {\tensorintroatXwithYandZ b s t} \pipe
@@
   \underline \Phi \vdash R \isaSet$.  A transformation $s$ has a context
-  $\Phi$ of transformation variable and a single set output $R$. To be
-  well formed, the context and set must be parameterized by the same
+  $\Phi$ of transformation variables and a single output set $R$. To be
+  well-formed, the context and set must be parameterized by the same
   contravariant and covariant object variables. To ensure this, we use a
@@
 %% The type theory has 4 central concepts, which we will refer to
-%% sometimes using type theoretic/set theoretic and at other times
+%% sometimes using type-theoretic/set theoretic and at other times
 %% category-theoretic terminology.
 %% %
-%% In type theoretic terminology, judgment theory has \emph{sorts} $\cat
+%% In type-theoretic terminology, judgment theory has \emph{sorts} $\cat
 %% C, \cat D, \cat E$, and \emph{types} $A,B,C$ which have a given sort
@@
 \[
-\inferrule*{}{ \Gamma \pipe \alpha : \cat C \vdash \alpha : \cat C}
+\inferrule*{~}{ \Gamma \pipe \alpha : \cat C \vdash \alpha : \cat C}
 \qquad
-\inferrule*{}{ \Gamma \pipe \alpha : \cat C, x : R, \beta : \cat D \vdash x : R}
+\inferrule*{~}{ \Gamma \pipe \alpha : \cat C, x : R, \beta : \cat D \vdash x : R}
 \]
-As in linear logic, the latter rule applies only when the context
+As when using variables in linear logic, the latter rule applies only when the context
 contains a single set $R$.  All syntactic forms typed in context admit
@@
 %
-First we introduce the additives: the unit category $1$ and product
-category $\cat C \times \cat D$ have the usual introduction and
-elimination rules defining functors to/from them.  
+First we introduce the additive connectives: the unit category $1$ and
+product category $\cat C \times \cat D$ have the usual introduction
+and elimination rules defining functors to/from them.
 %
@@
 as a subset $\{ (a,b) \in A\times B | R(a,b)\}$, any profunctor $P :
-C_-^o\times C_+ \to \Set$ can be viewed as a category with a functor
-to $C_- \times C_+$ (no op), specifically a two-sided discrete
-fibration. In set-based category theory, the objects are triples
+{\cat C}_-^o\times {\cat D}_+ \to \Set$ can be viewed as a category with a functor
+to ${\cat C}_- \times {\cat D}_+$ (no op), specifically a two-sided discrete
+fibration. In set-based category theory, the objects of $\graphProf \alpha
+\beta P$ are triples
 $(a_-, a_+, s : P(a_-,a_+))$ and morphisms from $(a_-,a_+, s)$ to
@@
 With various choices of $P$, this connective can be used to define the
-arrow category, slice category and comma category constructions.
+arrow category, slice category, comma category and category of
+elements.
 %
@@
 it. We use the notation $\negPresheafApp a p$ for the elements of the
-induced profunctor.
+induced profunctor. Since $a$ occurs in a negative position, it must
+depend only on the contravariant variable $d^-\Xi$ and vice-versa for
+$p$.
 %
@@
   \begin{array}{l}
-    \text{Unit:} \quad
+    \text{Unit:}
+    \qquad
     \inferrule{ ~ }{\Gamma \vdash 1 \isaCat}
-    \quad
-    \inferrule*{ ~ }{\Gamma\pipe\alpha:C \vdash () : 1}\\\\
+    \qquad
+    \inferrule{ ~ }{\Gamma\pipe\alpha:C \vdash () : 1}\\\\
 
@@
     \inferrule
-    {\defaultObCtx \vdash a_- : C_-\and
-      \defaultObCtx \vdash a_+ : C_+\and
+    {\defaultObCtx \vdash a_- : {\cat C}_-\and
+      \defaultObCtx \vdash a_+ : {\cat C}_+\and
       \defaultObCtx \vdash s : P[a_-/\alpha;a_+/\beta]}
@@
     \\\\
-    \text{Negative Presehaf:} \\
+    \text{Negative Presheaf:} \\
     \inferrule
@@
 %% %
-First, the unit set is our syntax for the profunctor of morphisms of a
-category $\cat C$, which is well-formed when given two objects of the
-same category.  
-%
-Its introduction and elimination rules are analogous to the usual rules
-for equality in intensional Martin-L\"of type theory.  The introduction
-rule is the identity morphism (reflexivity) and the elimination rule is
-an induction principle: we can use a term of $s : \punitinXfromYtoZ
-{\cat C} a b$ by specifying the behavior when $s$ is of the form
-$\punitrefl \alpha$ in the form of a continuation $\alpha. t$.  Like the
-$J$ elimination rule for equality in Martin-L\"of type theory, $P$ must
-be ``fully general'', i.e. well-typed for variables $\alpha$ and
-$\beta$; this is because for distinct variables $\alpha$ and $\beta$,
-$\punitinXfromYtoZ{\cat C}{\alpha}{\beta}$ denotes the unit, which has a
-universal property, but $\punitinXfromYtoZ{\cat C} a b$ denotes a
-restriction of a unit, which in general does not.  Those familiar with
-linear logic as in e.g.~\cite{polakow-pfenning} might expect a more
-general rule, with the continuation $t$ typed in a context $\Phi_l
-\jnctx \Phi_r$, and $\Phi_l \jnctx \Phi \jnctx \Phi_r$ as the context of
-the conclusion of the rule.  Because of dependency, this is not
+First, the unit set $\punitinXfromYtoZ {\cat C} a b$ is our syntax for
+the profunctor of morphisms in $\cat C$ instantiated at generalized
+objects $a$ and $b$.
+%
+Its introduction and elimination rules are analogous to the usual
+rules for equality in intensional Martin-L\"of type theory.  The
+introduction rule is the identity morphism (reflexivity) and the
+elimination rule is an induction principle: we can use a term of $s :
+\punitinXfromYtoZ {\cat C} a b$ by specifying the behavior when $s$ is
+of the form $\punitrefl \alpha$ in the form of a continuation
+$\alpha. t$.  Like the $J$ elimination rule for equality in
+Martin-L\"of type theory, $P$ must be ``fully general'',
+i.e. well-typed for variables $\alpha$ and $\beta$. This is because
+for distinct variables $\alpha$ and $\beta$, $\punitinXfromYtoZ{\cat
+  C}{\alpha}{\beta}$ denotes the unit in a virtual double category,
+which has a universal property, but $\punitinXfromYtoZ{\cat C} a b$
+denotes a restriction of the unit, which in general does not.  Those
+familiar with linear logic as in e.g.~\cite{polakow-pfenning} might
+expect a more general rule, where the continuation $t$ is allowed to
+use variables that are not used in $s$, i.e., have a context $\Phi_l
+\jnctx \Phi_r$ and the conclusion of the rule to have a context
+$\Phi_l \jnctx \Phi \jnctx \Phi_r$. Because of dependency, this is not
 necessarily well-formed in cases where the endpoints $a$ and $b$ of
 $\punitinXfromYtoZ{\cat C} a b$ are not distinct variables.  However,
-the instances of this more general rule that do type check are derivable
-from our more restricted rule using right/left-hom types.
+the instances of this more general rule that do type check are
+derivable from our more restricted rule using right/left-hom types.
 
@@
 %
+Similar to ordered lambda calculus, the covariant hom is
+right-associative while the contravariant hom is left-associative.
+%
 Then the
@@
 Semantically, in ordinary category theory these are known as the
-\emph{hom} of profunctors and are constructed using a combination of
-an end and a function set.
+\emph{hom} of profunctors and are adjoint to the composition of
+profunctors \cite{benabou2000distributors}.
 %
@@
     {\cat D \isSmall\\\\
-      \Gamma \pipe d^-\Xi; \beta:D\vdash P \isaSet\\\\
-     \Gamma \pipe \beta:D;d^+\Xi \vdash Q \isaSet}
-    {\Gamma \pipe \Xi \vdash \tensorexistsXwithYandZ {\beta:D} P Q \isaSet}
+      \Gamma \pipe d^-\Xi; \beta:{\cat D}\vdash P \isaSet\\\\
+     \Gamma \pipe \beta:{\cat D};d^+\Xi \vdash Q \isaSet}
+    {\Gamma \pipe \Xi \vdash \tensorexistsXwithYandZ {\beta:{\cat D}} P Q \isaSet}
     \quad
@@
    \Gamma\pipe \Phi_l \jnctx x:P, \beta:\cat D, y:Q \jnctx \Phi_r \vdash t : R\\\\
-   \Gamma \pipe\Phi_m \vdash s : \tensorexistsXwithYandZ {\beta} P Q}
-  {\Gamma\pipe\Phi_l\jnctx \Phi_m \jnctx \Phi_r \vdash \tensorelimWkontZ s {x.\beta.y. t} : R}
+   \Gamma \pipe\Phi_m \vdash s : \tensorexistsXwithYandZ {\beta:\cat D} P Q}
+  {\Gamma\pipe\Phi_l\jnctx \Phi_m \jnctx \Phi_r \vdash \tensorelimWkontZ s {x,\beta,y. t} : R}
   \\
@@
   {\Gamma \pipe \Phi \vdash
-   \homrlambdaXatYdotZ {x : R} {\alpha : \cat C} t : \homrallXYtoZ \alpha R P}
+   \homrlambdaXatYdotZ {x : R} {\alpha : \cat C} t : \homrallXYtoZ {\alpha:\cat C} R P}
   \quad
@@
   {\Gamma\pipe\alpha:\cat C, x : R, \Phi \vdash t : P}
-  {\Gamma\pipe\Phi \vdash \homllambdaXatYdotZ {x : R} {\alpha : \cat C} t : \homlallXYtoZ \alpha R P}
+  {\Gamma\pipe\Phi \vdash \homllambdaXatYdotZ {x : R} {\alpha : \cat C} t : \homlallXYtoZ {\alpha:\cat C} R P}
   \quad
@@
   \begin{enumerate}
-  \item Identity: $\pendallXdotY {\alpha} {\punitinXfromYtoZ {\cat C}{\alpha}{\alpha}}$
-  \item Composition: $\hombinary {\alpha_1} {\punitinXfromYtoZ {}{\alpha_1}{\alpha_2}} {\alpha_2} {\punitinXfromYtoZ {}{\alpha_1}{\alpha_2}} {\alpha_3} {\punitinXfromYtoZ {}{\alpha_1}{\alpha_3}}$
-  \item Functoriality: for any $F : \varr {\cat C}{\cat D}$, $\homunary {\alpha_1} {\punitinXfromYtoZ {}{\alpha_1}{\alpha_2}} {\alpha_2} {\punitinXfromYtoZ {} {F(\alpha_1)}{F(\alpha_2)}}$.
+  \item Identity: $\pendallXdotY {\alpha:\cat C} {\punitinXfromYtoZ {\cat C}{\alpha}{\alpha}}$
+  \item Composition: $\hombinary {\alpha_1:\cat C} {(\punitinXfromYtoZ {\cat C}{\alpha_1}{\alpha_2})} {\alpha_2:\cat C} {(\punitinXfromYtoZ {\cat C}{\alpha_2}{\alpha_3})} {\alpha_3:\cat C} {(\punitinXfromYtoZ {\cat C}{\alpha_1}{\alpha_3})}$
+  \item Functoriality: for any $F : \varr {\cat C}{\cat D}$, $\homunary {\alpha_1:\cat C} {(\punitinXfromYtoZ {\cat C}{\alpha_1}{\alpha_2})} {\alpha_2:\cat C} {(\punitinXfromYtoZ {\cat D} {F(\alpha_1)}{F(\alpha_2)})}$.
    \item Profunctoriality: for any $R : \harr {\cat C}{\cat D}$ if
-     $\cat D$ is small then \\ $\homtrinary {\alpha_1}
-     {\punitinXfromYtoZ{} {\alpha_1}{\alpha_2}} {\alpha_2}
-     {R{\alpha_2}{\beta_2}} {\beta_2} {\punitinXfromYtoZ{} {\beta_2}
-       {\beta_1}} {\beta_1} {R\alpha_1\beta_1}$
+     $\cat D$ is small then \\ $\homtrinary {\alpha_1:\cat C}
+     {(\punitinXfromYtoZ{\cat C} {\alpha_1}{\alpha_2})} {\alpha_2:\cat C}
+     {R{\alpha_2}{\beta_2}} {\beta_2:\cat D} {(\punitinXfromYtoZ{\cat D} {\beta_2}
+       {\beta_1})} {\beta_1:\cat D} {R\alpha_1\beta_1}$
   \end{enumerate}
 \end{construction}
-Reflexivity and transitivity generalize the corresponding properties
-of equality, with the lack of symmetry being a key feature of the
-generalization.
+Identity and Composition generalize the reflexivity and transitivity
+properties of equality, respectively, with the lack of symmetry being
+a key feature of the generalization.
 %
@@
 Naturality is more complex to state, and it is a statement about the
-\emph{proofs} so it has no analogue in ordinary higher-order
+\emph{proofs} so it has no analog in ordinary higher-order
 logic. The following version is stated for any \emph{profunctor}, with
@@
     \alpha}$, by composing with profunctoriality, we can construct
-  terms $\alpha_1, f : \punitinXfromYtoZ {\cat C} {\alpha_1}
-  {\alpha_2}, \alpha_2 \vdash \textrm{lcomp}(f,\pendappXtoY t
+  terms $\alpha_1:{\cat C}, f : \punitinXfromYtoZ {\cat C} {\alpha_1}
+  {\alpha_2}, \alpha_2:{\cat C} \vdash \textrm{lcomp}(f,\pendappXtoY t
   {\alpha_2})$ and $\textrm{rcomp}(\pendappXtoY t {\alpha_1}, f) :
@@
 Yoneda and Co-Yoneda lemmas. Despite being ultimately quite
-elementary, these are notoriously abstract. In \ohol{}, we view
-these as ordered generalizations of some very simple tautologies about
+elementary, these are notoriously abstract. In \ohol{}, we view these
+as ordered generalizations of some very simple tautologies about
 equality. For instance, the Yoneda lemma generalizes the equivalence
-between the formulae $\forall x. \forall y. x = y \Rightarrow P y$ and
-$P x$.
+between the formulae $\forall y. x = y \Rightarrow P y$ and $P x$ for
+any $x$.
 \begin{lemma} \label{lem:yoneda}
@@
   \item (Yoneda) The profunctor 
-    $\homrallXYtoZ {\alpha'} {\punitinXfromYtoZ {\cat C} {\alpha}{\alpha'}} {\posPresheafAppPtoX \pi {\alpha'}}$ is isomorphic to $\posPresheafAppPtoX \pi \alpha$
-  \item (Co-Yoneda) The profunctor $\tensorexistsXwithYandZ {\alpha'}{\posPresheafAppPtoX \pi \alpha'} {\punitinXfromYtoZ{}{\alpha'}{\alpha}}$ is isomorphic to $\posPresheafAppPtoX \pi \alpha$
+    $\homrallXYtoZ {\alpha'} {(\punitinXfromYtoZ {\cat C} {\alpha}{\alpha'})} {(\posPresheafAppPtoX \pi {\alpha'})}$ is isomorphic to $\posPresheafAppPtoX \pi \alpha$
+  \item (Co-Yoneda) The profunctor $\tensorexistsXwithYandZ {\alpha'}{(\posPresheafAppPtoX \pi \alpha')} {(\punitinXfromYtoZ{}{\alpha'}{\alpha})}$ is isomorphic to $\posPresheafAppPtoX \pi \alpha$
   \end{enumerate}
@@
 
-Next, we have the ``Fubini'' theorems, which are analogous to simple
-theorems relating tensor and hom types in ordered logic, and whose proofs are 
+Next, we have the ``Fubini'' theorems, which relate the tensor and hom
+types. The statement and proofs for these theorems are analogous to
+proofs relating tensor and hom in ordered logic. For instance, the
+second isomorphism below is analogous to the equivalence $(P \odot Q)
+\multimap R \cong P \multimap Q \multimap R$ in ordered logic.
 \begin{lemma}[{Fubini}]
@@
 \begin{proof}
-  Straightforward, for instance the forward direction of (1) is given
-  by $\lambdaunary \alpha {x} \delta {\tensorelimWkontZ x
-    {p.\beta.y. \tensorelimWkontZ y
-      {q.\gamma.r. \tensorintroatXwithYandZ {\gamma}
+  We show one case as an example, the forward direction of (1) is
+  given by $\lambdaunary \alpha {x} \delta {\tensorelimWkontZ x
+    {p,\beta,y. \tensorelimWkontZ y
+      {q,\gamma,r. \tensorintroatXwithYandZ {\gamma}
         {\tensorintroatXwithYandZ \beta p q} {r}}}}$
@@
 
-Next, we can prove that two definitions of an adjunction are equivalent:
+Next, we can prove that two definitions of an adjunction are
+equivalent:
 \begin{lemma} \label{lem:adjunction}
@@
 %% \subsection{Duality Principle}
-%% One common source of category theoretic arguments is the duality
+%% One common source of category-theoretic arguments is the duality
 %% principle: since the opposite of any category is also a category, any
@@
 %% %
-%% A similar duality principle holds in SCT.
+%% A similar duality principle holds in VETT.
 %% %
@@
 %% %
-%% We can define this as a translation $\cdot^* : SCT \to SCT$ defined on
-%% syntactic forms of SCT in Figure~\ref{fig:duality}.
+%% We can define this as a translation $\cdot^* : VETT \to VETT$ defined on
+%% syntactic forms of VETT in Figure~\ref{fig:duality}.
 %% \begin{figure}
@@
 %% The presheaf extensionality principle is independent from the existing
-%% axioms of \ohol{}: it is true in the ordinary category theoretic
+%% axioms of \ohol{}: it is true in the ordinary category-theoretic
 %% model, but it is false in the set theoretic model. That is, in the set
@@
 %
-First, we demonstrate that a hyperdoctrine of virtual equipments
-provides a sound and complete notion of categorical model.
+First, we define a sound and complete notion of categorical model
+based on hyperdoctrines of virtual equipments.
 %
@@
 First, we can model the judgmental structure of the unary type theory
-and predicate logic in certain \emph{virtual equipments}
-\cite{LEINSTER2002391,Cruttwell2010}, i.e., virtual double categories with
-restrictions. We briefly recount the structure present in a virtual
-double category, but see \citet{Cruttwell2010} for a precise
-definition of the composition rules for 2-cells and functor of virtual
-double categories.
-%
-Our notion here differs from \citet{Cruttwell2010} in several ways
-to more closely match our judgmental structure: (1) we require that
-restrictions are given as coherent structure rather than mere
-existence (2) we include a notion of ``small'' object and (3) we don't
-yet require the existence of unit horizontal arrows, as these are
-modeled by a connective rather than judgmental structure.
-
-\begin{definition}[Virtual Equipment]
-  A virtual equipment $\mathcal V$ consists of
+and predicate logic in \emph{virtual double categories} that are
+\emph{split fibrant} and have a notion of \emph{small object}
+\cite{LEINSTER2002391,Cruttwell2010}. We briefly recount the structure
+present in a virtual double category, but see \citet{Cruttwell2010}
+for a precise definition of the composition rules for 2-cells and
+functor of virtual double categories.
+
+\begin{definition}
+  A virtual double category $\mathcal V$ consists of
   \begin{enumerate}
   \item A category $V_o$ of ``objects and vertical arrows''
-  \item A chosen subset of the objects $V_s \subseteq V_o$ of ``small objects''
   \item A set ${\mathcal V}_h$ of ``horizontal arrows'' with source
@@
     composition:
-\[\begin{tikzcd}% https://q.uiver.app/?q=WzAsNixbMCwwLCJDXzAiXSxbMCwyLCJEXzAiXSxbMiwwLCJDX3tufSJdLFsxLDAsIlxcY2RvdHMiXSxbMiwyLCJEXzEiXSxbMSwxLCJcXHBoaSJdLFszLDIsIlJfbiIsMCx7InN0eWxlIjp7ImJvZHkiOnsibmFtZSI6ImJhcnJlZCJ9fX1dLFswLDEsIkYiLDJdLFsyLDQsIkciXSxbMSw0LCJTIiwyLHsic3R5bGUiOnsiYm9keSI6eyJuYW1lIjoiYmFycmVkIn19fV0sWzAsMywiUl8wIiwwLHsic3R5bGUiOnsiYm9keSI6eyJuYW1lIjoiYmFycmVkIn19fV1d
+\[\begin{tikzcd}[row sep=tiny]% https://q.uiver.app/?q=WzAsNixbMCwwLCJDXzAiXSxbMCwyLCJEXzAiXSxbMiwwLCJDX3tufSJdLFsxLDAsIlxcY2RvdHMiXSxbMiwyLCJEXzEiXSxbMSwxLCJcXHBoaSJdLFszLDIsIlJfbiIsMCx7InN0eWxlIjp7ImJvZHkiOnsibmFtZSI6ImJhcnJlZCJ9fX1dLFswLDEsIkYiLDJdLFsyLDQsIkciXSxbMSw0LCJTIiwyLHsic3R5bGUiOnsiYm9keSI6eyJuYW1lIjoiYmFycmVkIn19fV0sWzAsMywiUl8wIiwwLHsic3R5bGUiOnsiYm9keSI6eyJuYW1lIjoiYmFycmVkIn19fV1d
 	{C_0} & \cdots & {C_{n}} \\
@@
   \end{enumerate}
-  We say a virtual double category has \emph{split restrictions} when
-  for any horizontal arrow $R : C \pto D$ and vertical arrows $f : C'
-  \to C$ and $g : D' \to D$ there is a chosen horizontal arrow $R(f,g)
-  : C' \pto D'$ with a cartesian 2-cell to $R$ framed by $f,g$ and
-  these chosen cartesian lifts are functorial in $f,g$.
-
-  A functor of virtual equipments is a functor of the virtual double
-  categories that additionally preserves the restrictions and
-  smallness of objects.
+  We say a virtual double category is \emph{split fibrant} when it has
+  a choice of \emph{restrictions}, that is, for any horizontal arrow
+  $R : C \pto D$ and vertical arrows $f : C' \to C$ and $g : D' \to D$
+  there is a chosen horizontal arrow $R(f,g) : C' \pto D'$ with a
+  cartesian 2-cell to $R$ framed by $f,g$ and these chosen cartesian
+  lifts are functorial in $f,g$ (\cite{Shulman08}).  A \emph{choice of
+  small objects} is a subset of the objects $V_s \subseteq V_o$.  A
+  \emph{morphism} of split fibrant virtual double categories with
+  small objects is a functor of the virtual double categories that
+  additionally preserves the restrictions and smallness of objects.
+  This defines a category $\fVDC$.
 \end{definition}
@@
 In the presence of restrictions, every 2-cell can be represented as a
-``globular'' 2-cell where the left and right frame are identities. For
-example the 2-cell $\phi$ above can be represented as one with the
-same domain but whose codomain is $S(f,g)$. This property is crucial
-for the completeness of our semantics as we only include a syntax for these globular terms.
+``globular'' 2-cell where the left and right frame are identities
+\cite{Shulman08}. For example the 2-cell $\phi$ above can be
+represented as one with the same domain but whose codomain is
+$S(f,g)$. This property is crucial for the completeness of our
+semantics as we only include a syntax for these globular terms (proof
+of Construction~\ref{cons:syn-model}).
 %
@@
 have taken the restriction $S(F,G)$ of the codomain.
-
-To model the dependent type theory and fact that the categorical
-judgments are all typed under a context $\Gamma$ with an action of
-substitution, we use a variation on Lawvere's notion of
-\emph{hyperdoctrine} for modeling predicate logic\cite{Lawvere69}:
-\begin{definition}[Hyperdoctrine of Virtual Equipments]
-  A hyperdoctrine of virtual equipments (HVE) is a pair of a category with
+%
+Note that Cruttwell and Shulman define a \emph{virtual equipment} to
+be a virtual double category with all restrictions and all units. The
+units are the model of the unit of profunctors connective and so all
+of our models with the unit will be virtual equipments, hence the name
+VETT.
+
+To model the dependent type theory and indexing of category-theoretic
+judgments by a $\Gamma$ with an action of substitution, we use a
+variation on Lawvere's notion of \emph{hyperdoctrine} for modeling
+predicate logic\cite{Lawvere69}\footnote{note that unlike in
+hyperdoctrines, we do not require \emph{quantifiers} adjoint to
+substitution}:
+\begin{definition}[VETT Judgmental model]
+  A VETT judgmental model ($\VMJ$) is a pair of a category with
   families $\mathcal C$ and a functor $V^{(-)} : \mathcal C^o \to
-  \text{vEq}$ to the category of virtual equipments and functors.
+  \fVDC$.
 \end{definition}
@@
 well-behaved action of substitution.
-
-An HVE is then precisely the structure corresponding to the
-\emph{judgments} and actions of substitution in VETT.
+A $\VMJ$ is then precisely the structure corresponding to the
+judgments and actions of substitution in VETT.
 \begin{construction}[Syntactic Model]
   \label{cons:syn-model}
-  The syntax of VETT with any choice of which subset of connectives
-  are included presents an HVE.
+  The syntax of VETT with with any subset of connectives are included
+  presents a $\VMJ$.
 \end{construction}
 \begin{proof}
-  Straightforward, defining the category of families using the dependent
+  Define the category of families using the dependent
   type structure and the virtual equipment structure having
@@
   functors/sets as vertical/horizontal arrows and interpreting
-  compositions/restrictions as substitutions.  The biggest gap is in the
+  compositions/restrictions as substitutions.  The biggest gap between syntax and semantics is in the
   definition of the 2-cells.  A 2-cell from\\ $(\alpha_1:\cat
@@
 Then the \emph{connectives} of VETT each precisely correspond to a
-universal construction of structures in an HVE. The
-$\Pi,\Sigma,\text{Id}$ types correspond to their standard semantics in
-a CwF and the types of categories, functors etc as representing
-objects for the virtual equipment structures. Products of categories
-are interpreted as products in the vertical category, and products of
-sets as products in the category of pro-arrows and 2-cells. The units,
+universal construction in a $\VMJ$. The $\Pi,\Sigma,\text{Id}$ types
+correspond to their standard semantics in a CwF and the connectives
+for categories and profunctors correspond to universal constructions
+in the virtual double categories. Products of categories are
+interpreted as products in the vertical category, and products of sets
+as products in the category of pro-arrows and 2-cells. The units,
 tensor and covariant and contravariant homs are modeled by the
-universal properties of the same names, as described for instance in
+universal properties of the same names, as described in
 \citet{Shulman08}. The graph of a profunctor is modeled by
@@
 contravariant presheaf categories can be described as a weakening of
-the definition of a Yoneda equipment from \citet{diliberti-loregian} to
-virtual double categories. More detailed descriptions of these
-universal properties are included in the appendix.
-%
-Then we can package up the soundness and completeness of this notion
-of categorical model in the following modular initiality theorem.
-
+the definition of a Yoneda equipment from \citet{diliberti-loregian}
+to virtual double categories. More detailed descriptions of these
+universal properties are included in the extended version
+\cite{extended-version}.
+%
+Then the soundness and completeness of this notion of categorical
+model is formalized by the following initiality theorem.
 \begin{theorem}[Initiality]
-  The syntax of VETT with almost any choice of connectives presents an
-  HVE that is initial in the category of HVEs with the chosen
-  instances of the universal properties and functors that preserve
-  such chosen instances. The only exception is that if units are
-  included, then contravariant and covariant hom sets must be included
-  as well.
+  The syntax of VETT with any subset of connectives that includes the
+  hom types presents a $\VMJ$ that is initial in the category of
+  $\VMJ$ with the chosen instances of the universal properties and
+  functors that preserve such chosen instances.
 \end{theorem}
 \begin{proof}
-  It is straightforward to extend the construction
-  \ref{cons:syn-model} for any connective modularly, with the
-  exception that the unit relies on the presence of hom sets in order
-  to satisfy the ``distributivity'' requirement that its elimination
-  can occur in any context. Then we can construct the unique morphism
-  to any HVE by straightforward induction on syntax.
+  The construction \ref{cons:syn-model} can be extended for any
+  connective modularly, with the exception that the unit relies on the
+  presence of hom sets in order to satisfy the ``distributivity''
+  requirement that its elimination can occur in any context. Then we
+  can construct the unique morphism to any HVE induction on syntax.
 \end{proof}
@@
 %
-First, we can extend any virtual equipment to an HVE as follows:
+First, we can extend any set-theoretic model of the category theoretic
+judgments to a hyperdoctrine of models where the category of families
+is the category of sets:
 \begin{construction}
-  Given a virtual equipment $\mathcal V$, we
-  can construct a functor $\mathcal V^{-}: \Set \to \text{vDbl}_r$ by
-  defining the objects of $\mathcal V^\Gamma$ to be functions $\Gamma
-  \to \mathcal V_o$, and similarly for morphisms and 2-cells with all
-  operations given pointwise. 
+  Given a $\mathcal V \in \fVDC$, we can construct a $\VMJ$ $\mathcal
+  V^{-}: \Set \to \text{vDbl}_r$ by defining of $(\mathcal
+  V^\Gamma)_o$ to be functions $\mathcal V_o^\Gamma$, and similarly
+  for morphisms and 2-cells with all operations given pointwise.
 \end{construction}
@@
 %
+The ``standard model'' is the virtual double category of locally
+small categories where the small objects are the small categories.
 \begin{construction}
-  Fix a cardinal $\kappa$. The virtual equipment $\Cat_{\kappa}$ is
-  defined to have as objects locally $\kappa$-small categories, small
-  objects as $\kappa$-small categories, vertical morphisms as functors,
-  horizontal arrows as functors $\cat C^o \times \cat D \to \kappa Set$ and
-  2-cells as morphisms of profunctors. Restriction of profunctors is
-  given by composition, which is strictly associative and unital.
-  $\Cat_{U}$ has objects satisfying the universal properties of all
-  connectives in VETT.
+  Fix a cardinal $\kappa$. The virtual double category $\Cat_{\kappa}$
+  is defined to have as objects locally $\kappa$-small categories,
+  small objects as $\kappa$-small categories, vertical morphisms as
+  functors, horizontal arrows as functors $\cat C^o \times \cat D \to
+  \kappa Set$ and 2-cells as morphisms of profunctors. Restriction of
+  profunctors is given by composition, which is strictly associative
+  and unital.  $\Cat_{U}$ has objects satisfying the universal
+  properties of all connectives in VETT.
 \end{construction}
@@
 %
-We leave strictification theorems to future work.
-
-\begin{construction}[(\citet{Shulman13})]
+Shulman's construction carries over when the functor is strict but
+some of their example instances would require a strictification
+theorem.
+\begin{construction}[Shulman \citet{Shulman13}]
   Given any functor $\mathcal V : S^o \to \SymMonCat$ such that $S$
@@
 
+A final model that uses a CwF that is not $\Set$ would be given by
+taking extensional dependent type theory as the CwF and interpreting
+the category-theoretic constructions by their definitions inside type
+theory.
+
 \section{Related and Future Work}
@@
 
+C\'accamo and Winskel~\cite{caccamo_winskel_2001} develop a formal
+language for defining categories, functors (of many variables) and
+proving existence of natural equivalences between them. Their system
+can encode profunctors as functors into $\Set$. Their natural
+equivalence judgment does not have proof terms or equality between
+equivalences and they do not support natural
+transformations. Additionally, they only consider ordinary categories
+as the intended model and do not develop a more general semantics.
 Riehl and Verity~\cite{riehl_verity_2022} use a formal language of virtual
@@
 diagrams for monoidal categories. These are an
-alternative approach to type theoretic calculi, with the string
+alternative approach to type-theoretic calculi, with the string
 diagrams typically making tensor products simpler to work with, while
-a type theoretic calculus like SCT makes the closed structure
+a type-theoretic calculus like VETT makes the closed structure
 $\homrallXYtoZ \alpha P Q$ simpler to work with by using bound
@@
 
-One alternative to proarrow equipments is to axiomatize the bicategory
-of categories, profunctors and natural transformations rather than the
-virtual double category of categories, functors, profunctors and natural
-transformations. An abstract structure along these lines is a
-\emph{Cartesian
-  bicategory}~\citet{CARBONI198711}.
-Frey~\citet{frey} describes some preliminary work on
-a proof system for Cartesian bicateogires.  This calculus is more
-general than \vett{} in that their profunctors may have 0, 1 or more
-covariant or contravariant variables, but they do not have a term syntax
-or equational theory for the natural transformations/elements, and Cartesian bicategories do
-not have functors and restriction of profunctors by
-functors. Semantically, Cartesian bicategories are less general
-than virtual equipments because in general the set of functors between
-two categories can only be recovered from the set of profunctors if the
-codomain category is Cauchy complete, in which case functors are
-equivalent to adjoint pairs of profunctors.
+\emph{Cartesian bicategories} are similar to equipments but they
+axiomatize the bicategory of profunctors rather than the full double
+category of functors and profunctors \citet{CARBONI198711}.
+Frey~\citet{frey} describes preliminary work on a proof system for
+Cartesian bicateogires. Their profunctors are more general than in
+\vett{} in as they may have 0, 1 or more covariant or contravariant
+variables. But they do not have a term syntax for functors or natural
+transformations.
 
@@
 definable property of certain types.  Other type
-theories~\cite{north18,ahrens+22semantics-2dtt} define morphism types via an induction
-principle, corresponding to the lifting properties of certain kinds of
-fibrations of categories.  While these previous works can express some
-constructions on $\Cat$ that are not expressible in \vett{}, because
-\vett{} is more restricted, \vett{} contrariwise has more models, for
-instance categories enriched in non-cartesian monoidal categories, so
-the theorems that are provable in \vett{} apply in more settings.
+theories~\cite{north18,ahrens+22semantics-2dtt} define morphism types
+via an induction principle, corresponding to the lifting properties of
+certain kinds of fibrations of categories.  While these previous works
+can express some constructions on $\Cat$ that are not expressible in
+\vett{}, because \vett{} is more restricted, \vett{} contrariwise has
+more models, for instance categories enriched in non-cartesian
+monoidal categories, so the theorems that are provable in \vett{}
+apply in more settings.
 
@@
 logics. GTT~\cite{double-cats-gradual-typing} is a logic for
-\emph{vertically thin} pro-arrow equipments, i.e., pro-arrow equipments
-where the functor and natural transformation judgments are posetal. This
-calculus is quite different from ours in that the horizontal morphisms
-are terms of a programming language, and so they are not presented as
-relations with co-variant and contravariant variables, and the equipment
-structure is modeled by explicit coercions rather than by syntactic
-substitution of vertical arrows in horizontal arrows. Another similar
-calculus is System P~\cite{dunphyreddy} which is an internal language
-of \emph{reflexive graph categories}, which are like double categories
+\emph{vertically thin} pro-arrow equipments, where there is at most
+one vertical arrow or 2-cell of any tyepe, so their calculus does not
+include functor or transformation judgments. Another similar calculus
+is System P~\cite{dunphyreddy} which is an internal language of
+\emph{reflexive graph categories}, which are like double categories
 without horizontal composition.
@@
 define functor categories as function types, and incorporate
-multi-variable profunctors as in~\cite{frey}.  Similarly, ideas from
-coeffects and enriched category theory could be used for defining
-opposite categories \cite{Shulman18,coeffect}.
+multi-variable profunctors as in~\cite{frey}. This would require to
+the models to have a monoidal structure. Ideas from coeffects and
+enriched category theory may be useful for defining opposite
+categories \cite{Shulman18,coeffect}.
 
@@
 %% theory can be expressed using the simple combination of homs, tensors
-%% and units, there are several that are out of reach for SCT due to its
+%% and units, there are several that are out of reach for VETT due to its
 %% restrictive judgmental structure. These require extensions to the
@@
 
+\ifextended
 \appendix
@@
     (\punitrefl b)[a/\alpha] &= \punitrefl {b[a/\alpha]}\\
-    {\tensorelimWkontZ s {x.\beta.y.r}[\phi_l\jnctx\phi_m\jnctx\phi_r]} &=
-    \tensorelimWkontZ {s[\phi_m]} {x.\beta.y. r[\phi_l,x/x,\beta/\beta \jnctx \beta/\beta,y/y,\phi_r]}\\
+    {\tensorelimWkontZ s {x,\beta,y.r}[\phi_l\jnctx\phi_m\jnctx\phi_r]} &=
+    \tensorelimWkontZ {s[\phi_m]} {x,\beta,y. r[\phi_l,x/x,\beta/\beta \jnctx \beta/\beta,y/y,\phi_r]}\\
     {\tensorintroatXwithYandZ b s t}[\phi_s \jnctx \phi_t] &= \tensorintroatXwithYandZ {b[d^+\phi_s]} {s[\phi_s]} {t[\phi_t]}\\
@@
     {~}
-    {\tensorelimWkontZ {\tensorintroatXwithYandZ b s t} {x.\beta.y. r} = t[s/x;b/\beta;t/y]}
+    {\tensorelimWkontZ {\tensorintroatXwithYandZ b s t} {x,\beta,y. r} = t[s/x;b/\beta;t/y]}
 
@@
     {\Gamma\pipe\Phi,z:\tensorexistsXwithYandZ \beta P Q, \Psi \vdash s : R}
-    {\Gamma\pipe\Phi,z:\tensorexistsXwithYandZ \beta P Q, \Psi \vdash s = \tensorelimWkontZ s {x.\beta.y. s[\tensorintroatXwithYandZ \beta x y/z]}: \tensorexistsXwithYandZ \beta P Q}
+    {\Gamma\pipe\Phi,z:\tensorexistsXwithYandZ \beta P Q, \Psi \vdash s = \tensorelimWkontZ s {x,\beta,y. s[\tensorintroatXwithYandZ \beta x y/z]}: \tensorexistsXwithYandZ \beta P Q}
 
@@
     We can also define left and right composition for profunctors by
-    applying the profunctorial action to a reflexivity on one side or
+    applying the profunctorial action to an identity morphism on one side or
     the other:
@@
     \item Right to Left
-      \[ \lambdaunary {\alpha} k \delta \homllambdaXatYdotZ w \gamma \tensorelimWkontZ {w} {p.\beta.q. \homlappXtoYatZ {\homlappXtoYatZ k p \gamma} q \beta } \]
+      \[ \lambdaunary {\alpha} k \delta \homllambdaXatYdotZ w \gamma \tensorelimWkontZ {w} {p,\beta,q. \homlappXtoYatZ {\homlappXtoYatZ k p \gamma} q \beta } \]
     \item Left to Right to Left
       \begin{align*}
-        &\lambdaunary \alpha h \delta \homllambdaXatYdotZ w \gamma {\tensorelimWkontZ {w} {p.\beta.q.
+        &\lambdaunary \alpha h \delta \homllambdaXatYdotZ w \gamma {\tensorelimWkontZ {w} {p,\beta,q.
 (\homlappXtoYatZ {\homlappXtoYatZ {(\homllambdaXatYdotZ q \beta \homllambdaXatYdotZ p \gamma {\homlappXtoYatZ h {(p,\beta,q)} \gamma})} q \beta }  p \gamma) }}\\
-        &=\lambdaunary \alpha h \delta \homllambdaXatYdotZ w \gamma {\tensorelimWkontZ {w} {p.\beta.q.
+        &=\lambdaunary \alpha h \delta \homllambdaXatYdotZ w \gamma {\tensorelimWkontZ {w} {p,\beta,q.
 (\homlappXtoYatZ {{(\homllambdaXatYdotZ p \gamma {\homlappXtoYatZ h {(p,\beta,q)} \gamma})} }  p \gamma) }}\tag{contrahom$\beta$}\\
-        &=\lambdaunary \alpha h \delta \homllambdaXatYdotZ w \gamma {\tensorelimWkontZ {w} {p.\beta.q. {{({\homlappXtoYatZ h {(p,\beta,q)} \gamma})} } }}\tag{contrahom$\beta$}\\
+        &=\lambdaunary \alpha h \delta \homllambdaXatYdotZ w \gamma {\tensorelimWkontZ {w} {p,\beta,q. {{({\homlappXtoYatZ h {(p,\beta,q)} \gamma})} } }}\tag{contrahom$\beta$}\\
         &=\lambdaunary \alpha h \delta \homllambdaXatYdotZ w \gamma \homlappXtoYatZ h w \gamma\tag{tensor$\eta$}\\
@@
       \begin{align*}
-        &\lambdaunary \alpha k \delta \homllambdaXatYdotZ q \beta \homllambdaXatYdotZ p \gamma \homlappXtoYatZ {(\homllambdaXatYdotZ w \gamma {\tensorelimWkontZ {w} {p.\beta.q.
+        &\lambdaunary \alpha k \delta \homllambdaXatYdotZ q \beta \homllambdaXatYdotZ p \gamma \homlappXtoYatZ {(\homllambdaXatYdotZ w \gamma {\tensorelimWkontZ {w} {p,\beta,q.
 \homlappXtoYatZ {\homlappXtoYatZ k q \beta}  p \gamma }})} {(p,\beta,q)} \gamma\\
-        &=\lambdaunary \alpha k \delta \homllambdaXatYdotZ q \beta \homllambdaXatYdotZ p \gamma { {\tensorelimWkontZ { {(p,\beta,q)}} {p.\beta.q.
+        &=\lambdaunary \alpha k \delta \homllambdaXatYdotZ q \beta \homllambdaXatYdotZ p \gamma { {\tensorelimWkontZ { {(p,\beta,q)}} {p,\beta,q.
 \homlappXtoYatZ {\homlappXtoYatZ k q \beta } p \gamma }}}\\
@@
 \end{definition}
-
+\fi
 \end{document}
